### 2021-05-04 paper/authors.tex
stage manuscript

\correspondingauthor{L.\,G.\,Bouma}
\email{luke@astro.princeton.edu}

\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\author[0000-0002-2792-134X]{J. L. Curtis}
\affiliation{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY 10027, USA}
\affiliation{Department of Astrophysics, American Museum of Natural History, Central Park West, New York, NY 10024, USA}

### 2021-05-04 paper/ms.tex
stage manuscript

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\documentclass[12pt,twocolumn,tighten]{aastex63}
\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

\newcommand{\cn}{$\delta$~Lyr\ cluster} % cluster name
\newcommand{\sn}{Kepler\,1627} % star system name (binary)
\newcommand{\pn}{Kepler\,1627Ab} % plane name

% gaia target sample numbers
\newcommand{\nkinematic}{3{,}298} % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nnbhd}{13{,}843} % make_compstar_NGC_2516_sourcelist.py
\newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nhalo}{2{,}192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes

%
% Symbols
%
\newcommand{\kms}{\,km\,s$^{-1}$}
\newcommand{\ms}{\,m\,s$^{-1}$}
\newcommand{\bpmrpo}{(G_{\rm BP}-G_{\rm RP})_0}
\newcommand{\bpmrp}{G_{\rm BP}-G_{\rm RP}}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{---}
\revised{---}
\accepted{---}
\submitjournal{Nature}
\shorttitle{Kepler\,1627}

\begin{document}

\defcitealias{cantatgaudin_gaia_2018}{CG18}
\defcitealias{kounkel_untangling_2019}{KC19}
\defcitealias{meingast_2021}{M21}

% Cluster Difference Imaging Photometric Survey. III.
% Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs
\title{
  A 30 Million Year Old Mini-Neptune in the Kepler Field
}

%\suppressAffiliations
%\NewPageAfterKeywords
\input{authors.tex}

\begin{abstract}
  The Gaia satellite is revitalizing our understanding of nearby
  open clusters and moving groups.  Here, we focus on the
  underappreciated \cn.  Based on rotation periods from TESS and lithium
  measurements from ground-based spectrographs, we find the age of the
  cluster to be $30\pm XX$\,Myr.  \sn\ is a binary system in the
  cluster, serendipitously observed by the Kepler satellite because
  the primary is nearby and Sun-like.  Kepler\,1627A was found to host
  a $3.7\pm X.X\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  We
  re-validate the existence of \pn, and cement it as the youngest
  planet with a well-measured age observed by the main Kepler mission.
  Newly derived ages from Gaia offer the opportunity to significantly
  expand the census of age-dated planets.  The properties of \pn\ are
  may also help clarify how the orbits and atmospheres of the
  mini-Neptune planets evolve.
\end{abstract}

\keywords{
  planetary evolution (XXXX),
  stellar associations (1582),
  open star clusters (1160),
	stellar ages (1581),
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

At the time of the main Kepler
mission (2009--2013), only four open clusters were known in the Kepler
field: NGC\,6866, NGC\,6811, NGC\,6819,
and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.

Section~\ref{sec:first}.  Section~\ref{sec:second}.  In
Section~\ref{sec:discussion}, we discuss.
Section~\ref{sec:conclusion} gives our conclusions.

\section{first}
\label{sec:first}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{
  {\bf Galactic position and tangential velocities of the
  $\delta$\,Lyr cluster (also known as Theia 73 and Stephenson\,1).}  Points are
  candidate cluster members with $\varpi/\sigma_\varpi > 20$, reported
  to be in the group by \citet{kounkel_untangling_2019}.  We focus
  on stars in a small region (black points) in the kinematic
  vicinity of Kepler\,1627 (yellow star).  The other candidate cluster
  members (gray points) may or may not share the ages of the selected
  kinematic group.  The location of the Sun is ($\odot$) is shown.
  \label{fig:XYZvtang}
  }
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{
  {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored points are
  kinematically selected members of the $\delta$\,Lyr cluster (black points in
  Figure~\ref{fig:XYZvtang}).  Both Kepler (gray panels) and TESS (colored
  points) observed members of the cluster.  Gray points are naked-eye stars
  ($m_{\rm V}<6.5$), three of which are annotated.  Kepler\,1627 (green star) was
  observed for two TESS sectors, and during the entirety of the Kepler mission.
  \label{fig:skychart}
  }
\end{figure*}

\section{second}
\label{sec:second}

\section{Discussion}
\label{sec:discussion}

\section{Conclusion}
\label{sec:conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\clearpage
\acknowledgements
\raggedbottom

L.G.B{.} is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
K{.}~Hawkins for the discussions on young stars, rotation, and lithium
that encouraged this analysis.
%
L.G.B{.} and J.H{.} acknowledge support by the TESS GI Program, programs
G011103 and G022117, through NASA grants 80NSSC19K0386 and
80NSSC19K1728.
%
L.G.B{.} was also supported by a Charlotte Elizabeth Procter Fellowship
from Princeton University.
%
This study was based in part on observations at Cerro Tololo
Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
Association of Universities for Research in Astronomy (AURA) under a
cooperative agreement with the National Science Foundation.
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%

%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

% %
% This research made use of the NASA Exoplanet Archive, which is
% operated by the California Institute of Technology, under contract
% with the National Aeronautics and Space Administration under the
% Exoplanet Exploration Program.
% %

% Resources supporting this work were provided by the NASA High-End
% Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
% Division at Ames Research Center for the production of the SPOC data
% products.
%

% A.J.\ and R.B.\ acknowledge support from project IC120009 ``Millennium
% Institute of Astrophysics (MAS)'' of the Millenium Science Initiative,
% Chilean Ministry of Economy. A.J.\ acknowledges additional support
% from FONDECYT project 1171208.  J.I.V\ acknowledges support from
% CONICYT-PFCHA/Doctorado Nacional-21191829.  R.B.\ acknowledges support
% from FONDECYT Post-doctoral Fellowship Project 3180246.
% %
% C.T.\ and C.B\ acknowledge support from Australian Research Council
% grants LE150100087, LE160100014, LE180100165, DP170103491 and
% DP190103688.
% %
% C.Z.\ is supported by a Dunlap Fellowship at the Dunlap Institute for
% Astronomy \& Astrophysics, funded through an endowment established by
% the Dunlap family and the University of Toronto.
% %
% D.D.\ acknowledges support through the TESS Guest Investigator Program
% Grant 80NSSC19K1727.
%
%
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

\software{
  %\texttt{arviz} \citep{arviz_2019},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  %\texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
   	exoplanet:theano},
	%\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  \texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  \texttt{TESS-point}  \citep{burke_2020},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}
\

\facilities{
 	{\it Astrometry}:
 	Gaia \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2020_edr3}.
 	{\it Imaging}:
    Second Generation Digitized Sky Survey. %,
    %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
 	%Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	{\it Spectroscopy}:
	CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
  %PFS ({\bf CITE}),
  %  MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
	AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
 	%Keck:I~(HIRES; \citealt{vogt_hires_1994}).
 	VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
%	  ASTEP:0.40$\,$m (ASTEP400),
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
%	  El Sauce:0.356$\,$m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
% 	NTT,
% 	SOAR (SOI),
 	TESS \citep{ricker_transiting_2015}.
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
}

% \input{maintable_header.tex}
% \input{TOI837_phot_table.tex}
% \input{TOI837_rv_table.tex}
% \input{ic2602_ages.tex}
% \input{starparams.tex}
% \input{compparams.tex}
% \input{model_posterior_table.tex}

\clearpage
\bibliographystyle{yahapj}                            
\bibliography{bibliography}

\appendix
\section{first appendix}
\label{app:first}

\listofchanges

%\allauthors
\end{document}

### 2021-05-04 paper/collect.sh

@@ -1,0 +1,3 @@
+cp /Users/luke/Dropbox/proj/rudolf/results/XYZvtang/XYZvtang_show1627_showcomovers_showsun.pdf f1.pdf
+
+cp /Users/luke/Dropbox/proj/rudolf/results/skychart/skychart_showkepler_showtess_shownakedeye_showcomovers.pdf f2.pdf

### 2021-05-11 paper/ms.tex
add metadata table writer

@@
   open clusters and moving groups.  Here, we focus on the
-  underappreciated \cn.  Based on rotation periods from TESS and lithium
-  measurements from ground-based spectrographs, we find the age of the
-  cluster to be $30\pm XX$\,Myr.  \sn\ is a binary system in the
-  cluster, serendipitously observed by the Kepler satellite because
-  the primary is nearby and Sun-like.  Kepler\,1627A was found to host
-  a $3.7\pm X.X\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  We
-  re-validate the existence of \pn, and cement it as the youngest
-  planet with a well-measured age observed by the main Kepler mission.
-  Newly derived ages from Gaia offer the opportunity to significantly
-  expand the census of age-dated planets.  The properties of \pn\ are
-  may also help clarify how the orbits and atmospheres of the
-  mini-Neptune planets evolve.
+  underappreciated \cn.  Based on rotation periods from TESS and
+  lithium measurements from ground-based spectrographs, we find the
+  age of the cluster to be $30\pm XX$\,Myr.  \sn\ is a binary system
+  in the cluster, serendipitously observed by the Kepler satellite
+  because the primary is nearby and Sun-like.  Kepler\,1627A was
+  previously found to host a $3.7\pm X.X\,R_\oplus$ mini-Neptune on a
+  7.2\,day orbit.  We re-validate the existence of \pn, and cement it
+  as the youngest planet with a well-measured age observed by the main
+  Kepler mission.  Newly derived ages from Gaia offer the opportunity
+  to significantly expand the census of age-dated planets -- we offer
+  a literature compilation of young stars to enable this expansion.
+  The properties of \pn\ are may also help clarify how the orbits and
+  atmospheres of the mini-Neptune planets evolve.
 \end{abstract}
@@
 
-% \input{maintable_header.tex}
 % \input{TOI837_phot_table.tex}
@@
 \appendix
-\section{first appendix}
-\label{app:first}
+\section{Young, Age-Dated, and Age-Dateable Star Compilation}
+\label{app:targetlist}
+
+\input{table1.tex}
+
+The \texttt{v0.5} CDIPS target catalog includes some important updates
+from previous versions.  As in \citet{bouma_cdipsI_2019}, we collected
+membership information for young, age-dated, or age-dateable stars
+from across the literature.  Table~\ref{tab:metadata} gives a list of
+the sources included, and some brief summary statistics.
+
+The first major important change is that the extent of analyses
+performed on the Gaia data at the time of our compilation was wide and
+deep enough that we opted to neglect pre-Gaia analyses, except in
+cases for which spectroscopically confirmed samples of stars had been
+collected.  The membership lists for instance of
+\citet{Kharchenko_et_al_2013} and \citet{dias_proper_2014} (MWSC and
+DAML) were no longer required, especially given their relatively high
+field-star contamination rates compared to Gaia-derived membership
+catalogs.
+
+For any of the catalogs for which Gaia DR2 identifiers were not
+immediately available, we either followed the spatial (plus
+proper-motion) crossmatching procedures described in
+\citet{bouma_cdipsI_2019}, or else we pulled the Gaia DR2 source
+identifiers associated with the catalog from SIMBAD.  We consequently
+opted to drop the $\texttt{ext\_catalog\_name}$ and $\texttt{dist}$
+columns maintained in \citet{bouma_cdipsI_2019}, as these were only
+populated for a small number of stars.
+
+The most crucial parameters of a given star for our purposes are the
+Gaia DR2 source identifier ($\texttt{source\_id}$), the cluster name
+($\texttt{cluster}$), and the ($\texttt{age}$).  Given the
+hierarchical nature of many stellar associations, we do not attempt to
+resolve the cluster names to a single unique string.  The Orion
+complex for instance, can be divided into almost one hundred kinematic
+subgroups \citep{kounkel_apogee2_2018}.  Similar complexity applies to
+the problem of determining homogeneous ages, which we do not attempt
+to resolve.  Instead, we simply merged the cluster names and ages
+reported by various authors together.
+
+This means that our ``age'' column can be null, for cases in which the
+original authors did not report an age, and a reference literature age
+was not readily available.  Nonetheless, since we do generally prefer
+stars with known ages, we made a few additional efforts to populate
+this column.  When available, the age provenance is from the original
+analysis of the cluster.  However, in a few cases we adopted other
+ages when the string-based crossmatches on the ``cluster'' name was
+straightforward.  In particular, we used the ages determined by
+\citet{CantatGaudin2020b} to assign ages to the catalogs from
+\citet{GaiaCollaboration2018}, \citet{CantatGaudin2018a},
+\citet{CastroGinard2020}, and \citet{CantatGaudin2020a}.
+
+The catalogs we included for which ages were not immediately available
+were those of \citet{CottenSong2016}, \citet{Oh2017},
+\citet{Zari2018}, \citet{Gagne2018a},
+\citet{Gagne2018b},\citet{Gagne2018c}, and \citet{Ujjwal2020}.  While
+in principle the moving group members discussed by
+\citet{Gagne2018a,Gagne2018b,Gagne2018c} and \citet{Ujjwal2020} have
+easily associated ages, our SIMBAD cross-matching lost the moving
+group association from those studies, which should therefore be
+recovered tools such as BANYAN
+$\Sigma$.\footnote{\url{http://www.exoplanetes.umontreal.ca/banyan/banyansigma.php}}.
+We also included the SIMBAD object identifiers $\texttt{TT*}$,
+$\texttt{Y*O}, $\texttt{Y*?}, $\texttt{TT?}$, and $\texttt{pMS*}$.
+Finally, we also included every star in the NASA Exoplanet Archive
+$\texttt{ps}$ table that had a Gaia identifier available
+\citep{NASAExoArchive_ps_20210506}.  If the age had finite
+uncertainties, we also included it, since stellar ages determined
+through the combination of isochrone-fitting and transit-derived
+stellar densities typically have higher precision than from isochrones
+alone.
+
+The technical manipulations for the merging, cleaning, and joining
+were performed using $\texttt{pandas}$
+\citep{mckinney-proc-scipy-2010}.  The eventual crossmatch (using the
+Gaia DR2 $\texttt{source\_id}$) against the Gaia DR2 archive was performed
+asychronously on the Gaia archive
+website\footnote{\url{https://gea.esac.esa.int/archive/}}.
 

### 2021-05-11 drivers/write_metadata_table.py

@@ -1,0 +1,104 @@
+import numpy as np, pandas as pd
+import re, os
+# https://ads.readthedocs.io/en/latest/#getting-started
+import ads
+
+from rudolf.paths import PAPERDIR
+
+csvpath = '/Users/luke/Dropbox/proj/cdips/data/cluster_data/list_of_lists_keys_paths_assembled_v0.5_gaiasources.csv'
+df = pd.read_csv(csvpath)
+
+outcols = (
+    'reference_id,bibcode,N_gaia,Nstars_with_age,N_Rplt16_orclose'
+)
+outcols = outcols.split(',')
+
+sdf = df[outcols]
+
+# Don't mention that the table actually includes HATS and HATN candidates.
+sel = (sdf.reference_id != 'HATSandHATNcandidates20210505')
+sdf = sdf[sel].reset_index()
+
+#
+# First get all the bibtex information from ADS.
+# replace "ARTICLE{2014AJ...}" style with reference_ids
+#
+assert np.all(~pd.isnull(sdf.bibcode))
+bibcodes = list(sdf.bibcode)
+bibcodes = [b.replace("A%26A","A&A") for b in bibcodes]
+
+print('beginning ads query')
+r = ads.ExportQuery(bibcodes).execute()
+
+MAPDICT = {
+    'candYSO':r'$\texttt{Y*?}$',
+    'YSO':r'$\texttt{Y*O}$',
+    'PMS':r'$\texttt{pMS*}$',
+    'TTS':r'$\texttt{TT*}$',
+    'candTTS':r'$\texttt{TT?}$',
+    'gt250':r'$d>250\,{\rm pc}$',
+    'lt250':r'$d>250\,{\rm pc}$',
+    'ums':'UMS',
+    'pms':'PMS',
+    'PscEri': 'Psc-Eri',
+    'Pleiades': 'Pleiades',
+    't1': 'Coma-Ber',
+    't2': 'Neighbor Group',
+    '':''
+}
+
+ref_ids, scnds = [], []
+
+for _b, _id in zip(bibcodes, list(sdf.reference_id)):
+
+    if _id.startswith("SIMBAD_"):
+        scnd = MAPDICT[_id.split("SIMBAD_")[1]]
+        _id = "SIMBAD"
+    elif _id.startswith("GaiaCollaboration2018"):
+        scnd = MAPDICT[_id.split("GaiaCollaboration2018")[1]]
+        _id = "GaiaCollaboration2018"
[…]

### 2021-05-11 paper/ms.tex
update ms

@@
 
-%\documentclass[12pt,twocolumn,tighten]{aastex63}
-\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
+\documentclass[12pt,twocolumn,tighten]{aastex63}
+%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
@@
 		\leavevmode
-		\includegraphics[width=1\textwidth]{f1.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{
-  {\bf Galactic position and tangential velocities of the
-  $\delta$\,Lyr cluster (also known as Theia 73 and Stephenson\,1).}  Points are
-  candidate cluster members with $\varpi/\sigma_\varpi > 20$, reported
-  to be in the group by \citet{kounkel_untangling_2019}.  We focus
-  on stars in a small region (black points) in the kinematic
-  vicinity of Kepler\,1627 (yellow star).  The other candidate cluster
-  members (gray points) may or may not share the ages of the selected
-  kinematic group.  The location of the Sun is ($\odot$) is shown.
-  \label{fig:XYZvtang}
-  }
-\end{figure*}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f2.pdf}
+		\includegraphics[width=0.99\textwidth]{f2.pdf}
 	\end{center}
@@
   Figure~\ref{fig:XYZvtang}).  Both Kepler (gray panels) and TESS (colored
-  points) observed members of the cluster.  Gray points are naked-eye stars
+  points) observed portions of the cluster.  Gray points are naked-eye stars
   ($m_{\rm V}<6.5$), three of which are annotated.  Kepler\,1627 (green star) was
-  observed for two TESS sectors, and during the entirety of the Kepler mission.
+  observed during the entirety of the Kepler mission, and has (so far) been observed
+  for two lunar months by TESS.
   \label{fig:skychart}
@@
 
-\input{table1.tex}
+\input{t1_appendix.tex}
+\input{t2_appendix.tex}
 
-The \texttt{v0.5} CDIPS target catalog includes some important updates
-from previous versions.  As in \citet{bouma_cdipsI_2019}, we collected
-membership information for young, age-dated, or age-dateable stars
-from across the literature.  Table~\ref{tab:metadata} gives a list of
-the sources included, and some brief summary statistics.
+The \texttt{v0.5} CDIPS target catalog (Table~\ref{tab:v05}) includes
+some important updates from previous versions.  As in
+\citet{bouma_cdipsI_2019}, we collected membership information for
+young, age-dated, or age-dateable stars from across the literature.
+Table~\ref{tab:metadata} gives a list of the sources included, and
+some brief summary statistics.
 
@@
 
+\section{Kinematic Selection of $\delta$\,Lyr Cluster Members}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f1.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Galactic position and tangential velocities of the
+			$\delta$\,Lyr cluster (also known as Theia 73 and Stephenson\,1).}  Points are
+		candidate cluster members with $\varpi/\sigma_\varpi > 20$, reported
+		to be in the group by \citet{kounkel_untangling_2019}.  We focus
+		on stars in a small region (black points) in the kinematic
+		vicinity of Kepler\,1627 (yellow star).  The other candidate cluster
+		members (gray points) may or may not share the ages of the selected
+		kinematic group.  The location of the Sun is ($\odot$) is shown.
+		\label{fig:XYZvtang}
+	}
+\end{figure*}
+
+Figure~\ref{fig:XYZvtang} shows members of the \cn\
+reported
+by \citet{kounkel_untangling_2019} to be in the group.
+Galactic positions are determined and plotted only for stars with parallax
+signal-to-noise exceeding 20.
+The location of the Sun is shown on the plots.
+The non-uniform ``clumps'' might be
+an artifact of the data processing steps performed by \citet{kounkel_untangling_2019}.
+We therefore only consider stars in the immediate kinematic group
+around \sn. 
+The tangential velocities relative to \sn\ are shown in the bottom right panel.
+These are computed by assuming that every star has the same three-dimensional
+spatial velocity as \sn, where we assume a systemic radial velocity of
+$-16.7 \pm 0.2$\,\kms\ based on the reconnaissance spectra obtained by
+A.~Howard on HIRES and D.~Latham on TRES.
+The relevant projection effects are then taken into account,
+as discussed by {\it e.g.}, \citet{Meingast2021} and {\bf L.~Bouma et al (2021, submitted)}.
+
 \listofchanges

### 2021-05-11 paper/t1_appendix.tex
update ms

%% \begin{deluxetable}{} command tell LaTeX how many columns
%% there are and how to align them.
%\startlongtable
\begin{deluxetable*}{lll}

%% Keep a portrait orientation

%% Over-ride the default font size
%% Use Default (12pt)
\tabletypesize{\scriptsize}
%\tabletypesize{\small}

%% Use \tablewidth{?pt} to over-ride the default table width.
%% If you are unhappy with the default look at the end of the
%% *.log file to see what the default was set at before adjusting
%% this value.

%% This is the title of the table.
\tablecaption{Young, Age-dated, and Age-dateable Stars Within the
  Nearest Few Kiloparsecs ($\texttt{v0.5}$ of the CDIPS Target
  List).}
\label{tab:v05}

%% This command over-rides LaTeX's natural table count
%% and replaces it with this number.  LaTeX will increment 
%% all other tables after this table based on this number
%\tablenum{3}

%% The \tablehead gives provides the column headers.  It
%% is currently set up so that the column labels are on the
%% top line and the units surrounded by ()s are in the 
%% bottom line.  You may add more header information by writing
%% another line between these lines. For each column that requries
%% extra information be sure to include a \colhead{text} command
%% and remember to end any extra lines with \\ and include the 
%% correct number of &s.
\tablehead{
  \colhead{Parameter} &
  \colhead{Example Value} &
  \colhead{Description}
}

%% All data must appear between the \startdata and \enddata commands
%
% paste from
% /Users/luke/Dropbox/proj/rudolf/results/tables/v05_main_tableheader.tex
% via drivers/write_v05_main_tableheader.py
\startdata
         \texttt{source\_id} &                                         1709456705329541504 &                                             Gaia DR2 source identifier. \\
                 \texttt{ra} &                                                     247.826 &                                         Gaia DR2 right ascension [deg]. \\
                \texttt{dec} &                                                      79.789 &                                             Gaia DR2 declination [deg]. \\
           \texttt{parallax} &                                                      35.345 &                                                Gaia DR2 parallax [mas]. \\
    \texttt{parallax\_error} &                                                       0.028 &                                    Gaia DR2 parallax uncertainty [mas]. \\
               \texttt{pmra} &                                                      94.884 &     Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr${^-1}$]. \\
              \texttt{pmdec} &                                                     -86.971 &                 Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr${^-1}$]. \\
 \texttt{phot\_g\_mean\_mag} &                                                        6.85 &                                                 Gaia DR2 $G$ magnitude. \\
\texttt{phot\_bp\_mean\_mag} &                                                       6.409 &                                     Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
\texttt{phot\_rp\_mean\_mag} &                                                       7.189 &                                     Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
            \texttt{cluster} &                 Uma,IR\_excess,NASAExoArchive\_ps\_20210506 &                                  Comma-separated cluster or group name. \\
                \texttt{age} &                                                nan,nan,9.48 & Comma-separated logarithm (base-10) of reported$^{\rm a}$ age in years. \\
          \texttt{mean\_age} &                                                        9.48 &                          Mean (ignoring NaNs) of $\texttt{age}$ column. \\
      \texttt{reference\_id} &      Ujjwal2020,CottenSong2016,NASAExoArchive\_ps\_20210506 &                          Comma-separted provenance of group membership. \\
 \texttt{reference\_bibcode} & 2020AJ....159..166U,2016ApJS..225...15C,2013PASP..125..989A &                  ADS bibcode corresponding to $\texttt{reference\_id}$. \\
\enddata

%% Include any \tablenotetext{key}{text}, \tablerefs{ref list},
%% or \tablecomments{text} between the \enddata and 
%% \end{deluxetable} commands

%% General table comment marker
\tablecomments{
Table~\ref{tab:v05} is
published in its entirety in a machine-readable format.   This table is a
concatenation of the studies listed in Table~\ref{tab:metadata}.
One entry is
shown for guidance regarding form and content. 
In this particular example, the star has a
cold Jupiter on a 16 year orbit, HD 150706b \citep{2012AA...545A..55B}.
An infrared excess has been reported \citep{CottenSong2016}, and the star was identified by \citet{Ujjwal2020} as a candidate UMa moving group
member ($\approx 400\,{\rm Myr}$; \citealt{mann_tess_2020}).
The star's RV activity and TESS rotation period corroborate its youth.
}
\vspace{-0.5cm}
\end{deluxetable*}

### 2021-05-11 drivers/write_metadata_table.py

@@ -3,7 +3,7 @@ import re, os
 # https://ads.readthedocs.io/en/latest/#getting-started
 import ads
 
-from rudolf.paths import PAPERDIR
+from rudolf.paths import RESULTSDIR
 
 csvpath = '/Users/luke/Dropbox/proj/cdips/data/cluster_data/list_of_lists_keys_paths_assembled_v0.5_gaiasources.csv'
 df = pd.read_csv(csvpath)
@@ -98,7 +98,7 @@ outdf = pd.DataFrame({
     'N_{\mathrm{Age}}': sdf.Nstars_with_age,
     'N_{G_\mathrm{RP}<16}': sdf.N_Rplt16_orclose
 })
-outpath = os.path.join(PAPERDIR, 'metadata_table_data.tex')
+outpath = os.path.join(RESULTSDIR, 'tables', 'metadata_table_data.tex')
 outdf.to_latex(outpath, index=False, escape=False)
 print(f'Wrote {outpath}')
 

### 2021-05-11 drivers/write_v05_main_tableheader.py

@@ -1,0 +1,63 @@
+"""
+Write the header for the v05 CDIPS target list.
+"""
+import os
+from numpy import array as nparr
+import numpy as np, pandas as pd
+
+from cdips.utils.catalogs import get_cdips_pub_catalog
+from rudolf.paths import RESULTSDIR
+
+df = get_cdips_pub_catalog(ver=0.5)
+
+#
+# make the header table for the paper
+#
+namedesc_dict = {
+'source_id': "Gaia DR2 source identifier.",
+'ra': "Gaia DR2 right ascension [deg].",
+'dec': "Gaia DR2 declination [deg].",
+'parallax': "Gaia DR2 parallax [mas].",
+'parallax_error': "Gaia DR2 parallax uncertainty [mas].",
+'pmra': r"Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr$^{-1}$].",
+'pmdec': "Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr$^{-1}$].",
+'phot_g_mean_mag': "Gaia DR2 $G$ magnitude.",
+'phot_bp_mean_mag': "Gaia DR2 $G_\mathrm{BP}$ magnitude.",
+'phot_rp_mean_mag': "Gaia DR2 $G_\mathrm{RP}$ magnitude.",
+'cluster': "Comma-separated cluster or group name.",
+'age': r"Comma-separated logarithm (base-10) of reported$^{\rm a}$ age in years.",
+'mean_age': r"Mean (ignoring NaNs) of $\texttt{age}$ column.",
+'reference_id': "Comma-separted provenance of group membership.",
+'reference_bibcode': r"ADS bibcode corresponding to $\texttt{reference\_id}$.",
+}
+
+keys = list(namedesc_dict.keys())
+keys = ["\\texttt{"+k.replace("_", "\_")+"}" for k in keys]
+
+#sel = df.source_id == np.int64(3311804515502788352)
+sel = df.source_id == np.int64(1709456705329541504)
+_vals = df[sel].T.values.flatten()
+vals = []
+for v in _vals:
+    if isinstance(v, float):
+        v = np.round(v,3)
+    elif isinstance(v, str):
+        v = str(v).replace("_", "\_").replace("%","\%")
+    else:
+        v = v
+    vals.append(v)
+
+descrs = list(namedesc_dict.values())
+
+df_totex = pd.DataFrame({
+    'Parameter': keys,
[…]

### 2021-05-13 paper/ms.tex
MAP soln looking good, begin sampling

@@
 
-L.G.B{.} is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
-E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
-K{.}~Hawkins for the discussions on young stars, rotation, and lithium
-that encouraged this analysis.
+L.G.B{.} and J.L.C{.} are grateful to T{.}~David for help with the
+transit fitting.
+% is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
+% E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
+% K{.}~Hawkins for the discussions on young stars, rotation, and lithium
+% that encouraged this analysis.
 %
-L.G.B{.} and J.H{.} acknowledge support by the TESS GI Program, programs
+L.G.B{.} acknowledges support by the TESS GI Program, programs
 G011103 and G022117, through NASA grants 80NSSC19K0386 and
@@
 %
-L.G.B{.} was also supported by a Charlotte Elizabeth Procter Fellowship
-from Princeton University.
+L.G.B{.} was also supported by a Charlotte Elizabeth Procter
+Fellowship from Princeton University.
 %

### 2021-05-13 drivers/run_gptransit.py

@@ -53,10 +53,6 @@ def run_gptransit(starid='Kepler_1627', N_samples=1000):
     # flags.
     sel = (qual == 0)
 
-    # FIXME: to make things manageable
-    sel &= ( (time > 200) & (time < 350) )
-
-
     datasets['keplerllc'] = [time[sel], flux[sel], flux_err[sel], texp]
 
     priorpath = join(DATADIR, 'priors', f'{starid}_priors.py')

### 2021-05-15 drivers/run_gptransit.py
posterior table: looks decent

@@ -74,7 +74,7 @@ def run_gptransit(starid='Kepler_1627', N_samples=500):
                     N_cores=os.cpu_count())
 
     var_names = [
-        'mean','logg_star','t0','period','log_jitter',
+        'mean','logg_star','t0','period','log_r','log_jitter',
         'log_prot','log_Q0','log_dQ','r_star','rho_star','u[0]','u[1]', 'r',
         'b', 'ecc', 'omega', 'sigma', 'rho', 'sigma_rot', 'prot', 'f',
         'r_planet', 'a_Rs', 'cosi', 'sini','T_14','T_13'
@@ -86,11 +86,6 @@ def run_gptransit(starid='Kepler_1627', N_samples=500):
                         kind='stats', stat_funcs={'median':np.nanmedian},
                         extend=True)
 
-    #FIXME FIXME gotta fix posterior table!
-    #FIXME FIXME gotta fix posterior table!
-    #FIXME FIXME gotta fix posterior table!
-    #FIXME FIXME gotta fix posterior table!
-
     fitindiv = 1
     phaseplot = 1
     cornerplot = 1
@@ -98,7 +93,8 @@ def run_gptransit(starid='Kepler_1627', N_samples=500):
 
     if posttable:
         outpath = join(PLOTDIR, f'{starid}_{modelid}_posteriortable.tex')
-        make_posterior_table(pklpath, priordict, outpath, modelid, makepdf=1)
+        make_posterior_table(pklpath, priordict, outpath, modelid, makepdf=1,
+                             var_names=var_names)
 
     if phaseplot:
         outpath = join(PLOTDIR, f'{starid}_{modelid}_phaseplot.png')

### 2021-05-15 paper/ms.tex
posterior table: looks decent
update ms

@@
   because the primary is nearby and Sun-like.  Kepler\,1627A was
-  previously found to host a $3.7\pm X.X\,R_\oplus$ mini-Neptune on a
-  7.2\,day orbit.  We re-validate the existence of \pn, and cement it
+  previously found to host a $3.56\pm 0.04\,R_\oplus$ mini-Neptune on a
+  7.2\,day orbit.  We re-validate the existence of \pn, cementing it
   as the youngest planet with a well-measured age observed by the main
@@
 \label{sec:second}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f5.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Phase-folded transit of Kepler 1627b, with stellar
+    variability removed.}  
+    The 1-$\sigma$ model uncertainties (faint purple band) and the
+    median model are shown.
+		\label{fig:phasefold}
+	}
+\end{figure*}
 
@@
 % \input{compparams.tex}
-% \input{model_posterior_table.tex}
+
+\input{model_posterior_table.tex}
 

### 2021-05-18 paper/ms.tex
update ms

@@
 \newcommand{\sn}{Kepler\,1627} % star system name (binary)
-\newcommand{\pn}{Kepler\,1627Ab} % plane name
+\newcommand{\pn}{Kepler\,1627Ab} % planet name
 
@@
 \label{sec:second}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f4.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf \sn\ light curve over 100 days.}  
+    The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
+    30 minute cadence.  The {\it top panel} shows the largest
+    signal, induced by starspots and plages.  The model for the
+    stellar variability (orange line) is subtracted in the {\it
+    bottom panel}, highlighting the transits of \pn, as well as other
+    deviations from the stellar variability model.  The 
+    Figure Set available online shows the entire 3.9 years of
+    available data.
+		\label{fig:lightcurve}
+	}
+\end{figure*}
 

### 2021-05-18 paper/ms.tex
fine-tune del Lyr skyplot

@@
   \caption{
-  {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored points are
-  kinematically selected members of the $\delta$\,Lyr cluster (black points in
-  Figure~\ref{fig:XYZvtang}).  Both Kepler (gray panels) and TESS (colored
-  points) observed portions of the cluster.  Gray points are naked-eye stars
-  ($m_{\rm V}<6.5$), three of which are annotated.  Kepler\,1627 (green star) was
-  observed during the entirety of the Kepler mission, and has (so far) been observed
-  for two lunar months by TESS.
+  {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
+  points are kinematically selected members of the $\delta$\,Lyr
+  cluster (black points in Figure~\ref{fig:XYZvtang}).  Both Kepler
+  (gray panels) and TESS (colored points) observed portions of the
+  cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray; four
+  of them (orange points) have their names shown.  Kepler\,1627 (green
+  star) was observed during the entirety of the Kepler mission, and
+  has been observed  by TESS for two lunar months to date.
   \label{fig:skychart}

### 2021-05-20 paper/ms.tex
update ms

@@
     The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
-    30 minute cadence.  The {\it top panel} shows the largest
-    signal, induced by starspots and plages.  The model for the
+    30 minute cadence.  The {\it top panel} shows the
+    \texttt{PDCSAP} median-subtracted flux in units of
+    parts-per-thousand ($\times 10^{-3}$).  The dominant signal
+    is induced by starspots and plages.  The model for the
     stellar variability (orange line) is subtracted in the {\it
-    bottom panel}, highlighting the transits of \pn, as well as other
+    bottom panel}, revealing the transits of \pn, as well as other
     deviations from the stellar variability model.  The 

### 2021-05-21 paper/ms.tex
gptransit v2
* updated LD
* rho prior fixed to avoid the nonperiodic SHOTerm competing with the
periodic one

@@
     variability removed.}  
-    The 1-$\sigma$ model uncertainties (faint purple band) and the
-    median model are shown.
+    The 1-$\sigma$ model uncertainties and the median model are shown
+    as the faint purple band, and the dark purple line.
+    Gray points are individual \texttt{PDCSAP} flux measurements;
+    black points are binned to 20 minutes, with a representative error
+    bar shown.
 		\label{fig:phasefold}
@@
 \end{figure*}
+
+We assumed a quadratic limb-darkening law, with the uninformative prior
+advocated by CITET Kipping2013.
+
+% In the first approach, we adopted uniform priors centered on the
+% interpolated CITET Claret+2011 coefficients in the Kepler-band, with
+% with $\pm 0.2$ in either parameter.  The fit was not as good.
 

### 2021-05-21 data/priors/Kepler_1627_priors.py

@@ -1,5 +1,10 @@
 import numpy as np
 
+#
+# Note: the names of these parameters do matter for posterior_table creation.
+# So does the extra comma in singletons.
+#
+
 # normal: mu, sd
 # uniform: lower, upper, testval
 # ImpactParameter: testval
@@ -9,21 +14,22 @@ priordict = {
 't0': ('Normal', 120.790531, 0.02), # Holczer+16, 2454953.790531 - 2454833 for Kepler timestamp
 'log_r': ('Uniform', np.log(1e-2), np.log(1), np.log(0.0433)),
 'b': ('ImpactParameter', 0.5),
-#'u_star': ('QuadLimbDark'),
-'u[0]': ('Uniform', 0.51-0.2, 0.51+0.2, 0.51), # Claret+2011, Teff 5500K, logg 4.5, solar metallicity, V-band
-'u[1]': ('Uniform', 0.24-0.2, 0.24+0.2, 0.240),
+'u_star': ('QuadLimbDark',),
+#'u[0]': ('Uniform', 0.51-0.2, 0.51+0.2, 0.51), # Claret+2011, Teff 5500K, logg 4.5, solar metallicity, V-band
+#'u[1]': ('Uniform', 0.24-0.2, 0.24+0.2, 0.240),
 'r_star': ('TruncatedNormal', 0.910, 0.052), # HIRES recon spec
 'logg_star': ('Normal', 4.6, 0.10), # HIRES recon spec
-'mean': ('Normal', 1, 0.1),
-'ecc': ('EccentricityVanEylen19'),
+'mean': ('Normal', 0.5, 0.1),
+'ecc': ('EccentricityVanEylen19',),
 'omega': ('Uniform', 0, 2*np.pi),
 'log_jitter': ('Normal', r"\log\langle \sigma_f \rangle", 2.0),
 # SHO term
-'rho': ("InverseGamma", 0.5, 2),
+#'rho': ("InverseGamma", 0.5, 2),
+'rho': ("Uniform", 1, 10),
 'sigma': ("InverseGamma", 1, 5),
 # Rotation term
 'sigma_rot': ("InverseGamma", 1, 5),
-'log_prot': ('Normal', np.log(2.606418), 0.02), # LGB LombScargle, +/-2% systematic (on the log) assumed.
[…]

### 2021-05-21 drivers/run_gptransit.py
update ms

@@ -76,7 +76,7 @@ def run_gptransit(starid='Kepler_1627', N_samples=500):
 
     if phaseplot:
         outpath = join(PLOTDIR, f'{starid}_{modelid}_posterior_phaseplot.png')
-        ylimd = {'A':[-3.5, 2.5], 'B':[-2,2]}
+        ylimd = {'A':[-3.5, 2.5], 'B':[-1,1]}
         bp.plot_phased_light_curve(datasets, m.trace.posterior, outpath,
                                    from_trace=True, ylimd=ylimd)
 

### 2021-05-21 paper/collect.sh

@@ -6,4 +6,4 @@ cp /Users/luke/Dropbox/proj/rudolf/results/keplerlc/keplerlc_Nsample500_xlim0500
 
 cp /Users/luke/Dropbox/proj/rudolf/results/run_gptransit/Kepler_1627_gptransit_posterior_phaseplot.pdf f5.pdf
 
-
+cp /Users/luke/Dropbox/proj/rudolf/results/run_gptransit/Kepler_1627_gptransit_cornerplot.png dontinclude0.png

### 2021-05-27 paper/ms.tex
update ms

@@
     variability removed.}  
-    The 1-$\sigma$ model uncertainties and the median model are shown
+    The 1-$\sigma$ model uncertainties and the mean model are shown
     as the faint purple band, and the dark purple line.
@@
     bar shown.
+    {\bf LGB FIXME: seems like the uncertainty is bigger than the RMS
+    in the out of transit...}
 		\label{fig:phasefold}
@@
 % with $\pm 0.2$ in either parameter.  The fit was not as good.
+
+Figure~\ref{fig:phasefold} shows X, Y, Z.  The residuals during the
+transit may hint at some small degree of unfitted signal -- in the
+sense that the observations are systematically high in the first half
+of transit, and low in the second half.  
+One possible explanation could be that we are seeing the effects of
+gravity darkening (e.g., CITE Masuda 2015).
+The expected amplitude of the effect is XXX, based on CITE.
+However, there is also a dip of comparable amplitude shortly before
+the transit. 
+Given the uncertainties in the stellar variability model, we expect
+that these patterns may correspond to more noise than signal.
 
@@
 
+\section{Bonus figures that wont be included}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{dontinclude0.png}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf corner plot of gptransit fit}
+	}
+\end{figure*}
+
 \listofchanges

### 2021-05-27 drivers/run_iterative_bls.py

@@ -7,7 +7,7 @@ from rudolf.paths import DATADIR, RESULTSDIR
 
 outdir = os.path.join(RESULTSDIR, 'iterative_bls')
 # made by plot_keplerlc.py, in a dumper under _plot_zoom_light_curve
-incsv = os.path.join(outdir, 'gptransit_dtr_20210518.csv')
+incsv = os.path.join(outdir, 'gptransit_dtr_20210521.csv')
 df = pd.read_csv(incsv)
 
 x,y,yerr,gp_mod = nparr(df.x), nparr(df.y), nparr(df.yerr), nparr(df.gp_mod)

### 2021-05-27 paper/ms.tex
update ms

@@
 
-\documentclass[12pt,twocolumn,tighten]{aastex63}
+\documentclass[12pt,modern,tighten]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
@@
 \begin{abstract}
-  The Gaia satellite is revitalizing our understanding of nearby
-  open clusters and moving groups.  Here, we focus on the
-  underappreciated \cn.  Based on rotation periods from TESS and
-  lithium measurements from ground-based spectrographs, we find the
-  age of the cluster to be $30\pm XX$\,Myr.  \sn\ is a binary system
-  in the cluster, serendipitously observed by the Kepler satellite
-  because the primary is nearby and Sun-like.  Kepler\,1627A was
-  previously found to host a $3.56\pm 0.04\,R_\oplus$ mini-Neptune on a
-  7.2\,day orbit.  We re-validate the existence of \pn, cementing it
-  as the youngest planet with a well-measured age observed by the main
-  Kepler mission.  Newly derived ages from Gaia offer the opportunity
-  to significantly expand the census of age-dated planets -- we offer
-  a literature compilation of young stars to enable this expansion.
-  The properties of \pn\ are may also help clarify how the orbits and
-  atmospheres of the mini-Neptune planets evolve.
+  Kepler\,1627 is an active Sun-like star previously found to host a
+  $3.56\pm 0.04\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
+  selected for observation by the Kepler satellite because it is
+  nearby and Sun-like.  Here, combining data from the Gaia satellite
+  with rotation periods from TESS, we show that Kepler\,1627 is a
+  member of the underappreciated \cn.  Based on isochronal and
+  gyrochronal evidence, the cluster is $30\pm
+  XX$\,Myr old.  Based on archival NIRC2 imaging, we also highlight the existence of 
+  a previously unknown stellar companion. Our analysis advances Kepler\,1627Ab as the youngest planet with a
+  precise age observed by the main Kepler mission.  Newly
+  derived ages from Gaia offer the opportunity to significantly expand
+  the census of age-dated planets -- we offer a literature compilation
+  of young stars to support this effort.  The properties of \pn\
+  may also help clarify how the orbits and atmospheres of the
+  mini-Neptune planets evolve.
 \end{abstract}
@@
 
-\section{Introduction}
+\section{Manuscript}
 
-At the time of the main Kepler
-mission (2009--2013), only four open clusters were known in the Kepler
-field: NGC\,6866, NGC\,6811, NGC\,6819,
-and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
-\citep{meibom_kepler_2011}.
+While thousands of exoplanets have been discovered orbiting nearby
+stars, the vast majority of them are several billion years old.  This
+results in difficulty testing theories for the origins of the
+different families of planets, since many expected evolutionary
+processes operate on timescales of less than 100 million years.
 
-Section~\ref{sec:first}.  Section~\ref{sec:second}.  In
-Section~\ref{sec:discussion}, we discuss.
-Section~\ref{sec:conclusion} gives our conclusions.
+For instance, the ``mini-Neptune'' planets, thought to be made of
+rocky cores and extended atmospheric envelopes of hydrogen and helium,
+are expected to shrink in size by factors of several over their first
+$10^8$ years \citep{owen_constraining_2020}.  Specifically, they start
+with sizes of 4--12\,$R_\oplus$ at the time of disk dispersal
+($\approx$$10^7$\,years), and shrink to sizes of 2--4\,$R_\oplus$ by
+10$^8$ years.  This change is driven by a combination of stellar
+irradiation and internal heat that powers an outflow, which eventually
+depletes or entirely strips the envelope from the rocky core
+\citep{Owen_Wu_2013,gupta_sculpting_2019}.  Discovering young planets
+and measuring their compositions and atmospheric outflows are key
+steps toward testing this paradigm, which is often invoked to explain
+the observed radius distribution of both mature exoplanets
+\citep{Fulton_et_al_2017}, as well as younger planets
+\citep{bouma_cluster_2020}.
 
-\section{first}
-\label{sec:first}
+% The orbits of the planets can also change early on, and can become
+% significantly misaligned from the equators of their host stars.
+% Measurements of the sky-projected angle between the planet orbit and
+% the stellar spin (the stellar obliquity) have shown that high
+% obliquities characterize hot field stars hosting both hot Jupiters and
+% smaller planets \citep{winn10,louden21}.  Conversely, the obliquities
+% of planets around cool stars tend to be low, but this could be due to
+% tidal realignment \citep[{\it e.g.},][]{anderson21}.  Obliquity
+% measurements for young planets are therefore important, because they
+% provide an unobstructed view of the primordial obliquity distribution.
 
@@
 		cluster (black points in Figure~\ref{fig:XYZvtang}).  Both Kepler
-  (gray panels) and TESS (colored points) observed portions of the
+		(gray squares) and TESS (colored points) observed portions of the
 		cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray; four
-  of them (orange points) have their names shown.  Kepler\,1627 (green
+		of them (orange points) have their names annotated.  Kepler\,1627 (green
 		star) was observed during the entirety of the Kepler mission, and
@@
 \end{figure*}
-
-\section{second}
-\label{sec:second}
 
@@
 	\caption{
-    {\bf \sn\ light curve over 100 days.}  
+		{\bf 100 day window of the \sn\ light curve.}  
 		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
@@
 
-We assumed a quadratic limb-darkening law, with the uninformative prior
-advocated by CITET Kipping2013.
+\input{model_posterior_table.tex}
 
-% In the first approach, we adopted uniform priors centered on the
-% interpolated CITET Claret+2011 coefficients in the Kepler-band, with
-% with $\pm 0.2$ in either parameter.  The fit was not as good.
+At the time of the main Kepler mission (2009--2013), only four open
+clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
+NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
+\citep{meibom_kepler_2011}.  Since that time, analyses of the
+kinematic, photometric, and astrometric Gaia data have expanded our
+knowledge of open cluster and moving group memberships \citep[{\it
+e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021}.
+As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
+\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
+from the literature, which yielded a list of candidate young and
+age-dated stars (see Section~\ref{app:targetlist}).
 
-Figure~\ref{fig:phasefold} shows X, Y, Z.  The residuals during the
-transit may hint at some small degree of unfitted signal -- in the
-sense that the observations are systematically high in the first half
-of transit, and low in the second half.  
-One possible explanation could be that we are seeing the effects of
-gravity darkening (e.g., CITE Masuda 2015).
-The expected amplitude of the effect is XXX, based on CITE.
-However, there is also a dip of comparable amplitude shortly before
-the transit. 
-Given the uncertainties in the stellar variability model, we expect
-that these patterns may correspond to more noise than signal.
+Comparing our young star list against the Kepler field yielded two
+notable discoveries.  The first, to be discussed in an upcoming analysis
+by J. Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
+open cluster Theia\,520 (UBC\,1).  At least five Kepler planets are
+known in the cluster, including the Kepler-52 and Kepler-968 systems.
+The second discovery, and the focus of this work, is that Kepler
+observed a portion of the $\delta$\,Lyr cluster, shown in
+Figure~\ref{fig:skychart}.
 
-\section{Discussion}
-\label{sec:discussion}
+While we have yet to perform an in-depth search of the {\bf XXX} light
+curves Kepler acquired for the \cn, there is already a known planet
+that has been detected in the cluster: Kepler\,1627b.  A portion of
+the light curve is shown in Figure~\ref{fig:lightcurve}, and the
+phase-folded transit is shown in Figure~\ref{fig:phasefold}.  The
+dominant signal in the \texttt{PDCSAP} photometry is a quasi-period
+starspot signal with a peak-to-peak amplitude that varies between
+roughly 2\% and 8\%, depending on the degree of asymmetry in spot
+coverage between the stellar hemispheres.  The starspot signal repeats
+roughly every 2.606\,days, though this value varies by a few percent
+depending on which portion of the light curve is analyzed {\bf (FIXME:
+	do!)}.  The details of how we model this signal to derive the
+parameters of the planet are discussed in Section~\ref{app:gptransit}.
+The results, given in Table~\ref{tab:posterior}, are consistent with a
+mini-Neptune sized planet ($3.56\pm 0.04\,R_\oplus$) on a circular
+7.2\,day orbit, around a G8V star ($0.91 \pm 0.05 R_\odot$) {\bf
+	fixme: age}.
 
-\section{Conclusion}
-\label{sec:conclusion}
+Could the transit signal be produced by anything other than a planet
+orbiting this near-solar analog?  \citet{morton_false_2016} considered
+the various false positive scenarios, including those of a foreground
+eclipsing binary, a hierarchical triple-star binary system, and a
+background eclipsing binary.  \citet{morton_false_2016} reported the latter
+scenario to be the only possible case, with a (model-depending)
+probability of $\approx10^{-5}$.  While this statistical validation is
+reassuring, ``validated" planets have been demonstrated to be false
+positives \citep{shporer_three_2017}.  We are not particularly
+concerned in this case however, since these cases almost always
+originate when attempting to separate the smallest stars from
+identically-sized gas giant planets.  With a size of $\approx
+3.6\,R_\oplus$, Kepler\,1627b is well below the regime of concern.
+
+Nonetheless, our analysis of archival Keck-NIRC2 Kp-band (2.12\,$\mu
+$m) imaging did reveal the existence of a neighboring star, unresolved
+in the Gaia source catalog {\bf FIXME: uncertainties}.  The NIRC2
+images yield a projected separation $\rho=0\farcs17$, with $\Delta
+{\rm Kp} = 2.5$.  Using the measured Gaia EDR3 parallax for the
+system, this implies a projected separation of 54\,AU.  The presence
+of this star is consistent with the excess noise reported to be
+present in the Gaia astrometric time-series (see
+Section~\ref{app:companionstar}).  Given the low chance of a star
+imaged within {\bf X.X arcseconds} to be a chance unbound companion
+along this line of sight, we dub the companion star Kepler\,1627B, and
+proceed under the assumption that it is bound.
+
+Unforuntately, we do not have any reliable color information about the
+secondary star.  Based on the
+tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
+version \texttt{2021/03/02}.} by \citet{pecaut_mamajek_2013}, the
+NIR-contrast measured for a main sequence G8V star corresponds to a
+spectral type for the companion of $\approx$M2V ($M_\star \approx
+0.44\,M_\odot$).  However, the companion should therefore have a
+significantly longer pre-main-sequence contraction phase, which would
+imply that this mass is an overestimate by $\approx20\%$.
+
+Could the companion host the planet, or be creating a false positive
+signal?  The star contributes $\approx$1\% of the total flux observed
+in the Kepler aperture(0.7-2.0\% V-band to G-band dmag difference from
+EEM table; todo is fix using model spectra).  The observed transit has
+a depth of 0.18\%.  An 18\% deep eclipse of the secondary star could
+produce such a signal.  Given the shape of the observed signal
+(allowed impact parameters span 0.02 to 0.73, see
+Table~\ref{tab:posterior}), the body transiting the secondary would
+need to be non-grazing with $R/R_2 \approx 0.42$, {\it i.e.}, with
+radius $\approx 0.2R_\odot$.
+The duration of the transit 
+($2.823 \pm 0.057 hr$) is ultimately what makes this scenario
+implausible.
+
+% {\bf TODO:
+% uhhh what about the pre-MS M dwarf star being not very dense?
+% What is the density of the G dwarf expected to be?
+% }
+% 
+% $T_{14}$ & hr & -- & 2.822 & 2.823 & 0.057 & 2.710 & 2.917 & 1086.008 & 0.002 \\
+% $\rho_\star$ & g$\ $cm$^{-3}$ & -- & 2.27 & 2.312 & 0.516 & 1.408 & 3.272 & 910.746 & 0.001 \\
+
+%\section{Discussion}
+%\label{sec:discussion}
+
+%\section{Conclusion}
+%\label{sec:conclusion}
 
@@
 
-\input{model_posterior_table.tex}
-
 \clearpage
@@
 
-\section{Bonus figures that wont be included}
+\section{Transit and Stellar Variability Model}
+\label{app:gptransit}
 
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=1\textwidth]{dontinclude0.png}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf corner plot of gptransit fit}
-	}
-\end{figure*}
+We assumed a quadratic limb-darkening law, with the uninformative prior
+advocated by CITET Kipping2013.
+
+% In the first approach, we adopted uniform priors centered on the
+% interpolated CITET Claret+2011 coefficients in the Kepler-band, with
+% with $\pm 0.2$ in either parameter.  The fit was not as good.
+
+Figure~\ref{fig:phasefold} shows X, Y, Z.  The residuals during the
+transit may hint at some small degree of unfitted signal -- in the
+sense that the observations are systematically high in the first half
+of transit, and low in the second half.  
+One possible explanation could be that we are seeing the effects of
+gravity darkening (e.g., CITE Masuda 2015).
+The expected amplitude of the effect is XXX, based on CITE.
+However, there is also a dip of comparable amplitude shortly before
+the transit. 
+Given the uncertainties in the stellar variability model, we expect
+that these patterns may correspond to more noise than signal.
+
+\section{Companion Star}
+\label{app:companionstar}
+
+{\bf TODO: include the RUWE vs Rp mag.}
+
+% \section{Bonus figures that wont be included}
+% 
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\includegraphics[width=1\textwidth]{dontinclude0.png}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+% 	\caption{
+%     {\bf corner plot of gptransit fit}
+% 	}
+% \end{figure*}
 

### 2021-05-27 paper/ms.tex
wow, nature papers are short

@@
 \begin{abstract}
-  Kepler\,1627 is an active Sun-like star previously found to host a
-  $3.56\pm 0.04\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
-  selected for observation by the Kepler satellite because it is
-  nearby and Sun-like.  Here, combining data from the Gaia satellite
-  with rotation periods from TESS, we show that Kepler\,1627 is a
-  member of the underappreciated \cn.  Based on isochronal and
-  gyrochronal evidence, the cluster is $30\pm
-  XX$\,Myr old.  Based on archival NIRC2 imaging, we also highlight the existence of 
-  a previously unknown stellar companion. Our analysis advances Kepler\,1627Ab as the youngest planet with a
-  precise age observed by the main Kepler mission.  Newly
-  derived ages from Gaia offer the opportunity to significantly expand
-  the census of age-dated planets -- we offer a literature compilation
-  of young stars to support this effort.  The properties of \pn\
-  may also help clarify how the orbits and atmospheres of the
+  Kepler\,1627 is an active G8V star previously found to host a
+  $3.56\pm 0.04\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star
+  was selected for observation by the Kepler satellite because it is
+  nearby and it resembles the Sun.  Combining data from the Gaia
+  satellite with rotation periods from TESS, we show that Kepler\,1627
+  is a member of the \cn.  Based on isochronal and gyrochronal
+  evidence, the cluster is $30\pm XX$\,Myr old.  Based on archival
+  NIRC2 imaging, we also note the existence of a low-mass stellar
+  companion. Our analysis advances Kepler\,1627Ab as the youngest
+  planet with a precise age detected by the main Kepler mission.
+  Newly derived ages from Gaia offer the opportunity to significantly
+  expand the census of age-dated planets -- we offer a literature
+  concatenation of young stars to support this effort.  The properties
+  of \pn\ may also help clarify how the orbits and atmospheres of the
   mini-Neptune planets evolve.
@@
 		\leavevmode
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{TEMP_Prot_vs_BpmRp.png}
+			\includegraphics[width=0.49\textwidth]{TEMP_hrd.png}
+		}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf \cn is $35\pm15$\,Myr old.} {\it Left:} TESS and
+    Kepler rotations periods and dereddened Gaia colors, with the
+    Pleiades \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
+    \citep[650\,Myr;][]{douglas_poking_2017} shown for reference.
+    Stars of a given color (mass) move up through time due to magnetic
+    braking.  The \cn\ has a 40 Myr gyrochronal age, as calibrated
+    against the LDB-age of IC\,2602.  {\it Right:} HR diagram of the
+    same comparison. {\bf todo: caption improve}.
+   \label{fig:age}
+	}
+\end{figure*}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
 		\includegraphics[width=1\textwidth]{f4.pdf}
@@
 	\caption{
-		{\bf 100 day window of the \sn\ light curve.}  
+		{\bf Zoomed view of the \sn\ light curve.}  
 		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
@@
 			variability removed.}  
-		The 1-$\sigma$ model uncertainties and the mean model are shown
-		as the faint purple band, and the dark purple line.
-		Gray points are individual \texttt{PDCSAP} flux measurements;
-		black points are binned to 20 minutes, with a representative error
-		bar shown.
+    The 1-$\sigma$ model uncertainties and the maximum {\it a
+    posteriori} model are shown as the faint purple band, and the dark
+    purple line.  Gray points are individual \texttt{PDCSAP} flux
+    measurements; black points are binned to 20 minutes, with a
+    representative error bar shown.
     {\bf LGB FIXME: seems like the uncertainty is bigger than the RMS
@@
 \end{figure*}
-
-\input{model_posterior_table.tex}
 
@@
 Comparing our young star list against the Kepler field yielded two
-notable discoveries.  The first, to be discussed in an upcoming analysis
+discoveries.  The first, to be discussed in an upcoming analysis
 by J. Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
@@
 
-While we have yet to perform an in-depth search of the {\bf XXX} light
-curves Kepler acquired for the \cn, there is already a known planet
-that has been detected in the cluster: Kepler\,1627b.  A portion of
-the light curve is shown in Figure~\ref{fig:lightcurve}, and the
-phase-folded transit is shown in Figure~\ref{fig:phasefold}.  The
-dominant signal in the \texttt{PDCSAP} photometry is a quasi-period
-starspot signal with a peak-to-peak amplitude that varies between
-roughly 2\% and 8\%, depending on the degree of asymmetry in spot
-coverage between the stellar hemispheres.  The starspot signal repeats
-roughly every 2.606\,days, though this value varies by a few percent
-depending on which portion of the light curve is analyzed {\bf (FIXME:
-	do!)}.  The details of how we model this signal to derive the
-parameters of the planet are discussed in Section~\ref{app:gptransit}.
-The results, given in Table~\ref{tab:posterior}, are consistent with a
-mini-Neptune sized planet ($3.56\pm 0.04\,R_\oplus$) on a circular
-7.2\,day orbit, around a G8V star ($0.91 \pm 0.05 R_\odot$) {\bf
-	fixme: age}.
+While we have yet to perform an in-depth planet search of the {\bf
+XXX} light curves Kepler acquired for the \cn, there is already a
+known planet in the cluster: Kepler\,1627b.  We considered the
+planet's previous statistical validation by \citet{morton_false_2016}
+to be sufficient cause to examine the age of the star and its cluster more closely.
+We collected the previously measured {\bf XXX} Kepler
+rotation periods (CITEP MCQUILLAN2014), and measured {\bf YYY} new stellar
+rotation periods for cluster members observed by TESS.
+We also performed an isochronal analysis.
+The results are shown in Figure~\ref{fig:age}.
+From the ensemble perspective, which provides the strongest
+constraints (SEE CITET SODERBLOM 2014),
+the age of the star is $XX \pm YY$\,Myr.
+Individual measurements of the star's rotation period and lithium
+abundance corroborate this conclusion.
+
+If the Kepler\,1627b transit signal truly corresponds to a planet,
+this age would make it the youngest planet yet found by the main
+Kepler mission.  A portion of the light curve is shown in
+Figure~\ref{fig:lightcurve}, and the phase-folded transit is shown in
+Figure~\ref{fig:phasefold}.  The dominant signal in the
+\texttt{PDCSAP} photometry is a quasi-period starspot signal with a
+peak-to-peak amplitude that varies between roughly 2\% and 8\%,
+depending on the degree of asymmetry in spot coverage between the
+stellar hemispheres.  The starspot signal repeats roughly every
+2.606\,days, though this value varies by a few percent depending on
+which portion of the light curve is analyzed {\bf (FIXME: do!)}.  The
+details of how we model this signal to derive the parameters of the
+planet are discussed in Section~\ref{app:gptransit}.  The results,
+given in Table~\ref{tab:posterior}, are consistent with a mini-Neptune
+sized planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit,
+around a G8V star ($0.91 \pm 0.05 R_\odot$) {\bf fixme: age}.
 
@@
 eclipsing binary, a hierarchical triple-star binary system, and a
-background eclipsing binary.  \citet{morton_false_2016} reported the latter
-scenario to be the only possible case, with a (model-depending)
+background eclipsing binary.  \citet{morton_false_2016} reported the
+latter scenario to be the only possible case, with a (model-dependent)
 probability of $\approx10^{-5}$.  While this statistical validation is
 reassuring, ``validated" planets have been demonstrated to be false
-positives \citep{shporer_three_2017}.  We are not particularly
-concerned in this case however, since these cases almost always
-originate when attempting to separate the smallest stars from
-identically-sized gas giant planets.  With a size of $\approx
-3.6\,R_\oplus$, Kepler\,1627b is well below the regime of concern.
+positives \citep{shporer_three_2017}.  Most failures of the
+statistical validation approach arise when attempting to separate the
+smallest stars from identically-sized gas giant planets (CITE).  With a size
+of $\approx 3.6\,R_\oplus$, Kepler\,1627b does not meet this
+concern.  The fact that we detect a low-mass stellar companion in
+archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does not change
+the conclusion: the companion contributes 1\% of the flux
+observed by Kepler, which combined with the observed transit shape
+cannot be invoked as an explanation for the observed data
+(Section~\ref{app:companionstar}).
 
-Nonetheless, our analysis of archival Keck-NIRC2 Kp-band (2.12\,$\mu
-$m) imaging did reveal the existence of a neighboring star, unresolved
-in the Gaia source catalog {\bf FIXME: uncertainties}.  The NIRC2
-images yield a projected separation $\rho=0\farcs17$, with $\Delta
-{\rm Kp} = 2.5$.  Using the measured Gaia EDR3 parallax for the
-system, this implies a projected separation of 54\,AU.  The presence
-of this star is consistent with the excess noise reported to be
-present in the Gaia astrometric time-series (see
-Section~\ref{app:companionstar}).  Given the low chance of a star
-imaged within {\bf X.X arcseconds} to be a chance unbound companion
-along this line of sight, we dub the companion star Kepler\,1627B, and
-proceed under the assumption that it is bound.
+Kepler\,1627Ab therefore provides an interesting extremum in the ages of the
+Kepler planets, and opens multiple avenues for detailed
+study.  In order of ascending difficulty, observing its
+Doppler transit ($\Delta v_{\rm RM} \approx 20$\ms) with HIRES or the
+upcoming KPF spectrograph could yield a measurement of its orbital
+obliquity (CITEP,CITEP).  A transmission
+spectrum focused on either hydrogen or helium absorption could yield
+insight into its atmosphere, which is expected to be outflowing at a
+rate of {\bf X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  Finally, the
+most challenging measurement that may currently be in reach would be a
+measurement of its mass, through either detailed timing analyses of
+the Kepler data, or through a high-cadence multi-color radial velocity campaign.  The mass would yield constraints on both the planet's
+composition (do the mini-Neptunes form with gaseous envelopes?), as
+well as its initial entropy (``hot'' or ``cold''? CITE).
 
-Unforuntately, we do not have any reliable color information about the
-secondary star.  Based on the
-tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
-version \texttt{2021/03/02}.} by \citet{pecaut_mamajek_2013}, the
-NIR-contrast measured for a main sequence G8V star corresponds to a
-spectral type for the companion of $\approx$M2V ($M_\star \approx
-0.44\,M_\odot$).  However, the companion should therefore have a
-significantly longer pre-main-sequence contraction phase, which would
-imply that this mass is an overestimate by $\approx20\%$.
-
-Could the companion host the planet, or be creating a false positive
-signal?  The star contributes $\approx$1\% of the total flux observed
-in the Kepler aperture(0.7-2.0\% V-band to G-band dmag difference from
-EEM table; todo is fix using model spectra).  The observed transit has
-a depth of 0.18\%.  An 18\% deep eclipse of the secondary star could
-produce such a signal.  Given the shape of the observed signal
-(allowed impact parameters span 0.02 to 0.73, see
-Table~\ref{tab:posterior}), the body transiting the secondary would
-need to be non-grazing with $R/R_2 \approx 0.42$, {\it i.e.}, with
-radius $\approx 0.2R_\odot$.
-The duration of the transit 
-($2.823 \pm 0.057 hr$) is ultimately what makes this scenario
-implausible.
-
-% {\bf TODO:
-% uhhh what about the pre-MS M dwarf star being not very dense?
-% What is the density of the G dwarf expected to be?
-% }
-% 
-% $T_{14}$ & hr & -- & 2.822 & 2.823 & 0.057 & 2.710 & 2.917 & 1086.008 & 0.002 \\
-% $\rho_\star$ & g$\ $cm$^{-3}$ & -- & 2.27 & 2.312 & 0.516 & 1.408 & 3.272 & 910.746 & 0.001 \\
-
-%\section{Discussion}
-%\label{sec:discussion}
-
-%\section{Conclusion}
-%\label{sec:conclusion}
+Ultimately, our age measurement for Kepler\,6127Ab was enabled by
+identifying its connection to the \cn\ through the Gaia data.  The
+Supplementary Table~\ref{tab:v05} enables similar crossmatches to
+already known planets, as well as forthcoming planets from the TESS
+mission (CITE RickerJATIS2014).  An alternative approach also
+likely to bear fruit is to identify kinematic associations
+around exoplanet host stars using the Gaia data, and to then verify
+these associations with ancilliary rotation periods and spectroscopy
+(CITEP Tofflemire).
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\input{model_posterior_table.tex}
 
@@
 
-\section{Companion Star}
+\section{Companion Star and False Positive Assessment}
 \label{app:companionstar}
 
-{\bf TODO: include the RUWE vs Rp mag.}
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.6\textwidth]{f6.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Gaia EDR3 renormalized unit weight error (RUWE) point
+    estimates for Kepler\,1627A and other members of
+    the $\delta$\,Lyr cluster.}  Since other members of the cluster
+    with similar colors have comparable degrees of photometric
+    variability, the high RUWE estimate suggests that Kepler\,1627A is
+    a binary. 
+		\label{fig:ruwe}
+	}
+\end{figure*}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.6\textwidth]{TEMP_binarity_ao.png}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Keck NIRC2 AO image of Kepler 1627A and Kepler 1627B.}  
+    {\bf SCALEBAR DENOTES...}
+    {\bf North is up, East is left}.
+    \label{fig:ao}
+	}
+\end{figure*}
+
+Our analysis of archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging
+revealed the existence of a previously unreported stellar neighbor,
+unresolved in the Gaia source catalog {\bf FIXME: uncertainties}.  The
+NIRC2 images yield a projected separation $\rho=0\farcs17$, with
+$\Delta {\rm Kp} = 2.5$.  Using the measured Gaia EDR3 parallax for
+the system, this implies a projected separation of 54\,AU.  The
+presence of this star is consistent with the excess noise in the Gaia
+astrometric time-series (see Section~\ref{app:companionstar}).  Given
+the low chance of a star imaged within {\bf X.X arcseconds} to be a
+chance companion along the line of sight, we proceed under the
+assumption that it is bound, and that the Kepler\,1627 system is
+binary.
+
+Unforuntately, we do not have any reliable color information about
+Kepler\,1627B.  Based on the
+tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
+version \texttt{2021/03/02}.} by \citet{pecaut_mamajek_2013}, the
+measured NIR-contrast for a main sequence G8V star corresponds to a
+spectral type for the companion of $\approx$M2V ($M_\star \approx
+0.44\,M_\odot$).  However, the companion should have a longer
+pre-main-sequence contraction phase than the primary, which would
+imply that this mass is overestimated by $\sim10$ to $20\%$.
+
+Could the companion be creating a false positive signal?  The
+companion star contributes $\approx$1\% of the total flux observed in
+the Kepler aperture (0.7-2.0\% V-band to G-band dmag difference from
+EEM table; todo is fix using model spectra).  The observed transit has
+a depth of 0.18\%.  An 18\% deep eclipse of the secondary star would
+therefore be needed to produce a deep enough signal.  The shape of the
+observed signal requires allowed impact parameters to span 0.02 to
+0.73 (Table~\ref{tab:posterior}); the body transiting the secondary
+would therefore need to be non-grazing with $R_3/R_2 \approx 0.42$.
+Assuming a $\approx 0.42R_\odot$ radius of the imaged secondary, this
+would imply a tertiary stellar radius of $\approx 0.2R_\odot$.  This
+scenario ultimately yields a contradiction, because it would require
+an ingress and egress phase that each span $\approx$40\% of the
+transit duration ($\approx 65\,{\rm minutes}$).  The actual measured
+ingress and egress duration is $\approx 15\,{\rm minutes}$),
+4.4$\times$ shorter.
+
+\paragraph{Stellar density implied by transit duration}
+The duration of the transit ($2.823 \pm 0.057 hr$) and the implied
+stellar density ($2.31 \pm 0.52\,{\rm g\,cm}^{-3}$) could in theory
+help rule between whether the transiting body orbits the primary or
+secondary star.  Ultimately, our stellar density measurement is not
+precise enough to render the blend scenario implausible.  At 30 Myr, a
+0.40\,$M_\odot$ solar-metallicity dwarf is $\approx 26\%$ larger than
+when it is fully contracted on the main sequence (0.40\,$R_\odot$
+vs{.} 0.50\,$R_\odot$; CITEALT: Choi et al, MIST grids).  The
+theoretically implied companion density of $3.2\,{\rm g\,cm}^{-3}$ is
+indeed larger than the primary star's density of $1.80\,{\rm
+g\,cm}^{-3}$ (measured through the HIRES reconaissance spectroscopy).
+However the stellar density measured from the transit fitting ($2.31
+\pm 0.52\,{\rm g\,cm}^{-3}$) is only discrepant at the
+$\approx$2-$\sigma$ level from the M-dwarf blend scenario.  It is
+instead the combination of the flux contrast, transit depth, and
+ingress duration that rule out this scenario.
 

### 2021-05-27 paper/ms.tex
update ms

@@
 upcoming KPF spectrograph could yield a measurement of its orbital
-obliquity (CITEP,CITEP).  A transmission
+obliquity (CITEP,CITEP).  
+(A simpler path to the same result might be to determine whether the
+$\approx80\,{\rm ppm}$ residual apparent in our transit fit is caused
+by gravity darkening.)
+A transmission
 spectrum focused on either hydrogen or helium absorption could yield

### 2021-06-03 paper/ms.tex
minor

@@
 \title{
-  A 30 Million Year Old Mini-Neptune in the Kepler Field
+  An Adolescent Mini-Neptune in the Kepler Field
 }
@@
 \begin{abstract}
-  Kepler\,1627 is an active G8V star previously found to host a
-  $3.56\pm 0.04\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star
-  was selected for observation by the Kepler satellite because it is
-  nearby and it resembles the Sun.  Combining data from the Gaia
-  satellite with rotation periods from TESS, we show that Kepler\,1627
-  is a member of the \cn.  Based on isochronal and gyrochronal
-  evidence, the cluster is $30\pm XX$\,Myr old.  Based on archival
-  NIRC2 imaging, we also note the existence of a low-mass stellar
-  companion. Our analysis advances Kepler\,1627Ab as the youngest
-  planet with a precise age detected by the main Kepler mission.
-  Newly derived ages from Gaia offer the opportunity to significantly
-  expand the census of age-dated planets -- we offer a literature
-  concatenation of young stars to support this effort.  The properties
-  of \pn\ may also help clarify how the orbits and atmospheres of the
-  mini-Neptune planets evolve.
+  Kepler\,1627A is an active G8V star previously found to host a
+  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
+  selected for observation by the Kepler satellite because it is
+  nearby ($d\approx 333\,{\rm pc}$) and resembles the Sun.  Combining
+  data from the Gaia satellite with rotation periods from TESS, we
+  show that Kepler\,1627 is a member of the $35\pm 10$\,Myr old
+  $\delta$~Lyr cluster.  Our analysis advances Kepler\,1627Ab as the
+  youngest planet with a precise age detected by the main Kepler
+  mission.  New ages from the combination of Gaia and TESS offer the
+  opportunity to significantly expand the census of age-dated planets
+  -- we offer a literature concatenation of young stars to support
+  this effort.  Future studies of \pn\ may also help clarify how the
+  orbits and atmospheres of the mini-Neptune planets evolve.
 \end{abstract}
@@
 		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
-		30 minute cadence.  The {\it top panel} shows the
+		30 minute cadence.  A one hundred day segment is shown here.  The {\it top panel} shows the
 		\texttt{PDCSAP} median-subtracted flux in units of
@@
 (Section~\ref{app:companionstar}).
+
+Could the planet be on a high obliquity orbit?
+% * how do the spectroscopic vsini and the measured v_rot=2πRstar/Prot compare?
+%   * Rstar=0.91+-0.05, 2.606 days gives vsini=17.6km/s (+/-0.97km/s)
+% * spectroscopic vsini is 18.8 +/- 4.7km/s from Keck HIRES... "20km/s" from
+%   TRES.
 

### 2021-06-05 paper/ms.tex
add Q15 2min fitting

@@
 
+Our default stellar variability model (RotGPtransit) allows for a
+\texttt{RotTerm} GP kernel, with a logjitter term to inflate the
+uncertainties to account for excess white noise.
+(This results in an inflation of the uncertainties by a factor
+of about three).
+
+We considered including an additive \texttt{SHOTerm} kernel to account
+for stochastic noise (RotStochGPtransit).  This didn't seem to affect the results much, so
+we opted for the simpler model.
+
 Figure~\ref{fig:phasefold} shows X, Y, Z.  The residuals during the

### 2021-06-14 data/priors/Kepler_1627_RotGPtransit_priors.py
minor

@@ -32,5 +32,5 @@ priordict = {
 'log_prot': ('Normal', np.log(2.606418), 0.02), # LGB LombScargle, +/-2% sys (on the log) assumed.
 'log_Q0': ('Normal', 0, 2),
 'log_dQ': ('Normal', 0, 2),
-'f': ('Uniform', 0.1, 1),
+'f': ('Uniform', 0.01, 1),
 }

### 2021-06-15 paper/ms.tex
minor
update ms

@@
 probability of $\approx10^{-5}$.  While this statistical validation is
-reassuring, ``validated" planets have been demonstrated to be false
+reassuring, ``validated'' planets have been demonstrated to be false
 positives \citep{shporer_three_2017}.  Most failures of the
@@
 
-Could the planet be on a high obliquity orbit?
+% Could the planet be on a high obliquity orbit?
 % * how do the spectroscopic vsini and the measured v_rot=2πRstar/Prot compare?
@@
 Kepler planets, and opens multiple avenues for detailed
-study.  In order of ascending difficulty, observing its
-Doppler transit ($\Delta v_{\rm RM} \approx 20$\ms) with HIRES or the
-upcoming KPF spectrograph could yield a measurement of its orbital
-obliquity (CITEP,CITEP).  
+study.  Ground-based observations of its Doppler transit ($\Delta
+v_{\rm RM} \approx 20$\ms) with HIRES or the upcoming KPF spectrograph
+could yield a measurement of its orbital obliquity (CITEP,CITEP).
+Similarly, transit spectroscopy aimed at detecting either hydrogen or
+helium absorption could yield insight into the planet's atmosphere,
+which is expected to be outflowing at a rate of {\bf
+X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  A more challenging
+measurement would be a measurement of the planet's mass, through
+either detailed timing analyses of the Kepler data, or through a
+high-cadence multi-color radial velocity campaign.  The mass would
+yield constraints on both the planet's composition (``what is the
+extent of the initial gaseous envelope for a typical mini-Neptune?''),
+as well as its initial entropy (``hot'' or ``cold''? CITE).
+
+The precision of the Kepler data itself may enable detailed analyses
+without even needing to acquire more data.
+These could include studies transit timing variations, distortion of
+the transit signal due to the star's gravity darkening, 
+and interactions between the star and planet exhibited by patterns in
+the star's flaring.
+On the latter point,
+Appendix REFERENCE highlights an odd coincidence: 
+in the 97.7 days of available 1-minute cadence Kepler observations
+(``short-cadence''), 
+24 flares were observed to exceed $0.5\%$ in relative flux.
+Cumulatively, these flares spanned a total of 0.27 days.
+The brightest flare, highlighted in FIGURE CITE, had a flare occur
+after it to within 0.2\% of an orbital period.
+This seems {\it a priori} unlikely; in fact, 4 of the 24 flare events
+have candidate ``successors'' (events that follow to within 2\% of the
+planet's orbital period).
+Further analysis would be needed to test the hypothesis of whether or
+not the flare time
+distribution deviated from a Poisson process...
+[like: compare Poisson model versus Poisson+periodic mixture?]
+
+FIXME
 (A simpler path to the same result might be to determine whether the
@@
 by gravity darkening.)
-A transmission
-spectrum focused on either hydrogen or helium absorption could yield
-insight into its atmosphere, which is expected to be outflowing at a
-rate of {\bf X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  Finally, the
-most challenging measurement that may currently be in reach would be a
-measurement of its mass, through either detailed timing analyses of
-the Kepler data, or through a high-cadence multi-color radial velocity campaign.  The mass would yield constraints on both the planet's
-composition (do the mini-Neptunes form with gaseous envelopes?), as
-well as its initial entropy (``hot'' or ``cold''? CITE).
 
@@
 L.G.B{.} and J.L.C{.} are grateful to T{.}~David for help with the
-transit fitting.
+transit fitting, and to R{.}~Kerr for kindly providing us with the
+\citet{Kerr2021} membership list prior to its publication.
 % is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,

### 2021-06-15 rudolf/plotting.py

@@ -890,6 +890,8 @@ def _get_detrended_flare_data(cachepath, method):
             flat_flux, trend_flux = flatten_starspots(
                 time, flux, flux_err, P_rotation, flare_iterate=True
             )
+        else:
+            raise NotImplementedError(f'Got unknown method {method}.')
 
         #  then, find positive outliers using 
         resid = (flat_flux - np.nanmedian(flat_flux))
@@ -1013,7 +1015,7 @@ def plot_flare_checker(outdir, method=None):
 
 
     _pd = phase_magseries(nparr(fl_df.tstart), nparr(fl_df.ampl_rec), period,
-                          t0, wrap=False, sort=True)
+                          t0, wrap=False, sort=False)
     x_fold = _pd['phase']
     y_fold = _pd['mags']
 
@@ -1026,7 +1028,6 @@ def plot_flare_checker(outdir, method=None):
     axd['B'].scatter(
         x_fold, y_fold, c='k', s=2, rasterized=True, linewidths=0,
         zorder=1
-
     )
 
     #axd['B'].set_ylim([0,1])

### 2021-06-15 paper/ms.tex
update ms flares

@@
 
-Kepler\,1627Ab therefore provides an interesting extremum in the ages of the
-Kepler planets, and opens multiple avenues for detailed
-study.  Ground-based observations of its Doppler transit ($\Delta
-v_{\rm RM} \approx 20$\ms) with HIRES or the upcoming KPF spectrograph
-could yield a measurement of its orbital obliquity (CITEP,CITEP).
+Kepler\,1627Ab therefore provides an interesting extremum in the ages
+of the Kepler planets, and opens multiple avenues for study from the
+ground.  Observations of its Doppler transit ($\Delta v_{\rm RM}
+\approx 20$\ms) with HIRES or the upcoming KPF spectrograph could
+yield a measurement of its orbital obliquity (CITEP,CITEP).
 Similarly, transit spectroscopy aimed at detecting either hydrogen or
@@
 
-The precision of the Kepler data itself may enable detailed analyses
-without even needing to acquire more data.
-These could include studies transit timing variations, distortion of
-the transit signal due to the star's gravity darkening, 
-and interactions between the star and planet exhibited by patterns in
-the star's flaring.
-On the latter point,
-Appendix REFERENCE highlights an odd coincidence: 
-in the 97.7 days of available 1-minute cadence Kepler observations
-(``short-cadence''), 
-24 flares were observed to exceed $0.5\%$ in relative flux.
-Cumulatively, these flares spanned a total of 0.27 days.
-The brightest flare, highlighted in FIGURE CITE, had a flare occur
-after it to within 0.2\% of an orbital period.
-This seems {\it a priori} unlikely; in fact, 4 of the 24 flare events
-have candidate ``successors'' (events that follow to within 2\% of the
-planet's orbital period).
-Further analysis would be needed to test the hypothesis of whether or
-not the flare time
-distribution deviated from a Poisson process...
-[like: compare Poisson model versus Poisson+periodic mixture?]
+The precision of the Kepler data itself may enable more detailed
+analyses without even needing to acquire more data.  These could
+include studies transit timing variations, distortion of the transit
+signal due to the star's gravity darkening, and interactions between
+the star and planet exhibited by patterns in the star's flaring.
+The gravity darkening might be interesting as an explanation for the
+$\approx80\,{\rm ppm}$ residual apparent in our transit fit.
+On the point of planet-induced flaring, Appendix REFERENCE
+highlights an odd timing coincidence that may be worthy of further
+examination.
 
-FIXME
-(A simpler path to the same result might be to determine whether the
-$\approx80\,{\rm ppm}$ residual apparent in our transit fit is caused
-by gravity darkening.)
-
-Ultimately, our age measurement for Kepler\,6127Ab was enabled by
+Ultimately though, our age measurement for Kepler\,6127Ab was enabled by
 identifying its connection to the \cn\ through the Gaia data.  The
@@
 
+\section{Flare Analysis}
+\label{app:flare}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f7.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Flares in Kepler\,1627}.  
+    {\it Top:}
+    The full short-cadence Kepler dataset, acquired at 1-minute
+    sampling (black points) is shown with a stellar variability model
+    (blue line).
+    {\it Middle:}
+    Residual after subtracting the stellar variability model.  Flares
+    appear as spikes.
+    {\it Bottom:}
+    Zooms of the brightest, and third-brightest flares.  A timing
+    coincidence -- that both flares has ``successors'' approximately
+    one orbital period after the initial event -- is emphasized.
+		\label{fig:flarezoom}
+	}
+\end{figure*}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1.0\textwidth]{f8.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Phase-folded flares in Kepler\,1627}.  
+    {\it Top:}
+    As in the middle of Figure~\ref{fig:flarezoom}, phase-folded at
+    the planet's ephemeris.
+    {\it Bottom:}
+    Fitted flare amplitudes and orbital phases.
+    Colors indiate relative time, from the beginning of the
+    98-day short-cadence Q15 dataset (dark blue) to the end (light
+    yellow).
+    Lower amplitude flares likely exist in the data, and were not
+    examined.
+		\label{fig:flarephase}
+	}
+\end{figure*}
+
+The 1-minute cadence Kepler observations span 98 days, and cover 24
+flares exceeding $0.5\%$ in relative flux.  These 24 flares spanned a
+total of 6.5 hours ($\sim$15 minutes per flare).  The coincidence is
+that despite the low flare duty cycle, one orbital period after the
+the brightest flare, a second flare followed.  This and a similar
+event are shown in Figure~\ref{fig:flarezoom}.  The timing error is
+good to a $\approx0.2\%$ difference from the orbital period, which
+seems {\it a priori} unlikely.  If we consider flares falling within
+2\% of the planet's orbital period after a previous flare, then 4 of
+the 24 flare events have candidate ``successors''.
+
+A brief note on how we cleaned the light curve, and identified the
+flares.
+For cleaning, we performed the following iterative detrending procedure.
+\begin{itemize}
+  \item Step 1: Build a two-term mixed SHOTerm GP model with
+    quasi-periodic kernels at Prot and 0.5$\times$Prot. Fit the model
+    to the time, flux, and flux uncertainty.
+  \item Step 2: Select points more than twice the median absolute
+    deviation from the residual, and exclude them from the light
+    curve.  Repeat Step 1.
+  \item Step 3: On the residual from Step 2, identify all flares,
+    requiring them to be at least 20 cadences apart, at least 7 median
+    absolute deviations above the median baseline, and lasting at least
+    2 cadences in duration.  Build the mask spanning these times, from
+    5 minutes before each flare begins to 2.5 minutes after the final
+    flare cadence.  Repeat Step 1 a final time.
+\end{itemize}
+
+The flares were identified and fitted using ALTAIPONY (CITE CITE).
+The fitted model is that of CITE XXX, which parametrizes the flare
+with a start time, a lag time, and an amplitude (CHECK).
+Figure~\ref{fig:flarephase} shows the resulting flares, amplitudes,
+and phases.
+
+We considered two hypotheses: {\it i)} the flare arrival times are
+distributed as a Poisson process, and {\it ii)} they are better
+explained as a Poisson process with a periodic mixture component.
+
+TODO: explore!  Fit the flare inter-arrival times...
+[even the 2-separated ones?]
+
 \section{Companion Star and False Positive Assessment}

### 2021-06-15 rudolf/plotting.py

@@ -19,6 +19,7 @@ from importlib.machinery import SourceFileLoader
 
 import matplotlib as mpl
 from mpl_toolkits.axes_grid1 import make_axes_locatable
+from mpl_toolkits.axes_grid1.inset_locator import inset_axes
 
 from astropy import units as u, constants as const
 from astropy.coordinates import SkyCoord
@@ -992,14 +993,14 @@ def plot_flare_checker(outdir, method=None):
     flcd = flcd.find_flares(20, **{'N1':7,'N2':7,'N3':2,'sigma':c.mad})
     fl_df = flcd.flares
 
-    fig = plt.figure(figsize=(4,3))
+    fig = plt.figure(figsize=(6,4))
     axd = fig.subplot_mosaic(
         """
         A
         B
         """,
         gridspec_kw={
-            "height_ratios": [3, 1.5]
+            "height_ratios": [1,1]
         }
     )
 
@@ -1025,10 +1026,24 @@ def plot_flare_checker(outdir, method=None):
     #axd['B'].vlines(
     #    x_fold, 0, 1, ls='-', lw=0.5, colors='k'
     #)
-    axd['B'].scatter(
-        x_fold, y_fold, c='k', s=2, rasterized=True, linewidths=0,
-        zorder=1
+    _p = axd['B'].scatter(
+        x_fold, y_fold, s=15, rasterized=True, linewidths=0.5,
+        edgecolors='k',
+        zorder=1, #c='k',
+        c=nparr(fl_df.tstart)-min(c.time), cmap='viridis'
     )
+
+    fig.tight_layout()
+
+    axins1 = inset_axes(axd['B'], width="25%", height="5%", loc='upper right')
+    cb = fig.colorbar(_p, cax=axins1, orientation="horizontal")
+    cb.ax.tick_params(labelsize='xx-small')
+    #cb.ax.set_title('Days from start', fontsize='x-small')
+    axd['B'].text(0.725,0.955, '$t$ [days]',
+            transform=axd['B'].transAxes,
+            ha='right',va='top', color='k', fontsize='xx-small')
+    cb.ax.tick_params(size=0, which='both') # remove the ticks
+    #axins1.xaxis.set_ticks_position("bottom")
 
     #axd['B'].set_ylim([0,1])
     axd['B'].set_yscale('log')
@@ -1040,7 +1055,6 @@ def plot_flare_checker(outdir, method=None):
[…]

### 2021-06-15 paper/ms.tex
minor

@@
 For instance, the ``mini-Neptune'' planets, thought to be made of
-rocky cores and extended atmospheric envelopes of hydrogen and helium,
-are expected to shrink in size by factors of several over their first
-$10^8$ years \citep{owen_constraining_2020}.  Specifically, they start
-with sizes of 4--12\,$R_\oplus$ at the time of disk dispersal
-($\approx$$10^7$\,years), and shrink to sizes of 2--4\,$R_\oplus$ by
-10$^8$ years.  This change is driven by a combination of stellar
-irradiation and internal heat that powers an outflow, which eventually
-depletes or entirely strips the envelope from the rocky core
-\citep{Owen_Wu_2013,gupta_sculpting_2019}.  Discovering young planets
-and measuring their compositions and atmospheric outflows are key
-steps toward testing this paradigm, which is often invoked to explain
-the observed radius distribution of both mature exoplanets
-\citep{Fulton_et_al_2017}, as well as younger planets
-\citep{bouma_cluster_2020}.
+molten rocky cores and extended atmospheric envelopes of hydrogen and
+helium, are expected to shrink in size by factors of several over
+their first $10^8$ years \citep{owen_constraining_2020}.
+Specifically, they start with sizes of 4--12\,$R_\oplus$ at the time
+of disk dispersal ($\approx$$10^7$\,years), and shrink to sizes of
+2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
+combination of stellar irradiation and internal heat that powers an
+outflow, which eventually depletes or entirely strips the envelope
+from the rocky core \citep{Owen_Wu_2013,gupta_sculpting_2019}.
+Discovering young planets and measuring their compositions and
+atmospheric outflows are key steps toward testing this paradigm, which
+is often invoked to explain the observed radius distribution of both
+mature exoplanets \citep{Fulton_et_al_2017}, as well as younger
+planets \citep{bouma_cluster_2020}.
 

### 2021-06-15 rudolf/plotting.py

@@ -389,7 +389,7 @@ def plot_skychart(outdir, narrowlims=0, showkepler=0, showtess=0,
 
         cax = ax.scatter(
             mdf.ra, mdf.dec, c=mdf.n_tess_sector, cmap=cmap,
-            alpha=0.9, zorder=40, s=2, rasterized=True, linewidths=0,
+            alpha=0.9, zorder=40, s=4, rasterized=True, linewidths=0,
             marker='.', norm=norm,
         )
 
@@ -484,9 +484,6 @@ def plot_skychart(outdir, narrowlims=0, showkepler=0, showtess=0,
                     connectionstyle='angle3', linewidth=1
                 )
             )
-
-
-
 
     if not showtess:
         leg = ax.legend(loc='upper left', handletextpad=0.1,

### 2021-06-16 paper/ms.tex
update ms

@@
   nearby ($d\approx 333\,{\rm pc}$) and resembles the Sun.  Combining
-  data from the Gaia satellite with rotation periods from TESS, we
-  show that Kepler\,1627 is a member of the $35\pm 10$\,Myr old
+  data from Gaia with rotation periods from TESS, we
+  show that Kepler\,1627 is a member of the $35\pm XX$\,Myr old
   $\delta$~Lyr cluster.  Our analysis advances Kepler\,1627Ab as the
-  youngest planet with a precise age detected by the main Kepler
-  mission.  New ages from the combination of Gaia and TESS offer the
-  opportunity to significantly expand the census of age-dated planets
-  -- we offer a literature concatenation of young stars to support
-  this effort.  Future studies of \pn\ may also help clarify how the
-  orbits and atmospheres of the mini-Neptune planets evolve.
+  youngest planet with a precise age found by the main Kepler
+  mission.  
+  Leveraging the corresponding photometry, we explore the possibility
+  that star-planet interactions induce flares in the system, and find
+  evidence that starspot crossings distort the average transit shape.
+  More broadly though, the merging of Gaia and TESS data offers an
+  opportunity to significantly expand the census of age-dated planets;
+  we provide a literature concatenation of young stars to support this
+  effort.
+  In the future, spectroscopic studies of \pn\ may also help clarify
+  how the orbits and atmospheres of the mini-Neptune planets evolve.
 \end{abstract}
@@
 
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f9.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Transit model residuals through time}.  
+    {\it Left:}
+		Phase-folded transit of Kepler 1627b, with stellar
+    variability removed, binned by Kepler quarter.
+    Black points are binned to 20 minute intervals.
+    The 1-$\sigma$ model uncertainties and the maximum {\it a
+    posteriori} model are shown as the faint purple band, and the dark
+    purple line.
+    {\it Right:}
+    As on the left, with the transit removed.
+    Quarters 6 and 7 show a consistent deviation in the second half of
+    the transit.
+		\label{fig:phasequarter}
+	}
+\end{figure*}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f10.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Transit model residuals through time (Part 2)}.  
+    {\it Left:}
+		Phase-folded transit of Kepler 1627b, with stellar
+    variability removed, binned by year of observation.
+    Points and models are as in Figure~\ref{fig:phasequarter}.
+    {\it Right:}
+    As on the left, with the transit removed.
+		\label{fig:phaseyear}
+	}
+\end{figure*}
+
 We assumed a quadratic limb-darkening law, with the uninformative prior
@@
 \texttt{RotTerm} GP kernel, with a logjitter term to inflate the
-uncertainties to account for excess white noise.
-(This results in an inflation of the uncertainties by a factor
-of about three).
+uncertainties to account for excess white noise.  (This results in an
+inflation of the uncertainties by a factor of about three).
 
 We considered including an additive \texttt{SHOTerm} kernel to account
-for stochastic noise (RotStochGPtransit).  This didn't seem to affect the results much, so
-we opted for the simpler model.
+for stochastic noise (RotStochGPtransit).  This didn't seem to affect
+the results much, so we opted for the simpler model.
 
@@
 sense that the observations are systematically high in the first half
-of transit, and low in the second half.  
-One possible explanation could be that we are seeing the effects of
-gravity darkening (e.g., CITE Masuda 2015).
-The expected amplitude of the effect is XXX, based on CITE.
-However, there is also a dip of comparable amplitude shortly before
-the transit. 
-Given the uncertainties in the stellar variability model, we expect
-that these patterns may correspond to more noise than signal.
+of transit, and low in the second half.  This is the phase of maximal
+scatter during each orbital period (TODO: in vs out of transit
+scatter).  The amplitude of the anomaly is $\approx$30\,ppm.
+
+To explore the origin of the anomaly, we binned the Kepler data over
+quarters (Figure~\ref{fig:phasequarter}) and years
+(Figure~\ref{fig:phaseyear}).  In Figure~\ref{fig:phasequarter}
+Quarter 6 shows the strongest asymmetry out of any of the quarters: a
+deviation of about 3\,ppt from expectation.  Quarter 7 shows an
+anomaly at roughly the same transit phase.  Year 2 correspondingly
+shows the strongest anomaly out of any year in
+Figure~\ref{fig:phaseyear}.
+
+We considered three possible explanations for the anomaly: gravity
+darkening, transit timing variations, and spot-crossing events.
+
+Gravity darkening (e.g., CITE Masuda 2015) is based on the premise
+that the rapidly rotating star becomes oblate, and brighter near the
+poles than the equator.  The fractional shape change due to gravity
+darkening is on the order of $(P_{\rm break}/P_{\rm rot})^2$, for
+$P_{\rm break}$ the break-up rotation period, and $P_{\rm rot}$ the
+rotation period.  Using the parameters from Table~\ref{tab:posterior},
+this yields an expected 1.6\% distortion of the $\approx$1.8\,ppt
+transit depth: {\it i.e.}, an absolute deviation of $\approx$3\,ppm.
+This is smaller than the observed anomaly by an order of magnitude,
+and therefore seems unlikely.
+
+The scenario of transit timing variations (TTVs) producing the
+asymmetry seems unlikely, since the analysis by
+\citet{holczer_transit_2016} implies that any such variations in \sn\
+need to be less than of order minutes.  Figures~\ref{fig:phasequarter}
+and~\ref{fig:phaseyear} also provide little evidence in support of
+this possibility.
+
+The final possibility is that of starspot crossings.  Given the high
+expected spot-covering fractions for young stars ({\it e.g.},
+\citealt{morris_relationship_2020}, \citealt{plavchan_planet_2020}),
+\pn\ may cross spot groups on the stellar surface in projection.
+Spot-crossing anomalies often reach amplitudes exceeding 100\,ppm
+({\it e.g.}, CITE Dai 2018, maybe Southworth's recent stuff).  For our
+system,  $P_{\rm orb}/P_{\rm rot} \approx 2.76$.  This means that
+every 4 transits (and 11 stellar rotations), the planet crosses over
+roughly the same stellar longitude.  Over a given quarter, this could
+occur at most three times, assuming the spot groups are persistent.
+
+This could provide a plausible path to creating the distorted transit
+signal, based on the amplitudes involved.  However, the typical S/N
+per Kepler transit is $\approx8$, and so the individual spot-crossing
+events cannot be resolved.  {\it A priori}, one would also expect the
+spot-crossing events to not have a preferred phase, so that they would
+average out over the $\approx200$ Kepler transits.  Nonetheless, given
+the available data, it is our best explanation for what we believe is
+real additional scatter during the transits, relative to the
+out-of-transit observations.
 

### 2021-06-16 drivers/run_RotGPtransit.py

@@ -26,8 +26,8 @@ from betty.paths import BETTYDIR
 
 # NOTE: change starid as desired based on the dataset to use.
 # Kepler_1627_Q15slc, or Kepler_1627
-#def run_RotGPtransit(starid='Kepler_1627', N_samples=1000):
-def run_RotGPtransit(starid='Kepler_1627_Q15slc', N_samples=1000):
+#def run_RotGPtransit(starid='Kepler_1627_Q15slc', N_samples=1000):
+def run_RotGPtransit(starid='Kepler_1627', N_samples=1000):
 
     assert starid in ['Kepler_1627', 'Kepler_1627_Q15slc']
 
@@ -85,6 +85,35 @@ def run_RotGPtransit(starid='Kepler_1627_Q15slc', N_samples=1000):
     phaseplot = 1
     cornerplot = 1
     posttable = 1
+    phasedsubsets = 1
+
+    if phasedsubsets:
+        outpath = join(PLOTDIR, f'{starid}_{modelid}_phasedsubsets_yearchunk.png')
+        timesubsets = [(50,450,'Y1'), (450,850,'Y2'), (850,1250,'Y3'),
+                       (1250,1650,'Y4') ]
+        ylimd = {'0':[-4.5,0.5], '1':[-4.5,0.5]}
+        bp.plot_phased_subsets(
+            datasets, m.trace.posterior, outpath, timesubsets, from_trace=True,
+            map_estimate=m.map_estimate, yoffsetNsigma=5, ylimd=ylimd,
+            inch_per_subset=0.75
+        )
+
+        timepath = os.path.join(
+            DATADIR,'phot','time_to_quarter_conversion.csv'
+        )
+        tdf = pd.read_csv(timepath)
+        outpath = join(PLOTDIR, f'{starid}_{modelid}_phasedsubsets_quarterchunk.png')
+        timesubsets = [(r['tstart'], r['tstop'], 'Q'+str(int(r['quarter'])))
+                       for _,r in tdf.iterrows() if int(r['quarter'])!=0]
+        ylimd = {'0':[-10.5,0.5], '1':[-10.5,0.5]}
+        bp.plot_phased_subsets(
+            datasets, m.trace.posterior, outpath, timesubsets, from_trace=True,
+            map_estimate=m.map_estimate, yoffsetNsigma=3.5, ylimd=ylimd,
+            inch_per_subset=0.35
+        )
+        #FIXME
+        assert 0
+
 
[…]

### 2021-06-21 paper/ms.tex
add ttv vs local slope analysis

@@
 \begin{abstract}
-  Kepler\,1627A is an active G8V star previously found to host a
+  Kepler\,1627A is a G8V star previously found to host a
   $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
@@
   nearby ($d\approx 333\,{\rm pc}$) and resembles the Sun.  Combining
-  data from Gaia with rotation periods from TESS, we
-  show that Kepler\,1627 is a member of the $35\pm XX$\,Myr old
-  $\delta$~Lyr cluster.  Our analysis advances Kepler\,1627Ab as the
-  youngest planet with a precise age found by the main Kepler
-  mission.  
-  Leveraging the corresponding photometry, we explore the possibility
-  that star-planet interactions induce flares in the system, and find
-  evidence that starspot crossings distort the average transit shape.
-  More broadly though, the merging of Gaia and TESS data offers an
-  opportunity to significantly expand the census of age-dated planets;
-  we provide a literature concatenation of young stars to support this
-  effort.
-  In the future, spectroscopic studies of \pn\ may also help clarify
-  how the orbits and atmospheres of the mini-Neptune planets evolve.
+  data from Gaia with rotation periods from TESS, we show that
+  Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
+  cluster.  Our analysis advances Kepler\,1627Ab as the youngest
+  planet with a precise age found by the main Kepler mission.
+  Analyzing the photometry, we find evidence that starspot crossings
+  {\it i)} distort the average transit shape, and {\it ii)} induce
+  transit timing variations that suggest the planet is on a prograde
+  orbit.  The possibility that star-planet interactions could induce
+  flares in the system is also explored.  Future spectroscopic studies
+  of \pn\ may help clarify the early evolution of the orbit and
+  atmosphere for this mini-Neptune.  More generally, the merging of
+  Gaia and TESS data offers an opportunity to significantly expand the
+  census of age-dated planets; we provide a literature concatenation
+  of young stars to support this effort.
 \end{abstract}
@@
 	\caption{
-    {\bf \cn is $35\pm15$\,Myr old.} {\it Left:} TESS and
+    {\bf The \cn\ is $35\pm15$\,Myr old.} {\it Left:} TESS and
     Kepler rotations periods and dereddened Gaia colors, with the
@@
 
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=1\textwidth]{f4.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Zoomed view of the \sn\ light curve.}  
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\includegraphics[width=1\textwidth]{f4.pdf}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+% 	\caption{
+% 		{\bf Zoomed view of the \sn\ light curve.}  
+% 		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
+% 		30 minute cadence.  A one hundred day segment is shown here.  The {\it top panel} shows the
+% 		\texttt{PDCSAP} median-subtracted flux in units of
+% 		parts-per-thousand ($\times 10^{-3}$).  The dominant signal
+% 		is induced by starspots.  The model for the
+% 		stellar variability (orange line) is subtracted in the {\it
+% 			bottom panel}, revealing the transits of \pn, as well as other
+% 		deviations from the stellar variability model.  The 
+% 		Figure Set available online shows the entire 3.9 years of
+% 		available data.
+% 		\label{fig:lightcurve}
+% 	}
+% \end{figure*}
+% 
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\includegraphics[width=1\textwidth]{f5.pdf}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+% 	\caption{
+% 		{\bf Phase-folded transit of Kepler 1627b, with stellar
+% 			variability removed.}  
+%     The 1-$\sigma$ model uncertainties and the maximum {\it a
+%     posteriori} model are shown as the faint purple band, and the dark
+%     purple line.  Gray points are individual \texttt{PDCSAP} flux
+%     measurements; black points are binned to 20 minutes, with a
+%     representative error bar shown.
+%     {\bf LGB FIXME: seems like the uncertainty is bigger than the RMS
+%     in the out of transit...}
+% 		\label{fig:phasefold}
+% 	}
+% \end{figure*}
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\subfloat{
+			\includegraphics[width=0.9\textwidth]{f4.pdf}
+			\includegraphics[width=0.9\textwidth]{f5.pdf}
+		}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf The light curve of Kepler\,1627.}
+    {\it Top}: 
 		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
@@
 		parts-per-thousand ($\times 10^{-3}$).  The dominant signal
-		is induced by starspots and plages.  The model for the
-		stellar variability (orange line) is subtracted in the {\it
-			bottom panel}, revealing the transits of \pn, as well as other
-		deviations from the stellar variability model.  The 
-		Figure Set available online shows the entire 3.9 years of
-		available data.
-		\label{fig:lightcurve}
-	}
-\end{figure*}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=1\textwidth]{f5.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Phase-folded transit of Kepler 1627b, with stellar
-			variability removed.}  
-    The 1-$\sigma$ model uncertainties and the maximum {\it a
+		is induced by starspots.  The model for the
+    stellar variability (orange line) is subtracted below, revealing
+    the transits of \pn, as well as other deviations from the stellar
+    variability model.  The Figure Set available online shows the
+    entire 3.9 years of available data.
+    {\it Bottom}:
+		Phase-folded transit of Kepler 1627b, with stellar
+			variability removed.
+    The 2-$\sigma$ model uncertainties and the maximum {\it a
     posteriori} model are shown as the faint purple band, and the dark
@@
     representative error bar shown.
-    {\bf LGB FIXME: seems like the uncertainty is bigger than the RMS
-    in the out of transit...}
-		\label{fig:phasefold}
+    The in-transit variation is robust against detrending methods; we
+    believe it is due to starspot crossings (see text).
+    \label{fig:lc}
   }
@@
 XXX} light curves Kepler acquired for the \cn, there is already a
-known planet in the cluster: Kepler\,1627b.  We considered the
-planet's previous statistical validation by \citet{morton_false_2016}
-to be sufficient cause to examine the age of the star and its cluster more closely.
+known planet in the cluster: Kepler\,1627b.  Given the
+planet's previous statistical validation by \citet{morton_false_2016},
+we decided to examine the age of the star and its cluster more closely.
 We collected the previously measured {\bf XXX} Kepler
@@
 
-If the Kepler\,1627b transit signal truly corresponds to a planet,
+If the Kepler\,1627b transit signal is created by a bonafide planet,
 this age would make it the youngest planet yet found by the main
@@
 orbiting this near-solar analog?  \citet{morton_false_2016} considered
-the various false positive scenarios, including those of a foreground
+the false positive scenarios of a foreground
 eclipsing binary, a hierarchical triple-star binary system, and a
@@
 statistical validation approach arise when attempting to separate the
-smallest stars from identically-sized gas giant planets (CITE).  With a size
+smallest stars from identically-sized gas giant planets.  With a size
 of $\approx 3.6\,R_\oplus$, Kepler\,1627b does not meet this
-concern.  The fact that we detect a low-mass stellar companion in
+concern.
+The fact that we detect a low-mass stellar companion in
 archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does not change
@@
 
+Two additional lines of evidence ultimately persuaded us in favor of
+the planetary interpretation.  First, the $\approx$100 days of
+1-minute short-cadence data enables a more precise measurement of the
+transit impact parameter: {\bf todo numbers} it is constrained to be
+$<0.62$ (3-$\sigma$) from the 1-minute cadenced data, compared to
+$<0.XX$ (3-$\sigma$) from the 30-minute data.  The second, and
+ultimately convincing line of evidence came from our transit timing
+analysis.  {\bf todo: KM pls check numbers} We isolated each of the
+144 observed transits to within $\pm4.5$\,hr of each transit, and
+fitted each window with both {\it i)} a local second-order polynomial
+and transit, and {\it ii)} a local linear trend.  We let the mid-time
+of each transit float, and then calculated the residual between the
+measured mid-time and that of a perfectly periodic orbit.  This
+residual, the ``transit timing variation'' (TTV), is plotted in
+Figure~\ref{fig:ttvslope} against the local linear slope.  Only a
+handful of Kepler Objects of Interest have shown this correlation
+\citep{holczer_time_2015}, which is most readily interpreted as a TTV
+induced by unresolved starspot crossings \citep{mazeh_time_2015}.  
+This is of course only possible if the planet is transiting the
+primary star, which excludes the any of the aforementioned false
+positive scenarios.
+It also implies that the planet's orbit is prograde.  The latter
+interpretation assumes that the dominant
+photometric variability is induced by dark spots, rather than bright
+faculae.  Given the observed transition of Sun-like stellar
+variability from spot to faculae-dominated regimes, we expect this
+inference to be reasonably secure
+\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f11.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Evidence for a prograde orbit from transit timing variations.}
+    The time of each transit was measured, along with the local linear
+    slope.  The two quantities are anti-correlated, which is most
+    easily explained by starspot crossings during the first
+    (second) half of transit inducing a positive (negative) TTV,
+    provided that the orbit is prograde \citep{mazeh_time_2015}.
+		\label{fig:ttvslope}
+	}
+\end{figure*}
+
 % Could the planet be on a high obliquity orbit?
@@
 
-Kepler\,1627Ab therefore provides an interesting extremum in the ages
+In short, Kepler\,1627Ab provides an interesting extremum in the ages
 of the Kepler planets, and opens multiple avenues for study from the
@@
 \approx 20$\ms) with HIRES or the upcoming KPF spectrograph could
-yield a measurement of its orbital obliquity (CITEP,CITEP).
-Similarly, transit spectroscopy aimed at detecting either hydrogen or
-helium absorption could yield insight into the planet's atmosphere,
-which is expected to be outflowing at a rate of {\bf
+yield a quantification of the stellar obliquity, and a means for
+verifying the suggestion of a prograde order from transit timing
+(CITEP,CITEP).  Similar transit spectroscopy aimed at detecting either
+hydrogen or helium absorption could yield insight into the planet's
+atmosphere, which is expected to be outflowing at a rate of {\bf
 X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  A more challenging
@@
 The precision of the Kepler data itself may enable more detailed
-analyses without even needing to acquire more data.  These could
-include studies transit timing variations, distortion of the transit
-signal due to the star's gravity darkening, and interactions between
-the star and planet exhibited by patterns in the star's flaring.
-The gravity darkening might be interesting as an explanation for the
-$\approx80\,{\rm ppm}$ residual apparent in our transit fit.
-On the point of planet-induced flaring, Appendix REFERENCE
-highlights an odd timing coincidence that may be worthy of further
-examination.
-
-Ultimately though, our age measurement for Kepler\,6127Ab was enabled by
-identifying its connection to the \cn\ through the Gaia data.  The
+analyses without even needing to acquire more data.  While our transit
+timing study did not yield the detection of any clearly significant
+planet-induced TTVs, the star's flaring may be suggestive of
+star-planet interactions.  Appendix REFERENCE highlights an odd flare
+timing coincidence that may be worthy of further examination.
+
+Ultimately though, our age measurement for \pn\ was enabled
+by identifying its connection to the \cn\ through the Gaia data.  The
 Supplementary Table~\ref{tab:v05} enables similar crossmatches to
 already known planets, as well as forthcoming planets from the TESS
-mission (CITE RickerJATIS2014).  An alternative approach also
-likely to bear fruit is to identify kinematic associations
-around exoplanet host stars using the Gaia data, and to then verify
-these associations with ancilliary rotation periods and spectroscopy
-(CITEP Tofflemire).
+mission (CITE RickerJATIS2014).  An alternative approach also likely
+to bear fruit is to identify kinematic associations around exoplanet
+host stars using the Gaia data, and to then verify these associations
+with ancilliary rotation periods and spectroscopy (CITEP Tofflemire).
 
@@
 L.G.B{.} and J.L.C{.} are grateful to T{.}~David for help with the
-transit fitting, and to R{.}~Kerr for kindly providing us with the
-\citet{Kerr2021} membership list prior to its publication.
+transit fitting, to R{.}~Kerr for kindly providing us with the
+\citet{Kerr2021} membership list prior to its publication, and to
+K.~Masuda for advice on gravity darkening.
 % is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
@@
 
+\subsection{Long Cadence Data}
 \begin{figure*}[t]
@@
 signal, based on the amplitudes involved.  However, the typical S/N
-per Kepler transit is $\approx8$, and so the individual spot-crossing
-events cannot be resolved.  {\it A priori}, one would also expect the
-spot-crossing events to not have a preferred phase, so that they would
-average out over the $\approx200$ Kepler transits.  Nonetheless, given
-the available data, it is our best explanation for what we believe is
-real additional scatter during the transits, relative to the
-out-of-transit observations.
+per Kepler transit is $\approx8$, making individual spot-crossing
+unresolved, and this scenario challenging to test.  {\it A priori},
+one would also expect the spot-crossing events to not have a preferred
+orbital phase, so that they would average out over the $\approx200$
+Kepler transits.  Nonetheless, given the available data, it is our
+best explanation for what we believe is real additional photometric
+scatter observed during the transits.
+
+\subsection{Short Cadence Data}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f12.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Phase-folded short cadence transit of Kepler\,1627 Ab}.  
+    The precision with which the impact parameter can be measured is
+    higher than from the long cadence data.
+		\label{fig:phase_slc}
+	}
+\end{figure*}
+
+Figure~\ref{fig:phase_slc} shows the result from analyzing 100 days of
+short cadence (1-minute) data acquired during Quarter 15.
+%TODO: get quantitative on better impact parameter...
 
@@
     Zooms of the brightest, and third-brightest flares.  A timing
-    coincidence -- that both flares has ``successors'' approximately
+    coincidence -- that both flares have ``successors'' approximately
     one orbital period after the initial event -- is emphasized.
@@
     {\it Bottom:}
-    Fitted flare amplitudes and orbital phases.
-    Colors indiate relative time, from the beginning of the
-    98-day short-cadence Q15 dataset (dark blue) to the end (light
+    Fitted flare amplitudes and orbital phases -- each point
+    represents one flare.
+    Colors indicate relative time, from the beginning of the
+    98-day short-cadence Quarter 15 dataset (dark blue) to the end (light
     yellow).
@@
 that despite the low flare duty cycle, one orbital period after the
-the brightest flare, a second flare followed.  This and a similar
+brightest flare, a second flare followed.  This and a similar
 event are shown in Figure~\ref{fig:flarezoom}.  The timing error is

### 2021-06-21 rudolf/plotting.py

@@ -22,6 +22,8 @@ import matplotlib as mpl
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
 
+from scipy.optimize import curve_fit
+
 from astropy import units as u, constants as const
 from astropy.coordinates import SkyCoord
 from astropy.io import fits
@@ -1197,4 +1199,100 @@ def plot_ttv(outdir, narrowylim=0):
     savefig(fig, outpath, dpi=400)
 
 
+def _linear_model(xdata, m, b):
+    return m*xdata + b
 
+
+def plot_ttv_vs_local_slope(outdir):
+
+    # get data
+    datadir = os.path.join(DATADIR, 'ttv', 'Bouma', 'K05245.01/')
+    tnum, tc, tc_err, tc_lin = np.loadtxt(datadir+"tc.dat").T
+
+    ttv = tc - tc_lin
+
+    slopes = []
+    for _tnum in tnum:
+        _t, _f, _ferr, _fmodel = np.loadtxt(datadir+"transit_%04d.dat"%int(_tnum)).T
+        coeffs = np.polyfit(_t, _f, deg=1)
+        slopes.append(coeffs[0])
+
+    slopes = np.array(slopes)
+
+    ttv_mean, ttv_std = np.mean(ttv), np.std(ttv)
+
+    ttv_cut = 0.02 # doesn't mean match
+
+    z1 = np.poly1d(
+        np.polyfit(slopes[np.abs(ttv)<ttv_cut],
+                   ttv[np.abs(ttv)<ttv_cut], deg=1)
+    )
+
+    slope_guess = -5e-2
+    intercept_guess = 0
+    sel = np.abs(ttv)<ttv_cut
+    # ignore the error bars... because they are probably wrong.
+    p_opt, p_cov = curve_fit(
+        _linear_model, slopes[sel], ttv[sel],
+        p0=(slope_guess, intercept_guess)
+    )
+    lsfit_slope = p_opt[0]
+    lsfit_slope_err = p_cov[0,0]**0.5
+    lsfit_int = p_opt[1]
+    lsfit_int_err = p_cov[1,1]**0.5
+
+
+    # make plot
+    plt.close('all')
+    set_style()
+
+    fig, ax = plt.subplots(figsize=(4,3))
+
+    ax.set_ylabel("TTV [minutes]")
[…]

### 2021-06-21 paper/ms.tex
update ms

@@
 			\includegraphics[width=0.9\textwidth]{f4.pdf}
-			\includegraphics[width=0.9\textwidth]{f5.pdf}
+		}
+
+		\subfloat{
+			\includegraphics[width=0.87\textwidth]{f5.pdf}
 		}
@@
     representative error bar shown.
-    The in-transit variation is robust against detrending methods; we
-    believe it is due to starspot crossings (see text).
+    The in-transit residual is robust against detrending methods; we
+    believe it is caused by starspot crossings (see text).
     \label{fig:lc}
@@
 		{\bf Evidence for a prograde orbit from transit timing variations.}
-    The time of each transit was measured, along with the local linear
-    slope.  The two quantities are anti-correlated, which is most
+    The time of each transit was measured, along with the local
+    slope of the light curve.  The two quantities appear anti-correlated, which is most
     easily explained by starspot crossings during the first

### 2021-07-02 paper/ms.tex
add Prot window slider and unc calibration

@@
   Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
-  cluster.  Our analysis advances Kepler\,1627Ab as the youngest
-  planet with a precise age found by the main Kepler mission.
+  cluster.  Kepler\,1627Ab is therefore the youngest planet with a
+  precise age found so far by the main Kepler mission.
   Analyzing the photometry, we find evidence that starspot crossings
-  {\it i)} distort the average transit shape, and {\it ii)} induce
-  transit timing variations that suggest the planet is on a prograde
-  orbit.  The possibility that star-planet interactions could induce
+  distort the average transit shape, and induce transit timing
+  variations that suggest the planet is on a prograde orbit.  The
+  possibility that star-planet interactions could induce
   flares in the system is also explored.  Future spectroscopic studies
-  of \pn\ may help clarify the early evolution of the orbit and
-  atmosphere for this mini-Neptune.  More generally, the merging of
-  Gaia and TESS data offers an opportunity to significantly expand the
-  census of age-dated planets; we provide a literature concatenation
-  of young stars to support this effort.
+  of \pn\ may help clarify how the orbit and atmosphere of this
+  mini-Neptune evolve.  More generally, the combination of Gaia and TESS
+  offers an opportunity to significantly expand the census of
+  age-dated planets; we provide a literature concatenation of young
+  stars to support this effort.
 \end{abstract}
@@
 stars, the vast majority of them are several billion years old.  This
-results in difficulty testing theories for the origins of the
+makes it difficult to test theories for the origins of the
 different families of planets, since many expected evolutionary
@@
 Specifically, they start with sizes of 4--12\,$R_\oplus$ at the time
-of disk dispersal ($\approx$$10^7$\,years), and shrink to sizes of
+of disk dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
 2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
@@
 		points are kinematically selected members of the $\delta$\,Lyr
-		cluster (black points in Figure~\ref{fig:XYZvtang}).  Both Kepler
+		cluster.  Both Kepler
 		(gray squares) and TESS (colored points) observed portions of the
@@
     representative error bar shown.
-    The in-transit residual is robust against detrending methods; we
-    believe it is caused by starspot crossings (see text).
+    The asymmetric residual visible during the transit is robust
+    against detrending methods; we believe it is caused by starspot
+    crossings.
     \label{fig:lc}
@@
 by J. Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
-open cluster Theia\,520 (UBC\,1).  At least five Kepler planets are
+open cluster Theia\,520 (UBC\,1).  At least six Kepler planets are
 known in the cluster, including the Kepler-52 and Kepler-968 systems.
@@
 
-While we have yet to perform an in-depth planet search of the {\bf
-XXX} light curves Kepler acquired for the \cn, there is already a
-known planet in the cluster: Kepler\,1627b.  Given the
+Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was previosuly
+reported to host a close-in mini-Neptune planet candidate,
+Kepler\,1627Ab
+\citep{2012ApJS..199...24T,thompson_planetary_2018}
+Given the
 planet's previous statistical validation by \citet{morton_false_2016},
@@
 
-If the Kepler\,1627b transit signal is created by a bonafide planet,
-this age would make it the youngest planet yet found by the main
-Kepler mission.  A portion of the light curve is shown in
-Figure~\ref{fig:lightcurve}, and the phase-folded transit is shown in
-Figure~\ref{fig:phasefold}.  The dominant signal in the
+% rotation period from plot_rotation_period_windowslider.py
+
+If the \pn\ transit signal is created by a bonafide planet, this age
+would make it the youngest planet yet found by the main Kepler
+mission\footnote{The re-purposed K2 mission found two slightly younger
+systems: K2-33b \citep{David_et_al_2017,Mann_K2_33b_2016} and V1298
+Tau \citep{david_four_2019}}.  A portion of the light curve is shown
+in Figure~\ref{fig:lightcurve}, and the phase-folded transit is shown
+in Figure~\ref{fig:phasefold}.  The dominant signal in the
 \texttt{PDCSAP} photometry is a quasi-period starspot signal with a
@@
 depending on the degree of asymmetry in spot coverage between the
-stellar hemispheres.  The starspot signal repeats roughly every
-2.606\,days, though this value varies by a few percent depending on
-which portion of the light curve is analyzed {\bf (FIXME: do!)}.  The
-details of how we model this signal to derive the parameters of the
-planet are discussed in Section~\ref{app:gptransit}.  The results,
-given in Table~\ref{tab:posterior}, are consistent with a mini-Neptune
-sized planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit,
-around a G8V star ($0.91 \pm 0.05 R_\odot$) {\bf fixme: age}.
+stellar hemispheres.  The starspot signal repeats every
+$2.642\pm0.042$\,days, where we have adopted an uncertainty based on
+the scatter observed between different windows (quarters) of the
+time-series data.  The details of how we model this starspot signal to
+derive the parameters of the planet are discussed in
+Section~\ref{app:gptransit}.  The results, given in
+Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
+planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit
+($e<0.XX$), around
+a G8V star ($0.91 \pm 0.05 R_\odot$). 
+{\bf TODO quote the eccentricity limit}
 

### 2021-07-02 rudolf/helpers.py

@@ -46,7 +46,9 @@ def get_kep1627_kepler_lightcurve(lctype='longcadence'):
     each quater.
     """
 
-    if lctype == 'longcadence':
+    assert lctype in ['longcadence', 'shortcadence', 'longcadence_byquarter']
+
+    if lctype in ['longcadence', 'longcadence_byquarter']:
         lcfiles = glob(os.path.join(DATADIR, 'phot', 'kplr*_llc.fits'))
     elif lctype == 'shortcadence':
         lcfiles = glob(os.path.join(DATADIR, 'phot', 'full_MAST_sc', 'MAST_*',
@@ -79,11 +81,20 @@ def get_kep1627_kepler_lightcurve(lctype='longcadence'):
         qual_list.append(qual[sel])
         texp_list.append(texp)
 
-    time = np.hstack(timelist)
-    flux = np.hstack(f_list)
-    flux_err = np.hstack(ferr_list)
-    qual = np.hstack(qual_list)
-    texp = np.nanmedian(texp_list)
+    if lctype == 'longcadence':
+        time = np.hstack(timelist)
+        flux = np.hstack(f_list)
+        flux_err = np.hstack(ferr_list)
+        qual = np.hstack(qual_list)
+        texp = np.nanmedian(texp_list)
+    elif lctype == 'longcadence_byquarter':
+        return (
+            timelist,
+            f_list,
+            ferr_list,
+            qual_list,
+            texp_list
+        )
 
     # require ascending time
     s = np.argsort(time)

### 2021-07-02 rudolf/plotting.py

@@ -10,6 +10,7 @@ plot_keplerlc
 plot_flare_checker
 plot_ttv
 plot_ttv_vs_local_slope
+plot_rotation_period_windowslider
 """
 import os, corner, pickle, inspect
 from glob import glob
@@ -1297,3 +1298,43 @@ def plot_ttv_vs_local_slope(outdir):
 
     outpath = os.path.join(outdir, f'ttv_vs_local_slope{s}.png')
     savefig(fig, outpath, dpi=400)
+
+
+def plot_rotation_period_windowslider(outdir):
+
+    from timmy.rotationperiod import measure_rotation_period_and_unc
+
+    # get data
+    modelid, starid = 'gptransit', 'Kepler_1627'
+    datasets = OrderedDict()
+    time, flux, flux_err, qual, texp = (
+        get_kep1627_kepler_lightcurve(lctype='longcadence_byquarter')
+    )
+
+    N_quarters = len(time)
+
+    periods, period_uncs = [], []
+    for i in range(N_quarters):
+        print(i)
+
+        t,f = time[i],flux[i]
+
+        plotpath = os.path.join(outdir, f'kep1627_rotationperiodslider_quarter_ix{i}.png')
+        p, p_unc = measure_rotation_period_and_unc(t, f, 1, 10,
+                                                   period_fit_cut=0.5, nterms=1,
+                                                   samples_per_peak=50,
+                                                   plotpath=plotpath)
+
+        periods.append(p)
+        period_uncs.append(p_unc)
+
+    outdf = pd.DataFrame(
+        {'period': periods, 'period_unc': period_uncs}
+    )
+    outpath = os.path.join(outdir, 'rotation_period_windowslider_QUARTERS.csv')
+    outdf.to_csv(outpath, index=False)
+
+    print(42*'-')
+    print(f'P: {np.mean(outdf.period):.4f}, scatter: {np.std(outdf.period):.4f}')
+    print(outdf)
+    print(42*'-')

### 2021-07-02 paper/authors.tex
wow, nature papers are short
add ttv vs local slope analysis
update ms

@@
 \correspondingauthor{L.\,G.\,Bouma}
-\email{luke@astro.princeton.edu}
+\email{bouma.luke@gmail.com}
 
@@
 \affiliation{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY 10027, USA}
-\affiliation{Department of Astrophysics, American Museum of Natural History, Central Park West, New York, NY 10024, USA}
+\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
+
+\author[0000-0003-1298-9699]{K. Masuda}
+\affiliation{Department of Earth and Space Science, Osaka University, Osaka 560-0043, Japan}

### 2021-07-02 paper/ms.tex
update ms

@@
 stars, the vast majority of them are several billion years old.  This
-makes it difficult to test theories for the origins of the
-different families of planets, since many expected evolutionary
-processes operate on timescales of less than 100 million years.
+makes it difficult to test theories for the origins of the different
+families of planets, since many evolutionary processes are expected to
+operate on timescales of less than 100 million years.
 
@@
 helium, are expected to shrink in size by factors of several over
-their first $10^8$ years \citep{owen_constraining_2020}.
-Specifically, they start with sizes of 4--12\,$R_\oplus$ at the time
-of disk dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
+their first $10^8$ years
+\citep{owen_constraining_2020,kite_atmosphere_2020}.  Specifically,
+they start with sizes of 4--12\,$R_\oplus$ at the time of disk
+dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
 2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
 combination of stellar irradiation and internal heat that powers an
-outflow, which eventually depletes or entirely strips the envelope
+outflow, which can eventually deplete or entirely strip the envelope
 from the rocky core \citep{Owen_Wu_2013,gupta_sculpting_2019}.
@@
 planets \citep{bouma_cluster_2020}.
+
+% TODO: decide whether a paragraph about young planet obliquities is
+% needed.
 
@@
     points are kinematically selected members of the $\delta$\,Lyr
-		cluster.  Both Kepler
-		(gray squares) and TESS (colored points) observed portions of the
-		cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray; four
-		of them (orange points) have their names annotated.  Kepler\,1627 (green
-		star) was observed during the entirety of the Kepler mission, and
-		has been observed  by TESS for two lunar months to date.
+    cluster.  Both Kepler (gray squares) and TESS (colored points)
+    observed portions of the cluster.  Naked-eye stars ($m_{\rm
+    V}<6.5$) are shown in gray; four of them (orange points) have
+    their names annotated.  Kepler\,1627 (green star) was observed
+    during the entirety of the Kepler mission, and has been observed
+    by TESS for two lunar months to date.
 		\label{fig:skychart}
@@
 	\caption{
-    {\bf The \cn\ is $35\pm15$\,Myr old.} {\it Left:} TESS and
-    Kepler rotations periods and dereddened Gaia colors, with the
-    Pleiades \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
+    {\bf The \cn\ is $35\pm15$\,Myr old.} {\it Left:} TESS and Kepler
+    rotations periods and dereddened Gaia colors, with the Pleiades
+    \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
     \citep[650\,Myr;][]{douglas_poking_2017} shown for reference.
@@
     against the LDB-age of IC\,2602.  {\it Right:} HR diagram of the
-    same comparison. {\bf todo: caption improve}.
+    same comparison. {\bf todo: caption improve. fix figures.}.
    \label{fig:age}
@@
 \end{figure*}
-
-% \begin{figure*}[t]
-% 	\begin{center}
-% 		\leavevmode
-% 		\includegraphics[width=1\textwidth]{f4.pdf}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-% 	\caption{
-% 		{\bf Zoomed view of the \sn\ light curve.}  
-% 		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
-% 		30 minute cadence.  A one hundred day segment is shown here.  The {\it top panel} shows the
-% 		\texttt{PDCSAP} median-subtracted flux in units of
-% 		parts-per-thousand ($\times 10^{-3}$).  The dominant signal
-% 		is induced by starspots.  The model for the
-% 		stellar variability (orange line) is subtracted in the {\it
-% 			bottom panel}, revealing the transits of \pn, as well as other
-% 		deviations from the stellar variability model.  The 
-% 		Figure Set available online shows the entire 3.9 years of
-% 		available data.
-% 		\label{fig:lightcurve}
-% 	}
-% \end{figure*}
-% 
-% \begin{figure*}[t]
-% 	\begin{center}
-% 		\leavevmode
-% 		\includegraphics[width=1\textwidth]{f5.pdf}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-% 	\caption{
-% 		{\bf Phase-folded transit of Kepler 1627b, with stellar
-% 			variability removed.}  
-%     The 1-$\sigma$ model uncertainties and the maximum {\it a
-%     posteriori} model are shown as the faint purple band, and the dark
-%     purple line.  Gray points are individual \texttt{PDCSAP} flux
-%     measurements; black points are binned to 20 minutes, with a
-%     representative error bar shown.
-%     {\bf LGB FIXME: seems like the uncertainty is bigger than the RMS
-%     in the out of transit...}
-% 		\label{fig:phasefold}
-% 	}
-% \end{figure*}
 
@@
     The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
-		30 minute cadence.  A one hundred day segment is shown here.  The {\it top panel} shows the
-		\texttt{PDCSAP} median-subtracted flux in units of
-		parts-per-thousand ($\times 10^{-3}$).  The dominant signal
-		is induced by starspots.  The model for the
-    stellar variability (orange line) is subtracted below, revealing
-    the transits of \pn, as well as other deviations from the stellar
+    30 minute cadence.  A one hundred day segment is shown here.  The
+    {\it top panel} shows the \texttt{PDCSAP} median-subtracted flux
+    in units of parts-per-thousand ($\times 10^{-3}$).  The dominant
+    signal is induced by starspots.  The model for the stellar
+    variability (orange line) is subtracted below, revealing the
+    transits of \pn, as well as other deviations from the stellar
     variability model.  The Figure Set available online shows the
@@
     {\it Bottom}:
-		Phase-folded transit of Kepler 1627b, with stellar
-			variability removed.
-    The 2-$\sigma$ model uncertainties and the maximum {\it a
-    posteriori} model are shown as the faint purple band, and the dark
-    purple line.  Gray points are individual \texttt{PDCSAP} flux
+    Phase-folded transit of Kepler 1627b, with stellar variability
+    removed.  The 2-$\sigma$ model uncertainties and the maximum {\it
+    a posteriori} model are shown as the faint purple band, and the
+    dark purple line.  Gray points are individual \texttt{PDCSAP} flux
     measurements; black points are binned to 20 minutes, with a
-    representative error bar shown.
-    The asymmetric residual visible during the transit is robust
-    against detrending methods; we believe it is caused by starspot
-    crossings.
+    representative error bar shown.  The asymmetric residual visible
+    during the transit is robust against detrending methods; we
+    believe it is caused by starspot crossings.
     \label{fig:lc}
@@
 from the literature, which yielded a list of candidate young and
-age-dated stars (see Section~\ref{app:targetlist}).
+age-dated stars (see Appendix~\ref{app:targetlist}).
 
 Comparing our young star list against the Kepler field yielded two
-discoveries.  The first, to be discussed in an upcoming analysis
-by J. Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
-open cluster Theia\,520 (UBC\,1).  At least six Kepler planets are
-known in the cluster, including the Kepler-52 and Kepler-968 systems.
-The second discovery, and the focus of this work, is that Kepler
-observed a portion of the $\delta$\,Lyr cluster, shown in
+discoveries.  The first, to be discussed in an upcoming analysis by J.
+Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
+cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
+the cluster, from the Kepler-52 and Kepler-968 systems
+\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
+discovery, and the focus of this work, is that Kepler observed a
+portion of the $\delta$\,Lyr cluster, shown in
 Figure~\ref{fig:skychart}.
 
-Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was previosuly
-reported to host a close-in mini-Neptune planet candidate,
-Kepler\,1627Ab
+{\bf Jason Curtis todo}
+The \cn\ contains
+Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470), which 
+was previously reported to host a close-in mini-Neptune candidate planet
 \citep{2012ApJS..199...24T,thompson_planetary_2018}
-Given the
-planet's previous statistical validation by \citet{morton_false_2016},
-we decided to examine the age of the star and its cluster more closely.
-We collected the previously measured {\bf XXX} Kepler
-rotation periods (CITEP MCQUILLAN2014), and measured {\bf YYY} new stellar
-rotation periods for cluster members observed by TESS.
-We also performed an isochronal analysis.
-The results are shown in Figure~\ref{fig:age}.
-From the ensemble perspective, which provides the strongest
-constraints (SEE CITET SODERBLOM 2014),
-the age of the star is $XX \pm YY$\,Myr.
-Individual measurements of the star's rotation period and lithium
-abundance corroborate this conclusion.
+Given the planet's previous statistical validation by
+\citet{morton_false_2016}, we decided to examine the age of the star
+and its cluster more closely.  We collected the previously measured
+{\bf XXX} Kepler rotation periods (CITEP MCQUILLAN2014), and measured
+{\bf YYY} new stellar rotation periods for cluster members observed by
+TESS.  We also performed an isochronal analysis.  The results are
+shown in Figure~\ref{fig:age}.  From the ensemble perspective, which
+provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
+of the star is $XX \pm YY$\,Myr.  Individual measurements of the
+star's rotation period and lithium abundance corroborate this
+conclusion.
 
@@
 Tau \citep{david_four_2019}}.  A portion of the light curve is shown
-in Figure~\ref{fig:lightcurve}, and the phase-folded transit is shown
-in Figure~\ref{fig:phasefold}.  The dominant signal in the
+with the phase-folded transit
+in Figure~\ref{fig:lc}.  The dominant signal in the
 \texttt{PDCSAP} photometry is a quasi-period starspot signal with a
@@
 $2.642\pm0.042$\,days, where we have adopted an uncertainty based on
-the scatter observed between different windows (quarters) of the
+the scatter observed between different portions (quarters) of the
 time-series data.  The details of how we model this starspot signal to
 derive the parameters of the planet are discussed in
-Section~\ref{app:gptransit}.  The results, given in
+Appendix~\ref{app:gptransit}.  The results, given in
 Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
 planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit
-($e<0.XX$), around
-a G8V star ($0.91 \pm 0.05 R_\odot$). 
-{\bf TODO quote the eccentricity limit}
+($e<0.XX$), around a G8V star ($0.91 \pm 0.05 R_\odot$).  {\bf TODO
+quote the eccentricity limit}
 
@@
 orbiting this near-solar analog?  \citet{morton_false_2016} considered
-the false positive scenarios of a foreground
-eclipsing binary, a hierarchical triple-star binary system, and a
-background eclipsing binary.  \citet{morton_false_2016} reported the
-latter scenario to be the only possible case, with a (model-dependent)
-probability of $\approx10^{-5}$.  While this statistical validation is
-reassuring, ``validated'' planets have been demonstrated to be false
-positives \citep{shporer_three_2017}.  Most failures of the
-statistical validation approach arise when attempting to separate the
-smallest stars from identically-sized gas giant planets.  With a size
-of $\approx 3.6\,R_\oplus$, Kepler\,1627b does not meet this
-concern.
-The fact that we detect a low-mass stellar companion in
-archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does not change
-the conclusion: the companion contributes 1\% of the flux
+the false positive scenarios of a foreground eclipsing binary, a
+hierarchical triple-star binary system, and a background eclipsing
+binary.  Based on the transit shape, \citet{morton_false_2016} found
+the latter scenario to be the only possible case, with a
+(model-dependent) probability of $\approx10^{-5}$.  While this
+statistical validation is reassuring, ``validated'' planets have been
+demonstrated to be false positives \citep{shporer_three_2017}.  Most
+failures of the statistical validation approach arise when attempting
+to separate the smallest stars from identically-sized gas giant
+planets.  With a size of $\approx 3.6\,R_\oplus$, Kepler\,1627b does
+not meet this concern.  The fact that we detect a low-mass stellar
+companion in archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does
+not change the conclusion: the companion contributes 1\% of the flux
 observed by Kepler, which combined with the observed transit shape
 cannot be invoked as an explanation for the observed data
-(Section~\ref{app:companionstar}).
-
-Two additional lines of evidence ultimately persuaded us in favor of
-the planetary interpretation.  First, the $\approx$100 days of
-1-minute short-cadence data enables a more precise measurement of the
-transit impact parameter: {\bf todo numbers} it is constrained to be
-$<0.62$ (3-$\sigma$) from the 1-minute cadenced data, compared to
-$<0.XX$ (3-$\sigma$) from the 30-minute data.  The second, and
-ultimately convincing line of evidence came from our transit timing
-analysis.  {\bf todo: KM pls check numbers} We isolated each of the
-144 observed transits to within $\pm4.5$\,hr of each transit, and
-fitted each window with both {\it i)} a local second-order polynomial
-and transit, and {\it ii)} a local linear trend.  We let the mid-time
-of each transit float, and then calculated the residual between the
-measured mid-time and that of a perfectly periodic orbit.  This
-residual, the ``transit timing variation'' (TTV), is plotted in
-Figure~\ref{fig:ttvslope} against the local linear slope.  Only a
-handful of Kepler Objects of Interest have shown this correlation
-\citep{holczer_time_2015}, which is most readily interpreted as a TTV
-induced by unresolved starspot crossings \citep{mazeh_time_2015}.  
-This is of course only possible if the planet is transiting the
-primary star, which excludes the any of the aforementioned false
-positive scenarios.
-It also implies that the planet's orbit is prograde.  The latter
-interpretation assumes that the dominant
-photometric variability is induced by dark spots, rather than bright
-faculae.  Given the observed transition of Sun-like stellar
-variability from spot to faculae-dominated regimes, we expect this
-inference to be reasonably secure
+(Appendix~\ref{app:companionstar}).
+
+{\bf todo: LGB to check numbers on full-dataset transit ecc and b} 
+Two additional new lines of evidence support the planetary
+interpretation.  First, the $\approx$100 days of 1-minute
+short-cadence data reduces the cadence-induced smearing enables a more
+precise measurement of the transit impact parameter
+\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
+(3-$\sigma$) from the 1-minute cadenced data, compared to $<0.XX$
+(3-$\sigma$) from the 30-minute data.
+
+{\bf todo: Kento to check numbers on per-transit analysis} 
+The second line of evidence came from our transit timing analysis. We
+isolated each of the 144 observed transits to within $\pm4.5$\,hr of
+each transit, and fitted each window with both {\it i)} a local
+second-order polynomial and transit, and {\it ii)} a local linear
+trend.  We let the mid-time of each transit float, and then calculated
+the residual between the measured mid-time and that of a perfectly
+periodic orbit.  This residual, the ``transit timing variation''
+(TTV), is plotted in Figure~\ref{fig:ttvslope} against the local
+linear slope.  Only a handful of Kepler Objects of Interest have shown
+this correlation \citep{holczer_time_2015}, which is most readily
+interpreted as a TTV induced by unresolved starspot crossings
+\citep{mazeh_time_2015}.  This is of course only possible if the
+planet is transiting the primary star, which excludes the any of the
+aforementioned false positive scenarios.  It also implies that the
+planet's orbit is prograde.  The latter interpretation assumes that
+the dominant photometric variability is induced by dark spots, rather
+than bright faculae.  Given the observed transition of Sun-like
+stellar variability from spot to faculae-dominated regimes, we expect
+this inference to be reasonably secure
 \citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
@@
 		{\bf Evidence for a prograde orbit from transit timing variations.}
-    The time of each transit was measured, along with the local
-    slope of the light curve.  The two quantities appear anti-correlated, which is most
-    easily explained by starspot crossings during the first
-    (second) half of transit inducing a positive (negative) TTV,
+    The time of each transit was measured, along with the local slope
+    of the light curve.  The two quantities appear anti-correlated,
+    which is most easily explained by starspot crossings during the
+    first (second) half of transit inducing a positive (negative) TTV,
     provided that the orbit is prograde \citep{mazeh_time_2015}.
@@
 
-In short, Kepler\,1627Ab provides an interesting extremum in the ages
-of the Kepler planets, and opens multiple avenues for study from the
-ground.  Observations of its Doppler transit ($\Delta v_{\rm RM}
-\approx 20$\ms) with HIRES or the upcoming KPF spectrograph could
-yield a quantification of the stellar obliquity, and a means for
-verifying the suggestion of a prograde order from transit timing
-(CITEP,CITEP).  Similar transit spectroscopy aimed at detecting either
-hydrogen or helium absorption could yield insight into the planet's
-atmosphere, which is expected to be outflowing at a rate of {\bf
+{\bf LGB TODO: ask James Owen what he thinks the mass loss rate is}
+Kepler\,1627Ab therefore provides a new extremum in the ages of the
+Kepler planets, and opens multiple avenues for study from the ground.
+Observations of its Doppler transit are expected to induce a
+Rossiter-McLaughlin anomaly with an amplitude of $\approx 20$\ms, and
+could yield a quantification of the stellar obliquity, and a means for
+verifying our transit-timing based suggestion of a prograde orbit.
+Similar transit spectroscopy aimed at detecting either hydrogen or
+helium absorption could yield insight into the planet's atmosphere,
+which is expected to be outflowing at a rate of {\bf
 X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  A more challenging
@@
 either detailed timing analyses of the Kepler data, or through a
-high-cadence multi-color radial velocity campaign.  The mass would
-yield constraints on both the planet's composition (``what is the
-extent of the initial gaseous envelope for a typical mini-Neptune?''),
-as well as its initial entropy (``hot'' or ``cold''? CITE).
+high-cadence multi-color radial velocity campaign.  The mass could
+yield constraints on both the planet's composition as well as its
+initial entropy \citep{owen_constraining_2020}.
 
@@
 planet-induced TTVs, the star's flaring may be suggestive of
-star-planet interactions.  Appendix REFERENCE highlights an odd flare
-timing coincidence that may be worthy of further examination.
-
-Ultimately though, our age measurement for \pn\ was enabled
-by identifying its connection to the \cn\ through the Gaia data.  The
+star-planet interactions.  Appendix~\ref{app:flare} highlights an odd flare
+timing coincidence that may be worthy of further examination by those
+more knowledgeable than us.
+
+Ultimately however, our age measurement for \pn\ was enabled by
+identifying its connection to the \cn\ using the Gaia data.  The
 Supplementary Table~\ref{tab:v05} enables similar crossmatches to
 already known planets, as well as forthcoming planets from the TESS
-mission (CITE RickerJATIS2014).  An alternative approach also likely
-to bear fruit is to identify kinematic associations around exoplanet
-host stars using the Gaia data, and to then verify these associations
-with ancilliary rotation periods and spectroscopy (CITEP Tofflemire).
+mission \citep{ricker_transiting_2015,guerrero_tess_2021}.  An
+alternative approach also likely to bear fruit is to identify
+kinematic associations around exoplanet host stars using the Gaia
+data, and to then verify these associations with ancilliary rotation
+periods and spectroscopy \citep{tofflemire_tess_2021}.
 

### 2021-07-02 paper/ms.tex
update ttv slope units

@@
     first (second) half of transit inducing a positive (negative) TTV,
-    provided that the orbit is prograde \citep{mazeh_time_2015}.
+    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
+    units along the abscissa are most easily understood by considering
+    that the flux from the star changes by $\sim$60\,ppt per half
+    rotation period ($\sim$1.3\,days).
 		\label{fig:ttvslope}
@@
 The precision of the Kepler data itself may enable more detailed
-analyses without even needing to acquire more data.  While our transit
+analyses without needing to acquire more data.  While our transit
 timing study did not yield the detection of any clearly significant
-planet-induced TTVs, the star's flaring may be suggestive of
+planet-planet induced TTVs, the star's flaring may be suggestive of
 star-planet interactions.  Appendix~\ref{app:flare} highlights an odd flare
@@
 
-Ultimately however, our age measurement for \pn\ was enabled by
-identifying its connection to the \cn\ using the Gaia data.  The
-Supplementary Table~\ref{tab:v05} enables similar crossmatches to
-already known planets, as well as forthcoming planets from the TESS
-mission \citep{ricker_transiting_2015,guerrero_tess_2021}.  An
-alternative approach also likely to bear fruit is to identify
-kinematic associations around exoplanet host stars using the Gaia
-data, and to then verify these associations with ancilliary rotation
-periods and spectroscopy \citep{tofflemire_tess_2021}.
+Ultimately, our age measurement for \pn\ was enabled by identifying
+its connection to the \cn\ using the Gaia data.  The Supplementary
+Table~\ref{tab:v05} enables similar crossmatches to already known
+planets, as well as forthcoming planets from the TESS mission
+\citep{ricker_transiting_2015,guerrero_tess_2021}.  An alternative
+approach would be to identify kinematic associations around exoplanet
+host stars using the Gaia data, and to then verify these associations
+with ancilliary rotation periods and spectroscopy
+\citep{tofflemire_tess_2021}.
 
@@
 transit fitting, to R{.}~Kerr for kindly providing us with the
-\citet{Kerr2021} membership list prior to its publication, and to
-K.~Masuda for advice on gravity darkening.
+\citet{kerr_stars_2021} membership list prior to its publication.
 % is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,

### 2021-07-02 rudolf/plotting.py

@@ -1251,13 +1251,18 @@ def plot_ttv_vs_local_slope(outdir):
     fig, ax = plt.subplots(figsize=(4,3))
 
     ax.set_ylabel("TTV [minutes]")
-    ax.set_xlabel("Local light curve slope [hour$^{-1}$]")
+    ax.set_xlabel("Local light curve slope [ppt$\,$day$^{-1}$]")
+
+    # 1000 will give ppt per day...
+    # for context, the light curve changes by ~50 ppt per half rotation period
+    # (which is ~1.3 days).
+    XMULT = 1e3
 
     sel = np.abs(ttv)<ttv_cut
     slopes,ttv,tc_err = slopes[sel],ttv[sel],tc_err[sel]
 
     ax.errorbar(
-        24*slopes, ttv*24*60, tc_err*24*60,
+        XMULT*slopes, ttv*24*60, tc_err*24*60,
         marker='o', elinewidth=0.5, capsize=4, lw=0, mew=0.5, color='k',
         markersize=3, zorder=5
     )
@@ -1266,30 +1271,28 @@ def plot_ttv_vs_local_slope(outdir):
 
     #ax.plot(24*t0, z1(t0)*24*60, '-', label='linear fit', color='C1', alpha=0.6)
 
-    ax.plot(24*t0, _linear_model(t0, lsfit_slope, lsfit_int)*24*60, '-',
+    ax.plot(XMULT*t0, _linear_model(t0, lsfit_slope, lsfit_int)*24*60, '-',
             label='Best fit', color='C2', alpha=1, lw=1)
 
-    ax.fill_between(24*t0,
+    ax.fill_between(XMULT*t0,
                     _linear_model(t0, lsfit_slope-lsfit_slope_err, lsfit_int)*24*60,
                     _linear_model(t0, lsfit_slope+lsfit_slope_err, lsfit_int)*24*60,
                     alpha=0.6,
                     color='C2', lw=0, label='1$\sigma$',zorder=-1)
 
-    ax.fill_between(24*t0,
+    ax.fill_between(XMULT*t0,
                     _linear_model(t0, lsfit_slope-3*lsfit_slope_err, lsfit_int)*24*60,
                     _linear_model(t0, lsfit_slope+3*lsfit_slope_err, lsfit_int)*24*60,
                     alpha=0.2,
[…]

### 2021-07-03 drivers/get_flare_pair_time_distribution.py
add get_flare_pair_time_distribution

@@ -1,0 +1,66 @@
+import os
+import pandas as pd, numpy as np, matplotlib.pyplot as plt
+from aesthetic.plot import set_style, savefig
+from rudolf.paths import RESULTSDIR
+from rudolf.plotting import _get_detrended_flare_data
+
+flaredir = os.path.join(RESULTSDIR, 'flares')
+
+# read data
+method = 'itergp'
+cachepath = os.path.join(flaredir, f'flare_checker_cache_{method}.pkl')
+c = _get_detrended_flare_data(cachepath, method)
+flpath = os.path.join(flaredir, f'fldict_{method}.csv')
+df = pd.read_csv(flpath)
+
+FL_AMP_CUTOFF = 5e-3
+sel = df.ampl_rec > FL_AMP_CUTOFF
+sdf = df[sel]
+
+# get pairwise distances...
+t_arr = np.array(sdf.tstart)
+dists = np.abs(t_arr - t_arr[:, None])
+
+# remove duplicates
+uniq_dists = np.unique(np.round(dists,3).flatten())
+
+plt.close('all')
+set_style()
+fig, ax = plt.subplots(figsize=(4,3))
+bins = np.logspace(-1, 2, 100)
+ax.hist(uniq_dists, bins=bins, cumulative=False, color='k',
+        fill=False, histtype='step', linewidth=0.5)
+
+P_orb = 7.2028041 # pm 0.0000074 
+P_rot = 2.642 # pm 0.042
+P_syn = (1/P_rot - 1/P_orb)**(-1) # ~=4.172 day
+
+ylim = ax.get_ylim()
+ax.vlines(P_orb, min(ylim), max(ylim), ls='--', lw=0.5, colors='C0',
+          zorder=-1, label='P$_\mathrm{orb}$'+f' ({P_orb:.3f} d)')
+
+#ax.vlines(2*P_orb, min(ylim), max(ylim), ls='--', lw=0.5, colors='C0',
+#          zorder=-1, label=r'2$\times$P$_\mathrm{orb}$')
+ax.vlines(P_syn, min(ylim), max(ylim), ls='--', lw=0.5, colors='C1',
+          zorder=-1, label='P$_\mathrm{syn}$'+f' ({P_syn:.3f} d)')
+ax.vlines(2*P_syn, min(ylim), max(ylim), ls='--', lw=0.5, colors='C1',
+          zorder=-1, label=r'2$\times$P$_\mathrm{syn}$')
[…]

### 2021-07-08 paper/ms.tex
add get_flare_pair_time_distribution
update ms
flare analysis

@@
   selected for observation by the Kepler satellite because it is
-  nearby ($d\approx 333\,{\rm pc}$) and resembles the Sun.  Combining
+  nearby ($d\approx 333\,{\rm pc}$) and it resembles the Sun.  Combining
   data from Gaia with rotation periods from TESS, we show that
@@
   cluster.  Kepler\,1627Ab is therefore the youngest planet with a
-  precise age found so far by the main Kepler mission.
+  precise age found to date by the main Kepler mission.
   Analyzing the photometry, we find evidence that starspot crossings
@@
   variations that suggest the planet is on a prograde orbit.  The 
-  possibility that star-planet interactions could induce
-  flares in the system is also explored.  Future spectroscopic studies
-  of \pn\ may help clarify how the orbit and atmosphere of this
-  mini-Neptune evolve.  More generally, the combination of Gaia and TESS
-  offers an opportunity to significantly expand the census of
-  age-dated planets; we provide a literature concatenation of young
-  stars to support this effort.
+  flare arrival times may show hints of being
+  non-Poissonian, but the statistical evidence is not overwhelming ($<3$-$\sigma$).  In the future,
+  spectroscopic studies of \pn\ may help clarify the evolution of this
+  mini-Neptune's orbit and atmosphere.  More broadly, the
+  combination of Gaia and TESS data offers an opportunity to
+  expand the census of age-dated planets; we provide a literature
+  concatenation of young stars to support this effort.
 \end{abstract}
@@
 \end{figure*}
-
-% Could the planet be on a high obliquity orbit?
-% * how do the spectroscopic vsini and the measured v_rot=2πRstar/Prot compare?
-%   * Rstar=0.91+-0.05, 2.606 days gives vsini=17.6km/s (+/-0.97km/s)
-% * spectroscopic vsini is 18.8 +/- 4.7km/s from Keck HIRES... "20km/s" from
-%   TRES.
 
@@
 
-The 1-minute cadence Kepler observations span 98 days, and cover 24
-flares exceeding $0.5\%$ in relative flux.  These 24 flares spanned a
-total of 6.5 hours ($\sim$15 minutes per flare).  The coincidence is
-that despite the low flare duty cycle, one orbital period after the
-brightest flare, a second flare followed.  This and a similar
-event are shown in Figure~\ref{fig:flarezoom}.  The timing error is
-good to a $\approx0.2\%$ difference from the orbital period, which
-seems {\it a priori} unlikely.  If we consider flares falling within
-2\% of the planet's orbital period after a previous flare, then 4 of
-the 24 flare events have candidate ``successors''.
-
-A brief note on how we cleaned the light curve, and identified the
-flares.
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1.0\textwidth]{f13.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Statistics of inter-flare arrival times}.  
+    24 flares were recorded with amplitudes exceeding 0.5\% over the
+    97.7 days of short cadence observations.  The histogram of the
+    time intervals between every possible pair of flares is shown in
+    black.  Some plausibly important timescales for star-planet
+    interactions, namely the planetary orbital period and synodic
+    period (the orbital period as seen from the rotating stellar frame) are
+    shown along with their linear combinations.  Monte Carlo
+    draws from a Poisson distribution are shown
+    with the gray bands.  While peaks in the observed distribution do
+    coincide with the locations of these ``special periods'', the
+    statistical evidence for a non-Poissonian process driving the
+    flares does not reach the $5$-$\sigma$ threshold.
+		\label{fig:flarestats}
+	}
+\end{figure*}
+
+The short cadence (1-minute) Kepler observations span 97.7 days.
+The light curve shows a significant number of flares
+(Figure~\ref{fig:flarezoom}).
+To quantify clean the light curve and identify the flares, we
+performed the following procedure.
 For cleaning, we performed the following iterative detrending procedure.
@@
 \end{itemize}
-
-The flares were identified and fitted using ALTAIPONY (CITE CITE).
+The flares were then identified and fitted using ALTAIPONY (CITE CITE).
 The fitted model is that of CITE XXX, which parametrizes the flare
@@
 
-We considered two hypotheses: {\it i)} the flare arrival times are
-distributed as a Poisson process, and {\it ii)} they are better
-explained as a Poisson process with a periodic mixture component.
-
-TODO: explore!  Fit the flare inter-arrival times...
-[even the 2-separated ones?]
+There were $N_{\rm f}=24$ flares exceeding $0.5\%$ in relative flux during the
+short-cadence observations.  These 24 flares spanned a total of 6.5
+hours ($\sim$15 minutes per flare).  For better or worse, we noticed a
+coincidence in the flare arrival times.  The coincidence is that
+despite the low flare duty cycle, one orbital period after the
+brightest flare, a second flare followed.  This and a similar event
+are shown in Figure~\ref{fig:flarezoom}.  The timing error is good to
+a $\approx0.2\%$ difference from the orbital period, which seems {\it
+a priori} unlikely.  If we consider flares falling within 2\% of the
+planet's orbital period after a previous flare, then 4 of the 24 flare
+events have candidate ``successors''.
+
+As with any coincidence, when one does not have a firm prediction, it
+is difficult to assess the statistical significance of a surprise.
+Since our surprise was specifically at the inter-arrival time of
+certain flares coinciding with special time intervals, we performed
+the following analysis.  First, we considered all unordered pairs of
+flares.  For $N$ flares there are ${n \choose 2}$ such pairs (for our
+case, 276 pairs).  We then compared the distribution of the pair
+separations against that from a Poisson distribution.  Specifically,
+we drew $N_{\rm f}=24$ samples from a Poisson distribution with
+$\lambda = \Delta t / N_{\rm f}$, for $\Delta t=97.7\,{\rm days}$ the
+duration of the observations, and repeated the draw $10^3$ times with
+unique random seeds.
+
+Figure~\ref{fig:flarestats} shows the results.  The vertical lines in
+the figure show the planetary orbital period, the synodic period
+$P_{\rm syn} = (P_{\rm rot}^{-1} - P_{\rm orb}^{-1})^{-1}$, and linear
+combinations thereof.  Note that the tidal period (half the synodic
+period) is not shown.  The bins are logarithmically spaced to give 100
+bins between the minimum and maximum ordinate values.  The gray bands
+express the range of values observed from the Poissonian draws.  While
+it does seem like a rather odd coincidence for peaks in the observed
+flare arrival time distribution to coincide with the locations of
+these ``special intervals'', the statistical evidence for a
+non-Poissonian process driving the flares does not seem particularly
+strong.  More quantitatively, the peaks observed at the orbital and
+synodic periods are within the $\pm 2$-$\sigma$ range of a
+Poissonian process, and those at $P_{\rm orb}+P_{\rm syn}$ and $P_{\rm
+orb}+2P_{\rm syn}$ are only slightly above this range. 
+With that said,
+future analyses of these data by investigators more versed in the
+topic than ourselves could very well yield new insights.
 

### 2021-07-08 paper/collect.sh

@@ -14,9 +14,14 @@ cp /Users/luke/Dropbox/proj/rudolf/results/flares/flarezoom_itergp.pdf f7.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/flares/flarephase_itergp_altailabel.pdf f8.pdf
 
+cp /Users/luke/Dropbox/proj/rudolf/results/flares/flare_pair_separation_histogram_poisson_real_and_1000sample_average.pdf f13.pdf
+
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_phasedsubsets_quarterchunk.pdf f9.pdf
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_phasedsubsets_yearchunk.pdf f10.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/ttv_vs_local_slope/ttv_vs_local_slope.pdf f11.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit_Kepler_1627_Q15slc/Kepler_1627_Q15slc_RotGPtransit_posterior_phaseplot.pdf f12.pdf
+
+
+

### 2021-07-08 rudolf/helpers.py

@@ -102,7 +102,7 @@ def get_kep1627_kepler_lightcurve(lctype='longcadence'):
         qual_list.append(qual[sel])
         texp_list.append(texp)
 
-    if lctype == 'longcadence':
+    if lctype == 'longcadence' or 'shortcadence':
         time = np.hstack(timelist)
         flux = np.hstack(f_list)
         flux_err = np.hstack(ferr_list)

### 2021-07-08 paper/ms.tex
update ms

@@
 planet-planet induced TTVs, the star's flaring may be suggestive of
-star-planet interactions.  Appendix~\ref{app:flare} highlights an odd flare
-timing coincidence that may be worthy of further examination by those
-more knowledgeable than us.
+star-planet interactions.  Appendix~\ref{app:flare} highlights an odd
+flare timing coincidence that may be worthy of further examination.
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=\textwidth]{f14.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Radii, orbital periods, and ages of known transiting
+    exoplanets}.  Planets older than
+    $10^9$ years are shown with gray markers.  Planets younger than a
+    gigayear, with $t/\sigma_t > 3$, are shown in color.  
+    \pn\ is shown with a star.
+    The abnormally large radii of the youngest transiting planets
+    could be explained by their primordial atmospheres not yet
+    having evaporated.
+    Selection effects could also be important.
+    Parameters were acquired from the NASA Exoplanet Archive on July
+    2, 2021.
+		\label{fig:rp_period_age}
+	}
+\end{figure*}
+
+In the context of the known transiting planets, \pn\ is among the
+youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
+with precisely measured ages spanning 10 to 100\,Myr include K2-33b,
+DS~Tuc~Ab, HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet
+V1298~Tau system
+\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020}.
+Of these systems, \pn\ inhabits a similar range of planetary size and
+orbital period.  It is the smallest though ($3.6\,R_\oplus$), which
+may hint at selection effects in the sample of young planets
+\citep{zhou_2021_tois}.  Assuming that these planets have masses
+between $\approx$10 and 20$\,M_\oplus$, their large sizes and presumed
+low densities are in accord with the expectation that mini-Neptunes
+start their lives with large primordial atmopsheres that are are lost
+over the first gigayear
+\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_core-powered_2018}.
+The observation from starspot-induced TTVs that the planetary orbit
+could be aligned with the stellar spin differs from a number of other
+close-in
+mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
+Overall however, it is consistent with a quiescent disk-driven
+migration history for the planet, while the in-situ formation scenario
+is disfavored because the necessary disk mass would always drive
+planets to migrate
+\citep{inamdar_formation_2015,ogihara_reassessment_2015}.
 

### 2021-07-12 paper/ms.tex
update ms

@@
 
-\defcitealias{cantatgaudin_gaia_2018}{CG18}
-\defcitealias{kounkel_untangling_2019}{KC19}
-\defcitealias{meingast_2021}{M21}
-
-% Cluster Difference Imaging Photometric Survey. III.
-% Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs
 \title{
@@
   variations that suggest the planet is on a prograde orbit.  The 
-  flare arrival times may show hints of being
-  non-Poissonian, but the statistical evidence is not overwhelming ($<3$-$\sigma$).  In the future,
-  spectroscopic studies of \pn\ may help clarify the evolution of this
-  mini-Neptune's orbit and atmosphere.  More broadly, the
-  combination of Gaia and TESS data offers an opportunity to
-  expand the census of age-dated planets; we provide a literature
-  concatenation of young stars to support this effort.
+  flare arrival time distribution is most likely Poissonian, though
+  there may be hints ($<3$-$\sigma$) of additional flares at the
+  orbital and synodic periods.  In the future, spectroscopic studies
+  of \pn\ may help clarify the evolution of this mini-Neptune's orbit
+  and atmosphere.  More broadly, the combination of Gaia and TESS data
+  offers an opportunity to expand the census of age-dated planets; we
+  provide a literature concatenation of young stars to support this
+  effort.
 \end{abstract}
@@
 atmospheric outflows are key steps toward testing this paradigm, which
-is often invoked to explain the observed radius distribution of both
-mature exoplanets \citep{Fulton_et_al_2017}, as well as younger
-planets \citep{bouma_cluster_2020}.
+is often invoked to explain the observed radius distribution of 
+mature exoplanets \citep{Fulton_et_al_2017}.
 
@@
 knowledge of open cluster and moving group memberships \citep[{\it
-e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021}.
+e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,kerr_stars_2021}.
 As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
@@
 cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
-the cluster, from the Kepler-52 and Kepler-968 systems
+the cluster from the Kepler-52 and Kepler-968 systems
 \citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
 discovery, and the focus of this work, is that Kepler observed a
-portion of the $\delta$\,Lyr cluster, shown in
+portion of the $\delta$\,Lyr cluster \citep{stephenson_possible_1959,eggen_photometric_1968}.
+The overlap between the Kepler field and the cluster is shown in
 Figure~\ref{fig:skychart}.
@@
 statistical validation is reassuring, ``validated'' planets have been
-demonstrated to be false positives \citep{shporer_three_2017}.  Most
+shown to be false positives \citep{shporer_three_2017}.  Most
 failures of the statistical validation approach arise when attempting
@@
 companion in archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does
-not change the conclusion: the companion contributes 1\% of the flux
-observed by Kepler, which combined with the observed transit shape
-cannot be invoked as an explanation for the observed data
-(Appendix~\ref{app:companionstar}).
+not change the conclusion: the companion star contributes 1\% of the
+flux observed by Kepler.  Combined with the observed transit depth and
+shape (Appendix~\ref{app:companionstar}), a planet transiting the
+primary star is the only plausible interpretation of the
+data.
 
 {\bf todo: LGB to check numbers on full-dataset transit ecc and b} 
-Two additional new lines of evidence support the planetary
+Two additional new lines of evidence confirm the planetary
 interpretation.  First, the $\approx$100 days of 1-minute
-short-cadence data reduces the cadence-induced smearing enables a more
-precise measurement of the transit impact parameter
+cadenced Kepler observations reduce the cadence-induced smearing,
+enabling a more precise measurement of the transit impact parameter
 \citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
@@
 {\bf todo: Kento to check numbers on per-transit analysis} 
-The second line of evidence came from our transit timing analysis. We
+The second line of evidence comes from our transit timing analysis. We
 isolated each of the 144 observed transits to within $\pm4.5$\,hr of

### 2021-07-17 paper/ms.tex
update ms

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[12pt,modern,tighten]{aastex63}
+\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
@@
 \accepted{---}
-\submitjournal{Nature}
+\submitjournal{AAS Journals}
 \shorttitle{Kepler\,1627}
@@
 
+% \begin{abstract}
+%   Kepler\,1627A is a G8V star previously found to host a
+%   $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
+%   selected for observation by the Kepler satellite because it is
+%   nearby ($d\approx 333\,{\rm pc}$) and it resembles the Sun.  Combining
+%   data from Gaia with rotation periods from TESS, we show that
+%   Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
+%   cluster.  Kepler\,1627Ab is therefore the youngest planet with a
+%   precise age found to date by the main Kepler mission.
+%   Analyzing the photometry, we find evidence that starspot crossings
+%   distort the average transit shape, and induce transit timing
+%   variations that suggest the planet is on a prograde orbit.  The 
+%   flare arrival time distribution is most likely Poissonian, though
+%   there may be hints ($<3$-$\sigma$) of additional flares at the
+%   orbital and synodic periods.  In the future, spectroscopic studies
+%   of \pn\ may help clarify the evolution of this mini-Neptune's orbit
+%   and atmosphere.  More broadly, the combination of Gaia and TESS data
+%   offers an opportunity to expand the census of age-dated planets; we
+%   provide a literature concatenation of young stars to support this
+%   effort.
+% \end{abstract}
+
+% 195 words (250 max)
 \begin{abstract}
@@
   $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
-  selected for observation by the Kepler satellite because it is
-  nearby ($d\approx 333\,{\rm pc}$) and it resembles the Sun.  Combining
-  data from Gaia with rotation periods from TESS, we show that
-  Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
-  cluster.  Kepler\,1627Ab is therefore the youngest planet with a
-  precise age found to date by the main Kepler mission.
-  Analyzing the photometry, we find evidence that starspot crossings
-  distort the average transit shape, and induce transit timing
-  variations that suggest the planet is on a prograde orbit.  The 
-  flare arrival time distribution is most likely Poissonian, though
-  there may be hints ($<3$-$\sigma$) of additional flares at the
-  orbital and synodic periods.  In the future, spectroscopic studies
-  of \pn\ may help clarify the evolution of this mini-Neptune's orbit
-  and atmosphere.  More broadly, the combination of Gaia and TESS data
-  offers an opportunity to expand the census of age-dated planets; we
-  provide a literature concatenation of young stars to support this
-  effort.
+  observed by the Kepler satellite because it is nearby ($d\approx
+  333\,{\rm pc}$) and it resembles the Sun.  Combining Gaia kinematics
+  with rotation periods from TESS, we show that Kepler\,1627 is a
+  member of the $35\pm XX$\,Myr old $\delta$~Lyr cluster.
+  Kepler\,1627Ab is therefore the youngest planet with a precise age
+  yet found by the main Kepler mission.  The precision of the Kepler
+  photometry lets us detect two anomalies.  First, unresolved
+  starspot-crossing events distort the average transit shape, and
+  induce transit timing variations that imply the planet is likely on
+  a prograde orbit.  Second, although the flare arrival time
+  distribution is probably Poissonian, there may be an excess of
+  flares separated by the orbital and synodic periods, and linear
+  combinations thereof.  The $\delta$~Lyr cluster is one of many
+  stellar groups whose properties have been clarified using the Gaia
+  data; many of the other new clusters also overlap with confirmed and
+  candidate exoplanets.  Further studies that combine Gaia kinematics
+  and TESS rotation periods are therefore well-poised to expand the
+  census of age-dated planets, and our understanding of planetary
+  evolution.
 \end{abstract}
 
+% While thousands of exoplanets have been discovered orbiting nearby
+% stars, the vast majority of them are several billion years old.  This
+% makes it difficult to test theories for the origins of the different
+% families of planets, since many evolutionary processes are expected to
+% operate on timescales of less than 100 million years.  The K2 and TESS
+% missions have enabled advances on this front [1,2,3,4,5,6].  The
+% Kepler mission however has not yielded any planets with precise ages
+% below one gigayear [7,8].  Here, we combine kinematic data from Gaia
+% with stellar rotation periods from TESS to show that Kepler 1627 is a
+% member of the 35 ± 10 million year old δ Lyr cluster.  A
+% $3.6\,R_\oplus$ mini-Neptune transits the G8V primary on a 7.2 day
+% orbit, and it is therefore the youngest planet yet found by the main
+% Kepler mission.  The precision of the Kepler photometry lets us detect
+% two anomalies.  First, unresolved starspot-crossing events distort the
+% average transit shape, and induce transit timing variations that imply
+% the planet is likely on a prograde orbit [9].  Second, although the
+% flare arrival time distribution is most likely Poissonian, there may
+% be hints of an excess of flares at the orbital and synodic periods.
+% The δ Lyr cluster is one of many stellar groups whose nature has been
+% clarified using the Gaia data; many other new clusters are expected to
+% overlap datasets of confirmed and candidate planets.  Further studies
+% that combine Gaia kinematics and TESS rotation periods are therefore
+% well-poised to expand the census of age-dated planets, and our
+% understanding of planetary evolution.
+
 \keywords{
-  planetary evolution (XXXX),
+  exoplanet evolution (491),
   stellar associations (1582),
@@
 
-\section{Manuscript}
+\section{Introduction}
 
@@
 stars, the vast majority of them are several billion years old.  This
-makes it difficult to test theories for the origins of the different
+makes it difficult to test origin theories for the different
 families of planets, since many evolutionary processes are expected to
@@
 
-For instance, the ``mini-Neptune'' planets, thought to be made of
-molten rocky cores and extended atmospheric envelopes of hydrogen and
+%TODO: better citations on radius evolution?
+For instance, the ``mini-Neptunes'', thought to be made of
+molten rocky cores \citep{kite_atmosphere_2020} and extended atmospheric envelopes of hydrogen and
 helium, are expected to shrink in size by factors of several over
 their first $10^8$ years
-\citep{owen_constraining_2020,kite_atmosphere_2020}.  Specifically,
+\citep{owen_constraining_2020}.  Specifically,
 they start with sizes of 4--12\,$R_\oplus$ at the time of disk
@@
 outflow, which can eventually deplete or entirely strip the envelope
-from the rocky core \citep{Owen_Wu_2013,gupta_sculpting_2019}.
+from the rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
 Discovering young planets and measuring their compositions and
 atmospheric outflows are key steps toward testing this paradigm, which
-is often invoked to explain the observed radius distribution of 
-mature exoplanets \citep{Fulton_et_al_2017}.
-
-% TODO: decide whether a paragraph about young planet obliquities is
-% needed.
-
-% The orbits of the planets can also change early on, and can become
-% significantly misaligned from the equators of their host stars.
-% Measurements of the sky-projected angle between the planet orbit and
-% the stellar spin (the stellar obliquity) have shown that high
-% obliquities characterize hot field stars hosting both hot Jupiters and
-% smaller planets \citep{winn10,louden21}.  Conversely, the obliquities
-% of planets around cool stars tend to be low, but this could be due to
-% tidal realignment \citep[{\it e.g.},][]{anderson21}.  Obliquity
-% measurements for young planets are therefore important, because they
-% provide an unobstructed view of the primordial obliquity distribution.
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.99\textwidth]{f2.pdf}
+is often invoked to explain the observed radius distribution of mature
+exoplanets \citep{Fulton_et_al_2017}.
+
+The K2 and TESS missions have now enabled the detection of about ten
+close-in planets younger than 100 million years, all smaller than
+Jupiter
+\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020}.
+The Kepler mission however has not yielded any planets with precise
+ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
+at the time of the main Kepler mission (2009--2013), only four open
+clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
+NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
+\citep{meibom_kepler_2011}.  Since that time, analyses of the
+kinematic, photometric, and astrometric Gaia data have expanded our
+knowledge of open cluster and moving group memberships \citep[{\it
+e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,kerr_stars_2021}.
+As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
+\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
+from the literature, which yielded a list of candidate young and
+age-dated stars (see Appendix~\ref{app:targetlist}).
+
+Comparing our young star list against the Kepler field yielded two
+discoveries.  The first, to be discussed in an upcoming analysis by J.
+Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
+cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
+the cluster from the Kepler-52 and Kepler-968 systems
+\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
+discovery, and the focus of this work, is that Kepler observed a
+portion of the $\delta$\,Lyr cluster \citep[Stephenson-1;][]{stephenson_possible_1959,eggen_photometric_1968}.
+The overlap between the Kepler field and the cluster is shown in
+Figure~\ref{fig:skychart}.
+
+A previous clustering analysis of the Gaia data
+reported that
+Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was a member of the
+\cn\ \citep{kounkel_untangling_2019}.
+Given the statistical validation of the close-in mini-Neptune
+\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018}, we therefore began by examining the age of
+the star and its putative host cluster (Section~\ref{sec:cluster}). 
+We find that the $\delta$ Lyr cluster is $\approx35$\,Myr old, and
+that \sn\ is indeed a member of the cluster.
+Focusing on the planet itself (Section~\ref{sec:planet}), we confirm
+that despite a previously unreported binary companion, it is indeed a
+mini-Neptune orbiting the roughly Sun-like primary star.
+The presence of a correlation between transit timing variations and
+the local light curve slope renders any other explanation implausible.
+The broader implications for our ability to age-date a statistical
+sample of planets are highlighted in closing (Section~\ref{sec:conc}).
+
+\section{The Cluster}
+\label{sec:cluster}
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.48\textwidth]{f2.pdf}
 	\end{center}
@@
     by TESS for two lunar months to date.
+    {\bf todo: move colorbar...}
 		\label{fig:skychart}
 	}
-\end{figure*}
+\end{figure}
 
@@
 \end{figure*}
+
+{\bf Jason Curtis todo}
+We collected the previously measured
+{\bf XXX} Kepler rotation periods (CITEP MCQUILLAN2014), and measured
+{\bf YYY} new stellar rotation periods for cluster members observed by
+TESS.  We also performed an isochronal analysis.  The results are
+shown in Figure~\ref{fig:age}.  From the ensemble perspective, which
+provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
+of the star is $XX \pm YY$\,Myr.  Individual measurements of the
+star's rotation period and lithium abundance corroborate this
+conclusion.
+
+\section{The Planet}
+\label{sec:planet}
 
@@
 
-At the time of the main Kepler mission (2009--2013), only four open
-clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
-NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
-\citep{meibom_kepler_2011}.  Since that time, analyses of the
-kinematic, photometric, and astrometric Gaia data have expanded our
-knowledge of open cluster and moving group memberships \citep[{\it
-e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,kerr_stars_2021}.
-As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
-\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
-from the literature, which yielded a list of candidate young and
-age-dated stars (see Appendix~\ref{app:targetlist}).
-
-Comparing our young star list against the Kepler field yielded two
-discoveries.  The first, to be discussed in an upcoming analysis by J.
-Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
-cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
-the cluster from the Kepler-52 and Kepler-968 systems
-\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
-discovery, and the focus of this work, is that Kepler observed a
-portion of the $\delta$\,Lyr cluster \citep{stephenson_possible_1959,eggen_photometric_1968}.
-The overlap between the Kepler field and the cluster is shown in
-Figure~\ref{fig:skychart}.
-
-{\bf Jason Curtis todo}
-The \cn\ contains
-Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470), which 
-was previously reported to host a close-in mini-Neptune candidate planet
-\citep{2012ApJS..199...24T,thompson_planetary_2018}
-Given the planet's previous statistical validation by
-\citet{morton_false_2016}, we decided to examine the age of the star
-and its cluster more closely.  We collected the previously measured
-{\bf XXX} Kepler rotation periods (CITEP MCQUILLAN2014), and measured
-{\bf YYY} new stellar rotation periods for cluster members observed by
-TESS.  We also performed an isochronal analysis.  The results are
-shown in Figure~\ref{fig:age}.  From the ensemble perspective, which
-provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
-of the star is $XX \pm YY$\,Myr.  Individual measurements of the
-star's rotation period and lithium abundance corroborate this
-conclusion.
+\begin{figure}[tp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.48\textwidth]{f11.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Evidence for a prograde orbit from transit timing variations.}
+    The time of each transit was measured, along with the local slope
+    of the light curve.  The two quantities appear anti-correlated,
+    which is most easily explained by starspot crossings during the
+    first (second) half of transit inducing a positive (negative) TTV,
+    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
+    units along the abscissa are most easily understood by considering
+    that the flux from the star changes by $\sim$60\,ppt per half
+    rotation period ($\sim$1.3\,days).
+		\label{fig:ttvslope}
+	}
+\end{figure}
 
@@
 
+\section{Discussion \& Conclusions}
+\label{sec:conc}
+
 \begin{figure*}[tp]
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f11.pdf}
+		\includegraphics[width=\textwidth]{f14.pdf}
 	\end{center}
@@
 	\caption{
-		{\bf Evidence for a prograde orbit from transit timing variations.}
-    The time of each transit was measured, along with the local slope
-    of the light curve.  The two quantities appear anti-correlated,
-    which is most easily explained by starspot crossings during the
-    first (second) half of transit inducing a positive (negative) TTV,
-    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
-    units along the abscissa are most easily understood by considering
-    that the flux from the star changes by $\sim$60\,ppt per half
-    rotation period ($\sim$1.3\,days).
-		\label{fig:ttvslope}
+    {\bf Radii, orbital periods, and ages of known transiting
+    exoplanets}.  Planets older than
+    $10^9$ years are shown with gray markers.  Planets younger than a
+    gigayear, with $t/\sigma_t > 3$, are shown in color.  
+    \pn\ is shown with a star.
+    The abnormally large radii of the youngest transiting planets
+    could be explained by their primordial atmospheres not yet
+    having evaporated.
+    Selection effects could also be important.
+    Parameters were acquired from the NASA Exoplanet Archive on July
+    2, 2021.
+		\label{fig:rp_period_age}
 	}
@@
 
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=\textwidth]{f14.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Radii, orbital periods, and ages of known transiting
-    exoplanets}.  Planets older than
-    $10^9$ years are shown with gray markers.  Planets younger than a
-    gigayear, with $t/\sigma_t > 3$, are shown in color.  
-    \pn\ is shown with a star.
-    The abnormally large radii of the youngest transiting planets
-    could be explained by their primordial atmospheres not yet
-    having evaporated.
-    Selection effects could also be important.
-    Parameters were acquired from the NASA Exoplanet Archive on July
-    2, 2021.
-		\label{fig:rp_period_age}
-	}
-\end{figure*}
-
 In the context of the known transiting planets, \pn\ is among the
@@
 over the first gigayear
-\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_core-powered_2018}.
+\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.
 The observation from starspot-induced TTVs that the planetary orbit
@@
 \citep{inamdar_formation_2015,ogihara_reassessment_2015}.
+
+% TODO: decide whether a paragraph about young planet obliquities is
+% needed.
+
+% The orbits of the planets can also change early on, and can become
+% significantly misaligned from the equators of their host stars.
+% Measurements of the sky-projected angle between the planet orbit and
+% the stellar spin (the stellar obliquity) have shown that high
+% obliquities characterize hot field stars hosting both hot Jupiters and
+% smaller planets \citep{winn10,louden21}.  Conversely, the obliquities
+% of planets around cool stars tend to be low, but this could be due to
+% tidal realignment \citep[{\it e.g.},][]{anderson21}.  Obliquity
+% measurements for young planets are therefore important, because they
+% provide an unobstructed view of the primordial obliquity distribution.
 
@@
 % 	Euler1.2m (EulerCam),
+	  Kepler \citep{borucki_kepler_2010},
 % 	Magellan:Baade (MagIC),

### 2021-07-27 paper/ms.tex
update ms

@@
 
-% \begin{abstract}
-%   Kepler\,1627A is a G8V star previously found to host a
-%   $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
-%   selected for observation by the Kepler satellite because it is
-%   nearby ($d\approx 333\,{\rm pc}$) and it resembles the Sun.  Combining
-%   data from Gaia with rotation periods from TESS, we show that
-%   Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
-%   cluster.  Kepler\,1627Ab is therefore the youngest planet with a
-%   precise age found to date by the main Kepler mission.
-%   Analyzing the photometry, we find evidence that starspot crossings
-%   distort the average transit shape, and induce transit timing
-%   variations that suggest the planet is on a prograde orbit.  The 
-%   flare arrival time distribution is most likely Poissonian, though
-%   there may be hints ($<3$-$\sigma$) of additional flares at the
-%   orbital and synodic periods.  In the future, spectroscopic studies
-%   of \pn\ may help clarify the evolution of this mini-Neptune's orbit
-%   and atmosphere.  More broadly, the combination of Gaia and TESS data
-%   offers an opportunity to expand the census of age-dated planets; we
-%   provide a literature concatenation of young stars to support this
-%   effort.
-% \end{abstract}
-
 % 195 words (250 max)
@@
   with rotation periods from TESS, we show that Kepler\,1627 is a
-  member of the $35\pm XX$\,Myr old $\delta$~Lyr cluster.
+  member of the {\bf $35\pm XX$\,Myr old} $\delta$~Lyr cluster.
   Kepler\,1627Ab is therefore the youngest planet with a precise age
-  yet found by the main Kepler mission.  The precision of the Kepler
-  photometry lets us detect two anomalies.  First, unresolved
+  yet found by the main Kepler mission.  The Kepler
+  photometry seems to show two anomalies.  First, unresolved
   starspot-crossing events distort the average transit shape, and
@@
   candidate exoplanets.  Further studies that combine Gaia kinematics
-  and TESS rotation periods are therefore well-poised to expand the
-  census of age-dated planets, and our understanding of planetary
-  evolution.
+  and TESS rotation periods are well-poised to expand the census of
+  age-dated planets, and our understanding of planetary evolution.
 \end{abstract}
@@
 from the rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
-Discovering young planets and measuring their compositions and
+Discovering young planets, measuring their masses, and detecting their
 atmospheric outflows are key steps toward testing this paradigm, which
@@
 discovery, and the focus of this work, is that Kepler observed a
-portion of the $\delta$\,Lyr cluster \citep[Stephenson-1;][]{stephenson_possible_1959,eggen_photometric_1968}.
-The overlap between the Kepler field and the cluster is shown in
-Figure~\ref{fig:skychart}.
-
-A previous clustering analysis of the Gaia data
-reported that
-Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was a member of the
-\cn\ \citep{kounkel_untangling_2019}.
-Given the statistical validation of the close-in mini-Neptune
-\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018}, we therefore began by examining the age of
-the star and its putative host cluster (Section~\ref{sec:cluster}). 
-We find that the $\delta$ Lyr cluster is $\approx35$\,Myr old, and
-that \sn\ is indeed a member of the cluster.
-Focusing on the planet itself (Section~\ref{sec:planet}), we confirm
-that despite a previously unreported binary companion, it is indeed a
-mini-Neptune orbiting the roughly Sun-like primary star.
-The presence of a correlation between transit timing variations and
-the local light curve slope renders any other explanation implausible.
-The broader implications for our ability to age-date a statistical
-sample of planets are highlighted in closing (Section~\ref{sec:conc}).
+portion of the $\delta$\,Lyr cluster
+(Stephenson-1; Theia~73; \citealt{stephenson_possible_1959}).
+Figure~\ref{fig:skychart} shows the overlap between the Kepler field
+and the cluster.
+
+A previous HDBScan-based clustering analysis of the Gaia data found
+that Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was a member
+of the \cn\ \citep{kounkel_untangling_2019}.  Given the statistical
+validation of close-in mini-Neptune
+\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
+we opted to examine the age of the star and its putative
+host cluster (Section~\ref{sec:cluster}).  We find that the $\delta$
+Lyr cluster is {\bf $\approx35$\,Myr old}, and that \sn\ is indeed a member
+of the cluster.  Focusing on the planet itself
+(Section~\ref{sec:planet}), we confirm that despite a previously
+unreported binary companion, it is indeed a mini-Neptune orbiting the
+roughly Solar-type primary star.  The presence of a correlation
+between transit timing variations and the local light curve slope
+renders any other explanation implausible.  We conclude by
+highlighting some broader implications for our ability to age-date a
+statistical sample of planets (Section~\ref{sec:conc}).
 
@@
     by TESS for two lunar months to date.
-    {\bf todo: move colorbar...}
 		\label{fig:skychart}
@@
 		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f15a.pdf}
 			\includegraphics[width=0.49\textwidth]{TEMP_Prot_vs_BpmRp.png}
-			\includegraphics[width=0.49\textwidth]{TEMP_hrd.png}
 		}
@@
 	\caption{
-    {\bf The \cn\ is $35\pm15$\,Myr old.} {\it Left:} TESS and Kepler
-    rotations periods and dereddened Gaia colors, with the Pleiades
+    {\bf The \cn\ is $35\pm15$\,Myr old.}  {\it Left:} HR diagram of 
+    the \cn, IC\,2602 ($\approx35$\,Myr), the Pleiades ($\approx 115$\,Myr), and the field.
+    The \cn\ and IC\,2602 are approximately the same isochronal age.
+    {\it Right:} TESS and Kepler rotations periods and dereddened Gaia
+    colors, with the Pleiades
     \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
@@
     braking.  The \cn\ has a 40 Myr gyrochronal age, as calibrated
-    against the LDB-age of IC\,2602.  {\it Right:} HR diagram of the
-    same comparison. {\bf todo: caption improve. fix figures.}.
+    against the LDB-age of IC\,2602.  {\bf todo: caption improve. fix
+    figures.}.
    \label{fig:age}
@@
 
+% paragraph 1: HR 
+To assess the age of the cluster as well as our star of interest,
+we used a combination of the isochronal, gyrochronal, and
+lithium-based age-dating.
+
+\subsection{HR Diagram}
+We meaured the isochrone age using an empirical approach, and also
+compared against theoretical models.
+The left panel of Figure~\ref{fig:age} shows the HR diagram of 
+stars in the \cn, the Pleiades, IC\,2602, and the field.
+The stars shown in the \cn\ are selected based on their positions and
+kinematics, as discussed in Appendix~\ref{app:kinematicselection}.
+Those from the Pleiades and IC\,2602 are adopted from
+\citet{cantatgaudin_gaia_2018}, and the field stars are from the Gaia
+EDR3 Catalog of Nearby Stars \citep{gaia_gcns_2021}.
+% FIXME FIXME
+We corrected for reddening following {\bf procedure X,Y,Z}.
+
+The immediate impression from the left panel of Figure~\ref{fig:age}
+is that the \cn\ and IC\,2602 share approximately the same isochronal
+age, since they overlap.  In our exploration, we also compared against
+$\mu$-Tau ($\approx 60$\,Myr; \citealt{gagne_mutau_2020}) and subsets
+of the Sco OB2 association ($\approx10$-$15$\,Myr;
+\citealt{damiani_stellar_2019}), and confirmed that the
+pre-main-sequence M dwarfs were intermediate between these two
+clusters.
+
+% FIXME FIXME
+We quantified this impression following the empirical fitting procedure
+discussed by \citet{gagne_mutau_2020}.
+
+\subsection{Rotation Periods}
 {\bf Jason Curtis todo}
@@
 \section{Kinematic Selection of $\delta$\,Lyr Cluster Members}
+\label{app:kinematicselection}
 
@@
     {\bf Galactic position and tangential velocities of the
-			$\delta$\,Lyr cluster (also known as Theia 73 and Stephenson\,1).}  Points are
-		candidate cluster members with $\varpi/\sigma_\varpi > 20$, reported
-		to be in the group by \citet{kounkel_untangling_2019}.  We focus
-		on stars in a small region (black points) in the kinematic
-		vicinity of Kepler\,1627 (yellow star).  The other candidate cluster
-		members (gray points) may or may not share the ages of the selected
-		kinematic group.  The location of the Sun is ($\odot$) is shown.
+    $\delta$\,Lyr cluster.}
+    Points are candidate cluster members with $\varpi/\sigma_\varpi >
+    20$, reported to be in the group by
+    \citet{kounkel_untangling_2019}.  We focus on stars in a small
+    region (black points) in the kinematic vicinity of Kepler\,1627
+    (yellow star).  The other candidate cluster members (gray points)
+    may or may not share the ages of the selected kinematic group.
+    The location of the Sun is ($\odot$) is shown.
 		\label{fig:XYZvtang}
@@
 
-Figure~\ref{fig:XYZvtang} shows members of the \cn\
-reported
-by \citet{kounkel_untangling_2019} to be in the group.
-Galactic positions are determined and plotted only for stars with parallax
-signal-to-noise exceeding 20.
-The location of the Sun is shown on the plots.
-The non-uniform ``clumps'' might be
-an artifact of the data processing steps performed by \citet{kounkel_untangling_2019}.
-We therefore only consider stars in the immediate kinematic group
-around \sn. 
-The tangential velocities relative to \sn\ are shown in the bottom right panel.
-These are computed by assuming that every star has the same three-dimensional
-spatial velocity as \sn, where we assume a systemic radial velocity of
-$-16.7 \pm 0.2$\,\kms\ based on the reconnaissance spectra obtained by
-A.~Howard on HIRES and D.~Latham on TRES.
-The relevant projection effects are then taken into account,
-as discussed by {\it e.g.}, \citet{Meingast2021} and {\bf L.~Bouma et al (2021, submitted)}.
+Figure~\ref{fig:XYZvtang} shows stars reported by
+\citet{kounkel_untangling_2019} to be in the group ``Theia 73'', which
+was cross-matched by \citealt{kounkel_untangling_2019} as being
+``Stephenson 1''.  For our Figure, galactic positions were calculated
+and plotted only for stars with parallax signal-to-noise exceeding 20.
+The location of the Sun is shown on the plots.  The smattering of
+reported cluster members (gray and black points) has a significantly
+different structure that the cluster initially identified by
+\citet{stephenson_possible_1959} and corroborated by
+\citet{eggen_photometric_1968}.  While the non-uniform ``clumps''
+might be part of a real structure, they could also be an artifact of
+the data processing steps performed by
+\citet{kounkel_untangling_2019}.  We therefore opted to only consider
+stars in the immediate kinematic and spatial vicinity of \sn.  The
+tangential velocities relative to \sn\ are shown in the bottom right
+panel.  These are computed by assuming that every star has the same
+three-dimensional spatial velocity as \sn, where we assume a systemic
+radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the reconnaissance
+spectra obtained by A.~Howard on HIRES and D.~Latham on TRES.  The
+relevant projection effects are then taken into account, as discussed
+by {\it e.g.}, \citet{Meingast2021} and \citet{bouma_2021_ngc2516}.
+We performed the actual selection by then manually drawing lassos with
+the interactive \texttt{glue} visualization tool
+\citep{beaumont_2014_13866} in the four projections shown in
+Figure~\ref{fig:XYZvtang}.  While bonafide members likely exist
+outside of our selection region (and our selection also includes some
+field star interlopers), our aim is to verify the existence of the
+cluster in the vicinity of Kepler\,1627, and to measure its age.  The
+procedure we have adopted enables both tasks.
 

### 2021-07-27 paper/collect.sh

@@ -25,3 +25,5 @@ cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit_Kepler_1627_Q15slc/K
 
 cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20210702_showss_colorbyage_showkep1627.pdf f14.pdf
 
+
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades.pdf f15a.pdf

### 2021-07-30 paper/ms.tex
update extinction correction to Gaia2018

@@
     points are kinematically selected members of the $\delta$\,Lyr
-    cluster.  Both Kepler (gray squares) and TESS (colored points)
+    cluster.  Both Kepler (gray squares) and TESS 
     observed portions of the cluster.  Naked-eye stars ($m_{\rm
@@
     by TESS for two lunar months to date.
+    The colors of the other cluster members indicate the number of
+    TESS sectors during which they have been observed.
 		\label{fig:skychart}
@@
 We meaured the isochrone age using an empirical approach, and also
-compared against theoretical models.
-The left panel of Figure~\ref{fig:age} shows the HR diagram of 
-stars in the \cn, the Pleiades, IC\,2602, and the field.
-The stars shown in the \cn\ are selected based on their positions and
-kinematics, as discussed in Appendix~\ref{app:kinematicselection}.
-Those from the Pleiades and IC\,2602 are adopted from
-\citet{cantatgaudin_gaia_2018}, and the field stars are from the Gaia
-EDR3 Catalog of Nearby Stars \citep{gaia_gcns_2021}.
-% FIXME FIXME
-We corrected for reddening following {\bf procedure X,Y,Z}.
+compared against theoretical models.  The left panel of
+Figure~\ref{fig:age} shows the HR diagram of stars in the \cn, the
+Pleiades, IC\,2602, and the field.  The stars shown in the \cn\ are
+selected based on their positions and kinematics, as discussed in
+Appendix~\ref{app:kinematicselection}.  Those from the Pleiades and
+IC\,2602 are adopted from \citet{cantatgaudin_gaia_2018}, and the
+field stars are from the Gaia EDR3 Catalog of Nearby Stars
+\citep{gaia_gcns_2021}.  We cleaned these following identical data
+filtering criteria to those discussed in Appendix~B of
+\citet{GaiaCollaboration2018}, except that we loosened the parallax
+precision requirement to $\texttt{parallax\_over\_error}>5$.  These
+filters were designed to include genuine binaries while omitting
+instrumental artifacts.  We then corrected for extinction by querying
+the 3-D maps of \citet{capitanio_threedimensional_2017} and
+\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}},
+and by then using the extinction coefficients $k_X\equiv A_X/A_0$
+computed by \citet{GaiaCollaboration2018}, while assuming that $A_0 =
+3.1 E(B-V)$.  For IC\,2602, the Pleaides, and the \cn, this procedure
+yielded a respective mean and standard deviation for the reddening of
+$E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$.  These
+values agree reasonably well with previously reported values from the
+literature \citep[{\it
+e.g.},][]{GaiaCollaboration2018,kounkel_untangling_2019,bossini_age_2019},
+and have the benefit of being sensitive to effects such as
+differential reddening that may be present in the dust maps.
+
+% Gaia+2018 E(B-V):
+% Pleiades 0.045, IC2602 0.031
+%
+% Randich+2018 E(B-V):
+% Pleiades ??, IC2602 0.032 - 0.092
+%
+% KC2019: AV->E(B-V)
+% Pleiades 0.220 AV -> 0.071
+% IC2602 0.243297 -> 0.078
+% Stephenson1 0.184 -> 0.059
+%
+% Bossini+2019 AV->E(B-V)
+% PleiadesMelotte22 0.140 -> 0.045
+% IC2602 0.096 -> 0.031
+% Steph1: 0.157 -> 0.051
+
+following {\bf procedure X,Y,Z}.
 

### 2021-07-30 rudolf/extinction.py

@@ -16,6 +16,68 @@ from scipy.interpolate import interp1d
 
 from rudolf.paths import DATADIR
 
+def append_corrected_gaia_phot_Gaia2018(df):
+    """
+    Using the coefficients calculated by GaiaCollaboration+2018 Table
+    1, and the STILISM reddening values, calculate corrected Gaia
+    photometric magnitudes.  Assumes you have acquired STILISM
+    reddening estimates per retrieve_stilism_reddening below.
+
+    Args:
+        df (DataFrame): contains Gaia photometric magnitudes, and STILISM
+        reddening columns.
+
+    Returns:
+        Same DataFrame, with 'phot_g_mean_mag_corr', 'phot_rp_mean_mag_corr',
+        'phot_bp_mean_mag_corr' columns.
+    """
+    #FIXME
+    corr_path = os.path.join(DATADIR, 'extinction',
+                             'GaiaCollaboration_2018_table1.csv')
+    cdf = pd.read_csv(corr_path, sep=',')
+
+    bandpasses = ['G','BP','RP']
+
+    E_BmV = df['reddening[mag][stilism]']
+    A_0 = 3.1 * E_BmV
+
+    for bp in bandpasses:
+
+        c1 = float(cdf.loc[cdf.bandpass==f'k{bp}', 'c1'])
+        c2 = float(cdf.loc[cdf.bandpass==f'k{bp}', 'c2'])
+        c3 = float(cdf.loc[cdf.bandpass==f'k{bp}', 'c3'])
+        c4 = float(cdf.loc[cdf.bandpass==f'k{bp}', 'c4'])
+        c5 = float(cdf.loc[cdf.bandpass==f'k{bp}', 'c5'])
+        c6 = float(cdf.loc[cdf.bandpass==f'k{bp}', 'c6'])
+        c7 = float(cdf.loc[cdf.bandpass==f'k{bp}', 'c7'])
+
+        BpmRp = df.phot_bp_mean_mag - df.phot_rp_mean_mag
+
+        # nb Eq 1 of the paper has BpmRp0 ... which presents a bit of
+        # a self-consistency issue
+        k_X = (
+            c1
+            + c2*(BpmRp)
+            + c3*(BpmRp)**2
+            + c4*(BpmRp)**3
+            + c5*A_0
+            + c6*A_0**2
+            + c7*A_0*BpmRp
+        )
+
+        A_X = k_X * A_0
+
+        # each "corrected magnitude" is whatever the observed
+        # magnitude was, MINUS the extinction (because it needs to
[…]

### 2021-07-30 rudolf/helpers.py

@@ -530,15 +530,73 @@ def get_keplerfieldfootprint_dict():
     return kep_d
 
 
-def supplement_gaia_stars_extinctions_corrected_photometry(df):
+def supplement_gaia_stars_extinctions_corrected_photometry(
+    df, extinctionmethod='gaia2018', savpath=None
+):
 
     from cdips.utils.gaiaqueries import parallax_to_distance_highsn
     from rudolf.extinction import (
-        retrieve_stilism_reddening, append_corrected_gaia_phot_Gagne2020
+        retrieve_stilism_reddening,
+        append_corrected_gaia_phot_Gagne2020,
+        append_corrected_gaia_phot_Gaia2018
     )
 
-    df['distance'] = parallax_to_distance_highsn(df['parallax'])
-    df = retrieve_stilism_reddening(df, verbose=False)
-    df = append_corrected_gaia_phot_Gagne2020(df)
+    # cache the slow step
+    if not os.path.exists(savpath):
+        df['distance'] = parallax_to_distance_highsn(df['parallax'])
+        df = retrieve_stilism_reddening(df, verbose=False, outpath=savpath)
+    else:
+        df = pd.read_csv(savpath)
+
+    if extinctionmethod == 'gaia2018':
+        df = append_corrected_gaia_phot_Gaia2018(df)
+    if extinctionmethod == 'gagne2020':
+        df = append_corrected_gaia_phot_Gagne2020(df)
 
     return df
+
+
+def get_cleansel(df):
+    """
+    Given a dataframe of Gaia DR2 columns, apply the "cleaning
+    selection" described on page 3 / Appendix B of GaiaCollab+2018 HR
+    diagram paper.
+    """
+    chisq = nparr(df.astrometric_chi2_al)
+    nuprime = nparr(df.astrometric_n_good_obs_al)
+    Gmag = nparr(df.phot_g_mean_mag)
+
+    sel0 = np.sqrt(chisq/(nuprime-5)) < 1.2*np.maximum(
+        1,
+        np.exp(-0.2*(Gmag-19.5))
+    )
+
+    sel1 = (df.parallax_over_error > 5)
+
+    sel2 = (
+        (df.phot_g_mean_flux_over_error > 50)
+        &
+        (df.phot_rp_mean_flux_over_error > 20)
+        &
+        (df.phot_bp_mean_flux_over_error > 20)
+    )
+
+    sel3 = (
+        (
[…]

### 2021-07-30 paper/ms.tex
rotation periods

@@
 			\includegraphics[width=0.49\textwidth]{f15a.pdf}
-			\includegraphics[width=0.49\textwidth]{TEMP_Prot_vs_BpmRp.png}
+			\includegraphics[width=0.49\textwidth]{f15b.pdf}
+		}
+
+		\vspace{-0.5cm}
+		\subfloat{
+			\includegraphics[width=0.8\textwidth]{f15c.pdf}
 		}
@@
 	\caption{
-    {\bf The \cn\ is $35\pm15$\,Myr old.}  {\it Left:} HR diagram of 
+    {\bf The \cn\ is $35\pm15$\,Myr old.}  {\it Top:} HR diagram of 
     the \cn, IC\,2602 ($\approx35$\,Myr), the Pleiades ($\approx 115$\,Myr), and the field.
+    The right panel provides a closer view of stars on the
+    pre-main-sequence.
     The \cn\ and IC\,2602 are approximately the same isochronal age.
-    {\it Right:} TESS and Kepler rotations periods and dereddened Gaia
+    {\it Bottom:} TESS and Kepler rotations periods and dereddened Gaia
     colors, with the Pleiades
@@
     braking.  The \cn\ has a 40 Myr gyrochronal age, as calibrated
-    against the LDB-age of IC\,2602.  {\bf todo: caption improve. fix
-    figures.}.
+    against the LDB-age of IC\,2602.  {\bf todo: caption improve.}.
    \label{fig:age}
@@
 
-following {\bf procedure X,Y,Z}.
+%FIXME FIXME TODO TODO get the ages measured
 
@@
 \subsection{Rotation Periods}
+%FIXME TODO REFERENCE LOWER PANELS
 {\bf Jason Curtis todo}
@@
 conclusion.
+
+For photometric and astrometric binaries, see the appendix!
 
@@
 
+\section{Binarity and Stellar Rotation for $\delta$\,Lyr Cluster Members}
+
+f15supp.pdf
+
 \section{Transit and Stellar Variability Model}

### 2021-07-30 drivers/plot_hr.py

@@ -6,27 +6,20 @@ isochrones = [None]#, 'parsec', 'mist'] # could be None, mist, parsec.
 colors = ['phot_bp_mean_mag']#, 'phot_g_mean_mag']
 
 #clusters = ['δ Lyr cluster', 'IC 2602', 'Pleiades', 'μ Tau', 'UCL']
-clusters = ['δ Lyr cluster', 'IC 2602', 'Pleiades']#, 'μ Tau', 'UCL']
 
 PLOTDIR = os.path.join(RESULTSDIR, 'hr')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
-rp.plot_hr(PLOTDIR, isochrone=None, color0='phot_bp_mean_mag', show100pc=1,
-           clusters=['δ Lyr cluster', 'IC 2602', 'Pleiades'], reddening_corr=1)
-
-rp.plot_hr(PLOTDIR, isochrone=None, color0='phot_bp_mean_mag', show100pc=1,
-           clusters=['δ Lyr cluster', 'IC 2602', 'Pleiades'], reddening_corr=1,
-           smalllims=1)
-
-rp.plot_hr(PLOTDIR, isochrone=None, color0='phot_bp_mean_mag', show100pc=1,
-           clusters=['δ Lyr cluster', 'IC 2602', 'Pleiades', 'UCL'],
-           reddening_corr=1, smalllims=1)
-
-assert 0
-for isochrone in isochrones:
-    for c in colors:
-        for o in [1]:
-            for reddening_corr in [1,0]:
-                rp.plot_hr(PLOTDIR, isochrone=isochrone, color0=c, show100pc=o,
-                           clusters=clusters, reddening_corr=reddening_corr)
+# subsets actually used in paper with & w/out reddening
+for r in [1,0]:
+    rp.plot_hr(PLOTDIR, isochrone=None, color0='phot_bp_mean_mag', show100pc=0,
+               clusters=['δ Lyr cluster'], reddening_corr=r)
+    rp.plot_hr(PLOTDIR, isochrone=None, color0='phot_bp_mean_mag', show100pc=1,
+               clusters=['δ Lyr cluster', 'IC 2602', 'Pleiades'], reddening_corr=r)
+    rp.plot_hr(PLOTDIR, isochrone=None, color0='phot_bp_mean_mag', show100pc=1,
[…]

### 2021-07-30 paper/collect.sh

@@ -26,4 +26,10 @@ cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit_Kepler_1627_Q15slc/K
 cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20210702_showss_colorbyage_showkep1627.pdf f14.pdf
 
 
-cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades.pdf f15a.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018.pdf f15a.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_smalllims.pdf f15b.pdf
+
+
+cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_harderlsp.pdf f15c.pdf
+
+cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_emphbinaries_talkaspect_kinematicspatialselected_linear_harderlsp.pdf f15supp.pdf

### 2021-07-30 rudolf/helpers.py

@@ -4,17 +4,21 @@ Data getters:
         get_deltalyr_kc19_gaia_data
         get_deltalyr_kc19_comovers
         get_deltalyr_kc19_cleansubset
+        get_autorotation_dataframe
     Kepler 1627:
         get_kep1627_kepler_lightcurve
         get_keplerfieldfootprint_dict
         get_flare_df
 
-    Other stellar and cluster datasets:
+    Get other stellar and cluster datasets:
         get_gaia_catalog_of_nearby_stars
         get_clustermembers_cg18_subset
 
     Supplement a set of Gaia stars with extinctions and corrected photometry:
         supplement_gaia_stars_extinctions_corrected_photometry
+
+    Clean Gaia sources based on photometry:
+        get_clean_gaia_photometric_sources
 
 Proposal/RM-related:
     get_simulated_RM_data
@@ -27,6 +31,7 @@ import os, collections, pickle
 import numpy as np, pandas as pd
 from glob import glob
 from copy import deepcopy
+from datetime import datetime
 
 from numpy import array as nparr
 
@@ -556,7 +561,7 @@ def supplement_gaia_stars_extinctions_corrected_photometry(
     return df
 
 
-def get_cleansel(df):
+def get_clean_gaia_photometric_sources(df):
     """
     Given a dataframe of Gaia DR2 columns, apply the "cleaning
     selection" described on page 3 / Appendix B of GaiaCollab+2018 HR
@@ -600,3 +605,301 @@ def get_cleansel(df):
     return sel
 
 
+def get_autorotation_dataframe(runid='deltaLyrCluster', verbose=1,
+                               cleaning='defaultcleaning'):
+    """
+    runid = 'deltaLyrCluster', for example
+
+    Cleaning options:
+        'defaultcleaning' P<15d, LSP>0.1, Nequal==0, Nclose<=1.
+        'harderlsp' P<15d, LSP>0.15, Nequal==0, Nclose<=1.
+        'nocleaning': P<99d.
+        'defaultcleaning_cutProtColor': add Prot-color plane cut to
+            defaultcleaning.
+    """
+
+    assert isinstance(cleaning, str)
+
+    rotdir = os.path.join(DATADIR, 'rotation')
+
+    #FIXME need to create this file for the deltaLyrCluster...
+    # contains all the GAIA 
[…]

### 2021-07-30 rudolf/plotting.py

@@ -13,6 +13,7 @@ plot_ttv_vs_local_slope
 plot_rotation_period_windowslider
 plot_flare_pair_time_distribution
 plot_hr
+plot_rotationperiod_vs_color
 """
 import os, corner, pickle, inspect
 from glob import glob
@@ -58,7 +59,7 @@ from rudolf.helpers import (
     get_gaia_catalog_of_nearby_stars, get_clustermembers_cg18_subset,
     get_mutau_members, get_ScoOB2_members,
     supplement_gaia_stars_extinctions_corrected_photometry,
-    get_cleansel
+    get_clean_gaia_photometric_sources
 )
 from rudolf.extinction import (
     retrieve_stilism_reddening, append_corrected_gaia_phot_Gagne2020
@@ -1499,7 +1500,7 @@ def plot_hr(
         df.to_csv(outpath, index=False)
     df = pd.read_csv(outpath)
     if cleanhrcut:
-        df = df[get_cleansel(df)]
+        df = df[get_clean_gaia_photometric_sources(df)]
     if reddening_corr:
         print('delta Lyr cluster')
         print(df['reddening[mag][stilism]'].describe())
@@ -1580,7 +1581,7 @@ def plot_hr(
             _df.to_csv(outpath, index=False)
         _df = pd.read_csv(outpath)
         if cleanhrcut:
-            _df = _df[get_cleansel(_df)]
+            _df = _df[get_clean_gaia_photometric_sources(_df)]
         if reddening_corr:
             print('UCL')
             print(_df['reddening[mag][stilism]'].describe())
@@ -1612,7 +1613,7 @@ def plot_hr(
             _df.to_csv(outpath, index=False)
         _df = pd.read_csv(outpath)
         if cleanhrcut:
-            _df = _df[get_cleansel(_df)]
+            _df = _df[get_clean_gaia_photometric_sources(_df)]
         if reddening_corr:
             print('IC2602')
             print(_df['reddening[mag][stilism]'].describe())
@@ -1644,7 +1645,7 @@ def plot_hr(
             _df.to_csv(outpath, index=False)
         _df = pd.read_csv(outpath)
         if cleanhrcut:
-            _df = _df[get_cleansel(_df)]
+            _df = _df[get_clean_gaia_photometric_sources(_df)]
         if reddening_corr:
             print('Pleaides')
             print(_df['reddening[mag][stilism]'].describe())
@@ -1955,6 +1956,216 @@ def plot_hr(
[…]

### 2021-07-30 paper/ms.tex
update ms for rotation

@@
 % gaia target sample numbers
-\newcommand{\nkinematic}{3{,}298} % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
-\newcommand{\nnbhd}{13{,}843} % make_compstar_NGC_2516_sourcelist.py
-\newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
-\newcommand{\nhalo}{2{,}192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes
+\newcommand{\nkinematic}{1{,}201} % deltalyr_kc19_cleansubset_withreddening.csv
+% all from oneoff_calculate_rotation_numbers.py
+\newcommand{\noriginal}{3{,}071} % Vizier; KC19 Theia 73 member count
+\newcommand{\nwithtess}{924} % How many stars have TESS data?
+\newcommand{\nkinwithtess}{391} % How many kinematically selected stars have TESS data?
+\newcommand{\nkinwithtessandcrowding}{192} % How many kinematically selected stars have TESS data and meeting the crowding cutoff (nequal<=0, nclose<=1)?
+\newcommand{\nkindefaultcleaning}{157} % How many of these meet the "defaultcleaning" criteria (LSP>0.1, Prot<15, and crowding cut)?
+\newcommand{\nrotgood}{147} % nkindefaultcleaning, minus outliers.
 
@@
 			\includegraphics[width=0.49\textwidth]{f15a.pdf}
-			\includegraphics[width=0.49\textwidth]{f15b.pdf}
+			\includegraphics[width=0.469\textwidth]{f15b.pdf}
 		}
@@
     {\it Bottom:} TESS and Kepler rotations periods and dereddened Gaia
-    colors, with the Pleiades
-    \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
-    \citep[650\,Myr;][]{douglas_poking_2017} shown for reference.
-    Stars of a given color (mass) move up through time due to magnetic
-    braking.  The \cn\ has a 40 Myr gyrochronal age, as calibrated
-    against the LDB-age of IC\,2602.  {\bf todo: caption improve.}.
+    colors, with the Pleiades (125\,Myr)
+    and Praesepe (650\,Myr)
+    shown for reference
+    \citep{rebull_rotation_2016a,douglas_poking_2017}.
+    Kepler\,1627 is shown with a yellow star.
+     {\bf todo: caption improve.}.
    \label{fig:age}
@@
 
-\subsection{Rotation Periods}
-%FIXME TODO REFERENCE LOWER PANELS
-{\bf Jason Curtis todo}
-We collected the previously measured
-{\bf XXX} Kepler rotation periods (CITEP MCQUILLAN2014), and measured
-{\bf YYY} new stellar rotation periods for cluster members observed by
-TESS.  We also performed an isochronal analysis.  The results are
-shown in Figure~\ref{fig:age}.  From the ensemble perspective, which
+\subsection{Stellar Rotation Periods}
+
+Of the \noriginal\ candidate \cn\ members reported by
+\citet{kounkel_untangling_2019}, TESS data were available for
+\nwithtess\ stars.  We extracted light curves using the nearest pixel
+to each star, and regressed them against systematics using the causal
+pixel model implemented by \citet{hattorio_2021_cpm} in the
+\texttt{unpopular} package.  We then measured candidate rotation
+periods using a Lomb-Scargle periodogram
+\citep{lomb_1976,scargle_studies_1982,astropy_2018}.  To enable cuts
+on crowding, we queried the Gaia source catalog for stars within a
+$21\farcs0$ radius of the target star (1 TESS pixel).  We recorded the
+number of stars within the aperture with greater brightness than the
+target star, and with brightness within 1.25 and 2.5 TESS magnitudes
+of the target star.  {\bf We then supplemented these measurements with XXX Kepler
+rotation periods (CITEP MCQUILLAN2014)}.
+
+We cleaned the candidate TESS rotation period measurements through a
+combination of automated and manual steps.  As a matter of scope, we
+restricted our attention to the \nkinwithtess\ stars in
+the spatial and kinematic proximity of Kepler\,1627, as discussed in
+Appendix~\ref{app:kinematicselection}.  We also only focused on
+stars for which no companions were known with a brightness exceeding one-tenth of
+the target star within a $21\farcs0$ radius: there were 
+\nkinwithtessandcrowding\ stars that met these requirements, and that
+had TESS data available.
+For plotting purposes we then imposed a selection based on the strength of the
+rotation period signal itself: we required the Lomb Scargle power to
+exceed 0.1, and the period to be below 15\,days.
+
+The lower panel of Figure~\ref{fig:age} shows the resulting
+\nkindefaultcleaning\ stars.  The majority of these stars fall below
+the ``slow sequence'' of the Pleiades, consistent with a
+gyrochronological age for the \cn\ below 100\,Myr.  Approximately 10 stars 
+appear as outliers above the ``slow sequence''.  Assuming that they
+are all false positives ({\it i.e.}, field interlopers), the rotation
+period detection fraction would be
+\nrotgood/\nkinwithtessandcrowding$\approx 77\%$.
+
+Binarity is known to be an important factor in the density of stars
+falling underneath the ``slow sequence''
+\citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}.
+Appendix~\ref{app:rotationbinarity} discusses this topic, and finds
+evidence for similar effects to those that have been previously
+discussed in the literature.
+
+\subsection{Age of Kepler\,1627}
+
+%FIXME FIXME FIXME
+From the ensemble perspective, which
 provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
@@
 conclusion.
-
-For photometric and astrometric binaries, see the appendix!
 
@@
 \citet{kounkel_untangling_2019} to be in the group ``Theia 73'', which
-was cross-matched by \citealt{kounkel_untangling_2019} as being
-``Stephenson 1''.  For our Figure, galactic positions were calculated
-and plotted only for stars with parallax signal-to-noise exceeding 20.
-The location of the Sun is shown on the plots.  The smattering of
-reported cluster members (gray and black points) has a significantly
-different structure that the cluster initially identified by
-\citet{stephenson_possible_1959} and corroborated by
+was cross-matched by Kounkel \& Covey as being ``Stephenson 1''.
+\citet{kounkel_untangling_2019} reported \noriginal\ stars to be
+present in this cluster.  For our Figure, galactic positions were
+calculated and plotted only for stars with parallax signal-to-noise
+exceeding 20.  The location of the Sun is shown on the plots.  The
+smattering of reported cluster members (gray and black points) has a
+significantly different structure that the cluster initially
+identified by \citet{stephenson_possible_1959} and corroborated by
 \citet{eggen_photometric_1968}.  While the non-uniform ``clumps''
@@
 \section{Binarity and Stellar Rotation for $\delta$\,Lyr Cluster Members}
-
-f15supp.pdf
+\label{app:rotationbinarity}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f15supp.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+  \caption{ {\bf Binarity indicators for rotators in \cn.} 
+    Data are as in the lower panel of
+    Figure~\ref{fig:age}.  Astrometric binaries ($\mathrm{RUWE}>1.2$)
+    appear in red; photometric binaries ($>$0.3\,mag above an
+    empirical isochrone; 20\% of stars) are orange.  
+    Over $0.6<\bpmrpo<1.5$, the detected rotation periods in binaries tend to 
+    fall below the slow sequence.
+    %FIXME!
+		% Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has \replaced{281}{523} stars.
+		% Dividing these stars into ``slow'' and ``fast'' sequences by eye,
+		% the fraction of stars showing signs of binarity in the slow and fast
+		% subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and
+		% \replaced{56\% (40/71)}{51\% (68/134)}, respectively.
+		\label{fig:binarity}
+	}
+\end{figure*}
+
+Figure~\ref{fig:binarity} shows the rotation-color diagram for \cn\
+members, with the points colored according to indicators of binarity.
+The binaries tend to be redder and have shorter rotation periods.  We
+know of two possible explanations based on selection effects, and two
+based on physics.
+
+Possible selection effects include {\it i)} binaries have a component
+that contributes additional red light to the system, which could skew
+the color measurement of the primary; and {\it ii)} the unresolved
+binary companions could contaminate the rotation period measurement
+\citep[{\it e.g.},][Section~5.1]{stauffer_rotation_2016}.
+
+The two possible physical effects that could be relevant are tidal
+locking and pre-main-sequence disk locking.  Tidal locking has been
+argued to be an unlikely explanation for the frequency of rapid
+rotators due to the population statistics; a more likely scenario
+would be that the presence of the binary leads to faster disk
+dispersal, enabling the primary to contract to more rapid rotation
+periods than possible for single stars
+\citep{meibom_effect_2007,bouma_cluster_2020}.
 

### 2021-07-30 paper/collect.sh

@@ -29,7 +29,6 @@ cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20210
 cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018.pdf f15a.pdf
 cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_smalllims.pdf f15b.pdf
 
+cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_defaultcleaning_overplotkep1627.pdf f15c.pdf
 
-cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_harderlsp.pdf f15c.pdf
-
-cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_emphbinaries_talkaspect_kinematicspatialselected_linear_harderlsp.pdf f15supp.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_emphbinaries_talkaspect_kinematicspatialselected_linear_defaultcleaning.pdf f15supp.pdf

### 2021-07-30 rudolf/plotting.py

@@ -1961,7 +1961,8 @@ def plot_hr(
 def plot_rotationperiod_vs_color(outdir, runid, yscale='linear', cleaning=None,
                                  emph_binaries=False, refcluster_only=False,
                                  talk_aspect=0, xval_absmag=0,
-                                 kinematic_selection=0):
+                                 kinematic_selection=0,
+                                 overplotkep1627=0):
     """
     Plot rotation periods that satisfy the automated selection criteria
     (specified in helpers.get_autorotation_dataframe)
@@ -2013,7 +2014,7 @@ def plot_rotationperiod_vs_color(outdir, runid, yscale='linear', cleaning=None,
 
         else:
             from rudolf.helpers import get_autorotation_dataframe
-            auto_df = get_autorotation_dataframe(
+            auto_df, base_df = get_autorotation_dataframe(
                 runid, cleaning=cleaning
             )
 
@@ -2097,6 +2098,16 @@ def plot_rotationperiod_vs_color(outdir, runid, yscale='linear', cleaning=None,
                 marker='o', linewidths=_lw, label="Astrometric binary"
             )
 
+    if overplotkep1627:
+        sel = (df.dr2_source_id == 2103737241426734336)
+        sdf = df[sel]
+        ax.plot(
+            xval[sel], df[sel]['period'], alpha=1, mew=0.5,
+            zorder=9001, label='Kepler 1627',
+            markerfacecolor='yellow', markersize=14, marker='*',
+            color='black', lw=0
+        )
+
     ax.set_ylabel('Rotation Period [days]', fontsize='medium')
 
     if not xval_absmag:
@@ -2111,9 +2122,9 @@ def plot_rotationperiod_vs_color(outdir, runid, yscale='linear', cleaning=None,
         ax.set_xlim((1.5, 10))
 
     if yscale == 'linear':
-        ax.set_ylim((0,15))
+        ax.set_ylim((0,13))
     elif yscale == 'log':
-        ax.set_ylim((0.05,15))
+        ax.set_ylim((0.05,13))
     else:
         raise NotImplementedError
     ax.set_yscale(yscale)
@@ -2162,8 +2173,12 @@ def plot_rotationperiod_vs_color(outdir, runid, yscale='linear', cleaning=None,
         outstr += '_xvalAbsG'
     if kinematic_selection:
         outstr += '_kinematicspatialselected'
+    else:
+        outstr += '_allKC19'
[…]

### 2021-08-04 paper/ms.tex
age now has uncertainties

@@
   with rotation periods from TESS, we show that Kepler\,1627 is a
-  member of the {\bf $35\pm XX$\,Myr old} $\delta$~Lyr cluster.
+  member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr cluster.
   Kepler\,1627Ab is therefore the youngest planet with a precise age
@@
 host cluster (Section~\ref{sec:cluster}).  We find that the $\delta$
-Lyr cluster is {\bf $\approx35$\,Myr old}, and that \sn\ is indeed a member
+Lyr cluster is $35^{+6}_{-5}$\,Myr old, and that \sn\ is indeed a member
 of the cluster.  Focusing on the planet itself
@@
 
-\subsection{HR Diagram}
+\subsection{Color-Magnitude Diagram}
 We meaured the isochrone age using an empirical approach, and also
 compared against theoretical models.  The left panel of
-Figure~\ref{fig:age} shows the HR diagram of stars in the \cn, the
+Figure~\ref{fig:age} shows the color-magnitude diagram (CMD) of stars in the \cn, the
 Pleiades, IC\,2602, and the field.  The stars shown in the \cn\ are
@@
 \citet{GaiaCollaboration2018}, except that we loosened the parallax
-precision requirement to $\texttt{parallax\_over\_error}>5$.  These
+precision requirement to $\varpi/\sigma_\varpi>5$.  These
 filters were designed to include genuine binaries while omitting
@@
 literature \citep[{\it
-e.g.},][]{GaiaCollaboration2018,kounkel_untangling_2019,bossini_age_2019},
-and have the benefit of being sensitive to effects such as
-differential reddening that may be present in the dust maps.
+e.g.},][]{GaiaCollaboration2018,kounkel_untangling_2019,bossini_age_2019}.
 
@@
 age, since they overlap.  In our exploration, we also compared against
-$\mu$-Tau ($\approx 60$\,Myr; \citealt{gagne_mutau_2020}) and subsets
-of the Sco OB2 association ($\approx10$-$15$\,Myr;
-\citealt{damiani_stellar_2019}), and confirmed that the
-pre-main-sequence M dwarfs were intermediate between these two
-clusters.
-
-% FIXME FIXME
-We quantified this impression following the empirical fitting procedure
-discussed by \citet{gagne_mutau_2020}.
+$\mu$-Tau ($\approx 60$\,Myr; \citealt{gagne_mutau_2020}) and the
+Upper-Centaurus-Lupus component of the Sco OB2 association (UCL;
+$\approx10$-$15$\,Myr; \citealt{damiani_stellar_2019}).  We confirmed
+that the pre-main-sequence M dwarfs of the \cn\ were intermediate
+between these two clusters.  To quantify this into an age measurement
+with an uncertainty, we adopted the empirical procedure
+advanced by \citet{gagne_mutau_2020} in their Section~6.3.  In brief,
+we parametrized the pre-main-sequence locus of each of a set of
+reference clusters using splines, and the locus of the target \cn\ was
+then fitted as a piecewise linear combination of these reference
+clusters.  As reference clusters, we adopted members of UCL, IC\,2602, and the
+Pleaides from \citet{damiani_stellar_2019} and \cite{CantatGaudin2018a}.
+We removed binaries from each set by requiring
+$\texttt{RUWE}<1.3$, $\texttt{radial\_velocity\_error}$ below the
+80$^{\rm th}$ percentile of each cluster's distribution, and by
+manually excluding stars that were obvious photometric binaries.  We
+then passed a moving box average and standard deviation across the
+CMD in 0.10\,mag bins, fitted a univariate spline to the binned
+values, and assembled a piecewise grid of hybrid isochrones spanning
+the ages between UCL to the Pleiades following Equations~6 and~7 from
+\citet{gagne_mutau_2020}.
+
+The ages calculated from this procedure depend on the ages assumed for
+each reference cluster.
+In our case, the IC\,2602 cluster is most important.
+As noted by \citet{bouma_cluster_2020}, plausible ages for this
+cluster span 30 to 46\,Myr, with slightly older ages being
+preferred by the lithium-depletion-boundary (LDB) measurements and younger
+ages by the main-sequence turn-off
+\citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019}.
+If we were to adopt the lower 30\,Myr age for IC\,2602, the \cn\ would be
+$31^{+5}_{-4}$\,Myr old.
+For the converse case of the 46\,Myr boundary, the \cn\ would be
+$44^{+8}_{-7}$\,Myr old.
+Taking an intermediate 35\,Myr age as our reference, 
+the resulting age for the \cn\ is
+$35^{+6}_{-5}$\, Myr, though it should be understood based on the
+systematic uncertainties that an age of up to $\approx50$\,Myr remains
+plausible.  A detailed analysis of the LDB and main-sequence
+turn-off in the \cn\ could help clarify the issue, but both are beyond
+our scope.
 

### 2021-08-04 drivers/calc_empirical_isochrone_age.py

@@ -1,0 +1,342 @@
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+import os, pickle
+from numpy import array as nparr
+
+from rudolf.paths import DATADIR, RESULTSDIR
+
+from rudolf.helpers import (
+    get_deltalyr_kc19_gaia_data, get_simulated_RM_data,
+    get_keplerfieldfootprint_dict, get_deltalyr_kc19_comovers,
+    get_deltalyr_kc19_cleansubset, get_kep1627_kepler_lightcurve,
+    get_gaia_catalog_of_nearby_stars, get_clustermembers_cg18_subset,
+    get_mutau_members, get_ScoOB2_members,
+    supplement_gaia_stars_extinctions_corrected_photometry,
+    get_clean_gaia_photometric_sources
+)
+from cdips.utils.gaiaqueries import given_source_ids_get_gaia_data
+
+from aesthetic.plot import savefig, format_ax, set_style
+
+from scipy.ndimage.filters import uniform_filter
+from scipy.interpolate import UnivariateSpline, interp1d
+
+from astropy.modeling import models, fitting
+
+extinctionmethod = 'gaia2018'
+
+#
+# getters made by plot_hr
+#
+def _get_deltalyr():
+    deltalyrpath = os.path.join(
+        RESULTSDIR, 'tables', 'deltalyr_kc19_cleansubset_withreddening.csv'
+    )
+    df = pd.read_csv(deltalyrpath)
+    df = df[get_clean_gaia_photometric_sources(df)]
+    return df
+
+def _get_ucl():
+    outpath = os.path.join(
+        RESULTSDIR, 'tables', f'UCL_withreddening_{extinctionmethod}.csv'
+    )
+    _df = pd.read_csv(outpath)
+    _df = _df[get_clean_gaia_photometric_sources(_df)]
+    ruwe_df =  given_source_ids_get_gaia_data(
+        nparr(_df.source_id).astype(np.int64), 'UCL_rudolf_dr2_ruwe',
+        n_max=10000, overwrite=False, gaia_datarelease='gaiadr2',
+        getdr2ruwe=True
+    )
+    mdf = _df.merge(ruwe_df, how='left', on='source_id')
+    return mdf
+
+def _get_ic2602():
+    outpath = os.path.join(
[…]

### 2021-08-04 paper/collect.sh

@@ -26,7 +26,7 @@ cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit_Kepler_1627_Q15slc/K
 cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20210702_showss_colorbyage_showkep1627.pdf f14.pdf
 
 
-cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018.pdf f15a.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_overplotkep1627.pdf f15a.pdf
 cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_smalllims.pdf f15b.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_defaultcleaning_overplotkep1627.pdf f15c.pdf

### 2021-08-06 paper/ms.tex
update ms

--- 2021-07-30 paper/ms.tex
@@
 \begin{abstract}
-  Kepler\,1627A is a G8V star previously found to host a
+  Kepler\,1627A is a G8V star previously known to host a
   $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
@@
   with rotation periods from TESS, we show that Kepler\,1627 is a
-  member of the {\bf $35\pm XX$\,Myr old} $\delta$~Lyr cluster.
-  Kepler\,1627Ab is therefore the youngest planet with a precise age
+  member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr cluster.
+  To our knowledge, Kepler\,1627Ab is therefore the youngest planet with a precise age
   yet found by the main Kepler mission.  The Kepler
@@
 
+%TODO: might want to revise this text to better frame the boil-off phase...
 %TODO: better citations on radius evolution?
@@
 their first $10^8$ years
-\citep{owen_constraining_2020}.  Specifically,
-they start with sizes of 4--12\,$R_\oplus$ at the time of disk
+\citep{owen_atmospheres_2016,owen_constraining_2020}.  Specifically,
+they start with sizes of 4--12\,$R_\oplus$ near the time of disk
 dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
@@
 
-A previous HDBScan-based clustering analysis of the Gaia data found
-that Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was a member
+A previous clustering analysis of the Gaia data reported
+that Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member
 of the \cn\ \citep{kounkel_untangling_2019}.  Given the statistical
-validation of close-in mini-Neptune
+validation of its close-in mini-Neptune
 \citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
-we opted to examine the age of the star and its putative
-host cluster (Section~\ref{sec:cluster}).  We find that the $\delta$
-Lyr cluster is {\bf $\approx35$\,Myr old}, and that \sn\ is indeed a member
+we examined the age of the star and its putative
+host cluster most closely (Section~\ref{sec:cluster}).  We find that the $\delta$
+Lyr cluster is $35^{+6}_{-5}$\,Myr old, and that \sn\ is indeed a member
 of the cluster.  Focusing on the planet itself
@@
 unreported binary companion, it is indeed a mini-Neptune orbiting the
-roughly Solar-type primary star.  The presence of a correlation
+G-dwarf primary star.  The presence of a correlation
 between transit timing variations and the local light curve slope
@@
 
-\subsection{HR Diagram}
-We meaured the isochrone age using an empirical approach, and also
-compared against theoretical models.  The left panel of
-Figure~\ref{fig:age} shows the HR diagram of stars in the \cn, the
-Pleiades, IC\,2602, and the field.  The stars shown in the \cn\ are
-selected based on their positions and kinematics, as discussed in
-Appendix~\ref{app:kinematicselection}.  Those from the Pleiades and
-IC\,2602 are adopted from \citet{cantatgaudin_gaia_2018}, and the
-field stars are from the Gaia EDR3 Catalog of Nearby Stars
-\citep{gaia_gcns_2021}.  We cleaned these following identical data
-filtering criteria to those discussed in Appendix~B of
-\citet{GaiaCollaboration2018}, except that we loosened the parallax
-precision requirement to $\texttt{parallax\_over\_error}>5$.  These
+\subsection{Color-Magnitude Diagram}
+\label{sec:cmd}
+
+We meaured the isochrone age using an empirical approach.  The left
+panel of Figure~\ref{fig:age} shows the color-magnitude diagram (CMD)
+of stars in the \cn, IC\,2602, the Pleiades, and the field.  The stars
+shown in the \cn\ are selected based on their positions and
+kinematics, as discussed in Appendix~\ref{app:kinematicselection}.
+Those from the Pleiades and IC\,2602 are adopted from
+\citet{cantatgaudin_gaia_2018}, and the field stars are from the Gaia
+EDR3 Catalog of Nearby Stars \citep{gaia_gcns_2021}.  We cleaned these
+following identical data filtering criteria discussed in
+Appendix~B of \citet{GaiaCollaboration2018}, except that we weakened
+the parallax precision requirement to $\varpi/\sigma_\varpi>5$.  These
 filters were designed to include genuine binaries while omitting
 instrumental artifacts.  We then corrected for extinction by querying
-the 3-D maps of \citet{capitanio_threedimensional_2017} and
+the 3-dimensional maps of \citet{capitanio_threedimensional_2017} and
 \citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}},
-and by then using the extinction coefficients $k_X\equiv A_X/A_0$
-computed by \citet{GaiaCollaboration2018}, while assuming that $A_0 =
+and by then applying the extinction coefficients $k_X\equiv A_X/A_0$
+computed by \citet{GaiaCollaboration2018} while assuming that $A_0 =
 3.1 E(B-V)$.  For IC\,2602, the Pleaides, and the \cn, this procedure
@@
 
-The immediate impression from the left panel of Figure~\ref{fig:age}
-is that the \cn\ and IC\,2602 share approximately the same isochronal
-age, since they overlap.  In our exploration, we also compared against
-$\mu$-Tau ($\approx 60$\,Myr; \citealt{gagne_mutau_2020}) and subsets
-of the Sco OB2 association ($\approx10$-$15$\,Myr;
-\citealt{damiani_stellar_2019}), and confirmed that the
-pre-main-sequence M dwarfs were intermediate between these two
-clusters.
-
-% FIXME FIXME
-We quantified this impression following the empirical fitting procedure
-discussed by \citet{gagne_mutau_2020}.
+The immediate impression from Figure~\ref{fig:age}
+is that the \cn\ and IC\,2602 overlap, and therefore are approximately the same isochronal
+age.  In our exploration, we also compared against
+$\mu$-Tau ($\approx 60$\,Myr; \citealt{gagne_mutau_2020}) and the
+Upper-Centaurus-Lupus component of the Sco OB2 association (UCL;
+$\approx10$-$15$\,Myr; \citealt{damiani_stellar_2019}).  
+The pre-main-sequence M dwarfs of the \cn\ were intermediate
+between these latter two clusters.  To quantify this into an age measurement
+with an uncertainty, we adopted the empirical procedure
+developed by \citet{gagne_mutau_2020} in their Section~6.  In brief,
+we fitted the pre-main-sequence loci of a set of
+reference clusters, and the locus of the target \cn\ was
+then modeled as a piecewise linear combination of these reference
+clusters.  For our reference clusters, we adopted members of UCL, IC\,2602, and the
+Pleaides from \citet{damiani_stellar_2019} and
+\cite{CantatGaudin2018a} respectively.
+We removed binaries by requiring
+$\texttt{RUWE}<1.3$, $\texttt{radial\_velocity\_error}$ below the
+80$^{\rm th}$ percentile of each cluster's distribution, and by
+manually excluding stars that were obvious photometric binaries in the
+CMD.  We then passed a moving box average and standard deviation
+across the CMD in 0.10\,mag bins, fitted a univariate spline to the
+binned values, and assembled a piecewise grid of hybrid isochrones
+spanning the ages between UCL to the Pleiades following Equations~6
+and~7 from \citet{gagne_mutau_2020}.
+
+The ages returned by this procedure depend on the ages assumed for
+each reference cluster.
+In our case, the IC\,2602 cluster is most important.
+As noted by \citet{bouma_cluster_2020}, plausible ages for IC\,2602
+span 30\,Myr to 46\,Myr, with slightly older ages being preferred by the
+lithium-depletion-boundary (LDB) measurements and younger ages by the
+main-sequence turn-off
+\citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019}.
+If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would be
+$31^{+5}_{-4}$\,Myr old.
+For the converse case of 46\,Myr, the \cn\ would be
+$44^{+8}_{-7}$\,Myr old.
+We adopt an intermediate 35\,Myr age as our reference, 
+yielding an age for the \cn\ of
+$35^{+6}_{-5}$\, Myr, though it should be understood based on the
+systematic uncertainties that an age of up to $\approx50$\,Myr remains
+plausible.  Detailed analyses of the LDB or main-sequence
+turn-off in the \cn\ could help clarify the issue, but both are beyond
+our scope.
 
@@
 to each star, and regressed them against systematics using the causal
-pixel model implemented by \citet{hattorio_2021_cpm} in the
-\texttt{unpopular} package.  We then measured candidate rotation
+pixel model implemented in the \texttt{unpopular} package
+\citep{hattorio_2021_cpm}.  We then measured candidate rotation
 periods using a Lomb-Scargle periodogram
@@
 \subsection{Age of Kepler\,1627}
+
+Kinematics are reasonably compelling
+(Appendix~\ref{app:kinematicselection}).
+HR diagram in and of itself doesn't give any age information about the
+star.   (Could quote funny literature isochrone ages).
+Rotation is informative, and is consistent with cluster.
+Final point: Lithium.  Quote your EW. 
+All together, pretty good.  We adopt the age based on the pre-main-sequence
+empirical fitting procedure from Section~\ref{sec:cmd}:
+$35^{+6}_{-5}$\, Myr.
 

### 2021-08-07 paper/ms.tex
update ms

@@
 % paragraph 1: HR 
-To assess the age of the cluster as well as our star of interest,
+To assess the age of the cluster and our star of interest,
 we used a combination of the isochronal, gyrochronal, and
@@
 
-Kinematics are reasonably compelling
-(Appendix~\ref{app:kinematicselection}).
-HR diagram in and of itself doesn't give any age information about the
-star.   (Could quote funny literature isochrone ages).
-Rotation is informative, and is consistent with cluster.
-Final point: Lithium.  Quote your EW. 
-All together, pretty good.  We adopt the age based on the pre-main-sequence
-empirical fitting procedure from Section~\ref{sec:cmd}:
-$35^{+6}_{-5}$\, Myr.
-
-%FIXME FIXME FIXME
-From the ensemble perspective, which
-provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
-of the star is $XX \pm YY$\,Myr.  Individual measurements of the
-star's rotation period and lithium abundance corroborate this
-conclusion.
+To summarize, the age of the \cn\ cluster based on the empirical
+fitting procedure described in Section~\ref{sec:cmd} is
+$35^{+6}_{-5}$\, Myr.  This procedure hinges on the age adopted
+for IC\,2602, since it is the closest reference cluster in the color-magnitude diagram.
+Gyrochronologically, while most of the G and K
+dwarfs in the \cn\ have not yet converged to form a slow sequence,
+their rotation distribution and that of the F dwarfs is consistent
+with the same age.  The rotation periods of other
+30\,Myr to 50\,Myr clusters ({\it e.g.}, IC\,2602, Collinder 135)
+are indistinguishable
+(S.~Douglas, priv{.}~comm{.}).  The most precise age for \sn\ and its
+planet are therefore also $35^{+6}_{-5}$\, Myr, based on the star's spatial,
+kinematic, and rotational connection to this larger ensemble, and the assumption that
+\pn\ formed very quickly after the star itself.
+
+% TODO:  update this based on HIRES instead of TRES
+% TODO:  cite Stephanie instead of priv comm? Send her your Prot
+Beyond the kinematic and rotational lines of evidence for the youth of \sn,
+there is a final consistency check: the abundance of Li\,\textsc{I} as
+inferred from the 6708\,\AA\ doublet.  This connection of this element
+to age measurements for young stars has been reviewed by
+\citet{soderblom_ages_2014}.  Following the equivalent width
+measurement procedure described by \citet{bouma_2021_ngc2516}, we find
+${\rm EW}_{\rm Li} = 232^{+23}_{-13}\,$m\AA, which does not
+correct for the Fe\,\textsc{I} blend at 6707.44\AA\.  Given $T_{\rm eff} \approx
+5500\,{\rm K}$, this measurement is also in agreement with an age for
+the star between 30\,Myr and 50\,Myr (see {\it e.g.}, Figure~9 of
+\citealt{bouma_2021_ngc2516}, or the data
+from \citealt{randich_gaiaeso_2018}).
 
@@
 
-If the \pn\ transit signal is created by a bonafide planet, this age
-would make it the youngest planet yet found by the main Kepler
-mission\footnote{The re-purposed K2 mission found two slightly younger
-systems: K2-33b \citep{David_et_al_2017,Mann_K2_33b_2016} and V1298
-Tau \citep{david_four_2019}}.  A portion of the light curve is shown
-with the phase-folded transit
-in Figure~\ref{fig:lc}.  The dominant signal in the
-\texttt{PDCSAP} photometry is a quasi-period starspot signal with a
-peak-to-peak amplitude that varies between roughly 2\% and 8\%,
+If the \pn\ transit signal is created by a bonafide planet,  then to
+our knowledge its age would make it the youngest planet yet found by
+the main Kepler mission\footnote{The re-purposed K2 mission found two
+slightly younger systems: K2-33b
+\citep{David_et_al_2017,Mann_K2_33b_2016} and V1298 Tau
+\citep{david_four_2019}}.  A portion of the light curve is shown with
+the phase-folded transit in Figure~\ref{fig:lc}.  The dominant signal
+in the \texttt{PDCSAP} photometry is a quasi-periodic starspot signal
+with a peak-to-peak amplitude that varies between roughly 2\% and 8\%,
 depending on the degree of asymmetry in spot coverage between the

### 2021-08-12 paper/ms.tex
update ms

@@
 	\caption{
-    {\bf The \cn\ is $35\pm15$\,Myr old.}  {\it Top:} HR diagram of 
+    {\bf The \cn\ is $35\pm15$\,Myr old.}  {\it Top:} Color-magnitude diagram of 
     the \cn, IC\,2602 ($\approx35$\,Myr), the Pleiades ($\approx 115$\,Myr), and the field.
@@
 %
+% ACKNOWLEDGE PFS / CAMPANAS.
+%
+This paper also includes data collected by the TESS mission, which are
+publicly available from the Mikulski Archive for Space Telescopes
+(MAST).
+%
+Funding for the TESS mission is provided by NASA's Science Mission
+directorate.
+%
+We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
+S.~Seager, J.~Jenkins) and the many TESS team members for their
+efforts to make the mission a continued success.
+%
+
+%
 This study was based in part on observations at Cerro Tololo
@@
 %
-% ACKNOWLEDGE PFS / CAMPANAS.
-%
-This paper also includes data collected by the TESS mission, which are
-publicly available from the Mikulski Archive for Space Telescopes
-(MAST).
-%
-Funding for the TESS mission is provided by NASA's Science Mission
-directorate.
-%
-We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
-S.~Seager, J.~Jenkins) and the many TESS team members for their
-efforts to make the mission a continued success.
+%
+This research has made use of the Keck Observatory Archive (KOA),
+which is operated by the W. M. Keck Observatory and the NASA Exoplanet
+Science Institute (NExScI), under contract with the National
+Aeronautics and Space Administration.
 %

### 2021-08-12 paper/ms.tex
update ms and CMD age

@@
 
-% 195 words (250 max)
+% 208 words (250 max)
 \begin{abstract}
   Kepler\,1627A is a G8V star previously known to host a
-  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
+  $3.5\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
   observed by the Kepler satellite because it is nearby ($d\approx
-  333\,{\rm pc}$) and it resembles the Sun.  Combining Gaia kinematics
-  with rotation periods from TESS, we show that Kepler\,1627 is a
-  member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr cluster.
-  To our knowledge, Kepler\,1627Ab is therefore the youngest planet with a precise age
-  yet found by the main Kepler mission.  The Kepler
-  photometry seems to show two anomalies.  First, unresolved
-  starspot-crossing events distort the average transit shape, and
-  induce transit timing variations that imply the planet is likely on
-  a prograde orbit.  Second, although the flare arrival time
-  distribution is probably Poissonian, there may be an excess of
-  flares separated by the orbital and synodic periods, and linear
-  combinations thereof.  The $\delta$~Lyr cluster is one of many
-  stellar groups whose properties have been clarified using the Gaia
-  data; many of the other new clusters also overlap with confirmed and
-  candidate exoplanets.  Further studies that combine Gaia kinematics
-  and TESS rotation periods are well-poised to expand the census of
-  age-dated planets, and our understanding of planetary evolution.
+  333\,{\rm pc}$) and it resembles the Sun.  Combining kinematic
+  information from Gaia, stellar rotation periods from TESS, and Li
+  abundances from ground-based spectra, we show that Kepler\,1627 is a
+  member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr cluster.  To our
+  knowledge, Kepler\,1627Ab is the youngest planet with a precise age
+  yet found by the main Kepler mission.  From both unresolved
+  starspot-crossing events in the Kepler photometry and from a
+  ground-based Keck/HIRES transit we find that the planet is likely on
+  an orbit nearly aligned with the stellar spin axis.  We also find
+  from the short-cadence Kepler photometry that although the flare
+  arrival time distribution is probably Poissonian, there may be an
+  excess of flares separated by the orbital and synodic periods, and
+  linear combinations thereof.  The $\delta$~Lyr cluster is one of
+  many stellar groups whose properties have been clarified using the
+  Gaia data; many of the other new clusters also overlap with
+  confirmed and candidate exoplanets.  Further studies that combine
+  Gaia kinematics and TESS rotation periods are poised to expand
+  the census of age-dated planets, and our understanding of planetary
+  evolution.
 \end{abstract}
-
-% While thousands of exoplanets have been discovered orbiting nearby
-% stars, the vast majority of them are several billion years old.  This
-% makes it difficult to test theories for the origins of the different
-% families of planets, since many evolutionary processes are expected to
-% operate on timescales of less than 100 million years.  The K2 and TESS
-% missions have enabled advances on this front [1,2,3,4,5,6].  The
-% Kepler mission however has not yielded any planets with precise ages
-% below one gigayear [7,8].  Here, we combine kinematic data from Gaia
-% with stellar rotation periods from TESS to show that Kepler 1627 is a
-% member of the 35 ± 10 million year old δ Lyr cluster.  A
-% $3.6\,R_\oplus$ mini-Neptune transits the G8V primary on a 7.2 day
-% orbit, and it is therefore the youngest planet yet found by the main
-% Kepler mission.  The precision of the Kepler photometry lets us detect
-% two anomalies.  First, unresolved starspot-crossing events distort the
-% average transit shape, and induce transit timing variations that imply
-% the planet is likely on a prograde orbit [9].  Second, although the
-% flare arrival time distribution is most likely Poissonian, there may
-% be hints of an excess of flares at the orbital and synodic periods.
-% The δ Lyr cluster is one of many stellar groups whose nature has been
-% clarified using the Gaia data; many other new clusters are expected to
-% overlap datasets of confirmed and candidate planets.  Further studies
-% that combine Gaia kinematics and TESS rotation periods are therefore
-% well-poised to expand the census of age-dated planets, and our
-% understanding of planetary evolution.
 
@@
 stars, the vast majority of them are several billion years old.  This
-makes it difficult to test origin theories for the different
-families of planets, since many evolutionary processes are expected to
-operate on timescales of less than 100 million years.
+makes it difficult to test origin theories for the different families
+of planets, since many evolutionary processes are expected to operate
+on timescales of less than 100 million years.
 
@@
 %TODO: better citations on radius evolution?
-For instance, the ``mini-Neptunes'', thought to be made of
-molten rocky cores \citep{kite_atmosphere_2020} and extended atmospheric envelopes of hydrogen and
-helium, are expected to shrink in size by factors of several over
-their first $10^8$ years
+For instance, the ``mini-Neptunes'', thought to be made of molten
+rocky cores \citep{kite_atmosphere_2020} and extended atmospheric
+envelopes of hydrogen and helium, are expected to shrink in size by
+factors of several over their first $10^8$ years
 \citep{owen_atmospheres_2016,owen_constraining_2020}.  Specifically,
@@
 dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
-2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
-combination of stellar irradiation and internal heat that powers an
-outflow, which can eventually deplete or entirely strip the envelope
-from the rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
-Discovering young planets, measuring their masses, and detecting their
-atmospheric outflows are key steps toward testing this paradigm, which
-is often invoked to explain the observed radius distribution of mature
+2--4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
+is expected to occur within the first few million years of disk
+dispersal \citep{owen_atmospheres_2016}, the combination of stellar
+irradiation and internal heat may also power a more
+gradual outflow, which can eventually deplete or entirely strip the
+envelope from the rocky core
+\citep{Owen_Wu_2013,ginzburg_corepowered_2018}.  Discovering young
+planets, measuring their masses, and detecting their atmospheric
+outflows are key steps toward testing this paradigm, which is often
+invoked to explain the observed radius distribution of mature
 exoplanets \citep{Fulton_et_al_2017}.
@@
 Comparing our young star list against the Kepler field yielded two
-discoveries.  The first, to be discussed in an upcoming analysis by J.
-Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
+discoveries.  The first, to be discussed in an upcoming analysis by
+J.~Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
 cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
@@
 discovery, and the focus of this work, is that Kepler observed a
-portion of the $\delta$\,Lyr cluster
-(Stephenson-1; Theia~73; \citealt{stephenson_possible_1959}).
-Figure~\ref{fig:skychart} shows the overlap between the Kepler field
-and the cluster.
-
-A previous clustering analysis of the Gaia data reported
-that Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member
-of the \cn\ \citep{kounkel_untangling_2019}.  Given the statistical
+portion of the $\delta$\,Lyr cluster (Stephenson-1; Theia~73;
+\citealt{stephenson_possible_1959}).  Figure~\ref{fig:skychart} shows
+the overlap between the Kepler field and the cluster.
+
+A previous clustering analysis of the Gaia data reported that
+Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member of the
+\cn\ \citep{kounkel_untangling_2019}.  Given the statistical
 validation of its close-in mini-Neptune
 \citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
-we examined the age of the star and its putative
-host cluster most closely (Section~\ref{sec:cluster}).  We find that the $\delta$
-Lyr cluster is $35^{+6}_{-5}$\,Myr old, and that \sn\ is indeed a member
+we examined the age of the star and its putative host cluster more
+closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
+cluster is $35^{+6}_{-5}$\,Myr old, and that \sn\ is indeed a member
 of the cluster.  Focusing on the planet itself
-(Section~\ref{sec:planet}), we confirm that despite a previously
-unreported binary companion, it is indeed a mini-Neptune orbiting the
-G-dwarf primary star.  The presence of a correlation
-between transit timing variations and the local light curve slope
-renders any other explanation implausible.  We conclude by
+(Section~\ref{sec:planet}), we confirm through two independent
+approaches that despite a previously unreported binary companion, it
+is a mini-Neptune orbiting the G-dwarf primary star.  We conclude by
 highlighting some broader implications for our ability to age-date a
@@
     points are kinematically selected members of the $\delta$\,Lyr
-    cluster.  Both Kepler (gray squares) and TESS 
-    observed portions of the cluster.  Naked-eye stars ($m_{\rm
-    V}<6.5$) are shown in gray; four of them (orange points) have
-    their names annotated.  Kepler\,1627 (green star) was observed
-    during the entirety of the Kepler mission, and has been observed
-    by TESS for two lunar months to date.
-    The colors of the other cluster members indicate the number of
-    TESS sectors during which they have been observed.
+    cluster.  Both Kepler (gray squares) and TESS observed portions of
+    the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray;
+    four of them (orange points) have their names annotated.
+    Kepler\,1627 (green star) was observed during the entirety of the
+    Kepler mission, and has been observed by TESS for two lunar months
+    to date.  The colors of the other cluster members indicate the
+    number of TESS sectors during which they have been observed.
     \label{fig:skychart}
@@
 
-\begin{figure*}[t]
+\begin{figure*}[tp]
 	\begin{center}
@@
 	\caption{
-    {\bf The \cn\ is $35\pm15$\,Myr old.}  {\it Top:} Color-magnitude diagram of 
-    the \cn, IC\,2602 ($\approx35$\,Myr), the Pleiades ($\approx 115$\,Myr), and the field.
-    The right panel provides a closer view of stars on the
-    pre-main-sequence.
-    The \cn\ and IC\,2602 are approximately the same isochronal age.
-    {\it Bottom:} TESS and Kepler rotations periods and dereddened Gaia
-    colors, with the Pleiades (125\,Myr)
-    and Praesepe (650\,Myr)
-    shown for reference
-    \citep{rebull_rotation_2016a,douglas_poking_2017}.
+    {\bf The \cn\ is $35^{+6}_{-5}$\,Myr old.}  {\it Top:}
+    Color-magnitude diagram of the \cn, IC\,2602 ($\approx35$\,Myr),
+    the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
+    field (background contours).  The right panel provides a closer
+    view of stars on the pre-main-sequence.  The \cn\ and IC\,2602 are
+    approximately the same isochronal age.  {\it Bottom:} TESS and
+    Kepler stellar rotation period versus dereddened Gaia color, with
+    the Pleiades (115\,Myr) and Praesepe (650\,Myr) shown for
+    reference \citep{rebull_rotation_2016a,douglas_poking_2017}.
     Kepler\,1627 is shown with a yellow star.
-     {\bf todo: caption improve.}.
    \label{fig:age}
@@
 
-% paragraph 1: HR 
-To assess the age of the cluster and our star of interest,
-we used a combination of the isochronal, gyrochronal, and
-lithium-based age-dating.
+To assess the age of the cluster and our star of interest, we used a
+combination of the isochronal, gyrochronal, and lithium-based
+age-dating.
 
@@
 EDR3 Catalog of Nearby Stars \citep{gaia_gcns_2021}.  We cleaned these
-following identical data filtering criteria discussed in
-Appendix~B of \citet{GaiaCollaboration2018}, except that we weakened
-the parallax precision requirement to $\varpi/\sigma_\varpi>5$.  These
-filters were designed to include genuine binaries while omitting
-instrumental artifacts.  We then corrected for extinction by querying
-the 3-dimensional maps of \citet{capitanio_threedimensional_2017} and
+following identical data filtering criteria discussed in Appendix~B of
+\citet{GaiaCollaboration2018}, except that we weakened the parallax
+precision requirement to $\varpi/\sigma_\varpi>5$.  These filters were
+designed to include genuine binaries while omitting instrumental
+artifacts.  We then corrected for extinction by querying the
+3-dimensional maps of \citet{capitanio_threedimensional_2017} and
 \citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}},
@@
 
-%FIXME FIXME TODO TODO get the ages measured
-
-The immediate impression from Figure~\ref{fig:age}
-is that the \cn\ and IC\,2602 overlap, and therefore are approximately the same isochronal
-age.  In our exploration, we also compared against
-$\mu$-Tau ($\approx 60$\,Myr; \citealt{gagne_mutau_2020}) and the
-Upper-Centaurus-Lupus component of the Sco OB2 association (UCL;
-$\approx10$-$15$\,Myr; \citealt{damiani_stellar_2019}).  
-The pre-main-sequence M dwarfs of the \cn\ were intermediate
-between these latter two clusters.  To quantify this into an age measurement
-with an uncertainty, we adopted the empirical procedure
-developed by \citet{gagne_mutau_2020} in their Section~6.  In brief,
-we fitted the pre-main-sequence loci of a set of
-reference clusters, and the locus of the target \cn\ was
-then modeled as a piecewise linear combination of these reference
-clusters.  For our reference clusters, we adopted members of UCL, IC\,2602, and the
-Pleaides from \citet{damiani_stellar_2019} and
-\cite{CantatGaudin2018a} respectively.
-We removed binaries by requiring
-$\texttt{RUWE}<1.3$, $\texttt{radial\_velocity\_error}$ below the
-80$^{\rm th}$ percentile of each cluster's distribution, and by
-manually excluding stars that were obvious photometric binaries in the
-CMD.  We then passed a moving box average and standard deviation
-across the CMD in 0.10\,mag bins, fitted a univariate spline to the
-binned values, and assembled a piecewise grid of hybrid isochrones
-spanning the ages between UCL to the Pleiades following Equations~6
-and~7 from \citet{gagne_mutau_2020}.
+Figure~\ref{fig:age} shows that the \cn\ and IC\,2602 overlap, and are
+therefore approximately the same age.  In our exploration, we also
+compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
+and the Upper-Centaurus-Lupus (UCL) component of the Sco OB2
+association ($\approx$-$16$\,Myr; \citealt{pecaut_star_2016}).  The
+pre-main-sequence M dwarfs of the \cn\ were intermediate between these
+latter two groups.  To turn this heuristic interpolation into an
+age measurement with an uncertainty, we used the empirical procedure
+developed by \citet{gagne_mutau_2020}.  In brief, we fitted the
+pre-main-sequence loci of a set of reference clusters, and the locus
+of the target \cn\ was then modeled as a piecewise linear combination
+of these reference clusters.  For our reference clusters, we adopted
+members of UCL, IC\,2602, and the Pleaides from
+\citet{damiani_stellar_2019} and \cite{CantatGaudin2018a}
+respectively.  We removed binaries by requiring $\texttt{RUWE}<1.3$,
+$\texttt{radial\_velocity\_error}$ below the 80$^{\rm th}$ percentile
+of each cluster's distribution, and by manually excluding stars that
+were obvious photometric binaries in the CMD.  We then passed a moving
+box average and standard deviation across the CMD in 0.10\,mag bins,
+fitted a univariate spline to the binned values, and assembled a
+piecewise grid of hybrid isochrones spanning the ages between UCL to
+the Pleiades following Equations~6 and~7 from
+\citet{gagne_mutau_2020}.
 
 The ages returned by this procedure depend on the ages assumed for
-each reference cluster.
-In our case, the IC\,2602 cluster is most important.
+each reference cluster.  We adopted a 115\,Myr age for the Pleiades,
+and a 16\,Myr age for UCL \citep{dahm_2015,pecaut_star_2016}.
+The age of the IC\,2602 cluster however is the most important
+ingredient, since it receives the most weight in the interpolation.
 As noted by \citet{bouma_cluster_2020}, plausible ages for IC\,2602
-span 30\,Myr to 46\,Myr, with slightly older ages being preferred by the
-lithium-depletion-boundary (LDB) measurements and younger ages by the
-main-sequence turn-off
+span 30\,Myr to 46\,Myr, with slightly older ages being preferred by
+the lithium-depletion-boundary (LDB) measurements and younger ages by
+the main-sequence turn-off
 \citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019}.
-If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would be
-$31^{+5}_{-4}$\,Myr old.
-For the converse case of 46\,Myr, the \cn\ would be
-$44^{+8}_{-7}$\,Myr old.
-We adopt an intermediate 35\,Myr age as our reference, 
-yielding an age for the \cn\ of
+If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would
+be $31^{+5}_{-4}$\,Myr old.  For the converse case of 46\,Myr, the
+\cn\ would be $44^{+8}_{-7}$\,Myr old.  We adopt an intermediate
+35\,Myr age as our reference, yielding an age for the \cn\ of
 $35^{+6}_{-5}$\, Myr, though it should be understood based on the
 systematic uncertainties that an age of up to $\approx50$\,Myr remains
-plausible.  Detailed analyses of the LDB or main-sequence
-turn-off in the \cn\ could help clarify the issue, but both are beyond
-our scope.
+plausible.  Detailed analyses of the LDB or main-sequence turn-off in
+the \cn\ could help clarify the issue, but both are beyond our scope.
 

### 2021-08-12 drivers/calc_empirical_isochrone_age.py

@@ -251,9 +251,9 @@ def collect_isochrone_data():
 
     # NOTE: need to assume hard-coded ages for the reference clusters
     age_dict = {
-        'UCL': 15e6,
+        'UCL': 16e6, # Preibisch & Mamajek 2008, Table 11, UCL
         'IC 2602': 35e6,
-        'Pleiades': 120e6
+        'Pleiades': 112e6 # Dahn 2015
     }
 
     # now make the grid!

### 2021-08-13 paper/ms.tex
update ms

@@
 \newcommand{\nkinwithtessandcrowding}{192} % How many kinematically selected stars have TESS data and meeting the crowding cutoff (nequal<=0, nclose<=1)?
-\newcommand{\nkindefaultcleaning}{157} % How many of these meet the "defaultcleaning" criteria (LSP>0.1, Prot<15, and crowding cut)?
-\newcommand{\nrotgood}{147} % nkindefaultcleaning, minus outliers.
+\newcommand{\nkindefaultcleaning}{145} % How many of these meet the "defaultcleaning" criteria (LSP>0.1, Prot<15, and crowding cut)?
+\newcommand{\nrotgood}{135} % nkindefaultcleaning, minus outliers.
+\newcommand{\nfracprot}{70} % fraction nrotgood/nkinwithtessandcrowding
 
@@
 \citet{kounkel_untangling_2019}, TESS data were available for
-\nwithtess\ stars.  We extracted light curves using the nearest pixel
-to each star, and regressed them against systematics using the causal
-pixel model implemented in the \texttt{unpopular} package
-\citep{hattorio_2021_cpm}.  We then measured candidate rotation
-periods using a Lomb-Scargle periodogram
+\nwithtess\ stars.  \footnote{Kepler rotation periods were derived by
+\citet{mcquillan_rotation_2014} for 56 stars; in these cases, we
+adopted the Kepler rotation period.  } We extracted light curves from
+the TESS images using the nearest pixel to each star, and regressed
+them against systematics with the causal pixel model implemented in
+the \texttt{unpopular} package \citep{hattorio_2021_cpm}.  We then
+measured candidate rotation periods using a Lomb-Scargle periodogram
 \citep{lomb_1976,scargle_studies_1982,astropy_2018}.  To enable cuts
@@
 target star, and with brightness within 1.25 and 2.5 TESS magnitudes
-of the target star.  {\bf We then supplemented these measurements with XXX Kepler
-rotation periods (CITEP MCQUILLAN2014)}.
+of the target star.
 
@@
 combination of automated and manual steps.  As a matter of scope, we
-restricted our attention to the \nkinwithtess\ stars in
-the spatial and kinematic proximity of Kepler\,1627, as discussed in
-Appendix~\ref{app:kinematicselection}.  We also only focused on
-stars for which no companions were known with a brightness exceeding one-tenth of
-the target star within a $21\farcs0$ radius: there were 
+restricted our attention to the \nkinwithtess\ stars in the spatial
+and kinematic proximity of Kepler\,1627, as discussed in
+Appendix~\ref{app:kinematicselection}.  We also only focused on stars
+for which no companions were known with a brightness exceeding
+one-tenth of the target star within a $21\farcs0$ radius: there were
 \nkinwithtessandcrowding\ stars that met these requirements, and that
-had TESS data available.
-For plotting purposes we then imposed a selection based on the strength of the
-rotation period signal itself: we required the Lomb Scargle power to
-exceed 0.1, and the period to be below 15\,days.
+had TESS data available.  For plotting purposes we then imposed a
+selection based on the strength of the rotation period signal itself:
+we required the Lomb Scargle power to exceed 0.2, and the period to be
+below 15\,days.
 
@@
 the ``slow sequence'' of the Pleiades, consistent with a
-gyrochronological age for the \cn\ below 100\,Myr.  Approximately 10 stars 
-appear as outliers above the ``slow sequence''.  Assuming that they
-are all false positives ({\it i.e.}, field interlopers), the rotation
-period detection fraction would be
-\nrotgood/\nkinwithtessandcrowding$\approx 77\%$.
+gyrochronological age for the \cn\ below 100\,Myr.  Approximately 10
+stars appear as outliers above the ``slow sequence''.  Assuming that
+they are all false positives ({\it i.e.}, field interlopers), the
+rotation period detection fraction would be
+\nrotgood/\nkinwithtessandcrowding$\approx \nfracprot\%$.
 
@@
 fitting procedure described in Section~\ref{sec:cmd} is
-$35^{+6}_{-5}$\, Myr.  This procedure hinges on the age adopted
-for IC\,2602, since it is the closest reference cluster in the color-magnitude diagram.
-Gyrochronologically, while most of the G and K
-dwarfs in the \cn\ have not yet converged to form a slow sequence,
-their rotation distribution and that of the F dwarfs is consistent
-with the same age.  The rotation periods of other
-30\,Myr to 50\,Myr clusters ({\it e.g.}, IC\,2602, Collinder 135)
-are indistinguishable
+$35^{+6}_{-5}$\,Myr.  This procedure hinges on the age adopted for
+IC\,2602, since it is the closest reference cluster in the
+color-magnitude diagram.  Gyrochronologically, while most of the G and
+K dwarfs in the \cn\ have not yet converged to a slow sequence, their
+rotation distribution and that of the F dwarfs is consistent with the
+same age.  The rotation periods of other 30\,Myr to 50\,Myr clusters
+({\it e.g.}, IC\,2602 and Collinder 135) are indistinguishable
 (S.~Douglas, priv{.}~comm{.}).  The most precise age for \sn\ and its
-planet are therefore also $35^{+6}_{-5}$\, Myr, based on the star's spatial,
-kinematic, and rotational connection to this larger ensemble, and the assumption that
-\pn\ formed very quickly after the star itself.
-
-% TODO:  update this based on HIRES instead of TRES
-% TODO:  cite Stephanie instead of priv comm? Send her your Prot
-Beyond the kinematic and rotational lines of evidence for the youth of \sn,
-there is a final consistency check: the abundance of Li\,\textsc{I} as
-inferred from the 6708\,\AA\ doublet.  This connection of this element
-to age measurements for young stars has been reviewed by
-\citet{soderblom_ages_2014}.  Following the equivalent width
-measurement procedure described by \citet{bouma_2021_ngc2516}, we find
-${\rm EW}_{\rm Li} = 232^{+23}_{-13}\,$m\AA, which does not
-correct for the Fe\,\textsc{I} blend at 6707.44\AA\.  Given $T_{\rm eff} \approx
-5500\,{\rm K}$, this measurement is also in agreement with an age for
-the star between 30\,Myr and 50\,Myr (see {\it e.g.}, Figure~9 of
-\citealt{bouma_2021_ngc2516}, or the data
-from \citealt{randich_gaiaeso_2018}).
+planet are therefore also $35^{+6}_{-5}$\, Myr, based on the star's
+spatial, kinematic, and rotational connection to this larger ensemble,
+and the assumption that the planet formed very quickly after the star.
+
+Beyond the kinematic and rotational lines of evidence for the youth of
+\sn, there is a final consistency check: the abundance of
+Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet.  The utility
+of this element for age dating young stars has been reviewed by
+\citet{soderblom_ages_2014}.  We acquired an iodine-free spectrum from
+Keck/HIRES on the night of 26 March 2021 using the standard setup and
+reduction techniques of the California Planet Survey
+\citep{howard_cps_2010}.  Following the equivalent width measurement
+procedure described by \citet{bouma_2021_ngc2516}, we measure ${\rm
+EW}_{\rm Li} = 233^{+5}_{-7}\,$m\AA.   Note that this value does not
+correct for the Fe\,\textsc{I} blend at 6707.44\AA\.  Given $T_{\rm
+eff} \approx 5500\,{\rm K}$, this measurement is also in agreement
+with an age for the star between 30\,Myr and 50\,Myr (see {\it e.g.},
+Figure~9 of \citealt{bouma_2021_ngc2516}, or the data from
+\citealt{randich_gaiaeso_2018}).
 

### 2021-08-13 paper/ms.tex
update ms

@@
 
-\newcommand{\cn}{$\delta$~Lyr\ cluster} % cluster name
+\newcommand{\cn}{$\delta$\ Lyr\ cluster} % cluster name
 \newcommand{\sn}{Kepler\,1627} % star system name (binary)
@@
 \citep{owen_atmospheres_2016,owen_constraining_2020}.  Specifically,
-they start with sizes of 4--12\,$R_\oplus$ near the time of disk
+they start with sizes of 4--12\,$R_\oplus$ shortly after the time of disk
 dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
 2--4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
-is expected to occur within the first few million years of disk
-dispersal \citep{owen_atmospheres_2016}, the combination of stellar
+is expected to occur within the first few million years of the disk's pressure
+support being removed \citep{owen_atmospheres_2016}, the combination of stellar
 irradiation and internal heat may also power a more
-gradual outflow, which can eventually deplete or entirely strip the
+gradual outflow which can eventually deplete or entirely strip the
 envelope from the rocky core
@@
 Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member of the
-\cn\ \citep{kounkel_untangling_2019}.  Given the statistical
-validation of its close-in mini-Neptune
+\cn\ \citep{kounkel_untangling_2019}.  Given the previous statistical
+validation the close-in mini-Neptune Kepler 1627b
 \citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
@@
 cluster is $35^{+6}_{-5}$\,Myr old, and that \sn\ is indeed a member
-of the cluster.  Focusing on the planet itself
-(Section~\ref{sec:planet}), we confirm through two independent
-approaches that despite a previously unreported binary companion, it
-is a mini-Neptune orbiting the G-dwarf primary star.  We conclude by
-highlighting some broader implications for our ability to age-date a
+of the cluster.  Focusing on the planet itself,
+we confirm that despite the presence of a previously unreported binary companion,
+the planet orbits the G-dwarf primary star on an orbit that appears aligned
+with the stellar rotation axis
+(Section~\ref{sec:planet}).  We conclude by
+highlighting broader implications for our ability to age-date a
 statistical sample of planets (Section~\ref{sec:conc}).
@@
     the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray;
-    four of them (orange points) have their names annotated.
+    four of them (orange circles) have their names annotated.
     Kepler\,1627 (green star) was observed during the entirety of the
@@
     the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
-    field (background contours).  The right panel provides a closer
+    field (gray).  The right panel provides a zoomed
     view of stars on the pre-main-sequence.  The \cn\ and IC\,2602 are
     approximately the same isochronal age.  {\it Bottom:} TESS and
-    Kepler stellar rotation period versus dereddened Gaia color, with
-    the Pleiades (115\,Myr) and Praesepe (650\,Myr) shown for
+    Kepler stellar rotation periods versus dereddened Gaia colors, with
+    the Pleiades and Praesepe (650\,Myr) shown for
     reference \citep{rebull_rotation_2016a,douglas_poking_2017}.
-    Kepler\,1627 is shown with a yellow star.
+    The \cn\ is gyrochronally younger than the Pleiades.
    \label{fig:age}
@@
 To assess the age of the cluster and our star of interest, we used a
-combination of the isochronal, gyrochronal, and lithium-based
+combination of isochronal, gyrochronal, and lithium-based
 age-dating.
@@
 computed by \citet{GaiaCollaboration2018} while assuming that $A_0 =
-3.1 E(B-V)$.  For IC\,2602, the Pleaides, and the \cn, this procedure
+3.1 E(B-V)$.  For IC\,2602, the Pleiades, and the \cn, this procedure
 yielded a respective mean and standard deviation for the reddening of
@@
 and the Upper-Centaurus-Lupus (UCL) component of the Sco OB2
-association ($\approx$-$16$\,Myr; \citealt{pecaut_star_2016}).  The
+association ($\approx$$16$\,Myr; \citealt{pecaut_star_2016}).  The
 pre-main-sequence M dwarfs of the \cn\ were intermediate between these
 latter two groups.  To turn this heuristic interpolation into an
-age measurement with an uncertainty, we used the empirical procedure
+age measurement, we used the empirical procedure
 developed by \citet{gagne_mutau_2020}.  In brief, we fitted the
@@
 of these reference clusters.  For our reference clusters, we adopted
-members of UCL, IC\,2602, and the Pleaides from
+members of UCL, IC\,2602, and the Pleiades from
 \citet{damiani_stellar_2019} and \cite{CantatGaudin2018a}
@@
 and a 16\,Myr age for UCL \citep{dahm_2015,pecaut_star_2016}.
-The age of the IC\,2602 cluster however is the most important
+The age of IC\,2602 however is the most important
 ingredient, since it receives the most weight in the interpolation.
-As noted by \citet{bouma_cluster_2020}, plausible ages for IC\,2602
+Plausible ages for IC\,2602
 span 30\,Myr to 46\,Myr, with slightly older ages being preferred by
@@
 the main-sequence turn-off
-\citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019}.
+\citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019,bouma_cluster_2020}.
 If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would
@@
 plausible.  Detailed analyses of the LDB or main-sequence turn-off in
-the \cn\ could help clarify the issue, but both are beyond our scope.
+the \cn\ could help resolve the issue, but both are beyond our scope.
 
@@
 \citet{kounkel_untangling_2019}, TESS data were available for
-\nwithtess\ stars.  \footnote{Kepler rotation periods were derived by
+\nwithtess\ stars.\footnote{Kepler rotation periods were derived by
 \citet{mcquillan_rotation_2014} for 56 stars; in these cases, we
@@
 rotation period detection fraction would be
-\nrotgood/\nkinwithtessandcrowding$\approx \nfracprot\%$.
+$\nrotgood/\nkinwithtessandcrowding \approx \nfracprot\%$.
 
 Binarity is known to be an important factor in the density of stars
-falling underneath the ``slow sequence''
+falling underneath the slow sequence
 \citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}.
@@
 
-To summarize, the age of the \cn\ cluster based on the empirical
+In brief , the age of the \cn\ cluster based on the empirical
 fitting procedure described in Section~\ref{sec:cmd} is
@@
 same age.  The rotation periods of other 30\,Myr to 50\,Myr clusters
-({\it e.g.}, IC\,2602 and Collinder 135) are indistinguishable
-(S.~Douglas, priv{.}~comm{.}).  The most precise age for \sn\ and its
+({\it e.g.}, IC\,2602 and IC\,2391) are indistinguishable
+\citep{douglas_stephanie_t_2021_5131306}.  The most precise age for \sn\ and its
 planet are therefore also $35^{+6}_{-5}$\, Myr, based on the star's
@@
 
+%TODO:  check Teff here once you have it.
 Beyond the kinematic and rotational lines of evidence for the youth of
@@
 EW}_{\rm Li} = 233^{+5}_{-7}\,$m\AA.   Note that this value does not
-correct for the Fe\,\textsc{I} blend at 6707.44\AA\.  Given $T_{\rm
+correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given $T_{\rm
 eff} \approx 5500\,{\rm K}$, this measurement is also in agreement
@@
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f11.pdf}
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f11.pdf}
+		}
+
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f16.pdf}
+		}
 	\end{center}
@@
 	\caption{
-		{\bf Evidence for a prograde orbit from transit timing variations.}
-    The time of each transit was measured, along with the local slope
+		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
+    {\it Top:} 
+    The time of each Kepler transit was measured, along with the local slope
     of the light curve.  The two quantities appear anti-correlated,
@@
     rotation period ($\sim$1.3\,days).
-		\label{fig:ttvslope}
+        {\it Bottom:} A Keck/HIRES transit was observed on 7 Aug 2021 UT,
+    simultaneous with MUSCAT3.  A quadratic baseline trend was allowed
+    in the fit.
+    The resulting best-fit sky-projected obliquity is listed in red.
+    \label{fig:obliquity}
 	}
@@
 
+%TODO: fix sizes and numbers
 If the \pn\ transit signal is created by a bonafide planet,  then to
 our knowledge its age would make it the youngest planet yet found by
-the main Kepler mission\footnote{The re-purposed K2 mission found two
+the main Kepler mission.\footnote{The re-purposed K2 mission found two
 slightly younger systems: K2-33b
 \citep{David_et_al_2017,Mann_K2_33b_2016} and V1298 Tau
-\citep{david_four_2019}}.  A portion of the light curve is shown with
+\citep{david_four_2019}}  A portion of the light curve is shown with
 the phase-folded transit in Figure~\ref{fig:lc}.  The dominant signal
@@
 the latter scenario to be the only possible case, with a
-(model-dependent) probability of $\approx10^{-5}$.  While this
-statistical validation is reassuring, ``validated'' planets have been
-shown to be false positives \citep{shporer_three_2017}.  Most
-failures of the statistical validation approach arise when attempting
+(model-dependent) probability of $\approx10^{-5}$. Most
+reported failures of the statistical validation approach have arisen when attempting
 to separate the smallest stars from identically-sized gas giant
-planets.  With a size of $\approx 3.6\,R_\oplus$, Kepler\,1627b does
+planets \citep[{\it e.g.},][]{shporer_three_2017}.  With a size of $\approx 3.5\,R_\oplus$, Kepler\,1627b does
 not meet this concern.  The fact that we detect a low-mass stellar
@@
 {\bf todo: LGB to check numbers on full-dataset transit ecc and b} 
-Two additional new lines of evidence confirm the planetary
+Three additional new lines of evidence confirm the planetary
 interpretation.  First, the $\approx$100 days of 1-minute
@@
 periodic orbit.  This residual, the ``transit timing variation''
-(TTV), is plotted in Figure~\ref{fig:ttvslope} against the local
+(TTV), is plotted in Figure~\ref{fig:obliquity} against the local
 linear slope.  Only a handful of Kepler Objects of Interest have shown
@@
 interpreted as a TTV induced by unresolved starspot crossings
-\citep{mazeh_time_2015}.  This is of course only possible if the
-planet is transiting the primary star, which excludes the any of the
+\citep{mazeh_time_2015}.  This is only possible if the
+planet transits the primary star, which excludes the
 aforementioned false positive scenarios.  It also implies that the
@@
 \citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
+
+The final line of evidence confirming \pn\ comes from a spectroscopic
+transit we measured from Keck/HIRES during the night of 7 Aug 2021
+(UT) using the iodine cell for the wavelength calibration.  We
+scheduled the transit using the ephemeris of
+\citet{holczer_transit_2016}, which we verified using the available
+TESS data.  We also acquired simultaneous multi-band photometry using
+MUSCAT3, to help interpret the radial velocities \citep{Narita_2020}.
+The lower panels in Figure~\ref{fig:obliquity} show the results.  The
+photometry shows the expected starspot trend, along with the transit
+and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
+The spectroscopic data show significant scatter, though a significant
+increase in radial velocity is seen shortly after the expected
+ingress.  We fitted the RVs using the \citet{hirano_analytic_2010} and
+\citet{hirano_2011} models for the Rossiter-McLaughlin effect, and
+considered the cases of a linear and quadratic trend.
+% TODO finish this paragraph
+We found $\chi^2_{\rm red} = X.XX$ for the best-fitting linear case, 
+and
+$\chi^2_{\rm red} = X.XX$ for the best-fitting quadratic, with a
+$\Delta {\rm BIC} = Y.YY$ between the two, providing {\bf (very)
+strong} evidence to prefer the latter.
 

### 2021-08-13 rudolf/plotting.py

@@ -1669,7 +1669,7 @@ def plot_hr(
         if cleanhrcut:
             _df = _df[get_clean_gaia_photometric_sources(_df)]
         if reddening_corr:
-            print('Pleaides')
+            print('Pleiades')
             print(_df['reddening[mag][stilism]'].describe())
 
         ax.scatter(
@@ -2017,7 +2017,7 @@ def plot_rotationperiod_vs_color(outdir, runid, yscale='linear', cleaning=None,
         l = '$\delta$ Lyr cluster'
     else:
         raise NotImplementedError
-    labels = ['Pleaides', 'Praesepe', l]
+    labels = ['Pleiades', 'Praesepe', l]
 
     # plot vals
     for _cls, _col, z, m, l, _lw, s, mew in zip(

### 2021-08-13 paper/ms.tex
update ms

@@
   $3.5\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
-  observed by the Kepler satellite because it is nearby ($d\approx
-  333\,{\rm pc}$) and it resembles the Sun.  Combining kinematic
-  information from Gaia, stellar rotation periods from TESS, and Li
-  abundances from ground-based spectra, we show that Kepler\,1627 is a
-  member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr cluster.  To our
-  knowledge, Kepler\,1627Ab is the youngest planet with a precise age
-  yet found by the main Kepler mission.  From both unresolved
-  starspot-crossing events in the Kepler photometry and from a
-  ground-based Keck/HIRES transit we find that the planet is likely on
-  an orbit nearly aligned with the stellar spin axis.  We also find
-  from the short-cadence Kepler photometry that although the flare
-  arrival time distribution is probably Poissonian, there may be an
-  excess of flares separated by the orbital and synodic periods, and
-  linear combinations thereof.  The $\delta$~Lyr cluster is one of
+  observed by the Kepler satellite because it is nearby
+  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Combining
+  kinematic information from Gaia, stellar rotation periods from TESS,
+  and Li abundances from ground-based spectra, we show that
+  Kepler\,1627 is a member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr
+  cluster.  To our knowledge, Kepler\,1627Ab is the youngest planet
+  with a precise age yet found by the main Kepler mission.  From both
+  unresolved starspot-crossing events in the Kepler photometry and
+  from a ground-based Keck/HIRES transit we find that the planet is
+  likely on an orbit nearly aligned with the stellar spin axis.  We
+  also find from the short-cadence Kepler photometry that although the
+  flare arrival time distribution is probably Poissonian, there may be
+  an excess of flares separated by the orbital and synodic periods,
+  and linear combinations thereof.  The $\delta$~Lyr cluster is one of
   many stellar groups whose properties have been clarified using the
@@
   confirmed and candidate exoplanets.  Further studies that combine
-  Gaia kinematics and TESS rotation periods are poised to expand
-  the census of age-dated planets, and our understanding of planetary
+  Gaia kinematics and TESS rotation periods are poised to expand the
+  census of age-dated planets, and our understanding of planetary
   evolution.
@@
 
-In brief , the age of the \cn\ cluster based on the empirical
+In brief, the age of the \cn\ cluster based on the empirical
 fitting procedure described in Section~\ref{sec:cmd} is
@@
 
-%TODO:  check Teff here once you have it.
 Beyond the kinematic and rotational lines of evidence for the youth of
@@
 EW}_{\rm Li} = 233^{+5}_{-7}\,$m\AA.   Note that this value does not
-correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given $T_{\rm
-eff} \approx 5500\,{\rm K}$, this measurement is also in agreement
-with an age for the star between 30\,Myr and 50\,Myr (see {\it e.g.},
-Figure~9 of \citealt{bouma_2021_ngc2516}, or the data from
+correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given the stellar
+parameters of \sn\ (Table~\ref{tab:starparams}), most notably its
+effective temperature, this measurement is also in agreement with an
+age for the star between 30\,Myr and 50\,Myr (see {\it e.g.}, Figure~9
+of \citealt{bouma_2021_ngc2516}, or the data from
 \citealt{randich_gaiaeso_2018}).
@@
     rotation period ($\sim$1.3\,days).
-        {\it Bottom:} A Keck/HIRES transit was observed on 7 Aug 2021 UT,
-    simultaneous with MUSCAT3.  A quadratic baseline trend was allowed
+        {\it Bottom:} Keck/HIRES transit of 7 Aug 2021 UT,
+    observed simultaneously with MUSCAT3.  A quadratic baseline trend was allowed
     in the fit.
@@
 to separate the smallest stars from identically-sized gas giant
-planets \citep[{\it e.g.},][]{shporer_three_2017}.  With a size of $\approx 3.5\,R_\oplus$, Kepler\,1627b does
-not meet this concern.  The fact that we detect a low-mass stellar
-companion in archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does
-not change the conclusion: the companion star contributes 1\% of the
-flux observed by Kepler.  Combined with the observed transit depth and
-shape (Appendix~\ref{app:companionstar}), a planet transiting the
-primary star is the only plausible interpretation of the
-data.
+planets \citep[{\it e.g.},][]{shporer_three_2017}.  With a size of
+$\approx 3.5\,R_\oplus$, Kepler\,1627b does not meet this concern.
+The fact that we detect a low-mass stellar companion in archival
+Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does not change the
+conclusion: the companion star contributes 1\% of the flux observed by
+Kepler.  Combined with the observed transit depth and shape
+(Appendix~\ref{app:companionstar}), a planet transiting the primary
+star is the only plausible interpretation of the data.
 
@@
 
-{\bf todo: Kento to check numbers on per-transit analysis} 
+% TODO: Kento Masuda, please verify
 The second line of evidence comes from our transit timing analysis. We
@@
 (TTV), is plotted in Figure~\ref{fig:obliquity} against the local
-linear slope.  Only a handful of Kepler Objects of Interest have shown
-this correlation \citep{holczer_time_2015}, which is most readily
-interpreted as a TTV induced by unresolved starspot crossings
-\citep{mazeh_time_2015}.  This is only possible if the
-planet transits the primary star, which excludes the
-aforementioned false positive scenarios.  It also implies that the
-planet's orbit is prograde.  The latter interpretation assumes that
-the dominant photometric variability is induced by dark spots, rather
-than bright faculae.  Given the observed transition of Sun-like
-stellar variability from spot to faculae-dominated regimes, we expect
-this inference to be reasonably secure
+linear slope.  A significant negative slope of $-0.0564 \pm 0.0098\
+{\rm ppt\,day}^{-1}$ is observed.  Only a handful of Kepler Objects of
+Interest have shown this correlation \citep{holczer_time_2015}, which
+is most readily interpreted as a TTV induced by unresolved starspot
+crossings \citep{mazeh_time_2015}.  This is only possible if the
+planet transits the primary star, which excludes the aforementioned
+false positive scenarios.  It also implies that the planet's orbit is
+prograde.  The latter interpretation assumes that the dominant
+photometric variability is induced by dark spots, rather than bright
+faculae.  Given the observed transition of Sun-like stellar
+variability from spot to faculae-dominated regimes, we expect this
+inference to be reasonably secure
 \citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
@@
 transit we measured from Keck/HIRES during the night of 7 Aug 2021
-(UT) using the iodine cell for the wavelength calibration.  We
+(UT), using the iodine cell for the wavelength calibration.  We
 scheduled the transit using the ephemeris of
@@
 approach would be to identify kinematic associations around exoplanet
-host stars using the Gaia data, and to then verify these associations
+host stars using tangential velocities from Gaia, and to then verify these associations
 with ancilliary rotation periods and spectroscopy
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{model_posterior_table.tex}
 
@@
 \citet{kerr_stars_2021} membership list prior to its publication.
+L.G.B{.}, L.H{.}, and N.N{.} are also grateful to K{.}~Collins for
+helping resolve the scheduling conflict that would have otherwise
+prevented the MUSCAT3 observations.
 % is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
@@
 %
-This study was based in part on observations at Cerro Tololo
-Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
-2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
-Association of Universities for Research in Astronomy (AURA) under a
-cooperative agreement with the National Science Foundation.
+%This study was based in part on observations at Cerro Tololo
+%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
+%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
+%Association of Universities for Research in Astronomy (AURA) under a
+%cooperative agreement with the National Science Foundation.
 %
@@
 Aeronautics and Space Administration.
+We also thank
+the Keck Observatory staff, particularly {\bf A.~Doe},
+for support of HIRES and remote observing.  We recognize
+the importance
+that the summit of Maunakea has within the indigenous
+Hawaiian community, and are deeply grateful to
+have the opportunity to conduct observations from this
+mountain.
 %
@@
 
+\input{starparams.tex}
+\input{model_posterior_table.tex}
+
 \clearpage
@@
 		\leavevmode
-		\includegraphics[width=0.6\textwidth]{TEMP_binarity_ao.png}
+		\includegraphics[width=0.6\textwidth]{Kepler1627_20210624_562_832_final.pdf}
 	\end{center}

### 2021-08-13 rudolf/plotting.py

@@ -1273,7 +1273,7 @@ def _linear_model(xdata, m, b):
 def plot_ttv_vs_local_slope(outdir):
 
     # get data
-    datadir = os.path.join(DATADIR, 'ttv', 'Bouma', 'K05245.01/')
+    datadir = os.path.join(DATADIR, 'ttv', 'Masuda', 'K05245.01/')
     tnum, tc, tc_err, tc_lin = np.loadtxt(datadir+"tc.dat").T
 
     ttv = tc - tc_lin
@@ -1301,7 +1301,7 @@ def plot_ttv_vs_local_slope(outdir):
     # ignore the error bars... because they are probably wrong.
     p_opt, p_cov = curve_fit(
         _linear_model, slopes[sel], ttv[sel],
-        p0=(slope_guess, intercept_guess)
+        p0=(slope_guess, intercept_guess), #sigma=tc_err[sel]
     )
     lsfit_slope = p_opt[0]
     lsfit_slope_err = p_cov[0,0]**0.5

### 2021-08-13 paper/ms.tex
update RM model comparison

@@
 
+% from /drivers/logs/rm.log
+% quad: 
+% chi2red 4.45
+% BIC 272.4
+%
+% linear:
+% chi2red 17.1
+% BIC 291.8
+%
+% trendonly:
+% chi2red 13.8
+% BIC 271.3
+
 The final line of evidence confirming \pn\ comes from a spectroscopic
@@
 and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
-The spectroscopic data show significant scatter, though a significant
-increase in radial velocity is seen shortly after the expected
-ingress.  We fitted the RVs using the \citet{hirano_analytic_2010} and
-\citet{hirano_2011} models for the Rossiter-McLaughlin effect, and
-considered the cases of a linear and quadratic trend.
-% TODO finish this paragraph
-We found $\chi^2_{\rm red} = X.XX$ for the best-fitting linear case, 
-and
-$\chi^2_{\rm red} = X.XX$ for the best-fitting quadratic, with a
-$\Delta {\rm BIC} = Y.YY$ between the two, providing {\bf (very)
-strong} evidence to prefer the latter.
+The radial velocities show significant scatter, though a significant
+increase in RV is seen shortly after the expected ingress.  The
+internal uncertainties on the RVs ($\approx 15\,{\rm m\,s}^{-1}$)is
+dominated by the $v\sin i$ broadening.
+
+We fitted the RVs using the \citet{hirano_analytic_2010,hirano_2011}
+models for the Rossiter-McLaughlin (RM) effect, and allowed for a
+linear and optionally quadratic trend, along with a white-noise jitter
+term.  We found $\chi^2_{\rm red} = 17.1$ for the best-fitting linear
+case, and $\chi^2_{\rm red} = 16.7$ for the best-fitting quadratic,
+with a $\Delta {\rm BIC} = {\rm BIC}_{\rm lin} - {\rm BIC}_{\rm quad}
+= 19.3$.  Nominally, this provides strong statistical evidence to
+prefer the quadratic model over the linear one.  However the
+$\chi^2_{\rm red}$ indicate that neither fit is particularly good.
+Considering a quadratic trend alone without any RM effect yields
+$\chi^2_{\rm red} = 13.8$, and $\Delta {\rm BIC} = {\rm BIC}_{\rm
+trend} - {\rm BIC}_{\rm quad} = 1.2$; from a model-comparison
+standpoint, there is no statistical justification to prefer the
+best-fitting ``RM + quadratic trend'' model over the best-fitting
+``quadratic trend'' model, or vice-versa.  Given the expected
+semi-amplitude of the velocity anomaly ($\Delta v_{\rm RM}\approx
+20\,{\rm m\,s}^{-1}$) and the achieved internal precision, the claim
+of having ``detected the Doppler transit'' is perhaps not yet
+justified.  Observationally however, we know from the MUSCAT3 data
+that the transit was observed, and the increase in radial velocity
+seen shortly after ingress also seems consistent with a prograde or
+polar orbit, and not a retrograde one.  More precise data will likely
+be required to make a more precise measurement of the obliquity.
 

### 2021-08-13 drivers/plot_RM.py

@@ -6,6 +6,8 @@ PLOTDIR = os.path.join(RESULTSDIR, 'RM')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
-rp.plot_RM(PLOTDIR)
+models = ['quadratic','linear','trendonly']
 
-rp.plot_RM_and_phot(PLOTDIR)
+for model in models:
+    rp.plot_RM(PLOTDIR, model=model)
+    rp.plot_RM_and_phot(PLOTDIR, model=model)

### 2021-08-14 paper/ms.tex
update RM

--- 2021-08-13 paper/ms.tex
@@
   unresolved starspot-crossing events in the Kepler photometry and
-  from a ground-based Keck/HIRES transit we find that the planet is
+  from a Keck/HIRES transit we find that the planet is
   likely on an orbit nearly aligned with the stellar spin axis.  We
@@
     rotation period ($\sim$1.3\,days).
-        {\it Bottom:} Keck/HIRES transit of 7 Aug 2021 UT,
-    observed simultaneously with MUSCAT3.  A quadratic baseline trend was allowed
-    in the fit.
-    The resulting best-fit sky-projected obliquity is listed in red.
+    {\it Bottom:} Keck/HIRES transit of 7 Aug 2021 UT, observed
+    simultaneously with MUSCAT3.  A quadratic baseline trend was
+    allowed in the fit.  The best-fit sky-projected obliquity in this
+    model is listed in red; the detection however has marginal
+    statistical significance compared to a ``trend-only'' model.
     \label{fig:obliquity}
@@
 (3-$\sigma$) from the 30-minute data.
+This implies an even stronger constraint on the transit shape in an
+analysis of the false positive probability.
 
@@
 and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
-The spectroscopic data show significant scatter, though a significant
-increase in radial velocity is seen shortly after the expected
-ingress.  We fitted the RVs using the \citet{hirano_analytic_2010} and
-\citet{hirano_2011} models for the Rossiter-McLaughlin effect, and
-considered the cases of a linear and quadratic trend.
-% TODO finish this paragraph
-We found $\chi^2_{\rm red} = X.XX$ for the best-fitting linear case, 
-and
-$\chi^2_{\rm red} = X.XX$ for the best-fitting quadratic, with a
-$\Delta {\rm BIC} = Y.YY$ between the two, providing {\bf (very)
-strong} evidence to prefer the latter.
+The radial velocities show significant scatter, though a significant
+increase in RV is seen shortly after the expected ingress.  The
+internal uncertainties on the RVs ($\approx 15\,{\rm m\,s}^{-1}$) is
+dominated by the $v\sin i$ broadening.
+
+% from /drivers/logs/rm.log
+% quad: 
+% chi2red: 4.452
+% BIC: 272.434
+% AIC: 264.796
+%
+% linear:
+% chi2red: 5.425
+% BIC: 291.987
+% AIC: 285.441
+%
+% trendonly:
+% chi2red: 4.256
+% BIC: 274.383
+% AIC: 270.018
+%
+% retrograde:
+% chi2red: 5.439
+% BIC: 292.415
+% AIC: 285.869
+% 
+% prograde:
+% chi2red: 4.320
+% BIC: 272.039
+% AIC: 265.493
+
+We fitted the RVs using the \citet{hirano_analytic_2010,hirano_2011}
+models for the Rossiter-McLaughlin (RM) effect, which we parametrized
+with three parameters: the sky-projected obliquity, the stellar
+equatorial velocity, and the gaussian dispersion of spectral lines.
+We then allowed for a linear and optionally quadratic trend, along
+with a white-noise jitter term.  The quadratic model with the RM
+effect is shown in Figure~\ref{fig:obliquity}.  We found $\chi^2_{\rm
+red} = 5.4$ for the best-fitting linear case, and $\chi^2_{\rm red} =
+4.4$ for the best-fitting quadratic, with a $\Delta {\rm BIC} = {\rm
+BIC}_{\rm lin} - {\rm BIC}_{\rm quad} = 19.6$ and $\Delta {\rm AIC} =
+{\rm AIC}_{\rm lin} - {\rm AIC}_{\rm quad} = 20.6$.  While this
+provides strong statistical evidence to prefer the quadratic model
+over the linear one, the $\chi^2_{\rm red}$ indicate that neither fit
+is particularly good.  Considering a quadratic trend alone without any
+RM effect yields $\chi^2_{\rm red} = 4.3$, and $\Delta {\rm BIC} =
+{\rm BIC}_{\rm trend} - {\rm BIC}_{\rm quad} = 1.9$ (similarly $\Delta
+{\rm AIC}=5.2$); from a model-comparison standpoint, there is only
+weak statistical evidence to prefer the best-fitting ``RM and
+quadratic trend'' model over the best-fitting ``quadratic trend only''
+model.  Given the expected semi-amplitude of the velocity anomaly
+($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) and the actual achieved
+precision, the claim of having ``detected the Doppler
+transit'' is therefore marginal.  Observationally however, we
+know from the MUSCAT3 data that the transit was observed, and the
+increase in radial velocity observed shortly after ingress is more
+consistent with a prograde or polar orbit than a retrograde one.
+Fixing the orbit to be either exactly prograde or exactly retrograde
+and allowing for a quadratic trend, we find $\Delta {\rm BIC} = {\rm
+BIC}_{\rm retrograde} - {\rm BIC}_{\rm prograde} = 20.4$; the
+retrograde case is ruled out on this basis.  More precise data will
+however still be useful to make a more precise obliquity measurement.
 

### 2021-08-16 paper/ms.tex
update ms

@@
 \citep{howard_cps_2010}.  Following the equivalent width measurement
-procedure described by \citet{bouma_2021_ngc2516}, we measure ${\rm
+procedure described by \citet{bouma_2021_ngc2516}, we find ${\rm
 EW}_{\rm Li} = 233^{+5}_{-7}\,$m\AA.   Note that this value does not
@@
     rotation period ($\sim$1.3\,days).
-    {\it Bottom:} Keck/HIRES transit of 7 Aug 2021 UT, observed
-    simultaneously with MUSCAT3.  A quadratic baseline trend was
-    allowed in the fit.  The best-fit sky-projected obliquity in this
+    {\it Bottom:} Radial velocities from
+    Keck/HIRES during a transit on 7 Aug 2021 UT, observed
+    simultaneously with MUSCAT3 ({\it griz} photometry).  A quadratic baseline trend was
+    allowed in the RV fit.  The best-fit sky-projected obliquity in this
     model is listed in red; the detection however has marginal
@@
 We fitted the RVs using the \citet{hirano_analytic_2010,hirano_2011}
-models for the Rossiter-McLaughlin (RM) effect, which we parametrized
-with three parameters: the sky-projected obliquity, the stellar
-equatorial velocity, and the gaussian dispersion of spectral lines.
-We then allowed for a linear and optionally quadratic trend, along
-with a white-noise jitter term.  The quadratic model with the RM
+models for the Rossiter-McLaughlin (RM) effect, and allowed three parameters
+to vary: the sky-projected obliquity, the stellar
+equatorial velocity, and the Gaussian dispersion of spectral lines.
+We also allowed for an optional linear and quadratic trend, and
+a white-noise jitter term.  The quadratic model with the RM
 effect is shown in Figure~\ref{fig:obliquity}.  We found $\chi^2_{\rm
@@
 4.4$ for the best-fitting quadratic, with a $\Delta {\rm BIC} = {\rm
-BIC}_{\rm lin} - {\rm BIC}_{\rm quad} = 19.6$ and $\Delta {\rm AIC} =
-{\rm AIC}_{\rm lin} - {\rm AIC}_{\rm quad} = 20.6$.  While this
+BIC}_{\rm line} - {\rm BIC}_{\rm quad} = 19.6$ and $\Delta {\rm AIC} =
+{\rm AIC}_{\rm line} - {\rm AIC}_{\rm quad} = 20.6$.  While this
 provides strong statistical evidence to prefer the quadratic model
@@
 {\rm BIC}_{\rm trend} - {\rm BIC}_{\rm quad} = 1.9$ (similarly $\Delta
-{\rm AIC}=5.2$); from a model-comparison standpoint, there is only
+{\rm AIC}=5.2$); from a model-comparison standpoint, there is therefore only
 weak statistical evidence to prefer the best-fitting ``RM and
@@
 ($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) and the actual achieved
-precision, the claim of having ``detected the Doppler
-transit'' is therefore marginal.  Observationally however, we
-know from the MUSCAT3 data that the transit was observed, and the
-increase in radial velocity observed shortly after ingress is more
+precision, the claim of having detected the Doppler
+transit is therefore marginal.  
+From the MUSCAT3 data however, we do know that the transit was observed, and the
+increase in radial velocity seen shortly after ingress is more
 consistent with a prograde or polar orbit than a retrograde one.

### 2021-08-24 paper/ms.tex
update ms

@@
 
+\newcommand{\clusterage}{$38^{+6}_{-5}$\,Myr} %
+
 % gaia target sample numbers
@@
 \title{
-  An Adolescent Mini-Neptune in the Kepler Field
+  %Kepler 1627Ab: A 38 Million Year Old
+  Kepler 1627 Ab: A 38$\,$Myr Mini-Neptune in the Kepler Field
 }
@@
   and Li abundances from ground-based spectra, we show that
-  Kepler\,1627 is a member of the $35^{+6}_{-5}$\,Myr old $\delta$~Lyr
+  Kepler\,1627 is a member of the \clusterage\ old $\delta$~Lyr
   cluster.  To our knowledge, Kepler\,1627Ab is the youngest planet
@@
   many stellar groups whose properties have been clarified using the
-  Gaia data; many of the other new clusters also overlap with
-  confirmed and candidate exoplanets.  Further studies that combine
-  Gaia kinematics and TESS rotation periods are poised to expand the
-  census of age-dated planets, and our understanding of planetary
-  evolution.
+  Gaia data; many other new clusters also overlap with confirmed and
+  candidate exoplanets.  Further studies that combine Gaia kinematics
+  and TESS rotation periods are poised to expand the census of
+  age-dated planets, and our understanding of planetary evolution.
 \end{abstract}
@@
 closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
-cluster is $35^{+6}_{-5}$\,Myr old, and that \sn\ is indeed a member
+cluster is \clusterage\ old, and that \sn\ is indeed a member
 of the cluster.  Focusing on the planet itself,
@@
 	\caption{
-    {\bf The \cn\ is $35^{+6}_{-5}$\,Myr old.}  {\it Top:}
+    {\bf The \cn\ is \clusterage\ old.}  {\it Top:}
     Color-magnitude diagram of the \cn, IC\,2602 ($\approx35$\,Myr),
@@
 
+% Randich+18: 43.7 +4.3 - 3.9 LDB, for E(B-V) = 0.068 +/- 0.025
+% Dobbie+10: 46 +6 -5 LDB, for
+
 The ages returned by this procedure depend on the ages assumed for
 each reference cluster.  We adopted a 115\,Myr age for the Pleiades,
-and a 16\,Myr age for UCL \citep{dahm_2015,pecaut_star_2016}.
-The age of IC\,2602 however is the most important
-ingredient, since it receives the most weight in the interpolation.
-Plausible ages for IC\,2602
-span 30\,Myr to 46\,Myr, with slightly older ages being preferred by
-the lithium-depletion-boundary (LDB) measurements and younger ages by
-the main-sequence turn-off
+and a 16\,Myr age for UCL \citep{dahm_2015,pecaut_star_2016}.  The age
+of IC\,2602 however is the most important ingredient, since it
+receives the most weight in the interpolation.  Plausible ages for
+IC\,2602 span 30\,Myr to 46\,Myr, with slightly older ages being
+preferred by the lithium-depletion-boundary (LDB) measurements and
+younger ages by the main-sequence turn-off
 \citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019,bouma_cluster_2020}.
@@
 \cn\ would be $44^{+8}_{-7}$\,Myr old.  We adopt an intermediate
-35\,Myr age as our reference, yielding an age for the \cn\ of
-$35^{+6}_{-5}$\, Myr, though it should be understood based on the
-systematic uncertainties that an age of up to $\approx50$\,Myr remains
-plausible.  Detailed analyses of the LDB or main-sequence turn-off in
-the \cn\ could help resolve the issue, but both are beyond our scope.
+38\,Myr age as our reference, yielding an age for the \cn\ of
+\clusterage.\footnote{Our own exploration of the PARSEC and MIST
+isochrone models over a grid of ages, metallicities, and
+reddenings, we yielded the best agreement for this $\approx 38\,$Myr age
+as well, given ${\rm [Fe/H] = +0.1}$ and $A_V=0.2$
+\citep{bressan_parsec_2012,choi_mesa_2016}.
+} While detailed analyses of the LDB or main-sequence turn-off in the
+\cn\ could help resolve the tension between the LDB and isochrone
+ages, these are beyond our scope.
 
@@
 fitting procedure described in Section~\ref{sec:cmd} is
-$35^{+6}_{-5}$\,Myr.  This procedure hinges on the age adopted for
+\clusterage.  This procedure hinges on the age adopted for
 IC\,2602, since it is the closest reference cluster in the
@@
 \citep{douglas_stephanie_t_2021_5131306}.  The most precise age for \sn\ and its
-planet are therefore also $35^{+6}_{-5}$\, Myr, based on the star's
+planet are therefore also \clusterage\ Myr, based on the star's
 spatial, kinematic, and rotational connection to this larger ensemble,
@@
 a white-noise jitter term.  The quadratic model with the RM
-effect is shown in Figure~\ref{fig:obliquity}.  We found $\chi^2_{\rm
+effect is shown in Figure~\ref{fig:obliquity}, where the 
+prior on $v\sin i$ has been assumed from Table~\ref{tab:starparams}.  We found $\chi^2_{\rm
 red} = 5.4$ for the best-fitting linear case, and $\chi^2_{\rm red} =
@@
 over the linear one, the $\chi^2_{\rm red}$ indicate that neither fit
-is particularly good.  Considering a quadratic trend alone without any
-RM effect yields $\chi^2_{\rm red} = 4.3$, and $\Delta {\rm BIC} =
+is particularly good.  Considering a quadratic trend alone {\it without any
+RM effect} yields $\chi^2_{\rm red} = 4.3$, and $\Delta {\rm BIC} =
 {\rm BIC}_{\rm trend} - {\rm BIC}_{\rm quad} = 1.9$ (similarly $\Delta
-{\rm AIC}=5.2$); from a model-comparison standpoint, there is therefore only
+{\rm AIC}=5.2$).  From a model-comparison standpoint, there is therefore only
 weak statistical evidence to prefer the best-fitting ``RM and
@@
 precision, the claim of having detected the Doppler
-transit is therefore marginal.  
+transit is marginal at best.  
 From the MUSCAT3 data however, we do know that the transit was observed, and the
 increase in radial velocity seen shortly after ingress is more
-consistent with a prograde or polar orbit than a retrograde one.
-Fixing the orbit to be either exactly prograde or exactly retrograde
-and allowing for a quadratic trend, we find $\Delta {\rm BIC} = {\rm
-BIC}_{\rm retrograde} - {\rm BIC}_{\rm prograde} = 20.4$; the
+consistent with a prograde or polar orbit than a retrograde orbit.
+Fixing the orbit to be either exactly prograde or exactly retrograde,
+we find $\Delta {\rm BIC} = {\rm
+BIC}_{\rm retrograde} - {\rm BIC}_{\rm trend} = 18.0$; the
 retrograde case is ruled out on this basis.  More precise data will
-however still be useful to make a more precise obliquity measurement.
+however still be needed to make a precise obliquity measurement.
 

### 2021-08-24 paper/authors.tex
update ms
update ms and CMD age
update starparams

@@
 
+% Key authors:
+% ... stellar rotation & the initial crossmatch
 \author[0000-0002-2792-134X]{J. L. Curtis}
@@
 
+% ... Kepler correlations
 \author[0000-0003-1298-9699]{K. Masuda}
 \affiliation{Department of Earth and Space Science, Osaka University, Osaka 560-0043, Japan}
+
+% ... HIRES PI
+\author{L. A. Hillenbrand}
+\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
+
+% ... RM fitting
+\author[0000-0001-7409-5688]{G. Stefansson}
+\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+
+%
+% PFS Collaborators
+%
+\author[0000-0001-8638-0320]{A. W. Howard}
+\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
+%
+\author[0000-0002-0531-1073]{H. Isaacson}
+\affiliation{Astronomy Department, University of California, Berkeley,
+CA 94720, USA}
+
+%
+% MUSCAT3 Collaborators
+%
+\author[0000-0001-8511-2981]{N. Narita}
+\affiliation{Komaba Institute for Science, The University of Tokyo, 3-8-1 Komaba, Meguro, Tokyo 153-8902, Japan}
+\affiliation{Japan Science and Technology Agency, PRESTO, 3-8-1 Komaba, Meguro, Tokyo 153-8902, Japan}
+\affiliation{Astrobiology Center, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
+\affiliation{Instituto de Astrof\'{i}sica de Canarias (IAC), 38205 La Laguna, Tenerife, Spain}
+
+\author[0000-0002-4909-5763]{A. Fukui} % afukui@g.ecc.u-tokyo.ac.jp
+\affiliation{Komaba Institute for Science, The University of Tokyo, 3-8-1 Komaba, Meguro, Tokyo 153-8902, Japan}
+\affiliation{Instituto de Astrof\'{i}sica de Canarias (IAC), 38205 La Laguna, Tenerife, Spain}
+
+\author[0000-0002-5658-5971]{Masahiro Ikoma} % ikoma.masahiro@gmail.com
+\affiliation{Division of Science, National Astronomical Observatory of Japan, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
+
+\author[0000-0002-6510-0681]{M. Tamura} % motohide.tamura@nao.ac.jp
+\affiliation{Department of Astronomy, University of Tokyo, 7-3-1 Hongo, Bunkyo-ku, Tokyo 113-0033, Japan}
+\affiliation{Astrobiology Center, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
+\affiliation{National Astronomical Observatory of Japan, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
+
+% AO IMAGING
+\author[0000-0002-2532-2853]{S. B. Howell}
+\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}

### 2021-08-24 paper/ms.tex
update starparams

@@
   %Kepler 1627Ab: A 38 Million Year Old
-  Kepler 1627 Ab: A 38$\,$Myr Mini-Neptune in the Kepler Field
+  %Kepler 1627 Ab:
+  A 38 Million Year Old Mini-Neptune in the Kepler Field
 }
@@
 38\,Myr age as our reference, yielding an age for the \cn\ of
-\clusterage.\footnote{Our own exploration of the PARSEC and MIST
+\clusterage.\footnote{Our exploration of the PARSEC and MIST
 isochrone models over a grid of ages, metallicities, and
-reddenings, we yielded the best agreement for this $\approx 38\,$Myr age
+reddenings, yielded the best agreement for this $\approx 38\,$Myr age
 as well, given ${\rm [Fe/H] = +0.1}$ and $A_V=0.2$
-\citep{bressan_parsec_2012,choi_mesa_2016}.
+\citep{bressan_parsec_2012,choi_mesa_2016};  this preferred CMD reddening
+is higher than the \citet{lallement_threedimensional_2018} value by a factor of two.
 } While detailed analyses of the LDB or main-sequence turn-off in the
@@
 in the \texttt{PDCSAP} photometry is a quasi-periodic starspot signal
-with a peak-to-peak amplitude that varies between roughly 2\% and 8\%,
+with a peak-to-peak amplitude that varies between 2\% and 8\%,
 depending on the degree of asymmetry in spot coverage between the
@@
 $2.642\pm0.042$\,days, where we have adopted an uncertainty based on
-the scatter observed between different portions (quarters) of the
-time-series data.  The details of how we model this starspot signal to
+the scatter observed between different Kepler quarters.
+The details of how we model this starspot signal to
 derive the parameters of the planet are discussed in
-Appendix~\ref{app:gptransit}.  The results, given in
+Appendix~\ref{app:gptransit}.  The stellar parameters are listed
+in Table~\ref{tab:starparams}; the stellar mass, radius, and effective
+temperature are found by interpolating against a 38$\,$Myr MIST
+isochrone.  The systematic uncertainties are taken to be the
+difference in parameters from the PARSEC isochrones, and the reported
+uncertainties are a quadrature sum of the statistical and systematic
+components.  The transit-fitting results, given in
 Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
 planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit
-($e<0.XX$), around a G8V star ($0.91 \pm 0.05 R_\odot$).  {\bf TODO
+($e<0.XX$), around a G8V star ($0.88 \pm 0.02 R_\odot$).  {\bf TODO
 quote the eccentricity limit}
@@
     could be explained by their primordial atmospheres not yet
-    having evaporated.
-    Selection effects could also be important.
-    Parameters were acquired from the NASA Exoplanet Archive on July
-    2, 2021.
+    having evaporated; direct measurements of the atmospheric outflows
+    or planetary masses would help to confirm this picture.
+    Selection effects may also be important.
+    Parameters are from the NASA Exoplanet Archive (July
+    2, 2021).
 		\label{fig:rp_period_age}
@@
 %
-
+This paper is based in part on observations made with the MuSCAT3
+instrument, developed by the Astrobiology Center and under financial
+supports by JSPS KAKENHI (JP18H05439) and JST PRESTO (JPMJPR1775), at
+Faulkes Telescope North on Maui, HI, operated by the Las Cumbres
+Observatory.
+%
+This work is partly supported by JSPS KAKENHI Grant Numbers 22000005,
+JP15H02063, JP17H04574, JP18H05439, JP18H05442, JST PRESTO Grant
+Number JPMJPR1775, the Astrobiology Center of National Institutes of
+Natural Sciences (NINS) (Grant Number AB031010).
 %

### 2021-08-24 paper/collect.sh

@@ -23,7 +23,7 @@ cp /Users/luke/Dropbox/proj/rudolf/results/ttv_vs_local_slope/ttv_vs_local_slope
 
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit_Kepler_1627_Q15slc/Kepler_1627_Q15slc_RotGPtransit_posterior_phaseplot.pdf f12.pdf
 
-cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20210702_showss_colorbyage_showkep1627.pdf f14.pdf
+cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20210702_colorbyage_showkep1627.pdf f14.pdf
 
 
 cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_overplotkep1627.pdf f15a.pdf
@@ -33,4 +33,4 @@ cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_
 
 cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_emphbinaries_talkaspect_kinematicspatialselected_linear_curtiscleaning.pdf f15supp.pdf
 
-cp /Users/luke/Dropbox/proj/rudolf/results/RM/rm_RV_and_phot.pdf f16.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/RM/quadratic/rm_RV_and_phot.pdf f16.pdf

### 2021-08-30 paper/authors.tex
update ms

@@
 \author[0000-0001-8511-2981]{N. Narita}
-\affiliation{Komaba Institute for Science, The University of Tokyo, 3-8-1 Komaba, Meguro, Tokyo 153-8902, Japan}
-\affiliation{Japan Science and Technology Agency, PRESTO, 3-8-1 Komaba, Meguro, Tokyo 153-8902, Japan}
-\affiliation{Astrobiology Center, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
+\affiliation{Komaba Institute for Science, The University of Tokyo, Tokyo 153-8902, Japan}
+\affiliation{Japan Science and Technology Agency, PRESTO, Tokyo 153-8902, Japan}
+\affiliation{Astrobiology Center, Tokyo 181-8588, Japan}
 \affiliation{Instituto de Astrof\'{i}sica de Canarias (IAC), 38205 La Laguna, Tenerife, Spain}
@@
 \author[0000-0002-4909-5763]{A. Fukui} % afukui@g.ecc.u-tokyo.ac.jp
-\affiliation{Komaba Institute for Science, The University of Tokyo, 3-8-1 Komaba, Meguro, Tokyo 153-8902, Japan}
+\affiliation{Komaba Institute for Science, The University of Tokyo, Tokyo 153-8902, Japan}
 \affiliation{Instituto de Astrof\'{i}sica de Canarias (IAC), 38205 La Laguna, Tenerife, Spain}
@@
 \author[0000-0002-5658-5971]{Masahiro Ikoma} % ikoma.masahiro@gmail.com
-\affiliation{Division of Science, National Astronomical Observatory of Japan, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
+\affiliation{Division of Science, National Astronomical Observatory of Japan, Tokyo 181-8588, Japan}
 
 \author[0000-0002-6510-0681]{M. Tamura} % motohide.tamura@nao.ac.jp
-\affiliation{Department of Astronomy, University of Tokyo, 7-3-1 Hongo, Bunkyo-ku, Tokyo 113-0033, Japan}
-\affiliation{Astrobiology Center, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
-\affiliation{National Astronomical Observatory of Japan, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
+\affiliation{Department of Astronomy, University of Tokyo, Tokyo 113-0033, Japan}
+\affiliation{Astrobiology Center, Tokyo 181-8588, Japan}
+\affiliation{National Astronomical Observatory of Japan, Tokyo 181-8588, Japan}
 
 % AO IMAGING
+\author[0000-0001-9800-6248]{E. Furlan} % furlan@ipac.caltech.edu
+\affiliation{NASA Exoplanet Science Institute, Caltech/IPAC, Pasadena, CA 91125, USA}
+
+\author[0000-0003-2519-6161]{C.~L.~Gnilka} % clgnilka@gmail.com
+\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+
+\author[0000-0002-9903-9911]{K.~V.~Lester} % klester192@gmail.com
+\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+
 \author[0000-0002-2532-2853]{S. B. Howell}

### 2021-09-05 paper/ms.tex
minor ms updtaes

@@
   observed by the Kepler satellite because it is nearby
-  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Combining
-  kinematic information from Gaia, stellar rotation periods from TESS,
-  and Li abundances from ground-based spectra, we show that
+  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Here we show that
   Kepler\,1627 is a member of the \clusterage\ old $\delta$~Lyr
-  cluster.  To our knowledge, Kepler\,1627Ab is the youngest planet
-  with a precise age yet found by the main Kepler mission.  From both
-  unresolved starspot-crossing events in the Kepler photometry and
-  from a Keck/HIRES transit we find that the planet is
-  likely on an orbit nearly aligned with the stellar spin axis.  We
-  also find from the short-cadence Kepler photometry that although the
-  flare arrival time distribution is probably Poissonian, there may be
-  an excess of flares separated by the orbital and synodic periods,
-  and linear combinations thereof.  The $\delta$~Lyr cluster is one of
-  many stellar groups whose properties have been clarified using the
-  Gaia data; many other new clusters also overlap with confirmed and
-  candidate exoplanets.  Further studies that combine Gaia kinematics
-  and TESS rotation periods are poised to expand the census of
+  cluster using kinematic information from Gaia, stellar rotation
+  periods from TESS, and Li abundances from ground-based spectra.  To
+  our knowledge, Kepler\,1627Ab is the youngest planet with a precise
+  age yet found by the main Kepler mission.  From unresolved
+  starspot-crossing events in the Kepler photometry and
+  a Keck/HIRES transit we find that the planet is likely on an
+  orbit aligned with the stellar spin axis.  We also briefly explore the flare
+  arrival time distribution in the short-cadence Kepler
+  photometry.
+  The $\delta$~Lyr cluster is one of many stellar groups whose
+  properties have been clarified using the Gaia data; many other new
+  clusters also overlap with confirmed and candidate exoplanets.
+  Verifying the results of Gaia clustering analyses using TESS rotation
+  periods seems likely to be an effective path toward expanding the census of
   age-dated planets, and our understanding of planetary evolution.

### 2021-09-06 paper/ms.tex
update ms

@@
   Kepler\,1627A is a G8V star previously known to host a
-  $3.5\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
+  $3.3\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
   observed by the Kepler satellite because it is nearby
@@
 2--4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
-is expected to occur within the first few million years of the disk's pressure
-support being removed \citep{owen_atmospheres_2016}, the combination of stellar
+is expected to occur within the first few million years of disk
+dispersal \citep{owen_atmospheres_2016}, the combination of stellar
 irradiation and internal heat may also power a more
@@
 ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
-at the time of the main Kepler mission (2009--2013), only four open
+during the main Kepler mission (2009--2013), only four open
 clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
@@
 we confirm that despite the presence of a previously unreported binary companion,
-the planet orbits the G-dwarf primary star on an orbit that appears aligned
+the planet orbits the G-dwarf primary on an orbit that appears aligned
 with the stellar rotation axis
@@
     the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray;
-    four of them (orange circles) have their names annotated.
+    four of them (orange circles) are annotated.
     Kepler\,1627 (green star) was observed during the entirety of the
@@
     {\bf The \cn\ is \clusterage\ old.}  {\it Top:}
-    Color-magnitude diagram of the \cn, IC\,2602 ($\approx35$\,Myr),
+    Color-magnitude diagram of the \cn, IC\,2602 ($\approx38$\,Myr),
     the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
-    field (gray).  The right panel provides a zoomed
-    view of stars on the pre-main-sequence.  The \cn\ and IC\,2602 are
-    approximately the same isochronal age.  {\it Bottom:} TESS and
-    Kepler stellar rotation periods versus dereddened Gaia colors, with
-    the Pleiades and Praesepe (650\,Myr) shown for
-    reference \citep{rebull_rotation_2016a,douglas_poking_2017}.
-    The \cn\ is gyrochronally younger than the Pleiades.
+    Gaia EDR3 Catalog of Nearby Stars (gray background;
+    log-stretch density).  The right panel provides a zoomed view of stars on
+    the pre-main-sequence.  The \cn\ and IC\,2602 are approximately
+    the same isochronal age.  {\it Bottom:} TESS and Kepler stellar
+    rotation periods versus dereddened Gaia colors, with the Pleiades
+    and Praesepe (650\,Myr) shown for reference
+    \citep{rebull_rotation_2016a,douglas_poking_2017}.  The \cn\ is
+    gyrochronally younger than the Pleiades.
    \label{fig:age}
@@
 
-We meaured the isochrone age using an empirical approach.  The left
+We measured the isochrone age through an empirical approach.  The left
 panel of Figure~\ref{fig:age} shows the color-magnitude diagram (CMD)
-of stars in the \cn, IC\,2602, the Pleiades, and the field.  The stars
-shown in the \cn\ are selected based on their positions and
+of stars in the \cn, IC\,2602, the Pleiades, and the field.  The
+candidate \cn\ members shown are
+selected based on their positions and
 kinematics, as discussed in Appendix~\ref{app:kinematicselection}.
@@
 EDR3 Catalog of Nearby Stars \citep{gaia_gcns_2021}.  We cleaned these
-following identical data filtering criteria discussed in Appendix~B of
+following the data filtering criteria discussed in Appendix~B of
 \citet{GaiaCollaboration2018}, except that we weakened the parallax
@@
 \citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}},
-and by then applying the extinction coefficients $k_X\equiv A_X/A_0$
-computed by \citet{GaiaCollaboration2018} while assuming that $A_0 =
+and then applied the extinction coefficients $k_X\equiv A_X/A_0$
+computed by \citet{GaiaCollaboration2018} and assumed that $A_0 =
 3.1 E(B-V)$.  For IC\,2602, the Pleiades, and the \cn, this procedure
@@
 latter two groups.  To turn this heuristic interpolation into an
-age measurement, we used the empirical procedure
+age measurement, we used the empirical method
 developed by \citet{gagne_mutau_2020}.  In brief, we fitted the
@@
 $\texttt{radial\_velocity\_error}$ below the 80$^{\rm th}$ percentile
-of each cluster's distribution, and by manually excluding stars that
+of each cluster's distribution, and manually excluded stars that
 were obvious photometric binaries in the CMD.  We then passed a moving
@@
 \cn\ would be $44^{+8}_{-7}$\,Myr old.  We adopt an intermediate
-38\,Myr age as our reference, yielding an age for the \cn\ of
+38\,Myr age for IC\,2602, which yields an age for the \cn\ of
 \clusterage.\footnote{Our exploration of the PARSEC and MIST
@@
 combination of automated and manual steps.  As a matter of scope, we
-restricted our attention to the \nkinwithtess\ stars in the spatial
-and kinematic proximity of Kepler\,1627, as discussed in
-Appendix~\ref{app:kinematicselection}.  We also only focused on stars
+restricted our attention to the \nkinwithtess\ stars discussed in
+Appendix~\ref{app:kinematicselection} in the spatial
+and kinematic proximity of Kepler\,1627.  To avoid blending, we also only focused on stars
 for which no companions were known with a brightness exceeding
-one-tenth of the target star within a $21\farcs0$ radius: there were
+one-tenth of the target star in a $21\farcs0$ radius.  There were
 \nkinwithtessandcrowding\ stars that met these requirements, and that
 had TESS data available.  For plotting purposes we then imposed a
-selection based on the strength of the rotation period signal itself:
+selection based on the strength of the signal itself:
 we required the Lomb Scargle power to exceed 0.2, and the period to be
@@
 $\nrotgood/\nkinwithtessandcrowding \approx \nfracprot\%$.
+The rotation periods of other 30\,Myr to 50\,Myr clusters
+({\it e.g.}, IC\,2602 and IC\,2391) are indistinguishable
+\citep{douglas_stephanie_t_2021_5131306}.
 
@@
 evidence for similar effects to those that have been previously
-discussed in the literature.
+noted in the literature.
 
@@
 
-In brief, the age of the \cn\ cluster based on the empirical
-fitting procedure described in Section~\ref{sec:cmd} is
+To summarize, the age of the \cn\ cluster based on the empirical
+procedure described in Section~\ref{sec:cmd} is
 \clusterage.  This procedure hinges on the age adopted for
@@
 K dwarfs in the \cn\ have not yet converged to a slow sequence, their
-rotation distribution and that of the F dwarfs is consistent with the
-same age.  The rotation periods of other 30\,Myr to 50\,Myr clusters
-({\it e.g.}, IC\,2602 and IC\,2391) are indistinguishable
-\citep{douglas_stephanie_t_2021_5131306}.  The most precise age for \sn\ and its
+rotation periods and those of the F dwarfs are also consistent with a
+30 to 50\,Myr age. The most precise age for \sn\ and its
 planet are therefore also \clusterage\ Myr, based on the star's
-spatial, kinematic, and rotational connection to this larger ensemble,
+spatial, kinematic, and rotational connection to the cluster,
 and the assumption that the planet formed very quickly after the star.
@@
 \sn, there is a final consistency check: the abundance of
-Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet.  The utility
-of this element for age dating young stars has been reviewed by
-\citet{soderblom_ages_2014}.  We acquired an iodine-free spectrum from
+Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet
+\citep[see {\it e.g.},]{soderblom_ages_2014}.
+We acquired an iodine-free spectrum from
 Keck/HIRES on the night of 26 March 2021 using the standard setup and
@@
 procedure described by \citet{bouma_2021_ngc2516}, we find ${\rm
-EW}_{\rm Li} = 233^{+5}_{-7}\,$m\AA.   Note that this value does not
+EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   Note that this value does not
 correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given the stellar
@@
 effective temperature, this measurement is also in agreement with an
-age for the star between 30\,Myr and 50\,Myr (see {\it e.g.}, Figure~9
-of \citealt{bouma_2021_ngc2516}, or the data from
-\citealt{randich_gaiaeso_2018}).
+age for the star between 30\,Myr and 50\,Myr ({\it e.g.}, 
+\citealt{randich_gaiaeso_2018}; \citealt{bouma_2021_ngc2516}).
 
@@
     variability model.  The Figure Set available online shows the
-    entire 3.9 years of available data.
+    entire 3.9 years of observations.
     {\it Bottom}:
@@
     {\it Top:} 
-    The time of each Kepler transit was measured, along with the local slope
-    of the light curve.  The two quantities appear anti-correlated,
-    which is most easily explained by starspot crossings during the
-    first (second) half of transit inducing a positive (negative) TTV,
-    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
-    units along the abscissa are most easily understood by considering
-    that the flux from the star changes by $\sim$60\,ppt per half
-    rotation period ($\sim$1.3\,days).
-    {\it Bottom:} Radial velocities from
-    Keck/HIRES during a transit on 7 Aug 2021 UT, observed
-    simultaneously with MUSCAT3 ({\it griz} photometry).  A quadratic baseline trend was
-    allowed in the RV fit.  The best-fit sky-projected obliquity in this
-    model is listed in red; the detection however has marginal
-    statistical significance compared to a ``trend-only'' model.
+    The time of each Kepler transit was measured, along with the local
+    slope of the light curve.  The two quantities appear
+    anti-correlated, which is most easily explained by starspot
+    crossings during the first (second) half of transit inducing a
+    positive (negative) TTV, provided that the orbit is prograde
+    \citep{mazeh_time_2015}.  The units along the abscissa are most
+    easily understood by considering that the flux from the star
+    changes by $\sim$60\,ppt per half rotation period
+    ($\sim$1.3\,days).
+    {\it Bottom:} Radial velocities from Keck/HIRES during a transit
+    on 7 Aug 2021 UT, observed simultaneously with MuSCAT3 ({\it griz}
+    photometry).  A quadratic baseline trend was allowed in the RV
+    fit.  The best-fit sky-projected obliquity in this model is listed
+    in red; the detection however has marginal statistical
+    significance compared to a ``trend-only'' model.
     \label{fig:obliquity}
@@
 
-%TODO: fix sizes and numbers
 If the \pn\ transit signal is created by a bonafide planet,  then to
 our knowledge its age would make it the youngest planet yet found by
-the main Kepler mission.\footnote{The re-purposed K2 mission found two
-slightly younger systems: K2-33b
-\citep{David_et_al_2017,Mann_K2_33b_2016} and V1298 Tau
-\citep{david_four_2019}}  A portion of the light curve is shown with
-the phase-folded transit in Figure~\ref{fig:lc}.  The dominant signal
-in the \texttt{PDCSAP} photometry is a quasi-periodic starspot signal
-with a peak-to-peak amplitude that varies between 2\% and 8\%,
-depending on the degree of asymmetry in spot coverage between the
-stellar hemispheres.  The starspot signal repeats every
-$2.642\pm0.042$\,days, where we have adopted an uncertainty based on
-the scatter observed between different Kepler quarters.
-The details of how we model this starspot signal to
-derive the parameters of the planet are discussed in
-Appendix~\ref{app:gptransit}.  The stellar parameters are listed
-in Table~\ref{tab:starparams}; the stellar mass, radius, and effective
+the main Kepler mission.\footnote{The re-purposed K2
+mission however has found two younger systems: K2-33b and V1298 Tau
+\citep{David_et_al_2017,Mann_K2_33b_2016,david_four_2019}}  A portion
+of the light curve is shown with the phase-folded transit in
+Figure~\ref{fig:lc}.  The dominant signal in the \texttt{PDCSAP}
+photometry is a quasi-periodic starspot signal with a peak-to-peak
+amplitude that varies between 2\% and 8\%, depending on the degree of
+asymmetry in spot coverage between the stellar hemispheres.  The
+starspot signal repeats every $2.642\pm0.042$\,days, where the
+uncertainty is based on the scatter observed between different Kepler
+quarters.  The details of how we model the starspot signal to derive
+the parameters of the planet are discussed in
+Appendix~\ref{app:gptransit}.  The stellar parameters are listed in
+Table~\ref{tab:starparams}; the stellar mass, radius, and effective
 temperature are found by interpolating against a 38$\,$Myr MIST
@@
 Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
-planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit
-($e<0.XX$), around a G8V star ($0.88 \pm 0.02 R_\odot$).  {\bf TODO
-quote the eccentricity limit}
+planet ($3.34\pm 0.18\,R_\oplus$) on a circular\footnote{Note $e<0.48$
+at 2-$\sigma$; the constraints on the eccentricity are not
+particularly strong.} 7.2\,day orbit, around a G8V star ($0.88 \pm
+0.02 R_\odot$).
 
 Could the transit signal be produced by anything other than a planet
-orbiting this near-solar analog?  \citet{morton_false_2016} considered
-the false positive scenarios of a foreground eclipsing binary, a
-hierarchical triple-star binary system, and a background eclipsing
-binary.  Based on the transit shape, \citet{morton_false_2016} found
-the latter scenario to be the only possible case, with a
-(model-dependent) probability of $\approx10^{-5}$. Most
-reported failures of the statistical validation approach have arisen when attempting
-to separate the smallest stars from identically-sized gas giant
-planets \citep[{\it e.g.},][]{shporer_three_2017}.  With a size of
-$\approx 3.5\,R_\oplus$, Kepler\,1627b does not meet this concern.
-The fact that we detect a low-mass stellar companion in archival
-Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does not change the
-conclusion: the companion star contributes 1\% of the flux observed by
-Kepler.  Combined with the observed transit depth and shape
-(Appendix~\ref{app:companionstar}), a planet transiting the primary
-star is the only plausible interpretation of the data.
-
-{\bf todo: LGB to check numbers on full-dataset transit ecc and b} 
-Three additional new lines of evidence confirm the planetary
-interpretation.  First, the $\approx$100 days of 1-minute
-cadenced Kepler observations reduce the cadence-induced smearing,
-enabling a more precise measurement of the transit impact parameter
-\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
-(3-$\sigma$) from the 1-minute cadenced data, compared to $<0.XX$
-(3-$\sigma$) from the 30-minute data.
-This implies an even stronger constraint on the transit shape in an
-analysis of the false positive probability.
-
-% TODO: Kento Masuda, please verify
-The second line of evidence comes from our transit timing analysis. We
-isolated each of the 144 observed transits to within $\pm4.5$\,hr of
-each transit, and fitted each window with both {\it i)} a local
-second-order polynomial and transit, and {\it ii)} a local linear
-trend.  We let the mid-time of each transit float, and then calculated
-the residual between the measured mid-time and that of a perfectly
-periodic orbit.  This residual, the ``transit timing variation''
+orbiting this near-solar analog?  \citet{morton_false_2016} validated
+the planet based on the transit shape, arguing that the only plausible
+false positive scenario was that of a background eclipsing binary,
+reported to have a model-dependent probability of $\approx10^{-5}$.
+As discussed in Appendix~\ref{app:companionstar},
+we discovered a low-mass stellar companion ($M_{\rm B}\approx
+0.44\,M_\odot$), and so here we reconsider this false positive scenario
+in some detail.
+
+% At 30 Myr, a 0.44\,$M_\odot$ solar-metallicity dwarf is $\approx 26\%$
+% larger than when it is fully contracted on the main sequence
+% (0.40\,$R_\odot$ vs{.} 0.50\,$R_\odot$; ).
+
+Three new lines of evidence support the planetary interpretation.
+First, the transit duration and the orbital period are inconsistent
+with an eclipsing body around the M-dwarf companion: we find
+$\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
+theoretically expected density for the companion at 38\,Myr is
+$\approx 3.2\,{\rm g\,cm}^{-3}$ (Table~\ref{tab:posterior};
+\citealt{choi_mesa_2016}).  The transit duration is therefore too long
+to explained by a star eclipsing the M dwarf secondary at
+$5$-$\sigma$ confidence.
+
+%NOTE to the arxiv readers:, the $\approx$100 days of short-cadence
+%Kepler observations also mitigate the cadence-induced smearing, and
+%enable a more precise measurement of the transit impact parameter
+%\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
+%(2-$\sigma$) from the 1-minute data, compared to $<0.76$ (2-$\sigma$)
+%from the 30-minute data.  This implies an even stronger constraint on
+%the transit shape in an analysis of the false positive probability.
+
+The second line of confirmatory evidence for the planetary
+interpretation comes by analyzing the individual transit times in the
+Kepler data. We isolated each of the 144 observed transits to within
+$\pm4.5$\,hr of each transit, and fitted each window with both {\it
+i)} a local second-order polynomial and transit, and {\it ii)} a local
+linear trend.  We let the mid-time of each transit float, and then
+calculated the residual between the measured mid-time and that of a
+perfectly periodic orbit.  This residual, the transit timing variation
 (TTV), is plotted in Figure~\ref{fig:obliquity} against the local
-linear slope.  A significant negative slope of $-0.0564 \pm 0.0098\
-{\rm ppt\,day}^{-1}$ is observed.  Only a handful of Kepler Objects of
+linear slope.  A significant correlation of $-0.0564 \pm 0.0098\ {\rm
+ppt\,day}^{-1}$ is observed.  Less than ten Kepler Objects of
 Interest have shown this correlation \citep{holczer_time_2015}, which
@@
 crossings \citep{mazeh_time_2015}.  This is only possible if the
-planet transits the primary star, which excludes the aforementioned
-false positive scenarios.  It also implies that the planet's orbit is
-prograde.  The latter interpretation assumes that the dominant
-photometric variability is induced by dark spots, rather than bright
-faculae.  Given the observed transition of Sun-like stellar
-variability from spot to faculae-dominated regimes, we expect this
-inference to be reasonably secure
+planet transits the primary star, which excludes a background
+eclipsing binary scenario, and therefore confirms that \pn\ is a
+planet.  It also implies that the planet's orbit is prograde.  The
+latter interpretation assumes that the dominant photometric
+variability is induced by dark spots, rather than bright faculae
+\citep{mazeh_time_2015}.  Given the observed transition of Sun-like
+stellar variability from spot to faculae-dominated regimes, we expect
+this inference to be reasonably secure
 \citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
 
-The final line of evidence confirming \pn\ comes from a spectroscopic
+The final piece of evidence \pn\ comes from a spectroscopic
 transit we measured from Keck/HIRES during the night of 7 Aug 2021
@@
 TESS data.  We also acquired simultaneous multi-band photometry using
-MUSCAT3, to help interpret the radial velocities \citep{Narita_2020}.
+MuSCAT3 to help interpret the radial velocities \citep{Narita_2020}.
 The lower panels in Figure~\ref{fig:obliquity} show the results.  The
@@
 and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
-The radial velocities show significant scatter, though a significant
+The radial velocities show significant scatter, and a significant
 increase in RV is seen shortly after the expected ingress.  The
@@
 We fitted the RVs using the \citet{hirano_analytic_2010,hirano_2011}
-models for the Rossiter-McLaughlin (RM) effect, and allowed three parameters
-to vary: the sky-projected obliquity, the stellar
-equatorial velocity, and the Gaussian dispersion of spectral lines.
-We also allowed for an optional linear and quadratic trend, and
-a white-noise jitter term.  The quadratic model with the RM
-effect is shown in Figure~\ref{fig:obliquity}, where the 
-prior on $v\sin i$ has been assumed from Table~\ref{tab:starparams}.  We found $\chi^2_{\rm
-red} = 5.4$ for the best-fitting linear case, and $\chi^2_{\rm red} =
-4.4$ for the best-fitting quadratic, with a $\Delta {\rm BIC} = {\rm
+models for the Rossiter-McLaughlin (RM) effect, and allowed the
+sky-projected obliquity, the stellar equatorial velocity, and the
+Gaussian dispersion of the spectral lines to vary.  We also allowed
+for an optional linear and quadratic trend, and a white-noise jitter
+term.  The quadratic model with the RM effect is shown in
+Figure~\ref{fig:obliquity}; the prior on $v\sin i$ has been
+assumed from Table~\ref{tab:starparams}.  We found $\chi^2_{\rm red} =
+5.4$ for the best-fitting linear case, and $\chi^2_{\rm red} = 4.4$
+for the best-fitting quadratic, with a $\Delta {\rm BIC} = {\rm
 BIC}_{\rm line} - {\rm BIC}_{\rm quad} = 19.6$ and $\Delta {\rm AIC} =
@@
 over the linear one, the $\chi^2_{\rm red}$ indicate that neither fit
-is particularly good.  Considering a quadratic trend alone {\it without any
-RM effect} yields $\chi^2_{\rm red} = 4.3$, and $\Delta {\rm BIC} =
-{\rm BIC}_{\rm trend} - {\rm BIC}_{\rm quad} = 1.9$ (similarly $\Delta
-{\rm AIC}=5.2$).  From a model-comparison standpoint, there is therefore only
-weak statistical evidence to prefer the best-fitting ``RM and
+is very good.  Considering a quadratic trend alone without any RM
+effect yields $\chi^2_{\rm red} = 4.3$, and $\Delta {\rm BIC} = {\rm
+BIC}_{\rm trend} - {\rm BIC}_{\rm quad} = 1.9$ ($\Delta {\rm
+AIC}=5.2$).  From a model-comparison standpoint, there is therefore
+only weak statistical evidence to prefer the best-fitting ``RM and
 quadratic trend'' model over the best-fitting ``quadratic trend only''
 model.  Given the expected semi-amplitude of the velocity anomaly
-($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) and the actual achieved
-precision, the claim of having detected the Doppler
-transit is marginal at best.  
-From the MUSCAT3 data however, we do know that the transit was observed, and the
-increase in radial velocity seen shortly after ingress is more
-consistent with a prograde or polar orbit than a retrograde orbit.
-Fixing the orbit to be either exactly prograde or exactly retrograde,
-we find $\Delta {\rm BIC} = {\rm
-BIC}_{\rm retrograde} - {\rm BIC}_{\rm trend} = 18.0$; the
-retrograde case is ruled out on this basis.  More precise data will
-however still be needed to make a precise obliquity measurement.
+($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) and the actual
+achieved precision, the claim of having detected the Doppler transit
+is marginal at best.  From the MuSCAT3 data however, we do know that
+the transit was observed, and the increase in radial velocity seen
+shortly after ingress is more consistent with a prograde or polar
+orbit than a retrograde orbit.  Fixing the orbit to be either exactly
+prograde or exactly retrograde, we find $\Delta {\rm BIC} = {\rm
+BIC}_{\rm retrograde} - {\rm BIC}_{\rm trend} = 18.0$; the retrograde
+case is ruled out on this basis.  More precise data will be needed
+however to make a more precise obliquity measurement.
 
@@
 		\leavevmode
-		\includegraphics[width=\textwidth]{f14.pdf}
+		\includegraphics[width=0.9\textwidth]{f14.pdf}
 	\end{center}
@@
     {\bf Radii, orbital periods, and ages of known transiting
-    exoplanets}.  Planets older than
-    $10^9$ years are shown with gray markers.  Planets younger than a
-    gigayear, with $t/\sigma_t > 3$, are shown in color.  
-    \pn\ is shown with a star.
-    The abnormally large radii of the youngest transiting planets
-    could be explained by their primordial atmospheres not yet
-    having evaporated; direct measurements of the atmospheric outflows
-    or planetary masses would help to confirm this picture.
-    Selection effects may also be important.
-    Parameters are from the NASA Exoplanet Archive (July
+    exoplanets}.  Planets older than $10^9$ years are shown with gray
+    markers.  Planets younger than a gigayear, with $t/\sigma_t > 3$,
+    are shown in color.  \pn\ is shown with a star.  The large radii
+    of the youngest transiting planets could be explained by their
+    primordial atmospheres not yet having evaporated; direct
+    measurements of the atmospheric outflows or planetary masses would
+    help to confirm this expectation.  Selection effects may also be
+    important.  Parameters are from the NASA Exoplanet Archive (July
     2, 2021).
@@
 
-{\bf LGB TODO: ask James Owen what he thinks the mass loss rate is}
-Kepler\,1627Ab therefore provides a new extremum in the ages of the
-Kepler planets, and opens multiple avenues for study from the ground.
-Observations of its Doppler transit are expected to induce a
-Rossiter-McLaughlin anomaly with an amplitude of $\approx 20$\ms, and
-could yield a quantification of the stellar obliquity, and a means for
-verifying our transit-timing based suggestion of a prograde orbit.
-Similar transit spectroscopy aimed at detecting either hydrogen or
-helium absorption could yield insight into the planet's atmosphere,
-which is expected to be outflowing at a rate of {\bf
-X.X}\,$M_\oplus\,{\rm yr}^{-1}$ (CITE).  A more challenging
-measurement would be a measurement of the planet's mass, through
-either detailed timing analyses of the Kepler data, or through a
-high-cadence multi-color radial velocity campaign.  The mass could
-yield constraints on both the planet's composition as well as its
-initial entropy \citep{owen_constraining_2020}.
+%, which is expected to be outflowing at a rate of
+%$\sim$$10^{-2}$\,$M_\oplus\,{\rm Myr}^{-1}$ \citep[{\it
+%e.g.},][]{owen_constraining_2020}.
+
+Kepler\,1627Ab provides a new extremum in the ages of the Kepler
+planets, and opens multiple avenues for further study.  More
+precise observations of the Doppler transit could yield a more precise
+stellar obliquity, and a means for verifying the expected prograde
+orbit.  Similar transit
+spectroscopy aimed at detecting either hydrogen or helium absorption
+could yield insight into the stability of the planet's atmosphere
+\citep[{\it e.g.},][]{spake_helium_2018,vissapragada_2020}.  A more
+challenging measurement would be to determine the planet's mass,
+either through detailed timing analyses of the Kepler data, or through
+a high-cadence multi-color radial velocity campaign.  The mass,
+measured at sufficient precision, could yield constraints on both the
+planet's composition as well as its initial entropy
+\citep{owen_constraining_2020}.
 
@@
 youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
-with precisely measured ages spanning 10 to 100\,Myr include K2-33b,
+with precisely measured ages below 100\,Myr include K2-33b,
 DS~Tuc~Ab, HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet
@@
 \citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020}.
-Of these systems, \pn\ inhabits a similar range of planetary size and
-orbital period.  It is the smallest though ($3.6\,R_\oplus$), which
-may hint at selection effects in the sample of young planets
-\citep{zhou_2021_tois}.  Assuming that these planets have masses
-between $\approx$10 and 20$\,M_\oplus$, their large sizes and presumed
-low densities are in accord with the expectation that mini-Neptunes
-start their lives with large primordial atmopsheres that are are lost
-over the first gigayear
-\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.
-The observation from starspot-induced TTVs that the planetary orbit
-could be aligned with the stellar spin differs from a number of other
-close-in
+\pn\ is currently the smallest of these systems ($\approx
+3.3\,R_\oplus$), which may hint at selection effects in the 
+young planet sample \citep{zhou_2021_tois}.  Assuming that these planets
+have masses between $\approx$5 and 20$\,M_\oplus$, their large sizes
+and presumed low densities are in accord with the expectation that
+mini-Neptunes start their lives with large primordial atmospheres that
+are shed over the first gigayear
+\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
+suggestion of an aligned orbit differs from observed misalignments in number of other close-in
 mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
-Overall however, it is consistent with a quiescent disk-driven
-migration history for the planet, while the in-situ formation scenario
-is disfavored because the necessary disk mass would always drive
-planets to migrate
+Unlike these older planets, the properties of \pn\
+seem to be consistent with a quiescent disk-driven migration history.
+The alternative scenario of in-situ formation is disfavored on
+theoretical grounds: the required disk mass would likely drive
+migration, and so lacks internal consistency
 \citep{inamdar_formation_2015,ogihara_reassessment_2015}.
 
-% TODO: decide whether a paragraph about young planet obliquities is
-% needed.
-
-% The orbits of the planets can also change early on, and can become
-% significantly misaligned from the equators of their host stars.
-% Measurements of the sky-projected angle between the planet orbit and
-% the stellar spin (the stellar obliquity) have shown that high
-% obliquities characterize hot field stars hosting both hot Jupiters and
-% smaller planets \citep{winn10,louden21}.  Conversely, the obliquities
-% of planets around cool stars tend to be low, but this could be due to
-% tidal realignment \citep[{\it e.g.},][]{anderson21}.  Obliquity
-% measurements for young planets are therefore important, because they
-% provide an unobstructed view of the primordial obliquity distribution.
-
-Ultimately, our age measurement for \pn\ was enabled by identifying
-its connection to the \cn\ using the Gaia data.  The Supplementary
-Table~\ref{tab:v05} enables similar crossmatches to already known
-planets, as well as forthcoming planets from the TESS mission
-\citep{ricker_transiting_2015,guerrero_tess_2021}.  An alternative
-approach would be to identify kinematic associations around exoplanet
-host stars using tangential velocities from Gaia, and to then verify these associations
-with ancilliary rotation periods and spectroscopy
-\citep{tofflemire_tess_2021}.
+Ultimately though, the main advance of this work was a precise measurement of
+the age of \pn.   The age measurement was enabled by identifying the
+connection of the star to the \cn\ using the Gaia data.  Table~\ref{tab:v05}
+enables similar crossmatches to
+already known planets, as well as forthcoming planets from the TESS
+mission \citep{ricker_transiting_2015,guerrero_tess_2021}.  An
+alternative path toward age-dating more planets could be to identify kinematic associations
+around known exoplanet host stars using positions and tangential velocities
+from Gaia, and to then verify these associations with ancilliary
+rotation periods and spectroscopy \citep[{\it e.g.},][]{tofflemire_tess_2021}.
 
@@
 helping resolve the scheduling conflict that would have otherwise
-prevented the MUSCAT3 observations.
+prevented the MuSCAT3 observations.
 % is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
@@
 %
+The archival Keck/NIRC2 imaging was acquired by programs 2015A/N301N2L
+(PI: A.~Kraus) and 2019A/N069 (PI: E.~Petigura).
+%
+In addition, this paper is based in part on observations made with the
+MuSCAT3 instrument, developed by the Astrobiology Center and under
+financial supports by JSPS KAKENHI (JP18H05439) and JST PRESTO
+(JPMJPR1775), at Faulkes Telescope North on Maui, HI, operated by the
+Las Cumbres Observatory.
+%
+This work is partly supported by JSPS KAKENHI Grant Numbers 22000005,
+JP15H02063, JP17H04574, JP18H05439, JP18H05442, JST PRESTO Grant
+Number JPMJPR1775, the Astrobiology Center of National Institutes of
+Natural Sciences (NINS) (Grant Number AB031010).
+%
 % ACKNOWLEDGE PFS / CAMPANAS.
@@
 %
-
 %
@@
 %
-This research has made use of the Keck Observatory Archive (KOA),
+Finally, this research has made use of the Keck Observatory Archive (KOA),
 which is operated by the W. M. Keck Observatory and the NASA Exoplanet
 Science Institute (NExScI), under contract with the National
-Aeronautics and Space Administration.
-We also thank
-the Keck Observatory staff, particularly {\bf A.~Doe},
-for support of HIRES and remote observing.  We recognize
-the importance
-that the summit of Maunakea has within the indigenous
-Hawaiian community, and are deeply grateful to
-have the opportunity to conduct observations from this
-mountain.
-%
-This paper is based in part on observations made with the MuSCAT3
-instrument, developed by the Astrobiology Center and under financial
-supports by JSPS KAKENHI (JP18H05439) and JST PRESTO (JPMJPR1775), at
-Faulkes Telescope North on Maui, HI, operated by the Las Cumbres
-Observatory.
-%
-This work is partly supported by JSPS KAKENHI Grant Numbers 22000005,
-JP15H02063, JP17H04574, JP18H05439, JP18H05442, JST PRESTO Grant
-Number JPMJPR1775, the Astrobiology Center of National Institutes of
-Natural Sciences (NINS) (Grant Number AB031010).
+Aeronautics and Space Administration.  We also thank the Keck
+Observatory staff for their support of HIRES and remote observing.  We
+recognize the importance that the summit of Maunakea has within the
+indigenous Hawaiian community, and are deeply grateful to have the
+opportunity to conduct observations from this mountain.
 %
@@
     %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
- 	%Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
+ 	Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
  	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
+ 	Gemini:North~(`Alopeke; \citealt{scott_nessi_2018}.
  	{\it Spectroscopy}:
-	CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
+	%CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
   %PFS ({\bf CITE}),
@@
 	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
-	AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
- 	%Keck:I~(HIRES; \citealt{vogt_hires_1994}).
- 	VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
+	%AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
+ 	Keck:I~(HIRES; \citealt{vogt_hires_1994}).
+ 	%VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
 % 	Euler1.2m~(CORALIE),
@@
 % 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
+    MuSCAT3 \citep{Narita_2020},
 % 	NTT,
@@
 }
-
-% \input{TOI837_phot_table.tex}
-% \input{TOI837_rv_table.tex}
-% \input{ic2602_ages.tex}
-% \input{starparams.tex}
-% \input{compparams.tex}
 
@@
 
+\section{Companion Star and False Positive Assessment}
+\label{app:companionstar}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.6\textwidth]{Kepler1627_20210624_562_832_final.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Keck NIRC2 AO image of Kepler 1627A and Kepler 1627B.}  
+    {\bf SCALEBAR DENOTES...}
+    {\bf North is up, East is left}.
+    \label{fig:ao}
+	}
+\end{figure*}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.6\textwidth]{f6.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Gaia EDR3 renormalized unit weight error (RUWE) point
+    estimates for Kepler\,1627A and other members of the $\delta$\,Lyr
+    cluster.}  Since other members of the cluster with similar colors
+    have comparable degrees of photometric variability, the high RUWE
+    suggests that Kepler\,1627A is an astrometric binary. 
+		\label{fig:ruwe}
+	}
+\end{figure*}
+
+We first noted the presence of a stellar neighbor in archival
+Keck/NIRC2 Kp-band adaptive optics imaging (2.12\,$\mu $m), from the
+Spring semesters of both 2015 and 2019.  We acquired additional
+Gemini-North/`Alopeke speckle imaging on 24 June 2021, which showed
+the same companion in the 0.83\,$\mu$m notch filter, but not the
+$0.56\,\mu$m filter.  Figure~\ref{fig:ao} summarizes the results of
+the high-resolution imaging.  Although the companion is unresolved in
+the Gaia source catalog, its existence was also suggested by the
+primary star's large RUWE, relative to other members of the \cn.  The
+images yield a projected separation $\rho \approx 0\farcs16$, with
+$\Delta m_{832} \approx 3.1$ and $\Delta m_{\rm Kp} = 2.5$.  Using the
+measured Gaia EDR3 parallax for the system, this implies a projected
+separation of $53 \pm 7$.  Given the low geometric probably of a star
+imaged within $\rho \approx 0\farcs16$ to be a chance companion along
+the line of sight, and the large observed RUWE, we proceed under the
+assumption that the companion is bound, and that the Kepler\,1627
+system is binary.  Based on the
+tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
+version \texttt{2021/03/02}.} by \citet{pecaut_mamajek_2013}, the
+measured magnitude differences for a G8V primary correspond to a
+spectral type for the companion of $\approx$M2V ($M_\star \approx
+0.44\,M_\odot$).  However, the companion should have a longer
+pre-main-sequence contraction phase than the primary, which would
+imply that this mass is overestimated by $\sim10$ to $20\%$.
+
+% (0.7-2.0\% V-band to G-band dmag difference from EEM table; NOTE
+% could determine using model spectra)
+Could the companion be creating a false positive signal?  The
+companion star contributes $\approx$1\% of the total flux observed in
+the Kepler aperture.  The observed transit has a depth of 0.18\%.  An
+18\% deep eclipse of the secondary star would therefore be needed to
+produce a signal with the appropriate depth.  The shape of the
+observed signal requires the impact parameter to be below 0.74
+(Table~\ref{tab:posterior}); the body transiting the secondary would
+therefore need to be non-grazing with $R_3/R_2 \approx 0.42$.
+Assuming a $\approx 0.42R_\odot$ radius of the imaged secondary, this
+would imply a tertiary stellar radius of $\approx 0.2R_\odot$.  This
+yields a contradiction:  this scenario would require an ingress and
+egress phase that each span $\approx$40\% of the transit duration
+($\approx 65\,{\rm minutes}$).  The actual measured ingress and egress
+duration is $\approx 15\,{\rm minutes}$), 4.4$\times$ shorter.  The
+combination of the companion's brightness, transit depth, and ingress
+duration therefore rule out the scenario that it could explain the
+transit signal.
+
 \section{Flare Analysis}
@@
 
-\section{Companion Star and False Positive Assessment}
-\label{app:companionstar}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.6\textwidth]{f6.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Gaia EDR3 renormalized unit weight error (RUWE) point
-    estimates for Kepler\,1627A and other members of
-    the $\delta$\,Lyr cluster.}  Since other members of the cluster
-    with similar colors have comparable degrees of photometric
-    variability, the high RUWE estimate suggests that Kepler\,1627A is
-    a binary. 
-		\label{fig:ruwe}
-	}
-\end{figure*}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.6\textwidth]{Kepler1627_20210624_562_832_final.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Keck NIRC2 AO image of Kepler 1627A and Kepler 1627B.}  
-    {\bf SCALEBAR DENOTES...}
-    {\bf North is up, East is left}.
-    \label{fig:ao}
-	}
-\end{figure*}
-
-Our analysis of archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging
-revealed the existence of a previously unreported stellar neighbor,
-unresolved in the Gaia source catalog {\bf FIXME: uncertainties}.  The
-NIRC2 images yield a projected separation $\rho=0\farcs17$, with
-$\Delta {\rm Kp} = 2.5$.  Using the measured Gaia EDR3 parallax for
-the system, this implies a projected separation of 54\,AU.  The
-presence of this star is consistent with the excess noise in the Gaia
-astrometric time-series (see Section~\ref{app:companionstar}).  Given
-the low chance of a star imaged within {\bf X.X arcseconds} to be a
-chance companion along the line of sight, we proceed under the
-assumption that it is bound, and that the Kepler\,1627 system is
-binary.
-
-Unforuntately, we do not have any reliable color information about
-Kepler\,1627B.  Based on the
-tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
-version \texttt{2021/03/02}.} by \citet{pecaut_mamajek_2013}, the
-measured NIR-contrast for a main sequence G8V star corresponds to a
-spectral type for the companion of $\approx$M2V ($M_\star \approx
-0.44\,M_\odot$).  However, the companion should have a longer
-pre-main-sequence contraction phase than the primary, which would
-imply that this mass is overestimated by $\sim10$ to $20\%$.
-
-Could the companion be creating a false positive signal?  The
-companion star contributes $\approx$1\% of the total flux observed in
-the Kepler aperture (0.7-2.0\% V-band to G-band dmag difference from
-EEM table; todo is fix using model spectra).  The observed transit has
-a depth of 0.18\%.  An 18\% deep eclipse of the secondary star would
-therefore be needed to produce a deep enough signal.  The shape of the
-observed signal requires allowed impact parameters to span 0.02 to
-0.73 (Table~\ref{tab:posterior}); the body transiting the secondary
-would therefore need to be non-grazing with $R_3/R_2 \approx 0.42$.
-Assuming a $\approx 0.42R_\odot$ radius of the imaged secondary, this
-would imply a tertiary stellar radius of $\approx 0.2R_\odot$.  This
-scenario ultimately yields a contradiction, because it would require
-an ingress and egress phase that each span $\approx$40\% of the
-transit duration ($\approx 65\,{\rm minutes}$).  The actual measured
-ingress and egress duration is $\approx 15\,{\rm minutes}$),
-4.4$\times$ shorter.
-
-\paragraph{Stellar density implied by transit duration}
-The duration of the transit ($2.823 \pm 0.057 hr$) and the implied
-stellar density ($2.31 \pm 0.52\,{\rm g\,cm}^{-3}$) could in theory
-help rule between whether the transiting body orbits the primary or
-secondary star.  Ultimately, our stellar density measurement is not
-precise enough to render the blend scenario implausible.  At 30 Myr, a
-0.40\,$M_\odot$ solar-metallicity dwarf is $\approx 26\%$ larger than
-when it is fully contracted on the main sequence (0.40\,$R_\odot$
-vs{.} 0.50\,$R_\odot$; CITEALT: Choi et al, MIST grids).  The
-theoretically implied companion density of $3.2\,{\rm g\,cm}^{-3}$ is
-indeed larger than the primary star's density of $1.80\,{\rm
-g\,cm}^{-3}$ (measured through the HIRES reconaissance spectroscopy).
-However the stellar density measured from the transit fitting ($2.31
-\pm 0.52\,{\rm g\,cm}^{-3}$) is only discrepant at the
-$\approx$2-$\sigma$ level from the M-dwarf blend scenario.  It is
-instead the combination of the flux contrast, transit depth, and
-ingress duration that rule out this scenario.
-
-% \section{Bonus figures that wont be included}
-% 
-% \begin{figure*}[t]
-% 	\begin{center}
-% 		\leavevmode
-% 		\includegraphics[width=1\textwidth]{dontinclude0.png}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-% 	\caption{
-%     {\bf corner plot of gptransit fit}
-% 	}
-% \end{figure*}
-
 \listofchanges

### 2021-09-06 paper/collect.sh

@@ -4,6 +4,8 @@ cp /Users/luke/Dropbox/proj/rudolf/results/skychart/skychart_showkepler_showtess
 
 cp /Users/luke/Dropbox/proj/rudolf/results/keplerlc/keplerlc_Nsample500_xlim0500_0600.pdf f4.pdf
 
+cp /Users/luke/Dropbox/proj/rudolf/results/keplerlc/*.pdf FigureSetKeplerLC/.
+
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_posterior_phaseplot.pdf f5.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_cornerplot.png dontinclude0.png

### 2021-09-06 paper/ms.tex
update ms / Rp|P|age

@@
 closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
-cluster is \clusterage\ old, and that \sn\ is indeed a member
-of the cluster.  Focusing on the planet itself,
-we confirm that despite the presence of a previously unreported binary companion,
-the planet orbits the G-dwarf primary on an orbit that appears aligned
-with the stellar rotation axis
-(Section~\ref{sec:planet}).  We conclude by
+cluster is \clusterage\ old, and that \sn\ is indeed a member of the
+cluster.  Focusing on the planet itself, we confirm that despite the
+presence of a previously unreported binary companion, the planet
+orbits the G-dwarf primary on an orbit that appears aligned with the
+stellar rotation axis (Section~\ref{sec:planet}).  We conclude by
 highlighting broader implications for our ability to age-date a
@@
     {\bf The \cn\ is \clusterage\ old.}  {\it Top:}
-    Color-magnitude diagram of the \cn, IC\,2602 ($\approx38$\,Myr),
+    Color-magnitude diagram of the \cn, IC\,2602 ($\approx35$\,Myr),
     the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
-    Gaia EDR3 Catalog of Nearby Stars (gray background;
-    log-stretch density).  The right panel provides a zoomed view of stars on
+    Gaia EDR3 Catalog of Nearby Stars (gray background density;
+    log-stretch).  The right panel provides a zoomed view of stars on
     the pre-main-sequence.  The \cn\ and IC\,2602 are approximately
@@
 To summarize, the age of the \cn\ cluster based on the empirical
-procedure described in Section~\ref{sec:cmd} is
-\clusterage.  This procedure hinges on the age adopted for
-IC\,2602, since it is the closest reference cluster in the
-color-magnitude diagram.  Gyrochronologically, while most of the G and
-K dwarfs in the \cn\ have not yet converged to a slow sequence, their
-rotation periods and those of the F dwarfs are also consistent with a
-30 to 50\,Myr age. The most precise age for \sn\ and its
-planet are therefore also \clusterage\ Myr, based on the star's
-spatial, kinematic, and rotational connection to the cluster,
+pre-main-sequence interpolation is \clusterage.  This procedure hinges
+on the age adopted for IC\,2602, since it is the closest reference
+cluster in the color-magnitude diagram.  Gyrochronologically, while
+most of the G and K dwarfs in the \cn\ have not yet converged to a
+slow sequence, their rotation periods and those of the F dwarfs are
+also consistent with a 30 to 50\,Myr age. The most precise age for
+\sn\ and its planet are therefore also \clusterage\ Myr, based on the
+star's spatial, kinematic, and rotational connection to the cluster,
 and the assumption that the planet formed very quickly after the star.
@@
 of the light curve is shown with the phase-folded transit in
-Figure~\ref{fig:lc}.  The dominant signal in the \texttt{PDCSAP}
-photometry is a quasi-periodic starspot signal with a peak-to-peak
+Figure~\ref{fig:lc}.  The texttt{PDCSAP}
+photometry is dominated by a quasi-periodic starspot signal with a peak-to-peak
 amplitude that varies between 2\% and 8\%, depending on the degree of
@@
 Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
-planet ($3.34\pm 0.18\,R_\oplus$) on a circular\footnote{Note $e<0.48$
-at 2-$\sigma$; the constraints on the eccentricity are not
-particularly strong.} 7.2\,day orbit, around a G8V star ($0.88 \pm
-0.02 R_\odot$).
+planet ($3.34\pm 0.18\,R_\oplus$) on a close-in circular\footnote{Our
+transit fitting yields $e<0.48$ at 2-$\sigma$; the constraints on the
+eccentricity are not particularly strong.} orbit around a G8V star
+($0.88 \pm 0.02 R_\odot$).
 
@@
 orbiting this near-solar analog?  \citet{morton_false_2016} validated
-the planet based on the transit shape, arguing that the only plausible
+the planet based on the transit shape, arguing that the most probable 
 false positive scenario was that of a background eclipsing binary,
-reported to have a model-dependent probability of $\approx10^{-5}$.
+which had a model-dependent probability of $\approx10^{-5}$.
 As discussed in Appendix~\ref{app:companionstar},
 we discovered a low-mass stellar companion ($M_{\rm B}\approx
-0.44\,M_\odot$), and so here we reconsider this false positive scenario
+0.44\,M_\odot$), and so we reconsidered this false positive scenario
 in some detail.
@@
 
-Three new lines of evidence support the planetary interpretation.
-First, the transit duration and the orbital period are inconsistent
-with an eclipsing body around the M-dwarf companion: we find
-$\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
-theoretically expected density for the companion at 38\,Myr is
+We identified three new lines of evidence that support the planetary
+interpretation.  First, the transit duration and the orbital period
+are inconsistent with an eclipsing body around the M-dwarf companion:
+we find $\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
+theoretically expected density for the companion is
 $\approx 3.2\,{\rm g\,cm}^{-3}$ (Table~\ref{tab:posterior};
@@
 to explained by a star eclipsing the M dwarf secondary at
-$5$-$\sigma$ confidence.
+$\approx$$5$-$\sigma$.
 
@@
 
-The second line of confirmatory evidence for the planetary
-interpretation comes by analyzing the individual transit times in the
-Kepler data. We isolated each of the 144 observed transits to within
-$\pm4.5$\,hr of each transit, and fitted each window with both {\it
-i)} a local second-order polynomial and transit, and {\it ii)} a local
-linear trend.  We let the mid-time of each transit float, and then
-calculated the residual between the measured mid-time and that of a
-perfectly periodic orbit.  This residual, the transit timing variation
-(TTV), is plotted in Figure~\ref{fig:obliquity} against the local
-linear slope.  A significant correlation of $-0.0564 \pm 0.0098\ {\rm
-ppt\,day}^{-1}$ is observed.  Less than ten Kepler Objects of
-Interest have shown this correlation \citep{holczer_time_2015}, which
-is most readily interpreted as a TTV induced by unresolved starspot
-crossings \citep{mazeh_time_2015}.  This is only possible if the
-planet transits the primary star, which excludes a background
-eclipsing binary scenario, and therefore confirms that \pn\ is a
-planet.  It also implies that the planet's orbit is prograde.  The
-latter interpretation assumes that the dominant photometric
-variability is induced by dark spots, rather than bright faculae
+The second line of confirmatory evidence comes by analyzing the
+individual Kepler transit times. We isolated each of the
+144 observed transits to within $\pm4.5$\,hr of each transit, and
+fitted each window with both {\it i)} a local second-order polynomial
+and transit, and {\it ii)} a local linear trend.  We let the mid-time
+of each transit float, and then calculated the residual between the
+measured mid-time and that of a perfectly periodic orbit.  This
+residual, the transit timing variation (TTV), is plotted in
+Figure~\ref{fig:obliquity} against the local linear slope.  A
+significant correlation of $-0.0564 \pm 0.0098\ {\rm ppt\,day}^{-1}$
+is observed.  Less than ten Kepler Objects of Interest have shown this
+correlation \citep{holczer_time_2015}, which is most readily
+interpreted as a TTV induced by unresolved starspot crossings
+\citep{mazeh_time_2015}.  This is only possible if the planet transits
+the primary star, which excludes a background eclipsing binary
+scenario, and therefore confirms that \pn\ is a planet.  It also
+implies that the planet's orbit is prograde.  The latter
+interpretation assumes that the dominant photometric variability is
+induced by dark spots, rather than bright faculae
 \citep{mazeh_time_2015}.  Given the observed transition of Sun-like
@@
 
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f14.pdf}
+		\includegraphics[width=0.49\textwidth]{f14.pdf}
 	\end{center}
@@
     exoplanets}.  Planets older than $10^9$ years are shown with gray
-    markers.  Planets younger than a gigayear, with $t/\sigma_t > 3$,
+    markers.  Planets younger than a gigayear with ${\rm
+    Age}/\sigma_{\rm Age} > 3$
     are shown in color.  \pn\ is shown with a star.  The large radii
@@
 	}
-\end{figure*}
+\end{figure}
 
@@
 youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
-with precisely measured ages below 100\,Myr include K2-33b,
-DS~Tuc~Ab, HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet
-V1298~Tau system
+with precisely measured ages below 100\,Myr include K2-33b, DS~Tuc~Ab,
+HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet V1298~Tau system
 \citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020}.
 \pn\ is currently the smallest of these systems ($\approx
-3.3\,R_\oplus$), which may hint at selection effects in the 
-young planet sample \citep{zhou_2021_tois}.  Assuming that these planets
-have masses between $\approx$5 and 20$\,M_\oplus$, their large sizes
-and presumed low densities are in accord with the expectation that
-mini-Neptunes start their lives with large primordial atmospheres that
-are shed over the first gigayear
+3.3\,R_\oplus$), which may hint at selection effects in the young
+planet sample \citep{zhou_2021_tois}.  Assuming that these planets
+have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$, their
+large sizes and presumed low densities are in accord with the
+expectation that mini-Neptunes start their lives with large primordial
+atmospheres that are shed over the first gigayear
 \citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
-suggestion of an aligned orbit differs from observed misalignments in number of other close-in
+suggestion of an aligned orbit differs from observed misalignments in
+number of other close-in
 mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
-Unlike these older planets, the properties of \pn\
-seem to be consistent with a quiescent disk-driven migration history.
-The alternative scenario of in-situ formation is disfavored on
-theoretical grounds: the required disk mass would likely drive
-migration, and so lacks internal consistency
+Unlike these older planets, the properties of \pn\ seem to be
+consistent with a quiescent disk-driven migration history.  The
+alternative scenario of in-situ formation is disfavored on theoretical
+grounds: the required disk mass would likely drive migration, and so
+lacks internal consistency
 \citep{inamdar_formation_2015,ogihara_reassessment_2015}.

### 2021-09-06 drivers/plot_rp_vs_period_scatter.py

@@ -132,22 +132,21 @@ def plot_rp_vs_period_scatter(
         s1 = (has_factorN_age) & (age <= 1e9)
 
         ax.scatter(period[s0], rp[s0],
-                   color='darkgray', s=1.5, zorder=1, marker='o', linewidth=0,
+                   color='darkgray', s=2.5, zorder=1, marker='o', linewidth=0,
                    alpha=1, rasterized=True)
 
         # draw the colored points
         axins1 = inset_axes(ax, width="3%", height="33%", loc='lower right',
                             borderpad=0.7)
 
-        cmapname = 'viridis'
-        cmap = mpl.cm.viridis
-        bounds = np.arange(6.9,9.1,0.01)
+        cmap = mpl.cm.magma_r
+        bounds = np.arange(7.0,9.0,0.01)
         norm = mpl.colors.LogNorm(vmin=1e7, vmax=1e9)
         #norm = mpl.colors.BoundaryNorm(bounds, cmap.N, extend='neither')
 
         _p = ax.scatter(
             period[s1], rp[s1],
-            c=age[s1], alpha=1, zorder=2, s=17, edgecolors='k',
+            c=age[s1], alpha=1, zorder=2, s=24, edgecolors='k',
             marker='o', cmap=cmap, linewidths=0.3, norm=norm
         )
 
@@ -169,7 +168,7 @@ def plot_rp_vs_period_scatter(
 
                 ax.scatter(
                     _per, _rp,
-                    c=_age, alpha=1, zorder=2, s=150, edgecolors='k',
+                    c=_age, alpha=1, zorder=2, s=170, edgecolors='k',
                     marker='*', cmap=cmap, linewidths=0.3, norm=norm
                 )
 
@@ -179,7 +178,7 @@ def plot_rp_vs_period_scatter(
             namelist = ['Kepler-52', 'Kepler-968', 'Kepler-1627']
             ages = [3.5e8, 3.5e8, 3.5e7]
             markers = ['o','d','*']
-            sizes = [80, 80, 150]
+            sizes = [80, 80, 170]
 
             for n, a, m, _s in zip(namelist, ages, markers, sizes):
                 sel = ea_df.hostname == n

### 2021-09-06 paper/ms.tex
update ms

@@
   orbit aligned with the stellar spin axis.  We also briefly explore the flare
-  arrival time distribution in the short-cadence Kepler
+  arrival time distribution in the short cadence Kepler
   photometry.
@@
     Age}/\sigma_{\rm Age} > 3$
-    are shown in color.  \pn\ is shown with a star.  The large radii
+    are emphasized.  \pn\ is shown with a star.  The large radii
     of the youngest transiting planets could be explained by their
@@
 youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
-with precisely measured ages below 100\,Myr include K2-33b, DS~Tuc~Ab,
+with precise ages below 100\,Myr include K2-33b, DS~Tuc~Ab,
 HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet V1298~Tau system
@@
 mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
-Unlike these older planets, the properties of \pn\ seem to be
+Unlike these older planets, \pn\ seems to have properties
 consistent with a quiescent disk-driven migration history.  The
@@
 
-Ultimately though, the main advance of this work was a precise measurement of
-the age of \pn.   The age measurement was enabled by identifying the
+Ultimately, the main advance of this work was a precise measurement of
+the age of \pn.   This age measurement was enabled by identifying the
 connection of the star to the \cn\ using the Gaia data.  Table~\ref{tab:v05}
-enables similar crossmatches to
+should enable similar crossmatches to
 already known planets, as well as forthcoming planets from the TESS
@@
 
-\input{t1_appendix.tex}
-\input{t2_appendix.tex}
-
 The \texttt{v0.5} CDIPS target catalog (Table~\ref{tab:v05}) includes
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f15supp.pdf}
+		\includegraphics[width=0.49\textwidth]{f15supp.pdf}
 	\end{center}
@@
 \subsection{Long Cadence Data}
-\begin{figure*}[t]
+\begin{figure}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=1\textwidth]{f9.pdf}
+		\includegraphics[width=0.49\textwidth]{f9.pdf}
 	\end{center}
@@
 	\caption{
-		{\bf Transit model residuals through time}.  
+		{\bf Transit model residuals through time (binned by Kepler quarter)}.  
     {\it Left:}
-		Phase-folded transit of Kepler 1627b, with stellar
-    variability removed, binned by Kepler quarter.
-    Black points are binned to 20 minute intervals.
-    The 1-$\sigma$ model uncertainties and the maximum {\it a
-    posteriori} model are shown as the faint purple band, and the dark
-    purple line.
+    Phase-folded transit of Kepler 1627b, with stellar variability
+    removed.  Black points are binned to 20
+    minute intervals.  The 2-$\sigma$ model uncertainties and the
+    maximum {\it a posteriori} model are shown as the faint purple
+    band, and the dark purple line.
     {\it Right:}
-    As on the left, with the transit removed.
-    Quarters 6 and 7 show a consistent deviation in the second half of
-    the transit.
+    As on the left, with the transit removed.  Quarters 6 and 7 show a
+    consistent deviation in the second half of the transit.
 		\label{fig:phasequarter}
 	}
-\end{figure*}
-
-\begin{figure*}[t]
+\end{figure}
+
+\begin{figure}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=1\textwidth]{f10.pdf}
+		\includegraphics[width=0.49\textwidth]{f10.pdf}
 	\end{center}
@@
 	\caption{
-    {\bf Transit model residuals through time (Part 2)}.  
+    {\bf Transit model residuals through time (binned by year of observation)}.  
     {\it Left:}
-		Phase-folded transit of Kepler 1627b, with stellar
-    variability removed, binned by year of observation.
-    Points and models are as in Figure~\ref{fig:phasequarter}.
+    Phase-folded transit of Kepler 1627b, with stellar variability
+    removed.  Points and models are as
+    in Figure~\ref{fig:phasequarter}.
     {\it Right:}
@@
 	}
-\end{figure*}
-
-We assumed a quadratic limb-darkening law, with the uninformative prior
-advocated by CITET Kipping2013.
-
-% In the first approach, we adopted uniform priors centered on the
-% interpolated CITET Claret+2011 coefficients in the Kepler-band, with
-% with $\pm 0.2$ in either parameter.  The fit was not as good.
-
-Our default stellar variability model (RotGPtransit) allows for a
-\texttt{RotTerm} GP kernel, with a logjitter term to inflate the
-uncertainties to account for excess white noise.  (This results in an
-inflation of the uncertainties by a factor of about three).
-
-We considered including an additive \texttt{SHOTerm} kernel to account
-for stochastic noise (RotStochGPtransit).  This didn't seem to affect
-the results much, so we opted for the simpler model.
-
-Figure~\ref{fig:phasefold} shows X, Y, Z.  The residuals during the
-transit may hint at some small degree of unfitted signal -- in the
-sense that the observations are systematically high in the first half
-of transit, and low in the second half.  This is the phase of maximal
-scatter during each orbital period (TODO: in vs out of transit
-scatter).  The amplitude of the anomaly is $\approx$30\,ppm.
-
-To explore the origin of the anomaly, we binned the Kepler data over
-quarters (Figure~\ref{fig:phasequarter}) and years
-(Figure~\ref{fig:phaseyear}).  In Figure~\ref{fig:phasequarter}
-Quarter 6 shows the strongest asymmetry out of any of the quarters: a
-deviation of about 3\,ppt from expectation.  Quarter 7 shows an
-anomaly at roughly the same transit phase.  Year 2 correspondingly
-shows the strongest anomaly out of any year in
-Figure~\ref{fig:phaseyear}.
+\end{figure}
+
+We fitted the Kepler nce time series with a model that
+simultaneously included the planetary transit and the stellar
+variability.  The stellar variability was modelled with the
+\texttt{RotationTerm} Gaussian Process kernel in \texttt{exoplanet}
+\citep{exoplanet:exoplanet}.  This kernel assumes that the variability
+is generated by a mixture of two damped simple harmonic oscillators
+with characteristic frequencies set by 1/$P_{\rm rot}$ and its first
+harmonic.  We additionally incorporated a ``jitter'' term to inflate
+the flux uncertainties in a manner that accounted for otherwise
+unmodelled excess white noise, and let the eccentricity float.  For
+the limb-darkening, we assumed a quadratic law, and sampled using the
+uninformative prior suggested by \citet{exoplanet:kipping13}.
+
+Our adopted model therefore included 10 free parameters for the
+transit ($\{P,t_0,\log R_{\rm p}/R_\star,b,u_1,u_2,R_\star, \log g,
+e\cos\omega, e\sin\omega \}$), 2 parameters for the light curve
+normalization and uncertainty calibration ($\{\langle f \rangle,  \log
+\sigma_f \}$), and 5 hyperparameters for the GP
+($\{\sigma_{\mathrm{rot}}, \log P_{\mathrm{rot}}, \log Q_0, \log
+\mathrm{d}Q, f \}$).  We also considered including an additive
+\texttt{SHOTerm} kernel to account for stochastic noise, but found
+that this did not significantly affect the results, and so opted for
+the simpler GP kernel.  We fitted the models using \texttt{PyMC3}
+\citep{salvatier_2016_PyMC3,exoplanet:theano}.  After initializing
+each model with the parameters of the maximum {\it a posteriori}
+model, we assumed a Gaussian likelihood, and sampled using
+\texttt{PyMC3}'s gradient-based No-U-Turn Sampler
+\citep{hoffman_no-u-turn_2014}. We used $\hat{R}$ as our convergence
+diagnostic \citep{gelman_inference_1992}.
+
+Figure~\ref{fig:lc} shows the resulting model in orange (top) and
+purple (bottom).  While the fit appears reasonably good, an asymmetry
+is present during the transit: the measured flux is systematically
+high during the first half, and low in the second half.  We verified
+that this effect is not caused by our stellar variability model by
+exploring an alternative approach in which we isolated each transit
+window and locally fitted out polynomial trends, and then binned all
+the observed transits; the asymmetry was still present, at a
+comparable amplitude.
+
+Figures~\ref{fig:phasequarter} and~~\ref{fig:phaseyear} show the
+Kepler data binned over quarters and years.  In
+Figure~\ref{fig:phasequarter} Quarter 6 shows the strongest asymmetry
+out of any of the quarters: a deviation of about 3\,ppt from
+expectation.  Quarter 7 shows an anomaly at roughly the same transit
+phase.  Year 2 correspondingly shows the strongest anomaly out of any
+year in Figure~\ref{fig:phaseyear}.
 
@@
 darkening, transit timing variations, and spot-crossing events.
-
-Gravity darkening (e.g., CITE Masuda 2015) is based on the premise
-that the rapidly rotating star becomes oblate, and brighter near the
-poles than the equator.  The fractional shape change due to gravity
-darkening is on the order of $(P_{\rm break}/P_{\rm rot})^2$, for
-$P_{\rm break}$ the break-up rotation period, and $P_{\rm rot}$ the
-rotation period.  Using the parameters from Table~\ref{tab:posterior},
-this yields an expected 1.6\% distortion of the $\approx$1.8\,ppt
-transit depth: {\it i.e.}, an absolute deviation of $\approx$3\,ppm.
-This is smaller than the observed anomaly by an order of magnitude,
-and therefore seems unlikely.
+Gravity darkening is based on the premise that the rapidly rotating
+star becomes oblate, and brighter near the poles than the equator
+\citep[{\it e.g.},][]{masuda_spin-orbit_2015}.  The fractional shape
+change due to gravity darkening is on the order of $(P_{\rm
+break}/P_{\rm rot})^2$, for $P_{\rm break}$ the break-up rotation
+period, and $P_{\rm rot}$ the rotation period.  Using the parameters
+from Table~\ref{tab:posterior}, this yields an expected 1.6\%
+distortion of the $\approx$1.8\,ppt transit depth: {\it i.e.}, an
+absolute deviation of $\approx$3\,ppm.  The observed residual has an
+amplitude of $\approx 30\,{\rm ppm}$.  Since the expected signal is
+smaller than the observed anomaly by an order of magnitude, gravity
+darkening seems to be an unlikely explanation.
 
@@
 
-The final possibility is that of starspot crossings.  Given the high
-expected spot-covering fractions for young stars ({\it e.g.},
-\citealt{morris_relationship_2020}, \citealt{plavchan_planet_2020}),
-\pn\ may cross spot groups on the stellar surface in projection.
-Spot-crossing anomalies often reach amplitudes exceeding 100\,ppm
-({\it e.g.}, CITE Dai 2018, maybe Southworth's recent stuff).  For our
-system,  $P_{\rm orb}/P_{\rm rot} \approx 2.76$.  This means that
-every 4 transits (and 11 stellar rotations), the planet crosses over
-roughly the same stellar longitude.  Over a given quarter, this could
-occur at most three times, assuming the spot groups are persistent.
-
-This could provide a plausible path to creating the distorted transit
-signal, based on the amplitudes involved.  However, the typical S/N
-per Kepler transit is $\approx8$, making individual spot-crossing
-unresolved, and this scenario challenging to test.  {\it A priori},
-one would also expect the spot-crossing events to not have a preferred
-orbital phase, so that they would average out over the $\approx200$
-Kepler transits.  Nonetheless, given the available data, it is our
-best explanation for what we believe is real additional photometric
-scatter observed during the transits.
-
-\subsection{Short Cadence Data}
-
-\begin{figure*}[t]
+The final possibility is that of starspot crossings.  Young stars are
+thought to have higher spot-covering fractions than old stars ({\it
+e.g.}, \citealt{morris_relationship_2020},
+\citealt{plavchan_planet_2020}).  Young solar-type stars may also host
+dark starspots at high stellar latitudes \citep[{\it e.g.},
+EK~Dra][]{strassmeier_starspots_2009}, though interferometric imaging
+of spotted giant stars has found different starspot latitude distributions than
+those inferred from Doppler imaging
+\citep{roettenbacher_contemporaneous_2017}.  Regardless, for any
+spot-crossing anomalies to add coherently over the 144 Kepler
+transits, it seems likely that we would need either for spots to be
+persistent at a particular latitude (and for a slightly misaligned
+orbit), or for a ``stroboscopic'' longitudinal phasing \citep[{\it
+e.g.},][]{dai_stellar_2018}.  For our system, $P_{\rm orb}/P_{\rm rot}
+\approx 2.76$, which means that every 4 transits and 11 stellar
+rotations, the planet crosses over roughly the same stellar longitude,
+which could enable the necessary phasing if the spot-groups are
+long-lived.  Unforunately, the typical S/N per Kepler transit is
+$\approx8$, which renders individual spot-crossing events unresolved.
+Therefore, although among the possible ideas starspot crossings seem
+the most likely to provide the additional photometric scatter observed
+during transits, our reasoning is essentially that the expected
+spot-crossing anomaly amplitudes ($\approx 100\,{\rm ppm}$) could be
+correct; we are not able to definitely show that this is the correct
+explanation, and the issue could deserve further exploration.
+
+%  \subsection{Short Cadence Data}
+%  
+%  \begin{figure*}[t]
+%  	\begin{center}
+%  		\leavevmode
+%  		\includegraphics[width=1\textwidth]{f12.pdf}
+%  	\end{center}
+%  	\vspace{-0.7cm}
+%  	\caption{
+%  		{\bf Phase-folded short cadence transit of Kepler\,1627 Ab}.  
+%      The precision with which the impact parameter can be measured is
+%      higher than from the long cadence data.
+%  		\label{fig:phase_slc}
+%  	}
+%  \end{figure*}
+%  
+%  Figure~\ref{fig:phase_slc} shows the result from analyzing 100 days of
+%  short cadence (1-minute) data acquired during Quarter 15.
+%  %TODO: get quantitative on better impact parameter...
+
+\section{Companion Star and False Positive Assessment}
+\label{app:companionstar}
+
+\begin{figure*}[tp]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=1\textwidth]{f12.pdf}
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f17a.pdf}
+			\includegraphics[width=0.37\textwidth]{f17b.jpg}
+		}
+
+		\vspace{-0.1cm}
+		\subfloat{
+			\includegraphics[width=0.55\textwidth]{f6.pdf}
+		}
 	\end{center}
@@
 	\caption{
-		{\bf Phase-folded short cadence transit of Kepler\,1627 Ab}.  
-    The precision with which the impact parameter can be measured is
-    higher than from the long cadence data.
-		\label{fig:phase_slc}
-	}
-\end{figure*}
-
-Figure~\ref{fig:phase_slc} shows the result from analyzing 100 days of
-short cadence (1-minute) data acquired during Quarter 15.
-%TODO: get quantitative on better impact parameter...
-
-\section{Companion Star and False Positive Assessment}
-\label{app:companionstar}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.6\textwidth]{Kepler1627_20210624_562_832_final.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Keck NIRC2 AO image of Kepler 1627A and Kepler 1627B.}  
-    {\bf SCALEBAR DENOTES...}
-    {\bf North is up, East is left}.
-    \label{fig:ao}
-	}
-\end{figure*}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.6\textwidth]{f6.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Gaia EDR3 renormalized unit weight error (RUWE) point
-    estimates for Kepler\,1627A and other members of the $\delta$\,Lyr
-    cluster.}  Since other members of the cluster with similar colors
-    have comparable degrees of photometric variability, the high RUWE
-    suggests that Kepler\,1627A is an astrometric binary. 
-		\label{fig:ruwe}
+    {\bf Kepler 1627 is a binary.} {\it Top:} High-resolution imaging
+    from Gemini-North/`Alopeke ({\it left}) and Keck/NIRC2 ({\it
+    right}) shows an $\approx$M2V companion at $\rho \approx
+    0\farcs16$, which corresponds to a projected separation of
+    $53\pm7$\,AU.  {\it Bottom:} Gaia EDR3 renormalized unit weight
+    error (RUWE) point estimates for Kepler\,1627A and other members
+    of the $\delta$\,Lyr cluster.  Since other members of the cluster
+    with similar colors have comparable degrees of photometric
+    variability, the high RUWE suggests that Kepler\,1627 is an
+    astrometric binary. 
+    \label{fig:kep1627binary}
 	}
@@
 the same companion in the 0.83\,$\mu$m notch filter, but not the
-$0.56\,\mu$m filter.  Figure~\ref{fig:ao} summarizes the results of
-the high-resolution imaging.  Although the companion is unresolved in
-the Gaia source catalog, its existence was also suggested by the
-primary star's large RUWE, relative to other members of the \cn.  The
-images yield a projected separation $\rho \approx 0\farcs16$, with
-$\Delta m_{832} \approx 3.1$ and $\Delta m_{\rm Kp} = 2.5$.  Using the
-measured Gaia EDR3 parallax for the system, this implies a projected
-separation of $53 \pm 7$.  Given the low geometric probably of a star
-imaged within $\rho \approx 0\farcs16$ to be a chance companion along
-the line of sight, and the large observed RUWE, we proceed under the
-assumption that the companion is bound, and that the Kepler\,1627
-system is binary.  Based on the
+$0.56\,\mu$m filter.  Figure~\ref{fig:kep1627binary} summarizes the
+results of the high-resolution imaging.  Although the companion is
+unresolved in the Gaia source catalog, its existence was also
+suggested by the primary star's large RUWE, relative to other members
+of the \cn.  The images yield a projected separation $\rho \approx
+0\farcs16$, with $\Delta m_{832} \approx 3.1$ and $\Delta m_{\rm Kp} =
+2.5$.  Using the measured Gaia EDR3 parallax for the system, this
+implies a projected separation of $53 \pm 7$.  Given the low geometric
+probably of a star imaged within $\rho \approx 0\farcs16$ to be a
+chance companion along the line of sight, and the large observed RUWE,
+we proceed under the assumption that the companion is bound, and that
+the Kepler\,1627 system is binary.  Based on the
 tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
@@
     {\it Top:}
-    The full short-cadence Kepler dataset, acquired at 1-minute
+    The full short cadence Kepler dataset, acquired at 1-minute
     sampling (black points) is shown with a stellar variability model
@@
     Colors indicate relative time, from the beginning of the
-    98-day short-cadence Quarter 15 dataset (dark blue) to the end (light
+    98-day short cadence Quarter 15 dataset (dark blue) to the end (light
     yellow).
@@
 
-The short cadence (1-minute) Kepler observations span 97.7 days.
-The light curve shows a significant number of flares
-(Figure~\ref{fig:flarezoom}).
-To quantify clean the light curve and identify the flares, we
-performed the following procedure.
+In addition to the 3.9 years of long cadence data, short cadence
+(1-minute) Kepler observations were acquired over 97.7 days during
+Quarter 15.  The short cadence light curve shows a higher rate of
+flaring than visible in the long cadence data
+(Figure~\ref{fig:flarezoom}).  We analyzed the short cadence light
+curve and its flares according to the following procedure.
+
 For cleaning, we performed the following iterative detrending procedure.
-\begin{itemize}
-  \item Step 1: Build a two-term mixed SHOTerm GP model with
-    quasi-periodic kernels at Prot and 0.5$\times$Prot. Fit the model
-    to the time, flux, and flux uncertainty.
-  \item Step 2: Select points more than twice the median absolute
+\begin{enumerate}
+  \item Model the starspot-induced variability using a
+    Gaussian Process with a \texttt{SHOTerm} kernel (see
+    Section~\ref{app:gptransit}), a white-noise jitter term, and the
+    mean flux.  Fit the model to the time, flux, and flux uncertainty.
+  \item Select points more than twice the median absolute
     deviation from the residual, and exclude them from the light
-    curve.  Repeat Step 1.
-  \item Step 3: On the residual from Step 2, identify all flares,
+    curve (these points include the flares).  Repeat Step~1.
+  \item Using the residual from Step 2, identify all flares,
     requiring them to be at least 20 cadences apart, at least 7 median
-    absolute deviations above the median baseline, and lasting at least
-    2 cadences in duration.  Build the mask spanning these times, from
-    5 minutes before each flare begins to 2.5 minutes after the final
-    flare cadence.  Repeat Step 1 a final time.
-\end{itemize}
-The flares were then identified and fitted using ALTAIPONY (CITE CITE).
-The fitted model is that of CITE XXX, which parametrizes the flare
-with a start time, a lag time, and an amplitude (CHECK).
+    absolute deviations above the median baseline, and lasting at
+    least 2 cadences in duration.  Build the mask spanning these
+    times, from 5 minutes before each flare begins to 2.5 minutes
+    after the final flare cadence.  Repeat Step 1 a final time.
+\end{enumerate}
+The flares were identified and fitted using \texttt{altaipony}
+\citep{davenport_2016,ilin_flares_2021}.  The analytic flare model is
+from \citet{davenport_2014} and it parametrizes the flare with a start
+time, an exponential lag time, and an amplitude.
 Figure~\ref{fig:flarephase} shows the resulting flares, amplitudes,
@@
 
-There were $N_{\rm f}=24$ flares exceeding $0.5\%$ in relative flux during the
-short-cadence observations.  These 24 flares spanned a total of 6.5
-hours ($\sim$15 minutes per flare).  For better or worse, we noticed a
-coincidence in the flare arrival times.  The coincidence is that
-despite the low flare duty cycle, one orbital period after the
-brightest flare, a second flare followed.  This and a similar event
-are shown in Figure~\ref{fig:flarezoom}.  The timing error is good to
-a $\approx0.2\%$ difference from the orbital period, which seems {\it
-a priori} unlikely.  If we consider flares falling within 2\% of the
-planet's orbital period after a previous flare, then 4 of the 24 flare
-events have candidate ``successors''.
-
-As with any coincidence, when one does not have a firm prediction, it
-is difficult to assess the statistical significance of a surprise.
-Since our surprise was specifically at the inter-arrival time of
-certain flares coinciding with special time intervals, we performed
-the following analysis.  First, we considered all unordered pairs of
+There were $N_{\rm f}=24$ flares that exceeded $0.5\%$ in relative
+flux during the short cadence observations.  These 24 flares spanned a
+total of 6.5 hours ($\sim$15 minutes per flare).  Whether it is
+meaningful or not, we noticed a coincidence in the flare arrival
+times.  The coincidence is that despite the low flare duty cycle, one
+orbital period after the brightest flare, a second flare followed.
+This and a similar event are shown in Figure~\ref{fig:flarezoom}.  The
+timing error is good to a $\approx0.2\%$ difference from the orbital
+period, which given the duty cycle seems {\it a priori} unlikely.  If
+we consider flares falling within 2\% of the planet's orbital period
+after a previous flare, then 4 of the 24 flare events have candidate
+``successors''.
+
+As with any coincidence, if one does not have a firm prediction, it is
+difficult to assess the statistical significance of a surprise.  Since
+our surprise was specifically at the inter-arrival time of certain
+flares coinciding with special time intervals, we performed the
+following analysis.  First, we considered all unordered pairs of
 flares.  For $N$ flares there are ${n \choose 2}$ such pairs (for our
@@
 these ``special intervals'', the statistical evidence for a
-non-Poissonian process driving the flares does not seem particularly
+non-Poissonian process driving the flares does not seem especially
 strong.  More quantitatively, the peaks observed at the orbital and
-synodic periods are within the $\pm 2$-$\sigma$ range of a
-Poissonian process, and those at $P_{\rm orb}+P_{\rm syn}$ and $P_{\rm
-orb}+2P_{\rm syn}$ are only slightly above this range. 
-With that said,
-future analyses of these data by investigators more versed in the
-topic than ourselves could very well yield new insights.
+synodic periods are within the $\pm 2$-$\sigma$ range of a Poissonian
+process, and those at $P_{\rm orb}+P_{\rm syn}$ and $P_{\rm
+orb}+2P_{\rm syn}$ are only slightly above this range.  With that
+said, future analyses of these data by investigators more versed in
+this topic than ourselves could very well yield more detailed insights.
+
+\input{t1_appendix.tex}
+\input{t2_appendix.tex}
 

### 2021-09-06 paper/collect.sh

@@ -36,3 +36,7 @@ cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_
 cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_emphbinaries_talkaspect_kinematicspatialselected_linear_curtiscleaning.pdf f15supp.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/RM/quadratic/rm_RV_and_phot.pdf f16.pdf
+
+
+cp /Users/luke/Dropbox/proj/rudolf/data/imaging/Kepler1627_20210624_562_832_final.pdf f17a.pdf
+cp /Users/luke/Dropbox/proj/rudolf/data/imaging/KOA_104524/NIRC2/calibrated/N2.20150722.39302_drp.jpg f17b.jpg

### 2021-09-07 paper/ms.tex
update ms

@@
   Verifying the results of Gaia clustering analyses using TESS rotation
-  periods seems likely to be an effective path toward expanding the census of
+  periods seems likely to be a path toward expanding the census of
   age-dated planets, and our understanding of planetary evolution.
@@
 Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet
-\citep[see {\it e.g.},]{soderblom_ages_2014}.
+\citep[see {\it e.g.},][]{soderblom_ages_2014}.
 We acquired an iodine-free spectrum from
@@
 		\subfloat{
-			\includegraphics[width=0.9\textwidth]{f4.pdf}
+			\includegraphics[width=\textwidth]{f4.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.87\textwidth]{f5.pdf}
+			\includegraphics[width=\textwidth]{f5.pdf}
 		}
@@
     variability (orange line) is subtracted below, revealing the
-    transits of \pn, as well as other deviations from the stellar
-    variability model.  The Figure Set available online shows the
+    transits of \pn, as well as other deviations from the
+    model.  The Figure Set online shows the
     entire 3.9 years of observations.
     {\it Bottom}:
-    Phase-folded transit of Kepler 1627b, with stellar variability
+    Phase-folded transit of Kepler 1627Ab, with stellar variability
     removed.  The 2-$\sigma$ model uncertainties and the maximum {\it
@@
 of the light curve is shown with the phase-folded transit in
-Figure~\ref{fig:lc}.  The texttt{PDCSAP}
+Figure~\ref{fig:lc}.  The \texttt{PDCSAP}
 photometry is dominated by a quasi-periodic starspot signal with a peak-to-peak

### 2021-09-07 paper/ms.tex
update star table

@@
 knowledge of open cluster and moving group memberships \citep[{\it
-e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,kerr_stars_2021}.
+e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,Kerr2021}.
 As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
@@
 the age of \pn.   This age measurement was enabled by identifying the
-connection of the star to the \cn\ using the Gaia data.  Table~\ref{tab:v05}
+connection of the star to the \cn\ using the Gaia data.  Table~\ref{tab:v06}
 should enable similar crossmatches to
@@
 transit fitting, to R{.}~Kerr for kindly providing us with the
-\citet{kerr_stars_2021} membership list prior to its publication.
+\citet{Kerr2021} membership list prior to its publication.
 L.G.B{.}, L.H{.}, and N.N{.} are also grateful to K{.}~Collins for
@@
 \input{model_posterior_table.tex}
+\input{t1_appendix.tex}
+\input{t2_appendix.tex}
 
@@
 
-The \texttt{v0.5} CDIPS target catalog (Table~\ref{tab:v05}) includes
+The \texttt{v0.6} CDIPS target catalog (Table~\ref{tab:v06}) includes
 some important updates from previous versions.  As in
@@
 
-\input{t1_appendix.tex}
-\input{t2_appendix.tex}
-
 \listofchanges

### 2021-09-07 paper/t1_appendix.tex
update star table

@@
 \tablecaption{Young, Age-dated, and Age-dateable Stars Within the
-  Nearest Few Kiloparsecs ($\texttt{v0.5}$ of the CDIPS Target
-  List).}
-\label{tab:v05}
+  Nearest Few Kiloparsecs.}
+\label{tab:v06}
 
@@
 % paste from
-% /Users/luke/Dropbox/proj/rudolf/results/tables/v05_main_tableheader.tex
-% via drivers/write_v05_main_tableheader.py
+% /Users/luke/Dropbox/proj/rudolf/results/tables/v06_main_tableheader.tex
+% via drivers/write_v06_main_tableheader.py
 \startdata
@@
      \texttt{parallax\_error} &                                                        0.028 &                                     Gaia DR2 parallax uncertainty [mas]. \\
-               \texttt{pmra} &                                                      94.884 &     Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr${^-1}$]. \\
-              \texttt{pmdec} &                                                     -86.971 &                 Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr${^-1}$]. \\
+                \texttt{pmra} &                                                       94.884 &      Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr$^{-1}$]. \\
+               \texttt{pmdec} &                                                      -86.971 &                  Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr$^{-1}$]. \\
   \texttt{phot\_g\_mean\_mag} &                                                         6.85 &                                                  Gaia DR2 $G$ magnitude. \\
@@
  \texttt{phot\_rp\_mean\_mag} &                                                        7.189 &                                      Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
-            \texttt{cluster} &                 Uma,IR\_excess,NASAExoArchive\_ps\_20210506 &                                  Comma-separated cluster or group name. \\
-                \texttt{age} &                                                nan,nan,9.48 & Comma-separated logarithm (base-10) of reported$^{\rm a}$ age in years. \\
+             \texttt{cluster} &                  NASAExoArchive\_ps\_20210506,Uma,IR\_excess &                                   Comma-separated cluster or group name. \\
+                 \texttt{age} &                                                 9.48,nan,nan &  Comma-separated logarithm (base-10) of reported$^{\rm a}$ age in years. \\
            \texttt{mean\_age} &                                                         9.48 &                           Mean (ignoring NaNs) of $\texttt{age}$ column. \\
-      \texttt{reference\_id} &      Ujjwal2020,CottenSong2016,NASAExoArchive\_ps\_20210506 &                          Comma-separted provenance of group membership. \\
- \texttt{reference\_bibcode} & 2020AJ....159..166U,2016ApJS..225...15C,2013PASP..125..989A &                  ADS bibcode corresponding to $\texttt{reference\_id}$. \\
+       \texttt{reference\_id} &       NASAExoArchive\_ps\_20210506,Ujjwal2020,CottenSong2016 &                           Comma-separted provenance of group membership. \\
+  \texttt{reference\_bibcode} &  2013PASP..125..989A,2020AJ....159..166U,2016ApJS..225...15C &                   ADS bibcode corresponding to $\texttt{reference\_id}$. \\
 \enddata
@@
 \tablecomments{
-Table~\ref{tab:v05} is
-published in its entirety in a machine-readable format.   This table is a
-concatenation of the studies listed in Table~\ref{tab:metadata}.
-One entry is
-shown for guidance regarding form and content. 
-In this particular example, the star has a
-cold Jupiter on a 16 year orbit, HD 150706b \citep{2012AA...545A..55B}.
-An infrared excess has been reported \citep{CottenSong2016}, and the star was identified by \citet{Ujjwal2020} as a candidate UMa moving group
-member ($\approx 400\,{\rm Myr}$; \citealt{mann_tess_2020}).
-The star's RV activity and TESS rotation period corroborate its youth.
+Table~\ref{tab:v06} is published in its entirety in a machine-readable
+format.   This table is a concatenation of the studies listed in
+Table~\ref{tab:metadata}.  One entry is shown for guidance regarding
+form and content.  In this particular example, the star has a cold
+Jupiter on a 16 year orbit, HD 150706b \citep{2012AA...545A..55B}.  An
+infrared excess has been reported \citep{CottenSong2016}, and the star
+was identified by \citet{Ujjwal2020} as a candidate UMa moving group
+member ($\approx 400\,{\rm Myr}$; \citealt{mann_tess_2020}).  The
+star's RV activity and TESS rotation period corroborate its youth.
 }

### 2021-09-07 drivers/write_metadata_table.py

@@ -3,9 +3,9 @@ import re, os
 # https://ads.readthedocs.io/en/latest/#getting-started
 import ads
 
-from rudolf.paths import RESULTSDIR
+from rudolf.paths import RESULTSDIR, PAPERDIR
 
-csvpath = '/Users/luke/Dropbox/proj/cdips/data/cluster_data/list_of_lists_keys_paths_assembled_v0.5_gaiasources.csv'
+csvpath = '/Users/luke/Dropbox/proj/cdips/data/cluster_data/list_of_lists_keys_paths_assembled_v0.6_gaiasources.csv'
 df = pd.read_csv(csvpath)
 
 outcols = (
@@ -37,7 +37,7 @@ MAPDICT = {
     'TTS':r'$\texttt{TT*}$',
     'candTTS':r'$\texttt{TT?}$',
     'gt250':r'$d>250\,{\rm pc}$',
-    'lt250':r'$d>250\,{\rm pc}$',
+    'lt250':r'$d<250\,{\rm pc}$',
     'ums':'UMS',
     'pms':'PMS',
     'PscEri': 'Psc-Eri',

### 2021-09-07 paper/ms.tex
update ms

@@
   Kepler\,1627 is a member of the \clusterage\ old $\delta$~Lyr
-  cluster using kinematic information from Gaia, stellar rotation
-  periods from TESS, and Li abundances from ground-based spectra.  To
+  cluster, based on Gaia kinematics, TESS stellar rotation
+  periods, and spectroscopic lithium abundances.  To
   our knowledge, Kepler\,1627Ab is the youngest planet with a precise
@@
   The $\delta$~Lyr cluster is one of many stellar groups whose
-  properties have been clarified using the Gaia data; many other new
+  properties have been clarified using Gaia; many other new
   clusters also overlap with confirmed and candidate exoplanets.
   Verifying the results of Gaia clustering analyses using TESS rotation
-  periods seems likely to be a path toward expanding the census of
-  age-dated planets, and our understanding of planetary evolution.
+  periods seems to be a promising path toward expanding the census of
+  age-dated planets, and our corresponding understanding of planetary evolution.
 \end{abstract}
@@
     {\bf The \cn\ is \clusterage\ old.}  {\it Top:}
-    Color-magnitude diagram of the \cn, IC\,2602 ($\approx35$\,Myr),
+    Color-magnitude diagram of the \cn, IC\,2602 ($\approx38$\,Myr),
     the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
-    Gaia EDR3 Catalog of Nearby Stars (gray background density;
-    log-stretch).  The right panel provides a zoomed view of stars on
+    Gaia EDR3 Catalog of Nearby Stars (gray background;
+    log-stretch).  The right panel is a zoomed view of 
     the pre-main-sequence.  The \cn\ and IC\,2602 are approximately
@@
     and Praesepe (650\,Myr) shown for reference
-    \citep{rebull_rotation_2016a,douglas_poking_2017}.  The \cn\ is
-    gyrochronally younger than the Pleiades.
+    \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most of the
+    candidate \cn\ members are gyrochronally younger than the
+    Pleiades.
    \label{fig:age}
@@
     30 minute cadence.  A one hundred day segment is shown here.  The
-    {\it top panel} shows the \texttt{PDCSAP} median-subtracted flux
-    in units of parts-per-thousand ($\times 10^{-3}$).  The dominant
+    top panel shows the \texttt{PDCSAP} median-subtracted flux in
+    units of parts-per-thousand ($\times 10^{-3}$).  The dominant
     signal is induced by starspots.  The model for the stellar
     variability (orange line) is subtracted below, revealing the
-    transits of \pn, as well as other deviations from the
-    model.  The Figure Set online shows the
-    entire 3.9 years of observations.
+    transits of \pn, as well as other deviations from the model.  The
+    Figure Set online shows the entire 3.9 years of observations.
     {\it Bottom}:
@@
     during the transit is robust against detrending methods; we
-    believe it is caused by starspot crossings.
+    believe that it is caused by starspot crossings.
     \label{fig:lc}
@@
     \citep{mazeh_time_2015}.  The units along the abscissa are most
-    easily understood by considering that the flux from the star
+    easily understood by considering that the stellar flux
     changes by $\sim$60\,ppt per half rotation period
     ($\sim$1.3\,days).
-    {\it Bottom:} Radial velocities from Keck/HIRES during a transit
-    on 7 Aug 2021 UT, observed simultaneously with MuSCAT3 ({\it griz}
-    photometry).  A quadratic baseline trend was allowed in the RV
-    fit.  The best-fit sky-projected obliquity in this model is listed
-    in red; the detection however has marginal statistical
+    {\it Bottom:} Radial velocities acquired using Keck/HIRES during
+    the transit of 7 Aug 2021 UT, observed simultaneously with MuSCAT3
+    ({\it griz} photometry).  A quadratic baseline trend was allowed
+    in the RV fit.  The best-fit sky-projected obliquity in this model
+    is listed in red; the detection however has marginal statistical
     significance compared to a ``trend-only'' model.
@@
 Kepler\,1627Ab provides a new extremum in the ages of the Kepler
-planets, and opens multiple avenues for further study.  More
-precise observations of the Doppler transit could yield a more precise
-stellar obliquity, and a means for verifying the expected prograde
-orbit.  Similar transit
-spectroscopy aimed at detecting either hydrogen or helium absorption
-could yield insight into the stability of the planet's atmosphere
-\citep[{\it e.g.},][]{spake_helium_2018,vissapragada_2020}.  A more
-challenging measurement would be to determine the planet's mass,
-either through detailed timing analyses of the Kepler data, or through
-a high-cadence multi-color radial velocity campaign.  The mass,
-measured at sufficient precision, could yield constraints on both the
+planets, and opens multiple avenues for further study.  More precise
+observations of the Doppler transit could yield a more precise stellar
+obliquity, and a means for verifying the expected prograde orbit.
+Similarly, transit spectroscopy aimed at detecting either hydrogen or
+helium absorption could yield insight into the stability of the
+planet's atmosphere \citep[{\it
+e.g.},][]{spake_helium_2018,vissapragada_2020}.  A more challenging
+measurement would be to determine the planet's mass, either through
+detailed timing analyses of the Kepler data, or through a high-cadence
+multi-color radial velocity campaign.  Measured at
+sufficient precision, the mass could yield constraints on both the
 planet's composition as well as its initial entropy
@@
 
-The precision of the Kepler data itself may enable more detailed
-analyses without needing to acquire more data.  While our transit
-timing study did not yield the detection of any clearly significant
-planet-planet induced TTVs, the star's flaring may be suggestive of
-star-planet interactions.  Appendix~\ref{app:flare} highlights an odd
-flare timing coincidence that may be worthy of further examination.
+Separately, the Kepler data may still contain further information.
+Although our transit timing study did not yield the detection of any
+clearly significant planet-planet TTVs, the star's flaring may be
+suggestive of star-planet interactions.  Appendix~\ref{app:flare}
+highlights an odd flare timing coincidence that may be worthy of
+further examination.
 
@@
 planet sample \citep{zhou_2021_tois}.  Assuming that these planets
-have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$, their
-large sizes and presumed low densities are in accord with the
+have masses between $\approx$5$\,M_\oplus$ and 20$\,M_\oplus$, their
+large sizes and presumed low densities would be in accord with the
 expectation that mini-Neptunes start their lives with large primordial
@@
 connection of the star to the \cn\ using the Gaia data.  Table~\ref{tab:v06}
-should enable similar crossmatches to
+should enable similar cross-matches to
 already known planets, as well as forthcoming planets from the TESS
@@
   %\texttt{arviz} \citep{arviz_2019},
+  \texttt{altaipony} \citep{ilin_flares_2021},
   \texttt{astrobase} \citep{bhatti_astrobase_2018},
@@
 The \texttt{v0.6} CDIPS target catalog (Table~\ref{tab:v06}) includes
-some important updates from previous versions.  As in
-\citet{bouma_cdipsI_2019}, we collected membership information for
-young, age-dated, or age-dateable stars from across the literature.
-Table~\ref{tab:metadata} gives a list of the sources included, and
-some brief summary statistics.
-
-The first major important change is that the extent of analyses
-performed on the Gaia data at the time of our compilation was wide and
-deep enough that we opted to neglect pre-Gaia analyses, except in
-cases for which spectroscopically confirmed samples of stars had been
-collected.  The membership lists for instance of
-\citet{Kharchenko_et_al_2013} and \citet{dias_proper_2014} (MWSC and
-DAML) were no longer required, especially given their relatively high
-field-star contamination rates compared to Gaia-derived membership
+stars that are young, age-dated, and age-dateable.  By ``age-dateable'',
+we mean that the age should be measurable at greater
+precision than a typical FGK field star.
+As in
+\citet{bouma_cdipsI_2019}, we collected stars that met these criteria
+from across the literature.  Table~\ref{tab:metadata} gives a list of
+the studies included, and some brief summary statistics.
+The age measurement approaches adopted by each study differ: in many,
+spatial and kinematic clustering has been performed on the Gaia data,
+and ensemble isochrone fitting of the cluster has been performed
+(typically focusing on the turn-off).
+In others however, the claim of youth is based on the location of a
+single star in the color-magnitude diagram, or on spectroscopic
+information.
+
+One major change in Table~\ref{tab:v06} relative to the analogous
+table from \citet{bouma_cdipsI_2019} is that the extent of Gaia-based
+analyses has now matured to the point that we can neglect pre-Gaia
+cluster memberships, except for a few cases with spectroscopically
+confirmed samples of age-dated stars.  The membership lists for
+instance of \citet{Kharchenko_et_al_2013} and \citet{dias_proper_2014}
+(MWSC and DAML) are no longer required.  This is helpful for various
+post-processing projects,  since the field-star contamination rates
+were typically much higher than analogous Gaia-based cluster membership
 catalogs.
 
-For any of the catalogs for which Gaia DR2 identifiers were not
-immediately available, we either followed the spatial (plus
-proper-motion) crossmatching procedures described in
-\citet{bouma_cdipsI_2019}, or else we pulled the Gaia DR2 source
-identifiers associated with the catalog from SIMBAD.  We consequently
-opted to drop the $\texttt{ext\_catalog\_name}$ and $\texttt{dist}$
-columns maintained in \citet{bouma_cdipsI_2019}, as these were only
-populated for a small number of stars.
-
 The most crucial parameters of a given star for our purposes are the
-Gaia DR2 source identifier ($\texttt{source\_id}$), the cluster name
-($\texttt{cluster}$), and the ($\texttt{age}$).  Given the
+Gaia DR2 $\texttt{source\_id}$, the cluster or
+group name ($\texttt{cluster}$), and the $\texttt{age}$.  Given the
 hierarchical nature of many stellar associations, we do not attempt to
 resolve the cluster names to a single unique string.  The Orion
-complex for instance, can be divided into almost one hundred kinematic
+complex for instance can be divided into almost one hundred kinematic
 subgroups \citep{kounkel_apogee2_2018}.  Similar complexity applies to
@@
 to resolve.  Instead, we simply merged the cluster names and ages
-reported by various authors together.
-
-This means that our ``age'' column can be null, for cases in which the
-original authors did not report an age, and a reference literature age
-was not readily available.  Nonetheless, since we do generally prefer
-stars with known ages, we made a few additional efforts to populate
-this column.  When available, the age provenance is from the original
-analysis of the cluster.  However, in a few cases we adopted other
-ages when the string-based crossmatches on the ``cluster'' name was
-straightforward.  In particular, we used the ages determined by
-\citet{CantatGaudin2020b} to assign ages to the catalogs from
-\citet{GaiaCollaboration2018}, \citet{CantatGaudin2018a},
+reported by various authors into a comma-separated string.
+
+This means that our \texttt{age} column can be null, for cases in
+which the original authors did not report an age, and a reference
+literature age was not readily available.  Nonetheless, since we do
+generally prefer stars with known ages, we made a few additional
+efforts to populate this column.  When available, the age provenance
+is from the original analysis of the cluster.  In a few cases
+however we adopted other ages when string-based cross-matches on the
+cluster name was straightforward.  In particular, we used the ages
+determined by \citet{CantatGaudin2020b} to assign ages to the catalogs
+from \citet{GaiaCollaboration2018}, \citet{CantatGaudin2018a},
 \citet{CastroGinard2020}, and \citet{CantatGaudin2020a}.
@@
 were those of \citet{CottenSong2016}, \citet{Oh2017},
-\citet{Zari2018}, \citet{Gagne2018a},
-\citet{Gagne2018b},\citet{Gagne2018c}, and \citet{Ujjwal2020}.  While
+\citet{Zari2018}, \citet{Gagne2018b},
+\citet{Gagne2018a}, \citet{Gagne2018c}, and \citet{Ujjwal2020}.  While
 in principle the moving group members discussed by
-\citet{Gagne2018a,Gagne2018b,Gagne2018c} and \citet{Ujjwal2020} have
-easily associated ages, our SIMBAD cross-matching lost the moving
-group association from those studies, which should therefore be
-recovered tools such as BANYAN
+\citet{Gagne2018b,Gagne2018a,Gagne2018c} and \citet{Ujjwal2020} have
+easily associated ages, our SIMBAD cross-match lost the moving group
+identifiers given by those studies, which should therefore be recovered
+tools such as BANYAN
 $\Sigma$.\footnote{\url{http://www.exoplanetes.umontreal.ca/banyan/banyansigma.php}}.
@@
 $\texttt{Y*O}, $\texttt{Y*?}, $\texttt{TT?}$, and $\texttt{pMS*}$.
-Finally, we also included every star in the NASA Exoplanet Archive
+Finally, we  included every star in the NASA Exoplanet Archive
 $\texttt{ps}$ table that had a Gaia identifier available
@@
 
+For any of the catalogs for which Gaia DR2 identifiers were not
+immediately available, we either followed the spatial (plus
+proper-motion) cross-matching procedures described in
+\citet{bouma_cdipsI_2019}, or else we pulled the Gaia DR2 source
+identifiers associated with the catalog from SIMBAD.  We consequently
+opted to drop the $\texttt{ext\_catalog\_name}$ and $\texttt{dist}$
+columns maintained in \citet{bouma_cdipsI_2019}, as these were only
+populated for a small number of stars.
 The technical manipulations for the merging, cleaning, and joining
 were performed using $\texttt{pandas}$
-\citep{mckinney-proc-scipy-2010}.  The eventual crossmatch (using the
-Gaia DR2 $\texttt{source\_id}$) against the Gaia DR2 archive was performed
-asychronously on the Gaia archive
+\citep{mckinney-proc-scipy-2010}.  The eventual cross-match (using the
+Gaia DR2 $\texttt{source\_id}$) against the Gaia DR2 archive was
+performed asychronously on the Gaia archive
 website\footnote{\url{https://gea.esac.esa.int/archive/}}.
@@
 		\leavevmode
-		\includegraphics[width=0.49\textwidth]{f15supp.pdf}
+		\includegraphics[width=0.7\textwidth]{f15supp.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.49\textwidth]{f9.pdf}
+		\includegraphics[width=0.64\textwidth]{f9.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.49\textwidth]{f10.pdf}
+		\includegraphics[width=0.64\textwidth]{f10.pdf}
 	\end{center}
@@
 
-Our adopted model therefore included 10 free parameters for the
-transit ($\{P,t_0,\log R_{\rm p}/R_\star,b,u_1,u_2,R_\star, \log g,
-e\cos\omega, e\sin\omega \}$), 2 parameters for the light curve
-normalization and uncertainty calibration ($\{\langle f \rangle,  \log
+Our model therefore included 10 free parameters for the
+transit ($\{P, t_0, R_{\rm p}/R_\star, b, u_1 ,u_2 ,R_\star, \log g,
+e, \omega \}$), 2 parameters for the light curve
+normalization and uncertainty calibration ($\{\langle f \rangle, 
 \sigma_f \}$), and 5 hyperparameters for the GP
-($\{\sigma_{\mathrm{rot}}, \log P_{\mathrm{rot}}, \log Q_0, \log
+($\{\sigma_{\mathrm{rot}}, P_{\mathrm{rot}}, Q_0,
 \mathrm{d}Q, f \}$).  We also considered including an additive
@@
 \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
-\citep{hoffman_no-u-turn_2014}. We used $\hat{R}$ as our convergence
+\citep{hoffman_no-u-turn_2014} in these bases indicated in Table~\ref{tab:posterior}.
+We used $\hat{R}$ as our convergence
 diagnostic \citep{gelman_inference_1992}.
@@
 Kepler data binned over quarters and years.  In
-Figure~\ref{fig:phasequarter} Quarter 6 shows the strongest asymmetry
-out of any of the quarters: a deviation of about 3\,ppt from
+Figure~\ref{fig:phasequarter}, Quarter 6 shows the strongest asymmetry
+out of the quarters: a deviation of about 3\,ppt from
 expectation.  Quarter 7 shows an anomaly at roughly the same transit
@@
 thought to have higher spot-covering fractions than old stars ({\it
-e.g.}, \citealt{morris_relationship_2020},
-\citealt{plavchan_planet_2020}).  Young solar-type stars may also host
+e.g.}, \citealt{morris_relationship_2020}).  Young solar-type stars may also host
 dark starspots at high stellar latitudes \citep[{\it e.g.},
-EK~Dra][]{strassmeier_starspots_2009}, though interferometric imaging
+EK~Dra;][]{strassmeier_starspots_2009}, though interferometric imaging
 of spotted giant stars has found different starspot latitude distributions than
@@
 the most likely to provide the additional photometric scatter observed
-during transits, our reasoning is essentially that the expected
+during transits, our reasoning is that the expected
 spot-crossing anomaly amplitudes ($\approx 100\,{\rm ppm}$) could be
-correct; we are not able to definitely show that this is the correct
+correct; we have not been able to definitely show that this is the correct
 explanation, and the issue could deserve further exploration.
@@
     error (RUWE) point estimates for Kepler\,1627A and other members
-    of the $\delta$\,Lyr cluster.  Since other members of the cluster
+    of the $\delta$\,Lyr cluster reported by \citet{KounkelCovey2019}.  Since other members of the cluster
     with similar colors have comparable degrees of photometric
@@
 transit signal.
+Two other arguments presented in the main text -- the inferred stellar
+density, and the TTV-local slope correlation -- confirm this point.
 
@@
     with the gray bands.  While peaks in the observed distribution do
-    coincide with the locations of these ``special periods'', the
+    coincide with some of the ``special periods'', the
     statistical evidence for a non-Poissonian process driving the
@@
 curve and its flares according to the following procedure.
-
-For cleaning, we performed the following iterative detrending procedure.
 \begin{enumerate}
@@
 said, future analyses of these data by investigators more versed in
-this topic than ourselves could very well yield more detailed insights.
+this topic than ourselves could very well yield more quantitative insights.
 

### 2021-09-07 rudolf/plotting.py

@@ -132,12 +132,12 @@ def plot_ruwe_vs_apparentmag(outdir):
 
     ax.scatter(
         get_xval(df_edr3), get_yval(df_edr3), c='k', alpha=0.9,
-        zorder=4, s=5, rasterized=True, linewidths=0, label='Theia 73 (KC19)', marker='.'
+        zorder=4, s=5, rasterized=True, linewidths=0, label='$\delta$ Lyr cluster', marker='.'
     )
     ax.plot(
         get_xval(trgt_df), get_yval(trgt_df), alpha=1, mew=0.5,
-        zorder=8, label='Kepler 1627', markerfacecolor='yellow',
-        markersize=20, marker='*', color='black', lw=0
+        zorder=8, label='Kepler 1627A', markerfacecolor='yellow',
+        markersize=12, marker='*', color='black', lw=0
     )
 
     leg = ax.legend(loc='upper left', handletextpad=0.1, fontsize='x-small',

### 2021-09-09 paper/ms.tex
update ms

@@
 J.~Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
-cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
+cluster Theia~520 (UBC~1).  At least six Kepler planets are known in
 the cluster from the Kepler-52 and Kepler-968 systems
@@
 presence of a previously unreported binary companion, the planet
-orbits the G-dwarf primary on an orbit that appears aligned with the
+orbits the G-dwarf primary on an orbit that is likely aligned with the
 stellar rotation axis (Section~\ref{sec:planet}).  We conclude by
@@
 
-The final piece of evidence \pn\ comes from a spectroscopic
-transit we measured from Keck/HIRES during the night of 7 Aug 2021
-(UT), using the iodine cell for the wavelength calibration.  We
-scheduled the transit using the ephemeris of
-\citet{holczer_transit_2016}, which we verified using the available
-TESS data.  We also acquired simultaneous multi-band photometry using
-MuSCAT3 to help interpret the radial velocities \citep{Narita_2020}.
+The final piece of evidence comes from a spectroscopic transit we
+observed using Keck/HIRES during the night of 7 Aug 2021 (UT).  We
+used the iodine cell for the wavelength calibration, and scheduled the
+transit using the ephemeris of \citet{holczer_transit_2016}.  We also acquired
+simultaneous multi-band photometry using MuSCAT3 \citep{Narita_2020}.
 The lower panels in Figure~\ref{fig:obliquity} show the results.  The
@@
 and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
-The radial velocities show significant scatter, and a significant
-increase in RV is seen shortly after the expected ingress.  The
-internal uncertainties on the RVs ($\approx 15\,{\rm m\,s}^{-1}$) is
+The radial velocities show significant scatter, and an
+increase in RV is seen shortly after the expected ingress time.  The
+internal uncertainty on the RVs ($\approx 15\,{\rm m\,s}^{-1}$) is
 dominated by the $v\sin i$ broadening.
@@
 term.  The quadratic model with the RM effect is shown in
-Figure~\ref{fig:obliquity}; the prior on $v\sin i$ has been
-assumed from Table~\ref{tab:starparams}.  We found $\chi^2_{\rm red} =
+Figure~\ref{fig:obliquity}; the prior on $v\sin i$ has been assumed
+from Table~\ref{tab:starparams}, and the jitter term is included.
+We found $\chi^2_{\rm red} =
 5.4$ for the best-fitting linear case, and $\chi^2_{\rm red} = 4.4$
@@
 over the linear one, the $\chi^2_{\rm red}$ indicate that neither fit
-is very good.  Considering a quadratic trend alone without any RM
+is very good.  Considering a quadratic trend without any RM
 effect yields $\chi^2_{\rm red} = 4.3$, and $\Delta {\rm BIC} = {\rm
@@
 AIC}=5.2$).  From a model-comparison standpoint, there is therefore
-only weak statistical evidence to prefer the best-fitting ``RM and
+only marginal statistical evidence to prefer the best-fitting ``RM and
 quadratic trend'' model over the best-fitting ``quadratic trend only''
 model.  Given the expected semi-amplitude of the velocity anomaly
-($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) and the actual
-achieved precision, the claim of having detected the Doppler transit
-is marginal at best.  From the MuSCAT3 data however, we do know that
-the transit was observed, and the increase in radial velocity seen
-shortly after ingress is more consistent with a prograde or polar
-orbit than a retrograde orbit.  Fixing the orbit to be either exactly
-prograde or exactly retrograde, we find $\Delta {\rm BIC} = {\rm
-BIC}_{\rm retrograde} - {\rm BIC}_{\rm trend} = 18.0$; the retrograde
-case is ruled out on this basis.  More precise data will be needed
-however to make a more precise obliquity measurement.
+($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) and the achieved precision,
+we cannot claim a detection of the Doppler transit.
+From the MuSCAT3 data however, we do know that the transit
+was observed, and the increase in radial velocity seen shortly after
+ingress is more consistent with a prograde or polar orbit than a
+retrograde orbit.  Fixing the orbit to be either exactly prograde or
+exactly retrograde, we find $\Delta {\rm BIC} = {\rm BIC}_{\rm
+retrograde} - {\rm BIC}_{\rm trend} = 18.0$; the retrograde case is
+ruled out on this basis.  More data will be needed however to
+make a more precise measurement of the stellar obliquity.
 
@@
     markers.  Planets younger than a gigayear with ${\rm
-    Age}/\sigma_{\rm Age} > 3$
-    are emphasized.  \pn\ is shown with a star.  The large radii
-    of the youngest transiting planets could be explained by their
-    primordial atmospheres not yet having evaporated; direct
-    measurements of the atmospheric outflows or planetary masses would
-    help to confirm this expectation.  Selection effects may also be
-    important.  Parameters are from the NASA Exoplanet Archive (July
-    2, 2021).
+    Age}/\sigma_{\rm Age} > 3$ are emphasized.  \pn\ is shown with a
+    star.  The large radii of the youngest transiting planets could be
+    explained by their primordial atmospheres not yet having
+    evaporated; direct measurements of the atmospheric outflows or
+    planetary masses would help to confirm this expectation.
+    Selection effects may also be important.  Parameters are from the
+    NASA Exoplanet Archive (July 2, 2021).
 		\label{fig:rp_period_age}
@@
 Kepler\,1627Ab provides a new extremum in the ages of the Kepler
-planets, and opens multiple avenues for further study.  More precise
-observations of the Doppler transit could yield a more precise stellar
-obliquity, and a means for verifying the expected prograde orbit.
-Similarly, transit spectroscopy aimed at detecting either hydrogen or
-helium absorption could yield insight into the stability of the
-planet's atmosphere \citep[{\it
-e.g.},][]{spake_helium_2018,vissapragada_2020}.  A more challenging
-measurement would be to determine the planet's mass, either through
-detailed timing analyses of the Kepler data, or through a high-cadence
-multi-color radial velocity campaign.  Measured at
-sufficient precision, the mass could yield constraints on both the
-planet's composition as well as its initial entropy
-\citep{owen_constraining_2020}.
-
-Separately, the Kepler data may still contain further information.
+planets, and opens multiple avenues for further study.  Further
+observations of Doppler transits could yield a more precise stellar
+obliquity, and could confirm or refute the expected prograde orbital
+geometry.  Similarly, transit spectroscopy aimed at detecting outflows
+of hydrogen or helium could yield insight into the evolutionary state
+of the atmosphere \citep[{\it
+e.g.},][]{spake_helium_2018,vissapragada_2020}.  Determining the
+precise level of high-energy irradiation incident on the planet would
+also help in clarifying the expected outflow rate \citep[{\it
+e.g.},][]{poppenhaeger_2021}.  Finally, perhaps the most challenging,
+but also the most informative measurement would be to determine the
+planet's mass.  This might be possible either through detailed timing
+analyses of the Kepler data, or through a high-cadence multi-color
+radial velocity campaign.  Measured at sufficient precision, the mass
+could yield constraints on both the planet's composition and its
+initial entropy \citep{owen_constraining_2020}.
+
+Separately, the Kepler data may yet contain further information.
 Although our transit timing study did not yield the detection of any
@@
 planet sample \citep{zhou_2021_tois}.  Assuming that these planets
-have masses between $\approx$5$\,M_\oplus$ and 20$\,M_\oplus$, their
+have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$, their
 large sizes and presumed low densities would be in accord with the
@@
 suggestion of an aligned orbit differs from observed misalignments in
-number of other close-in
+number of close-in
 mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
-Unlike these older planets, \pn\ seems to have properties
-consistent with a quiescent disk-driven migration history.  The
-alternative scenario of in-situ formation is disfavored on theoretical
-grounds: the required disk mass would likely drive migration, and so
-lacks internal consistency
+Unlike these older planets, \pn\ seems to have properties consistent
+with a quiescent disk-driven migration history.  The alternative
+scenario of in-situ formation is disfavored on theoretical grounds:
+the required disk mass would likely drive migration, and so lacks
+internal consistency
 \citep{inamdar_formation_2015,ogihara_reassessment_2015}.
@@
 Ultimately, the main advance of this work was a precise measurement of
-the age of \pn.   This age measurement was enabled by identifying the
-connection of the star to the \cn\ using the Gaia data.  Table~\ref{tab:v06}
-should enable similar cross-matches to
-already known planets, as well as forthcoming planets from the TESS
-mission \citep{ricker_transiting_2015,guerrero_tess_2021}.  An
-alternative path toward age-dating more planets could be to identify kinematic associations
-around known exoplanet host stars using positions and tangential velocities
-from Gaia, and to then verify these associations with ancilliary
-rotation periods and spectroscopy \citep[{\it e.g.},][]{tofflemire_tess_2021}.
+the age of \pn.   The age measurement was enabled by identifying the
+connection of the star to the \cn\ using the Gaia data.
+Table~\ref{tab:v06} should enable similar cross-matches to already
+known planets, as well as forthcoming planets from the TESS mission
+\citep{ricker_transiting_2015,guerrero_tess_2021}.  An alternative
+path toward age-dating more planets could be to identify kinematic
+associations around known exoplanet host stars using positions and
+tangential velocities from Gaia, and to then verify these associations
+with ancilliary rotation periods and spectroscopy \citep[{\it
+e.g.},][]{tofflemire_tess_2021}.
 
@@
 
-L.G.B{.} and J.L.C{.} are grateful to T{.}~David for help with the
-transit fitting, to R{.}~Kerr for kindly providing us with the
-\citet{Kerr2021} membership list prior to its publication.
+The authors are grateful to T{.}~David for help with the transit
+fitting, to R{.}~Kerr for kindly providing us with the
+\citet{Kerr2021} membership list prior to its publication, and to
+J{.}~Winn for helpful discussions on the RM observations and flare
+timing.
 L.G.B{.}, L.H{.}, and N.N{.} are also grateful to K{.}~Collins for
@@
  	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
- 	Gemini:North~(`Alopeke; \citealt{scott_nessi_2018}.
+ 	Gemini:North~(`Alopeke; \citealt{scott_nessi_2018,scott_twin_2021}.
  	{\it Spectroscopy}:
@@
     0\farcs16$, which corresponds to a projected separation of
-    $53\pm7$\,AU.  {\it Bottom:} Gaia EDR3 renormalized unit weight
+    $53\pm4$\,AU.  {\it Bottom:} Gaia EDR3 renormalized unit weight
     error (RUWE) point estimates for Kepler\,1627A and other members
-    of the $\delta$\,Lyr cluster reported by \citet{KounkelCovey2019}.  Since other members of the cluster
-    with similar colors have comparable degrees of photometric
-    variability, the high RUWE suggests that Kepler\,1627 is an
-    astrometric binary. 
+    of the $\delta$\,Lyr cluster reported by \citet{KounkelCovey2019}.
+    Since other members of the cluster with similar colors have
+    comparable degrees of photometric variability, the high RUWE
+    suggests that Kepler\,1627 is an astrometric binary. 
     \label{fig:kep1627binary}
@@
 2.5$.  Using the measured Gaia EDR3 parallax for the system, this
-implies a projected separation of $53 \pm 7$.  Given the low geometric
-probably of a star imaged within $\rho \approx 0\farcs16$ to be a
-chance companion along the line of sight, and the large observed RUWE,
-we proceed under the assumption that the companion is bound, and that
-the Kepler\,1627 system is binary.  Based on the
+implies a projected separation of $53 \pm 4$.  Given the large
+observed RUWE of the primary, and the low geometric probability that a
+companion imaged within $\rho \approx 0\farcs16$ is a chance
+line-of-sight companion, we proceed under the assumption that the
+companion is bound, and that Kepler\,1627 system is a binary.  Based
+on the
 tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},

### 2021-09-09 drivers/calc_apparent_separation.py

@@ -1,6 +1,6 @@
 import numpy as np
 rho_arcsec = 0.16
-e_rho_arcsec = 0.02
+e_rho_arcsec = 0.01
 d_pc = 329.5
 e_d_pc = 3.5
 

### 2021-09-09 rudolf/plotting.py

@@ -2560,9 +2560,17 @@ def plot_RM_and_phot(outdir, model=None):
 
     ax = axs[0]
 
-    # first data
+    ## first data
+    # ax.errorbar(
+    #     scale_x(TF.lpf.data['x']), TF.lpf.data['y'], TF.lpf.data['error'],
+    #     marker='o', elinewidth=0.5, capsize=4, lw=0, mew=0.5, color='k',
+    #     markersize=3, zorder=5
+    # )
+    # NOTE: just show the jitter included
+    sigma_rv = float(df_medvals[df_medvals.Labels == 'sigma_rv'].medvals)
     ax.errorbar(
-        scale_x(TF.lpf.data['x']), TF.lpf.data['y'], TF.lpf.data['error'],
+        scale_x(TF.lpf.data['x']), TF.lpf.data['y'],
+        np.sqrt(TF.lpf.data['error']**2 + sigma_rv**2),
         marker='o', elinewidth=0.5, capsize=4, lw=0, mew=0.5, color='k',
         markersize=3, zorder=5
     )

### 2021-09-13 paper/ms.tex
update ms

@@
 %\suppressAffiliations
-%\NewPageAfterKeywords
+\NewPageAfterKeywords
 \input{authors.tex}
@@
     {\bf The \cn\ is \clusterage\ old.}  {\it Top:}
-    Color-magnitude diagram of the \cn, IC\,2602 ($\approx38$\,Myr),
-    the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
-    Gaia EDR3 Catalog of Nearby Stars (gray background;
-    log-stretch).  The right panel is a zoomed view of 
-    the pre-main-sequence.  The \cn\ and IC\,2602 are approximately
-    the same isochronal age.  {\it Bottom:} TESS and Kepler stellar
-    rotation periods versus dereddened Gaia colors, with the Pleiades
-    and Praesepe (650\,Myr) shown for reference
-    \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most of the
+    Color-absolute magnitude diagram of candidate \cn\ members, in addition to
+    stars in IC\,2602 ($\approx38$\,Myr), the Pleiades ($\approx
+    115$\,Myr; \citealt{dahm_2015}), and the Gaia EDR3 Catalog of
+    Nearby Stars (gray background).  The right panel is a
+    zoomed view of the pre-main-sequence.  The \cn\ and IC\,2602 are
+    approximately the same isochronal age.  {\it Bottom:} TESS and
+    Kepler stellar rotation periods versus dereddened Gaia colors,
+    with the Pleiades and Praesepe (650\,Myr) shown for reference
+    \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most
     candidate \cn\ members are gyrochronally younger than the
-    Pleiades.
+    Pleiades; outliers are probably field interlopers.
    \label{fig:age}
@@
 
-\subsection{Color-Magnitude Diagram}
-\label{sec:cmd}
+\subsection{Color-Absolute Magnitude Diagram}
+\label{sec:camd}
 
 We measured the isochrone age through an empirical approach.  The left
-panel of Figure~\ref{fig:age} shows the color-magnitude diagram (CMD)
+panel of Figure~\ref{fig:age} shows the color-absolute magnitude
+diagram (CAMD)
 of stars in the \cn, IC\,2602, the Pleiades, and the field.  The
@@
 of each cluster's distribution, and manually excluded stars that
-were obvious photometric binaries in the CMD.  We then passed a moving
-box average and standard deviation across the CMD in 0.10\,mag bins,
+were obvious photometric binaries in the CAMD.  We then passed a moving
+box average and standard deviation across the CAMD in 0.10\,mag bins,
 fitted a univariate spline to the binned values, and assembled a
@@
 as well, given ${\rm [Fe/H] = +0.1}$ and $A_V=0.2$
-\citep{bressan_parsec_2012,choi_mesa_2016};  this preferred CMD reddening
+\citep{bressan_parsec_2012,choi_mesa_2016};  this preferred CAMD reddening
 is higher than the \citet{lallement_threedimensional_2018} value by a factor of two.
@@
 \citep{douglas_stephanie_t_2021_5131306}. 
-
 Binarity is known to be an important factor in the density of stars
@@
 
-To summarize, the age of the \cn\ cluster based on the empirical
-pre-main-sequence interpolation is \clusterage.  This procedure hinges
-on the age adopted for IC\,2602, since it is the closest reference
-cluster in the color-magnitude diagram.  Gyrochronologically, while
-most of the G and K dwarfs in the \cn\ have not yet converged to a
-slow sequence, their rotation periods and those of the F dwarfs are
-also consistent with a 30 to 50\,Myr age. The most precise age for
-\sn\ and its planet are therefore also \clusterage\ Myr, based on the
-star's spatial, kinematic, and rotational connection to the cluster,
-and the assumption that the planet formed very quickly after the star.
+In summary, the age of the \cn\ cluster based on the empirical
+pre-main-sequence interpolation method is \clusterage.  This method
+hinges on the age adopted for IC\,2602, since it is the closest
+reference cluster in the color-absolute magnitude diagram.
+Gyrochronologically, while most of the FGK dwarfs in the \cn\ have
+not yet converged to a slow sequence, their rotation periods 
+are generally consistent with an $\approx$38\,Myr age. The most
+precise age for \sn\ and its planet are therefore identical to the cluster,
+based on the star's spatial and kinematic association with the
+cluster, and the assumption that the planet formed shortly after
+the star.
 
@@
 \sn, there is a final consistency check: the abundance of
-Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet
-\citep[see {\it e.g.},][]{soderblom_ages_2014}.
-We acquired an iodine-free spectrum from
-Keck/HIRES on the night of 26 March 2021 using the standard setup and
-reduction techniques of the California Planet Survey
+Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet \citep[see {\it
+e.g.},][]{soderblom_ages_2014}.  We acquired an iodine-free spectrum
+from Keck/HIRES on the night of 26 March 2021 using the standard setup
+and reduction techniques of the California Planet Survey
 \citep{howard_cps_2010}.  Following the equivalent width measurement
@@
 EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   Note that this value does not
-correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given the stellar
-parameters of \sn\ (Table~\ref{tab:starparams}), most notably its
-effective temperature, this measurement is also in agreement with an
-age for the star between 30\,Myr and 50\,Myr ({\it e.g.}, 
-\citealt{randich_gaiaeso_2018}; \citealt{bouma_2021_ngc2516}).
+correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given the
+effective temperature of \sn\ (Table~\ref{tab:starparams}), this
+measurement is also in line with expectations for a $\approx$38\,Myr
+star ({\it e.g.}, \citealt{randich_gaiaeso_2018};
+\citealt{bouma_2021_ngc2516}).
 
@@
 which had a model-dependent probability of $\approx10^{-5}$.
-As discussed in Appendix~\ref{app:companionstar},
-we discovered a low-mass stellar companion ($M_{\rm B}\approx
+As we describe in Appendix~\ref{app:companionstar} however,
+the system has a low-mass stellar companion ($M_{\rm B}\approx
 0.44\,M_\odot$), and so we reconsidered this false positive scenario
@@
 
-We identified three new lines of evidence that support the planetary
+There are three new lines of evidence that support the planetary
 interpretation.  First, the transit duration and the orbital period
@@
 \citealt{choi_mesa_2016}).  The transit duration is therefore too long
-to explained by a star eclipsing the M dwarf secondary at
+to be explained by a star eclipsing the M dwarf secondary at
 $\approx$$5$-$\sigma$.
@@
 of each transit float, and then calculated the residual between the
-measured mid-time and that of a perfectly periodic orbit.  This
+measured mid-time and that of a periodic orbit.  This
 residual, the transit timing variation (TTV), is plotted in
@@
 The final piece of evidence comes from a spectroscopic transit we
-observed using Keck/HIRES during the night of 7 Aug 2021 (UT).  We
+observed using Keck/HIRES during the night of 7 Aug 2021.  We
 used the iodine cell for the wavelength calibration, and scheduled the
@@
 Figure~\ref{fig:obliquity}; the prior on $v\sin i$ has been assumed
-from Table~\ref{tab:starparams}, and the jitter term is included.
+from Table~\ref{tab:starparams}, and the jitter term has been
+incorpoarted.
 We found $\chi^2_{\rm red} =
@@
 retrograde} - {\rm BIC}_{\rm trend} = 18.0$; the retrograde case is
-ruled out on this basis.  More data will be needed however to
-make a more precise measurement of the stellar obliquity.
+ruled out on this basis.
+% More data will be needed however to
+% make a more precise measurement of the stellar obliquity.
 
@@
 e.g.},][]{spake_helium_2018,vissapragada_2020}.  Determining the
-precise level of high-energy irradiation incident on the planet would
+precise amount of high-energy irradiation incident on the planet would
 also help in clarifying the expected outflow rate \citep[{\it
-e.g.},][]{poppenhaeger_2021}.  Finally, perhaps the most challenging,
-but also the most informative measurement would be to determine the
-planet's mass.  This might be possible either through detailed timing
-analyses of the Kepler data, or through a high-cadence multi-color
-radial velocity campaign.  Measured at sufficient precision, the mass
-could yield constraints on both the planet's composition and its
-initial entropy \citep{owen_constraining_2020}.
+e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
+quantity to measure would be the planet's mass.  This might be
+possible either through detailed timing analyses of the Kepler data,
+or through a high-cadence multi-color radial velocity campaign.
+Measured at sufficient precision, the mass could yield constraints on
+both the planet's composition and its initial entropy
+\citep{owen_constraining_2020}.
 
@@
 Although our transit timing study did not yield the detection of any
-clearly significant planet-planet TTVs, the star's flaring may be
-suggestive of star-planet interactions.  Appendix~\ref{app:flare}
-highlights an odd flare timing coincidence that may be worthy of
-further examination.
-
-In the context of the known transiting planets, \pn\ is among the
+clearly significant planet-planet TTVs, Appendix~\ref{app:flare}
+highlights an odd flare timing coincidence that may be suggestive of
+star-planet interactions.  We mention it here to encourage its
+exploration by investigators more versed in the topic than ourselves.
+
+In the context of the transiting planet population, \pn\ is among the
 youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
@@
 \citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020}.
-\pn\ is currently the smallest of these systems ($\approx
-3.3\,R_\oplus$), which may hint at selection effects in the young
+\pn\ is currently the smallest of these systems ($3.34\pm
+0.18\,R_\oplus$), which may hint at selection effects in the young
 planet sample \citep{zhou_2021_tois}.  Assuming that these planets
@@
 
-Ultimately, the main advance of this work was a precise measurement of
-the age of \pn.   The age measurement was enabled by identifying the
+Ultimately, the main advance of this work is a precise measurement of
+the age of \pn.   This age measurement was enabled by identifying the
 connection of the star to the \cn\ using the Gaia data.
@@
 tangential velocities from Gaia, and to then verify these associations
-with ancilliary rotation periods and spectroscopy \citep[{\it
+with ancillary rotation periods and spectroscopy \citep[{\it
 e.g.},][]{tofflemire_tess_2021}.
@@
 timing.
-L.G.B{.}, L.H{.}, and N.N{.} are also grateful to K{.}~Collins for
+The authors are also grateful to K{.}~Collins for
 helping resolve the scheduling conflict that would have otherwise
@@
 %
-L.G.B{.} acknowledges support by the TESS GI Program, programs
+L.G.B{.} acknowledges support from a Charlotte Elizabeth Procter
+Fellowship from Princeton University, as well as from the TESS GI Program, programs
 G011103 and G022117, through NASA grants 80NSSC19K0386 and
@@
 %
-L.G.B{.} was also supported by a Charlotte Elizabeth Procter
-Fellowship from Princeton University.
-%
-The archival Keck/NIRC2 imaging was acquired by programs 2015A/N301N2L
+%
+Keck/NIRC2 imaging was acquired by programs 2015A/N301N2L
 (PI: A.~Kraus) and 2019A/N069 (PI: E.~Petigura).
@@
 from across the literature.  Table~\ref{tab:metadata} gives a list of
-the studies included, and some brief summary statistics.
-The age measurement approaches adopted by each study differ: in many,
+the studies included, and brief summary statistics.
+The age measurement methodologies adopted by each study differ: in many,
 spatial and kinematic clustering has been performed on the Gaia data,
-and ensemble isochrone fitting of the cluster has been performed
+and ensemble isochrone fitting of the resulting clusters has been performed
 (typically focusing on the turn-off).
-In others however, the claim of youth is based on the location of a
-single star in the color-magnitude diagram, or on spectroscopic
+In other studies however, the claim of youth is based on the location of a
+single star in the color-absolute magnitude diagram, or on spectroscopic
 information.
@@
 
-This means that our \texttt{age} column can be null, for cases in
-which the original authors did not report an age, and a reference
+This means that the \texttt{age} column can be null, for cases in
+which the original authors did not report an age, or for which a reference
 literature age was not readily available.  Nonetheless, since we do
@@
 is from the original analysis of the cluster.  In a few cases
-however we adopted other ages when string-based cross-matches on the
+however we adopted other ages when string-based cross-matching on the
 cluster name was straightforward.  In particular, we used the ages
-determined by \citet{CantatGaudin2020b} to assign ages to the catalogs
+determined by \citet{CantatGaudin2020b} to assign ages to the clusters
 from \citet{GaiaCollaboration2018}, \citet{CantatGaudin2018a},
@@
 \citet{Gagne2018b,Gagne2018a,Gagne2018c} and \citet{Ujjwal2020} have
-easily associated ages, our SIMBAD cross-match lost the moving group
+easily associated ages, our SIMBAD cross-match did not retain the moving group
 identifiers given by those studies, which should therefore be recovered
@@
 Finally, we  included every star in the NASA Exoplanet Archive
-$\texttt{ps}$ table that had a Gaia identifier available
+planetary system ($\texttt{ps}$) table that had a Gaia identifier available
 \citep{NASAExoArchive_ps_20210506}.  If the age had finite
@@
 three-dimensional spatial velocity as \sn, where we assume a systemic
-radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the reconnaissance
-spectra obtained by A.~Howard on HIRES and D.~Latham on TRES.  The
+radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the HIRES
+spectra.  The
 relevant projection effects are then taken into account, as discussed
@@
 
-We fitted the Kepler nce time series with a model that
+We fitted the Kepler long cadence time series with a model that
 simultaneously included the planetary transit and the stellar
@@
 with characteristic frequencies set by 1/$P_{\rm rot}$ and its first
-harmonic.  We additionally incorporated a ``jitter'' term to inflate
+harmonic.  We additionally included a jitter term to inflate
 the flux uncertainties in a manner that accounted for otherwise
@@
 transit ($\{P, t_0, R_{\rm p}/R_\star, b, u_1 ,u_2 ,R_\star, \log g,
-e, \omega \}$), 2 parameters for the light curve
+e, \omega \}$), 2 free parameters for the light curve
 normalization and uncertainty calibration ($\{\langle f \rangle, 
@@
 the simpler GP kernel.  We fitted the models using \texttt{PyMC3}
-\citep{salvatier_2016_PyMC3,exoplanet:theano}.  After initializing
+\citep{salvatier_2016_PyMC3,exoplanet:theano}.  
+We assumed a Gaussian likelihood, and
+after initializing
 each model with the parameters of the maximum {\it a posteriori}
-model, we assumed a Gaussian likelihood, and sampled using
+model, we sampled using
 \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
@@
 is present during the transit: the measured flux is systematically
-high during the first half, and low in the second half.  We verified
-that this effect is not caused by our stellar variability model by
-exploring an alternative approach in which we isolated each transit
+high during the first half, and low in the second half.  To determine
+whether this effect was not caused by our stellar variability model, we 
+explored an alternative approach in which we isolated each transit
 window and locally fitted out polynomial trends, and then binned all
@@
 
-Figures~\ref{fig:phasequarter} and~~\ref{fig:phaseyear} show the
+Figures~\ref{fig:phasequarter} and~\ref{fig:phaseyear} show the
 Kepler data binned over quarters and years.  In
@@
 \citet{holczer_transit_2016} implies that any such variations in \sn\
-need to be less than of order minutes.  Figures~\ref{fig:phasequarter}
+need to be less than of order minutes.  
+Our independent TTV analysis found similar results.
+Figures~\ref{fig:phasequarter}
 and~\ref{fig:phaseyear} also provide little evidence in support of
@@
 which could enable the necessary phasing if the spot-groups are
-long-lived.  Unforunately, the typical S/N per Kepler transit is
+large and long-lived.  Unfortunately, the  S/N per Kepler transit is
 $\approx8$, which renders individual spot-crossing events unresolved.
-Therefore, although among the possible ideas starspot crossings seem
-the most likely to provide the additional photometric scatter observed
+Therefore, although starspot crossings seem to be
+the most likely source of additional photometric scatter
 during transits, our reasoning is that the expected
 spot-crossing anomaly amplitudes ($\approx 100\,{\rm ppm}$) could be
-correct; we have not been able to definitely show that this is the correct
-explanation, and the issue could deserve further exploration.
+correct; we have not been able to definitely show that this is the case, and 
+the issue could deserve further exploration.
 
@@
 		\leavevmode
-		\includegraphics[width=1\textwidth]{f7.pdf}
+		\includegraphics[width=0.9\textwidth]{f7.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=1.0\textwidth]{f8.pdf}
+		\includegraphics[width=0.9\textwidth]{f8.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=1.0\textwidth]{f13.pdf}
+		\includegraphics[width=0.9\textwidth]{f13.pdf}
 	\end{center}
@@
 \begin{enumerate}
-  \item Model the starspot-induced variability using a
+  \item Fit the starspot-induced variability using a
     Gaussian Process with a \texttt{SHOTerm} kernel (see
     Section~\ref{app:gptransit}), a white-noise jitter term, and the
-    mean flux.  Fit the model to the time, flux, and flux uncertainty.
+    mean flux.
   \item Select points more than twice the median absolute
@@
 \end{enumerate}
-The flares were identified and fitted using \texttt{altaipony}
+The final step of flare identification and fitting was performed using \texttt{altaipony}
 \citep{davenport_2016,ilin_flares_2021}.  The analytic flare model is
@@
 flux during the short cadence observations.  These 24 flares spanned a
-total of 6.5 hours ($\sim$15 minutes per flare).  Whether it is
-meaningful or not, we noticed a coincidence in the flare arrival
+total of 6.5 hours ($\sim$15 minutes per flare).  Inspecting the data,
+we noticed a coincidence in the flare arrival
 times.  The coincidence is that despite the low flare duty cycle, one
@@
 As with any coincidence, if one does not have a firm prediction, it is
-difficult to assess the statistical significance of a surprise.  Since
+difficult to assess whether a surprise is statistically significant.  Since
 our surprise was specifically at the inter-arrival time of certain
@@
 case, 276 pairs).  We then compared the distribution of the pair
-separations against that from a Poisson distribution.  Specifically,
+separations against that of a Poisson distribution.  Specifically,
 we drew $N_{\rm f}=24$ samples from a Poisson distribution with
 $\lambda = \Delta t / N_{\rm f}$, for $\Delta t=97.7\,{\rm days}$ the
-duration of the observations, and repeated the draw $10^3$ times with
+total duration of the observations, and repeated the draw $10^3$ times with
 unique random seeds.
@@
 express the range of values observed from the Poissonian draws.  While
-it does seem like a rather odd coincidence for peaks in the observed
+it does seem like an odd coincidence for peaks in the observed
 flare arrival time distribution to coincide with the locations of
@@
 non-Poissonian process driving the flares does not seem especially
-strong.  More quantitatively, the peaks observed at the orbital and
+overwhelming.  More quantitatively, the peaks observed at the orbital and
 synodic periods are within the $\pm 2$-$\sigma$ range of a Poissonian
@@
 orb}+2P_{\rm syn}$ are only slightly above this range.  With that
-said, future analyses of these data by investigators more versed in
-this topic than ourselves could very well yield more quantitative insights.
+said, future analyses of these data by investigators with more
+knowledge of this topic could very well yield more quantitative
+insights.
 

### 2021-09-14 paper/ms.tex
update ms
Co-authored-by: Motohide Tamura <motohide.tamura@nao.ac.jp>

@@
 cluster.  Focusing on the planet itself, we confirm that despite the
-presence of a previously unreported binary companion, the planet
-orbits the G-dwarf primary on an orbit that is likely aligned with the
-stellar rotation axis (Section~\ref{sec:planet}).  We conclude by
-highlighting broader implications for our ability to age-date a
-statistical sample of planets (Section~\ref{sec:conc}).
+presence of a previously unreported binary companion, hereafter
+Kepler\,1627B, the planet orbits the G-dwarf primary on an orbit that
+is likely aligned with the stellar rotation axis
+(Section~\ref{sec:planet}).  We conclude by highlighting broader
+implications for our ability to age-date a statistical sample of
+planets (Section~\ref{sec:conc}).
 
@@
     115$\,Myr; \citealt{dahm_2015}), and the Gaia EDR3 Catalog of
-    Nearby Stars (gray background).  The right panel is a
-    zoomed view of the pre-main-sequence.  The \cn\ and IC\,2602 are
+    Nearby Stars (gray background).  The zoomed right panel
+    highlights the pre-main-sequence.  The \cn\ and IC\,2602 are
     approximately the same isochronal age.  {\it Bottom:} TESS and
-    Kepler stellar rotation periods versus dereddened Gaia colors,
+    Kepler stellar rotation period versus dereddened Gaia color,
     with the Pleiades and Praesepe (650\,Myr) shown for reference
@@
 
-In summary, the age of the \cn\ cluster based on the empirical
+In summary, the age of the \cn\ based on the empirical
 pre-main-sequence interpolation method is \clusterage.  This method
@@
 
-If the \pn\ transit signal is created by a bonafide planet,  then to
+If the \pn\ transit signal is created by a genuine planet,  then to
 our knowledge its age would make it the youngest planet yet found by
-the main Kepler mission.\footnote{The re-purposed K2
-mission however has found two younger systems: K2-33b and V1298 Tau
-\citep{David_et_al_2017,Mann_K2_33b_2016,david_four_2019}}  A portion
-of the light curve is shown with the phase-folded transit in
-Figure~\ref{fig:lc}.  The \texttt{PDCSAP}
-photometry is dominated by a quasi-periodic starspot signal with a peak-to-peak
-amplitude that varies between 2\% and 8\%, depending on the degree of
-asymmetry in spot coverage between the stellar hemispheres.  The
-starspot signal repeats every $2.642\pm0.042$\,days, where the
-uncertainty is based on the scatter observed between different Kepler
-quarters.  The details of how we model the starspot signal to derive
-the parameters of the planet are discussed in
-Appendix~\ref{app:gptransit}.  The stellar parameters are listed in
-Table~\ref{tab:starparams}; the stellar mass, radius, and effective
-temperature are found by interpolating against a 38$\,$Myr MIST
-isochrone.  The systematic uncertainties are taken to be the
-difference in parameters from the PARSEC isochrones, and the reported
-uncertainties are a quadrature sum of the statistical and systematic
-components.  The transit-fitting results, given in
+the main Kepler mission.\footnote{The re-purposed K2 mission however
+has found two younger systems containing five planets: K2-33b ($9\pm1\,{\rm Myr}$;
+\citealt{Mann_K2_33b_2016,David_et_al_2017}) and V1298 Tau
+($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  A portion of the
+light curve is shown with the phase-folded transit in
+Figure~\ref{fig:lc}.  The \texttt{PDCSAP} photometry is dominated by a
+quasi-periodic starspot signal with a peak-to-peak amplitude that
+varies between 2\% and 8\%, depending on the degree of asymmetry in
+spot coverage between the stellar hemispheres.  The starspot signal
+repeats every $2.642\pm0.042$\,days, where the uncertainty is based on
+the scatter observed between different Kepler quarters.  The details
+of how we model the starspot signal to derive the parameters of the
+planet are discussed in Appendix~\ref{app:gptransit}.  The stellar
+parameters are listed in Table~\ref{tab:starparams}; the stellar mass,
+radius, and effective temperature are found by interpolating against a
+38$\,$Myr MIST isochrone.  The systematic uncertainties are taken to
+be the difference in parameters from the PARSEC isochrones, and the
+reported uncertainties are a quadrature sum of the statistical and
+systematic components.  The transit-fitting results, given in
 Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
@@
 As we describe in Appendix~\ref{app:companionstar} however,
-the system has a low-mass stellar companion ($M_{\rm B}\approx
-0.44\,M_\odot$), and so we reconsidered this false positive scenario
-in some detail.
+Kepler\,1627A is accompanied by a low-mass stellar companion ($M_{\rm
+B}\approx 0.44\,M_\odot$).  We therefore reconsidered this false
+positive scenario in some detail.
 
@@
 observed using Keck/HIRES during the night of 7 Aug 2021.  We
-used the iodine cell for the wavelength calibration, and scheduled the
+used the iodine cell for wavelength calibration, and scheduled the
 transit using the ephemeris of \citet{holczer_transit_2016}.  We also acquired
@@
 from Table~\ref{tab:starparams}, and the jitter term has been
-incorpoarted.
+incorporated.
 We found $\chi^2_{\rm red} =
@@
 
-The authors are grateful to T{.}~David for help with the transit
-fitting, to R{.}~Kerr for kindly providing us with the
-\citet{Kerr2021} membership list prior to its publication, and to
-J{.}~Winn for helpful discussions on the RM observations and flare
-timing.
+The authors are grateful to T{.}~David, J{.}~Winn, and A{.}~Howard for
+illuminating discussions and guidance, and to R{.}~Kerr for kindly
+providing us with the \citet{Kerr2021} membership list prior to its
+publication.
 The authors are also grateful to K{.}~Collins for
@@
 \citep{beaumont_2014_13866} in the four projections shown in
-Figure~\ref{fig:XYZvtang}.  While bonafide members likely exist
+Figure~\ref{fig:XYZvtang}.  While cluster members likely exist
 outside of our selection region (and our selection also includes some
@@
     0\farcs16$, which corresponds to a projected separation of
-    $53\pm4$\,AU.  {\it Bottom:} Gaia EDR3 renormalized unit weight
+    $53\pm4$\,au.  {\it Bottom:} Gaia EDR3 renormalized unit weight
     error (RUWE) point estimates for Kepler\,1627A and other members
@@
 2.5$.  Using the measured Gaia EDR3 parallax for the system, this
-implies a projected separation of $53 \pm 4$.  Given the large
+implies a projected separation of $53 \pm 4$\,au.  Given the large
 observed RUWE of the primary, and the low geometric probability that a

### 2021-09-14 drivers/write_v06_main_tableheader.py

@@ -27,7 +27,7 @@ namedesc_dict = {
 'cluster': "Comma-separated cluster or group name.",
 'age': r"Comma-separated logarithm (base-10) of reported$^{\rm a}$ age in years.",
 'mean_age': r"Mean (ignoring NaNs) of $\texttt{age}$ column.",
-'reference_id': "Comma-separted provenance of group membership.",
+'reference_id': "Comma-separated provenance of group membership.",
 'reference_bibcode': r"ADS bibcode corresponding to $\texttt{reference\_id}$.",
 }
 

### 2021-09-14 paper/collect.sh

@@ -29,7 +29,7 @@ cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20210
 
 
 cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_overplotkep1627.pdf f15a.pdf
-cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_smalllims.pdf f15b.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_smalllims_overplotkep1627.pdf f15b.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_curtiscleaning_overplotkep1627.pdf f15c.pdf
 

### 2021-09-14 rudolf/plotting.py

@@ -2008,8 +2008,8 @@ def plot_hr(
         ax.set_ylim((16, -3))
 
     if smalllims and 'phot_bp_mean_mag' in color0:
-        ax.set_xlim([0.95,3.45])
-        ax.set_ylim([12.5,5.5])
+        ax.set_xlim([0.85,3.45])
+        ax.set_ylim([12.5,5.0])
     elif smalllims and 'phot_bp_mean_mag' not in color0:
         raise NotImplementedError
 

### 2021-09-14 paper/ms.tex
update ms
Co-authored-by: gummiks <gstefansson@astro.princeton.edu>

@@
   Kepler\,1627 is a member of the \clusterage\ old $\delta$~Lyr
-  cluster, based on Gaia kinematics, TESS stellar rotation
-  periods, and spectroscopic lithium abundances.  To
-  our knowledge, Kepler\,1627Ab is the youngest planet with a precise
-  age yet found by the main Kepler mission.  From unresolved
-  starspot-crossing events in the Kepler photometry and
-  a Keck/HIRES transit we find that the planet is likely on an
-  orbit aligned with the stellar spin axis.  We also briefly explore the flare
-  arrival time distribution in the short cadence Kepler
-  photometry.
-  The $\delta$~Lyr cluster is one of many stellar groups whose
-  properties have been clarified using Gaia; many other new
-  clusters also overlap with confirmed and candidate exoplanets.
-  Verifying the results of Gaia clustering analyses using TESS rotation
-  periods seems to be a promising path toward expanding the census of
-  age-dated planets, and our corresponding understanding of planetary evolution.
+  cluster, based on Gaia kinematics, TESS stellar rotation periods,
+  and spectroscopic lithium abundances.  To our knowledge,
+  Kepler\,1627Ab is the youngest planet with a precise age yet found
+  by the main Kepler mission.  From unresolved starspot-crossing
+  events in the Kepler photometry and a Keck/HIRES transit we find
+  that the planet is likely on an orbit aligned with the stellar spin
+  axis.  We also briefly explore the flare arrival time distribution
+  in the short cadence Kepler photometry.  The $\delta$~Lyr cluster is
+  one of many stellar groups whose properties have been clarified
+  using Gaia; many other new clusters also overlap with confirmed and
+  candidate exoplanets.  Verifying the results of Gaia clustering
+  analyses using TESS rotation periods offers a promising path toward
+  expanding the census of age-dated planets, and our corresponding
+  understanding of planetary evolution.
 \end{abstract}
@@
 envelopes of hydrogen and helium, are expected to shrink in size by
-factors of several over their first $10^8$ years
-\citep{owen_atmospheres_2016,owen_constraining_2020}.  Specifically,
-they start with sizes of 4--12\,$R_\oplus$ shortly after the time of disk
+factors of several over their first $10^8$ years.  Specifically, in
+the models of \citet{owen_atmospheres_2016} and
+\citet{owen_constraining_2020}, the $\approx5\,M_\oplus$ planets start
+with sizes of 4--12\,$R_\oplus$ shortly after the time of disk
 dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
@@
 dispersal \citep{owen_atmospheres_2016}, the combination of stellar
-irradiation and internal heat may also power a more
-gradual outflow which can eventually deplete or entirely strip the
-envelope from the rocky core
-\citep{Owen_Wu_2013,ginzburg_corepowered_2018}.  Discovering young
-planets, measuring their masses, and detecting their atmospheric
-outflows are key steps toward testing this paradigm, which is often
-invoked to explain the observed radius distribution of mature
+irradiation and internal heat may also power a more gradual outflow
+which can eventually deplete or entirely strip the envelope from the
+rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
+Discovering young planets, measuring their masses, and detecting their
+atmospheric outflows are key steps toward testing this paradigm, which
+is often invoked to explain the observed radius distribution of mature
 exoplanets \citep{Fulton_et_al_2017}.
@@
 \cn\ \citep{kounkel_untangling_2019}.  Given the previous statistical
-validation the close-in mini-Neptune Kepler 1627b
+validation of the close-in mini-Neptune Kepler 1627b
 \citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
@@
     {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
-    points are kinematically selected members of the $\delta$\,Lyr
-    cluster.  Both Kepler (gray squares) and TESS observed portions of
-    the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray;
-    four of them (orange circles) are annotated.
+    points are kinematically selected candidate $\delta$\,Lyr
+    cluster members.  Both Kepler (gray panels) and TESS observed portions of
+    the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray or orange.
     Kepler\,1627 (green star) was observed during the entirety of the
-    Kepler mission, and has been observed by TESS for two lunar months
-    to date.  The colors of the other cluster members indicate the
-    number of TESS sectors during which they have been observed.
+    Kepler mission.  Colors of candidate cluster members indicate the
+    duration of TESS observations to date.  So far, Kepler\,1627 has been
+    observed by TESS for two lunar months.
     \label{fig:skychart}
@@
 We fitted the RVs using the \citet{hirano_analytic_2010,hirano_2011}
-models for the Rossiter-McLaughlin (RM) effect, and allowed the
-sky-projected obliquity, the stellar equatorial velocity, and the
-Gaussian dispersion of the spectral lines to vary.  We also allowed
+models for the Rossiter-McLaughlin (RM) effect, 
+following the methodology developed by \citet{stefansson_2020}.
+We allowed the sky-projected obliquity, the projected stellar equatorial
+velocity, and the Gaussian dispersion of the spectral lines to vary, and fixed the limb-darkening
+using the $V$-band tabulation from \citet{claret_gravity_2011}.  We also
+allowed
 for an optional linear and quadratic trend, and a white-noise jitter
@@
 quadratic trend'' model over the best-fitting ``quadratic trend only''
-model.  Given the expected semi-amplitude of the velocity anomaly
-($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) and the achieved precision,
+model.  Comparing the expected semi-amplitude of the velocity anomaly
+for a well-aligned orbit ($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) with the achieved precision,
 we cannot claim a detection of the Doppler transit.
@@
 	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
-  \texttt{IPython} \citep{perez_2007},
+  %\texttt{IPython} \citep{perez_2007},
 	%\texttt{isochrones} \citep{morton_2015_isochrones},
 	%\texttt{lightkurve} \citep{lightkurve_2018},
-  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
+  %\texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
   %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
-  \texttt{numpy} \citep{walt_numpy_2011}, 
-  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
+  %\texttt{numpy} \citep{walt_numpy_2011}, 
+  %\texttt{pandas} \citep{mckinney-proc-scipy-2010},
   %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
@@
 
-The \texttt{v0.6} CDIPS target catalog (Table~\ref{tab:v06}) includes
-stars that are young, age-dated, and age-dateable.  By ``age-dateable'',
-we mean that the age should be measurable at greater
-precision than a typical FGK field star.
+The \texttt{v0.6} CDIPS target catalog (Table~\ref{tab:v06}) includes stars
+that are young, age-dated, and age-dateable.  By ``age-dateable'', we mean that
+the age should be measurable at greater precision than a typical FGK field star
+using isochronal, gyrochronal, or spectroscopic techniques.
 As in
@@
 \citep{beaumont_2014_13866} in the four projections shown in
-Figure~\ref{fig:XYZvtang}.  While cluster members likely exist
+Figure~\ref{fig:XYZvtang}.  While other cluster members likely exist
 outside of our selection region (and our selection also includes some
@@
     of the $\delta$\,Lyr cluster reported by \citet{KounkelCovey2019}.
-    Since other members of the cluster with similar colors have
+    Since other members of the cluster with similar brightnesses have
     comparable degrees of photometric variability, the high RUWE
@@
 There were $N_{\rm f}=24$ flares that exceeded $0.5\%$ in relative
-flux during the short cadence observations.  These 24 flares spanned a
+flux during the short cadence observations.  
+These 24 flares spanned a
 total of 6.5 hours ($\sim$15 minutes per flare).  Inspecting the data,
@@
 $P_{\rm syn} = (P_{\rm rot}^{-1} - P_{\rm orb}^{-1})^{-1}$, and linear
-combinations thereof.  Note that the tidal period (half the synodic
+combinations thereof.  The tidal period (half the synodic
 period) is not shown.  The bins are logarithmically spaced to give 100
@@
 insights.
+Such analyses should keep in mind an important caveat:
+the amplitude
+distribution of M-dwarf flares extends up to many times the quiescent
+flux \citep[see Figure~7 of][]{gunther_2020}.  A flare on
+Kepler\,1627B producing double its quiescent white-light flux would yield
+a $\approx$1\% apparent amplitude.
+Such flares could represent a significant fraction
+of those in the Kepler observations.
 

### 2021-09-14 rudolf/plotting.py

@@ -28,6 +28,7 @@ RM:
 
 General for re-use:
     multiline: iterate through a colormap with many lines.
+    truncate_colormap: extract a subset of a colormap
 """
 import os, corner, pickle, inspect
 from glob import glob
@@ -420,8 +421,9 @@ def plot_skychart(outdir, narrowlims=0, showkepler=0, showtess=0,
         mdf = pd.read_csv(outcsv)
 
         cmap = mpl.cm.viridis
-        bounds = np.arange(-0.5, 4.5, 1)
-        ticks = (np.arange(-1,4)+1)
+        cmap = truncate_colormap(cmap, 0., 0.75)
+        bounds = np.arange(-0.5, 3.5, 1)
+        ticks = (np.arange(-1,3)+1)
         norm = mpl.colors.BoundaryNorm(bounds, cmap.N)
 
         cax = ax.scatter(
@@ -435,9 +437,15 @@ def plot_skychart(outdir, narrowlims=0, showkepler=0, showtess=0,
 
         _p = ax.scatter(
             trgt_mdf.ra, trgt_mdf.dec, c=trgt_mdf.n_tess_sector, cmap=cmap,
-            alpha=1, zorder=42, s=60, linewidths=0.2,
+            alpha=1, zorder=42, s=80, linewidths=0.2,
             marker='*', norm=norm, edgecolors='k'
         )
+
+        bbox = dict(facecolor='white', alpha=0.9, pad=0, edgecolor='white')
+        deltay=1.2
+        ax.text(trgt_mdf.ra, trgt_mdf.dec+deltay, 'Kepler 1627', ha='center',
+                va='bottom', fontsize='xx-small', bbox=bbox, zorder=49)
+
 
         # standard colorbar...
         # cb = f.colorbar(cax, extend='max', ticks=ticks)
@@ -452,7 +460,7 @@ def plot_skychart(outdir, narrowlims=0, showkepler=0, showtess=0,
                             loc='lower left',
                             bbox_transform=ax.transAxes)
         cb = f.colorbar(_p, cax=axins1, orientation="vertical",
-                        extend='max', ticks=ticks)
+                        ticks=ticks)
         cb.ax.minorticks_off()
         cb.ax.tick_params(labelsize='x-small')
         cb.ax.set_title("$N_{\mathrm{TESS}}$", fontsize='x-small')
[…]

### 2021-09-15 paper/ms.tex
update ms (binary companion appendix)
Co-authored-by: adamkraus <alk@astro.as.utexas.edu>

@@
 e.g.},][]{soderblom_ages_2014}.  We acquired an iodine-free spectrum
-from Keck/HIRES on the night of 26 March 2021 using the standard setup
+from Keck/HIRES on the night of 2021 March 26 using the standard setup
 and reduction techniques of the California Planet Survey
@@
     {\it Bottom:} Radial velocities acquired using Keck/HIRES during
-    the transit of 7 Aug 2021 UT, observed simultaneously with MuSCAT3
+    the transit of 2021 Aug 7, observed simultaneously with MuSCAT3
     ({\it griz} photometry).  A quadratic baseline trend was allowed
@@
 \end{figure}
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.8\textwidth]{f14.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Radii, orbital periods, and ages of transiting
+			exoplanets}.  Planets older than $10^9$ years are shown with gray
+		markers.  Planets younger than a gigayear with ${\rm
+			\tau}/\sigma_{\rm \tau} > 3$ are emphasized, for $\tau$ the age.
+		\pn\ is shown with a
+		star.  The large sizes of the youngest transiting planets could be
+		explained by their primordial atmospheres not yet having
+		evaporated; direct measurements of the atmospheric outflows or
+		planetary masses would help to confirm this expectation.
+		Selection effects may also be important.  Parameters are from the
+		NASA Exoplanet Archive (2021 July 2).
+		\label{fig:rp_period_age}
+	}
+\end{figure*}
 
@@
 The final piece of evidence comes from a spectroscopic transit we
-observed using Keck/HIRES during the night of 7 Aug 2021.  We
+observed using Keck/HIRES during the night of 2021 Aug 7.  We
 used the iodine cell for wavelength calibration, and scheduled the
@@
 
-\begin{figure}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.49\textwidth]{f14.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Radii, orbital periods, and ages of known transiting
-    exoplanets}.  Planets older than $10^9$ years are shown with gray
-    markers.  Planets younger than a gigayear with ${\rm
-    Age}/\sigma_{\rm Age} > 3$ are emphasized.  \pn\ is shown with a
-    star.  The large radii of the youngest transiting planets could be
-    explained by their primordial atmospheres not yet having
-    evaporated; direct measurements of the atmospheric outflows or
-    planetary masses would help to confirm this expectation.
-    Selection effects may also be important.  Parameters are from the
-    NASA Exoplanet Archive (July 2, 2021).
-		\label{fig:rp_period_age}
-	}
-\end{figure}
-
 %, which is expected to be outflowing at a rate of
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f17a.pdf}
-			\includegraphics[width=0.37\textwidth]{f17b.jpg}
-		}
-
-		\vspace{-0.1cm}
-		\subfloat{
-			\includegraphics[width=0.55\textwidth]{f6.pdf}
+			\includegraphics[width=0.49\textwidth]{f17.pdf}
+			\includegraphics[width=0.48\textwidth]{f6.pdf}
 		}
@@
 	\caption{
-    {\bf Kepler 1627 is a binary.} {\it Top:} High-resolution imaging
-    from Gemini-North/`Alopeke ({\it left}) and Keck/NIRC2 ({\it
-    right}) shows an $\approx$M2V companion at $\rho \approx
-    0\farcs16$, which corresponds to a projected separation of
-    $53\pm4$\,au.  {\it Bottom:} Gaia EDR3 renormalized unit weight
-    error (RUWE) point estimates for Kepler\,1627A and other members
-    of the $\delta$\,Lyr cluster reported by \citet{KounkelCovey2019}.
-    Since other members of the cluster with similar brightnesses have
-    comparable degrees of photometric variability, the high RUWE
-    suggests that Kepler\,1627 is an astrometric binary. 
+    {\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
+    from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2V
+    companion at $\rho \approx 0\farcs16$, which corresponds to a
+    projected separation of $53\pm4$\,AU.  The inset shows a cutout of
+    the NIRC2 image (North is up, East is left, scale is set by the
+    separation of the binary).  The lines show 5-$\sigma$ contrast
+    limits for the `Alopeke filters, and 6-$\sigma$ contrast limits
+    for NIRC2.  {\it Left:} Gaia EDR3 renormalized
+    unit weight error (RUWE) point estimates for Kepler\,1627A and
+    other members of the $\delta$\,Lyr cluster reported by
+    \citet{KounkelCovey2019}.  Since other members of the cluster with
+    similar brightnesses have comparable degrees of photometric
+    variability, the high RUWE suggests that Kepler\,1627 is an
+    astrometric binary. 
     \label{fig:kep1627binary}
@@
 
-We first noted the presence of a stellar neighbor in archival
-Keck/NIRC2 Kp-band adaptive optics imaging (2.12\,$\mu $m), from the
-Spring semesters of both 2015 and 2019.  We acquired additional
-Gemini-North/`Alopeke speckle imaging on 24 June 2021, which showed
-the same companion in the 0.83\,$\mu$m notch filter, but not the
-$0.56\,\mu$m filter.  Figure~\ref{fig:kep1627binary} summarizes the
-results of the high-resolution imaging.  Although the companion is
-unresolved in the Gaia source catalog, its existence was also
-suggested by the primary star's large RUWE, relative to other members
-of the \cn.  The images yield a projected separation $\rho \approx
-0\farcs16$, with $\Delta m_{832} \approx 3.1$ and $\Delta m_{\rm Kp} =
-2.5$.  Using the measured Gaia EDR3 parallax for the system, this
-implies a projected separation of $53 \pm 4$\,au.  Given the large
-observed RUWE of the primary, and the low geometric probability that a
-companion imaged within $\rho \approx 0\farcs16$ is a chance
+We first noted the presence of a close neighbor on 2015 July 22 when
+we acquired adaptive optics imaging of \sn\ using the NIRC2 imager on
+Keck-II. We used the narrow camera (FOV = 10.2\arcsec) to obtain 8
+images in the $K'$ filter ($\lambda = 2.12\,\mu$m) with a total
+exposure time of 160\,s. We analyzed these data following
+\citet{kraus_impact_2016}, beginning by applying standard reduction
+steps (dark-subtraction, flat-fielding, linearization, and bad pixel
+correction). We then used PSF-fitting to measure the separation,
+position angle, and contrast of the candidate companion, identifying
+the best-fitting empirical PSF template from among the
+near-contemporaneous observations of single stars in the same filter.
+Finally, we used the optical distortion solution of \citet{yelda_2010}
+to convert pixel-space relative positions to on-sky relative
+astrometry. The mean values inferred from the 8 images are reported in
+Table~\ref{tab:starparams}.
+
+We also estimated detection limits following
+\citet{kraus_impact_2016}, analyzing the residuals after subtracting
+the empirical PSF templates, as well as the residuals after
+subtracting an azimuthally smoothed PSF profile. Within each residual
+image, the flux was measured through 40\,mas apertures centered on
+every pixel, and then the noise as a function of radius was estimated
+from the RMS within concentric rings. Finally, the detection limits
+were estimated from the strehl-weighted sum of the detection
+significances in the image stack, and we adopted the $+6\sigma$
+threshold as the detection limit for ruling out additional companions.
+We find contrast limits of $\Delta m_{K'} = 5.1$ mag at $\rho = 150$ mas,
+$\Delta m_{K'} = 7.0$ mag at $\rho = 300$ mas, and $\Delta m_{K'} = 9.0$ mag
+at $\rho > 1000$ mas.  Given the distance ($d = 330$\,pc) and age
+($\tau \sim 40$\,Myr), the models of \citet{chabrier_2000} imply
+corresponding physical limits on tertiary companions of $M_{\rm ter} < 50
+M_{\rm Jup}$ at $\rho = 50$\,AU, $M_{\rm ter} < 20 M_{\rm Jup}$ at $\rho =
+100$\,AU, and $M_{\rm ter} < 10 M_{\rm Jup}$ at $\rho = 330$\,AU.
+
+We also observed \sn\ on Gemini-North using the `Alopeke speckle
+imager on 2021 June 24.  `Alopeke is a dual-channel speckle
+interferometer that uses narrow-band filters centered at 0.83\,$\mu$m
+and $0.56\,\mu$m.  We acquired three sets of $1000\times 60$$\,$msec
+exposures during good seeing (0.45$''$), and used the autocorrelation
+function of these images to reconstruct a single image and 5-$\sigma$
+detection limits (see \citealt{howell_speckle_2011}).  This procedure
+yielded a detection of the companion in the 0.83\,$\mu$m notch filter,
+but not the $0.56\,\mu$m filter.
+The measured projected separation and magnitude difference
+are given in Table~\ref{tab:starparams}.
+
+Figure~\ref{fig:kep1627binary} summarizes the results of the
+high-resolution imaging.  
+The Gaia EDR3 parallax for the primary implies a projected separation
+of $53 \pm 4$\,AU, assuming the companion is bound. 
+Although the companion is unresolved in the
+Gaia source catalog (there are no candidates brighter than $G < 20.5$ mag within $\rho < 120\arcsec$), its existence was also suggested by the primary
+star's large RUWE, relative to other members of the \cn.  Given the
+large observed RUWE of the primary, and the low geometric probability
+that a companion imaged at $\rho \approx 0\farcs16$ is a chance
 line-of-sight companion, we proceed under the assumption that the
-companion is bound, and that Kepler\,1627 system is a binary.  Based
-on the
-tabulation\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
-version \texttt{2021/03/02}.} by \citet{pecaut_mamajek_2013}, the
-measured magnitude differences for a G8V primary correspond to a
-spectral type for the companion of $\approx$M2V ($M_\star \approx
-0.44\,M_\odot$).  However, the companion should have a longer
-pre-main-sequence contraction phase than the primary, which would
-imply that this mass is overestimated by $\sim10$ to $20\%$.
+companion is bound, and that Kepler\,1627 is a binary. 
+Given the distance and age, the models of \citet{baraffe_new_2015} imply
+a companion mass of $M_{\rm B} \approx 0.33 M_{\odot}$ and companion
+temperature of $T_{\rm eff,B} \approx 3450$ K.
+The corresponding spectral type is roughly M2.5V \citep{pecaut_mamajek_2013}.
+% \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
 
@@
 duration therefore rule out the scenario that it could explain the
-transit signal.
-Two other arguments presented in the main text -- the inferred stellar
-density, and the TTV-local slope correlation -- confirm this point.
+transit signal.  Two other arguments presented in the main text -- the
+inferred stellar density, and the TTV-local slope correlation --
+confirm this point.
 

### 2021-09-15 drivers/plot_contrast_imgs.py

@@ -1,0 +1,72 @@
+import os, pickle
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from numpy import array as nparr
+from aesthetic.plot import savefig, format_ax, set_style
+import matplotlib as mpl
+from astropy import units as u, constants as c
+import matplotlib.patheffects as pe
+import matplotlib.image as mpimg
+from rudolf.paths import DATADIR, RESULTSDIR
+
+def plot_zorro_speckle():
+
+    datapaths = [
+        os.path.join(DATADIR, 'imaging', 'Alopeke_562_deltaM_vs_sep.csv'),
+        os.path.join(DATADIR, 'imaging', 'Alopeke_832_deltaM_vs_sep.csv'),
+        os.path.join(DATADIR, 'imaging', 'NIRC2_deltaM_vs_sep.csv'),
+    ]
+    labels = [
+        '`Alopeke 0.56$\,\mu$m',
+        '`Alopeke 0.83$\,\mu$m',
+        'NIRC2 2.12$\,\mu$m',
+    ]
+    colors = [
+        'dodgerblue',
+        'tomato',
+        'darkred'
+    ]
+
+    fig, ax = plt.subplots(figsize=(4,3))
+    set_style()
+
+    for d,l,c in zip(datapaths, labels, colors):
+        df = pd.read_csv(d)
+        ax.plot(
+            df.sep, df.deltamag, color=c, label=l
+        )
+
+    leg = ax.legend(
+        loc='upper center', handletextpad=0.2, fontsize='x-small',
+        framealpha=0.9, bbox_to_anchor=(0.4,0.9)
+    )
+
+    img = mpimg.imread(
+        os.path.join(DATADIR, 'imaging',
+                     'NIRC2_img_20150722_39302_rot.png')
+    )
+
+    # [left, bottom, width, height]
+    inset = fig.add_axes([0.66, 0.63, .3, .3])
+    inset.imshow(img)
+    inset.axis('off')
+    plt.setp(inset, xticks=[], yticks=[])
+
+    ax.set_ylabel('$\Delta$mag')
+    ax.set_xlabel('Angular separation [arcsec]')
+
+    ylim = ax.get_ylim()
+    ax.set_ylim(ylim[::-1])
+
+    ax.set_xlim([0,1.25])
+
+    fig.tight_layout(h_pad=0, w_pad=0)
+
+    figpath = (
[…]

### 2021-09-15 paper/collect.sh

@@ -38,5 +38,7 @@ cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_
 cp /Users/luke/Dropbox/proj/rudolf/results/RM/quadratic/rm_RV_and_phot.pdf f16.pdf
 
 
+cp /Users/luke/Dropbox/proj/rudolf/results/imaging/imaging_summary.pdf f17.pdf
+
 cp /Users/luke/Dropbox/proj/rudolf/data/imaging/Kepler1627_20210624_562_832_final.pdf f17a.pdf
 cp /Users/luke/Dropbox/proj/rudolf/data/imaging/KOA_104524/NIRC2/calibrated/N2.20150722.39302_drp.jpg f17b.jpg

### 2021-09-15 paper/ms.tex
update ms (binary companion appendix)
Co-authored-by: adamkraus <alk@astro.as.utexas.edu>

@@
   \caption{ {\bf Binarity indicators for rotators in \cn.} 
-    Data are as in the lower panel of
-    Figure~\ref{fig:age}.  Astrometric binaries ($\mathrm{RUWE}>1.2$)
-    appear in red; photometric binaries ($>$0.3\,mag above an
-    empirical isochrone; 20\% of stars) are orange.  
-    Over $0.6<\bpmrpo<1.5$, the detected rotation periods in binaries tend to 
-    fall below the slow sequence.
-    %FIXME!
-		% Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has \replaced{281}{523} stars.
-		% Dividing these stars into ``slow'' and ``fast'' sequences by eye,
-		% the fraction of stars showing signs of binarity in the slow and fast
-		% subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and
-		% \replaced{56\% (40/71)}{51\% (68/134)}, respectively.
+    Data are as in the lower panel of Figure~\ref{fig:age}.  Sources
+    with $\mathrm{RUWE}>1.2$ appear in red; these include astrometric binaries as well
+    as sources poorly fit by a single-star PSF.  Photometric binaries
+    ($>$0.3\,mag above an empirical isochrone; 20\% of stars) are
+    orange.  Over $0.6<\bpmrpo<1.5$, the detected rotation periods in
+    binaries tend to fall below the slow sequence.
 		\label{fig:binarity}
@@
 	\end{center}
-	\vspace{-0.7cm}
+	\vspace{-0.5cm}
 	\caption{
@@
     similar brightnesses have comparable degrees of photometric
-    variability, the high RUWE suggests that Kepler\,1627 is an
-    astrometric binary. 
+    variability, the high RUWE suggests that Kepler\,1627 is a
+    binary. 
     \label{fig:kep1627binary}
@@
 yielded a detection of the companion in the 0.83\,$\mu$m notch filter,
-but not the $0.56\,\mu$m filter.
-The measured projected separation and magnitude difference
-are given in Table~\ref{tab:starparams}.
+but not the $0.56\,\mu$m filter.  The measured projected separation
+and magnitude difference are given in Table~\ref{tab:starparams}.
 
 Figure~\ref{fig:kep1627binary} summarizes the results of the
-high-resolution imaging.  
-The Gaia EDR3 parallax for the primary implies a projected separation
-of $53 \pm 4$\,AU, assuming the companion is bound. 
-Although the companion is unresolved in the
-Gaia source catalog (there are no candidates brighter than $G < 20.5$ mag within $\rho < 120\arcsec$), its existence was also suggested by the primary
-star's large RUWE, relative to other members of the \cn.  Given the
-large observed RUWE of the primary, and the low geometric probability
-that a companion imaged at $\rho \approx 0\farcs16$ is a chance
-line-of-sight companion, we proceed under the assumption that the
-companion is bound, and that Kepler\,1627 is a binary. 
-Given the distance and age, the models of \citet{baraffe_new_2015} imply
-a companion mass of $M_{\rm B} \approx 0.33 M_{\odot}$ and companion
-temperature of $T_{\rm eff,B} \approx 3450$ K.
-The corresponding spectral type is roughly M2.5V \citep{pecaut_mamajek_2013}.
+high-resolution imaging.  The Gaia EDR3 parallax for the primary
+implies a projected separation of $53 \pm 4$\,AU, assuming the
+companion is bound.  Although the companion is unresolved in the Gaia
+source catalog (there are no candidates brighter than $G < 20.5$ mag
+within $\rho < 120\arcsec$), its existence was also suggested by the
+primary star's large RUWE, relative to other members of the \cn.
+Based on the apparent separation, the binary orbital period is of
+order hundreds of years.  The large RUWE is therefore more likely to
+be caused by a PSF-mismatch skewing the Gaia centroiding during
+successive scans, rather than true astrometric motion.  Regardless,
+given the large observed RUWE of the primary, and the low geometric
+probability that a companion imaged at $\rho \approx 0\farcs16$ is a
+chance line-of-sight companion, we proceed under the assumption that
+the companion is bound, and that Kepler\,1627 is a binary.  Given the
+distance and age, the models of \citet{baraffe_new_2015} imply a
+companion mass of $M_{\rm B} \approx 0.33 M_{\odot}$ and companion
+temperature of $T_{\rm eff,B} \approx 3450$ K.  The corresponding
+spectral type is roughly M2.5V \citep{pecaut_mamajek_2013}.
 % \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}

### 2021-09-15 rudolf/plotting.py

@@ -2226,7 +2226,7 @@ def plot_rotationperiod_vs_color(outdir, runid, yscale='linear', cleaning=None,
                 nparr(xval)[nparr(df.is_astrometric_binary)],
                 df[nparr(df.is_astrometric_binary)][ykey],
                 c='red', alpha=1, zorder=9, s=s, edgecolors='k',
-                marker='o', linewidths=_lw, label="Astrometric binary"
+                marker='o', linewidths=_lw, label="RUWE>1.2"
             )
 
     if overplotkep1627:

### 2021-09-15 paper/ms.tex
update ms

@@
     Kepler mission.  Colors of candidate cluster members indicate the
-    duration of TESS observations to date.  So far, Kepler\,1627 has been
-    observed by TESS for two lunar months.
+    duration of TESS observations to date, in units of lunar months.
     \label{fig:skychart}
@@
 		\subfloat{
-			\includegraphics[width=0.8\textwidth]{f15c.pdf}
+			\includegraphics[width=0.7\textwidth]{f15c.pdf}
 		}
@@
     transits of \pn, as well as other deviations from the model.  The
-    Figure Set online shows the entire 3.9 years of observations.
+    online Figure Set shows the entire 3.9 years of observations.
     {\it Bottom}:
@@
     in the RV fit.  The best-fit sky-projected obliquity in this model
-    is listed in red; the detection however has marginal statistical
+    is listed in red; the spectroscopic detection however has marginal statistical
     significance compared to a ``trend-only'' model.
@@
 Kepler\,1627A is accompanied by a low-mass stellar companion ($M_{\rm
-B}\approx 0.44\,M_\odot$).  We therefore reconsidered this false
+B}\approx 0.33\,M_\odot$).  We therefore reconsidered this false
 positive scenario in some detail.
+
+% Baraffe+15: M_B 0.33 Mdot
+% Rstar = 0.455 --> 0.465 Rsun
+%
 
@@
 theoretically expected density for the companion is
-$\approx 3.2\,{\rm g\,cm}^{-3}$ (Table~\ref{tab:posterior};
+$\approx 4.6\,{\rm g\,cm}^{-3}$ (Table~\ref{tab:posterior};
 \citealt{choi_mesa_2016}).  The transit duration is therefore too long
 to be explained by a star eclipsing the M dwarf secondary at
-$\approx$$5$-$\sigma$.
+$10$-$\sigma$.
 
@@
     {\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
-    from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2V
+    from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2.5V
     companion at $\rho \approx 0\farcs16$, which corresponds to a
@@
 therefore need to be non-grazing with $R_3/R_2 \approx 0.42$.
-Assuming a $\approx 0.42R_\odot$ radius of the imaged secondary, this
+Assuming a $\approx 0.46R_\odot$ radius of the imaged secondary
+from the \citet{baraffe_new_2015} models, this
 would imply a tertiary stellar radius of $\approx 0.2R_\odot$.  This

### 2021-09-15 drivers/calc_companion_density.py

@@ -1,0 +1,13 @@
+from astropy import units as u, constants as c
+import numpy as np
+
+M = 0.33*u.Msun
+R = 0.465*u.Rsun
+# from http://perso.ens-lyon.fr/isabelle.baraffe/BHAC15dir/BHAC15_tracks+structure
+# using 0.300 Msun, logt 7.571495, teff 3411 and rounding up
+
+V = (4/3)*np.pi*R**3
+
+rho = M/V
+
+print(rho.cgs)

### 2021-09-15 paper/ms.tex
update ms

@@
 		Selection effects may also be important.  Parameters are from the
-		NASA Exoplanet Archive (2021 July 2).
+		NASA Exoplanet Archive (2021 Sept 15).
 		\label{fig:rp_period_age}
@@
 MuSCAT3 instrument, developed by the Astrobiology Center and under
-financial supports by JSPS KAKENHI (JP18H05439) and JST PRESTO
+financial support by JSPS KAKENHI (JP18H05439) and JST PRESTO
 (JPMJPR1775), at Faulkes Telescope North on Maui, HI, operated by the

### 2021-09-15 drivers/plot_rp_vs_period_scatter.py

@@ -14,14 +14,14 @@ from cdips.utils import today_YYYYMMDD
 from cdips.utils.catalogs import get_nasa_exoplanet_archive_pscomppars
 from aesthetic.plot import savefig, format_ax, set_style
 
-VER = '20210702' # could be today_YYYYMMDD()
+VER = '20210915' # could be today_YYYYMMDD()
 
 def arr(x):
     return np.array(x)
 
 def plot_rp_vs_period_scatter(
     showlegend=1, colorbydisc=1, showarchetypes=1, showss=1, colorbyage=0,
-    verbose=0, add_kep1627=0, add_allkep=0
+    verbose=0, add_kep1627=0, add_allkep=0, add_plnames=0
 ):
 
     set_style()
@@ -114,6 +114,7 @@ def plot_rp_vs_period_scatter(
     age = sdf['st_age']*1e9
     period = sdf['pl_orbper']
     discoverymethod = sdf['discoverymethod']
+    pl_name = sdf['pl_name']
 
     #
     # plot age vs rp. (age is on y axis b/c it has the error bars, and I at
@@ -150,6 +151,14 @@ def plot_rp_vs_period_scatter(
             marker='o', cmap=cmap, linewidths=0.3, norm=norm
         )
 
+        if add_plnames:
+            bbox = dict(facecolor='white', alpha=0.7, pad=0, edgecolor='white',
+                       lw=0)
+            for _x,_y,_s in zip(period[s1],rp[s1],pl_name[s1]):
+                ax.text(_x, _y, _s, ha='right', va='bottom', fontsize=2,
+                        bbox=bbox, zorder=49)
+
+
         if add_kep1627:
             namelist = ['Kepler-1627']
             ages = [3.5e7]
@@ -172,6 +181,10 @@ def plot_rp_vs_period_scatter(
                     marker='*', cmap=cmap, linewidths=0.3, norm=norm
                 )
 
+                if add_plnames:
+                    ax.text(_per, _rp, 'Kepler-1627 b', ha='right',
+                            va='bottom', fontsize=2, bbox=bbox, zorder=49)
+
 
         if add_allkep:
             # Kepler-52 and Kepler-968
@@ -189,13 +202,23 @@ def plot_rp_vs_period_scatter(
                 _age = np.ones(len(_sdf))*a
 
                 if n == 'Kepler-1627':
-                    _rp = (0.298*u.Rjup).to(u.Rearth).value
+                    _rp = [(0.298*u.Rjup).to(u.Rearth).value]
 
[…]

### 2021-09-15 paper/collect.sh

@@ -25,7 +25,7 @@ cp /Users/luke/Dropbox/proj/rudolf/results/ttv_vs_local_slope/ttv_vs_local_slope
 
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit_Kepler_1627_Q15slc/Kepler_1627_Q15slc_RotGPtransit_posterior_phaseplot.pdf f12.pdf
 
-cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20210702_colorbyage_showkep1627.pdf f14.pdf
+cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20210915_colorbyage_showkep1627.pdf f14.pdf
 
 
 cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_overplotkep1627.pdf f15a.pdf

### 2021-09-15 paper/ms.tex
update ms
AU Mic c exists

@@
 of hydrogen or helium could yield insight into the evolutionary state
-of the atmosphere \citep[{\it
-e.g.},][]{spake_helium_2018,vissapragada_2020}.  Determining the
+of the atmosphere \citep{spake_helium_2018,vissapragada_2020}.  Determining the
 precise amount of high-energy irradiation incident on the planet would
@@
 quantity to measure would be the planet's mass.  This might be
-possible either through detailed timing analyses of the Kepler data,
+possible through timing analyses of the Kepler data,
 or through a high-cadence multi-color radial velocity campaign.
@@
 highlights an odd flare timing coincidence that may be suggestive of
-star-planet interactions.  We mention it here to encourage its
+star-planet interactions.  We encourage its
 exploration by investigators more versed in the topic than ourselves.
@@
 youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
-with precise ages below 100\,Myr include K2-33b, DS~Tuc~Ab,
-HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet V1298~Tau system
-\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020}.
-\pn\ is currently the smallest of these systems ($3.34\pm
-0.18\,R_\oplus$), which may hint at selection effects in the young
-planet sample \citep{zhou_2021_tois}.  Assuming that these planets
-have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$, their
-large sizes and presumed low densities would be in accord with the
-expectation that mini-Neptunes start their lives with large primordial
-atmospheres that are shed over the first gigayear
+with precise ages below 100\,Myr include K2-33, DS~Tuc~A, HIP~67522,
+TOI~837, the two-planet AU~Mic system, and the four-planet V1298~Tau
+system
+\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020,martioli_aumicbc_2021}.
+\pn\ is one of the smallest planets in this sample ($3.34\pm
+0.18\,R_\oplus$), which could be linked to the selection effects
+imposed by the spot-induced photometric variability at very young ages
+\citep[{\it e.g.},][]{zhou_2021_tois}.  Assuming that these young
+planets have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$,
+their large sizes and presumed low densities would be in accord with
+the expectation that mini-Neptunes start their lives with large
+primordial atmospheres that are shed over the first gigayear
 \citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
-suggestion of an aligned orbit differs from observed misalignments in
-number of close-in
+suggestion of an aligned orbit is consistent with a quiescent
+disk-driven migration history, and differs from misalignments that
+have been observed for some older
 mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
-Unlike these older planets, \pn\ seems to have properties consistent
-with a quiescent disk-driven migration history.  The alternative
-scenario of in-situ formation is disfavored on theoretical grounds:
-the required disk mass would likely drive migration, and so lacks
-internal consistency
-\citep{inamdar_formation_2015,ogihara_reassessment_2015}.
+A more detailed spectroscopic analysis of the transit would however be
+worthwhile to verify whether the TTV-slope correlation is yielding the
+correct orbital geometry.
+%The alternative
+%scenario of in-situ formation is disfavored on theoretical grounds:
+%the required disk mass would likely drive migration, and so lacks
+%internal consistency
+%\citep{inamdar_formation_2015,ogihara_reassessment_2015}.
 
@@
 \citep{ricker_transiting_2015,guerrero_tess_2021}.  An alternative
-path toward age-dating more planets could be to identify kinematic
+approach for age-dating more planets could be to identify kinematic
 associations around known exoplanet host stars using positions and
@@
 %
-Keck/NIRC2 imaging was acquired by programs 2015A/N301N2L
-(PI: A.~Kraus) and 2019A/N069 (PI: E.~Petigura).
+Keck/NIRC2 imaging was acquired by program 2015A/N301N2L
+(PI: A.~Kraus). % and 2019A/N069 (PI: E.~Petigura).
 %
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f13.pdf}
+		\includegraphics[width=0.7\textwidth]{f13.pdf}
 	\end{center}

### 2021-09-17 paper/ms.tex
update ms & add GD driver

@@
 period, and $P_{\rm rot}$ the rotation period.  Using the parameters
-from Table~\ref{tab:posterior}, this yields an expected 1.6\%
+from Table~\ref{tab:posterior}, this yields an expected 1.4\%
 distortion of the $\approx$1.8\,ppt transit depth: {\it i.e.}, an
-absolute deviation of $\approx$3\,ppm.  The observed residual has an
-amplitude of $\approx 30\,{\rm ppm}$.  Since the expected signal is
-smaller than the observed anomaly by an order of magnitude, gravity
+absolute deviation of $\approx$2.5\,ppm.  The observed residual has a
+semi-amplitude of $\approx 50\,{\rm ppm}$.  Since the expected signal is
+smaller than the observed anomaly by over an order of magnitude, gravity
 darkening seems to be an unlikely explanation.

### 2021-09-17 drivers/calc_breakup_gravity_darkening.py

@@ -1,0 +1,25 @@
+import numpy as np
+from astropy import units as u, constants as c
+
+P_rot = 2.642*u.day
+
+Rstar = 0.881*u.Rsun
+Mstar = 0.953*u.Msun
+
+P_break = (
+    2*np.pi*Rstar**(3/2) / ((c.G*Mstar)**(1/2))
+).to(u.day)
+
+print(f'P_break: {P_break:.3f} days')
+
+frac = ((P_break/P_rot)**2).cgs
+
+print(f'frac transit change: {frac*100:.4f}%')
+
+delta = 1800 # ppm
+delta_GD = 1800*frac
+
+print(f'delta_GD: {delta_GD:.2f} ppm')
+
+delta_obs = 100 # 0.1 ppt = 0.1 e-3 = 100 e-6
+print(f'delta_obs: {delta_obs:.2f} ppm')

### 2021-09-21 paper/ms.tex
update ms for RM discussion

@@
 \newcommand{\noriginal}{3{,}071} % Vizier; KC19 Theia 73 member count
-\newcommand{\nwithtess}{924} % How many stars have TESS data?
+\newcommand{\nwithtess}{924} % How many stars have TESS data with G<17 and (Bp-Rp)0 > 0.5?
+\newcommand{\nforgyro}{1842} % How many stars have G<17 and (Bp-Rp)0 > 0.5?
 \newcommand{\nkinwithtess}{391} % How many kinematically selected stars have TESS data?
@@
   observed by the Kepler satellite because it is nearby
-  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Here we show that
-  Kepler\,1627 is a member of the \clusterage\ old $\delta$~Lyr
-  cluster, based on Gaia kinematics, TESS stellar rotation periods,
-  and spectroscopic lithium abundances.  To our knowledge,
+  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Here we show
+  using Gaia kinematics, TESS stellar rotation periods, and
+  spectroscopic lithium abundances that Kepler\,1627 is a member of
+  the \clusterage\ old $\delta$~Lyr cluster.  To our knowledge,
   Kepler\,1627Ab is the youngest planet with a precise age yet found
@@
   one of many stellar groups whose properties have been clarified
-  using Gaia; many other new clusters also overlap with confirmed and
+  using Gaia; many other new clusters overlap with confirmed and
   candidate exoplanets.  Verifying the results of Gaia clustering
@@
 Jupiter
-\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020}.
+\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020,martioli_aumicbc_2021}.
 The Kepler mission however has not yielded any planets with precise
@@
     {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
-    points are kinematically selected candidate $\delta$\,Lyr
-    cluster members.  Both Kepler (gray panels) and TESS observed portions of
-    the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray or orange.
-    Kepler\,1627 (green star) was observed during the entirety of the
-    Kepler mission.  Colors of candidate cluster members indicate the
-    duration of TESS observations to date, in units of lunar months.
+    points are kinematically selected candidate $\delta$\,Lyr cluster
+    members.  Both Kepler (gray panels) and TESS observed portions of
+    the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray
+    or orange.  Kepler\,1627 (green star) was observed during the
+    entirety of the Kepler mission.  Colors of candidate cluster
+    members indicate the duration of TESS observations to date, in
+    units of lunar months.
     \label{fig:skychart}
@@
 	\caption{
-    {\bf The \cn\ is \clusterage\ old.}  {\it Top:}
-    Color-absolute magnitude diagram of candidate \cn\ members, in addition to
-    stars in IC\,2602 ($\approx38$\,Myr), the Pleiades ($\approx
-    115$\,Myr; \citealt{dahm_2015}), and the Gaia EDR3 Catalog of
-    Nearby Stars (gray background).  The zoomed right panel
-    highlights the pre-main-sequence.  The \cn\ and IC\,2602 are
-    approximately the same isochronal age.  {\it Bottom:} TESS and
-    Kepler stellar rotation period versus dereddened Gaia color,
-    with the Pleiades and Praesepe (650\,Myr) shown for reference
-    \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most
-    candidate \cn\ members are gyrochronally younger than the
-    Pleiades; outliers are probably field interlopers.
+    {\bf The \cn\ is \clusterage\ old.}  {\it Top:} Color-absolute
+    magnitude diagram of candidate \cn\ members, in addition to stars
+    in IC\,2602 ($\approx38$\,Myr), the Pleiades ($\approx 115$\,Myr;
+    \citealt{dahm_2015}), and the Gaia EDR3 Catalog of Nearby Stars
+    (gray background).  The zoomed right panel highlights the
+    pre-main-sequence.  The \cn\ and IC\,2602 are approximately the
+    same isochronal age.  {\it Bottom:} TESS and Kepler stellar
+    rotation period versus dereddened Gaia color, with the Pleiades
+    and Praesepe (650\,Myr) shown for reference
+    \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most candidate
+    \cn\ members are gyrochronally younger than the Pleiades; outliers
+    are probably field interlopers.
    \label{fig:age}
@@
 panel of Figure~\ref{fig:age} shows the color-absolute magnitude
-diagram (CAMD)
-of stars in the \cn, IC\,2602, the Pleiades, and the field.  The
-candidate \cn\ members shown are
-selected based on their positions and
-kinematics, as discussed in Appendix~\ref{app:kinematicselection}.
+diagram (CAMD) of stars in the \cn, IC\,2602, the Pleiades, and the
+field.  The candidate \cn\ members shown are selected based on their
+positions and kinematics (see Appendix~\ref{app:kinematicselection}).
 Those from the Pleiades and IC\,2602 are adopted from
@@
 and then applied the extinction coefficients $k_X\equiv A_X/A_0$
-computed by \citet{GaiaCollaboration2018} and assumed that $A_0 =
-3.1 E(B-V)$.  For IC\,2602, the Pleiades, and the \cn, this procedure
+computed by \citet{GaiaCollaboration2018} and assumed that $A_0 = 3.1
+E(B-V)$.  For IC\,2602, the Pleiades, and the \cn, this procedure
 yielded a respective mean and standard deviation for the reddening of
@@
 
-Figure~\ref{fig:age} shows that the \cn\ and IC\,2602 overlap, and are
-therefore approximately the same age.  In our exploration, we also
+Figure~\ref{fig:age} shows that the \cn\ and IC\,2602 overlap, and
+therefore are approximately the same age.  In our exploration, we also
 compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
@@
 association ($\approx$$16$\,Myr; \citealt{pecaut_star_2016}).  The
-pre-main-sequence M dwarfs of the \cn\ were intermediate between these
-latter two groups.  To turn this heuristic interpolation into an
-age measurement, we used the empirical method
-developed by \citet{gagne_mutau_2020}.  In brief, we fitted the
-pre-main-sequence loci of a set of reference clusters, and the locus
-of the target \cn\ was then modeled as a piecewise linear combination
-of these reference clusters.  For our reference clusters, we adopted
-members of UCL, IC\,2602, and the Pleiades from
-\citet{damiani_stellar_2019} and \cite{CantatGaudin2018a}
-respectively.  We removed binaries by requiring $\texttt{RUWE}<1.3$,
-$\texttt{radial\_velocity\_error}$ below the 80$^{\rm th}$ percentile
-of each cluster's distribution, and manually excluded stars that
-were obvious photometric binaries in the CAMD.  We then passed a moving
-box average and standard deviation across the CAMD in 0.10\,mag bins,
-fitted a univariate spline to the binned values, and assembled a
-piecewise grid of hybrid isochrones spanning the ages between UCL to
-the Pleiades following Equations~6 and~7 from
-\citet{gagne_mutau_2020}.
+pre-main-sequence M dwarfs of the \cn\ were intermediate between the
+latter two clusters.  To turn this heuristic interpolation into an age
+measurement, we used the empirical method developed by
+\citet{gagne_mutau_2020}.  In brief, we fitted the pre-main-sequence
+loci of a set of reference clusters, and the locus of the target \cn\
+was then modeled as a piecewise linear combination of these reference
+clusters.  For our reference clusters, we adopted members of UCL,
+IC\,2602, and the Pleiades from \citet{damiani_stellar_2019} and
+\cite{CantatGaudin2018a} respectively.  We removed binaries by
+requiring $\texttt{RUWE}<1.3$, $\texttt{radial\_velocity\_error}$
+below the 80$^{\rm th}$ percentile of each cluster's distribution, and
+manually excluded stars that were obvious photometric binaries in the
+CAMD.  We then passed a moving box average and standard deviation
+across the CAMD in 0.10\,mag bins, fitted a univariate spline to the
+binned values, and assembled a piecewise grid of hybrid isochrones
+spanning the ages between UCL to the Pleiades using Equations~6 and~7
+from \citet{gagne_mutau_2020}.
 
@@
 If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would
-be $31^{+5}_{-4}$\,Myr old.  For the converse case of 46\,Myr, the
+be $31^{+5}_{-4}$\,Myr old.  For the converse extreme of 46\,Myr, the
 \cn\ would be $44^{+8}_{-7}$\,Myr old.  We adopt an intermediate
 38\,Myr age for IC\,2602, which yields an age for the \cn\ of
-\clusterage.\footnote{Our exploration of the PARSEC and MIST
-isochrone models over a grid of ages, metallicities, and
-reddenings, yielded the best agreement for this $\approx 38\,$Myr age
-as well, given ${\rm [Fe/H] = +0.1}$ and $A_V=0.2$
-\citep{bressan_parsec_2012,choi_mesa_2016};  this preferred CAMD reddening
-is higher than the \citet{lallement_threedimensional_2018} value by a factor of two.
-} While detailed analyses of the LDB or main-sequence turn-off in the
-\cn\ could help resolve the tension between the LDB and isochrone
-ages, these are beyond our scope.
+\clusterage.\footnote{Our exploration of the PARSEC and MIST isochrone
+models over a grid of ages, metallicities, and reddenings, yielded the
+best agreement for this $\approx 38\,$Myr age as well, given ${\rm
+[Fe/H] = +0.1}$ and $A_V=0.2$
+\citep{bressan_parsec_2012,choi_mesa_2016};  this preferred CAMD
+reddening is higher than the \citet{lallement_threedimensional_2018}
+value by a factor of two.  } While detailed analyses of the LDB or
+main-sequence turn-off in the \cn\ could help resolve the tension
+between the LDB and isochrone ages, these are beyond our scope.
 
@@
 Of the \noriginal\ candidate \cn\ members reported by
-\citet{kounkel_untangling_2019}, TESS data were available for
-\nwithtess\ stars.\footnote{Kepler rotation periods were derived by
+\citet{kounkel_untangling_2019}, \nwithtess\ stars were amenable to
+rotation period measurements ($G<17$ and $\bpmrpo>0.5$) and had TESS
+data available.\footnote{Kepler rotation periods were derived by
 \citet{mcquillan_rotation_2014} for 56 stars; in these cases, we
@@
 restricted our attention to the \nkinwithtess\ stars discussed in
-Appendix~\ref{app:kinematicselection} in the spatial
-and kinematic proximity of Kepler\,1627.  To avoid blending, we also only focused on stars
-for which no companions were known with a brightness exceeding
+Appendix~\ref{app:kinematicselection} in the spatial and kinematic
+proximity of Kepler\,1627.  To avoid blending, we also only focused on
+stars for which no companions were known with a brightness exceeding
 one-tenth of the target star in a $21\farcs0$ radius.  There were
@@
 had TESS data available.  For plotting purposes we then imposed a
-selection based on the strength of the signal itself:
-we required the Lomb Scargle power to exceed 0.2, and the period to be
-below 15\,days.
+selection based on the strength of the signal itself: we required the
+Lomb Scargle power to exceed 0.2, and the period to be below 15\,days.
 
@@
 rotation period detection fraction would be
-$\nrotgood/\nkinwithtessandcrowding \approx \nfracprot\%$.
-The rotation periods of other 30\,Myr to 50\,Myr clusters
+$\nrotgood/\nkinwithtessandcrowding \approx \nfracprot\%$.  The
+rotation period distributions of other 30\,Myr to 50\,Myr clusters
 ({\it e.g.}, IC\,2602 and IC\,2391) are indistinguishable
-\citep{douglas_stephanie_t_2021_5131306}. 
-Binarity is known to be an important factor in the density of stars
-falling underneath the slow sequence
+\citep{douglas_stephanie_t_2021_5131306}.  Binarity is known to be an
+important factor in the density of stars falling underneath the slow
+sequence
 \citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}.
 Appendix~\ref{app:rotationbinarity} discusses this topic, and finds
-evidence for similar effects to those that have been previously
-noted in the literature.
+evidence for similar effects to those previously noted in the
+literature.
 
@@
 
-In summary, the age of the \cn\ based on the empirical
+To summarize, the age of the \cn\ based on the empirical
 pre-main-sequence interpolation method is \clusterage.  This method
@@
 reference cluster in the color-absolute magnitude diagram.
-Gyrochronologically, while most of the FGK dwarfs in the \cn\ have
-not yet converged to a slow sequence, their rotation periods 
-are generally consistent with an $\approx$38\,Myr age. The most
-precise age for \sn\ and its planet are therefore identical to the cluster,
-based on the star's spatial and kinematic association with the
-cluster, and the assumption that the planet formed shortly after
-the star.
+Gyrochronologically, while most of the FGK dwarfs in the \cn\ have not
+yet converged to a slow sequence, their rotation periods are generally
+consistent with a 38\,Myr age. The most precise age for \sn\ and its
+planet are therefore identical to the cluster, based on the star's
+spatial and kinematic association with the cluster, and the assumption
+that the planet formed shortly after the star.
 
@@
 \sn, there is a final consistency check: the abundance of
-Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet \citep[see {\it
+Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet \citep[{\it
 e.g.},][]{soderblom_ages_2014}.  We acquired an iodine-free spectrum
@@
 procedure described by \citet{bouma_2021_ngc2516}, we find ${\rm
-EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   Note that this value does not
-correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Given the
-effective temperature of \sn\ (Table~\ref{tab:starparams}), this
-measurement is also in line with expectations for a $\approx$38\,Myr
-star ({\it e.g.}, \citealt{randich_gaiaeso_2018};
+EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   %Note that this value does not
+%correct for the Fe\,\textsc{I} blend at 6707.44\AA. 
+Given the effective temperature of \sn\ (Table~\ref{tab:starparams}),
+this measurement is also in line with expectations for a 38\,Myr star
+({\it e.g.}, \citealt{randich_gaiaeso_2018};
 \citealt{bouma_2021_ngc2516}).
@@
     \citep{mazeh_time_2015}.  The units along the abscissa are most
-    easily understood by considering that the stellar flux
-    changes by $\sim$60\,ppt per half rotation period
+    easily understood by considering that the stellar flux changes by
+    $\sim$60\,ppt per half rotation period
     ($\sim$1.3\,days).
     {\it Bottom:} Radial velocities acquired using Keck/HIRES during
-    the transit of 2021 Aug 7, observed simultaneously with MuSCAT3
-    ({\it griz} photometry).  A quadratic baseline trend was allowed
-    in the RV fit.  The best-fit sky-projected obliquity in this model
-    is listed in red; the spectroscopic detection however has marginal statistical
-    significance compared to a ``trend-only'' model.
+    the transit of 2021 Aug 7, observed simultaneously with MuSCAT3.
+    The best-fit sky-projected obliquity for this model is listed in
+    red.  While the intra-exposure RV scatter precludes a detection of
+    the RM effect, the photometric transit depths are consistent
+    across the {\it griz} bandpasses.
     \label{fig:obliquity}
@@
 	\caption{
-		{\bf Radii, orbital periods, and ages of transiting
-			exoplanets}.  Planets older than $10^9$ years are shown with gray
-		markers.  Planets younger than a gigayear with ${\rm
-			\tau}/\sigma_{\rm \tau} > 3$ are emphasized, for $\tau$ the age.
-		\pn\ is shown with a
+    {\bf Radii, orbital periods, and ages of transiting exoplanets}.
+    Planets older than $10^9$ years are shown with gray markers.
+    Planets younger than a gigayear with ${\rm \tau}/\sigma_{\rm \tau}
+    > 3$ are emphasized, for $\tau$ the age.  \pn\ is shown with a
     star.  The large sizes of the youngest transiting planets could be
@@
 the main Kepler mission.\footnote{The re-purposed K2 mission however
-has found two younger systems containing five planets: K2-33b ($9\pm1\,{\rm Myr}$;
-\citealt{Mann_K2_33b_2016,David_et_al_2017}) and V1298 Tau
-($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  A portion of the
-light curve is shown with the phase-folded transit in
+has found two younger systems containing five planets: K2-33b
+($9\pm1\,{\rm Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and
+V1298 Tau ($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  A
+portion of the light curve is shown with the phase-folded transit in
 Figure~\ref{fig:lc}.  The \texttt{PDCSAP} photometry is dominated by a
@@
 false positive scenario was that of a background eclipsing binary,
-which had a model-dependent probability of $\approx10^{-5}$.
-As we describe in Appendix~\ref{app:companionstar} however,
-Kepler\,1627A is accompanied by a low-mass stellar companion ($M_{\rm
-B}\approx 0.33\,M_\odot$).  We therefore reconsidered this false
-positive scenario in some detail.
+which had a model-dependent probability of $\approx10^{-5}$.  As we
+describe in Appendix~\ref{app:companionstar} however, Kepler\,1627A is
+accompanied by a low-mass stellar companion ($M_{\rm B}\approx
+0.33\,M_\odot$).  We therefore reconsidered this false positive
+scenario in some detail.
 
@@
 
-There are three new lines of evidence that support the planetary
+There are two new lines of evidence that confirm the planetary
 interpretation.  First, the transit duration and the orbital period
@@
 we find $\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
-theoretically expected density for the companion is
-$\approx 4.6\,{\rm g\,cm}^{-3}$ (Table~\ref{tab:posterior};
-\citealt{choi_mesa_2016}).  The transit duration is therefore too long
-to be explained by a star eclipsing the M dwarf secondary at
-$10$-$\sigma$.
+theoretically expected density for the companion is $\approx 4.6\,{\rm
+g\,cm}^{-3}$ (Table~\ref{tab:posterior}; \citealt{choi_mesa_2016}).
+The transit duration is therefore too long to be explained by a star
+eclipsing the M dwarf secondary at $10$-$\sigma$.
 
@@
 The second line of confirmatory evidence comes by analyzing the
-individual Kepler transit times. We isolated each of the
-144 observed transits to within $\pm4.5$\,hr of each transit, and
-fitted each window with both {\it i)} a local second-order polynomial
-and transit, and {\it ii)} a local linear trend.  We let the mid-time
-of each transit float, and then calculated the residual between the
-measured mid-time and that of a periodic orbit.  This
-residual, the transit timing variation (TTV), is plotted in
-Figure~\ref{fig:obliquity} against the local linear slope.  A
-significant correlation of $-0.0564 \pm 0.0098\ {\rm ppt\,day}^{-1}$
-is observed.  Less than ten Kepler Objects of Interest have shown this
-correlation \citep{holczer_time_2015}, which is most readily
-interpreted as a TTV induced by unresolved starspot crossings
-\citep{mazeh_time_2015}.  This is only possible if the planet transits
-the primary star, which excludes a background eclipsing binary
-scenario, and therefore confirms that \pn\ is a planet.  It also
-implies that the planet's orbit is prograde.  The latter
-interpretation assumes that the dominant photometric variability is
-induced by dark spots, rather than bright faculae
-\citep{mazeh_time_2015}.  Given the observed transition of Sun-like
-stellar variability from spot to faculae-dominated regimes, we expect
-this inference to be reasonably secure
+individual Kepler transit times. We isolated each of the 144 observed
+transits to within $\pm4.5$\,hr of each transit, and fitted each
+window with both {\it i)} a local second-order polynomial and transit,
+and {\it ii)} a local linear trend.  We let the mid-time of each
+transit float, and then calculated the residual between the measured
+mid-time and that of a periodic orbit.  This residual, the transit
+timing variation (TTV), is plotted in Figure~\ref{fig:obliquity}
+against the local linear slope.  A significant correlation of $-0.0564
+\pm 0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Less than ten Kepler
+Objects of Interest have shown this correlation
+\citep{holczer_time_2015}, which is most readily interpreted as a TTV
+induced by unresolved starspot crossings \citep{mazeh_time_2015}.
+This is only possible if the planet transits the primary star, which
+excludes a background eclipsing binary scenario, and therefore
+confirms that \pn\ is a planet.  It also implies that the planet's
+orbit is prograde.  The latter interpretation assumes that the
+dominant photometric variability is induced by dark spots, rather than
+bright faculae.  Given the observed transition of Sun-like stellar
+variability from spot to faculae-dominated regimes, we expect this
+inference to be reasonably secure
 \citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
 
-The final piece of evidence comes from a spectroscopic transit we
-observed using Keck/HIRES during the night of 2021 Aug 7.  We
-used the iodine cell for wavelength calibration, and scheduled the
-transit using the ephemeris of \citet{holczer_transit_2016}.  We also acquired
-simultaneous multi-band photometry using MuSCAT3 \citep{Narita_2020}.
-The lower panels in Figure~\ref{fig:obliquity} show the results.  The
-photometry shows the expected starspot trend, along with the transit
-and a likely starspot crossing event at ${\rm JD} - 2459433 = 0.955$.
-The radial velocities show significant scatter, and an
-increase in RV is seen shortly after the expected ingress time.  The
-internal uncertainty on the RVs ($\approx 15\,{\rm m\,s}^{-1}$) is
-dominated by the $v\sin i$ broadening.
-
-% from /drivers/logs/rm.log
-% quad: 
-% chi2red: 4.452
-% BIC: 272.434
-% AIC: 264.796
-%
-% linear:
-% chi2red: 5.425
-% BIC: 291.987
-% AIC: 285.441
-%
-% trendonly:
-% chi2red: 4.256
-% BIC: 274.383
-% AIC: 270.018
-%
-% retrograde:
-% chi2red: 5.439
-% BIC: 292.415
-% AIC: 285.869
-% 
-% prograde:
-% chi2red: 4.320
-% BIC: 272.039
-% AIC: 265.493
-
-We fitted the RVs using the \citet{hirano_analytic_2010,hirano_2011}
-models for the Rossiter-McLaughlin (RM) effect, 
-following the methodology developed by \citet{stefansson_2020}.
-We allowed the sky-projected obliquity, the projected stellar equatorial
-velocity, and the Gaussian dispersion of the spectral lines to vary, and fixed the limb-darkening
-using the $V$-band tabulation from \citet{claret_gravity_2011}.  We also
-allowed
-for an optional linear and quadratic trend, and a white-noise jitter
-term.  The quadratic model with the RM effect is shown in
-Figure~\ref{fig:obliquity}; the prior on $v\sin i$ has been assumed
-from Table~\ref{tab:starparams}, and the jitter term has been
-incorporated.
-We found $\chi^2_{\rm red} =
-5.4$ for the best-fitting linear case, and $\chi^2_{\rm red} = 4.4$
-for the best-fitting quadratic, with a $\Delta {\rm BIC} = {\rm
-BIC}_{\rm line} - {\rm BIC}_{\rm quad} = 19.6$ and $\Delta {\rm AIC} =
-{\rm AIC}_{\rm line} - {\rm AIC}_{\rm quad} = 20.6$.  While this
-provides strong statistical evidence to prefer the quadratic model
-over the linear one, the $\chi^2_{\rm red}$ indicate that neither fit
-is very good.  Considering a quadratic trend without any RM
-effect yields $\chi^2_{\rm red} = 4.3$, and $\Delta {\rm BIC} = {\rm
-BIC}_{\rm trend} - {\rm BIC}_{\rm quad} = 1.9$ ($\Delta {\rm
-AIC}=5.2$).  From a model-comparison standpoint, there is therefore
-only marginal statistical evidence to prefer the best-fitting ``RM and
-quadratic trend'' model over the best-fitting ``quadratic trend only''
-model.  Comparing the expected semi-amplitude of the velocity anomaly
-for a well-aligned orbit ($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$) with the achieved precision,
-we cannot claim a detection of the Doppler transit.
-From the MuSCAT3 data however, we do know that the transit
-was observed, and the increase in radial velocity seen shortly after
-ingress is more consistent with a prograde or polar orbit than a
-retrograde orbit.  Fixing the orbit to be either exactly prograde or
-exactly retrograde, we find $\Delta {\rm BIC} = {\rm BIC}_{\rm
-retrograde} - {\rm BIC}_{\rm trend} = 18.0$; the retrograde case is
-ruled out on this basis.
-% More data will be needed however to
-% make a more precise measurement of the stellar obliquity.
+A third supporting line of evidence for the planetary interpretation
+also exists, though we consider it less definitive than the stellar
+density and TTV-local slope correlation.  We observed a transit of
+\pn\ on the night of 2021 Aug 7 simultaneously with Keck/HIRES and
+MuSCAT3.  We scheduled the observations using the ephemeris of
+\citet{holczer_transit_2016}.  Although we did not detect the
+Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
+show the transit, and it appears to be achromatic
+(Figure~\ref{fig:obliquity}, lower panels).  Fitting the MuSCAT3
+photometry with a model that lets the transit depths vary across
+bandpasses, we find {\it griz} depths consistent with the Kepler depth
+at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  The MuSCAT3
+observations also suggest a transit duration $17.4\pm3.6\,{\rm
+minutes}$ shorter than the Kepler transits; further photometric
+follow-up could help confirm whether such changes are indeed
+happening.  For the RM analysis, the details are discussed in
+Appendix~\ref{app:rm}.  While our RM data are marginally more
+consistent with a prograde or polar orbit than a retrograde orbit,
+given that the expected velocity anomaly ($\Delta v_{\rm RM}\approx
+20\,{\rm m\,s}^{-1}$) is a factor of two smaller than the
+spot-corrected exposure-to-exposure scatter, we are not in a position
+to claim a spectroscopic detection of the RM effect, nor a precise
+quantification of the stellar obliquity.
 
@@
 
-%, which is expected to be outflowing at a rate of
-%$\sim$$10^{-2}$\,$M_\oplus\,{\rm Myr}^{-1}$ \citep[{\it
-%e.g.},][]{owen_constraining_2020}.
-
 Kepler\,1627Ab provides a new extremum in the ages of the Kepler
 planets, and opens multiple avenues for further study.  Further
-observations of Doppler transits could yield a more precise stellar
-obliquity, and could confirm or refute the expected prograde orbital
-geometry.  Similarly, transit spectroscopy aimed at detecting outflows
-of hydrogen or helium could yield insight into the evolutionary state
-of the atmosphere \citep{spake_helium_2018,vissapragada_2020}.  Determining the
-precise amount of high-energy irradiation incident on the planet would
-also help in clarifying the expected outflow rate \citep[{\it
+observations of spectroscopic transits could yield a more precise
+stellar obliquity, and could confirm or refute the expected prograde
+orbital geometry.  Similarly, transit spectroscopy aimed at detecting
+outflows of hydrogen or helium could yield insight into the
+evolutionary state of the atmosphere
+\citep{spake_helium_2018,vissapragada_2020}.  Determining the precise
+amount of high-energy irradiation incident on the planet would also
+help in clarifying the expected outflow rate \citep[{\it
 e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
 quantity to measure would be the planet's mass.  This might be
-possible through timing analyses of the Kepler data,
-or through a high-cadence multi-color radial velocity campaign.
-Measured at sufficient precision, the mass could yield constraints on
-both the planet's composition and its initial entropy
+possible through timing analyses of the Kepler data, or through a
+high-cadence multi-color radial velocity campaign.  Measured at
+sufficient precision, the mass could yield constraints on both the
+planet's composition and its initial entropy
 \citep{owen_constraining_2020}.
@@
 Although our transit timing study did not yield the detection of any
-clearly significant planet-planet TTVs, Appendix~\ref{app:flare}
-highlights an odd flare timing coincidence that may be suggestive of
-star-planet interactions.  We encourage its
+planet-induced transit timing or transit duration variations,
+Appendix~\ref{app:flare} highlights an odd flare timing coincidence
+that may be suggestive of star-planet interactions.  We encourage its
 exploration by investigators more versed in the topic than ourselves.
@@
 mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
-A more detailed spectroscopic analysis of the transit would however be
+Spectroscopic follow-up observations of the transit would however be
 worthwhile to verify whether the TTV-slope correlation is yielding the
 correct orbital geometry.
-%The alternative
-%scenario of in-situ formation is disfavored on theoretical grounds:
-%the required disk mass would likely drive migration, and so lacks
-%internal consistency
+
+%NB to arxiv readers: The alternative scenario of in-situ formation is
+%disfavored on theoretical grounds: the required disk mass would
+%likely drive migration, and so lacks internal consistency
 %\citep{inamdar_formation_2015,ogihara_reassessment_2015}.
@@
 
-The authors are grateful to T{.}~David, J{.}~Winn, and A{.}~Howard for
-illuminating discussions and guidance, and to R{.}~Kerr for kindly
-providing us with the \citet{Kerr2021} membership list prior to its
-publication.
-The authors are also grateful to K{.}~Collins for
-helping resolve the scheduling conflict that would have otherwise
-prevented the MuSCAT3 observations.
+The authors are grateful to T{.}~David, A{.}~Howard, J{.}~Spake, and
+J{.}~Winn for illuminating discussions and suggestions, and to
+R{.}~Kerr for kindly providing us with the \citet{Kerr2021} membership
+list prior to its publication.
+The authors are also grateful to K{.}~Collins for helping resolve the
+scheduling conflict that would have otherwise prevented the MuSCAT3
+observations.
 % is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
@@
 L.G.B{.} acknowledges support from a Charlotte Elizabeth Procter
-Fellowship from Princeton University, as well as from the TESS GI Program, programs
-G011103 and G022117, through NASA grants 80NSSC19K0386 and
-80NSSC19K1728.
+Fellowship from Princeton University, as well as from the TESS GI
+Program, programs G011103 and G022117, through NASA grants
+80NSSC19K0386 and 80NSSC19K1728.
 %
@@
  	{\it Astrometry}:
- 	Gaia \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2020_edr3}.
+ 	Gaia \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2021_edr3}.
  	{\it Imaging}:
@@
 
+\section{Spectroscopic Transit Analysis}
+\label{app:rm}
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\subfloat{
+			\includegraphics[width=\textwidth]{f18.pdf}
+		}
+	\end{center}
+	\vspace{-0.5cm}
+	\caption{
+    {\bf Spectroscopic activity indicators during the transit of 2021
+    Aug 7.} The {\it top panels} show the median line profiles Ca K,
+    Ca H, and H$\alpha$ line profiles from the HIRES spectra.  The
+    {\it lower panels} show the differences of each individual
+    spectrum relative to the median spectrum.  The bump in the red
+    wing of Ca H is H$\epsilon$.
+    The spectra in the lower panels are smoothed for visualization
+    purposes.
+    \label{fig:rvactivity}
+	}
+\end{figure*}
+
+We monitored \sn\ with Keck/HIRES before, during, and after transit on
+the night of 2021 Aug 7.  We used the iodine cell for wavelength
+calibration\footnote{We considered a line-profile based analysis in
+the regions without iodine, but the line profile stability of HIRES
+precludes such an approach.}, extracted the 1-D spectra using the
+standard California Planet Survey pipeline \citep{howard_cps_2010},
+and measured the velocities by cross-correlating against a template
+found via spectral classification with \texttt{SpecMatch-Emp}
+\citep{yee_SM_2017}.  The airmass ranged between 1.1 and 2.2 from the
+start through the end of observations; the seeing ranged from
+$1\farcs1$ at the beginning to $1\farcs5$ at the end.  We also
+simultaneously observed across {\it griz} bands using MuSCAT3 at
+Faulkes Telescope North on Maui, HI.
+
+The lower panels in Figure~\ref{fig:obliquity} show the results.  The
+MuSCAT3 photometry shows the expected starspot trend, along with the
+transit and what is likely a chromatic starspot crossing event at
+${\rm JD} - 2459433 = 0.955$.  The radial velocities decrease by
+800\ms\ over the six hour window.  A $\approx50$\ms\  increase in RV
+is seen shortly after the expected ingress time.
+
+Overall, we expect the dominant trends in both the photometry and
+radial velocities to be caused by starspots on the stellar photosphere
+rotating into and out of view.  The plasma in the leading and receding
+limbs of the stellar disk has an apparent line-of-sight velocity of
+$\pm 20$\kms.  Over 10\% of a rotation cycle ($P_{\rm
+rot}=2.6\,{\rm days}$), spots near these limbs come into and out of
+view, modulate the stellar velocity profile, and can thereby produce
+the overall 800\ms\ trend.  The Ca HK and H$\alpha$ emission
+profiles support this interpretation;
+Figure~\ref{fig:rvactivity} shows that each line gradually decreases in
+intensity over the course of the six hour sequence.
+The overall RVs are also heavily correlated against the S-indices
+derived from the Ca HK lines.
+
+The expectation however is for the spot-induced signals to be smooth,
+at worst with contributions at $0.5\,P_{\rm rot}$ or $0.25\,P_{\rm
+rot}$ \citep{klein_simulated_2020}.  We therefore fitted the RVs using
+the \citet{hirano_analytic_2010,hirano_2011} models for the
+Rossiter-McLaughlin (RM) effect, and allowed for an optional linear
+and quadratic trend in time to fit the spot-induced signal.  We
+followed the methodology developed by \citet{stefansson_2020}.  We
+allowed the sky-projected obliquity, the projected stellar equatorial
+velocity, and the Gaussian dispersion of the spectral lines to vary,
+and fixed the limb-darkening using the $V$-band tabulation from
+\citet{claret_gravity_2011}.  We assumed a prior on $v\sin i$ and
+$a/R_\star$ from Table~\ref{tab:starparams}, and also allowed for a
+white-noise jitter term to be added in quadrature to the measurement
+uncertainties.
+
+The quadratic model with the RM effect is shown in
+Figure~\ref{fig:obliquity}; the jitter term is incorporated in the
+model uncertainties, but not the plotted measurement uncertainties.
+The plotted measurement uncertainties are the internal uncertainties
+on the RVs ($\approx 15\,{\rm m\,s}^{-1}$), and are dominated by the
+$v\sin i$ broadening.  However, between exposures, the RVs show
+significant additional scatter that is not captured by the slow quadratic
+trend.  The white-noise jitter for this
+particular model is $\sigma_{\rm RV} = 50^{+11}_{-8}$\ms, which is
+larger than the expected RM anomaly of $\Delta v_{\rm RM}
+\approx f_{\rm LD} \cdot \delta \cdot v\sin i \cdot \sqrt{1-b^2}
+\approx 20\,{\rm m\,s}^{-1}$.
+
+The presence of this additional scatter prevents a clear detection of
+the RM effect.  The reason can be understood through a
+model-comparison analysis.  If we compare the model with a quadratic trend and
+the RM effect (Figure~\ref{fig:obliquity}) against a model with a
+linear trend and the RM effect, or even a model with no RM effect at
+all, then the usual goodness of fit diagnostics are as follows.
+% from logs/rm_freejitter.log
+\begin{align}
+  \chi^2_{\rm red} &= 1.45 \quad ({\rm Quadratic+RM}) \nonumber \\
+  \chi^2_{\rm red} &= 1.37 \quad ({\rm Linear+RM}) \nonumber \\
+  \chi^2_{\rm red} &= 1.22 \quad ({\rm Only\ Quadratic}),
+\end{align}
+and
+\begin{align}
+  {\rm BIC} &= 251.7  \quad ({\rm Quadratic+RM}) \nonumber \\
+  {\rm BIC} &= 257.7  \quad ({\rm Linear+RM}) \nonumber \\
+  {\rm BIC} &= 242.8  \quad ({\rm Only\ Quadratic}).
+\end{align}
+The ``only quadratic'' model does well in these metrics
+because it can inflate the jitter term to account for any additional
+scatter during the transit.  However, we cannot justify a physical
+prior on the jitter term, because we do not understand the origin of
+the exposure-to-exposure scatter.  As noted above, the velocity
+deviations from starspots are expected to have contributions at the
+stellar rotation frequency, or harmonics thereof.  This jitter is
+present on the exposure timescale (15 minutes), which is only 0.4\% of
+the stellar rotation period; it is not obvious that starspots could be
+the culprit.
+
+The amplitude of both the spot-induced trend and
+the jitter are larger than recent comparable measurements
+in systems such as AU~Mic
+\citep{palle_transmission_2020}, DS~Tuc
+\citep{montet_young_2020,zhou_well_2020} and TOI~942
+\citep{wirth_2021_toi942}.  One possible explanation for the jitter is that it is
+astrophysical in origin, and that it is caused by some novel process
+operating on the surface of Kepler 1627A.  Another possibility is that our RV
+analysis underestimates the measurement uncertainties; the S/N per
+resolution element for each of our spectra (70 to 80) was somewhat low
+due to the observing requirements, and the star is a rapid rotator.
+We currently favor the latter possibility.  Observations at higher
+precision remain worthwhile to clarify the orbital geometry of \pn.
+
 \section{Flare Analysis}
 \label{app:flare}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f7.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Flares in Kepler\,1627}.  
-    {\it Top:}
-    The full short cadence Kepler dataset, acquired at 1-minute
-    sampling (black points) is shown with a stellar variability model
-    (blue line).
-    {\it Middle:}
-    Residual after subtracting the stellar variability model.  Flares
-    appear as spikes.
-    {\it Bottom:}
-    Zooms of the brightest, and third-brightest flares.  A timing
-    coincidence -- that both flares have ``successors'' approximately
-    one orbital period after the initial event -- is emphasized.
-		\label{fig:flarezoom}
-	}
-\end{figure*}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f8.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Phase-folded flares in Kepler\,1627}.  
-    {\it Top:}
-    As in the middle of Figure~\ref{fig:flarezoom}, phase-folded at
-    the planet's ephemeris.
-    {\it Bottom:}
-    Fitted flare amplitudes and orbital phases -- each point
-    represents one flare.
-    Colors indicate relative time, from the beginning of the
-    98-day short cadence Quarter 15 dataset (dark blue) to the end (light
-    yellow).
-    Lower amplitude flares likely exist in the data, and were not
-    examined.
-		\label{fig:flarephase}
-	}
-\end{figure*}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.7\textwidth]{f13.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Statistics of inter-flare arrival times}.  
-    24 flares were recorded with amplitudes exceeding 0.5\% over the
-    97.7 days of short cadence observations.  The histogram of the
-    time intervals between every possible pair of flares is shown in
-    black.  Some plausibly important timescales for star-planet
-    interactions, namely the planetary orbital period and synodic
-    period (the orbital period as seen from the rotating stellar frame) are
-    shown along with their linear combinations.  Monte Carlo
-    draws from a Poisson distribution are shown
-    with the gray bands.  While peaks in the observed distribution do
-    coincide with some of the ``special periods'', the
-    statistical evidence for a non-Poissonian process driving the
-    flares does not reach the $5$-$\sigma$ threshold.
-		\label{fig:flarestats}
-	}
-\end{figure*}
 
@@
 time, an exponential lag time, and an amplitude.
-Figure~\ref{fig:flarephase} shows the resulting flares, amplitudes,
-and phases.
+%Figure~\ref{fig:flarephase} shows the resulting flares, amplitudes,
+%and phases.
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f7.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Flares in Kepler\,1627}.  
+		{\it Top:}
+		The full short cadence Kepler dataset, acquired at 1-minute
+		sampling (black points) is shown with a stellar variability model
+		(blue line).
+		{\it Middle:}
+		Residual after subtracting the stellar variability model.  Flares
+		appear as spikes.
+		{\it Bottom:}
+		Zooms of the brightest, and third-brightest flares.  A timing
+		coincidence -- that both flares have ``successors'' approximately
+		one orbital period after the initial event -- is emphasized.
+		\label{fig:flarezoom}
+	}
+\end{figure*}
+
+%\begin{figure*}[t]
+%	\begin{center}
+%		\leavevmode
+%		\includegraphics[width=0.9\textwidth]{f8.pdf}
+%	\end{center}
+%	\vspace{-0.7cm}
+%	\caption{
+%		{\bf Phase-folded flares in Kepler\,1627}.  
+%    {\it Top:}
+%    As in the middle of Figure~\ref{fig:flarezoom}, phase-folded at
+%    the planet's ephemeris.
+%    {\it Bottom:}
+%    Fitted flare amplitudes and orbital phases -- each point
+%    represents one flare.
+%    Colors indicate relative time, from the beginning of the
+%    98-day short cadence Quarter 15 dataset (dark blue) to the end (light
+%    yellow).
+%    Lower amplitude flares likely exist in the data, and were not
+%    examined.
+%		\label{fig:flarephase}
+%	}
+%\end{figure*}
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.8\textwidth]{f13.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Statistics of inter-flare arrival times}.  
+		24 flares were recorded with amplitudes exceeding 0.5\% over the
+		97.7 days of short cadence observations.  The histogram of the
+		time intervals between every possible pair of flares is shown in
+		black.  Some plausibly important timescales for star-planet
+		interactions, namely the planetary orbital period and synodic
+		period (the orbital period as seen from the rotating stellar frame) are
+		shown along with their linear combinations.  Monte Carlo
+		draws from a Poisson distribution are shown
+		with the gray bands.  While peaks in the observed distribution do
+		coincide with some of the ``special periods'', the
+		statistical evidence for a non-Poissonian process driving the
+		flares does not reach the $5$-$\sigma$ threshold.
+		\label{fig:flarestats}
+	}
+\end{figure*}
 

### 2021-09-21 drivers/calc_tdur_difference_MUSCAT_Kepler.py

@@ -1,3 +1,8 @@
+"""
+Compare fitted value from MUSCAT routines
+"""
+import numpy as np
+
 Tdur_0 = 2.538
 eTdur_0 = 0.030
 
@@ -8,3 +13,20 @@ e = (eTdur_1**2 + eTdur_0**2)**(1/2)
 
 N_sigma = (Tdur_1 - Tdur_0)/e
 print(f'Difference = {Tdur_1 - Tdur_0:.3f} +/- {e:.3f} hr ({N_sigma:.1f}σ)')
+
+
+colors = 'g,r,i,z'.split(',')
+rps = [0.315, 0.291, 0.307, 0.270]
+rp_uncs = [0.025, 0.021, 0.024, 0.020]
+
+rp_0 = 0.298
+erp_0 = 0.016
+
+for _r, _er, c in zip(rps, rp_uncs, colors):
+
+    e = (erp_0**2 + _er**2)**(1/2)
+
+    N_sigma = np.abs((_r - rp_0)/e)
+    print(f'{c}-band: Difference = {_r - rp_0:.3f} +/- {e:.3f} Rjup ({N_sigma:.1f}σ)')
+
+

### 2021-09-21 paper/ms.tex
update ms

@@
 consistent with a prograde or polar orbit than a retrograde orbit,
-given that the expected velocity anomaly ($\Delta v_{\rm RM}\approx
+the expected velocity anomaly ($\Delta v_{\rm RM}\approx
 20\,{\rm m\,s}^{-1}$) is a factor of two smaller than the
-spot-corrected exposure-to-exposure scatter, we are not in a position
-to claim a spectroscopic detection of the RM effect, nor a precise
-quantification of the stellar obliquity.
+spot-corrected exposure-to-exposure scatter.  We are therefore not in
+a position to claim a spectroscopic detection of the RM effect, nor a
+precise quantification of the stellar obliquity.
 
@@
 quantity to measure would be the planet's mass.  This might be
-possible through timing analyses of the Kepler data, or through a
-high-cadence multi-color radial velocity campaign.  Measured at
-sufficient precision, the mass could yield constraints on both the
-planet's composition and its initial entropy
+possible through long-term monitoring of the transit times and
+durations, or through a high-cadence multi-color radial velocity
+campaign.  Measured at sufficient precision, the mass could yield
+constraints on both the planet's composition and its initial entropy
 \citep{owen_constraining_2020}.
@@
 Although our transit timing study did not yield the detection of any
-planet-induced transit timing or transit duration variations,
+planet-induced transit timing or duration variations,
 Appendix~\ref{app:flare} highlights an odd flare timing coincidence
@@
 youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
-with precise ages below 100\,Myr include K2-33, DS~Tuc~A, HIP~67522,
+with precise ages below 100\,Myr include K2-33, DS~Tuc, HIP~67522,
 TOI~837, the two-planet AU~Mic system, and the four-planet V1298~Tau
@@
 mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
-Spectroscopic follow-up observations of the transit would however be
-worthwhile to verify whether the TTV-slope correlation is yielding the
+Spectroscopic follow-up observations of the transit remain necessary
+however to verify whether the TTV-slope correlation is yielding the
 correct orbital geometry.

### 2021-09-21 paper/ms.tex
update ms

@@
 
-The amplitude of both the spot-induced trend and
-the jitter are larger than recent comparable measurements
-in systems such as AU~Mic
+The amplitude of both the spot-induced trend and the jitter are larger
+than recent comparable measurements in systems such as AU~Mic
 \citep{palle_transmission_2020}, DS~Tuc
 \citep{montet_young_2020,zhou_well_2020} and TOI~942
-\citep{wirth_2021_toi942}.  One possible explanation for the jitter is that it is
-astrophysical in origin, and that it is caused by some novel process
-operating on the surface of Kepler 1627A.  Another possibility is that our RV
-analysis underestimates the measurement uncertainties; the S/N per
-resolution element for each of our spectra (70 to 80) was somewhat low
-due to the observing requirements, and the star is a rapid rotator.
-We currently favor the latter possibility.  Observations at higher
-precision remain worthwhile to clarify the orbital geometry of \pn.
+\citep{wirth_2021_toi942}.  One possible explanation for the jitter is
+that it is astrophysical in origin, and that it is caused by some
+novel process operating on the surface of Kepler 1627A.  Another
+possibility is that our RV analysis underestimates the measurement
+uncertainties; the S/N per resolution element for each of our spectra
+(70 to 80) was moderated by the observing requirements, and the rapid
+rotation of the star could affect the velocity extraction.
+Observations at higher precision are necessary to distinguish these
+two possibilities, and remain worthwhile as a means toward clarifying
+the orbital geometry of \pn.
 
@@
     interactions, namely the planetary orbital period and synodic
-		period (the orbital period as seen from the rotating stellar frame) are
-		shown along with their linear combinations.  Monte Carlo
-		draws from a Poisson distribution are shown
-		with the gray bands.  While peaks in the observed distribution do
-		coincide with some of the ``special periods'', the
-		statistical evidence for a non-Poissonian process driving the
-		flares does not reach the $5$-$\sigma$ threshold.
+    period (the orbital period as seen from the rotating stellar
+    frame) are shown along with their linear combinations.  Monte
+    Carlo draws from a Poisson distribution are shown with the gray
+    bands.  While peaks in the observed distribution do coincide with
+    some of the ``special periods'', the statistical evidence for a
+    non-Poissonian process driving the flares does not reach the
+    $5$-$\sigma$ threshold.
 		\label{fig:flarestats}
@@
 There were $N_{\rm f}=24$ flares that exceeded $0.5\%$ in relative
-flux during the short cadence observations.  
-These 24 flares spanned a
+flux during the short cadence observations.  These 24 flares spanned a
 total of 6.5 hours ($\sim$15 minutes per flare).  Inspecting the data,
-we noticed a coincidence in the flare arrival
-times.  The coincidence is that despite the low flare duty cycle, one
-orbital period after the brightest flare, a second flare followed.
-This and a similar event are shown in Figure~\ref{fig:flarezoom}.  The
-timing error is good to a $\approx0.2\%$ difference from the orbital
-period, which given the duty cycle seems {\it a priori} unlikely.  If
-we consider flares falling within 2\% of the planet's orbital period
-after a previous flare, then 4 of the 24 flare events have candidate
-``successors''.
+we noticed a coincidence in the flare arrival times.  The coincidence
+is that despite the low flare duty cycle, one orbital period after the
+brightest flare, a second flare followed.  This and a similar event
+are shown in Figure~\ref{fig:flarezoom}.  The timing error is good to
+a $\approx0.2\%$ difference from the orbital period, which given the
+duty cycle seems {\it a priori} unlikely.  If we consider flares
+falling within 2\% of the planet's orbital period after a previous
+flare, then 4 of the 24 flare events have candidate ``successors''.
 
 As with any coincidence, if one does not have a firm prediction, it is
-difficult to assess whether a surprise is statistically significant.  Since
-our surprise was specifically at the inter-arrival time of certain
-flares coinciding with special time intervals, we performed the
-following analysis.  First, we considered all unordered pairs of
+difficult to assess whether a surprise is statistically significant.
+Since our surprise was specifically at the inter-arrival time of
+certain flares coinciding with special time intervals, we performed
+the following analysis.  First, we considered all unordered pairs of
 flares.  For $N$ flares there are ${n \choose 2}$ such pairs (for our
 case, 276 pairs).  We then compared the distribution of the pair
-separations against that of a Poisson distribution.  Specifically,
-we drew $N_{\rm f}=24$ samples from a Poisson distribution with
-$\lambda = \Delta t / N_{\rm f}$, for $\Delta t=97.7\,{\rm days}$ the
-total duration of the observations, and repeated the draw $10^3$ times with
+separations against that of a Poisson distribution.  Specifically, we
+drew $N_{\rm f}=24$ samples from a Poisson distribution with $\lambda
+= \Delta t / N_{\rm f}$, for $\Delta t=97.7\,{\rm days}$ the total
+duration of the observations, and repeated the draw $10^3$ times with
 unique random seeds.
@@
 $P_{\rm syn} = (P_{\rm rot}^{-1} - P_{\rm orb}^{-1})^{-1}$, and linear
-combinations thereof.  The tidal period (half the synodic
-period) is not shown.  The bins are logarithmically spaced to give 100
-bins between the minimum and maximum ordinate values.  The gray bands
+combinations thereof.  The tidal period (half the synodic period) is
+not shown.  The bins are logarithmically spaced to give 100 bins
+between the minimum and maximum ordinate values.  The gray bands
 express the range of values observed from the Poissonian draws.  While
-it does seem like an odd coincidence for peaks in the observed
-flare arrival time distribution to coincide with the locations of
-these ``special intervals'', the statistical evidence for a
-non-Poissonian process driving the flares does not seem especially
-overwhelming.  More quantitatively, the peaks observed at the orbital and
-synodic periods are within the $\pm 2$-$\sigma$ range of a Poissonian
-process, and those at $P_{\rm orb}+P_{\rm syn}$ and $P_{\rm
-orb}+2P_{\rm syn}$ are only slightly above this range.  With that
-said, future analyses of these data by investigators with more
-knowledge of this topic could very well yield more quantitative
-insights.
-Such analyses should keep in mind an important caveat:
-the amplitude
-distribution of M-dwarf flares extends up to many times the quiescent
-flux \citep[see Figure~7 of][]{gunther_2020}.  A flare on
-Kepler\,1627B producing double its quiescent white-light flux would yield
-a $\approx$1\% apparent amplitude.
-Such flares could represent a significant fraction
-of those in the Kepler observations.
+it does seem like an odd coincidence for peaks in the observed flare
+arrival time distribution to coincide with the locations of these
+``special intervals'', the statistical evidence for a non-Poissonian
+process driving the flares does not seem especially overwhelming.
+More quantitatively, the peaks observed at the orbital and synodic
+periods are within the $\pm 2$-$\sigma$ range of a Poissonian process,
+and those at $P_{\rm orb}+P_{\rm syn}$ and $P_{\rm orb}+2P_{\rm syn}$
+are only slightly above this range.  With that said, future analyses
+of these data by investigators with more knowledge of this topic could
+very well yield more quantitative insights.  Such analyses should keep
+in mind an important caveat: the amplitude distribution of M-dwarf
+flares extends up to many times the quiescent flux \citep[see Figure~7
+of][]{gunther_2020}.  A flare on Kepler\,1627B producing double its
+quiescent white-light flux would yield a $\approx$1\% apparent
+amplitude.  Such flares could represent a significant fraction of
+those in the Kepler observations.
 

### 2021-09-21 paper/ms.tex
update ms

@@
 linear trend and the RM effect, or even a model with no RM effect at
-all, then the usual goodness of fit diagnostics are as follows.
+all, then the respective Bayesian Information Criterion (BIC) values
+are as follows.
 % from logs/rm_freejitter.log
-\begin{align}
-  \chi^2_{\rm red} &= 1.45 \quad ({\rm Quadratic+RM}) \nonumber \\
-  \chi^2_{\rm red} &= 1.37 \quad ({\rm Linear+RM}) \nonumber \\
-  \chi^2_{\rm red} &= 1.22 \quad ({\rm Only\ Quadratic}),
-\end{align}
-and
+% \begin{align}
+%   \chi^2_{\rm red} &= 1.45 \quad ({\rm Quadratic+RM}) \nonumber \\
+%   \chi^2_{\rm red} &= 1.37 \quad ({\rm Linear+RM}) \nonumber \\
+%   \chi^2_{\rm red} &= 1.22 \quad ({\rm Only\ Quadratic}),
+% \end{align}
+% and
 \begin{align}
@@
 \end{align}
-The ``only quadratic'' model does well in these metrics
-because it can inflate the jitter term to account for any additional
-scatter during the transit.  However, we cannot justify a physical
-prior on the jitter term, because we do not understand the origin of
-the exposure-to-exposure scatter.  As noted above, the velocity
-deviations from starspots are expected to have contributions at the
-stellar rotation frequency, or harmonics thereof.  This jitter is
-present on the exposure timescale (15 minutes), which is only 0.4\% of
-the stellar rotation period; it is not obvious that starspots could be
-the culprit.
+There is therefore no evidence to prefer the model with the RM effect
+against a model that only accounts for the stellar variability.
+The ``only quadratic'' model does particularly well because it can
+inflate the jitter term to account for any additional scatter during
+the transit, and it has fewer free parameters.  However, we cannot
+justify a physical prior on the jitter term, because we do not
+understand the origin of the exposure-to-exposure scatter.  As noted
+above, the velocity deviations from starspots are expected to have
+contributions at the stellar rotation frequency, or harmonics thereof.
+This jitter is present on the exposure timescale (15 minutes), which
+is only 0.4\% of the stellar rotation period; it is not obvious that
+starspots could be the culprit.
 

### 2021-09-21 paper/ms.tex
update ms
Co-authored-by: Kento Masuda <kmasuda@ess.sci.osaka-u.ac.jp>

@@
   $3.3\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
-  observed by the Kepler satellite because it is nearby
+  observed by the Kepler space telescope because it is nearby
   ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Here we show
@@
   events in the Kepler photometry and a Keck/HIRES transit we find
-  that the planet is likely on an orbit aligned with the stellar spin
-  axis.  We also briefly explore the flare arrival time distribution
-  in the short cadence Kepler photometry.  The $\delta$~Lyr cluster is
-  one of many stellar groups whose properties have been clarified
-  using Gaia; many other new clusters overlap with confirmed and
-  candidate exoplanets.  Verifying the results of Gaia clustering
-  analyses using TESS rotation periods offers a promising path toward
-  expanding the census of age-dated planets, and our corresponding
-  understanding of planetary evolution.
+  that the planet is likely on a prograde orbit.  We also briefly
+  explore the flare arrival time distribution in the short cadence
+  Kepler photometry.  The $\delta$~Lyr cluster is one of many stellar
+  groups whose properties have been clarified using Gaia; many other
+  exoplanet hosts belong to these new clusters.  Verifying the results
+  of Gaia clustering analyses using TESS rotation periods offers a
+  promising path toward expanding the census of age-dated planets, and
+  our corresponding understanding of planetary evolution.
 \end{abstract}
@@
 ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
-during the main Kepler mission (2009--2013), only four open
-clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
-NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
-\citep{meibom_kepler_2011}.  Since that time, analyses of the
-kinematic, photometric, and astrometric Gaia data have expanded our
-knowledge of open cluster and moving group memberships \citep[{\it
+during the main Kepler mission (2009--2013), only four open clusters
+were known in the Kepler field: NGC\,6866, NGC\,6811, NGC\,6819, and
+NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
+\citep{meibom_kepler_2011}.  Though isochronal and gyrochronal
+analyses suggest that younger Kepler planets do exist
+\citep{berger_identifying_2018,david_sizes_2021}, accurate and precise
+age measurements typically require an ensemble of stars.  Fortunately,
+recent analyses of the kinematic, photometric, and astrometric Gaia
+data have significantly expanded our knowledge of cluster memberships
+\citep[{\it
 e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,Kerr2021}.
@@
 Kepler\,1627B, the planet orbits the G-dwarf primary on an orbit that
-is likely aligned with the stellar rotation axis
+is likely prograde
 (Section~\ref{sec:planet}).  We conclude by highlighting broader
@@
 \citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
-suggestion of an aligned orbit is consistent with a quiescent
+suggestion of a prograde is consistent with a quiescent
 disk-driven migration history, and differs from misalignments that
@@
 were those of \citet{CottenSong2016}, \citet{Oh2017},
-\citet{Zari2018}, \citet{Gagne2018b},
-\citet{Gagne2018a}, \citet{Gagne2018c}, and \citet{Ujjwal2020}.  While
-in principle the moving group members discussed by
+\citet{Zari2018}, \citet{Gagne2018b}, \citet{Gagne2018a},
+\citet{Gagne2018c}, and \citet{Ujjwal2020}.  While in principle the
+moving group members discussed by
 \citet{Gagne2018b,Gagne2018a,Gagne2018c} and \citet{Ujjwal2020} have
-easily associated ages, our SIMBAD cross-match did not retain the moving group
-identifiers given by those studies, which should therefore be recovered
-tools such as BANYAN
+easily associated ages, our SIMBAD cross-match did not retain the
+moving group identifiers given by those studies, which should
+therefore be recovered using tools such as BANYAN
 $\Sigma$.\footnote{\url{http://www.exoplanetes.umontreal.ca/banyan/banyansigma.php}}.
@@
 Finally, we  included every star in the NASA Exoplanet Archive
-planetary system ($\texttt{ps}$) table that had a Gaia identifier available
-\citep{NASAExoArchive_ps_20210506}.  If the age had finite
+planetary system ($\texttt{ps}$) table that had a Gaia identifier
+available \citep{NASAExoArchive_ps_20210506}.  If the age had finite
 uncertainties, we also included it, since stellar ages determined
@@
 smattering of reported cluster members (gray and black points) has a
-significantly different structure that the cluster initially
+significantly different structure than the cluster initially
 identified by \citet{stephenson_possible_1959} and corroborated by
@@
 are as follows.
-% from logs/rm_freejitter.log
-% \begin{align}
-%   \chi^2_{\rm red} &= 1.45 \quad ({\rm Quadratic+RM}) \nonumber \\
-%   \chi^2_{\rm red} &= 1.37 \quad ({\rm Linear+RM}) \nonumber \\
-%   \chi^2_{\rm red} &= 1.22 \quad ({\rm Only\ Quadratic}),
-% \end{align}
-% and
+% from logs/rm_freejitter_FINAL.log
 \begin{align}
@@
   {\rm BIC} &= 257.7  \quad ({\rm Linear+RM}) \nonumber \\
-  {\rm BIC} &= 242.8  \quad ({\rm Only\ Quadratic}).
+  {\rm BIC} &= 247.1  \quad ({\rm Only\ Quadratic}).
 \end{align}

### 2021-09-22 paper/ms.tex
update ms

@@
   exoplanet evolution (491),
-  stellar associations (1582),
+  %stellar associations (1582),
   open star clusters (1160),
-	stellar ages (1581),
+	stellar ages (1581)
 }
@@
 during the main Kepler mission (2009--2013), only four open clusters
-were known in the Kepler field: NGC\,6866, NGC\,6811, NGC\,6819, and
-NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
-\citep{meibom_kepler_2011}.  Though isochronal and gyrochronal
+were known in the Kepler field,
+%: NGC\,6866, NGC\,6811, NGC\,6819, and NGC\,6791,
+with ages spanning 0.7\,Gyr to 9\,Gyr
+\citep{meibom_kepler_2011}.  Though isochronal, gyrochronal, and lithium-based
 analyses suggest that younger Kepler planets do exist
@@
 recent analyses of the kinematic, photometric, and astrometric Gaia
-data have significantly expanded our knowledge of cluster memberships
+data have since expanded our knowledge of cluster memberships
 \citep[{\it
@@
 age-dated stars (see Appendix~\ref{app:targetlist}).
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.5\textwidth]{f2.pdf}
+	\end{center}
+	\vspace{-0.6cm}
+	\caption{
+		{\bf Kepler and TESS views of the $\delta$~Lyr cluster.} Colored
+		points are kinematically selected candidate $\delta$\,Lyr cluster
+		members.  Both Kepler (gray panels) and TESS observed portions of
+		the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray;
+		Deneb, Vega, and $\delta$~Lyr are orange.  Kepler\,1627 (green star) was observed during the
+		entirety of the Kepler mission.  Colors of candidate cluster
+		members indicate the duration of TESS observations to date, in
+		units of lunar months.
+		\label{fig:skychart}
+	}
+\end{figure}
 
@@
 \label{sec:cluster}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f2.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
-    points are kinematically selected candidate $\delta$\,Lyr cluster
-    members.  Both Kepler (gray panels) and TESS observed portions of
-    the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray
-    or orange.  Kepler\,1627 (green star) was observed during the
-    entirety of the Kepler mission.  Colors of candidate cluster
-    members indicate the duration of TESS observations to date, in
-    units of lunar months.
-    \label{fig:skychart}
-	}
-\end{figure}
 
@@
 		\subfloat{
-			\includegraphics[width=\textwidth]{f5.pdf}
+			\includegraphics[width=0.5\textwidth]{f5a.pdf}
+			\includegraphics[width=0.5\textwidth]{f5b.pdf}
 		}
@@
     {\it Top}: 
-    The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
-    30 minute cadence.  A one hundred day segment is shown here.  The
+    The Kepler data span 1{,}437 days (3.9 years), sampled at
+    30 minute cadence;  a 100 day segment is shown.  The
     top panel shows the \texttt{PDCSAP} median-subtracted flux in
     units of parts-per-thousand ($\times 10^{-3}$).  The dominant
-    signal is induced by starspots.  The model for the stellar
-    variability (orange line) is subtracted below, revealing the
+    signal is induced by starspots.  The stellar
+    variability model (orange line) is subtracted below, revealing the
     transits of \pn, as well as other deviations from the model.  The
-    online Figure Set shows the entire 3.9 years of observations.
+    online Figure Set spans the entire 3.9 years of observations.
     {\it Bottom}:
-    Phase-folded transit of Kepler 1627Ab, with stellar variability
-    removed.  The 2-$\sigma$ model uncertainties and the maximum {\it
-    a posteriori} model are shown as the faint purple band, and the
-    dark purple line.  Gray points are individual \texttt{PDCSAP} flux
-    measurements; black points are binned to 20 minutes, with a
-    representative error bar shown.  The asymmetric residual visible
-    during the transit is robust against detrending methods; we
-    believe that it is caused by starspot crossings.
+    Phase-folded transit of Kepler 1627Ab with stellar variability
+    removed.  Windows over 20 hours ({\it left}) and the entire orbit
+    ({\it right}) are shown.  The 2-$\sigma$ uncertainties and the
+    maximum best-fit model are the faint purple band, and the dark
+    purple line.  Gray points are individual flux measurements; black
+    points bin these every 20 minutes, and have a representative error bar
+    shown.  The asymmetric residual during transit appears to be robust when using
+    different
+    detrending methods.
     \label{fig:lc}
@@
     {\bf Radii, orbital periods, and ages of transiting exoplanets}.
-    Planets older than $10^9$ years are shown with gray markers.
+    %Planets older than $10^9$ years are shown with gray markers.
     Planets younger than a gigayear with ${\rm \tau}/\sigma_{\rm \tau}
@@
 ($9\pm1\,{\rm Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and
-V1298 Tau ($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  A
-portion of the light curve is shown with the phase-folded transit in
-Figure~\ref{fig:lc}.  The \texttt{PDCSAP} photometry is dominated by a
+V1298 Tau ($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  The light
+curve and the phase-folded transit are shown in
+Figure~\ref{fig:lc}.  The photometry is dominated by a
 quasi-periodic starspot signal with a peak-to-peak amplitude that
-varies between 2\% and 8\%, depending on the degree of asymmetry in
-spot coverage between the stellar hemispheres.  The starspot signal
-repeats every $2.642\pm0.042$\,days, where the uncertainty is based on
-the scatter observed between different Kepler quarters.  The details
-of how we model the starspot signal to derive the parameters of the
-planet are discussed in Appendix~\ref{app:gptransit}.  The stellar
-parameters are listed in Table~\ref{tab:starparams}; the stellar mass,
+varies between 2\% and 8\%.  The signal
+repeats every $2.642\pm0.042$\,days, where the quoted uncertainty is based on
+the scatter in rotation periods measured from each individual Kepler quarter. 
+Our model for the starspot signal is discussed in detail
+in Appendix~\ref{app:gptransit}, and the adopted stellar
+parameters are listed in Table~\ref{tab:starparams}.  The stellar mass,
 radius, and effective temperature are found by interpolating against a
-38$\,$Myr MIST isochrone.  The systematic uncertainties are taken to
-be the difference in parameters from the PARSEC isochrones, and the
+38$\,$Myr MIST isochrone.  The systematic uncertainties are taken
+to be the difference between the PARSEC and MIST isochrones, and the
 reported uncertainties are a quadrature sum of the statistical and
-systematic components.  The transit-fitting results, given in
+systematic components. 
+The transit fitting results, given in
 Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
@@
 
-Could the transit signal be produced by anything other than a planet
+Could this transit signal be produced by anything other than a planet
 orbiting this near-solar analog?  \citet{morton_false_2016} validated
@@
 accompanied by a low-mass stellar companion ($M_{\rm B}\approx
-0.33\,M_\odot$).  We therefore reconsidered this false positive
-scenario in some detail.
+0.33\,M_\odot$).  We therefore reassessed false positive
+scenarios in some detail.
 
@@
 
-The expectation however is for the spot-induced signals to be smooth,
+The expectation however is for the starspot-induced signals to be smooth,
 at worst with contributions at $0.5\,P_{\rm rot}$ or $0.25\,P_{\rm
@@
 Rossiter-McLaughlin (RM) effect, and allowed for an optional linear
-and quadratic trend in time to fit the spot-induced signal.  We
+and quadratic trend in time to fit the 800\ms\ spot-induced trend.  We
 followed the methodology developed by \citet{stefansson_2020}.  We
@@
 The presence of this additional scatter prevents a clear detection of
-the RM effect.  The reason can be understood through a
-model-comparison analysis.  If we compare the model with a quadratic trend and
+the RM effect.  The reason can be understood via
+model comparison.  If we compare the model with a quadratic trend and
 the RM effect (Figure~\ref{fig:obliquity}) against a model with a
@@
 The ``only quadratic'' model does particularly well because it can
-inflate the jitter term to account for any additional scatter during
-the transit, and it has fewer free parameters.  However, we cannot
+inflate the jitter term to account for scatter during
+the transit (even if the scatter contains astrophysics!), and it has fewer free parameters.  However, we cannot
 justify a physical prior on the jitter term, because we do not
@@
 is only 0.4\% of the stellar rotation period; it is not obvious that
-starspots could be the culprit.
+starspots would be the culprit.
 
@@
 uncertainties; the S/N per resolution element for each of our spectra
-(70 to 80) was moderated by the observing requirements, and the rapid
-rotation of the star could affect the velocity extraction.
-Observations at higher precision are necessary to distinguish these
-two possibilities, and remain worthwhile as a means toward clarifying
-the orbital geometry of \pn.
+was moderated by the observing requirements to 70 to 80, and the rapid
+rotation of the star could affect accuracy of the uncertainties from
+the velocity extraction.  Observations at higher precision are
+necessary to distinguish these two possibilities, and remain
+worthwhile as a means toward clarifying the orbital geometry of \pn.
 

### 2021-09-22 drivers/plot_RM.py

@@ -6,10 +6,10 @@ PLOTDIR = os.path.join(RESULTSDIR, 'RM')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
-# models = ['quadraticfreejitter', 'linearfreejitter', 'trendonlyfreejitter',
-#           'quadraticretrogradefreejitter','quadraticprogradefreejitter']
-models = ['trendonlyfreejitter',
+models = ['quadraticfreejitter', 'linearfreejitter', 'trendonlyfreejitter',
           'quadraticretrogradefreejitter','quadraticprogradefreejitter']
+#models = ['trendonlyfreejitter',
+#          'quadraticretrogradefreejitter','quadraticprogradefreejitter']
 
 
 #models = ['quadraticfreejitter', 'quadratic']

### 2021-09-22 paper/collect.sh

@@ -6,7 +6,8 @@ cp /Users/luke/Dropbox/proj/rudolf/results/keplerlc/keplerlc_Nsample500_xlim0500
 
 cp /Users/luke/Dropbox/proj/rudolf/results/keplerlc/*.pdf FigureSetKeplerLC/.
 
-cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_posterior_phaseplot.pdf f5.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_posterior_phaseplot.pdf f5a.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_posterior_phaseplot_fullxlim.pdf f5b.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_cornerplot.png dontinclude0.png
 

### 2021-09-22 paper/ms.tex
update ms

@@
   events in the Kepler photometry and a Keck/HIRES transit we find
-  that the planet is likely on a prograde orbit.  We also briefly
-  explore the flare arrival time distribution in the short cadence
-  Kepler photometry.  The $\delta$~Lyr cluster is one of many stellar
-  groups whose properties have been clarified using Gaia; many other
-  exoplanet hosts belong to these new clusters.  Verifying the results
-  of Gaia clustering analyses using TESS rotation periods offers a
-  promising path toward expanding the census of age-dated planets, and
-  our corresponding understanding of planetary evolution.
+  that the planet is likely on a prograde orbit.  The average Kepler
+  transit profile also appears asymmetric; our current best
+  explanation is that of a dusty atmospheric outflow being present,
+  but this remains to be confirmed.  Broadly speaking, the
+  $\delta$~Lyr cluster is one of many stellar groups whose properties
+  have been clarified using Gaia; many other exoplanet hosts belong to
+  these new clusters.  Verifying the results of Gaia clustering
+  analyses using TESS rotation periods offers a promising path toward
+  expanding the census of age-dated planets, and our corresponding
+  understanding of planetary evolution.
 \end{abstract}
@@
 age measurements typically require an ensemble of stars.  Fortunately,
-recent analyses of the kinematic, photometric, and astrometric Gaia
+recent analyses of the Gaia
 data have since expanded our knowledge of cluster memberships
@@
 
-Comparing our young star list against the Kepler field yielded two
-discoveries.  The first, to be discussed in an upcoming analysis by
-J.~Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
-cluster Theia~520 (UBC~1).  At least six Kepler planets are known in
-the cluster from the Kepler-52 and Kepler-968 systems
+Matching our young star list against stars observed by Kepler yielded
+two discoveries.  The first, to be discussed in an upcoming analysis
+by J.~Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
+open cluster Theia~520 (UBC~1).  At least six Kepler planets are known
+in the cluster from the Kepler-52 and Kepler-968 systems
 \citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
@@
 ({\it e.g.}, \citealt{randich_gaiaeso_2018};
-\citealt{bouma_2021_ngc2516}).
+\citealt{bouma_cluster_2020}).
 
@@
     variability model (orange line) is subtracted below, revealing the
-    transits of \pn, as well as other deviations from the model.  The
+    transits of \pn.  The
     online Figure Set spans the entire 3.9 years of observations.
@@
     removed.  Windows over 20 hours ({\it left}) and the entire orbit
-    ({\it right}) are shown.  The 2-$\sigma$ uncertainties and the
-    maximum best-fit model are the faint purple band, and the dark
-    purple line.  Gray points are individual flux measurements; black
-    points bin these every 20 minutes, and have a representative error bar
-    shown.  The asymmetric residual during transit appears to be robust when using
-    different
-    detrending methods.
+    ({\it right}) are shown.  The 2-$\sigma$ model uncertainties and
+    the best-fit model are the light purple band and the dark purple
+    line.  Gray points are individual flux measurements; black points
+    bin these to 20 minute intervals, and have a representative
+    1-$\sigma$ error bar in the lower right of each panel.  The
+    asymmetric residual during transit appears to be robust to
+    different detrending methods.
     \label{fig:lc}
@@
 
+		\vspace{-0.1cm}
 		\subfloat{
@@
     easily understood by considering that the stellar flux changes by
-    $\sim$60\,ppt per half rotation period
-    ($\sim$1.3\,days).
-    {\it Bottom:} Radial velocities acquired using Keck/HIRES during
-    the transit of 2021 Aug 7, observed simultaneously with MuSCAT3.
-    The best-fit sky-projected obliquity for this model is listed in
-    red.  While the intra-exposure RV scatter precludes a detection of
-    the RM effect, the photometric transit depths are consistent
-    across the {\it griz} bandpasses.
+    $\sim$60\,ppt per half rotation period ($\sim$1.3\,days).
+    {\it Bottom:}
+    Radial velocities acquired using Keck/HIRES during the transit of
+    2021 Aug 7, observed simultaneously with MuSCAT3.  Shaded bands
+    show 2-$\sigma$ model uncertainties.  The photometric transit
+    depths are consistent across the {\it griz} bandpasses.  The
+    best-fitting model including the RM effect and a quadratic trend
+    in time is shown with the red line.  The radial velocity scatter
+    between exposures prevents a significant detection of the RM
+    effect.
     \label{fig:obliquity}
@@
     %Planets older than $10^9$ years are shown with gray markers.
-    Planets younger than a gigayear with ${\rm \tau}/\sigma_{\rm \tau}
-    > 3$ are emphasized, for $\tau$ the age.  \pn\ is shown with a
-    star.  The large sizes of the youngest transiting planets could be
-    explained by their primordial atmospheres not yet having
-    evaporated; direct measurements of the atmospheric outflows or
-    planetary masses would help to confirm this expectation.
-    Selection effects may also be important.  Parameters are from the
-    NASA Exoplanet Archive (2021 Sept 15).
+    Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
+    3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
+    its uncertainty.  \pn\ is shown with a star.  The large sizes of
+    the youngest transiting planets could be explained by their
+    primordial atmospheres not yet having evaporated; direct
+    measurements of the atmospheric outflows or planetary masses would
+    help to confirm this expectation.  Selection effects may also be
+    important.  Parameters are from the NASA Exoplanet Archive (2021
+    Sept 15).
 		\label{fig:rp_period_age}
@@
 radius, and effective temperature are found by interpolating against a
-38$\,$Myr MIST isochrone.  The systematic uncertainties are taken
+38$\,$Myr MIST isochrone.  Their systematic uncertainties are taken
 to be the difference between the PARSEC and MIST isochrones, and the
@@
 The transit fitting results, given in
-Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
+Table~\ref{tab:posterior}, are broadly consistent with a mini-Neptune sized
 planet ($3.34\pm 0.18\,R_\oplus$) on a close-in circular\footnote{Our
@@
 against the local linear slope.  A significant correlation of $-0.0564
-\pm 0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Less than ten Kepler
+\pm 0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
 Objects of Interest have shown this correlation
@@
 confirms that \pn\ is a planet.  It also implies that the planet's
-orbit is prograde.  The latter interpretation assumes that the
-dominant photometric variability is induced by dark spots, rather than
-bright faculae.  Given the observed transition of Sun-like stellar
+orbit is prograde.  The latter point assumes that the dominant
+photometric variability is induced by dark spots, and not bright
+faculae.  Given the observed transition of Sun-like stellar
 variability from spot to faculae-dominated regimes, we expect this
-inference to be reasonably secure
+assumption to be reasonably secure
 \citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
@@
 Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
-show the transit, and it appears to be achromatic
-(Figure~\ref{fig:obliquity}, lower panels).  Fitting the MuSCAT3
-photometry with a model that lets the transit depths vary across
-bandpasses, we find {\it griz} depths consistent with the Kepler depth
-at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  The MuSCAT3
-observations also suggest a transit duration $17.4\pm3.6\,{\rm
+show the transit (Figure~\ref{fig:obliquity}, lower panels).  Fitting
+the MuSCAT3 photometry with a model that lets the transit depths vary
+across each bandpass, we find {\it griz} depths consistent with the
+Kepler depth at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  The
+MuSCAT3 observations also suggest a transit duration $17.4\pm3.6\ {\rm
 minutes}$ shorter than the Kepler transits; further photometric
-follow-up could help confirm whether such changes are indeed
-happening.  For the RM analysis, the details are discussed in
-Appendix~\ref{app:rm}.  While our RM data are marginally more
-consistent with a prograde or polar orbit than a retrograde orbit,
-the expected velocity anomaly ($\Delta v_{\rm RM}\approx
-20\,{\rm m\,s}^{-1}$) is a factor of two smaller than the
-spot-corrected exposure-to-exposure scatter.  We are therefore not in
-a position to claim a spectroscopic detection of the RM effect, nor a
-precise quantification of the stellar obliquity.
+follow-up could help confirm whether the transit duration is indeed
+changing.  For our RM analysis, the details are discussed in
+Appendix~\ref{app:rm}.  While the velocities are marginally more
+consistent with a prograde or polar orbit than a retrograde orbit, the
+spot-corrected exposure-to-exposure scatter ($\sigma_{\rm RV}\approx
+50$\ms) is significantly larger than the expected velocity anomaly
+($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$).  We are therefore
+not in a position to claim a spectroscopic detection of the RM effect,
+nor a precise quantification of the stellar obliquity.
 
@@
 planets, and opens multiple avenues for further study.  Further
-observations of spectroscopic transits could yield a more precise
-stellar obliquity, and could confirm or refute the expected prograde
-orbital geometry.  Similarly, transit spectroscopy aimed at detecting
-outflows of hydrogen or helium could yield insight into the
-evolutionary state of the atmosphere
-\citep{spake_helium_2018,vissapragada_2020}.  Determining the precise
-amount of high-energy irradiation incident on the planet would also
-help in clarifying the expected outflow rate \citep[{\it
-e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
-quantity to measure would be the planet's mass.  This might be
-possible through long-term monitoring of the transit times and
-durations, or through a high-cadence multi-color radial velocity
-campaign.  Measured at sufficient precision, the mass could yield
-constraints on both the planet's composition and its initial entropy
+observations of spectroscopic transits could yield a quantification of
+the stellar obliquity, and would confirm or refute the expected
+prograde orbital geometry.  Similarly, transit spectroscopy aimed at
+detecting atmospheric outflows could yield insight into the
+evolutionary state of the atmosphere \citep[{\it
+e.g.},][]{ehrenreich_giant_2015,spake_helium_2018,vissapragada_2020}.
+Observations aimed at quantifying the amount of high-energy
+irradiation incident on the planet would also help in clarifying the
+expected outflow rate \citep[{\it e.g.},][]{poppenhaeger_2021}.
+Finally, a challenging but informative quantity to measure would be
+the planet's mass.  This might be possible through long-term
+monitoring of the transit times and durations, or through a
+high-cadence multi-color radial velocity campaign.  Measured at
+sufficient precision, the mass could yield constraints on both the
+planet's composition and its initial entropy
 \citep{owen_constraining_2020}.
 
-Separately, the Kepler data may yet contain further information.
-Although our transit timing study did not yield the detection of any
-planet-induced transit timing or duration variations,
-Appendix~\ref{app:flare} highlights an odd flare timing coincidence
-that may be suggestive of star-planet interactions.  We encourage its
-exploration by investigators more versed in the topic than ourselves.
+Separately, the Kepler data may yet contain additional information.
+For instance, one possible explanation for the transit asymmetry shown
+in Figure~\ref{fig:lc} is that of a dusty asymmetric outflow.  Other
+possible explanations such as favored longitudes for starspot
+crossings seem strained, for reasons discussed in
+Appendix~\ref{app:gptransit}.  Dusty outflows are theoretically
+expected for young mini-Neptunes, and the amplitude of the observed
+asymmetry is also roughly consistent with predictions
+\citep{wang_dai_2019}.  Beyond the planetary transits,
+Appendix~\ref{app:flare} highlights a separate odd coincidence in the
+Kepler data, in the arrival time distribution of stellar flares.  We
+encourage its exploration by investigators more versed in the topic
+than ourselves.
 
@@
 youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
-with precise ages below 100\,Myr include K2-33, DS~Tuc, HIP~67522,
+with precise ages include K2-33, DS~Tuc, HIP~67522,
 TOI~837, the two-planet AU~Mic system, and the four-planet V1298~Tau
@@
 \citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
-suggestion of a prograde is consistent with a quiescent
+suggestion of a prograde orbit is consistent with a quiescent
 disk-driven migration history, and differs from misalignments that
@@
 
-Ultimately, the main advance of this work is a precise measurement of
-the age of \pn.   This age measurement was enabled by identifying the
-connection of the star to the \cn\ using the Gaia data.
+Ultimately, the main advance from this work is a precise measurement of
+the age of \pn.   This measurement was enabled by identifying the
+connection of the star to the \cn\ using Gaia kinematics, and then by
+using the Gaia HR diagram and TESS stellar rotation periods to verify the
+cluster's existence.
 Table~\ref{tab:v06} should enable similar cross-matches to already
-known planets, as well as forthcoming planets from the TESS mission
-\citep{ricker_transiting_2015,guerrero_tess_2021}.  An alternative
+known planets, as well as forthcoming planets from TESS
+\citep{guerrero_tess_2021}.  An alternative
 approach for age-dating more planets could be to identify kinematic
@@
 
-The authors are grateful to T{.}~David, A{.}~Howard, J{.}~Spake, and
-J{.}~Winn for illuminating discussions and suggestions, and to
+The authors are grateful to J{.}~Winn, J{.}~Spake, A{.}~Howard, and
+T{.}~David for illuminating discussions and suggestions, and to
 R{.}~Kerr for kindly providing us with the \citet{Kerr2021} membership
@@
 
-Our model therefore included 10 free parameters for the
-transit ($\{P, t_0, R_{\rm p}/R_\star, b, u_1 ,u_2 ,R_\star, \log g,
-e, \omega \}$), 2 free parameters for the light curve
-normalization and uncertainty calibration ($\{\langle f \rangle, 
-\sigma_f \}$), and 5 hyperparameters for the GP
-($\{\sigma_{\mathrm{rot}}, P_{\mathrm{rot}}, Q_0,
-\mathrm{d}Q, f \}$).  We also considered including an additive
-\texttt{SHOTerm} kernel to account for stochastic noise, but found
-that this did not significantly affect the results, and so opted for
-the simpler GP kernel.  We fitted the models using \texttt{PyMC3}
-\citep{salvatier_2016_PyMC3,exoplanet:theano}.  
-We assumed a Gaussian likelihood, and
-after initializing
-each model with the parameters of the maximum {\it a posteriori}
-model, we sampled using
-\texttt{PyMC3}'s gradient-based No-U-Turn Sampler
-\citep{hoffman_no-u-turn_2014} in these bases indicated in Table~\ref{tab:posterior}.
-We used $\hat{R}$ as our convergence
+Our model therefore included 10 free parameters for the transit ($\{P,
+t_0, R_{\rm p}/R_\star, b, u_1 ,u_2 ,R_\star, \log g, e, \omega \}$),
+2 free parameters for the light curve normalization and uncertainty
+calibration ($\{\langle f \rangle, \sigma_f \}$), and 5
+hyperparameters for the GP ($\{\sigma_{\mathrm{rot}},
+P_{\mathrm{rot}}, Q_0, \mathrm{d}Q, f \}$).  We also considered
+including an additive \texttt{SHOTerm} kernel to account for
+stochastic noise, but found that this did not significantly affect the
+results, and so opted for the simpler GP kernel.  We fitted the models
+using \texttt{PyMC3} \citep{salvatier_2016_PyMC3,exoplanet:theano}.
+We assumed a Gaussian likelihood, and after initializing each model
+with the parameters of the maximum {\it a posteriori} model, we
+sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
+\citep{hoffman_no-u-turn_2014} in these bases indicated in
+Table~\ref{tab:posterior}.  We used $\hat{R}$ as our convergence
 diagnostic \citep{gelman_inference_1992}.
@@
 high during the first half, and low in the second half.  To determine
-whether this effect was not caused by our stellar variability model, we 
+whether this effect was caused by our stellar variability model, we 
 explored an alternative approach in which we isolated each transit
@@
 
-Figures~\ref{fig:phasequarter} and~\ref{fig:phaseyear} show the
-Kepler data binned over quarters and years.  In
+Figures~\ref{fig:phasequarter} and~\ref{fig:phaseyear} show the Kepler
+data binned over quarters and years.  In
 Figure~\ref{fig:phasequarter}, Quarter 6 shows the strongest asymmetry
-out of the quarters: a deviation of about 3\,ppt from
-expectation.  Quarter 7 shows an anomaly at roughly the same transit
-phase.  Year 2 correspondingly shows the strongest anomaly out of any
-year in Figure~\ref{fig:phaseyear}.
-
-We considered three possible explanations for the anomaly: gravity
-darkening, transit timing variations, and spot-crossing events.
+out of the quarters: a deviation of about 3\,ppt from expectation.
+Quarter 7 shows an anomaly at roughly the same transit phase.  Year 2
+correspondingly shows the strongest anomaly out of any year in
+Figure~\ref{fig:phaseyear}; the asymmetry is visually apparent however
+in each of Years 2, 3, and 4.
+
+We therefore considered four possible explanations for the anomaly:
+gravity darkening, transit timing variations, spot-crossing events,
+and a persistent asymmetric dusty outflow.
+
 Gravity darkening is based on the premise that the rapidly rotating
-star becomes oblate, and brighter near the poles than the equator
-\citep[{\it e.g.},][]{masuda_spin-orbit_2015}.  The fractional shape
-change due to gravity darkening is on the order of $(P_{\rm
+star is oblate, and brighter near the poles than the equator
+\citep[{\it e.g.},][]{masuda_spin-orbit_2015}.  The fractional transit
+shape change due to gravity darkening is on the order of $(P_{\rm
 break}/P_{\rm rot})^2$, for $P_{\rm break}$ the break-up rotation
 period, and $P_{\rm rot}$ the rotation period.  Using the parameters
-from Table~\ref{tab:posterior}, this yields an expected 1.4\%
+from Table~\ref{tab:posterior}, this yields an expected 0.14\%
 distortion of the $\approx$1.8\,ppt transit depth: {\it i.e.}, an
 absolute deviation of $\approx$2.5\,ppm.  The observed residual has a
-semi-amplitude of $\approx 50\,{\rm ppm}$.  Since the expected signal is
-smaller than the observed anomaly by over an order of magnitude, gravity
-darkening seems to be an unlikely explanation.
+semi-amplitude of $\approx 50\,{\rm ppm}$.  Since the expected signal
+is smaller than the observed anomaly by over an order of magnitude,
+gravity darkening seems to be an unlikely explanation.
 
 The scenario of transit timing variations (TTVs) producing the
-asymmetry seems unlikely, since the analysis by
-\citet{holczer_transit_2016} implies that any such variations in \sn\
-need to be less than of order minutes.  
-Our independent TTV analysis found similar results.
-Figures~\ref{fig:phasequarter}
-and~\ref{fig:phaseyear} also provide little evidence in support of
-this possibility.
-
-The final possibility is that of starspot crossings.  Young stars are
-thought to have higher spot-covering fractions than old stars ({\it
-e.g.}, \citealt{morris_relationship_2020}).  Young solar-type stars may also host
-dark starspots at high stellar latitudes \citep[{\it e.g.},
-EK~Dra;][]{strassmeier_starspots_2009}, though interferometric imaging
-of spotted giant stars has found different starspot latitude distributions than
-those inferred from Doppler imaging
+asymmetry seems unlikely because our analysis showed that transit
+timing variations in \sn\ have amplitudes smaller than $\approx$10
+minutes (see Figure~\ref{fig:obliquity}).  This is consistent with
+previous results from \citet{holczer_transit_2016}.  Moreover, the
+transit timing variations we do observe are correlated with the local
+light curve slope, which on average increases as much as it decreases.
+From our analysis, the mean TTV and its standard deviation are
+$0.00\pm12.05$ minutes; similarly the mean local slope and its
+standard deviation are $0.49\pm44.49$\,ppt\,day$^{-1}$.  There is
+therefore no expectation for TTVs to produce the asymmetry.
+
+The third possibility is that of starspot crossings.  Young stars 
+have higher spot-covering fractions than old stars ({\it
+e.g.}, \citealt{morris_relationship_2020}).  Young solar-type stars
+may also host dark starspots at high stellar latitudes \citep[{\it
+e.g.}, EK~Dra;][]{strassmeier_starspots_2009}, though interferometric
+imaging of spotted giant stars has shown different starspot latitude
+distributions than those inferred from Doppler imaging
 \citep{roettenbacher_contemporaneous_2017}.  Regardless, for any
@@
 transits, it seems likely that we would need either for spots to be
-persistent at a particular latitude (and for a slightly misaligned
-orbit), or for a ``stroboscopic'' longitudinal phasing \citep[{\it
-e.g.},][]{dai_stellar_2018}.  For our system, $P_{\rm orb}/P_{\rm rot}
-\approx 2.76$, which means that every 4 transits and 11 stellar
-rotations, the planet crosses over roughly the same stellar longitude,
-which could enable the necessary phasing if the spot-groups are
-large and long-lived.  Unfortunately, the  S/N per Kepler transit is
-$\approx8$, which renders individual spot-crossing events unresolved.
-Therefore, although starspot crossings seem to be
-the most likely source of additional photometric scatter
-during transits, our reasoning is that the expected
-spot-crossing anomaly amplitudes ($\approx 100\,{\rm ppm}$) could be
-correct; we have not been able to definitely show that this is the case, and 
-the issue could deserve further exploration.
+persistent at a particular latitude (and for the planetary orbit to be
+quite misaligned), or for a ``stroboscopic'' longitudinal phasing
+\citep[{\it e.g.},][]{dai_stellar_2018}.  For our system, $P_{\rm
+orb}/P_{\rm rot} \approx 2.76$, which means that every 4 transits and
+11 stellar rotations, the planet crosses over roughly the same stellar
+longitude, which might enable the necessary phasing if the spot-groups
+are large and long-lived.  Unfortunately, the  S/N per Kepler transit
+is $\approx8$, which renders individual spot-crossing events
+unresolved.  This explanation seems marginally plausible, because the
+expected spot-crossing anomaly amplitudes ($\approx 100\,{\rm ppm}$)
+resemble the observed amplitude of the asymmetry ($\approx 50\,{\rm
+ppm}$).  We ultimately disfavor it though, because {\it a)} it seems
+implausible that the starspot crossings could add coherently, and {\it
+b)} there is no reason to expect starspot crossing events to last
+exactly half the transit
+duration.
+
+A persistent feature of the planet itself might therefore be needed to
+explain the transit asymmetry.  An asymmetric outflow from the
+planet's atmosphere could at least geometrically meet the requirements
+\citep[{\it e.g.},][]{mccann_2019}.  To explain the asymmetric
+transit, a small, dense component would lead the planet, and a long,
+more rarefied (and variable) component would trail it.  This might
+also explain the slight flux decrement visible for $\sim$1 hour
+pre-ingress (Figure~\ref{fig:lc}).  The amplitude of the asymmetry is
+roughly in line with theoretical expectations for dusty outflows
+\citep{wang_dai_2019}, and based on the planet's size, its mass is
+likely in a regime where such outflows are possible.  Out of the four
+explanations discussed, this one currently seems the most plausible.
+By composition, the expectation would be that the envelope is mostly
+hydrogen and helium gas, with a dust or haze component providing the
+broadband opacity in the Kepler bandpass.  A natural path
+toward confirming or refuting this idea would be to observe additional
+transits of the planet in hydrogen absorption, 
+metastable helium absorption, or across a broad wavelength range in
+the near-infrared.
 

### 2021-09-22 rudolf/plotting.py

@@ -1308,6 +1308,9 @@ def plot_ttv_vs_local_slope(outdir):
 
     ttv_mean, ttv_std = np.mean(ttv), np.std(ttv)
 
+    import IPython; IPython.embed()
+    assert 0
+
     ttv_cut = 0.02 # doesn't mean match
 
     z1 = np.poly1d(
@@ -2615,15 +2618,15 @@ def plot_RM_and_phot(outdir, model=None, showmodelbands=0, showmodel=0):
         mmodel1.append(m1)
     mmodel1 = np.array(mmodel1)
 
-    ax.fill_between(scale_x(t1_mod), np.quantile(mmodel1,0.16,axis=0),
-                    np.quantile(mmodel1,0.84,axis=0), alpha=0.1,
-                    color=shadecolor, lw=0, label='1$\sigma$',zorder=-1)
+    #ax.fill_between(scale_x(t1_mod), np.quantile(mmodel1,0.16,axis=0),
+    #                np.quantile(mmodel1,0.84,axis=0), alpha=0.1,
+    #                color=shadecolor, lw=0, label='1$\sigma$',zorder=-1)
     ax.fill_between(scale_x(t1_mod), np.quantile(mmodel1,0.02,axis=0),
                     np.quantile(mmodel1,0.98,axis=0), alpha=0.1,
                     color=shadecolor, lw=0, label='2$\sigma$', zorder=-1)
-    ax.fill_between(scale_x(t1_mod), np.quantile(mmodel1,0.0015,axis=0),
-                    np.quantile(mmodel1,0.9985,axis=0), alpha=0.1,
-                    color=shadecolor, lw=0, label='3$\sigma$', zorder=-1)
+    #ax.fill_between(scale_x(t1_mod), np.quantile(mmodel1,0.0015,axis=0),
+    #                np.quantile(mmodel1,0.9985,axis=0), alpha=0.1,
+    #                color=shadecolor, lw=0, label='3$\sigma$', zorder=-1)
 
     sdf = df_medvals[df_medvals.Labels == 'lam_p1']
     #from rudolf.helpers import ORIENTATIONTRUTHDICT
@@ -2663,7 +2666,7 @@ def plot_RM_and_phot(outdir, model=None, showmodelbands=0, showmodel=0):
     colors = 'blue,green,orange,red'.split(',')
 
     shift = 0
-    delta = 0.007
+    delta = 0.0055
[…]

### 2021-09-23 paper/ms.tex
update ms
Co-authored-by: adamkraus <alk@astro.as.utexas.edu>

@@
 2--4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
-is expected to occur within the first few million years of disk
+is expected to occur within the first few million years after disk
 dispersal \citep{owen_atmospheres_2016}, the combination of stellar
@@
 Given the effective temperature of \sn\ (Table~\ref{tab:starparams}),
-this measurement is also in line with expectations for a 38\,Myr star
-({\it e.g.}, \citealt{randich_gaiaeso_2018};
-\citealt{bouma_cluster_2020}).
+this measurement is in agreement with expectations for a 38\,Myr star
+({\it e.g.}, as measured in IC~2602 by
+\citealt{randich_gaiaeso_2018}).  It is also larger than any ${\rm
+EW}_{\rm Li}$ found by \citet{berger_identifying_2018} in their
+analysis of 1{,}301 Kepler spectra.
 
@@
     removed.  Windows over 20 hours ({\it left}) and the entire orbit
-    ({\it right}) are shown.  The 2-$\sigma$ model uncertainties and
-    the best-fit model are the light purple band and the dark purple
-    line.  Gray points are individual flux measurements; black points
-    bin these to 20 minute intervals, and have a representative
+    ({\it right}) are shown, and the residual after subtracting the
+    transit in the bottom-most row.  The 2-$\sigma$ model uncertainties
+    and the best-fit model are the light purple band and the dark
+    purple line.  Gray points are individual flux measurements; black
+    points bin these to 20 minute intervals, and have a representative
     1-$\sigma$ error bar in the lower right of each panel.  The
@@
 faculae.  Given the observed transition of Sun-like stellar
-variability from spot to faculae-dominated regimes, we expect this
-assumption to be reasonably secure
+variability from spot to faculae-dominated regimes between young and
+old ages, we expect this assumption to be reasonably secure
 \citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
@@
 Finally, a challenging but informative quantity to measure would be
-the planet's mass.  This might be possible through long-term
-monitoring of the transit times and durations, or through a
-high-cadence multi-color radial velocity campaign.  Measured at
-sufficient precision, the mass could yield constraints on both the
-planet's composition and its initial entropy
-\citep{owen_constraining_2020}.
+the planet's mass.  This might be possible through a high-cadence
+multi-color radial velocity campaign, or through long-term monitoring
+of the transit times and durations.  Measured at sufficient precision,
+the mass could yield constraints on both the planet's composition and
+its initial entropy \citep{owen_constraining_2020}.
 
@@
 \citep{wang_dai_2019}.  Beyond the planetary transits,
-Appendix~\ref{app:flare} highlights a separate odd coincidence in the
-Kepler data, in the arrival time distribution of stellar flares.  We
+Appendix~\ref{app:flare} highlights a separate oddity in the Kepler
+data, in the arrival time distribution of stellar flares.  We
 encourage its exploration by investigators more versed in the topic
@@
     projected separation of $53\pm4$\,AU.  The inset shows a cutout of
-    the NIRC2 image (North is up, East is left, scale is set by the
+    the stacked NIRC2 image (North is up, East is left, scale is set by the
     separation of the binary).  The lines show 5-$\sigma$ contrast
     limits for the `Alopeke filters, and 6-$\sigma$ contrast limits
-    for NIRC2.  {\it Left:} Gaia EDR3 renormalized
-    unit weight error (RUWE) point estimates for Kepler\,1627A and
-    other members of the $\delta$\,Lyr cluster reported by
-    \citet{KounkelCovey2019}.  Since other members of the cluster with
-    similar brightnesses have comparable degrees of photometric
-    variability, the high RUWE suggests that Kepler\,1627 is a
-    binary. 
+    for NIRC2.  Inside the diffraction limit (estimated as the gray
+    band), the detection limits are less reliable.  {\it Left:} Gaia
+    EDR3 renormalized unit weight error (RUWE) point estimates for
+    Kepler\,1627A and other members of the $\delta$\,Lyr cluster
+    reported by \citet{KounkelCovey2019}.  Since other members of the
+    cluster with similar brightnesses have comparable degrees of
+    photometric variability, the high RUWE suggests that Kepler\,1627
+    is a binary. 
     \label{fig:kep1627binary}
@@
 companion is bound.  Although the companion is unresolved in the Gaia
-source catalog (there are no candidates brighter than $G < 20.5$ mag
+source catalog (there are no comoving, codistant candidate companions brighter than $G < 20.5$ mag
 within $\rho < 120\arcsec$), its existence was also suggested by the

### 2021-09-23 drivers/plot_contrast_imgs.py

@@ -31,9 +31,23 @@ def plot_zorro_speckle():
 
     for d,l,c in zip(datapaths, labels, colors):
         df = pd.read_csv(d)
+        sel = df.sep>0.15
         ax.plot(
-            df.sep, df.deltamag, color=c, label=l
+            df[sel].sep, df[sel].deltamag, color=c, label=l
         )
+        ax.plot(
+            df[~sel].sep, df[~sel].deltamag, color=c, ls=':'
+        )
+
+    xf = np.linspace(0,0.15,100)
+    ylim = ax.get_ylim()
+
+    ax.fill_between(xf, min(ylim),
+                    max(ylim), alpha=0.3,
+                    color='gray', lw=0,
+                    label=r'Diffraction limit',zorder=-10)
+    ax.set_ylim(ylim[::-1])
+
 
     leg = ax.legend(
         loc='upper center', handletextpad=0.2, fontsize='x-small',
@@ -42,7 +56,7 @@ def plot_zorro_speckle():
 
     img = mpimg.imread(
         os.path.join(DATADIR, 'imaging',
-                     'NIRC2_img_20150722_39302_rot.png')
+                     'NIRC2_img_2015stack_rot.png')
     )
 
     # [left, bottom, width, height]
@@ -54,8 +68,6 @@ def plot_zorro_speckle():
     ax.set_ylabel('$\Delta$mag')
     ax.set_xlabel('Angular separation [arcsec]')
 
-    ylim = ax.get_ylim()
-    ax.set_ylim(ylim[::-1])
 
     ax.set_xlim([0,1.25])
 

### 2021-09-23 rudolf/plotting.py

@@ -140,7 +140,8 @@ def plot_ruwe_vs_apparentmag(outdir):
 
     ax.scatter(
         get_xval(df_edr3), get_yval(df_edr3), c='k', alpha=0.9,
-        zorder=4, s=5, rasterized=True, linewidths=0, label='$\delta$ Lyr cluster', marker='.'
+        zorder=4, s=5, rasterized=True, linewidths=0,
+        label='$\delta$ Lyr candidates', marker='.'
     )
     ax.plot(
         get_xval(trgt_df), get_yval(trgt_df), alpha=1, mew=0.5,

### 2021-09-24 paper/ms.tex
update ms; consider slope quartile phasedlc
Co-authored-by: adamkraus <alk@astro.as.utexas.edu>

@@
 ($9\pm1\,{\rm Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and
-V1298 Tau ($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  The light
-curve and the phase-folded transit are shown in
-Figure~\ref{fig:lc}.  The photometry is dominated by a
-quasi-periodic starspot signal with a peak-to-peak amplitude that
-varies between 2\% and 8\%.  The signal
-repeats every $2.642\pm0.042$\,days, where the quoted uncertainty is based on
-the scatter in rotation periods measured from each individual Kepler quarter. 
-Our model for the starspot signal is discussed in detail
-in Appendix~\ref{app:gptransit}, and the adopted stellar
-parameters are listed in Table~\ref{tab:starparams}.  The stellar mass,
-radius, and effective temperature are found by interpolating against a
-38$\,$Myr MIST isochrone.  Their systematic uncertainties are taken
-to be the difference between the PARSEC and MIST isochrones, and the
-reported uncertainties are a quadrature sum of the statistical and
-systematic components. 
-The transit fitting results, given in
-Table~\ref{tab:posterior}, are broadly consistent with a mini-Neptune sized
-planet ($3.34\pm 0.18\,R_\oplus$) on a close-in circular\footnote{Our
-transit fitting yields $e<0.48$ at 2-$\sigma$; the constraints on the
-eccentricity are not particularly strong.} orbit around a G8V star
+V1298 Tau ($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  The
+light curve and the phase-folded transit are shown in
+Figure~\ref{fig:lc}.  The photometry is dominated by a quasi-periodic
+starspot signal with a peak-to-peak amplitude that varies between 2\%
+and 8\%.  The signal repeats every $2.642\pm0.042$\,days, where the
+quoted uncertainty is based on the scatter in rotation periods
+measured from each individual Kepler quarter.  Our model for the
+starspot signal is discussed in detail in
+Appendix~\ref{app:gptransit}, and the adopted stellar parameters are
+listed in Table~\ref{tab:starparams}.  The stellar mass, radius, and
+effective temperature are found by interpolating against a 38$\,$Myr
+MIST isochrone.  Their systematic uncertainties are taken to be the
+difference between the PARSEC and MIST isochrones, and the reported
+uncertainties are a quadrature sum of the statistical and systematic
+components.  The transit fitting results, given in
+Table~\ref{tab:posterior}, are broadly consistent with a mini-Neptune
+sized planet ($3.34\pm 0.18\,R_\oplus$) on a close-in
+circular\footnote{Our transit fitting yields $e<0.48$ at 2-$\sigma$;
+the constraints on the eccentricity are not particularly strong.}
+orbit around a G8V star
 ($0.88 \pm 0.02 R_\odot$).
@@
 the stellar obliquity, and would confirm or refute the expected
-prograde orbital geometry.  Similarly, transit spectroscopy aimed at
+prograde orbital geometry.  Separately, transit spectroscopy aimed at
 detecting atmospheric outflows could yield insight into the
@@
 Observations aimed at quantifying the amount of high-energy
-irradiation incident on the planet would also help in clarifying the
-expected outflow rate \citep[{\it e.g.},][]{poppenhaeger_2021}.
-Finally, a challenging but informative quantity to measure would be
-the planet's mass.  This might be possible through a high-cadence
-multi-color radial velocity campaign, or through long-term monitoring
-of the transit times and durations.  Measured at sufficient precision,
-the mass could yield constraints on both the planet's composition and
-its initial entropy \citep{owen_constraining_2020}.
-
-Separately, the Kepler data may yet contain additional information.
-For instance, one possible explanation for the transit asymmetry shown
-in Figure~\ref{fig:lc} is that of a dusty asymmetric outflow.  Other
-possible explanations such as favored longitudes for starspot
-crossings seem strained, for reasons discussed in
+irradiation incident on the planet would complement these efforts, by
+helping to clarify the expected outflow rate \citep[{\it
+e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
+quantity to measure would be the planet's mass.  This might be
+possible through a high-cadence multi-color radial velocity campaign,
+or through long-term monitoring of the transit times and durations.
+Measured at sufficient precision, the mass, combined with the known
+age and size, would yield constraints on both the planet's composition
+and its initial entropy \citep{owen_constraining_2020}.
+
+More immediately, the Kepler data may yet contain additional
+information.  For instance, one possible explanation for the transit
+asymmetry shown in Figure~\ref{fig:lc} is that of a dusty asymmetric
+outflow.  Other possible explanations such as favored longitudes for
+starspot crossings seem strained for reasons discussed in
 Appendix~\ref{app:gptransit}.  Dusty outflows are theoretically
@@
 
-\subsection{Long Cadence Data}
 \begin{figure}[t]
@@
 results, and so opted for the simpler GP kernel.  We fitted the models
-using \texttt{PyMC3} \citep{salvatier_2016_PyMC3,exoplanet:theano}.
-We assumed a Gaussian likelihood, and after initializing each model
-with the parameters of the maximum {\it a posteriori} model, we
-sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
+using \texttt{PyMC3} \citep{salvatier_2016_PyMC3,exoplanet:theano},
+and accounted for the finite integration time of each exposure in the
+numerical integration when evaluating the model light curve
+\citep[see][]{kipping_binning_2010}.  We assumed a Gaussian
+likelihood, and after initializing each model with the parameters of
+the maximum {\it a posteriori} model, we sampled using
+\texttt{PyMC3}'s gradient-based No-U-Turn Sampler
 \citep{hoffman_no-u-turn_2014} in these bases indicated in

### 2021-09-24 drivers/plot_phasedlc_slope_quartiles.py

@@ -1,0 +1,13 @@
+import os
+import rudolf.plotting as rp
+from rudolf.paths import RESULTSDIR
+import numpy as np
+
+PLOTDIR = os.path.join(RESULTSDIR, 'phasedlc_slope_quartiles')
+if not os.path.exists(PLOTDIR):
+    os.mkdir(PLOTDIR)
+
+ylimd = {'0':[-3.5, 2.5],'1':[-3.5, 2.5],'2':[-3.5, 2.5],'3':[-3.5, 2.5],
+         '4':[-0.19,0.19],'5':[-0.19,0.19],'6':[-0.19,0.19],'7':[-0.19,0.19],}
+rp.plot_phasedlc_slope_quartiles(PLOTDIR, from_trace=True, ylimd=ylimd,
+                                 fullxlim=False, do_hacky_reprerror=True)

### 2021-09-24 drivers/run_RotGPtransit.py

@@ -97,25 +97,20 @@ def run_RotGPtransit(starid='Kepler_1627', N_samples=2000):
 
     if phaseplot:
         outpath = join(PLOTDIR, f'{starid}_{modelid}_posterior_phaseplot.png')
-        ylimd = {'A':[-3.5, 2.5], 'B':[-0.2,0.2]}
+        ylimd = {'A':[-3.5, 2.5], 'B':[-0.19,0.19]}
         bp.plot_phased_light_curve(datasets, m.trace.posterior, outpath,
                                    from_trace=True, ylimd=ylimd,
                                    map_estimate=m.map_estimate,
-                                   do_hacky_reprerror=True)
+                                   do_hacky_reprerror=True,
+                                   rebinthemodel=False, binsize_minutes=20)
         outpath = join(PLOTDIR,
                        f'{starid}_{modelid}_posterior_phaseplot_fullxlim.png')
-        ylimd = {'A':[-3.5, 2.5], 'B':[-0.2,0.2]}
+        ylimd = {'A':[-3.5, 2.5], 'B':[-0.19,0.19]}
         bp.plot_phased_light_curve(datasets, m.trace.posterior, outpath,
                                    from_trace=True, ylimd=ylimd,
                                    map_estimate=m.map_estimate, fullxlim=True,
-                                   BINMS=0.5, do_hacky_reprerror=True)
-        assert 0
-
-    if posttable:
-        outpath = join(PLOTDIR, f'{starid}_{modelid}_posteriortable.tex')
-        make_posterior_table(pklpath, priordict, outpath, modelid, makepdf=1,
-                             var_names=var_names)
-        assert 0
+                                   BINMS=0.5, do_hacky_reprerror=True,
+                                   rebinthemodel=False, binsize_minutes=20)
 
     if phasedsubsets:
         outpath = join(PLOTDIR, f'{starid}_{modelid}_phasedsubsets_yearchunk.png')
@@ -125,7 +120,7 @@ def run_RotGPtransit(starid='Kepler_1627', N_samples=2000):
         bp.plot_phased_subsets(
             datasets, m.trace.posterior, outpath, timesubsets, from_trace=True,
             map_estimate=m.map_estimate, yoffsetNsigma=5, ylimd=ylimd,
-            inch_per_subset=0.75
+            inch_per_subset=0.75, binsize_minutes=20
         )
 
         timepath = os.path.join(
[…]

### 2021-09-24 paper/ms.tex
update ms (add quartile phased lc)

@@
   Kepler\,1627Ab is the youngest planet with a precise age yet found
-  by the main Kepler mission.  From unresolved starspot-crossing
-  events in the Kepler photometry and a Keck/HIRES transit we find
-  that the planet is likely on a prograde orbit.  The average Kepler
-  transit profile also appears asymmetric; our current best
-  explanation is that of a dusty atmospheric outflow being present,
-  but this remains to be confirmed.  Broadly speaking, the
-  $\delta$~Lyr cluster is one of many stellar groups whose properties
-  have been clarified using Gaia; many other exoplanet hosts belong to
-  these new clusters.  Verifying the results of Gaia clustering
-  analyses using TESS rotation periods offers a promising path toward
-  expanding the census of age-dated planets, and our corresponding
-  understanding of planetary evolution.
+  by the main Kepler mission.  The Kepler data show two notable
+  anomalies.  First, after removal of the stellar variability, the
+  average Kepler transit profile appears to be asymmetric.  One
+  possible explanation is that a dusty atmospheric outflow is present,
+  but this remains to be confirmed.  Second, the individual transits
+  show a correlation between the transit time and the local slope of
+  the light curve.  Interpreted in a framework in which unresolved
+  starspot-crossing events cause the transit timing variations, this
+  suggests that the planet is on a prograde orbit.  We present
+  Keck/HIRES observations that are consistent with the latter point, though
+  the data do not have the precision necessary to claim a detection of the
+  spectroscopic transit.  More broadly, the $\delta$~Lyr cluster
+  is one of many stellar groups whose properties have been clarified
+  using Gaia; many other exoplanet hosts belong to these new clusters.
+  Verifying the results of Gaia clustering analyses using TESS
+  rotation periods offers a promising path toward expanding the census
+  of age-dated planets, and our corresponding understanding of
+  planetary evolution.
 \end{abstract}
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.5\textwidth]{f2.pdf}
+		\includegraphics[width=0.8\textwidth]{f2.pdf}
 	\end{center}
-	\vspace{-0.6cm}
+	\vspace{-0.7cm}
 	\caption{
@@
 		members.  Both Kepler (gray panels) and TESS observed portions of
-		the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray;
+		the cluster.  Stars visible by eye ($m_{\rm V}<6.5$) are shown in gray;
 		Deneb, Vega, and $\delta$~Lyr are orange.  Kepler\,1627 (green star) was observed during the
-		entirety of the Kepler mission.  Colors of candidate cluster
-		members indicate the duration of TESS observations to date, in
-		units of lunar months.
+		entirety of the Kepler mission, and to date has been observed for
+		two lunar months by TESS.%.  Colors of candidate cluster
+		%members indicate the duration of TESS observations to date, in
+		%units of lunar months.
 		\label{fig:skychart}
 	}
-\end{figure}
+\end{figure*}
 
@@
 
-		\vspace{-0.5cm}
+		\vspace{-0.6cm}
 		\subfloat{
@@
 
+		\vspace{-0.2cm}	
 		\subfloat{
@@
 
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=\textwidth]{f19.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Transit models and residuals, binned by local slope of the
+    raw light curve}.  Local linear slopes are measured over each
+    transit window (144 transits total).  Four outlier transits were
+    removed, leaving 140 transits.  These were then divided into
+    quartiles, so that each panel shows 35 transits binned together.
+    The exact light curve slopes are listed in the lower left panels.
+    Representative uncertainties for the black points (binned at 20
+    minute intervals) are shown in the lower right of each panel.  A
+    similar transit asymmetry to that shown in Figure~\ref{fig:lc}
+    seems to be present in three of the four bins.
+		\label{fig:phaseslope}
+	}
+\end{figure}
+
 We fitted the Kepler long cadence time series with a model that
@@
 
-Figures~\ref{fig:phasequarter} and~\ref{fig:phaseyear} show the Kepler
-data binned over quarters and years.  In
-Figure~\ref{fig:phasequarter}, Quarter 6 shows the strongest asymmetry
-out of the quarters: a deviation of about 3\,ppt from expectation.
+As a means of exploring the robustness of the asymmetry,
+Figures~\ref{fig:phasequarter},~\ref{fig:phaseyear},
+and~\ref{fig:phaseslope} show the Kepler data binned in three ways:
+over quarters, years, and local slope quartiles.  Over quarters
+(Figure~\ref{fig:phasequarter}), Quarter 6 shows the strongest asymmetry
+out of any quarter: a deviation of about 3\,ppt from expectation.
 Quarter 7 shows an anomaly at roughly the same transit phase.  Year 2
@@
 in each of Years 2, 3, and 4.
-
-We therefore considered four possible explanations for the anomaly:
+Binned by local slope quartiles (Figure~\ref{fig:phaseslope}), the
+asymmetry is visually present in three of the four quartiles: the only
+bin for which it appears to not be present is ${\rm d}f/{\rm
+d}t\in[0.001,0.033]\,{\rm ppt\,day}^{-1}$.
+
+We considered four possible interpretations of the transit asymmetry:
 gravity darkening, transit timing variations, spot-crossing events,
@@
 standard deviation are $0.49\pm44.49$\,ppt\,day$^{-1}$.  There is
-therefore no expectation for TTVs to produce the asymmetry.
-
-The third possibility is that of starspot crossings.  Young stars 
-have higher spot-covering fractions than old stars ({\it
+therefore no expectation for TTVs to produce the asymmetry.  A
+separate line of argument comes from Figure~\ref{fig:phaseslope}.  If
+the local slope were essential to producing the transit asymmetry, we
+would expect that in the largest ${\rm d}f/{\rm d}t$ bin, ${\rm
+d}f/{\rm d}t\in[0.033,0.092]\,{\rm ppt\,day}^{-1}$, the sign of the
+asymmetry would reverse.  We do not see evidence for that being the
+case.
+
+The third and related possibility is that of starspot crossings.
+Young stars have higher spot-covering fractions than old stars ({\it
 e.g.}, \citealt{morris_relationship_2020}).  Young solar-type stars
@@
 b)} there is no reason to expect starspot crossing events to last
-exactly half the transit
-duration.
+exactly half the transit duration.
 

### 2021-09-24 paper/collect.sh

@@ -21,6 +21,7 @@ cp /Users/luke/Dropbox/proj/rudolf/results/flares/flare_pair_separation_histogra
 
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_phasedsubsets_quarterchunk.pdf f9.pdf
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_phasedsubsets_yearchunk.pdf f10.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/phasedlc_slope_quartiles/Kepler_1627_RotGPtransit_phasedlc_slope_quartiles.pdf f19.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/ttv_vs_local_slope/ttv_vs_local_slope.pdf f11.pdf
 

### 2021-09-24 paper/ms.tex
update ms

@@
 ({\it e.g.}, as measured in IC~2602 by
-\citealt{randich_gaiaeso_2018}).  It is also larger than any ${\rm
-EW}_{\rm Li}$ found by \citet{berger_identifying_2018} in their
-analysis of 1{,}301 Kepler spectra.
+\citealt{randich_gaiaeso_2018}).  It is also larger than any
+lithium equivalent widths measured by \citet{berger_identifying_2018}
+in their analysis of 1{,}301 Kepler-star spectra.
 
@@
     ({\it right}) are shown, and the residual after subtracting the
-    transit in the bottom-most row.  The 2-$\sigma$ model uncertainties
+    transit is in the bottom-most row.  The 2-$\sigma$ model uncertainties
     and the best-fit model are the light purple band and the dark
@@
 excludes a background eclipsing binary scenario, and therefore
-confirms that \pn\ is a planet.  It also implies that the planet's
+confirms that \pn\ is a planet.  It also suggests that the planet's
 orbit is prograde.  The latter point assumes that the dominant
@@
 spot-corrected exposure-to-exposure scatter ($\sigma_{\rm RV}\approx
-50$\ms) is significantly larger than the expected velocity anomaly
+50$\ms) is larger than the expected velocity anomaly
 ($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$).  We are therefore
 not in a position to claim a spectroscopic detection of the RM effect,
-nor a precise quantification of the stellar obliquity.
+nor to quantify the stellar obliquity.
 
@@
 planets, and opens multiple avenues for further study.  Further
-observations of spectroscopic transits could yield a quantification of
-the stellar obliquity, and would confirm or refute the expected
-prograde orbital geometry.  Separately, transit spectroscopy aimed at
-detecting atmospheric outflows could yield insight into the
-evolutionary state of the atmosphere \citep[{\it
+observations of spectroscopic transits could yield a stellar obliquity
+measurement, which would confirm or refute the prograde orbital
+geometry suggested by the TTV-local slope correlation.  Separately,
+transit spectroscopy aimed at detecting atmospheric outflows could
+yield insight into the evolutionary state of the atmosphere
+\citep[{\it
 e.g.},][]{ehrenreich_giant_2015,spake_helium_2018,vissapragada_2020}.
@@
 helping to clarify the expected outflow rate \citep[{\it
-e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
+e.g.},][]{poppenhaeger_2021}.  Finally, a challenging and informative
 quantity to measure would be the planet's mass.  This might be
@@
 
-More immediately, the Kepler data may yet contain additional
+More immediately, the Kepler data may contain additional
 information.  For instance, one possible explanation for the transit
@@
 primordial atmospheres that are shed over the first gigayear
-\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
-suggestion of a prograde orbit is consistent with a quiescent
-disk-driven migration history, and differs from misalignments that
+\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  The
+prograde orbit, if confirmed, would be consistent with a quiescent
+disk-driven migration history, and would differ from misalignments that
 have been observed for some older
-mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
-Spectroscopic follow-up observations of the transit remain necessary
-however to verify whether the TTV-slope correlation is yielding the
-correct orbital geometry.
+hot Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
+%Spectroscopic follow-up observations of the transit remain necessary
+%however to verify whether the TTV-slope correlation is yielding the
+%correct orbital geometry.
 
@@
 
-Ultimately, the main advance from this work is a precise measurement of
+Ultimately, the main advance of this work is a precise measurement of
 the age of \pn.   This measurement was enabled by identifying the

### 2021-09-24 rudolf/plotting.py

@@ -1312,9 +1312,6 @@ def plot_ttv_vs_local_slope(outdir):
 
     ttv_mean, ttv_std = np.mean(ttv), np.std(ttv)
 
-    import IPython; IPython.embed()
-    assert 0
-
     ttv_cut = 0.02 # doesn't mean match
 
     z1 = np.poly1d(
@@ -1384,7 +1381,7 @@ def plot_ttv_vs_local_slope(outdir):
 
 
     ax.set_ylim([-0.03*24*60,0.03*24*60])
-    ax.set_xlim([-0.15*XMULT,0.15*XMULT])
+    ax.set_xlim([-0.149*XMULT,0.149*XMULT])
 
     ax.legend(fontsize='x-small')
 

### 2021-09-24 paper/ms.tex
update abstract

@@
 \begin{abstract}
+  % VERSION A
   Kepler\,1627A is a G8V star previously known to host a
@@
   by the main Kepler mission.  The Kepler data show two notable
-  anomalies.  First, after removal of the stellar variability, the
-  average Kepler transit profile appears to be asymmetric.  One
-  possible explanation is that a dusty atmospheric outflow is present,
-  but this remains to be confirmed.  Second, the individual transits
-  show a correlation between the transit time and the local slope of
-  the light curve.  Interpreted in a framework in which unresolved
-  starspot-crossing events cause the transit timing variations, this
-  suggests that the planet is on a prograde orbit.  We present
-  Keck/HIRES observations that are consistent with the latter point, though
-  the data do not have the precision necessary to claim a detection of the
-  spectroscopic transit.  More broadly, the $\delta$~Lyr cluster
-  is one of many stellar groups whose properties have been clarified
-  using Gaia; many other exoplanet hosts belong to these new clusters.
+  anomalies: the average Kepler transit profile appears to be
+  asymmetric, and the individual transits show a correlation between
+  the transit time and the local slope of the light curve.  The former
+  might be explained by a dusty atmospheric outflow; the latter could
+  be explained by unresolved starspot crossings of a planet on a
+  prograde orbit.  Broadly speaking, the $\delta$~Lyr cluster is one
+  of many stellar groups whose properties have been clarified using
+  Gaia; many other exoplanet hosts belong to these new clusters.
   Verifying the results of Gaia clustering analyses using TESS
@@
   planetary evolution.
+  %
+  % % VERSION B
+  % Kepler\,1627A is a G8V star previously known to host a
+  % $3.3\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
+  % observed by the Kepler space telescope because it is nearby
+  % ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Here we show
+  % using Gaia kinematics, TESS stellar rotation periods, and
+  % spectroscopic lithium abundances that Kepler\,1627 is a member of
+  % the \clusterage\ old $\delta$~Lyr cluster.  To our knowledge,
+  % Kepler\,1627Ab is the youngest planet with a precise age yet found
+  % by the main Kepler mission.  The Kepler data show two notable
+  % anomalies.  First, after removal of the stellar variability, the
+  % average Kepler transit profile appears to be asymmetric.  One
+  % possible explanation is that a dusty atmospheric outflow is present,
+  % but this remains to be confirmed.  Second, the individual transits
+  % show a correlation between the transit time and the local slope of
+  % the light curve.  Interpreted in a framework in which unresolved
+  % starspot-crossing events cause the transit timing variations, this
+  % suggests that the planet is on a prograde orbit.  We present
+  % Keck/HIRES observations that are consistent with the latter point, though
+  % the data do not have the precision necessary to claim a detection of the
+  % spectroscopic transit.  More broadly, the $\delta$~Lyr cluster
+  % is one of many stellar groups whose properties have been clarified
+  % using Gaia; many other exoplanet hosts belong to these new clusters.
+  % Verifying the results of Gaia clustering analyses using TESS
+  % rotation periods offers a promising path toward expanding the census
+  % of age-dated planets, and our corresponding understanding of
+  % planetary evolution.
 \end{abstract}

### 2021-09-28 paper/ms.tex
clean up paper directory
update ms (Furlan comments)
Co-authored-by: Elise Furlan <furlan@ipac.caltech.edu>

--- 2021-09-24 paper/ms.tex
@@
   Kepler\,1627A is a G8V star previously known to host a
-  $3.3\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
+  $3.4\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
   observed by the Kepler space telescope because it is nearby
@@
   exoplanet evolution (491),
-  %stellar associations (1582),
   open star clusters (1160),
@@
 recent analyses of the Gaia
-data have since expanded our knowledge of cluster memberships
+data have greatly expanded our knowledge of cluster memberships
 \citep[{\it
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f2.pdf}
+		\includegraphics[width=0.49\textwidth]{f1.pdf}
 	\end{center}
@@
 	\caption{
-		{\bf Kepler and TESS views of the $\delta$~Lyr cluster.} Colored
-		points are kinematically selected candidate $\delta$\,Lyr cluster
-		members.  Both Kepler (gray panels) and TESS observed portions of
-		the cluster.  Stars visible by eye ($m_{\rm V}<6.5$) are shown in gray;
-		Deneb, Vega, and $\delta$~Lyr are orange.  Kepler\,1627 (green star) was observed during the
-		entirety of the Kepler mission, and to date has been observed for
-		two lunar months by TESS.%.  Colors of candidate cluster
-		%members indicate the duration of TESS observations to date, in
-		%units of lunar months.
+    {\bf Kepler's view of the $\delta$~Lyr cluster.} Each black circle is a
+    candidate cluster member, selected from the catalog of \citet{KounkelCovey2019} based on its
+    position and kinematics (see Appendix~\ref{app:kinematicselection}).
+    Of the \nkinematic\ candidate cluster members, 58 have at least one
+    quarter of Kepler data.  TESS has also observed most of the
+    cluster, for one to two lunar months to date.
 		\label{fig:skychart}
 	}
-\end{figure*}
+\end{figure}
 
@@
 presence of a previously unreported binary companion, hereafter
-Kepler\,1627B, the planet orbits the G-dwarf primary on an orbit that
-is likely prograde
-(Section~\ref{sec:planet}).  We conclude by highlighting broader
-implications for our ability to age-date a statistical sample of
-planets (Section~\ref{sec:conc}).
+Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A (Section~\ref{sec:planet}).
+Based on a correlation between the planetary transit times and the local
+light curve slope, the orbit also seems to be prograde.  We
+conclude by highlighting broader implications for our ability to
+age-date a larger sample of planets (Section~\ref{sec:conc}).
 
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f15a.pdf}
-			\includegraphics[width=0.469\textwidth]{f15b.pdf}
+			\includegraphics[width=0.49\textwidth]{f2a.pdf}
+			\includegraphics[width=0.469\textwidth]{f2b.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.7\textwidth]{f15c.pdf}
+			\includegraphics[width=0.7\textwidth]{f2c.pdf}
 		}
@@
 3-dimensional maps of \citet{capitanio_threedimensional_2017} and
-\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}},
+\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}, 2021/09/25},
 and then applied the extinction coefficients $k_X\equiv A_X/A_0$
@@
 reddening is higher than the \citet{lallement_threedimensional_2018}
-value by a factor of two.  } While detailed analyses of the LDB or
-main-sequence turn-off in the \cn\ could help resolve the tension
-between the LDB and isochrone ages, these are beyond our scope.
+value by a factor of two.  } Follow-up studies of the LDB or
+main-sequence turn-off in the \cn\ could help determine a more precise
+and accurate age for the cluster, however these are beyond our scope.
 
@@
 \citet{kounkel_untangling_2019}, \nwithtess\ stars were amenable to
-rotation period measurements ($G<17$ and $\bpmrpo>0.5$) and had TESS
-data available.\footnote{Kepler rotation periods were derived by
-\citet{mcquillan_rotation_2014} for 56 stars; in these cases, we
+rotation period measurements ($G<17$ and $\bpmrpo>0.5$) using the TESS
+full frame image data.\footnote{Kepler rotation periods were derived
+by \citet{mcquillan_rotation_2014} for 28 stars; for these cases, we
 adopted the Kepler rotation period.  } We extracted light curves from
@@
 ({\it e.g.}, IC\,2602 and IC\,2391) are indistinguishable
-\citep{douglas_stephanie_t_2021_5131306}.  Binarity is known to be an
-important factor in the density of stars falling underneath the slow
-sequence
-\citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}.
-Appendix~\ref{app:rotationbinarity} discusses this topic, and finds
-evidence for similar effects to those previously noted in the
-literature.
+\citep{douglas_stephanie_t_2021_5131306}.  While binarity is known to
+affect the density of stars that fall beneath the slow sequence
+\citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}, it is
+not expected to change the central conclusion regarding the cluster's
+age.
+%%VERSION1
+%Based on an analysis of photometric binaries and stars with high RUWE,
+%we found evidence for similar effects to those previously noted in the
+%literature.  However, since the purpose of our rotation period
+%analysis is to determine the age of the \cn, we do not dwell on this
+%topic.
+% %VERSION0
+% Appendix~\ref{app:rotationbinarity} discusses this topic, and finds
+% evidence for similar effects to those previously noted in the
+% literature.
 
@@
 yet converged to a slow sequence, their rotation periods are generally
-consistent with a 38\,Myr age. The most precise age for \sn\ and its
+consistent with a $\approx40$\,Myr age. The most precise age for \sn\ and its
 planet are therefore identical to the cluster, based on the star's
@@
 Given the effective temperature of \sn\ (Table~\ref{tab:starparams}),
-this measurement is in agreement with expectations for a 38\,Myr star
+this measurement is in agreement with expectations for a $\approx40$\,Myr star
 ({\it e.g.}, as measured in IC~2602 by
@@
 		\subfloat{
-			\includegraphics[width=\textwidth]{f4.pdf}
+			\includegraphics[width=\textwidth]{f3a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.5\textwidth]{f5a.pdf}
-			\includegraphics[width=0.5\textwidth]{f5b.pdf}
+			\includegraphics[width=0.5\textwidth]{f3b.pdf}
+			\includegraphics[width=0.5\textwidth]{f3c.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f11.pdf}
+			\includegraphics[width=0.49\textwidth]{f4a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f16.pdf}
+			\includegraphics[width=0.49\textwidth]{f4b.pdf}
 		}
@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f14.pdf}
+		\includegraphics[width=0.8\textwidth]{f5.pdf}
 	\end{center}
@@
     {\bf Radii, orbital periods, and ages of transiting exoplanets}.
-    %Planets older than $10^9$ years are shown with gray markers.
     Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
@@
 If the \pn\ transit signal is created by a genuine planet,  then to
-our knowledge its age would make it the youngest planet yet found by
-the main Kepler mission.\footnote{The re-purposed K2 mission however
-has found two younger systems containing five planets: K2-33b
-($9\pm1\,{\rm Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and
-V1298 Tau ($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  The
-light curve and the phase-folded transit are shown in
-Figure~\ref{fig:lc}.  The photometry is dominated by a quasi-periodic
-starspot signal with a peak-to-peak amplitude that varies between 2\%
-and 8\%.  The signal repeats every $2.642\pm0.042$\,days, where the
-quoted uncertainty is based on the scatter in rotation periods
-measured from each individual Kepler quarter.  Our model for the
-starspot signal is discussed in detail in
-Appendix~\ref{app:gptransit}, and the adopted stellar parameters are
-listed in Table~\ref{tab:starparams}.  The stellar mass, radius, and
-effective temperature are found by interpolating against a 38$\,$Myr
-MIST isochrone.  Their systematic uncertainties are taken to be the
-difference between the PARSEC and MIST isochrones, and the reported
-uncertainties are a quadrature sum of the statistical and systematic
-components.  The transit fitting results, given in
-Table~\ref{tab:posterior}, are broadly consistent with a mini-Neptune
-sized planet ($3.34\pm 0.18\,R_\oplus$) on a close-in
-circular\footnote{Our transit fitting yields $e<0.48$ at 2-$\sigma$;
-the constraints on the eccentricity are not particularly strong.}
-orbit around a G8V star
-($0.88 \pm 0.02 R_\odot$).
-
-Could this transit signal be produced by anything other than a planet
+our knowledge it would be the youngest planet yet found by the main
+Kepler mission.\footnote{The re-purposed K2 mission however has found
+two younger systems containing five planets: K2-33b ($9\pm1\,{\rm
+Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and V1298 Tau
+($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  The light curve
+and the phase-folded transit are shown in Figure~\ref{fig:lc}.  The
+photometry is dominated by a quasi-periodic starspot signal with a
+peak-to-peak amplitude that varies between 2\% and 8\%.  The signal
+repeats every $2.642\pm0.042$\,days, where the quoted uncertainty is
+based on the scatter in rotation periods measured from each individual
+Kepler quarter.  Our model for the starspot and transit signals is
+discussed in detail in Appendix~\ref{app:gptransit}, and the adopted
+stellar parameters are listed in Table~\ref{tab:starparams}.  The
+stellar mass, radius, and effective temperature are found by
+interpolating against a 38$\,$Myr MIST isochrone.  Their systematic
+uncertainties are taken to be the difference between the PARSEC and
+MIST isochrones, and the reported uncertainties are a quadrature sum
+of the statistical and systematic components.
+
+The transit fitting results, given in Table~\ref{tab:posterior}, are
+broadly consistent with a mini-Neptune sized planet ($3.4\pm
+0.2\,R_\oplus$) on a close-in circular\footnote{ Our transit fitting
+yields $e<0.48$ at 2-$\sigma$; the constraints on the eccentricity are
+not particularly strong.} orbit around a G8V host star ($0.88 \pm 0.02
+R_\odot$).  Our best-fit planet size is smaller than those
+previously reported by \citet{morton_false_2016} and
+\citet{berger_identifying_2018}, at 0.7-$\sigma$ and 1.2-$\sigma$
+confidence respectively.  The former is explained by differing
+assumptions about the stellar radius; the latter appears to be due to
+differences in the measured transit depth, which could be linked to
+the different methods used to remove the stellar rotation signal.
+
+Could the transit be produced by anything other than a planet
 orbiting this near-solar analog?  \citet{morton_false_2016} validated
@@
 accompanied by a low-mass stellar companion ($M_{\rm B}\approx
-0.33\,M_\odot$).  We therefore reassessed false positive
-scenarios in some detail.
+0.33\,M_\odot$; $\rho\approx0\farcs16$).  We therefore reassessed
+false positive scenarios in some detail.
 
@@
 \citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020,martioli_aumicbc_2021}.
-\pn\ is one of the smallest planets in this sample ($3.34\pm
-0.18\,R_\oplus$), which could be linked to the selection effects
+\pn\ is one of the smallest planets in this sample ($3.4\pm
+0.2\,R_\oplus$), which could be linked to the selection effects
 imposed by the spot-induced photometric variability at very young ages
@@
 
-The \texttt{v0.6} CDIPS target catalog (Table~\ref{tab:v06}) includes stars
-that are young, age-dated, and age-dateable.  By ``age-dateable'', we mean that
-the age should be measurable at greater precision than a typical FGK field star
-using isochronal, gyrochronal, or spectroscopic techniques.
+The \texttt{v0.6} CDIPS target catalog (Table~\ref{tab:v06}) includes
+stars that are young, age-dated, and age-dateable.  By
+``age-dateable'', we mean that the stellar age should be measurable at
+greater precision than that of a typical FGK field star, through
+either isochronal, gyrochronal, or spectroscopic techniques.
 As in
@@
 
-One major change in Table~\ref{tab:v06} relative to the analogous
-table from \citet{bouma_cdipsI_2019} is that the extent of Gaia-based
+One major change in Table~\ref{tab:v06} relative to the earlier
+iteration from \citet{bouma_cdipsI_2019} is that the extent of Gaia-based
 analyses has now matured to the point that we can neglect pre-Gaia
@@
 (MWSC and DAML) are no longer required.  This is helpful for various
-post-processing projects,  since the field-star contamination rates
-were typically much higher than analogous Gaia-based cluster membership
-catalogs.
+post-processing projects,  since the field star contamination rates
+were typically much higher in these catalogs than in the newer
+Gaia-based catalogs.
 
@@
 performed asychronously on the Gaia archive
-website\footnote{\url{https://gea.esac.esa.int/archive/}}.
+website.\footnote{\url{https://gea.esac.esa.int/archive/}}
 
@@
 		\leavevmode
-		\includegraphics[width=1\textwidth]{f1.pdf}
+		\includegraphics[width=1\textwidth]{f6.pdf}
 	\end{center}
@@
 
-\section{Binarity and Stellar Rotation for $\delta$\,Lyr Cluster Members}
-\label{app:rotationbinarity}
+% \section{Binarity and Stellar Rotation for $\delta$\,Lyr Cluster Members}
+% \label{app:rotationbinarity}
+% 
+% \begin{figure}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\includegraphics[width=0.7\textwidth]{f7.pdf}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+%   \caption{ {\bf Binarity indicators for rotators in \cn.} 
+%     Data are as in the lower panel of Figure~\ref{fig:age}.  Sources
+%     with $\mathrm{RUWE}>1.2$ appear in red; these include astrometric binaries as well
+%     as sources poorly fit by a single-star PSF.  Photometric binaries
+%     ($>$0.3\,mag above an empirical isochrone; 20\% of stars) are
+%     orange.  Over $0.6<\bpmrpo<1.5$, the detected rotation periods in
+%     binaries tend to fall below the slow sequence.
+% 		\label{fig:binarity}
+% 	}
+% \end{figure}
+% 
+% Figure~\ref{fig:binarity} shows the rotation-color diagram for \cn\
+% members, with the points colored according to indicators of binarity.
+% The binaries tend to be redder and have shorter rotation periods.  We
+% know of two possible explanations based on selection effects, and two
+% based on physics.
+% 
+% Possible selection effects include {\it i)} binaries have a component
+% that contributes additional red light to the system, which could skew
+% the color measurement of the primary; and {\it ii)} the unresolved
+% binary companions could contaminate the rotation period measurement
+% \citep[{\it e.g.},][Section~5.1]{stauffer_rotation_2016}.
+% 
+% The two possible physical effects that could be relevant are tidal
+% locking and pre-main-sequence disk locking.  Tidal locking has been
+% argued to be an unlikely explanation for the frequency of rapid
+% rotators due to the population statistics; a more likely scenario
+% would be that the presence of the binary leads to faster disk
+% dispersal, enabling the primary to contract to more rapid rotation
+% periods than possible for single stars
+% \citep{meibom_effect_2007,bouma_cluster_2020}.
+
+\section{Transit and Stellar Variability Model}
+\label{app:gptransit}
 
@@
 		\leavevmode
-		\includegraphics[width=0.7\textwidth]{f15supp.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf Binarity indicators for rotators in \cn.} 
-    Data are as in the lower panel of Figure~\ref{fig:age}.  Sources
-    with $\mathrm{RUWE}>1.2$ appear in red; these include astrometric binaries as well
-    as sources poorly fit by a single-star PSF.  Photometric binaries
-    ($>$0.3\,mag above an empirical isochrone; 20\% of stars) are
-    orange.  Over $0.6<\bpmrpo<1.5$, the detected rotation periods in
-    binaries tend to fall below the slow sequence.
-		\label{fig:binarity}
-	}
-\end{figure}
-
-Figure~\ref{fig:binarity} shows the rotation-color diagram for \cn\
-members, with the points colored according to indicators of binarity.
-The binaries tend to be redder and have shorter rotation periods.  We
-know of two possible explanations based on selection effects, and two
-based on physics.
-
-Possible selection effects include {\it i)} binaries have a component
-that contributes additional red light to the system, which could skew
-the color measurement of the primary; and {\it ii)} the unresolved
-binary companions could contaminate the rotation period measurement
-\citep[{\it e.g.},][Section~5.1]{stauffer_rotation_2016}.
-
-The two possible physical effects that could be relevant are tidal
-locking and pre-main-sequence disk locking.  Tidal locking has been
-argued to be an unlikely explanation for the frequency of rapid
-rotators due to the population statistics; a more likely scenario
-would be that the presence of the binary leads to faster disk
-dispersal, enabling the primary to contract to more rapid rotation
-periods than possible for single stars
-\citep{meibom_effect_2007,bouma_cluster_2020}.
-
-\section{Transit and Stellar Variability Model}
-\label{app:gptransit}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.64\textwidth]{f9.pdf}
+		\includegraphics[width=0.64\textwidth]{f7.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.64\textwidth]{f10.pdf}
+		\includegraphics[width=0.64\textwidth]{f8.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=\textwidth]{f19.pdf}
+		\includegraphics[width=\textwidth]{f9.pdf}
 	\end{center}
@@
 is present during the transit: the measured flux is systematically
-high during the first half, and low in the second half.  To determine
+high during the first half, and low in the second half.  
+To determine
 whether this effect was caused by our stellar variability model, we 
@@
 comparable amplitude.
+An important note is that even though this asymmetric feature is
+within the plotted 2-$\sigma$ model uncertainties, this is within a
+model that has a jitter term that grows to account for otherwise unmodelled 
+white noise in the flux.
+The significance of the asymmetry is therefore better assessed in
+comparison against the intrinsic out-of-transit scatter, rather than the
+model uncertainties.
 
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f17.pdf}
-			\includegraphics[width=0.48\textwidth]{f6.pdf}
+			\includegraphics[width=0.49\textwidth]{f10a.pdf}
+			\includegraphics[width=0.48\textwidth]{f10b.pdf}
 		}
@@
     limits for the `Alopeke filters, and 6-$\sigma$ contrast limits
-    for NIRC2.  Inside the diffraction limit (estimated as the gray
-    band), the detection limits are less reliable.  {\it Left:} Gaia
+    for NIRC2 outside of $0\farcs15$. {\it Right:} Gaia
     EDR3 renormalized unit weight error (RUWE) point estimates for
-    Kepler\,1627A and other members of the $\delta$\,Lyr cluster
+    Kepler\,1627A and all members of the $\delta$\,Lyr cluster
     reported by \citet{KounkelCovey2019}.  Since other members of the
@@
 companion is bound.  Although the companion is unresolved in the Gaia
-source catalog (there are no comoving, codistant candidate companions brighter than $G < 20.5$ mag
-within $\rho < 120\arcsec$), its existence was also suggested by the
-primary star's large RUWE, relative to other members of the \cn.
-Based on the apparent separation, the binary orbital period is of
-order hundreds of years.  The large RUWE is therefore more likely to
-be caused by a PSF-mismatch skewing the Gaia centroiding during
-successive scans, rather than true astrometric motion.  Regardless,
-given the large observed RUWE of the primary, and the low geometric
-probability that a companion imaged at $\rho \approx 0\farcs16$ is a
-chance line-of-sight companion, we proceed under the assumption that
-the companion is bound, and that Kepler\,1627 is a binary.  Given the
-distance and age, the models of \citet{baraffe_new_2015} imply a
-companion mass of $M_{\rm B} \approx 0.33 M_{\odot}$ and companion
-temperature of $T_{\rm eff,B} \approx 3450$ K.  The corresponding
-spectral type is roughly M2.5V \citep{pecaut_mamajek_2013}.
+source catalog (there are no comoving, codistant candidate companions
+brighter than $G < 20.5$ mag within $\rho < 120\arcsec$), its
+existence was also suggested by the primary star's large RUWE,
+relative to other members of the \cn.  Based on the apparent
+separation, the binary orbital period is of order hundreds of years.
+The large RUWE is therefore more likely to be caused by a PSF-mismatch
+skewing the Gaia centroiding during successive scans, rather than true
+astrometric motion.  Regardless, given the large observed RUWE of the
+primary, and the low geometric probability that a companion imaged at
+$\rho \approx 0\farcs16$ is a chance line-of-sight companion, we
+proceed under the assumption that the companion is bound, and that
+Kepler\,1627 is a binary.  Given the distance and age, the models of
+\citet{baraffe_new_2015} imply a companion mass of $M_{\rm B} \approx
+0.33 M_{\odot}$ and companion temperature of $T_{\rm eff,B} \approx
+3450$ K.  The corresponding spectral type is roughly M2.5V
+\citep{pecaut_mamajek_2013}.
 % \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
@@
 		\subfloat{
-			\includegraphics[width=\textwidth]{f18.pdf}
+			\includegraphics[width=\textwidth]{f11.pdf}
 		}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f7.pdf}
+		\includegraphics[width=0.9\textwidth]{f12.pdf}
 	\end{center}

### 2021-09-28 drivers/plot_skychart.py

@@ -6,6 +6,15 @@ PLOTDIR = os.path.join(RESULTSDIR, 'skychart')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
+rp.plot_skychart(PLOTDIR, narrowlims=0, showkepler=1, showtess=0,
+                 shownakedeye=0, showcomovers=1, showkepstars=0)
+
+rp.plot_skychart(PLOTDIR, narrowlims=0, showkepler=1, showtess=0,
+                 shownakedeye=0, showcomovers=1, showkepstars=1)
+
+rp.plot_skychart(PLOTDIR, narrowlims=0, showkepler=1, showtess=1,
+                 shownakedeye=0, showcomovers=1)
+
 rp.plot_skychart(PLOTDIR, narrowlims=0, showkepler=1, showtess=1,
                  shownakedeye=1, showcomovers=1)
 rp.plot_skychart(PLOTDIR, narrowlims=1, showkepler=1, showtess=1,

### 2021-09-28 rudolf/plotting.py

@@ -341,16 +341,17 @@ def plot_simulated_RM(outdir, orientation, N_mcmc=10000):
 
 
 def plot_skychart(outdir, narrowlims=0, showkepler=0, showtess=0,
-                  shownakedeye=0, showcomovers=0):
+                  shownakedeye=0, showcomovers=0, showkepstars=0):
 
     set_style()
 
     df_dr2, df_edr3, trgt_df = get_deltalyr_kc19_gaia_data()
     if showcomovers:
-        df_edr3 = get_deltalyr_kc19_comovers()
+        df_edr3 = get_deltalyr_kc19_cleansubset()
+        #df_edr3 = get_deltalyr_kc19_comovers()
 
     plt.close('all')
-    f, ax = plt.subplots(figsize=(4,3), constrained_layout=True)
+    f, ax = plt.subplots(figsize=(3,2.5), constrained_layout=True)
 
     xkey, ykey = 'ra', 'dec'
     get_yval = (
@@ -367,14 +368,27 @@ def plot_skychart(outdir, narrowlims=0, showkepler=0, showtess=0,
     if not showtess:
         ax.scatter(
             get_xval(df_edr3), get_yval(df_edr3), c='k', alpha=0.9,
-            zorder=4, s=5, rasterized=True, linewidths=0,
-            label='$\delta$ Lyr Comovers', marker='.'
+            zorder=4, s=8, rasterized=True, linewidths=0,
+            label='$\delta$ Lyr candidates', marker='.'
         )
         ax.plot(
             get_xval(trgt_df), get_yval(trgt_df), alpha=1, mew=0.5,
             zorder=8, label='Kepler 1627', markerfacecolor='yellow',
-            markersize=10, marker='*', color='black', lw=0
+            markersize=15, marker='*', color='black', lw=0
         )
+    if showkepstars:
+        _df = pd.read_csv('/Users/luke/Dropbox/proj/cdips_followup/results/KNOWNPLANET_YOUNGSTAR_XMATCH/20210916_kepgaiafun_X_cdips0pt6/20210916_kepgaiafun_X_cdipsv0pt6_short.csv')
+        sel = ~pd.isnull(_df.cluster)
+        _sdf = _df[sel]
+        sel = (_sdf.cluster.str.contains('Stephenson'))
+        _sdf0 = _sdf[sel]
+        sel = (df_edr3.source_id.astype(str).isin(_sdf0.source_id.astype(str)))
+        ax.scatter(
[…]

### 2021-09-28 paper/ms.tex
update ms (LAH init comments)
Co-authored-by: Lynne Hillenbrand <lah@astro.caltech.edu>

@@
 %\suppressAffiliations
-\NewPageAfterKeywords
+%\NewPageAfterKeywords
 \input{authors.tex}
@@
   Kepler\,1627Ab is the youngest planet with a precise age yet found
-  by the main Kepler mission.  The Kepler data show two notable
+  by the main Kepler mission.  The Kepler photometry shows two notable
   anomalies: the average Kepler transit profile appears to be
   asymmetric, and the individual transits show a correlation between
-  the transit time and the local slope of the light curve.  The former
-  might be explained by a dusty atmospheric outflow; the latter could
-  be explained by unresolved starspot crossings of a planet on a
-  prograde orbit.  Broadly speaking, the $\delta$~Lyr cluster is one
-  of many stellar groups whose properties have been clarified using
-  Gaia; many other exoplanet hosts belong to these new clusters.
-  Verifying the results of Gaia clustering analyses using TESS
-  rotation periods offers a promising path toward expanding the census
-  of age-dated planets, and our corresponding understanding of
-  planetary evolution.
+  the transit time and the local light curve slope.
+  % The former could be explained by a dusty atmospheric outflow; the
+  % latter could be explained by unresolved starspot crossings of a
+  % planet on a prograde orbit.  
+  We discuss plausible explanations for the light curve anomalies, and
+  also highlight broader implications for our ability to expand the census
+  of age-dated planets.
 \end{abstract}
@@
     candidate cluster member, selected from the catalog of \citet{KounkelCovey2019} based on its
-    position and kinematics (see Appendix~\ref{app:kinematicselection}).
+    position and kinematics.
     Of the \nkinematic\ candidate cluster members, 58 have at least one
@@
 Matching our young star list against stars observed by Kepler yielded
-two discoveries.  The first, to be discussed in an upcoming analysis
-by J.~Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
-open cluster Theia~520 (UBC~1).  At least six Kepler planets are known
-in the cluster from the Kepler-52 and Kepler-968 systems
-\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
-discovery, and the focus of this work, is that Kepler observed a
-portion of the $\delta$\,Lyr cluster (Stephenson-1; Theia~73;
-\citealt{stephenson_possible_1959}).  Figure~\ref{fig:skychart} shows
-the overlap between the Kepler field and the cluster.
-
-A previous clustering analysis of the Gaia data reported that
-Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member of the
-\cn\ \citep{kounkel_untangling_2019}.  Given the previous statistical
-validation of the close-in mini-Neptune Kepler 1627b
+%two discoveries.  The first, to be discussed in an upcoming analysis
+%by J.~Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
+%open cluster Theia~520 (UBC~1).  At least six Kepler planets are known
+%in the cluster from the Kepler-52 and Kepler-968 systems
+%\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
+%discovery, and the focus of this work, is
+the discovery that Kepler observed a portion of the \cn\ (Stephenson-1; Theia~73; \citealt{stephenson_possible_1959}).
+Figure~\ref{fig:skychart} shows the overlap between the Kepler field
+and the cluster.  A previous clustering analysis of the Gaia
+data by \citet{kounkel_untangling_2019}
+reported that Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a
+member of the \cn.  Given the
+previous statistical validation of the close-in mini-Neptune Kepler
+1627b
 \citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
@@
 presence of a previously unreported binary companion, hereafter
-Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A (Section~\ref{sec:planet}).
-Based on a correlation between the planetary transit times and the local
-light curve slope, the orbit also seems to be prograde.  We
-conclude by highlighting broader implications for our ability to
-age-date a larger sample of planets (Section~\ref{sec:conc}).
+Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A
+(Section~\ref{sec:planet}).  Based on a correlation between the
+planetary transit times and the local light curve slope, the orbit
+also seems to be prograde.  We conclude by highlighting broader
+implications for our ability to age-date a larger sample of planets
+(Section~\ref{sec:conc}).
 
@@
     magnitude diagram of candidate \cn\ members, in addition to stars
-    in IC\,2602 ($\approx38$\,Myr), the Pleiades ($\approx 115$\,Myr;
-    \citealt{dahm_2015}), and the Gaia EDR3 Catalog of Nearby Stars
-    (gray background).  The zoomed right panel highlights the
-    pre-main-sequence.  The \cn\ and IC\,2602 are approximately the
-    same isochronal age.  {\it Bottom:} TESS and Kepler stellar
-    rotation period versus dereddened Gaia color, with the Pleiades
-    and Praesepe (650\,Myr) shown for reference
-    \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most candidate
-    \cn\ members are gyrochronally younger than the Pleiades; outliers
-    are probably field interlopers.
+    in IC\,2602 ($\approx38$\,Myr), the Pleiades ($\approx 115$\,Myr),
+    and the Gaia EDR3 Catalog of Nearby Stars (gray background).  The
+    zoomed right panel highlights the pre-main-sequence.  The \cn\ and
+    IC\,2602 are approximately the same isochronal age.  {\it Bottom:}
+    TESS and Kepler stellar rotation period versus dereddened Gaia
+    color, with the Pleiades and Praesepe (650\,Myr) shown for
+    reference \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most
+    candidate \cn\ members are gyrochronally younger than the Pleiades; outliers are
+    probably field interlopers.
    \label{fig:age}
@@
 The ages returned by this procedure depend on the ages assumed for
-each reference cluster.  We adopted a 115\,Myr age for the Pleiades,
-and a 16\,Myr age for UCL \citep{dahm_2015,pecaut_star_2016}.  The age
+each reference cluster.  We adopted a 115\,Myr age for the Pleiades \citep{dahm_2015},
+and a 16\,Myr age for UCL \citep{pecaut_star_2016}.  The age
 of IC\,2602 however is the most important ingredient, since it
@@
 IC\,2602 span 30\,Myr to 46\,Myr, with slightly older ages being
-preferred by the lithium-depletion-boundary (LDB) measurements and
+preferred by the lithium-depletion-boundary (LDB) measurements 
+\citep{dobbie_ic_2010,randich_gaiaeso_2018}
+and
 younger ages by the main-sequence turn-off
-\citep{stauffer_rotational_1997,dobbie_ic_2010,david_ages_2015,randich_gaiaeso_2018,bossini_age_2019,bouma_cluster_2020}.
+\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.
 If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would
@@
 on crowding, we queried the Gaia source catalog for stars within a
-$21\farcs0$ radius of the target star (1 TESS pixel).  We recorded the
-number of stars within the aperture with greater brightness than the
-target star, and with brightness within 1.25 and 2.5 TESS magnitudes
-of the target star.
-
-We cleaned the candidate TESS rotation period measurements through a
-combination of automated and manual steps.  As a matter of scope, we
-restricted our attention to the \nkinwithtess\ stars discussed in
-Appendix~\ref{app:kinematicselection} in the spatial and kinematic
-proximity of Kepler\,1627.  To avoid blending, we also only focused on
-stars for which no companions were known with a brightness exceeding
-one-tenth of the target star in a $21\farcs0$ radius.  There were
-\nkinwithtessandcrowding\ stars that met these requirements, and that
-had TESS data available.  For plotting purposes we then imposed a
+$21\farcs0$ radius of the target star (a radius of 1 TESS pixel).
+Within this radius, we recorded the number of stars with greater
+brightness than the target star, and with brightness within 1.25 and
+2.5 TESS magnitudes of the target star.
+
+We then cleaned the candidate TESS rotation period measurements
+through a combination of automated and manual steps.  As a matter of
+scope, we restricted our attention to the \nkinwithtess\ stars
+discussed in Appendix~\ref{app:kinematicselection} in the spatial and
+kinematic proximity of Kepler\,1627.  To avoid blending, we also only
+focused on stars for which no companions were known with a brightness
+exceeding one-tenth of the target star in a $21\farcs0$ radius.  There
+were \nkinwithtessandcrowding\ stars that met these requirements, and
+that had TESS data available.  For plotting purposes we then imposed a
 selection based on the strength of the signal itself: we required the
@@
 
-The second line of confirmatory evidence comes by analyzing the
+The second line of confirming evidence comes by analyzing the
 individual Kepler transit times. We isolated each of the 144 observed
@@
 A third supporting line of evidence for the planetary interpretation
-also exists, though we consider it less definitive than the stellar
+exists, though we consider it less definitive than the stellar
 density and TTV-local slope correlation.  We observed a transit of
@@
 connection of the star to the \cn\ using Gaia kinematics, and then by
-using the Gaia HR diagram and TESS stellar rotation periods to verify the
-cluster's existence.
-Table~\ref{tab:v06} should enable similar cross-matches to already
-known planets, as well as forthcoming planets from TESS
-\citep{guerrero_tess_2021}.  An alternative
-approach for age-dating more planets could be to identify kinematic
+using the Gaia HR diagram and TESS stellar rotation periods to verify
+the cluster's existence.  Table~\ref{tab:v06} enables similar
+cross-matches for both known and forthcoming exoplanet
+systems \citep[{\it e.g.},][]{guerrero_tess_2021}.  The entry shown in
+the header of Table~\ref{tab:v06}, HD 150706b, is an example: the
+host star appears to be a member of the Ursa Major moving group
+($\approx 400\,{\rm Myr}$; \citealt{mann_tess_2020}).  Confirming
+such kinematic associations with independent age indicators is
+essential, because the false positive rates are uncharted.
+Alternatively, one could instead opt to identify new kinematic
 associations around known exoplanet host stars using positions and
 tangential velocities from Gaia, and to then verify these associations
-with ancillary rotation periods and spectroscopy \citep[{\it
+with stellar rotation periods and spectroscopy \citep[{\it
 e.g.},][]{tofflemire_tess_2021}.  
+Each path seems likely to expand the census of planets with precisely
+measured ages over the coming years.
 
@@
     20$, reported to be in the group by
-    \citet{kounkel_untangling_2019}.  We focus on stars in a small
-    region (black points) in the kinematic vicinity of Kepler\,1627
-    (yellow star).  The other candidate cluster members (gray points)
-    may or may not share the ages of the selected kinematic group.
-    The location of the Sun is ($\odot$) is shown.
+    \citet{kounkel_untangling_2019}.  We focus on stars (black points)
+    in the spatial and kinematic vicinity of Kepler\,1627 (yellow
+    star).  The other candidate cluster members (gray points) may or
+    may not share the ages of the selected kinematic group.  The
+    location of the Sun is ($\odot$) is shown.
 		\label{fig:XYZvtang}
@@
 is present during the transit: the measured flux is systematically
-high during the first half, and low in the second half.  
-To determine
-whether this effect was caused by our stellar variability model, we 
-explored an alternative approach in which we isolated each transit
-window and locally fitted out polynomial trends, and then binned all
-the observed transits; the asymmetry was still present, at a
-comparable amplitude.
-An important note is that even though this asymmetric feature is
-within the plotted 2-$\sigma$ model uncertainties, this is within a
-model that has a jitter term that grows to account for otherwise unmodelled 
-white noise in the flux.
-The significance of the asymmetry is therefore better assessed in
-comparison against the intrinsic out-of-transit scatter, rather than the
-model uncertainties.
+high during the first half, and low in the second half.  The
+semi-amplitude of the deviation is $\approx50\,{\rm ppm}$, which
+represents a $\approx 3\%$ distortion of the transit depth
+($\approx1700\,{\rm ppm}$).  To determine whether this effect was
+caused by our stellar variability model, we explored an alternative
+approach in which we isolated each transit window and locally fitted
+out polynomial trends, and then binned all the observed transits; the
+asymmetry was still present, at a comparable amplitude.  Note that
+even though this asymmetric feature is within the 2-$\sigma$ model
+uncertainties, this model itself has a jitter term that grows to
+account for otherwise unmodelled white noise in the flux.  The
+significance of the asymmetry is therefore best assessed in comparison
+against the intrinsic out-of-transit scatter, not the model
+uncertainties.
 

### 2021-09-28 paper/ms.tex
update ms

@@
   observed by the Kepler space telescope because it is nearby
-  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Here we show
+  ($d=329\,{\rm pc}$) and it resembles the Sun.  Here we show
   using Gaia kinematics, TESS stellar rotation periods, and
   spectroscopic lithium abundances that Kepler\,1627 is a member of
-  the \clusterage\ old $\delta$~Lyr cluster.  To our knowledge,
-  Kepler\,1627Ab is the youngest planet with a precise age yet found
-  by the main Kepler mission.  The Kepler photometry shows two notable
-  anomalies: the average Kepler transit profile appears to be
-  asymmetric, and the individual transits show a correlation between
-  the transit time and the local light curve slope.
+  the \clusterage\ old $\delta$~Lyr cluster.  To our knowledge, this makes
+  Kepler\,1627Ab the youngest planet with a precise age yet found
+  by the main Kepler mission.
+  The Kepler photometry shows two 
+  peculiarities: the average transit profile appears to be
+  asymmetric, and the individual transit times are correlated with the local light curve slope.
   % The former could be explained by a dusty atmospheric outflow; the
@@
   % planet on a prograde orbit.  
-  We discuss plausible explanations for the light curve anomalies, and
-  also highlight broader implications for our ability to expand the census
+  We discuss possible explanations for these anomalies, and
+  highlight broader implications for our ability to expand the census
   of age-dated planets.
@@
 youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
-with precise ages include K2-33, DS~Tuc, HIP~67522,
-TOI~837, the two-planet AU~Mic system, and the four-planet V1298~Tau
-system
-\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020,martioli_aumicbc_2021}.
-\pn\ is one of the smallest planets in this sample ($3.4\pm
+with precise ages include K2-33
+\citep{Mann_K2_33b_2016,David_et_al_2017}, DS~Tuc
+\citep{benatti_possibly_2019,newton_tess_2019} , HIP~67522
+\citep{rizzuto_tess_2020}, TOI~837
+\citep{bouma_cluster_2020}, the two-planet AU~Mic system
+\citep{plavchan_planet_2020,martioli_aumicbc_2021} and the four-planet
+V1298~Tau system \citep{david_four_2019}.  \pn\ is one of the smallest
+planets in this sample ($3.4\pm
 0.2\,R_\oplus$), which could be linked to the selection effects

### 2021-09-29 paper/ms.tex
update ms
Moved appendix material around, per LAH suggestion.

Co-authored-by: Lynne Hillenbrand <lah@astro.caltech.edu>

@@
   %Kepler 1627 Ab:
+  %A Mini-Neptune in the Kepler Field Only 38 Million Years Old
   A 38 Million Year Old Mini-Neptune in the Kepler Field
@@
   peculiarities: the average transit profile appears to be
-  asymmetric, and the individual transit times are correlated with the local light curve slope.
-  % The former could be explained by a dusty atmospheric outflow; the
-  % latter could be explained by unresolved starspot crossings of a
-  % planet on a prograde orbit.  
-  We discuss possible explanations for these anomalies, and
-  highlight broader implications for our ability to expand the census
-  of age-dated planets.
+  asymmetric, and the individual transit times are correlated with the
+  local light curve slope.
+  We discuss possible explanations for each anomaly.
+  More widely applicable though is the fact that the $\delta$~Lyr
+  cluster is one of $\sim$10$^3$ clusters whose properties have been
+  clarified by Gaia.  Many other exoplanet hosts are candidate
+  members of these clusters based on kinematics; their ages can
+  be confirmed by using the TESS full frame images to
+  measure stellar rotation periods.
 \end{abstract}
@@
 
+Matching our young star list against stars observed by Kepler revealed that
+%two discoveries.  The first, to be discussed in an upcoming analysis
+%by J.~Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
+%open cluster Theia~520 (UBC~1).  At least six Kepler planets are known
+%in the cluster from the Kepler-52 and Kepler-968 systems
+%\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
+%discovery, and the focus of this work, is
+Kepler observed a portion of the \cn\ (Stephenson-1;
+Theia~73).
+More specifically, a clustering analysis of the Gaia
+data by \citet{kounkel_untangling_2019}
+reported that Kepler~1627 (KIC 6184894; KOI 5245; TIC 120105470) is a
+\cn\ member.  Given the
+previous statistical validation of the close-in mini-Neptune Kepler
+1627b
+\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
+we begin by examining the properties of the cluster more
+closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
+cluster is \clusterage\ old, and 
+in 
+Section~\ref{sec:stars} show
+that \sn\ is both a binary and also a member of the cluster.
+Focusing on the planet (Section~\ref{sec:planet}), we confirm that despite the
+presence of the previously unreported M2.5V companion, hereafter
+Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A.  Based on a correlation between the
+planetary transit times and the local light curve slope, the orbit
+may also be prograde.  We conclude by highlighting broader
+implications for our ability to age-date a larger sample of planets
+(Section~\ref{sec:conc}).
+
+\section{The Cluster}
+\label{sec:cluster}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=1\textwidth]{f6.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Galactic positions and tangential velocities of stars in the
+			$\delta$\,Lyr cluster.} Points are reported cluster members from
+		\citet{kounkel_untangling_2019}.  
+		The tangential velocities relative to
+		\sn\ (lower-right) are computed assuming that every star has the same
+		three-dimensional spatial velocity as \sn.
+		Our analysis uses stars (black points)
+		in the spatial and kinematic vicinity of Kepler\,1627 (yellow
+		star).  The question of whether the other candidate cluster members
+		(gray points) are part of the cluster is outside our scope.  The
+		location of the Sun is ($\odot$) is shown.
+		\label{fig:XYZvtang}
+	}
+\end{figure*}
+
+To measure the age of the $\delta$ Lyr cluster, we first selected
+a set of candidate cluster members
+(Section~\ref{sec:kinematicselection}), and then analyzed these stars using a
+combination of the isochronal and gyrochronal techniques
+(Section~\ref{sec:clusterage}).
+
+\subsection{Selecting Cluster Members}
+\label{sec:kinematicselection}
+
+\citet{kounkel_untangling_2019} applied an unsupervised clustering
+algorithm to Gaia DR2 on-sky positions, proper motions, and
+parallaxes for stars within the nearest kiloparsec.  For the \cn\
+(Theia 73), they reported \noriginal\ candidate members.  We
+matched these stars against the latest Gaia EDR3 observations using
+the \texttt{dr2\_neighbourhood} table from the ESA archive\footnote{See
+\url{https://gea.esac.esa.int/archive/documentation/GEDR3/Gaia_archive/chap_datamodel/sec_dm_auxiliary_tables/ssec_dm_dr2_neighbourhood.html}.},
+and took the closest proper motion and epoch-corrected angular
+distance as the presumed match
+\citep{gaia_collaboration_2021_edr3}.  In Figure~\ref{fig:XYZvtang},
+have shown galactic positions only for the stars with
+parallax signal-to-noise exceeding 20.  The reported cluster members
+(gray and black points) extend over a much larger volume than the
+cluster previously identified by \citet{stephenson_possible_1959} and
+later corroborated by \citet{eggen_photometric_1968}.  While the
+non-uniform ``clumps'' of stars might comprise a {\it bona fide}
+cluster of identically-aged stars, they could also be significantly
+contaminated by field stars.  We therefore considered stars only in
+the immediate kinematic and spatial vicinity of \sn\ as candidate
+cluster members.  We performed the selection cuts manually, by
+drawing lassos with the interactive \texttt{glue} visualization tool
+\citep{beaumont_2014_13866} in the four projections shown in
+Figure~\ref{fig:XYZvtang}.  The overlap between the Kepler
+field and the resulting candidate cluster members is shown in
+Figure~\ref{fig:skychart}.  While this method will
+include some field interlopers in the ``cluster star'' sample, and
+vice-versa, it should suffice for our aim of verifying the existence
+of the cluster in the vicinity of \sn.
+
 \begin{figure}[t]
@@
 		\leavevmode
-		\includegraphics[width=0.49\textwidth]{f1.pdf}
+		\includegraphics[width=0.47\textwidth]{f1.pdf}
 	\end{center}
@@
 
-Matching our young star list against stars observed by Kepler yielded
-%two discoveries.  The first, to be discussed in an upcoming analysis
-%by J.~Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
-%open cluster Theia~520 (UBC~1).  At least six Kepler planets are known
-%in the cluster from the Kepler-52 and Kepler-968 systems
-%\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
-%discovery, and the focus of this work, is
-the discovery that Kepler observed a portion of the \cn\ (Stephenson-1; Theia~73; \citealt{stephenson_possible_1959}).
-Figure~\ref{fig:skychart} shows the overlap between the Kepler field
-and the cluster.  A previous clustering analysis of the Gaia
-data by \citet{kounkel_untangling_2019}
-reported that Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a
-member of the \cn.  Given the
-previous statistical validation of the close-in mini-Neptune Kepler
-1627b
-\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
-we examined the age of the star and its putative host cluster more
-closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
-cluster is \clusterage\ old, and that \sn\ is indeed a member of the
-cluster.  Focusing on the planet itself, we confirm that despite the
-presence of a previously unreported binary companion, hereafter
-Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A
-(Section~\ref{sec:planet}).  Based on a correlation between the
-planetary transit times and the local light curve slope, the orbit
-also seems to be prograde.  We conclude by highlighting broader
-implications for our ability to age-date a larger sample of planets
-(Section~\ref{sec:conc}).
-
-\section{The Cluster}
-\label{sec:cluster}
+\subsection{The Cluster's Age}
+\label{sec:clusterage}
 
@@
 
-To assess the age of the cluster and our star of interest, we used a
-combination of isochronal, gyrochronal, and lithium-based
-age-dating.
-
-\subsection{Color-Absolute Magnitude Diagram}
+\subsubsection{Color-Absolute Magnitude Diagram}
 \label{sec:camd}
 
-We measured the isochrone age through an empirical approach.  The left
+We measured the isochrone age using an empirical approach.  The left
 panel of Figure~\ref{fig:age} shows the color-absolute magnitude
-diagram (CAMD) of stars in the \cn, IC\,2602, the Pleiades, and the
-field.  The candidate \cn\ members shown are selected based on their
-positions and kinematics (see Appendix~\ref{app:kinematicselection}).
-Those from the Pleiades and IC\,2602 are adopted from
+diagram (CAMD) of candidate \cn\ members, IC\,2602, the Pleiades, and the
+field.  
+The stars from the Pleiades and IC\,2602 were adopted from
 \citet{cantatgaudin_gaia_2018}, and the field stars are from the Gaia
 EDR3 Catalog of Nearby Stars \citep{gaia_gcns_2021}.  We cleaned these
-following the data filtering criteria discussed in Appendix~B of
-\citet{GaiaCollaboration2018}, except that we weakened the parallax
+following the data filtering criteria from
+\citet[][Appendix B]{GaiaCollaboration2018}, except that we weakened the parallax
 precision requirement to $\varpi/\sigma_\varpi>5$.  These filters were
@@
 \citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}, 2021/09/25},
-and then applied the extinction coefficients $k_X\equiv A_X/A_0$
-computed by \citet{GaiaCollaboration2018} and assumed that $A_0 = 3.1
+and applied the extinction coefficients $k_X\equiv A_X/A_0$
+computed by \citet{GaiaCollaboration2018} assuming that $A_0 = 3.1
 E(B-V)$.  For IC\,2602, the Pleiades, and the \cn, this procedure
@@
 main-sequence turn-off in the \cn\ could help determine a more precise
-and accurate age for the cluster, however these are beyond our scope.
-
-\subsection{Stellar Rotation Periods}
+and accurate age for the cluster, and are left for future work.
+
+\subsubsection{Stellar Rotation Periods}
 
@@
 scope, we restricted our attention to the \nkinwithtess\ stars
-discussed in Appendix~\ref{app:kinematicselection} in the spatial and
+discussed in Section~\ref{sec:kinematicselection} in the spatial and
 kinematic proximity of Kepler\,1627.  To avoid blending, we also only
@@
 \nkindefaultcleaning\ stars.  The majority of these stars fall below
-the ``slow sequence'' of the Pleiades, consistent with a
-gyrochronological age for the \cn\ below 100\,Myr.  Approximately 10
-stars appear as outliers above the ``slow sequence''.  Assuming that
-they are all false positives ({\it i.e.}, field interlopers), the
-rotation period detection fraction would be
-$\nrotgood/\nkinwithtessandcrowding \approx \nfracprot\%$.  The
-rotation period distributions of other 30\,Myr to 50\,Myr clusters
-({\it e.g.}, IC\,2602 and IC\,2391) are indistinguishable
-\citep{douglas_stephanie_t_2021_5131306}.  While binarity is known to
-affect the density of stars that fall beneath the slow sequence
-\citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}, it is
-not expected to change the central conclusion regarding the cluster's
+the ``slow sequence'' of the Pleiades, consistent with a gyrochronal
+age for the \cn\ below 100\,Myr.  In fact, the rotation-color
+distributions of other 30\,Myr to 50\,Myr clusters ({\it e.g.},
+IC\,2602 and IC\,2391) are indistinguishable
+\citep{douglas_stephanie_t_2021_5131306}. 
+Approximately 10 of the \cn\ stars appear as outliers above
+the ``slow sequence''.  Assuming that they are all false positives
+({\it i.e.}, field interlopers), our rotation period detection
+fraction would be $\nrotgood/\nkinwithtessandcrowding \approx
+\nfracprot\%$.   
+The other stars are likely to be field contaminants.
+A final possible confounding factor -- binarity -- is known to affect
+the density of stars that fall beneath the slow sequence
+\citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}.  We do
+not expect it to change the central conclusion regarding the cluster's
 age.
-%%VERSION1
-%Based on an analysis of photometric binaries and stars with high RUWE,
-%we found evidence for similar effects to those previously noted in the
-%literature.  However, since the purpose of our rotation period
-%analysis is to determine the age of the \cn, we do not dwell on this
-%topic.
-% %VERSION0
-% Appendix~\ref{app:rotationbinarity} discusses this topic, and finds
-% evidence for similar effects to those previously noted in the
-% literature.
-
-\subsection{Age of Kepler\,1627}
-
-To summarize, the age of the \cn\ based on the empirical
-pre-main-sequence interpolation method is \clusterage.  This method
-hinges on the age adopted for IC\,2602, since it is the closest
-reference cluster in the color-absolute magnitude diagram.
-Gyrochronologically, while most of the FGK dwarfs in the \cn\ have not
-yet converged to a slow sequence, their rotation periods are generally
-consistent with a $\approx40$\,Myr age. The most precise age for \sn\ and its
-planet are therefore identical to the cluster, based on the star's
-spatial and kinematic association with the cluster, and the assumption
-that the planet formed shortly after the star.
-
-Beyond the kinematic and rotational lines of evidence for the youth of
-\sn, there is a final consistency check: the abundance of
-Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet \citep[{\it
-e.g.},][]{soderblom_ages_2014}.  We acquired an iodine-free spectrum
-from Keck/HIRES on the night of 2021 March 26 using the standard setup
-and reduction techniques of the California Planet Survey
-\citep{howard_cps_2010}.  Following the equivalent width measurement
-procedure described by \citet{bouma_2021_ngc2516}, we find ${\rm
-EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   %Note that this value does not
-%correct for the Fe\,\textsc{I} blend at 6707.44\AA. 
-Given the effective temperature of \sn\ (Table~\ref{tab:starparams}),
-this measurement is in agreement with expectations for a $\approx40$\,Myr star
-({\it e.g.}, as measured in IC~2602 by
-\citealt{randich_gaiaeso_2018}).  It is also larger than any
-lithium equivalent widths measured by \citet{berger_identifying_2018}
-in their analysis of 1{,}301 Kepler-star spectra.
+
+\section{The Stars}
+\label{sec:stars}
+
+\subsection{Kepler\,1627A}
+\subsubsection{Age}
+Based on the spatial and kinematic association of \sn\ with the \cn,
+and the assumption that the planet formed shortly after the star, it
+seems likely that \sn\ is the same age as the cluster. There are
+two consistency checks on whether this is true: rotation and lithium.  Based on the Kepler
+light curve, the rotation period is $2.642\pm0.042$\,days, where the
+quoted uncertainty is based on the scatter in rotation periods
+measured from each individual Kepler quarter.  This is consistent with
+comparable cluster members (Figure~\ref{fig:age}).
+
+To infer the abundance of Li\,\textsc{I} from the 6708\,\AA\ doublet
+\citep[{\it e.g.},][]{soderblom_ages_2014}, we acquired an iodine-free
+spectrum from Keck/HIRES on the night of 2021 March 26 using the
+standard setup and reduction techniques of the California Planet
+Survey \citep{howard_cps_2010}.  Following the equivalent width
+measurement procedure described by \citet{bouma_2021_ngc2516}, we find
+${\rm EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   This value does not
+correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Nonetheless,
+given the stellar effective temperature (Table~\ref{tab:starparams}),
+this measurement is in agreement with expectations for a
+$\approx40$\,Myr star ({\it e.g.}, as measured in IC~2602 by
+\citealt{randich_gaiaeso_2018}).  It is also larger than any lithium
+equivalent widths measured by \citet{berger_identifying_2018} in their
+analysis of 1{,}301 Kepler-star spectra.
+
+\subsubsection{Stellar Properties}
+
+The adopted stellar parameters are listed in
+Table~\ref{tab:starparams}.  The stellar mass, radius, and effective
+temperature are found by interpolating against a 38$\,$Myr MIST
+isochrone \citep{choi_mesa_2016}.  
+The statistical uncertainties are propagated from the absolute
+magnitude (mostly originating from the the parallax uncertainty) and the color; the systematic
+uncertainties are taken to be the difference between the PARSEC
+\citep{bressan_parsec_2012} and MIST isochrones.  Reported
+uncertainties are a quadrature sum of the statistical and systematic
+components.
+As a consistency check, 
+we analyzed the aforementioned Keck/HIRES spectrum from
+the night of 2021 March 26 using a combination of
+\texttt{SpecMatch-Emp} for stellar properties, and
+\texttt{SpecMatch-Synth} for $v\sin i$
+\citep{yee_SM_2017}.
+This procedure yielded
+$T_{\rm eff}=5498\pm100\,{\rm K}$,
+$\log g=4.6\pm0.1$, $[{\rm Fe/H}]=0.15\pm0.10$ 
+from \texttt{SpecMatch-Emp}, and $v\sin i = 18.9\pm1.0$ from
+\texttt{SpecMatch-Synth}.
+These values are within the $1$-$\sigma$ uncertainties of our adopted values from
+the isochrone interpolation.
+
+\subsection{Kepler\,1627B}
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f10a.pdf}
+			\includegraphics[width=0.48\textwidth]{f10b.pdf}
+		}
+	\end{center}
+	\vspace{-0.5cm}
+	\caption{
+    {\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
+    from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2.5V
+    companion at $\rho \approx 0\farcs16$, which corresponds to a
+    projected separation of $53\pm4$\,AU.  The inset shows a cutout of
+    the stacked NIRC2 image (North is up, East is left, scale is set by the
+    separation of the binary).  The lines show 5-$\sigma$ contrast
+    limits for the `Alopeke filters, and 6-$\sigma$ contrast limits
+    for NIRC2 outside of $0\farcs15$. {\it Right:} Gaia
+    EDR3 renormalized unit weight error (RUWE) point estimates for
+    candidate $\delta$\,Lyr cluster
+    members from Figure~\ref{fig:XYZvtang}.  Since other members of the
+    cluster with similar brightnesses have comparable degrees of
+    photometric variability, the high RUWE independently suggests that Kepler\,1627
+    is a binary. 
+    \label{fig:kep1627binary}
+	}
+\end{figure*}
+
+We first noted the presence of a close neighbor in the \sn\ system on 2015 July 22 when
+we acquired adaptive optics imaging using the NIRC2 imager on
+Keck-II.
+We used the narrow camera (FOV = 10.2\arcsec) to obtain 8
+images in the $K'$ filter ($\lambda = 2.12\,\mu$m) with a total
+exposure time of 160\,s. We analyzed these data following
+\citet{kraus_impact_2016}, which entailed using PSF-fitting to measure the separation,
+position angle, and contrast of the candidate companion. 
+The best-fitting empirical PSF template was identified from among the
+near-contemporaneous observations of single stars in the same filter.
+The mean values inferred from the 8 images are reported in
+Table~\ref{tab:starparams}.
+To estimate the detection limits, we analyzed the residuals after subtracting
+the empirical PSF template. Within each residual
+image, the flux was measured through 40\,mas apertures centered on
+every pixel, and then the noise as a function of radius was estimated
+from the RMS within concentric rings. Finally, the detection limits
+were estimated from the strehl-weighted sum of the detection
+significances in the image stack, and we adopted the $6$-$\sigma$
+threshold as the detection limit for ruling out additional companions.
+
+We also observed \sn\ on Gemini-North using the `Alopeke speckle
+imager on 2021 June 24.  `Alopeke is a dual-channel speckle
+interferometer that uses narrow-band filters centered at 0.83\,$\mu$m
+and $0.56\,\mu$m.  We acquired three sets of $1000\times 60$$\,$msec
+exposures during good seeing (0.45$''$), and used the autocorrelation
+function of these images to reconstruct a single image and 5-$\sigma$
+detection limits (see \citealt{howell_speckle_2011}).  This procedure
+yielded a detection of the companion in the 0.83\,$\mu$m notch filter,
+but not the $0.56\,\mu$m filter.  The measured projected separation
+and magnitude difference are given in Table~\ref{tab:starparams}.
+
+Figure~\ref{fig:kep1627binary} summarizes the results of the
+high-resolution imaging.  The Gaia EDR3 parallax for the primary
+implies a projected separation of $53 \pm 4$\,AU, assuming the
+companion is bound.  Although the companion is unresolved in the Gaia
+source catalog (there are no comoving, codistant candidate companions
+brighter than $G < 20.5$ mag within $\rho < 120\arcsec$), its
+existence was also suggested by the primary star's large renormalized unit weight error (RUWE),
+relative to other members of the \cn.  Based on the apparent
+separation, the binary orbital period is of order hundreds of years.
+The large RUWE is therefore more likely to be caused by a PSF-mismatch
+skewing the Gaia centroiding during successive scans, rather than true
+astrometric motion.  Regardless, given the low geometric probability that a companion imaged at
+$\rho \approx 0\farcs16$ is a chance line-of-sight companion, we
+proceed under the assumption that the companion is bound, and that
+Kepler\,1627 is a binary.  
+Given the distance and age, the models of
+\citet{baraffe_new_2015} imply a companion mass of $M_{\rm B} \approx
+0.33 M_{\odot}$ and companion temperature of $T_{\rm eff,B} \approx
+3450$ K.  The corresponding spectral type is roughly M2.5V
+\citep{pecaut_mamajek_2013}.
+% \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
+These models combined with the NIRC2 contrast limits similarly imply
+physical limits on tertiary companions of $M_{\rm ter} < 50
+M_{\rm Jup}$ at $\rho = 50$\,AU, $M_{\rm ter} < 20 M_{\rm Jup}$ at $\rho =
+100$\,AU, and $M_{\rm ter} < 10 M_{\rm Jup}$ at $\rho = 330$\,AU.
 
@@
     points bin these to 20 minute intervals, and have a representative
-    1-$\sigma$ error bar in the lower right of each panel.  The
-    asymmetric residual during transit appears to be robust to
-    different detrending methods.
+    1-$\sigma$ error bar in the lower right of each panel.  The asymmetric
+    residual during transit is significantly larger than the out-of-transit scatter.
     \label{fig:lc}
@@
 \end{figure*}
+
+\subsection{Kepler Light Curve}
+
+The Kepler space telescope observed \sn\ at a 30-minute cadence from
+2009 May 2 until 2013 April 8.  Data gaps during quarters 4, 9, and 13
+led to an average duty cycle over the 3.9~year interval of 78\%.  \sn\
+was also observed at 1-minute cadence from 2012 Oct 5 until 2013 Jan
+11.  The top panel of Figure~\ref{fig:lc} shows a portion of the 30-minute
+cadence \texttt{PDCSAP} light curve.  Nonastrophysical variability has
+been removed using the methods discussed by
+\citet{smith_kepler_PDC_2017}; the default optimal aperture was
+assumed \citep{smith_finding_2016}. 
+Cadences with non-zero quality flags (9\% of the data) have been omitted.
+ The resulting photometry is dominated by
+a quasi-periodic starspot signal with a peak-to-peak amplitude that
+varies between 2\% and 8\%.  
+Previous analyses have identified and
+characterized the hidden transit signal
+\citep{2012ApJS..199...24T,thompson_planetary_2018}, validated its
+planetary nature \citep{morton_false_2016}, and even searched the
+system for transit timing variations \citep{holczer_transit_2016}.
+Nonetheless, since the cluster membership provides us
+with more precise stellar parameters than those previously available,
+we opted to reanalyze the
+light curve.
+
+We fitted the Kepler long cadence time series with a model that
+simultaneously included the planetary transit and the stellar
+variability.  The stellar variability was modeled with the
+\texttt{RotationTerm} Gaussian Process kernel in \texttt{exoplanet}
+\citep{exoplanet:exoplanet}.  This kernel assumes that the variability
+is generated by a mixture of two damped simple harmonic oscillators
+with characteristic frequencies set by 1/$P_{\rm rot}$ and its first
+harmonic.  We additionally included a jitter term to inflate
+the flux uncertainties in a manner that accounted for otherwise
+unmodeled excess white noise, and let the eccentricity float.  For
+the limb-darkening, we assumed a quadratic law, and sampled using the
+uninformative prior suggested by \citet{exoplanet:kipping13}.
+
+Our model therefore included 10 free parameters for the transit ($\{P,
+t_0, R_{\rm p}/R_\star, b, u_1 ,u_2 ,R_\star, \log g, e, \omega \}$),
+2 free parameters for the light curve normalization and a white noise 
+jitter ($\{\langle f \rangle, \sigma_f \}$), and 5
+hyperparameters for the GP ($\{\sigma_{\mathrm{rot}},
+P_{\mathrm{rot}}, Q_0, \mathrm{d}Q, f \}$).  We also considered
+including an additive \texttt{SHOTerm} kernel to account for
+stochastic noise, but found that this did not significantly affect the
+results, and so opted for the simpler GP kernel.  We fitted the models
+using \texttt{PyMC3} \citep{salvatier_2016_PyMC3,exoplanet:theano},
+and accounted for the finite integration time of each exposure in the
+numerical integration when evaluating the model light curve
+\citep[see][]{kipping_binning_2010}.  We assumed a Gaussian
+likelihood, and after initializing each model with the parameters of
+the maximum {\it a posteriori} model, we sampled using
+\texttt{PyMC3}'s gradient-based No-U-Turn Sampler
+\citep{hoffman_no-u-turn_2014} in the bases indicated in
+Table~\ref{tab:posterior}.  We used $\hat{R}$ as our convergence
+diagnostic \citep{gelman_inference_1992}.
+
+Figure~\ref{fig:lc} shows the resulting best-fit model in orange (top) and
+purple (bottom).
+The model parameters and their uncertainties, given in Table~\ref{tab:posterior}, are
+broadly consistent with a mini-Neptune sized planet ($3.36\pm0.18\,R_\oplus$) on a close-in circular\footnote{ Our transit fitting
+	yields $e<0.48$ at 2-$\sigma$; the constraints on the eccentricity are
+	not particularly strong.} orbit around a G8V host star ($0.88 \pm 0.02
+R_\odot$).  Our best-fit planet size is smaller than that
+previously reported by \citet{morton_false_2016} and
+\citet{berger_identifying_2018}, at 0.7-$\sigma$ and 1.2-$\sigma$
+confidence respectively.  The former is explained by differing
+assumptions about the stellar radius; the latter appears to be due to
+differences in the measured transit depth, which could be linked to
+the different methods used to account for the stellar rotation signal.
+
+The transit fit however is not perfect: the lower panels of
+Figure~\ref{fig:lc} show an asymmetric
+residual in the data relative to the model: the measured flux is systematically
+high during the first half of transit, and low in the second half.  The
+semi-amplitude of this deviation is $\approx50\,{\rm ppm}$, which
+represents a $\approx 3\%$ distortion of the transit depth
+($\approx1700\,{\rm ppm}$).  
+Note that
+while this asymmetry is within the 2-$\sigma$ model
+uncertainties, the model has a jitter term that grows to
+account for otherwise unmodeled white noise in the flux.  The
+significance of the asymmetry is therefore best assessed in comparison
+against the intrinsic out-of-transit scatter in the data, not the model
+uncertainties.
+
+To determine whether the asymmetry is a systematic caused by our stellar variability model,
+we explored an alternative
+approach in which we isolated each transit window and locally fitted
+out polynomial trends, and then binned all the observed transits; the
+asymmetry was still present at a comparable amplitude.  
+Appendix~\ref{app:asymmetry} includes a deeper analysis, and
+finds that the asymmetry also seems to be robust to different methods of
+data binning.  Plausible astrophysical explanations are given in
+Section~\ref{sec:conc}.
+
+\subsection{Planet Confirmation}
 
@@
 		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f4a.pdf}
-		}
-
-		\vspace{-0.1cm}
-		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f4b.pdf}
-		}
+		\includegraphics[width=0.47\textwidth]{f4a.pdf}
 	\end{center}
@@
 		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
-    {\it Top:} 
 		The time of each Kepler transit was measured, along with the local
@@
 		$\sim$60\,ppt per half rotation period ($\sim$1.3\,days).
-    {\it Bottom:}
-    Radial velocities acquired using Keck/HIRES during the transit of
-    2021 Aug 7, observed simultaneously with MuSCAT3.  Shaded bands
+		\label{fig:corr}
+	}
+\end{figure}
+
+% rotation period from plot_rotation_period_windowslider.py
+
+If the \pn\ transit signal is created by a genuine planet,  then to
+our knowledge it would be the youngest planet yet found by the main
+Kepler mission.\footnote{The re-purposed K2 mission however has found
+two younger systems containing five planets: K2-33b ($9\pm1\,{\rm
+Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and V1298 Tau
+($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}   
+Could the transit be produced by anything other than a planet
+orbiting this near-solar analog?  \citet{morton_false_2016} validated
+the planet based on the transit shape, arguing that the most probable
+false positive scenario was that of a background eclipsing binary,
+which had a model-dependent probability of $\approx10^{-5}$.  However,
+this calculation was performed without knowledge of the low-mass stellar companion ($M_{\rm B}\approx
+0.33\,M_\odot$).  
+Validated planets have also previously been refuted \citep[{\it e.g.},][]{shporer_three_2017}.
+We therefore reassessed
+false positive scenarios in some detail.
+
+% (0.7-2.0\% V-band to G-band dmag difference from EEM table; NOTE
+% could determine using model spectra)
+As an initial plausibility check,
+Kepler\,1627B contributes 1\% to 2\% of the total flux observed in
+the Kepler aperture.  
+For the sake of argument, assume the former value.
+The observed transit has a depth of $\approx$0.17\%.  A
+17\% deep eclipse of the secondary star would therefore be needed to
+produce a signal with the appropriate depth.  The shape of the
+transit signal however requires the impact parameter to be below 0.74
+(Table~\ref{tab:posterior}); the tertiary transiting the secondary would
+therefore need to be non-grazing with $R_3/R_2 \approx 0.4$.
+%Assuming a $\approx 0.46R_\odot$ radius of the imaged secondary
+%from the \citet{baraffe_new_2015} models, this
+%would imply a tertiary stellar radius of $\approx 0.2R_\odot$.
+This
+yields a contradiction:  this scenario would require an ingress and
+egress phase that each span $\approx$40\% of the transit duration
+($\approx 65\,{\rm minutes}$).  The actual measured ingress and egress
+duration is $\approx 15\,{\rm minutes}$, 4.4$\times$ shorter.  The
+combination of Kepler 1627B's brightness, the transit depth, and the ingress
+duration therefore disfavor the scenario that Kepler 1627B might host the
+transit signal.
+
+% Baraffe+15: M_B 0.33 Mdot
+% Rstar = 0.455 --> 0.465 Rsun
+%
+
+% At 30 Myr, a 0.44\,$M_\odot$ solar-metallicity dwarf is $\approx 26\%$
+% larger than when it is fully contracted on the main sequence
+% (0.40\,$R_\odot$ vs{.} 0.50\,$R_\odot$; ).
+
+There are two other new lines of evidence that confirm the planetary
+interpretation.  First, the transit duration and the orbital period
+are inconsistent with an eclipsing body around the M-dwarf companion:
+we find $\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
+theoretically expected density for the companion is $\approx 4.6\,{\rm
+g\,cm}^{-3}$ (Table~\ref{tab:posterior}; \citealt{choi_mesa_2016}).
+The transit duration is therefore too long to be explained by a star
+eclipsing the M dwarf secondary at $10$-$\sigma$.
+
+%NOTE to the arxiv readers:, the $\approx$100 days of short-cadence
+%Kepler observations also mitigate the cadence-induced smearing, and
+%enable a more precise measurement of the transit impact parameter
+%\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
+%(2-$\sigma$) from the 1-minute data, compared to $<0.76$ (2-$\sigma$)
+%from the 30-minute data.  This implies an even stronger constraint on
+%the transit shape in an analysis of the false positive probability.
+
+The second line of confirming evidence comes by analyzing the
+individual Kepler transit times. We isolated each of the 144 observed
+transits to within $\pm4.5$\,hr of each transit, and fitted each
+window with both {\it i)} a local second-order polynomial and transit,
+and {\it ii)} a local linear trend.  We let the mid-time of each
+transit float, and then calculated the residual between the measured
+mid-time and that of a periodic orbit.  This residual, the transit
+timing variation (TTV), is plotted in Figure~\ref{fig:corr}
+against the local linear slope.  A significant correlation of $-0.0564
+\pm 0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
+Objects of Interest have shown this correlation
+\citep{holczer_time_2015}, which is most readily interpreted as a TTV
+induced by unresolved starspot crossings \citep{mazeh_time_2015}.
+This is only possible if the planet transits the primary star, which
+excludes a background eclipsing binary scenario, and therefore
+confirms that \pn\ is a planet.  It also suggests that the planet's
+orbit is prograde.  The latter point assumes that the dominant
+photometric variability is induced by dark spots, and not bright
+faculae.  Given the observed transition of Sun-like stellar
+variability from spot to faculae-dominated regimes between young and
+old ages, we expect this assumption to be reasonably secure
+\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
+
+\begin{figure}[t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.47\textwidth]{f4b.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Radial velocities acquired using Keck/HIRES during the transit of
+			2021 Aug 7, observed simultaneously with MuSCAT3.}  Shaded bands
 		show 2-$\sigma$ model uncertainties.  The photometric transit
 		depths are consistent across the {\it griz} bandpasses.  The
-    best-fitting model including the RM effect and a quadratic trend
-    in time is shown with the red line.  The radial velocity scatter
-    between exposures prevents a significant detection of the RM
+		best-fitting model in the top panel includes the RM effect and a quadratic trend
+		in time.  The radial velocity scatter
+		between exposures ($\sigma_{\rm RV}\approx50$\ms) prevented us from detecting the RM
 		effect.
-    \label{fig:obliquity}
+		\label{fig:ground}
 	}
 \end{figure}
+
+A third supporting line of evidence for the planetary interpretation
+exists, though we consider it less definitive than the stellar
+density and TTV-local slope correlation.  We observed a transit of
+\pn\ on the night of 2021 Aug 7 simultaneously with Keck/HIRES and
+MuSCAT3.  We scheduled the observations using the ephemeris of
+\citet{holczer_transit_2016}.  Although we did not detect the
+Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
+show the transit (Figure~\ref{fig:ground}).  Fitting
+the MuSCAT3 photometry with a model that lets the transit depths vary
+across each bandpass, we find {\it griz} depths consistent with the
+Kepler depth at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  The
+MuSCAT3 observations also suggest a transit duration $17.4\pm3.6\ {\rm
+minutes}$ shorter than the Kepler transits; further photometric
+follow-up could help confirm whether the transit duration is indeed
+changing.  For our RM analysis, the details are discussed in
+Appendix~\ref{app:rm}.  While the velocities are marginally more
+consistent with a prograde or polar orbit than a retrograde orbit, the
+spot-corrected exposure-to-exposure scatter ($\sigma_{\rm RV}\approx
+50$\ms) is larger than the expected velocity anomaly
+($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$).  We are therefore
+not in a position to claim a spectroscopic detection of the RM effect,
+nor to quantify the stellar obliquity.
+
+\section{Discussion \& Conclusions}
+\label{sec:conc}
+
 \begin{figure*}[tp]
@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f5.pdf}
+		\includegraphics[width=0.7\textwidth]{f5.pdf}
 	\end{center}
@@
 
-% rotation period from plot_rotation_period_windowslider.py
-
-If the \pn\ transit signal is created by a genuine planet,  then to
-our knowledge it would be the youngest planet yet found by the main
-Kepler mission.\footnote{The re-purposed K2 mission however has found
-two younger systems containing five planets: K2-33b ($9\pm1\,{\rm
-Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and V1298 Tau
-($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  The light curve
-and the phase-folded transit are shown in Figure~\ref{fig:lc}.  The
-photometry is dominated by a quasi-periodic starspot signal with a
-peak-to-peak amplitude that varies between 2\% and 8\%.  The signal
-repeats every $2.642\pm0.042$\,days, where the quoted uncertainty is
-based on the scatter in rotation periods measured from each individual
-Kepler quarter.  Our model for the starspot and transit signals is
-discussed in detail in Appendix~\ref{app:gptransit}, and the adopted
-stellar parameters are listed in Table~\ref{tab:starparams}.  The
-stellar mass, radius, and effective temperature are found by
-interpolating against a 38$\,$Myr MIST isochrone.  Their systematic
-uncertainties are taken to be the difference between the PARSEC and
-MIST isochrones, and the reported uncertainties are a quadrature sum
-of the statistical and systematic components.
-
-The transit fitting results, given in Table~\ref{tab:posterior}, are
-broadly consistent with a mini-Neptune sized planet ($3.4\pm
-0.2\,R_\oplus$) on a close-in circular\footnote{ Our transit fitting
-yields $e<0.48$ at 2-$\sigma$; the constraints on the eccentricity are
-not particularly strong.} orbit around a G8V host star ($0.88 \pm 0.02
-R_\odot$).  Our best-fit planet size is smaller than those
-previously reported by \citet{morton_false_2016} and
-\citet{berger_identifying_2018}, at 0.7-$\sigma$ and 1.2-$\sigma$
-confidence respectively.  The former is explained by differing
-assumptions about the stellar radius; the latter appears to be due to
-differences in the measured transit depth, which could be linked to
-the different methods used to remove the stellar rotation signal.
-
-Could the transit be produced by anything other than a planet
-orbiting this near-solar analog?  \citet{morton_false_2016} validated
-the planet based on the transit shape, arguing that the most probable
-false positive scenario was that of a background eclipsing binary,
-which had a model-dependent probability of $\approx10^{-5}$.  As we
-describe in Appendix~\ref{app:companionstar} however, Kepler\,1627A is
-accompanied by a low-mass stellar companion ($M_{\rm B}\approx
-0.33\,M_\odot$; $\rho\approx0\farcs16$).  We therefore reassessed
-false positive scenarios in some detail.
-
-% Baraffe+15: M_B 0.33 Mdot
-% Rstar = 0.455 --> 0.465 Rsun
-%
-
-% At 30 Myr, a 0.44\,$M_\odot$ solar-metallicity dwarf is $\approx 26\%$
-% larger than when it is fully contracted on the main sequence
-% (0.40\,$R_\odot$ vs{.} 0.50\,$R_\odot$; ).
-
-There are two new lines of evidence that confirm the planetary
-interpretation.  First, the transit duration and the orbital period
-are inconsistent with an eclipsing body around the M-dwarf companion:
-we find $\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
-theoretically expected density for the companion is $\approx 4.6\,{\rm
-g\,cm}^{-3}$ (Table~\ref{tab:posterior}; \citealt{choi_mesa_2016}).
-The transit duration is therefore too long to be explained by a star
-eclipsing the M dwarf secondary at $10$-$\sigma$.
-
-%NOTE to the arxiv readers:, the $\approx$100 days of short-cadence
-%Kepler observations also mitigate the cadence-induced smearing, and
-%enable a more precise measurement of the transit impact parameter
-%\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
-%(2-$\sigma$) from the 1-minute data, compared to $<0.76$ (2-$\sigma$)
-%from the 30-minute data.  This implies an even stronger constraint on
-%the transit shape in an analysis of the false positive probability.
-
-The second line of confirming evidence comes by analyzing the
-individual Kepler transit times. We isolated each of the 144 observed
-transits to within $\pm4.5$\,hr of each transit, and fitted each
-window with both {\it i)} a local second-order polynomial and transit,
-and {\it ii)} a local linear trend.  We let the mid-time of each
-transit float, and then calculated the residual between the measured
-mid-time and that of a periodic orbit.  This residual, the transit
-timing variation (TTV), is plotted in Figure~\ref{fig:obliquity}
-against the local linear slope.  A significant correlation of $-0.0564
-\pm 0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
-Objects of Interest have shown this correlation
-\citep{holczer_time_2015}, which is most readily interpreted as a TTV
-induced by unresolved starspot crossings \citep{mazeh_time_2015}.
-This is only possible if the planet transits the primary star, which
-excludes a background eclipsing binary scenario, and therefore
-confirms that \pn\ is a planet.  It also suggests that the planet's
-orbit is prograde.  The latter point assumes that the dominant
-photometric variability is induced by dark spots, and not bright
-faculae.  Given the observed transition of Sun-like stellar
-variability from spot to faculae-dominated regimes between young and
-old ages, we expect this assumption to be reasonably secure
-\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
-
-A third supporting line of evidence for the planetary interpretation
-exists, though we consider it less definitive than the stellar
-density and TTV-local slope correlation.  We observed a transit of
-\pn\ on the night of 2021 Aug 7 simultaneously with Keck/HIRES and
-MuSCAT3.  We scheduled the observations using the ephemeris of
-\citet{holczer_transit_2016}.  Although we did not detect the
-Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
-show the transit (Figure~\ref{fig:obliquity}, lower panels).  Fitting
-the MuSCAT3 photometry with a model that lets the transit depths vary
-across each bandpass, we find {\it griz} depths consistent with the
-Kepler depth at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  The
-MuSCAT3 observations also suggest a transit duration $17.4\pm3.6\ {\rm
-minutes}$ shorter than the Kepler transits; further photometric
-follow-up could help confirm whether the transit duration is indeed
-changing.  For our RM analysis, the details are discussed in
-Appendix~\ref{app:rm}.  While the velocities are marginally more
-consistent with a prograde or polar orbit than a retrograde orbit, the
-spot-corrected exposure-to-exposure scatter ($\sigma_{\rm RV}\approx
-50$\ms) is larger than the expected velocity anomaly
-($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$).  We are therefore
-not in a position to claim a spectroscopic detection of the RM effect,
-nor to quantify the stellar obliquity.
-
-\section{Discussion \& Conclusions}
-\label{sec:conc}
-
 Kepler\,1627Ab provides a new extremum in the ages of the Kepler
-planets, and opens multiple avenues for further study.  Further
-observations of spectroscopic transits could yield a stellar obliquity
-measurement, which would confirm or refute the prograde orbital
+planets, and opens multiple avenues for further study.  
+Observations of spectroscopic transits at greater precision should yield a
+measurement of the stellar obliquity, which would confirm or refute the prograde orbital
 geometry suggested by the TTV-local slope correlation.  Separately,
-transit spectroscopy aimed at detecting atmospheric outflows could
+transit spectroscopy aimed at detecting atmospheric outflows might
 yield insight into the evolutionary state of the atmosphere
@@
 helping to clarify the expected outflow rate \citep[{\it
-e.g.},][]{poppenhaeger_2021}.  Finally, a challenging and informative
-quantity to measure would be the planet's mass.  This might be
-possible through a high-cadence multi-color radial velocity campaign,
-or through long-term monitoring of the transit times and durations.
-Measured at sufficient precision, the mass, combined with the known
+e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
+quantity to measure would be the planet's mass.  The most feasible
+approach would likely be
+long-term monitoring of the transit times and durations,
+given the technical difficulty of performing a high-cadence multi-color radial velocity campaign
+on a $V=13.1$ target.
+Measured at sufficient precision though, the mass, combined with the known
 age and size, would yield constraints on both the planet's composition
@@
 
-More immediately, the Kepler data may contain additional
+More immediately, the Kepler data may yet contain additional
 information.  For instance, one possible explanation for the transit
 asymmetry shown in Figure~\ref{fig:lc} is that of a dusty asymmetric
-outflow.  Other possible explanations such as favored longitudes for
-starspot crossings seem strained for reasons discussed in
-Appendix~\ref{app:gptransit}.  Dusty outflows are theoretically
+outflow.  
+A second possibility is that the planetary orbit is slightly misaligned
+from the stellar spin axis, and tends to
+transit starspot groups at
+favored stellar latitudes.
+Other possibilities such as gravity darkening or TTVs causing
+the asymmetry are disfavored
+(see Appendix~\ref{app:asymmetry}).  Dusty outflows are theoretically
 expected for young mini-Neptunes, and the amplitude of the observed
 asymmetry is also roughly consistent with predictions
-\citep{wang_dai_2019}.  Beyond the planetary transits,
-Appendix~\ref{app:flare} highlights a separate oddity in the Kepler
+\citep{wang_dai_2019}.  Beyond the asymmetric transits,
+Appendix~\ref{app:flare} highlights an additional abnormality in the Kepler
 data, in the arrival time distribution of stellar flares.  We
@@
 V1298~Tau system \citep{david_four_2019}.  \pn\ is one of the smallest
-planets in this sample ($3.4\pm
-0.2\,R_\oplus$), which could be linked to the selection effects
+planets in this sample ($3.36\pm0.18\,R_\oplus$), which could be linked to the selection effects
 imposed by the spot-induced photometric variability at very young ages
-\citep[{\it e.g.},][]{zhou_2021_tois}.  Assuming that these young
+\citep[{\it e.g.},][]{zhou_2021_tois}.  If these young
 planets have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$,
-their large sizes and presumed low densities would be in accord with
+then their low densities would be in accord with
 the expectation that mini-Neptunes start their lives with large
@@
 connection of the star to the \cn\ using Gaia kinematics, and then by
-using the Gaia HR diagram and TESS stellar rotation periods to verify
+using the Gaia color-absolute magnitude diagram and TESS stellar rotation periods to verify
 the cluster's existence.  Table~\ref{tab:v06} enables similar
@@
 systems \citep[{\it e.g.},][]{guerrero_tess_2021}.  The entry shown in
-the header of Table~\ref{tab:v06}, HD 150706b, is an example: the
+the header, HD 150706b, is one example: the
 host star appears to be a member of the Ursa Major moving group
-($\approx 400\,{\rm Myr}$; \citealt{mann_tess_2020}).  Confirming
-such kinematic associations with independent age indicators is
-essential, because the false positive rates are uncharted.
+($400\,{\rm Myr}$; \citealt{mann_tess_2020}).  Confirming
+this suggested kinematic association with independent age indicators is
+essential because the false positive rates are not known.
 Alternatively, one could instead opt to identify new kinematic
@@
 
-%\clearpage
+\clearpage
 \acknowledgements
@@
 
-\section{Kinematic Selection of $\delta$\,Lyr Cluster Members}
-\label{app:kinematicselection}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=1\textwidth]{f6.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Galactic position and tangential velocities of the
-    $\delta$\,Lyr cluster.}
-    Points are candidate cluster members with $\varpi/\sigma_\varpi >
-    20$, reported to be in the group by
-    \citet{kounkel_untangling_2019}.  We focus on stars (black points)
-    in the spatial and kinematic vicinity of Kepler\,1627 (yellow
-    star).  The other candidate cluster members (gray points) may or
-    may not share the ages of the selected kinematic group.  The
-    location of the Sun is ($\odot$) is shown.
-		\label{fig:XYZvtang}
-	}
-\end{figure*}
-
-Figure~\ref{fig:XYZvtang} shows stars reported by
-\citet{kounkel_untangling_2019} to be in the group ``Theia 73'', which
-was cross-matched by Kounkel \& Covey as being ``Stephenson 1''.
-\citet{kounkel_untangling_2019} reported \noriginal\ stars to be
-present in this cluster.  For our Figure, galactic positions were
-calculated and plotted only for stars with parallax signal-to-noise
-exceeding 20.  The location of the Sun is shown on the plots.  The
-smattering of reported cluster members (gray and black points) has a
-significantly different structure than the cluster initially
-identified by \citet{stephenson_possible_1959} and corroborated by
-\citet{eggen_photometric_1968}.  While the non-uniform ``clumps''
-might be part of a real structure, they could also be an artifact of
-the data processing steps performed by
-\citet{kounkel_untangling_2019}.  We therefore opted to only consider
-stars in the immediate kinematic and spatial vicinity of \sn.  The
-tangential velocities relative to \sn\ are shown in the bottom right
-panel.  These are computed by assuming that every star has the same
-three-dimensional spatial velocity as \sn, where we assume a systemic
-radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the HIRES
-spectra.  The
-relevant projection effects are then taken into account, as discussed
-by {\it e.g.}, \citet{Meingast2021} and \citet{bouma_2021_ngc2516}.
-We performed the actual selection by then manually drawing lassos with
-the interactive \texttt{glue} visualization tool
-\citep{beaumont_2014_13866} in the four projections shown in
-Figure~\ref{fig:XYZvtang}.  While other cluster members likely exist
-outside of our selection region (and our selection also includes some
-field star interlopers), our aim is to verify the existence of the
-cluster in the vicinity of Kepler\,1627, and to measure its age.  The
-procedure we have adopted enables both tasks.
+% \section{Kinematic Selection of $\delta$\,Lyr Cluster Members}
+% \label{app:kinematicselection}
+% 
+% \begin{figure*}[t]
+% 	\begin{center}
+% 		\leavevmode
+% 		\includegraphics[width=1\textwidth]{f6.pdf}
+% 	\end{center}
+% 	\vspace{-0.7cm}
+% 	\caption{
+%     {\bf Galactic position and tangential velocities of the
+%     $\delta$\,Lyr cluster.}
+%     Points are candidate cluster members with $\varpi/\sigma_\varpi >
+%     20$, reported to be in the group by
+%     \citet{kounkel_untangling_2019}.  We focus on stars (black points)
+%     in the spatial and kinematic vicinity of Kepler\,1627 (yellow
+%     star).  The other candidate cluster members (gray points) may or
+%     may not share the ages of the selected kinematic group.  The
+%     location of the Sun is ($\odot$) is shown.
+% 		\label{fig:XYZvtang}
+% 	}
+% \end{figure*}
+% 
+% Figure~\ref{fig:XYZvtang} shows stars reported by
+% \citet{kounkel_untangling_2019} to be in the group ``Theia 73'', which
+% was cross-matched by Kounkel \& Covey as being ``Stephenson 1''.
+% \citet{kounkel_untangling_2019} reported \noriginal\ stars to be
+% present in this cluster.  For our Figure, galactic positions were
+% calculated and plotted only for stars with parallax signal-to-noise
+% exceeding 20.  The location of the Sun is shown on the plots.  The
+% smattering of reported cluster members (gray and black points) has a
+% significantly different structure than the cluster initially
+% identified by \citet{stephenson_possible_1959} and corroborated by
+% \citet{eggen_photometric_1968}.  While the non-uniform ``clumps''
+% might be part of a real structure, they could also be an artifact of
+% the data processing steps performed by
+% \citet{kounkel_untangling_2019}.  We therefore opted to only consider
+% stars in the immediate kinematic and spatial vicinity of \sn.  The
+% tangential velocities relative to \sn\ are shown in the bottom right
+% panel.  These are computed by assuming that every star has the same
+% three-dimensional spatial velocity as \sn, where we assume a systemic
+% radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the HIRES
+% spectra.  The
+% relevant projection effects are then taken into account, as discussed
+% by {\it e.g.}, \citet{Meingast2021} and \citet{bouma_2021_ngc2516}.
+% We performed the actual selection by then manually drawing lassos with
+% the interactive \texttt{glue} visualization tool
+% \citep{beaumont_2014_13866} in the four projections shown in
+% Figure~\ref{fig:XYZvtang}.  While other cluster members likely exist
+% outside of our selection region (and our selection also includes some
+% field star interlopers), our aim is to verify the existence of the
+% cluster in the vicinity of Kepler\,1627, and to measure its age.  The
+% procedure we have adopted enables both tasks.
 
@@
 
-\section{Transit and Stellar Variability Model}
-\label{app:gptransit}
+\section{The Transit Asymmetry}
+\label{app:asymmetry}
 
@@
 
-We fitted the Kepler long cadence time series with a model that
-simultaneously included the planetary transit and the stellar
-variability.  The stellar variability was modelled with the
-\texttt{RotationTerm} Gaussian Process kernel in \texttt{exoplanet}
-\citep{exoplanet:exoplanet}.  This kernel assumes that the variability
-is generated by a mixture of two damped simple harmonic oscillators
-with characteristic frequencies set by 1/$P_{\rm rot}$ and its first
-harmonic.  We additionally included a jitter term to inflate
-the flux uncertainties in a manner that accounted for otherwise
-unmodelled excess white noise, and let the eccentricity float.  For
-the limb-darkening, we assumed a quadratic law, and sampled using the
-uninformative prior suggested by \citet{exoplanet:kipping13}.
-
-Our model therefore included 10 free parameters for the transit ($\{P,
-t_0, R_{\rm p}/R_\star, b, u_1 ,u_2 ,R_\star, \log g, e, \omega \}$),
-2 free parameters for the light curve normalization and uncertainty
-calibration ($\{\langle f \rangle, \sigma_f \}$), and 5
-hyperparameters for the GP ($\{\sigma_{\mathrm{rot}},
-P_{\mathrm{rot}}, Q_0, \mathrm{d}Q, f \}$).  We also considered
-including an additive \texttt{SHOTerm} kernel to account for
-stochastic noise, but found that this did not significantly affect the
-results, and so opted for the simpler GP kernel.  We fitted the models
-using \texttt{PyMC3} \citep{salvatier_2016_PyMC3,exoplanet:theano},
-and accounted for the finite integration time of each exposure in the
-numerical integration when evaluating the model light curve
-\citep[see][]{kipping_binning_2010}.  We assumed a Gaussian
-likelihood, and after initializing each model with the parameters of
-the maximum {\it a posteriori} model, we sampled using
-\texttt{PyMC3}'s gradient-based No-U-Turn Sampler
-\citep{hoffman_no-u-turn_2014} in these bases indicated in
-Table~\ref{tab:posterior}.  We used $\hat{R}$ as our convergence
-diagnostic \citep{gelman_inference_1992}.
-
-Figure~\ref{fig:lc} shows the resulting model in orange (top) and
-purple (bottom).  While the fit appears reasonably good, an asymmetry
-is present during the transit: the measured flux is systematically
-high during the first half, and low in the second half.  The
-semi-amplitude of the deviation is $\approx50\,{\rm ppm}$, which
-represents a $\approx 3\%$ distortion of the transit depth
-($\approx1700\,{\rm ppm}$).  To determine whether this effect was
-caused by our stellar variability model, we explored an alternative
-approach in which we isolated each transit window and locally fitted
-out polynomial trends, and then binned all the observed transits; the
-asymmetry was still present, at a comparable amplitude.  Note that
-even though this asymmetric feature is within the 2-$\sigma$ model
-uncertainties, this model itself has a jitter term that grows to
-account for otherwise unmodelled white noise in the flux.  The
-significance of the asymmetry is therefore best assessed in comparison
-against the intrinsic out-of-transit scatter, not the model
-uncertainties.
-
-As a means of exploring the robustness of the asymmetry,
+As a means of exploring the robustness of the transit asymmetry,
 Figures~\ref{fig:phasequarter},~\ref{fig:phaseyear},
@@
 timing variations in \sn\ have amplitudes smaller than $\approx$10
-minutes (see Figure~\ref{fig:obliquity}).  This is consistent with
+minutes (see Figure~\ref{fig:corr}).  This is consistent with
 previous results from \citet{holczer_transit_2016}.  Moreover, the
@@
 d}f/{\rm d}t\in[0.033,0.092]\,{\rm ppt\,day}^{-1}$, the sign of the
-asymmetry would reverse.  We do not see evidence for that being the
+asymmetry would reverse.  We do not see evidence for this being the
 case.
@@
 persistent at a particular latitude (and for the planetary orbit to be
-quite misaligned), or for a ``stroboscopic'' longitudinal phasing
+somewhat misaligned), or for a ``stroboscopic'' longitudinal phasing
 \citep[{\it e.g.},][]{dai_stellar_2018}.  For our system, $P_{\rm
@@
 is $\approx8$, which renders individual spot-crossing events
-unresolved.  This explanation seems marginally plausible, because the
+unresolved.  This explanation seems marginally plausible, mainly because the
 expected spot-crossing anomaly amplitudes ($\approx 100\,{\rm ppm}$)
 resemble the observed amplitude of the asymmetry ($\approx 50\,{\rm
-ppm}$).  We ultimately disfavor it though, because {\it a)} it seems
-implausible that the starspot crossings could add coherently, and {\it
-b)} there is no reason to expect starspot crossing events to last
+ppm}$).
+One issue with this explanation however is that
+there is no reason to expect starspot crossing events to last
 exactly half the transit duration.
@@
 likely in a regime where such outflows are possible.  Out of the four
-explanations discussed, this one currently seems the most plausible.
+explanations discussed, this one at least theoretically seems the most plausible.
 By composition, the expectation would be that the envelope is mostly
@@
 
-\section{Companion Star and False Positive Assessment}
-\label{app:companionstar}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f10a.pdf}
-			\includegraphics[width=0.48\textwidth]{f10b.pdf}
-		}
-	\end{center}
-	\vspace{-0.5cm}
-	\caption{
-    {\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
-    from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2.5V
-    companion at $\rho \approx 0\farcs16$, which corresponds to a
-    projected separation of $53\pm4$\,AU.  The inset shows a cutout of
-    the stacked NIRC2 image (North is up, East is left, scale is set by the
-    separation of the binary).  The lines show 5-$\sigma$ contrast
-    limits for the `Alopeke filters, and 6-$\sigma$ contrast limits
-    for NIRC2 outside of $0\farcs15$. {\it Right:} Gaia
-    EDR3 renormalized unit weight error (RUWE) point estimates for
-    Kepler\,1627A and all members of the $\delta$\,Lyr cluster
-    reported by \citet{KounkelCovey2019}.  Since other members of the
-    cluster with similar brightnesses have comparable degrees of
-    photometric variability, the high RUWE suggests that Kepler\,1627
-    is a binary. 
-    \label{fig:kep1627binary}
-	}
-\end{figure*}
-
-We first noted the presence of a close neighbor on 2015 July 22 when
-we acquired adaptive optics imaging of \sn\ using the NIRC2 imager on
-Keck-II. We used the narrow camera (FOV = 10.2\arcsec) to obtain 8
-images in the $K'$ filter ($\lambda = 2.12\,\mu$m) with a total
-exposure time of 160\,s. We analyzed these data following
-\citet{kraus_impact_2016}, beginning by applying standard reduction
-steps (dark-subtraction, flat-fielding, linearization, and bad pixel
-correction). We then used PSF-fitting to measure the separation,
-position angle, and contrast of the candidate companion, identifying
-the best-fitting empirical PSF template from among the
-near-contemporaneous observations of single stars in the same filter.
-Finally, we used the optical distortion solution of \citet{yelda_2010}
-to convert pixel-space relative positions to on-sky relative
-astrometry. The mean values inferred from the 8 images are reported in
-Table~\ref{tab:starparams}.
-
-We also estimated detection limits following
-\citet{kraus_impact_2016}, analyzing the residuals after subtracting
-the empirical PSF templates, as well as the residuals after
-subtracting an azimuthally smoothed PSF profile. Within each residual
-image, the flux was measured through 40\,mas apertures centered on
-every pixel, and then the noise as a function of radius was estimated
-from the RMS within concentric rings. Finally, the detection limits
-were estimated from the strehl-weighted sum of the detection
-significances in the image stack, and we adopted the $+6\sigma$
-threshold as the detection limit for ruling out additional companions.
-We find contrast limits of $\Delta m_{K'} = 5.1$ mag at $\rho = 150$ mas,
-$\Delta m_{K'} = 7.0$ mag at $\rho = 300$ mas, and $\Delta m_{K'} = 9.0$ mag
-at $\rho > 1000$ mas.  Given the distance ($d = 330$\,pc) and age
-($\tau \sim 40$\,Myr), the models of \citet{chabrier_2000} imply
-corresponding physical limits on tertiary companions of $M_{\rm ter} < 50
-M_{\rm Jup}$ at $\rho = 50$\,AU, $M_{\rm ter} < 20 M_{\rm Jup}$ at $\rho =
-100$\,AU, and $M_{\rm ter} < 10 M_{\rm Jup}$ at $\rho = 330$\,AU.
-
-We also observed \sn\ on Gemini-North using the `Alopeke speckle
-imager on 2021 June 24.  `Alopeke is a dual-channel speckle
-interferometer that uses narrow-band filters centered at 0.83\,$\mu$m
-and $0.56\,\mu$m.  We acquired three sets of $1000\times 60$$\,$msec
-exposures during good seeing (0.45$''$), and used the autocorrelation
-function of these images to reconstruct a single image and 5-$\sigma$
-detection limits (see \citealt{howell_speckle_2011}).  This procedure
-yielded a detection of the companion in the 0.83\,$\mu$m notch filter,
-but not the $0.56\,\mu$m filter.  The measured projected separation
-and magnitude difference are given in Table~\ref{tab:starparams}.
-
-Figure~\ref{fig:kep1627binary} summarizes the results of the
-high-resolution imaging.  The Gaia EDR3 parallax for the primary
-implies a projected separation of $53 \pm 4$\,AU, assuming the
-companion is bound.  Although the companion is unresolved in the Gaia
-source catalog (there are no comoving, codistant candidate companions
-brighter than $G < 20.5$ mag within $\rho < 120\arcsec$), its
-existence was also suggested by the primary star's large RUWE,
-relative to other members of the \cn.  Based on the apparent
-separation, the binary orbital period is of order hundreds of years.
-The large RUWE is therefore more likely to be caused by a PSF-mismatch
-skewing the Gaia centroiding during successive scans, rather than true
-astrometric motion.  Regardless, given the large observed RUWE of the
-primary, and the low geometric probability that a companion imaged at
-$\rho \approx 0\farcs16$ is a chance line-of-sight companion, we
-proceed under the assumption that the companion is bound, and that
-Kepler\,1627 is a binary.  Given the distance and age, the models of
-\citet{baraffe_new_2015} imply a companion mass of $M_{\rm B} \approx
-0.33 M_{\odot}$ and companion temperature of $T_{\rm eff,B} \approx
-3450$ K.  The corresponding spectral type is roughly M2.5V
-\citep{pecaut_mamajek_2013}.
-% \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
-
-% (0.7-2.0\% V-band to G-band dmag difference from EEM table; NOTE
-% could determine using model spectra)
-Could the companion be creating a false positive signal?  The
-companion star contributes $\approx$1\% of the total flux observed in
-the Kepler aperture.  The observed transit has a depth of 0.18\%.  An
-18\% deep eclipse of the secondary star would therefore be needed to
-produce a signal with the appropriate depth.  The shape of the
-observed signal requires the impact parameter to be below 0.74
-(Table~\ref{tab:posterior}); the body transiting the secondary would
-therefore need to be non-grazing with $R_3/R_2 \approx 0.42$.
-Assuming a $\approx 0.46R_\odot$ radius of the imaged secondary
-from the \citet{baraffe_new_2015} models, this
-would imply a tertiary stellar radius of $\approx 0.2R_\odot$.  This
-yields a contradiction:  this scenario would require an ingress and
-egress phase that each span $\approx$40\% of the transit duration
-($\approx 65\,{\rm minutes}$).  The actual measured ingress and egress
-duration is $\approx 15\,{\rm minutes}$), 4.4$\times$ shorter.  The
-combination of the companion's brightness, transit depth, and ingress
-duration therefore rule out the scenario that it could explain the
-transit signal.  Two other arguments presented in the main text -- the
-inferred stellar density, and the TTV-local slope correlation --
-confirm this point.
-
 \section{Spectroscopic Transit Analysis}
@@
 
-The lower panels in Figure~\ref{fig:obliquity} show the results.  The
+Figure~\ref{fig:ground} shows the results.  The
 MuSCAT3 photometry shows the expected starspot trend, along with the
@@
 The quadratic model with the RM effect is shown in
-Figure~\ref{fig:obliquity}; the jitter term is incorporated in the
+Figure~\ref{fig:ground}; the jitter term is incorporated in the
 model uncertainties, but not the plotted measurement uncertainties.
@@
 
-The presence of this additional scatter prevents a clear detection of
+The presence of this additional scatter prevents a convincing detection of
 the RM effect.  The reason can be understood via
 model comparison.  If we compare the model with a quadratic trend and
-the RM effect (Figure~\ref{fig:obliquity}) against a model with a
+the RM effect against a model with a
 linear trend and the RM effect, or even a model with no RM effect at
@@
   \item Fit the starspot-induced variability using a
-    Gaussian Process with a \texttt{SHOTerm} kernel (see
-    Section~\ref{app:gptransit}), a white-noise jitter term, and the
+    Gaussian Process with a \texttt{SHOTerm} kernel, a white-noise jitter term, and the
     mean flux.

### 2021-09-29 paper/ms.tex
update ms

@@
 non-uniform ``clumps'' of stars might comprise a {\it bona fide}
-cluster of identically-aged stars, they could also be significantly
+cluster of identically-aged stars, they could also be heavily
 contaminated by field stars.  We therefore considered stars only in
@@
     1-$\sigma$ error bar in the lower right of each panel.  The asymmetric
-    residual during transit is significantly larger than the out-of-transit scatter.
+    residual during transit is larger than the out-of-transit scatter.
     \label{fig:lc}
@@
 including an additive \texttt{SHOTerm} kernel to account for
-stochastic noise, but found that this did not significantly affect the
+stochastic noise, but found that this did not affect the
 results, and so opted for the simpler GP kernel.  We fitted the models
@@
 and fixed the limb-darkening using the $V$-band tabulation from
-\citet{claret_gravity_2011}.  We assumed a prior on $v\sin i$ and
+\citet{claret_gravity_2011}.  We assumed a Gaussian prior on $v\sin i$ and
 $a/R_\star$ from Table~\ref{tab:starparams}, and also allowed for a
@@
 novel process operating on the surface of Kepler 1627A.  Another
-possibility is that our RV analysis underestimates the measurement
+possibility is that our RV analysis underestimates our measurement
 uncertainties; the S/N per resolution element for each of our spectra
-was moderated by the observing requirements to 70 to 80, and the rapid
-rotation of the star could affect accuracy of the uncertainties from
-the velocity extraction.  Observations at higher precision are
-necessary to distinguish these two possibilities, and remain
-worthwhile as a means toward clarifying the orbital geometry of \pn.
+was 70 to 80, which is somewhat low for deriving high-precision
+velocities.  In addition, the rapid rotation of the star could affect
+accuracy of the uncertainties from the velocity extraction.
+Observations at higher precision are necessary to distinguish these
+two possibilities, and remain worthwhile as a means toward clarifying
+the orbital geometry of \pn.
 

### 2021-10-01 paper/ms.tex
update ms
* Improve phased quartile appendix discussion
* Improve RM appendix discussion
* Push updated RM plots

@@
 		best-fitting model in the top panel includes the RM effect and a quadratic trend
-		in time.  The radial velocity scatter
-		between exposures ($\sigma_{\rm RV}\approx50$\ms) prevented us from detecting the RM
+		in time to model the spot-induced $\approx$\,$250$\ms\ trend (see Appendix~\ref{app:rm}).  The radial velocity jitter
+		over the 15~minute exposure timescale ($\sigma_{\rm RV}\approx30$\ms) prevented us from detecting the RM
 		effect.
+		The RV variations are strongly correlated with varying emission in the Ca H and K lines (middle panel).
 		\label{fig:ground}
@@
 spot-corrected exposure-to-exposure scatter ($\sigma_{\rm RV}\approx
-50$\ms) is larger than the expected velocity anomaly
-($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$).  We are therefore
-not in a position to claim a spectroscopic detection of the RM effect,
-nor to quantify the stellar obliquity.
+30$\ms) is comparable to the expected velocity anomaly assuming an
+aligned orbit ($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$).  We
+are therefore not in a position to claim a spectroscopic detection of
+the RM effect, nor to quantify the stellar obliquity.
 
@@
 		\leavevmode
-		\includegraphics[width=0.7\textwidth]{f5.pdf}
+		\includegraphics[width=0.75\textwidth]{f5.pdf}
 	\end{center}
@@
 
-\begin{figure}[t]
+\begin{figure}[tp]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=\textwidth]{f9.pdf}
+		\includegraphics[width=0.9\textwidth]{f9.pdf}
 	\end{center}
@@
 	\caption{
-    {\bf Transit models and residuals, binned by local slope of the
-    raw light curve}.  Local linear slopes are measured over each
-    transit window (144 transits total).  Four outlier transits were
-    removed, leaving 140 transits.  These were then divided into
-    quartiles, so that each panel shows 35 transits binned together.
-    The exact light curve slopes are listed in the lower left panels.
+    {\bf Transit models and residuals, binned by quartiles in the local slope of the
+    light curve}.  
     Representative uncertainties for the black points (binned at 20
@@
     seems to be present in three of the four bins.
+    \vspace{-0.5cm}
 		\label{fig:phaseslope}
@@
 \end{figure}
+\begin{figure}[tp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f9b.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+    {\bf Transit models and residuals, binned by quartiles of transit
+    timing variations}.  The data are otherwise the same as in
+    Figure~\ref{fig:phaseslope}.
+		\label{fig:phasettv}
+	}
+\end{figure}
+
+\subsection{Is the Transit Asymmetry Robust?}
 
 As a means of exploring the robustness of the transit asymmetry,
-Figures~\ref{fig:phasequarter},~\ref{fig:phaseyear},
-and~\ref{fig:phaseslope} show the Kepler data binned in three ways:
-over quarters, years, and local slope quartiles.  Over quarters
-(Figure~\ref{fig:phasequarter}), Quarter 6 shows the strongest asymmetry
-out of any quarter: a deviation of about 3\,ppt from expectation.
+Figures~\ref{fig:phasequarter},~\ref{fig:phaseyear},~\ref{fig:phaseslope},
+and~\ref{fig:phasettv} show the Kepler data binned in four ways:
+over Kepler quarters, Julian years, quartiles of local slope, and quartiles of
+transit timing variation.  Over Kepler quarters
+(Figure~\ref{fig:phasequarter}), Quarter 6 shows the strongest asymmetry: a deviation of about 3\,ppt from expectation.
 Quarter 7 shows an anomaly at roughly the same transit phase.  Year 2
@@
 in each of Years 2, 3, and 4.
-Binned by local slope quartiles (Figure~\ref{fig:phaseslope}), the
-asymmetry is visually present in three of the four quartiles: the only
-bin for which it appears to not be present is ${\rm d}f/{\rm
-d}t\in[0.001,0.033]\,{\rm ppt\,day}^{-1}$.
-
-We considered four possible interpretations of the transit asymmetry:
-gravity darkening, transit timing variations, spot-crossing events,
-and a persistent asymmetric dusty outflow.
+
+To bin by quartiles in local slope and TTV, we used our measurements
+of the local linear slopes in each of the observed transit windows
+(144 transits total).  Four outlier transits were removed, leaving 140
+transits.  These were then divided into quartiles, so that each panel
+shows 35 transits binned together.  The exact light curve slope and
+TTV intervals are listed in the lower left panels of
+Figures~\ref{fig:phaseslope} and~\ref{fig:phasettv}.  Binned by local
+slope quartiles (Figure~\ref{fig:phaseslope}), the asymmetry is
+visually present in three of the four quartiles: the only bin in
+which it does not appear is ${\rm d}f/{\rm
+d}t\in[0.001,0.033]\,{\rm ppt\,day}^{-1}$.  The corresponding bins in TTV
+(largest TTV corresponding to smallest ${\rm d}f/{\rm d}t$)
+show the same asymmetries, though with somewhat greater scatter
+(Figure~\ref{fig:phasettv}).
+
+Within the theory presented by \citet{mazeh_time_2015}, unresolved
+starspot crossings cause the correlation between TTVs and the local
+light curve slope (Figure~\ref{fig:corr}).  In this model, we would
+expect the light curves with the most negative local slopes
+to have the largest positive TTVs, due to spot crossing events during the latter
+half of transit.  The lower-right panel of Figure~\ref{fig:phasettv}
+and upper-left panel of Figure~\ref{fig:phaseslope} agree with this
+expectation.  However, we would also expect the sign of the effect to
+reverse when considering the most positive local slopes, and the most
+negative TTVs.  The upper-left panel of Figure~\ref{fig:phasettv} and
+lower-right panel of Figure~\ref{fig:phaseslope} contradict this
+expectation: the residual in both cases maintains the same
+parity!  On the one hand, this shows us that the residual is not
+strongly dependent on the local light curve slope, which mitigates the
+possibility that it might be an artifact of our detrending methods.
+On the other, it raises the question of whether
+unresolved starspot crossings are indeed the root cause of the correlation shown in
+Figure~\ref{fig:corr}.
+
+% FIXME: KM to discuss injection recovery...
+
+\subsection{Possible Explanations}
+
+Since the transit asymmetry seems robust against most methods of
+binning the data, we considered four possible astrophysical
+interpretations: gravity darkening, transit timing variations,
+spot-crossing events, and a persistent asymmetric dusty outflow.
 
@@
 the regions without iodine, but the line profile stability of HIRES
-precludes such an approach.}, extracted the 1-D spectra using the
-standard California Planet Survey pipeline \citep{howard_cps_2010},
-and measured the velocities by cross-correlating against a template
-found via spectral classification with \texttt{SpecMatch-Emp}
-\citep{yee_SM_2017}.  The airmass ranged between 1.1 and 2.2 from the
+precludes such an approach.}, and extracted the 1-D spectra using the
+standard California Planet Survey pipeline \citep{howard_cps_2010}.
+The airmass ranged between 1.1 and 2.2 from the
 start through the end of observations; the seeing ranged from
@@
 
+We considered two approaches to measuring the velocities: in the
+first, hereafter ``Method 1'', we cross-correlated against a template found via
+spectral classification with \texttt{SpecMatch-Emp}
+\citep{yee_SM_2017}.  In ``Method 2'', we used a high S/N template of
+V1298~Tau.  Although V1298 Tau is cooler than Kepler~1627A by
+$\approx$500\,K, it has a comparable amount of line-broadening ($v\sin
+i = 23$\kms), and a comparable level of stellar activity.  The mean
+and standard deviation of the internal RV uncertainties averaged over
+all epochs were $16.2\pm1.1$\ms\ from Method 1, and $12.6\pm0.6$\ms\ from Method 2.
+The corresponding time-averaged reduced $\chi^2$ from the template
+match was $1.57\pm0.04$ (Method 1) and $1.30\pm0.02$ (Method 2).
+Given these diagnostics, we adopted the velocities from the second
+approach.
+
 Figure~\ref{fig:ground} shows the results.  The
@@
 ${\rm JD} - 2459433 = 0.955$.  The radial velocities decrease by
-800\ms\ over the six hour window.  A $\approx50$\ms\  increase in RV
-is seen shortly after the expected ingress time.
+$\approx$250\ms\ over the six hour window.  
+This decrease in RV is heavily with a decrease in the S-indices
+derived from the Ca HK lines.
+One outlying RV point is apparent shortly before egress; it is
+temporally coincident with an outlying value in the S-index time series.
 
@@
 view, modulate the stellar velocity profile, and can thereby produce
-the overall 800\ms\ trend.  The Ca HK and H$\alpha$ emission
+the overall $\approx$250\ms\ trend.  The Ca HK and H$\alpha$ emission
 profiles support this interpretation;
@@
 intensity over the course of the six hour sequence.
-The overall RVs are also heavily correlated against the S-indices
-derived from the Ca HK lines.
 
@@
 Rossiter-McLaughlin (RM) effect, and allowed for an optional linear
-and quadratic trend in time to fit the 800\ms\ spot-induced trend.  We
+and quadratic trend in time to fit the $\approx$250\ms\ spot-induced trend.  We
 followed the methodology developed by \citet{stefansson_2020}.  We
@@
 uncertainties.
+We used the 15\,minute exposure time when numerically evaluating the
+model.
 
@@
 The plotted measurement uncertainties are the internal uncertainties
-on the RVs ($\approx 15\,{\rm m\,s}^{-1}$), and are dominated by the
+on the RVs ($\approx 13\,{\rm m\,s}^{-1}$), and are dominated by the
 $v\sin i$ broadening.  However, between exposures, the RVs show
@@
 trend.  The white-noise jitter for this
-particular model is $\sigma_{\rm RV} = 50^{+11}_{-8}$\ms, which is
-larger than the expected RM anomaly of $\Delta v_{\rm RM}
+particular model is $\sigma_{\rm RV} = 27^{+6}_{-5}$\ms, which is
+comparable to the expected RM anomaly of $\Delta v_{\rm RM}
 \approx f_{\rm LD} \cdot \delta \cdot v\sin i \cdot \sqrt{1-b^2}
-\approx 20\,{\rm m\,s}^{-1}$.
+\approx 20\,{\rm m\,s}^{-1}$, assuming a perfectly aligned orbit.
 
@@
 are as follows.
-% from logs/rm_freejitter_FINAL.log
+% from rm_TEMPLATE_V1298TAU_quadraticfreejitter_v0.log and
+% rm_TEMPLATE_V1298TAU_v1_linearfreejitter_trendonlyfreejitter.log
 \begin{align}
-  {\rm BIC} &= 251.7  \quad ({\rm Quadratic+RM}) \nonumber \\
-  {\rm BIC} &= 257.7  \quad ({\rm Linear+RM}) \nonumber \\
-  {\rm BIC} &= 247.1  \quad ({\rm Only\ Quadratic}).
+  {\rm BIC} &= 227.1  \quad ({\rm Quadratic+RM}) \nonumber \\
+  {\rm BIC} &= 231.1  \quad ({\rm Linear+RM}) \nonumber \\
+  {\rm BIC} &= 221.4  \quad ({\rm Only\ Quadratic}).
 \end{align}
 There is therefore no evidence to prefer the model with the RM effect
-against a model that only accounts for the stellar variability.
-The ``only quadratic'' model does particularly well because it can
-inflate the jitter term to account for scatter during
-the transit (even if the scatter contains astrophysics!), and it has fewer free parameters.  However, we cannot
-justify a physical prior on the jitter term, because we do not
-understand the origin of the exposure-to-exposure scatter.  As noted
-above, the velocity deviations from starspots are expected to have
-contributions at the stellar rotation frequency, or harmonics thereof.
-This jitter is present on the exposure timescale (15 minutes), which
-is only 0.4\% of the stellar rotation period; it is not obvious that
-starspots would be the culprit.
-
-The amplitude of both the spot-induced trend and the jitter are larger
-than recent comparable measurements in systems such as AU~Mic
-\citep{palle_transmission_2020}, DS~Tuc
+against a model that only accounts for the stellar variability.  The
+``only quadratic'' model does particularly well because it can inflate
+the jitter term to account for scatter during the transit (even if the
+scatter contains astrophysics!), and it has fewer free parameters.
+However, we cannot justify a physical prior on the jitter term,
+because we do not understand the origin of the exposure-to-exposure
+scatter.  As noted above, the velocity deviations from starspots are
+expected to have contributions at the stellar rotation frequency, or
+harmonics thereof.  This jitter is present on the exposure timescale
+(15 minutes), which is only 0.4\% of the stellar rotation period; it
+is not obvious that starspots would be the culprit.
+
+The amplitude of both the spot-induced trend and the jitter are
+somewhat larger than recent comparable measurements in systems such as
+AU~Mic \citep{palle_transmission_2020}, DS~Tuc
 \citep{montet_young_2020,zhou_well_2020} and TOI~942
@@
 possibility is that our RV analysis underestimates our measurement
-uncertainties; the S/N per resolution element for each of our spectra
-was 70 to 80, which is somewhat low for deriving high-precision
-velocities.  In addition, the rapid rotation of the star could affect
-accuracy of the uncertainties from the velocity extraction.
-Observations at higher precision are necessary to distinguish these
-two possibilities, and remain worthwhile as a means toward clarifying
-the orbital geometry of \pn.
+uncertainties; in order to achieve the requisite time-sampling the S/N
+per resolution element in our spectra was 70 to 80, which is lower
+than desired for deriving high-precision velocities.  In addition, the
+rapid rotation of the star could affect accuracy of the uncertainties
+from the velocity extraction.  Observations at higher S/N are
+necessary to distinguish these two possibilities, and remain
+worthwhile in order to clarify the orbital geometry of \pn.  Useful
+next steps would include transit observations with a stabilized
+optical spectrograph, or potentially also observations in the
+near-infrared \citep[{\it e.g.},][]{feinstein_halpha_2021}.
 

### 2021-10-01 paper/collect.sh

@@ -29,7 +29,9 @@ cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtra
 
 cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_phasedsubsets_yearchunk.pdf f8.pdf
 
-cp /Users/luke/Dropbox/proj/rudolf/results/phasedlc_slope_quartiles/Kepler_1627_RotGPtransit_phasedlc_slope_quartiles.pdf f9.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/phasedlc_quartiles/Kepler_1627_RotGPtransit_phasedlc_slope_quartiles.pdf f9.pdf
+
+cp /Users/luke/Dropbox/proj/rudolf/results/phasedlc_quartiles/Kepler_1627_RotGPtransit_phasedlc_ttv_quartiles.pdf f9b.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/imaging/imaging_summary.pdf f10a.pdf
 cp /Users/luke/Dropbox/proj/rudolf/results/gaia_ruwe_vs_apparentmag/ruwe_vs_apparentmag.pdf f10b.pdf

### 2021-10-02 paper/ms.tex
update ms

@@
 \citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
-we begin by examining the properties of the cluster more
-closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
+we begin by examining the properties of the cluster (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
 cluster is \clusterage\ old, and 
@@
 below the 80$^{\rm th}$ percentile of each cluster's distribution, and
-manually excluded stars that were obvious photometric binaries in the
+excluded stars that were obvious photometric binaries in the
 CAMD.  We then passed a moving box average and standard deviation
@@
 receives the most weight in the interpolation.  Plausible ages for
-IC\,2602 span 30\,Myr to 46\,Myr, with slightly older ages being
+IC\,2602 span 30\,Myr to 46\,Myr, with older ages being
 preferred by the lithium-depletion-boundary (LDB) measurements 
@@
 The statistical uncertainties are propagated from the absolute
-magnitude (mostly originating from the the parallax uncertainty) and the color; the systematic
+magnitude (mostly originating from the parallax uncertainty) and the color; the systematic
 uncertainties are taken to be the difference between the PARSEC
@@
 % \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
-These models combined with the NIRC2 contrast limits similarly imply
+These models combined with the NIRC2 contrast limits imply
 physical limits on tertiary companions of $M_{\rm ter} < 50
@@
 Figure~\ref{fig:lc} show an asymmetric
-residual in the data relative to the model: the measured flux is systematically
+residual in the data relative to the model: the measured flux is 
 high during the first half of transit, and low in the second half.  The
@@
 finds that the asymmetry also seems to be robust to different methods of
-data binning.  Plausible astrophysical explanations are given in
+data binning.  The most likely astrophysical explanations are
+discussed in
 Section~\ref{sec:conc}.
@@
 expected for young mini-Neptunes, and the amplitude of the observed
-asymmetry is also roughly consistent with predictions
+asymmetry is also consistent with predictions
 \citep{wang_dai_2019}.  Beyond the asymmetric transits,
@@
 T{.}~David for illuminating discussions and suggestions, and to
-R{.}~Kerr for kindly providing us with the \citet{Kerr2021} membership
+R{.}~Kerr for providing us with the \citet{Kerr2021} membership
 list prior to its publication.
@@
 literature age was not readily available.  Nonetheless, since we do
-generally prefer stars with known ages, we made a few additional
+prefer stars with known ages, we made a few additional
 efforts to populate this column.  When available, the age provenance
@@
 For any of the catalogs for which Gaia DR2 identifiers were not
-immediately available, we either followed the spatial (plus
+available, we either followed the spatial (plus
 proper-motion) cross-matching procedures described in
@@
 parity!  On the one hand, this shows us that the residual is not
-strongly dependent on the local light curve slope, which mitigates the
+dependent on the local light curve slope, which mitigates the
 possibility that it might be an artifact of our detrending methods.
@@
 We considered two approaches to measuring the velocities: in the
-first, hereafter ``Method 1'', we cross-correlated against a template found via
-spectral classification with \texttt{SpecMatch-Emp}
+first, hereafter ``Method 1'', we cross-correlated against a template
+found via spectral classification with \texttt{SpecMatch-Emp}
 \citep{yee_SM_2017}.  In ``Method 2'', we used a high S/N template of
@@
 and standard deviation of the internal RV uncertainties averaged over
-all epochs were $16.2\pm1.1$\ms\ from Method 1, and $12.6\pm0.6$\ms\ from Method 2.
-The corresponding time-averaged reduced $\chi^2$ from the template
-match was $1.57\pm0.04$ (Method 1) and $1.30\pm0.02$ (Method 2).
-Given these diagnostics, we adopted the velocities from the second
-approach.
+all epochs were $16.2\pm1.1$\ms\ from Method 1, and $12.6\pm0.6$\ms\
+from Method 2.  The corresponding time-averaged reduced $\chi^2$ from
+the template match was $1.57\pm0.04$ (Method 1) and $1.30\pm0.02$
+(Method 2).  Given these diagnostics, we adopted the velocities from
+the second approach.
 
@@
 $\approx$250\ms\ over the six hour window.  
-This decrease in RV is heavily with a decrease in the S-indices
+This decrease in RV is correlated with a decrease in the S-indices
 derived from the Ca HK lines.

### 2021-10-02 paper/ms.tex
update ms

@@
 \citep[{\it
-e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,Kerr2021}.
+e.g.},][]{CantatGaudin2018a,Zari2018,KounkelCovey2019,Meingast2021,Kerr2021}.
 As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
@@
 More specifically, a clustering analysis of the Gaia
-data by \citet{kounkel_untangling_2019}
+data by \citet{KounkelCovey2019}
 reported that Kepler~1627 (KIC 6184894; KOI 5245; TIC 120105470) is a
@@
 			$\delta$\,Lyr cluster.} Points are reported cluster members from
-		\citet{kounkel_untangling_2019}.  
+		\citet{KounkelCovey2019}.  
 		The tangential velocities relative to
@@
 
-\citet{kounkel_untangling_2019} applied an unsupervised clustering
+\citet{KounkelCovey2019} applied an unsupervised clustering
 algorithm to Gaia DR2 on-sky positions, proper motions, and
@@
 The stars from the Pleiades and IC\,2602 were adopted from
-\citet{cantatgaudin_gaia_2018}, and the field stars are from the Gaia
+\citet{CantatGaudin2018a}, and the field stars are from the Gaia
 EDR3 Catalog of Nearby Stars \citep{gaia_gcns_2021}.  We cleaned these
@@
 literature \citep[{\it
-e.g.},][]{GaiaCollaboration2018,kounkel_untangling_2019,bossini_age_2019}.
+e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}.
 
@@
 clusters.  For our reference clusters, we adopted members of UCL,
-IC\,2602, and the Pleiades from \citet{damiani_stellar_2019} and
+IC\,2602, and the Pleiades from \citet{Damiani2019} and
 \cite{CantatGaudin2018a} respectively.  We removed binaries by
@@
 Of the \noriginal\ candidate \cn\ members reported by
-\citet{kounkel_untangling_2019}, \nwithtess\ stars were amenable to
+\citet{KounkelCovey2019}, \nwithtess\ stars were amenable to
 rotation period measurements ($G<17$ and $\bpmrpo>0.5$) using the TESS
@@
 %     20$, reported to be in the group by
-%     \citet{kounkel_untangling_2019}.  We focus on stars (black points)
+%     \citet{KounkelCovey2019}.  We focus on stars (black points)
 %     in the spatial and kinematic vicinity of Kepler\,1627 (yellow
@@
 % Figure~\ref{fig:XYZvtang} shows stars reported by
-% \citet{kounkel_untangling_2019} to be in the group ``Theia 73'', which
+% \citet{KounkelCovey2019} to be in the group ``Theia 73'', which
 % was cross-matched by Kounkel \& Covey as being ``Stephenson 1''.
-% \citet{kounkel_untangling_2019} reported \noriginal\ stars to be
+% \citet{KounkelCovey2019} reported \noriginal\ stars to be
 % present in this cluster.  For our Figure, galactic positions were
@@
 % the data processing steps performed by
-% \citet{kounkel_untangling_2019}.  We therefore opted to only consider
+% \citet{KounkelCovey2019}.  We therefore opted to only consider
 % stars in the immediate kinematic and spatial vicinity of \sn.  The
@@
 
-\subsection{Is the Transit Asymmetry Robust?}
+\subsection{How Robust is the Asymmetric Transit?}
 
@@
 
-\subsection{Possible Explanations}
+\subsection{Interpretation}
 
@@
 uncertainties.
-We used the 15\,minute exposure time when numerically evaluating the
+We used a 15\,minute exposure time to numerically evaluate the
 model.
@@
 next steps would include transit observations with a stabilized
-optical spectrograph, or potentially also observations in the
-near-infrared \citep[{\it e.g.},][]{feinstein_halpha_2021}.
+spectrograph in the optical \citep[{\it
+e.g.},][]{gibson_2016_kpf,seifahrt_maroon-x_2018}, or in the near-infrared
+\citep[{\it e.g.},][]{feinstein_halpha_2021}.
 

### 2021-10-07 paper/ms.tex
update ms
* debugged RotGPtransit
* clarified discussion points

@@
   Kepler\,1627A is a G8V star previously known to host a
-  $3.4\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
+  $3.8\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
   observed by the Kepler space telescope because it is nearby
@@
   The Kepler photometry shows two 
-  peculiarities: the average transit profile appears to be
+  peculiarities: the average transit profile is
   asymmetric, and the individual transit times are correlated with the
@@
   We discuss possible explanations for each anomaly.
-  More widely applicable though is the fact that the $\delta$~Lyr
+  More broadly applicable though is the fact that the $\delta$~Lyr
   cluster is one of $\sim$10$^3$ clusters whose properties have been
   clarified by Gaia.  Many other exoplanet hosts are candidate
-  members of these clusters based on kinematics; their ages can
-  be confirmed by using the TESS full frame images to
-  measure stellar rotation periods.
+  kinematic members of these clusters; their cluster memberships can
+  be verified by measuring stellar rotation periods in the TESS full
+  frame images.
 \end{abstract}
@@
 Focusing on the planet (Section~\ref{sec:planet}), we confirm that despite the
-presence of the previously unreported M2.5V companion, hereafter
+existence of the previously unreported M2.5V companion, hereafter
 Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A.  Based on a correlation between the
@@
 		\leavevmode
-		\includegraphics[width=1\textwidth]{f6.pdf}
+		\includegraphics[width=1\textwidth]{f1.pdf}
 	\end{center}
@@
 		three-dimensional spatial velocity as \sn.
-		Our analysis uses stars (black points)
+		Our analysis considers stars (black points)
 		in the spatial and kinematic vicinity of Kepler\,1627 (yellow
@@
 matched these stars against the latest Gaia EDR3 observations using
-the \texttt{dr2\_neighbourhood} table from the ESA archive\footnote{See
-\url{https://gea.esac.esa.int/archive/documentation/GEDR3/Gaia_archive/chap_datamodel/sec_dm_auxiliary_tables/ssec_dm_dr2_neighbourhood.html}.},
+the \texttt{dr2\_neighbourhood} table from the ESA archive,\footnote{See
+\url{https://gea.esac.esa.int/archive/documentation/GEDR3/Gaia_archive/chap_datamodel/sec_dm_auxiliary_tables/ssec_dm_dr2_neighbourhood.html}.}
 and took the closest proper motion and epoch-corrected angular
@@
 		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f1.pdf}
+		\includegraphics[width=0.47\textwidth]{f2.pdf}
 	\end{center}
@@
 		{\bf Kepler's view of the $\delta$~Lyr cluster.} Each black circle is a
-		candidate cluster member, selected from the catalog of \citet{KounkelCovey2019} based on its
+		candidate cluster member selected from the catalog of \citet{KounkelCovey2019} based on its
 		position and kinematics.
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f2a.pdf}
-			\includegraphics[width=0.469\textwidth]{f2b.pdf}
+			\includegraphics[width=0.49\textwidth]{f3a.pdf}
+			\includegraphics[width=0.469\textwidth]{f3b.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.7\textwidth]{f2c.pdf}
+			\includegraphics[width=0.7\textwidth]{f3c.pdf}
 		}
@@
 A final possible confounding factor -- binarity -- is known to affect
-the density of stars that fall beneath the slow sequence
+the ``fast sequence'' of stars beneath the slow sequence
 \citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}.  We do
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f10a.pdf}
-			\includegraphics[width=0.48\textwidth]{f10b.pdf}
+			\includegraphics[width=0.49\textwidth]{f4a.pdf}
+			\includegraphics[width=0.48\textwidth]{f4b.pdf}
 		}
@@
     candidate $\delta$\,Lyr cluster
-    members from Figure~\ref{fig:XYZvtang}.  Since other members of the
+    members.  Since other members of the
     cluster with similar brightnesses have comparable degrees of
@@
 		\subfloat{
-			\includegraphics[width=\textwidth]{f3a.pdf}
+			\includegraphics[width=\textwidth]{f5a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.5\textwidth]{f3b.pdf}
-			\includegraphics[width=0.5\textwidth]{f3c.pdf}
+			\includegraphics[width=0.5\textwidth]{f5b.pdf}
+			\includegraphics[width=0.5\textwidth]{f5c.pdf}
 		}
@@
 Previous analyses have identified and
-characterized the hidden transit signal
+characterized the smaller transit signal
 \citep{2012ApJS..199...24T,thompson_planetary_2018}, validated its
@@
 Our model therefore included 10 free parameters for the transit ($\{P,
-t_0, R_{\rm p}/R_\star, b, u_1 ,u_2 ,R_\star, \log g, e, \omega \}$),
+t_0, \delta, b, u_1 ,u_2 ,R_\star, \log g, e, \omega \}$),
 2 free parameters for the light curve normalization and a white noise 
@@
 The model parameters and their uncertainties, given in Table~\ref{tab:posterior}, are
-broadly consistent with a mini-Neptune sized planet ($3.36\pm0.18\,R_\oplus$) on a close-in circular\footnote{ Our transit fitting
+broadly consistent with a mini-Neptune sized planet ($3.82\pm0.16\,R_\oplus$) on a close-in circular\footnote{ Our transit fitting
 	yields $e<0.48$ at 2-$\sigma$; the constraints on the eccentricity are
 	not particularly strong.} orbit around a G8V host star ($0.88 \pm 0.02
-R_\odot$).  Our best-fit planet size is smaller than that
+R_\odot$).  This best-fit planet size is consistent with those
 previously reported by \citet{morton_false_2016} and
-\citet{berger_identifying_2018}, at 0.7-$\sigma$ and 1.2-$\sigma$
-confidence respectively.  The former is explained by differing
-assumptions about the stellar radius; the latter appears to be due to
-differences in the measured transit depth, which could be linked to
-the different methods used to account for the stellar rotation signal.
+\citet{berger_identifying_2018}, and corrects for the small
+amount of flux dilution from Kepler\,1627B.
+%at 0.3-$\sigma$ and 0.2-$\sigma$ respectively.
 
@@
 represents a $\approx 3\%$ distortion of the transit depth
-($\approx1700\,{\rm ppm}$).  
+($1759\pm62\,{\rm ppm}$).  
 Note that
@@
 finds that the asymmetry also seems to be robust to different methods of
-data binning.  The most likely astrophysical explanations are
-discussed in
-Section~\ref{sec:conc}.
+data binning in time and by local light curve slope.  The most likely
+astrophysical explanations are discussed in Section~\ref{sec:conc}.
 
@@
 		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f4a.pdf}
+		\includegraphics[width=0.47\textwidth]{f6.pdf}
 	\end{center}
@@
 		positive (negative) TTV, provided that the orbit is prograde
-		\citep{mazeh_time_2015}.  The units along the abscissa are most
-		easily understood by considering that the stellar flux changes by
+		\citep{mazeh_time_2015}.  The units along the abscissa can
+		be understood by considering that the stellar flux changes by
 		$\sim$60\,ppt per half rotation period ($\sim$1.3\,days).
@@
 false positive scenario was that of a background eclipsing binary,
-which had a model-dependent probability of $\approx10^{-5}$.  However,
+which had a model-dependent probability of $\approx$$10^{-5}$.  However,
 this calculation was performed without knowledge of the low-mass stellar companion ($M_{\rm B}\approx
@@
 For the sake of argument, assume the former value.
-The observed transit has a depth of $\approx$0.17\%.  A
-17\% deep eclipse of the secondary star would therefore be needed to
+The observed transit has a depth of $\approx$0.18\%.  A
+18\% deep eclipse of the secondary star would therefore be needed to
 produce a signal with the appropriate depth.  The shape of the
-transit signal however requires the impact parameter to be below 0.74
-(Table~\ref{tab:posterior}); the tertiary transiting the secondary would
+transit signal however requires the impact parameter to be below 0.77
+(2-$\sigma$); the tertiary transiting the secondary would
 therefore need to be non-grazing with $R_3/R_2 \approx 0.4$.
@@
 egress phase that each span $\approx$40\% of the transit duration
-($\approx 65\,{\rm minutes}$).  The actual measured ingress and egress
-duration is $\approx 15\,{\rm minutes}$, 4.4$\times$ shorter.  The
-combination of Kepler 1627B's brightness, the transit depth, and the ingress
-duration therefore disfavor the scenario that Kepler 1627B might host the
-transit signal.
+($\approx 68\,{\rm minutes}$).  The actual measured ingress and egress
+duration is $\approx 17\,{\rm minutes}$, a factor of four times too
+short.  The combination of Kepler 1627B's brightness, the transit
+depth, and the ingress duration therefore disfavor the scenario that
+Kepler 1627B might host the transit signal.
 
@@
 eclipsing the M dwarf secondary at $10$-$\sigma$.
+While the planet might hypothetically still orbit a hidden close and
+bright companion, such scenarios are ruled out by {\it i)} a lack of
+secondary lines in the HIRES spectra and {\it ii)} a lack of
+secondary rotation signals in the Kepler photometry.
 
@@
 		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f4b.pdf}
+		\includegraphics[width=0.47\textwidth]{f7.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.75\textwidth]{f5.pdf}
+		\includegraphics[width=0.74\textwidth]{f8.pdf}
 	\end{center}
@@
 geometry suggested by the TTV-local slope correlation.  Separately,
-transit spectroscopy aimed at detecting atmospheric outflows might
+transit spectroscopy aimed at detecting atmospheric outflows could
 yield insight into the evolutionary state of the atmosphere
@@
 long-term monitoring of the transit times and durations,
-given the technical difficulty of performing a high-cadence multi-color radial velocity campaign
-on a $V=13.1$ target.
+given the technical difficulty of performing a multi-color radial velocity campaign
+on a $V\approx13$ star.
 Measured at sufficient precision though, the mass, combined with the known
@@
 asymmetry is also consistent with predictions
-\citep{wang_dai_2019}.  Beyond the asymmetric transits,
+\citep{wang_dai_2019}.  
+Beyond the asymmetric transits,
 Appendix~\ref{app:flare} highlights an additional abnormality in the Kepler
@@
 \citep{rizzuto_tess_2020}, TOI~837
-\citep{bouma_cluster_2020}, the two-planet AU~Mic system
+\citep{bouma_cluster_2020}, the two-planet AU~Mic
 \citep{plavchan_planet_2020,martioli_aumicbc_2021} and the four-planet
-V1298~Tau system \citep{david_four_2019}.  \pn\ is one of the smallest
-planets in this sample ($3.36\pm0.18\,R_\oplus$), which could be linked to the selection effects
-imposed by the spot-induced photometric variability at very young ages
-\citep[{\it e.g.},][]{zhou_2021_tois}.  If these young
-planets have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$,
-then their low densities would be in accord with
-the expectation that mini-Neptunes start their lives with large
-primordial atmospheres that are shed over the first gigayear
-\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  The
-prograde orbit, if confirmed, would be consistent with a quiescent
-disk-driven migration history, and would differ from misalignments that
-have been observed for some older
-hot Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
+V1298~Tau \citep{david_four_2019}.  \pn\ is one of the smaller
+planets in this sample ($3.82\pm0.16\,R_\oplus$), which could be
+linked to the selection effects imposed by spot-induced
+photometric variability at very young ages \citep[{\it
+e.g.},][]{zhou_2021_tois}.  
+However,
+smaller planets could have been detected:
+the Kepler pipeline's median completeness extended down to
+$1.6\,R_\oplus$ at 10~day orbital periods, and $3.3\,R_\oplus$ at
+100~days \citep{2021ascl.soft07027B}.
+The large size of \pn\ relative to the typical Kepler mini-Neptune might
+therefore support a picture in which the typical mini-Neptune
+($5\,M_\oplus$ to 10$\,M_\oplus$; \citealt{wu_mass_2019}) loses a
+significant fraction of its primordial atmosphere over its first gigayear
+\citep{Owen_Wu_2013,ginzburg_corepowered_2018}.  
+% The
+% prograde orbit, if confirmed, would be consistent with a quiescent
+% disk-driven migration history, and would differ from misalignments that
+% have been observed for some older
+% hot Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
 %Spectroscopic follow-up observations of the transit remain necessary
@@
 %correct orbital geometry.
-
-%NB to arxiv readers: The alternative scenario of in-situ formation is
+% The alternative scenario of in-situ formation is
 %disfavored on theoretical grounds: the required disk mass would
@@
 cross-matches for both known and forthcoming exoplanet
-systems \citep[{\it e.g.},][]{guerrero_tess_2021}.  The entry shown in
-the header, HD 150706b, is one example: the
-host star appears to be a member of the Ursa Major moving group
-($400\,{\rm Myr}$; \citealt{mann_tess_2020}).  Confirming
-this suggested kinematic association with independent age indicators is
-essential because the false positive rates are not known.
-Alternatively, one could instead opt to identify new kinematic
+systems \citep[{\it e.g.},][]{guerrero_tess_2021}. Confirming
+these candidate associations using independent age indicators is
+essential because their false positive rates are not known.
+A related path is to identify new kinematic
 associations around known exoplanet host stars using positions and
@@
 e.g.},][]{tofflemire_tess_2021}.  
-Each path seems likely to expand the census of planets with precisely
-measured ages over the coming years.
+Each approach seems likely to expand the census of planets with precisely
+measured ages over the coming years, which will help in
+deciphering the early stages of exoplanet evolution.
 
@@
 		\leavevmode
-		\includegraphics[width=0.64\textwidth]{f7.pdf}
+		\includegraphics[width=0.64\textwidth]{f9.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.64\textwidth]{f8.pdf}
+		\includegraphics[width=0.64\textwidth]{f10.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f9.pdf}
+		\includegraphics[width=0.9\textwidth]{f11.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f9b.pdf}
+		\includegraphics[width=0.9\textwidth]{f12.pdf}
 	\end{center}
@@
 		\subfloat{
-			\includegraphics[width=\textwidth]{f11.pdf}
+			\includegraphics[width=\textwidth]{f13.pdf}
 		}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f12.pdf}
+		\includegraphics[width=0.9\textwidth]{f14.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f13.pdf}
+		\includegraphics[width=0.8\textwidth]{f15.pdf}
 	\end{center}

### 2021-10-07 paper/collect.sh

@@ -1,17 +1,23 @@
-cp /Users/luke/Dropbox/proj/rudolf/results/skychart/skychart_showkepler_showcomovers.pdf f1.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/XYZvtang/XYZvtang_show1627_showcomovers_showsun.pdf f1.pdf
 
-cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_overplotkep1627.pdf f2a.pdf
-cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_smalllims_overplotkep1627.pdf f2b.pdf
-cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_curtiscleaning_overplotkep1627.pdf f2c.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/skychart/skychart_showkepler_showcomovers.pdf f2.pdf
 
-cp /Users/luke/Dropbox/proj/rudolf/results/keplerlc/keplerlc_Nsample500_xlim0500_0600.pdf f3a.pdf
-cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_posterior_phaseplot.pdf f3b.pdf
-cp /Users/luke/Dropbox/proj/rudolf/results/run_RotGPtransit/Kepler_1627_RotGPtransit_posterior_phaseplot_fullxlim.pdf f3c.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_overplotkep1627.pdf f3a.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_IC_2602_Pleiades_cleanhrcut_gaia2018_smalllims_overplotkep1627.pdf f3b.pdf
[…]

### 2021-10-08 paper/ms.tex
update ms

@@
 \title{
-  %Kepler 1627Ab: A 38 Million Year Old
-  %Kepler 1627 Ab:
-  %A Mini-Neptune in the Kepler Field Only 38 Million Years Old
   A 38 Million Year Old Mini-Neptune in the Kepler Field
@@
   observed by the Kepler space telescope because it is nearby
-  ($d=329\,{\rm pc}$) and it resembles the Sun.  Here we show
-  using Gaia kinematics, TESS stellar rotation periods, and
-  spectroscopic lithium abundances that Kepler\,1627 is a member of
-  the \clusterage\ old $\delta$~Lyr cluster.  To our knowledge, this makes
-  Kepler\,1627Ab the youngest planet with a precise age yet found
-  by the main Kepler mission.
-  The Kepler photometry shows two 
-  peculiarities: the average transit profile is
-  asymmetric, and the individual transit times are correlated with the
-  local light curve slope.
-  We discuss possible explanations for each anomaly.
-  More broadly applicable though is the fact that the $\delta$~Lyr
-  cluster is one of $\sim$10$^3$ clusters whose properties have been
-  clarified by Gaia.  Many other exoplanet hosts are candidate
-  kinematic members of these clusters; their cluster memberships can
-  be verified by measuring stellar rotation periods in the TESS full
-  frame images.
+  ($d=329\,{\rm pc}$) and it resembles the Sun.  Here we show using
+  Gaia kinematics, TESS stellar rotation periods, and spectroscopic
+  lithium abundances that Kepler\,1627 is a member of the \clusterage\
+  old $\delta$~Lyr cluster.  To our knowledge, this makes
+  Kepler\,1627Ab the youngest planet with a precise age yet found by
+  the main Kepler mission.  The Kepler photometry shows two
+  peculiarities: the average transit profile is asymmetric, and the
+  individual transit times are correlated with the local light curve
+  slope.  We discuss possible explanations for each anomaly.  More
+  broadly, the $\delta$~Lyr cluster
+  is one of $\sim$10$^3$ coeval groups whose properties have been clarified
+  by Gaia.  Many other exoplanet hosts are potentially kinematic members
+  of these clusters; their candidate memberships can be verified by
+  measuring stellar rotation periods in the TESS full frame images.
 \end{abstract}
@@
 
-%TODO: might want to revise this text to better frame the boil-off phase...
-%TODO: better citations on radius evolution?
-For instance, the ``mini-Neptunes'', thought to be made of molten
-rocky cores \citep{kite_atmosphere_2020} and extended atmospheric
-envelopes of hydrogen and helium, are expected to shrink in size by
-factors of several over their first $10^8$ years.  Specifically, in
-the models of \citet{owen_atmospheres_2016} and
+For instance, the ``mini-Neptunes'', thought to be made of metal
+cores, silicate mantles \citep{kite_atmosphere_2020}, and extended
+hydrogen-dominated atmospheres, are expected to shrink
+in size by factors of several over their first $10^8$ years.
+Specifically, in the models of \citet{owen_atmospheres_2016} and
 \citet{owen_constraining_2020}, the $\approx5\,M_\oplus$ planets start
@@
 2--4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
-is expected to occur within the first few million years after disk
-dispersal \citep{owen_atmospheres_2016}, the combination of stellar
-irradiation and internal heat may also power a more gradual outflow
-which can eventually deplete or entirely strip the envelope from the
-rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
+is expected to occur within the first few million years after the disk
+disperses \citep{ikoma_situ_2012}, stellar
+irradiation and internal heat can also power gradual outflows
+which, if strong enough, can deplete or entirely strip the atmospheric
+envelope \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
 Discovering young planets, measuring their masses, and detecting their
@@
 %: NGC\,6866, NGC\,6811, NGC\,6819, and NGC\,6791,
-with ages spanning 0.7\,Gyr to 9\,Gyr
-\citep{meibom_kepler_2011}.  Though isochronal, gyrochronal, and lithium-based
-analyses suggest that younger Kepler planets do exist
+with ages spanning 0.7\,Gyr to 9\,Gyr \citep{meibom_kepler_2011}.
+Though isochronal, gyrochronal, and lithium-based analyses suggest
+that younger Kepler planets do exist
 \citep{berger_identifying_2018,david_sizes_2021}, accurate and precise
 age measurements typically require an ensemble of stars.  Fortunately,
-recent analyses of the Gaia
-data have greatly expanded our knowledge of cluster memberships
-\citep[{\it
+recent analyses of the Gaia data have greatly expanded our knowledge
+of cluster memberships \citep[{\it
 e.g.},][]{CantatGaudin2018a,Zari2018,KounkelCovey2019,Meingast2021,Kerr2021}.
@@
 
-Matching our young star list against stars observed by Kepler revealed that
-%two discoveries.  The first, to be discussed in an upcoming analysis
-%by J.~Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
-%open cluster Theia~520 (UBC~1).  At least six Kepler planets are known
-%in the cluster from the Kepler-52 and Kepler-968 systems
-%\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
-%discovery, and the focus of this work, is
-Kepler observed a portion of the \cn\ (Stephenson-1;
-Theia~73).
-More specifically, a clustering analysis of the Gaia
-data by \citet{KounkelCovey2019}
-reported that Kepler~1627 (KIC 6184894; KOI 5245; TIC 120105470) is a
-\cn\ member.  Given the
-previous statistical validation of the close-in mini-Neptune Kepler
-1627b
+Matching our young star list against stars observed by Kepler revealed
+that Kepler observed a portion of the \cn\ (Stephenson-1; Theia~73).
+More specifically, a clustering analysis of the Gaia data by
+\citet{KounkelCovey2019} reported that Kepler~1627 (KIC 6184894; KOI
+5245; TIC 120105470) is a \cn\ member.  Given the previous statistical
+validation of the close-in mini-Neptune Kepler 1627b
 \citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
-we begin by examining the properties of the cluster (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
-cluster is \clusterage\ old, and 
-in 
-Section~\ref{sec:stars} show
-that \sn\ is both a binary and also a member of the cluster.
-Focusing on the planet (Section~\ref{sec:planet}), we confirm that despite the
+we begin by examining the properties of the cluster
+(Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr cluster is
+\clusterage\ old, and in Section~\ref{sec:stars} show that \sn\ is
+both a binary and also a member of the cluster.  Focusing on the
+planet (Section~\ref{sec:planet}), we confirm that despite the
 existence of the previously unreported M2.5V companion, hereafter
-Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A.  Based on a correlation between the
-planetary transit times and the local light curve slope, the orbit
-may also be prograde.  We conclude by highlighting broader
-implications for our ability to age-date a larger sample of planets
-(Section~\ref{sec:conc}).
+Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A.
+Based on a correlation between the planetary transit times and the
+local light curve slope, the orbit may also be prograde.  We conclude
+by highlighting broader implications for our ability to age-date a
+larger sample of planets (Section~\ref{sec:conc}).
 
@@
     $\delta$\,Lyr cluster.} Points are reported cluster members from
-		\citet{KounkelCovey2019}.  
-		The tangential velocities relative to
-		\sn\ (lower-right) are computed assuming that every star has the same
-		three-dimensional spatial velocity as \sn.
-		Our analysis considers stars (black points)
-		in the spatial and kinematic vicinity of Kepler\,1627 (yellow
-		star).  The question of whether the other candidate cluster members
-		(gray points) are part of the cluster is outside our scope.  The
-		location of the Sun is ($\odot$) is shown.
+    \citet{KounkelCovey2019}.  The tangential velocities relative to
+    \sn\ (lower-right) are computed assuming that every star has the
+    same three-dimensional spatial velocity as \sn.  Our analysis
+    considers stars (black points) in the spatial and kinematic
+    vicinity of Kepler\,1627 (yellow star).  The question of whether
+    the other candidate cluster members (gray points) are part of the
+    cluster is outside our scope.  The location of the Sun is
+    ($\odot$) is shown.
 		\label{fig:XYZvtang}
@@
 
-To measure the age of the $\delta$ Lyr cluster, we first selected
-a set of candidate cluster members
-(Section~\ref{sec:kinematicselection}), and then analyzed these stars using a
-combination of the isochronal and gyrochronal techniques
+To measure the age of the $\delta$ Lyr cluster, we first selected a
+set of candidate cluster members
+(Section~\ref{sec:kinematicselection}), and then analyzed these stars
+using a combination of the isochronal and gyrochronal techniques
 (Section~\ref{sec:clusterage}).
@@
 
-\citet{KounkelCovey2019} applied an unsupervised clustering
-algorithm to Gaia DR2 on-sky positions, proper motions, and
-parallaxes for stars within the nearest kiloparsec.  For the \cn\
-(Theia 73), they reported \noriginal\ candidate members.  We
-matched these stars against the latest Gaia EDR3 observations using
-the \texttt{dr2\_neighbourhood} table from the ESA archive,\footnote{See
+\citet{KounkelCovey2019} applied an unsupervised clustering algorithm
+to Gaia DR2 on-sky positions, proper motions, and parallaxes for stars
+within the nearest kiloparsec.  For the \cn\ (Theia 73), they reported
+\noriginal\ candidate members.  We matched these stars against the
+latest Gaia EDR3 observations using the \texttt{dr2\_neighbourhood}
+table from the ESA archive,\footnote{See
 \url{https://gea.esac.esa.int/archive/documentation/GEDR3/Gaia_archive/chap_datamodel/sec_dm_auxiliary_tables/ssec_dm_dr2_neighbourhood.html}.}
-and took the closest proper motion and epoch-corrected angular
-distance as the presumed match
-\citep{gaia_collaboration_2021_edr3}.  In Figure~\ref{fig:XYZvtang},
-have shown galactic positions only for the stars with
-parallax signal-to-noise exceeding 20.  The reported cluster members
-(gray and black points) extend over a much larger volume than the
-cluster previously identified by \citet{stephenson_possible_1959} and
-later corroborated by \citet{eggen_photometric_1968}.  While the
-non-uniform ``clumps'' of stars might comprise a {\it bona fide}
-cluster of identically-aged stars, they could also be heavily
-contaminated by field stars.  We therefore considered stars only in
-the immediate kinematic and spatial vicinity of \sn\ as candidate
-cluster members.  We performed the selection cuts manually, by
-drawing lassos with the interactive \texttt{glue} visualization tool
-\citep{beaumont_2014_13866} in the four projections shown in
-Figure~\ref{fig:XYZvtang}.  The overlap between the Kepler
-field and the resulting candidate cluster members is shown in
-Figure~\ref{fig:skychart}.  While this method will
-include some field interlopers in the ``cluster star'' sample, and
-vice-versa, it should suffice for our aim of verifying the existence
-of the cluster in the vicinity of \sn.
+taking the stars closest in proper motion and epoch-corrected angular
+distance as the presumed match \citep{gaia_collaboration_2021_edr3}.
+In Figure~\ref{fig:XYZvtang}, have shown galactic positions only for
+the stars with parallax signal-to-noise exceeding 20.  The reported
+cluster members (gray and black points) extend over a much larger
+volume than the cluster previously identified by
+\citet{stephenson_possible_1959} and later corroborated by
+\citet{eggen_photometric_1968}.  While the non-uniform ``clumps'' of
+stars might comprise a {\it bona fide} cluster of identically-aged
+stars, they could also be heavily contaminated by field stars.  We
+therefore considered stars only in the immediate kinematic and spatial
+vicinity of \sn\ as candidate cluster members.  We performed the
+selection cuts manually, by drawing lassos with the interactive
+\texttt{glue} visualization tool \citep{beaumont_2014_13866} in the
+four projections shown in Figure~\ref{fig:XYZvtang}.  The overlap
+between the Kepler field and the resulting candidate cluster members
+is shown in Figure~\ref{fig:skychart}.  While this method will include
+some field interlopers in the ``cluster star'' sample, and vice-versa,
+it should suffice for our aim of verifying the existence of the
+cluster in the vicinity of \sn.
 
@@
 	\caption{
-		{\bf Kepler's view of the $\delta$~Lyr cluster.} Each black circle is a
-		candidate cluster member selected from the catalog of \citet{KounkelCovey2019} based on its
-		position and kinematics.
-		Of the \nkinematic\ candidate cluster members, 58 have at least one
+    {\bf Kepler's view of the $\delta$~Lyr cluster.} Each black circle
+    is a candidate cluster member selected from the catalog of
+    \citet{KounkelCovey2019} based on its position and kinematics.  Of
+    the \nkinematic\ candidate cluster members, 58 have at least one
     quarter of Kepler data.  TESS has also observed most of the
@@
     reference \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most
-		candidate \cn\ members are gyrochronally younger than the Pleiades; outliers are
-		probably field interlopers.
+    candidate \cn\ members are gyrochronally younger than the
+    Pleiades; outliers are probably field interlopers.
 		\label{fig:age}
@@
 panel of Figure~\ref{fig:age} shows the color-absolute magnitude
-diagram (CAMD) of candidate \cn\ members, IC\,2602, the Pleiades, and the
-field.  
-The stars from the Pleiades and IC\,2602 were adopted from
-\citet{CantatGaudin2018a}, and the field stars are from the Gaia
-EDR3 Catalog of Nearby Stars \citep{gaia_gcns_2021}.  We cleaned these
-following the data filtering criteria from
-\citet[][Appendix B]{GaiaCollaboration2018}, except that we weakened the parallax
+diagram (CAMD) of candidate \cn\ members, IC\,2602, the Pleiades, and
+the field.  The stars from the Pleiades and IC\,2602 were adopted from
+\citet{CantatGaudin2018a}, and the field stars are from the Gaia EDR3
+Catalog of Nearby Stars \citep{gaia_gcns_2021}.  We cleaned these
+following the data filtering criteria from \citet[][Appendix
+B]{GaiaCollaboration2018}, except that we weakened the parallax
 precision requirement to $\varpi/\sigma_\varpi>5$.  These filters were
@@
 3-dimensional maps of \citet{capitanio_threedimensional_2017} and
-\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}, 2021/09/25},
-and applied the extinction coefficients $k_X\equiv A_X/A_0$
-computed by \citet{GaiaCollaboration2018} assuming that $A_0 = 3.1
-E(B-V)$.  For IC\,2602, the Pleiades, and the \cn, this procedure
-yielded a respective mean and standard deviation for the reddening of
-$E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$.  These
-values agree reasonably well with previously reported values from the
-literature \citep[{\it
+\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/},
+2021/09/25}, and applied the extinction coefficients $k_X\equiv
+A_X/A_0$ computed by \citet{GaiaCollaboration2018} assuming that $A_0
+= 3.1 E(B-V)$.  For IC\,2602, the Pleiades, and the \cn, this
+procedure yielded a respective mean and standard deviation for the
+reddening of $E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$.
+These values agree reasonably well with previously reported values
+from the literature \citep[{\it
 e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}.
@@
 below the 80$^{\rm th}$ percentile of each cluster's distribution, and
-excluded stars that were obvious photometric binaries in the
-CAMD.  We then passed a moving box average and standard deviation
-across the CAMD in 0.10\,mag bins, fitted a univariate spline to the
-binned values, and assembled a piecewise grid of hybrid isochrones
-spanning the ages between UCL to the Pleiades using Equations~6 and~7
-from \citet{gagne_mutau_2020}.
-
-% Randich+18: 43.7 +4.3 - 3.9 LDB, for E(B-V) = 0.068 +/- 0.025
-% Dobbie+10: 46 +6 -5 LDB, for
+excluded stars that were obvious photometric binaries in the CAMD.  We
+then passed a moving box average and standard deviation across the
+CAMD in 0.10\,mag bins, fitted a univariate spline to the binned
+values, and assembled a piecewise grid of hybrid isochrones spanning
+the ages between UCL to the Pleiades using Equations~6 and~7 from
+\citet{gagne_mutau_2020}.
+
+% Randich+18: 43.7 +4.3 - 3.9 LDB
+% Dobbie+10: 46 +6 -5 LDB
 
 The ages returned by this procedure depend on the ages assumed for
-each reference cluster.  We adopted a 115\,Myr age for the Pleiades \citep{dahm_2015},
-and a 16\,Myr age for UCL \citep{pecaut_star_2016}.  The age
-of IC\,2602 however is the most important ingredient, since it
+each reference cluster.  We adopted a 115\,Myr age for the Pleiades
+\citep{dahm_2015}, and a 16\,Myr age for UCL \citep{pecaut_star_2016}.
+The age of IC\,2602 however is the most important ingredient, since it
 receives the most weight in the interpolation.  Plausible ages for
-IC\,2602 span 30\,Myr to 46\,Myr, with older ages being
-preferred by the lithium-depletion-boundary (LDB) measurements 
-\citep{dobbie_ic_2010,randich_gaiaeso_2018}
-and
-younger ages by the main-sequence turn-off
-\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.
-If we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would
-be $31^{+5}_{-4}$\,Myr old.  For the converse extreme of 46\,Myr, the
+IC\,2602 span 30\,Myr to 46\,Myr, with older ages being preferred by
+the lithium-depletion-boundary (LDB) measurements
+\citep{dobbie_ic_2010,randich_gaiaeso_2018} and younger ages by the
+main-sequence turn-off
+\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.  If
+we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would be
+$31^{+5}_{-4}$\,Myr old.  For the converse extreme of 46\,Myr, the
 \cn\ would be $44^{+8}_{-7}$\,Myr old.  We adopt an intermediate
@@
 Of the \noriginal\ candidate \cn\ members reported by
-\citet{KounkelCovey2019}, \nwithtess\ stars were amenable to
-rotation period measurements ($G<17$ and $\bpmrpo>0.5$) using the TESS
-full frame image data.\footnote{Kepler rotation periods were derived
-by \citet{mcquillan_rotation_2014} for 28 stars; for these cases, we
+\citet{KounkelCovey2019}, \nwithtess\ stars were amenable to rotation
+period measurements ($G<17$ and $\bpmrpo>0.5$) using the TESS full
+frame image data.\footnote{Kepler rotation periods were derived by
+\citet{mcquillan_rotation_2014} for 28 stars; for these cases, we
 adopted the Kepler rotation period.  } We extracted light curves from
@@
 IC\,2602 and IC\,2391) are indistinguishable
-\citep{douglas_stephanie_t_2021_5131306}. 
-Approximately 10 of the \cn\ stars appear as outliers above
-the ``slow sequence''.  Assuming that they are all false positives
-({\it i.e.}, field interlopers), our rotation period detection
-fraction would be $\nrotgood/\nkinwithtessandcrowding \approx
-\nfracprot\%$.   
-The other stars are likely to be field contaminants.
-A final possible confounding factor -- binarity -- is known to affect
-the ``fast sequence'' of stars beneath the slow sequence
+\citep{douglas_stephanie_t_2021_5131306}.  Approximately 10 of the
+\cn\ stars appear as outliers above the ``slow sequence''.  Assuming
+that they are all false positives ({\it i.e.}, field interlopers), our
+rotation period detection fraction would be
+$\nrotgood/\nkinwithtessandcrowding \approx \nfracprot\%$.   The other
+stars are likely to be field contaminants.  A final possible
+confounding factor -- binarity -- is known to affect the ``fast
+sequence'' of stars beneath the slow sequence
 \citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}.  We do
@@
 \subsubsection{Age}
+% rotation period from plot_rotation_period_windowslider.py
 Based on the spatial and kinematic association of \sn\ with the \cn,
 and the assumption that the planet formed shortly after the star, it
-seems likely that \sn\ is the same age as the cluster. There are
-two consistency checks on whether this is true: rotation and lithium.  Based on the Kepler
-light curve, the rotation period is $2.642\pm0.042$\,days, where the
-quoted uncertainty is based on the scatter in rotation periods
-measured from each individual Kepler quarter.  This is consistent with
-comparable cluster members (Figure~\ref{fig:age}).
-
-To infer the abundance of Li\,\textsc{I} from the 6708\,\AA\ doublet
+seems likely that \sn\ is the same age as the cluster. There are two
+consistency checks on whether this is true: rotation and lithium.
+Based on the Kepler light curve, the rotation period is
+$2.642\pm0.042$\,days, where the quoted uncertainty is based on the
+scatter in rotation periods measured from each individual Kepler
+quarter.  This is consistent with comparable cluster members
+(Figure~\ref{fig:age}).
+
+To infer the amount of Li\,\textsc{I} from the 6708\,\AA\ doublet
 \citep[{\it e.g.},][]{soderblom_ages_2014}, we acquired an iodine-free
@@
 temperature are found by interpolating against a 38$\,$Myr MIST
-isochrone \citep{choi_mesa_2016}.  
-The statistical uncertainties are propagated from the absolute
-magnitude (mostly originating from the parallax uncertainty) and the color; the systematic
-uncertainties are taken to be the difference between the PARSEC
+isochrone \citep{choi_mesa_2016}.  The statistical uncertainties are
+propagated from the absolute magnitude (mostly originating from the
+parallax uncertainty) and the color; the systematic uncertainties are
+taken to be the difference between the PARSEC
 \citep{bressan_parsec_2012} and MIST isochrones.  Reported
 uncertainties are a quadrature sum of the statistical and systematic
-components.
-As a consistency check, 
-we analyzed the aforementioned Keck/HIRES spectrum from
-the night of 2021 March 26 using a combination of
-\texttt{SpecMatch-Emp} for stellar properties, and
-\texttt{SpecMatch-Synth} for $v\sin i$
-\citep{yee_SM_2017}.
-This procedure yielded
-$T_{\rm eff}=5498\pm100\,{\rm K}$,
-$\log g=4.6\pm0.1$, $[{\rm Fe/H}]=0.15\pm0.10$ 
-from \texttt{SpecMatch-Emp}, and $v\sin i = 18.9\pm1.0$ from
-\texttt{SpecMatch-Synth}.
-These values are within the $1$-$\sigma$ uncertainties of our adopted values from
-the isochrone interpolation.
+components.  As a consistency check, we analyzed the aforementioned
+Keck/HIRES spectrum from the night of 2021 March 26 using a
+combination of \texttt{SpecMatch-Emp} for stellar properties, and
+\texttt{SpecMatch-Synth} for $v\sin i$ \citep{yee_SM_2017}.  This
+procedure yielded $T_{\rm eff}=5498\pm100\,{\rm K}$, $\log
+g=4.6\pm0.1$, $[{\rm Fe/H}]=0.15\pm0.10$ from \texttt{SpecMatch-Emp},
+and $v\sin i = 18.9\pm1.0$ from \texttt{SpecMatch-Synth}.  These
+values are within the $1$-$\sigma$ uncertainties of our adopted values
+from the isochrone interpolation.
 
@@
     projected separation of $53\pm4$\,AU.  The inset shows a cutout of
-    the stacked NIRC2 image (North is up, East is left, scale is set by the
-    separation of the binary).  The lines show 5-$\sigma$ contrast
-    limits for the `Alopeke filters, and 6-$\sigma$ contrast limits
-    for NIRC2 outside of $0\farcs15$. {\it Right:} Gaia
-    EDR3 renormalized unit weight error (RUWE) point estimates for
-    candidate $\delta$\,Lyr cluster
-    members.  Since other members of the
-    cluster with similar brightnesses have comparable degrees of
-    photometric variability, the high RUWE independently suggests that Kepler\,1627
-    is a binary. 
+    the stacked NIRC2 image (North is up, East is left, scale is set
+    by the separation of the binary).  The lines show 5-$\sigma$
+    contrast limits for the `Alopeke filters, and 6-$\sigma$ contrast
+    limits for NIRC2 outside of $0\farcs15$. {\it Right:} Gaia EDR3
+    renormalized unit weight error (RUWE) point estimates for
+    candidate $\delta$\,Lyr cluster members.  Since other members of
+    the cluster with similar brightnesses have comparable degrees of
+    photometric variability, the high RUWE independently suggests that
+    Kepler\,1627 is a binary. 
     \label{fig:kep1627binary}
@@
 
-We first noted the presence of a close neighbor in the \sn\ system on 2015 July 22 when
-we acquired adaptive optics imaging using the NIRC2 imager on
-Keck-II.
-We used the narrow camera (FOV = 10.2\arcsec) to obtain 8
-images in the $K'$ filter ($\lambda = 2.12\,\mu$m) with a total
-exposure time of 160\,s. We analyzed these data following
-\citet{kraus_impact_2016}, which entailed using PSF-fitting to measure the separation,
-position angle, and contrast of the candidate companion. 
-The best-fitting empirical PSF template was identified from among the
-near-contemporaneous observations of single stars in the same filter.
-The mean values inferred from the 8 images are reported in
-Table~\ref{tab:starparams}.
-To estimate the detection limits, we analyzed the residuals after subtracting
-the empirical PSF template. Within each residual
-image, the flux was measured through 40\,mas apertures centered on
-every pixel, and then the noise as a function of radius was estimated
-from the RMS within concentric rings. Finally, the detection limits
-were estimated from the strehl-weighted sum of the detection
-significances in the image stack, and we adopted the $6$-$\sigma$
-threshold as the detection limit for ruling out additional companions.
+We first noted the presence of a close neighbor in the \sn\ system on
+2015 July 22 when we acquired adaptive optics imaging using the NIRC2
+imager on Keck-II.  We used the narrow camera (FOV = 10.2\arcsec) to
+obtain 8 images in the $K'$ filter ($\lambda = 2.12\,\mu$m) with a
+total exposure time of 160\,s. We analyzed these data following
+\citet{kraus_impact_2016}, which entailed using PSF-fitting to measure
+the separation, position angle, and contrast of the candidate
+companion.  The best-fitting empirical PSF template was identified
+from among the near-contemporaneous observations of single stars in
+the same filter.  The mean values inferred from the 8 images are
+reported in Table~\ref{tab:starparams}.  To estimate the detection
+limits, we analyzed the residuals after subtracting the empirical PSF
+template. Within each residual image, the flux was measured through
+40\,mas apertures centered on every pixel, and then the noise as a
+function of radius was estimated from the RMS within concentric rings.
+Finally, the detection limits were estimated from the strehl-weighted
+sum of the detection significances in the image stack, and we adopted
+the $6$-$\sigma$ threshold as the detection limit for ruling out
+additional companions.
 
@@
 
+% \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
 Figure~\ref{fig:kep1627binary} summarizes the results of the
@@
 brighter than $G < 20.5$ mag within $\rho < 120\arcsec$), its
-existence was also suggested by the primary star's large renormalized unit weight error (RUWE),
-relative to other members of the \cn.  Based on the apparent
-separation, the binary orbital period is of order hundreds of years.
-The large RUWE is therefore more likely to be caused by a PSF-mismatch
-skewing the Gaia centroiding during successive scans, rather than true
-astrometric motion.  Regardless, given the low geometric probability that a companion imaged at
-$\rho \approx 0\farcs16$ is a chance line-of-sight companion, we
-proceed under the assumption that the companion is bound, and that
-Kepler\,1627 is a binary.  
-Given the distance and age, the models of
+existence was also suggested by the primary star's large renormalized
+unit weight error (RUWE), relative to other members of the \cn.  Based
+on the apparent separation, the binary orbital period is of order
+hundreds of years.  The large RUWE is therefore more likely to be
+caused by a PSF-mismatch skewing the Gaia centroiding during
+successive scans, rather than true astrometric motion.  Regardless,
+given the low geometric probability that a companion imaged at $\rho
+\approx 0\farcs16$ is a chance line-of-sight companion, we proceed
+under the assumption that the companion is bound, and that
+Kepler\,1627 is a binary.  Given the distance and age, the models of
 \citet{baraffe_new_2015} imply a companion mass of $M_{\rm B} \approx
@@
 3450$ K.  The corresponding spectral type is roughly M2.5V
-\citep{pecaut_mamajek_2013}.
-% \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
-These models combined with the NIRC2 contrast limits imply
-physical limits on tertiary companions of $M_{\rm ter} < 50
-M_{\rm Jup}$ at $\rho = 50$\,AU, $M_{\rm ter} < 20 M_{\rm Jup}$ at $\rho =
-100$\,AU, and $M_{\rm ter} < 10 M_{\rm Jup}$ at $\rho = 330$\,AU.
+\citep{pecaut_mamajek_2013}.  These models combined with the NIRC2
+contrast limits imply physical limits on tertiary companions of
+$M_{\rm ter} < 50 M_{\rm Jup}$ at $\rho = 50$\,AU, $M_{\rm ter} < 20
+M_{\rm Jup}$ at $\rho = 100$\,AU, and $M_{\rm ter} < 10 M_{\rm Jup}$
+at $\rho = 330$\,AU.
 
@@
     {\it Top}: 
-    The Kepler data span 1{,}437 days (3.9 years), sampled at
-    30 minute cadence;  a 100 day segment is shown.  The
-    top panel shows the \texttt{PDCSAP} median-subtracted flux in
-    units of parts-per-thousand ($\times 10^{-3}$).  The dominant
-    signal is induced by starspots.  The stellar
-    variability model (orange line) is subtracted below, revealing the
-    transits of \pn.  The
-    online Figure Set spans the entire 3.9 years of observations.
+    The Kepler data span 1{,}437 days (3.9 years), sampled at 30
+    minute cadence;  a 100 day segment is shown.  The top panel shows
+    the \texttt{PDCSAP} median-subtracted flux in units of
+    parts-per-thousand ($\times 10^{-3}$).  The dominant signal is
+    induced by starspots.  The stellar variability model (orange line)
+    is subtracted below, revealing the transits of \pn.  The online
+    Figure Set spans the entire 3.9 years of observations.
     {\it Bottom}:
@@
 2009 May 2 until 2013 April 8.  Data gaps during quarters 4, 9, and 13
-led to an average duty cycle over the 3.9~year interval of 78\%.  \sn\
+led to an average duty cycle over the 3.9~year interval of 67\%.  \sn\
 was also observed at 1-minute cadence from 2012 Oct 5 until 2013 Jan
-11.  The top panel of Figure~\ref{fig:lc} shows a portion of the 30-minute
-cadence \texttt{PDCSAP} light curve.  Nonastrophysical variability has
-been removed using the methods discussed by
+11.  The top panel of Figure~\ref{fig:lc} shows a portion of the
+30-minute cadence \texttt{PDCSAP} light curve.  Nonastrophysical
+variability has been removed using the methods discussed by
 \citet{smith_kepler_PDC_2017}; the default optimal aperture was
-assumed \citep{smith_finding_2016}. 
-Cadences with non-zero quality flags (9\% of the data) have been omitted.
- The resulting photometry is dominated by
-a quasi-periodic starspot signal with a peak-to-peak amplitude that
-varies between 2\% and 8\%.  
-Previous analyses have identified and
-characterized the smaller transit signal
+assumed \citep{smith_finding_2016}.  Cadences with non-zero quality
+flags (9\% of the data) have been omitted.  The resulting photometry
+is dominated by a quasi-periodic starspot signal with a peak-to-peak
+amplitude that varies between 2\% and 8\%.  Previous analyses have
+identified and characterized the smaller transit signal
 \citep{2012ApJS..199...24T,thompson_planetary_2018}, validated its
@@
 system for transit timing variations \citep{holczer_transit_2016}.
-Nonetheless, since the cluster membership provides us
-with more precise stellar parameters than those previously available,
-we opted to reanalyze the
-light curve.
+Nonetheless, since the cluster membership provides us with more
+precise stellar parameters than those previously available, we opted
+to reanalyze the light curve.
 
@@
 with characteristic frequencies set by 1/$P_{\rm rot}$ and its first
-harmonic.  We additionally included a jitter term to inflate
-the flux uncertainties in a manner that accounted for otherwise
-unmodeled excess white noise, and let the eccentricity float.  For
-the limb-darkening, we assumed a quadratic law, and sampled using the
+harmonic.  We additionally included a jitter term to inflate the flux
+uncertainties in a manner that accounted for otherwise unmodeled
+excess white noise, and let the eccentricity float.  For the
+limb-darkening, we assumed a quadratic law, and sampled using the
 uninformative prior suggested by \citet{exoplanet:kipping13}.
@@
 Our model therefore included 10 free parameters for the transit ($\{P,
-t_0, \delta, b, u_1 ,u_2 ,R_\star, \log g, e, \omega \}$),
-2 free parameters for the light curve normalization and a white noise 
-jitter ($\{\langle f \rangle, \sigma_f \}$), and 5
-hyperparameters for the GP ($\{\sigma_{\mathrm{rot}},
-P_{\mathrm{rot}}, Q_0, \mathrm{d}Q, f \}$).  We also considered
-including an additive \texttt{SHOTerm} kernel to account for
-stochastic noise, but found that this did not affect the
+t_0, \delta, b, u_1 ,u_2 ,R_\star, \log g, e, \omega \}$), 2 free
+parameters for the light curve normalization and a white noise jitter
+($\{\langle f \rangle, \sigma_f \}$), and 5 hyperparameters for the GP
+($\{\sigma_{\mathrm{rot}}, P_{\mathrm{rot}}, Q_0, \mathrm{d}Q, f \}$).
+We also considered including an additive \texttt{SHOTerm} kernel to
+account for stochastic noise, but found that this did not affect the
 results, and so opted for the simpler GP kernel.  We fitted the models
@@
 
-Figure~\ref{fig:lc} shows the resulting best-fit model in orange (top) and
-purple (bottom).
-The model parameters and their uncertainties, given in Table~\ref{tab:posterior}, are
-broadly consistent with a mini-Neptune sized planet ($3.82\pm0.16\,R_\oplus$) on a close-in circular\footnote{ Our transit fitting
-	yields $e<0.48$ at 2-$\sigma$; the constraints on the eccentricity are
-	not particularly strong.} orbit around a G8V host star ($0.88 \pm 0.02
-R_\odot$).  This best-fit planet size is consistent with those
-previously reported by \citet{morton_false_2016} and
-\citet{berger_identifying_2018}, and corrects for the small
-amount of flux dilution from Kepler\,1627B.
-%at 0.3-$\sigma$ and 0.2-$\sigma$ respectively.
+Figure~\ref{fig:lc} shows the resulting best-fit model in orange (top)
+and purple (bottom).  The model parameters and their uncertainties,
+given in Table~\ref{tab:posterior}, are broadly consistent with a
+mini-Neptune sized planet ($3.82\pm0.16\,R_\oplus$) on a close-in
+circular\footnote{ Our transit fitting yields $e<0.48$ at 2-$\sigma$;
+the constraints on the eccentricity are not particularly strong.}
+orbit around a G8V host star ($0.88 \pm 0.02 R_\odot$).  This best-fit
+planet size is consistent with those previously reported by
+\citet{morton_false_2016} and \citet{berger_identifying_2018}, and
+corrects for the small amount of flux dilution from Kepler\,1627B.
 
 The transit fit however is not perfect: the lower panels of
-Figure~\ref{fig:lc} show an asymmetric
-residual in the data relative to the model: the measured flux is 
-high during the first half of transit, and low in the second half.  The
-semi-amplitude of this deviation is $\approx50\,{\rm ppm}$, which
-represents a $\approx 3\%$ distortion of the transit depth
-($1759\pm62\,{\rm ppm}$).  
-Note that
-while this asymmetry is within the 2-$\sigma$ model
-uncertainties, the model has a jitter term that grows to
-account for otherwise unmodeled white noise in the flux.  The
-significance of the asymmetry is therefore best assessed in comparison
-against the intrinsic out-of-transit scatter in the data, not the model
-uncertainties.
-
-To determine whether the asymmetry is a systematic caused by our stellar variability model,
-we explored an alternative
-approach in which we isolated each transit window and locally fitted
-out polynomial trends, and then binned all the observed transits; the
+Figure~\ref{fig:lc} show an asymmetric residual in the data relative
+to the model: the measured flux is high during the first half of
+transit, and low in the second half.  The semi-amplitude of this
+deviation is $\approx50\,{\rm ppm}$, which represents a $\approx 3\%$
+distortion of the transit depth ($1759\pm62\,{\rm ppm}$).  Note that
+while this asymmetry is within the 2-$\sigma$ model uncertainties, the
+model has a jitter term that grows to account for otherwise unmodeled
+white noise in the flux.  The significance of the asymmetry is
+therefore best assessed in comparison against the intrinsic
+out-of-transit scatter in the data, not the model uncertainties.
+
+To determine whether the asymmetry is a systematic caused by our
+stellar variability model, we explored an alternative approach in
+which we isolated each transit window and locally fitted out
+polynomial trends, and then binned all the observed transits; the
 asymmetry was still present at a comparable amplitude.
-Appendix~\ref{app:asymmetry} includes a deeper analysis, and
-finds that the asymmetry also seems to be robust to different methods of
-data binning in time and by local light curve slope.  The most likely
+Appendix~\ref{app:asymmetry} includes a deeper analysis, and finds
+that the asymmetry also seems to be robust to different methods of
+data binning in time and by local light curve slope.  Possible
 astrophysical explanations are discussed in Section~\ref{sec:conc}.
@@
 		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f6.pdf}
+		\includegraphics[width=0.49\textwidth]{f6.pdf}
 	\end{center}
-	\vspace{-0.7cm}
+	\vspace{-0.6cm}
 	\caption{
-		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
-		The time of each Kepler transit was measured, along with the local
-		slope of the light curve.  The two quantities appear
-		anti-correlated, which is most easily explained by starspot
-		crossings during the first (second) half of transit inducing a
-		positive (negative) TTV, provided that the orbit is prograde
-		\citep{mazeh_time_2015}.  The units along the abscissa can
-		be understood by considering that the stellar flux changes by
-		$\sim$60\,ppt per half rotation period ($\sim$1.3\,days).
+    {\bf Evidence for a prograde orbit of Kepler 1627 Ab.} The time of
+    each Kepler transit was measured, along with the local slope of
+    the light curve.  The two quantities appear anti-correlated, which
+    is most easily explained by starspot crossings during the first
+    (second) half of transit inducing a positive (negative) TTV,
+    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
+    units along the abscissa can be understood by considering that the
+    stellar flux changes by $\sim$60\,ppt per half rotation period
+    ($\sim$1.3\,days).
 		\label{fig:corr}
@@
 \end{figure}
-
-% rotation period from plot_rotation_period_windowslider.py
 
@@
 Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and V1298 Tau
-($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}   
-Could the transit be produced by anything other than a planet
-orbiting this near-solar analog?  \citet{morton_false_2016} validated
-the planet based on the transit shape, arguing that the most probable
-false positive scenario was that of a background eclipsing binary,
-which had a model-dependent probability of $\approx$$10^{-5}$.  However,
-this calculation was performed without knowledge of the low-mass stellar companion ($M_{\rm B}\approx
-0.33\,M_\odot$).  
-Validated planets have also previously been refuted \citep[{\it e.g.},][]{shporer_three_2017}.
-We therefore reassessed
-false positive scenarios in some detail.
-
-% (0.7-2.0\% V-band to G-band dmag difference from EEM table; NOTE
-% could determine using model spectra)
-As an initial plausibility check,
-Kepler\,1627B contributes 1\% to 2\% of the total flux observed in
-the Kepler aperture.  
-For the sake of argument, assume the former value.
-The observed transit has a depth of $\approx$0.18\%.  A
-18\% deep eclipse of the secondary star would therefore be needed to
-produce a signal with the appropriate depth.  The shape of the
-transit signal however requires the impact parameter to be below 0.77
-(2-$\sigma$); the tertiary transiting the secondary would
-therefore need to be non-grazing with $R_3/R_2 \approx 0.4$.
-%Assuming a $\approx 0.46R_\odot$ radius of the imaged secondary
-%from the \citet{baraffe_new_2015} models, this
-%would imply a tertiary stellar radius of $\approx 0.2R_\odot$.
-This
-yields a contradiction:  this scenario would require an ingress and
-egress phase that each span $\approx$40\% of the transit duration
+($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}   Could the transit
+be produced by anything other than a planet orbiting this near-solar
+analog?  \citet{morton_false_2016} validated the planet based on the
+transit shape, arguing that the most probable false positive scenario
+was that of a background eclipsing binary, which had a model-dependent
+probability of $\approx$$10^{-5}$.  However, this calculation was
+performed without knowledge of the low-mass stellar companion ($M_{\rm
+B}\approx 0.33\,M_\odot$).  Validated planets have also previously
+been refuted \citep[{\it e.g.},][]{shporer_three_2017}.  We therefore
+reassessed false positive scenarios in some detail.
+
+% (0.7-2.0\% V-band to G-band dmag difference from EEM table)
+
+As an initial plausibility check, Kepler\,1627B contributes 1\% to 2\%
+of the total flux observed in the Kepler aperture.  For the sake of
+argument, assume the former value.  The observed transit has a depth
+of $\approx$0.18\%.  A 18\% deep eclipse of Kepler\,1627B would
+therefore be needed to produce a signal with the appropriate depth.
+The shape of the transit signal however requires the impact parameter
+to be below 0.77 (2-$\sigma$); the tertiary transiting the secondary
+would therefore need to be non-grazing with $R_3/R_2 \approx 0.4$.
+This yields a contradiction:  this scenario would require an ingress
+and egress phase that each span $\approx$40\% of the transit duration
 ($\approx 68\,{\rm minutes}$).  The actual measured ingress and egress
@@
 % Rstar = 0.455 --> 0.465 Rsun
-%
-
-% At 30 Myr, a 0.44\,$M_\odot$ solar-metallicity dwarf is $\approx 26\%$
-% larger than when it is fully contracted on the main sequence
-% (0.40\,$R_\odot$ vs{.} 0.50\,$R_\odot$; ).
 
 There are two other new lines of evidence that confirm the planetary
-interpretation.  First, the transit duration and the orbital period
-are inconsistent with an eclipsing body around the M-dwarf companion:
+interpretation.  First, the stellar density implied by the transit duration and orbital period
+is inconsistent with an eclipsing body around the M-dwarf companion:
 we find $\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
-theoretically expected density for the companion is $\approx 4.6\,{\rm
+theoretically expected density for Kepler\,1627B is $\approx 4.6\,{\rm
 g\,cm}^{-3}$ (Table~\ref{tab:posterior}; \citealt{choi_mesa_2016}).
 The transit duration is therefore too long to be explained by a star
-eclipsing the M dwarf secondary at $10$-$\sigma$.
-While the planet might hypothetically still orbit a hidden close and
-bright companion, such scenarios are ruled out by {\it i)} a lack of
-secondary lines in the HIRES spectra and {\it ii)} a lack of
-secondary rotation signals in the Kepler photometry.
+eclipsing the M dwarf secondary at $10$-$\sigma$.  While the planet
+might hypothetically still orbit a hidden close and bright companion,
+this possibility is ruled out by {\it i)} a lack of secondary lines in
+the HIRES spectra and {\it ii)} a lack of secondary rotation signals
+in the Kepler photometry.
 
@@
 %Kepler observations also mitigate the cadence-induced smearing, and
-%enable a more precise measurement of the transit impact parameter
-%\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
-%(2-$\sigma$) from the 1-minute data, compared to $<0.76$ (2-$\sigma$)
-%from the 30-minute data.  This implies an even stronger constraint on
-%the transit shape in an analysis of the false positive probability.
+%enable a more precise measurement of the transit impact parameter.
 
@@
 mid-time and that of a periodic orbit.  This residual, the transit
-timing variation (TTV), is plotted in Figure~\ref{fig:corr}
-against the local linear slope.  A significant correlation of $-0.0564
-\pm 0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
+timing variation (TTV), is plotted in Figure~\ref{fig:corr} against
+the local linear slope.  A significant correlation of $-0.0564 \pm
+0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
 Objects of Interest have shown this correlation
@@
 		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f7.pdf}
+		\includegraphics[width=0.49\textwidth]{f7.pdf}
 	\end{center}
@@
 	\caption{
-		{\bf Radial velocities acquired using Keck/HIRES during the transit of
-			2021 Aug 7, observed simultaneously with MuSCAT3.}  Shaded bands
-		show 2-$\sigma$ model uncertainties.  The photometric transit
-		depths are consistent across the {\it griz} bandpasses.  The
-		best-fitting model in the top panel includes the RM effect and a quadratic trend
-		in time to model the spot-induced $\approx$\,$250$\ms\ trend (see Appendix~\ref{app:rm}).  The radial velocity jitter
-		over the 15~minute exposure timescale ($\sigma_{\rm RV}\approx30$\ms) prevented us from detecting the RM
-		effect.
-		The RV variations are strongly correlated with varying emission in the Ca H and K lines (middle panel).
+    {\bf Radial velocities acquired using Keck/HIRES during the
+    transit of 2021 Aug 7, observed simultaneously with MuSCAT3.}
+    Shaded bands show 2-$\sigma$ model uncertainties.  The photometric
+    transit depths are consistent across the {\it griz} bandpasses.
+    The best-fitting model in the top panel includes the RM effect and
+    a quadratic trend in time to model the spot-induced
+    $\approx$\,$250$\ms\ trend (see Appendix~\ref{app:rm}).  The
+    radial velocity jitter over the 15~minute exposure timescale
+    ($\sigma_{\rm RV}\approx30$\ms) prevented us from detecting the RM
+    effect.  The RV variations are strongly correlated with varying
+    emission in the Ca H and K lines (middle panel).
 		\label{fig:ground}
@@
 A third supporting line of evidence for the planetary interpretation
-exists, though we consider it less definitive than the stellar
-density and TTV-local slope correlation.  We observed a transit of
-\pn\ on the night of 2021 Aug 7 simultaneously with Keck/HIRES and
-MuSCAT3.  We scheduled the observations using the ephemeris of
+exists, though we consider it less definitive than the stellar density
+and TTV-local slope correlation.  We observed a transit of \pn\ on the
+night of 2021 Aug 7 simultaneously with Keck/HIRES and MuSCAT3.  We
+scheduled the observations using the ephemeris of
 \citet{holczer_transit_2016}.  Although we did not detect the
 Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
-show the transit (Figure~\ref{fig:ground}).  Fitting
-the MuSCAT3 photometry with a model that lets the transit depths vary
-across each bandpass, we find {\it griz} depths consistent with the
-Kepler depth at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  The
-MuSCAT3 observations also suggest a transit duration $17.4\pm3.6\ {\rm
-minutes}$ shorter than the Kepler transits; further photometric
-follow-up could help confirm whether the transit duration is indeed
-changing.  For our RM analysis, the details are discussed in
-Appendix~\ref{app:rm}.  While the velocities are marginally more
-consistent with a prograde or polar orbit than a retrograde orbit, the
-spot-corrected exposure-to-exposure scatter ($\sigma_{\rm RV}\approx
-30$\ms) is comparable to the expected velocity anomaly assuming an
-aligned orbit ($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$).  We
-are therefore not in a position to claim a spectroscopic detection of
-the RM effect, nor to quantify the stellar obliquity.
+show the transit (Figure~\ref{fig:ground}).  Fitting the MuSCAT3
+photometry with a model that lets the transit depths vary across each
+bandpass, we find {\it griz} depths consistent with the Kepler depth
+at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  Conditioned on the
+Kepler ephemeris and transit depth, the MuSCAT3 observations may also
+show a transit duration $17.3\pm4.0\ {\rm minutes}$ shorter than the
+Kepler transits; further photometric follow-up would help confirm
+whether the transit duration is indeed changing.  For our RM analysis,
+the details are discussed in Appendix~\ref{app:rm}.  While the
+velocities are marginally more consistent with a prograde or polar
+orbit than a retrograde orbit, the spot-corrected exposure-to-exposure
+scatter ($\sigma_{\rm RV}\approx 30$\ms) is comparable to the expected
+RM anomaly assuming an aligned orbit ($\Delta v_{\rm RM}\approx
+20\,{\rm m\,s}^{-1}$).  We are therefore not in a position to claim a
+spectroscopic detection of the RM effect, nor to quantify the stellar
+obliquity.
 
@@
 Kepler\,1627Ab provides a new extremum in the ages of the Kepler
-planets, and opens multiple avenues for further study.  
-Observations of spectroscopic transits at greater precision should yield a
-measurement of the stellar obliquity, which would confirm or refute the prograde orbital
-geometry suggested by the TTV-local slope correlation.  Separately,
-transit spectroscopy aimed at detecting atmospheric outflows could
-yield insight into the evolutionary state of the atmosphere
-\citep[{\it
+planets, and opens multiple avenues for further study.  Observations
+of spectroscopic transits at greater precision should yield a
+measurement of the stellar obliquity, which would confirm or refute
+the prograde orbital geometry suggested by the TTV-local slope
+correlation.  Separately, transit spectroscopy aimed at detecting
+atmospheric outflows could yield insight into the evolutionary state
+of the atmosphere \citep[{\it
 e.g.},][]{ehrenreich_giant_2015,spake_helium_2018,vissapragada_2020}.
@@
 quantity to measure would be the planet's mass.  The most feasible
-approach would likely be
-long-term monitoring of the transit times and durations,
-given the technical difficulty of performing a multi-color radial velocity campaign
-on a $V\approx13$ star.
-Measured at sufficient precision though, the mass, combined with the known
-age and size, would yield constraints on both the planet's composition
-and its initial entropy \citep{owen_constraining_2020}.
+approach would likely be long-term monitoring of the transit times and
+durations, given the technical difficulty of performing a multi-color
+radial velocity campaign on a $V=13$ star.  Measured at
+sufficient precision though, the mass, combined with the known age and
+size, would yield constraints on both the planet's composition and its
+initial entropy \citep{owen_constraining_2020}.
 
@@
 asymmetry shown in Figure~\ref{fig:lc} is that of a dusty asymmetric
-outflow.  
-A second possibility is that the planetary orbit is slightly misaligned
-from the stellar spin axis, and tends to
-transit starspot groups at
-favored stellar latitudes.
-Other possibilities such as gravity darkening or TTVs causing
-the asymmetry are disfavored
-(see Appendix~\ref{app:asymmetry}).  Dusty outflows are theoretically
+outflow.  A second possibility is that the planetary orbit is slightly
+misaligned from the stellar spin axis, and tends to transit starspot
+groups at favored stellar latitudes.  Other possibilities such as
+gravity darkening or TTVs causing the asymmetry are disfavored (see
+Appendix~\ref{app:asymmetry}).  Dusty outflows are theoretically
 expected for young mini-Neptunes, and the amplitude of the observed
-asymmetry is also consistent with predictions
-\citep{wang_dai_2019}.  
-Beyond the asymmetric transits,
-Appendix~\ref{app:flare} highlights an additional abnormality in the Kepler
-data, in the arrival time distribution of stellar flares.  We
-encourage its exploration by investigators more versed in the topic
-than ourselves.
+asymmetry is also consistent with predictions \citep{wang_dai_2019}.
+Beyond the asymmetric transits, Appendix~\ref{app:flare} highlights an
+additional abnormality in the Kepler data, in the arrival time
+distribution of stellar flares.  We encourage its exploration by
+investigators more versed in the topic than ourselves.
 
@@
 \citep{benatti_possibly_2019,newton_tess_2019} , HIP~67522
-\citep{rizzuto_tess_2020}, TOI~837
-\citep{bouma_cluster_2020}, the two-planet AU~Mic
-\citep{plavchan_planet_2020,martioli_aumicbc_2021} and the four-planet
-V1298~Tau \citep{david_four_2019}.  \pn\ is one of the smaller
-planets in this sample ($3.82\pm0.16\,R_\oplus$), which could be
-linked to the selection effects imposed by spot-induced
+\citep{rizzuto_tess_2020}, TOI~837 \citep{bouma_cluster_2020}, the
+two-planet AU~Mic \citep{plavchan_planet_2020,martioli_aumicbc_2021}
+and the four-planet V1298~Tau \citep{david_four_2019}.  \pn\ is one of
+the smaller planets in this sample ($3.82\pm0.16\,R_\oplus$), which
+could be linked to the selection effects imposed by spot-induced
 photometric variability at very young ages \citep[{\it
-e.g.},][]{zhou_2021_tois}.  
-However,
-smaller planets could have been detected:
-the Kepler pipeline's median completeness extended down to
+e.g.},][]{zhou_2021_tois}.  However, smaller planets could have been
+detected: the Kepler pipeline's median completeness extended to
 $1.6\,R_\oplus$ at 10~day orbital periods, and $3.3\,R_\oplus$ at
-100~days \citep{2021ascl.soft07027B}.
-The large size of \pn\ relative to the typical Kepler mini-Neptune might
-therefore support a picture in which the typical mini-Neptune
-($5\,M_\oplus$ to 10$\,M_\oplus$; \citealt{wu_mass_2019}) loses a
-significant fraction of its primordial atmosphere over its first gigayear
+100~days \citep{2021ascl.soft07027B}.  The large size of \pn\ relative
+to most Kepler mini-Neptunes might therefore support a picture
+in which the typical $5\,M_\oplus$ to 10$\,M_\oplus$ mini-Neptune
+\citep{wu_mass_2019} loses a significant fraction of its primordial
+atmosphere over its first gigayear
 \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
-% The
-% prograde orbit, if confirmed, would be consistent with a quiescent
-% disk-driven migration history, and would differ from misalignments that
-% have been observed for some older
-% hot Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
+
+% The prograde orbit, if confirmed, would be consistent with a
+% quiescent disk-driven migration history, and would differ from
+% misalignments that have been observed for some older hot
+% Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
 %Spectroscopic follow-up observations of the transit remain necessary
 %however to verify whether the TTV-slope correlation is yielding the
-%correct orbital geometry.
-% The alternative scenario of in-situ formation is
-%disfavored on theoretical grounds: the required disk mass would
-%likely drive migration, and so lacks internal consistency
+%correct orbital geometry.  The alternative scenario of in-situ
+%formation is disfavored on theoretical grounds: the required disk
+%mass would likely drive migration, and so lacks internal consistency
 %\citep{inamdar_formation_2015,ogihara_reassessment_2015}.
@@
 connection of the star to the \cn\ using Gaia kinematics, and then by
-using the Gaia color-absolute magnitude diagram and TESS stellar rotation periods to verify
-the cluster's existence.  Table~\ref{tab:v06} enables similar
-cross-matches for both known and forthcoming exoplanet
-systems \citep[{\it e.g.},][]{guerrero_tess_2021}. Confirming
-these candidate associations using independent age indicators is
-essential because their false positive rates are not known.
-A related path is to identify new kinematic
-associations around known exoplanet host stars using positions and
-tangential velocities from Gaia, and to then verify these associations
-with stellar rotation periods and spectroscopy \citep[{\it
-e.g.},][]{tofflemire_tess_2021}.  
-Each approach seems likely to expand the census of planets with precisely
-measured ages over the coming years, which will help in
-deciphering the early stages of exoplanet evolution.
+using the Gaia color-absolute magnitude diagram and TESS stellar
+rotation periods to verify the cluster's existence.
+Table~\ref{tab:v06} enables similar cross-matches for both known and
+forthcoming exoplanet systems \citep[{\it
+e.g.},][]{guerrero_tess_2021}. Confirming these candidate associations
+using independent age indicators is essential because their false
+positive rates are not known.  A related path is to identify new
+kinematic associations around known exoplanet host stars using
+positions and tangential velocities from Gaia, and to verify
+these associations with stellar rotation periods and spectroscopy
+\citep[{\it e.g.},][]{tofflemire_tess_2021}.  Each approach seems
+likely to expand the census of planets with precisely measured ages
+over the coming years, which will help in deciphering the early stages
+of exoplanet evolution.
 
@@
 T{.}~David for illuminating discussions and suggestions, and to
-R{.}~Kerr for providing us with the \citet{Kerr2021} membership
-list prior to its publication.
-The authors are also grateful to K{.}~Collins for helping resolve the
-scheduling conflict that would have otherwise prevented the MuSCAT3
-observations.
-% is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
-% E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
-% K{.}~Hawkins for the discussions on young stars, rotation, and lithium
-% that encouraged this analysis.
+R{.}~Kerr for providing us with the \citet{Kerr2021} membership list
+prior to its publication.  The authors are also grateful to
+K{.}~Collins for helping resolve the scheduling conflict that would
+have otherwise prevented the MuSCAT3 observations.
 %
@@
 greater precision than that of a typical FGK field star, through
-either isochronal, gyrochronal, or spectroscopic techniques.
-As in
+either isochronal, gyrochronal, or spectroscopic techniques.  As in
 \citet{bouma_cdipsI_2019}, we collected stars that met these criteria
 from across the literature.  Table~\ref{tab:metadata} gives a list of
-the studies included, and brief summary statistics.
-The age measurement methodologies adopted by each study differ: in many,
+the studies included, and brief summary statistics.  The age
+measurement methodologies adopted by each study differ: in many,
 spatial and kinematic clustering has been performed on the Gaia data,
-and ensemble isochrone fitting of the resulting clusters has been performed
-(typically focusing on the turn-off).
-In other studies however, the claim of youth is based on the location of a
-single star in the color-absolute magnitude diagram, or on spectroscopic
+and ensemble isochrone fitting of the resulting clusters has been
+performed (typically focusing on the turn-off).  In other studies
+however, the claim of youth is based on the location of a single star
+in the color-absolute magnitude diagram, or on spectroscopic
 information.
@@
 One major change in Table~\ref{tab:v06} relative to the earlier
-iteration from \citet{bouma_cdipsI_2019} is that the extent of Gaia-based
-analyses has now matured to the point that we can neglect pre-Gaia
-cluster memberships, except for a few cases with spectroscopically
-confirmed samples of age-dated stars.  The membership lists for
-instance of \citet{Kharchenko_et_al_2013} and \citet{dias_proper_2014}
-(MWSC and DAML) are no longer required.  This is helpful for various
-post-processing projects,  since the field star contamination rates
-were typically much higher in these catalogs than in the newer
-Gaia-based catalogs.
+iteration from \citet{bouma_cdipsI_2019} is that the extent of
+Gaia-based analyses has now matured to the point that we can neglect
+pre-Gaia cluster memberships, except for a few cases with
+spectroscopically confirmed samples of age-dated stars.  The
+membership lists for instance of \citet{Kharchenko_et_al_2013} and
+\citet{dias_proper_2014} (MWSC and DAML) are no longer required.  This
+is helpful for various post-processing projects,  since the field star
+contamination rates were typically much higher in these catalogs than
+in the newer Gaia-based catalogs.
 
 The most crucial parameters of a given star for our purposes are the
-Gaia DR2 $\texttt{source\_id}$, the cluster or
-group name ($\texttt{cluster}$), and the $\texttt{age}$.  Given the
-hierarchical nature of many stellar associations, we do not attempt to
-resolve the cluster names to a single unique string.  The Orion
-complex for instance can be divided into almost one hundred kinematic
-subgroups \citep{kounkel_apogee2_2018}.  Similar complexity applies to
-the problem of determining homogeneous ages, which we do not attempt
-to resolve.  Instead, we simply merged the cluster names and ages
+Gaia DR2 $\texttt{source\_id}$, the cluster or group name
+($\texttt{cluster}$), and the $\texttt{age}$.  Given the hierarchical
+nature of many stellar associations, we do not attempt to resolve the
+cluster names to a single unique string.  The Orion complex for
+instance can be divided into almost one hundred kinematic subgroups
+\citep{kounkel_apogee2_2018}.  Similar complexity applies to the
+problem of determining homogeneous ages, which we do not attempt to
+resolve.  Instead, we simply merged the cluster names and ages
 reported by various authors into a comma-separated string.
@@
 This means that the \texttt{age} column can be null, for cases in
-which the original authors did not report an age, or for which a reference
-literature age was not readily available.  Nonetheless, since we do
-prefer stars with known ages, we made a few additional
+which the original authors did not report an age, or for which a
+reference literature age was not readily available.  Nonetheless,
+since we do prefer stars with known ages, we made a few additional
 efforts to populate this column.  When available, the age provenance
-is from the original analysis of the cluster.  In a few cases
-however we adopted other ages when string-based cross-matching on the
-cluster name was straightforward.  In particular, we used the ages
-determined by \citet{CantatGaudin2020b} to assign ages to the clusters
-from \citet{GaiaCollaboration2018}, \citet{CantatGaudin2018a},
+is from the original analysis of the cluster.  In a few cases however
+we adopted other ages when string-based cross-matching on the cluster
+name was straightforward.  In particular, we used the ages determined
+by \citet{CantatGaudin2020b} to assign ages to the clusters from
+\citet{GaiaCollaboration2018}, \citet{CantatGaudin2018a},
 \citet{CastroGinard2020}, and \citet{CantatGaudin2020a}.
@@
 For any of the catalogs for which Gaia DR2 identifiers were not
-available, we either followed the spatial (plus
-proper-motion) cross-matching procedures described in
-\citet{bouma_cdipsI_2019}, or else we pulled the Gaia DR2 source
-identifiers associated with the catalog from SIMBAD.  We consequently
-opted to drop the $\texttt{ext\_catalog\_name}$ and $\texttt{dist}$
-columns maintained in \citet{bouma_cdipsI_2019}, as these were only
-populated for a small number of stars.
-The technical manipulations for the merging, cleaning, and joining
-were performed using $\texttt{pandas}$
+available, we either followed the spatial (plus proper-motion)
+cross-matching procedures described in \citet{bouma_cdipsI_2019}, or
+else we pulled the Gaia DR2 source identifiers associated with the
+catalog from SIMBAD.  We consequently opted to drop the
+$\texttt{ext\_catalog\_name}$ and $\texttt{dist}$ columns maintained
+in \citet{bouma_cdipsI_2019}, as these were only populated for a small
+number of stars.  The technical manipulations for the merging,
+cleaning, and joining were performed using $\texttt{pandas}$
 \citep{mckinney-proc-scipy-2010}.  The eventual cross-match (using the
@@
 website.\footnote{\url{https://gea.esac.esa.int/archive/}}
-
-% \section{Kinematic Selection of $\delta$\,Lyr Cluster Members}
-% \label{app:kinematicselection}
-% 
-% \begin{figure*}[t]
-% 	\begin{center}
-% 		\leavevmode
-% 		\includegraphics[width=1\textwidth]{f6.pdf}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-% 	\caption{
-%     {\bf Galactic position and tangential velocities of the
-%     $\delta$\,Lyr cluster.}
-%     Points are candidate cluster members with $\varpi/\sigma_\varpi >
-%     20$, reported to be in the group by
-%     \citet{KounkelCovey2019}.  We focus on stars (black points)
-%     in the spatial and kinematic vicinity of Kepler\,1627 (yellow
-%     star).  The other candidate cluster members (gray points) may or
-%     may not share the ages of the selected kinematic group.  The
-%     location of the Sun is ($\odot$) is shown.
-% 		\label{fig:XYZvtang}
-% 	}
-% \end{figure*}
-% 
-% Figure~\ref{fig:XYZvtang} shows stars reported by
-% \citet{KounkelCovey2019} to be in the group ``Theia 73'', which
-% was cross-matched by Kounkel \& Covey as being ``Stephenson 1''.
-% \citet{KounkelCovey2019} reported \noriginal\ stars to be
-% present in this cluster.  For our Figure, galactic positions were
-% calculated and plotted only for stars with parallax signal-to-noise
-% exceeding 20.  The location of the Sun is shown on the plots.  The
-% smattering of reported cluster members (gray and black points) has a
-% significantly different structure than the cluster initially
-% identified by \citet{stephenson_possible_1959} and corroborated by
-% \citet{eggen_photometric_1968}.  While the non-uniform ``clumps''
-% might be part of a real structure, they could also be an artifact of
-% the data processing steps performed by
-% \citet{KounkelCovey2019}.  We therefore opted to only consider
-% stars in the immediate kinematic and spatial vicinity of \sn.  The
-% tangential velocities relative to \sn\ are shown in the bottom right
-% panel.  These are computed by assuming that every star has the same
-% three-dimensional spatial velocity as \sn, where we assume a systemic
-% radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the HIRES
-% spectra.  The
-% relevant projection effects are then taken into account, as discussed
-% by {\it e.g.}, \citet{Meingast2021} and \citet{bouma_2021_ngc2516}.
-% We performed the actual selection by then manually drawing lassos with
-% the interactive \texttt{glue} visualization tool
-% \citep{beaumont_2014_13866} in the four projections shown in
-% Figure~\ref{fig:XYZvtang}.  While other cluster members likely exist
-% outside of our selection region (and our selection also includes some
-% field star interlopers), our aim is to verify the existence of the
-% cluster in the vicinity of Kepler\,1627, and to measure its age.  The
-% procedure we have adopted enables both tasks.
 
@@
     Phase-folded transit of Kepler 1627b, with stellar variability
-    removed.  Black points are binned to 20
-    minute intervals.  The 2-$\sigma$ model uncertainties and the
-    maximum {\it a posteriori} model are shown as the faint purple
-    band, and the dark purple line.
+    removed.  Black points are binned to 20 minute intervals.  The
+    2-$\sigma$ model uncertainties and the maximum {\it a posteriori}
+    model are shown as the faint purple band, and the dark purple
+    line.
     {\it Right:}
@@
     Phase-folded transit of Kepler 1627b, with stellar variability
-    removed.  Points and models are as
-    in Figure~\ref{fig:phasequarter}.
+    removed.  Points and models are as in
+    Figure~\ref{fig:phasequarter}.
     {\it Right:}
@@
 	\caption{
-    {\bf Transit models and residuals, binned by quartiles in the local slope of the
-    light curve}.  
+    {\bf Transit models and residuals, binned by quartiles in the
+    local slope of the light curve}.  
     Representative uncertainties for the black points (binned at 20
@@
 Figures~\ref{fig:phasequarter},~\ref{fig:phaseyear},~\ref{fig:phaseslope},
-and~\ref{fig:phasettv} show the Kepler data binned in four ways:
-over Kepler quarters, Julian years, quartiles of local slope, and quartiles of
-transit timing variation.  Over Kepler quarters
-(Figure~\ref{fig:phasequarter}), Quarter 6 shows the strongest asymmetry: a deviation of about 3\,ppt from expectation.
-Quarter 7 shows an anomaly at roughly the same transit phase.  Year 2
+and~\ref{fig:phasettv} show the Kepler data binned in four ways: over
+Kepler quarters, Julian years, quartiles of local slope, and quartiles
+of transit timing variation.  Over Kepler quarters
+(Figure~\ref{fig:phasequarter}), Quarter 6 shows the strongest
+asymmetry: a deviation of about 3\,ppt from expectation.  Quarter 7
+shows an anomaly at roughly the same transit phase.  Year 2
 correspondingly shows the strongest anomaly out of any year in
@@
 slope quartiles (Figure~\ref{fig:phaseslope}), the asymmetry is
-visually present in three of the four quartiles: the only bin in
-which it does not appear is ${\rm d}f/{\rm
-d}t\in[0.001,0.033]\,{\rm ppt\,day}^{-1}$.  The corresponding bins in TTV
-(largest TTV corresponding to smallest ${\rm d}f/{\rm d}t$)
-show the same asymmetries, though with somewhat greater scatter
+visually present in three of the four quartiles: the only bin in which
+it does not appear is ${\rm d}f/{\rm d}t\in[0.001,0.033]\,{\rm
+ppt\,day}^{-1}$.  The corresponding bins in TTV (largest TTV
+corresponding to smallest ${\rm d}f/{\rm d}t$) show the same
+asymmetries, though with somewhat greater scatter
 (Figure~\ref{fig:phasettv}).
@@
 light curve slope (Figure~\ref{fig:corr}).  In this model, we would
-expect the light curves with the most negative local slopes
-to have the largest positive TTVs, due to spot crossing events during the latter
-half of transit.  The lower-right panel of Figure~\ref{fig:phasettv}
-and upper-left panel of Figure~\ref{fig:phaseslope} agree with this
-expectation.  However, we would also expect the sign of the effect to
-reverse when considering the most positive local slopes, and the most
-negative TTVs.  The upper-left panel of Figure~\ref{fig:phasettv} and
-lower-right panel of Figure~\ref{fig:phaseslope} contradict this
-expectation: the residual in both cases maintains the same
-parity!  On the one hand, this shows us that the residual is not
-dependent on the local light curve slope, which mitigates the
-possibility that it might be an artifact of our detrending methods.
-On the other, it raises the question of whether
-unresolved starspot crossings are indeed the root cause of the correlation shown in
-Figure~\ref{fig:corr}.
+expect the light curves with the most negative local slopes to have
+the largest positive TTVs, due to spot crossing events during the
+latter half of transit.  The lower-right panel of
+Figure~\ref{fig:phasettv} and upper-left panel of
+Figure~\ref{fig:phaseslope} agree with this expectation.  However, we
+would also expect the sign of the effect to reverse when considering
+the most positive local slopes, and the most negative TTVs.  The
+upper-left panel of Figure~\ref{fig:phasettv} and lower-right panel of
+Figure~\ref{fig:phaseslope} contradict this expectation: the residual
+in both cases maintains the same parity!  On the one hand, this shows
+that the residual is not dependent on the local light curve slope,
+which mitigates the possibility that it might be an artifact of our
+detrending methods.  On the other, it raises the question of whether
+unresolved starspot crossings are indeed the root cause of the
+correlation shown in Figure~\ref{fig:corr}.
 
@@
 
-Since the transit asymmetry seems robust against most methods of
-binning the data, we considered four possible astrophysical
-interpretations: gravity darkening, transit timing variations,
-spot-crossing events, and a persistent asymmetric dusty outflow.
+The transit asymmetry seems robust against most methods of binning the
+data, though with some caveats ({\it e.g.}, the ``middle quartile'' in
+local flux (${\rm d}f/{\rm d}t\in[0.001,0.033]\,{\rm ppt\,day}^{-1}$),
+where the asymmetry does not appear).  Nonetheless, if the effect were
+entirely systematic we would expect to see it change as a function of
+the sign of the local slope or TTV.  We therefore considered four
+possible astrophysical interpretations: gravity darkening, transit
+timing variations, spot-crossing events, and a persistent asymmetric
+dusty outflow.
 
@@
 timing variations in \sn\ have amplitudes smaller than $\approx$10
-minutes (see Figure~\ref{fig:corr}).  This is consistent with
-previous results from \citet{holczer_transit_2016}.  Moreover, the
-transit timing variations we do observe are correlated with the local
-light curve slope, which on average increases as much as it decreases.
-From our analysis, the mean TTV and its standard deviation are
+minutes (see Figure~\ref{fig:corr}).  This is consistent with previous
+results from \citet{holczer_transit_2016}.  Moreover, the transit
+timing variations we do observe are correlated with the local light
+curve slope, which on average increases as much as it decreases.  From
+our analysis, the mean TTV and its standard deviation are
 $0.00\pm12.05$ minutes; similarly the mean local slope and its
@@
 is $\approx8$, which renders individual spot-crossing events
-unresolved.  This explanation seems marginally plausible, mainly because the
-expected spot-crossing anomaly amplitudes ($\approx 100\,{\rm ppm}$)
-resemble the observed amplitude of the asymmetry ($\approx 50\,{\rm
-ppm}$).
-One issue with this explanation however is that
-there is no reason to expect starspot crossing events to last
+unresolved.  This explanation seems marginally plausible, mainly
+because the expected spot-crossing anomaly amplitudes ($\approx
+100\,{\rm ppm}$) resemble the observed amplitude of the asymmetry
+($\approx 50\,{\rm ppm}$).  One issue with this explanation however is
+that there is no reason to expect starspot crossing events to last
 exactly half the transit duration.
@@
 likely in a regime where such outflows are possible.  Out of the four
-explanations discussed, this one at least theoretically seems the most plausible.
-By composition, the expectation would be that the envelope is mostly
-hydrogen and helium gas, with a dust or haze component providing the
-broadband opacity in the Kepler bandpass.  A natural path
-toward confirming or refuting this idea would be to observe additional
-transits of the planet in hydrogen absorption, 
-metastable helium absorption, or across a broad wavelength range in
-the near-infrared.
+explanations discussed, this one at least theoretically seems the most
+plausible.  By composition, the expectation would be that the envelope
+is mostly hydrogen and helium gas, with a dust or haze component
+providing the broadband opacity in the Kepler bandpass.  A natural
+path toward confirming or refuting this idea would be to observe
+additional transits of the planet in hydrogen absorption, metastable
+helium absorption, or across a broad wavelength range in the
+near-infrared.
 
@@
     spectrum relative to the median spectrum.  The bump in the red
-    wing of Ca H is H$\epsilon$.
-    The spectra in the lower panels are smoothed for visualization
-    purposes.
+    wing of Ca H is H$\epsilon$.  The spectra in the lower panels are
+    smoothed for visualization purposes.
     \label{fig:rvactivity}
@@
 standard California Planet Survey pipeline \citep{howard_cps_2010}.
-The airmass ranged between 1.1 and 2.2 from the
-start through the end of observations; the seeing ranged from
-$1\farcs1$ at the beginning to $1\farcs5$ at the end.  We also
-simultaneously observed across {\it griz} bands using MuSCAT3 at
-Faulkes Telescope North on Maui, HI.
+The airmass ranged between 1.1 and 2.2 from the start through the end
+of observations; the seeing ranged from $1\farcs1$ at the beginning to
+$1\farcs5$ at the end.  We also simultaneously observed across {\it
+griz} bands using MuSCAT3 at Faulkes Telescope North on Maui, HI.
 
@@
 
-Figure~\ref{fig:ground} shows the results.  The
-MuSCAT3 photometry shows the expected starspot trend, along with the
-transit and what is likely a chromatic starspot crossing event at
-${\rm JD} - 2459433 = 0.955$.  The radial velocities decrease by
-$\approx$250\ms\ over the six hour window.  
-This decrease in RV is correlated with a decrease in the S-indices
-derived from the Ca HK lines.
-One outlying RV point is apparent shortly before egress; it is
-temporally coincident with an outlying value in the S-index time series.
+Figure~\ref{fig:ground} shows the results.  The MuSCAT3 photometry
+shows the expected starspot trend, along with the transit and what is
+likely a chromatic starspot crossing event at ${\rm JD} - 2459433 =
+0.955$.  The radial velocities decrease by $\approx$250\ms\ over the
+six hour window.  This decrease in RV is correlated with a decrease in
+the S-indices derived from the Ca HK lines.  One outlying RV point is
+apparent shortly before egress; it is temporally coincident with an
+outlying value in the S-index time series.
 
@@
 limbs of the stellar disk has an apparent line-of-sight velocity of
-$\pm 20$\kms.  Over 10\% of a rotation cycle ($P_{\rm
-rot}=2.6\,{\rm days}$), spots near these limbs come into and out of
-view, modulate the stellar velocity profile, and can thereby produce
-the overall $\approx$250\ms\ trend.  The Ca HK and H$\alpha$ emission
-profiles support this interpretation;
-Figure~\ref{fig:rvactivity} shows that each line gradually decreases in
-intensity over the course of the six hour sequence.
-
-The expectation however is for the starspot-induced signals to be smooth,
-at worst with contributions at $0.5\,P_{\rm rot}$ or $0.25\,P_{\rm
-rot}$ \citep{klein_simulated_2020}.  We therefore fitted the RVs using
-the \citet{hirano_analytic_2010,hirano_2011} models for the
-Rossiter-McLaughlin (RM) effect, and allowed for an optional linear
-and quadratic trend in time to fit the $\approx$250\ms\ spot-induced trend.  We
-followed the methodology developed by \citet{stefansson_2020}.  We
-allowed the sky-projected obliquity, the projected stellar equatorial
-velocity, and the Gaussian dispersion of the spectral lines to vary,
-and fixed the limb-darkening using the $V$-band tabulation from
-\citet{claret_gravity_2011}.  We assumed a Gaussian prior on $v\sin i$ and
-$a/R_\star$ from Table~\ref{tab:starparams}, and also allowed for a
-white-noise jitter term to be added in quadrature to the measurement
-uncertainties.
-We used a 15\,minute exposure time to numerically evaluate the
-model.
+$\pm 20$\kms.  Over 10\% of a rotation cycle ($P_{\rm rot}=2.6\,{\rm
+days}$), spots near these limbs come into and out of view, modulate
+the stellar velocity profile, and can thereby produce the overall
+$\approx$250\ms\ trend.  The Ca HK and H$\alpha$ emission profiles
+support this interpretation; Figure~\ref{fig:rvactivity} shows that
+each line gradually decreases in intensity over the course of the six
+hour sequence.
+
+The expectation however is for the starspot-induced signals to be
+smooth, at worst with contributions at $0.5\,P_{\rm rot}$ or
+$0.25\,P_{\rm rot}$ \citep{klein_simulated_2020}.  We therefore fitted
+the RVs using the \citet{hirano_analytic_2010,hirano_2011} models for
+the Rossiter-McLaughlin (RM) effect, and allowed for an optional
+linear and quadratic trend in time to fit the $\approx$250\ms\
+spot-induced trend.  We followed the methodology developed by
+\citet{stefansson_2020}.  We allowed the sky-projected obliquity, the
+projected stellar equatorial velocity, and the Gaussian dispersion of
+the spectral lines to vary, and fixed the limb-darkening using the
+$V$-band tabulation from \citet{claret_gravity_2011}.  We assumed a
+Gaussian prior on $v\sin i$ and $a/R_\star$ from
+Table~\ref{tab:starparams}, and also allowed for a white-noise jitter
+term to be added in quadrature to the measurement uncertainties.  We
+used a 15\,minute exposure time to numerically evaluate the model.
 
 The quadratic model with the RM effect is shown in
-Figure~\ref{fig:ground}; the jitter term is incorporated in the
-model uncertainties, but not the plotted measurement uncertainties.
-The plotted measurement uncertainties are the internal uncertainties
-on the RVs ($\approx 13\,{\rm m\,s}^{-1}$), and are dominated by the
+Figure~\ref{fig:ground}; the jitter term is incorporated in the model
+uncertainties, but not the plotted measurement uncertainties.  The
+plotted measurement uncertainties are the internal uncertainties on
+the RVs ($\approx 13\,{\rm m\,s}^{-1}$), and are dominated by the
 $v\sin i$ broadening.  However, between exposures, the RVs show
-significant additional scatter that is not captured by the slow quadratic
-trend.  The white-noise jitter for this
-particular model is $\sigma_{\rm RV} = 27^{+6}_{-5}$\ms, which is
-comparable to the expected RM anomaly of $\Delta v_{\rm RM}
-\approx f_{\rm LD} \cdot \delta \cdot v\sin i \cdot \sqrt{1-b^2}
-\approx 20\,{\rm m\,s}^{-1}$, assuming a perfectly aligned orbit.
-
-The presence of this additional scatter prevents a convincing detection of
-the RM effect.  The reason can be understood via
-model comparison.  If we compare the model with a quadratic trend and
-the RM effect against a model with a
-linear trend and the RM effect, or even a model with no RM effect at
-all, then the respective Bayesian Information Criterion (BIC) values
-are as follows.
+significant additional scatter that is not captured by the slow
+quadratic trend.  The white-noise jitter for this particular model is
+$\sigma_{\rm RV} = 27^{+6}_{-5}$\ms, which is comparable to the
+expected RM anomaly of $\Delta v_{\rm RM} \approx f_{\rm LD} \cdot
+\delta \cdot v\sin i \cdot \sqrt{1-b^2} \approx 20\,{\rm m\,s}^{-1}$,
+assuming a perfectly aligned orbit.
+
+The presence of this additional scatter prevents a convincing
+detection of the RM effect.  The reason can be understood via model
+comparison.  If we compare the model with a quadratic trend and the RM
+effect against a model with a linear trend and the RM effect, or even
+a model with no RM effect at all, then the respective Bayesian
+Information Criterion (BIC) values are as follows.
 % from rm_TEMPLATE_V1298TAU_quadraticfreejitter_v0.log and
@@
 spectrograph in the optical \citep[{\it
-e.g.},][]{gibson_2016_kpf,seifahrt_maroon-x_2018}, or in the near-infrared
-\citep[{\it e.g.},][]{feinstein_halpha_2021}.
+e.g.},][]{gibson_2016_kpf,seifahrt_maroon-x_2018}, or in the
+near-infrared \citep[{\it e.g.},][]{feinstein_halpha_2021}.
 
@@
 \begin{enumerate}
-  \item Fit the starspot-induced variability using a
-    Gaussian Process with a \texttt{SHOTerm} kernel, a white-noise jitter term, and the
+  \item Fit the starspot-induced variability using a Gaussian Process
+    with a \texttt{SHOTerm} kernel, a white-noise jitter term, and the
     mean flux.
@@
     curve (these points include the flares).  Repeat Step~1.
-  \item Using the residual from Step 2, identify all flares,
-    requiring them to be at least 20 cadences apart, at least 7 median
-    absolute deviations above the median baseline, and lasting at
-    least 2 cadences in duration.  Build the mask spanning these
-    times, from 5 minutes before each flare begins to 2.5 minutes
-    after the final flare cadence.  Repeat Step 1 a final time.
+  \item Using the residual from Step 2, identify all flares, requiring
+    them to be at least 20 cadences apart, at least 7 median absolute
+    deviations above the median baseline, and lasting at least 2
+    cadences in duration.  Build the mask spanning these times, from 5
+    minutes before each flare begins to 2.5 minutes after the final
+    flare cadence.  Repeat Step 1 a final time.
 \end{enumerate}
-The final step of flare identification and fitting was performed using \texttt{altaipony}
-\citep{davenport_2016,ilin_flares_2021}.  The analytic flare model is
-from \citet{davenport_2014} and it parametrizes the flare with a start
-time, an exponential lag time, and an amplitude.
+The final step of flare identification and fitting was performed using
+\texttt{altaipony} \citep{davenport_2016,ilin_flares_2021}.  The
+analytic flare model is from \citet{davenport_2014} and it
+parametrizes the flare with a start time, an exponential lag time, and
+an amplitude.
 %Figure~\ref{fig:flarephase} shows the resulting flares, amplitudes,

### 2021-10-09 paper/ms.tex
update ms

@@
 		\leavevmode
-		\includegraphics[width=0.49\textwidth]{f7.pdf}
+		\includegraphics[width=0.47\textwidth]{f7.pdf}
 	\end{center}
@@
 	\caption{
-    {\bf Radial velocities acquired using Keck/HIRES during the
-    transit of 2021 Aug 7, observed simultaneously with MuSCAT3.}
-    Shaded bands show 2-$\sigma$ model uncertainties.  The photometric
-    transit depths are consistent across the {\it griz} bandpasses.
-    The best-fitting model in the top panel includes the RM effect and
-    a quadratic trend in time to model the spot-induced
-    $\approx$\,$250$\ms\ trend (see Appendix~\ref{app:rm}).  The
-    radial velocity jitter over the 15~minute exposure timescale
+    {\bf Keck/HIRES radial velocities and MuSCAT3 photometry from the transit of 2021 Aug 7.}
+    {\it Top:} 
+    The radial velocity jitter across the 15~minute exposures
     ($\sigma_{\rm RV}\approx30$\ms) prevented us from detecting the RM
-    effect.  The RV variations are strongly correlated with varying
-    emission in the Ca H and K lines (middle panel).
+    effect;
+    a model including the RM anomaly and a quadratic trend in time to
+    fit the spot-induced $\approx$\,$250$\ms\ trend is shown (see
+    Appendix~\ref{app:rm}).  Shaded bands show 2-$\sigma$ model
+    uncertainties.  
+    {\it Middle:} The RV variations are strongly correlated with varying
+    emission in the Ca H and K lines.
+    {\it Bottom}: The photometric transit depths are consistent across
+    the {\it griz} bandpasses. The photometry is binned at 10~minute
+    intervals.
 		\label{fig:ground}
@@
 Kepler ephemeris and transit depth, the MuSCAT3 observations may also
-show a transit duration $17.3\pm4.0\ {\rm minutes}$ shorter than the
-Kepler transits; further photometric follow-up would help confirm
+suggest a transit duration $17.3\pm4.3\ {\rm minutes}$ shorter than the
+Kepler transits; further photometric follow-up could help confirm
 whether the transit duration is indeed changing.  For our RM analysis,
@@
 		\leavevmode
-		\includegraphics[width=0.74\textwidth]{f8.pdf}
+		\includegraphics[width=0.8\textwidth]{f8.pdf}
 	\end{center}
@@
 e.g.},][]{ehrenreich_giant_2015,spake_helium_2018,vissapragada_2020}.
-Observations aimed at quantifying the amount of high-energy
+Observations that quantify the amount of high-energy
 irradiation incident on the planet would complement these efforts, by
@@
 
-\clearpage
+% \clearpage
 \acknowledgements

### 2021-10-09 paper/ms.tex
add data tables (rvs + phot)

@@
     the light curve.  The two quantities appear anti-correlated, which
-    is most easily explained by starspot crossings during the first
+    could be explained by starspot crossings during the first
     (second) half of transit inducing a positive (negative) TTV,
@@
 might hypothetically still orbit a hidden close and bright companion,
-this possibility is ruled out by {\it i)} a lack of secondary lines in
-the HIRES spectra and {\it ii)} a lack of secondary rotation signals
+this possibility is implausible given {\it i)} the lack of secondary lines in
+the HIRES spectra and {\it ii)} the lack of secondary rotation signals
 in the Kepler photometry.
@@
 Objects of Interest have shown this correlation
-\citep{holczer_time_2015}, which is most readily interpreted as a TTV
+\citep{holczer_time_2015}, which can be explained as a TTV
 induced by unresolved starspot crossings \citep{mazeh_time_2015}.
@@
 
-\section{Spectroscopic Transit Analysis}
+\section{Spectroscopy and Photometry During the Transit of 2021 Aug 7}
 \label{app:rm}
@@
 $1\farcs5$ at the end.  We also simultaneously observed across {\it
-griz} bands using MuSCAT3 at Faulkes Telescope North on Maui, HI.
+griz} bands using MuSCAT3 at Haleakal\=a Observatory on Maui, HI.
+Performing aperture photometry on the latter image stack yielded the
+data given in Table~\ref{tab:phot}.
 
@@
 (Method 2).  Given these diagnostics, we adopted the velocities from
-the second approach.
+the second approach, which are reported in Table~\ref{tab:rv}.
 
@@
 near-infrared \citep[{\it e.g.},][]{feinstein_halpha_2021}.
+
+\input{phot_table.tex}
+\input{rv_table.tex}
 

### 2021-10-09 drivers/write_muscat_table.py

@@ -1,0 +1,25 @@
+import os
+from numpy import array as nparr
+import numpy as np, pandas as pd
+from rudolf.paths import RESULTSDIR, PAPERDIR
+from rudolf.helpers import get_kep1627_muscat_lightcuve
+
+datasets = get_kep1627_muscat_lightcuve()
+
+bps = 'g,r,i,z'.split(',')
+
+dfs = [
+    pd.DataFrame({
+        'BJDTDB': datasets[f"muscat3_{bp}"][0],
+        'FLUX': datasets[f"muscat3_{bp}"][1],
+        'E_FLUX': datasets[f"muscat3_{bp}"][2],
+        'BANDPASS': str(bp)
+    }) for bp in bps
+]
+
+df = pd.concat(dfs)
+
+outpath = os.path.join(PAPERDIR, 'muscat_phot_table.csv')
+
+df.to_csv(outpath, index=False)
+print(f"Wrote {outpath}")

### 2021-10-10 paper/ms.tex
update ms
revised TTV analysis

@@
   peculiarities: the average transit profile is asymmetric, and the
-  individual transit times are correlated with the local light curve
+  individual transit times might be correlated with the local light curve
   slope.  We discuss possible explanations for each anomaly.  More
-  broadly, the $\delta$~Lyr cluster
+  importantly, the $\delta$~Lyr cluster
   is one of $\sim$10$^3$ coeval groups whose properties have been clarified
@@
 Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A.
-Based on a correlation between the planetary transit times and the
-local light curve slope, the orbit may also be prograde.  We conclude
-by highlighting broader implications for our ability to age-date a
-larger sample of planets (Section~\ref{sec:conc}).
+We also analyze an asymmetry in the average transit profile, and a
+possible correlation between the individual transit times and the
+local light curve slope.  We conclude by discussing broader
+implications for our ability to age-date a larger sample of planets
+(Section~\ref{sec:conc}).
 
@@
 
+\subsection{Kepler Light Curve}
+
 \begin{figure*}[tp]
@@
 
-\subsection{Kepler Light Curve}
-
 The Kepler space telescope observed \sn\ at a 30-minute cadence from
@@
 
+\subsubsection{Transit and Stellar Variability Model}
 We fitted the Kepler long cadence time series with a model that
@@
 
+\subsubsection{Transit Asymmetry}
+\label{subsec:asymmetry}
 The transit fit however is not perfect: the lower panels of
@@
 deviation is $\approx50\,{\rm ppm}$, which represents a $\approx 3\%$
-distortion of the transit depth ($1759\pm62\,{\rm ppm}$).  Note that
-while this asymmetry is within the 2-$\sigma$ model uncertainties, the
+distortion of the transit depth ($\delta=1759\pm62\,{\rm ppm}$).  Note that
+although this asymmetry is within the 2-$\sigma$ model uncertainties, the
 model has a jitter term that grows to account for otherwise unmodeled
@@
 therefore best assessed in comparison against the intrinsic
-out-of-transit scatter in the data, not the model uncertainties.
-
-To determine whether the asymmetry is a systematic caused by our
+out-of-transit scatter in the data ($\approx16\,{\rm ppm}$), not the model uncertainties.  The
+lower right panel of Figure~\ref{fig:lc} demonstrates that the scatter during
+transit is higher than during all other phases of the planet's orbit.
+
+To determine whether the asymmetry could be a systematic caused by our
 stellar variability model, we explored an alternative approach in
-which we isolated each transit window and locally fitted out
+which we isolated each transit window, locally fitted out
 polynomial trends, and then binned all the observed transits; the
 asymmetry was still present at a comparable amplitude.
-Appendix~\ref{app:asymmetry} includes a deeper analysis, and finds
+Appendix~\ref{app:asymmetry} describes a more detailed analysis, which finds
 that the asymmetry also seems to be robust to different methods of
@@
 
-\subsection{Planet Confirmation}
+\subsubsection{Transit Timing and the Local Slope}
+\label{subsec:ttvslope}
 
@@
 	\caption{
-    {\bf Evidence for a prograde orbit of Kepler 1627 Ab.} The time of
+    {\bf Possible evidence for a prograde orbit of Kepler 1627 Ab.} The time of
     each Kepler transit was measured, along with the local slope of
-    the light curve.  The two quantities appear anti-correlated, which
+    the light curve.  The two quantities are weakly anti-correlated
+    ($\approx$2-$\sigma$), which
     could be explained by starspot crossings during the first
@@
 
-If the \pn\ transit signal is created by a genuine planet,  then to
-our knowledge it would be the youngest planet yet found by the main
-Kepler mission.\footnote{The re-purposed K2 mission however has found
-two younger systems containing five planets: K2-33b ($9\pm1\,{\rm
-Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and V1298 Tau
-($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}   Could the transit
-be produced by anything other than a planet orbiting this near-solar
-analog?  \citet{morton_false_2016} validated the planet based on the
-transit shape, arguing that the most probable false positive scenario
-was that of a background eclipsing binary, which had a model-dependent
-probability of $\approx$$10^{-5}$.  However, this calculation was
-performed without knowledge of the low-mass stellar companion ($M_{\rm
-B}\approx 0.33\,M_\odot$).  Validated planets have also previously
-been refuted \citep[{\it e.g.},][]{shporer_three_2017}.  We therefore
-reassessed false positive scenarios in some detail.
-
-% (0.7-2.0\% V-band to G-band dmag difference from EEM table)
-
-As an initial plausibility check, Kepler\,1627B contributes 1\% to 2\%
-of the total flux observed in the Kepler aperture.  For the sake of
-argument, assume the former value.  The observed transit has a depth
-of $\approx$0.18\%.  A 18\% deep eclipse of Kepler\,1627B would
-therefore be needed to produce a signal with the appropriate depth.
-The shape of the transit signal however requires the impact parameter
-to be below 0.77 (2-$\sigma$); the tertiary transiting the secondary
-would therefore need to be non-grazing with $R_3/R_2 \approx 0.4$.
-This yields a contradiction:  this scenario would require an ingress
-and egress phase that each span $\approx$40\% of the transit duration
-($\approx 68\,{\rm minutes}$).  The actual measured ingress and egress
-duration is $\approx 17\,{\rm minutes}$, a factor of four times too
-short.  The combination of Kepler 1627B's brightness, the transit
-depth, and the ingress duration therefore disfavor the scenario that
-Kepler 1627B might host the transit signal.
-
-% Baraffe+15: M_B 0.33 Mdot
-% Rstar = 0.455 --> 0.465 Rsun
-
-There are two other new lines of evidence that confirm the planetary
-interpretation.  First, the stellar density implied by the transit duration and orbital period
-is inconsistent with an eclipsing body around the M-dwarf companion:
-we find $\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
-theoretically expected density for Kepler\,1627B is $\approx 4.6\,{\rm
-g\,cm}^{-3}$ (Table~\ref{tab:posterior}; \citealt{choi_mesa_2016}).
-The transit duration is therefore too long to be explained by a star
-eclipsing the M dwarf secondary at $10$-$\sigma$.  While the planet
-might hypothetically still orbit a hidden close and bright companion,
-this possibility is implausible given {\it i)} the lack of secondary lines in
-the HIRES spectra and {\it ii)} the lack of secondary rotation signals
-in the Kepler photometry.
-
-%NOTE to the arxiv readers:, the $\approx$100 days of short-cadence
-%Kepler observations also mitigate the cadence-induced smearing, and
-%enable a more precise measurement of the transit impact parameter.
-
-The second line of confirming evidence comes by analyzing the
-individual Kepler transit times. We isolated each of the 144 observed
-transits to within $\pm4.5$\,hr of each transit, and fitted each
-window with both {\it i)} a local second-order polynomial and transit,
-and {\it ii)} a local linear trend.  We let the mid-time of each
+The previous analysis by \citet{holczer_transit_2016} did not
+find any significant long-term transit timing or duration variations
+for \sn\ (TTVs or TDVs).  Quantitatively, the mean and standard deviation of the
+TTVs and TDVs they measured were $-1.1\pm13.8\,{\rm
+min}$ and $-3.3\pm22.1\,{\rm min}$.  In an
+earlier analysis however, \citet{holczer_time_2015} studied correlations
+between TTVs and local light curve slopes, and for \sn\
+found a weak correlation of $-29\pm13\,{\rm min}\,{\rm day}^{-1}$
+between the two quantities.  While this constitutes only weak
+statistical evidence, given the possible connection between such
+correlations and the unresolved starspot crossings that we expect to
+be present in the data for \sn\ \citep{mazeh_time_2015}, we opted to re-examine the
+individual transit times.
+
+We therefore isolated each of the 144 observed transits to within
+$\pm4.5$\,hr of each transit, and fitted each window with both {\it
+i)} a local polynomial baseline plus the transit, and {\it ii)} a local linear
+trend.  We considered the results both for a second and fourth-order
+time-dependence in the local baseline.  We let the mid-time of each
 transit float, and then calculated the residual between the measured
 mid-time and that of a periodic orbit.  This residual, the transit
-timing variation (TTV), is plotted in Figure~\ref{fig:corr} against
-the local linear slope.  A significant correlation of $-0.0564 \pm
-0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
-Objects of Interest have shown this correlation
-\citep{holczer_time_2015}, which can be explained as a TTV
-induced by unresolved starspot crossings \citep{mazeh_time_2015}.
-This is only possible if the planet transits the primary star, which
-excludes a background eclipsing binary scenario, and therefore
-confirms that \pn\ is a planet.  It also suggests that the planet's
-orbit is prograde.  The latter point assumes that the dominant
-photometric variability is induced by dark spots, and not bright
-faculae.  Given the observed transition of Sun-like stellar
-variability from spot to faculae-dominated regimes between young and
-old ages, we expect this assumption to be reasonably secure
-\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
+timing variation, is plotted in Figure~\ref{fig:corr} against
+the local linear slope for the fourth-order polynomial baseline.  The
+slope of $-21 \pm 10\,{\rm min\,day}^{-1}$ is similar to that
+found by \citet{holczer_time_2015}.
+
+One concern we had in this analysis was whether our transit fitting
+procedure might induce spurious correlations between the slope and
+transit time.  In particular, using the second-order polynomial
+baseline yielded a larger anti-correlation between the TTVs and local
+slopes, of $-79 \pm 14\,{\rm min\,day}^{-1}$.  We therefore
+performed an injection-recovery procedure in which we injected
+transits at different phases in the \sn\ light curve and repeated the
+TTV analysis.  This was done at $\approx$50 phases, each separated by
+$0.02\,P_{\rm orb}$ while omitting the phases in transit.  For
+the second-order polynomial baseline, this procedure yielded
+similar anti-correlations in the injected transits to that present in the real
+transit;
+using this baseline would therefore bias the
+result.  However for the fourth-order baseline, the correlation
+present in the data was stronger than in all but one of the injected
+transits.
+Possible interpretations are discussed below, however given the
+statistical significance, this analysis should be interpreted as
+suggestive at best.
+
+\subsection{Planet Confirmation}
 
@@
 		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f7.pdf}
+		\includegraphics[width=0.45\textwidth]{f7.pdf}
 	\end{center}
@@
 
+If the \pn\ transit signal is created by a genuine planet,  then to
+our knowledge it would be the youngest planet yet found by the main
+Kepler mission.\footnote{The re-purposed K2 mission however has found
+two younger systems containing five planets: K2-33b ($9\pm1\,{\rm
+Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and V1298 Tau
+($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}   Could the transit
+be produced by anything other than a planet orbiting this near-solar
+analog?  \citet{morton_false_2016} validated the planet based on the
+transit shape, arguing that the most probable false positive scenario
+was that of a background eclipsing binary, which had a model-dependent
+probability of $\approx$$10^{-5}$.  However, this calculation was
+performed without knowledge of the low-mass stellar companion ($M_{\rm
+B}\approx 0.33\,M_\odot$).  Validated planets have also previously
+been refuted \citep[{\it e.g.},][]{shporer_three_2017}.  We therefore
+reassessed false positive scenarios in some detail.
+
+% (0.7-2.0\% V-band to G-band dmag difference from EEM table)
+
+As an initial plausibility check, Kepler\,1627B contributes 1\% to 2\%
+of the total flux observed in the Kepler aperture.  For the sake of
+argument, assume the former value.  The observed transit has a depth
+of $\approx$0.18\%.  A 18\% deep eclipse of Kepler\,1627B would
+therefore be needed to produce a signal with the appropriate depth.
+The shape of the transit signal however requires the impact parameter
+to be below 0.77 (2-$\sigma$); the tertiary transiting the secondary
+would therefore need to be non-grazing with $R_3/R_2 \approx 0.4$.
+This yields a contradiction:  this scenario would require an ingress
+and egress phase that each span $\approx$40\% of the transit duration
+($\approx 68\,{\rm minutes}$).  The actual measured ingress and egress
+duration is $\approx 17\,{\rm minutes}$, a factor of four times too
+short.  The combination of Kepler 1627B's brightness, the transit
+depth, and the ingress duration therefore disfavor the scenario that
+Kepler 1627B might host the transit signal.
+
+% Baraffe+15: M_B 0.33 Mdot
+% Rstar = 0.455 --> 0.465 Rsun
+
+Beyond this simple test, a line of evidence that
+effectively confirms the planetary interpretation is that the stellar density
+implied by the transit duration and orbital period is inconsistent
+with an eclipsing body around the M-dwarf companion.  We find
+$\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
+theoretically expected density for Kepler\,1627B is $\approx 4.6\,{\rm
+g\,cm}^{-3}$ (Table~\ref{tab:posterior}; \citealt{choi_mesa_2016}).
+The transit duration is therefore too long to be explained by a star
+eclipsing the M dwarf secondary at $10$-$\sigma$.  While the planet
+might hypothetically still orbit a hidden close and bright companion,
+this possibility is implausible given {\it i)} the lack of secondary
+lines in the HIRES spectra and {\it ii)} the lack of secondary
+rotation signals in the Kepler photometry.
+
+%NOTE to the arxiv readers:, the $\approx$100 days of short-cadence
+%Kepler observations also mitigate the cadence-induced smearing, and
+%enable a more precise measurement of the transit impact parameter.
+
+The correlation noted in Section~\ref{subsec:ttvslope} between the
+TTVs and the local light curve slope could be a further line of evidence
+in support of the planetary interpretation.  Unless it is a
+statistical fluke (a $\approx$5\% possibility), then its most likely
+cause is unresolved starspot
+crossings \citep{mazeh_time_2015}.  This would only be possible if the
+planet transits the primary star, which excludes a background
+eclipsing binary scenario.  It would also suggest that the planet's
+orbit is prograde.  The latter point assumes that the dominant
+photometric variability is induced by dark spots, and not bright
+faculae.  Given the observed transition of Sun-like stellar
+variability from spot to faculae-dominated regimes between young and
+old ages, we expect this latter assumption to be reasonably secure
+\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
+
 A third supporting line of evidence for the planetary interpretation
-exists, though we consider it less definitive than the stellar density
-and TTV-local slope correlation.  We observed a transit of \pn\ on the
+also exists.  We observed a transit of \pn\ on the
 night of 2021 Aug 7 simultaneously with Keck/HIRES and MuSCAT3.  We
@@
 Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
-show the transit (Figure~\ref{fig:ground}).  Fitting the MuSCAT3
+show that the transit is achromatic (Figure~\ref{fig:ground}).
+Quantitatively, when we fitted the MuSCAT3
 photometry with a model that lets the transit depths vary across each
-bandpass, we find {\it griz} depths consistent with the Kepler depth
+bandpass, we found {\it griz} depths consistent with the Kepler depth
 at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  Conditioned on the
-Kepler ephemeris and transit depth, the MuSCAT3 observations may also
-suggest a transit duration $17.3\pm4.3\ {\rm minutes}$ shorter than the
-Kepler transits; further photometric follow-up could help confirm
-whether the transit duration is indeed changing.  For our RM analysis,
+Kepler ephemeris and transit depth, the MuSCAT3 observations also
+suggested a transit duration $17.3\pm4.3\ {\rm minutes}$ shorter than the
+Kepler transits.
+However, given both the lack of transit duration variations in the Kepler data
+and the relatively low signal-to-noise of the MuSCAT3 transit,
+further photometric follow-up would be necessary to determine
+whether the transit duration is indeed changing.  
+For our RM analysis,
 the details are discussed in Appendix~\ref{app:rm}.  While the
@@
 
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f8.pdf}
+		\includegraphics[width=0.48\textwidth]{f8.pdf}
 	\end{center}
@@
 asymmetry shown in Figure~\ref{fig:lc} is that of a dusty asymmetric
-outflow.  A second possibility is that the planetary orbit is slightly
-misaligned from the stellar spin axis, and tends to transit starspot
-groups at favored stellar latitudes.  Other possibilities such as
-gravity darkening or TTVs causing the asymmetry are disfavored (see
-Appendix~\ref{app:asymmetry}).  Dusty outflows are theoretically
+outflow.  
+Dusty outflows are theoretically
 expected for young mini-Neptunes, and the amplitude of the observed
 asymmetry is also consistent with predictions \citep{wang_dai_2019}.
-Beyond the asymmetric transits, Appendix~\ref{app:flare} highlights an
-additional abnormality in the Kepler data, in the arrival time
-distribution of stellar flares.  We encourage its exploration by
-investigators more versed in the topic than ourselves.
+A second possibility is that the planetary orbit is slightly
+misaligned from the stellar spin axis, and tends to transit starspot
+groups at favored stellar latitudes.  
+This somewhat contrived geometry would be necessary in order to
+explain why the starspot crossings would add up coherently.
+Other possible explanations such as
+gravity darkening or TTVs causing the asymmetry are disfavored (see
+Appendix~\ref{app:asymmetry}).  
+Aside from the asymmetric transits, Appendix~\ref{app:flare}
+highlights an additional abnormality in the short-cadence Kepler data, in the
+arrival time distribution of stellar flares.  We encourage its
+exploration by investigators more versed in the topic than ourselves.
 
@@
 \end{figure}
-\begin{figure}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f12.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Transit models and residuals, binned by quartiles of transit
-    timing variations}.  The data are otherwise the same as in
-    Figure~\ref{fig:phaseslope}.
-		\label{fig:phasettv}
-	}
-\end{figure}
 
@@
 As a means of exploring the robustness of the transit asymmetry,
-Figures~\ref{fig:phasequarter},~\ref{fig:phaseyear},~\ref{fig:phaseslope},
-and~\ref{fig:phasettv} show the Kepler data binned in four ways: over
-Kepler quarters, Julian years, quartiles of local slope, and quartiles
-of transit timing variation.  Over Kepler quarters
+Figures~\ref{fig:phasequarter},~\ref{fig:phaseyear}, and~\ref{fig:phaseslope}
+show the Kepler data binned in three ways: over
+Kepler quarters, Julian years, and quartiles of local slope.  Over Kepler quarters
 (Figure~\ref{fig:phasequarter}), Quarter 6 shows the strongest
@@
 Figure~\ref{fig:phaseyear}; the asymmetry is visually apparent however
-in each of Years 2, 3, and 4.
-
-To bin by quartiles in local slope and TTV, we used our measurements
+in each of the four years.
+
+To bin by quartiles in local slope, we used our measurements
 of the local linear slopes in each of the observed transit windows
@@
 transits.  These were then divided into quartiles, so that each panel
-shows 35 transits binned together.  The exact light curve slope and
-TTV intervals are listed in the lower left panels of
-Figures~\ref{fig:phaseslope} and~\ref{fig:phasettv}.  Binned by local
+shows 35 transits binned together.  The exact light curve slope intervals are listed in the lower left panels of
+Figure~\ref{fig:phaseslope}.  Binned by local
 slope quartiles (Figure~\ref{fig:phaseslope}), the asymmetry is
 visually present in three of the four quartiles: the only bin in which
-it does not appear is ${\rm d}f/{\rm d}t\in[0.001,0.033]\,{\rm
-ppt\,day}^{-1}$.  The corresponding bins in TTV (largest TTV
-corresponding to smallest ${\rm d}f/{\rm d}t$) show the same
-asymmetries, though with somewhat greater scatter
-(Figure~\ref{fig:phasettv}).
+it does not appear is ${\rm d}f/{\rm d}t\in[3.0,32.5]\,{\rm
+ppt\,day}^{-1}$.
 
 Within the theory presented by \citet{mazeh_time_2015}, unresolved
-starspot crossings cause the correlation between TTVs and the local
+starspot crossings cause the weak correlation between TTVs and the local
 light curve slope (Figure~\ref{fig:corr}).  In this model, we would
@@
 the largest positive TTVs, due to spot crossing events during the
-latter half of transit.  The lower-right panel of
-Figure~\ref{fig:phasettv} and upper-left panel of
-Figure~\ref{fig:phaseslope} agree with this expectation.  However, we
+latter half of transit.  The upper-left panel of
+Figure~\ref{fig:phaseslope} agrees with this expectation.  However, we
 would also expect the sign of the effect to reverse when considering
-the most positive local slopes, and the most negative TTVs.  The
-upper-left panel of Figure~\ref{fig:phasettv} and lower-right panel of
-Figure~\ref{fig:phaseslope} contradict this expectation: the residual
+the most positive local slopes (most negative TTVs).  The
+lower-right panel of
+Figure~\ref{fig:phaseslope} contradicts this expectation: the residual
 in both cases maintains the same parity!  On the one hand, this shows
 that the residual is not dependent on the local light curve slope,
-which mitigates the possibility that it might be an artifact of our
+which lowers the likelihood that it might be an artifact of our
 detrending methods.  On the other, it raises the question of whether
@@
 correlation shown in Figure~\ref{fig:corr}.
-
-% FIXME: KM to discuss injection recovery...
+While we do not have a solution to this contradiction, 
+the injection-recovery tests discussed in
+Section~\ref{subsec:ttvslope} provide some assurance that the
+TTV-slope correlation is not simply a systematic artifact.
 
@@
 data, though with some caveats ({\it e.g.}, the ``middle quartile'' in
-local flux (${\rm d}f/{\rm d}t\in[0.001,0.033]\,{\rm ppt\,day}^{-1}$),
-where the asymmetry does not appear).  Nonetheless, if the effect were
-entirely systematic we would expect to see it change as a function of
-the sign of the local slope or TTV.  We therefore considered four
-possible astrophysical interpretations: gravity darkening, transit
+local flux, ${\rm d}f/{\rm d}t\in[3.0,32.5]\,{\rm ppt\,day}^{-1}$,
+where the asymmetry does not appear).  Nonetheless, if the asymmetric were
+systematic we might expect its parity to reverse as a function of
+the sign of the local slope, and it does not.  We therefore entertained four
+possible astrophysical explanations: gravity darkening, transit
 timing variations, spot-crossing events, and a persistent asymmetric
@@
 The scenario of transit timing variations (TTVs) producing the
-asymmetry seems unlikely because our analysis showed that transit
-timing variations in \sn\ have amplitudes smaller than $\approx$10
-minutes (see Figure~\ref{fig:corr}).  This is consistent with previous
-results from \citet{holczer_transit_2016}.  Moreover, the transit
+asymmetry seems unlikely because the transit
 timing variations we do observe are correlated with the local light
-curve slope, which on average increases as much as it decreases.  From
+curve slope, which increases roughly as much as it decreases.  From
 our analysis, the mean TTV and its standard deviation are
-$0.00\pm12.05$ minutes; similarly the mean local slope and its
-standard deviation are $0.49\pm44.49$\,ppt\,day$^{-1}$.  There is
-therefore no expectation for TTVs to produce the asymmetry.  A
+$0.66\pm5.53$ minutes; similarly the mean local slope and its
+standard deviation are $0.59\pm45.50$\,ppt\,day$^{-1}$.  There is
+therefore little expectation for TTVs to produce the asymmetry.  A
 separate line of argument comes from Figure~\ref{fig:phaseslope}.  If
@@
 would expect that in the largest ${\rm d}f/{\rm d}t$ bin, ${\rm
-d}f/{\rm d}t\in[0.033,0.092]\,{\rm ppt\,day}^{-1}$, the sign of the
+d}f/{\rm d}t\in[3.0,32.5]\,{\rm ppt\,day}^{-1}$, the sign of the
 asymmetry would reverse.  We do not see evidence for this being the
@@
 \citep{roettenbacher_contemporaneous_2017}.  Regardless, for any
-spot-crossing anomalies to add coherently over the 144 Kepler
+spot-crossing anomalies to add coherently over the 140 Kepler
 transits, it seems likely that we would need either for spots to be
@@
 providing the broadband opacity in the Kepler bandpass.  A natural
-path toward confirming or refuting this idea would be to observe
+path for testing this idea would be to observe
 additional transits of the planet in hydrogen absorption, metastable
@@
 		\subfloat{
-			\includegraphics[width=\textwidth]{f13.pdf}
+			\includegraphics[width=\textwidth]{f12.pdf}
 		}
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f14.pdf}
+		\includegraphics[width=0.9\textwidth]{f13.pdf}
 	\end{center}
@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f15.pdf}
+		\includegraphics[width=0.8\textwidth]{f14.pdf}
 	\end{center}

### 2021-10-11 paper/ms.tex
update ms
figure formatting

minor text edits

@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=1\textwidth]{f1.pdf}
+		\includegraphics[width=0.48\textwidth]{f1.pdf}
 	\end{center}
@@
     \citet{KounkelCovey2019}.  The tangential velocities relative to
-    \sn\ (lower-right) are computed assuming that every star has the
+    \sn\ (bottom row) are computed assuming that every star has the
     same three-dimensional spatial velocity as \sn.  Our analysis
@@
 latest Gaia EDR3 observations using the \texttt{dr2\_neighbourhood}
-table from the ESA archive,\footnote{See
-\url{https://gea.esac.esa.int/archive/documentation/GEDR3/Gaia_archive/chap_datamodel/sec_dm_auxiliary_tables/ssec_dm_dr2_neighbourhood.html}.}
+table from the ESA archive,
 taking the stars closest in proper motion and epoch-corrected angular
@@
 	\caption{
-    {\bf Kepler's view of the $\delta$~Lyr cluster.} Each black circle
+    {\bf Kepler's view of the $\delta$~Lyr cluster, shown in equatorial
+    	coordinates.} Each black circle
     is a candidate cluster member selected from the catalog of
@@
 find any significant long-term transit timing or duration variations
-for \sn\ (TTVs or TDVs).  Quantitatively, the mean and standard deviation of the
+(TTVs or TDVs) for \sn.  Quantitatively, the mean and standard deviation of the
 TTVs and TDVs they measured were $-1.1\pm13.8\,{\rm
@@
 found a weak correlation of $-29\pm13\,{\rm min}\,{\rm day}^{-1}$
-between the two quantities.  While this constitutes only weak
-statistical evidence, given the possible connection between such
+between the two quantities.  Given the possible connection between such
 correlations and the unresolved starspot crossings that we expect to
@@
 the second-order polynomial baseline, this procedure yielded
-similar anti-correlations in the injected transits to that present in the real
+a similar anti-correlation in the injected transits as that present in the real
 transit;
@@
 transits.
-Possible interpretations are discussed below, however given the
+Possible interpretations are discussed below. Given the
 statistical significance, this analysis should be interpreted as
@@
 
-\begin{figure}[t]
+\begin{figure}[tp]
 	\begin{center}
@@
 would therefore need to be non-grazing with $R_3/R_2 \approx 0.4$.
-This yields a contradiction:  this scenario would require an ingress
+This yields a contradiction:  this scenario requires an ingress
 and egress phase that each span $\approx$40\% of the transit duration
@@
 this possibility is implausible given {\it i)} the lack of secondary
-lines in the HIRES spectra and {\it ii)} the lack of secondary
-rotation signals in the Kepler photometry.
+lines in the HIRES spectra, {\it ii)} the lack of secondary
+rotation signals in the Kepler photometry, and {\it iii)} the proximity
+of \sn\ to the main cluster isochrone on the Gaia CAMD (Figure~\ref{fig:age}).
 
@@
 The correlation noted in Section~\ref{subsec:ttvslope} between the
-TTVs and the local light curve slope could be a further line of evidence
+TTVs and the local light curve slope might be an additional line of evidence
 in support of the planetary interpretation.  Unless it is a
-statistical fluke (a $\approx$5\% possibility), then its most likely
-cause is unresolved starspot
-crossings \citep{mazeh_time_2015}.  This would only be possible if the
+statistical fluke (a $\approx$5\% possibility), then the most likely
+cause of the correlation is unresolved starspot
+crossings \citep{mazeh_time_2015}.  These would only be possible if the
 planet transits the primary star, which excludes a background
-eclipsing binary scenario.  It would also suggest that the planet's
+eclipsing binary scenario.  The correlation would also suggest that the planet's
 orbit is prograde.  The latter point assumes that the dominant
@@
 at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  Conditioned on the
-Kepler ephemeris and transit depth, the MuSCAT3 observations also
+ephemeris and transit depth from the Kepler data, the MuSCAT3 observations also
 suggested a transit duration $17.3\pm4.3\ {\rm minutes}$ shorter than the
@@
 
-\begin{figure}[tp]
+\begin{figure*}[tp]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f8.pdf}
+		\includegraphics[width=0.8\textwidth]{f8.pdf}
 	\end{center}
@@
 	}
-\end{figure}
+\end{figure*}
 
@@
 e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
-quantity to measure would be the planet's mass.  The most feasible
-approach would likely be long-term monitoring of the transit times and
-durations, given the technical difficulty of performing a multi-color
-radial velocity campaign on a $V=13$ star.  Measured at
-sufficient precision though, the mass, combined with the known age and
-size, would yield constraints on both the planet's composition and its
+quantity to measure would be the planet's mass. Measured at
+sufficient precision,
+for instance through a multi-wavelength radial velocity campaign,
+the combination of the size, mass, and age 
+would yield constraints on both the planet's composition and its
 initial entropy \citep{owen_constraining_2020}.
@@
 groups at favored stellar latitudes.  
-This somewhat contrived geometry would be necessary in order to
-explain why the starspot crossings would add up coherently.
-Other possible explanations such as
+This fairly contrived geometry would be needed in order to
+explain why the starspot crossings could add up coherently.
+Other possibilities including
 gravity darkening or TTVs causing the asymmetry are disfavored (see
 Appendix~\ref{app:asymmetry}).
-Aside from the asymmetric transits, Appendix~\ref{app:flare}
+
+Beyond the asymmetric transits, Appendix~\ref{app:flare}
 highlights an additional abnormality in the short-cadence Kepler data, in the
@@
 performed asychronously on the Gaia archive
-website.\footnote{\url{https://gea.esac.esa.int/archive/}}
+website.
 
@@
 the night of 2021 Aug 7.  We used the iodine cell for wavelength
-calibration\footnote{We considered a line-profile based analysis in
-the regions without iodine, but the line profile stability of HIRES
-precludes such an approach.}, and extracted the 1-D spectra using the
+calibration, and extracted the 1-D spectra using the
 standard California Planet Survey pipeline \citep{howard_cps_2010}.

### 2021-10-11 paper/collect.sh

@@ -1,4 +1,4 @@
-cp /Users/luke/Dropbox/proj/rudolf/results/XYZvtang/XYZvtang_show1627_showcomovers_showsun.pdf f1.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/XYZvtang/XYZvtang_show1627_showcomovers_showsun_portrait.pdf f1.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/skychart/skychart_showkepler_showcomovers.pdf f2.pdf
 

### 2021-10-11 paper/ms.tex
update ms

@@
 period measurements ($G<17$ and $\bpmrpo>0.5$) using the TESS full
-frame image data.\footnote{Kepler rotation periods were derived by
-\citet{mcquillan_rotation_2014} for 28 stars; for these cases, we
-adopted the Kepler rotation period.  } We extracted light curves from
+frame image data. 
+We extracted light curves from
 the TESS images using the nearest pixel to each star, and regressed
@@
 Within this radius, we recorded the number of stars with greater
-brightness than the target star, and with brightness within 1.25 and
-2.5 TESS magnitudes of the target star.
+brightness than the target star, and with brightness within 1.25 TESS magnitudes of the target star.
 
@@
 discussed in Section~\ref{sec:kinematicselection} in the spatial and
-kinematic proximity of Kepler\,1627.  To avoid blending, we also only
-focused on stars for which no companions were known with a brightness
+kinematic proximity of Kepler\,1627.  
+Kepler rotation periods were derived by
+\citet{mcquillan_rotation_2014} for 28 of these stars; for these cases, we
+simply adopted the Kepler rotation period.
+For the remaining stars with only TESS data, we
+focused only on the stars for which no companions were known with a brightness
 exceeding one-tenth of the target star in a $21\farcs0$ radius.  There
-were \nkinwithtessandcrowding\ stars that met these requirements, and
+were \nkinwithtessandcrowding\ stars that met these crowding requirements, and
 that had TESS data available.  For plotting purposes we then imposed a
@@
 The lower panel of Figure~\ref{fig:age} shows the resulting
-\nkindefaultcleaning\ stars.  The majority of these stars fall below
+\nkindefaultcleaning\ stars.  
+The majority of these stars fall below
 the ``slow sequence'' of the Pleiades, consistent with a gyrochronal
@@
 \label{sec:stars}
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f4a.pdf}
+			\includegraphics[width=0.48\textwidth]{f4b.pdf}
+		}
+	\end{center}
+	\vspace{-0.5cm}
+	\caption{
+		{\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
+		from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2.5V
+		companion at $\rho \approx 0\farcs16$, which corresponds to a
+		projected separation of $53\pm4$\,AU.  The inset shows a cutout of
+		the stacked NIRC2 image (North is up, East is left, scale is set
+		by the separation of the binary).  The lines show 5-$\sigma$
+		contrast limits for the `Alopeke filters, and 6-$\sigma$ contrast
+		limits for NIRC2 outside of $0\farcs15$. {\it Right:} Gaia EDR3
+		renormalized unit weight error (RUWE) point estimates for
+		candidate $\delta$\,Lyr cluster members.  Since other members of
+		the cluster with similar brightnesses have comparable degrees of
+		photometric variability, the high RUWE independently suggests that
+		Kepler\,1627 is a binary. 
+		\label{fig:kep1627binary}
+	}
+\end{figure*}
 
@@
 \subsection{Kepler\,1627B}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f4a.pdf}
-			\includegraphics[width=0.48\textwidth]{f4b.pdf}
-		}
-	\end{center}
-	\vspace{-0.5cm}
-	\caption{
-    {\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
-    from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2.5V
-    companion at $\rho \approx 0\farcs16$, which corresponds to a
-    projected separation of $53\pm4$\,AU.  The inset shows a cutout of
-    the stacked NIRC2 image (North is up, East is left, scale is set
-    by the separation of the binary).  The lines show 5-$\sigma$
-    contrast limits for the `Alopeke filters, and 6-$\sigma$ contrast
-    limits for NIRC2 outside of $0\farcs15$. {\it Right:} Gaia EDR3
-    renormalized unit weight error (RUWE) point estimates for
-    candidate $\delta$\,Lyr cluster members.  Since other members of
-    the cluster with similar brightnesses have comparable degrees of
-    photometric variability, the high RUWE independently suggests that
-    Kepler\,1627 is a binary. 
-    \label{fig:kep1627binary}
-	}
-\end{figure*}
 
@@
 
-% \clearpage
+\clearpage
 \acknowledgements
@@
 therefore be recovered using tools such as BANYAN
-$\Sigma$.\footnote{\url{http://www.exoplanetes.umontreal.ca/banyan/banyansigma.php}}.
+$\Sigma$ \citep{Gagne2018a}.
 We also included the SIMBAD object identifiers $\texttt{TT*}$,
@@
     seems to be present in three of the four bins.
-    \vspace{-0.5cm}
 		\label{fig:phaseslope}
@@
 \label{app:flare}
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f13.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Flares in Kepler\,1627}.  
+		{\it Top:}
+		The full short cadence Kepler dataset, acquired at 1-minute
+		sampling (black points) is shown with a stellar variability model
+		(blue line).
+		{\it Middle:}
+		Residual after subtracting the stellar variability model.  Flares
+		appear as spikes.
+		{\it Bottom:}
+		Zooms of the brightest, and third-brightest flares.  A timing
+		coincidence -- that both flares have ``successors'' approximately
+		one orbital period after the initial event -- is emphasized.
+		\label{fig:flarezoom}
+	}
+\end{figure*}
+
+%\begin{figure*}[t]
+%	\begin{center}
+%		\leavevmode
+%		\includegraphics[width=0.9\textwidth]{f8.pdf}
+%	\end{center}
+%	\vspace{-0.7cm}
+%	\caption{
+%		{\bf Phase-folded flares in Kepler\,1627}.  
+%    {\it Top:}
+%    As in the middle of Figure~\ref{fig:flarezoom}, phase-folded at
+%    the planet's ephemeris.
+%    {\it Bottom:}
+%    Fitted flare amplitudes and orbital phases -- each point
+%    represents one flare.
+%    Colors indicate relative time, from the beginning of the
+%    98-day short cadence Quarter 15 dataset (dark blue) to the end (light
+%    yellow).
+%    Lower amplitude flares likely exist in the data, and were not
+%    examined.
+%		\label{fig:flarephase}
+%	}
+%\end{figure*}
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.8\textwidth]{f14.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Statistics of inter-flare arrival times}.  
+		24 flares were recorded with amplitudes exceeding 0.5\% over the
+		97.7 days of short cadence observations.  The histogram of the
+		time intervals between every possible pair of flares is shown in
+		black.  Some plausibly important timescales for star-planet
+		interactions, namely the planetary orbital period and synodic
+		period (the orbital period as seen from the rotating stellar
+		frame) are shown along with their linear combinations.  Monte
+		Carlo draws from a Poisson distribution are shown with the gray
+		bands.  While peaks in the observed distribution do coincide with
+		some of the ``special periods'', the statistical evidence for a
+		non-Poissonian process driving the flares does not reach the
+		$5$-$\sigma$ threshold.
+		\label{fig:flarestats}
+	}
+\end{figure*}
 
@@
 
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f13.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Flares in Kepler\,1627}.  
-		{\it Top:}
-		The full short cadence Kepler dataset, acquired at 1-minute
-		sampling (black points) is shown with a stellar variability model
-		(blue line).
-		{\it Middle:}
-		Residual after subtracting the stellar variability model.  Flares
-		appear as spikes.
-		{\it Bottom:}
-		Zooms of the brightest, and third-brightest flares.  A timing
-		coincidence -- that both flares have ``successors'' approximately
-		one orbital period after the initial event -- is emphasized.
-		\label{fig:flarezoom}
-	}
-\end{figure*}
-
-%\begin{figure*}[t]
-%	\begin{center}
-%		\leavevmode
-%		\includegraphics[width=0.9\textwidth]{f8.pdf}
-%	\end{center}
-%	\vspace{-0.7cm}
-%	\caption{
-%		{\bf Phase-folded flares in Kepler\,1627}.  
-%    {\it Top:}
-%    As in the middle of Figure~\ref{fig:flarezoom}, phase-folded at
-%    the planet's ephemeris.
-%    {\it Bottom:}
-%    Fitted flare amplitudes and orbital phases -- each point
-%    represents one flare.
-%    Colors indicate relative time, from the beginning of the
-%    98-day short cadence Quarter 15 dataset (dark blue) to the end (light
-%    yellow).
-%    Lower amplitude flares likely exist in the data, and were not
-%    examined.
-%		\label{fig:flarephase}
-%	}
-%\end{figure*}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f14.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Statistics of inter-flare arrival times}.  
-    24 flares were recorded with amplitudes exceeding 0.5\% over the
-    97.7 days of short cadence observations.  The histogram of the
-    time intervals between every possible pair of flares is shown in
-    black.  Some plausibly important timescales for star-planet
-    interactions, namely the planetary orbital period and synodic
-    period (the orbital period as seen from the rotating stellar
-    frame) are shown along with their linear combinations.  Monte
-    Carlo draws from a Poisson distribution are shown with the gray
-    bands.  While peaks in the observed distribution do coincide with
-    some of the ``special periods'', the statistical evidence for a
-    non-Poissonian process driving the flares does not reach the
-    $5$-$\sigma$ threshold.
-		\label{fig:flarestats}
-	}
-\end{figure*}
-
 There were $N_{\rm f}=24$ flares that exceeded $0.5\%$ in relative

### 2021-10-11 rudolf/plotting.py

@@ -674,15 +674,23 @@ def plot_XYZvtang(outdir, show_1627=0, save_candcomovers=1, save_allphys=1,
         )
     elif orientation == 'portrait':
         factor=1
-        fig = plt.figure(figsize=(factor*3,factor*5.5))
+        fig = plt.figure(figsize=(factor*3,factor*6.5))
         axd = fig.subplot_mosaic(
             """
-            AAAA
-            BBCC
-            DDDD
-            """,
+            A
+            B
+            C
+            D
+            """
+            #"""
+            #AAAA
+            #BBCC
+            #DDDD
+            #"""
+            ,
             gridspec_kw={
-                "height_ratios": [2,1,1],
+                #"height_ratios": [2,1,1],
+                "height_ratios": [2.5,1,1,2.5],
             }
         )
 
@@ -760,32 +768,36 @@ def plot_XYZvtang(outdir, show_1627=0, save_candcomovers=1, save_allphys=1,
 
         elif k == 'B':
             delta_x = 0.1
-            axd['B'].arrow(0.20, 0.69, delta_x, 0, length_includes_head=True,
-                           head_width=1e-2, head_length=1e-2,
+            x0 = 0.15
+            axd['B'].arrow(x0, 0.69, delta_x, 0, length_includes_head=True,
+                           head_width=3e-2, head_length=1e-2,
                            transform=axd['B'].transAxes)
-            axd['B'].text(0.20+delta_x/2, 0.71, 'Galactic\ncenter',
+            axd['B'].text(x0+delta_x/2, 0.71, 'Galactic center',
                           va='bottom', ha='center',
                           transform=axd['B'].transAxes, fontsize='xx-small')
 
         elif k == 'C':
             delta_x = 0.1
-            axd['C'].arrow(0.20, 0.69, delta_x, 0,
-                         length_includes_head=True, head_width=1e-2,
+            x0 = 0.15
+            axd['C'].arrow(x0, 0.69, delta_x, 0,
+                         length_includes_head=True, head_width=3e-2,
                          head_length=1e-2,
                          transform=axd['C'].transAxes)
-            axd['C'].text(0.20+delta_x/2, 0.71, 'Galactic\nrotation',
+            axd['C'].text(x0+delta_x/2, 0.71, 'Galactic rotation',
                           va='bottom', ha='center',
[…]

### 2021-10-12 paper/ms.tex
update ms, and AAS journals submitted

@@
-\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
+%\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
+\documentclass[12pt,modern,twocolumn,tighten,linenumbers]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
@@
 %% Add "Submitted to " argument.
-\received{---}
+\received{2021 Oct 12}
 \revised{---}
@@
   is one of $\sim$10$^3$ coeval groups whose properties have been clarified
-  by Gaia.  Many other exoplanet hosts are potentially kinematic members
-  of these clusters; their candidate memberships can be verified by
-  measuring stellar rotation periods in the TESS full frame images.
+  by Gaia.  Many other exoplanet hosts are candidate members
+  of these clusters; these memberships can be verified with the
+  trifecta of Gaia, TESS, and ground-based spectroscopy.
 \end{abstract}
@@
 Specifically, in the models of \citet{owen_atmospheres_2016} and
-\citet{owen_constraining_2020}, the $\approx5\,M_\oplus$ planets start
-with sizes of 4--12\,$R_\oplus$ shortly after the time of disk
+\citet{owen_constraining_2020}, the $\approx$$5\,M_\oplus$ planets start
+with sizes of 4\,--\,12\,$R_\oplus$ shortly after the time of disk
 dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
-2--4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
+2\,--\,4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
 is expected to occur within the first few million years after the disk
@@
 irradiation and internal heat can also power gradual outflows
-which, if strong enough, can deplete or entirely strip the atmospheric
-envelope \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
+which, if strong enough, can deplete or entirely strip the atmosphere \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
 Discovering young planets, measuring their masses, and detecting their
@@
 is often invoked to explain the observed radius distribution of mature
-exoplanets \citep{Fulton_et_al_2017}.
+exoplanets \citep{Fulton_et_al_2017,van_eylen_asteroseismic_2018}.
 
@@
 \citet{KounkelCovey2019} reported that Kepler~1627 (KIC 6184894; KOI
-5245; TIC 120105470) is a \cn\ member.  Given the previous statistical
+5245) is a \cn\ member.  Given the previous statistical
 validation of the close-in mini-Neptune Kepler 1627b
 \citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
-we begin by examining the properties of the cluster
+we begin by scrutinizing the properties of the cluster
 (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr cluster is
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.48\textwidth]{f1.pdf}
+		\includegraphics[width=0.95\textwidth]{f1.pdf}
 	\end{center}
@@
     \citet{KounkelCovey2019}.  The tangential velocities relative to
-    \sn\ (bottom row) are computed assuming that every star has the
+    \sn\ (bottom right) are computed assuming that every star has the
     same three-dimensional spatial velocity as \sn.  Our analysis
@@
     	coordinates.} Each black circle
-    is a candidate cluster member selected from the catalog of
-    \citet{KounkelCovey2019} based on its position and kinematics.  Of
+    is a candidate cluster member selected based on its position and kinematics (Figure~\ref{fig:XYZvtang}).  Of
     the \nkinematic\ candidate cluster members, 58 have at least one
@@
 planet size is consistent with those previously reported by
-\citet{morton_false_2016} and \citet{berger_identifying_2018}, and
+\citet{morton_false_2016} and \citet{berger_2018_radii}, and
 corrects for the small amount of flux dilution from Kepler\,1627B.
@@
 although this asymmetry is within the 2-$\sigma$ model uncertainties, the
-model has a jitter term that grows to account for otherwise unmodeled
+model has a jitter term that grows to account for excess
 white noise in the flux.  The significance of the asymmetry is
@@
     ($\approx$2-$\sigma$), which
-    could be explained by starspot crossings during the first
+    might be caused by starspot crossings during the first
     (second) half of transit inducing a positive (negative) TTV,
@@
 correlations and the unresolved starspot crossings that we expect to
-be present in the data for \sn\ \citep{mazeh_time_2015}, we opted to re-examine the
+be present in the \sn\ light curve \citep{mazeh_time_2015}, we opted to re-examine the
 individual transit times.
@@
 \end{figure}
+\begin{figure*}[!t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f8.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
+		Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
+		3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
+		its uncertainty.  \pn\ is shown with a star.  The large sizes of
+		the youngest transiting planets could be explained by their
+		primordial atmospheres not yet having evaporated; direct
+		measurements of the atmospheric outflows or planetary masses would
+		help to confirm this expectation.  Selection effects may also be
+		important.  Parameters are from the NASA Exoplanet Archive (2021
+		Sept 15).
+		\label{fig:rp_period_age}
+	}
+\end{figure*}
 
@@
 and egress phase that each span $\approx$40\% of the transit duration
-($\approx 68\,{\rm minutes}$).  The actual measured ingress and egress
-duration is $\approx 17\,{\rm minutes}$, a factor of four times too
+($\approx 68\,{\rm min}$).  The actual measured ingress and egress
+duration is $\approx 17\,{\rm min}$, a factor of four times too
 short.  The combination of Kepler 1627B's brightness, the transit
@@
 rotation signals in the Kepler photometry, and {\it iii)} the proximity
-of \sn\ to the main cluster isochrone on the Gaia CAMD (Figure~\ref{fig:age}).
+of \sn\ to the \cn\ locus on the Gaia CAMD (Figure~\ref{fig:age}).
 
@@
 ephemeris and transit depth from the Kepler data, the MuSCAT3 observations also
-suggested a transit duration $17.3\pm4.3\ {\rm minutes}$ shorter than the
+suggested a transit duration $17.3\pm4.3\ {\rm min}$ shorter than the
 Kepler transits.
-However, given both the lack of transit duration variations in the Kepler data
+However, given both the lack of TDVs in the Kepler data
 and the relatively low signal-to-noise of the MuSCAT3 transit,
@@
 \label{sec:conc}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f8.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
-    Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
-    3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
-    its uncertainty.  \pn\ is shown with a star.  The large sizes of
-    the youngest transiting planets could be explained by their
-    primordial atmospheres not yet having evaporated; direct
-    measurements of the atmospheric outflows or planetary masses would
-    help to confirm this expectation.  Selection effects may also be
-    important.  Parameters are from the NASA Exoplanet Archive (2021
-    Sept 15).
-		\label{fig:rp_period_age}
-	}
-\end{figure*}
 
@@
 expected for young mini-Neptunes, and the amplitude of the observed
-asymmetry is also consistent with predictions \citep{wang_dai_2019}.
+asymmetry is consistent with predictions \citep{wang_dai_2019}.
 A second possibility is that the planetary orbit is slightly
@@
 groups at favored stellar latitudes.  
-This fairly contrived geometry would be needed in order to
-explain why the starspot crossings could add up coherently.
+This geometry would be necessary in order to
+explain how the starspot crossings could add up coherently.
 Other possibilities including
@@
 the age of \pn.   This measurement was enabled by identifying the
-connection of the star to the \cn\ using Gaia kinematics, and then by
+connection of the star to the \cn\ using Gaia kinematics, and by then 
 using the Gaia color-absolute magnitude diagram and TESS stellar
@@
 
-\clearpage
+%\clearpage
 \acknowledgements
@@
 Fellowship from Princeton University, as well as from the TESS GI
-Program, programs G011103 and G022117, through NASA grants
-80NSSC19K0386 and 80NSSC19K1728.
+Program (NASA grants 80NSSC19K0386 and 80NSSC19K1728) and the
+Heising-Simons Foundation (51 Pegasi~b Fellowship).
+% programs G011103 and G022117, through 
 %
@@
     {\it Right:}
-    As on the left, with the transit removed.  Quarters 6 and 7 show a
-    consistent deviation in the second half of the transit.
+    As on the left, with the transit removed.  
 		\label{fig:phasequarter}
@@
 our analysis, the mean TTV and its standard deviation are
-$0.66\pm5.53$ minutes; similarly the mean local slope and its
+$0.66\pm5.53$\,min; similarly the mean local slope and its
 standard deviation are $0.59\pm45.50$\,ppt\,day$^{-1}$.  There is
@@
 		{\it Top:}
-		The full short cadence Kepler dataset, acquired at 1-minute
-		sampling (black points) is shown with a stellar variability model
+		The 1-minute cadence Kepler data (black points) is shown with a stellar variability model superposed
 		(blue line).
@@
 		some of the ``special periods'', the statistical evidence for a
-		non-Poissonian process driving the flares does not reach the
+		non-Poissonian process driving the flares does not clear the
 		$5$-$\sigma$ threshold.

### 2021-10-12 paper/clean.sh

@@ -1,0 +1,1 @@
+rm ms.synctex.gz ms.log ms.out ms.aux ms.blg 

### 2021-10-12 paper/collect.sh

@@ -1,4 +1,4 @@
-cp /Users/luke/Dropbox/proj/rudolf/results/XYZvtang/XYZvtang_show1627_showcomovers_showsun_portrait.pdf f1.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/XYZvtang/XYZvtang_show1627_showcomovers_showsun.pdf f1.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/skychart/skychart_showkepler_showcomovers.pdf f2.pdf
 

### 2021-11-24 paper_review/response_1.txt
stage paper_review

Dear Dr. Jackson,

We thank you for organizing the review of our paper, and also thank
the referee for constructive criticism.  We have considered the
referee's remarks carefully and revised our manuscript accordingly.

Below, we have reproduced the relevant portions of the referee's
report, alongside our responses. A list of changes is appended to
the manuscript using the trackchanges AASTeX macros.

Sincerely,

Luke Bouma

******************************************
Reviewer #1:

REVIEWER COMMENT
----------
Line 57 (evaporation): Consider adding citation to Lopez et al. (2012) or
similar

RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Line 78 (Kepler ages): Consider adding citation to Walkowicz & Basri (2013)

RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Line 93: Clarify origin of name delta Lyr, given there are two pre-existing
names, and how they overlap. It might also be worth mentioning Kerr's Cep-Her
association as in prep just to save future readers the headache of
cross-identifying them, either here or in Sec 2.1

RESPONSE
----------
> delta Lyr cluster was the first name suggested by Stephenson (1958),
> and it is the default name listed in SIMBAD.  It was also the name
> used in the title of Eggen's 1967 paper.
>
> The cluster was called Stephenson-1 by people other than Stephenson,
> including by Kounkel & Covey 2019 -- their Theia catalog identifiers
> are simply given to any cluster in their catalog, in addition to any 
> previous names that existed.
>
> The overlap between the earlier lists (Stephenson 1958, Eggen 1968,
> Kharchenko et al. 2012) and the Theia 73 list from Kounkel & Covey
> 2019 is only present in the stars closest to delta Lyr itself.  For
> instance, see the 30 stars shown in Figure 1 of Eggen 1968.  The
> broader "complex" shown in Figure 1 of our paper, which to our
> knowledge first reported to be a set of kinematically associated
> stars by Kounkel & Covey (inasmuch as an "HDBScan group"
> implies kinematic association).
>
> Regarding Ronan Kerr's view of the delta Lyr cluster as a component
> of the larger Cep-Her complex, we would prefer to leave this as
> Ronan's result to present!  There's significant work here in
> designing the clustering method and verifying that it works.
> There's also work in understanding the overall structure of the
> broader association (which is different in nuanced ways from the
> gray Theia 73 points in Figure 1).  Given the complexity, and the
> potential importance of the complex given its proximity and age, we
> really would prefer to leave this for the in prep R. Kerr work.  And
> similarly, the introduction of the term "Cep-Her complex" is
> probably best left for that future work.

REVIEWER COMMENT
----------
Line 165: Consider mentioning the axes on which data was cleaned (e.g. color
precision)

RESPONSE
----------
> Done - thank you.  We have added the following sentence: "This also
> involved cuts on the photometric signal to noise ratio, the number
> of visibility periods used, the astrometric $\chi^2$ of the
> single-source solution, and the $G_{\rm BP}-G_{\rm RP}$ color excess
> factor."

REVIEWER COMMENT
----------
Line 179 (reddening): What does "reasonably well" mean? Something like 1-2
sigma?

RESPONSE
----------
> Thank you for catching the weasel words.  The revised text now reads
> as follows: "These values \replaced{agree reasonably well with}{are
> within a factor of two of} previously reported values
> \replaced{from}{in} the literature \citep[{\it
> e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}\added{,
> and are all small enough that the choice of whether to use them {\it
> vs.} other extinction estimates does not affect our primary
> conclusions}."
>
> More quantitatively, from the Lallement maps we report
> $E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$ for IC
> 2602, the Pleiades, and the delta Lyr cluster.  The relevant
> literature reports:
>
>    Gaia+2018 E(B-V):
>    Pleiades 0.045
>    IC2602 0.031
>   
>    Randich+2018 E(B-V):
>    IC2602 0.032 - 0.092
>   
>    KC2019: E(B-V)
>    Pleiades 0.071
>    IC2602 0.078
>    del Lyr cluster 0.059
>   
>    Bossini+2019 E(B-V)
>    Pleiades/Melotte22 0.045
>    IC2602 0.031
>    del Lyr cluster 0.051
>
> hence our assessment of "it's within a factor of two, and doesn't
> matter much."  Given that all these clusters are within ~400
> parsecs, this perhaps isn't too surprising.

REVIEWER COMMENT
----------
Fig 3: It may be helpful to include one of the even younger clusters on the CMD
(i.e. to bracket delta Lyra), though perhaps it makes the plot too busy.

RESPONSE
----------
> Please see camd_with_UCL.png and camd_with_UCL_no_redcorr.png
> attached.  This approach indeed makes the plot a bit too busy, in
> our opinion.

REVIEWER COMMENT
----------
Page 3: Consider re-organizing. Based on the 1st paragraph, I thought the
quantitative analysis would only deal with Pleiades and IC 2602 and one might
infer reddening was only calculated for these clusters (whereas I assume the
latter are also corrected for reddening prior to the analysis described). My
suggestion would be to introduce all the clusters first. It would also be
helpful to list the age of IC 2602 at some point in the text.

RESPONSE
----------
> Regarding the main re-organization point, FIXME TODO REORGANIZE?
>
> Regarding the last point, the final paragraph of Section 2.2.1
> discusses the age of IC 2602 in the text.  There's a fairly wide
> spread of reported literature ages between 30 and 46 Myr -- we
> ultimately adopted the half-way point between the isochrone and LDB
> ages.

REVIEWER COMMENT
----------
Line 202 (binaries): Add RUWE citation(s). Nice approach to remove
radial_velocity_error based on each cluster's distribution. It should be noted
somewhere in this discussion that the target itself does *not* meet the RUWE
limit and so wouldn't appear on the plot by default.

RESPONSE
----------
> We've added a footnote to the Gaia DPAC memo by Lindegren on the
> RUWE.
>
> The stars shown in Figure 3 are those discussed in the first
> paragraph of Section 2.2.1.  (The binarity cut has not yet been
> applied -- which is why the binary sequence is visible for the δ Lyr
> cluster).  So for the latter point, we have not made any changes.

REVIEWER COMMENT
----------
Line 209: I don't follow starting "assembled a piecewise grid of hybrid
isochrones..." What does "hybrid" mean here? Once the splines for each cluster
are defined, how does one interpolate between them? Are the ages then
determined for each star and combined?

RESPONSE
----------
> The manuscript relies on the citation to Gagne+2020 for the full
> description of this method.  The "hybrid" isochrone I_i is the
> linear combination of any two reference isochrones.  The cited
> Equation 6 of Gagne+2020's mu-Tau paper gives the definition:
>
>   I_i = α_i * I_45 + (1-α_i)*I_112,
>
> for α_i a coefficient between 0 and 1, I_45 the reference 45 Myr
> isochrone spline (from Tuc-Hr, Columba, and Carina in Gagne's
> paper), and I_112 the Pleiades isochrone spline. 
>
> In our case, given our different reference clusters, the definition
> is modified to
>
>   I_i = α_i * I_16 + (1-α_i)*I_38,   (age from 16-38 Myr)
>   I_i = α_i * I_38 + (1-α_i)*I_112,  (age from 38-112 Myr)
>
> where I_38 is the IC 2602 isochrone spline, and I_16 is the UCL
> isochrone spline.
>
> The main assumption of this approach is that between the reference
> times set by the adopted clusters, the isochrones evolve linearly in
> time through color-absolute magnitude space.  If we were to attempt
> this linear interpolation all the way from UCL to the Pleiades
> without putting any intermediate point, the validity of this
> assumption would be crucial to test.  Since we include IC 2602,
> and it overlaps almost directly on the CAMD with the δ Lyr cluster,
> the only useful quantity that this method really gives us is a
> formal uncertainty on the age.  (Since the age the method returns
> will be very close to whatever age we adopt for IC 2602).
>
> To better illustrate the method, we've attached a figure to the
> response, "age_interpolation_example.png". This shows the CAMD data
> from the clusters after all the cleaning steps, their reference
> splines (constructed from bins over a select color range), and the
> interpolated splines spaced every 4.4 Myr between UCL and IC2602,
> and every 14.8 Myr betwen IC2602 and Pleiades.  While we do interpolate
> over a much finer grid, this is not shown to improve the clarity of
> the figure.  The age of δ Lyr is determined by using all the stars
> in the displayed color range, through the referenced Equation 7 of
> Gange+2020, which is a Gaussian likelihood that treats the
> interpolated isochrone as the "model" and the δ Lyr cluster's
> isochrone as the "data".  The statistical uncertainty on the age is
> then the standard deviation of a Gaussian fit to the posterior
> probability (in log-age; it becomes asymmetric when transformed to
> linear age).  We've added a few sentences to the manuscript to
> clarify this latter point.

REVIEWER COMMENT
----------
Line 230 (age): Consider adopting a systematic error (perhaps corresponding to
the range noted for the extremes of the ages) in addition to the random
(assumed to be the quoted +6/-5 error). Relatedly, how are the quoted errors
derived?

RESPONSE
----------
> We're comfortable quoting the statistical error derived from the
> standard deviation of the logarithmic age posterior since this
> distribution is roughly Gaussian.  The worry with quoting a
> systematic error, e.g., "The age of the δ Lyr cluster is
> $38^{+6}_{-5}$ (stat) $^{+6}_{-8}$ (sys)" is that the systematic
> uncertainties are not Gaussian and are not based on any probability
> distribution that we can easily assess -- they're estimated based on
> the extremes of previous literature age measurements.  Quoting this
> systematic range next to a number that has been derived based on a
> more formal approach (and that has statistical meaning within that
> approach) somehow misses the point.  Given the roughly comparable
> amplitude of the two sources of uncertainty, does it matter?  Open
> cluster ages are not H_0.

REVIEWER COMMENT
----------
Line 235: May be worth mentioning the restriction to the 391 stars in the
beginning of sec 2.2.2.

RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Line 278: "Indistinguishable" -- from each other or from the Pleiades?

RESPONSE
----------
> From each other (and the manuscript has been updated; thank you!).

REVIEWER COMMENT
----------
Line 286: I thought I remember Curtis or Douglas citing binarity as a
possibility for stars above to slow sequence as well, but I don't have a
reference on-hand.

RESPONSE
----------
> Thanks for this point -- we've updated the text with a reference to
> Stauffer+16.  In their Section 5.1, they discuss the case of some
> unresolved F+K binaries in the Pleiades which show anomalously long
> K2 rotation periods.

REVIEWER COMMENT
----------
Section 3.2: The text mentions statistical and systematic uncertainties but I
only see one set of errors listed in Table 1. Were these uncertainties combined
or was only one adopted in the end?

RESPONSE
----------
> They were combined in quadrature (see Line 331 of the original
> manuscript, "Reported uncertainties are a quadrature sum of the
> statistical and systematic components."

REVIEWER COMMENT
----------
Line 391: It would be helpful to state the star's RUWE in the text here.
Additionally, I believe RUWE acornym was already defined previously.

RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Section 3.2: Comment on whether the binary companion should effect the rotation
measurement of the host star (either here, or in the rotation section).

RESPONSE
----------
> We've added the following sentence at the beginning of Section 4.1,
> when the Kepler light curve is introduced
>
> """
>The resulting photometry is dominated by a quasi-periodic starspot
>signal with a peak-to-peak amplitude that varies between 2\% and 8\%.
>\added{Given that the secondary companion's brightness in the Kepler
>band is 1\% to 2\% that of the primary, source confusion for the
>rotation signal is not expected to be an issue.}
 """
> The rationale here is that the rotation amplitude on the secondary
> would need to be a factor of >two times its intrinsic brightness for
> any confusion to happen, and the primary would need to not show any
> rotation signal.  Such large variability amplitudes have not been
> observed except for disked stars (e.g., Rebull+2020, Figure 10), and
> there is no evidence for a secondary period in the Lomb Scargle
> periodograms.

REVIEWER COMMENT
----------
One note about the transit analysis: The MuSCAT3 data was introduced as
supporting evidence but the conclusion (that the transits appear achromatic)
was never explicitly stated.

RESPONSE
----------
> Line 642 of the original manuscript notes "Although we did not
> detect the RM anomaly, the multi-band MuSCAT3 light curves show that
> the transit is achromatic (Figure 7)."

REVIEWER COMMENT
----------
Line 725: The fast rotation period and flares could plausibly effect the
completeness.

RESPONSE
----------
> These completeness statistics were computed using KeplerPORTS (the
> Burke & Catanzarite 2021 citation).  KeplerPORTS does correct for
> the red noise in the light curves by using the CDPP over different
> timescales -- see Section 3.3 of Burke, C. J. & Catanzarite, J.
> 2017, Planet Detection Metrics: Per-Target Detection Contours for
> Data Release 25, KSCI-19111-002.  Nonetheless, we do agree that this
> is a worry -- perhaps a full injection-recovery analysis of the
> light curve with a pipeline of our own might help verify whether
> these detection limits are plausible.  This effort is outside our
> scope, for now.

REVIEWER COMMENT
----------
Line 729: Considering Owen 2020, contraction could also contribute to some
extent to the larger radii of small planets; might this be significant?

RESPONSE
----------
> TODO WHAT?

REVIEWER COMMENT
----------
Appendix A paragraph beginning at line 1158: May be worth mentioning that delta
Lyr may also be part of one of these heirarchical associations.

RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Figure 13 caption: Specify planetary orbital period since technically there is
also a stellar orbital period involved.

RESPONSE
----------
> Done, thank you.

******************************************
Reviewer #2:

REVIEWER COMMENT
----------
Abstract
line 24: Kepl-1627Ab cannot be a "mini-Neptune" if it is same size of Neptune.
Replace with "Neptune-size" Yes, it falls within a distribution of objects in a
period-radius diagram with an average that is smaller than Neptune, but
claiming affiliation with smaller objects implies more than we know about this
planet, e.g. its mass.

RESPONSE
----------
> Point well-taken. 
> R_p = 3.79 +/- 0.16 R_earth, so yes, this is the same size as
> Neptune within 1-sigma.

REVIEWER COMMENT
----------
line 25: Kepler target stars were chosen for their suitability for transiting
detection of planets in general, and for planets that are Earth-analogs in
paraticular. That involves radius, brightess. Distance and spectral types were
not direct selection criteria; this was before Gaia.

RESPONSE
----------
>

REVIEWER COMMENT
----------
line 28: "main" --> "prime"

RESPONSE
----------
>

REVIEWER COMMENT
----------
Intro

Line 47: (First paragraph on p 2) the time-scale of 1e8 is associated with one
particular set of photoevaporation models; core-powered mass loss proposes
timescales of Gyr(s) (see the Ginzburg paper that was cited). More recent
photoevaporation models also propose 1 Gyr timescales. It would be good to
revise/update this paragraph.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 2

Line 65: it is convention to italicize names of space missions like TESS and K2

RESPONSE
----------
>

REVIEWER COMMENT
----------
Lines 142-144: perhaps rephrasing this would clarify that the authors did not
use Kepler-1627 as a fiducial star about which to select cluster members, and
that they restricted the analysis to one of several sub-clusters or clumps
which contained Kepler-1627.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Figure 1: the first three panels would be more effective if they were plotted
at the same scale and with appropriately aligned axes. As such the upper right
panels are too small to be useful. Consider leaving out the Sun (which is
irrelevant) so as to have a more appropriate choice of scale to how the details
of the cluster. The purpose of the fourth panel (tangential velocities) and the
second line in the caption are unclear to me. Aren't the tangential velocities
known from Gaia proper motions and parallaxes? Why calculate them with the
assumption of common space motion? If that is done, what does it show, anyway?

RESPONSE
----------
>

REVIEWER COMMENT
----------
Figure 3 bottom: Can the IC 2602 and 2391 rotation periods be added to this
plot? Lines 276-278 mention that these and cite Douglas et al. 2021 (although
the reference does not seem to be complete). Isn't it interesting that the
interlopers all fall on the Praesepe rotation sequence?

RESPONSE
----------
>

REVIEWER COMMENT
----------
p 5 lines 258-261: rather than simply throw the TESS values out in favor of the
Kepler ones, the authors should compare them to evaluate the accuracy and
reliability of the TESS values.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 3

page 7 line 391: report the RUWE value (yes its in the Table but don't make the
reader work)

RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 4.1.3 and Figure 6

The significance of this slope is unclear and a sigma is not a good way to
evaluate it. Perform a ratio of chi-squared or F-test to evaluate it vs. a flat
line.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Lines 530-542: does option (ii) include a transit fit as well? I assume it does
but it would be good to clarify. And is option (i) the 2nd and 4th order
polynomial? Perhaps just state this?

RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 561-562: do the authors mean "lack of statistical significance"?

RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 4.2

Analysis of the motion of the photocentroid has been widely used to help
validate Kepler planets. Can this be used here to distinguish between the
possible A and B host stars?

RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 604-608: The statement that the density from the transit duration is
inconsistent with that expected for the M dwarf "B" cites Table 2 and
Choi+2016, but that Table does not appear to provide any information on "B",
and the Choi+2016 reference is just for the Mesa models. A density of 2 g/cm3
(or lower) is possible for an M dwarf on the pre-main sequence, depending on
the mass, age and model used.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Lines 613-614: it would be good to quantify this, i.e. how faint an unresolved
companion or background star must be based on the star's position in the CMD.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Lines 635-638: Here, new observations appear for the first time but are not
described. While the details are given in the appendix, and the outcome of the
observations doesn't define this work, the authors need to give some basic
description of the observations, e.g. passbands, SNR, resolution, observation
times, etc. The authors should not assume the reader actually know what HIRES
and MuSCAT are!

RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 5

line 670: assuming that SNR is not to blame, the problem is not precision, the
problem is either measuremnt systematics or astrophysical noise.

RESPONSE
----------
>

REVIEWER COMMENT
----------
line 678-683: The Ca II HK activity indicator "S" as well as the H-alpha
equivalent width could be obtained from the HIRES spectrum. Have the authors
checked fora X-ray or UV (Galex) detection of this star? The star is likely not
detected but its worth a look. In the same vein, the authors may wish to
comment on whether the star's infrared emission indicates or rule out a
substantial debris disk -- the star is detected in WISE (and perhaps Spitzer?)

RESPONSE
----------
>

REVIEWER COMMENT
----------
lines 699-701 It could be mentioned here that the planet orbital period and
stellar rotational period are not a rational combination.

RESPONSE
----------
>

REVIEWER COMMENT
----------
line 723-726: The caveat that should be mentioned here is that this is for
older, main sequence stars.

RESPONSE
----------
>

******************************************

Other minor changes made to the manuscript include:

* Updated binning interval on lower panels of Figure 5 from 20 to 15 minutes.

* L.G.B.'s affiliation order was reversed, to reflect the primary
  institution at which this work was performed.

### 2021-11-24 paper/authors.tex
minor ms updtaes
update ms
update ms
Co-authored-by: Motohide Tamura <motohide.tamura@nao.ac.jp>
update ms
Co-authored-by: gummiks <gstefansson@astro.princeton.edu>
update ms
Co-authored-by: adamkraus <alk@astro.as.utexas.edu>
update ms (LAH init comments)
Co-authored-by: Lynne Hillenbrand <lah@astro.caltech.edu>
[WIP] review #1 most comments addressed

@@
 \correspondingauthor{L.\,G.\,Bouma}
-\email{bouma.luke@gmail.com}
+\email{luke@astro.caltech.edu}
 
 \author[0000-0002-0514-5538]{L. G. Bouma}
+\altaffiliation{51 Pegasi b Fellow}
+\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
 \affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
@@
 \affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
+% pending
 
@@
 \author[0000-0001-7409-5688]{G. Stefansson}
+\altaffiliation{Henry Norris Russell Fellow}
 \affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
@@
 %
-\author[0000-0001-8638-0320]{A. W. Howard}
-\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
+% \author[0000-0001-8638-0320]{A. W. Howard} % preferred acknowledgements
+% \affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
 %
 \author[0000-0002-0531-1073]{H. Isaacson}
-\affiliation{Astronomy Department, University of California, Berkeley,
-CA 94720, USA}
+\affiliation{Astronomy Department, University of California, Berkeley, CA 94720, USA}
+% pending
 
@@
 
-\author[0000-0002-5658-5971]{Masahiro Ikoma} % ikoma.masahiro@gmail.com
+\author[0000-0002-5658-5971]{M. Ikoma} % ikoma.masahiro@gmail.com
 \affiliation{Division of Science, National Astronomical Observatory of Japan, Tokyo 181-8588, Japan}
@@
 \author[0000-0002-6510-0681]{M. Tamura} % motohide.tamura@nao.ac.jp
-\affiliation{Department of Astronomy, University of Tokyo, Tokyo 113-0033, Japan}
+\affiliation{Department of Astronomy, The University of Tokyo, Tokyo 113-0033, Japan}
 \affiliation{Astrobiology Center, Tokyo 181-8588, Japan}
-\affiliation{National Astronomical Observatory of Japan, Tokyo 181-8588, Japan}
+\affiliation{Division of Science, National Astronomical Observatory of Japan, Tokyo 181-8588, Japan}
 
 % AO IMAGING
+\author[0000-0001-9811-568X]{A.~L.~Kraus}
+\affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}
+
 \author[0000-0001-9800-6248]{E. Furlan} % furlan@ipac.caltech.edu
@@
 \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+\affiliation{NASA Exoplanet Science Institute, Caltech/IPAC, Pasadena, CA 91125, USA}
 
@@
 \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+% pending
 

### 2021-11-24 paper/ms.tex
[WIP] review #1 most comments addressed

@@
 %\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
-\documentclass[12pt,modern,twocolumn,tighten,linenumbers]{aastex63}
+%\documentclass[12pt,modern,twocolumn,tighten,linenumbers]{aastex63}
+\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
@@
   by Gaia.  Many other exoplanet hosts are candidate members
-  of these clusters; these memberships can be verified with the
+  of these clusters; their ages can be verified \replaced{through}{with} the
   trifecta of Gaia, TESS, and ground-based spectroscopy.
@@
 irradiation and internal heat can also power gradual outflows
-which, if strong enough, can deplete or entirely strip the atmosphere \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
+which, if strong enough, can deplete or entirely strip the atmosphere \citep{lopez_how_2012,Owen_Wu_2013,ginzburg_corepowered_2018}.
 Discovering young planets, measuring their masses, and detecting their
@@
 that younger Kepler planets do exist
-\citep{berger_identifying_2018,david_sizes_2021}, accurate and precise
+\citep{walkowicz_rotation_2013,berger_identifying_2018,david_sizes_2021}, accurate and precise
 age measurements typically require an ensemble of stars.  Fortunately,
@@
 cluster in the vicinity of \sn.
+%\added{The application of a more careful membership selection procedure to the $\delta$\ Lyr\ cluster (as part of a larger complex spanning Cepheus and Hercules) is a topic under current investigation by by R.~Kerr and collaborators.}
 
@@
 B]{GaiaCollaboration2018}, except that we weakened the parallax
-precision requirement to $\varpi/\sigma_\varpi>5$.  These filters were
+precision requirement to $\varpi/\sigma_\varpi>5$.  \added{This also involved cuts
+on the photometric signal to noise ratio, the number of visibility
+periods used, the astrometric $\chi^2$ of the single-source solution, and the
+$G_{\rm BP}-G_{\rm RP}$ color excess factor.} These filters were
 designed to include genuine binaries while omitting instrumental
@@
 reddening of $E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$.
-These values agree reasonably well with previously reported values
-from the literature \citep[{\it
-e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}.
+These values \replaced{agree reasonably well with}{are within a
+factor of two of} previously reported values
+\replaced{from}{in} the literature \citep[{\it
+e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}\added{,
+and are all small enough that the choice of whether to use them {\it vs.} other
+extinction estimates does not affect our primary conclusions}.
 
@@
 below the 80$^{\rm th}$ percentile of each cluster's distribution, and
-excluded stars that were obvious photometric binaries in the CAMD.  We
-then passed a moving box average and standard deviation across the
-CAMD in 0.10\,mag bins, fitted a univariate spline to the binned
-values, and assembled a piecewise grid of hybrid isochrones spanning
-the ages between UCL to the Pleiades using Equations~6 and~7 from
+excluded stars that were obvious photometric binaries in the CAMD.\footnote{
+  For a description of the renormalized unit weight error (RUWE), see the GAIA DPAC technical note
+  \url{http://www.rssd.esa.int/doc_fetch.php?id=3757412}.}  We then passed a
+moving box average and standard deviation across the CAMD in 0.10\,mag bins,
+fitted a univariate spline to the binned values, and assembled a
+piecewise grid
+of \deleted{hybrid }isochrones spanning the ages between UCL \replaced{to}{and} the Pleiades using
+Equation\replaced{s~6 and~7}{6} from
 \citet{gagne_mutau_2020}.  
+\added{To derive a probability distribution function for the age of \cn, we
+then assumed a Gaussian likelihood that treated the interpolated isochrones as
+the ``model'' and the $\delta$\ Lyr\ cluster's isochrone as the ``data''
+(Equation~7 from \citealt{gagne_mutau_2020}).  The cluster's age and its
+statistical uncertainty are then quoted as the the mean and standard deviation
+of this age posterior.}
 
@@
 frame image data. 
+\added{As a matter of
+scope, we restricted our attention to the \nkinwithtess\ stars
+discussed in Section~\ref{sec:kinematicselection} in the spatial and
+kinematic proximity of Kepler\,1627. }
 We extracted light curves from
@@
 We then cleaned the candidate TESS rotation period measurements
-through a combination of automated and manual steps.  As a matter of
+through a combination of automated and manual steps.  \deleted{As a matter of
 scope, we restricted our attention to the \nkinwithtess\ stars
 discussed in Section~\ref{sec:kinematicselection} in the spatial and
-kinematic proximity of Kepler\,1627.  
+kinematic proximity of Kepler\,1627.}  
 Kepler rotation periods were derived by
-\citet{mcquillan_rotation_2014} for 28 of these stars; for these cases, we
+\citet{mcquillan_rotation_2014} for 28 of \replaced{these }{our \nkinwithtess\ }stars; for these cases, we
 simply adopted the Kepler rotation period.
@@
 distributions of other 30\,Myr to 50\,Myr clusters ({\it e.g.},
-IC\,2602 and IC\,2391) are indistinguishable
+IC\,2602 and IC\,2391) are indistinguishable\added{ from each other}
 \citep{douglas_stephanie_t_2021_5131306}.  Approximately 10 of the
@@
 rotation period detection fraction would be
-$\nrotgood/\nkinwithtessandcrowding \approx \nfracprot\%$.   The other
-stars are likely to be field contaminants.  A final possible
+$\nrotgood/\nkinwithtessandcrowding \approx \nfracprot\%$.   \replaced{The other
+stars are likely to be field contaminants.}{Although some of these outlier stars might be
+unresolved F+K binaries that are in the cluster
+\citep{stauffer_rotation_2016}, assuming that they are field contaminants
+provides a more secure lower bound of the rotation period detection fraction.}  A final possible
 confounding factor -- binarity -- is known to affect the ``fast
@@
 brighter than $G < 20.5$ mag within $\rho < 120\arcsec$), its
-existence was also suggested by the primary star's large renormalized
-unit weight error (RUWE), relative to other members of the \cn.  Based
+existence was also suggested by the primary star's large \replaced{renormalized
+unit weight error (RUWE),}{RUWE} relative to other members of the \cn\
+\added{(RUWE$\approx$2.9; roughly the $98^{\rm th}$ percentile of the cluster's
+distribution)}.  Based
 on the apparent separation, the binary orbital period is of order
@@
     purple line.  Gray points are individual flux measurements; black
-    points bin these to 20 minute intervals, and have a representative
+    points bin these to 15 minute intervals, and have a representative
     1-$\sigma$ error bar in the lower right of each panel.  The asymmetric
@@
 is dominated by a quasi-periodic starspot signal with a peak-to-peak
-amplitude that varies between 2\% and 8\%.  Previous analyses have
+amplitude that varies between 2\% and 8\%.  \added{Given that the secondary
+companion's brightness in the Kepler band is $\approx$1.5\% that of the primary,
+source confusion for the rotation signal is not expected to be
+an issue.}
+Previous analyses have
 identified and characterized the smaller transit signal
@@
 instance can be divided into almost one hundred kinematic subgroups
-\citep{kounkel_apogee2_2018}.  Similar complexity applies to the
+\citep{kounkel_apogee2_2018}.  \added{Based on Figure~\ref{fig:XYZvtang}, the
+\cn\ may also be part of a similar hierarchical association.} Similar complexity applies to the
 problem of determining homogeneous ages, which we do not attempt to
@@
 		coincidence -- that both flares have ``successors'' approximately
-		one orbital period after the initial event -- is emphasized.
+		one \added{planetary }orbital period after the initial event -- is emphasized.
 		\label{fig:flarezoom}

### 2021-11-28 paper_review/response_1.txt
[WIP] reviewer #2 staging

@@
 ----------
-> Point well-taken. 
-> R_p = 3.79 +/- 0.16 R_earth, so yes, this is the same size as
-> Neptune within 1-sigma.
+> Point well-taken.  R_p = 3.82 +/- 0.16 R_earth, so yes, Kepler-1627 Ab is the
+> same size as Neptune within 1-sigma.  All instances in the text that referred
+> to Kepler-1627 Ab as a mini-Neptune have been modified to call it a
+> "Neptune-sized planet" (including the manuscript's title).
 
@@
 ----------
->
+> Indeed.  Batalha+2010 (ApJL) discuss the selection function, which favored
+> stars that were both small and bright since these stars enable the detection
+> of Earth-size planets.  M dwarfs thus were generally excluded unless they
+> happened to be very close.  F, G, and K stars were favored, particularly when
+> they were nearby.  The sentence in L25 currently states that "The star was
+> observed by the Kepler space telescope because it is nearby (d=329pc) and it
+> resembles the Sun".  More directly, yes, the star was chosen because it was
+> thought that it might enable the detection of small planets in the habitable
+> zone.  However the underlying reason for this given how the transit S/N
+> scales is that the star was relatively bright, and it also fell within a
+> particular color range.  That color range (given the Batalha+2010 logg cuts)
+> corresponds to Sun-like stars.  And the brightness range for Sun-like stars
+> corresponds to stars within a particular distance limit.  Our claim therefore
+> holds, albeit indirectly.  It also serves the more important purpose of
+> informing the reader what kind of star this paper is about: a star that is at
+> a distance of ~330 parsecs, and that resembles the Sun (except for its age!).
 
@@
 ----------
->
+> Done here and throughout the manuscript, thank you.
 
@@
 ----------
->
+> FIXME FIXME TODO
 
@@
 ----------
->
+> Yes, and we leave the imposition of this convention to the copy-editors at
+> the journal.
 
@@
 ----------
->
+> We have slightly rephrased Lines 142-144 to clarify the meaning, which is
+> indeed what the reviewer has stated.
 
@@
 ----------
+> Thank you for these important points of feedback.
 >
+> The Sun has been included to clarify the direction along which parallax
+> uncertainties are expected to produce erroneous clusterings (analogous to the
+> "Fingers of God" effect in the extragalactic literature).  This is relevant
+> because the HDBScan clustering method used by Kounkel & Covey in their 2019
+> paper likely was susceptible to this systematic effect.  A sentence has been
+> added to the caption of Figure 1 to clarify this point.
+> 
+> The {X,Y}, {Z,X}, and {Z,Y} panels are plotted at different scales in order
+> to keep the aspect ratios roughly correct.  The reported KC2019 cluster
+> members span {ΔX, ΔY, ΔZ} of roughly {400, 600, 200} parsecs,  which grows to
+> ~{600, 600, 200} parsecs when including the Sun.  The {Z,X} and {Z,Y}
+> panels should therefore roughly have aspect ratios of 1:3 in order for the
+> parsec to not be stretched too much in either dimension.  The top-down {X,Y}
+> view is the most informative dimension, which is why it is enlarged related
+> to the edge-on views (and its 1:1 aspect ratio is set by the same
+> consideration).
+>
+> Finally, the tangential velocities are important because they tell us whether
+> the structures are kinematically coherent.  Open clusters typically have
+> velocity dispersions of order ~1km/s.  Meingast et al. 2021 (A&A 645 84) for
+> instance suggest tangential velocities <2 km/s (from the mean cluster
+> velocity) as a plausible cutoff for selecting candidate cluster members.  One
+> needs to compute these as relative tangential velocities instead of just
+> looking at the proper motions because the proper motions are changed by
+> line-of-sight projection effects, especially when the clusters appear
+> stretched over tens of degrees on-sky.  See for instance Figure 2 of
+> Meingast+2021, or Figure 14 of Bouma+2021 (AJ, 162, 197).  In our case, the
+> tangential velocities indicate that the spread in Kounkel & Covey 2019's
+> reported cluster members is ~10km/s, which is large enough that it does not
+> make obvious kinematic sense to consider all of these stars to be part of the
+> same structure.  We've added a few sentences in the manuscript to clarify
+> these latter points.
 
@@
 ----------
+> Thank you for pointing out the missing link in the reference to Douglas et
+> al. 2021, which should have pointed to
+> https://ui.adsabs.harvard.edu/abs/2021tsc2.confE.157D/abstract, and has been
+> updated in the revision.  The rotation periods from this work, which includes
+> IC 2602, IC 2391, and a few other 30-50 Myr clusters, are not yet public.  
 >
+> The interlopers falling on the Praesepe sequence is indeed a bit odd, but
+> it's really only ~3-5 stars. A much bigger set of interlopers didn't have
+> detected rotation periods at all! (See final paragraph of Section 2.2.2).
+> At the other reviewer's request, we've also added a sentence noting that some
+> of these stars might be unresolved F+K binaries.
 
@@
 ----------
+> Thank you for this point -- please see kepler_vs_tess_prot.png in the
+> attached, which shows the 28 referenced stars that overlap.  5 of the 28
+> stars have Kepler rotation periods above 10 days, of which none were
+> correctly recovered by TESS (3 are chosen at the half-period harmonic).  Of
+> the 23 stars with periods below 10 days, 21 of the TESS periods agree with
+> the Kepler rotation periods, and 2 are at the double-period harmonic.
 >
+> We've updated the text to reflect this point, which doesn't affect any of the
+> conclusions given that almost all the δ Lyr cluster stars have rotation
+> periods below below 10 days.
 
@@
 ----------
->
+> Done, thank you.
 
@@
 ----------
->
+> TODO ??? WHY ISN'T SIGMA OK?  THE POINT IS THAT IT'S 2-SIGMA...
 
@@
 ----------
+> Yes, option (ii) includes the transit -- we've updated this in the text for
+> clarity.
 >
+> Yes, option (i) is the 2nd or 4th order polynomial.  We've updated the text
+> to state this more directly as well.  Thank you!
 
@@
 ----------
->
+> Yes -- we've clarified this in the text.  Thank you!
 

### 2021-11-28 paper/ms.tex
[WIP] reviewer #2 staging

@@
 \title{
-  A 38 Million Year Old Mini-Neptune in the Kepler Field
+  A 38 Million Year Old Neptune-Sized Planet in the Kepler Field
 }
@@
   Kepler\,1627A is a G8V star previously known to host a
-  $3.8\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
+  $3.8\,R_\oplus$ \replaced{mini-Neptune}{planet} on a 7.2\,day orbit.  The star was
   observed by the Kepler space telescope because it is nearby
@@
   Kepler\,1627Ab the youngest planet with a precise age yet found by
-  the main Kepler mission.  The Kepler photometry shows two
+  the \replaced{main}{prime} Kepler mission.  The Kepler photometry shows two
   peculiarities: the average transit profile is asymmetric, and the
@@
 ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
-during the main Kepler mission (2009--2013), only four open clusters
+during the \replaced{main}{prime} Kepler mission (2009--2013), only four open clusters
 were known in the Kepler field,
@@
 5245) is a \cn\ member.  Given the previous statistical
-validation of the close-in mini-Neptune Kepler 1627b
+validation of the close-in \replaced{mini-Neptune}{Neptune-sized
+planet} Kepler 1627b
 \citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
@@
     cluster is outside our scope.  The location of the Sun is
-    ($\odot$) is shown.
+    ($\odot$) is shown \added{to clarify the direction along which
+    parallax uncertainties are expected to produce erroneous
+    clusters}.
 		\label{fig:XYZvtang}
@@
 distance as the presumed match \citep{gaia_collaboration_2021_edr3}.
-In Figure~\ref{fig:XYZvtang}, have shown galactic positions only for
-the stars with parallax signal-to-noise exceeding 20.  The reported
+In Figure~\ref{fig:XYZvtang}, \added{we }have shown galactic positions only for
+the stars with parallax signal-to-noise exceeding 20.  \added{We
+calculated the tangential velocities for each of these stars relative to \sn\
+($\Delta \mu^{*}$) by subtracting the observed proper motion from what the proper motion at the star's
+position would be if it were co-moving with \sn. }The reported
 cluster members (gray and black points) extend over a much larger
-volume than the cluster previously identified by
+volume \added{in both physical and kinematic space }than the cluster previously identified by
 \citet{stephenson_possible_1959} and later corroborated by
@@
 stars might comprise a {\it bona fide} cluster of identically-aged
-stars, they could also be heavily contaminated by field stars.  We
+stars, they could also be heavily contaminated by field stars.
+\added{One reason to suspect this is that the spread in tangential
+velocities exceeds typical limits for kinematic coherence
+\citep{Meingast2021}}.
+We
 therefore considered stars only in the immediate kinematic and spatial
-vicinity of \sn\ as candidate cluster members.  We performed the
-selection cuts manually, by drawing lassos with the interactive
+vicinity of \sn\ as candidate cluster members.  We performed
+\replaced{the}{this}
+selection cut\deleted{s} manually, by drawing lassos with the interactive
 \texttt{glue} visualization tool \citep{beaumont_2014_13866} in the
@@
 We then cleaned the candidate TESS rotation period measurements
-through a combination of automated and manual steps.  \deleted{As a matter of
-scope, we restricted our attention to the \nkinwithtess\ stars
-discussed in Section~\ref{sec:kinematicselection} in the spatial and
-kinematic proximity of Kepler\,1627.}  
-Kepler rotation periods were derived by
-\citet{mcquillan_rotation_2014} for 28 of \replaced{these }{our \nkinwithtess\ }stars; for these cases, we
-simply adopted the Kepler rotation period.
+through a combination of automated and manual steps.  \deleted{As a
+matter of scope, we restricted our attention to the \nkinwithtess\
+stars discussed in Section~\ref{sec:kinematicselection} in the spatial
+and kinematic proximity of Kepler\,1627.}  \deleted{Kepler rotation
+periods were derived by \citet{mcquillan_rotation_2014} for 28 of
+\replaced{these }{our \nkinwithtess\ }stars; for these cases, we
+simply adopted the Kepler rotation period.} \added{First, to validate
+the TESS rotation periods, we compared against 28 stars
+from \citet{mcquillan_rotation_2014} that were also observed by Kepler.  Of the 23 stars with Kepler
+periods below 10 days, 21 of the TESS periods agreed with the Kepler
+rotation periods; the other 2 were measured at the double-period
+harmonic.  Of the remaining 5 stars with Kepler rotation periods above
+10 days, none were correctly recovered by TESS (3 were near the
+half-period harmonic).  We therefore adopted a general stance of
+caution for TESS rotation periods exceeding 10 days, and used the
+Kepler rotation periods for the stars for which they were available.}
 For the remaining stars with only TESS data, we
@@
 given in Table~\ref{tab:posterior}, are broadly consistent with a
-mini-Neptune sized planet ($3.82\pm0.16\,R_\oplus$) on a close-in
+\replaced{mini-Neptune }{Neptune-}sized planet ($3.82\pm0.16\,R_\oplus$) on a close-in
 circular\footnote{ Our transit fitting yields $e<0.48$ at 2-$\sigma$;
@@
 $\pm4.5$\,hr of each transit, and fitted each window with both {\it
-i)} a local polynomial baseline plus the transit, and {\it ii)} a local linear
-trend.  We considered the results both for a second and fourth-order
-time-dependence in the local baseline.  We let the mid-time of each
+i)} a local \added{second or fourth-order }polynomial baseline plus the transit, and {\it ii)} a local linear
+trend\added{ plus the transit}.  \deleted{We considered the results both for a second and fourth-order
+time-dependence in the local baseline.  }We let the mid-time of each
 transit float, and then calculated the residual between the measured
@@
 Possible interpretations are discussed below. Given the
-statistical significance, this analysis should be interpreted as
+\added{lack of }statistical significance, this analysis should be interpreted as
 suggestive at best.
@@
 If the \pn\ transit signal is created by a genuine planet,  then to
-our knowledge it would be the youngest planet yet found by the main
+our knowledge it would be the youngest planet yet found by the
+\replaced{main}{prime}
 Kepler mission.\footnote{The re-purposed K2 mission however has found
@@
 \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.  
+\added{This of course depends on the mass of \pn; an alternative
+explanation for its large size could simply be that it is more massive
+than the typical Kepler mini-Neptune, or that it has an abnormally
+large envelope to core mass ratio.}
 

### 2021-11-28 drivers/plot_kepler_vs_tess_prot.py

@@ -1,0 +1,61 @@
+import os
+import numpy as np, matplotlib.pyplot as plt, pandas as pd
+from rudolf.paths import DATADIR, RESULTSDIR
+from rudolf.helpers import (
+    get_autorotation_dataframe, get_clean_gaia_photometric_sources
+)
+
+# GET DATA
+auto_df, base_df = get_autorotation_dataframe(
+    'deltaLyrCluster', cleaning='curtiscleaning'
+)
+
+csvpath = os.path.join(DATADIR, 'rotation',
+                       'Theia73-Prot_Results-Prelim-v2.csv')
+curtis_df = pd.read_csv(csvpath)
+
+# kinematic bit
+csvpath = os.path.join(
+    RESULTSDIR, 'tables', 'deltalyr_kc19_cleansubset_withreddening.csv'
+)
+kinematic_df = pd.read_csv(csvpath)
+N_kin = len(kinematic_df)
+
+mdf0 = curtis_df.merge(base_df, how='inner', left_on='Gaia_DR2_Source',
+                      right_on='dr2_source_id')
+
+mdf = kinematic_df.merge(mdf0, how='left', left_on='source_id',
+                         right_on='dr3_source_id')
+
+
+selcols = ['source_id', 'phot_bp_mean_mag_corr',
+           'phot_rp_mean_mag_corr', 'phot_g_mean_mag_corr']
+kinematic_df = kinematic_df[get_clean_gaia_photometric_sources(kinematic_df)]
+mdf1 = pd.DataFrame(kinematic_df[selcols]).merge(
+    auto_df, left_on='source_id', right_on='dr3_source_id',
+    how='inner'
+)
+
+# make plot
+
+plt.figure(figsize=(6,6))
+plt.scatter(mdf.Prot_Kepler_y, mdf.Prot_LS_Auto_y, s=1, c='k')
+_ = np.linspace(0,40,1000)
+plt.plot(_,_,ls='-',zorder=-1,color='gray',lw=0.5,
+         label='Prot TESS = Prot Kepler')
+plt.plot(_,0.5*_,ls=':',zorder=-1,color='gray',lw=0.5,
+         label='Prot TESS = 0.5*Prot Kepler')
+plt.plot(_,2*_,ls='--',zorder=-1,color='gray',lw=0.5,
+         label='Prot TESS = 2*Prot Kepler')
[…]

### 2021-11-30 paper_review/response_1.txt
[WIP] reviewer #2 staging

@@
 ----------
-> These completeness statistics were computed using KeplerPORTS (the
-> Burke & Catanzarite 2021 citation).  KeplerPORTS does correct for
-> the red noise in the light curves by using the CDPP over different
-> timescales -- see Section 3.3 of Burke, C. J. & Catanzarite, J.
-> 2017, Planet Detection Metrics: Per-Target Detection Contours for
-> Data Release 25, KSCI-19111-002.  Nonetheless, we do agree that this
-> is a worry -- perhaps a full injection-recovery analysis of the
-> light curve with a pipeline of our own might help verify whether
-> these detection limits are plausible.  This effort is outside our
+> These completeness statistics were computed using KeplerPORTS (the Burke &
+> Catanzarite 2021 citation).  KeplerPORTS corrects for red noise in the light
+> curves by using the CDPP over different timescales -- see Section 3.3 of
+> Burke, C. J. & Catanzarite, J.  2017, Planet Detection Metrics: Per-Target
+> Detection Contours for Data Release 25, KSCI-19111-002.  This correction is
+> performed using both a) the slope of the RMS CDPP versus transit duration for
+> for "long" transit durations (the average of 7.5, 9.0, 10.5, 12.0, 12.5, and
+> 15.0 hours) and b) the same slope for "short" transit durations (2.0, 2.5,
+> 3.0, 3.5, and 4.5 hours). We retrieved these values from the DR25 Kepler
+> occurrence rate products hosted on the NASA Exoplanet Archive.  For Kepler
+> 1627, they were CDPP_long = 0.25478 and CDPP_short=0.79505.  Per Figure 2 of
+> the KSCI-19111-002 memo, these do indeed indicate a noisy star -- no
+> surprises.
+>
+> We've added a sentence to the manuscript to clarify that we are talking about
+> per-target detection contours, which should correct for the impact of at
+> least the fast rotation period, inasmuch as it manifests in the slope of the
+> CDPP versus transit duration.  We have also modified the wording for clarity,
+> and have added a citation to Burke & Catanzarite's 2017 KSCI-19111-002 memo.
+>
+> To determine whether this CDPP-based detection efficiency correction is
+> *accurate* given the exact red-noise properties of Kepler-1627 (i.e., the
+> rotation and flares), a full injection-recovery analysis of the light curve
+> with a pipeline of our own would be necessary. This effort is outside our
 > scope, for now.
@@
 ----------
-> Yes -- we've clarified this in the text.  Thank you!
+> Yes -- we've added this to the text.  Thank you.
 
@@
 ----------
+> Kepler-1627B is ~0.16 arcseconds East of Kepler-1627A.  For the sake of
+> argument, assume it contributes 2% of the flux to the aperture (an
+> overestimate), and neglect other stars in the scene. 
 >
+> Taking the origin at Kepler-1627A, the center of light outside of transit is
+> then roughly (0.98*0 arcsec + 0.02*0.16 arcsec) = 0.0032 arcsec due East of
+> the primary.
+>
+> During transit, if the transit happens on the primary, the center of light is
+> roughly ( (0.98-0.002)*0 arcsec + (1-(0.98-0.002))*0.16 arcsec ) = 0.0035
+> arcsec due East of the primary -- i.e., it is closer to the secondary by 0.3
+> milliarcseconds, since the primary is getting fainter.
+>
+> If the transit instead happened on the secondary, the center of light would
+> be roughly (0.98*0 arcsec + (0.9*0.02)*0.16 arcsec) ~= 0.00288 arcsec due
+> East of the primary, where a 10% intrinsic eclipse depth on the secondary is
+> required to fit the observed transit depth.  In this scenario, the centroid
+> would shift closer to the *primary* by 0.3 milliarcseconds.
+>
+> Examining the NASA SPOC Data Validation Report (DVR), there are two available
+> photocentroid tests, the first based on the flux-weighted centroid analysis,
+> and the second based on the difference image centroid offset test.  The
+> latter test yields offset uncertainties of ~100 milliarcsec for Kepler-1627,
+> which is far above the relevant signal.  The former test yields centroid peak
+> RA and DEC offsets consistent with zero, with uncertainties of 0.9 and 1.1
+> milliarcsec respectively (page 19 of 552 in the DR25 DVR).  So, given that
+> the quoted centroid precision uncertainties are a factor of ~3 larger than
+> the expected shift, it is not possible to use the test in a way that
+> distinguishes the two host stars in a statistically convincing way.
 
@@
 ----------
+> Thank you for catching this -- this citation is indeed deprecated and
+> erroneously was not updated during the writing process.  The final adopted
+> calculation was based on the masses and radii from the Baraffe+2015 models,
+> rather than the MESA models (Choi+2016).  The Baraffe models are available
+> from
+> http://perso.ens-lyon.fr/isabelle.baraffe/BHAC15dir/BHAC15_tracks+structure.
 >
+> We believe that the companion's mass is roughly between 0.3 and 0.4 Msun, and
+> that the age of the system is ~38 +/- 6 Myr.  The Baraffe models give:
+>
+>    32 Myr 0.300 Msun, 0.481 Rsun, 3.80 g / cm3
+>    35 Myr 0.300 Msun, 0.465 Rsun, 4.21 g / cm3
+>    40 Myr 0.300 Msun, 0.443 Rsun, 4.86 g / cm3
+>    32 Myr 0.400 Msun, 0.555 Rsun, 3.30 g / cm3
+>    35 Myr 0.400 Msun, 0.537 Rsun, 3.64 g / cm3
+>    40 Myr 0.400 Msun, 0.514 Rsun, 4.15 g / cm3
+>    32 Myr 0.330 Msun, 0.506 Rsun, 3.60 g / cm3
+>    35 Myr 0.330 Msun, 0.489 Rsun, 3.98 g / cm3
+>    40 Myr 0.330 Msun, 0.467 Rsun, 4.58 g / cm3
+>
+> i.e., our default assumption was 40 Myr and 0.33 Msun.  The age is adopted
+> for the reasons discussed in the text.  The mass is based off
+> the observed deltaMag in 832nm and K' (3.14 +/- 0.15 and 2.37 +/- 0.02
+> respectively), which implies absolute magnitudes for the secondary of K'=XXX and
+> [832]=YYY.  Based on a set of absolute magnitude and color relations from a
+> submitted paper by A. Kraus, these gave ~=0.33 Msun.  (TODO: NEED TO BEEF
+> THIS UP!)
+>
+> The plausible range of companion densities is therefore 3.30 g/cc to
+> 4.86 g/cc.  We've updated the text to reflect this uncertainty, but
+> it doesn't change the overall point that our measured stellar
+> density (2.00+/-0.24 g/cc) is still a fair bit lower than the
+> expected range.
+
+TODO TODO REMOVE THIS
+> As a sanity check, these calibrations find that for a companion spectral type
+> of M2.5, one expects colors of G-K'=3.39 mag and G-[832] = 1.51 mag, though
+> we cannot test this since Gaia DR2/EDR3 does not resolve the secondary.
+>
+> For the primary spectral type of G8, one expects G-K'=1.58 mag and G-[832] =
+> 0.71 mag (i.e., K'=13.02-1.58=11.44, and [832]=G-0.71=12.31 for the primary).
+> So it holds together (bolstering the case for it being a bound companion) to
+> find DeltaK' = 2.37 mag and Delta[832] = 3.14 mag.
+
+G8V: M_G = 5.098
+M2.5V: M_G = 9.67 (ZAMS).
+-> delta G = 9.67 - 5.098 = 4.57, so G =13.02+4.57= 17.592 for secondary
+-> G-K'=3.39mag implies K' = 14.20 for secondary
+-> G-[832]=1.51mag implies [832] = 16.08 for secondary
+->
+-> expected delta K' = 2.76
+-> expected delta 832 = 3.78
+.. (both are observed a bit lower!  i.e., it's not red enough)
 
@@
 ----------
->
+> Visually, one can see that the width of the single-star sequence of δ Lyr
+> candidates (black points) in the upper-right panel of Figure 3 is <~0.2 mag.
+> So, if we saw a star with the same (BP-RP) color as Kepler 1627 that were
+> >~0.3 mag brighter in M_G, we would classify it as a likely photometric
+> binary.  This means we are able to exclude stars ~0.32* as bright as the
+> primary (in Gaia G-band). This corresponds to excluding bound companions more
+> massive than ~0.85 Msun (the primary's mass is ~0.95 Msun).  The SB2 tests do
+> much better if the velocity separation exceeds the CCF width (down to 1% as
+> bright as the primary;
+> https://ui.adsabs.harvard.edu/abs/2015AJ....149...18K/abstract).  For
+> brevity, we would prefer to not present this information in the text.
 
@@
 ----------
+> Thank you for this suggestion.  The main text now reads
 >
+> "We observed a transit of Kepler-1627 Ab on the night of 2021 Aug 7
+> spectroscopically with HIRES at the Keck-I telescope and photometrically in
+> $griz$ bands with MuSCAT3 at Haleakal\=a Observatory.  Details of the
+> observation sequence are discussed in Appendix C; Figure 7 shows the
+> results."
+>
+> The first paragraph of Appendix C has been significantly modified to include
+> the requested additional information, including the passbands, exposure
+> times, S/N, and instrument resolution.  Regarding observation times, "the
+> night of 2021 Aug 7" is sufficiently precise for the purposes of the text;
+> Tables 5 and 6 have the BJD-TDB timestamps for those who desire greater
+> precision.
 
@@
 ----------
->
+> Thank you for this point; we've updated the text to say "signal-to-noise"
+> rather than "precision".
 
@@
 ----------
+> Regarding the S-values, they are shown in the second panel of Figure 7, and
+> are also reported in Table 6.  The line profile changes shown in Figure 12
+> suggested that the majority of the RV variation observed was due to
+> starspot-induced variations.  Quantitative measurements of the H-alpha EWs
+> are not expected to change this picture, and so we opt to not include them.
 >
+> Regarding the NIR, UV, and X-ray detections, we searched Vizier within a
+> radius of 4 arcseconds.
+>
+> For the NIR and the possibility of a debris disk, Kepler-1627 is detected in
+> W1, W2, W3, and W4 (see kepler1627_vizier_photometry.png attached).  While W4
+> might be in excess of the usual Rayleigh-Jeans tail (and this could be
+> quantified by fitting synthetic stellar atmospheres), quantifying this excess
+> is best done at the population level, especially given the calibration
+> uncertainties that are present in W4 (e.g., Cotten & Song 2016, ApJS 225 15).
+> Since Kepler-1627 has not been identified as even a candidate debris disk
+> host in the previous literature, we opt to not speculate on this topic.
+>
+> For the UV, Olmedo+2015 reported a coincident source with NUVmag of 19.052
+> +/- 0.017 [AB mags], with a NUV S/N of 65 (2015, ApJ, 813, 100). Loyd et al
+> 2020 report a NUV flux of 4e-13 +/- 4e-15 mW/m^2.  We can use this detection
+> as a youth indicator (e.g., Figure 9 of Newton+2021, which Newton et al
+> attribute from the idea in Shkolnik+2011, ApJ 727 6. Since for Kepler 1627
+> 2MASS J = 11.690+/-0.021 (Cutri+2003), we have m_NUV-m_J = 7.36.  Given the
+> (BP-RP)0 of ~0.95... FIXME TODO ARE YOU GOING TO MAKE A WHOLE NEW PLOT FOR
+> THIS DAMN THING?  MIGHT BE WORTHWHILE... QUICK VETTING CHECK FOR ANY G AND K
+> TYPE STARS IN THE CLUSTER...
+>
+> For the X-ray (i.e., ROSAT), there were no reported detections in Vizier.
 
@@
 ----------
+> The first reviewer had a similar comment -- our response is the same:
 >
+> These completeness statistics were computed using KeplerPORTS (the Burke &
+> Catanzarite 2021 citation).  KeplerPORTS corrects for red noise in the light
+> curves by using the CDPP over different timescales -- see Section 3.3 of
+> Burke, C. J. & Catanzarite, J.  2017, Planet Detection Metrics: Per-Target
+> Detection Contours for Data Release 25, KSCI-19111-002.  This correction is
+> performed using both a) the slope of the RMS CDPP versus transit duration for
+> for "long" transit durations (the average of 7.5, 9.0, 10.5, 12.0, 12.5, and
+> 15.0 hours) and b) the same slope for "short" transit durations (2.0, 2.5,
+> 3.0, 3.5, and 4.5 hours). We retrieved these values from the DR25 Kepler
+> occurrence rate products hosted on the NASA Exoplanet Archive.  For Kepler
+> 1627, they were CDPP_long = 0.25478 and CDPP_short=0.79505.  Per Figure 2 of
+> the KSCI-19111-002 memo, these do indeed indicate a noisy star -- no
+> surprises.
+>
+> We've added a sentence to the manuscript to clarify that we are talking about
+> per-target detection contours, which should correct for the impact of at
+> least the fast rotation period, inasmuch as it manifests in the slope of the
+> CDPP versus transit duration.  We have also modified the wording for clarity,
+> and have added a citation to Burke & Catanzarite's 2017 KSCI-19111-002 memo.
+>
+> To determine whether this CDPP-slope detection efficiency correction is
+> *accurate* given the exact red-noise properties of Kepler-1627 (i.e., the
+> rotation and flares), a full injection-recovery analysis of the light curve
+> with a pipeline of our own would be necessary. This effort is outside our
+> scope, for now.
 

### 2021-11-30 paper/ms.tex
[WIP] reviewer #2 staging

@@
 periods were derived by \citet{mcquillan_rotation_2014} for 28 of
-\replaced{these }{our \nkinwithtess\ }stars; for these cases, we
+our \nkinwithtess\ stars; for these cases, we
 simply adopted the Kepler rotation period.} \added{First, to validate
@@
 		fit the spot-induced $\approx$\,$250$\ms\ trend is shown (see
-		Appendix~\ref{app:rm}).  Shaded bands show 2-$\sigma$ model
+		Appendix~\ref{app:groundobs}).  Shaded bands show 2-$\sigma$ model
 		uncertainties.  
@@
 theoretically expected density for Kepler\,1627B is $\approx 4.6\,{\rm
-g\,cm}^{-3}$ (Table~\ref{tab:posterior}; \citealt{choi_mesa_2016}).
+g\,cm}^{-3}$ \citep{baraffe_new_2015}.
 The transit duration is therefore too long to be explained by a star
@@
 also exists.  We observed a transit of \pn\ on the
-night of 2021 Aug 7 simultaneously with Keck/HIRES and MuSCAT3.  We
+night of 2021 Aug 7 \replaced{simultaneously with Keck/HIRES and
+MuSCAT3}{spectroscopically with HIRES at the Keck-I telescope and photometrically in
+$griz$ bands with MuSCAT3 at Haleakal\=a Observatory}.  \added{
+Details of the observation sequence are discussed in
+Appendix~\ref{app:groundobs}; Figure~\ref{fig:ground} shows the
+results.}\deleted{.  We
 scheduled the observations using the ephemeris of
-\citet{holczer_transit_2016}.  Although we did not detect the
+\citet{holczer_transit_2016}.}  Although we did not detect the
 Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
-show that the transit is achromatic (Figure~\ref{fig:ground}).
+show that the transit is achromatic\deleted{
+  (Figure~\ref{fig:ground})}.
 Quantitatively, when we fitted the MuSCAT3
@@
 bandpass, we found {\it griz} depths consistent with the Kepler depth
-at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  Conditioned on the
+at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  \added{The
+achromatic transits strongly favor Kepler\ 1627A as the transit host,
+since Kepler\ 1627B is much redder.}  Conditioned on the
 ephemeris and transit depth from the Kepler data, the MuSCAT3 observations also
@@
 For our RM analysis,
-the details are discussed in Appendix~\ref{app:rm}.  While the
+the details are discussed in Appendix~\ref{app:groundobs}.  While the
 velocities are marginally more consistent with a prograde or polar
@@
 planets, and opens multiple avenues for further study.  Observations
-of spectroscopic transits at greater precision should yield a
+of spectroscopic transits at greater
+\replaced{precision}{signal-to-noise} should yield a
 measurement of the stellar obliquity, which would confirm or refute
@@
 This geometry would be necessary in order to
-explain how the starspot crossings could add up coherently.
+explain how the starspot crossings could add up coherently\added{,
+given that the planetary orbital period (7.203\,days) and the stellar
+rotation period (2.642\,days) are not a rational combination}.
 Other possibilities including
@@
 \citep{Mann_K2_33b_2016,David_et_al_2017}, DS~Tuc
-\citep{benatti_possibly_2019,newton_tess_2019} , HIP~67522
+\citep{benatti_possibly_2019,newton_tess_2019}, HIP~67522
 \citep{rizzuto_tess_2020}, TOI~837 \citep{bouma_cluster_2020}, the
@@
 photometric variability at very young ages \citep[{\it
-e.g.},][]{zhou_2021_tois}.  However, smaller planets could have been
-detected: the Kepler pipeline's median completeness extended to
+e.g.},][]{zhou_2021_tois}.  \replaced{However,}{However, it seems that} smaller planets could have been
+detected: \added{based on the per-target detection contours, }the Kepler pipeline's median completeness extended to
 $1.6\,R_\oplus$ at 10~day orbital periods, and $3.3\,R_\oplus$ at
-100~days \citep{2021ascl.soft07027B}.  The large size of \pn\ relative
+100~days
+\citep{2021ascl.soft07027B}.
+\added{These limits account for the spot-induced variability in the
+system through
+a correction based on the Combined Differential Photometric Precision
+(CDPP) in the \sn\ light curve over the relevant transit timescales
+\citep{2017ksci.rept...19B}.} The large size of \pn\ relative
 to most Kepler mini-Neptunes might therefore support a picture
-in which the typical $5\,M_\oplus$ to 10$\,M_\oplus$ mini-Neptune
+in which the typical $5\,M_\oplus$ \deleted{to 10$\,M_\oplus$} mini-Neptune
 \citep{wu_mass_2019} loses a significant fraction of its primordial
@@
 \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.  
-\added{This of course depends on the mass of \pn; an alternative
-explanation for its large size could simply be that it is more massive
-than the typical Kepler mini-Neptune, or that it has an abnormally
+\added{This of course would need to be confirmed through a measurement
+of the planetary mass; alternative explanations for its large size
+could be that it is abnormally massive, or that it has an abnormally
 large envelope to core mass ratio.}
@@
 \section{Spectroscopy and Photometry During the Transit of 2021 Aug 7}
-\label{app:rm}
+\label{app:groundobs}
 
@@
 
-We monitored \sn\ with Keck/HIRES before, during, and after transit on
-the night of 2021 Aug 7.  We used the iodine cell for wavelength
-calibration, and extracted the 1-D spectra using the
-standard California Planet Survey pipeline \citep{howard_cps_2010}.
-The airmass ranged between 1.1 and 2.2 from the start through the end
-of observations; the seeing ranged from $1\farcs1$ at the beginning to
-$1\farcs5$ at the end.  We also simultaneously observed across {\it
-griz} bands using MuSCAT3 at Haleakal\=a Observatory on Maui, HI.
-Performing aperture photometry on the latter image stack yielded the
-data given in Table~\ref{tab:phot}.
+\added{We used the ephemeris of \citet{holczer_transit_2016} to
+observe a transit of \pn\ on the night of 2021 Aug 7 both
+spectroscopically and photometrically.  We used the HIRES echelle
+spectrograph at the Keck-I telescope and MuSCAT3 at Haleakal\=a
+Observatory on Maui, HI \citep{Narita_2020}.} \deleted{ We monitored
+\sn\ with Keck/HIRES before, during, and after transit on the night of
+2021 Aug 7.} \replaced{We used the iodine cell for wavelength
+calibration}{For the HIRES wavelength calibration, we used the iodine
+cell}, and extracted the 1-D spectra using the standard California
+Planet Survey pipeline \citep{howard_cps_2010}.  \added{Given the
+faintness of the target ($V=13.1$), we observed using the C2 decker,
+which yielded an instrument resolution of $\approx$50{,}000.} The
+airmass ranged between 1.1 and 2.2 from the start through the end of
+observations; the seeing ranged from $1\farcs1$ at the beginning to
+$1\farcs5$ at the end.  \added{The HIRES exposure time was set at
+$\approx$15 minutes in order to resolve the 2.8 hour transit event,
+which yielded a S/N per resolution element of $\approx 75$ (low for
+precision radial velocity standards).} \replaced{We also
+simultaneously observed across {\it griz} bands using MuSCAT3 at
+Haleakal\=a Observatory on Maui, HI.}{For the MuSCAT3 observations, we
+observed simultaneously across the {\it griz} bands.} \added{The
+exposure times in each bandpass ranged between 23 and 46 seconds, and
+were chosen in order to yield a S/N in the peak pixel that exceeded
+130 while also preventing saturation.} Performing aperture photometry
+on the latter image stack yielded the data given in
+Table~\ref{tab:phot}.
 

### 2021-12-01 paper_review/response_1.txt
reviewer #1 complete; a few reviewer #2 items todo

--- 2021-11-24 paper_review/response_1.txt
@@
 Below, we have reproduced the relevant portions of the referee's
-report, alongside our responses. A list of changes is appended to
-the manuscript using the trackchanges AASTeX macros.
+report, alongside our responses. A list of changes is appended to the
+manuscript using the trackchanges AASTeX macros.
 
@@
 ----------
-Line 57 (evaporation): Consider adding citation to Lopez et al. (2012) or
-similar
+Line 57 (evaporation): Consider adding citation to Lopez et al.
+(2012) or similar
 
@@
 ----------
-Line 78 (Kepler ages): Consider adding citation to Walkowicz & Basri (2013)
+Line 78 (Kepler ages): Consider adding citation to Walkowicz & Basri
+(2013)
 
@@
 ----------
-Line 93: Clarify origin of name delta Lyr, given there are two pre-existing
-names, and how they overlap. It might also be worth mentioning Kerr's Cep-Her
-association as in prep just to save future readers the headache of
-cross-identifying them, either here or in Sec 2.1
+Line 93: Clarify origin of name delta Lyr, given there are two
+pre-existing names, and how they overlap. It might also be worth
+mentioning Kerr's Cep-Her association as in prep just to save future
+readers the headache of cross-identifying them, either here or in Sec
+2.1
 
@@
 ----------
-> delta Lyr cluster was the first name suggested by Stephenson (1958),
-> and it is the default name listed in SIMBAD.  It was also the name
-> used in the title of Eggen's 1967 paper.
+> delta Lyr cluster was the first name suggested by Stephenson
+> (1958), and it is the default name listed in SIMBAD.  It was also
+> the name used in the title of Eggen's 1967 paper.
 >
-> The cluster was called Stephenson-1 by people other than Stephenson,
-> including by Kounkel & Covey 2019 -- their Theia catalog identifiers
-> are simply given to any cluster in their catalog, in addition to any 
-> previous names that existed.
+> The cluster was called Stephenson-1 by people other than
+> Stephenson, including by Kounkel & Covey 2019.  The Kounkel & Covey
+> 2019 "Theia" catalog identifiers are simply given to any cluster in
+> their catalog, in addition to any previous names that existed.
 >
@@
 > instance, see the 30 stars shown in Figure 1 of Eggen 1968.  The
-> broader "complex" shown in Figure 1 of our paper, which to our
+> broader "complex" shown in Figure 1 of our manuscript was to our
 > knowledge first reported to be a set of kinematically associated
-> stars by Kounkel & Covey (inasmuch as an "HDBScan group"
-> implies kinematic association).
+> stars by Kounkel & Covey -- inasmuch as an "HDBScan group" implies
+> kinematic association.
 >
@@
 > potential importance of the complex given its proximity and age, we
-> really would prefer to leave this for the in prep R. Kerr work.  And
-> similarly, the introduction of the term "Cep-Her complex" is
-> probably best left for that future work.
+> really would prefer to leave this for the in prep R. Kerr work.
+> And similarly, the introduction of the term "Cep-Her complex" is
+> probably best left for that future work, where it can be mentioned
+> that the "del Lyr cluster" is a subset of this complex.
 
@@
 ----------
-Line 165: Consider mentioning the axes on which data was cleaned (e.g. color
-precision)
+Line 165: Consider mentioning the axes on which data was cleaned
+(e.g. color precision)
 
@@
 ----------
-> Done - thank you.  We have added the following sentence: "This also
-> involved cuts on the photometric signal to noise ratio, the number
-> of visibility periods used, the astrometric $\chi^2$ of the
-> single-source solution, and the $G_{\rm BP}-G_{\rm RP}$ color excess
-> factor."
+> Done - thank you.  We have added the sentence: "This also involved
+> cuts on the photometric signal to noise ratio, the number of
+> visibility periods used, the astrometric $\chi^2$ of the
+> single-source solution, and the $G_{\rm BP}-G_{\rm RP}$ color
+> excess factor."
 
@@
 ----------
-Line 179 (reddening): What does "reasonably well" mean? Something like 1-2
-sigma?
+Line 179 (reddening): What does "reasonably well" mean? Something
+like 1-2 sigma?
 
@@
 ----------
-> Thank you for catching the weasel words.  The revised text now reads
-> as follows: "These values \replaced{agree reasonably well with}{are
-> within a factor of two of} previously reported values
-> \replaced{from}{in} the literature \citep[{\it
-> e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}\added{,
-> and are all small enough that the choice of whether to use them {\it
-> vs.} other extinction estimates does not affect our primary
-> conclusions}."
+> Thank you for catching the imprecise wording.  The revised text now
+> reads as follows: "These values are within a factor of two of
+> previously reported values in the literature (citations), and are
+> all small enough that the choice of whether to use them {\it vs.}
+> other extinction estimates does not affect our primary
+> conclusions."
 >
-> More quantitatively, from the Lallement maps we report
-> $E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$ for IC
-> 2602, the Pleiades, and the delta Lyr cluster.  The relevant
-> literature reports:
+> More quantitatively, from the Lallement maps we find
+> $E(B-V)=\{0.084\pm0.041, 0.020\pm0.003, 0.045\pm0.008,
+> 0.032\pm0.006\}$ for UCL, IC 2602, the Pleiades, and the delta Lyr
+> cluster respectively.  The relevant literature reports:
 >
@@
 >    IC2602 0.031
->   
->    Randich+2018 E(B-V):
->    IC2602 0.032 - 0.092
 >   
@@
 >
+>    Randich+2018 E(B-V):
+>    IC2602 0.032 - 0.092
+>
+>    Pecaut & Mamajek 2016 Table 7 UCL members E(B-V)
+>    UCL 0.059
+>
 > hence our assessment of "it's within a factor of two, and doesn't
@@
 ----------
-Fig 3: It may be helpful to include one of the even younger clusters on the CMD
-(i.e. to bracket delta Lyra), though perhaps it makes the plot too busy.
+Fig 3: It may be helpful to include one of the even younger clusters
+on the CMD (i.e. to bracket delta Lyra), though perhaps it makes the
+plot too busy.
 
@@
 ----------
-> Please see camd_with_UCL.png and camd_with_UCL_no_redcorr.png
-> attached.  This approach indeed makes the plot a bit too busy, in
-> our opinion.
+> Please see camd_with_UCL.png attached.  This does make the plot a
+> bit too busy, in our opinion.
 
@@
 ----------
-Page 3: Consider re-organizing. Based on the 1st paragraph, I thought the
-quantitative analysis would only deal with Pleiades and IC 2602 and one might
-infer reddening was only calculated for these clusters (whereas I assume the
-latter are also corrected for reddening prior to the analysis described). My
-suggestion would be to introduce all the clusters first. It would also be
-helpful to list the age of IC 2602 at some point in the text.
+Page 3: Consider re-organizing. Based on the 1st paragraph, I thought
+the quantitative analysis would only deal with Pleiades and IC 2602
+and one might infer reddening was only calculated for these clusters
+(whereas I assume the latter are also corrected for reddening prior
+to the analysis described). My suggestion would be to introduce all
+the clusters first. It would also be helpful to list the age of IC
+2602 at some point in the text.
 
@@
 ----------
-> Regarding the main re-organization point, FIXME TODO REORGANIZE?
+> Thank you for this suggestion.
 >
-> Regarding the last point, the final paragraph of Section 2.2.1
-> discusses the age of IC 2602 in the text.  There's a fairly wide
-> spread of reported literature ages between 30 and 46 Myr -- we
-> ultimately adopted the half-way point between the isochrone and LDB
-> ages.
+> The first paragraph of Section 2.2.1 has been reorganized to now
+> introduce all five of the relevant clusters: UCL, IC 2602, del Lyr
+> cluster, mu-Tau, and the Pleaides.  The second paragraph has been
+> modified to account for this hcnage.
+>
+> The extinction for UCL is now also listed in the text.  So is the
+> adopted age of IC 2602.
 
@@
 Line 202 (binaries): Add RUWE citation(s). Nice approach to remove
-radial_velocity_error based on each cluster's distribution. It should be noted
-somewhere in this discussion that the target itself does *not* meet the RUWE
-limit and so wouldn't appear on the plot by default.
+radial_velocity_error based on each cluster's distribution. It should
+be noted somewhere in this discussion that the target itself does
+*not* meet the RUWE limit and so wouldn't appear on the plot by
+default.
 
@@
 ----------
-Line 209: I don't follow starting "assembled a piecewise grid of hybrid
-isochrones..." What does "hybrid" mean here? Once the splines for each cluster
-are defined, how does one interpolate between them? Are the ages then
-determined for each star and combined?
+Line 209: I don't follow starting "assembled a piecewise grid of
+hybrid isochrones..." What does "hybrid" mean here? Once the splines
+for each cluster are defined, how does one interpolate between them?
+Are the ages then determined for each star and combined?
 
@@
 > The main assumption of this approach is that between the reference
-> times set by the adopted clusters, the isochrones evolve linearly in
-> time through color-absolute magnitude space.  If we were to attempt
-> this linear interpolation all the way from UCL to the Pleiades
-> without putting any intermediate point, the validity of this
-> assumption would be crucial to test.  Since we include IC 2602,
-> and it overlaps almost directly on the CAMD with the δ Lyr cluster,
-> the only useful quantity that this method really gives us is a
-> formal uncertainty on the age.  (Since the age the method returns
-> will be very close to whatever age we adopt for IC 2602).
+> times set by the adopted clusters, the isochrones evolve linearly
+> in time through color-absolute magnitude space.  If we were to
+> attempt this linear interpolation all the way from UCL to the
+> Pleiades without putting any intermediate point, the validity of
+> this assumption would be crucial to test.  Since we include IC
+> 2602, and it overlaps almost directly on the CAMD with the δ Lyr
+> cluster, the only useful quantity that this method really gives us
+> is a formal uncertainty on the age.  (Since the age the method
+> returns will be very close to whatever age we adopt for IC 2602).
 >
@@
 > interpolated splines spaced every 4.4 Myr between UCL and IC2602,
-> and every 14.8 Myr betwen IC2602 and Pleiades.  While we do interpolate
-> over a much finer grid, this is not shown to improve the clarity of
-> the figure.  The age of δ Lyr is determined by using all the stars
-> in the displayed color range, through the referenced Equation 7 of
-> Gange+2020, which is a Gaussian likelihood that treats the
-> interpolated isochrone as the "model" and the δ Lyr cluster's
-> isochrone as the "data".  The statistical uncertainty on the age is
-> then the standard deviation of a Gaussian fit to the posterior
-> probability (in log-age; it becomes asymmetric when transformed to
-> linear age).  We've added a few sentences to the manuscript to
-> clarify this latter point.
+> and every 14.8 Myr betwen IC2602 and Pleiades.  While we do
+> interpolate over a much finer grid, this is not shown to improve
+> the clarity of the figure.  The age of δ Lyr is determined by using
+> all the stars in the displayed color range, through the referenced
+> Equation 7 of Gange+2020, which is a Gaussian likelihood that
+> treats the interpolated isochrone as the "model" and the δ Lyr
+> cluster's isochrone as the "data".  The statistical uncertainty on
+> the age is then the standard deviation of a Gaussian fit to the
+> posterior probability (in log-age; it becomes asymmetric when
+> transformed to linear age).  We've added a few sentences to the
+> manuscript to clarify this latter point.
 
@@
 ----------
-Line 230 (age): Consider adopting a systematic error (perhaps corresponding to
-the range noted for the extremes of the ages) in addition to the random
-(assumed to be the quoted +6/-5 error). Relatedly, how are the quoted errors
-derived?
+Line 230 (age): Consider adopting a systematic error (perhaps
+corresponding to the range noted for the extremes of the ages) in
+addition to the random (assumed to be the quoted +6/-5 error).
+Relatedly, how are the quoted errors derived?
 
@@
 > standard deviation of the logarithmic age posterior since this
-> distribution is roughly Gaussian.  The worry with quoting a
+> distribution is roughly Gaussian (see the attached
+> del_lyr_cluster_age_posterior.png).  The worry with quoting a
 > systematic error, e.g., "The age of the δ Lyr cluster is
@@
 > uncertainties are not Gaussian and are not based on any probability
-> distribution that we can easily assess -- they're estimated based on
-> the extremes of previous literature age measurements.  Quoting this
-> systematic range next to a number that has been derived based on a
-> more formal approach (and that has statistical meaning within that
-> approach) somehow misses the point.  Given the roughly comparable
-> amplitude of the two sources of uncertainty, does it matter?  Open
-> cluster ages are not H_0.
+> distribution that we can easily assess -- they're estimated based
+> on the extremes of previous literature age measurements.  Quoting
+> this systematic range next to a number that has been derived based
+> on a more formal approach (and that has statistical meaning within
+> that approach) somehow misses the point.  Given the comparable
+> amplitude of the two sources of uncertainty, does this matter?
+> Discussing the range of extremes in the text does communicate the
+> point that this is a source of systematic uncertainty.
 
@@
 ----------
-Line 235: May be worth mentioning the restriction to the 391 stars in the
-beginning of sec 2.2.2.
+Line 235: May be worth mentioning the restriction to the 391 stars in
+the beginning of sec 2.2.2.
 
@@
 Line 286: I thought I remember Curtis or Douglas citing binarity as a
-possibility for stars above to slow sequence as well, but I don't have a
-reference on-hand.
+possibility for stars above to slow sequence as well, but I don't
+have a reference on-hand.
 
@@
 ----------
-Section 3.2: The text mentions statistical and systematic uncertainties but I
-only see one set of errors listed in Table 1. Were these uncertainties combined
-or was only one adopted in the end?
+Section 3.2: The text mentions statistical and systematic
+uncertainties but I only see one set of errors listed in Table 1.
+Were these uncertainties combined or was only one adopted in the end?
 
@@
 ----------
-Section 3.2: Comment on whether the binary companion should effect the rotation
-measurement of the host star (either here, or in the rotation section).
+Section 3.2: Comment on whether the binary companion should effect
+the rotation measurement of the host star (either here, or in the
+rotation section).
 
@@
 > """
->The resulting photometry is dominated by a quasi-periodic starspot
->signal with a peak-to-peak amplitude that varies between 2\% and 8\%.
->\added{Given that the secondary companion's brightness in the Kepler
->band is 1\% to 2\% that of the primary, source confusion for the
->rotation signal is not expected to be an issue.}
+> The resulting photometry is dominated by a quasi-periodic starspot
+> signal with a peak-to-peak amplitude that varies between 2\% and 8\%.
+> \added{Given that the secondary companion's brightness in the Kepler
+> band is 1\% to 2\% that of the primary, source confusion for the
+> rotation signal is not expected to be an issue.}
   """
@@
 ----------
-One note about the transit analysis: The MuSCAT3 data was introduced as
-supporting evidence but the conclusion (that the transits appear achromatic)
-was never explicitly stated.
+One note about the transit analysis: The MuSCAT3 data was introduced
+as supporting evidence but the conclusion (that the transits appear
+achromatic) was never explicitly stated.
 
@@
 > Line 642 of the original manuscript notes "Although we did not
-> detect the RM anomaly, the multi-band MuSCAT3 light curves show that
-> the transit is achromatic (Figure 7)."
+> detect the RM anomaly, the multi-band MuSCAT3 light curves show
+> that the transit is achromatic (Figure 7)."  We have nonetheless
+> reworded this section to improve the clarity with which this point
+> is made.
 
@@
 ----------
-Line 725: The fast rotation period and flares could plausibly effect the
-completeness.
+Line 725: The fast rotation period and flares could plausibly effect
+the completeness.
 
@@
 > These completeness statistics were computed using KeplerPORTS (the
-> Burke & Catanzarite 2021 citation).  KeplerPORTS does correct for
-> the red noise in the light curves by using the CDPP over different
+> Burke & Catanzarite 2021 citation).  KeplerPORTS corrects for red
+> noise in the light curves by using the CDPP over different
 > timescales -- see Section 3.3 of Burke, C. J. & Catanzarite, J.
 > 2017, Planet Detection Metrics: Per-Target Detection Contours for
-> Data Release 25, KSCI-19111-002.  Nonetheless, we do agree that this
-> is a worry -- perhaps a full injection-recovery analysis of the
-> light curve with a pipeline of our own might help verify whether
-> these detection limits are plausible.  This effort is outside our
-> scope, for now.
+> Data Release 25, KSCI-19111-002.  This correction is performed
+> using both a) the slope of the RMS CDPP versus transit duration for
+> for "long" transit durations (the average of 7.5, 9.0, 10.5, 12.0,
+> 12.5, and 15.0 hours) and b) the same slope for "short" transit
+> durations (2.0, 2.5, 3.0, 3.5, and 4.5 hours). We retrieved these
+> values from the DR25 Kepler occurrence rate products hosted on the
+> NASA Exoplanet Archive.  For Kepler 1627, they were CDPP_long =
+> 0.25478 and CDPP_short=0.79505.  Per Figure 2 of the KSCI-19111-002
+> memo, these do indeed indicate a noisy star -- no surprises.
+>
+> We've added a sentence to the manuscript to clarify that we are
+> talking about per-target detection contours, which should correct
+> for the impact of at least the fast rotation period, inasmuch as it
+> manifests in the slope of the CDPP versus transit duration.  We
+> have also modified the wording for clarity, and have added a
+> citation to Burke & Catanzarite's 2017 KSCI-19111-002 memo.
+>
+> To determine whether this CDPP-based detection efficiency
+> correction is accurate given the exact red-noise properties of
+> Kepler-1627 (i.e., the rotation and flares), a full
+> injection-recovery analysis of the light curve with a pipeline of
+> our own would be necessary. This effort is outside our scope, for
+> now.
 
@@
 ----------
-Line 729: Considering Owen 2020, contraction could also contribute to some
-extent to the larger radii of small planets; might this be significant?
+Line 729: Considering Owen 2020, contraction could also contribute to
+some extent to the larger radii of small planets; might this be
+significant?
 
@@
 ----------
-> TODO WHAT?
+> Thank you for this point -- we've added a sentence in the
+> discussion to mention this possibility.
 
@@
 ----------
-Appendix A paragraph beginning at line 1158: May be worth mentioning that delta
-Lyr may also be part of one of these heirarchical associations.
+Appendix A paragraph beginning at line 1158: May be worth mentioning
+that delta Lyr may also be part of one of these heirarchical
+associations.
 
@@
 ----------
-Figure 13 caption: Specify planetary orbital period since technically there is
-also a stellar orbital period involved.
+Figure 13 caption: Specify planetary orbital period since technically
+there is also a stellar orbital period involved.
 
@@
 Abstract
-line 24: Kepl-1627Ab cannot be a "mini-Neptune" if it is same size of Neptune.
-Replace with "Neptune-size" Yes, it falls within a distribution of objects in a
-period-radius diagram with an average that is smaller than Neptune, but
-claiming affiliation with smaller objects implies more than we know about this
-planet, e.g. its mass.
+line 24: Kepl-1627Ab cannot be a "mini-Neptune" if it is same size of
+Neptune.  Replace with "Neptune-size" Yes, it falls within a
+distribution of objects in a period-radius diagram with an average
+that is smaller than Neptune, but claiming affiliation with smaller
+objects implies more than we know about this planet, e.g. its mass.
 
@@
 ----------
-> Point well-taken. 
-> R_p = 3.79 +/- 0.16 R_earth, so yes, this is the same size as
-> Neptune within 1-sigma.
+> Point well-taken.  R_p = 3.82 +/- 0.16 R_earth, so yes, Kepler-1627
+> Ab is the same size as Neptune within 1-sigma.  All instances in
+> the text that referred to Kepler-1627 Ab as a mini-Neptune have
+> been modified to call it a "Neptune-sized planet", including the
+> manuscript's title.
 
@@
 ----------
-line 25: Kepler target stars were chosen for their suitability for transiting
-detection of planets in general, and for planets that are Earth-analogs in
-paraticular. That involves radius, brightess. Distance and spectral types were
-not direct selection criteria; this was before Gaia.
+line 25: Kepler target stars were chosen for their suitability for
+transiting detection of planets in general, and for planets that are
+Earth-analogs in paraticular. That involves radius, brightess.
+Distance and spectral types were not direct selection criteria; this
+was before Gaia.
 
@@
 ----------
->
+> Indeed.  Batalha+2010 (ApJL) discuss the selection function, which
+> favored stars that were both small and bright since these stars
+> enable the detection of Earth-size planets.  M dwarfs thus were
+> generally excluded unless they happened to be very close.  F, G,
+> and K stars were favored, particularly when they were nearby.  The
+> sentence in L25 currently states that "The star was observed by the
+> Kepler space telescope because it is nearby (d=329pc) and it
+> resembles the Sun".  More directly, yes, the star was chosen
+> because it was thought that it might enable the detection of small
+> planets in the habitable zone.  However the underlying reason for
+> why the star enables such detections is based on the star's
+> brightness and assumed radius, which both factor into the transit
+> SNR.  Our claim therefore holds, albeit indirectly.  It also serves
+> the more important purpose of informing the reader what kind of
+> star this paper is about: a star that is at a distance of ~330
+> parsecs, and that resembles the Sun (except for its age!).
 
@@
 
-Line 47: (First paragraph on p 2) the time-scale of 1e8 is associated with one
-particular set of photoevaporation models; core-powered mass loss proposes
-timescales of Gyr(s) (see the Ginzburg paper that was cited). More recent
-photoevaporation models also propose 1 Gyr timescales. It would be good to
-revise/update this paragraph.
+Line 47: (First paragraph on p 2) the time-scale of 1e8 is associated
+with one particular set of photoevaporation models; core-powered mass
+loss proposes timescales of Gyr(s) (see the Ginzburg paper that was
+cited). More recent photoevaporation models also propose 1 Gyr
+timescales. It would be good to revise/update this paragraph.
 
@@
 ----------
+> The sentence corresponds to a claim that mini-Neptunes are
+> "expected to shrink in size by factors of several over their first
+> 10^8 years".   The Owen 2020 models are then cited because this is
+> their main claim: e.g., Figure 1 top panel of Owen 2020.  The range
+> in initial raddi from that study come from differences in
+> assumptions about whether or not the boil-off phase (Owen & Wu
+> 2016) happens, and differences in assumption about the envelope to
+> core mass ratio.
 >
+> Nonetheless, the point is well-taken.  We've added a sentence that
+> clarifies that the outflows are expected to persist for 1e8 to 1e9
+> years, and have added citations to Gupta & Schlichting 2020, Rogers
+> & Owen 2021, and King & Wheatley 2021.
 
@@
 
-Line 65: it is convention to italicize names of space missions like TESS and K2
+Line 65: it is convention to italicize names of space missions like
+TESS and K2
 
@@
 ----------
->
+> Yes, and we leave the imposition of this convention to the
+> copy-editors at the journal.
 
@@
 ----------
-Lines 142-144: perhaps rephrasing this would clarify that the authors did not
-use Kepler-1627 as a fiducial star about which to select cluster members, and
-that they restricted the analysis to one of several sub-clusters or clumps
-which contained Kepler-1627.
+Lines 142-144: perhaps rephrasing this would clarify that the authors
+did not use Kepler-1627 as a fiducial star about which to select
+cluster members, and that they restricted the analysis to one of
+several sub-clusters or clumps which contained Kepler-1627.
 
@@
 ----------
->
+> We have rephrased Lines 142-144 to clarify the meaning, which is
+> indeed what the reviewer has stated.
 
@@
 ----------
-Figure 1: the first three panels would be more effective if they were plotted
-at the same scale and with appropriately aligned axes. As such the upper right
-panels are too small to be useful. Consider leaving out the Sun (which is
-irrelevant) so as to have a more appropriate choice of scale to how the details
-of the cluster. The purpose of the fourth panel (tangential velocities) and the
-second line in the caption are unclear to me. Aren't the tangential velocities
-known from Gaia proper motions and parallaxes? Why calculate them with the
-assumption of common space motion? If that is done, what does it show, anyway?
+Figure 1: the first three panels would be more effective if they were
+plotted at the same scale and with appropriately aligned axes. As
+such the upper right panels are too small to be useful. Consider
+leaving out the Sun (which is irrelevant) so as to have a more
+appropriate choice of scale to how the details of the cluster. The
+purpose of the fourth panel (tangential velocities) and the second
+line in the caption are unclear to me. Aren't the tangential
+velocities known from Gaia proper motions and parallaxes? Why
+calculate them with the assumption of common space motion? If that is
+done, what does it show, anyway?
 
@@
 ----------
+> Thank you for these important points of feedback.
 >
+> The Sun has been included to clarify the direction along which
+> parallax uncertainties are expected to produce erroneous
+> clusterings (analogous to the "Fingers of God" effect in the
+> extragalactic literature).  This is relevant because the HDBScan
+> clustering method used by Kounkel & Covey in their 2019 paper
+> likely was susceptible to this systematic effect.  A sentence has
+> been added to the caption of Figure 1 to clarify this point.
+> 
+> The {X,Y}, {Z,X}, and {Z,Y} panels are plotted at different scales
+> in order to keep the aspect ratios roughly correct.  The reported
+> KC2019 cluster members span {ΔX, ΔY, ΔZ} of roughly {400, 600, 200}
+> parsecs,  which grows to ~{600, 600, 200} parsecs when including
+> the Sun.  The {Z,X} and {Z,Y} panels should therefore roughly have
+> aspect ratios of 1:3 in order for the parsec to not be stretched
+> too much in either dimension.  The top-down {X,Y} view is the most
+> informative dimension, which is why it is enlarged related to the
+> edge-on views (and its 1:1 aspect ratio is set by the same
+> consideration).
+>
+> Finally, the tangential velocities are important because they tell
+> us whether the structures are kinematically coherent.  Open
+> clusters typically have velocity dispersions of order ~1km/s.
+> Meingast et al. 2021 (A&A 645 84) for instance suggest tangential
+> velocities <2 km/s (from the mean cluster velocity) as a plausible
+> cutoff for selecting candidate cluster members.  One needs to
+> compute these as relative tangential velocities instead of just
+> looking at the proper motions because the proper motions are
+> changed by line-of-sight projection effects, especially when the
+> clusters appear stretched over tens of degrees on-sky.  See for
+> instance Figure 2 of Meingast+2021, or Figure 14 of Bouma+2021 (AJ,
+> 162, 197).  In our case, the tangential velocities indicate that
+> the spread in Kounkel & Covey 2019's reported cluster members is
+> ~10km/s, which is large enough that it does not make obvious
+> kinematic sense to consider all of these stars to be part of the
+> same structure.
+>
+> We've added a few sentences in the manuscript to clarify these
+> latter points.
 
@@
 ----------
-Figure 3 bottom: Can the IC 2602 and 2391 rotation periods be added to this
-plot? Lines 276-278 mention that these and cite Douglas et al. 2021 (although
-the reference does not seem to be complete). Isn't it interesting that the
-interlopers all fall on the Praesepe rotation sequence?
+Figure 3 bottom: Can the IC 2602 and 2391 rotation periods be added
+to this plot? Lines 276-278 mention that these and cite Douglas et
+al. 2021 (although the reference does not seem to be complete). Isn't
+it interesting that the interlopers all fall on the Praesepe rotation
+sequence?
 
@@
 ----------
+> Thank you for pointing out the missing link in the reference to
+> Douglas et al. 2021, which should have pointed to
+> https://ui.adsabs.harvard.edu/abs/2021tsc2.confE.157D/abstract, and
+> has been updated in the revision.  The rotation periods from this
+> work, which includes IC 2602, IC 2391, and a few other 30-50 Myr
+> clusters, are not yet public.  
 >
+> The interlopers falling on the Praesepe sequence is indeed a bit
+> odd, but it's really only ~3-5 stars. A much bigger set of
+> interlopers didn't have detected rotation periods at all! (See
+> final paragraph of Section 2.2.2).  At the other reviewer's
+> request, we've also added a sentence noting that some of these
+> stars might be unresolved F+K binaries.
 
@@
 ----------
-p 5 lines 258-261: rather than simply throw the TESS values out in favor of the
-Kepler ones, the authors should compare them to evaluate the accuracy and
-reliability of the TESS values.
+p 5 lines 258-261: rather than simply throw the TESS values out in
+favor of the Kepler ones, the authors should compare them to evaluate
+the accuracy and reliability of the TESS values.
 
@@
 ----------
+> Thank you for this point -- please see kepler_vs_tess_prot.png in
+> the attached, which shows the 28 referenced stars that overlap.  5
+> of the 28 stars have Kepler rotation periods above 10 days, of
+> which none were correctly recovered by TESS (3 are chosen at the
+> half-period harmonic).  Of the 23 stars with periods below 10 days,
+> 21 of the TESS periods agree with the Kepler rotation periods, and
+> 2 are at the double-period harmonic.
 >
+> We've updated the text to reflect this point, which doesn't affect
+> any of the conclusions given that almost all the δ Lyr cluster
+> stars have rotation periods below below 10 days.
 
@@
 
-page 7 line 391: report the RUWE value (yes its in the Table but don't make the
-reader work)
+page 7 line 391: report the RUWE value (yes its in the Table but
+don't make the reader work)
 
@@
 
-The significance of this slope is unclear and a sigma is not a good way to
-evaluate it. Perform a ratio of chi-squared or F-test to evaluate it vs. a flat
-line.
+The significance of this slope is unclear and a sigma is not a good
+way to evaluate it. Perform a ratio of chi-squared or F-test to
+evaluate it vs. a flat line.
 
@@
 ----------
-Lines 530-542: does option (ii) include a transit fit as well? I assume it does
-but it would be good to clarify. And is option (i) the 2nd and 4th order
-polynomial? Perhaps just state this?
+Lines 530-542: does option (ii) include a transit fit as well? I
+assume it does but it would be good to clarify. And is option (i) the
+2nd and 4th order polynomial? Perhaps just state this?
 
@@
 ----------
+> Yes, option (ii) includes the transit -- we've updated this in the
+> text for clarity.
 >
+> Yes, option (i) is the 2nd or 4th order polynomial.  We've updated
+> the text to state this more directly as well.  Thank you!
 
@@
 ----------
->
+> Yes -- we've added this to the text.  Thank you.
 
@@
 
-Analysis of the motion of the photocentroid has been widely used to help
-validate Kepler planets. Can this be used here to distinguish between the
-possible A and B host stars?
+Analysis of the motion of the photocentroid has been widely used to
+help validate Kepler planets. Can this be used here to distinguish
+between the possible A and B host stars?
 
@@
 ----------
+> Kepler-1627B is ~0.16 arcseconds East of Kepler-1627A.  For the
+> sake of argument, assume it contributes 2% of the flux to the
+> aperture (an overestimate), and neglect other stars in the scene. 
 >
+> Taking the origin at Kepler-1627A, the center of light outside of
+> transit is then roughly (0.98*0 arcsec + 0.02*0.16 arcsec) = 0.0032
+> arcsec due East of the primary.
+>
+> During transit, if the transit happens on the primary, the center
+> of light is roughly ( (0.98-0.002)*0 arcsec + (1-(0.98-0.002))*0.16
+> arcsec ) = 0.0035 arcsec due East of the primary -- i.e., it is
+> closer to the secondary by 0.3 milliarcseconds, since the primary
+> is getting fainter.
+>
+> If the transit instead happened on the secondary, the center of
+> light would be roughly (0.98*0 arcsec + (0.9*0.02)*0.16 arcsec) ~=
+> 0.00288 arcsec due East of the primary, where a 10% intrinsic
+> eclipse depth on the secondary is required to fit the observed
+> transit depth.  In this scenario, the centroid would shift closer
+> to the *primary* by 0.3 milliarcseconds.
+>
+> Examining the NASA SPOC Data Validation Report (DVR), there are two
+> available photocentroid tests, the first based on the flux-weighted
+> centroid analysis, and the second based on the difference image
+> centroid offset test.  The latter test yields offset uncertainties
+> of ~100 milliarcsec for Kepler-1627, which is far above the
+> relevant signal.  The former test yields centroid peak RA and DEC
+> offsets consistent with zero, with uncertainties of 0.9 and 1.1
+> milliarcsec respectively (page 19 of 552 in the DR25 DVR).  So,
+> given that the quoted centroid precision uncertainties are a factor
+> of ~3 larger than the expected shift, it is not possible to use the
+> test in a way that distinguishes the two host stars in a
+> statistically convincing way.
 
@@
 ----------
-Line 604-608: The statement that the density from the transit duration is
-inconsistent with that expected for the M dwarf "B" cites Table 2 and
-Choi+2016, but that Table does not appear to provide any information on "B",
-and the Choi+2016 reference is just for the Mesa models. A density of 2 g/cm3
-(or lower) is possible for an M dwarf on the pre-main sequence, depending on
-the mass, age and model used.
+Line 604-608: The statement that the density from the transit
+duration is inconsistent with that expected for the M dwarf "B" cites
+Table 2 and Choi+2016, but that Table does not appear to provide any
+information on "B", and the Choi+2016 reference is just for the Mesa
+models. A density of 2 g/cm3 (or lower) is possible for an M dwarf on
+the pre-main sequence, depending on the mass, age and model used.
 
@@
 ----------
+> Thank you for catching this -- this citation is indeed deprecated
+> and erroneously was not updated during the writing process.  The
+> final adopted calculation was based on the masses and radii from
+> the Baraffe+2015 models, rather than the MESA models (Choi+2016).
+> The Baraffe models are available from
+> http://perso.ens-lyon.fr/isabelle.baraffe/BHAC15dir/BHAC15_tracks+structure.
 >
+> We believe that the companion's mass is roughly between 0.3 and 0.4
+> Msun, and that the age of the system is ~38 +/- 6 Myr.  The Baraffe
+> models give:
+>
+>    32 Myr 0.300 Msun, 0.481 Rsun, 3.80 g / cm3
+>    35 Myr 0.300 Msun, 0.465 Rsun, 4.21 g / cm3
+>    40 Myr 0.300 Msun, 0.443 Rsun, 4.86 g / cm3
+>    32 Myr 0.400 Msun, 0.555 Rsun, 3.30 g / cm3
+>    35 Myr 0.400 Msun, 0.537 Rsun, 3.64 g / cm3
+>    40 Myr 0.400 Msun, 0.514 Rsun, 4.15 g / cm3
+>    32 Myr 0.330 Msun, 0.506 Rsun, 3.60 g / cm3
+>    35 Myr 0.330 Msun, 0.489 Rsun, 3.98 g / cm3
+>    40 Myr 0.330 Msun, 0.467 Rsun, 4.58 g / cm3
+>
+> i.e., our default assumption was 40 Myr and 0.33 Msun.  The age is
+> adopted for the reasons discussed in the text.  The mass is based
+> off the observed deltaMag in 832nm and K' (3.14 +/- 0.15 and 2.37
+> +/- 0.02 respectively), which implies absolute magnitudes for the
+> secondary of K'=XXX and [832]=YYY.  Based on a set of absolute
+> magnitude and color relations from a submitted paper by A. Kraus,
+> these gave ~=0.33 Msun.  (TODO: NEED TO BEEF THIS UP!)
+>
+> The plausible range of companion densities is therefore 3.30 g/cc
+> to 4.86 g/cc.  We've updated the text to reflect this uncertainty,
+> but it doesn't change the overall point that our measured stellar
+> density (2.00+/-0.24 g/cc) is still a fair bit lower than the
+> expected range.
+
+TODO TODO REMOVE THIS
+> As a sanity check, these calibrations find that for a companion
+> spectral type of M2.5, one expects colors of G-K'=3.39 mag and
+> G-[832] = 1.51 mag, though we cannot test this since Gaia DR2/EDR3
+> does not resolve the secondary.
+>
+> For the primary spectral type of G8, one expects G-K'=1.58 mag and
+> G-[832] = 0.71 mag (i.e., K'=13.02-1.58=11.44, and
+> [832]=G-0.71=12.31 for the primary).  So it holds together
+> (bolstering the case for it being a bound companion) to find
+> DeltaK' = 2.37 mag and Delta[832] = 3.14 mag.
+
+G8V: M_G = 5.098
+M2.5V: M_G = 9.67 (ZAMS).
+-> delta G = 9.67 - 5.098 = 4.57, so G =13.02+4.57= 17.592 for secondary
+-> G-K'=3.39mag implies K' = 14.20 for secondary
+-> G-[832]=1.51mag implies [832] = 16.08 for secondary
+->
+-> expected delta K' = 2.76
+-> expected delta 832 = 3.78
+.. (both are observed a bit lower!  i.e., it's not red enough)
 
@@
 ----------
-Lines 613-614: it would be good to quantify this, i.e. how faint an unresolved
-companion or background star must be based on the star's position in the CMD.
+Lines 613-614: it would be good to quantify this, i.e. how faint an
+unresolved companion or background star must be based on the star's
+position in the CMD.
 
@@
 ----------
->
+> Visually, one can see that the width of the single-star sequence of
+> δ Lyr candidates (black points) in the upper-right panel of Figure
+> 3 is <~0.2 mag.  So, if we saw a star with the same (BP-RP) color
+> as Kepler 1627 that were
+> >~0.3 mag brighter in M_G, we would classify it as a likely
+> >photometric
+> binary.  This means we are able to exclude stars ~0.32* as bright
+> as the primary (in Gaia G-band). This corresponds to excluding
+> bound companions more massive than ~0.85 Msun (the primary's mass
+> is ~0.95 Msun).  The SB2 tests do much better if the velocity
+> separation exceeds the CCF width (down to 1% as bright as the
+> primary;
+> https://ui.adsabs.harvard.edu/abs/2015AJ....149...18K/abstract).
+> For brevity, we would prefer to not present this information in the
+> text.
 
@@
 ----------
-Lines 635-638: Here, new observations appear for the first time but are not
-described. While the details are given in the appendix, and the outcome of the
-observations doesn't define this work, the authors need to give some basic
-description of the observations, e.g. passbands, SNR, resolution, observation
-times, etc. The authors should not assume the reader actually know what HIRES
-and MuSCAT are!
+Lines 635-638: Here, new observations appear for the first time but
+are not described. While the details are given in the appendix, and
+the outcome of the observations doesn't define this work, the authors
+need to give some basic description of the observations, e.g.
+passbands, SNR, resolution, observation times, etc. The authors
+should not assume the reader actually know what HIRES and MuSCAT are!
 
@@
 ----------
+> Thank you for this suggestion.  The main text now reads
 >
+> "We observed a transit of Kepler-1627 Ab on the night of 2021 Aug 7
+> spectroscopically with HIRES at the Keck-I telescope and
+> photometrically in $griz$ bands with MuSCAT3 at Haleakal\=a
+> Observatory.  Details of the observation sequence are discussed in
+> Appendix C; Figure 7 shows the results."
+>
+> The first paragraph of Appendix C has been significantly modified
+> to include the requested additional information, including the
+> passbands, exposure times, S/N, and instrument resolution.
+> Regarding observation times, "the night of 2021 Aug 7" is
+> sufficiently precise for the purposes of the text; Tables 5 and 6
+> have the BJD-TDB timestamps for those who desire greater precision.
 
@@
 
-line 670: assuming that SNR is not to blame, the problem is not precision, the
-problem is either measuremnt systematics or astrophysical noise.
+line 670: assuming that SNR is not to blame, the problem is not
+precision, the problem is either measuremnt systematics or
+astrophysical noise.
 
@@
 ----------
->
+> Thank you for this point; we've updated the text to say
+> "signal-to-noise" rather than "precision".
 
@@
 ----------
-line 678-683: The Ca II HK activity indicator "S" as well as the H-alpha
-equivalent width could be obtained from the HIRES spectrum. Have the authors
-checked fora X-ray or UV (Galex) detection of this star? The star is likely not
-detected but its worth a look. In the same vein, the authors may wish to
-comment on whether the star's infrared emission indicates or rule out a
-substantial debris disk -- the star is detected in WISE (and perhaps Spitzer?)
+line 678-683: The Ca II HK activity indicator "S" as well as the
+H-alpha equivalent width could be obtained from the HIRES spectrum.
+Have the authors checked fora X-ray or UV (Galex) detection of this
+star? The star is likely not detected but its worth a look. In the
+same vein, the authors may wish to comment on whether the star's
+infrared emission indicates or rule out a substantial debris disk --
+the star is detected in WISE (and perhaps Spitzer?)
 
@@
 ----------
+> Regarding the S-values, they are shown in the second panel of
+> Figure 7, and are also reported in Table 6.  The line profile
+> changes shown in Figure 12 suggested that the majority of the RV
+> variation observed was due to starspot-induced variations.
+> Quantitative measurements of the H-alpha EWs are not expected to
+> change this picture, and so we opt to not include them.
 >
+> Regarding the NIR, UV, and X-ray detections, we searched Vizier
+> within a radius of 4 arcseconds.
+>
+> For the X-ray (i.e., ROSAT), there were no reported detections in
+> Vizier.
+>
+> For the NIR and the possibility of a debris disk, Kepler-1627 is
+> detected in W1, W2, W3, and W4 (see
+> kepler1627_vizier_photometry.png attached).  While W4 might be in
+> excess of the usual Rayleigh-Jeans tail, and this could be
+> quantified by fitting synthetic stellar atmospheres, this type of
+> analysis is best done at the population level, particularly given
+> the calibration uncertainties that are present in W4 (e.g., Cotten
+> & Song 2016, ApJS 225 15).  Since Kepler-1627 has not been
+> identified as even a candidate debris disk host in the previous
+> literature that performs these population-level infrared excess
+> analyses, we opt to not speculate on this topic.
+>
+> For the UV, Olmedo+2015 reported a coincident source with NUVmag of
+> 19.052 +/- 0.017 [AB mags], with a NUV S/N of 65 (2015, ApJ, 813,
+> 100). Loyd et al 2020 similarly report a NUV flux of 4e-13 +/-
+> 4e-15 mW/m^2, though the latter value is based on the same
+> GALEX-CAUSE Kepler survey data presented by Olmedo+2015.  UV
+> detections can sometimes be used as a youth indicator (e.g., Figure
+> 9 of Newton+2021 AJ 161 65, which Newton et al attribute from the
+> idea in Shkolnik+2011, ApJ 727 6).  For Kepler-1627, we have 2MASS
+> J = 11.690+/-0.021, and therefore m_NUV-m_J = 7.36.  Given the Gaia
+> (BP-RP)0 of 0.977, comparing to Figure 9 of Newton+2021 we find
+> that Kepler-1627 overlaps with late-G type stars in Psc-Eri, the
+> Pleiades, and the Hyades.  In other words, the UV flux is not a
+> particularly strong diagnostic of the age for Kepler-1627 as an
+> individual star.  It may well be worth exploring the GALEX-CAUSE
+> Kepler survey data for the delta Lyr cluster generally, since the
+> later type K and M dwarfs are expected to be elevated with respect
+> to the Pleiades and Hyades sequences.  However, this goes beyond
+> the scope of the present work, and the combination of the cluster
+> membership, stellar rotation, stellar lithium content, and stellar
+> activity all point to an age consistent with that reported in our
+> work.
 
@@
 ----------
-lines 699-701 It could be mentioned here that the planet orbital period and
-stellar rotational period are not a rational combination.
+lines 699-701 It could be mentioned here that the planet orbital
+period and stellar rotational period are not a rational combination.
 
@@
 ----------
-line 723-726: The caveat that should be mentioned here is that this is for
-older, main sequence stars.
+line 723-726: The caveat that should be mentioned here is that this
+is for older, main sequence stars.
 
@@
 ----------
+> The first reviewer had a similar comment -- our response is the
+> same:
 >
+> These completeness statistics were computed using KeplerPORTS (the
+> Burke & Catanzarite 2021 citation).  KeplerPORTS corrects for red
+> noise in the light curves by using the CDPP over different
+> timescales -- see Section 3.3 of Burke, C. J. & Catanzarite, J.
+> 2017, Planet Detection Metrics: Per-Target Detection Contours for
+> Data Release 25, KSCI-19111-002.  This correction is performed
+> using both a) the slope of the RMS CDPP versus transit duration for
+> for "long" transit durations (the average of 7.5, 9.0, 10.5, 12.0,
+> 12.5, and 15.0 hours) and b) the same slope for "short" transit
+> durations (2.0, 2.5, 3.0, 3.5, and 4.5 hours). We retrieved these
+> values from the DR25 Kepler occurrence rate products hosted on the
+> NASA Exoplanet Archive.  For Kepler 1627, they were CDPP_long =
+> 0.25478 and CDPP_short=0.79505.  Per Figure 2 of the KSCI-19111-002
+> memo, these do indeed indicate a noisy star -- no surprises.
+>
+> We've added a sentence to the manuscript to clarify that we are
+> talking about per-target detection contours, which should correct
+> for the impact of at least the fast rotation period, inasmuch as it
+> manifests in the slope of the CDPP versus transit duration.  We
+> have also modified the wording for clarity, and have added a
+> citation to Burke & Catanzarite's 2017 KSCI-19111-002 memo.
+>
+> To determine whether this CDPP-slope detection efficiency
+> correction is *accurate* given the exact red-noise properties of
+> Kepler-1627 (i.e., the rotation and flares), a full
+> injection-recovery analysis of the light curve with a pipeline of
+> our own would be necessary. This effort is outside our scope, for
+> now.
 
@@
 
-* Updated binning interval on lower panels of Figure 5 from 20 to 15 minutes.
+* Updated binning interval on lower panels of Figure 5 from 20 to 15
+  minutes.
 

### 2021-12-01 paper/ms.tex
reviewer #1 complete; a few reviewer #2 items todo

@@
 %\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
-%\documentclass[12pt,modern,twocolumn,tighten,linenumbers]{aastex63}
-\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
+\documentclass[12pt,modern,twocolumn,tighten,linenumbers]{aastex63}
+%\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
@@
 irradiation and internal heat can also power gradual outflows
-which, if strong enough, can deplete or entirely strip the atmosphere \citep{lopez_how_2012,Owen_Wu_2013,ginzburg_corepowered_2018}.
-Discovering young planets, measuring their masses, and detecting their
+which, if strong enough, can deplete or entirely strip the atmosphere \citep{lopez_how_2012,Owen_Wu_2013,ginzburg_corepowered_2018}.\added{ The latter photoevaporative and core-powered
+	outflows are thought to persist for $\approx$$10^8$ to $\approx$$10^9$ years,
+	though the details depend on the planetary masses,
+	the irradiation environments, and the initial atmospheric
+	mass fractions \citep{owen_evaporation_2017,gupta_signatures_2020,2021MNRAS.503.1526R,2021MNRAS.501L..28K}.} Discovering young planets, measuring their masses, and detecting their
 atmospheric outflows are key steps toward testing this paradigm, which
@@
 distance as the presumed match \citep{gaia_collaboration_2021_edr3}.
-In Figure~\ref{fig:XYZvtang}, \added{we }have shown galactic positions only for
-the stars with parallax signal-to-noise exceeding 20.  \added{We
+\replaced{In Figure~\ref{fig:XYZvtang}, \added{we }have shown galactic positions only for
+the stars with parallax signal-to-noise exceeding 20. }{For plotting purposes, we focused only
+on the stars with parallax signal-to-noise exceeding 20.} \added{We
 calculated the tangential velocities for each of these stars relative to \sn\
-($\Delta \mu^{*}$) by subtracting the observed proper motion from what the proper motion at the star's
-position would be if it were co-moving with \sn. }The reported
+($\Delta \mu^{*}$) by subtracting the observed proper motion from what the proper motion at each star's
+position would be if it were co-moving with \sn. }
+
+\replaced{The}{Figure~\ref{fig:XYZvtang} shows that the} reported
 cluster members (gray and black points) extend over a much larger
@@
 
-We measured the isochrone age using an empirical approach.  The left
+We measured the isochrone age using an empirical approach.  The \added{upper }left
 panel of Figure~\ref{fig:age} shows the color-absolute magnitude
-diagram (CAMD) of candidate \cn\ members, IC\,2602, the Pleiades, and
+diagram (CAMD) of candidate \cn\ members, IC\,2602\added{ ($\approx 38$\,Myr)}, the Pleiades\added{ ($\approx 115$\,Myr)}, and
 the field. The stars from the Pleiades and IC\,2602 were adopted from
 \citet{CantatGaudin2018a}, and the field stars are from the Gaia EDR3
-Catalog of Nearby Stars \citep{gaia_gcns_2021}.  We cleaned these
+Catalog of Nearby Stars \citep{gaia_gcns_2021}.  
+\added{We also
+compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
+and the Upper-Centaurus-Lupus (UCL) component of the Sco OB2
+association ($\approx$$16$\,Myr; \citealt{pecaut_star_2016}).
+We adopted the UCL members from \citet{Damiani2019}.
+For visual clarity, the latter two clusters are not shown in Figure~\ref{fig:age}.
+}
+We cleaned \replaced{these}{the membership lists}
 following the data filtering criteria from \citet[][Appendix
@@
 designed to include genuine binaries while omitting instrumental
-artifacts.  We then corrected for extinction by querying the
+artifacts.
+
+\replaced{We then corrected for extinction by querying}{To correct for extinction, we queried} the
 3-dimensional maps of \citet{capitanio_threedimensional_2017} and
@@
 A_X/A_0$ computed by \citet{GaiaCollaboration2018} assuming that $A_0
-= 3.1 E(B-V)$.  For IC\,2602, the Pleiades, and the \cn, this
+= 3.1 E(B-V)$.  For \added{UCL, }IC\,2602, the Pleiades, and the \cn, this
 procedure yielded a respective mean and standard deviation for the
-reddening of $E(B-V)=\{0.020\pm0.003, 0.045\pm0.008, 0.032\pm0.006\}$.
+reddening of \replaced{$E(B-V)=\{0.020\pm0.003, 0.045\pm0.008,
+0.032\pm0.006\}$}{$E(B-V)=\{0.084\pm0.041, 0.020\pm0.003, 0.045\pm0.008,
+0.032\pm0.006\}$}.
 These values \replaced{agree reasonably well with}{are within a
 factor of two of} previously reported values
-\replaced{from}{in} the literature \citep[{\it
-e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}\added{,
+\replaced{from}{in} the literature \replaced{\citep[{\it
+e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}}{\citep{pecaut_star_2016,GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}}\added{,
 and are all small enough that the choice of whether to use them {\it vs.} other
@@
 Figure~\ref{fig:age} shows that the \cn\ and IC\,2602 overlap, and
-therefore are approximately the same age.  In our exploration, we also
+therefore are approximately the same age.  \replaced{In our exploration, we also
 compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
@@
 pre-main-sequence M dwarfs of the \cn\ were intermediate between the
-latter two clusters.  To turn this heuristic interpolation into an age
+latter two clusters.}{The
+pre-main-sequence M dwarfs of the \cn\ were also seen to be ``below''
+UCL and ``above'' the Pleiades.}  To turn this heuristic interpolation
+into \replaced{an age}{a quantitative age}
 measurement, we used the empirical method developed by
@@
 was then modeled as a piecewise linear combination of these reference
-clusters.  For our reference clusters, we adopted members of UCL,
+clusters.  For our reference clusters, \replaced{we adopted members of UCL,
 IC\,2602, and the Pleiades from \citet{Damiani2019} and
-\cite{CantatGaudin2018a} respectively.  We removed binaries by
+\cite{CantatGaudin2018a} respectively}{we used UCL, IC\,2602, and the
+Pleiades}.  We removed binaries by
 requiring $\texttt{RUWE}<1.3$, $\texttt{radial\_velocity\_error}$
@@
 of \deleted{hybrid }isochrones spanning the ages between UCL \replaced{to}{and} the Pleiades using
-Equation\replaced{s~6 and~7}{6} from
+Equation\replaced{s~6 and~7}{~6} from
 \citet{gagne_mutau_2020}.  
@@
 models over a grid of ages, metallicities, and reddenings, yielded the
-best agreement for this $\approx 38\,$Myr age as well, given ${\rm
+best agreement for this $\approx$$38\,$Myr age as well, given ${\rm
 [Fe/H] = +0.1}$ and $A_V=0.2$
@@
 We then cleaned the candidate TESS rotation period measurements
-through a combination of automated and manual steps.  \deleted{As a
+through a combination of automated and manual steps.\deleted{ As a
 matter of scope, we restricted our attention to the \nkinwithtess\
 stars discussed in Section~\ref{sec:kinematicselection} in the spatial
-and kinematic proximity of Kepler\,1627.}  \deleted{Kepler rotation
+and kinematic proximity of Kepler\,1627.}\deleted{ Kepler rotation
 periods were derived by \citet{mcquillan_rotation_2014} for 28 of
 our \nkinwithtess\ stars; for these cases, we
-simply adopted the Kepler rotation period.} \added{First, to validate
+simply adopted the Kepler rotation period.}\added{ First, to validate
 the TESS rotation periods, we compared against 28 stars
@@
 harmonic.  Of the remaining 5 stars with Kepler rotation periods above
-10 days, none were correctly recovered by TESS (3 were near the
-half-period harmonic).  We therefore adopted a general stance of
-caution for TESS rotation periods exceeding 10 days, and used the
-Kepler rotation periods for the stars for which they were available.}
+10 days, none were correctly recovered by TESS, and 3 were near the
+half-period harmonic.  We therefore adopted a general stance of
+skepticism for TESS rotation periods exceeding 10 days, and used the
+Kepler rotation periods whenever possible.}
 For the remaining stars with only TESS data, we
@@
 distributions of other 30\,Myr to 50\,Myr clusters ({\it e.g.},
-IC\,2602 and IC\,2391) are indistinguishable\added{ from each other}
+IC\,2602 and IC\,2391) are indistinguishable
 \citep{douglas_stephanie_t_2021_5131306}.  Approximately 10 of the
@@
 achromatic transits strongly favor Kepler\ 1627A as the transit host,
-since Kepler\ 1627B is much redder.}  Conditioned on the
+since Kepler\ 1627B is a much redder star.}  Conditioned on the
 ephemeris and transit depth from the Kepler data, the MuSCAT3 observations also
@@
 further photometric follow-up would be necessary to determine
-whether the transit duration is indeed changing.  
+whether the transit duration is \replaced{indeed}{actually} changing.
+
 For our RM analysis,
@@
 \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.  
-\added{This of course would need to be confirmed through a measurement
-of the planetary mass; alternative explanations for its large size
-could be that it is abnormally massive, or that it has an abnormally
+\added{
+It could also be consistent with a scenario in which an earlier ``boil-off'' of
+the planet's atmosphere during disk dispersal decreases the entropy of
+the planetary interior, leading to a $\sim$10$^8$ year Kelvin-Helmholtz cooling and
+contraction timescale
+\citep{owen_constraining_2020}.
+Confirming either of these scenarios would require a measurement
+of the planetary mass; otherwise, alternative explanations for its large size
+also include that it is abnormally massive, or that it has an abnormally
 large envelope to core mass ratio.}

### 2021-12-01 drivers/calc_empirical_isochrone_age.py

@@ -277,31 +277,33 @@ def collect_isochrone_data():
     outpath = os.path.join(RESULTSDIR, 'empirical_isochrone_age',
                            f"all_clusters_CMD_binning_interp.png")
     set_style()
-    f,ax = plt.subplots(figsize=(1.5*6,1.5*4))
+    f,ax = plt.subplots(figsize=(4,4))
 
-    colors = 'k,orange,deepskyblue,C3'.split(',')
+    colors = 'k,orange,deepskyblue,C1'.split(',')
 
     for ix, c, color in zip(range(len(clusters)), clusters, colors):
 
-        ax.scatter(hr_dict[c]['(Bp-Rp)0'], hr_dict[c]['(MG)0'], s=1,
-                   c=color, label=c, zorder=ix)
+        #label = c+' spline'
+        label = c
+        ax.scatter(hr_dict[c]['(Bp-Rp)0'], hr_dict[c]['(MG)0'], s=0.25,
+                   c=color, zorder=ix)
         #ax.errorbar(hr_dict[c]['binmids'], hr_dict[c]['ymean'],
         #            hr_dict[c]['ystdev'], marker='o', elinewidth=0.5,
         #            capsize=4, lw=0, mew=0.5, color=f'k', markersize=3,
         #            zorder=5)
-        ax.plot(hr_dict[c]['x_interp'], hr_dict[c]['y_interp'], lw=1,
-                c=color, zorder=42, label=c+' spline')
+        ax.plot(hr_dict[c]['x_interp'], hr_dict[c]['y_interp'], lw=0.5,
+                c=color, zorder=42, label=label)
 
     for i in range(0,I_grid.shape[1],200):
         if i == 0:
-            ax.plot(x_eval, I_grid[:,i], color='gray', lw=0.5,
+            ax.plot(x_eval, I_grid[:,i], color='gray', lw=0.25,
                     label='Interpolated splines')
         else:
-            ax.plot(x_eval, I_grid[:,i], color='gray', lw=0.5)
+            ax.plot(x_eval, I_grid[:,i], color='gray', lw=0.25)
 
     ax.update({'xlabel': '$(G_{\mathrm{BP}}-G_{\mathrm{RP}})_0$ [mag]',
                'ylabel': '$M_{\mathrm{G},0}$ [mag]'})
-    ax.legend(loc='best')
+    ax.legend(loc='best', fontsize='x-small')
[…]

### 2021-12-06 paper_review/response_1.txt
update paper and review

@@
 Below, we have reproduced the relevant portions of the referee's
-report, alongside our responses. A list of changes is appended to the
-manuscript using the trackchanges AASTeX macros.
+report, alongside our responses. The changes are bolded in the 
+attached ms_boldchanges.pdf file, using the trackchanges AASTeX
+macros.
 
@@
 ----------
-> TODO ??? WHY ISN'T SIGMA OK?  THE POINT IS THAT IT'S 2-SIGMA...
+> The second paragraph of Section 4.1.3 has been modified to mention
+> the difference in BIC, and the corresponding approximate Bayes
+> factor. 
+>
+> The slope of $-21 \pm 10\,{\rm min\,day}^{-1}$ is similar to that
+> found by \citet{holczer_time_2015}. \added{The $\chi^2$ of the
+> best-fit line is $306.1$, with $n=140$ data points.  An alternative
+> model of a flat line yields $chi^2=315.6$.  The difference in the
+> Bayesian information criterion (BIC) between the two models is ${\rm
+> BIC}_{\rm flat} - {\rm BIC}_{\rm line} = 4.5$, which corresponds to a
+> Bayes factor of $\approx$9.4. According to the usual
+> \citet{kass_bayes_1995} criteria, this is ``positive'' evidence for
+> the model with a finite slope.  We view it as suggestive evidence at
+> best, particularly given the excess scatter in the transit timing
+> measurement uncertainties.}
+>
+> We have also modified the caption of Figure 6 to clarify that this
+> is not a detection.
 
@@
 ----------
+
+(NOTE: TODO: THIS RESPONSE HAS YET TO BE UPDATED WITH A MORE LUCID
+DESCRIPTION OF THE DERIVATION OF THE COMPANION MASS)
+
 > Thank you for catching this -- this citation is indeed deprecated
@@
 > The Baraffe models are available from
-> http://perso.ens-lyon.fr/isabelle.baraffe/BHAC15dir/BHAC15_tracks+structure.
+> http://perso.ens-lyon.fr/isabelle.baraffe/BHAC15dir/.
 >
@@
 ----------
+
+(NOTE: TODO: THIS RESPONSE STILL MIGHT BE UPDATED WITH A POSSIBLE
+DEEPER ANALYSIS OF THE UV)
+
 > Regarding the S-values, they are shown in the second panel of

### 2021-12-06 paper/ms.tex
update paper and review

@@
 factor of two of} previously reported values
-\replaced{from}{in} the literature \replaced{\citep[{\it
-e.g.},][]{GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}}{\citep{pecaut_star_2016,GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}}\added{,
-and are all small enough that the choice of whether to use them {\it vs.} other
+\replaced{from}{in} the literature \citep{pecaut_star_2016,GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}\added{, and are all small enough that the choice of whether to use them {\it vs.} other
 extinction estimates does not affect our primary conclusions}.
@@
 	\caption{
-    {\bf Possible evidence for a prograde orbit of Kepler 1627 Ab.} The time of
+    {\bf \replaced{Possible}{Weak} evidence for a prograde orbit of Kepler 1627 Ab.} The time of
     each Kepler transit was measured, along with the local slope of
-    the light curve.  The two quantities are weakly anti-correlated
+    the light curve.  The two quantities \replaced{are weakly}{might
+    be} anti-correlated
     ($\approx$2-$\sigma$), which
-    might be caused by starspot crossings during the first
+    \replaced{might}{could} be caused by starspot crossings during the first
     (second) half of transit inducing a positive (negative) TTV,
@@
 slope of $-21 \pm 10\,{\rm min\,day}^{-1}$ is similar to that
-found by \citet{holczer_time_2015}.
-
-One concern we had in this analysis was whether our transit fitting
+found by \citet{holczer_time_2015}. \added{The $\chi^2$ of the
+best-fit line is $306.1$, with $n=140$ data points.  An alternative
+model of a flat line yields $\chi^2=315.6$.  The difference in the
+Bayesian information criterion (BIC) between the two models is ${\rm
+BIC}_{\rm flat} - {\rm BIC}_{\rm line} = 4.5$,
+which corresponds to a Bayes factor of $\approx$9.4. According to the
+usual \citet{kass_bayes_1995} criteria, this is ``positive'' evidence
+for the model with a finite slope.  We view it as suggestive evidence
+at best, particularly given the excess scatter in the transit timing
+measurement uncertainties.}
+
+\replaced{One}{A separate} concern we had in this analysis was whether our transit fitting
 procedure might induce spurious correlations between the slope and
@@
 spectroscopically and photometrically.  We used the HIRES echelle
-spectrograph at the Keck-I telescope and MuSCAT3 at Haleakal\=a
+spectrograph at the Keck-I telescope and the MuSCAT3 photometer at Haleakal\=a
 Observatory on Maui, HI \citep{Narita_2020}.} \deleted{ We monitored
@@
 
-\listofchanges
+%\listofchanges
 

### 2021-12-06 paper_review/response_1.txt
update response

@@
 
-(NOTE: TODO: THIS RESPONSE HAS YET TO BE UPDATED WITH A MORE LUCID
-DESCRIPTION OF THE DERIVATION OF THE COMPANION MASS)
-
 > Thank you for catching this -- this citation is indeed deprecated
 > and erroneously was not updated during the writing process.  The
-> final adopted calculation was based on the masses and radii from
-> the Baraffe+2015 models, rather than the MESA models (Choi+2016).
-> The Baraffe models are available from
-> http://perso.ens-lyon.fr/isabelle.baraffe/BHAC15dir/.
+> final adopted calculation was based on the stellar masses and radii
+> from the Baraffe+2015 models, rather than the MESA models
+> (Choi+2016).  These models are available from
+> http://perso.ens-lyon.fr/isabelle.baraffe/BHAC15dir/.  We used
+> their 2MASS synthetic photometry to estimate the companion's mass,
+> as follows.
 >
-> We believe that the companion's mass is roughly between 0.3 and 0.4
-> Msun, and that the age of the system is ~38 +/- 6 Myr.  The Baraffe
-> models give:
+>     m_K = 11.194 (2MASS)
+>     m_K,A = 11.31
+>     m_K,B = 13.68 (NIRC2)
+>     m-M = 5*log10(329.5) - 5 = 7.59
+>     -> M_K,A = 3.72
+>     -> M_K,B = 6.09
 >
+> Assuming an age of 40 Myr, this yields M_comp = 0.31 Msun, by
+> interpolating between the Baraffe 0.3 and 0.4 Msun values (M_K =
+> 6.12 and 5.71, respectively).  Assuming an age of 30 Myr yields
+> M_comp = 0.27 Msun (M_K=6.48, and 5.90 for 0.20 and 0.30 Msun).
+> The bilinear interpolation at 38 Myr (our nominal age) yields
+> M_comp=0.30 Msun, Teff=3408K, R=0.451 Rsun, and log(g)=4.60.
+> In our original manuscript, we quoted a companion mass of 0.33
+> Msun; this has been corrected in the revision.
+>
+> Given our age uncertainies (38 +/- 6 Myr), a plausible mass range
+> based on these models for the companion is roughly 0.27 Msun to
+> 0.33 Msun.  The model densities over the relevant ages and masses
+> are
+>
+> 	age    mass        radius      density
 >   32 Myr 0.300 Msun, 0.481 Rsun, 3.80 g / cm3
->    35 Myr 0.300 Msun, 0.465 Rsun, 4.21 g / cm3
 >   40 Myr 0.300 Msun, 0.443 Rsun, 4.86 g / cm3
->    32 Myr 0.400 Msun, 0.555 Rsun, 3.30 g / cm3
->    35 Myr 0.400 Msun, 0.537 Rsun, 3.64 g / cm3
->    40 Myr 0.400 Msun, 0.514 Rsun, 4.15 g / cm3
 >   32 Myr 0.330 Msun, 0.506 Rsun, 3.60 g / cm3
->    35 Myr 0.330 Msun, 0.489 Rsun, 3.98 g / cm3
 >   40 Myr 0.330 Msun, 0.467 Rsun, 4.58 g / cm3
+>   32 Myr 0.350 Msun, 0.518 Rsun, 3.55 g / cm3
+>   40 Myr 0.350 Msun, 0.479 Rsun, 4.50 g / cm3
+>   38 Myr 0.300 Msun, 0.451 Rsun, 4.61 g / cm3
 > 
-> i.e., our default assumption was 40 Myr and 0.33 Msun.  The age is
-> adopted for the reasons discussed in the text.  The mass is based
-> off the observed deltaMag in 832nm and K' (3.14 +/- 0.15 and 2.37
-> +/- 0.02 respectively), which implies absolute magnitudes for the
-> secondary of K'=XXX and [832]=YYY.  Based on a set of absolute
-> magnitude and color relations from a submitted paper by A. Kraus,
-> these gave ~=0.33 Msun.  (TODO: NEED TO BEEF THIS UP!)
->
-> The plausible range of companion densities is therefore 3.30 g/cc
-> to 4.86 g/cc.  We've updated the text to reflect this uncertainty,
-> but it doesn't change the overall point that our measured stellar
-> density (2.00+/-0.24 g/cc) is still a fair bit lower than the
-> expected range.
-
-TODO TODO REMOVE THIS
-> As a sanity check, these calibrations find that for a companion
-> spectral type of M2.5, one expects colors of G-K'=3.39 mag and
-> G-[832] = 1.51 mag, though we cannot test this since Gaia DR2/EDR3
-> does not resolve the secondary.
->
-> For the primary spectral type of G8, one expects G-K'=1.58 mag and
-> G-[832] = 0.71 mag (i.e., K'=13.02-1.58=11.44, and
-> [832]=G-0.71=12.31 for the primary).  So it holds together
-> (bolstering the case for it being a bound companion) to find
-> DeltaK' = 2.37 mag and Delta[832] = 3.14 mag.
-
-G8V: M_G = 5.098
-M2.5V: M_G = 9.67 (ZAMS).
--> delta G = 9.67 - 5.098 = 4.57, so G =13.02+4.57= 17.592 for secondary
--> G-K'=3.39mag implies K' = 14.20 for secondary
--> G-[832]=1.51mag implies [832] = 16.08 for secondary
-->
--> expected delta K' = 2.76
--> expected delta 832 = 3.78
-.. (both are observed a bit lower!  i.e., it's not red enough)
+> This means that for the argument concerning the stellar density
+> inferred from the transit, the worst-case scenario would be that of
+> a) the companion being more massive than expected  (~0.35 Msun),
+> simultaneously with b) the system being younger than expected (~32
+> Myr).  In this scenario, the companion star's density would be ~=
+> 3.5 g/cc.  We've updated the text to reflect this uncertainty.
 
@@
 > 19.052 +/- 0.017 [AB mags], with a NUV S/N of 65 (2015, ApJ, 813,
-> 100). Loyd et al 2020 similarly report a NUV flux of 4e-13 +/-
-> 4e-15 mW/m^2, though the latter value is based on the same
-> GALEX-CAUSE Kepler survey data presented by Olmedo+2015.  UV
-> detections can sometimes be used as a youth indicator (e.g., Figure
-> 9 of Newton+2021 AJ 161 65, which Newton et al attribute from the
-> idea in Shkolnik+2011, ApJ 727 6).  For Kepler-1627, we have 2MASS
-> J = 11.690+/-0.021, and therefore m_NUV-m_J = 7.36.  Given the Gaia
+> 100), through the GALEX-CAUSE Kepler survey.  UV detections can
+> sometimes be used as a youth indicator (e.g., Figure 9 of
+> Newton+2021 AJ 161 65, which Newton et al attribute from the idea
+> in Shkolnik+2011, ApJ 727 6).  For Kepler-1627, we have 2MASS J =
+> 11.690+/-0.021, and therefore m_NUV-m_J = 7.36.  Given the Gaia
 > (BP-RP)0 of 0.977, comparing to Figure 9 of Newton+2021 we find
@@
 > particularly strong diagnostic of the age for Kepler-1627 as an
-> individual star.  It may well be worth exploring the GALEX-CAUSE
-> Kepler survey data for the delta Lyr cluster generally, since the
-> later type K and M dwarfs are expected to be elevated with respect
-> to the Pleiades and Hyades sequences.  However, this goes beyond
-> the scope of the present work, and the combination of the cluster
-> membership, stellar rotation, stellar lithium content, and stellar
-> activity all point to an age consistent with that reported in our
-> work.
+> individual star.  
+>
+> To substantiate the latter claim, we cross-matched our lists of
+> Pleaides and δ Lyr cluster members against 2MASS and GALEX.  The
+> results are in galex_jmk.png and galex_bpmrp.png, attached.  We
+> attempted including IC2602 as well, but only got a few matches
+> because it was not observed.  We see that the earliest A and F
+> stars in the del Lyr cluster (38 Myr) are brighter in the UV than
+> the Pleiades sequence (112 Myr).  However, there's not really any
+> information in this statement independent of the work that's
+> already done in the paper using Gaia isochrones and the stellar
+> rotation.
 

### 2021-12-06 paper/ms.tex
update response

@@
 planet (Section~\ref{sec:planet}), we confirm that despite the
-existence of the previously unreported M2.5V companion, hereafter
+existence of the previously unreported \replaced{M2.5}{M3}V companion, hereafter
 Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A.
@@
 		\leavevmode
-		\includegraphics[width=0.95\textwidth]{f1.pdf}
+		\includegraphics[width=0.99\textwidth]{f1.pdf}
 	\end{center}
@@
 
+\subsubsection{Color-Absolute Magnitude Diagram}
+\label{sec:camd}
+
+We measured the isochrone age using an empirical approach.  The \added{upper }left
+panel of Figure~\ref{fig:age} shows the color-absolute magnitude
+diagram (CAMD) of candidate \cn\ members, IC\,2602\added{ ($\approx 38$\,Myr)}, the Pleiades\added{ ($\approx 115$\,Myr)}, and
+the field. The stars from the Pleiades and IC\,2602 were adopted from
+\citet{CantatGaudin2018a}, and the field stars are from the Gaia EDR3
+Catalog of Nearby Stars \citep{gaia_gcns_2021}.  
+\added{We also
+compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
+and the Upper-Centaurus-Lupus (UCL) component of the Sco OB2
+association ($\approx$$16$\,Myr; \citealt{pecaut_star_2016}).
+We adopted the UCL members from \citet{Damiani2019}.
+For visual clarity, the latter two clusters are not shown in Figure~\ref{fig:age}.
+}
+We cleaned \replaced{these}{the membership lists}
+following the data filtering criteria from \citet[][Appendix
+B]{GaiaCollaboration2018}, except that we weakened the parallax
+precision requirement to $\varpi/\sigma_\varpi>5$.  \added{This also involved cuts
+on the photometric signal to noise ratio, the number of visibility
+periods used, the astrometric $\chi^2$ of the single-source solution, and the
+$G_{\rm BP}-G_{\rm RP}$ color excess factor.} These filters were
+designed to include genuine binaries while omitting instrumental
+artifacts.
+
+\replaced{We then corrected for extinction by querying}{To correct for extinction, we queried} the
+3-dimensional maps of \citet{capitanio_threedimensional_2017} and
+\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/},
+2021/09/25}, and applied the extinction coefficients $k_X\equiv
+A_X/A_0$ computed by \citet{GaiaCollaboration2018} assuming that $A_0
+= 3.1 E(B-V)$.  For \added{UCL, }IC\,2602, the Pleiades, and the \cn, this
+procedure yielded a respective mean and standard deviation for the
+reddening of \replaced{$E(B-V)=\{0.020\pm0.003, 0.045\pm0.008,
+0.032\pm0.006\}$}{$E(B-V)=\{0.084\pm0.041, 0.020\pm0.003, 0.045\pm0.008,
+0.032\pm0.006\}$}.
+These values \replaced{agree reasonably well with}{are within a
+factor of two of} previously reported values
+\replaced{from}{in} the literature \citep{pecaut_star_2016,GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}\added{, and are all small enough that the choice of whether to use them {\it vs.} other
+extinction estimates does not affect our primary conclusions}.
+
+% Gaia+2018 E(B-V):
+% Pleiades 0.045, IC2602 0.031
+%
+% Randich+2018 E(B-V):
+% Pleiades ??, IC2602 0.032 - 0.092
+%
+% KC2019: AV->E(B-V)
+% Pleiades 0.220 AV -> 0.071
+% IC2602 0.243297 -> 0.078
+% Stephenson1 0.184 -> 0.059
+%
+% Bossini+2019 AV->E(B-V)
+% PleiadesMelotte22 0.140 -> 0.045
+% IC2602 0.096 -> 0.031
+% Steph1: 0.157 -> 0.051
+
 \begin{figure*}[tp]
@@
 \end{figure*}
-
-\subsubsection{Color-Absolute Magnitude Diagram}
-\label{sec:camd}
-
-We measured the isochrone age using an empirical approach.  The \added{upper }left
-panel of Figure~\ref{fig:age} shows the color-absolute magnitude
-diagram (CAMD) of candidate \cn\ members, IC\,2602\added{ ($\approx 38$\,Myr)}, the Pleiades\added{ ($\approx 115$\,Myr)}, and
-the field. The stars from the Pleiades and IC\,2602 were adopted from
-\citet{CantatGaudin2018a}, and the field stars are from the Gaia EDR3
-Catalog of Nearby Stars \citep{gaia_gcns_2021}.  
-\added{We also
-compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
-and the Upper-Centaurus-Lupus (UCL) component of the Sco OB2
-association ($\approx$$16$\,Myr; \citealt{pecaut_star_2016}).
-We adopted the UCL members from \citet{Damiani2019}.
-For visual clarity, the latter two clusters are not shown in Figure~\ref{fig:age}.
-}
-We cleaned \replaced{these}{the membership lists}
-following the data filtering criteria from \citet[][Appendix
-B]{GaiaCollaboration2018}, except that we weakened the parallax
-precision requirement to $\varpi/\sigma_\varpi>5$.  \added{This also involved cuts
-on the photometric signal to noise ratio, the number of visibility
-periods used, the astrometric $\chi^2$ of the single-source solution, and the
-$G_{\rm BP}-G_{\rm RP}$ color excess factor.} These filters were
-designed to include genuine binaries while omitting instrumental
-artifacts.
-
-\replaced{We then corrected for extinction by querying}{To correct for extinction, we queried} the
-3-dimensional maps of \citet{capitanio_threedimensional_2017} and
-\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/},
-2021/09/25}, and applied the extinction coefficients $k_X\equiv
-A_X/A_0$ computed by \citet{GaiaCollaboration2018} assuming that $A_0
-= 3.1 E(B-V)$.  For \added{UCL, }IC\,2602, the Pleiades, and the \cn, this
-procedure yielded a respective mean and standard deviation for the
-reddening of \replaced{$E(B-V)=\{0.020\pm0.003, 0.045\pm0.008,
-0.032\pm0.006\}$}{$E(B-V)=\{0.084\pm0.041, 0.020\pm0.003, 0.045\pm0.008,
-0.032\pm0.006\}$}.
-These values \replaced{agree reasonably well with}{are within a
-factor of two of} previously reported values
-\replaced{from}{in} the literature \citep{pecaut_star_2016,GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}\added{, and are all small enough that the choice of whether to use them {\it vs.} other
-extinction estimates does not affect our primary conclusions}.
-
-% Gaia+2018 E(B-V):
-% Pleiades 0.045, IC2602 0.031
-%
-% Randich+2018 E(B-V):
-% Pleiades ??, IC2602 0.032 - 0.092
-%
-% KC2019: AV->E(B-V)
-% Pleiades 0.220 AV -> 0.071
-% IC2602 0.243297 -> 0.078
-% Stephenson1 0.184 -> 0.059
-%
-% Bossini+2019 AV->E(B-V)
-% PleiadesMelotte22 0.140 -> 0.045
-% IC2602 0.096 -> 0.031
-% Steph1: 0.157 -> 0.051
 
@@
 		{\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
-		from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2.5V
+		from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$\replaced{M2.5}{M3}V
 		companion at $\rho \approx 0\farcs16$, which corresponds to a
@@
 Kepler\,1627 is a binary.  Given the distance and age, the models of
-\citet{baraffe_new_2015} imply a companion mass of $M_{\rm B} \approx
-0.33 M_{\odot}$ and companion temperature of $T_{\rm eff,B} \approx
-3450$ K.  The corresponding spectral type is roughly M2.5V
+\citet{baraffe_new_2015} imply a companion mass of \replaced{$M_{\rm B} \approx
+0.33 M_{\odot}$}{$M_{\rm B} \approx 0.30 M_{\odot}$} and companion
+temperature of \replaced{$T_{\rm eff,B} \approx 3450$}{$T_{\rm eff,B}
+\approx 3408$} K.  The corresponding spectral type is roughly \replaced{M2.5}{M3}V
 \citep{pecaut_mamajek_2013}.  These models combined with the NIRC2
@@
 		\leavevmode
-		\includegraphics[width=0.49\textwidth]{f6.pdf}
+		\includegraphics[width=0.47\textwidth]{f6.pdf}
 	\end{center}
@@
 slope of $-21 \pm 10\,{\rm min\,day}^{-1}$ is similar to that
-found by \citet{holczer_time_2015}. \added{The $\chi^2$ of the
-best-fit line is $306.1$, with $n=140$ data points.  An alternative
-model of a flat line yields $\chi^2=315.6$.  The difference in the
-Bayesian information criterion (BIC) between the two models is ${\rm
+found by \citet{holczer_time_2015}. \added{The
+best-fit line yields $\chi^2 = 306.1$, with $n=140$ data points.  An alternative
+model of just a flat line yields $\chi^2=315.6$.  The difference in the
+Bayesian information criterion between the two models is ${\rm
 BIC}_{\rm flat} - {\rm BIC}_{\rm line} = 4.5$,
@@
 		\leavevmode
-		\includegraphics[width=0.45\textwidth]{f7.pdf}
+		\includegraphics[width=0.47\textwidth]{f7.pdf}
 	\end{center}
@@
 \end{figure}
-\begin{figure*}[!t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f8.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
-		Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
-		3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
-		its uncertainty.  \pn\ is shown with a star.  The large sizes of
-		the youngest transiting planets could be explained by their
-		primordial atmospheres not yet having evaporated; direct
-		measurements of the atmospheric outflows or planetary masses would
-		help to confirm this expectation.  Selection effects may also be
-		important.  Parameters are from the NASA Exoplanet Archive (2021
-		Sept 15).
-		\label{fig:rp_period_age}
-	}
-\end{figure*}
 
@@
 our knowledge it would be the youngest planet yet found by the
-\replaced{main}{prime}
-Kepler mission.\footnote{The re-purposed K2 mission however has found
+\replaced{main}{prime} Kepler mission.\footnote{The re-purposed K2 mission however has found
 two younger systems containing five planets: K2-33b ($9\pm1\,{\rm
@@
 $\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
-theoretically expected density for Kepler\,1627B is $\approx 4.6\,{\rm
+theoretically expected density for Kepler\,1627B \added{given its
+nominal age of 38\,Myr and mass of 0.30\,$M_\odot$} is $\approx 4.6\,{\rm
 g\,cm}^{-3}$ \citep{baraffe_new_2015}.
 The transit duration is therefore too long to be explained by a star
-eclipsing the M dwarf secondary at $10$-$\sigma$.  While the planet
+eclipsing the M dwarf secondary at $10$-$\sigma$.\added{ Adversarially assuming a younger age (32\,Myr) and a larger companion
+mass (0.35\,$M_\odot$), the companion's density could be as low as $\approx 3.5\,{\rm
+g\,cm}^{-3}$.  This is still in tension with the data.} While the planet
 might hypothetically still orbit a hidden close and bright companion,
@@
 \label{sec:conc}
+
+\begin{figure*}[!t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f8.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
+		Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
+		3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
+		its uncertainty.  \pn\ is shown with a star.  The large sizes of
+		the youngest transiting planets could be explained by their
+		primordial atmospheres not yet having evaporated; direct
+		measurements of the atmospheric outflows or planetary masses would
+		help to confirm this expectation.  Selection effects may also be
+		important.  Parameters are from the NASA Exoplanet Archive (2021
+		Sept 15).
+		\label{fig:rp_period_age}
+	}
+\end{figure*}
 

### 2021-12-06 paper_review/response_1.txt
update response (ready to resubmit)

@@
 ----------
-> delta Lyr cluster was the first name suggested by Stephenson
+> delta Lyra cluster was the first name suggested by Stephenson
 > (1958), and it is the default name listed in SIMBAD.  It was also
@@
 > of the larger Cep-Her complex, we would prefer to leave this as
-> Ronan's result to present!  There's significant work here in
+> Ronan's result to present.  There's significant work here in
 > designing the clustering method and verifying that it works.
@@
 > broader association (which is different in nuanced ways from the
-> gray Theia 73 points in Figure 1).  Given the complexity, and the
-> potential importance of the complex given its proximity and age, we
-> really would prefer to leave this for the in prep R. Kerr work.
-> And similarly, the introduction of the term "Cep-Her complex" is
-> probably best left for that future work, where it can be mentioned
-> that the "del Lyr cluster" is a subset of this complex.
+> gray Theia 73 points in Figure 1).  Given these complications, and
+> the potential importance of the complex given its proximity and
+> age, we really would prefer to leave this for the in prep R. Kerr
+> work.  And similarly, the introduction of the term "Cep-Her
+> complex" is probably best left for that future work, where it can
+> be mentioned that the "del Lyr cluster" is a subset of this
+> complex.
 
@@
 ----------
-> Thank you for this suggestion.
->
-> The first paragraph of Section 2.2.1 has been reorganized to now
-> introduce all five of the relevant clusters: UCL, IC 2602, del Lyr
-> cluster, mu-Tau, and the Pleaides.  The second paragraph has been
-> modified to account for this hcnage.
->
-> The extinction for UCL is now also listed in the text.  So is the
-> adopted age of IC 2602.
+> Thank you for this suggestion.  The first paragraph of Section
+> 2.2.1 has been reorganized to now introduce all five of the
+> relevant clusters: UCL, IC 2602, del Lyr cluster, mu-Tau, and the
+> Pleaides.  The second paragraph has been modified to account for
+> this change.  The extinction for UCL is now also listed in the
+> text.  So is the adopted age of IC 2602.
 
@@
 > The resulting photometry is dominated by a quasi-periodic starspot
-> signal with a peak-to-peak amplitude that varies between 2\% and 8\%.
-> \added{Given that the secondary companion's brightness in the Kepler
-> band is 1\% to 2\% that of the primary, source confusion for the
-> rotation signal is not expected to be an issue.}
-  """
-> The rationale here is that the rotation amplitude on the secondary
-> would need to be a factor of >two times its intrinsic brightness for
-> any confusion to happen, and the primary would need to not show any
+> signal with a peak-to-peak amplitude that varies between 2\% and
+> 8\%.  \added{Given that the secondary companion's brightness in the
+> Kepler band is 1\% to 2\% that of the primary, source confusion for
+> the rotation signal is not expected to be an issue.}
+> """
+>
+> The rationale is that the rotation amplitude on the secondary would
+> need to be a factor of >two times its intrinsic brightness for any
+> confusion to happen, and the primary would need to not show any
 > rotation signal.  Such large variability amplitudes have not been
-> observed except for disked stars (e.g., Rebull+2020, Figure 10), and
-> there is no evidence for a secondary period in the Lomb Scargle
+> observed except for disked stars (e.g., Rebull+2020, Figure 10),
+> and there is no evidence for a secondary period in the Lomb Scargle
 > periodograms.
@@
 ----------
-> Line 642 of the original manuscript notes "Although we did not
-> detect the RM anomaly, the multi-band MuSCAT3 light curves show
-> that the transit is achromatic (Figure 7)."  We have nonetheless
-> reworded this section to improve the clarity with which this point
-> is made.
+> Thank you -- we have reworded this section to improve the clarity
+> with which this point is made.
 
@@
 > We've added a few sentences in the manuscript to clarify these
-> latter points.
+> points.
 
@@
 > model of a flat line yields $chi^2=315.6$.  The difference in the
-> Bayesian information criterion (BIC) between the two models is ${\rm
-> BIC}_{\rm flat} - {\rm BIC}_{\rm line} = 4.5$, which corresponds to a
-> Bayes factor of $\approx$9.4. According to the usual
-> \citet{kass_bayes_1995} criteria, this is ``positive'' evidence for
-> the model with a finite slope.  We view it as suggestive evidence at
-> best, particularly given the excess scatter in the transit timing
-> measurement uncertainties.}
+> Bayesian information criterion (BIC) between the two models is
+> ${\rm BIC}_{\rm flat} - {\rm BIC}_{\rm line} = 4.5$, which
+> corresponds to a Bayes factor of $\approx$9.4. According to the
+> usual \citet{kass_bayes_1995} criteria, this is ``positive''
+> evidence for the model with a finite slope.  We view it as
+> suggestive evidence at best, particularly given the excess scatter
+> in the transit timing measurement uncertainties.}
 >
@@
 > Myr).  In this scenario, the companion star's density would be ~=
-> 3.5 g/cc.  We've updated the text to reflect this uncertainty.
+> 3.5 g/cc.  We've updated the manuscript to reflect this point.
 
@@
 > 3 is <~0.2 mag.  So, if we saw a star with the same (BP-RP) color
-> as Kepler 1627 that were
-> >~0.3 mag brighter in M_G, we would classify it as a likely
-> >photometric
-> binary.  This means we are able to exclude stars ~0.32* as bright
-> as the primary (in Gaia G-band). This corresponds to excluding
-> bound companions more massive than ~0.85 Msun (the primary's mass
-> is ~0.95 Msun).  The SB2 tests do much better if the velocity
-> separation exceeds the CCF width (down to 1% as bright as the
-> primary;
+> as Kepler 1627 that were more than ~0.3 mag brighter in M_G, we
+> would classify it as a likely photometric binary.  This means we
+> are able to exclude stars ~0.32* as bright as the primary (in Gaia
+> G-band). This corresponds to excluding bound companions more
+> massive than ~0.85 Msun (the primary's mass is ~0.95 Msun).  The
+> SB2 tests do much better if the velocity separation exceeds the CCF
+> width (down to 1% as bright as the primary;
 > https://ui.adsabs.harvard.edu/abs/2015AJ....149...18K/abstract).
@@
 
-(NOTE: TODO: THIS RESPONSE STILL MIGHT BE UPDATED WITH A POSSIBLE
-DEEPER ANALYSIS OF THE UV)
-
 > Regarding the S-values, they are shown in the second panel of
@@
 > To substantiate the latter claim, we cross-matched our lists of
-> Pleaides and δ Lyr cluster members against 2MASS and GALEX.  The
-> results are in galex_jmk.png and galex_bpmrp.png, attached.  We
-> attempted including IC2602 as well, but only got a few matches
-> because it was not observed.  We see that the earliest A and F
+> IC2602, Pleaides, and δ Lyr cluster members against 2MASS and
+> GALEX.  The results are in galex_jmk.png and galex_bpmrp.png,
+> attached.  IC2602 only got a few matches because it was in a region
+> of sky not observed by GALEX.  We see that the earliest A and F
 > stars in the del Lyr cluster (38 Myr) are brighter in the UV than
 > the Pleiades sequence (112 Myr).  However, there's not really any
-> information in this statement independent of the work that's
-> already done in the paper using Gaia isochrones and the stellar
-> rotation.
+> new information in this statement that goes beyond the work that's
+> already been done in the manuscript using Gaia isochrones and the
+> stellar rotation.  So, for brevity, we would prefer to omit
+> this analysis from the manuscript -- but thank you for suggesting
+> it!
 

### 2021-12-06 paper/ms.tex
update response (ready to resubmit)

@@
 irradiation and internal heat can also power gradual outflows
-which, if strong enough, can deplete or entirely strip the atmosphere \citep{lopez_how_2012,Owen_Wu_2013,ginzburg_corepowered_2018}.\added{ The latter photoevaporative and core-powered
+which, if strong enough, can deplete or entirely strip the atmosphere \citep{lopez_how_2012,Owen_Wu_2013,ginzburg_corepowered_2018}.\added{ The photoevaporative and core-powered
 	outflows are thought to persist for $\approx$$10^8$ to $\approx$$10^9$ years,
@@
 \citet{CantatGaudin2018a}, and the field stars are from the Gaia EDR3
-Catalog of Nearby Stars \citep{gaia_gcns_2021}.  
-\added{We also
+Catalog of Nearby Stars \citep{gaia_gcns_2021}.\added{ We also
 compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
@@
 We adopted the UCL members from \citet{Damiani2019}.
-For visual clarity, the latter two clusters are not shown in Figure~\ref{fig:age}.
-}
-We cleaned \replaced{these}{the membership lists}
+For visual clarity, the latter two clusters are not shown in Figure~\ref{fig:age}.} We cleaned \replaced{these}{the membership lists}
 following the data filtering criteria from \citet[][Appendix
@@
 period measurements ($G<17$ and $\bpmrpo>0.5$) using the TESS full
-frame image data. 
-\added{As a matter of
+frame image data.\added{ As a matter of
 scope, we restricted our attention to the \nkinwithtess\ stars
 discussed in Section~\ref{sec:kinematicselection} in the spatial and
-kinematic proximity of Kepler\,1627. }
-We extracted light curves from
+kinematic proximity of Kepler\,1627.} We extracted light curves from
 the TESS images using the nearest pixel to each star, and regressed
@@
 usual \citet{kass_bayes_1995} criteria, this is ``positive'' evidence
-for the model with a finite slope.  We view it as suggestive evidence
-at best, particularly given the excess scatter in the transit timing
-measurement uncertainties.}
+for the model with a finite slope.  We view it as suggestive
+at best, particularly given the poor reduced $\chi^2$.}
 
@@
 a correction based on the Combined Differential Photometric Precision
-(CDPP) in the \sn\ light curve over the relevant transit timescales
+in the \sn\ light curve over the relevant transit timescales
 \citep{2017ksci.rept...19B}.} The large size of \pn\ relative
 to most Kepler mini-Neptunes might therefore support a picture
-in which the typical $5\,M_\oplus$ \deleted{to 10$\,M_\oplus$} mini-Neptune
+in which the typical $5\,M_\oplus$\deleted{ to 10$\,M_\oplus$} mini-Neptune
 \citep{wu_mass_2019} loses a significant fraction of its primordial
@@
 the planet's atmosphere during disk dispersal decreases the entropy of
-the planetary interior, leading to a $\sim$10$^8$ year Kelvin-Helmholtz cooling and
+the planetary interior, leading to a $\sim$10$^8$ year Kelvin-Helmholtz
 contraction timescale
@@
 of the planetary mass; otherwise, alternative explanations for its large size
-also include that it is abnormally massive, or that it has an abnormally
+could include that it is abnormally massive, or that it has an abnormally
 large envelope to core mass ratio.}

### 2021-12-07 paper_review/response_1.txt
AAS resubmitted

@@
 We thank you for organizing the review of our paper, and also thank
-the referee for constructive criticism.  We have considered the
-referee's remarks carefully and revised our manuscript accordingly.
+the referees for constructive criticism.  We have considered the
+remarks of both of the reviewers carefully and revised our manuscript
+accordingly.
 
-Below, we have reproduced the relevant portions of the referee's
-report, alongside our responses. The changes are bolded in the 
-attached ms_boldchanges.pdf file, using the trackchanges AASTeX
-macros.
+Below, we have reproduced the relevant portions of the reviews' reports,
+alongside our responses. The changes are bolded in the  attached
+ms_boldchanges.pdf file, using the trackchanges AASTeX macros.
 
@@
 > $38^{+6}_{-5}$ (stat) $^{+6}_{-8}$ (sys)" is that the systematic
-> uncertainties are not Gaussian and are not based on any probability
-> distribution that we can easily assess -- they're estimated based
-> on the extremes of previous literature age measurements.  Quoting
-> this systematic range next to a number that has been derived based
-> on a more formal approach (and that has statistical meaning within
-> that approach) somehow misses the point.  Given the comparable
-> amplitude of the two sources of uncertainty, does this matter?
-> Discussing the range of extremes in the text does communicate the
-> point that this is a source of systematic uncertainty.
+> uncertainties are not based on any probability distribution that
+> we can easily assess -- they're estimated based on the extremes of
+> previous literature age measurements.  Quoting this systematic
+> range next to a number that has been derived based on a more
+> formal approach (and that has statistical meaning within that
+> approach) somehow misses this point.  We would prefer to leave the
+> discussion in the manuscript as-is in order to communicate the
+> spread of ages from different measurement approaches (which
+> reflects a systematic uncertainty in the overall age scale),
+> rather than quoting separate statistical and systematic
+> uncertainties.
 
@@
 ----------
-> From each other (and the manuscript has been updated; thank you!).
+> From each other, and the manuscript has been updated; thank you!
 
@@
 ----------
-> Thanks for this point -- we've updated the text with a reference to
-> Stauffer+16.  In their Section 5.1, they discuss the case of some
-> unresolved F+K binaries in the Pleiades which show anomalously long
-> K2 rotation periods.
+> Thank you for this point -- we've updated the text with a
+> reference to Stauffer+16.  In their Section 5.1, they discuss the
+> case of some unresolved F+K binaries in the Pleiades which show
+> anomalously long K2 rotation periods.
 
@@
 ----------
-> Indeed.  Batalha+2010 (ApJL) discuss the selection function, which
-> favored stars that were both small and bright since these stars
-> enable the detection of Earth-size planets.  M dwarfs thus were
-> generally excluded unless they happened to be very close.  F, G,
-> and K stars were favored, particularly when they were nearby.  The
-> sentence in L25 currently states that "The star was observed by the
-> Kepler space telescope because it is nearby (d=329pc) and it
-> resembles the Sun".  More directly, yes, the star was chosen
-> because it was thought that it might enable the detection of small
-> planets in the habitable zone.  However the underlying reason for
-> why the star enables such detections is based on the star's
-> brightness and assumed radius, which both factor into the transit
-> SNR.  Our claim therefore holds, albeit indirectly.  It also serves
-> the more important purpose of informing the reader what kind of
-> star this paper is about: a star that is at a distance of ~330
-> parsecs, and that resembles the Sun (except for its age!).
+> Thank you for this point.  Batalha+2010 (ApJL) discuss the
+> selection function.  The sentence in L25 currently states that
+> "The star was observed by the Kepler space telescope because it is
+> nearby (d=329pc) and it resembles the Sun".  More directly, yes,
+> the star was chosen because it was thought that it might enable
+> the detection of small planets in the habitable zone.  However the
+> underlying reason for why the star enables such detections is
+> based on the star's brightness and assumed radius, which both
+> factor into the transit SNR ( SNR ~ δ/σ ~ (Rplanet/Rstar)^2 *
+> (1/σ(Kepmag)); for σ(Kepmag) the relative precision as a function
+> of Kepler magnitude ).  Our claim therefore holds, albeit
+> indirectly: a star of the same color that was too far away would
+> not have been selected because it would be too faint.  If it were
+> the wrong color, it also wouldn't have been selected.  But perhaps
+> we are getting a bit too pedantic: this sentence also serves the
+> more important purpose of informing the reader what kind of star
+> this paper is about: a star that is at a distance of ~330 parsecs,
+> and that resembles the Sun (except for its age!).
 
@@
 > Douglas et al. 2021, which should have pointed to
-> https://ui.adsabs.harvard.edu/abs/2021tsc2.confE.157D/abstract, and
-> has been updated in the revision.  The rotation periods from this
-> work, which includes IC 2602, IC 2391, and a few other 30-50 Myr
-> clusters, are not yet public.  
+> https://ui.adsabs.harvard.edu/abs/2021tsc2.confE.157D/abstract,
+> and has been updated in the revision.  The rotation periods from
+> this work, which includes IC 2602, IC 2391, and a few other 30-50
+> Myr clusters, are not yet public.  
 >
@@
 > Thank you for this point -- please see kepler_vs_tess_prot.png in
-> the attached, which shows the 28 referenced stars that overlap.  5
-> of the 28 stars have Kepler rotation periods above 10 days, of
-> which none were correctly recovered by TESS (3 are chosen at the
-> half-period harmonic).  Of the 23 stars with periods below 10 days,
-> 21 of the TESS periods agree with the Kepler rotation periods, and
-> 2 are at the double-period harmonic.
+> the attached response, which shows the 28 referenced stars that
+> overlap.  5 of the 28 stars have Kepler rotation periods above 10
+> days, of which none were correctly recovered by TESS (3 are chosen
+> at the half-period harmonic).  Of the 23 stars with periods below
+> 10 days, 21 of the TESS periods agree with the Kepler rotation
+> periods, and 2 are at the double-period harmonic.
 >

### 2021-12-07 paper/ms.tex
AAS resubmitted

@@
 \received{2021 Oct 12}
-\revised{---}
+\revised{2021 Dec 6}
 \accepted{---}
@@
 latter two clusters.}{The
-pre-main-sequence M dwarfs of the \cn\ were also seen to be ``below''
-UCL and ``above'' the Pleiades.}  To turn this heuristic interpolation
+pre-main-sequence M dwarfs of a fixed color in the \cn\ are more luminous than those of the Pleiades, and were also seen to be 
+less luminous than those in
+UCL.}  To turn this heuristic interpolation
 into \replaced{an age}{a quantitative age}
@@
 10 days, none were correctly recovered by TESS, and 3 were near the
-half-period harmonic.  We therefore adopted a general stance of
-skepticism for TESS rotation periods exceeding 10 days, and used the
+half-period harmonic.  We were therefore 
+wary of any TESS rotation periods exceeding 10 days, and used the
 Kepler rotation periods whenever possible.}
@@
 \citet{baraffe_new_2015} imply a companion mass of \replaced{$M_{\rm B} \approx
-0.33 M_{\odot}$}{$M_{\rm B} \approx 0.30 M_{\odot}$} and companion
+0.33 M_{\odot}$}{$M_{\rm B} \approx 0.30\,M_{\odot}$} and companion
 temperature of \replaced{$T_{\rm eff,B} \approx 3450$}{$T_{\rm eff,B}
-\approx 3408$} K.  The corresponding spectral type is roughly \replaced{M2.5}{M3}V
+\approx 3408$}\,K.  The corresponding spectral type is roughly \replaced{M2.5}{M3}V
 \citep{pecaut_mamajek_2013}.  These models combined with the NIRC2
@@
 timing variation, is plotted in Figure~\ref{fig:corr} against
-the local linear slope for the fourth-order polynomial baseline.  The
+the local linear slope \replaced{for}{assuming} the fourth-order polynomial baseline.  The
 slope of $-21 \pm 10\,{\rm min\,day}^{-1}$ is similar to that
@@
 procedure might induce spurious correlations between the slope and
-transit time.  In particular, using the second-order polynomial
+transit time.  In particular, assuming the second-order polynomial
 baseline yielded a larger anti-correlation between the TTVs and local
@@
 transit;
-using this baseline would therefore bias the
+assuming this baseline would therefore bias the
 result.  However for the fourth-order baseline, the correlation
@@
 performed without knowledge of the low-mass stellar companion ($M_{\rm
-B}\approx 0.33\,M_\odot$).  Validated planets have also previously
+B}\approx 0.30\,M_\odot$).  Validated planets have also previously
 been refuted \citep[{\it e.g.},][]{shporer_three_2017}.  We therefore
@@
 argument, assume the former value.  The observed transit has a depth
-of $\approx$0.18\%.  A 18\% deep eclipse of Kepler\,1627B would
+of $\approx$0.18\%.  An $\approx$18\% deep eclipse of Kepler\,1627B would
 therefore be needed to produce a signal with the appropriate depth.
@@
 ($\approx 68\,{\rm min}$).  The actual measured ingress and egress
-duration is $\approx 17\,{\rm min}$, a factor of four times too
+duration is $\approx$$17\,{\rm min}$, a factor of four times too
 short.  The combination of Kepler 1627B's brightness, the transit
@@
 Kepler 1627B might host the transit signal.
-
-% Baraffe+15: M_B 0.33 Mdot
-% Rstar = 0.455 --> 0.465 Rsun
 
@@
 theoretically expected density for Kepler\,1627B \added{given its
-nominal age of 38\,Myr and mass of 0.30\,$M_\odot$} is $\approx 4.6\,{\rm
+nominal age of 38\,Myr and mass of 0.30\,$M_\odot$} is $\approx$$4.6\,{\rm
 g\,cm}^{-3}$ \citep{baraffe_new_2015}.
@@
 mass (0.35\,$M_\odot$), the companion's density could be as low as $\approx 3.5\,{\rm
-g\,cm}^{-3}$.  This is still in tension with the data.} While the planet
+g\,cm}^{-3}$.  This is still in tension with the fitted stellar density.} While the planet
 might hypothetically still orbit a hidden close and bright companion,
@@
 photometric variability at very young ages \citep[{\it
-e.g.},][]{zhou_2021_tois}.  \replaced{However,}{However, it seems that} smaller planets could have been
-detected: \added{based on the per-target detection contours, }the Kepler pipeline's median completeness extended to
+e.g.},][]{zhou_2021_tois}.
+
+ \replaced{However,}{It seems though that} smaller planets could have been
+detected\added{ in the \sn\ system}: \added{based on the per-target detection contours, }the Kepler pipeline's median completeness extended to
 $1.6\,R_\oplus$ at 10~day orbital periods, and $3.3\,R_\oplus$ at
@@
 a correction based on the Combined Differential Photometric Precision
-in the \sn\ light curve over the relevant transit timescales
+in the light curve over the relevant transit timescales
 \citep{2017ksci.rept...19B}.} The large size of \pn\ relative
@@
 the planet's atmosphere during disk dispersal decreases the entropy of
-the planetary interior, leading to a $\sim$10$^8$ year Kelvin-Helmholtz
+the planetary interior, leading to a rather long $\sim$10$^8$ year Kelvin-Helmholtz
 contraction timescale
@@
 of the planetary mass; otherwise, alternative explanations for its large size
-could include that it is abnormally massive, or that it has an abnormally
+could include that it is just abnormally massive, or that it has an abnormally
 large envelope to core mass ratio.}

### 2022-02-27 papers/paper_kep1627/paper/ms.tex
mv paper -> papers/paper_kep1627

--- 2021-12-07 paper/ms.tex
@@
 %\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
-\documentclass[12pt,modern,twocolumn,tighten,linenumbers]{aastex63}
+\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
 %\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
@@
 \revised{2021 Dec 6}
-\accepted{---}
-\submitjournal{AAS Journals}
+\accepted{2021 Dec 27}
+%\submitjournal{AAS Journals}
 \shorttitle{Kepler\,1627}
@@
 %
-Finally, this research has made use of the Keck Observatory Archive (KOA),
-which is operated by the W. M. Keck Observatory and the NASA Exoplanet
-Science Institute (NExScI), under contract with the National
-Aeronautics and Space Administration.  We also thank the Keck
+%Finally, this research has made use of the Keck Observatory Archive (KOA),
+%which is operated by the W. M. Keck Observatory and the NASA Exoplanet
+%Science Institute (NExScI), under contract with the National
+%Aeronautics and Space Administration.
+Finally, we also thank the Keck
 Observatory staff for their support of HIRES and remote observing.  We

### 2022-02-28 papers/authors.tex
stage Cep-Her outline

--- 2021-08-24 paper/authors.tex
@@
 \correspondingauthor{L.\,G.\,Bouma}
-\email{bouma.luke@gmail.com}
+\email{luke@astro.caltech.edu}
 
 \author[0000-0002-0514-5538]{L. G. Bouma}
-\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+\altaffiliation{51 Pegasi b Fellow}
+\affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
+%\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
 
 % Key authors:
+% ... Kinematics
+\author[0000-0002-6549-9792]{R.~Kerr}
+\affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}% % ... Kepler correlations
+
 % ... stellar rotation & the initial crossmatch
@@
 
-% ... Kepler correlations
-\author[0000-0003-1298-9699]{K. Masuda}
-\affiliation{Department of Earth and Space Science, Osaka University, Osaka 560-0043, Japan}
-
-% ... HIRES PI
+% \author[0000-0003-1298-9699]{K. Masuda}
+% \affiliation{Department of Earth and Space Science, Osaka University, Osaka 560-0043, Japan}
+% 
+% % ... planet-fitting, kinematics, meta.
 \author{L. A. Hillenbrand}
 \affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
-
-% ... RM fitting
-\author[0000-0001-7409-5688]{G. Stefansson}
-\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
+% 
+% % ... RM fitting
+% \author[0000-0001-7409-5688]{G. Stefansson}
+% \altaffiliation{Henry Norris Russell Fellow}
+% \affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
 
@@
 \author[0000-0002-0531-1073]{H. Isaacson}
-\affiliation{Astronomy Department, University of California, Berkeley,
-CA 94720, USA}
-
-%
-% MUSCAT3 Collaborators
-%
-\author[0000-0001-8511-2981]{N. Narita}
-\affiliation{Komaba Institute for Science, The University of Tokyo, 3-8-1 Komaba, Meguro, Tokyo 153-8902, Japan}
-\affiliation{Japan Science and Technology Agency, PRESTO, 3-8-1 Komaba, Meguro, Tokyo 153-8902, Japan}
-\affiliation{Astrobiology Center, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
-\affiliation{Instituto de Astrof\'{i}sica de Canarias (IAC), 38205 La Laguna, Tenerife, Spain}
-
-\author[0000-0002-4909-5763]{A. Fukui} % afukui@g.ecc.u-tokyo.ac.jp
-\affiliation{Komaba Institute for Science, The University of Tokyo, 3-8-1 Komaba, Meguro, Tokyo 153-8902, Japan}
-\affiliation{Instituto de Astrof\'{i}sica de Canarias (IAC), 38205 La Laguna, Tenerife, Spain}
-
-\author[0000-0002-5658-5971]{Masahiro Ikoma} % ikoma.masahiro@gmail.com
-\affiliation{Division of Science, National Astronomical Observatory of Japan, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
-
-\author[0000-0002-6510-0681]{M. Tamura} % motohide.tamura@nao.ac.jp
-\affiliation{Department of Astronomy, University of Tokyo, 7-3-1 Hongo, Bunkyo-ku, Tokyo 113-0033, Japan}
-\affiliation{Astrobiology Center, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
-\affiliation{National Astronomical Observatory of Japan, 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
+\affiliation{Astronomy Department, University of California, Berkeley, CA 94720, USA}
+% pending
 
 % AO IMAGING
-\author[0000-0002-2532-2853]{S. B. Howell}
-\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
+\author[0000-0001-9811-568X]{A.~L.~Kraus}
+\affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}% % ... Kepler correlations

### 2022-02-28 papers/ms.tex
stage Cep-Her outline

--- 2022-02-27 papers/paper_kep1627/paper/ms.tex
@@
 %\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
-\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
+\documentclass[12pt,modern,tighten]{aastex63}
 %\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
@@
 
-\newcommand{\cn}{$\delta$\ Lyr\ cluster} % cluster name
-\newcommand{\sn}{Kepler\,1627} % star system name (binary)
-\newcommand{\pn}{Kepler\,1627Ab} % planet name
+\newcommand{\cn}{Cep-Her complex} % cluster name
+\newcommand{\sysone}{Kepler-1627} % star system name (binary)
+\newcommand{\stone}{Kepler-1627 A} % star system name (binary)
+\newcommand{\plone}{Kepler-1627 Ab} % planet name
+\newcommand{\systwo}{Kepler-1643} % star system name (binary)
+\newcommand{\sttwo}{Kepler-1643} % star system name (binary)
+\newcommand{\pltwo}{Kepler-1643 b} % planet name
+\newcommand{\systhree}{KOI-7368} % star system name (binary)
+\newcommand{\stthree}{KOI-7368} % star system name (binary)
+\newcommand{\plthree}{KOI-7368 b} % planet name
+\newcommand{\sysfour}{KOI-7913 } % star system name (binary)
+\newcommand{\stfour}{KOI-7913 A} % star system name (binary)
+\newcommand{\plfour}{KOI-7913 Ab} % planet name
 
 \newcommand{\clusterage}{$38^{+6}_{-5}$\,Myr} %
-
-% gaia target sample numbers
-\newcommand{\nkinematic}{1{,}201} % deltalyr_kc19_cleansubset_withreddening.csv
-% all from oneoff_calculate_rotation_numbers.py
-\newcommand{\noriginal}{3{,}071} % Vizier; KC19 Theia 73 member count
-\newcommand{\nwithtess}{924} % How many stars have TESS data with G<17 and (Bp-Rp)0 > 0.5?
-\newcommand{\nforgyro}{1842} % How many stars have G<17 and (Bp-Rp)0 > 0.5?
-\newcommand{\nkinwithtess}{391} % How many kinematically selected stars have TESS data?
-\newcommand{\nkinwithtessandcrowding}{192} % How many kinematically selected stars have TESS data and meeting the crowding cutoff (nequal<=0, nclose<=1)?
-\newcommand{\nkindefaultcleaning}{145} % How many of these meet the "defaultcleaning" criteria (LSP>0.1, Prot<15, and crowding cut)?
-\newcommand{\nrotgood}{135} % nkindefaultcleaning, minus outliers.
-\newcommand{\nfracprot}{70} % fraction nrotgood/nkinwithtessandcrowding
 
@@
 \received{2021 Oct 12}
-\revised{2021 Dec 6}
-\accepted{2021 Dec 27}
+\revised{---}
+\accepted{---}
 %\submitjournal{AAS Journals}
-\shorttitle{Kepler\,1627}
+\shorttitle{Kepler Mini-Neptunes in Cep-Her}
 
@@
 \title{
-  A 38 Million Year Old Neptune-Sized Planet in the Kepler Field
+  Three 38 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
 }
@@
 \begin{abstract}
-  Kepler\,1627A is a G8V star previously known to host a
-  $3.8\,R_\oplus$ \replaced{mini-Neptune}{planet} on a 7.2\,day orbit.  The star was
-  observed by the Kepler space telescope because it is nearby
-  ($d=329\,{\rm pc}$) and it resembles the Sun.  Here we show using
-  Gaia kinematics, TESS stellar rotation periods, and spectroscopic
-  lithium abundances that Kepler\,1627 is a member of the \clusterage\
-  old $\delta$~Lyr cluster.  To our knowledge, this makes
-  Kepler\,1627Ab the youngest planet with a precise age yet found by
-  the \replaced{main}{prime} Kepler mission.  The Kepler photometry shows two
-  peculiarities: the average transit profile is asymmetric, and the
-  individual transit times might be correlated with the local light curve
-  slope.  We discuss possible explanations for each anomaly.  More
-  importantly, the $\delta$~Lyr cluster
-  is one of $\sim$10$^3$ coeval groups whose properties have been clarified
-  by Gaia.  Many other exoplanet hosts are candidate members
-  of these clusters; their ages can be verified \replaced{through}{with} the
-  trifecta of Gaia, TESS, and ground-based spectroscopy.
+  Stellar positions and velocities from Gaia
+  are yielding a refined view of stellar clusters during the
+  first hundred million years of their lives.
+  Here we present an analysis of a group of $38 \pm 6$ million year old stars
+  in the Cepheus-Hercules region, hereafter the Cep-Her complex.
+  This group of stars includes four previously known Kepler Objects of
+  Interest:
+  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$),
+  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
+  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), and
+  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$).
+	Kepler-1627 is a previously validated young planet
+	in a component of the Cep-Her complex called the $\delta$\ Lyr\ cluster \citep{bouma_kep1627_2022}.
+	Here we focus on the latter three systems, which are in other sub-components
+	of the complex (RSG5 and CH-2).
+	Based on kinematic evidence from Gaia, stellar rotation periods from TESS, and spectroscopy,
+	these three systems are also $38 \pm 6$ million years old.
+	Based on the transit shapes and high resolution imaging, we
+	statistically validate that they are all most likely planets (false positive probabilities of $6\times10^{-9}$,
+	$5\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368, and
+	KOI-7913 respectively).
+	Supplemented by Gaia and TESS, the main Kepler mission is now contributing to the
+	census of young close-in planets, and Kepler-1643 and KOI-7913 are the first
+	empirical demonstration that mini-Neptunes with sizes of $\approx$2 Earth radii
+	exist at ages of roughly 40 million years.
 \end{abstract}
@@
 
+% * Main text <3500 words (not including acknowledgements, appendices, or other
+%   supplementary)
+
 \section{Introduction}
-
-While thousands of exoplanets have been discovered orbiting nearby
-stars, the vast majority of them are several billion years old.  This
-makes it difficult to test origin theories for the different families
-of planets, since many evolutionary processes are expected to operate
-on timescales of less than 100 million years.
-
-For instance, the ``mini-Neptunes'', thought to be made of metal
-cores, silicate mantles \citep{kite_atmosphere_2020}, and extended
-hydrogen-dominated atmospheres, are expected to shrink
-in size by factors of several over their first $10^8$ years.
-Specifically, in the models of \citet{owen_atmospheres_2016} and
-\citet{owen_constraining_2020}, the $\approx$$5\,M_\oplus$ planets start
-with sizes of 4\,--\,12\,$R_\oplus$ shortly after the time of disk
-dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
-2\,--\,4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
-is expected to occur within the first few million years after the disk
-disperses \citep{ikoma_situ_2012}, stellar
-irradiation and internal heat can also power gradual outflows
-which, if strong enough, can deplete or entirely strip the atmosphere \citep{lopez_how_2012,Owen_Wu_2013,ginzburg_corepowered_2018}.\added{ The photoevaporative and core-powered
-	outflows are thought to persist for $\approx$$10^8$ to $\approx$$10^9$ years,
-	though the details depend on the planetary masses,
-	the irradiation environments, and the initial atmospheric
-	mass fractions \citep{owen_evaporation_2017,gupta_signatures_2020,2021MNRAS.503.1526R,2021MNRAS.501L..28K}.} Discovering young planets, measuring their masses, and detecting their
-atmospheric outflows are key steps toward testing this paradigm, which
-is often invoked to explain the observed radius distribution of mature
-exoplanets \citep{Fulton_et_al_2017,van_eylen_asteroseismic_2018}.
-
-The K2 and TESS missions have now enabled the detection of about ten
-close-in planets younger than 100 million years, all smaller than
-Jupiter
-\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020,martioli_aumicbc_2021}.
-The Kepler mission however has not yielded any planets with precise
-ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
-during the \replaced{main}{prime} Kepler mission (2009--2013), only four open clusters
-were known in the Kepler field,
-%: NGC\,6866, NGC\,6811, NGC\,6819, and NGC\,6791,
-with ages spanning 0.7\,Gyr to 9\,Gyr \citep{meibom_kepler_2011}.
-Though isochronal, gyrochronal, and lithium-based analyses suggest
-that younger Kepler planets do exist
-\citep{walkowicz_rotation_2013,berger_identifying_2018,david_sizes_2021}, accurate and precise
-age measurements typically require an ensemble of stars.  Fortunately,
-recent analyses of the Gaia data have greatly expanded our knowledge
-of cluster memberships \citep[{\it
-e.g.},][]{CantatGaudin2018a,Zari2018,KounkelCovey2019,Meingast2021,Kerr2021}.
-As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
-\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
-from the literature, which yielded a list of candidate young and
-age-dated stars (see Appendix~\ref{app:targetlist}).
-
-Matching our young star list against stars observed by Kepler revealed
-that Kepler observed a portion of the \cn\ (Stephenson-1; Theia~73).
-More specifically, a clustering analysis of the Gaia data by
-\citet{KounkelCovey2019} reported that Kepler~1627 (KIC 6184894; KOI
-5245) is a \cn\ member.  Given the previous statistical
-validation of the close-in \replaced{mini-Neptune}{Neptune-sized
-planet} Kepler 1627b
-\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
-we begin by scrutinizing the properties of the cluster
-(Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr cluster is
-\clusterage\ old, and in Section~\ref{sec:stars} show that \sn\ is
-both a binary and also a member of the cluster.  Focusing on the
-planet (Section~\ref{sec:planet}), we confirm that despite the
-existence of the previously unreported \replaced{M2.5}{M3}V companion, hereafter
-Kepler\,1627B, the planet orbits the G-dwarf primary, Kepler\,1627A.
-We also analyze an asymmetry in the average transit profile, and a
-possible correlation between the individual transit times and the
-local light curve slope.  We conclude by discussing broader
-implications for our ability to age-date a larger sample of planets
-(Section~\ref{sec:conc}).
 
@@
     {\bf Galactic positions  and tangential velocities of stars in the
-    $\delta$\,Lyr cluster.} Points are reported cluster members from
-    \citet{KounkelCovey2019}.  The tangential velocities relative to
-    \sn\ (bottom right) are computed assuming that every star has the
-    same three-dimensional spatial velocity as \sn.  Our analysis
-    considers stars (black points) in the spatial and kinematic
-    vicinity of Kepler\,1627 (yellow star).  The question of whether
-    the other candidate cluster members (gray points) are part of the
-    cluster is outside our scope.  The location of the Sun is
-    ($\odot$) is shown \added{to clarify the direction along which
-    parallax uncertainties are expected to produce erroneous
-    clusters}.
+    $\delta$\,Lyr cluster.} 
 		\label{fig:XYZvtang}
@@
 
-To measure the age of the $\delta$ Lyr cluster, we first selected a
-set of candidate cluster members
-(Section~\ref{sec:kinematicselection}), and then analyzed these stars
-using a combination of the isochronal and gyrochronal techniques
-(Section~\ref{sec:clusterage}).
-
 \subsection{Selecting Cluster Members}
@@
 
-\citet{KounkelCovey2019} applied an unsupervised clustering algorithm
-to Gaia DR2 on-sky positions, proper motions, and parallaxes for stars
-within the nearest kiloparsec.  For the \cn\ (Theia 73), they reported
-\noriginal\ candidate members.  We matched these stars against the
-latest Gaia EDR3 observations using the \texttt{dr2\_neighbourhood}
-table from the ESA archive,
-taking the stars closest in proper motion and epoch-corrected angular
-distance as the presumed match \citep{gaia_collaboration_2021_edr3}.
-\replaced{In Figure~\ref{fig:XYZvtang}, \added{we }have shown galactic positions only for
-the stars with parallax signal-to-noise exceeding 20. }{For plotting purposes, we focused only
-on the stars with parallax signal-to-noise exceeding 20.} \added{We
-calculated the tangential velocities for each of these stars relative to \sn\
-($\Delta \mu^{*}$) by subtracting the observed proper motion from what the proper motion at each star's
-position would be if it were co-moving with \sn. }
-
-\replaced{The}{Figure~\ref{fig:XYZvtang} shows that the} reported
-cluster members (gray and black points) extend over a much larger
-volume \added{in both physical and kinematic space }than the cluster previously identified by
-\citet{stephenson_possible_1959} and later corroborated by
-\citet{eggen_photometric_1968}.  While the non-uniform ``clumps'' of
-stars might comprise a {\it bona fide} cluster of identically-aged
-stars, they could also be heavily contaminated by field stars.
-\added{One reason to suspect this is that the spread in tangential
-velocities exceeds typical limits for kinematic coherence
-\citep{Meingast2021}}.
-We
-therefore considered stars only in the immediate kinematic and spatial
-vicinity of \sn\ as candidate cluster members.  We performed
-\replaced{the}{this}
-selection cut\deleted{s} manually, by drawing lassos with the interactive
-\texttt{glue} visualization tool \citep{beaumont_2014_13866} in the
-four projections shown in Figure~\ref{fig:XYZvtang}.  The overlap
-between the Kepler field and the resulting candidate cluster members
-is shown in Figure~\ref{fig:skychart}.  While this method will include
-some field interlopers in the ``cluster star'' sample, and vice-versa,
-it should suffice for our aim of verifying the existence of the
-cluster in the vicinity of \sn.
-%\added{The application of a more careful membership selection procedure to the $\delta$\ Lyr\ cluster (as part of a larger complex spanning Cepheus and Hercules) is a topic under current investigation by by R.~Kerr and collaborators.}
-
-\begin{figure}[t]
+\begin{figure*}[tp]
 	\begin{center}
 		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f2.pdf}
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f2a.pdf}
+			\includegraphics[width=0.469\textwidth]{f2b.pdf}
+		}
+
+		\vspace{-0.6cm}
+		\subfloat{
+			\includegraphics[width=0.7\textwidth]{f2c.pdf}
+		}
 	\end{center}
@@
 	\caption{
-    {\bf Kepler's view of the $\delta$~Lyr cluster, shown in equatorial
-    	coordinates.} Each black circle
-    is a candidate cluster member selected based on its position and kinematics (Figure~\ref{fig:XYZvtang}).  Of
-    the \nkinematic\ candidate cluster members, 58 have at least one
-    quarter of Kepler data.  TESS has also observed most of the
-    cluster, for one to two lunar months to date.
-		\label{fig:skychart}
+		{\bf The \cn\ is \clusterage\ old.} 
+		\label{fig:age}
 	}
-\end{figure}
+\end{figure*}
 
@@
 
-We measured the isochrone age using an empirical approach.  The \added{upper }left
-panel of Figure~\ref{fig:age} shows the color-absolute magnitude
-diagram (CAMD) of candidate \cn\ members, IC\,2602\added{ ($\approx 38$\,Myr)}, the Pleiades\added{ ($\approx 115$\,Myr)}, and
-the field. The stars from the Pleiades and IC\,2602 were adopted from
-\citet{CantatGaudin2018a}, and the field stars are from the Gaia EDR3
-Catalog of Nearby Stars \citep{gaia_gcns_2021}.\added{ We also
-compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
-and the Upper-Centaurus-Lupus (UCL) component of the Sco OB2
-association ($\approx$$16$\,Myr; \citealt{pecaut_star_2016}).
-We adopted the UCL members from \citet{Damiani2019}.
-For visual clarity, the latter two clusters are not shown in Figure~\ref{fig:age}.} We cleaned \replaced{these}{the membership lists}
-following the data filtering criteria from \citet[][Appendix
-B]{GaiaCollaboration2018}, except that we weakened the parallax
-precision requirement to $\varpi/\sigma_\varpi>5$.  \added{This also involved cuts
-on the photometric signal to noise ratio, the number of visibility
-periods used, the astrometric $\chi^2$ of the single-source solution, and the
-$G_{\rm BP}-G_{\rm RP}$ color excess factor.} These filters were
-designed to include genuine binaries while omitting instrumental
-artifacts.
+\subsubsection{Stellar Rotation Periods}
 
-\replaced{We then corrected for extinction by querying}{To correct for extinction, we queried} the
-3-dimensional maps of \citet{capitanio_threedimensional_2017} and
-\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/},
-2021/09/25}, and applied the extinction coefficients $k_X\equiv
-A_X/A_0$ computed by \citet{GaiaCollaboration2018} assuming that $A_0
-= 3.1 E(B-V)$.  For \added{UCL, }IC\,2602, the Pleiades, and the \cn, this
-procedure yielded a respective mean and standard deviation for the
-reddening of \replaced{$E(B-V)=\{0.020\pm0.003, 0.045\pm0.008,
-0.032\pm0.006\}$}{$E(B-V)=\{0.084\pm0.041, 0.020\pm0.003, 0.045\pm0.008,
-0.032\pm0.006\}$}.
-These values \replaced{agree reasonably well with}{are within a
-factor of two of} previously reported values
-\replaced{from}{in} the literature \citep{pecaut_star_2016,GaiaCollaboration2018,KounkelCovey2019,bossini_age_2019}\added{, and are all small enough that the choice of whether to use them {\it vs.} other
-extinction estimates does not affect our primary conclusions}.
+\section{The Stars}
+\label{sec:stars}
 
-% Gaia+2018 E(B-V):
-% Pleiades 0.045, IC2602 0.031
-%
-% Randich+2018 E(B-V):
-% Pleiades ??, IC2602 0.032 - 0.092
-%
-% KC2019: AV->E(B-V)
-% Pleiades 0.220 AV -> 0.071
-% IC2602 0.243297 -> 0.078
-% Stephenson1 0.184 -> 0.059
-%
-% Bossini+2019 AV->E(B-V)
-% PleiadesMelotte22 0.140 -> 0.045
-% IC2602 0.096 -> 0.031
-% Steph1: 0.157 -> 0.051
+\subsection{Kepler\,1627A}
+\subsection{Kepler\,1643}
+\subsection{KOI-7368}
+\subsection{KOI-7913}
+Is a binary.
+
+\section{The Planets}
+\label{sec:planet}
 
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f3a.pdf}
-			\includegraphics[width=0.469\textwidth]{f3b.pdf}
+			\includegraphics[width=0.5\textwidth]{f3a.png}
+			\includegraphics[width=0.5\textwidth]{f3b.png}
 		}
 
-		\vspace{-0.6cm}
 		\subfloat{
-			\includegraphics[width=0.7\textwidth]{f3c.pdf}
+			\includegraphics[width=0.5\textwidth]{f3c.png}
+			\includegraphics[width=0.5\textwidth]{f3d.png}
 		}
@@
 	\caption{
-		{\bf The \cn\ is \clusterage\ old.}  {\it Top:} Color-absolute
-		magnitude diagram of candidate \cn\ members, in addition to stars
-		in IC\,2602 ($\approx38$\,Myr), the Pleiades ($\approx 115$\,Myr),
-		and the Gaia EDR3 Catalog of Nearby Stars (gray background).  The
-		zoomed right panel highlights the pre-main-sequence.  The \cn\ and
-		IC\,2602 are approximately the same isochronal age.  {\it Bottom:}
-		TESS and Kepler stellar rotation period versus dereddened Gaia
-		color, with the Pleiades and Praesepe (650\,Myr) shown for
-		reference \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most
-		candidate \cn\ members are gyrochronally younger than the
-		Pleiades; outliers are probably field interlopers.
-		\label{fig:age}
+		{\bf planets.} 
+		\label{fig:planets}
 	}
 \end{figure*}
-
-Figure~\ref{fig:age} shows that the \cn\ and IC\,2602 overlap, and
-therefore are approximately the same age.  \replaced{In our exploration, we also
-compared against $\mu$-Tau ($62\pm7$\,Myr; \citealt{gagne_mutau_2020})
-and the Upper-Centaurus-Lupus (UCL) component of the Sco OB2
-association ($\approx$$16$\,Myr; \citealt{pecaut_star_2016}). The
-pre-main-sequence M dwarfs of the \cn\ were intermediate between the
-latter two clusters.}{The
-pre-main-sequence M dwarfs of a fixed color in the \cn\ are more luminous than those of the Pleiades, and were also seen to be 
-less luminous than those in
-UCL.}  To turn this heuristic interpolation
-into \replaced{an age}{a quantitative age}
-measurement, we used the empirical method developed by
-\citet{gagne_mutau_2020}.  In brief, we fitted the pre-main-sequence
-loci of a set of reference clusters, and the locus of the target \cn\
-was then modeled as a piecewise linear combination of these reference
-clusters.  For our reference clusters, \replaced{we adopted members of UCL,
-IC\,2602, and the Pleiades from \citet{Damiani2019} and
-\cite{CantatGaudin2018a} respectively}{we used UCL, IC\,2602, and the
-Pleiades}.  We removed binaries by
-requiring $\texttt{RUWE}<1.3$, $\texttt{radial\_velocity\_error}$
-below the 80$^{\rm th}$ percentile of each cluster's distribution, and
-excluded stars that were obvious photometric binaries in the CAMD.\footnote{
-  For a description of the renormalized unit weight error (RUWE), see the GAIA DPAC technical note
-  \url{http://www.rssd.esa.int/doc_fetch.php?id=3757412}.}  We then passed a
-moving box average and standard deviation across the CAMD in 0.10\,mag bins,
-fitted a univariate spline to the binned values, and assembled a
-piecewise grid
-of \deleted{hybrid }isochrones spanning the ages between UCL \replaced{to}{and} the Pleiades using
-Equation\replaced{s~6 and~7}{~6} from
-\citet{gagne_mutau_2020}.  
-\added{To derive a probability distribution function for the age of \cn, we
-then assumed a Gaussian likelihood that treated the interpolated isochrones as
-the ``model'' and the $\delta$\ Lyr\ cluster's isochrone as the ``data''
-(Equation~7 from \citealt{gagne_mutau_2020}).  The cluster's age and its
-statistical uncertainty are then quoted as the the mean and standard deviation
-of this age posterior.}
-
-% Randich+18: 43.7 +4.3 - 3.9 LDB
-% Dobbie+10: 46 +6 -5 LDB
-
-The ages returned by this procedure depend on the ages assumed for
-each reference cluster.  We adopted a 115\,Myr age for the Pleiades
-\citep{dahm_2015}, and a 16\,Myr age for UCL \citep{pecaut_star_2016}.
-The age of IC\,2602 however is the most important ingredient, since it
-receives the most weight in the interpolation.  Plausible ages for
-IC\,2602 span 30\,Myr to 46\,Myr, with older ages being preferred by
-the lithium-depletion-boundary (LDB) measurements
-\citep{dobbie_ic_2010,randich_gaiaeso_2018} and younger ages by the
-main-sequence turn-off
-\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.  If
-we were to adopt the 30\,Myr age for IC\,2602, then the \cn\ would be
-$31^{+5}_{-4}$\,Myr old.  For the converse extreme of 46\,Myr, the
-\cn\ would be $44^{+8}_{-7}$\,Myr old.  We adopt an intermediate
-38\,Myr age for IC\,2602, which yields an age for the \cn\ of
-\clusterage.\footnote{Our exploration of the PARSEC and MIST isochrone
-models over a grid of ages, metallicities, and reddenings, yielded the
-best agreement for this $\approx$$38\,$Myr age as well, given ${\rm
-[Fe/H] = +0.1}$ and $A_V=0.2$
-\citep{bressan_parsec_2012,choi_mesa_2016};  this preferred CAMD
-reddening is higher than the \citet{lallement_threedimensional_2018}
-value by a factor of two.  } Follow-up studies of the LDB or
-main-sequence turn-off in the \cn\ could help determine a more precise
-and accurate age for the cluster, and are left for future work.
-
-\subsubsection{Stellar Rotation Periods}
-
-Of the \noriginal\ candidate \cn\ members reported by
-\citet{KounkelCovey2019}, \nwithtess\ stars were amenable to rotation
-period measurements ($G<17$ and $\bpmrpo>0.5$) using the TESS full
-frame image data.\added{ As a matter of
-scope, we restricted our attention to the \nkinwithtess\ stars
-discussed in Section~\ref{sec:kinematicselection} in the spatial and
-kinematic proximity of Kepler\,1627.} We extracted light curves from
-the TESS images using the nearest pixel to each star, and regressed
-them against systematics with the causal pixel model implemented in
-the \texttt{unpopular} package \citep{hattorio_2021_cpm}.  We then
-measured candidate rotation periods using a Lomb-Scargle periodogram
-\citep{lomb_1976,scargle_studies_1982,astropy_2018}.  To enable cuts
-on crowding, we queried the Gaia source catalog for stars within a
-$21\farcs0$ radius of the target star (a radius of 1 TESS pixel).
-Within this radius, we recorded the number of stars with greater
-brightness than the target star, and with brightness within 1.25 TESS magnitudes of the target star.
-
-We then cleaned the candidate TESS rotation period measurements
-through a combination of automated and manual steps.\deleted{ As a
-matter of scope, we restricted our attention to the \nkinwithtess\
-stars discussed in Section~\ref{sec:kinematicselection} in the spatial
-and kinematic proximity of Kepler\,1627.}\deleted{ Kepler rotation
-periods were derived by \citet{mcquillan_rotation_2014} for 28 of
-our \nkinwithtess\ stars; for these cases, we
-simply adopted the Kepler rotation period.}\added{ First, to validate
-the TESS rotation periods, we compared against 28 stars
-from \citet{mcquillan_rotation_2014} that were also observed by Kepler.  Of the 23 stars with Kepler
-periods below 10 days, 21 of the TESS periods agreed with the Kepler
-rotation periods; the other 2 were measured at the double-period
-harmonic.  Of the remaining 5 stars with Kepler rotation periods above
-10 days, none were correctly recovered by TESS, and 3 were near the
-half-period harmonic.  We were therefore 
-wary of any TESS rotation periods exceeding 10 days, and used the
-Kepler rotation periods whenever possible.}
-For the remaining stars with only TESS data, we
-focused only on the stars for which no companions were known with a brightness
-exceeding one-tenth of the target star in a $21\farcs0$ radius.  There
-were \nkinwithtessandcrowding\ stars that met these crowding requirements, and
-that had TESS data available.  For plotting purposes we then imposed a
-selection based on the strength of the signal itself: we required the
-Lomb Scargle power to exceed 0.2, and the period to be below 15\,days.
-
-The lower panel of Figure~\ref{fig:age} shows the resulting
-\nkindefaultcleaning\ stars.  
-The majority of these stars fall below
-the ``slow sequence'' of the Pleiades, consistent with a gyrochronal
-age for the \cn\ below 100\,Myr.  In fact, the rotation-color
-distributions of other 30\,Myr to 50\,Myr clusters ({\it e.g.},
-IC\,2602 and IC\,2391) are indistinguishable
-\citep{douglas_stephanie_t_2021_5131306}.  Approximately 10 of the
-\cn\ stars appear as outliers above the ``slow sequence''.  Assuming
-that they are all false positives ({\it i.e.}, field interlopers), our
-rotation period detection fraction would be
-$\nrotgood/\nkinwithtessandcrowding \approx \nfracprot\%$.   \replaced{The other
-stars are likely to be field contaminants.}{Although some of these outlier stars might be
-unresolved F+K binaries that are in the cluster
-\citep{stauffer_rotation_2016}, assuming that they are field contaminants
-provides a more secure lower bound of the rotation period detection fraction.}  A final possible
-confounding factor -- binarity -- is known to affect the ``fast
-sequence'' of stars beneath the slow sequence
-\citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}.  We do
-not expect it to change the central conclusion regarding the cluster's
-age.
-
-\section{The Stars}
-\label{sec:stars}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f4a.pdf}
-			\includegraphics[width=0.48\textwidth]{f4b.pdf}
-		}
-	\end{center}
-	\vspace{-0.5cm}
-	\caption{
-		{\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
-		from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$\replaced{M2.5}{M3}V
-		companion at $\rho \approx 0\farcs16$, which corresponds to a
-		projected separation of $53\pm4$\,AU.  The inset shows a cutout of
-		the stacked NIRC2 image (North is up, East is left, scale is set
-		by the separation of the binary).  The lines show 5-$\sigma$
-		contrast limits for the `Alopeke filters, and 6-$\sigma$ contrast
-		limits for NIRC2 outside of $0\farcs15$. {\it Right:} Gaia EDR3
-		renormalized unit weight error (RUWE) point estimates for
-		candidate $\delta$\,Lyr cluster members.  Since other members of
-		the cluster with similar brightnesses have comparable degrees of
-		photometric variability, the high RUWE independently suggests that
-		Kepler\,1627 is a binary. 
-		\label{fig:kep1627binary}
-	}
-\end{figure*}
-
-\subsection{Kepler\,1627A}
-\subsubsection{Age}
-% rotation period from plot_rotation_period_windowslider.py
-Based on the spatial and kinematic association of \sn\ with the \cn,
-and the assumption that the planet formed shortly after the star, it
-seems likely that \sn\ is the same age as the cluster. There are two
-consistency checks on whether this is true: rotation and lithium.
-Based on the Kepler light curve, the rotation period is
-$2.642\pm0.042$\,days, where the quoted uncertainty is based on the
-scatter in rotation periods measured from each individual Kepler
-quarter.  This is consistent with comparable cluster members
-(Figure~\ref{fig:age}).
-
-To infer the amount of Li\,\textsc{I} from the 6708\,\AA\ doublet
-\citep[{\it e.g.},][]{soderblom_ages_2014}, we acquired an iodine-free
-spectrum from Keck/HIRES on the night of 2021 March 26 using the
-standard setup and reduction techniques of the California Planet
-Survey \citep{howard_cps_2010}.  Following the equivalent width
-measurement procedure described by \citet{bouma_2021_ngc2516}, we find
-${\rm EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   This value does not
-correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Nonetheless,
-given the stellar effective temperature (Table~\ref{tab:starparams}),
-this measurement is in agreement with expectations for a
-$\approx40$\,Myr star ({\it e.g.}, as measured in IC~2602 by
-\citealt{randich_gaiaeso_2018}).  It is also larger than any lithium
-equivalent widths measured by \citet{berger_identifying_2018} in their
-analysis of 1{,}301 Kepler-star spectra.
-
-\subsubsection{Stellar Properties}
-
-The adopted stellar parameters are listed in
-Table~\ref{tab:starparams}.  The stellar mass, radius, and effective
-temperature are found by interpolating against a 38$\,$Myr MIST
-isochrone \citep{choi_mesa_2016}.  The statistical uncertainties are
-propagated from the absolute magnitude (mostly originating from the
-parallax uncertainty) and the color; the systematic uncertainties are
-taken to be the difference between the PARSEC
-\citep{bressan_parsec_2012} and MIST isochrones.  Reported
-uncertainties are a quadrature sum of the statistical and systematic
-components.  As a consistency check, we analyzed the aforementioned
-Keck/HIRES spectrum from the night of 2021 March 26 using a
-combination of \texttt{SpecMatch-Emp} for stellar properties, and
-\texttt{SpecMatch-Synth} for $v\sin i$ \citep{yee_SM_2017}.  This
-procedure yielded $T_{\rm eff}=5498\pm100\,{\rm K}$, $\log
-g=4.6\pm0.1$, $[{\rm Fe/H}]=0.15\pm0.10$ from \texttt{SpecMatch-Emp},
-and $v\sin i = 18.9\pm1.0$ from \texttt{SpecMatch-Synth}.  These
-values are within the $1$-$\sigma$ uncertainties of our adopted values
-from the isochrone interpolation.
-
-\subsection{Kepler\,1627B}
-
-We first noted the presence of a close neighbor in the \sn\ system on
-2015 July 22 when we acquired adaptive optics imaging using the NIRC2
-imager on Keck-II.  We used the narrow camera (FOV = 10.2\arcsec) to
-obtain 8 images in the $K'$ filter ($\lambda = 2.12\,\mu$m) with a
-total exposure time of 160\,s. We analyzed these data following
-\citet{kraus_impact_2016}, which entailed using PSF-fitting to measure
-the separation, position angle, and contrast of the candidate
-companion.  The best-fitting empirical PSF template was identified
-from among the near-contemporaneous observations of single stars in
-the same filter.  The mean values inferred from the 8 images are
-reported in Table~\ref{tab:starparams}.  To estimate the detection
-limits, we analyzed the residuals after subtracting the empirical PSF
-template. Within each residual image, the flux was measured through
-40\,mas apertures centered on every pixel, and then the noise as a
-function of radius was estimated from the RMS within concentric rings.
-Finally, the detection limits were estimated from the strehl-weighted
-sum of the detection significances in the image stack, and we adopted
-the $6$-$\sigma$ threshold as the detection limit for ruling out
-additional companions.
-
-We also observed \sn\ on Gemini-North using the `Alopeke speckle
-imager on 2021 June 24.  `Alopeke is a dual-channel speckle
-interferometer that uses narrow-band filters centered at 0.83\,$\mu$m
-and $0.56\,\mu$m.  We acquired three sets of $1000\times 60$$\,$msec
-exposures during good seeing (0.45$''$), and used the autocorrelation
-function of these images to reconstruct a single image and 5-$\sigma$
-detection limits (see \citealt{howell_speckle_2011}).  This procedure
-yielded a detection of the companion in the 0.83\,$\mu$m notch filter,
-but not the $0.56\,\mu$m filter.  The measured projected separation
-and magnitude difference are given in Table~\ref{tab:starparams}.
-
-% \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
-Figure~\ref{fig:kep1627binary} summarizes the results of the
-high-resolution imaging.  The Gaia EDR3 parallax for the primary
-implies a projected separation of $53 \pm 4$\,AU, assuming the
-companion is bound.  Although the companion is unresolved in the Gaia
-source catalog (there are no comoving, codistant candidate companions
-brighter than $G < 20.5$ mag within $\rho < 120\arcsec$), its
-existence was also suggested by the primary star's large \replaced{renormalized
-unit weight error (RUWE),}{RUWE} relative to other members of the \cn\
-\added{(RUWE$\approx$2.9; roughly the $98^{\rm th}$ percentile of the cluster's
-distribution)}.  Based
-on the apparent separation, the binary orbital period is of order
-hundreds of years.  The large RUWE is therefore more likely to be
-caused by a PSF-mismatch skewing the Gaia centroiding during
-successive scans, rather than true astrometric motion.  Regardless,
-given the low geometric probability that a companion imaged at $\rho
-\approx 0\farcs16$ is a chance line-of-sight companion, we proceed
-under the assumption that the companion is bound, and that
-Kepler\,1627 is a binary.  Given the distance and age, the models of
-\citet{baraffe_new_2015} imply a companion mass of \replaced{$M_{\rm B} \approx
-0.33 M_{\odot}$}{$M_{\rm B} \approx 0.30\,M_{\odot}$} and companion
-temperature of \replaced{$T_{\rm eff,B} \approx 3450$}{$T_{\rm eff,B}
-\approx 3408$}\,K.  The corresponding spectral type is roughly \replaced{M2.5}{M3}V
-\citep{pecaut_mamajek_2013}.  These models combined with the NIRC2
-contrast limits imply physical limits on tertiary companions of
-$M_{\rm ter} < 50 M_{\rm Jup}$ at $\rho = 50$\,AU, $M_{\rm ter} < 20
-M_{\rm Jup}$ at $\rho = 100$\,AU, and $M_{\rm ter} < 10 M_{\rm Jup}$
-at $\rho = 330$\,AU.
-
-\section{The Planet}
-\label{sec:planet}
-
-\subsection{Kepler Light Curve}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=\textwidth]{f5a.pdf}
-		}
-
-		\vspace{-0.2cm}	
-		\subfloat{
-			\includegraphics[width=0.5\textwidth]{f5b.pdf}
-			\includegraphics[width=0.5\textwidth]{f5c.pdf}
-		}
-	\end{center}
-	\vspace{-0.7cm}
-  \caption{ {\bf The light curve of Kepler\,1627.}
-    {\it Top}: 
-    The Kepler data span 1{,}437 days (3.9 years), sampled at 30
-    minute cadence;  a 100 day segment is shown.  The top panel shows
-    the \texttt{PDCSAP} median-subtracted flux in units of
-    parts-per-thousand ($\times 10^{-3}$).  The dominant signal is
-    induced by starspots.  The stellar variability model (orange line)
-    is subtracted below, revealing the transits of \pn.  The online
-    Figure Set spans the entire 3.9 years of observations.
-    {\it Bottom}:
-    Phase-folded transit of Kepler 1627Ab with stellar variability
-    removed.  Windows over 20 hours ({\it left}) and the entire orbit
-    ({\it right}) are shown, and the residual after subtracting the
-    transit is in the bottom-most row.  The 2-$\sigma$ model uncertainties
-    and the best-fit model are the light purple band and the dark
-    purple line.  Gray points are individual flux measurements; black
-    points bin these to 15 minute intervals, and have a representative
-    1-$\sigma$ error bar in the lower right of each panel.  The asymmetric
-    residual during transit is larger than the out-of-transit scatter.
-    \label{fig:lc}
-  }
-\end{figure*}
-
-The Kepler space telescope observed \sn\ at a 30-minute cadence from
-2009 May 2 until 2013 April 8.  Data gaps during quarters 4, 9, and 13
-led to an average duty cycle over the 3.9~year interval of 67\%.  \sn\
-was also observed at 1-minute cadence from 2012 Oct 5 until 2013 Jan
-11.  The top panel of Figure~\ref{fig:lc} shows a portion of the
-30-minute cadence \texttt{PDCSAP} light curve.  Nonastrophysical
-variability has been removed using the methods discussed by
-\citet{smith_kepler_PDC_2017}; the default optimal aperture was
-assumed \citep{smith_finding_2016}.  Cadences with non-zero quality
-flags (9\% of the data) have been omitted.  The resulting photometry
-is dominated by a quasi-periodic starspot signal with a peak-to-peak
-amplitude that varies between 2\% and 8\%.  \added{Given that the secondary
-companion's brightness in the Kepler band is $\approx$1.5\% that of the primary,
-source confusion for the rotation signal is not expected to be
-an issue.}
-Previous analyses have
-identified and characterized the smaller transit signal
-\citep{2012ApJS..199...24T,thompson_planetary_2018}, validated its
-planetary nature \citep{morton_false_2016}, and even searched the
-system for transit timing variations \citep{holczer_transit_2016}.
-Nonetheless, since the cluster membership provides us with more
-precise stellar parameters than those previously available, we opted
-to reanalyze the light curve.
-
-\subsubsection{Transit and Stellar Variability Model}
-We fitted the Kepler long cadence time series with a model that
-simultaneously included the planetary transit and the stellar
-variability.  The stellar variability was modeled with the
-\texttt{RotationTerm} Gaussian Process kernel in \texttt{exoplanet}
-\citep{exoplanet:exoplanet}.  This kernel assumes that the variability
-is generated by a mixture of two damped simple harmonic oscillators
-with characteristic frequencies set by 1/$P_{\rm rot}$ and its first
-harmonic.  We additionally included a jitter term to inflate the flux
-uncertainties in a manner that accounted for otherwise unmodeled
-excess white noise, and let the eccentricity float.  For the
-limb-darkening, we assumed a quadratic law, and sampled using the
-uninformative prior suggested by \citet{exoplanet:kipping13}.
-
-Our model therefore included 10 free parameters for the transit ($\{P,
-t_0, \delta, b, u_1 ,u_2 ,R_\star, \log g, e, \omega \}$), 2 free
-parameters for the light curve normalization and a white noise jitter
-($\{\langle f \rangle, \sigma_f \}$), and 5 hyperparameters for the GP
-($\{\sigma_{\mathrm{rot}}, P_{\mathrm{rot}}, Q_0, \mathrm{d}Q, f \}$).
-We also considered including an additive \texttt{SHOTerm} kernel to
-account for stochastic noise, but found that this did not affect the
-results, and so opted for the simpler GP kernel.  We fitted the models
-using \texttt{PyMC3} \citep{salvatier_2016_PyMC3,exoplanet:theano},
-and accounted for the finite integration time of each exposure in the
-numerical integration when evaluating the model light curve
-\citep[see][]{kipping_binning_2010}.  We assumed a Gaussian
-likelihood, and after initializing each model with the parameters of
-the maximum {\it a posteriori} model, we sampled using
-\texttt{PyMC3}'s gradient-based No-U-Turn Sampler
-\citep{hoffman_no-u-turn_2014} in the bases indicated in
-Table~\ref{tab:posterior}.  We used $\hat{R}$ as our convergence
-diagnostic \citep{gelman_inference_1992}.
-
-Figure~\ref{fig:lc} shows the resulting best-fit model in orange (top)
-and purple (bottom).  The model parameters and their uncertainties,
-given in Table~\ref{tab:posterior}, are broadly consistent with a
-\replaced{mini-Neptune }{Neptune-}sized planet ($3.82\pm0.16\,R_\oplus$) on a close-in
-circular\footnote{ Our transit fitting yields $e<0.48$ at 2-$\sigma$;
-the constraints on the eccentricity are not particularly strong.}
-orbit around a G8V host star ($0.88 \pm 0.02 R_\odot$).  This best-fit
-planet size is consistent with those previously reported by
-\citet{morton_false_2016} and \citet{berger_2018_radii}, and
-corrects for the small amount of flux dilution from Kepler\,1627B.
-
-\subsubsection{Transit Asymmetry}
-\label{subsec:asymmetry}
-The transit fit however is not perfect: the lower panels of
-Figure~\ref{fig:lc} show an asymmetric residual in the data relative
-to the model: the measured flux is high during the first half of
-transit, and low in the second half.  The semi-amplitude of this
-deviation is $\approx50\,{\rm ppm}$, which represents a $\approx 3\%$
-distortion of the transit depth ($\delta=1759\pm62\,{\rm ppm}$).  Note that
-although this asymmetry is within the 2-$\sigma$ model uncertainties, the
-model has a jitter term that grows to account for excess
-white noise in the flux.  The significance of the asymmetry is
-therefore best assessed in comparison against the intrinsic
-out-of-transit scatter in the data ($\approx16\,{\rm ppm}$), not the model uncertainties.  The
-lower right panel of Figure~\ref{fig:lc} demonstrates that the scatter during
-transit is higher than during all other phases of the planet's orbit.
-
-To determine whether the asymmetry could be a systematic caused by our
-stellar variability model, we explored an alternative approach in
-which we isolated each transit window, locally fitted out
-polynomial trends, and then binned all the observed transits; the
-asymmetry was still present at a comparable amplitude.
-Appendix~\ref{app:asymmetry} describes a more detailed analysis, which finds
-that the asymmetry also seems to be robust to different methods of
-data binning in time and by local light curve slope.  Possible
-astrophysical explanations are discussed in Section~\ref{sec:conc}.
-
-\subsubsection{Transit Timing and the Local Slope}
-\label{subsec:ttvslope}
-
-\begin{figure}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f6.pdf}
-	\end{center}
-	\vspace{-0.6cm}
-	\caption{
-    {\bf \replaced{Possible}{Weak} evidence for a prograde orbit of Kepler 1627 Ab.} The time of
-    each Kepler transit was measured, along with the local slope of
-    the light curve.  The two quantities \replaced{are weakly}{might
-    be} anti-correlated
-    ($\approx$2-$\sigma$), which
-    \replaced{might}{could} be caused by starspot crossings during the first
-    (second) half of transit inducing a positive (negative) TTV,
-    provided that the orbit is prograde \citep{mazeh_time_2015}.  The
-    units along the abscissa can be understood by considering that the
-    stellar flux changes by $\sim$60\,ppt per half rotation period
-    ($\sim$1.3\,days).
-		\label{fig:corr}
-	}
-\end{figure}
-
-The previous analysis by \citet{holczer_transit_2016} did not
-find any significant long-term transit timing or duration variations
-(TTVs or TDVs) for \sn.  Quantitatively, the mean and standard deviation of the
-TTVs and TDVs they measured were $-1.1\pm13.8\,{\rm
-min}$ and $-3.3\pm22.1\,{\rm min}$.  In an
-earlier analysis however, \citet{holczer_time_2015} studied correlations
-between TTVs and local light curve slopes, and for \sn\
-found a weak correlation of $-29\pm13\,{\rm min}\,{\rm day}^{-1}$
-between the two quantities.  Given the possible connection between such
-correlations and the unresolved starspot crossings that we expect to
-be present in the \sn\ light curve \citep{mazeh_time_2015}, we opted to re-examine the
-individual transit times.
-
-We therefore isolated each of the 144 observed transits to within
-$\pm4.5$\,hr of each transit, and fitted each window with both {\it
-i)} a local \added{second or fourth-order }polynomial baseline plus the transit, and {\it ii)} a local linear
-trend\added{ plus the transit}.  \deleted{We considered the results both for a second and fourth-order
-time-dependence in the local baseline.  }We let the mid-time of each
-transit float, and then calculated the residual between the measured
-mid-time and that of a periodic orbit.  This residual, the transit
-timing variation, is plotted in Figure~\ref{fig:corr} against
-the local linear slope \replaced{for}{assuming} the fourth-order polynomial baseline.  The
-slope of $-21 \pm 10\,{\rm min\,day}^{-1}$ is similar to that
-found by \citet{holczer_time_2015}. \added{The
-best-fit line yields $\chi^2 = 306.1$, with $n=140$ data points.  An alternative
-model of just a flat line yields $\chi^2=315.6$.  The difference in the
-Bayesian information criterion between the two models is ${\rm
-BIC}_{\rm flat} - {\rm BIC}_{\rm line} = 4.5$,
-which corresponds to a Bayes factor of $\approx$9.4. According to the
-usual \citet{kass_bayes_1995} criteria, this is ``positive'' evidence
-for the model with a finite slope.  We view it as suggestive
-at best, particularly given the poor reduced $\chi^2$.}
-
-\replaced{One}{A separate} concern we had in this analysis was whether our transit fitting
-procedure might induce spurious correlations between the slope and
-transit time.  In particular, assuming the second-order polynomial
-baseline yielded a larger anti-correlation between the TTVs and local
-slopes, of $-79 \pm 14\,{\rm min\,day}^{-1}$.  We therefore
-performed an injection-recovery procedure in which we injected
-transits at different phases in the \sn\ light curve and repeated the
-TTV analysis.  This was done at $\approx$50 phases, each separated by
-$0.02\,P_{\rm orb}$ while omitting the phases in transit.  For
-the second-order polynomial baseline, this procedure yielded
-a similar anti-correlation in the injected transits as that present in the real
-transit;
-assuming this baseline would therefore bias the
-result.  However for the fourth-order baseline, the correlation
-present in the data was stronger than in all but one of the injected
-transits.
-Possible interpretations are discussed below. Given the
-\added{lack of }statistical significance, this analysis should be interpreted as
-suggestive at best.
-
-\subsection{Planet Confirmation}
-
-\begin{figure}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.47\textwidth]{f7.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Keck/HIRES radial velocities and MuSCAT3 photometry from the transit of 2021 Aug 7.}
-		{\it Top:} 
-		The radial velocity jitter across the 15~minute exposures
-		($\sigma_{\rm RV}\approx30$\ms) prevented us from detecting the RM
-		effect;
-		a model including the RM anomaly and a quadratic trend in time to
-		fit the spot-induced $\approx$\,$250$\ms\ trend is shown (see
-		Appendix~\ref{app:groundobs}).  Shaded bands show 2-$\sigma$ model
-		uncertainties.  
-		{\it Middle:} The RV variations are strongly correlated with varying
-		emission in the Ca H and K lines.
-		{\it Bottom}: The photometric transit depths are consistent across
-		the {\it griz} bandpasses. The photometry is binned at 10~minute
-		intervals.
-		\label{fig:ground}
-	}
-\end{figure}
-
-If the \pn\ transit signal is created by a genuine planet,  then to
-our knowledge it would be the youngest planet yet found by the
-\replaced{main}{prime} Kepler mission.\footnote{The re-purposed K2 mission however has found
-two younger systems containing five planets: K2-33b ($9\pm1\,{\rm
-Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and V1298 Tau
-($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}   Could the transit
-be produced by anything other than a planet orbiting this near-solar
-analog?  \citet{morton_false_2016} validated the planet based on the
-transit shape, arguing that the most probable false positive scenario
-was that of a background eclipsing binary, which had a model-dependent
-probability of $\approx$$10^{-5}$.  However, this calculation was
-performed without knowledge of the low-mass stellar companion ($M_{\rm
-B}\approx 0.30\,M_\odot$).  Validated planets have also previously
-been refuted \citep[{\it e.g.},][]{shporer_three_2017}.  We therefore
-reassessed false positive scenarios in some detail.
-
-% (0.7-2.0\% V-band to G-band dmag difference from EEM table)
-
-As an initial plausibility check, Kepler\,1627B contributes 1\% to 2\%
-of the total flux observed in the Kepler aperture.  For the sake of
-argument, assume the former value.  The observed transit has a depth
-of $\approx$0.18\%.  An $\approx$18\% deep eclipse of Kepler\,1627B would
-therefore be needed to produce a signal with the appropriate depth.
-The shape of the transit signal however requires the impact parameter
-to be below 0.77 (2-$\sigma$); the tertiary transiting the secondary
-would therefore need to be non-grazing with $R_3/R_2 \approx 0.4$.
-This yields a contradiction:  this scenario requires an ingress
-and egress phase that each span $\approx$40\% of the transit duration
-($\approx 68\,{\rm min}$).  The actual measured ingress and egress
-duration is $\approx$$17\,{\rm min}$, a factor of four times too
-short.  The combination of Kepler 1627B's brightness, the transit
-depth, and the ingress duration therefore disfavor the scenario that
-Kepler 1627B might host the transit signal.
-
-Beyond this simple test, a line of evidence that
-effectively confirms the planetary interpretation is that the stellar density
-implied by the transit duration and orbital period is inconsistent
-with an eclipsing body around the M-dwarf companion.  We find
-$\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
-theoretically expected density for Kepler\,1627B \added{given its
-nominal age of 38\,Myr and mass of 0.30\,$M_\odot$} is $\approx$$4.6\,{\rm
-g\,cm}^{-3}$ \citep{baraffe_new_2015}.
-The transit duration is therefore too long to be explained by a star
-eclipsing the M dwarf secondary at $10$-$\sigma$.\added{ Adversarially assuming a younger age (32\,Myr) and a larger companion
-mass (0.35\,$M_\odot$), the companion's density could be as low as $\approx 3.5\,{\rm
-g\,cm}^{-3}$.  This is still in tension with the fitted stellar density.} While the planet
-might hypothetically still orbit a hidden close and bright companion,
-this possibility is implausible given {\it i)} the lack of secondary
-lines in the HIRES spectra, {\it ii)} the lack of secondary
-rotation signals in the Kepler photometry, and {\it iii)} the proximity
-of \sn\ to the \cn\ locus on the Gaia CAMD (Figure~\ref{fig:age}).
-
-%NOTE to the arxiv readers:, the $\approx$100 days of short-cadence
-%Kepler observations also mitigate the cadence-induced smearing, and
-%enable a more precise measurement of the transit impact parameter.
-
-The correlation noted in Section~\ref{subsec:ttvslope} between the
-TTVs and the local light curve slope might be an additional line of evidence
-in support of the planetary interpretation.  Unless it is a
-statistical fluke (a $\approx$5\% possibility), then the most likely
-cause of the correlation is unresolved starspot
-crossings \citep{mazeh_time_2015}.  These would only be possible if the
-planet transits the primary star, which excludes a background
-eclipsing binary scenario.  The correlation would also suggest that the planet's
-orbit is prograde.  The latter point assumes that the dominant
-photometric variability is induced by dark spots, and not bright
-faculae.  Given the observed transition of Sun-like stellar
-variability from spot to faculae-dominated regimes between young and
-old ages, we expect this latter assumption to be reasonably secure
-\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
-
-A third supporting line of evidence for the planetary interpretation
-also exists.  We observed a transit of \pn\ on the
-night of 2021 Aug 7 \replaced{simultaneously with Keck/HIRES and
-MuSCAT3}{spectroscopically with HIRES at the Keck-I telescope and photometrically in
-$griz$ bands with MuSCAT3 at Haleakal\=a Observatory}.  \added{
-Details of the observation sequence are discussed in
-Appendix~\ref{app:groundobs}; Figure~\ref{fig:ground} shows the
-results.}\deleted{.  We
-scheduled the observations using the ephemeris of
-\citet{holczer_transit_2016}.}  Although we did not detect the
-Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
-show that the transit is achromatic\deleted{
-  (Figure~\ref{fig:ground})}.
-Quantitatively, when we fitted the MuSCAT3
-photometry with a model that lets the transit depths vary across each
-bandpass, we found {\it griz} depths consistent with the Kepler depth
-at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  \added{The
-achromatic transits strongly favor Kepler\ 1627A as the transit host,
-since Kepler\ 1627B is a much redder star.}  Conditioned on the
-ephemeris and transit depth from the Kepler data, the MuSCAT3 observations also
-suggested a transit duration $17.3\pm4.3\ {\rm min}$ shorter than the
-Kepler transits.
-However, given both the lack of TDVs in the Kepler data
-and the relatively low signal-to-noise of the MuSCAT3 transit,
-further photometric follow-up would be necessary to determine
-whether the transit duration is \replaced{indeed}{actually} changing.
-
-For our RM analysis,
-the details are discussed in Appendix~\ref{app:groundobs}.  While the
-velocities are marginally more consistent with a prograde or polar
-orbit than a retrograde orbit, the spot-corrected exposure-to-exposure
-scatter ($\sigma_{\rm RV}\approx 30$\ms) is comparable to the expected
-RM anomaly assuming an aligned orbit ($\Delta v_{\rm RM}\approx
-20\,{\rm m\,s}^{-1}$).  We are therefore not in a position to claim a
-spectroscopic detection of the RM effect, nor to quantify the stellar
-obliquity.
 
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f8.pdf}
+		\includegraphics[width=0.9\textwidth]{f4.pdf}
 	\end{center}
@@
 		3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
-		its uncertainty.  \pn\ is shown with a star.  The large sizes of
+		its uncertainty. STAR, SQUARE, DIAMOND.  The large sizes of
 		the youngest transiting planets could be explained by their
@@
 
-Kepler\,1627Ab provides a new extremum in the ages of the Kepler
-planets, and opens multiple avenues for further study.  Observations
-of spectroscopic transits at greater
-\replaced{precision}{signal-to-noise} should yield a
-measurement of the stellar obliquity, which would confirm or refute
-the prograde orbital geometry suggested by the TTV-local slope
-correlation.  Separately, transit spectroscopy aimed at detecting
-atmospheric outflows could yield insight into the evolutionary state
-of the atmosphere \citep[{\it
-e.g.},][]{ehrenreich_giant_2015,spake_helium_2018,vissapragada_2020}.
-Observations that quantify the amount of high-energy
-irradiation incident on the planet would complement these efforts, by
-helping to clarify the expected outflow rate \citep[{\it
-e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
-quantity to measure would be the planet's mass.  Measured at
-sufficient precision,
-for instance through a multi-wavelength radial velocity campaign,
-the combination of the size, mass, and age 
-would yield constraints on both the planet's composition and its
-initial entropy \citep{owen_constraining_2020}.
-
-More immediately, the Kepler data may yet contain additional
-information.  For instance, one possible explanation for the transit
-asymmetry shown in Figure~\ref{fig:lc} is that of a dusty asymmetric
-outflow.  
-Dusty outflows are theoretically
-expected for young mini-Neptunes, and the amplitude of the observed
-asymmetry is consistent with predictions \citep{wang_dai_2019}.
-A second possibility is that the planetary orbit is slightly
-misaligned from the stellar spin axis, and tends to transit starspot
-groups at favored stellar latitudes.  
-This geometry would be necessary in order to
-explain how the starspot crossings could add up coherently\added{,
-given that the planetary orbital period (7.203\,days) and the stellar
-rotation period (2.642\,days) are not a rational combination}.
-Other possibilities including
-gravity darkening or TTVs causing the asymmetry are disfavored (see
-Appendix~\ref{app:asymmetry}).
-
-Beyond the asymmetric transits, Appendix~\ref{app:flare}
-highlights an additional abnormality in the short-cadence Kepler data, in the
-arrival time distribution of stellar flares.  We encourage its
-exploration by investigators more versed in the topic than ourselves.
-
-In the context of the transiting planet population, \pn\ is among the
-youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
-with precise ages include K2-33
-\citep{Mann_K2_33b_2016,David_et_al_2017}, DS~Tuc
-\citep{benatti_possibly_2019,newton_tess_2019}, HIP~67522
-\citep{rizzuto_tess_2020}, TOI~837 \citep{bouma_cluster_2020}, the
-two-planet AU~Mic \citep{plavchan_planet_2020,martioli_aumicbc_2021}
-and the four-planet V1298~Tau \citep{david_four_2019}.  \pn\ is one of
-the smaller planets in this sample ($3.82\pm0.16\,R_\oplus$), which
-could be linked to the selection effects imposed by spot-induced
-photometric variability at very young ages \citep[{\it
-e.g.},][]{zhou_2021_tois}.
-
- \replaced{However,}{It seems though that} smaller planets could have been
-detected\added{ in the \sn\ system}: \added{based on the per-target detection contours, }the Kepler pipeline's median completeness extended to
-$1.6\,R_\oplus$ at 10~day orbital periods, and $3.3\,R_\oplus$ at
-100~days
-\citep{2021ascl.soft07027B}.
-\added{These limits account for the spot-induced variability in the
-system through
-a correction based on the Combined Differential Photometric Precision
-in the light curve over the relevant transit timescales
-\citep{2017ksci.rept...19B}.} The large size of \pn\ relative
-to most Kepler mini-Neptunes might therefore support a picture
-in which the typical $5\,M_\oplus$\deleted{ to 10$\,M_\oplus$} mini-Neptune
-\citep{wu_mass_2019} loses a significant fraction of its primordial
-atmosphere over its first gigayear
-\citep{Owen_Wu_2013,ginzburg_corepowered_2018}.  
-\added{
-It could also be consistent with a scenario in which an earlier ``boil-off'' of
-the planet's atmosphere during disk dispersal decreases the entropy of
-the planetary interior, leading to a rather long $\sim$10$^8$ year Kelvin-Helmholtz
-contraction timescale
-\citep{owen_constraining_2020}.
-Confirming either of these scenarios would require a measurement
-of the planetary mass; otherwise, alternative explanations for its large size
-could include that it is just abnormally massive, or that it has an abnormally
-large envelope to core mass ratio.}
-
-% The prograde orbit, if confirmed, would be consistent with a
-% quiescent disk-driven migration history, and would differ from
-% misalignments that have been observed for some older hot
-% Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
-%Spectroscopic follow-up observations of the transit remain necessary
-%however to verify whether the TTV-slope correlation is yielding the
-%correct orbital geometry.  The alternative scenario of in-situ
-%formation is disfavored on theoretical grounds: the required disk
-%mass would likely drive migration, and so lacks internal consistency
-%\citep{inamdar_formation_2015,ogihara_reassessment_2015}.
-
-Ultimately, the main advance of this work is a precise measurement of
-the age of \pn.   This measurement was enabled by identifying the
-connection of the star to the \cn\ using Gaia kinematics, and by then 
-using the Gaia color-absolute magnitude diagram and TESS stellar
-rotation periods to verify the cluster's existence.
-Table~\ref{tab:v06} enables similar cross-matches for both known and
-forthcoming exoplanet systems \citep[{\it
-e.g.},][]{guerrero_tess_2021}. Confirming these candidate associations
-using independent age indicators is essential because their false
-positive rates are not known.  A related path is to identify new
-kinematic associations around known exoplanet host stars using
-positions and tangential velocities from Gaia, and to verify
-these associations with stellar rotation periods and spectroscopy
-\citep[{\it e.g.},][]{tofflemire_tess_2021}.  Each approach seems
-likely to expand the census of planets with precisely measured ages
-over the coming years, which will help in deciphering the early stages
-of exoplanet evolution.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \raggedbottom
-
-The authors are grateful to J{.}~Winn, J{.}~Spake, A{.}~Howard, and
-T{.}~David for illuminating discussions and suggestions, and to
-R{.}~Kerr for providing us with the \citet{Kerr2021} membership list
-prior to its publication.  The authors are also grateful to
-K{.}~Collins for helping resolve the scheduling conflict that would
-have otherwise prevented the MuSCAT3 observations.
 %
-L.G.B{.} acknowledges support from a Charlotte Elizabeth Procter
-Fellowship from Princeton University, as well as from the TESS GI
-Program (NASA grants 80NSSC19K0386 and 80NSSC19K1728) and the
-Heising-Simons Foundation (51 Pegasi~b Fellowship).
+L.G.B{.} acknowledges support from the TESS GI Program (NASA grants
+80NSSC19K0386 and 80NSSC19K1728) and the Heising-Simons Foundation (51 Pegasi~b
+Fellowship).
 % programs G011103 and G022117, through 
@@
 %
+%FIXME
 Keck/NIRC2 imaging was acquired by program 2015A/N301N2L
 (PI: A.~Kraus). % and 2019A/N069 (PI: E.~Petigura).
-%
-In addition, this paper is based in part on observations made with the
-MuSCAT3 instrument, developed by the Astrobiology Center and under
-financial support by JSPS KAKENHI (JP18H05439) and JST PRESTO
-(JPMJPR1775), at Faulkes Telescope North on Maui, HI, operated by the
-Las Cumbres Observatory.
-%
-This work is partly supported by JSPS KAKENHI Grant Numbers 22000005,
-JP15H02063, JP17H04574, JP18H05439, JP18H05442, JST PRESTO Grant
-Number JPMJPR1775, the Astrobiology Center of National Institutes of
-Natural Sciences (NINS) (Grant Number AB031010).
 %
@@
 %Aeronautics and Space Administration.
-Finally, we also thank the Keck
-Observatory staff for their support of HIRES and remote observing.  We
-recognize the importance that the summit of Maunakea has within the
-indigenous Hawaiian community, and are deeply grateful to have the
-opportunity to conduct observations from this mountain.
+Finally, we also thank the Keck Observatory staff for their support of
+HIRES and remote observing.  We recognize the importance that the
+summit of Maunakea has within the indigenous Hawaiian community, and
+are deeply grateful to have the opportunity to conduct observations
+from this mountain.
 %
@@
   %\texttt{arviz} \citep{arviz_2019},
-  \texttt{altaipony} \citep{ilin_flares_2021},
+  %\texttt{altaipony} \citep{ilin_flares_2021},
   \texttt{astrobase} \citep{bhatti_astrobase_2018},
@@
   \texttt{scipy} \citep{jones_scipy_2001},
-  \texttt{TESS-point}  \citep{burke_2020},
+  %\texttt{TESS-point}  \citep{burke_2020},
   %\texttt{tesscut} \citep{brasseur_astrocut_2019},
@@
   %\texttt{webplotdigitzer} \citep{rohatgi_2019},
-  \texttt{wotan} \citep{hippke_wotan_2019}.
+  %\texttt{wotan} \citep{hippke_wotan_2019}.
 }
@@
  	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
- 	Gemini:North~(`Alopeke; \citealt{scott_nessi_2018,scott_twin_2021}.
+ 	%Gemini:North~(`Alopeke; \citealt{scott_nessi_2018,scott_twin_2021}.
  	{\it Spectroscopy}:
@@
   %PFS ({\bf CITE}),
-  %  MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
+	Tillinghast:1.5m~(TRES; \citealt{furesz_tres_2008}).
+  %MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
 	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
@@
 % 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
-    MuSCAT3 \citep{Narita_2020},
+%   MuSCAT3 \citep{Narita_2020},
 % 	NTT,
@@
 
-\input{starparams.tex}
-\input{model_posterior_table.tex}
-\input{t1_appendix.tex}
-\input{t2_appendix.tex}
+% \input{starparams.tex}
+% \input{model_posterior_table.tex}
+% \input{t1_appendix.tex}
+% \input{t2_appendix.tex}
 
@@
 
-\appendix
-\section{Young, Age-Dated, and Age-Dateable Star Compilation}
-\label{app:targetlist}
-
-The \texttt{v0.6} CDIPS target catalog (Table~\ref{tab:v06}) includes
-stars that are young, age-dated, and age-dateable.  By
-``age-dateable'', we mean that the stellar age should be measurable at
-greater precision than that of a typical FGK field star, through
-either isochronal, gyrochronal, or spectroscopic techniques.  As in
-\citet{bouma_cdipsI_2019}, we collected stars that met these criteria
-from across the literature.  Table~\ref{tab:metadata} gives a list of
-the studies included, and brief summary statistics.  The age
-measurement methodologies adopted by each study differ: in many,
-spatial and kinematic clustering has been performed on the Gaia data,
-and ensemble isochrone fitting of the resulting clusters has been
-performed (typically focusing on the turn-off).  In other studies
-however, the claim of youth is based on the location of a single star
-in the color-absolute magnitude diagram, or on spectroscopic
-information.
-
-One major change in Table~\ref{tab:v06} relative to the earlier
-iteration from \citet{bouma_cdipsI_2019} is that the extent of
-Gaia-based analyses has now matured to the point that we can neglect
-pre-Gaia cluster memberships, except for a few cases with
-spectroscopically confirmed samples of age-dated stars.  The
-membership lists for instance of \citet{Kharchenko_et_al_2013} and
-\citet{dias_proper_2014} (MWSC and DAML) are no longer required.  This
-is helpful for various post-processing projects,  since the field star
-contamination rates were typically much higher in these catalogs than
-in the newer Gaia-based catalogs.
-
-The most crucial parameters of a given star for our purposes are the
-Gaia DR2 $\texttt{source\_id}$, the cluster or group name
-($\texttt{cluster}$), and the $\texttt{age}$.  Given the hierarchical
-nature of many stellar associations, we do not attempt to resolve the
-cluster names to a single unique string.  The Orion complex for
-instance can be divided into almost one hundred kinematic subgroups
-\citep{kounkel_apogee2_2018}.  \added{Based on Figure~\ref{fig:XYZvtang}, the
-\cn\ may also be part of a similar hierarchical association.} Similar complexity applies to the
-problem of determining homogeneous ages, which we do not attempt to
-resolve.  Instead, we simply merged the cluster names and ages
-reported by various authors into a comma-separated string.
-
-This means that the \texttt{age} column can be null, for cases in
-which the original authors did not report an age, or for which a
-reference literature age was not readily available.  Nonetheless,
-since we do prefer stars with known ages, we made a few additional
-efforts to populate this column.  When available, the age provenance
-is from the original analysis of the cluster.  In a few cases however
-we adopted other ages when string-based cross-matching on the cluster
-name was straightforward.  In particular, we used the ages determined
-by \citet{CantatGaudin2020b} to assign ages to the clusters from
-\citet{GaiaCollaboration2018}, \citet{CantatGaudin2018a},
-\citet{CastroGinard2020}, and \citet{CantatGaudin2020a}.
-
-The catalogs we included for which ages were not immediately available
-were those of \citet{CottenSong2016}, \citet{Oh2017},
-\citet{Zari2018}, \citet{Gagne2018b}, \citet{Gagne2018a},
-\citet{Gagne2018c}, and \citet{Ujjwal2020}.  While in principle the
-moving group members discussed by
-\citet{Gagne2018b,Gagne2018a,Gagne2018c} and \citet{Ujjwal2020} have
-easily associated ages, our SIMBAD cross-match did not retain the
-moving group identifiers given by those studies, which should
-therefore be recovered using tools such as BANYAN
-$\Sigma$ \citep{Gagne2018a}.
-We also included the SIMBAD object identifiers $\texttt{TT*}$,
-$\texttt{Y*O}, $\texttt{Y*?}, $\texttt{TT?}$, and $\texttt{pMS*}$.
-Finally, we  included every star in the NASA Exoplanet Archive
-planetary system ($\texttt{ps}$) table that had a Gaia identifier
-available \citep{NASAExoArchive_ps_20210506}.  If the age had finite
-uncertainties, we also included it, since stellar ages determined
-through the combination of isochrone-fitting and transit-derived
-stellar densities typically have higher precision than from isochrones
-alone.
-
-For any of the catalogs for which Gaia DR2 identifiers were not
-available, we either followed the spatial (plus proper-motion)
-cross-matching procedures described in \citet{bouma_cdipsI_2019}, or
-else we pulled the Gaia DR2 source identifiers associated with the
-catalog from SIMBAD.  We consequently opted to drop the
-$\texttt{ext\_catalog\_name}$ and $\texttt{dist}$ columns maintained
-in \citet{bouma_cdipsI_2019}, as these were only populated for a small
-number of stars.  The technical manipulations for the merging,
-cleaning, and joining were performed using $\texttt{pandas}$
-\citep{mckinney-proc-scipy-2010}.  The eventual cross-match (using the
-Gaia DR2 $\texttt{source\_id}$) against the Gaia DR2 archive was
-performed asychronously on the Gaia archive
-website.
-
-% \section{Binarity and Stellar Rotation for $\delta$\,Lyr Cluster Members}
-% \label{app:rotationbinarity}
-% 
-% \begin{figure}[t]
-% 	\begin{center}
-% 		\leavevmode
-% 		\includegraphics[width=0.7\textwidth]{f7.pdf}
-% 	\end{center}
-% 	\vspace{-0.7cm}
-%   \caption{ {\bf Binarity indicators for rotators in \cn.} 
-%     Data are as in the lower panel of Figure~\ref{fig:age}.  Sources
-%     with $\mathrm{RUWE}>1.2$ appear in red; these include astrometric binaries as well
-%     as sources poorly fit by a single-star PSF.  Photometric binaries
-%     ($>$0.3\,mag above an empirical isochrone; 20\% of stars) are
-%     orange.  Over $0.6<\bpmrpo<1.5$, the detected rotation periods in
-%     binaries tend to fall below the slow sequence.
-% 		\label{fig:binarity}
-% 	}
-% \end{figure}
-% 
-% Figure~\ref{fig:binarity} shows the rotation-color diagram for \cn\
-% members, with the points colored according to indicators of binarity.
-% The binaries tend to be redder and have shorter rotation periods.  We
-% know of two possible explanations based on selection effects, and two
-% based on physics.
-% 
-% Possible selection effects include {\it i)} binaries have a component
-% that contributes additional red light to the system, which could skew
-% the color measurement of the primary; and {\it ii)} the unresolved
-% binary companions could contaminate the rotation period measurement
-% \citep[{\it e.g.},][Section~5.1]{stauffer_rotation_2016}.
-% 
-% The two possible physical effects that could be relevant are tidal
-% locking and pre-main-sequence disk locking.  Tidal locking has been
-% argued to be an unlikely explanation for the frequency of rapid
-% rotators due to the population statistics; a more likely scenario
-% would be that the presence of the binary leads to faster disk
-% dispersal, enabling the primary to contract to more rapid rotation
-% periods than possible for single stars
-% \citep{meibom_effect_2007,bouma_cluster_2020}.
-
-\section{The Transit Asymmetry}
-\label{app:asymmetry}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.64\textwidth]{f9.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Transit model residuals through time (binned by Kepler quarter)}.  
-    {\it Left:}
-    Phase-folded transit of Kepler 1627b, with stellar variability
-    removed.  Black points are binned to 20 minute intervals.  The
-    2-$\sigma$ model uncertainties and the maximum {\it a posteriori}
-    model are shown as the faint purple band, and the dark purple
-    line.
-    {\it Right:}
-    As on the left, with the transit removed.  
-		\label{fig:phasequarter}
-	}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.64\textwidth]{f10.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Transit model residuals through time (binned by year of observation)}.  
-    {\it Left:}
-    Phase-folded transit of Kepler 1627b, with stellar variability
-    removed.  Points and models are as in
-    Figure~\ref{fig:phasequarter}.
-    {\it Right:}
-    As on the left, with the transit removed.
-		\label{fig:phaseyear}
-	}
-\end{figure}
-
-\begin{figure}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f11.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    {\bf Transit models and residuals, binned by quartiles in the
-    local slope of the light curve}.  
-    Representative uncertainties for the black points (binned at 20
-    minute intervals) are shown in the lower right of each panel.  A
-    similar transit asymmetry to that shown in Figure~\ref{fig:lc}
-    seems to be present in three of the four bins.
-		\label{fig:phaseslope}
-	}
-\end{figure}
-
-\subsection{How Robust is the Asymmetric Transit?}
-
-As a means of exploring the robustness of the transit asymmetry,
-Figures~\ref{fig:phasequarter},~\ref{fig:phaseyear}, and~\ref{fig:phaseslope}
-show the Kepler data binned in three ways: over
-Kepler quarters, Julian years, and quartiles of local slope.  Over Kepler quarters
-(Figure~\ref{fig:phasequarter}), Quarter 6 shows the strongest
-asymmetry: a deviation of about 3\,ppt from expectation.  Quarter 7
-shows an anomaly at roughly the same transit phase.  Year 2
-correspondingly shows the strongest anomaly out of any year in
-Figure~\ref{fig:phaseyear}; the asymmetry is visually apparent however
-in each of the four years.
-
-To bin by quartiles in local slope, we used our measurements
-of the local linear slopes in each of the observed transit windows
-(144 transits total).  Four outlier transits were removed, leaving 140
-transits.  These were then divided into quartiles, so that each panel
-shows 35 transits binned together.  The exact light curve slope intervals are listed in the lower left panels of
-Figure~\ref{fig:phaseslope}.  Binned by local
-slope quartiles (Figure~\ref{fig:phaseslope}), the asymmetry is
-visually present in three of the four quartiles: the only bin in which
-it does not appear is ${\rm d}f/{\rm d}t\in[3.0,32.5]\,{\rm
-ppt\,day}^{-1}$.
-
-Within the theory presented by \citet{mazeh_time_2015}, unresolved
-starspot crossings cause the weak correlation between TTVs and the local
-light curve slope (Figure~\ref{fig:corr}).  In this model, we would
-expect the light curves with the most negative local slopes to have
-the largest positive TTVs, due to spot crossing events during the
-latter half of transit.  The upper-left panel of
-Figure~\ref{fig:phaseslope} agrees with this expectation.  However, we
-would also expect the sign of the effect to reverse when considering
-the most positive local slopes (most negative TTVs).  The
-lower-right panel of
-Figure~\ref{fig:phaseslope} contradicts this expectation: the residual
-in both cases maintains the same parity!  On the one hand, this shows
-that the residual is not dependent on the local light curve slope,
-which lowers the likelihood that it might be an artifact of our
-detrending methods.  On the other, it raises the question of whether
-unresolved starspot crossings are indeed the root cause of the
-correlation shown in Figure~\ref{fig:corr}.
-While we do not have a solution to this contradiction, 
-the injection-recovery tests discussed in
-Section~\ref{subsec:ttvslope} provide some assurance that the
-TTV-slope correlation is not simply a systematic artifact.
-
-\subsection{Interpretation}
-
-The transit asymmetry seems robust against most methods of binning the
-data, though with some caveats ({\it e.g.}, the ``middle quartile'' in
-local flux, ${\rm d}f/{\rm d}t\in[3.0,32.5]\,{\rm ppt\,day}^{-1}$,
-where the asymmetry does not appear).  Nonetheless, if the asymmetric were
-systematic we might expect its parity to reverse as a function of
-the sign of the local slope, and it does not.  We therefore entertained four
-possible astrophysical explanations: gravity darkening, transit
-timing variations, spot-crossing events, and a persistent asymmetric
-dusty outflow.
-
-Gravity darkening is based on the premise that the rapidly rotating
-star is oblate, and brighter near the poles than the equator
-\citep[{\it e.g.},][]{masuda_spin-orbit_2015}.  The fractional transit
-shape change due to gravity darkening is on the order of $(P_{\rm
-break}/P_{\rm rot})^2$, for $P_{\rm break}$ the break-up rotation
-period, and $P_{\rm rot}$ the rotation period.  Using the parameters
-from Table~\ref{tab:posterior}, this yields an expected 0.14\%
-distortion of the $\approx$1.8\,ppt transit depth: {\it i.e.}, an
-absolute deviation of $\approx$2.5\,ppm.  The observed residual has a
-semi-amplitude of $\approx 50\,{\rm ppm}$.  Since the expected signal
-is smaller than the observed anomaly by over an order of magnitude,
-gravity darkening seems to be an unlikely explanation.
-
-The scenario of transit timing variations (TTVs) producing the
-asymmetry seems unlikely because the transit
-timing variations we do observe are correlated with the local light
-curve slope, which increases roughly as much as it decreases.  From
-our analysis, the mean TTV and its standard deviation are
-$0.66\pm5.53$\,min; similarly the mean local slope and its
-standard deviation are $0.59\pm45.50$\,ppt\,day$^{-1}$.  There is
-therefore little expectation for TTVs to produce the asymmetry.  A
-separate line of argument comes from Figure~\ref{fig:phaseslope}.  If
-the local slope were essential to producing the transit asymmetry, we
-would expect that in the largest ${\rm d}f/{\rm d}t$ bin, ${\rm
-d}f/{\rm d}t\in[3.0,32.5]\,{\rm ppt\,day}^{-1}$, the sign of the
-asymmetry would reverse.  We do not see evidence for this being the
-case.
-
-The third and related possibility is that of starspot crossings.
-Young stars have higher spot-covering fractions than old stars ({\it
-e.g.}, \citealt{morris_relationship_2020}).  Young solar-type stars
-may also host dark starspots at high stellar latitudes \citep[{\it
-e.g.}, EK~Dra;][]{strassmeier_starspots_2009}, though interferometric
-imaging of spotted giant stars has shown different starspot latitude
-distributions than those inferred from Doppler imaging
-\citep{roettenbacher_contemporaneous_2017}.  Regardless, for any
-spot-crossing anomalies to add coherently over the 140 Kepler
-transits, it seems likely that we would need either for spots to be
-persistent at a particular latitude (and for the planetary orbit to be
-somewhat misaligned), or for a ``stroboscopic'' longitudinal phasing
-\citep[{\it e.g.},][]{dai_stellar_2018}.  For our system, $P_{\rm
-orb}/P_{\rm rot} \approx 2.76$, which means that every 4 transits and
-11 stellar rotations, the planet crosses over roughly the same stellar
-longitude, which might enable the necessary phasing if the spot-groups
-are large and long-lived.  Unfortunately, the  S/N per Kepler transit
-is $\approx8$, which renders individual spot-crossing events
-unresolved.  This explanation seems marginally plausible, mainly
-because the expected spot-crossing anomaly amplitudes ($\approx
-100\,{\rm ppm}$) resemble the observed amplitude of the asymmetry
-($\approx 50\,{\rm ppm}$).  One issue with this explanation however is
-that there is no reason to expect starspot crossing events to last
-exactly half the transit duration.
-
-A persistent feature of the planet itself might therefore be needed to
-explain the transit asymmetry.  An asymmetric outflow from the
-planet's atmosphere could at least geometrically meet the requirements
-\citep[{\it e.g.},][]{mccann_2019}.  To explain the asymmetric
-transit, a small, dense component would lead the planet, and a long,
-more rarefied (and variable) component would trail it.  This might
-also explain the slight flux decrement visible for $\sim$1 hour
-pre-ingress (Figure~\ref{fig:lc}).  The amplitude of the asymmetry is
-roughly in line with theoretical expectations for dusty outflows
-\citep{wang_dai_2019}, and based on the planet's size, its mass is
-likely in a regime where such outflows are possible.  Out of the four
-explanations discussed, this one at least theoretically seems the most
-plausible.  By composition, the expectation would be that the envelope
-is mostly hydrogen and helium gas, with a dust or haze component
-providing the broadband opacity in the Kepler bandpass.  A natural
-path for testing this idea would be to observe
-additional transits of the planet in hydrogen absorption, metastable
-helium absorption, or across a broad wavelength range in the
-near-infrared.
-
-%  \subsection{Short Cadence Data}
-%  
-%  \begin{figure*}[t]
-%  	\begin{center}
-%  		\leavevmode
-%  		\includegraphics[width=1\textwidth]{f12.pdf}
-%  	\end{center}
-%  	\vspace{-0.7cm}
-%  	\caption{
-%  		{\bf Phase-folded short cadence transit of Kepler\,1627 Ab}.  
-%      The precision with which the impact parameter can be measured is
-%      higher than from the long cadence data.
-%  		\label{fig:phase_slc}
-%  	}
-%  \end{figure*}
-%  
-%  Figure~\ref{fig:phase_slc} shows the result from analyzing 100 days of
-%  short cadence (1-minute) data acquired during Quarter 15.
-%  %TODO: get quantitative on better impact parameter...
-
-\section{Spectroscopy and Photometry During the Transit of 2021 Aug 7}
-\label{app:groundobs}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=\textwidth]{f12.pdf}
-		}
-	\end{center}
-	\vspace{-0.5cm}
-	\caption{
-    {\bf Spectroscopic activity indicators during the transit of 2021
-    Aug 7.} The {\it top panels} show the median line profiles Ca K,
-    Ca H, and H$\alpha$ line profiles from the HIRES spectra.  The
-    {\it lower panels} show the differences of each individual
-    spectrum relative to the median spectrum.  The bump in the red
-    wing of Ca H is H$\epsilon$.  The spectra in the lower panels are
-    smoothed for visualization purposes.
-    \label{fig:rvactivity}
-	}
-\end{figure*}
-
-\added{We used the ephemeris of \citet{holczer_transit_2016} to
-observe a transit of \pn\ on the night of 2021 Aug 7 both
-spectroscopically and photometrically.  We used the HIRES echelle
-spectrograph at the Keck-I telescope and the MuSCAT3 photometer at Haleakal\=a
-Observatory on Maui, HI \citep{Narita_2020}.} \deleted{ We monitored
-\sn\ with Keck/HIRES before, during, and after transit on the night of
-2021 Aug 7.} \replaced{We used the iodine cell for wavelength
-calibration}{For the HIRES wavelength calibration, we used the iodine
-cell}, and extracted the 1-D spectra using the standard California
-Planet Survey pipeline \citep{howard_cps_2010}.  \added{Given the
-faintness of the target ($V=13.1$), we observed using the C2 decker,
-which yielded an instrument resolution of $\approx$50{,}000.} The
-airmass ranged between 1.1 and 2.2 from the start through the end of
-observations; the seeing ranged from $1\farcs1$ at the beginning to
-$1\farcs5$ at the end.  \added{The HIRES exposure time was set at
-$\approx$15 minutes in order to resolve the 2.8 hour transit event,
-which yielded a S/N per resolution element of $\approx 75$ (low for
-precision radial velocity standards).} \replaced{We also
-simultaneously observed across {\it griz} bands using MuSCAT3 at
-Haleakal\=a Observatory on Maui, HI.}{For the MuSCAT3 observations, we
-observed simultaneously across the {\it griz} bands.} \added{The
-exposure times in each bandpass ranged between 23 and 46 seconds, and
-were chosen in order to yield a S/N in the peak pixel that exceeded
-130 while also preventing saturation.} Performing aperture photometry
-on the latter image stack yielded the data given in
-Table~\ref{tab:phot}.
-
-We considered two approaches to measuring the velocities: in the
-first, hereafter ``Method 1'', we cross-correlated against a template
-found via spectral classification with \texttt{SpecMatch-Emp}
-\citep{yee_SM_2017}.  In ``Method 2'', we used a high S/N template of
-V1298~Tau.  Although V1298 Tau is cooler than Kepler~1627A by
-$\approx$500\,K, it has a comparable amount of line-broadening ($v\sin
-i = 23$\kms), and a comparable level of stellar activity.  The mean
-and standard deviation of the internal RV uncertainties averaged over
-all epochs were $16.2\pm1.1$\ms\ from Method 1, and $12.6\pm0.6$\ms\
-from Method 2.  The corresponding time-averaged reduced $\chi^2$ from
-the template match was $1.57\pm0.04$ (Method 1) and $1.30\pm0.02$
-(Method 2).  Given these diagnostics, we adopted the velocities from
-the second approach, which are reported in Table~\ref{tab:rv}.
-
-Figure~\ref{fig:ground} shows the results.  The MuSCAT3 photometry
-shows the expected starspot trend, along with the transit and what is
-likely a chromatic starspot crossing event at ${\rm JD} - 2459433 =
-0.955$.  The radial velocities decrease by $\approx$250\ms\ over the
-six hour window.  This decrease in RV is correlated with a decrease in
-the S-indices derived from the Ca HK lines.  One outlying RV point is
-apparent shortly before egress; it is temporally coincident with an
-outlying value in the S-index time series.
-
-Overall, we expect the dominant trends in both the photometry and
-radial velocities to be caused by starspots on the stellar photosphere
-rotating into and out of view.  The plasma in the leading and receding
-limbs of the stellar disk has an apparent line-of-sight velocity of
-$\pm 20$\kms.  Over 10\% of a rotation cycle ($P_{\rm rot}=2.6\,{\rm
-days}$), spots near these limbs come into and out of view, modulate
-the stellar velocity profile, and can thereby produce the overall
-$\approx$250\ms\ trend.  The Ca HK and H$\alpha$ emission profiles
-support this interpretation; Figure~\ref{fig:rvactivity} shows that
-each line gradually decreases in intensity over the course of the six
-hour sequence.
-
-The expectation however is for the starspot-induced signals to be
-smooth, at worst with contributions at $0.5\,P_{\rm rot}$ or
-$0.25\,P_{\rm rot}$ \citep{klein_simulated_2020}.  We therefore fitted
-the RVs using the \citet{hirano_analytic_2010,hirano_2011} models for
-the Rossiter-McLaughlin (RM) effect, and allowed for an optional
-linear and quadratic trend in time to fit the $\approx$250\ms\
-spot-induced trend.  We followed the methodology developed by
-\citet{stefansson_2020}.  We allowed the sky-projected obliquity, the
-projected stellar equatorial velocity, and the Gaussian dispersion of
-the spectral lines to vary, and fixed the limb-darkening using the
-$V$-band tabulation from \citet{claret_gravity_2011}.  We assumed a
-Gaussian prior on $v\sin i$ and $a/R_\star$ from
-Table~\ref{tab:starparams}, and also allowed for a white-noise jitter
-term to be added in quadrature to the measurement uncertainties.  We
-used a 15\,minute exposure time to numerically evaluate the model.
-
-The quadratic model with the RM effect is shown in
-Figure~\ref{fig:ground}; the jitter term is incorporated in the model
-uncertainties, but not the plotted measurement uncertainties.  The
-plotted measurement uncertainties are the internal uncertainties on
-the RVs ($\approx 13\,{\rm m\,s}^{-1}$), and are dominated by the
-$v\sin i$ broadening.  However, between exposures, the RVs show
-significant additional scatter that is not captured by the slow
-quadratic trend.  The white-noise jitter for this particular model is
-$\sigma_{\rm RV} = 27^{+6}_{-5}$\ms, which is comparable to the
-expected RM anomaly of $\Delta v_{\rm RM} \approx f_{\rm LD} \cdot
-\delta \cdot v\sin i \cdot \sqrt{1-b^2} \approx 20\,{\rm m\,s}^{-1}$,
-assuming a perfectly aligned orbit.
-
-The presence of this additional scatter prevents a convincing
-detection of the RM effect.  The reason can be understood via model
-comparison.  If we compare the model with a quadratic trend and the RM
-effect against a model with a linear trend and the RM effect, or even
-a model with no RM effect at all, then the respective Bayesian
-Information Criterion (BIC) values are as follows.
-% from rm_TEMPLATE_V1298TAU_quadraticfreejitter_v0.log and
-% rm_TEMPLATE_V1298TAU_v1_linearfreejitter_trendonlyfreejitter.log
-\begin{align}
-  {\rm BIC} &= 227.1  \quad ({\rm Quadratic+RM}) \nonumber \\
-  {\rm BIC} &= 231.1  \quad ({\rm Linear+RM}) \nonumber \\
-  {\rm BIC} &= 221.4  \quad ({\rm Only\ Quadratic}).
-\end{align}
-There is therefore no evidence to prefer the model with the RM effect
-against a model that only accounts for the stellar variability.  The
-``only quadratic'' model does particularly well because it can inflate
-the jitter term to account for scatter during the transit (even if the
-scatter contains astrophysics!), and it has fewer free parameters.
-However, we cannot justify a physical prior on the jitter term,
-because we do not understand the origin of the exposure-to-exposure
-scatter.  As noted above, the velocity deviations from starspots are
-expected to have contributions at the stellar rotation frequency, or
-harmonics thereof.  This jitter is present on the exposure timescale
-(15 minutes), which is only 0.4\% of the stellar rotation period; it
-is not obvious that starspots would be the culprit.
-
-The amplitude of both the spot-induced trend and the jitter are
-somewhat larger than recent comparable measurements in systems such as
-AU~Mic \citep{palle_transmission_2020}, DS~Tuc
-\citep{montet_young_2020,zhou_well_2020} and TOI~942
-\citep{wirth_2021_toi942}.  One possible explanation for the jitter is
-that it is astrophysical in origin, and that it is caused by some
-novel process operating on the surface of Kepler 1627A.  Another
-possibility is that our RV analysis underestimates our measurement
-uncertainties; in order to achieve the requisite time-sampling the S/N
-per resolution element in our spectra was 70 to 80, which is lower
-than desired for deriving high-precision velocities.  In addition, the
-rapid rotation of the star could affect accuracy of the uncertainties
-from the velocity extraction.  Observations at higher S/N are
-necessary to distinguish these two possibilities, and remain
-worthwhile in order to clarify the orbital geometry of \pn.  Useful
-next steps would include transit observations with a stabilized
-spectrograph in the optical \citep[{\it
-e.g.},][]{gibson_2016_kpf,seifahrt_maroon-x_2018}, or in the
-near-infrared \citep[{\it e.g.},][]{feinstein_halpha_2021}.
-
-\input{phot_table.tex}
-\input{rv_table.tex}
-
-\section{Flare Analysis}
-\label{app:flare}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f13.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Flares in Kepler\,1627}.  
-		{\it Top:}
-		The 1-minute cadence Kepler data (black points) is shown with a stellar variability model superposed
-		(blue line).
-		{\it Middle:}
-		Residual after subtracting the stellar variability model.  Flares
-		appear as spikes.
-		{\it Bottom:}
-		Zooms of the brightest, and third-brightest flares.  A timing
-		coincidence -- that both flares have ``successors'' approximately
-		one \added{planetary }orbital period after the initial event -- is emphasized.
-		\label{fig:flarezoom}
-	}
-\end{figure*}
-
-%\begin{figure*}[t]
-%	\begin{center}
-%		\leavevmode
-%		\includegraphics[width=0.9\textwidth]{f8.pdf}
-%	\end{center}
-%	\vspace{-0.7cm}
-%	\caption{
-%		{\bf Phase-folded flares in Kepler\,1627}.  
-%    {\it Top:}
-%    As in the middle of Figure~\ref{fig:flarezoom}, phase-folded at
-%    the planet's ephemeris.
-%    {\it Bottom:}
-%    Fitted flare amplitudes and orbital phases -- each point
-%    represents one flare.
-%    Colors indicate relative time, from the beginning of the
-%    98-day short cadence Quarter 15 dataset (dark blue) to the end (light
-%    yellow).
-%    Lower amplitude flares likely exist in the data, and were not
-%    examined.
-%		\label{fig:flarephase}
-%	}
-%\end{figure*}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f14.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-		{\bf Statistics of inter-flare arrival times}.  
-		24 flares were recorded with amplitudes exceeding 0.5\% over the
-		97.7 days of short cadence observations.  The histogram of the
-		time intervals between every possible pair of flares is shown in
-		black.  Some plausibly important timescales for star-planet
-		interactions, namely the planetary orbital period and synodic
-		period (the orbital period as seen from the rotating stellar
-		frame) are shown along with their linear combinations.  Monte
-		Carlo draws from a Poisson distribution are shown with the gray
-		bands.  While peaks in the observed distribution do coincide with
-		some of the ``special periods'', the statistical evidence for a
-		non-Poissonian process driving the flares does not clear the
-		$5$-$\sigma$ threshold.
-		\label{fig:flarestats}
-	}
-\end{figure*}
-
-In addition to the 3.9 years of long cadence data, short cadence
-(1-minute) Kepler observations were acquired over 97.7 days during
-Quarter 15.  The short cadence light curve shows a higher rate of
-flaring than visible in the long cadence data
-(Figure~\ref{fig:flarezoom}).  We analyzed the short cadence light
-curve and its flares according to the following procedure.
-\begin{enumerate}
-  \item Fit the starspot-induced variability using a Gaussian Process
-    with a \texttt{SHOTerm} kernel, a white-noise jitter term, and the
-    mean flux.
-  \item Select points more than twice the median absolute
-    deviation from the residual, and exclude them from the light
-    curve (these points include the flares).  Repeat Step~1.
-  \item Using the residual from Step 2, identify all flares, requiring
-    them to be at least 20 cadences apart, at least 7 median absolute
-    deviations above the median baseline, and lasting at least 2
-    cadences in duration.  Build the mask spanning these times, from 5
-    minutes before each flare begins to 2.5 minutes after the final
-    flare cadence.  Repeat Step 1 a final time.
-\end{enumerate}
-The final step of flare identification and fitting was performed using
-\texttt{altaipony} \citep{davenport_2016,ilin_flares_2021}.  The
-analytic flare model is from \citet{davenport_2014} and it
-parametrizes the flare with a start time, an exponential lag time, and
-an amplitude.
-%Figure~\ref{fig:flarephase} shows the resulting flares, amplitudes,
-%and phases.
-
-There were $N_{\rm f}=24$ flares that exceeded $0.5\%$ in relative
-flux during the short cadence observations.  These 24 flares spanned a
-total of 6.5 hours ($\sim$15 minutes per flare).  Inspecting the data,
-we noticed a coincidence in the flare arrival times.  The coincidence
-is that despite the low flare duty cycle, one orbital period after the
-brightest flare, a second flare followed.  This and a similar event
-are shown in Figure~\ref{fig:flarezoom}.  The timing error is good to
-a $\approx0.2\%$ difference from the orbital period, which given the
-duty cycle seems {\it a priori} unlikely.  If we consider flares
-falling within 2\% of the planet's orbital period after a previous
-flare, then 4 of the 24 flare events have candidate ``successors''.
-
-As with any coincidence, if one does not have a firm prediction, it is
-difficult to assess whether a surprise is statistically significant.
-Since our surprise was specifically at the inter-arrival time of
-certain flares coinciding with special time intervals, we performed
-the following analysis.  First, we considered all unordered pairs of
-flares.  For $N$ flares there are ${n \choose 2}$ such pairs (for our
-case, 276 pairs).  We then compared the distribution of the pair
-separations against that of a Poisson distribution.  Specifically, we
-drew $N_{\rm f}=24$ samples from a Poisson distribution with $\lambda
-= \Delta t / N_{\rm f}$, for $\Delta t=97.7\,{\rm days}$ the total
-duration of the observations, and repeated the draw $10^3$ times with
-unique random seeds.
-
-Figure~\ref{fig:flarestats} shows the results.  The vertical lines in
-the figure show the planetary orbital period, the synodic period
-$P_{\rm syn} = (P_{\rm rot}^{-1} - P_{\rm orb}^{-1})^{-1}$, and linear
-combinations thereof.  The tidal period (half the synodic period) is
-not shown.  The bins are logarithmically spaced to give 100 bins
-between the minimum and maximum ordinate values.  The gray bands
-express the range of values observed from the Poissonian draws.  While
-it does seem like an odd coincidence for peaks in the observed flare
-arrival time distribution to coincide with the locations of these
-``special intervals'', the statistical evidence for a non-Poissonian
-process driving the flares does not seem especially overwhelming.
-More quantitatively, the peaks observed at the orbital and synodic
-periods are within the $\pm 2$-$\sigma$ range of a Poissonian process,
-and those at $P_{\rm orb}+P_{\rm syn}$ and $P_{\rm orb}+2P_{\rm syn}$
-are only slightly above this range.  With that said, future analyses
-of these data by investigators with more knowledge of this topic could
-very well yield more quantitative insights.  Such analyses should keep
-in mind an important caveat: the amplitude distribution of M-dwarf
-flares extends up to many times the quiescent flux \citep[see Figure~7
-of][]{gunther_2020}.  A flare on Kepler\,1627B producing double its
-quiescent white-light flux would yield a $\approx$1\% apparent
-amplitude.  Such flares could represent a significant fraction of
-those in the Kepler observations.
+%\appendix
+%\section{Young, Age-Dated, and Age-Dateable Star Compilation}
+%\label{app:targetlist}
 
 %\listofchanges
-
 %\allauthors

### 2022-02-28 papers/collect.sh

@@ -1,0 +1,20 @@
+# Cep-Her XYZ
+cp /Users/luke/Dropbox/proj/rudolf/results/XYZvtang/XYZvtang_show1627_showallknown_showrsg5_showsun_square.pdf f1.pdf
+
+# rotation, CAMD, spectra (Li/H-alpha)
+# CAMD:
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_Set1_IC_2602_Pleiades_cleanhrcut_gaia2018_overplotkep1627_allknownkois.pdf f2a.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_Set1_IC_2602_Pleiades_cleanhrcut_gaia2018_smalllims_overplotkep1627_allknownkois.pdf f2b.pdf
+# rotation:
+cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_curtiscleaning_overplotkep1627.pdf f2c.pdf
+# spectra: Li EW vs Teff and Halpha EW vs Teff
+#TODO TODO
+
+# phased and sliced LCs
+cp ~/Dropbox/proj/rudolf/results/run_simpletransit_Kepler_1643/*phase*png f3a.png
+cp ~/Dropbox/proj/rudolf/results/run_simpletransit_KOI_7913/*phase*png f3b.png
+cp ~/Dropbox/proj/rudolf/results/run_simpletransit_KOI_7368/*phase*png f3c.png
+cp ~/Dropbox/proj/rudolf/results/run_simpletransit_Kepler_1627/*phase*png f3d.png
+
+# Rp, Period, Age
+cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20220227_colorbyage_showCepHer.pdf f4.pdf

### 2022-03-01 drivers/plot_koikeplerlc.py
add plot_koikeplerlc

@@ -1,0 +1,12 @@
+import os
+import rudolf.plotting as rp
+from rudolf.paths import RESULTSDIR
+import numpy as np
+
+PLOTDIR = os.path.join(RESULTSDIR, 'koikeplerlc')
+if not os.path.exists(PLOTDIR):
+    os.mkdir(PLOTDIR)
+
+for t0 in np.arange(0,1500,50):
+    t1 = t0+50
+    rp.plot_koikeplerlc(PLOTDIR, [t0,t1])

### 2022-03-01 papers/collect.sh

@@ -11,6 +11,7 @@ cp /Users/luke/Dropbox/proj/rudolf/results/rotation/deltaLyrCluster_rotation_vs_
 #TODO TODO
 
 # phased and sliced LCs
+cp /Users/luke/Dropbox/proj/rudolf/results/koikeplerlc/koikeplerlc_xlim0550_0600.pdf f3e.pdf
 cp ~/Dropbox/proj/rudolf/results/run_simpletransit_Kepler_1643/*phase*png f3a.png
 cp ~/Dropbox/proj/rudolf/results/run_simpletransit_KOI_7913/*phase*png f3b.png
 cp ~/Dropbox/proj/rudolf/results/run_simpletransit_KOI_7368/*phase*png f3c.png

### 2022-03-02 papers/ms.tex
add plot_koikeplerlc
ditto to ms
compile paper

@@
   Here we present an analysis of a group of $38 \pm 6$ million year old stars
-  in the Cepheus-Hercules region, hereafter the Cep-Her complex.
+  spanning Cepheus ($l=100^\circ$) to Hercules ($l=40^\circ$), hereafter the Cep-Her complex.
   This group of stars includes four previously known Kepler Objects of
@@
   KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$).
-	Kepler-1627 is a previously validated young planet
-	in a component of the Cep-Her complex called the $\delta$\ Lyr\ cluster \citep{bouma_kep1627_2022}.
-	Here we focus on the latter three systems, which are in other sub-components
-	of the complex (RSG5 and CH-2).
-	Based on kinematic evidence from Gaia, stellar rotation periods from TESS, and spectroscopy,
-	these three systems are also $38 \pm 6$ million years old.
-	Based on the transit shapes and high resolution imaging, we
-	statistically validate that they are all most likely planets (false positive probabilities of $6\times10^{-9}$,
-	$5\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368, and
-	KOI-7913 respectively).
-	Supplemented by Gaia and TESS, the main Kepler mission is now contributing to the
-	census of young close-in planets, and Kepler-1643 and KOI-7913 are the first
-	empirical demonstration that mini-Neptunes with sizes of $\approx$2 Earth radii
-	exist at ages of roughly 40 million years.
+  Kepler-1627 is a Neptune-sized planet in a component of the Cep-Her
+  complex called the $\delta$\ Lyr\ cluster
+  \citep{bouma_kep1627_2022}.  Here we focus on the latter three
+  systems, which are in other sub-components of the complex (RSG-5 and
+  CH-2).  Based on kinematic evidence from Gaia, stellar rotation
+  periods from TESS, and spectroscopy, these three systems are also
+  $38 \pm 6$ million years old.  Based on the transit shapes and high
+  resolution imaging, we statistically validate that they are all most
+  likely planets (false positive probabilities of $6\times10^{-9}$,
+  $5\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368,
+  and KOI-7913 respectively).  Supplemented by Gaia and TESS, the main
+  Kepler mission is now contributing to the census of young close-in
+  planets, and Kepler-1643 and KOI-7913 are the first empirical
+  demonstration that mini-Neptunes with sizes of $\approx$2 Earth
+  radii exist at ages of roughly 40 million years.
 \end{abstract}
@@
     {\bf Galactic positions  and tangential velocities of stars in the
-    $\delta$\,Lyr cluster.} 
+    complex.} 
+    Sub-clusters include the $\delta$ Lyr cluster, RSG-5, and the
+    worryingly diffuse ``CH-2''.
 		\label{fig:XYZvtang}
@@
 		\subfloat{
-			\includegraphics[width=0.7\textwidth]{f2c.pdf}
+			\includegraphics[width=0.49\textwidth]{f2c.pdf}
+			\includegraphics[width=0.49\textwidth]{TEMP_f2d.pdf}
 		}
@@
 		{\bf The \cn\ is \clusterage\ old.} 
+    The top row shows CAMDs.  Left shows CH-2, right shows RSG-5.
+    The bottom left shows gyro.
+    The bottom right shows lithium (and is a {\bf place-holder}).  (Somewhere we will have
+    H-alpha?).
 		\label{fig:age}
@@
 		\subfloat{
-			\includegraphics[width=0.5\textwidth]{f3a.png}
-			\includegraphics[width=0.5\textwidth]{f3b.png}
+			\includegraphics[width=0.99\textwidth]{f3e.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.5\textwidth]{f3c.png}
-			\includegraphics[width=0.5\textwidth]{f3d.png}
+			\includegraphics[width=0.5\textwidth]{f3a.pdf}
+			\includegraphics[width=0.5\textwidth]{f3b.pdf}
+		}
+
+		\subfloat{
+			\includegraphics[width=0.5\textwidth]{f3c.pdf}
+			\includegraphics[width=0.5\textwidth]{f3d.pdf}
 		}
@@
 	\caption{
-		{\bf planets.} 
+		{\bf Raw and processed light curves for the objects of
+    interest.} Top: raw.  Bottom: processed.
+    There increases scatter during transit is likely due to starspot
+    crossing events.
 		\label{fig:planets}
@@
 	\caption{
+    %    namelist = ['Kepler-1627 A', 'KOI-7368', 'KOI-7913 A', 'KOI-7913 B', 'Kepler-1643']
+    %    markers = ['P', 'v', 'X', 'X', 's']
 		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
@@
 		3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
-		its uncertainty. STAR, SQUARE, DIAMOND.  The large sizes of
+    its uncertainty. Kepler-1627 (+), KOI-7368 (down-triangle),
+    KOI-7913 (X), Kepler-1643 (diamond).  The large sizes of
 		the youngest transiting planets could be explained by their
@@
 		help to confirm this expectation.  Selection effects may also be
-		important.  Parameters are from the NASA Exoplanet Archive (2021
-		Sept 15).
+		important.  Parameters are from the NASA Exoplanet Archive (2022
+		Feb 27).
 		\label{fig:rp_period_age}

### 2022-03-02 papers/ms.tex
update rotation and lithium plots

@@
 %% Add "Submitted to " argument.
-\received{2021 Oct 12}
+\received{\today}
 \revised{---}
@@
     {\bf Galactic positions  and tangential velocities of stars in the
-    complex.} 
+    complex.}  {\bf This is a placeholder before we get real kinematics.}
     Sub-clusters include the $\delta$ Lyr cluster, RSG-5, and the
@@
 			\includegraphics[width=0.49\textwidth]{f2c.pdf}
-			\includegraphics[width=0.49\textwidth]{TEMP_f2d.pdf}
+			\includegraphics[width=0.49\textwidth]{f2d.pdf}
 		}
@@
     The top row shows CAMDs.  Left shows CH-2, right shows RSG-5.
-    The bottom left shows gyro.
-    The bottom right shows lithium (and is a {\bf place-holder}).  (Somewhere we will have
-    H-alpha?).
+    The bottom left shows gyro (and is a {\bf place-holder since we might want to add CH-2 and RSG5}).
+    The bottom right shows lithium (black points are NGC2547 and
+    IC2602 from Randich+18 and probably are not believable at the red
+    end; for Kepler-1643 I'm less confident, but it might be tied to the ``slow'' rotation period -- this is why we need RSG5 rotation periods).
+    Also, we might want an
+    H-alpha plot?).
 		\label{fig:age}

### 2022-03-02 drivers/plot_rotation_period_windowslider.py

@@ -6,6 +6,7 @@ PLOTDIR = os.path.join(RESULTSDIR, 'rotation_period_windowslider')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
+rp.plot_rotation_period_windowslider(PLOTDIR, 'KOI_7913_B')
 rp.plot_rotation_period_windowslider(PLOTDIR, 'KOI_7913')
 rp.plot_rotation_period_windowslider(PLOTDIR, 'Kepler_1643')
 rp.plot_rotation_period_windowslider(PLOTDIR, 'KOI_7368')

### 2022-03-02 drivers/plot_rotation_vs_color.py

@@ -10,17 +10,13 @@ from earhart.priors import AVG_EBpmRp
 
 runid = 'deltaLyrCluster'
 
-#TODO: do a cut in rotation periods?
-# c = 'defaultcleaning_cutProtColor'
-# rp.plot_rotationperiod_vs_color(
-#     PLOTDIR, runid, E_BpmRp, core_halo=1, yscale='linear', cleaning=c,
-#     emph_binaries=0, talk_aspect=1, showPleiadesQuad=0
-# )
-# rp.plot_rotationperiod_vs_color(
-#     PLOTDIR, runid, E_BpmRp, core_halo=0, yscale='linear', cleaning=c,
-#     emph_binaries=1, talk_aspect=1, showPleiadesQuad=0
-# )
-
+# add all KOIs
+rp.plot_rotationperiod_vs_color(
+    PLOTDIR, runid, yscale='linear', cleaning='defaultcleaning',
+    emph_binaries=0, talk_aspect=1, kinematic_selection=1,
+    overplotkep1627=0, show_allknown=1
+)
+assert 0
 # pleiades and praesepe only
 rp.plot_rotationperiod_vs_color(
     PLOTDIR, runid, yscale='linear',

### 2022-03-02 rudolf/plotting.py

@@ -25,6 +25,9 @@ Kepler phot:
     plot_flare_pair_time_distribution
     plot_phasedlc_quartiles
 
+Spec:
+    plot_lithium
+
 RM:
     plot_simulated_RM
     plot_RM
@@ -2628,7 +2631,7 @@ def plot_rotationperiod_vs_color(outdir, runid, yscale='linear', cleaning=None,
                                  emph_binaries=False, refcluster_only=False,
                                  talk_aspect=0, xval_absmag=0,
                                  kinematic_selection=0,
-                                 overplotkep1627=0):
+                                 overplotkep1627=0, show_allknown=0):
     """
     Plot rotation periods that satisfy the automated selection criteria
     (specified in helpers.get_autorotation_dataframe)
@@ -2717,9 +2720,10 @@ def plot_rotationperiod_vs_color(outdir, runid, yscale='linear', cleaning=None,
 
         else:
             if not xval_absmag:
-                xval = (
-                    df['phot_bp_mean_mag'+cstr] - df['phot_rp_mean_mag'+cstr]
+                get_xval = lambda _df: np.array(
+                    _df['phot_bp_mean_mag'+cstr] - _df['phot_rp_mean_mag'+cstr]
                 )
+                xval = get_xval(df)
             else:
                 get_xval = lambda _df: np.array(
                     _df['phot_g_mean_mag'] + 5*np.log10(_df['parallax']/1e3) + 5
@@ -2773,6 +2777,36 @@ def plot_rotationperiod_vs_color(outdir, runid, yscale='linear', cleaning=None,
             markerfacecolor='yellow', markersize=14, marker='*',
             color='black', lw=0
         )
+
+    if show_allknown:
+        _, _, _, koi_df_dict = get_deltalyr_kc19_gaia_data(return_all_targets=1)
+
+        namelist = ['Kepler-1627 A', 'KOI-7368', 'KOI-7913 A', 'KOI-7913 B', 'Kepler-1643']
+        markers = ['P', 'v', 'X', 'X', 's']
+        # lime: CH-2 (KOI-7913, KOI-7368)
+        # magenta: RSG5 (Kepler-1643)
[…]

### 2022-03-02 rudolf/starinfo.py

@@ -1,0 +1,53 @@
+starinfodict = {
+    "KOI-7368": {
+        # TRES / SPC
+        'Teff':5241.27,
+        'Teff_err':50,
+        # TRES / KOI_7368_Li_EW_shift-0.30_results.csv
+        'Li_EW':236.2, 'Li_EW_perr': 15.6, 'Li_EW_merr': 14.3,
+        # rotation_period_windowslider
+        'Prot': 2.606363,
+        'Prot_err': 0.051906,
+    },
+    "KOI-7913 A": {
+        # HIRES / SM-Emp Median
+        'Teff': 4324,
+        'Teff_err': 70,
+        # HIRES / KOI-7913_ij440.75_Li_EW_shift-0.05_results.csv
+        'Li_EW':65.058, 'Li_EW_perr': 8.125, 'Li_EW_merr': 6.323,
+        # rotation_period_windowslider
+        'Prot': 3.197631,
+        'Prot_err': 0.056841,
+    },
+    "KOI-7913 B": {
+        # HIRES / SM-Emp Median
+        'Teff': 4038,
+        'Teff_err': 70,
+        # HIRES / KOI-7913B_ij438.81_Li_EW_shift-0.05_results.csv
+        'Li_EW':34.613, 'Li_EW_perr': 12.683, 'Li_EW_merr': 13.974,
+        # rotation_period_windowslider
+        'Prot': 2.641824,
+        'Prot_err': 0.066772,
+    },
+    "Kepler-1643": {
+        # HIRES / SM-Emp Median
+        'Teff': 4916,
+        'Teff_err': 110,
+        # HIRES / Kepler1643_ij384.99_Li_EW_shift-0.30.png
+        # or Kepler1643_ij438.82_Li_EW_shift-0.05.png
+        'Li_EW': 126.4, 'Li_EW_perr': 7.9, 'Li_EW_merr': 4.2,
+        # rotation_period_windowslider
+        'Prot': 5.106414,
+        'Prot_err': 0.043881,
+    },
+    "Kepler-1627 A": {
+        # HIRES / SM-Emp Median / Bouma+22
+        'Teff': 5505,
+        'Teff_err': 60,
+        # HIRES
+        'Li_EW': 233, 'Li_EW_perr': 5, 'Li_EW_merr': 7,
+        # rotation_period_windowslider
+        'Prot': 2.641824,
+        'Prot_err': 0.043764,
+    },
+}

### 2022-03-06 papers/ms.tex
improve phased lc layout
paper: add bad intro

@@
 %\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
-\documentclass[12pt,modern,tighten]{aastex63}
+\documentclass[12pt,twocolumn,linenumbers]{aastex63}
 %\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
@@
 \begin{abstract}
-  Stellar positions and velocities from Gaia
-  are yielding a refined view of stellar clusters during the
-  first hundred million years of their lives.
-  Here we present an analysis of a group of $38 \pm 6$ million year old stars
-  spanning Cepheus ($l=100^\circ$) to Hercules ($l=40^\circ$), hereafter the Cep-Her complex.
-  This group of stars includes four previously known Kepler Objects of
-  Interest:
+  Stellar positions and velocities from Gaia are yielding a new and
+  refined view of stellar cluster evolution.
+  Here we present an analysis of a group of $\approx$40 million
+  year old stars spanning Cepheus ($l=100^\circ$) to Hercules
+  ($l=40^\circ$), hereafter the Cep-Her complex.  The group contains
+  four known Kepler Objects of Interest:
   Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$),
@@
   KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$).
-  Kepler-1627 is a Neptune-sized planet in a component of the Cep-Her
+  Kepler-1627 is a Neptune-sized planet in a sub-component of the
   complex called the $\delta$\ Lyr\ cluster
@@
   $38 \pm 6$ million years old.  Based on the transit shapes and high
-  resolution imaging, we statistically validate that they are all most
+  resolution imaging, they are all most
   likely planets (false positive probabilities of $6\times10^{-9}$,
   $5\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368,
-  and KOI-7913 respectively).  Supplemented by Gaia and TESS, the main
-  Kepler mission is now contributing to the census of young close-in
-  planets, and Kepler-1643 and KOI-7913 are the first empirical
-  demonstration that mini-Neptunes with sizes of $\approx$2 Earth
-  radii exist at ages of roughly 40 million years.
+  and KOI-7913 respectively).  Kepler-1643 and KOI-7913 are therefore
+  the first empirical demonstration that mini-Neptunes with sizes of
+  $\approx$2 Earth radii exist at ages of roughly 40 million years.
 \end{abstract}
@@
 
+The discovery of young transiting planets is a major frontier in
+current exoplanet research.  The reason is that the properties of
+these planets can help constrain the timescales for processes
+including hot Jupiter migration (CITE), the cooling and contraction of
+mini-Neptunes (CITE Rizzuto), the early evolution of close-in
+multiplanet systems (CITE David), and whether and how mass-loss
+explains the radius valley (e.g., CITE, CITE, Lee+21 on theory; CITE
+CITE CITE on observations).
+
+Assuming a constant star formation rate in the Galaxy, $\approx$1\% of
+stars are $\lesssim$10$^8$ years old.  While $\approx$5{,}000
+transiting exoplanets are known (CITE), 11 currently meet this age
+cut.  The discrepancy between the $\approx$50 expected and $\approx$10
+known likely has a contribution from selection effects in planetary
+detection, which include photometric and spectroscopic starspot
+modulation, and the rarity of young stars.
+
+While detecting planets around young and active stars is challenging,
+so is measuring stellar ages (see CITE Soderblom for a review).  It
+should therefore not be surprising that advances along either
+dimension can yield advances in the census of young transiting
+planets.
+
+The prime Kepler mission (CITE) found most of the currently known
+transiting exoplanets, and it was conducted before the positions,
+velocities, and stellar brightnesses measured by Gaia revolutionized
+our understanding of stellar clusters (CITE, CITE, CITE).  This
+revolution is slated to continue with the upcoming third Gaia data
+release (DR3 CITE).  Nonetheless, it seems sensible to revisit the
+Kepler field, given our new knowlege of the stellar ages.
+
+In this work, we expand on a previous study of a 38 million year (Myr)
+old Neptune-sized planet in the Kepler field (Kepler-1627Ab; CITE).
+The age of this planet was derived based on its host star's membership
+in the $\delta$\ Lyr\ cluster.  Our analysis of this cluster focused
+on the immediate kinematic vicinity of Kepler-1627A in order to
+reliably confirm the age of the planet.  However it also became clear
+that the $\delta$\ Lyr\ cluster is spatially and kinematically close
+to a much larger group of similarly aged stars.  This group, which is
+at a distance of roughly 300 parsecs, appears to span Cepheus
+($l=100^\circ$) to Hercules ($l=40^\circ$), at galactic latitudes,
+$b$, between roughly 0 and 20 degrees.  We therefore refer to it as
+the Cep-Her complex.  Viewed top-down in the plane of the Galaxy, the
+complex spans roughly 250 parsecs in the direction toward the galactic
+center, and 150 parsecs in the perpendicular direction of galactic
+rotation.  A detailed kinematic analysis and exploration is currently
+being prepared by R.~Kerr and collaborators.
+
+The subject of this study is the intersection of the Cep-Her complex
+with the Kepler field, and an analysis of the consequent transiting
+planets and planet candidates.  Cross-matching the full set of
+candidate Cep-Her members against all known Kepler Objects of Interest
+(CITE) yielded XX candidate matches.  Given the purported XX Myr age
+of the group, a visual inspection of the light curves quickly enabled
+a decision about whether the matches could be valid, based on the
+presence or non-presence of stellar rotation signals of the
+appropriate periods and amplitudes (CITE REBULL).  Four possible
+matches remained from this exercise: Kepler-1627, Kepler-1643,
+KOI-7368, and KOI-7913.  Given our previous analysis of Kepler-1627,
+we will focus for the most part on the latter three.
+
 \section{The Cluster}
 \label{sec:cluster}
+
+\subsection{Selecting Cluster Members}
+\label{sec:kinematicselection}
 
@@
 
-\subsection{Selecting Cluster Members}
-\label{sec:kinematicselection}
+\subsection{The Cluster's Age}
+\label{sec:clusterage}
 
@@
 		{\bf The \cn\ is \clusterage\ old.} 
-    The top row shows CAMDs.  Left shows CH-2, right shows RSG-5.
-    The bottom left shows gyro (and is a {\bf place-holder since we might want to add CH-2 and RSG5}).
-    The bottom right shows lithium (black points are NGC2547 and
-    IC2602 from Randich+18 and probably are not believable at the red
-    end; for Kepler-1643 I'm less confident, but it might be tied to the ``slow'' rotation period -- this is why we need RSG5 rotation periods).
-    Also, we might want an
-    H-alpha plot?).
+    The top row shows CAMDs.  Left shows CH-2, right shows RSG-5.  The
+    bottom left shows gyro (and is a {\bf place-holder since we might
+    want to add CH-2 and RSG5}).  The bottom right shows lithium
+    (black points are NGC2547 and IC2602 from Randich+18 and probably
+    are not believable at the red end; for Kepler-1643 I'm less
+    confident, but it might be tied to the ``slow'' rotation period --
+    this is why we need RSG5 rotation periods).  Also, we might want
+    an H-alpha plot?).
 		\label{fig:age}
@@
 
-\subsection{The Cluster's Age}
-\label{sec:clusterage}
+The evidence for the RSG-5 and CH-2 sub-groups at least roughly
+sharing the age of the $\delta$\ Lyra\ cluster is shown in
+Figure~\ref{fig:age}.
 
@@
 
-\subsection{Kepler\,1627A}
+\input{tab_sysparams.tex}
+
+Many of the relevant stellar parameters can be gleaned by inspecting
+Figure~\ref{fig:age}.  The known planet-hosting stars in Cep-Her span
+spectral types of G8V (Kepler-1627) to K8V (KOI-7913 A).  The
+secondary in the KOI-7913 system is only marginally cooler than the
+primary (SpType $\approx$M0V).
+The latter system is 3$\farcs$501 
+%FIXME FIXME FIXME TODO TODO TODO
+A Solar-mass star with solar metallicity arrives at the zero-age main
+sequence at $t\approx40$ million years (CITE CHOI MESA), and so these
+stars are in the end stages of their pre-main-sequence contraction.
+Their Gaia EDR3 parallaxes span X.X to Y.Y, corresponding to distances
+between X.X and Y.Y parsecs.
+
+To derive the spectroscopic parameters $(T_{\rm eff}, \log g_\star,
+[{\rm Fe/H}])$ and  the equivalent width of the Li\,\textsc{I}
+from the 6708\,\AA\ doublet, we acquired spectra.  Specifically, we
+acquired iodine-free HIRES spectra for Kepler-1643 (YYYY/MM/DD) and
+KOI-7913 (YYYY/MM/DD and YYYY/MM/DD), where for the latter the two
+different epochs corresponded to observations of the secondary and
+primary respectively.  The acquisition and analysis followed the
+standard reduction techniques of the California Planet Survey
+\citep{howard_cps_2010}.  For KOI-7368, we acquired TRES spectra on
+YYYY/MM/DD and YYYY/MM/DD.  CITE METHOD PAPER.
+The results are given in Table~\ref{tab:sysparams}.
+
+Table~\ref{tab:sysparams} summarizes
+%
+%from the 
+%
+%
+%\citep[{\it e.g.},][]{soderblom_ages_2014}, we acquired an iodine-free
+%spectrum from Keck/HIRES on the night of 2021 March 26 using the
+%standard setup and reduction techniques of the California Planet
+%Survey \citep{howard_cps_2010}.  Following the equivalent width
+%measurement procedure described by \citet{bouma_2021_ngc2516}, we find
+%${\rm EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   This value does not
+%correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Nonetheless,
+%given the stellar effective temperature (Table~\ref{tab:starparams}),
+%this measurement is in agreement with expectations for a
+%$\approx40$\,Myr star ({\it e.g.}, as measured in IC~2602 by
+%\citealt{randich_gaiaeso_2018}).  It is also larger than any lithium
+%equivalent widths measured by \citet{berger_identifying_2018} in their
+%analysis of 1{,}301 Kepler-star spectra.
+%
+%
+%% rotation period from plot_rotation_period_windowslider.py
+%Based on the spatial and kinematic association of \sn\ with the \cn,
+%and the assumption that the planet formed shortly after the star, it
+%seems likely that \sn\ is the same age as the cluster. There are two
+%consistency checks on whether this is true: rotation and lithium.
+%Based on the Kepler light curve, the rotation period is
+%$2.642\pm0.042$\,days, where the quoted uncertainty is based on the
+%scatter in rotation periods measured from each individual Kepler
+%quarter.  This is consistent with comparable cluster members
+%(Figure~\ref{fig:age}).
+
+%\subsection{Kepler\,1627A}
 \subsection{Kepler\,1643}
@@
 		\subfloat{
-			\includegraphics[width=0.99\textwidth]{f3e.pdf}
+			\includegraphics[width=0.85\textwidth]{f3e.pdf}
 		}
 
+		\vspace{-0.5cm}
 		\subfloat{
-			\includegraphics[width=0.5\textwidth]{f3a.pdf}
-			\includegraphics[width=0.5\textwidth]{f3b.pdf}
+			\includegraphics[width=0.41\textwidth]{f3a.pdf}
+			\includegraphics[width=0.41\textwidth]{f3b.pdf}
 		}
 
+		\vspace{-1.45cm}	
 		\subfloat{
-			\includegraphics[width=0.5\textwidth]{f3c.pdf}
-			\includegraphics[width=0.5\textwidth]{f3d.pdf}
+			\includegraphics[width=0.41\textwidth]{f3c.pdf}
+			\includegraphics[width=0.41\textwidth]{f3d.pdf}
 		}
@@
     interest.} Top: raw.  Bottom: processed.
-    There increases scatter during transit is likely due to starspot
-    crossing events.
+    The increased scatter during transit is likely due to starspot
+    crossing events.  KOI-7913 is janky, but P=24 days.
 		\label{fig:planets}

### 2022-03-08 papers/ms.tex
paper: add kinematic context

@@
 
+%FIXME WORDING OF THIS PARAGRAPH NEEDS WORK.  ALSO, WHERE DO YOU PUT
+%THE CONTEXT RE CLUSTERS?
+Understanding the formation and evolution of stellar clusters is
+similarly an emerging frontier in analyses of star formation and
+cluster evolution.  In particular, it is understood that most embedded
+clusters disperse within a few tens of millions of years after forming
+their first stars (CITE).  The discovery of tidal tails (CITE CITE)...
+streams (CITE Meingast18, Curtis19), etc...  Connect to moving groups
+(CITE Gagne 21).  What are the typical densities for star formation?
+How does gas expulsion and the SFE affect cluster morphology?  {\bf
+FIXME}
+
 Assuming a constant star formation rate in the Galaxy, $\approx$1\% of
@@
 \end{figure*}
+
+Our focus is on a region of the Galaxy roughly 200 to 500 parsecs from
+the Sun, above the galactic plane, and spanning galactic longitudes of
+roughly $40^\circ$ to $100^\circ$ degrees.  A direct query of the Gaia
+archive imposing these constraints yields $\approx3.4\times10^7$
+sources.  Imposing the cuts suggested by \citet[][Appendix
+B]{GaiaCollaboration2018} on the photometric signal to noise ratio,
+the number of visibility periods used, the astrometric $\chi^2$ of the
+single-source solution, the $G_{\rm BP}-G_{\rm RP}$ color excess
+factor and the parallax precision ($\varpi/\sigma_\varpi>5$) yields
+$\approx5.0\times10^6$ sources.
+
+Imposing a thin slice in color-absolute magnitude space around the
+pre-main-sequence of the $\delta$ Lyr cluster's previously established
+$\approx$40 Myr locus, and imposing kinematics cut for $-8 < v_b /
+{\rm km\,s^{-1}} < 0$ and $-15 < v_{l'} / {\rm km\,s^{-1}} < 15$,
+where $l'=l\cos b$, yields $\approx 5.2\times10^3$ sources.  These
+objects are all in the range BP-RP 1.5 to 3.2, so roughly K5V to
+M4.5V, or $\approx$0.2 to 0.7 solar masses.  Much more manageable, but
+according to IMF arguments, missing at least a factor of two in mass
+of any clusters.
+
+In XYZ, and in tangential velocity space, these five thousand sources
+are visually clustered in between a few to a dozen groups, depending
+on the characteristic size and velocity scales of interest.  The first
+suggestion that these objects could all be part of a monolithic
+structure, to our knowledge, came from the manual grouping that was
+performed between them by \citet{KounkelCovey2019} when defining the
+structure they dubbed ``Theia~73''.  Different subsets of these stars
+have however previously been noticed and studied.  They include the
+$\delta$ Lyr cluster, also known as Stephenson-1
+\citep{stephenson_possible_1959}, which comprises a major grouping
+around $(l,b)=(68^\circ,15^\circ)$, with typical velocities $(v_{l'},
+v_b)=(-4.5{\rm \kms},-4{\rm \kms})$.  Another individual component
+that has previously been recognized is RSG-5 \citep{roser_nine_2016}.
+This group is at roughly $(l,b)=(83^\circ,6^\circ)$, with velocities
+$(v_{l'}, v_b)=(+5.5{\rm \kms},-3.5{\rm \kms})$.  Most of the other
+subclusters, including CH-2, and those in Cep-Cyg and Cerberus are too
+small or dispersed to have previously been analyzed in great detail.
 

### 2022-03-14 papers/authors.tex
compile paper
update ms

@@
 \affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
-%\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
 
@@
 \affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}% % ... Kepler correlations
-
+%
 % ... stellar rotation & the initial crossmatch
@@
 \affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
-
-% \author[0000-0003-1298-9699]{K. Masuda}
-% \affiliation{Department of Earth and Space Science, Osaka University, Osaka 560-0043, Japan}
 %
-% % ... planet-fitting, kinematics, meta.
+% HIRES
+\author[0000-0002-0531-1073]{H. Isaacson}
+\affiliation{Astronomy Department, University of California, Berkeley, CA 94720, USA}
+%
+% planet-fitting, kinematics, cluster.
 \author{L. A. Hillenbrand}
@@
 %
-% % ... RM fitting
-% \author[0000-0001-7409-5688]{G. Stefansson}
-% \altaffiliation{Henry Norris Russell Fellow}
-% \affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
-
-%
-% PFS Collaborators
-%
+% HIRES Collaborators
 \author[0000-0001-8638-0320]{A. W. Howard}
@@
 %
-\author[0000-0002-0531-1073]{H. Isaacson}
-\affiliation{Astronomy Department, University of California, Berkeley, CA 94720, USA}
-% pending
-
 % AO IMAGING
@@
 \affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}% % ... Kepler correlations
+%
+% TRES
+\author[0000-0001-6637-5401]{A. Bieryla}
+\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
+%
+% TRES
+\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu
+\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
+%
+% AO / NIRC2
+\author[0000-0003-0967-2893]{E. A.~Petigura}
+\affiliation{Department of Physics \& Astronomy, University of California Los Angeles, Los Angeles, CA 90095, USA}
+%
+% AO / NIRC2
+\author[0000-0001-8832-4488]{D. Huber}
+\affiliation{Institute for Astronomy, University of Hawaii, 2680 Woodlawn Drive, Honolulu, HI 96822, USA}

### 2022-03-14 papers/ms.tex
update ms

@@
   Stellar positions and velocities from Gaia are yielding a new and
-  refined view of stellar cluster evolution.
+  refined view of how stellar clusters evolve.
   Here we present an analysis of a group of $\approx$40 million
@@
 including hot Jupiter migration (CITE), the cooling and contraction of
-mini-Neptunes (CITE Rizzuto), the early evolution of close-in
-multiplanet systems (CITE David), and whether and how mass-loss
-explains the radius valley (e.g., CITE, CITE, Lee+21 on theory; CITE
-CITE CITE on observations).
+planets with massive gaseous envelopes (CITE Rizzuto), the early
+evolution of close-in multiplanet systems (CITE David), and whether
+and how mass-loss explains the radius valley (e.g., CITE, CITE, Lee+21
+on theory; CITE CITE CITE on observations).
 
@@
 	\caption{
-    {\bf Galactic positions  and tangential velocities of stars in the
-    complex.}  {\bf This is a placeholder before we get real kinematics.}
-    Sub-clusters include the $\delta$ Lyr cluster, RSG-5, and the
-    worryingly diffuse ``CH-2''.
+    {\bf Galactic positions  and tangential velocities of candidate members of the
+    Cep-Her complex.}
+	{\it Top row}: On-sky positions in galactic coordinates.
+	Black points are stars for which membership is more likely than for
+	gray points (``weight'' cuts above 0.10 and 0.02 respectively; see text).
+	The $\delta$ Lyr cluster, at $\{ l, b\} \approx \{ 66^\circ, 12^\circ\}$, contains Kepler-1627  \citep{bouma_kep1627_2022}.
+	{\it Middle row}:
+	Galactic positions.
+	The Sun is at $\{X, Y, Z\} = \{-8122, 0, 20.8\}$ pc.
+	{\it Bottom row}:
+	Galactic tangential velocities.
+	The spatial and kinematic overlap between RSG-5 ($\{ l, b\} \approx \{ 82^\circ, 6^\circ\}$; $\{ X, Y, Z\} \approx \{-8070, 330, 50\} $ pc)
+	and Kepler-1643 is fairly obvious.
+	KOI-7368 and KOI-7913 however are candidate members of a diffuse population,
+	that we dub ``CH-2''.
+	Although CH-2 appears close to RSG-5 when projected in $(X,Y)$ and $(v_b, v_{l*})$,
+	it is spatially  distinct in $(Z,Y)$.
 	\label{fig:XYZvtang}
@@
 
+\subsection{General Context} % might cut!
+
 Our focus is on a region of the Galaxy roughly 200 to 500 parsecs from
 the Sun, above the galactic plane, and spanning galactic longitudes of
-roughly $40^\circ$ to $100^\circ$ degrees.  A direct query of the Gaia
-archive imposing these constraints yields $\approx3.4\times10^7$
-sources.  Imposing the cuts suggested by \citet[][Appendix
+roughly $40^\circ$ to $100^\circ$ degrees.  Querying the Gaia archive
+and imposing the cuts suggested by \citet[][Appendix
 B]{GaiaCollaboration2018} on the photometric signal to noise ratio,
@@
 around $(l,b)=(68^\circ,15^\circ)$, with typical velocities $(v_{l'},
-v_b)=(-4.5{\rm \kms},-4{\rm \kms})$.  Another individual component
+v_b)=(-4.5{\rm kms},-4{\rm kms})$.  Another individual component
 that has previously been recognized is RSG-5 \citep{roser_nine_2016}.
 This group is at roughly $(l,b)=(83^\circ,6^\circ)$, with velocities
-$(v_{l'}, v_b)=(+5.5{\rm \kms},-3.5{\rm \kms})$.  Most of the other
+$(v_{l'}, v_b)=(+5.5{\rm kms},-3.5{\rm kms})$.  Most of the other
 subclusters, including CH-2, and those in Cep-Cyg and Cerberus are too
 small or dispersed to have previously been analyzed in great detail.
+
+\subsection{Actual Membership Selection} % might cut!
+
+% TODO: add words to acknowledge that Cep-Her is not yet demonstrated
+% to be a thing
+
+To identify stars in this region of space that could be members of the
+Cep-Her complex, we follow the methods of \citet{Kerr2021}.  The
+analysis therefore broadly consists of two steps.  The first step is
+to select stars that photometrically appear young based on the Gaia
+EDR3 magnitudes $\{G, G_{\rm RP}, G_{\rm BP}\}$, in combination with
+the parallaxes and reddening estimates.  This method works best for
+low-mass stars due to their long pre-main-sequence contraction
+timescale, and for massive stars because of their fast evolutionary
+timescale.  The second step in the analysis is to use the
+photometrically-identified young stars as the ``seed'' for a candidate
+population in a spatial and kinematic clustering analysis in $\{ X, Y,
+Z, c v_b, c v_{l^*} \} $.  The latter step helps identify F, G, and
+K-type stars that cannot be identified as young based solely on their
+positions in color versus absolute magnitude.
+
+%FIXME TODO TODO TODO: this analysis!
 

### 2022-03-14 papers/collect.sh

@@ -1,5 +1,5 @@
 # Cep-Her XYZ
-cp /Users/luke/Dropbox/proj/rudolf/results/XYZvtang/XYZvtang_showallknown_showrsg5_showsun_square_showset1.pdf f1.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/CepHer_XYZvtang_sky/CepHer_XYZvtang_sky.pdf f1.pdf
 
 # rotation, CAMD, spectra (Li/H-alpha)
 # CAMD:

### 2022-03-15 papers/ms.tex
update ms

@@
 	\caption{
-    {\bf Galactic positions  and tangential velocities of candidate members of the
+    {\bf Positions  and velocities of candidate members of the
     Cep-Her complex.}
 	{\it Top row}: On-sky positions in galactic coordinates.
-	Black points are stars for which membership is more likely than for
-	gray points (``weight'' cuts above 0.10 and 0.02 respectively; see text).
-	The $\delta$ Lyr cluster, at $\{ l, b\} \approx \{ 66^\circ, 12^\circ\}$, contains Kepler-1627  \citep{bouma_kep1627_2022}.
+	Black points are stars for which group membership is more likely than for
+	gray points (``weight'' cuts above 0.10 and 0.02; see text).
+	Kepler-1627 is in the outskirts of the $\delta$ Lyr cluster  \citep{bouma_kep1627_2022}, which is centered
+	at $\{ l, b\} \approx \{ 66^\circ, 12^\circ\}$.
 	{\it Middle row}:
@@
 	{\it Bottom row}:
-	Galactic tangential velocities.
-	The spatial and kinematic overlap between RSG-5 ($\{ l, b\} \approx \{ 82^\circ, 6^\circ\}$; $\{ X, Y, Z\} \approx \{-8070, 330, 50\} $ pc)
-	and Kepler-1643 is fairly obvious.
+	Galactic tangential velocities (left) and longitudinal tangential velocity
+	versus galactic latitude (right).
+	There is a strong spatial and kinematic overlap between Kepler-1643
+  and RSG-5 (purple points; $\{ l, b\} \approx \{ 82^\circ, 6^\circ\}$; $\{ X, Y, Z\}
+  \approx \{-8070, 330, 50\} $ pc).
 	KOI-7368 and KOI-7913 however are candidate members of a diffuse population,
-	that we dub ``CH-2''.
-	Although CH-2 appears close to RSG-5 when projected in $(X,Y)$ and $(v_b, v_{l*})$,
-	it is spatially  distinct in $(Z,Y)$.
+	that we dub ``CH-2'' (green points).
+	The green and purple points are manually selected from the initial candidate
+	population based on galactic position and tangential velocity.
 	\label{fig:XYZvtang}
@@
 			\includegraphics[width=0.49\textwidth]{f2a.pdf}
-			\includegraphics[width=0.469\textwidth]{f2b.pdf}
+			\includegraphics[width=0.49\textwidth]{f2b.pdf}
 		}

### 2022-03-16 papers/ms.tex
update ms: Gaia selection procedure

--- 2022-03-08 papers/ms.tex
@@
 
+\newcommand{\npms}{1097} % 20220311_Kerr_SPYGLASS205_Members_All.csv
+
 %
@@
 
-The discovery of young transiting planets is a major frontier in
-current exoplanet research.  The reason is that the properties of
-these planets can help constrain the timescales for processes
-including hot Jupiter migration (CITE), the cooling and contraction of
-mini-Neptunes (CITE Rizzuto), the early evolution of close-in
-multiplanet systems (CITE David), and whether and how mass-loss
-explains the radius valley (e.g., CITE, CITE, Lee+21 on theory; CITE
-CITE CITE on observations).
+The discovery and characterization of transiting planets younger than
+a billion years is a major frontier in current exoplanet research.
+The reason is that the properties of young planets provide benchmarks
+for studies of planetary evolution.  For instance, there are the
+questions of when hot Jupiters arrive on their close-in orbits
+\citep{dawson_johnson_2018}, how the sizes of planets with massive
+gaseous envelopes evolve \citep{rizzuto_tess_2020}, when and if
+close-in multiplanet systems fall out of resonance
+\citep{arevalo_stability_2022,goldberg_architectures_2022},
+and whether and how mass-loss explains the radius valley
+\citep{lopez_how_2012,Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018,lee_primordial_2021}.
 
-%FIXME WORDING OF THIS PARAGRAPH NEEDS WORK.  ALSO, WHERE DO YOU PUT
-%THE CONTEXT RE CLUSTERS?
-Understanding the formation and evolution of stellar clusters is
-similarly an emerging frontier in analyses of star formation and
-cluster evolution.  In particular, it is understood that most embedded
-clusters disperse within a few tens of millions of years after forming
-their first stars (CITE).  The discovery of tidal tails (CITE CITE)...
-streams (CITE Meingast18, Curtis19), etc...  Connect to moving groups
-(CITE Gagne 21).  What are the typical densities for star formation?
-How does gas expulsion and the SFE affect cluster morphology?  {\bf
-FIXME}
+The discovery of a young planet requires two claims to be fulfilled:
+the planet must exist, and its age must be secured.  Spaced-based
+photometry from K2 and TESS has yielded a number of young planets for
+which the planetary evidence comes from transits, and the age evidence
+is based on either cluster membership \citep[{\it
+e.g.},]{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
+or else on correlates of youth such as stellar rotation, photospheric
+lithium abundances, x-ray activity, or emission line strength
+\citep[{\it e.g.},]{zhou_2021_tois,hedges_toi-2076_2021}.
 
-Assuming a constant star formation rate in the Galaxy, $\approx$1\% of
-stars are $\lesssim$10$^8$ years old.  While $\approx$5{,}000
-transiting exoplanets are known (CITE), 11 currently meet this age
-cut.  The discrepancy between the $\approx$50 expected and $\approx$10
-known likely has a contribution from selection effects in planetary
-detection, which include photometric and spectroscopic starspot
-modulation, and the rarity of young stars.
+In this work, we leverage recent analyses of the Gaia data, which have
+greatly expanded our knowledge of stellar group memberships
+\citep[{\it e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.
+Broadly speaking these analyses cluster on the stellar positions and
+on-sky velocities measured by Gaia, with varying degrees of filtering
+and supervision.  One important result is the identification of
+diffuse streams and tidal tails comparable in stellar mass to the
+previously known cores of many open clusters
+\citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  These
+spatially diffuse groups typically have velocity dispersions of
+$\sim$1\kms, and can be verified to be ``co-eval'' through analyses of
+color-absolute magnitude diagrams, stellar rotation periods
+\citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
+\citep{hawkins_2020}.  While there are many implications for our
+understanding of star formation and cluster evolution \citep[{\it
+e.g.}][]{dinnbier_tidal_2020}, a more immediate consequence is that we
+now know the ages of many more stars, including previously known
+planet hosts.
 
-While detecting planets around young and active stars is challenging,
-so is measuring stellar ages (see CITE Soderblom for a review).  It
-should therefore not be surprising that advances along either
-dimension can yield advances in the census of young transiting
-planets.
+%Assuming a constant star formation rate in the Galaxy, $\approx$1\% of
+%stars are $\lesssim$10$^8$ years old.  While $\approx$5{,}000
+%transiting exoplanets are known (CITE), 11 currently meet this age
+%cut.  The discrepancy between the $\approx$50 expected and $\approx$10
+%known likely has a contribution from selection effects in planetary
+%detection, which include photometric and spectroscopic starspot
+%modulation, and the rarity of young stars.
 
-The prime Kepler mission (CITE) found most of the currently known
-transiting exoplanets, and it was conducted before the positions,
-velocities, and stellar brightnesses measured by Gaia revolutionized
-our understanding of stellar clusters (CITE, CITE, CITE).  This
-revolution is slated to continue with the upcoming third Gaia data
-release (DR3 CITE).  Nonetheless, it seems sensible to revisit the
-Kepler field, given our new knowlege of the stellar ages.
+% While detecting planets around young and active stars is challenging,
+% so is measuring stellar ages (see CITE Soderblom for a review).  It
+% should therefore not be surprising that advances along either
+% dimension can yield advances in the census of young transiting
+% planets.
 
-In this work, we expand on a previous study of a 38 million year (Myr)
-old Neptune-sized planet in the Kepler field (Kepler-1627Ab; CITE).
-The age of this planet was derived based on its host star's membership
-in the $\delta$\ Lyr\ cluster.  Our analysis of this cluster focused
-on the immediate kinematic vicinity of Kepler-1627A in order to
-reliably confirm the age of the planet.  However it also became clear
-that the $\delta$\ Lyr\ cluster is spatially and kinematically close
-to a much larger group of similarly aged stars.  This group, which is
-at a distance of roughly 300 parsecs, appears to span Cepheus
-($l=100^\circ$) to Hercules ($l=40^\circ$), at galactic latitudes,
-$b$, between roughly 0 and 20 degrees.  We therefore refer to it as
-the Cep-Her complex.  Viewed top-down in the plane of the Galaxy, the
-complex spans roughly 250 parsecs in the direction toward the galactic
-center, and 150 parsecs in the perpendicular direction of galactic
-rotation.  A detailed kinematic analysis and exploration is currently
-being prepared by R.~Kerr and collaborators.
+The prime Kepler mission \citep{borucki_kepler_2010} found most of the
+currently known transiting exoplanets, and it was conducted before
+Gaia.  It therefore seems sensible to revisit the Kepler field, given
+our new knowlege of the stellar ages.
 
-The subject of this study is the intersection of the Cep-Her complex
-with the Kepler field, and an analysis of the consequent transiting
-planets and planet candidates.  Cross-matching the full set of
-candidate Cep-Her members against all known Kepler Objects of Interest
-(CITE) yielded XX candidate matches.  Given the purported XX Myr age
-of the group, a visual inspection of the light curves quickly enabled
-a decision about whether the matches could be valid, based on the
-presence or non-presence of stellar rotation signals of the
-appropriate periods and amplitudes (CITE REBULL).  Four possible
-matches remained from this exercise: Kepler-1627, Kepler-1643,
-KOI-7368, and KOI-7913.  Given our previous analysis of Kepler-1627,
-we will focus for the most part on the latter three.
+In this work, we expand on our previous study of a $38^{+7}_{-6}$
+million year old Neptune-sized planet in the Kepler field (Kepler-1627
+Ab; \citealt{bouma_kep1627_2022}).  The age of this planet was derived
+based on its host star's membership in the $\delta$\ Lyr\ cluster.
+Our analysis of the cluster focused on the immediate spatial and
+kinematic vicinity of Kepler-1627~A in order to reliably confirm the
+age of the planet.  However it became clear that the $\delta$\ Lyr\
+cluster seems to also be part of a much larger group of similarly aged
+stars.  This group, which is at a distance of $\approx$300 parsecs
+from the Sun, appears to span Cepheus to Hercules (galactic
+longitudes, $l$, between 40$^\circ$ and 100$^\circ$), with galactic
+latitudes spanning roughly 0 and 20 degrees.  We therefore refer to it
+as the Cep-Her complex.  It exhibits significant sub-structure over
+its $\approx$250 parsec length, and a detailed analysis that delves
+into the memberships, kinematics, and possible origin of the group is
+currently being prepared by R.~Kerr and collaborators.
+
+Here, our focus is on the intersection of the Cep-Her complex with the
+Kepler field.  Cross-matching the full set of candidate Cep-Her
+members against known Kepler Objects of Interest
+\citep{thompson_planetary_2018} yielded four candidate cluster
+members: Kepler-1627, Kepler-1643, KOI-7368, and KOI-7913.  Given our
+previous analysis of Kepler-1627, we will mostly focus on the latter
+three.  After analyzing the relevant properties of Cep-Her
+(Section~\ref{sec:cluster}), we discuss the stellar properties
+(Section~\ref{sec:stars}) and validate the planetary nature of each
+system using a combination of the Kepler photometry and
+high-resolution imaging (Section~\ref{sec:planets}).  We conclude with
+a discussion of implications for the size-evolution of close-in
+mini-Neptunes (Section~\ref{sec:disc}).
 
@@
 \label{sec:cluster}
-
-\subsection{Selecting Cluster Members}
-\label{sec:kinematicselection}
 
@@
 	\caption{
-    {\bf Galactic positions  and tangential velocities of stars in the
-    complex.}  {\bf This is a placeholder before we get real kinematics.}
-    Sub-clusters include the $\delta$ Lyr cluster, RSG-5, and the
-    worryingly diffuse ``CH-2''.
+  {\bf Positions and velocities of candidate members of the Cep-Her
+  complex.}
+  {\it Top row}: On-sky positions in galactic coordinates.  Black
+  points are stars for which group membership is more likely than for
+  gray points (see Section~\ref{subsec:members}).
+  Kepler-1627 is in the outskirts of the $\delta$ Lyr cluster
+  \citep{bouma_kep1627_2022}, which is centered at $\{ l, b\} \approx
+  \{ 66^\circ, 12^\circ\}$.
+  {\it Middle row}: Galactic positions.  The Sun is at $\{X, Y, Z\} =
+  \{0, 0, 20.8\}$ pc.
+  {\it Bottom row}: Galactic tangential velocities (left) and
+  longitudinal tangential velocity versus galactic latitude (right).
+  The gray band in the lower-right shows the $\pm$1-$\sigma$
+  projection of the Solar velocity with respect to the local standard
+  of rest.
+  There is a strong spatial and kinematic overlap between Kepler-1643
+  and RSG-5 (magenta points; $\{ l, b\} \approx \{ 82^\circ,
+  6^\circ\}$; $\{ X, Y, Z\} \approx \{-8070, 330, 50\} $ pc).  The
+  local population of candidate young stars around KOI-7368 and
+  KOI-7913 is more diffuse -- we call this region ``CH-2'' (lime green
+  points).  The green and purple points are manually selected from the
+  initial candidate population based on galactic position and
+  tangential velocity.
 	\label{fig:XYZvtang}
@@
 
+\subsection{Previous Related Work}
+
 Our focus is on a region of the Galaxy roughly 200 to 500 parsecs from
 the Sun, above the galactic plane, and spanning galactic longitudes of
-roughly $40^\circ$ to $100^\circ$ degrees.  A direct query of the Gaia
-archive imposing these constraints yields $\approx3.4\times10^7$
-sources.  Imposing the cuts suggested by \citet[][Appendix
-B]{GaiaCollaboration2018} on the photometric signal to noise ratio,
-the number of visibility periods used, the astrometric $\chi^2$ of the
-single-source solution, the $G_{\rm BP}-G_{\rm RP}$ color excess
-factor and the parallax precision ($\varpi/\sigma_\varpi>5$) yields
-$\approx5.0\times10^6$ sources.
+roughly $40^\circ$ to $100^\circ$ degrees.  Two of the richest
+clusters in this region are the $\delta$~Lyra cluster
+\citep{stephenson_possible_1959} and RSG-5 \citep{roser_nine_2016}.
+Both of these clusters were known before Gaia, and have reported ages
+between $\approx$30 and $\approx$60 million years.  Early
+empirical evidence that these two clusters might be part of a broader
+population was apparent in the Gaia-based photometric analysis of
+pre-main-sequence stars by \citet[][see their Figures~11 and~13]{Zari2018}.
+Further kinematic connections and complexity were highlighted by
+\citet{KounkelCovey2019}, when defining the structures they dubbed
+``Theia~73'' and ``Theia~96''\footnote{A useful visualization is
+available online at \url{http://mkounkel.com/mw3d/mw2d.html} (accessed
+15 March 2022)}.  This connection by \citet{KounkelCovey2019} was made
+as part of an unsupervised clustering analysis of the Gaia DR2
+positions and tangential velocities with a subsequent manual
+``stitching'' step, and generally supports the idea that there is an
+overdensity of $30$--$60\,{\rm Myr}$ old stars in this region of the
+Galaxy.  \citet{Kerr2021}, in a volume-limited analysis of the Gaia
+DR2 point-source catalog out to one third of a kiloparsec, identified
+three of the nearest sub-populations, dubbed ``Cepheus-Cygnus'',
+``Lyra'', and ``Cerberus''.  \citet{Kerr2021} reported ages for each
+of these subgroups between 30 and 35 million years.
 
-Imposing a thin slice in color-absolute magnitude space around the
-pre-main-sequence of the $\delta$ Lyr cluster's previously established
-$\approx$40 Myr locus, and imposing kinematics cut for $-8 < v_b /
-{\rm km\,s^{-1}} < 0$ and $-15 < v_{l'} / {\rm km\,s^{-1}} < 15$,
-where $l'=l\cos b$, yields $\approx 5.2\times10^3$ sources.  These
-objects are all in the range BP-RP 1.5 to 3.2, so roughly K5V to
-M4.5V, or $\approx$0.2 to 0.7 solar masses.  Much more manageable, but
-according to IMF arguments, missing at least a factor of two in mass
-of any clusters.
+\subsection{Member Selection}
+\label{subsec:members}
 
-In XYZ, and in tangential velocity space, these five thousand sources
-are visually clustered in between a few to a dozen groups, depending
-on the characteristic size and velocity scales of interest.  The first
-suggestion that these objects could all be part of a monolithic
-structure, to our knowledge, came from the manual grouping that was
-performed between them by \citet{KounkelCovey2019} when defining the
-structure they dubbed ``Theia~73''.  Different subsets of these stars
-have however previously been noticed and studied.  They include the
-$\delta$ Lyr cluster, also known as Stephenson-1
-\citep{stephenson_possible_1959}, which comprises a major grouping
-around $(l,b)=(68^\circ,15^\circ)$, with typical velocities $(v_{l'},
-v_b)=(-4.5{\rm \kms},-4{\rm \kms})$.  Another individual component
-that has previously been recognized is RSG-5 \citep{roser_nine_2016}.
-This group is at roughly $(l,b)=(83^\circ,6^\circ)$, with velocities
-$(v_{l'}, v_b)=(+5.5{\rm \kms},-3.5{\rm \kms})$.  Most of the other
-subclusters, including CH-2, and those in Cep-Cyg and Cerberus are too
-small or dispersed to have previously been analyzed in great detail.
+{\bf TODO RONAN: PLEASE VET ENTIRE SECTION, AND UPDATE WHERE
+APPROPRIATE!}.
+
+The possibility that the $\delta$~Lyr cluster, RSG-5, and a few nearby
+sub-populations in this quadrant of the sky share a common origin has
+yet to be fully substantiated.  To further explore the population of
+$10$--$100\,{\rm Myr}$ stars in this region, we select candidate
+members through four steps, the first three being identical to those
+described in Section~3 of \citet{Kerr2021}.  We briefly summarize them
+here.
+
+The first step is to select stars that are photometrically distinct
+from the field star population based on Gaia EDR3 magnitudes $\{G,
+G_{\rm RP}, G_{\rm BP}\}$, parallaxes and auxiliary reddening
+estimates \citep{lallement_gaia-2mass_2019}.  This step yielded \npms\
+stars, which are either  pre-main-sequence K and M dwarfs due to their
+long contraction timescales, or massive stars near the zero-age main
+sequence due to their rapid evolutionary timescales.  The second step
+is to then perform in an unsupervised HDBScan clustering on this
+photometrically-young population
+\citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The specific
+parameters we use in this clustering analysis are $\{ X, Y, Z, c v_b,
+c v_{l^*} \} $, where $c$ is the size-velocity corrective factor,
+which is taken as $c=6\,{\rm pc / km\,s}^{-1}$ to ensure that the
+spatial and velocity scales have identical standard deviations.  As
+input parameters, we set the minimum $\epsilon$ threshold past which
+clusters cannot be fragmented as $25$ in units of parsecs, and $c$
+\kms\ in units of velocity.  The minimum cluster size $N$ is set to
+10, as is $k$, the parameter used to define the ``core distance''
+density metric.  This unsupervised clustering in our case
+yielded 8 distinct groups.
+These groups are then used as the ``seed'' populations for the third
+step, which is to search for objects that are spatially and
+kinematically consistent with each of the groups, {\bf WORDING NEEDS
+WORK FOR CLARITY at least as close to the 10$^{\rm th}$ nearest HDBSCAN-identified
+member in space-velocity coordinates as the most peripheral
+HDBSCAN-identified member}.
+The latter step yields F, G, and K-type stars that are spatially and
+kinematically close to the photometrically-young stars, but which
+cannot be identified as young based on their positions in color versus
+absolute magnitude.
+
+The outcome of this analysis is shown in Figure~\ref{fig:XYZvtang}.
+We computed positions in these plots assuming the \texttt{astropy
+v4.0} coordinate standard \citep{astropy_2018}, which assumes that the
+Sun is $8122$ pc from the galactic center.  It also assumes
+the solar velocity with respect to the local standard of rest from \citet{schonrich_local_2010}.
+The previously known $\delta$~Lyr\ cluster ($l,b=68^\circ,15^\circ$;
+$v_{l'}, v_b=-4.5{\rm kms},-4{\rm kms}$) is visible, as is
+RSG-5 ($l,b=83^\circ,6^\circ$, $v_{l'}, v_b=+5.5{\rm kms},-3.5{\rm
+kms}$).  Most of the other subclusters, including CH-2, and those in
+Cep-Cyg ($l,b=90^\circ,7^\circ$) and Cerberus
+($l,b=48^\circ,18^\circ$) are too small or dispersed to have
+previously been analyzed in great detail.
+
+To enable a selection cut that filters out field-star contaminants, we
+also compute a weight metric, defined such that the group member
+with the smallest core distance has a weight of 1, the group member
+with the greatest core distance has a weight of 0, and weights for the
+other group members are linearly spaced between these two extremes.
+For Figure~\ref{fig:XYZvtang}, we show 26{,}960 objects with weight
+exceeding 0.02 as gray points, and overplot 7{,}560 objects with
+weights exceeding 0.10 as black points.
+
+% FIXME FIXME ANYTHING ELSE TO SAY?
+{\bf TODO: describe final step: manual selection of RSG-5 and CH-2 populations}
+
+%Querying the Gaia archive
+%and imposing the cuts suggested by \citet[][Appendix
+%B]{GaiaCollaboration2018} on the photometric signal to noise ratio,
+%the number of visibility periods used, the astrometric $\chi^2$ of the
+%single-source solution, the $G_{\rm BP}-G_{\rm RP}$ color excess
+%factor and the parallax precision ($\varpi/\sigma_\varpi>5$) yields
+%$\approx5.0\times10^6$ sources.
+%
+%Imposing a thin slice in color-absolute magnitude space around the
+%pre-main-sequence of the $\delta$ Lyr cluster's previously established
+%$\approx$40 Myr locus, and imposing kinematics cut for $-8 < v_b /
+%{\rm km\,s^{-1}} < 0$ and $-15 < v_{l'} / {\rm km\,s^{-1}} < 15$,
+%where $l'=l\cos b$, yields $\approx 5.2\times10^3$ sources.  These
+%objects are all in the range BP-RP 1.5 to 3.2, so roughly K5V to
+%M4.5V, or $\approx$0.2 to 0.7 solar masses.  Much more manageable, but
+%according to IMF arguments, missing at least a factor of two in mass
+%of any clusters.
+%
+%In XYZ, and in tangential velocity space, these five thousand sources
+%are visually clustered in between a few to a dozen groups, depending
+%on the characteristic size and velocity scales of interest.
 

### 2022-03-16 papers/ms.tex
update ms

@@
   {\it Top row}: On-sky positions in galactic coordinates.  Black
-  points are stars for which group membership is more likely than for
+  points are stars for which group membership is more secure than for
   gray points (see Section~\ref{subsec:members}).
@@
   There is a strong spatial and kinematic overlap between Kepler-1643
-  and RSG-5 (magenta points; $\{ l, b\} \approx \{ 82^\circ,
-  6^\circ\}$; $\{ X, Y, Z\} \approx \{-8070, 330, 50\} $ pc).  The
+  and RSG-5 ($\{ l, b\} \approx \{ 82^\circ,
+  6^\circ\}$; $\{ X, Y, Z\} \approx \{60, 330, 50\} $ pc).  The
   local population of candidate young stars around KOI-7368 and
-  KOI-7913 is more diffuse -- we call this region ``CH-2'' (lime green
-  points).  The green and purple points are manually selected from the
-  initial candidate population based on galactic position and
-  tangential velocity.
+  KOI-7913 is more diffuse -- we call this region ``CH-2'',
+  and treat it with additional skepticism
+  The {\bf interactive figure} enables overplotting the candidate
+  cluster members.
 	\label{fig:XYZvtang}
@@
 long contraction timescales, or massive stars near the zero-age main
-sequence due to their rapid evolutionary timescales.  The second step
-is to then perform in an unsupervised HDBScan clustering on this
-photometrically-young population
+sequence due to their rapid evolutionary timescales.
+
+The second step is to then perform in an unsupervised HDBScan
+clustering on the photometrically selected population
 \citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The specific
@@
 which is taken as $c=6\,{\rm pc / km\,s}^{-1}$ to ensure that the
-spatial and velocity scales have identical standard deviations.  As
-input parameters, we set the minimum $\epsilon$ threshold past which
-clusters cannot be fragmented as $25$ in units of parsecs, and $c$
-\kms\ in units of velocity.  The minimum cluster size $N$ is set to
-10, as is $k$, the parameter used to define the ``core distance''
-density metric.  This unsupervised clustering in our case
-yielded 8 distinct groups.
+spatial and velocity scales have identical standard deviations.
+Positions are computed assuming the \texttt{astropy v4.0} coordinate
+standard \citep{astropy_2018}, which places the Sun $8122$ pc from the
+galactic center, and assumes the solar velocity with respect to the
+local standard of rest from \citet{schonrich_local_2010}.  As input
+parameters to HDBScan, we set the minimum $\epsilon$ threshold past
+which clusters cannot be fragmented as $25$ parsecs in physical space,
+and $c$ \kms\ in velocity.  The minimum cluster size $N$ is set to 10,
+as is $k$, the parameter used to define the ``core distance'' density
+metric.
+
+This unsupervised clustering in our case yielded 8 distinct groups.
 These groups are then used as the ``seed'' populations for the third
@@
 kinematically consistent with each of the groups, {\bf WORDING NEEDS
-WORK FOR CLARITY at least as close to the 10$^{\rm th}$ nearest HDBSCAN-identified
-member in space-velocity coordinates as the most peripheral
-HDBSCAN-identified member}.
-The latter step yields F, G, and K-type stars that are spatially and
-kinematically close to the photometrically-young stars, but which
-cannot be identified as young based on their positions in color versus
-absolute magnitude.
+WORK FOR CLARITY at least as close to the 10$^{\rm th}$ nearest
+HDBSCAN-identified member in space-velocity coordinates as the most
+peripheral HDBSCAN-identified member}.  This third step yields F, G,
+and K-type stars that are spatially and kinematically close to the
+photometrically-young stars, but which cannot be identified as young
+based on their positions in color versus absolute magnitude.
 
-The outcome of this analysis is shown in Figure~\ref{fig:XYZvtang}.
-We computed positions in these plots assuming the \texttt{astropy
-v4.0} coordinate standard \citep{astropy_2018}, which assumes that the
-Sun is $8122$ pc from the galactic center.  It also assumes
-the solar velocity with respect to the local standard of rest from \citet{schonrich_local_2010}.
-The previously known $\delta$~Lyr\ cluster ($l,b=68^\circ,15^\circ$;
-$v_{l'}, v_b=-4.5{\rm kms},-4{\rm kms}$) is visible, as is
-RSG-5 ($l,b=83^\circ,6^\circ$, $v_{l'}, v_b=+5.5{\rm kms},-3.5{\rm
-kms}$).  Most of the other subclusters, including CH-2, and those in
-Cep-Cyg ($l,b=90^\circ,7^\circ$) and Cerberus
-($l,b=48^\circ,18^\circ$) are too small or dispersed to have
-previously been analyzed in great detail.
-
+The outcome of from the analysis up to the point of the third step is
+shown in Figure~\ref{fig:XYZvtang}.  
 To enable a selection cut that filters out field-star contaminants, we
@@
 other group members are linearly spaced between these two extremes.
-For Figure~\ref{fig:XYZvtang}, we show 26{,}960 objects with weight
+In Figure~\ref{fig:XYZvtang}, we show 26{,}960 objects with weight
 exceeding 0.02 as gray points, and overplot 7{,}560 objects with
@@
 
-% FIXME FIXME ANYTHING ELSE TO SAY?
-{\bf TODO: describe final step: manual selection of RSG-5 and CH-2 populations}
+The previously known $\delta$~Lyr\ cluster ($l,b=68^\circ,15^\circ$;
+$v_{l'}, v_b=-4.5{\rm kms},-4{\rm kms}$) is visible, as is RSG-5
+($l,b=83^\circ,6^\circ$, $v_{l'}, v_b=+5.5{\rm kms},-3.5{\rm kms}$).
+Most of the other subclusters, including in Cep-Cyg
+($l,b=90^\circ,7^\circ$) and Cerberus ($l,b=48^\circ,18^\circ$) are
+too small or dispersed to have previously been analyzed in great
+detail.
+
+Finally, there is a final step.
+This paper is actually about planets, not this insane stellar
+population.
+We cross-matched
+our candidate Cep-Her member list against 
+all known Kepler Objects of Interest, including known false positives
+\citep{thompson_planetary_2018}.
+This yields 32 matches, of which 5 are ``confirmed''
+planets according to the NASA Exoplanet Archive.
+A cursory inspection of the light curves and RoboVetter (CITE)
+classifications for these objects quickly revealed whether they were
+consistent with being members of a $\lesssim 10^8$ year old
+population.
+Four candidate cluster members remained after this inspection:
+Kepler-1627, Kepler-1643, KOI-7368, and KOI-7913.
+
+Figure~\ref{fig:XYZvtang}
+
+CH-2:
+X: 20 to 80
+Y: 230 to 270
+Z: 70 to 105
+vb: -1.5 to 3.5 kms
+vl: 1 to 6 kms
+
+RSG-5:
+X: 45 to 75
+Y: 320 to 350
+Z: 40 to 70
+vl: 4 to 6 kms
+vb: -3 to -4 kms
 

### 2022-03-16 papers/ms.tex
update ms for auto Cep-Her groups

@@
 Kepler field.  Cross-matching the full set of candidate Cep-Her
-members against known Kepler Objects of Interest
+members against known Kepler Objects of Interest (KOIs)
 \citep{thompson_planetary_2018} yielded four candidate cluster
@@
   local population of candidate young stars around KOI-7368 and
-  KOI-7913 is more diffuse -- we call this region ``CH-2'',
-  and treat it with additional skepticism
-  The {\bf interactive figure} enables overplotting the candidate
-  cluster members.
+  KOI-7913 is more diffuse -- we call this region ``CH-2''.
+  The {\bf interactive figure} enables a few different cuts to be
+  shown.
 	\label{fig:XYZvtang}
@@
 
-Our focus is on a region of the Galaxy roughly 200 to 500 parsecs from
-the Sun, above the galactic plane, and spanning galactic longitudes of
-roughly $40^\circ$ to $100^\circ$ degrees.  Two of the richest
+Our focus is on a region of the Galaxy approximately 200 to 500
+parsecs from the Sun, above the galactic plane, and spanning galactic
+longitudes of roughly $40^\circ$ to $100^\circ$ degrees.  Two rich
 clusters in this region are the $\delta$~Lyra cluster
@@
 Both of these clusters were known before Gaia, and have reported ages
-between $\approx$30 and $\approx$60 million years.  Early
-empirical evidence that these two clusters might be part of a broader
-population was apparent in the Gaia-based photometric analysis of
-pre-main-sequence stars by \citet[][see their Figures~11 and~13]{Zari2018}.
-Further kinematic connections and complexity were highlighted by
-\citet{KounkelCovey2019}, when defining the structures they dubbed
-``Theia~73'' and ``Theia~96''\footnote{A useful visualization is
-available online at \url{http://mkounkel.com/mw3d/mw2d.html} (accessed
-15 March 2022)}.  This connection by \citet{KounkelCovey2019} was made
-as part of an unsupervised clustering analysis of the Gaia DR2
-positions and tangential velocities with a subsequent manual
-``stitching'' step, and generally supports the idea that there is an
-overdensity of $30$--$60\,{\rm Myr}$ old stars in this region of the
-Galaxy.  \citet{Kerr2021}, in a volume-limited analysis of the Gaia
-DR2 point-source catalog out to one third of a kiloparsec, identified
+between $\approx$30 and $\approx$60 million years.  Early empirical
+evidence that these two clusters are likely part of a large and more
+diffuse population was apparent in the Gaia-based photometric analysis
+of pre-main-sequence stars by \citet[][see their Figures~11
+and~13]{Zari2018}.  Further kinematic connections and complexity were
+highlighted by \citet{KounkelCovey2019}, who included these previously
+known groups in the larger structures dubbed ``Theia~73'' and
+``Theia~96''\footnote{A useful visualization is available online at
+\url{http://mkounkel.com/mw3d/mw2d.html} (accessed 15 March 2022)}.
+The connection made by \citet{KounkelCovey2019} between the previously
+known open clusters and the other groups in the region was made as
+part of an unsupervised clustering analysis of the Gaia DR2 positions
+and tangential velocities with a subsequent manual ``stitching'' step,
+and generally supports the idea that there is an overdensity of
+$30$--$60\,{\rm Myr}$ old stars in this region of the Galaxy.
+\citet{Kerr2021}, in a volume-limited analysis of the Gaia DR2
+point-source catalog out to one third of a kiloparsec, identified
 three of the nearest sub-populations, dubbed ``Cepheus-Cygnus'',
@@
 
+The possibility that the $\delta$~Lyr cluster, RSG-5, and a few nearby
+sub-populations in this quadrant of the sky share a common origin has
+yet to be fully substantiated, and will be the subject of an upcoming
+study by R.~Kerr and collaborators.  Our primary interest in the
+region stems from the fact that a portion of it was observed by Kepler
+(Figure~\ref{fig:XYZvtang}, top panel).  To further explore the
+population of $10$--$100\,{\rm Myr}$ stars that were observed, we
+select candidate Cep-Her members through four steps, the first three
+being identical to those described in Section~3 of \citet{Kerr2021}.
+We briefly summarize them here.
+
 {\bf TODO RONAN: PLEASE VET ENTIRE SECTION, AND UPDATE WHERE
 APPROPRIATE!}.
-
-The possibility that the $\delta$~Lyr cluster, RSG-5, and a few nearby
-sub-populations in this quadrant of the sky share a common origin has
-yet to be fully substantiated.  To further explore the population of
-$10$--$100\,{\rm Myr}$ stars in this region, we select candidate
-members through four steps, the first three being identical to those
-described in Section~3 of \citet{Kerr2021}.  We briefly summarize them
-here.
 
@@
 estimates \citep{lallement_gaia-2mass_2019}.  This step yielded \npms\
-stars, which are either  pre-main-sequence K and M dwarfs due to their
-long contraction timescales, or massive stars near the zero-age main
-sequence due to their rapid evolutionary timescales.
+stars with high-quality photometric and astrometry, which are either
+pre-main-sequence K and M dwarfs due to their long contraction
+timescales, or massive stars near the zero-age main sequence due to
+their rapid evolutionary timescales.
 
@@
 clustering on the photometrically selected population
-\citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The specific
+\citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The
 parameters we use in this clustering analysis are $\{ X, Y, Z, c v_b,
@@
 step, which is to search for objects that are spatially and
-kinematically consistent with each of the groups, {\bf WORDING NEEDS
-WORK FOR CLARITY at least as close to the 10$^{\rm th}$ nearest
+kinematically consistent with each of the groups, {\bf IMPROVE
+CLARITY at least as close to the 10$^{\rm th}$ nearest
 HDBSCAN-identified member in space-velocity coordinates as the most
-peripheral HDBSCAN-identified member}.  This third step yields F, G,
-and K-type stars that are spatially and kinematically close to the
+peripheral HDBSCAN-identified member}.  This third step yields stars
+that are spatially and kinematically close to the
 photometrically-young stars, but which cannot be identified as young
@@
 
-The outcome of from the analysis up to the point of the third step is
-shown in Figure~\ref{fig:XYZvtang}.  
-To enable a selection cut that filters out field-star contaminants, we
-also compute a weight metric, defined such that the group member
-with the smallest core distance has a weight of 1, the group member
-with the greatest core distance has a weight of 0, and weights for the
-other group members are linearly spaced between these two extremes.
-In Figure~\ref{fig:XYZvtang}, we show 26{,}960 objects with weight
-exceeding 0.02 as gray points, and overplot 7{,}560 objects with
-weights exceeding 0.10 as black points.
+The outcome of the analysis up to the point of the third step is
+shown in Figure~\ref{fig:XYZvtang}.  To enable a selection cut that
+filters out field-star contaminants, we also compute a weight metric,
+defined such that the group member with the smallest core distance has
+a weight of 1, the group member with the greatest core distance has a
+weight of 0, and weights for the other group members are log-normally
+distributed between these two extremes.  In Figure~\ref{fig:XYZvtang}, we
+show 26{,}960 objects with weight exceeding 0.02 as gray points, and
+overplot 7{,}560 objects with weights exceeding 0.10 as black points.
+{\bf TODO: FIXME these object numbers impose pretty weak quality cuts and include junk at G>18... 
+use get\_clean\_gaia\_photometric\_sources too, since they come in for
+the CAMDs}.
 
@@
 
-Finally, there is a final step.
-This paper is actually about planets, not this insane stellar
-population.
-We cross-matched
-our candidate Cep-Her member list against 
-all known Kepler Objects of Interest, including known false positives
-\citep{thompson_planetary_2018}.
-This yields 32 matches, of which 5 are ``confirmed''
-planets according to the NASA Exoplanet Archive.
-A cursory inspection of the light curves and RoboVetter (CITE)
-classifications for these objects quickly revealed whether they were
-consistent with being members of a $\lesssim 10^8$ year old
-population.
-Four candidate cluster members remained after this inspection:
-Kepler-1627, Kepler-1643, KOI-7368, and KOI-7913.
+The fourth and final step was cross-matching our candidate Cep-Her
+member list against all known Kepler Objects of Interest, including
+known false positives \citep{thompson_planetary_2018}.  This yields 32
+matches, of which 5 are ``confirmed'' planets according to the NASA
+Exoplanet Archive.  A cursory inspection of the Kepler data validation
+summaries and Robovetter classifications for these objects quickly
+showed whether they were potentially consistent with being {\it i)}
+planets, and {\it ii)} $\lesssim 10^8$ years old population, based on
+the presence of rotational modulation at the expected period and
+amplitude \citep[{\it e.g.},][Figure~9]{rebull_rotation_2020}.  Four
+objects remained after this inspection: Kepler-1627, Kepler-1643,
+KOI-7368, and KOI-7913.
 
-Figure~\ref{fig:XYZvtang}
+Figure~\ref{fig:XYZvtang} shows the positions of the KOIs along
+various projections.  Kepler-1643 is near the core RSG-5 population
+both spatially and kinematically.  KOI-7368 and KOI-7913 are in a more
+diffuse region $\approx$40 parsecs above RSG-5 in $Z$ and $\approx$100
+parsecs closer to the Sun in $Y$.  In tangential galactic velocity
+space, there may be some kinematic overlap between the region the
+latter two KOIs are in, and the main RSG-5 group.
 
-CH-2:
-X: 20 to 80
-Y: 230 to 270
-Z: 70 to 105
-vb: -1.5 to 3.5 kms
-vl: 1 to 6 kms
+We define two sets of stars in the local vicinity of our objects of
+interest, in order to verify whether they are indeed the same age.
+For candidate RSG-5 members, we require:
+\begin{align}
+  45<X/{\rm pc}<75 \nonumber \\
+  320<Y/{\rm pc}<350 \nonumber \\
+  40<Z/{\rm pc}<70 \nonumber \\
+  -4<v_b/{\rm km\,s^{-1}}<-3 \nonumber \\
+  4<v_{l^*}/{\rm km\,s^{-1}}<6 \nonumber
+\end{align}
+For the diffuse stars near KOI-7368 and KOI-7913, we require
+\begin{align}
+	20<X/{\rm pc}<70 \nonumber \\
+	230<Y/{\rm pc}<270 \nonumber \\
+	75<Z/{\rm pc}<105 \nonumber \\
+	-3.5<v_b/{\rm km\,s^{-1}}<-1.5 \nonumber \\
+	2<v_{l^*}/{\rm km\,s^{-1}}<6, \nonumber
+\end{align}
+and we call this latter set of stars ``CH-2''.
+% FIXME TODO numbers are from RSG-5_auto_XYZ_vl_vb_cut.csv,
+% CH-2_auto_XYZ_vl_vb_cut.csv right now.  Can we do better?
+These cuts yield 141 candidate RSG-5 members, and 
+37 candidate CH-2 members.
+Given the larger volume of the latter group, and the more diffuse base
+sample of stars, a major concern is contamination by field stars.
+We assess this in the following section.
 
-RSG-5:
-X: 45 to 75
-Y: 320 to 350
-Z: 40 to 70
-vl: 4 to 6 kms
-vb: -3 to -4 kms
-
-%Querying the Gaia archive
-%and imposing the cuts suggested by \citet[][Appendix
-%B]{GaiaCollaboration2018} on the photometric signal to noise ratio,
-%the number of visibility periods used, the astrometric $\chi^2$ of the
-%single-source solution, the $G_{\rm BP}-G_{\rm RP}$ color excess
-%factor and the parallax precision ($\varpi/\sigma_\varpi>5$) yields
-%$\approx5.0\times10^6$ sources.
+%Querying the Gaia archive and imposing the cuts suggested by
+%\citet[][Appendix B]{GaiaCollaboration2018} on the photometric signal
+%to noise ratio, the number of visibility periods used, the
+%astrometric $\chi^2$ of the single-source solution, the $G_{\rm
+%BP}-G_{\rm RP}$ color excess factor and the parallax precision
+%($\varpi/\sigma_\varpi>5$) yields $\approx5.0\times10^6$ sources.
 %
 %Imposing a thin slice in color-absolute magnitude space around the
-%pre-main-sequence of the $\delta$ Lyr cluster's previously established
-%$\approx$40 Myr locus, and imposing kinematics cut for $-8 < v_b /
-%{\rm km\,s^{-1}} < 0$ and $-15 < v_{l'} / {\rm km\,s^{-1}} < 15$,
-%where $l'=l\cos b$, yields $\approx 5.2\times10^3$ sources.  These
-%objects are all in the range BP-RP 1.5 to 3.2, so roughly K5V to
-%M4.5V, or $\approx$0.2 to 0.7 solar masses.  Much more manageable, but
-%according to IMF arguments, missing at least a factor of two in mass
-%of any clusters.
+%pre-main-sequence of the $\delta$ Lyr cluster's previously
+%established $\approx$40 Myr locus, and imposing kinematics cut for
+%$-8 < v_b / {\rm km\,s^{-1}} < 0$ and $-15 < v_{l'} / {\rm
+%km\,s^{-1}} < 15$, where $l'=l\cos b$, yields $\approx 5.2\times10^3$
+%sources.  These objects are all in the range BP-RP 1.5 to 3.2, so
+%roughly K5V to M4.5V, or $\approx$0.2 to 0.7 solar masses.  Much more
+%manageable, but according to IMF arguments, missing at least a factor
+%of two in mass of any clusters.
 %
@@
 \label{sec:camd}
+
+% FIXME FIXME calculate for both!
 

### 2022-03-26 papers/ms.tex
ms: push CH-2 and RSG-5 rotation periods

@@
 Given the larger volume of the latter group, and the more diffuse base
-sample of stars, a major concern is contamination by field stars.
+sample of stars, an important concern is contamination by field stars.
 We assess this in the following section.
@@
     bottom left shows gyro (and is a {\bf place-holder since we might
-    want to add CH-2 and RSG5}).  The bottom right shows lithium
-    (black points are NGC2547 and IC2602 from Randich+18 and probably
-    are not believable at the red end; for Kepler-1643 I'm less
-    confident, but it might be tied to the ``slow'' rotation period --
-    this is why we need RSG5 rotation periods).  Also, we might want
-    an H-alpha plot?).
+    want to add CH-2 and RSG5}).  
+    %The bottom right shows lithium
+    %(black points are NGC2547 and IC2602 from Randich+18 and probably
+    %are not believable at the red end; for Kepler-1643 I'm less
+    %confident, but it might be tied to the ``slow'' rotation period --
+    %this is why we need RSG5 rotation periods).  Also, we might want
+    %an H-alpha plot?).
 		\label{fig:age}

### 2022-03-27 papers/ms.tex
update ms

--- 2022-03-16 papers/ms.tex
@@
 \title{
-  Three 38 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
+  Three $\approx$40 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
 }
@@
   year old stars spanning Cepheus ($l=100^\circ$) to Hercules
-  ($l=40^\circ$), hereafter the Cep-Her complex.  The group contains
-  four known Kepler Objects of Interest:
+  ($l=40^\circ$), hereafter the Cep-Her complex.
+  The group includes four known Kepler Objects of Interest:
   Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$),
@@
   complex called the $\delta$\ Lyr\ cluster
-  \citep{bouma_kep1627_2022}.  Here we focus on the latter three
-  systems, which are in other sub-components of the complex (RSG-5 and
-  CH-2).  Based on kinematic evidence from Gaia, stellar rotation
-  periods from TESS, and spectroscopy, these three systems are also
-  $38 \pm 6$ million years old.  Based on the transit shapes and high
-  resolution imaging, they are all most
-  likely planets (false positive probabilities of $6\times10^{-9}$,
-  $5\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368,
-  and KOI-7913 respectively).  Kepler-1643 and KOI-7913 are therefore
-  the first empirical demonstration that mini-Neptunes with sizes of
-  $\approx$2 Earth radii exist at ages of roughly 40 million years.
+  \citep{bouma_kep1627_2022}.
+  Here we focus on the latter three systems, which are in other
+  sub-components of the complex (RSG-5 and CH-2).
+  Based on kinematic evidence from Gaia, stellar rotation periods from
+  TESS, and spectroscopy, these three systems are also $\approx$40
+  million years old.
+  More specifically, we find that Kepler-1643 (in RSG-5) is
+  $46^{+9}_{-7}$\,Myr old, while KOI-7368 and KOI-7913 (in CH-2) are
+  $36^+{10}_{-8}$\,Myr old.
+  Based on the transit shapes and high resolution imaging, they are
+  all most likely planets, with false positive probabilities of
+  $6\times10^{-9}$, $5\times10^{-3}$, and $1\times10^{-4}$ for
+  Kepler-1643, KOI-7368, and KOI-7913 respectively.
+  Kepler-1643 and KOI-7913 are therefore the first empirical
+  demonstration that mini-Neptunes with sizes of $\approx$2 Earth
+  radii exist at ages of $\approx$40 million years.
 \end{abstract}
@@
 close-in multiplanet systems fall out of resonance
-\citep{arevalo_stability_2022,goldberg_architectures_2022},
-and whether and how mass-loss explains the radius valley
+\citep{arevalo_stability_2022,goldberg_architectures_2022}, and
+whether and how mass-loss explains the radius valley
 \citep{lopez_how_2012,Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018,lee_primordial_2021}.
@@
 diffuse streams and tidal tails comparable in stellar mass to the
-previously known cores of many open clusters
+previously known cores of nearby open clusters
 \citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  These
 spatially diffuse groups typically have velocity dispersions of
-$\sim$1\kms, and can be verified to be ``co-eval'' through analyses of
-color-absolute magnitude diagrams, stellar rotation periods
-\citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
-\citep{hawkins_2020}.  While there are many implications for our
-understanding of star formation and cluster evolution \citep[{\it
-e.g.}][]{dinnbier_tidal_2020}, a more immediate consequence is that we
-now know the ages of many more stars, including previously known
-planet hosts.
-
-%Assuming a constant star formation rate in the Galaxy, $\approx$1\% of
-%stars are $\lesssim$10$^8$ years old.  While $\approx$5{,}000
-%transiting exoplanets are known (CITE), 11 currently meet this age
-%cut.  The discrepancy between the $\approx$50 expected and $\approx$10
-%known likely has a contribution from selection effects in planetary
-%detection, which include photometric and spectroscopic starspot
-%modulation, and the rarity of young stars.
-
-% While detecting planets around young and active stars is challenging,
-% so is measuring stellar ages (see CITE Soderblom for a review).  It
-% should therefore not be surprising that advances along either
-% dimension can yield advances in the census of young transiting
-% planets.
+$\approx$1\kms, though they can be much higher due to both projection
+effects and internal dynamics.  As an extreme example, in the Hyades
+the apparent velocities of stars in the tidal tails are expected to
+span up to $\pm 40$\kms relative to the cluster center
+\citep{jerabkova_800_2021}.  The stars in such diffuse regions can be
+verified to be the same age as the core members ({\it i.e.}, co-eval)
+through analyses of color-absolute magnitude diagrams, stellar
+rotation periods \citep{curtis_tess_2019,bouma_2021_ngc2516}, and
+chemical abundances \citep{hawkins_2020}.  While there are many
+implications for our understanding of star formation and cluster
+evolution \citep[{\it e.g.}][]{dinnbier_tidal_2020}, a more immediate
+consequence is that we now know the ages of many more stars, including
+previously known planet hosts.
 
@@
 Our analysis of the cluster focused on the immediate spatial and
-kinematic vicinity of Kepler-1627~A in order to reliably confirm the
-age of the planet.  However it became clear that the $\delta$\ Lyr\
-cluster seems to also be part of a much larger group of similarly aged
-stars.  This group, which is at a distance of $\approx$300 parsecs
-from the Sun, appears to span Cepheus to Hercules (galactic
-longitudes, $l$, between 40$^\circ$ and 100$^\circ$), with galactic
-latitudes spanning roughly 0 and 20 degrees.  We therefore refer to it
-as the Cep-Her complex.  It exhibits significant sub-structure over
-its $\approx$250 parsec length, and a detailed analysis that delves
-into the memberships, kinematics, and possible origin of the group is
-currently being prepared by R.~Kerr and collaborators.
+kinematic vicinity of Kepler-1627~A in order to confirm the age of the
+planet.  However it became clear that the $\delta$\ Lyr\ cluster seems
+to also be part of a much larger group of similarly aged stars.  This
+group, which is at a distance of $\sim$330 parsecs from the Sun,
+appears to span Cepheus to Hercules (galactic longitudes, $l$, between
+40$^\circ$ and 100$^\circ$), at galactic latitudes roughly between
+0$^\circ$ and 20$^\circ$.  We therefore refer to it as the Cep-Her
+complex.  It exhibits significant sub-structure over its $\approx$250
+parsec length, and a detailed analysis of its memberships, kinematics,
+and possible origin is currently being prepared by R.~Kerr and
+collaborators.
 
@@
 a discussion of implications for the size-evolution of close-in
-mini-Neptunes (Section~\ref{sec:disc}).
+mini-Neptunes (Section~\ref{sec:disc_conc}).
 
-\section{The Cluster}
+\section{The Cep-Her Complex}
 \label{sec:cluster}
@@
   {\it Middle row}: Galactic positions.  The Sun is at $\{X, Y, Z\} =
-  \{0, 0, 20.8\}$ pc.
+  \{0, 0, 20.8\}$\,pc; lines of constant heliocentric distance are
+  shown between 250 and 400\,pc, spaced by 50\,pc.
   {\it Bottom row}: Galactic tangential velocities (left) and
@@
   projection of the Solar velocity with respect to the local standard
-  of rest.
-  There is a strong spatial and kinematic overlap between Kepler-1643
-  and RSG-5 ($\{ l, b\} \approx \{ 82^\circ,
-  6^\circ\}$; $\{ X, Y, Z\} \approx \{60, 330, 50\} $ pc).  The
-  local population of candidate young stars around KOI-7368 and
-  KOI-7913 is more diffuse -- we call this region ``CH-2''.
+  of rest.  There is a strong spatial and kinematic overlap between
+  Kepler-1643 and RSG-5 ($\{ l, b\} \approx \{ 82^\circ, 6^\circ\}$;
+  $\{ X, Y, Z\} \approx \{60, 330, 50\} $ pc).  The local population
+  of candidate young stars around KOI-7368 and KOI-7913 is more
+  diffuse -- we call this region ``CH-2''.
   The {\bf interactive figure} enables a few different cuts to be
@@
 \citep{stephenson_possible_1959} and RSG-5 \citep{roser_nine_2016}.
-Both of these clusters were known before Gaia, and have reported ages
+Each of these clusters was known before Gaia.  They have reported ages
 between $\approx$30 and $\approx$60 million years.  Early empirical
-evidence that these two clusters are likely part of a large and more
+evidence that these two clusters could be part of a large and more
 diffuse population was apparent in the Gaia-based photometric analysis
@@
 known groups in the larger structures dubbed ``Theia~73'' and
-``Theia~96''\footnote{A useful visualization is available online at
+``Theia~96''\footnote{See their visualization online at
 \url{http://mkounkel.com/mw3d/mw2d.html} (accessed 15 March 2022)}.
@@
 and generally supports the idea that there is an overdensity of
-$30$--$60\,{\rm Myr}$ old stars in this region of the Galaxy.
-\citet{Kerr2021}, in a volume-limited analysis of the Gaia DR2
-point-source catalog out to one third of a kiloparsec, identified
-three of the nearest sub-populations, dubbed ``Cepheus-Cygnus'',
-``Lyra'', and ``Cerberus''.  \citet{Kerr2021} reported ages for each
-of these subgroups between 30 and 35 million years.
+$\approx$30 and $\approx$60 million year old stars in this region of
+the Galaxy.  \citet{Kerr2021}, in a volume-limited analysis of the
+Gaia DR2 point-source catalog out to one third of a kiloparsec,
+identified three of the nearest sub-populations, dubbed
+``Cepheus-Cygnus'', ``Lyra'', and ``Cerberus''.  \citet{Kerr2021}
+reported ages for each of these subgroups between 30 and 35 million
+years.
 
@@
 
-The possibility that the $\delta$~Lyr cluster, RSG-5, and a few nearby
-sub-populations in this quadrant of the sky share a common origin has
-yet to be fully substantiated, and will be the subject of an upcoming
+% {\bf TODO RONAN: PLEASE VET ENTIRE SECTION, AND UPDATE WHERE
+% APPROPRIATE!}.
+
+The possibility that the $\delta$~Lyr cluster, RSG-5, and the
+sub-populations identified by \citet{Kerr2021} share a common origin
+has yet to be fully substantiated, and is the subject of the upcoming
 study by R.~Kerr and collaborators.  Our primary interest in the
@@
 (Figure~\ref{fig:XYZvtang}, top panel).  To further explore the
-population of $10$--$100\,{\rm Myr}$ stars that were observed, we
-select candidate Cep-Her members through four steps, the first three
-being identical to those described in Section~3 of \citet{Kerr2021}.
-We briefly summarize them here.
-
-{\bf TODO RONAN: PLEASE VET ENTIRE SECTION, AND UPDATE WHERE
-APPROPRIATE!}.
+population of stars that were observed, we select candidate Cep-Her
+members through four steps, the first three being identical to those
+described in Section~3 of \citet{Kerr2021}.  We briefly summarize them
+here.
 
@@
 
-The second step is to then perform in an unsupervised HDBScan
+The second step is to perform in an unsupervised HDBScan
 clustering on the photometrically selected population
@@
 These groups are then used as the ``seed'' populations for the third
-step, which is to search for objects that are spatially and
-kinematically consistent with each of the groups, {\bf IMPROVE
-CLARITY at least as close to the 10$^{\rm th}$ nearest
-HDBSCAN-identified member in space-velocity coordinates as the most
-peripheral HDBSCAN-identified member}.  This third step yields stars
-that are spatially and kinematically close to the
-photometrically-young stars, but which cannot be identified as young
-based on their positions in color versus absolute magnitude.
+step, which is to search for objects at least as close to the 10$^{\rm
+th}$ nearest HDBSCAN-identified member in space-velocity coordinates.
+This third step yields stars that are spatially and kinematically
+close to the photometrically-young stars, but which cannot be
+identified as young based on their positions in color versus absolute
+magnitude.
 
-The outcome of the analysis up to the point of the third step is
-shown in Figure~\ref{fig:XYZvtang}.  To enable a selection cut that
-filters out field-star contaminants, we also compute a weight metric,
-defined such that the group member with the smallest core distance has
-a weight of 1, the group member with the greatest core distance has a
+The outcome of the analysis up to the point of the third step is shown
+in Figure~\ref{fig:XYZvtang}.  To enable a selection cut that filters
+out field-star contaminants, we also compute a weight metric, defined
+such that the group member with the smallest core distance has a
+weight of 1, the group member with the greatest core distance has a
 weight of 0, and weights for the other group members are log-normally
-distributed between these two extremes.  In Figure~\ref{fig:XYZvtang}, we
-show 26{,}960 objects with weight exceeding 0.02 as gray points, and
-overplot 7{,}560 objects with weights exceeding 0.10 as black points.
-{\bf TODO: FIXME these object numbers impose pretty weak quality cuts and include junk at G>18... 
-use get\_clean\_gaia\_photometric\_sources too, since they come in for
-the CAMDs}.
+distributed between these two extremes.  In Figure~\ref{fig:XYZvtang},
+we show 26{,}960 objects with weight exceeding 0.02 as gray points,
+and overplot 7{,}560 objects with weights exceeding 0.10 as black
+points.  {\bf TODO: FIXME these object numbers impose pretty weak
+quality cuts and include junk at G>18...  use
+get\_clean\_gaia\_photometric\_sources too, since they come in for the
+CAMDs}.
 
@@
 
-The fourth and final step was cross-matching our candidate Cep-Her
-member list against all known Kepler Objects of Interest, including
-known false positives \citep{thompson_planetary_2018}.  This yields 32
-matches, of which 5 are ``confirmed'' planets according to the NASA
-Exoplanet Archive.  A cursory inspection of the Kepler data validation
-summaries and Robovetter classifications for these objects quickly
-showed whether they were potentially consistent with being {\it i)}
-planets, and {\it ii)} $\lesssim 10^8$ years old population, based on
-the presence of rotational modulation at the expected period and
-amplitude \citep[{\it e.g.},][Figure~9]{rebull_rotation_2020}.  Four
-objects remained after this inspection: Kepler-1627, Kepler-1643,
-KOI-7368, and KOI-7913.
+%
+% KOI match numbers from calc_CepHer_group_numbers.py
+%
+The fourth and final step was to cross-match our candidate Cep-Her
+member list against all known Kepler Objects of Interest.  We used the
+Cumulative KOI table from the NASA Exoplanet Archive from 27 March
+2022, and also compared against the \texttt{q1\_q17\_dr25} table
+\citep{thompson_planetary_2018}.  This yielded 32 matches, of which 18
+were known false positives, 7 were designated
+``confirmed'', and 8 were designated ``candidates''.  A cursory
+inspection of the Kepler data validation summaries and Robovetter
+classifications for these objects quickly showed whether they were
+potentially consistent with being {\it i)} planets, and {\it ii)}
+$\lesssim 10^8$ years old, based on the presence of rotational
+modulation at the expected period and amplitude \citep[{\it
+e.g.},][Figure~9]{rebull_rotation_2020}.  Four objects remained after
+this inspection: Kepler-1627, Kepler-1643, KOI-7368, and KOI-7913.
 
@@
 various projections.  Kepler-1643 is near the core RSG-5 population
-both spatially and kinematically.  KOI-7368 and KOI-7913 are in a more
+both spatially and kinematically.  KOI-7368 and KOI-7913 are in a 
 diffuse region $\approx$40 parsecs above RSG-5 in $Z$ and $\approx$100
@@
 
+%
+% Numbers are from RSG-5_auto_XYZ_vl_vb_cut.csv,
+% CH-2_auto_XYZ_vl_vb_cut.csv
+%
 We define two sets of stars in the local vicinity of our objects of
-interest, in order to verify whether they are indeed the same age.
-For candidate RSG-5 members, we require:
+interest.  For candidate RSG-5 members, we require:
 \begin{align}
-  45<X/{\rm pc}<75 \nonumber \\
-  320<Y/{\rm pc}<350 \nonumber \\
-  40<Z/{\rm pc}<70 \nonumber \\
-  -4<v_b/{\rm km\,s^{-1}}<-3 \nonumber \\
-  4<v_{l^*}/{\rm km\,s^{-1}}<6 \nonumber
+  X/{\rm pc} &\in [45, 75] \nonumber \\
+  Y/{\rm pc} &\in [320, 350] \nonumber \\
+  Z/{\rm pc} &\in [40, 70] \nonumber \\
+  v_b/{\rm km\,s^{-1}} &\in [-4, -3] \nonumber \\
+  v_{l^*}/{\rm km\,s^{-1}} &\in [4, 6] \nonumber
+  %45 < X/{\rm pc} < 75 \nonumber \\
+  %320 < Y/{\rm pc} < 350 \nonumber \\
+  %40 < Z/{\rm pc} < 70 \nonumber \\
+  %-4 < v_b/{\rm km\,s^{-1}} < -3 \nonumber \\
+  %4 < v_{l^*}/{\rm km\,s^{-1}} < 6 \nonumber
 \end{align}
@@
 \begin{align}
-	20<X/{\rm pc}<70 \nonumber \\
-	230<Y/{\rm pc}<270 \nonumber \\
-	75<Z/{\rm pc}<105 \nonumber \\
-	-3.5<v_b/{\rm km\,s^{-1}}<-1.5 \nonumber \\
-	2<v_{l^*}/{\rm km\,s^{-1}}<6, \nonumber
+  X/{\rm pc} &\in [20, 70] \nonumber \\
+  Y/{\rm pc} &\in [230, 270] \nonumber \\
+  Z/{\rm pc} &\in [75, 105] \nonumber \\
+  v_b/{\rm km\,s^{-1}} &\in [-3.5, -1.5] \nonumber \\
+  v_{l^*}/{\rm km\,s^{-1}} &\in [2, 6] \nonumber
+  %20 < X/{\rm pc} < 70 \nonumber \\
+  %230 < Y/{\rm pc} < 270 \nonumber \\
+  %75 < Z/{\rm pc} < 105 \nonumber \\
+  %-3.5 < v_b/{\rm km\,s^{-1}} < -1.5 \nonumber \\
+  %2 < v_{l^*}/{\rm km\,s^{-1}} < 6, \nonumber
 \end{align}
-and we call this latter set of stars ``CH-2''.
-% FIXME TODO numbers are from RSG-5_auto_XYZ_vl_vb_cut.csv,
-% CH-2_auto_XYZ_vl_vb_cut.csv right now.  Can we do better?
-These cuts yield 141 candidate RSG-5 members, and 
-37 candidate CH-2 members.
-Given the larger volume of the latter group, and the more diffuse base
-sample of stars, a major concern is contamination by field stars.
-We assess this in the following section.
-
-%Querying the Gaia archive and imposing the cuts suggested by
-%\citet[][Appendix B]{GaiaCollaboration2018} on the photometric signal
-%to noise ratio, the number of visibility periods used, the
-%astrometric $\chi^2$ of the single-source solution, the $G_{\rm
-%BP}-G_{\rm RP}$ color excess factor and the parallax precision
-%($\varpi/\sigma_\varpi>5$) yields $\approx5.0\times10^6$ sources.
-%
-%Imposing a thin slice in color-absolute magnitude space around the
-%pre-main-sequence of the $\delta$ Lyr cluster's previously
-%established $\approx$40 Myr locus, and imposing kinematics cut for
-%$-8 < v_b / {\rm km\,s^{-1}} < 0$ and $-15 < v_{l'} / {\rm
-%km\,s^{-1}} < 15$, where $l'=l\cos b$, yields $\approx 5.2\times10^3$
-%sources.  These objects are all in the range BP-RP 1.5 to 3.2, so
-%roughly K5V to M4.5V, or $\approx$0.2 to 0.7 solar masses.  Much more
-%manageable, but according to IMF arguments, missing at least a factor
-%of two in mass of any clusters.
-%
-%In XYZ, and in tangential velocity space, these five thousand sources
-%are visually clustered in between a few to a dozen groups, depending
-%on the characteristic size and velocity scales of interest.
+and we call this latter set of stars ``CH-2''.  These cuts yield 141
+candidate RSG-5 members, and 37 candidate CH-2 members.  An important
+consideration, especially for CH-2, is the
+contamination rate by field stars.  We assess this in the following
+section.
 
@@
 	\caption{
-		{\bf The \cn\ is \clusterage\ old.} 
-    The top row shows CAMDs.  Left shows CH-2, right shows RSG-5.  The
-    bottom left shows gyro (and is a {\bf place-holder since we might
-    want to add CH-2 and RSG5}).  The bottom right shows lithium
-    (black points are NGC2547 and IC2602 from Randich+18 and probably
-    are not believable at the red end; for Kepler-1643 I'm less
-    confident, but it might be tied to the ``slow'' rotation period --
-    this is why we need RSG5 rotation periods).  Also, we might want
-    an H-alpha plot?).
+		{\bf The stellar groups near KOI-7368, KOI-7913, and Kepler-1643
+    are $\approx$40 million years old.} 
+    {\it Top row}: 
+    Color-absolute magnitude diagram of candidate Cep-Her members, in
+    addition to candidate members of the $\delta$~Lyr~cluster
+    ($\approx38$\,Myr; \citealt{bouma_kep1627_2022}) and and the Gaia
+    EDR3 Catalog of Nearby Stars (gray background).  The left and
+    right columns shows stars in RSG-5 and CH-2, respectively.  The
+    range of colors displayed is truncated to show whether or not
+    candidate members fall on the pre-main-sequence.
+    {\it Bottom row}:
+    TESS and ZTF-derived stellar rotation periods, with the Pleiades
+    ($\approx 112$\,Myr), and Praesepe ($\approx 650$\,Myr) shown for reference
+    \citep{rebull_rotation_2016a,douglas_poking_2017}.
+    Our detection efficiency for reliable rotation periods falls off
+    beyond $\bpmrpo \gtrsim 2.6$.
+    {\bf todo: fix zorder markers}.  
 		\label{fig:age}
@@
 
-The evidence for the RSG-5 and CH-2 sub-groups at least roughly
-sharing the age of the $\delta$\ Lyra\ cluster is shown in
-Figure~\ref{fig:age}.
-
 \subsubsection{Color-Absolute Magnitude Diagram}
@@
 
-% FIXME FIXME calculate for both!
+Color-absolute magnitude diagrams (CAMDs) of the candidate RSG-5 and
+CH-2 members are shown in the upper row of Figure~\ref{fig:age}.  The
+stars from the $\delta$~Lyr cluster are from
+\citet{bouma_kep1627_2022}, and the field stars are from the Gaia EDR3
+Catalog of Nearby Stars \citep{gaia_gcns_2021}.  To make these
+diagrams, we imposed the data filtering criteria from
+\citet[][Appendix B]{GaiaCollaboration2018}, which are designed to
+include binaries while omitting instrumental artifacts from for
+instance low photometric signal to noise, or a small number of
+visibility periods.  We then corrected for extinction using the
+\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}}
+dust maps and the extinction coefficients $k_X\equiv A_X/A_0$from
+\citet{GaiaCollaboration2018}, assuming that $A_0 = 3.1 E(B-V)$.  This
+yielded a mean and standard deviation for the reddening of
+$E(B-V)=0.036\pm0.002$ for RSG-5, and $E(B-V)=0.017\pm0.001$
+for CH-2.  By way of comparison, in \citet{bouma_kep1627_2022} the
+same query for the $\delta$~Lyr cluster yielded
+$E(B-V)=0.032\pm0.006$.  Finally, for the plots we set the $\bpmrpo$
+color range to best visualize the region of maximal age information
+content: the pre-main-sequence.
+
+The CAMDs show that for RSG-5, our membership selection gives stars
+that photometrically appear to almost all be on a tight
+pre-main-sequence locus.  This implies a false positive rate of at
+most a few percent.  By comparison, our control sample (the $\delta$~Lyr
+candidates) has a false positive rate of $\approx$12\%, based on the
+number of stars that photometrically appear to be more consistent with
+the field population than the bulk cluster population.  For CH-2, our
+membership selection gives 27 objects in the color range displayed,
+and 23 of them appear to be consistent with being on the
+pre-main-sequence.  This implies a false positive rate of
+$\approx$15\%.
+
+In addition, Figure~\ref{fig:age} shows that the candidate RSG-5 and
+CH-2 members overlap with the $\delta$~Lyr cluster, and are therefore
+roughly the same age.  To quantify this, we use the empirical method
+introduced by \citet[][see their Section~6.3]{gagne_mutau_2020}.  This
+idea of the method is to fit the pre-main-sequence loci of a set of
+reference clusters, and to then model the locus of the target cluster
+as a linear combination of these reference cluster loci.  For our
+reference clusters, we used UCL, IC\,2602, and the Pleiades, from the
+memberships reported by \citet{Damiani2019} and
+\cite{CantatGaudin2018a} respectively.  We adopted ages of 16\,Myr for
+UCL \citep{pecaut_star_2016}, 38\,Myr for IC\,2602\footnote{ Ages for
+IC\,2602 vary from 40 to 46\,Myr based on lithium-depletion-boundary
+(LDB) measurements \citep{dobbie_ic_2010,randich_gaiaeso_2018}, and
+from 30 to 46\,Myr based on isochronal analyses
+\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.},
+and 112\,Myr for the Pleiades \citep{dahm_2015}.  These assumptions
+and the consequent processing steps were identical to those described
+in \citet{bouma_kep1627_2022}.  The mean and standard deviation of the
+resulting age posterior are $46^{+9}_{-7}$\,Myr for RSG-5, and
+$36^{+10}_{-8}$\,Myr for CH-2.  For comparison, the
+this procedure yields an age for the $\delta$~Lyr cluster of
+$38^{+6}_{-5}$\,Myr.  The slightly older isochronal age of RSG-5 is
+expected given that its locus is slightly bluer and less luminous in
+the upper left panel of Figure~\ref{fig:age} relative to the
+$\delta$~Lyr cluster.
 
@@
 \section{The Planets}
-\label{sec:planet}
+\label{sec:planets}
 
@@
 \section{Discussion \& Conclusions}
-\label{sec:conc}
+\label{sec:disc_conc}
 

### 2022-03-27 papers/collect.sh

@@ -4,12 +4,12 @@ cp /Users/luke/Dropbox/proj/rudolf/results/CepHer_XYZvtang_sky/CepHer_XYZvtang_s
 # rotation, CAMD, spectra (Li/H-alpha)
 # CAMD:
 
-cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_CH-2_cleanhrcut_gaia2018_smalllims_overplotkoi7368_overplotkoi7913.pdf f2a.pdf
-cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_RSG-5_cleanhrcut_gaia2018_smalllims_overplotkep1643.pdf f2b.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_RSG-5_cleanhrcut_gaia2018_smalllims_overplotkep1643.pdf f2a.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_CH-2_cleanhrcut_gaia2018_smalllims_overplotkoi7368_overplotkoi7913.pdf f2b.pdf
 
 # rotation:
-cp /Users/luke/Dropbox/proj/rudolf/results/rotation/CH-2_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_defaultcleaning_allknown.pdf f2c.pdf
-cp /Users/luke/Dropbox/proj/rudolf/results/rotation/RSG-5_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_defaultcleaning_allknown.pdf f2d.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/rotation/RSG-5_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_defaultcleaning_allknown.pdf f2c.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/rotation/CH-2_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_defaultcleaning_allknown.pdf f2d.pdf
 
 # spectra: Li EW vs Teff and Halpha EW vs Teff
 # TODO:

### 2022-03-28 papers/ms.tex
update ms

@@
   v_{l^*}/{\rm km\,s^{-1}} &\in [4, 6] \nonumber
-  %45 < X/{\rm pc} < 75 \nonumber \\
-  %320 < Y/{\rm pc} < 350 \nonumber \\
-  %40 < Z/{\rm pc} < 70 \nonumber \\
-  %-4 < v_b/{\rm km\,s^{-1}} < -3 \nonumber \\
-  %4 < v_{l^*}/{\rm km\,s^{-1}} < 6 \nonumber
 \end{align}
@@
   v_{l^*}/{\rm km\,s^{-1}} &\in [2, 6] \nonumber
-  %20 < X/{\rm pc} < 70 \nonumber \\
-  %230 < Y/{\rm pc} < 270 \nonumber \\
-  %75 < Z/{\rm pc} < 105 \nonumber \\
-  %-3.5 < v_b/{\rm km\,s^{-1}} < -1.5 \nonumber \\
-  %2 < v_{l^*}/{\rm km\,s^{-1}} < 6, \nonumber
 \end{align}
@@
     right columns shows stars in RSG-5 and CH-2, respectively.  The
-    range of colors displayed is truncated to show whether or not
-    candidate members fall on the pre-main-sequence.
+    range of colors is truncated to emphasize the pre-main-sequence.
     {\it Bottom row}:
     TESS and ZTF-derived stellar rotation periods, with the Pleiades
-    ($\approx 112$\,Myr), and Praesepe ($\approx 650$\,Myr) shown for reference
+    ($\approx 112$\,Myr) and Praesepe ($\approx 650$\,Myr) shown for reference
     \citep{rebull_rotation_2016a,douglas_poking_2017}.
-    Our detection efficiency for reliable rotation periods falls off
+    The detection efficiency for reliable rotation periods falls off
     beyond $\bpmrpo \gtrsim 2.6$.
-    {\bf todo: fix zorder markers}.  
 	\label{fig:age}
@@
 and 112\,Myr for the Pleiades \citep{dahm_2015}.  These assumptions
-and the consequent processing steps were identical to those described
-in \citet{bouma_kep1627_2022}.  The mean and standard deviation of the
-resulting age posterior are $46^{+9}_{-7}$\,Myr for RSG-5, and
-$36^{+10}_{-8}$\,Myr for CH-2.  For comparison, the
-this procedure yields an age for the $\delta$~Lyr cluster of
+and the consequent processing steps taken to exclude field stars as
+well as photometric and astrometric binaries were identical to those
+described in \citet{bouma_kep1627_2022}.  The mean and standard
+deviation of the resulting age posterior are $46^{+9}_{-7}$\,Myr for
+RSG-5, and $36^{+10}_{-8}$\,Myr for CH-2.  For comparison, the this
+procedure yields an age for the $\delta$~Lyr cluster of
 $38^{+6}_{-5}$\,Myr.  The slightly older isochronal age of RSG-5 is
@@
 \subsubsection{Stellar Rotation Periods}
+\label{sec:rotation}
+
+An independent way to assess the age of the candidate cluster members
+is to measure their stellar rotation periods using data available from
+surveys such as TESS \citep{ricker_transiting_2015} and ZTF
+\citep{bellm_zwicky_2019}.  This approach leverages a storied
+tradition of measuring rotation periods for open clusters of various
+ages, in order to determine the rate of stellar spin-down \citep[{\it
+e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data in
+our case are especially useful, since they provide 3 to 5 lunar months
+of photometry for all of our candidate CH-2 and RSG-5 members.
+
+We selected stars suitable for gyrochronology by requiring $\bpmrpo
+\geq 0.5$ and $G<16$.  The latter cut corresponds to $\bpmrpo \lesssim
+2.6$, at the relevant distances.  These cuts gave 19 stars in CH-2 and
+42 stars in RSG-5.  We extracted light curves from the TESS images for
+these stars using the \texttt{unpopular} package
+\citep{hattorio_2021_cpm}, and regressed them against systematics with
+its causal pixel model.  
+{\bf For the ZTF light curves, we used the default photometry?}.
+We then measured candidate rotation periods
+using a Lomb-Scargle periodogram, and visually inspected them
+following the methods discussed in \citet{curtis_rup147_2020}.
+
+The lower panels of Figure~\ref{fig:age} show the results.
+In RSG-5,
+36/42 stars have rotation consistent with ``sub-Pleiades'' (86\%).
+Given the nearly perfectly clean photometric CAMD for RSG-5, this
+probably gives some indication for the fraction of stars for which we
+should expect
+to detect rotation periods with TESS (due to e.g., pole-on stars
+having less detectable starspot modulation).
+
+Consider stars in CH-2 with 0.5 < BP-RP0, and G < 16.
+19 stars meet this criterion -- all were observed by TESS.
+13/19 (68\%) have rotation that is obviously consistent with a 40 Myr
+age.
+4 are M-dwarfs with rotation periods between 10 and 12.5 days, for which
+the age interpretation is not obvious.
+L.~Rebull for instance have found numerous M-dwarfs
+with 10-12 day rotation periods at ages of USco ($\sim 8$\,Myr); some
+might exist at ages of LCC ($\sim 16$\,Myr)
+(\citealt{rebull_rotation_2020}; Rebull et al., in prep -- {\bf see
+Skumanich conf slide deck -- though beware TESS stitching systematics}).
+Fr the sake of argument, let's say that this provides an independent
+verification that at least two thirds of our CH-2 candidates are both
+photometrically and rotationally consistent with being $\approx$40
+Myr.
+
+For RSG-5, 42 stars are in the relevant color and brgithness range.
+36 have rotation consistent with ``sub-Pleiades'', i.e., 86\%.
+Given the nearly perfectly clean photometric CAMD for RSG-5, this
+probably gives some indication for the fraction of stars for which we
+should expect
+to detect rotation periods with TESS (due to e.g., pole-on stars
+having less detectable starspot modulation).
 

### 2022-03-28 papers/ms.tex
update ms

@@
 \newcommand{\kms}{\,km\,s$^{-1}$}
+\newcommand{\mkms}{\,km\,s^{-1}}  % math mode
 \newcommand{\ms}{\,m\,s$^{-1}$}
@@
 \citep{jerabkova_800_2021}.  The stars in such diffuse regions can be
-verified to be the same age as the core members ({\it i.e.}, co-eval)
+verified to be the same age as the core members ({\it i.e.}, coeval)
 through analyses of color-absolute magnitude diagrams, stellar
@@
 implications for our understanding of star formation and cluster
-evolution \citep[{\it e.g.}][]{dinnbier_tidal_2020}, a more immediate
+evolution \citep[{\it e.g.},][]{dinnbier_tidal_2020}, a more immediate
 consequence is that we now know the ages of many more stars, including
@@
 The CAMDs show that for RSG-5, our membership selection gives stars
-that photometrically appear to almost all be on a tight
-pre-main-sequence locus.  This implies a false positive rate of at
-most a few percent.  By comparison, our control sample (the $\delta$~Lyr
+that photometrically seem to almost all be on a tight
+pre-main-sequence locus.  This implies a false positive rate of 
+a few percent, at most.  By comparison, our control sample (the $\delta$~Lyr
 candidates) has a false positive rate of $\approx$12\%, based on the
@@
 An independent way to assess the age of the candidate cluster members
-is to measure their stellar rotation periods using data available from
-surveys such as TESS \citep{ricker_transiting_2015} and ZTF
-\citep{bellm_zwicky_2019}.  This approach leverages a storied
-tradition of measuring rotation periods for open clusters of various
-ages, in order to determine the rate of stellar spin-down \citep[{\it
-e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data in
-our case are especially useful, since they provide 3 to 5 lunar months
-of photometry for all of our candidate CH-2 and RSG-5 members.
+is to measure their stellar rotation periods.  This approach can be
+achieved using surveys such as TESS \citep{ricker_transiting_2015} and
+ZTF \citep{bellm_zwicky_2019}, and it leverages a storied tradition of
+rotation period measurement for benchmark open clusters \citep[see
+{\it e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data
+in our case are especially useful, since they provide 3 to 5 lunar
+months of photometry for all of our candidate CH-2 and RSG-5 members.
 
@@
 \citep{hattorio_2021_cpm}, and regressed them against systematics with
-its causal pixel model.  
-{\bf For the ZTF light curves, we used the default photometry?}.
-We then measured candidate rotation periods
-using a Lomb-Scargle periodogram, and visually inspected them
-following the methods discussed in \citet{curtis_rup147_2020}.
+its causal pixel model.  {\bf For the ZTF light curves, we used the
+default photometry?  Ran aperture photometry on the image cutouts?}.
+We then measured candidate rotation periods using a Lomb-Scargle
+periodogram, and visually inspected them following the methods
+discussed in \citet{curtis_rup147_2020}.
 
-The lower panels of Figure~\ref{fig:age} show the results.
-In RSG-5,
-36/42 stars have rotation consistent with ``sub-Pleiades'' (86\%).
-Given the nearly perfectly clean photometric CAMD for RSG-5, this
-probably gives some indication for the fraction of stars for which we
-should expect
-to detect rotation periods with TESS (due to e.g., pole-on stars
-having less detectable starspot modulation).
+The lower panels of Figure~\ref{fig:age} show the results.  In RSG-5,
+36/42 stars have rotation periods faster than the Pleiades (86\%).
+This numerator omits the two stars with periods $>12$\,days visible in
+the lower-left panel of Figure~\ref{fig:age}.  The age interpretation
+for these latter stars, particularly the $\approx$M2.5 dwarf, is not
+obvious.  \citet{rebull_usco_2018} for instance have found numerous
+M-dwarfs with 10-12 day rotation periods at ages of USco ($\sim
+8$\,Myr), and some may still exist at ages of LCC ($\sim 16$\,Myr;
+L.~Rebull in preparation).  Regardless, given that nearly no field
+star outliers seem to be present on the RSG-5 CAMD, the fact that we
+do not detect rotation periods for $\approx$14\% of stars should
+perhaps be taken as an indication for the fraction of stars for
+rotation periods might not be detectable, due to {\it e.g.}, pole-on
+stars having lower amplitude starspot modulation.
 
-Consider stars in CH-2 with 0.5 < BP-RP0, and G < 16.
-19 stars meet this criterion -- all were observed by TESS.
-13/19 (68\%) have rotation that is obviously consistent with a 40 Myr
-age.
-4 are M-dwarfs with rotation periods between 10 and 12.5 days, for which
-the age interpretation is not obvious.
-L.~Rebull for instance have found numerous M-dwarfs
-with 10-12 day rotation periods at ages of USco ($\sim 8$\,Myr); some
-might exist at ages of LCC ($\sim 16$\,Myr)
-(\citealt{rebull_rotation_2020}; Rebull et al., in prep -- {\bf see
-Skumanich conf slide deck -- though beware TESS stitching systematics}).
-Fr the sake of argument, let's say that this provides an independent
-verification that at least two thirds of our CH-2 candidates are both
-photometrically and rotationally consistent with being $\approx$40
-Myr.
+%
+% CH-2
+% 5 stars with Prot>10 days (from the Prot versus color plot):
+%                      DESIGNATION  TESS_Data   bp_rp_x     period
+%25  Gaia EDR3 2129930258400157440          4  1.909693  15.309000
+%24  Gaia EDR3 2129417404945830784          4  2.040988  10.981582
+%4   Gaia EDR3 2107169165113864064          4  2.059904  10.580494
+%9   Gaia EDR3 2134851775526125696          4  2.326998  12.227900
+%1   Gaia EDR3 2127726081184588160          4  2.347244  11.543281
+%
+% CH-2
+% Photometric field stars (from Glue -- one of these is omitted due to phot S/N cut, 
+% presumably 2126008609657702784 )
+%
+%In [9]: df[['DESIGNATION', 'bp_rp', 'M_G']]
+%Out[9]:
+%                     DESIGNATION     bp_rp        M_G
+%0  Gaia EDR3 2126529468937923456  2.518077   9.844733 --> NaN period, G=16.9
+%1  Gaia EDR3 2126008609657702784  2.746735  11.409737 --> NaN, G=18.4
+%2  Gaia EDR3 2133330051428541952  2.499458   9.459546 --> 15.7 d (ZTF), G=16.5
+%3  Gaia EDR3 2133510921093964928  2.720177  10.923209 --> NaN, G=17.9
+%4  Gaia EDR3 2139369183468814464  2.929506  11.170221 --> NaN, G=18.2
+%5  Gaia EDR3 2129930258400157440  1.909693   8.200652 --> P=15.3 (ZTF), G=15.2
 
-For RSG-5, 42 stars are in the relevant color and brgithness range.
-36 have rotation consistent with ``sub-Pleiades'', i.e., 86\%.
-Given the nearly perfectly clean photometric CAMD for RSG-5, this
-probably gives some indication for the fraction of stars for which we
-should expect
-to detect rotation periods with TESS (due to e.g., pole-on stars
-having less detectable starspot modulation).
+For CH-2, 13/19 stars have rotation periods that are obviously faster
+than their counterparts in the Pleiades (68\%).  4 stars, not included
+in the preceding numerator, are M-dwarfs with rotation periods between
+10 and 12.5 days.  The age interpretation for these M-dwarfs is, as
+just discussed, not obvious.  Regardless, the $\approx$15\% false
+positive rate determined from the CAMD seems consistent with our
+fraction of detected rotation periods, given that RSG-5 was also
+missing rotation period detections for $\approx$15\% of its candidate
+members, which all seemed photometrically consistent with being part
+of a single pre-main-sequence locus.
+
+It is challenging to convert these stellar rotation periods
+to a precise age estimate, since on the pre-main-sequence
+the stars are spinning up due to thermal contraction
+rather than down due to magnetized braking.  Regardless, the rotation
+period distributions of both CH-2 and RSG-5 seem consistent with other
+30\,Myr to 50\,Myr clusters ({\it e.g.}, IC\,2602 and IC\,2391;
+\citealt{douglas_stephanie_t_2021_5131306}).
+They also seem consistent with the false positive rate estimates
+determined from the color-absolute magnitude diagrams.
 
@@
 \label{sec:disc_conc}
+
+\subsection{Is CH-2 really a star cluster?}
+
+RSG-5, and Kepler-1643's membership inside it, clearly meet the
+typical expectations of a star claimed to be in an open cluster.
+RSG-5 shows an obvious overdensity relative to the local field
+population ({\it e.g.}, Figure~\ref{fig:XYZvtang}), and our membership
+selection easily produces a clean pre-main-sequence locus in 
+color-absolute magnitude space (Figure~\ref{fig:age}).
+CH-2, and KOI-7913 and KOI-7368's membership inside it, do not meet
+those expectations in as obvious a manner.  This is because this
+association of stars is diffuse.
+
+%really 80:9
+To quantify the density discrepancy, we can compare the spatial and
+velocity volumes searched to select candidate members of each cluster.
+For RSG-5, we drew 141 candidate members from a $30\,{\rm
+pc}\times30\,{\rm pc}\times30\,{\rm pc}$ spatial cube, given a $1 {\rm \mkms} \times 2 {\rm \mkms }$ rectangle in apparent galactic velocity.
+For CH-2, our 37 candidate members came from a spatial cube of
+dimension $50\,{\rm pc}\times40\,{\rm pc}\times30\,{\rm pc}$, and the
+velocity rectangle of $2 {\rm \mkms} \times 4 {\rm \mkms}$.  If we
+define the ``searched volume'' in units of ${\rm pc}^3 {\rm
+(\mkms)}^2$, then the volume ratio of CH-2:RSG-5 is $\approx$9:1.  The
+density (number of stars per unit searched volume) within RSG-5 relative to
+CH-2 similarly comes out to 34 to 1.
+
+So what makes a star cluster?  Historic answers to this question have
+recently by reviewed by \citet{krumholz_star_2019}: some definitions
+that have been offered include criteria such as being gravitationally
+bound, and having a mass density that significantly exceeds the mean
+in a cluster's galactic neighborhood.  We prefer a modified version of
+the definition adopted by \citet{krumholz_star_2019}: for our purposes
+a star cluster is a group of at least 12 stars that was physically
+associated at its time of formation.  The somewhat arbitrary ``12'' is
+set to distinguish clusters from high-order multiple star systems.  We
+therefore explicitly include dissolved clusters and their tidal tails
+in our concept of clusters.  We also explicitly exclude the idea that
+a particular number of stars per unit spatial volume is required to
+define a cluster.  The latter point acknowledges the fact that an
+important factor in cluster identification is now also the number per
+unit velocity volume, whether in 2-dimensional tangential velocity, or
+when including the third radial component.  Perhaps
+once stellar rotation periods and chemical abundances reach the same
+level of ubiquity as stellar proper motions, they might enable further
+refinement of our ability in cluster discovery.
+
+From a data-driven perspective, how do we demonstrate that a star is
+in a cluster, {\it i.e.}, that it is part of a group of stars that was
+physically associated at its time of formation?
+Back-integrating the orbits is one convincing approach, but it does
+not always work {\bf (CITE)}.
+The relatively minimal approach suggested by \citet{tofflemire_2021}
+is intriguing: search for coeval, phase-space neighbors, measure their
+ages, and determine if they share a common age.
+This approach is more accurately be described as a method for
+determining whether a star is currently associated with a set of coeval
+stars, which is much easier to determine than what the association
+looked like in the past.
+From this standard, our analysis thus far of CH-2 has already
+demonstrated the existence of such an assocation.
+
+A crucial logical step in this method however is to ensure that the
+(automated) search process for coeval phase-space neighbors in fact
+produces neighbors at a rate different from what it would for field
+stars.
+
+% TODO
+And so we ask: what is the density of field stars in the CH-2 region?
+If we had applied Ronan's pre-main-sequence selection, + HDBScan, +
+local neighbors... what would we have gotten?
+Similarly... if you just applied Adam Kraus' Comove 
+What is the likelihood we are fooling ourselves?
 

### 2022-03-29 papers/ms.tex
update ms

@@
   $46^{+9}_{-7}$\,Myr old, while KOI-7368 and KOI-7913 (in CH-2) are
-  $36^+{10}_{-8}$\,Myr old.
+  $36^{+10}_{-8}$\,Myr old.
   Based on the transit shapes and high resolution imaging, they are
@@
   Kepler-1643, KOI-7368, and KOI-7913 respectively.
-  Kepler-1643 and KOI-7913 are therefore the first empirical
+  These planets are therefore the first empirical
   demonstration that mini-Neptunes with sizes of $\approx$2 Earth
@@
 which the planetary evidence comes from transits, and the age evidence
-is based on either cluster membership \citep[{\it
-e.g.},]{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
+is based on either cluster membership \citep{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
 or else on correlates of youth such as stellar rotation, photospheric
 lithium abundances, x-ray activity, or emission line strength
-\citep[{\it e.g.},]{zhou_2021_tois,hedges_toi-2076_2021}.
+\citep{zhou_2021_tois,hedges_toi-2076_2021}.
 
@@
 effects and internal dynamics.  As an extreme example, in the Hyades
-the apparent velocities of stars in the tidal tails are expected to
+the velocities of stars in the tidal tails are expected to
 span up to $\pm 40$\kms relative to the cluster center
@@
 implications for our understanding of star formation and cluster
-evolution \citep[{\it e.g.},][]{dinnbier_tidal_2020}, a more immediate
+evolution \citep{dinnbier_tidal_2020}, a more immediate
 consequence is that we now know the ages of many more stars, including
@@
 
-Many of the relevant stellar parameters can be gleaned by inspecting
-Figure~\ref{fig:age}.  The known planet-hosting stars in Cep-Her span
-spectral types of G8V (Kepler-1627) to K8V (KOI-7913 A).  The
-secondary in the KOI-7913 system is only marginally cooler than the
-primary (SpType $\approx$M0V).
-The latter system is 3$\farcs$501 
-%FIXME FIXME FIXME TODO TODO TODO
-A Solar-mass star with solar metallicity arrives at the zero-age main
-sequence at $t\approx40$ million years (CITE CHOI MESA), and so these
-stars are in the end stages of their pre-main-sequence contraction.
-Their Gaia EDR3 parallaxes span X.X to Y.Y, corresponding to distances
-between X.X and Y.Y parsecs.
+Some of the most salient stellar parameters of the KOIs in Cep-Her can
+be gleaned by inspecting Figure~\ref{fig:age}.  They span spectral
+types of G8V (Kepler-1627) to K8V (KOI-7913 A).  The secondary in the
+KOI-7913 system is marginally cooler than the primary.  Since a
+Solar-mass star with solar metallicity arrives at the zero-age main
+sequence at $t\approx40$ million years \citep{choi_mesa_2016}, these
+stars are all in the late stages of their pre-main-sequence
+contraction.  Their heliocentric distances are $\approx$260\,pc
+(KOI-7913 and KOI-7368, in CH-2) and $\approx$340\,pc (Kepler-1643, in
+RSG-5).
 
-To derive the spectroscopic parameters $(T_{\rm eff}, \log g_\star,
-[{\rm Fe/H}])$ and  the equivalent width of the Li\,\textsc{I}
-from the 6708\,\AA\ doublet, we acquired spectra.  Specifically, we
-acquired iodine-free HIRES spectra for Kepler-1643 (YYYY/MM/DD) and
-KOI-7913 (YYYY/MM/DD and YYYY/MM/DD), where for the latter the two
-different epochs corresponded to observations of the secondary and
-primary respectively.  The acquisition and analysis followed the
-standard reduction techniques of the California Planet Survey
-\citep{howard_cps_2010}.  For KOI-7368, we acquired TRES spectra on
-YYYY/MM/DD and YYYY/MM/DD.  CITE METHOD PAPER.
-The results are given in Table~\ref{tab:sysparams}.
+To derive spectroscopic parameters $(T_{\rm eff}, \log g_\star, [{\rm
+Fe/H}])$ and to analyze spectroscopic youth proxies such as the
+Li\,\textsc{I} 6708\,\AA\ doublet equivalent width and H$\alpha$
+emission, we acquired spectra.  The system-by-system details follow.
+The results are summarized in Table~\ref{tab:sysparams}.
 
-Table~\ref{tab:sysparams} summarizes
-%
-%from the 
-%
-%
 %\citep[{\it e.g.},][]{soderblom_ages_2014}, we acquired an iodine-free
@@
 %\subsection{Kepler\,1627A}
+
 \subsection{Kepler\,1643}
+
+For Kepler-1643,
+we acquired an iodine-free spectrum from Keck/HIRES
+...
+%FIMXE TODO TODO TODO WORDS LET'S GO!
+
+spectra  (YYYY/MM/DD) and KOI-7913 (YYYY/MM/DD and
+YYYY/MM/DD), where for the latter the two different epochs
+corresponded to observations of the secondary and primary
+respectively.  The acquisition and analysis followed the standard
+reduction techniques of the California Planet Survey
+\citep{howard_cps_2010}.  For KOI-7368, we acquired TRES spectra on
+YYYY/MM/DD and YYYY/MM/DD.  CITE METHOD PAPER.
+
 \subsection{KOI-7368}
@@
 Is a binary.
+The latter system is 3$\farcs$501 
+%FIXME FIXME FIXME TODO TODO TODO
 

### 2022-03-29 papers/ms.tex
paper: update ms

@@
 \title{
-  Three $\approx$40 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
+  Three 40 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
 }
@@
 
-Some of the most salient stellar parameters of the KOIs in Cep-Her can
-be gleaned by inspecting Figure~\ref{fig:age}.  They span spectral
-types of G8V (Kepler-1627) to K8V (KOI-7913 A).  The secondary in the
-KOI-7913 system is marginally cooler than the primary.  Since a
+Some salient stellar parameters of the KOIs in Cep-Her can be gleaned
+by inspecting Figure~\ref{fig:age}.  They span spectral types of G8V
+(Kepler-1627) to K8V (KOI-7913 A).  The secondary in the KOI-7913
+system is only marginally cooler than the primary.  And since a
 Solar-mass star with solar metallicity arrives at the zero-age main
@@
 
-To derive spectroscopic parameters $(T_{\rm eff}, \log g_\star, [{\rm
-Fe/H}])$ and to analyze spectroscopic youth proxies such as the
-Li\,\textsc{I} 6708\,\AA\ doublet equivalent width and H$\alpha$
-emission, we acquired spectra.  The system-by-system details follow.
-The results are summarized in Table~\ref{tab:sysparams}.
-
-%\citep[{\it e.g.},][]{soderblom_ages_2014}, we acquired an iodine-free
-%spectrum from Keck/HIRES on the night of 2021 March 26 using the
-%standard setup and reduction techniques of the California Planet
-%Survey \citep{howard_cps_2010}.  Following the equivalent width
-%measurement procedure described by \citet{bouma_2021_ngc2516}, we find
-%${\rm EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   This value does not
-%correct for the Fe\,\textsc{I} blend at 6707.44\AA.  Nonetheless,
-%given the stellar effective temperature (Table~\ref{tab:starparams}),
-%this measurement is in agreement with expectations for a
-%$\approx40$\,Myr star ({\it e.g.}, as measured in IC~2602 by
-%\citealt{randich_gaiaeso_2018}).  It is also larger than any lithium
-%equivalent widths measured by \citet{berger_identifying_2018} in their
-%analysis of 1{,}301 Kepler-star spectra.
-%
-%
-%% rotation period from plot_rotation_period_windowslider.py
-%Based on the spatial and kinematic association of \sn\ with the \cn,
-%and the assumption that the planet formed shortly after the star, it
-%seems likely that \sn\ is the same age as the cluster. There are two
-%consistency checks on whether this is true: rotation and lithium.
-%Based on the Kepler light curve, the rotation period is
-%$2.642\pm0.042$\,days, where the quoted uncertainty is based on the
-%scatter in rotation periods measured from each individual Kepler
-%quarter.  This is consistent with comparable cluster members
-%(Figure~\ref{fig:age}).
-
-%\subsection{Kepler\,1627A}
+The adopted stellar parameters are listed in
+Table~\ref{tab:starparams}.  The stellar surface gravity, radius,
+mass, and density are found by interpolating against the MIST
+isochrones \citep{choi_mesa_2016}.  The statistical uncertainties from
+this technique mostly originate from the parallax uncertaintes; the
+systematic uncertainties are taken to be the absolute difference
+between the PARSEC \citep{bressan_parsec_2012} and MIST isochrones.
+Reported uncertainties are a quadrature sum of the statistical and
+systematic components.
+
+To verify these parameters and to also analyze spectroscopic youth
+proxies such as the Li\,\textsc{I} 6708\,\AA\ doublet and H$\alpha$,
+we acquired spectra.  We also acquired high resolution imaging for
+each system, to constrain the existence of visual companions,
+including possible bound binaries.  The system-by-system details
+follow, and the the relevant results are also summarized in
+Table~\ref{tab:sysparams}.
 
@@
 
-For Kepler-1643,
-we acquired an iodine-free spectrum from Keck/HIRES
-...
-%FIMXE TODO TODO TODO WORDS LET'S GO!
-
-spectra  (YYYY/MM/DD) and KOI-7913 (YYYY/MM/DD and
-YYYY/MM/DD), where for the latter the two different epochs
-corresponded to observations of the secondary and primary
-respectively.  The acquisition and analysis followed the standard
-reduction techniques of the California Planet Survey
-\citep{howard_cps_2010}.  For KOI-7368, we acquired TRES spectra on
-YYYY/MM/DD and YYYY/MM/DD.  CITE METHOD PAPER.
+\paragraph{Spectra}
+For Kepler-1643, we acquired two iodine-free spectra from Keck/HIRES
+on the nights of 2020 August 16 and 2021 October 25.  The acquisition
+and analysis followed the standard reduction techniques of the
+California Planet Survey \citep{howard_cps_2010}.  We derived the
+stellar spectral parameters ($T_{\rm eff}, \log g, R_\star$) using
+\texttt{SpecMatch-Emp} \citep{yee_SM_2017}, which yielded values in
+$<1$-$\sigma$ agreement with those from the cluster-isochrone
+approach.  This approach also yielded $[Fe/H]=0.13 \pm 0.09$.  Using
+the broadened synthetic templates from \texttt{SpecMatch-Synth}
+\citep{petigura_cksi_2017}, we found $v\sin i = 9.3 \pm 1.0\,\mkms$.
+The systemic radial velocity at the two sequence epochs was $-9.1 \pm
+1.9\,\mkms$ and $-7.8\pm 1.2\,\mkms$ respectively.  To infer the
+equivalent width of the Li\,\textsc{I} 6708\,\AA\ doublet, we followed
+the procedure described by \citet{bouma_2021_ngc2516}.  This yielded a
+strong detection: ${\rm EW}_{\rm Li} = 126^{+8}_{-4}$\,m\AA, with
+values consistent at $<1$-$\sigma$ between the two epochs.   The
+quoted value does not correct for the Fe\,\textsc{I} blend at
+6707.44\AA.
+
+{\bf TODO verify continuum normalization}
+Given the reddening-corrected color of the star this is
+consistent(ish) with the Pleiades/NGC2516 based on my NGC2516 figure.
+But it looks sub-IC2602 based on the lithium.png figure, which uses
+the Randich2018 values.
+It's pretty clearly super-field.
+I really need to verify the continuum normalization.
+COMPARE vs:
+\citealt{randich_gaiaeso_2018}).  
+\citet{berger_identifying_2018}
+
+\paragraph{High-Resolution Imaging}
+%FIXME TODO LEFT OFF HERE -- DESCRIBE!
 
 \subsection{KOI-7368}
+\paragraph{Spectra}
+For KOI-7368, we acquired TRES spectra on 2015 June 1.  The stellar
+parameter classification pipeline for these spectra is described by
+\citet{2021tsc2.confE.124B}, and is based on the template library
+originally constructed by \citet{buchhave_hatp16b_class_2010}.
+
+%FIXME TODO LEFT OFF HERE -- DESCRIBE!
+vsini?  Fe/H?  Li EW?
+
+\paragraph{High-Resolution Imaging}
+
 \subsection{KOI-7913}
+
+\paragraph{Spectra}
+
+KOI-7913 (YYYY/MM/DD and YYYY/MM/DD), where for the latter the two
+different epochs corresponded to observations of the secondary and
+primary respectively.
+
+\paragraph{High-Resolution Imaging}
 Is a binary.
@@
 
-%\appendix
-%\section{Young, Age-Dated, and Age-Dateable Star Compilation}
-%\label{app:targetlist}
+\appendix
+\section{Table of Candidate RSG-5 and CH-2 Members}
+\label{app:members}
+
+Gaia source lists and abbreviated info.  Include rotation periods.
+
+\section{Table of Transit Fit Parameters}
+\label{app:transit}
+
+Table goes here
 

### 2022-03-29 papers/ms.tex
update ms: spectra description

@@
 California Planet Survey \citep{howard_cps_2010}.  We derived the
-stellar spectral parameters ($T_{\rm eff}, \log g, R_\star$) using
+stellar parameters ($T_{\rm eff}, \log g, R_\star$) using
 \texttt{SpecMatch-Emp} \citep{yee_SM_2017}, which yielded values in
 $<1$-$\sigma$ agreement with those from the cluster-isochrone
-approach.  This approach also yielded $[Fe/H]=0.13 \pm 0.09$.  Using
+approach.  This approach also yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.  Using
 the broadened synthetic templates from \texttt{SpecMatch-Synth}
@@
 \paragraph{High-Resolution Imaging}
-%FIXME TODO LEFT OFF HERE -- DESCRIBE!
+{\bf todo}
 
@@
 \paragraph{Spectra}
-For KOI-7368, we acquired TRES spectra on 2015 June 1.  The stellar
-parameter classification pipeline for these spectra is described by
-\citet{2021tsc2.confE.124B}, and is based on the template library
-originally constructed by \citet{buchhave_hatp16b_class_2010}.
-
-%FIXME TODO LEFT OFF HERE -- DESCRIBE!
-vsini?  Fe/H?  Li EW?
+For KOI-7368, we acquired a spectrum using the echelle spectrograph
+(TRES; \citealt{furesz_tres_2008}) mounted at the Tillinghast 1.5m at
+the Fred Lawrence Whipple Observatory on Mt. Hopkins, AZ.  The
+spectrum was acquired on 2015 June 1.  The stellar parameter
+classification pipeline for these spectra is described by
+\citet{2021tsc2.confE.124B}.  It is based on the synthetic template
+library originally constructed by \citet{buchhave_hatp16b_class_2010}.
+The resulting stellar parameters ($T_{\rm eff}, \log g, R_\star$)
+agreed with those from the cluster-isochrone method within
+$1$-$\sigma$, though the effective temperature is more precise by a
+factor of three.  Auxiliary parameters included the metallicity $[{\rm
+Fe/H}]= -0.02 \pm 0.08$, the equatorial velocity $v\sin i = 20.21 \pm
+0.50\,\mkms$, and the systemic velocity ${\rm RV}_{\rm sys} = -10.9
+\pm 0.2\,\mkms$.  The Li 6708\AA\ EW measurement procedure yielded
+${\rm EW}_{\rm Li} = 236^{+16}_{-14}$\,m\AA.
 
@@
 
+\paragraph{Binarity}
+
+%(e.g., Jenkins KSCI-19081-002 KPDH pp 129, or many other pages)
+KOI-7913 appears to be a binary.  The north-west
+primary\footnote{Relevant identifiers for the primary are KIC 8873450
+= Gaia DR2/EDR3 2106235301785454208.  For the secondary, they are KIC
+8873448 = Gaia DR2/EDR3 2106235301785453824.  } is approximately 0.5
+magnitudes brighter than the south-east secondary in optical passbands.
+The two stars are separated in Gaia EDR3 by $3\farcs54$ on-sky, and
+have parallaxes consistent within $1$-$\sigma$ (the average is
+$\varpi=3.66 \pm 0.01$\,mas).  This corresponds to an apparent on-sky
+separation of $959 \pm 2$ au.  Gaia EDR3 quotes
+$(\mu_{\alpha'},\mu_\delta)=(2.439\pm0.015, 1.009\pm 0.016)\,{\rm
+mas}\,{\rm yr}^{-1}$ for the primary, and
+$(\mu_{\alpha'},\mu_\delta)=(2.696\pm0.020, 0.650 \pm 0.021) \,{\rm
+mas}\,{\rm yr}^{-1}$ for the secondary.  Since two stars were resolved
+in the Kepler Input Catalog and are roughly one Kepler pixel apart, an
+accurate crowding metric has already been applied in the NASA Ames
+data products to correct the mean flux level
+\citep{2017ksci.rept....6M}.  This is important for deriving accurate
+transit depths.
+
 \paragraph{Spectra}
 
-KOI-7913 (YYYY/MM/DD and YYYY/MM/DD), where for the latter the two
-different epochs corresponded to observations of the secondary and
-primary respectively.
+We acquired Keck/HIRES spectra for KOI-7913 A on the night of 2021 Nov
+13, and KOI-7913 B on the night of 2021 Oct 26.  The usual
+\texttt{SpecMatch-Emp} \citep{yee_SM_2017} machinery yielded $T_{\rm
+eff,A} = 4324 \pm 70\,{\rm K}$, $T_{\rm eff,B} = 4038 \pm 70\,{\rm
+K}$.  The remaining parameters were in agreement with those from the
+cluster isochrone.  For the primary, we also found $[{\rm Fe/H}]=
+-0.06 \pm 0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm
+sys} = -17.8 \pm 1.1\,\mkms$.  These same parameters for the secondary
+were $[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm 1.0\,\mkms$,
+and ${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.
+
+Regarding indicators of youth, the primary has a spectral type
+$\approx$K6V, while the secondary is $\approx$M0V.  At $\approx$40
+Myr, 
+lithium 6708\,\AA\ absorption is not expected for the
+secondary, but might be present for the primary 
+\citep[{\it e.g.},][Figure~8]{soderblom_ages_2014}.
+Conversely,  {\bf todo finish}
+% TODO: make this figure?
+
+% For Li:
+%    ~~~(CITE Figure 10 of Barrado y Navasceus 2004?).
+%   ****USE FIG8 OF SODERBLOM+13****
+% 
+%   OR USE BAFFLES AS YOUR REFERENCE?
+%     B-V=1.48 (primary), IC2602 age has no data.  beta pic age has lots of scatter
+%     around this color, between 10mA and 100mA.
+%     B-V=redder... 1.8?  SpType ~M0V, really no Li expected.
+%   A: no... they really didn't collect very good data from what i see.
+%   especially at the red end...
+% 
+% NW component SpType~K5.5-K6V; HIRES spectrum shows weak Li (EW 65 +/-
+% 8 mA), Halpha in emission (double-peaked?!), and CaHK in emission.
+% Reamatch SB2 checks don't show evidence for it being SB2.
+% SW component SpType~K9V-M0V; HIRES spectrum shows H-alpha in emission
+% (consistent with <150 Myr) and no Li (expected given Teff).
+% Given the kinematics, the spectra paint a picture consistent with KOI
+% 7913 being part of the ~40 Myr Cep-Her group.
 
 \paragraph{High-Resolution Imaging}
-Is a binary.
-The latter system is 3$\farcs$501 
-%FIXME FIXME FIXME TODO TODO TODO
 

### 2022-03-30 papers/ms.tex
update ms

@@
 \newcommand{\kms}{\,km\,s$^{-1}$}
-\newcommand{\mkms}{\,km\,s^{-1}}  % math mode
+\newcommand{\mkms}{{\rm \,km\,s^{-1}}}  % math mode
 \newcommand{\ms}{\,m\,s$^{-1}$}
@@
 \citep[{\it e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.
-Broadly speaking these analyses cluster on the stellar positions and
+To date these analyses have mostly clustered on the stellar positions and
 on-sky velocities measured by Gaia, with varying degrees of filtering
@@
 effects and internal dynamics.  As an extreme example, in the Hyades
-the velocities of stars in the tidal tails are expected to
-span up to $\pm 40$\kms relative to the cluster center
+the velocities of stars in the tidal tails are thought to
+span up to $\pm 40\mkms$ relative to the cluster center
 \citep{jerabkova_800_2021}.  The stars in such diffuse regions can be
@@
 sub-populations identified by \citet{Kerr2021} share a common origin
-has yet to be fully substantiated, and is the subject of the upcoming
+has yet to be fully substantiated, and will be the subject of the upcoming
 study by R.~Kerr and collaborators.  Our primary interest in the
@@
 distributed between these two extremes.  In Figure~\ref{fig:XYZvtang},
-we show 26{,}960 objects with weight exceeding 0.02 as gray points,
-and overplot 7{,}560 objects with weights exceeding 0.10 as black
-points.  {\bf TODO: FIXME these object numbers impose pretty weak
-quality cuts and include junk at G>18...  use
-get\_clean\_gaia\_photometric\_sources too, since they come in for the
-CAMDs}.
-
+we show 12{,}436 objects with weight exceeding 0.02 as gray points,
+and overplot 4{,}763 objects with weights exceeding 0.10 as black
+points.\footnote{These counts only include objects with reliable astrometry
+	and photometry: $\varpi/\sigma_\varpi>5$;
+	$G/\sigma_{G}>50$;
+	$G_{\rm RP}/\sigma_{G_{RP}}>20$;
+	$G_{\rm BP}/\sigma_{G_{BP}}>20$.}
 The previously known $\delta$~Lyr\ cluster ($l,b=68^\circ,15^\circ$;
-$v_{l'}, v_b=-4.5{\rm kms},-4{\rm kms}$) is visible, as is RSG-5
-($l,b=83^\circ,6^\circ$, $v_{l'}, v_b=+5.5{\rm kms},-3.5{\rm kms}$).
+$v_{l'}, v_b=-4.5 \mkms,-4 \mkms$) is visible, as is RSG-5
+($l,b=83^\circ,6^\circ$; $v_{l'}, v_b=5.5 \mkms,-3.5 \mkms$).
 Most of the other subclusters, including in Cep-Cyg
@@
 2022, and also compared against the \texttt{q1\_q17\_dr25} table
-\citep{thompson_planetary_2018}.  This yielded 32 matches, of which 18
-were known false positives, 7 were designated
-``confirmed'', and 8 were designated ``candidates''.  A cursory
-inspection of the Kepler data validation summaries and Robovetter
-classifications for these objects quickly showed whether they were
+\citep{thompson_planetary_2018}.  From the candidate members
+with weights exceeding 0.02, this yielded
+ yielded 25 matches, of which 11
+were known false positives, 6 were designated
+``confirmed'', and 8 were designated ``candidates''.  
+Inspection of the Kepler data validation summaries and Robovetter
+classifications for these objects then showed whether they were
 potentially consistent with being {\it i)} planets, and {\it ii)}
@@
 \citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}}
-dust maps and the extinction coefficients $k_X\equiv A_X/A_0$from
+dust maps and the extinction coefficients $k_X\equiv A_X/A_0$ from
 \citet{GaiaCollaboration2018}, assuming that $A_0 = 3.1 E(B-V)$.  This

### 2022-03-30 papers/ms.tex
update ms: imaging

@@
 The adopted stellar parameters are listed in
-Table~\ref{tab:starparams}.  The stellar surface gravity, radius,
+Table~\ref{tab:sysparams}.  The stellar surface gravity, radius,
 mass, and density are found by interpolating against the MIST
@@
 $<1$-$\sigma$ agreement with those from the cluster-isochrone
-approach.  This approach also yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.  Using
-the broadened synthetic templates from \texttt{SpecMatch-Synth}
+approach.  This approach also yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.
+Using the broadened synthetic templates from \texttt{SpecMatch-Synth}
 \citep{petigura_cksi_2017}, we found $v\sin i = 9.3 \pm 1.0\,\mkms$.
@@
 
-{\bf TODO verify continuum normalization}
-Given the reddening-corrected color of the star this is
-consistent(ish) with the Pleiades/NGC2516 based on my NGC2516 figure.
-But it looks sub-IC2602 based on the lithium.png figure, which uses
-the Randich2018 values.
-It's pretty clearly super-field.
-I really need to verify the continuum normalization.
-COMPARE vs:
-\citealt{randich_gaiaeso_2018}).  
-\citet{berger_identifying_2018}
+% {\bf TODO verify continuum normalization in Li EW measurement...}
+% Given the reddening-corrected color of the star this is
+% consistent(ish) with the Pleiades/NGC2516 based on my NGC2516 figure.
+% But it looks sub-IC2602 based on the lithium.png figure, which uses
+% the Randich2018 values.
+% It's pretty clearly super-field.
+% I really need to verify the continuum normalization.
+% COMPARE vs:
+% \citealt{randich_gaiaeso_2018}).  
+% \citet{berger_identifying_2018}
 
 \paragraph{High-Resolution Imaging}
-{\bf todo}
+% KOI-6186 & 58662.62 &   Kp     &  4 &   80.00 &  4.1 &  4.8 &  5.4 &  5.8 &  6.5 &  6.8 &  7.6 &  8.3 &  8.2 &  8.3 &             Petigura \\ % N2.20190628.53389
+We acquired adaptive optics imaging of Kepler-1643 on the night of
+2019 June 28 using the NIRC2 imager on Keck-II (PI: E.~Petigura).
+Using the narrow camera (FOV = 10.2\arcsec), we obtained 4 images in
+the $K'$ filter ($\lambda = 2.12\,\mu$m) with a total exposure time of
+320\,s. We analyzed these data following \citet{kraus_impact_2016},
+and determined the detection limits to visual companions by analyzing
+the residuals after subtracting an empirical PSF template. 
+This procedure yielded contrast limits of $\Delta K' = 4.1$ mag at
+$\rho = 150$ mas, $\Delta K' = 5.8$ mag at $\rho = 300$ mas, and
+$\Delta K' = 8.3$ mag at $\rho > 1000$ mas.
 
@@
 \paragraph{Spectra}
-For KOI-7368, we acquired a spectrum using the echelle spectrograph
-(TRES; \citealt{furesz_tres_2008}) mounted at the Tillinghast 1.5m at
-the Fred Lawrence Whipple Observatory on Mt. Hopkins, AZ.  The
-spectrum was acquired on 2015 June 1.  The stellar parameter
-classification pipeline for these spectra is described by
-\citet{2021tsc2.confE.124B}.  It is based on the synthetic template
-library originally constructed by \citet{buchhave_hatp16b_class_2010}.
-The resulting stellar parameters ($T_{\rm eff}, \log g, R_\star$)
-agreed with those from the cluster-isochrone method within
-$1$-$\sigma$, though the effective temperature is more precise by a
-factor of three.  Auxiliary parameters included the metallicity $[{\rm
-Fe/H}]= -0.02 \pm 0.08$, the equatorial velocity $v\sin i = 20.21 \pm
-0.50\,\mkms$, and the systemic velocity ${\rm RV}_{\rm sys} = -10.9
-\pm 0.2\,\mkms$.  The Li 6708\AA\ EW measurement procedure yielded
-${\rm EW}_{\rm Li} = 236^{+16}_{-14}$\,m\AA.
+For KOI-7368, we acquired a spectrum on 2015 June 1 using the echelle
+spectrograph (TRES; \citealt{furesz_tres_2008}) mounted at the
+Tillinghast 1.5m at the Fred Lawrence Whipple Observatory on Mt.
+Hopkins, AZ.  The stellar parameter classification pipeline for the
+TRES spectra has been described by \citet{2021tsc2.confE.124B}.  It is
+based on the synthetic template library originally constructed by
+\citet{buchhave_hatp16b_class_2010}.  The resulting stellar parameters
+($T_{\rm eff}, \log g, R_\star$) agreed with those from the
+cluster-isochrone method within $1$-$\sigma$, though the effective
+temperature is more precise by a factor of three.  Auxiliary
+spectroscopic parameters included the metallicity $[{\rm Fe/H}]= -0.02
+\pm 0.08$, the equatorial velocity $v\sin i = 20.21 \pm 0.50\,\mkms$,
+and the systemic velocity ${\rm RV}_{\rm sys} = -10.9 \pm 0.2\,\mkms$.
+The Li 6708\AA\ EW measurement procedure yielded ${\rm EW}_{\rm Li} =
+236^{+16}_{-14}$\,m\AA.
 
 \paragraph{High-Resolution Imaging}
+% KOI-7368 & 58646.57 &   Kp     &  4 &   80.00 &  5.2 &  5.8 &  6.4 &  6.7 &  7.1 &  7.5 &  8.1 &  8.7 &  8.7 &  8.8 &                Huber \\ % N2.20190612.49290
+We acquired adaptive optics imaging of KOI-7368 on the night of 2019
+June 12 using the NIRC2 imager on Keck-II (PI: D.~Huber).  The
+observational configuration and reduction was identical as for
+Kepler-1643.  No companions were detected, and the analysis of the
+image residuals yielded contrast limits of $\Delta K' = 5.2$ mag at
+$\rho = 150$ mas, $\Delta K' = 6.7$ mag at $\rho = 300$ mas, and
+$\Delta K' = 8.7$ mag at $\rho > 1000$ mas.
 
@@
 %(e.g., Jenkins KSCI-19081-002 KPDH pp 129, or many other pages)
-KOI-7913 appears to be a binary.  The north-west
-primary\footnote{Relevant identifiers for the primary are KIC 8873450
-= Gaia DR2/EDR3 2106235301785454208.  For the secondary, they are KIC
-8873448 = Gaia DR2/EDR3 2106235301785453824.  } is approximately 0.5
-magnitudes brighter than the south-east secondary in optical passbands.
-The two stars are separated in Gaia EDR3 by $3\farcs54$ on-sky, and
-have parallaxes consistent within $1$-$\sigma$ (the average is
-$\varpi=3.66 \pm 0.01$\,mas).  This corresponds to an apparent on-sky
-separation of $959 \pm 2$ au.  Gaia EDR3 quotes
+KOI-7913 appears to be a binary.\footnote{Relevant identifiers for the
+primary are KIC 8873450 = Gaia DR2/EDR3 2106235301785454208.  For the
+secondary, they are KIC 8873448 = Gaia DR2/EDR3 2106235301785453824.}
+The north-west primary is $\approx$0.5 magnitudes brighter than the
+south-east secondary in optical passbands.  The two stars are
+separated in Gaia EDR3 by $3\farcs54$ on-sky, and have parallaxes
+consistent within $1$-$\sigma$ (with an average $\varpi=3.66 \pm
+0.01$\,mas).  This corresponds to an apparent on-sky separation of
+$959 \pm 2$ au.  Gaia EDR3 quotes
 $(\mu_{\alpha'},\mu_\delta)=(2.439\pm0.015, 1.009\pm 0.016)\,{\rm
@@
 $\approx$K6V, while the secondary is $\approx$M0V.  At $\approx$40
-Myr, 
-lithium 6708\,\AA\ absorption is not expected for the
-secondary, but might be present for the primary 
-\citep[{\it e.g.},][Figure~8]{soderblom_ages_2014}.
-Conversely,  {\bf todo finish}
-% TODO: make this figure?
+Myr, lithium 6708\,\AA\ absorption is not expected for the secondary,
+but might be present for the primary \citep[{\it
+e.g.},][Figure~8]{soderblom_ages_2014}.  Conversely, H$\alpha$ is
+expected to be in emission for for all M-dwarfs younger than
+$\approx150$\,Myr \citep{kiman_calibration_2021}.
+
+Turning to the actual data, we find {\bf X for lithium}.
+For H$\alpha$, {\bf we see Y}.
 
@@
 \paragraph{High-Resolution Imaging}
+
+We acquired adaptive optics imaging of KOI-7913 on the night of 2020
+Aug 27 using the NIRC2 imager on Keck-II (PI: D.~Huber).  The
+observational configuration and reduction was identical as before.
+The images showed KOI-7913 A, KOI-7913 B, and an additional faint star
+$\approx0\farcs99$ due East of KOI-7913 B.  Applying the PSF-fitting
+routines from \citet{kraus_impact_2016}, the tertiary star has a
+separation $\rho = 4397 \pm 3$\,mas from the primary, at a position
+angle $231.17^\circ \pm 0.02^\circ$, with a $\Delta K' = 6.97 \pm
+0.04$.  While it is too faint to affect the interpretation of the KOI
+itself, it would be amusing if the faint neighbor were also comoving
+and young -- at the age of the system, it would have a mass roughly
+between 10 and 15\,M$_{\rm Jup}$.  Additional imaging epochs will be
+required to tell.
 

### 2022-03-31 papers/ms.tex
update ms

@@
 		\subfloat{
-			\includegraphics[width=0.41\textwidth]{f3a.pdf}
-			\includegraphics[width=0.41\textwidth]{f3b.pdf}
+			\includegraphics[width=0.33\textwidth]{f3a.pdf}
+			\includegraphics[width=0.33\textwidth]{f3b.pdf}
 		}
 
-		\vspace{-1.45cm}	
+		\vspace{-1.27cm}	
 		\subfloat{
-			\includegraphics[width=0.41\textwidth]{f3c.pdf}
-			\includegraphics[width=0.41\textwidth]{f3d.pdf}
+			\includegraphics[width=0.33\textwidth]{f3c.pdf}
+			\includegraphics[width=0.33\textwidth]{f3d.pdf}
 		}
 	\end{center}
-	\vspace{-0.7cm}
+	\vspace{-0.5cm}
 	\caption{
-		{\bf Raw and processed light curves for the objects of
-    interest.} Top: raw.  Bottom: processed.
-    The increased scatter during transit is likely due to starspot
-    crossing events.  KOI-7913 is janky, but P=24 days.
+		{\bf Raw and processed light curves for the Kepler Objects of
+    Interest in Cep-Her.}   
+    {\it Top}:
+    50 day light curve segment from the 3.9 years of Kepler data.  The
+    ordinate shows the \texttt{PDCSAP} median-subtracted flux in units
+    of parts-per-thousand ($\times 10^{-3}$).  The dominant signal is
+    induced by starspots.  
+    {\it Bottom}:
+    Phase-folded transits of Kepler-1643, KOI-7913, KOI-7368, and
+    Kepler-1627 with stellar variability removed.  The maximum {\it a
+    posteriori} transit model is shown with the gray line, and the
+    residual after subtracting the transit model is vertially
+    displaced.  Windows over 10 hours are shown.  Gray points are
+    individual flux measurements; black points are binned to 20 minute
+    intervals, and have a representative 1-$\sigma$ error bar in the
+    center-right of each panel. 
 		\label{fig:planets}
@@
 \end{figure*}
+
+\subsection{Kepler Data}
+
+The Kepler space telescope observed Kepler-1643, KOI-7913, and
+KOI-7368 at a 30-minute cadence between May 2009 and April 2013.  For
+all three systems quarters 1 through 17 were observed without any data
+gaps other than the usual quarterly roll.  The top panel of
+Figure~\ref{fig:planets} shows a 50-day slice of the \texttt{PDCSAP}
+light curves for the three new Cep-Her candidates, along with
+Kepler-1627.  In \texttt{PDCSAP}, nonastrophysical variability is
+removed through a cotrending approach that uses a set of basis vectors
+derived using singular value decomposition of a set of
+systematics-dominated stellar light curves
+\citep{smith_kepler_PDC_2017}.  In our analysis, we used the
+\texttt{PDCSAP} light curves with the default optimal aperture
+\citep{smith_finding_2016}.  Cadences with non-zero quality flags were
+omitted.  In all cases, the stars are dominated by spot-induced
+modulation with peak-to-peak variability between 2\% and 10\%.  These
+signals are much larger than the transits, which have depth
+$\approx$0.1\%.
+To quantify the stellar rotation periods, we calculated the
+Lomb-Scargle periodogram for each Kepler quarter independently.  The
+resulting means and standard deviations are quoted in
+Table~\ref{tab:sysparams}.
+
+\subsection{Transit and Stellar Variability Model}
+
+Our goals in fitting the Kepler light curves are twofold.  First, we
+want to derive accurate planetary sizes and orbital properties.
+Second, we want to remove the stellar variability signal to enable a
+statistical assessment of the probability that the transit signals are
+planetary.
+
+Our adopted approach was as follows.  Given an initial guess for the
+transit ephemeris from the \texttt{q1\_q17\_dr25} analysis
+\citep{thompson_planetary_2018}, the light curve was first trimmed to
+a local window around each transit of duration $\pm 3\times t_{14}$.
+The local out-of-transit points were then fitted with a fourth-order
+polynomial, which was then divided out from the light curve.  The
+resulting flattened transits were then fitted with a 
+planetary transit under the assumption of quadratic limb darkening.
+Our model therefore included 8 free parameters for the transit ($\{P,
+t_0, \log R_{\rm p}/R_\star, b, u_1 ,u_2, R_\star, \log g\}$), 2 free
+parameters for the light curve normalization and a white noise jitter
+($\{\langle f \rangle, \sigma_f \}$), and an additional 5 implicit
+parameters per transit through the polynomial.
+
+We fitted the models using \texttt{exoplanet}
+\citep{exoplanet:exoplanet}.  We assumed a Gaussian likelihood, and
+sampled using \texttt{PyMC3}'s No-U-Turn Sampler
+\citep{hoffman_no-u-turn_2014}, after having initialized using the
+maximum {\it a posteriori} (MAP) model.  We used the
+\citet{gelman_inference_1992} statistic ($\hat{R}$) as our convergence
+diagnostic.  Abbreviated results are in Table~\ref{fig:sysparams}, the
+full results are in Appendix~\ref{app:transit}.
+
+One weakness of our fitting approach is that we have fixed 5 implicit
+parameters per individual transit to their MAP values to remove the
+stellar variability signal.  An alternative could be to fit the
+planetary transits simultaneously with the starspot-induced
+variability using, say, a quasiperiodic Gaussian process (GP).  We
+explored this approach, but found that it required careful fine-tuning
+of the hyperparameter priors, otherwise the GP would tend to
+incorporate anything that was not part of the transit, even if it was
+short-term variability attributable to for instance flares, or simply
+instrumental noise.  This failure mode is particularly pernicious, in
+that it yields an inflated and ill-founded sense of confidence in the
+transit model results.  While our model is simpler, it has the benefit
+that its white noise jitter never trades off with any parameter
+equivalent to a damping timescale for the coherence of the GP.  It is
+also computationally efficient, while capturing the planetary
+parameters that we actually care about.
+
+\subsection{Planet Validation}
+
+%TODO : write a paragraph or two about previous VESPA work, and how we
+% went further.
+
+% Previous analyses have identified and characterized the smaller
+% transit signal \citep{2012ApJS..199...24T,thompson_planetary_2018},
+% validated its planetary nature \citep{morton_false_2016}, and even
+% searched the system for transit timing variations
+% \citep{holczer_transit_2016}.  Nonetheless, since the cluster
+% membership provides us with more precise stellar parameters than those
+% previously available, we opted to reanalyze the light curve.
 

### 2022-04-04 papers/ms.tex
update ms

@@
   all most likely planets, with false positive probabilities of
-  $6\times10^{-9}$, $5\times10^{-3}$, and $1\times10^{-4}$ for
+  $6\times10^{-9}$, $4\times10^{-3}$, and $1\times10^{-4}$ for
   Kepler-1643, KOI-7368, and KOI-7913 respectively.
@@
 %(e.g., Jenkins KSCI-19081-002 KPDH pp 129, or many other pages)
-KOI-7913 appears to be a binary.\footnote{Relevant identifiers for the
+KOI-7913 is almost certainly a binary.\footnote{Relevant identifiers for the
 primary are KIC 8873450 = Gaia DR2/EDR3 2106235301785454208.  For the
@@
 \subsection{Transit and Stellar Variability Model}
+\label{sec:fitting}
 
@@
 \citet{gelman_inference_1992} statistic ($\hat{R}$) as our convergence
-diagnostic.  Abbreviated results are in Table~\ref{fig:sysparams}, the
-full results are in Appendix~\ref{app:transit}.
+diagnostic.  The results are shown in the lower panels of
+Figure~\ref{fig:planets}, and the important derived parameters are in
+Table~\ref{tab:sysparams}.  The set of full parameters with their
+priors are given in Appendix~\ref{app:transit}.
 
@@
 of the hyperparameter priors, otherwise the GP would tend to
-incorporate anything that was not part of the transit, even if it was
-short-term variability attributable to for instance flares, or simply
-instrumental noise.  This failure mode is particularly pernicious, in
-that it yields an inflated and ill-founded sense of confidence in the
-transit model results.  While our model is simpler, it has the benefit
-that its white noise jitter never trades off with any parameter
+incorporate any variability that was not part of the transit, even if it was
+short-term variability attributable to flares, or simply
+instrumental noise.  This failure mode is somewhat pernicious, in that
+it yields an ill-founded sense of confidence in an ``overly clean''
+transit model fit.  Although our model is simpler, it has the benefit
+that the white noise jitter never trades off with any parameter
 equivalent to a damping timescale for the coherence of the GP.  It is
-also computationally efficient, while capturing the planetary
+also computationally efficient, and it captures the planetary
 parameters that we actually care about.
@@
 \subsection{Planet Validation}
+
+In the future, it may be possible to obtain independent evidence for
+the Cep-Her planets that directly confirms their planetary nature.
+The most likely path for that would be through transit
+observations with a high-resolution high-throughput spectrograph
+\citep[see][]{bouma_kep1627_2022}.  In the interim, it is of some
+interest whether any type of astrophysical false positive is likely to
+explain the transit signals, or whether they can be argued to be
+planetary on a statistical basis.  We adopt the Bayesian framework
+implemented in \texttt{VESPA} for this purpose
+\citep{morton_efficient_2012,vespa_2015}.  Briefly summarized, the
+priors in \texttt{VESPA} take as given the binary star occurrence rate
+from \citet{raghavan_survey_2010}, direction-specific star counts from
+\citet{girardi_star_2005}, and planet occurrence rates as described by
+\citet[][Section~3.4]{morton_efficient_2012}.  The likelihoods are
+then evaluated by forward-modeling a synthetic population of eclipsing
+bodies for each possible astrophysical model class, in which each
+population member has a particular trapezoidal eclipse depth, total
+duration, and ingress duration.  These summary statistics are then
+compared against the actual photometric data to evaluate the 
+probabilities of false positive scenarios such as foreground eclipsing
+binaries, hierarchical eclipsing binaries, and background eclipsing
+binaries.
+
+\paragraph{Kepler-1643}
+Kepler-1643~b (KOI-6186.01) was already validated as a transiting
+planet by \citet{morton_false_2016}.  Based on the Q1-Q17 DR25 False
+Positive Probability table hosted at the NASA Exoplanet Archive, they
+found a probability for any of the considered false positive scenarios
+of 9$\times$10$^{-6}$.  Repeating the calculation with our own
+stellar-variability correction and the new NIRC2 imaging constraints,
+we find ${\rm FPP} = 6\times10^{-9}$, where the improvement is
+presumably due to the new NIRC2 contrast curves.
+Figure~\ref{fig:planets} shows the intuitive justification for the
+strong statistical result: the transit is flat and has a reasonably
+high S/N ($\approx 47$), which is difficult to reproduce with
+eclipsing binary models.
+
+An interesting note on Kepler-1643 is that it formally fails the
+default centroid shift test in both the \texttt{q1\_q17\_dr25\_koi}
+and \texttt{q1\_q17\_dr25\_sup\_koi} releases, with the angular
+distance between the target star position and the location of the
+transiting source being reported as $0\farcs99$ (4.4-$\sigma$).  This
+seems to be mostly due to two outlying quarters (Q2 and Q6) driving
+the offset -- the centroid locations from the other Kepler quarters
+are all consistent within $\lesssim 0\farcs5$.  Nonetheless, this is
+an instructive exercise in how highly variable stars will complicate
+centroid-based vetting tests.  The centroid shifts measured by these
+tests are determined from the in- and out-of-transit flux-weighted
+centroids, but for our stars there is no static baseline in either in-
+or out-of-transit phase.
+
+\paragraph{KOI-7368}
+KOI-7368.01 is listed on the NASA Exoplanet Archive as a ``CANDIDATE''
+planet.  The Q1-Q17 DR25 False Positive Probability table however
+lists its probability of being an astrophysical false positive as
+``1.0''. However, given that none of the other columns in this table
+are populated, this is most likely a database entry error ({\bf ask
+Jessie}).  Performing the validation calculation using our new NIRC2
+imaging constraints, we find ${\rm FPP} \approx 4\times10^{-3}$.
+Though not as convincing as Kepler-1643, this formally clears the
+usual threshold for calling the planet statistically validated
+\citep{morton_efficient_2012}.  The S/N of the transit itself is
+$\approx 32$, which indicates that it is unlikely to be caused by
+systematic noise in the light curve.  Visually, one can also inspect
+each individual transit window in the light curve itself.  This
+exercise shows a weak notch in each case; over the 198 observed
+transits, the resulting transit signal seems sensible.
+
+Why was KOI-7368 not previously validated?  One reason could be the
+increased stellar variability, or the lower S/N of the transit.
+Another reason could be that it also formally fails the centroid shift
+test, just like Kepler-1643.  In this case, the reported offset is
+somewhat less significant ($0\farcs22$; 3.0-$\sigma$).  Again,
+inspection of the data validation reports shows that this is caused by
+a few outlying quarters: in this case Q4, Q5, Q8, and Q12.
+The fact that the remaining quarters show a far more consistent
+scatter suggests that these outliers are caused by local stellar
+variability, rather than an astrophysical offset.
+Our imaging also shows that there are no neighboring sources that
+could cause the offset.
+
+\paragraph{KOI-7913}
+KOI-7913.01 is listed on the NASA Exoplanet Archive as a ``CANDIDATE''
+planet.  The Q1-Q17 DR25 False Positive Probability table lists its
+probability of being an astrophysical false positive as
+$1.4\times10^{-4}$, with the most likely false positive scenario being
+a blended eclipsing binary.  Repeating the calculation with our new
+detrending and contrast curves, we find a similar result: ${\rm FPP} =
+1.3\times10^{-4}$.  While the transit has the lowest S/N of any of the
+objects discusssed (${\rm S/N}=14$), its marginally lower FPP relative
+to KOI-7368 can be understood through its relatively flat-bottomed
+shape, combined with its long transit duration relative to plausible
+eclipsing binary models.
+
+The disposition of KOI-7913.01 seems to have been previously debated:
+in \texttt{q1\_q17\_dr25\_koi} it was flagged as a ``FALSE POSITIVE'',
+with the comment ``CENT\_KIC\_POS---HALO\_GHOST''.  This comment and
+disposition seem to have been removed in the
+\texttt{q1\_q17\_dr25\_sup\_koi} data release, which renamed it a
+``CANDIDATE'', without any flag given.  As described by
+\citet{thompson_planetary_2018}, the ``CENT\_KIC\_POS'' flag is an
+indication that the measured source centroid is offset from its
+expected location in the Kepler Input Catalog.  The final Data
+Validation Reports (generated 2016 Jan 30) do not seem to show this,
+and the statistical significance of any centroid offset seems to be
+lower than for KOI-7368 and Kepler-1643 (which, as we have argued,
+both show centroid offsets that seem to be explained by the stellar
+variability).
+
+What of the ``HALO\_GHOST'' flag?  This test measures the transit
+strength for the pixels inside the aperture, and the ring of pixels
+around that aperture (the ``halo'').  Generally one expects the
+transit signal to be strongest in the central aperture.  Two types
+of false positive scenarios can change this, triggering the flag: the first is
+when optical ghosts from bright eclipsing binaries reflect off the
+CCD, and contaminate the target star.  The second is when the PRF of
+nearby stars directly overlaps with the PRF of the target star (see
+\citealt{thompson_planetary_2018}, Section A.5.2).
+
+The most obvious explanation for KOI-7913 is the latter case, since
+KOI-7913 B is $\approx0.9$ Kepler pixels away from Kepler-7913 A.
+Due to the on-sky orientation of KOI-7913 A and
+KOI-7913 B, the default ``optimal
+aperture'' selected in Q3, Q7, Q11, and Q15 in fact included both
+stars, while for the remaining quarters KOI-7913 B was (correctly)
+excluded from the optimal aperture
+(see pages 35 through 71 of the Data Validation Report.)
+
+{\bf
+TODO: words!  Describe Halo ghost statistic.  Why should I, or
+shouldn't I, be worried?
+}
 
@@
   %\texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
-  \texttt{corner} \citep{corner_2016},
+  %\texttt{corner} \citep{corner_2016},
   %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
@@
   %\texttt{tesscut} \citep{brasseur_astrocut_2019},
-	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
+	\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
   %\texttt{webplotdigitzer} \citep{rohatgi_2019},
@@
 
-Table goes here
+Tables~\ref{tab:kepler1643}, \ref{tab:koi7368}, \ref{tab:koi7913} give
+the full set of fitted and derived parameters from the model described in
+Section~\ref{sec:fitting}.
+Priors and convergence statistics are also listed.
+
+\input{source_Kepler_1643_simpletransit_posteriortable.tex}
+\input{source_KOI_7368_simpletransit_posteriortable.tex}
+\input{source_KOI_7913_simpletransit_posteriortable.tex}
 

### 2022-04-05 papers/ms.tex
update ms

@@
 expected to be in emission for for all M-dwarfs younger than
-$\approx150$\,Myr \citep{kiman_calibration_2021}.
+$\approx150$\,Myr \citep{kiman_calibration_2021}.  At $\approx$M0V, if
+the secondary were in older than $\approx 500\,{\rm Myr}$, it would be
+more common for its H$\alpha$ component to be observed in absorption
+\citep[][Figure~4]{kiman_calibration_2021}.  H$\alpha$ emission in
+late K-dwarfs is rarer still. Based on the data from K.~Rocio et al's
+analysis\footnote{Available at
+\url{https://zenodo.org/record/4659293#.YkuKh27MKDV}}, a $\approx$K6V
+star should be emitting in H$\alpha$ at 40\,Myr, but not typically
+past $\approx200\,{\rm Myr}$.  This presents a fairly direct test for
+the star's age.
 
@@
 In the future, it may be possible to obtain independent evidence for
-the Cep-Her planets that directly confirms their planetary nature.
-The most likely path for that would be through transit
-observations with a high-resolution high-throughput spectrograph
-\citep[see][]{bouma_kep1627_2022}.  In the interim, it is of some
-interest whether any type of astrophysical false positive is likely to
-explain the transit signals, or whether they can be argued to be
+the Cep-Her planets to directly confirm their planetary nature.  The
+most likely path would be through transit observations using a
+high-resolution spectrograph \citep[see][]{bouma_kep1627_2022}.  In
+the interim, it is of interest whether the transit signals could be
+astrophysical false positives, or whether they can be argued to be
 planetary on a statistical basis.  We adopt the Bayesian framework
@@
 
-An interesting note on Kepler-1643 is that it formally fails the
-default centroid shift test in both the \texttt{q1\_q17\_dr25\_koi}
-and \texttt{q1\_q17\_dr25\_sup\_koi} releases, with the angular
-distance between the target star position and the location of the
-transiting source being reported as $0\farcs99$ (4.4-$\sigma$).  This
-seems to be mostly due to two outlying quarters (Q2 and Q6) driving
-the offset -- the centroid locations from the other Kepler quarters
-are all consistent within $\lesssim 0\farcs5$.  Nonetheless, this is
-an instructive exercise in how highly variable stars will complicate
-centroid-based vetting tests.  The centroid shifts measured by these
-tests are determined from the in- and out-of-transit flux-weighted
-centroids, but for our stars there is no static baseline in either in-
-or out-of-transit phase.
+An interesting aside on Kepler-1643 is that it formally failed the
+centroid shift test in the \texttt{q1\_q17\_dr25\_koi} data release,
+with the angular distance between the target star position and the
+location of the transiting source being reported as $0\farcs99$
+(4.4-$\sigma$).  This seems to be mostly due to two outlying quarters
+(Q2 and Q6) driving the offset -- the centroid locations from the
+other Kepler quarters are all consistent within $\lesssim 0\farcs5$.
+Nonetheless, this is an instructive exercise in how variable stars can
+complicate centroid-based vetting tests.  The centroid shifts measured
+by these tests are determined from the in- and out-of-transit
+flux-weighted centroids, but for our stars there is no static baseline
+in either in- or out-of-transit phase.
 
@@
 ``1.0''. However, given that none of the other columns in this table
-are populated, this is most likely a database entry error ({\bf ask
+are populated, this is likely a database entry error ({\bf ask
 Jessie}).  Performing the validation calculation using our new NIRC2
 imaging constraints, we find ${\rm FPP} \approx 4\times10^{-3}$.
-Though not as convincing as Kepler-1643, this formally clears the
+While not as convincing as Kepler-1643, this formally clears the
 usual threshold for calling the planet statistically validated
@@
 exercise shows a weak notch in each case; over the 198 observed
-transits, the resulting transit signal seems sensible.
+transits, the resulting signal seems sensible.
 
 Why was KOI-7368 not previously validated?  One reason could be the
-increased stellar variability, or the lower S/N of the transit.
-Another reason could be that it also formally fails the centroid shift
-test, just like Kepler-1643.  In this case, the reported offset is
-somewhat less significant ($0\farcs22$; 3.0-$\sigma$).  Again,
-inspection of the data validation reports shows that this is caused by
-a few outlying quarters: in this case Q4, Q5, Q8, and Q12.
-The fact that the remaining quarters show a far more consistent
-scatter suggests that these outliers are caused by local stellar
-variability, rather than an astrophysical offset.
-Our imaging also shows that there are no neighboring sources that
-could cause the offset.
+high stellar variability.  Another reason could be that it also
+formally fails the centroid shift test, similar to Kepler-1643.  In
+the case of KOI-7368, the reported offset is marginally less
+significant ($0\farcs22$; 3.0-$\sigma$).  Again, inspection of the
+data validation reports shows that this is caused by a few outlying
+quarters: in this case Q4, Q5, Q8, and Q12.  The fact that the
+remaining quarters show a more consistent scatter suggests that these
+outliers are again caused by stellar variability local to the
+transits, rather than by an astrophysical offset.  Our imaging also
+shows that there are no neighboring sources that could cause an
+offset of the observed amplitude.
 
@@
 1.3\times10^{-4}$.  While the transit has the lowest S/N of any of the
-objects discusssed (${\rm S/N}=14$), its marginally lower FPP relative
-to KOI-7368 can be understood through its relatively flat-bottomed
-shape, combined with its long transit duration relative to plausible
-eclipsing binary models.
-
-The disposition of KOI-7913.01 seems to have been previously debated:
-in \texttt{q1\_q17\_dr25\_koi} it was flagged as a ``FALSE POSITIVE'',
+objects discusssed (${\rm S/N}=14$), its lower FPP relative to
+KOI-7368 can be understood through its relatively flat-bottomed shape,
+combined with its long transit duration relative to plausible
+eclipsing binary models (Figure~\ref{fig:planets}).
+
+The disposition of KOI-7913.01 has previously been debated:
+in \texttt{q1\_q17\_dr25\_koi} the source was flagged as a ``FALSE POSITIVE'',
 with the comment ``CENT\_KIC\_POS---HALO\_GHOST''.  This comment and
@@
 expected location in the Kepler Input Catalog.  The final Data
-Validation Reports (generated 2016 Jan 30) do not seem to show this,
-and the statistical significance of any centroid offset seems to be
-lower than for KOI-7368 and Kepler-1643 (which, as we have argued,
+Validation Reports, generated 2016 Jan 30, do not seem to show this.
+Moreover, the statistical significance of any centroid offset seems to
+be lower than for KOI-7368 and Kepler-1643 (which, as we have argued,
 both show centroid offsets that seem to be explained by the stellar
@@
 What of the ``HALO\_GHOST'' flag?  This test measures the transit
-strength for the pixels inside the aperture, and the ring of pixels
-around that aperture (the ``halo'').  Generally one expects the
-transit signal to be strongest in the central aperture.  Two types
-of false positive scenarios can change this, triggering the flag: the first is
-when optical ghosts from bright eclipsing binaries reflect off the
-CCD, and contaminate the target star.  The second is when the PRF of
-nearby stars directly overlaps with the PRF of the target star (see
+strength for the pixels inside the aperture, and compares it to that
+measured in the ring of pixels around said aperture (the ``halo'').
+Generally one expects the transit signal to be strongest in the
+central aperture, rather than the halo.  Two types of false positive
+scenarios can change this and trigger the flag: the first is when
+optical ghosts from bright eclipsing binaries reflect off the CCD, and
+contaminate the target star.  The second is when the PRF of nearby
+stars directly overlaps with the PRF of the target star (see
 \citealt{thompson_planetary_2018}, Section A.5.2).
@@
 The most obvious explanation for KOI-7913 is the latter case, since
-KOI-7913 B is $\approx0.9$ Kepler pixels away from Kepler-7913 A.
-Due to the on-sky orientation of KOI-7913 A and
-KOI-7913 B, the default ``optimal
-aperture'' selected in Q3, Q7, Q11, and Q15 in fact included both
-stars, while for the remaining quarters KOI-7913 B was (correctly)
-excluded from the optimal aperture
-(see pages 35 through 71 of the Data Validation Report.)
-
-{\bf
-TODO: words!  Describe Halo ghost statistic.  Why should I, or
-shouldn't I, be worried?
-}
-
-%TODO : write a paragraph or two about previous VESPA work, and how we
-% went further.
-
-% Previous analyses have identified and characterized the smaller
-% transit signal \citep{2012ApJS..199...24T,thompson_planetary_2018},
-% validated its planetary nature \citep{morton_false_2016}, and even
-% searched the system for transit timing variations
-% \citep{holczer_transit_2016}.  Nonetheless, since the cluster
-% membership provides us with more precise stellar parameters than those
-% previously available, we opted to reanalyze the light curve.
+KOI-7913 B is $\approx0.9$ Kepler pixels away from Kepler-7913 A.  Due
+to the on-sky orientation of KOI-7913 A and KOI-7913 B, the default
+``optimal aperture'' selected in Q3, Q7, Q11, and Q15 in fact included
+both stars, while for the remaining quarters KOI-7913 B was
+(correctly) excluded from the optimal aperture (see pages 35 through
+71 of the Data Validation Report.)
+
+As a test, we opted to re-extract the light curves from the target
+pixel files for these particular cases.
+{\bf Comparing the pixels on-target, and found}
+%FIXME FIXME TODO TODO
+{\bf We found...}
 

### 2022-04-05 papers/ms.tex
update ms

@@
 \title{
-  Three 40 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
+  Three 40 to 50 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
 }
@@
   Stellar positions and velocities from Gaia are yielding a new and
-  refined view of how stellar clusters evolve.
+  refined view of open cluster dispersal.
   Here we present an analysis of a group of $\approx$40 million
@@
   Kepler-1643, KOI-7368, and KOI-7913 respectively.
-  These planets are therefore the first empirical
-  demonstration that mini-Neptunes with sizes of $\approx$2 Earth
+  These planets therefore empirically
+  demonstrate that mini-Neptunes with sizes of $\approx$2 Earth
   radii exist at ages of $\approx$40 million years.
@@
 three.  After analyzing the relevant properties of Cep-Her
-(Section~\ref{sec:cluster}), we discuss the stellar properties
+(Section~\ref{sec:cluster}), we derive the stellar properties
 (Section~\ref{sec:stars}) and validate the planetary nature of each
 system using a combination of the Kepler photometry and
-high-resolution imaging (Section~\ref{sec:planets}).  We conclude with
-a discussion of implications for the size-evolution of close-in
+high-resolution imaging (Section~\ref{sec:planets}).  
+We conclude with a few caveats, 
+and discuss implications for the size-evolution of close-in
 mini-Neptunes (Section~\ref{sec:disc_conc}).
@@
 		{\bf The stellar groups near KOI-7368, KOI-7913, and Kepler-1643
-    are $\approx$40 million years old.} 
+    are 40 to 50 million years old.} 
     {\it Top row}: 
@@
 Myr, lithium 6708\,\AA\ absorption is not expected for the secondary,
-but might be present for the primary \citep[{\it
+but might be present in the primary \citep[{\it
 e.g.},][Figure~8]{soderblom_ages_2014}.  Conversely, H$\alpha$ is
@@
 analysis\footnote{Available at
-\url{https://zenodo.org/record/4659293#.YkuKh27MKDV}}, a $\approx$K6V
+\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K6V
 star should be emitting in H$\alpha$ at 40\,Myr, but not typically
@@
 
-\section{Discussion \& Conclusions}
+\section{Discussion \& Conclusion}
 \label{sec:disc_conc}
+
+\begin{figure*}[!t]
+	\begin{center}
+		\leavevmode
+		\includegraphics[width=0.9\textwidth]{f4.pdf}
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{
+		%    namelist = ['Kepler-1627 A', 'KOI-7368', 'KOI-7913 A', 'KOI-7913 B', 'Kepler-1643']
+		%    markers = ['P', 'v', 'X', 'X', 's']
+		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
+		Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
+		3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
+		its uncertainty. Kepler-1627 (+), KOI-7368 (down-triangle),
+		KOI-7913 (X), Kepler-1643 (diamond).  The large sizes of the
+		youngest transiting planets could be explained by their primordial
+		atmospheres not yet having evaporated; direct measurements of the
+		atmospheric outflows or planetary masses would help to confirm
+		this expectation.  Selection effects may also be important.
+		Parameters are from the NASA Exoplanet Archive (2022 Feb 27).
+		\label{fig:rp_period_age}
+	}
+\end{figure*}
 
@@
 
-\begin{figure*}[!t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f4.pdf}
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{
-    %    namelist = ['Kepler-1627 A', 'KOI-7368', 'KOI-7913 A', 'KOI-7913 B', 'Kepler-1643']
-    %    markers = ['P', 'v', 'X', 'X', 's']
-		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
-		Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
-		3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
-    its uncertainty. Kepler-1627 (+), KOI-7368 (down-triangle),
-    KOI-7913 (X), Kepler-1643 (diamond).  The large sizes of
-		the youngest transiting planets could be explained by their
-		primordial atmospheres not yet having evaporated; direct
-		measurements of the atmospheric outflows or planetary masses would
-		help to confirm this expectation.  Selection effects may also be
-		important.  Parameters are from the NASA Exoplanet Archive (2022
-		Feb 27).
-		\label{fig:rp_period_age}
-	}
-\end{figure*}
+\subsection{Normal-Sized Mini-Neptunes Exist at 40$\,$Myr}
+
+Perhaps the most significant point about
+Kepler-1643~b, KOI-7368.01, and KOI-7913 is that they have normal
+sizes.
+{\bf TODO for LGB: write more about this}
+
+\subsection{Summary}
+
+We have argued that Kepler-1643~b, KOI-7368.01, and KOI-7913 {\it i)}
+are all $\approx$40 million years old, and {\it ii)} that each signal
+is most likely planetary.  The evidence for the planetary
+interpretation comes from an application of \texttt{VESPA} to the
+Kepler data, alongside new adaptive optics imaging from NIRC2.  The
+validity of the \texttt{VESPA} framework rests on the premise that
+non-astrophysical false positives, including false alarms and known
+neighboring eclipsing binaries, can be rejected.  Kepler-1643 and
+KOI-7368 both pass all but one of the canonical vetting checks: the
+check for centroid offsets between the in and out-of-transit phases.
+For both cases, we have argued the weak observed shifts are most
+likely caused by starspot-induced variability in specific quarters
+shifting the stellar center-of-light.
+
+Each system has multiple indicators of youth that support their
+reported ages.  For Kepler-1643, perhaps the strongest youth indicator
+is its physical and kinematic association with RSG-5.  Based on the
+color-absolute magnitude diagram, we seem to be able to select members
+of this cluster with a false positive rate of $\ll 5\%$
+(Figure~\ref{fig:age}), and Kepler-1643 is one such member.  Its
+rotation period period agrees with this assessment, and {\bf its
+lithium is a little low}.
+
+The spatial and kinematic arguments for the youth of KOI-7368 and
+KOI-7913 are necessarily weaker because they are in an association of
+stars, CH-2, that is much more tenuous.  For KOI-7913 this is not of
+particular concern, since it happens to be a binary.  Both components
+have isochronal ages consistent with $40$\,Myr, but the measurement
+for the secondary is more constraining since it is less massive.  Both
+components also show H$\alpha$ in emission, which for the $\approx$K6V
+primary is a good indicator that the star is $\lesssim$100$\,$Myr old.
+For KOI-7368, which is slightly more massive, the Li 6708\AA\ EW
+measurement and stellar rotation period are more constraining, and
+provide independent verification of the stellar youth.
 

### 2022-04-05 papers/ms.tex
update ms
mostly tables
some text

@@
 \title{
-  Three 40 to 50 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
+  A Trio of 40 to 50 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
 }
@@
   refined view of open cluster dispersal.
-  Here we present an analysis of a group of $\approx$40 million
-  year old stars spanning Cepheus ($l=100^\circ$) to Hercules
+  Here we present an analysis of a group 
+  stars spanning Cepheus ($l=100^\circ$) to Hercules
   ($l=40^\circ$), hereafter the Cep-Her complex.
-  The group includes four known Kepler Objects of Interest:
-  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$),
+  The group includes four Kepler Objects of Interest:
   Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
-  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), and
-  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$).
-  Kepler-1627 is a Neptune-sized planet in a sub-component of the
-  complex called the $\delta$\ Lyr\ cluster
+  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), 
+  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$), and
+  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
+  The latter Neptune-sized planet is in a sub-component of the
+  Cep-Her complex called the $\delta$\ Lyr\ cluster
   \citep{bouma_kep1627_2022}.
-  Here we focus on the latter three systems, which are in other
-  sub-components of the complex (RSG-5 and CH-2).
+  Here we focus on the former three systems, which are in other
+  sub-components of the complex.
   Based on kinematic evidence from Gaia, stellar rotation periods from
@@
   million years old.
-  More specifically, we find that Kepler-1643 (in RSG-5) is
-  $46^{+9}_{-7}$\,Myr old, while KOI-7368 and KOI-7913 (in CH-2) are
-  $36^{+10}_{-8}$\,Myr old.
-  Based on the transit shapes and high resolution imaging, they are
-  all most likely planets, with false positive probabilities of
+  More specifically, we find that Kepler-1643 is in a dense
+  sub-cluster of the complex called RSG-5, and is $46^{+9}_{-7}$\,Myr old.
+  KOI-7368 and KOI-7913 are in a diffuse region that we call
+  CH-2, and are $36^{+10}_{-8}$\,Myr old.
+  Based on the transit shapes and high resolution imaging, all three are
+  most likely planets, with false positive probabilities of
   $6\times10^{-9}$, $4\times10^{-3}$, and $1\times10^{-4}$ for
@@
 to also be part of a much larger group of similarly aged stars.  This
-group, which is at a distance of $\sim$330 parsecs from the Sun,
+group, which is at a distance of $\sim$330\,pc from the Sun,
 appears to span Cepheus to Hercules (galactic longitudes, $l$, between
@@
   points are stars for which group membership is more secure than for
-  gray points (see Section~\ref{subsec:members}).
+  gray points.
   Kepler-1627 is in the outskirts of the $\delta$ Lyr cluster
@@
   {\it Bottom row}: Galactic tangential velocities (left) and
-  longitudinal tangential velocity versus galactic latitude (right).
+  galactic longitudinal velocity versus galactic longitude (right).
   The gray band in the lower-right shows the $\pm$1-$\sigma$
@@
   of rest.  There is a strong spatial and kinematic overlap between
-  Kepler-1643 and RSG-5 ($\{ l, b\} \approx \{ 82^\circ, 6^\circ\}$;
-  $\{ X, Y, Z\} \approx \{60, 330, 50\} $ pc).  The local population
+  Kepler-1643 and RSG-5 (magenta).  The local population
   of candidate young stars around KOI-7368 and KOI-7913 is more
-  diffuse -- we call this region ``CH-2''.
-  The {\bf interactive figure} enables a few different cuts to be
-  shown.
+  diffuse -- we call this region ``CH-2'' (lime-green).
+  The selection method for the black, gray, purple, and green points
+  is described in Section~\ref{subsec:members}.
+  %The {\bf interactive figure} enables a few different cuts to be
+  %shown.
 	\label{fig:XYZvtang}
@@
 
-Our focus is on a region of the Galaxy approximately 200 to 500
-parsecs from the Sun, above the galactic plane, and spanning galactic
+Our focus is on a region of the Galaxy approximately 200 to 500\,pc
+from the Sun, above the galactic plane, and spanning galactic
 longitudes of roughly $40^\circ$ to $100^\circ$ degrees.  Two rich
@@
 
-The second step is to perform in an unsupervised HDBScan
+The second step is to perform an unsupervised HDBScan
 clustering on the photometrically selected population
@@
 parameters to HDBScan, we set the minimum $\epsilon$ threshold past
-which clusters cannot be fragmented as $25$ parsecs in physical space,
-and $c$ \kms\ in velocity.  The minimum cluster size $N$ is set to 10,
+which clusters cannot be fragmented as $25$\,pc in physical space,
+and $c$\,\kms\ in velocity.  The minimum cluster size $N$ is set to 10,
 as is $k$, the parameter used to define the ``core distance'' density
@@
 both spatially and kinematically.  KOI-7368 and KOI-7913 are in a 
-diffuse region $\approx$40 parsecs above RSG-5 in $Z$ and $\approx$100
-parsecs closer to the Sun in $Y$.  In tangential galactic velocity
-space, there may be some kinematic overlap between the region the
+diffuse region $\approx$40\,pc above RSG-5 in $Z$ and $\approx$100\,pc
+closer to the Sun in $Y$.  In tangential galactic velocity space,
+there may be some kinematic overlap between the region the
 latter two KOIs are in, and the main RSG-5 group.
@@
     KOI-7913 (X), Kepler-1643 (diamond).  The large sizes of the
-		youngest transiting planets could be explained by their primordial
-		atmospheres not yet having evaporated; direct measurements of the
-		atmospheric outflows or planetary masses would help to confirm
-		this expectation.  Selection effects may also be important.
-		Parameters are from the NASA Exoplanet Archive (2022 Feb 27).
+    youngest transiting planets could be tied to a selection effect.
+    Kepler-1643, KOI-7368, and KOI-7913 have normal mini-Neptune sizes
+    of 2 to 3\,$R_\oplus$, which is a novelty given their ages.
+    Parameters are from the NASA Exoplanet Archive (2022 Apr 5).
 		\label{fig:rp_period_age}
@@
 
-Perhaps the most significant point about
-Kepler-1643~b, KOI-7368.01, and KOI-7913 is that they have normal
-sizes.
-{\bf TODO for LGB: write more about this}
+In the context of the existing exoplanet population, the most significant
+novelty about Kepler-1643~b, KOI-7368.01, and KOI-7913 is that they have normal
+mini-Neptune sizes.
+More specifically, the intrinsic distribution of planet sizes measured
+by Kepler peaks at $\approx 2.4\,R_\oplus$, with a valley at
+1.8\,$R_\oplus$, and falling off past $\approx$3.5\,$R_\oplus$
+\citep[][Figure~7]{Fulton_et_al_2017}.
+
+Figure~\ref{fig:rp_period_age} shows the sizes, orbital periods, and ages of
+known transiting planets, emphasizing those with precise ages.
+Considering planets younger than $10^8$ years, the smallest previously
+known were AU~Mic~c ($\approx3.0\pm0.2\,R_\oplus$, averaging the
+results of \citealt{martioli_aumicbc_2021} and
+\citealt{gilbert_flares_2022}), Kepler-1627~Ab ($3.8\pm0.2\,R_\oplus$;
+\citealt{bouma_kep1627_2022}), and AU~Mic~d ($4.2\pm0.2\,R_\oplus$;
+\citealt{plavchan_planet_2020}).
+The other transiting planets younger than $10^8$ years are much
+larger, with sizes between 4 and 10$\,R_\oplus$
+\citep{Mann_K2_33b_2016,David_et_al_2017,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020}.
 
@@
 We have argued that Kepler-1643~b, KOI-7368.01, and KOI-7913 {\it i)}
-are all $\approx$40 million years old, and {\it ii)} that each signal
+are 40 to 50 million years old, and {\it ii)} that each signal
 is most likely planetary.  The evidence for the planetary
@@
 
-Each system has multiple indicators of youth that support their
+Each system has multiple indicators of youth that support the
 reported ages.  For Kepler-1643, perhaps the strongest youth indicator
 is its physical and kinematic association with RSG-5.  Based on the
-color-absolute magnitude diagram, we seem to be able to select members
+color-absolute magnitude diagram, we are able to select members
 of this cluster with a false positive rate of $\ll 5\%$
-(Figure~\ref{fig:age}), and Kepler-1643 is one such member.  Its
+(Figure~\ref{fig:age}); Kepler-1643 is one such member.  The stellar
 rotation period period agrees with this assessment, and {\bf its
@@
 
-The spatial and kinematic arguments for the youth of KOI-7368 and
-KOI-7913 are necessarily weaker because they are in an association of
-stars, CH-2, that is much more tenuous.  For KOI-7913 this is not of
-particular concern, since it happens to be a binary.  Both components
-have isochronal ages consistent with $40$\,Myr, but the measurement
-for the secondary is more constraining since it is less massive.  Both
-components also show H$\alpha$ in emission, which for the $\approx$K6V
-primary is a good indicator that the star is $\lesssim$100$\,$Myr old.
-For KOI-7368, which is slightly more massive, the Li 6708\AA\ EW
-measurement and stellar rotation period are more constraining, and
-provide independent verification of the stellar youth.
+The spatio-kinematic argument for the youth of KOI-7368 and
+KOI-7913 are weaker because they are in an association of stars, CH-2,
+that is more diffuse.  For KOI-7913, stronger indicators of its age
+perhaps come from its binary.  Both stellar components in KOI-7913
+have isochronal ages consistent with $40$\,Myr, but the isochronal
+measurement for the secondary is more constraining since it is less
+massive.  Both components also show H$\alpha$ in emission, which for
+the $\approx$K6V primary is a good indicator that the star is
+$\lesssim$100$\,$Myr old.  For KOI-7368, which is slightly more
+massive, the Li 6708\AA\ EW measurement and stellar rotation period
+are more constraining, and provide independent verification of the
+star's youth.
 

### 2022-04-05 papers/collect.sh

@@ -24,4 +24,4 @@ cp ~/Dropbox/proj/rudolf/results/run_simpletransit_KOI_7368/*singlepanel*pdf f3c
 cp ~/Dropbox/proj/rudolf/results/run_simpletransit_Kepler_1627/*singlepanel*pdf f3d.pdf
 
 # Rp, Period, Age
-cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20220227_colorbyage_showCepHer.pdf f4.pdf
+cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20220405_colorbyage_showCepHer.pdf f4.pdf

### 2022-04-05 papers/ms.tex
update ms

@@
 %\submitjournal{AAS Journals}
-\shorttitle{Kepler Mini-Neptunes in Cep-Her}
+\shorttitle{Mini-Neptune Trio in Cep-Her}
 
@@
   refined view of open cluster dispersal.
-  Here we present an analysis of a group 
-  stars spanning Cepheus ($l=100^\circ$) to Hercules
-  ($l=40^\circ$), hereafter the Cep-Her complex.
+  Here we present an analysis of a group stars spanning Cepheus
+  ($l=100^\circ$) to Hercules ($l=40^\circ$), hereafter the Cep-Her
+  complex.
   The group includes four Kepler Objects of Interest:
@@
   Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
-  The latter Neptune-sized planet is in a sub-component of the
-  Cep-Her complex called the $\delta$\ Lyr\ cluster
-  \citep{bouma_kep1627_2022}.
+  The latter Neptune-sized planet has previously been noted to be in
+  the $\delta$\ Lyr\ cluster \citep{bouma_kep1627_2022}, which we
+  argue here is a sub-component of the Cep-Her complex.
   Here we focus on the former three systems, which are in other
@@
   is described in Section~\ref{subsec:members}.
-  %The {\bf interactive figure} enables a few different cuts to be
-  %shown.
 	\label{fig:XYZvtang}
@@
 
-In the context of the existing exoplanet population, the most significant
-novelty about Kepler-1643~b, KOI-7368.01, and KOI-7913 is that they have normal
-mini-Neptune sizes.
-More specifically, the intrinsic distribution of planet sizes measured
-by Kepler peaks at $\approx 2.4\,R_\oplus$, with a valley at
-1.8\,$R_\oplus$, and falling off past $\approx$3.5\,$R_\oplus$
-\citep[][Figure~7]{Fulton_et_al_2017}.
-
+Current thinking about the origins and evolution of lose-in
+mini-Neptunes has recently been reviewed by \citet{bean_nature_2021}.
+In the context of the existing exoplanet population, the most
+significant novelty about Kepler-1643~b, KOI-7368.01, and KOI-7913 is
+that they have normal sizes.
+More specifically, the size distribution of mini-Neptunes spans
+$\approx$1.8\,$R_\oplus$ to $\approx$3.6\,$R_\oplus$, with a peak at
+$\approx 2.4\,R_\oplus$ \citep[][Figure~7]{Fulton_et_al_2017}.
 Figure~\ref{fig:rp_period_age} shows the sizes, orbital periods, and ages of
@@
 \citep{Mann_K2_33b_2016,David_et_al_2017,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020}.
+
+%TODO: I think we might need James Owen to contribute something
+%similar to the TOI-1227 models?
+{\bf Run some evolutionary models, make some comments about the core
+mass, KH timescale, etc}.
 

### 2022-04-05 papers/ms.tex
update ms
Co-authored-by: Jason Curtis <jasonleecurtis@gmail.com>

@@
   Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
-  The latter Neptune-sized planet has previously been noted to be in
-  the $\delta$\ Lyr\ cluster \citep{bouma_kep1627_2022}, which we
-  argue here is a sub-component of the Cep-Her complex.
+  The latter Neptune-sized planet is in a sub-component of the
+  Cep-Her complex called the $\delta$\ Lyr\ cluster
+  \citep{bouma_kep1627_2022}.
   Here we focus on the former three systems, which are in other
@@
 achieved using surveys such as TESS \citep{ricker_transiting_2015} and
-ZTF \citep{bellm_zwicky_2019}, and it leverages a storied tradition of
-rotation period measurement for benchmark open clusters \citep[see
-{\it e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data
-in our case are especially useful, since they provide 3 to 5 lunar
-months of photometry for all of our candidate CH-2 and RSG-5 members.
+the Zwicky Transient Facility (ZTF, \citealt{bellm_zwicky_2019}), and
+it leverages a storied tradition of rotation period measurement for
+benchmark open clusters \citep[see {\it
+e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data in
+our case are especially useful, since they provide 3 to 5 lunar months
+of photometry for all of our candidate CH-2 and RSG-5 members.
 
 We selected stars suitable for gyrochronology by requiring $\bpmrpo
-\geq 0.5$ and $G<16$.  The latter cut corresponds to $\bpmrpo \lesssim
-2.6$, at the relevant distances.  These cuts gave 19 stars in CH-2 and
-42 stars in RSG-5.  We extracted light curves from the TESS images for
-these stars using the \texttt{unpopular} package
+\geq 0.5$ to focus on FGKM stars that experience magnetic braking.
+For TESS, we also restricted our sample to $G<16$, to ensure the stars
+are bright enough to extract usable light curves from the full-frame
+images (FFIs).  The magnitude cut corresponds to $\bpmrpo < 2.6$
+($\sim$M3V) at the relevant distances.  These cuts gave 19 stars in
+CH-2 and 42 stars in RSG-5.  We extracted light curves from the TESS
+images for these stars using the \texttt{unpopular} package
 \citep{hattorio_2021_cpm}, and regressed them against systematics with
-its causal pixel model.  {\bf For the ZTF light curves, we used the
-default photometry?  Ran aperture photometry on the image cutouts?}.
-We then measured candidate rotation periods using a Lomb-Scargle
-periodogram, and visually inspected them following the methods
-discussed in \citet{curtis_rup147_2020}.
+its causal pixel model.  We measured rotation periods using
+Lomb-Scargle periodograms and visually vetted the results using an
+interactive program that allows us to switch between TESS Cycles,
+select particular sectors, flag stars with multiple periods, and
+correct half-period harmonics. For ZTF, we used the same color cut to
+focus on FGKM stars, but restricted the sample to $13 < G < 18$ to
+avoid the saturation limit on the bright end and ensure sufficient
+photometric precision at the faint end. We followed the procedure
+outlined in \citet{curtis_rup147_2020}: we downloaded $8'\times8'$
+image cutouts, ran aperture photometry for the target and neighboring
+stars identified with Gaia, and used them to define a systematics
+correction to refine the target light curves. 
+%Like for TESS, we also
+%calculated periods with Lomb-Scargle periodograms using an interactive
+%program that allows us to visually validate our measurements.
+{\bf How many stars with periods from TESS? ZTF? Overlap?
+Consistency? }
 

### 2022-04-06 papers/ms.tex
update Li6708 analysis
Mostly wrangling the literature, and reminding myself that the
uncertainties on the M-dwarfs from Randich+18 make no sense, as do their
lack of upper limits.  (Though the earlier Randich+01 work seems fine).
Adding the Tuc-Hor Kraus+14 sample is a good sanity check.
See simultaneous commits at `timmy`.

@@
 quoted value does not correct for the Fe\,\textsc{I} blend at
-6707.44\AA.
+6707.44\AA, because HIRES has sufficient resolution to resolve the two
+features.
 
@@
 
+% Randich+01: Li starts to become upper limits around K8V, V-I=1.6.
+% Around K6V, V-I=1.4, ~100mA seen, but few stars in this region!
+
 Regarding indicators of youth, the primary has a spectral type
@@
 expected to be in emission for for all M-dwarfs younger than
-$\approx150$\,Myr \citep{kiman_calibration_2021}.  At $\approx$M0V, if
-the secondary were in older than $\approx 500\,{\rm Myr}$, it would be
-more common for its H$\alpha$ component to be observed in absorption
-\citep[][Figure~4]{kiman_calibration_2021}.  H$\alpha$ emission in
-late K-dwarfs is rarer still. Based on the data from K.~Rocio et al's
-analysis\footnote{Available at
+$\approx150$\,Myr \citep[{\it e.g.},][]{kiman_calibration_2021}.  For
+the $\approx$K6V primary, H$\alpha$ emission is also expected based on
+previous spectroscopic studies of the Pleiades, IC~2602, and IC~2391
+\citep[{\it e.g.},][their Figure~8]{stauffer_rotational_1997}.  Based
+on the data from K.~Rocio et al's analysis\footnote{Available at
 \url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K6V

### 2022-04-06 drivers/plot_lithium.py

@@ -6,4 +6,7 @@ PLOTDIR = os.path.join(RESULTSDIR, 'lithium')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
-rp.plot_lithium(PLOTDIR)
+rp.plot_lithium(PLOTDIR, reference='Pleiades')
+rp.plot_lithium(PLOTDIR, reference='Randich01_TucHorK14M08')
+rp.plot_lithium(PLOTDIR, reference='Randich01')
+rp.plot_lithium(PLOTDIR, reference='Randich18')

### 2022-04-06 rudolf/starinfo.py

@@ -23,8 +23,8 @@ starinfodict = {
         # HIRES / SM-Emp Median
         'Teff': 4038,
         'Teff_err': 70,
-        # HIRES / KOI-7913B_ij438.81_Li_EW_shift-0.05_results.csv
-        'Li_EW':34.613, 'Li_EW_perr': 12.683, 'Li_EW_merr': 13.974,
+        # HIRES / KOI-7913B_ij438.81_Li_EW_shift-0.20_deltawav5.0.png
+        'Li_EW':41.6, 'Li_EW_perr': 11.5, 'Li_EW_merr': 18.6,
         # rotation_period_windowslider
         'Prot': 2.641824,
         'Prot_err': 0.066772,
@@ -33,9 +33,9 @@ starinfodict = {
         # HIRES / SM-Emp Median
         'Teff': 4916,
         'Teff_err': 110,
-        # HIRES / Kepler1643_ij384.99_Li_EW_shift-0.30.png
-        # or Kepler1643_ij438.82_Li_EW_shift-0.05.png
-        'Li_EW': 126.4, 'Li_EW_perr': 7.9, 'Li_EW_merr': 4.2,
+        # Kepler1643_ij384.99_Li_EW_shift-0.30_deltawav5.0.png
+        # Kepler1643_ij438.82_Li_EW_shift0.00_deltawav5.0.png
+        'Li_EW': 129.7, 'Li_EW_perr': 5.9, 'Li_EW_merr': 4.5,
         # rotation_period_windowslider
         'Prot': 5.106414,
         'Prot_err': 0.043881,

### 2022-04-07 papers/ms.tex
add lithium appendix

@@
 
+\section{Spectroscopic Youth Indicators}
+\label{app:spectra}
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f5a.pdf}
+			\includegraphics[width=0.49\textwidth]{f5b.pdf}
+		}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{
+    {\bf Lithium 6708\AA\ equivalent widths for Kepler-1643, KOI-7368,
+    KOI-7913 AB, and Kepler-1627 compared to those of young open
+    clusters and field stars. }
+    The statistical uncertainties on the newly measured equivalent
+    widths are shown, or else are smaller than the markers.
+    The field stars are stars with Kepler planets from \citet{berger_identifying_2018}.
+    The ``40-50 Myr'' reference stars (left) are from
+		IC\,2602 \citep{randich_gaiaeso_2018} and Tuc-Hor
+		\citep{kraus_stellar_2014}.
+    The ``112 Myr'' stars (right) are from the Pleiades
+    \citep{soderblom_evolution_1993,jones_evolution_1996,bouvier_pleiades_lirot_2018},
+    where we adopt the LDB age from \citet{dahm_2015}.
+    \label{fig:lithium}
+	}
+\end{figure*}
+
+\subsection{Lithium}
+
+Figure~\ref{fig:lithium} compares the measured lithium equivalent
+widths of the objects of interest against a few reference populations.
+We selected the reference populations from the literature only for
+cases in which upper limits were explicitly reported.  KOI-7368 and
+KOI-7913~A have secure lithium detections, and the KOI-7913~B
+detection is marginal (${\rm EW}_{\rm Li} = 42^{+12}_{-19}$).  For all
+three stars, as well as for Kepler-1627~A, the observed lithium
+equivalent width is consistent with the stellar effective temperatures
+and $\approx40$\,Myr ages.
+
+Kepler-1643, in RSG-5, is conspicuously below the 40-50$\,$Myr sequence
+in the left-panel of Figure~\ref{fig:lithium}, while still being above
+the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$).
+The right panel shows the comparison against the Pleiades, where
+Kepler-1643 is more consistent with the observed dispersion in lithium.
+
+One explanation for this could be that Kepler-1643 is a field
+interloper; another could be that RSG-5 is older than 50\,Myr.  We do
+not favor the former, given that {\it i)} the star was selected based
+on its spatio-kinematic proximity to other RSG-5 members, nearly all
+of which appear consistent with being $\approx$50\,Myr old
+(Figure~\ref{fig:age} upper left), and {\it ii)} the same RSG-5
+members display a gyrochronal sequence consistent with being younger
+than the Pleaides (Figure~\ref{fig:age} lower left).
+We do not have an explanation, but in the following paragraph
+speculate on one possible line of thinking that might help resolve the
+discrepancy.
+
+Analyzing Figure~\ref{fig:lithium}, it is remarkable that in just 50
+million years, stars between $4500$\,K and $5200$\,K go from having
+such a tight lithium sequence to one with a dispersion
+$\approx10\times$ greater.  The existence of such a dispersion in
+Pleiades-age K-dwarfs has been known for longer than some of the
+present authors have been alive; it has also been known that the stars
+with the most lithium are also the most rapidly rotating
+\citep{butler_pleiades_1987,soderblom_evolution_1993}.  More recent
+analyses of this correlation have been reviewed by
+\citet{bouvier_lithium-rotation_2020}.  The conclusion of that work
+was that the origin of the rotation-lithium correlation likely lies
+within pre-main-sequence stellar physics.  If so, one would expect the
+IC\,2062 and Tuc-Hor K-dwarfs to show a larger intrinsic lithium
+dispersion, which they do not.  It seems therefore that the exact
+onset time of the K-dwarf lithium dispersion might be worth
+clarifying.  \citet{messina_rotation-lithium_2016} for instance do not
+obviously see it in the $\beta$~Pic moving group ($25\pm3$\,Myr).
+Given that RSG-5 seems somewhat isochronally older than the
+$\delta$~Lyr cluster (and therefore IC\,2602
+\citealt{bouma_kep1627_2022}), it might be a worthy cluster for such
+an analysis.
+
 \section{Table of Transit Fit Parameters}

### 2022-04-07 papers/ms.tex
update ms
we have a good title

@@
 %\submitjournal{AAS Journals}
-\shorttitle{Mini-Neptune Trio in Cep-Her}
+\shorttitle{Trio of Mini-Neptunes in Cep-Her}
 
@@
 \title{
-  A Trio of 40 to 50 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia
+  %Three Mini-Neptunes in a 40 Million Year Old Behemoth
+  Kepler and the Behemoth: Three Mini-Neptunes in a 40 Million Year Old Association
+  %Three Mini-Neptunes, 40 to 50 Million Years Old 
+  %A Trio of 40 to 50 Million Year Old Mini-Neptunes
 }
@@
 \begin{abstract}
-  Stellar positions and velocities from Gaia are yielding a new and
+  Stellar positions and velocities from Gaia are yielding a 
   refined view of open cluster dispersal.
@@
   Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
-  The latter Neptune-sized planet is in a sub-component of the
+  The latter Neptune-sized planet is in a component of the
   Cep-Her complex called the $\delta$\ Lyr\ cluster
@@
   Here we focus on the former three systems, which are in other
-  sub-components of the complex.
+  components of the complex.
   Based on kinematic evidence from Gaia, stellar rotation periods from
   TESS, and spectroscopy, these three systems are also $\approx$40
-  million years old.
-  More specifically, we find that Kepler-1643 is in a dense
-  sub-cluster of the complex called RSG-5, and is $46^{+9}_{-7}$\,Myr old.
-  KOI-7368 and KOI-7913 are in a diffuse region that we call
-  CH-2, and are $36^{+10}_{-8}$\,Myr old.
-  Based on the transit shapes and high resolution imaging, all three are
-  most likely planets, with false positive probabilities of
-  $6\times10^{-9}$, $4\times10^{-3}$, and $1\times10^{-4}$ for
+  million years (Myr) old.
+  More specifically, we find that Kepler-1643 is $46^{+9}_{-7}$\,Myr
+  old, based on its membership in a dense
+  sub-cluster of the complex called RSG-5.
+  KOI-7368 and KOI-7913 are $36^{+10}_{-8}$\,Myr old, and are in a
+  diffuse region that we call CH-2.
+  Based on the transit shapes and high resolution imaging, all three
+  objects are most likely planets, with false positive probabilities
+  of $6\times10^{-9}$, $4\times10^{-3}$, and $1\times10^{-4}$ for
   Kepler-1643, KOI-7368, and KOI-7913 respectively.
-  These planets therefore empirically
-  demonstrate that mini-Neptunes with sizes of $\approx$2 Earth
-  radii exist at ages of $\approx$40 million years.
+  These planets empirically demonstrate that mini-Neptunes with sizes
+  of $\approx$2 Earth radii exist at ages of $\approx$40 million
+  years.
 \end{abstract}
@@
 which the planetary evidence comes from transits, and the age evidence
-is based on either cluster membership \citep{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
+is based on either cluster membership
+\citep{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
 or else on correlates of youth such as stellar rotation, photospheric
-lithium abundances, x-ray activity, or emission line strength
+lithium content, x-ray activity, or emission line strength
 \citep{zhou_2021_tois,hedges_toi-2076_2021}.
@@
 In this work, we leverage recent analyses of the Gaia data, which have
-greatly expanded our knowledge of stellar group memberships
-\citep[{\it e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.
-To date these analyses have mostly clustered on the stellar positions and
-on-sky velocities measured by Gaia, with varying degrees of filtering
+greatly expanded our knowledge of stellar groups \citep[{\it
+e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To date these
+analyses have mostly clustered on the stellar positions and on-sky
+velocities measured by Gaia, with varying methodologies fo filtering
 and supervision.  One important result is the identification of
 diffuse streams and tidal tails comparable in stellar mass to the
-previously known cores of nearby open clusters
+previously known cores of many nearby open clusters
 \citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  These
@@
 effects and internal dynamics.  As an extreme example, in the Hyades
-the velocities of stars in the tidal tails are thought to
-span up to $\pm 40\mkms$ relative to the cluster center
-\citep{jerabkova_800_2021}.  The stars in such diffuse regions can be
-verified to be the same age as the core members ({\it i.e.}, coeval)
-through analyses of color-absolute magnitude diagrams, stellar
-rotation periods \citep{curtis_tess_2019,bouma_2021_ngc2516}, and
-chemical abundances \citep{hawkins_2020}.  While there are many
-implications for our understanding of star formation and cluster
-evolution \citep{dinnbier_tidal_2020}, a more immediate
-consequence is that we now know the ages of many more stars, including
-previously known planet hosts.
+the velocities of stars in the tidal tails span up to $\pm 40\mkms$
+relative to the cluster center \citep{jerabkova_800_2021}.  The stars
+in such diffuse regions can be verified to be the same age as the core
+members ({\it i.e.}, coeval) through analyses of color-absolute
+magnitude diagrams, stellar rotation periods
+\citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
+\citep{hawkins_2020}.  While there are many implications for our
+understanding of star formation and cluster evolution
+\citep{dinnbier_tidal_2020}, a separate consequence is that we now
+know the ages of many more stars, including previously known planet
+hosts.
 
@@
 currently known transiting exoplanets, and it was conducted before
-Gaia.  It therefore seems sensible to revisit the Kepler field, given
-our new knowlege of the stellar ages.
-
-In this work, we expand on our previous study of a $38^{+7}_{-6}$
-million year old Neptune-sized planet in the Kepler field (Kepler-1627
-Ab; \citealt{bouma_kep1627_2022}).  The age of this planet was derived
+Gaia.  It is therefore sensible to revisit the Kepler field, given our
+new knowlege of the stellar ages.
+
+In this work, we expand on our previous study of a $38^{+7}_{-6}$ Myr
+old Neptune-sized planet in the Kepler field (Kepler-1627~Ab;
+\citealt{bouma_kep1627_2022}).  The age of this planet was derived
 based on its host star's membership in the $\delta$\ Lyr\ cluster.
 Our analysis of the cluster focused on the immediate spatial and
-kinematic vicinity of Kepler-1627~A in order to confirm the age of the
-planet.  However it became clear that the $\delta$\ Lyr\ cluster seems
-to also be part of a much larger group of similarly aged stars.  This
-group, which is at a distance of $\sim$330\,pc from the Sun,
-appears to span Cepheus to Hercules (galactic longitudes, $l$, between
+kinematic vicinity of Kepler-1627~A in order to secure the planet's
+age.  However it became clear that the $\delta$\ Lyr\ cluster seems to
+also be part of a much larger group of similarly aged stars.  This
+group, which is at a distance of $\sim$330\,pc from the Sun, appears
+to span Cepheus to Hercules (galactic longitudes, $l$, between
 40$^\circ$ and 100$^\circ$), at galactic latitudes roughly between
@@
 system using a combination of the Kepler photometry and
-high-resolution imaging (Section~\ref{sec:planets}).  
-We conclude with a few caveats, 
-and discuss implications for the size-evolution of close-in
-mini-Neptunes (Section~\ref{sec:disc_conc}).
+high-resolution imaging (Section~\ref{sec:planets}).  We conclude with
+a few caveats, and discuss implications for the size-evolution of
+close-in mini-Neptunes (Section~\ref{sec:disc_conc}).
 
@@
 Each of these clusters was known before Gaia.  They have reported ages
-between $\approx$30 and $\approx$60 million years.  Early empirical
+between 30 and 60\,Myr.  Early empirical
 evidence that these two clusters could be part of a large and more
@@
 part of an unsupervised clustering analysis of the Gaia DR2 positions
-and tangential velocities with a subsequent manual ``stitching'' step,
-and generally supports the idea that there is an overdensity of
-$\approx$30 and $\approx$60 million year old stars in this region of
+and tangential velocities with a subsequent manual ``stitching'' step.
+Their results support the idea that there is an overdensity of
+30 to 60\,Myr old stars in this region of
 the Galaxy.  \citet{Kerr2021}, in a volume-limited analysis of the
@@
 ``Cepheus-Cygnus'', ``Lyra'', and ``Cerberus''.  \citet{Kerr2021}
-reported ages for each of these subgroups between 30 and 35 million
-years.
+reported ages for each of these subgroups between 30 and 35 Myr.
 
@@
 sub-populations identified by \citet{Kerr2021} share a common origin
-has yet to be fully substantiated, and will be the subject of the upcoming
+has yet to be fully substantiated, and is the subject of the upcoming
 study by R.~Kerr and collaborators.  Our primary interest in the
@@
 estimates \citep{lallement_gaia-2mass_2019}.  This step yielded \npms\
-stars with high-quality photometric and astrometry, which are either
-pre-main-sequence K and M dwarfs due to their long contraction
+stars with high-quality photometry and astrometry.  These stars are
+either pre-main-sequence K and M dwarfs due to their long contraction
 timescales, or massive stars near the zero-age main sequence due to
@@
 \citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The
-parameters we use in this clustering analysis are $\{ X, Y, Z, c v_b,
+parameters we use in the clustering are $\{ X, Y, Z, c v_b,
 c v_{l^*} \} $, where $c$ is the size-velocity corrective factor,
@@
 This third step yields stars that are spatially and kinematically
-close to the photometrically-young stars, but which cannot be
-identified as young based on their positions in color versus absolute
-magnitude.
+close to the photometrically young stars, but which cannot be
+identified as young based on their positions in the color-absolute
+magnitude diagram.
 
@@
 	$G_{\rm BP}/\sigma_{G_{BP}}>20$.}
-The previously known $\delta$~Lyr\ cluster ($l,b=68^\circ,15^\circ$;
-$v_{l'}, v_b=-4.5 \mkms,-4 \mkms$) is visible, as is RSG-5
-($l,b=83^\circ,6^\circ$; $v_{l'}, v_b=5.5 \mkms,-3.5 \mkms$).
+The previously known $\delta$~Lyr\ cluster is visible at
+$(l,b)=(68^\circ,15^\circ$) and
+$(v_{l'}, v_b)=(-4.5 \mkms,-4 \mkms)$.  RSG-5 is visible at
+$(l,b)=(83^\circ,6^\circ)$,  $(v_{l'}, v_b)=(5.5 \mkms,-3.5 \mkms)$.
 Most of the other subclusters, including in Cep-Cyg
@@
 %
-The fourth and final step was to cross-match our candidate Cep-Her
+Our fourth and final step was to cross-match the candidate Cep-Her
 member list against all known Kepler Objects of Interest.  We used the
@@
 with weights exceeding 0.02, this yielded
- yielded 25 matches, of which 11
-were known false positives, 6 were designated
-``confirmed'', and 8 were designated ``candidates''.  
+ yielded 11 were known false positives, 6 ``confirmed'' planets, and 8
+ ``candidate'' planets.  
 Inspection of the Kepler data validation summaries and Robovetter
-classifications for these objects then showed whether they were
+classifications for these objects showed whether they were
 potentially consistent with being {\it i)} planets, and {\it ii)}
@@
 e.g.},][Figure~9]{rebull_rotation_2020}.  Four objects remained after
-this inspection: Kepler-1627, Kepler-1643, KOI-7368, and KOI-7913.
+this inspection: Kepler-1627, Kepler-1643, KOI-7368, and
+KOI-7913.\footnote{These figures
+in fact give an estimate of the size of Cep-Her: if our 4/14
+possible matches represent the true positive rate, then
+multiplying by the number of stars above our weight threshold
+(12{,}436) yields $\approx$3{,}500 stars in the complex.}
 
@@
 closer to the Sun in $Y$.  In tangential galactic velocity space,
-there may be some kinematic overlap between the region the
-latter two KOIs are in, and the main RSG-5 group.
+there is some kinematic overlap between the region the
+latter two KOIs are in and the main RSG-5 group.
 
@@
 and we call this latter set of stars ``CH-2''.  These cuts yield 141
-candidate RSG-5 members, and 37 candidate CH-2 members.  An important
-consideration, especially for CH-2, is the
+candidate RSG-5 members, and 37 candidate CH-2 members.
+Tables of these stars are enumerated in Appendix~\ref{app:members}.
+An important consideration, especially for CH-2, is the
 contamination rate by field stars.  We assess this in the following
@@
     {\it Top row}: 
-    Color-absolute magnitude diagram of candidate Cep-Her members, in
-    addition to candidate members of the $\delta$~Lyr~cluster
-    ($\approx38$\,Myr; \citealt{bouma_kep1627_2022}) and and the Gaia
+    Color-absolute magnitude diagram of candidate Cep-Her members, plotted
+    over candidate members of the $\delta$~Lyr~cluster
+    ($\approx38$\,Myr; \citealt{bouma_kep1627_2022}) and the Gaia
     EDR3 Catalog of Nearby Stars (gray background).  The left and
@@
     range of colors is truncated to emphasize the pre-main-sequence.
+    Stars that fall far below the cluster sequences are field
+    interlopers.
     {\it Bottom row}:
@@
 diagrams, we imposed the data filtering criteria from
-\citet[][Appendix B]{GaiaCollaboration2018}, which are designed to
+\citet[][Appendix B]{GaiaCollaboration2018}, which 
 include binaries while omitting instrumental artifacts from for
@@
 
-The CAMDs show that for RSG-5, our membership selection gives stars
-that photometrically seem to almost all be on a tight
-pre-main-sequence locus.  This implies a false positive rate of 
-a few percent, at most.  By comparison, our control sample (the $\delta$~Lyr
-candidates) has a false positive rate of $\approx$12\%, based on the
-number of stars that photometrically appear to be more consistent with
-the field population than the bulk cluster population.  For CH-2, our
-membership selection gives 27 objects in the color range displayed,
-and 23 of them appear to be consistent with being on the
-pre-main-sequence.  This implies a false positive rate of
-$\approx$15\%.
-
-In addition, Figure~\ref{fig:age} shows that the candidate RSG-5 and
-CH-2 members overlap with the $\delta$~Lyr cluster, and are therefore
+The CAMDs show that for RSG-5, almost all of our candidate members are
+on a tight pre-main-sequence locus.  This implies a false positive
+rate of a few percent, at most.  By comparison, our control sample
+(the $\delta$~Lyr candidates) has a false positive rate of
+$\approx$12\%, based on the number of stars that photometrically
+overlap with the field population, rather than the bulk cluster
+population.  For CH-2, our membership selection gives 27 objects in
+the color range displayed, and 23 of them appear to be consistent with
+being on the pre-main-sequence.  This implies a false positive rate in
+CH-2 of $\approx$15\%.
+
+Figure~\ref{fig:age} also shows that most RSG-5 and
+CH-2 members overlap with the $\delta$~Lyr cluster, and that the groups are therefore
 roughly the same age.  To quantify this, we use the empirical method
-introduced by \citet[][see their Section~6.3]{gagne_mutau_2020}.  This
-idea of the method is to fit the pre-main-sequence loci of a set of
+introduced by \citet[][their Section~6.3]{gagne_mutau_2020}.  This
+idea is to fit the pre-main-sequence loci of a set of
 reference clusters, and to then model the locus of the target cluster
 as a linear combination of these reference cluster loci.  For our
-reference clusters, we used UCL, IC\,2602, and the Pleiades, from the
+reference clusters, we used UCL, IC\,2602, and the Pleiades, with the
 memberships reported by \citet{Damiani2019} and
@@
 and 112\,Myr for the Pleiades \citep{dahm_2015}.  These assumptions
-and the consequent processing steps taken to exclude field stars as
-well as photometric and astrometric binaries were identical to those
-described in \citet{bouma_kep1627_2022}.  The mean and standard
-deviation of the resulting age posterior are $46^{+9}_{-7}$\,Myr for
+and the subsequent processing steps taken to exclude field stars and binaries
+were identical to those
+described in \citet{bouma_kep1627_2022}.  The mean and 
+uncertainty of the resulting age posterior are $46^{+9}_{-7}$\,Myr for
 RSG-5, and $36^{+10}_{-8}$\,Myr for CH-2.  For comparison, the this
 procedure yields an age for the $\delta$~Lyr cluster of
-$38^{+6}_{-5}$\,Myr.  The slightly older isochronal age of RSG-5 is
-expected given that its locus is slightly bluer and less luminous in
-the upper left panel of Figure~\ref{fig:age} relative to the
-$\delta$~Lyr cluster.
+$38^{+6}_{-5}$\,Myr.  
+The slightly older isochronal age of RSG-5 is
+consistent with its location
+relative to the $\delta$~Lyr cluster in
+the upper left panel of Figure~\ref{fig:age}.
 
@@
 achieved using surveys such as TESS \citep{ricker_transiting_2015} and
-the Zwicky Transient Facility (ZTF, \citealt{bellm_zwicky_2019}), and
-it leverages a storied tradition of rotation period measurement for
+the Zwicky Transient Facility (ZTF, \citealt{bellm_zwicky_2019});
+it leverages a storied tradition of measuring rotation periods of stars in
 benchmark open clusters \citep[see {\it
@@
 CH-2 and 42 stars in RSG-5.  We extracted light curves from the TESS
-images for these stars using the \texttt{unpopular} package
+images using the \texttt{unpopular} package
 \citep{hattorio_2021_cpm}, and regressed them against systematics with
@@
 %program that allows us to visually validate our measurements.
-{\bf How many stars with periods from TESS? ZTF? Overlap?
-Consistency? }
+%{\bf How many stars with periods from TESS? ZTF? Overlap?
+%Consistency? }
 
@@
 36/42 stars have rotation periods faster than the Pleiades (86\%).
-This numerator omits the two stars with periods $>12$\,days visible in
+This numerator omits the two stars with periods $>$$12$\,days visible in
 the lower-left panel of Figure~\ref{fig:age}.  The age interpretation
@@
 obvious.  \citet{rebull_usco_2018} for instance have found numerous
-M-dwarfs with 10-12 day rotation periods at ages of USco ($\sim
-8$\,Myr), and some may still exist at ages of LCC ($\sim 16$\,Myr;
+M-dwarfs with 10-12 day rotation periods at ages of USco ($\sim$$8$\,Myr),
+and some may still exist at ages of LCC ($\sim$$16$\,Myr;
 L.~Rebull in preparation).  Regardless, given that nearly no field
@@
 perhaps be taken as an indication for the fraction of stars for
-rotation periods might not be detectable, due to {\it e.g.}, pole-on
+which rotation periods might not be detectable, due to {\it e.g.}, pole-on
 stars having lower amplitude starspot modulation.
@@
 in the preceding numerator, are M-dwarfs with rotation periods between
-10 and 12.5 days.  The age interpretation for these M-dwarfs is, as
-just discussed, not obvious.  Regardless, the $\approx$15\% false
+10 and 12.5 days.  As just discussed, the age interpretation for these M-dwarfs
+is not obvious.  Regardless, the $\approx$15\% false
 positive rate determined from the CAMD seems consistent with our
@@
 \citealt{douglas_stephanie_t_2021_5131306}).
-They also seem consistent with the false positive rate estimates
+They also seem consistent with the estimated false positive rates
 determined from the color-absolute magnitude diagrams.
@@
 
-Some salient stellar parameters of the KOIs in Cep-Her can be gleaned
-by inspecting Figure~\ref{fig:age}.  They span spectral types of G8V
-(Kepler-1627) to K8V (KOI-7913 A).  The secondary in the KOI-7913
-system is only marginally cooler than the primary.  And since a
+Some salient stellar properties of the Kepler objects in Cep-Her can be gleaned
+from Figure~\ref{fig:age}.  They span spectral types of G8V
+(Kepler-1627) to K7V (KOI-7913 A).  The secondary in the KOI-7913
+system has spectral type $\approx$M0V.  And since a
 Solar-mass star with solar metallicity arrives at the zero-age main
-sequence at $t\approx40$ million years \citep{choi_mesa_2016}, these
+sequence at $t\approx40$ Myr \citep{choi_mesa_2016}, these
 stars are all in the late stages of their pre-main-sequence
@@
 including possible bound binaries.  The system-by-system details
-follow, and the the relevant results are also summarized in
+follow, and the relevant results are also summarized in
 Table~\ref{tab:sysparams}.
@@
 For Kepler-1643, we acquired two iodine-free spectra from Keck/HIRES
-on the nights of 2020 August 16 and 2021 October 25.  The acquisition
-and analysis followed the standard reduction techniques of the
+on the nights of 2020 Aug 16 and 2021 Oct 25.  The acquisition
+and analysis followed the usual techniques of the
 California Planet Survey \citep{howard_cps_2010}.  We derived the
@@
 $<1$-$\sigma$ agreement with those from the cluster-isochrone
-approach.  This approach also yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.
+method.  This approach also yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.
 Using the broadened synthetic templates from \texttt{SpecMatch-Synth}
@@
 the procedure described by \citet{bouma_2021_ngc2516}.  This yielded a
-strong detection: ${\rm EW}_{\rm Li} = 126^{+8}_{-4}$\,m\AA, with
+strong detection: ${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with
 values consistent at $<1$-$\sigma$ between the two epochs.   The
 quoted value does not correct for the Fe\,\textsc{I} blend at
-6707.44\AA, because HIRES has sufficient resolution to resolve the two
+6707.44\AA, because HIRES resolves the two
 features.
@@
 We acquired adaptive optics imaging of Kepler-1643 on the night of
-2019 June 28 using the NIRC2 imager on Keck-II (PI: E.~Petigura).
+2019 June 28 using the NIRC2 imager on Keck-II.
 Using the narrow camera (FOV = 10.2\arcsec), we obtained 4 images in
@@
 spectrograph (TRES; \citealt{furesz_tres_2008}) mounted at the
-Tillinghast 1.5m at the Fred Lawrence Whipple Observatory on Mt.
-Hopkins, AZ.  The stellar parameter classification pipeline for the
+Tillinghast 1.5m at the Fred Lawrence Whipple Observatory.
+The stellar parameter classification pipeline for the
 TRES spectra has been described by \citet{2021tsc2.confE.124B}.  It is
@@
 cluster-isochrone method within $1$-$\sigma$, though the effective
-temperature is more precise by a factor of three.  Auxiliary
+temperature was more precise by a factor of three.  Auxiliary
 spectroscopic parameters included the metallicity $[{\rm Fe/H}]= -0.02
@@
 We acquired adaptive optics imaging of KOI-7368 on the night of 2019
-June 12 using the NIRC2 imager on Keck-II (PI: D.~Huber).  The
+June 12, again using NIRC2.  The
 observational configuration and reduction was identical as for
@@
 %(e.g., Jenkins KSCI-19081-002 KPDH pp 129, or many other pages)
-KOI-7913 is almost certainly a binary.\footnote{Relevant identifiers for the
-primary are KIC 8873450 = Gaia DR2/EDR3 2106235301785454208.  For the
-secondary, they are KIC 8873448 = Gaia DR2/EDR3 2106235301785453824.}
+KOI-7913 is almost certainly a binary. %\footnote{Relevant identifiers for the
+%primary are KIC 8873450 = Gaia DR2/EDR3 2106235301785454208.  For the
+%secondary, they are KIC 8873448 = Gaia DR2/EDR3 2106235301785453824.}
 The north-west primary is $\approx$0.5 magnitudes brighter than the
 south-east secondary in optical passbands.  The two stars are
-separated in Gaia EDR3 by $3\farcs54$ on-sky, and have parallaxes
+separated in Gaia EDR3 by $3\farcs5$ on-sky, and have parallaxes
 consistent within $1$-$\sigma$ (with an average $\varpi=3.66 \pm
-0.01$\,mas).  This corresponds to an apparent on-sky separation of
-$959 \pm 2$ au.  Gaia EDR3 quotes
-$(\mu_{\alpha'},\mu_\delta)=(2.439\pm0.015, 1.009\pm 0.016)\,{\rm
-mas}\,{\rm yr}^{-1}$ for the primary, and
-$(\mu_{\alpha'},\mu_\delta)=(2.696\pm0.020, 0.650 \pm 0.021) \,{\rm
-mas}\,{\rm yr}^{-1}$ for the secondary.  Since two stars were resolved
+0.01$\,mas).  The apparent on-sky separation is
+$959 \pm 2$ au.  
+The Gaia EDR3 proper motions are also nearly identical.
+%$(\mu_{\alpha'},\mu_\delta)=(2.439\pm0.015, 1.009\pm 0.016)\,{\rm
+%mas}\,{\rm yr}^{-1}$ for the primary, and
+%$(\mu_{\alpha'},\mu_\delta)=(2.696\pm0.020, 0.650 \pm 0.021) \,{\rm
+%mas}\,{\rm yr}^{-1}$ for the secondary.  
+Since two stars were resolved
 in the Kepler Input Catalog and are roughly one Kepler pixel apart, an
@@
 We acquired Keck/HIRES spectra for KOI-7913 A on the night of 2021 Nov
-13, and KOI-7913 B on the night of 2021 Oct 26.  The usual
+13, and KOI-7913 B on the night of 2021 Oct 26.  The 
 \texttt{SpecMatch-Emp} \citep{yee_SM_2017} machinery yielded $T_{\rm
@@
 -0.06 \pm 0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm
-sys} = -17.8 \pm 1.1\,\mkms$.  These same parameters for the secondary
+sys} = -17.8 \pm 1.1\,\mkms$.  For the secondary, these same parameters
 were $[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm 1.0\,\mkms$,
@@
 Regarding indicators of youth, the primary has a spectral type
-$\approx$K6V, while the secondary is $\approx$M0V.  At $\approx$40
+$\approx$K7V, while the secondary is $\approx$M0V.  At $\approx$40
 Myr, lithium 6708\,\AA\ absorption is not expected for the secondary,
@@
 $\approx150$\,Myr \citep[{\it e.g.},][]{kiman_calibration_2021}.  For
-the $\approx$K6V primary, H$\alpha$ emission is also expected based on
+the primary, H$\alpha$ emission is also expected based on
 previous spectroscopic studies of the Pleiades, IC~2602, and IC~2391
@@
 on the data from K.~Rocio et al's analysis\footnote{Available at
-\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K6V
+\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K7V
 star should be emitting in H$\alpha$ at 40\,Myr, but not typically
@@
 
+Analyzing the HIRES spectra, we find
+...
 Turning to the actual data, we find {\bf X for lithium}.
@@
 We acquired adaptive optics imaging of KOI-7913 on the night of 2020
-Aug 27 using the NIRC2 imager on Keck-II (PI: D.~Huber).  The
+Aug 27 using the NIRC2 imager.  The
 observational configuration and reduction was identical as before.
 The images showed KOI-7913 A, KOI-7913 B, and an additional faint star
-$\approx0\farcs99$ due East of KOI-7913 B.  Applying the PSF-fitting
+$\approx$$0\farcs99$ due East of KOI-7913 B.  Applying the PSF-fitting
 routines from \citet{kraus_impact_2016}, the tertiary star has a
@@
 and young -- at the age of the system, it would have a mass roughly
-between 10 and 15\,M$_{\rm Jup}$.  Additional imaging epochs will be
-required to tell.
+between 10 and 15\,M$_{\rm Jup}$.  Additional imaging epochs will 
+tell.
 
@@
 KOI-7368 at a 30-minute cadence between May 2009 and April 2013.  For
-all three systems quarters 1 through 17 were observed without any data
-gaps other than the usual quarterly roll.  The top panel of
+all three systems quarters 1 through 17 were observed with minimal  data
+gaps.  The top panel of
 Figure~\ref{fig:planets} shows a 50-day slice of the \texttt{PDCSAP}
@@
 removed through a cotrending approach that uses a set of basis vectors
-derived using singular value decomposition of a set of
-systematics-dominated stellar light curves
+derived by applying singular value decomposition to a set of
+systematics-dominated light curves
 \citep{smith_kepler_PDC_2017}.  In our analysis, we used the
@@
 Lomb-Scargle periodogram for each Kepler quarter independently.  The
-resulting means and standard deviations are quoted in
+resulting means and standard deviations are in
 Table~\ref{tab:sysparams}.
@@
 \citep{thompson_planetary_2018}, the light curve was first trimmed to
-a local window around each transit of duration $\pm 3\times t_{14}$.
+a local window around each transit, spanning $\pm 3\times t_{14}$.
 The local out-of-transit points were then fitted with a fourth-order
@@
 \citet{gelman_inference_1992} statistic ($\hat{R}$) as our convergence
-diagnostic.  The results are shown in the lower panels of
+diagnostic.  The resulting fits are shown in the lower panels of
 Figure~\ref{fig:planets}, and the important derived parameters are in
@@
 of the hyperparameter priors, otherwise the GP would tend to
-incorporate any variability that was not part of the transit, even if it was
+incorporate variability that was not part of the transit, even if it was
 short-term variability attributable to flares, or simply
@@
 it yields an ill-founded sense of confidence in an ``overly clean''
-transit model fit.  Although our model is simpler, it has the benefit
+transit model fit.  While our model is simpler, it has the benefit
 that the white noise jitter never trades off with any parameter
@@
 
+We measured rotation periods for 53 stars with ZTF and 51 stars with TESS, for a total of 71 periods reported for unique stars and 32 overlapping between the surveys.
+We adopted the ZTF period over the TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the Lomb-Scargle periodogram favored 6.67 days, consistent with the ZTF period of 6.61 days; however, we flagged it as a candidate double-dipper, which appears to have inaccurately doubled the TESS period to 13.34 days; (2) Gaia EDR3 2081737529891330560: we found 3.06 days with TESS and 6.64 days with ZTF; we suspect that TESS captured the 1/2-period harmonic and adopt the approximately double value from ZTF; (3) 2134851775526125696: for this star, we measured 1.91 days with TESS from Cycle 2, but noted that the signal appeared to be missing in Cycle 4; ZTF found a strong signal at 12.23 days and we adopt this as the star's period.
+In the remaining overlap cases, we adopted the average between TESS and ZTF as the final period.
+For these overlap stars, the median absolute deviation is 0.01 days, showing remarkable consistency between the surveys.
+For three stars, we failed to detect a period in TESS but recovered one from ZTF; in all cases the periods appear to be 13--16 days, pushing the limit of what is detectable with TESS with our procedure: (1) Gaia EDR3 2129930258400157440, TESS showed a flat light curve while ZTF yielded a 15.3-day period; (2) Gaia EDR3 2082376861542398336, LS found a 7.6-day period which we rejected during visual validation; we found 15.4 days with ZTF, and we suspect that the weak/rejected singal form TESS might have been a 1/2 period harmonic; (3) Gaia EDR3 2082397099429013120, similar to the previous case, we rejected a 6.7-day signal from TESS and recovered a 12.8-day period with ZTF.
+
 \section{Spectroscopic Youth Indicators}

### 2022-04-08 papers/ms.tex
paper: add plots of spectra

@@
 
-We measured rotation periods for 53 stars with ZTF and 51 stars with TESS, for a total of 71 periods reported for unique stars and 32 overlapping between the surveys.
-We adopted the ZTF period over the TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the Lomb-Scargle periodogram favored 6.67 days, consistent with the ZTF period of 6.61 days; however, we flagged it as a candidate double-dipper, which appears to have inaccurately doubled the TESS period to 13.34 days; (2) Gaia EDR3 2081737529891330560: we found 3.06 days with TESS and 6.64 days with ZTF; we suspect that TESS captured the 1/2-period harmonic and adopt the approximately double value from ZTF; (3) 2134851775526125696: for this star, we measured 1.91 days with TESS from Cycle 2, but noted that the signal appeared to be missing in Cycle 4; ZTF found a strong signal at 12.23 days and we adopt this as the star's period.
-In the remaining overlap cases, we adopted the average between TESS and ZTF as the final period.
-For these overlap stars, the median absolute deviation is 0.01 days, showing remarkable consistency between the surveys.
-For three stars, we failed to detect a period in TESS but recovered one from ZTF; in all cases the periods appear to be 13--16 days, pushing the limit of what is detectable with TESS with our procedure: (1) Gaia EDR3 2129930258400157440, TESS showed a flat light curve while ZTF yielded a 15.3-day period; (2) Gaia EDR3 2082376861542398336, LS found a 7.6-day period which we rejected during visual validation; we found 15.4 days with ZTF, and we suspect that the weak/rejected singal form TESS might have been a 1/2 period harmonic; (3) Gaia EDR3 2082397099429013120, similar to the previous case, we rejected a 6.7-day signal from TESS and recovered a 12.8-day period with ZTF.
+We measured rotation periods for 53 stars with ZTF and 51 stars with
+TESS, for a total of 71 periods reported for unique stars and 32
+overlapping between the surveys. We adopted the ZTF period over the
+TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the
+Lomb-Scargle periodogram favored 6.67 days, consistent with the ZTF
+period of 6.61 days; however, we flagged it as a candidate
+double-dipper, which appears to have inaccurately doubled the TESS
+period to 13.34 days; (2) Gaia EDR3 2081737529891330560: we found 3.06
+days with TESS and 6.64 days with ZTF; we suspect that TESS captured
+the 1/2-period harmonic and adopt the approximately double value from
+ZTF; (3) 2134851775526125696: for this star, we measured 1.91 days
+with TESS from Cycle 2, but noted that the signal appeared to be
+missing in Cycle 4; ZTF found a strong signal at 12.23 days and we
+adopt this as the star's period. In the remaining overlap cases, we
+adopted the average between TESS and ZTF as the final period. For
+these overlap stars, the median absolute deviation is 0.01 days,
+showing remarkable consistency between the surveys. For three stars,
+we failed to detect a period in TESS but recovered one from ZTF; in
+all cases the periods appear to be 13--16 days, pushing the limit of
+what is detectable with TESS with our procedure: (1) Gaia EDR3
+2129930258400157440, TESS showed a flat light curve while ZTF yielded
+a 15.3-day period; (2) Gaia EDR3 2082376861542398336, LS found a
+7.6-day period which we rejected during visual validation; we found
+15.4 days with ZTF, and we suspect that the weak/rejected singal form
+TESS might have been a 1/2 period harmonic; (3) Gaia EDR3
+2082397099429013120, similar to the previous case, we rejected a
+6.7-day signal from TESS and recovered a 12.8-day period with ZTF.
 
@@
 \label{app:spectra}
+
+\begin{figure*}[tp]
+	\begin{center}
+		\leavevmode
+			\includegraphics[width=0.99\textwidth]{f5.pdf}
+	\end{center}
+	\vspace{-0.3cm}
+	\caption{
+    {\bf Key spectroscopic youth diagnostics for Kepler-1627, KOI-7368,
+    Kepler-1643, and KOI-7913 AB. }
+    The spectra are shown as in the observed frame, and
+    the stars are sorted left-to-right in order of decreasing effective temperature.
+    \label{fig:koiyouthindicators}
+	}
+\end{figure*}
+
+Figure~\ref{fig:koiyouthindicators} shows key portions of the HIRES
+and TRES spectra for the Kepler objects in Cep-Her.
+Lithium absorption is obvious at 6708\AA\ in all stars except KOI-7913 B, where it is marginal.
+H$\alpha$ is in emission for both components of KOI-7913,
+and in absorption for the hotter stars.
+In the following, 
+we compare these observations against stars in benchmark
+open clusters.
+
+\subsection{Lithium}
 
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f5a.pdf}
-			\includegraphics[width=0.49\textwidth]{f5b.pdf}
+			\includegraphics[width=0.49\textwidth]{f6a.pdf}
+			\includegraphics[width=0.49\textwidth]{f6b.pdf}
 		}
@@
 
-\subsection{Lithium}
-
 Figure~\ref{fig:lithium} compares the measured lithium equivalent
-widths of the objects of interest against a few reference populations.
-We selected the reference populations from the literature only for
-cases in which upper limits were explicitly reported.  KOI-7368 and
+widths of the Kepler objects against a few reference populations.
+We selected studies from the literature only where
+upper limits were explicitly reported.  KOI-7368 and
 KOI-7913~A have secure lithium detections, and the KOI-7913~B
@@
 We do not have an explanation, but in the following paragraph
-speculate on one possible line of thinking that might help resolve the
+speculate on one possible line of future work that might help resolve the
 discrepancy.

### 2022-04-08 papers/ms.tex
ms: add words about Hα

@@
 mass, KH timescale, etc}.
+
+\subsection{Further Work on Cep-Her}
+
+Disentangling its kinematics
+* Need an RV survey.
+
+Benchmarking its age(s):
+* Lithium analysis of FGK depletion (bright enough to be feasible).
+
+* Asteroseismology of the delta Sct stars (TESS data are there, group is
+rich enough).
+
+Finding more interesting objects in it:
+* Kepler data... presumably $\sim$200 stars in the cluster were
+observed, in order to have detected 4 planets.
+
+* TESS data
 
@@
 
-Gaia source lists and abbreviated info.  Include rotation periods.
+{\bf LGB TODO: add Gaia source lists and abbreviated info.  Include rotation periods.}
 
@@
 	\caption{
-    {\bf Key spectroscopic youth diagnostics for Kepler-1627, KOI-7368,
+    {\bf Spectroscopic youth diagnostics for Kepler-1627, KOI-7368,
     Kepler-1643, and KOI-7913 AB. }
-    The spectra are shown as in the observed frame, and
+    The spectra are shown in the observed frame, and
     the stars are sorted left-to-right in order of decreasing effective temperature.
@@
 		}
+
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f6a.pdf}
+		}
 	\end{center}
@@
 	\caption{
-    {\bf Lithium 6708\AA\ equivalent widths for Kepler-1643, KOI-7368,
-    KOI-7913 AB, and Kepler-1627 compared to those of young open
-    clusters and field stars. }
+    {\bf Lithium 6708\AA\ and H$\alpha$ equivalent widths for the
+    Kepler objects of interest compared to young open clusters and
+    field stars. }
     The statistical uncertainties on the newly measured equivalent
     widths are shown, or else are smaller than the markers.
+    {\it Top}:
     The field stars are stars with Kepler planets from \citet{berger_identifying_2018}.
@@
     \citep{soderblom_evolution_1993,jones_evolution_1996,bouvier_pleiades_lirot_2018},
-    where we adopt the LDB age from \citet{dahm_2015}.
+    where the quoted LDB age os from \citet{dahm_2015}.
+    {\it Bottom}:
+    The H$\alpha$ comparison is against Tuc-Hor
+    \citep[$\approx$$40$\,Myr;][]{kraus_stellar_2014}
     \label{fig:lithium}
@@
 
+\subsection{H$\alpha$}
+
+As shown in Figure~\ref{fig:koiyouthindicators}, H$\alpha$ is in
+emission for both components of KOI-7913, and in absorption for the
+hotter stars.  Additionally, the emission appears double-peaked for
+both of the KOI-7913 components.  We note that KOI-7913~A and
+KOI-7913~B were spatially resolved from each other during data
+acquisition.  Performing a cross-correlation between each of the stars
+and the nearest matches in the Keck/HIRES template library, we also
+found that the CCFs for both components of KOI-7913 showed no
+indications of double-lined binarity \citep{kolbl_detection_2015}.
+
+Balmer line emission, particularly for H$\alpha$, is expected for
+stars of this age.  \citet{kraus_stellar_2014} for instance, in their
+survey of Tuc-Hor ($\approx$40\,Myr), observed that essentially all
+cluster members with spectral type $>$${\rm K4.5V}$ had H$\alpha$ in
+emission.  The agrees with our observations that KOI-7913 shows
+H$\alpha$ in emission for both components, while it is in absorption
+for all of our other Kepler objects.  The double-peaked nature of the
+emission, though not always present, is also not particularly rare for
+active stars.  Proxima Centauri, for instance, has double-peaked
+H$\alpha$ emission \citep{collins_calculations_2017}.  A simple
+explanation is self-absorption: photons nearest the center of the line
+see a greater optical depth from higher layers of the chromosphere,
+while photons on the wings are on average too far from the
+rest-wavelength to excite electrons and be re-absorbed in the upper
+layers.  The exact details of when a star's atmosphere reaches the
+conditions for such self-absorption require non-local thermal
+equilibrium models of the chromosphere
+\citep{short_chromospheric_1998,2005A&A...439.1137F}.
+
 \section{Table of Transit Fit Parameters}

### 2022-04-08 papers/ms.tex
update ms: h-alpha

@@
 from Figure~\ref{fig:age}.  They span spectral types of G8V
-(Kepler-1627) to K7V (KOI-7913 A).  The secondary in the KOI-7913
-system has spectral type $\approx$M0V.  And since a
+(Kepler-1627) to K6V (KOI-7913 A).  The secondary in the KOI-7913
+system has spectral type $\approx$K8V.  And since a
 Solar-mass star with solar metallicity arrives at the zero-age main
@@
 Regarding indicators of youth, the primary has a spectral type
-$\approx$K7V, while the secondary is $\approx$M0V.  At $\approx$40
+$\approx$K6V, while the secondary is $\approx$K8V.  At $\approx$40
 Myr, lithium 6708\,\AA\ absorption is not expected for the secondary,
@@
 on the data from K.~Rocio et al's analysis\footnote{Available at
-\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K7V
+\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K6V
 star should be emitting in H$\alpha$ at 40\,Myr, but not typically
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f6a.pdf}
+			\includegraphics[width=0.49\textwidth]{f6c.pdf}
 		}
@@
     {\bf Lithium 6708\AA\ and H$\alpha$ equivalent widths for the
-    Kepler objects of interest compared to young open clusters and
+    objects of interest compared to young open clusters and
     field stars. }
+	Positive equivalent width means absorption; negative
+	equivalent width means emission.
     The statistical uncertainties on the newly measured equivalent
@@
     The ``112 Myr'' stars (right) are from the Pleiades
-    \citep{soderblom_evolution_1993,jones_evolution_1996,bouvier_pleiades_lirot_2018},
-    where the quoted LDB age os from \citet{dahm_2015}.
+    \citep{soderblom_evolution_1993,jones_evolution_1996,bouvier_pleiades_lirot_2018}.
     {\it Bottom}:

### 2022-04-08 papers/collect.sh

@@ -21,11 +21,10 @@ cp ~/Dropbox/proj/rudolf/results/run_simpletransit_Kepler_1627/*singlepanel*pdf 
 # Rp, Period, Age
 cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20220405_colorbyage_showCepHer.pdf f4.pdf
 
-# spectra: Li EW vs Teff and Halpha EW vs Teff
-# TODO:
-# Li, H-alpha?
+# spectra
 cp /Users/luke/Dropbox/proj/rudolf/results/lithium/lithium_Randich01_TucHorK14M08.pdf f6a.pdf
 cp /Users/luke/Dropbox/proj/rudolf/results/lithium/lithium_Pleiades.pdf f6b.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/halpha/halpha.pdf f6c.pdf
 
 cp /Users/luke/Dropbox/proj/rudolf/results/koiyouthlines/koiyouthlines.pdf f5.pdf
 

### 2022-04-08 rudolf/starinfo.py

@@ -8,6 +8,7 @@ starinfodict = {
         # rotation_period_windowslider
         'Prot': 2.606363,
         'Prot_err': 0.051906,
+        'SpT': ['K0V', -10],
     },
     "KOI-7913 A": {
         # HIRES / SM-Emp Median
@@ -18,6 +19,7 @@ starinfodict = {
         # rotation_period_windowslider
         'Prot': 3.197631,
         'Prot_err': 0.056841,
+        'SpT': ['K6V', -4],
     },
     "KOI-7913 B": {
         # HIRES / SM-Emp Median
@@ -28,6 +30,7 @@ starinfodict = {
         # rotation_period_windowslider
         'Prot': 2.641824,
         'Prot_err': 0.066772,
+        'SpT': ['K8V', -2],
     },
     "Kepler-1643": {
         # HIRES / SM-Emp Median
@@ -39,6 +42,7 @@ starinfodict = {
         # rotation_period_windowslider
         'Prot': 5.106414,
         'Prot_err': 0.043881,
+        'SpT': ['K3V', -7],
     },
     "Kepler-1627 A": {
         # HIRES / SM-Emp Median / Bouma+22
@@ -49,5 +53,6 @@ starinfodict = {
         # rotation_period_windowslider
         'Prot': 2.641824,
         'Prot_err': 0.043764,
+        'SpT': ['G8V', -12],
     },
 }

### 2022-04-09 papers/ms.tex
update ms

@@
 %\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
-\documentclass[12pt,twocolumn,linenumbers]{aastex63}
+\documentclass[12pt,twocolumn]{aastex63}
 %\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
@@
   Here we focus on the former three systems, which are in other
-  components of the complex.
+  regions of the association.
   Based on kinematic evidence from Gaia, stellar rotation periods from
@@
 photometry from K2 and TESS has yielded a number of young planets for
-which the planetary evidence comes from transits, and the age evidence
+which the planetary evidence comes from transits, and the age 
 is based on either cluster membership
@@
 e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To date these
-analyses have mostly clustered on the stellar positions and on-sky
-velocities measured by Gaia, with varying methodologies fo filtering
-and supervision.  One important result is the identification of
+analyses have mostly clustered on the stellar positions and 2-D
+velocities measured by Gaia.  One important result is the discovery of
 diffuse streams and tidal tails comparable in stellar mass to the
-previously known cores of many nearby open clusters
-\citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  These
-spatially diffuse groups typically have velocity dispersions of
-$\approx$1\kms, though they can be much higher due to both projection
-effects and internal dynamics.  As an extreme example, in the Hyades
-the velocities of stars in the tidal tails span up to $\pm 40\mkms$
-relative to the cluster center \citep{jerabkova_800_2021}.  The stars
+previously known cores of nearby open clusters
+\citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  
+Even though these streams can be spread out over tens to hundreds of parsecs,
+their velocity dispersions can remain coherent at the $\sim$1\kms\ level.
+They can also be much larger:
+as an extreme example, internal dynamics and projection effects
+in the Hyades drive stars in the tidal tails span up to $\pm 40\mkms$
+in relative velocity when projected against the cluster center \citep{jerabkova_800_2021}.  
+The stars
 in such diffuse regions can be verified to be the same age as the core
 members ({\it i.e.}, coeval) through analyses of color-absolute
-magnitude diagrams, stellar rotation periods
+magnitude diagrams \citep{KounkelCovey2019}, stellar rotation periods
 \citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
@@
 and ${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.
-
-% Randich+01: Li starts to become upper limits around K8V, V-I=1.6.
-% Around K6V, V-I=1.4, ~100mA seen, but few stars in this region!
-
-Regarding indicators of youth, the primary has a spectral type
-$\approx$K6V, while the secondary is $\approx$K8V.  At $\approx$40
-Myr, lithium 6708\,\AA\ absorption is not expected for the secondary,
-but might be present in the primary \citep[{\it
-e.g.},][Figure~8]{soderblom_ages_2014}.  Conversely, H$\alpha$ is
-expected to be in emission for for all M-dwarfs younger than
-$\approx150$\,Myr \citep[{\it e.g.},][]{kiman_calibration_2021}.  For
-the primary, H$\alpha$ emission is also expected based on
-previous spectroscopic studies of the Pleiades, IC~2602, and IC~2391
-\citep[{\it e.g.},][their Figure~8]{stauffer_rotational_1997}.  Based
-on the data from K.~Rocio et al's analysis\footnote{Available at
-\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K6V
-star should be emitting in H$\alpha$ at 40\,Myr, but not typically
-past $\approx200\,{\rm Myr}$.  This presents a fairly direct test for
-the star's age.
-
-Analyzing the HIRES spectra, we find
-...
-Turning to the actual data, we find {\bf X for lithium}.
-For H$\alpha$, {\bf we see Y}.
-
-% For Li:
-%    ~~~(CITE Figure 10 of Barrado y Navasceus 2004?).
-%   ****USE FIG8 OF SODERBLOM+13****
-% 
-%   OR USE BAFFLES AS YOUR REFERENCE?
-%     B-V=1.48 (primary), IC2602 age has no data.  beta pic age has lots of scatter
-%     around this color, between 10mA and 100mA.
-%     B-V=redder... 1.8?  SpType ~M0V, really no Li expected.
-%   A: no... they really didn't collect very good data from what i see.
-%   especially at the red end...
-% 
-% NW component SpType~K5.5-K6V; HIRES spectrum shows weak Li (EW 65 +/-
-% 8 mA), Halpha in emission (double-peaked?!), and CaHK in emission.
-% Reamatch SB2 checks don't show evidence for it being SB2.
-% SW component SpType~K9V-M0V; HIRES spectrum shows H-alpha in emission
-% (consistent with <150 Myr) and no Li (expected given Teff).
-% Given the kinematics, the spectra paint a picture consistent with KOI
-% 7913 being part of the ~40 Myr Cep-Her group.
+Neither of the KOI-7913 components shows strong lithium absorption,
+but this is expected given their $\approx$K6V and $\approx$K8V
+spectral types.
+They do however both show H$\alpha$ in emission.
+We discuss the significant in greater depth in Appendix~\ref{app:spectra},
+but broadly speaking both observations are consistent with
+a $\approx$40 Myr age for KOI-7913.
 
@@
 Kepler-1643~b (KOI-6186.01) was already validated as a transiting
-planet by \citet{morton_false_2016}.  Based on the Q1-Q17 DR25 False
-Positive Probability table hosted at the NASA Exoplanet Archive, they
+planet by \citet{morton_false_2016}.  %Based on the Q1-Q17 DR25 False
+%Positive Probability table hosted at the NASA Exoplanet Archive, 
+They
 found a probability for any of the considered false positive scenarios
@@
 Figure~\ref{fig:planets} shows the intuitive justification for the
-strong statistical result: the transit is flat and has a reasonably
+result: the transit is flat and has a reasonably
 high S/N ($\approx 47$), which is difficult to reproduce with
@@
 
-An interesting aside on Kepler-1643 is that it formally failed the
-centroid shift test in the \texttt{q1\_q17\_dr25\_koi} data release,
+Intriguingly, Kepler-1643 also formally failed the
+centroid shift test (see the \texttt{q1\_q17\_dr25\_koi} data release),
 with the angular distance between the target star position and the
-location of the transiting source being reported as $0\farcs99$
-(4.4-$\sigma$).  This seems to be mostly due to two outlying quarters
+location of the transiting source being measured as $1\farcs0$
+at 4.4-$\sigma$ confidence.  There seem to be two outlying quarters
 (Q2 and Q6) driving the offset -- the centroid locations from the
@@
 flux-weighted centroids, but for our stars there is no static baseline
-in either in- or out-of-transit phase.
+in either phase.
 
 \paragraph{KOI-7368}
-KOI-7368.01 is listed on the NASA Exoplanet Archive as a ``CANDIDATE''
+KOI-7368.01 is listed on the NASA Exoplanet Archive as a ``candidate''
 planet.  The Q1-Q17 DR25 False Positive Probability table however
@@
 
-We have argued that Kepler-1643~b, KOI-7368.01, and KOI-7913 {\it i)}
+We have argued that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab {\it i)}
 are 40 to 50 million years old, and {\it ii)} that each signal
@@
 check for centroid offsets between the in and out-of-transit phases.
-For both cases, we have argued the weak observed shifts are most
-likely caused by starspot-induced variability in specific quarters
+For both cases, we have argued the weak observed shifts seem to be
+caused by starspot-induced variability in specific quarters
 shifting the stellar center-of-light.
+In addition, our imaging rules out companions stars with the necessary
+brightness and position to explain the reported shifts.
 
@@
 of this cluster with a false positive rate of $\ll 5\%$
-(Figure~\ref{fig:age}); Kepler-1643 is one such member.  The stellar
-rotation period period agrees with this assessment, and {\bf its
-lithium is a little low}.
+(Figure~\ref{fig:age}); Kepler-1643 is one such member.  While the stellar
+rotation period period agrees with this assessment, its lithium
+content is somewhat low, which might motivate future exploration
+of lithium depletion across FGKM stars in RSG-5 (see Appendix~\ref{app:spectra})
 
 The spatio-kinematic argument for the youth of KOI-7368 and
-KOI-7913 are weaker because they are in an association of stars, CH-2,
+KOI-7913 is weaker because they are in an association of stars, CH-2,
 that is more diffuse.  For KOI-7913, stronger indicators of its age
-perhaps come from its binary.  Both stellar components in KOI-7913
-have isochronal ages consistent with $40$\,Myr, but the isochronal
-measurement for the secondary is more constraining since it is less
-massive.  Both components also show H$\alpha$ in emission, which for
+come from its binary.  Both stellar components in KOI-7913
+have isochronal ages consistent with $40$\,Myr.
+Both components also show H$\alpha$ in emission, which for
 the $\approx$K6V primary is a good indicator that the star is
@@
 
+%Regarding indicators of youth, the primary has a spectral type
+%$\approx$K6V, while the secondary is $\approx$K8V.  At $\approx$40
+%Myr, lithium 6708\,\AA\ absorption is not expected for the secondary,
+%but might be present in the primary \citep[{\it
+%	e.g.},][Figure~8]{soderblom_ages_2014}.  Conversely, H$\alpha$ is
+%expected to be in emission for for all M-dwarfs younger than
+%$\approx150$\,Myr \citep[{\it e.g.},][]{kiman_calibration_2021}.  For
+%the primary, H$\alpha$ emission is also expected based on
+%previous spectroscopic studies of the Pleiades, IC~2602, and IC~2391
+%\citep[{\it e.g.},][their Figure~8]{stauffer_rotational_1997}.  Based
+%on the data from K.~Rocio et al's analysis\footnote{Available at
+%	\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K6V
+%star should be emitting in H$\alpha$ at 40\,Myr, but not typically
+%past $\approx200\,{\rm Myr}$.  This presents a fairly direct test for
+%the star's age.
+%We
+
 \section{Table of Transit Fit Parameters}

### 2022-04-10 papers/ms.tex
update ms

@@
 \paragraph{Kepler-1643}
+  %Based on the Q1-Q17 DR25 False Positive Probability table hosted at
+%the NASA Exoplanet Archive, 
 Kepler-1643~b (KOI-6186.01) was already validated as a transiting
-planet by \citet{morton_false_2016}.  %Based on the Q1-Q17 DR25 False
-%Positive Probability table hosted at the NASA Exoplanet Archive, 
-They
-found a probability for any of the considered false positive scenarios
-of 9$\times$10$^{-6}$.  Repeating the calculation with our own
-stellar-variability correction and the new NIRC2 imaging constraints,
-we find ${\rm FPP} = 6\times10^{-9}$, where the improvement is
-presumably due to the new NIRC2 contrast curves.
-Figure~\ref{fig:planets} shows the intuitive justification for the
-result: the transit is flat and has a reasonably
-high S/N ($\approx 47$), which is difficult to reproduce with
-eclipsing binary models.
-
-Intriguingly, Kepler-1643 also formally failed the
-centroid shift test (see the \texttt{q1\_q17\_dr25\_koi} data release),
-with the angular distance between the target star position and the
-location of the transiting source being measured as $1\farcs0$
-at 4.4-$\sigma$ confidence.  There seem to be two outlying quarters
-(Q2 and Q6) driving the offset -- the centroid locations from the
-other Kepler quarters are all consistent within $\lesssim 0\farcs5$.
-Nonetheless, this is an instructive exercise in how variable stars can
-complicate centroid-based vetting tests.  The centroid shifts measured
-by these tests are determined from the in- and out-of-transit
-flux-weighted centroids, but for our stars there is no static baseline
-in either phase.
+planet by \citet{morton_false_2016}, who found a probability for any
+of the aforementioned false positive scenarios of 9$\times$10$^{-6}$.
+Repeating the calculation with our own stellar-variability correction
+and the new NIRC2 imaging constraints, we find ${\rm FPP} =
+6\times10^{-9}$, where the improvement is due to the new NIRC2
+contrast curves.  Figure~\ref{fig:planets} shows the justification:
+the transit is flat and has a high S/N ($\approx 47$), which is
+difficult to reproduce with eclipsing binary models.
+
+Intriguingly, Kepler-1643 formally failed one of the centroid shift
+tests (see the \texttt{q1\_q17\_dr25\_koi} data release): the angular
+distance between the target star's catalog position and the position
+of the transiting source was measured as $1\farcs0$ at 4.4-$\sigma$.
+The reports show however that two outlying quarters (2 and 6) 
+drive the offset -- the centroid locations from the other Kepler
+quarters are consistent at $\lesssim 0\farcs4$.  Nonetheless, this
+is an instructive exercise in how stellar variability complicates
+centroid-based vetting tests.  The centroid shifts measured by these
+tests are determined from the in- and out-of-transit flux-weighted
+centroids.  For stars with significant spot-induced
+variability there is no static baseline in either the in- or
+out-of-transit phases.
 
@@
 KOI-7368.01 is listed on the NASA Exoplanet Archive as a ``candidate''
-planet.  The Q1-Q17 DR25 False Positive Probability table however
-lists its probability of being an astrophysical false positive as
-``1.0''. However, given that none of the other columns in this table
-are populated, this is likely a database entry error ({\bf ask
-Jessie}).  Performing the validation calculation using our new NIRC2
-imaging constraints, we find ${\rm FPP} \approx 4\times10^{-3}$.
-While not as convincing as Kepler-1643, this formally clears the
-usual threshold for calling the planet statistically validated
-\citep{morton_efficient_2012}.  The S/N of the transit itself is
-$\approx 32$, which indicates that it is unlikely to be caused by
-systematic noise in the light curve.  Visually, one can also inspect
-each individual transit window in the light curve itself.  This
-exercise shows a weak notch in each case; over the 198 observed
-transits, the resulting signal seems sensible.
-
-Why was KOI-7368 not previously validated?  One reason could be the
-high stellar variability.  Another reason could be that it also
-formally fails the centroid shift test, similar to Kepler-1643.  In
-the case of KOI-7368, the reported offset is marginally less
-significant ($0\farcs22$; 3.0-$\sigma$).  Again, inspection of the
-data validation reports shows that this is caused by a few outlying
-quarters: in this case Q4, Q5, Q8, and Q12.  The fact that the
-remaining quarters show a more consistent scatter suggests that these
-outliers are again caused by stellar variability local to the
-transits, rather than by an astrophysical offset.  Our imaging also
-shows that there are no neighboring sources that could cause an
-offset of the observed amplitude.
+planet.  \citet{morton_false_2016} did not compute a false positive
+probability for the system because their default trapezoidal fitting
+routine failed, presumably due to the spot-induced variability.
+%The Q1-Q17 DR25 False Positive Probability table however lists its
+%probability of being an astrophysical false positive as ``1.0''. Given
+%that none of the other columns in the table are populated, this seems
+%likely to be a database entry error ({\bf ask Jessie}). 
+Our fitting approach rectifies this point, and our new NIRC2 images
+revealed no new stellar companions.  Performing the validation
+calculation, we find ${\rm FPP} \approx 4\times10^{-3}$.  Though not
+as convincing as Kepler-1643, this formally clears the usual threshold
+for calling the planet statistically validated
+\citep{morton_efficient_2012}.  The S/N of the transit is
+$\approx$$32$, which indicates that it is unlikely to be caused by
+systematic noise in the light curve (see Figure~\ref{fig:planets}).
+The positional probability score also meets the requirements
+for transiting sources thought to share positions with their
+target stars \citep{morton_false_2016,2017ksci.rept...16B}. 
+%A visual inspection of
+%each individual transit also shows a weak notch at the predicted
+%transit times; over the 198 observed transits, the resulting signal
+%seems sensible.
+
+It bears mentioning that KOI-7368 shows a weak centroid shift in the
+\texttt{q1\_q17\_dr25\_koi} validation reports, similar to
+Kepler-1643.  For KOI-7368, the reported offset is smaller, and less
+statistically significant ($0\farcs2$; 3.0-$\sigma$).  Again,
+the data validation reports show that the shift is caused by a few
+outlying quarters (4, 5, 8, and 12).  The fact that the remaining data
+show a consistent scatter between centroid locations suggests that
+these outliers are also caused by the stellar variability.  Independently, our
+imaging shows that there are no neighboring sources that could
+cause an offset of the observed amplitude.
 
 \paragraph{KOI-7913}
-KOI-7913.01 is listed on the NASA Exoplanet Archive as a ``CANDIDATE''
-planet.  The Q1-Q17 DR25 False Positive Probability table lists its
-probability of being an astrophysical false positive as
-$1.4\times10^{-4}$, with the most likely false positive scenario being
-a blended eclipsing binary.  Repeating the calculation with our new
-detrending and contrast curves, we find a similar result: ${\rm FPP} =
-1.3\times10^{-4}$.  While the transit has the lowest S/N of any of the
-objects discusssed (${\rm S/N}=14$), its lower FPP relative to
+KOI-7913.01 is also listed on the NASA Exoplanet Archive as a
+``candidate'' planet.  The canonical \citet{morton_false_2016} analysis
+was of Q1-Q17 KOIs from DR24, and therefore spanned KOI-1.01 to
+KOI-7620.01 and omitted KOI-7913.01.  However the results of the
+subsequent DR25 analysis by Morton et al.\ are listed at the NASA
+Exoplanet Archive.  The relevant table gives a probability for the
+system being an astrophysical false positive of $1.4\times10^{-4}$,
+with the most likely false positive scenario being a blended eclipsing
+binary.  Repeating the calculation with our new detrending and
+contrast curves, we find a similar result: ${\rm FPP} =
+1.3\times10^{-4}$.  Though the transit has the lowest S/N of any of
+the objects discusssed (${\rm S/N}=14$), its lower FPP relative to
 KOI-7368 can be understood through its relatively flat-bottomed shape,
 combined with its long transit duration relative to plausible
-eclipsing binary models (Figure~\ref{fig:planets}).
-
-The disposition of KOI-7913.01 has previously been debated:
-in \texttt{q1\_q17\_dr25\_koi} the source was flagged as a ``FALSE POSITIVE'',
-with the comment ``CENT\_KIC\_POS---HALO\_GHOST''.  This comment and
-disposition seem to have been removed in the
-\texttt{q1\_q17\_dr25\_sup\_koi} data release, which renamed it a
-``CANDIDATE'', without any flag given.  As described by
-\citet{thompson_planetary_2018}, the ``CENT\_KIC\_POS'' flag is an
-indication that the measured source centroid is offset from its
-expected location in the Kepler Input Catalog.  The final Data
-Validation Reports, generated 2016 Jan 30, do not seem to show this.
-Moreover, the statistical significance of any centroid offset seems to
-be lower than for KOI-7368 and Kepler-1643 (which, as we have argued,
-both show centroid offsets that seem to be explained by the stellar
-variability).
-
-What of the ``HALO\_GHOST'' flag?  This test measures the transit
-strength for the pixels inside the aperture, and compares it to that
-measured in the ring of pixels around said aperture (the ``halo'').
-Generally one expects the transit signal to be strongest in the
-central aperture, rather than the halo.  Two types of false positive
-scenarios can change this and trigger the flag: the first is when
-optical ghosts from bright eclipsing binaries reflect off the CCD, and
-contaminate the target star.  The second is when the PRF of nearby
-stars directly overlaps with the PRF of the target star (see
-\citealt{thompson_planetary_2018}, Section A.5.2).
-
-The most obvious explanation for KOI-7913 is the latter case, since
-KOI-7913 B is $\approx0.9$ Kepler pixels away from Kepler-7913 A.  Due
-to the on-sky orientation of KOI-7913 A and KOI-7913 B, the default
-``optimal aperture'' selected in Q3, Q7, Q11, and Q15 in fact included
-both stars, while for the remaining quarters KOI-7913 B was
-(correctly) excluded from the optimal aperture (see pages 35 through
-71 of the Data Validation Report.)
-
-As a test, we opted to re-extract the light curves from the target
-pixel files for these particular cases.
-{\bf Comparing the pixels on-target, and found}
-%FIXME FIXME TODO TODO
-{\bf We found...}
+eclipsing binary models (Figure~\ref{fig:planets}).  The host star
+probability usability score \citep{2017ksci.rept...16B} also meets the
+usual threshold, and so the planet seems to be validated.  Its
+disposition has however previously fluctuated from ``false positive''
+to ``candidate''.  The most likely explanation is the presence of
+KOI-7913~B, which is located approximately one Kepler pixel away from Kepler-7913 A (see
+Appendix~\ref{app:koi7913}).  Our pixel-level analysis of the target
+pixel files did not show any evidence to support the idea that
+KOI-7913~B could host the transiting planet.  Our NIRC2 imaging
+similarly ruled out the possibility of any other plausible host star
+for the planet other than KOI-7913~A.
 
@@
 
+\subsection{Normal-Sized Mini-Neptunes Exist at 40$\,$Myr}
+
+Current thinking about the origins and evolution of lose-in
+mini-Neptunes has recently been reviewed by \citet{bean_nature_2021}.
+In the context of the existing exoplanet population, the most
+significant novelty about Kepler-1643~b, KOI-7368.01, and KOI-7913 is
+that they have normal sizes.
+More specifically, the size distribution of mini-Neptunes spans
+$\approx$1.8\,$R_\oplus$ to $\approx$3.6\,$R_\oplus$, with a peak at
+$\approx 2.4\,R_\oplus$ \citep[][Figure~7]{Fulton_et_al_2017}.
+Figure~\ref{fig:rp_period_age} shows the sizes, orbital periods, and ages of
+known transiting planets, emphasizing those with precise ages.
+Considering planets younger than $10^8$ years, the smallest previously
+known were AU~Mic~c ($\approx3.0\pm0.2\,R_\oplus$, averaging the
+results of \citealt{martioli_aumicbc_2021} and
+\citealt{gilbert_flares_2022}), Kepler-1627~Ab ($3.8\pm0.2\,R_\oplus$;
+\citealt{bouma_kep1627_2022}), and AU~Mic~d ($4.2\pm0.2\,R_\oplus$;
+\citealt{plavchan_planet_2020}).
+The other transiting planets younger than $10^8$ years are much
+larger, with sizes between 4 and 10$\,R_\oplus$
+\citep{Mann_K2_33b_2016,David_et_al_2017,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020}.
+
+%TODO: I think we might need James Owen to contribute something
+%similar to the TOI-1227 models?
+{\bf Run some evolutionary models, make some comments about the core
+	mass, KH timescale, etc}.
+
 \subsection{Is CH-2 really a star cluster?}
@@
 What is the likelihood we are fooling ourselves?
-
-\subsection{Normal-Sized Mini-Neptunes Exist at 40$\,$Myr}
-
-Current thinking about the origins and evolution of lose-in
-mini-Neptunes has recently been reviewed by \citet{bean_nature_2021}.
-In the context of the existing exoplanet population, the most
-significant novelty about Kepler-1643~b, KOI-7368.01, and KOI-7913 is
-that they have normal sizes.
-More specifically, the size distribution of mini-Neptunes spans
-$\approx$1.8\,$R_\oplus$ to $\approx$3.6\,$R_\oplus$, with a peak at
-$\approx 2.4\,R_\oplus$ \citep[][Figure~7]{Fulton_et_al_2017}.
-Figure~\ref{fig:rp_period_age} shows the sizes, orbital periods, and ages of
-known transiting planets, emphasizing those with precise ages.
-Considering planets younger than $10^8$ years, the smallest previously
-known were AU~Mic~c ($\approx3.0\pm0.2\,R_\oplus$, averaging the
-results of \citealt{martioli_aumicbc_2021} and
-\citealt{gilbert_flares_2022}), Kepler-1627~Ab ($3.8\pm0.2\,R_\oplus$;
-\citealt{bouma_kep1627_2022}), and AU~Mic~d ($4.2\pm0.2\,R_\oplus$;
-\citealt{plavchan_planet_2020}).
-The other transiting planets younger than $10^8$ years are much
-larger, with sizes between 4 and 10$\,R_\oplus$
-\citep{Mann_K2_33b_2016,David_et_al_2017,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020}.
-
-%TODO: I think we might need James Owen to contribute something
-%similar to the TOI-1227 models?
-{\bf Run some evolutionary models, make some comments about the core
-mass, KH timescale, etc}.
 
@@
 
+\section{Disposition History of KOI-7913}
+\label{app:koi7913}
+
+The disposition of KOI-7913.01 has been debated: in
+\texttt{q1\_q17\_dr25\_koi} the source was flagged as a false
+positive, with the comment ``cent\_kic\_pos---halo\_ghost''.  This
+comment and disposition were removed in the
+\texttt{q1\_q17\_dr25\_sup\_koi} data release, which renamed the
+planet a ``candidate''.  In this note, we discuss the interpretation
+of these flags (which do not apply to the system, according to the
+latest analysis).  We also discuss how the relative locations of
+KOI-7913~A and KOI-7913~B affect the interpretation of the Kepler
+data.
+
+As described by \citet{thompson_planetary_2018}, the
+``cent\_kic\_pos'' flag is an indication that the measured source
+centroid is offset from its expected location in the Kepler Input
+Catalog.  The final Kepler data validation reports, generated 2016 Jan
+30, do not show this to be the case for KOI-7913.  Moreover, the
+statistical significance of any centroid offset is lower than for
+KOI-7368 and Kepler-1643 (which both show centroid offsets that are
+likely explained by the stellar variability).
+
+What of the ``halo\_ghost'' flag?  This test measures the transit
+strength for the pixels inside the aperture, and compares it to that
+measured in the ring of pixels around said aperture (the ``halo'').
+Generally one expects the transit signal to be strongest in the
+central aperture, rather than the halo.  Two types of false positive
+scenarios can change this and trigger the flag: the first is when
+optical ghosts from bright eclipsing binaries reflect off the CCD, and
+contaminate the target star.  The second is when the PRF of nearby
+stars directly overlaps with the PRF of the target star (see
+\citealt{thompson_planetary_2018}, Section A.5.2).
+The most obvious explanation for KOI-7913 is the latter case, given
+that KOI-7913 B is $\approx0.9$ Kepler pixels away from Kepler-7913 A
+and so it usually part of the ``halo''.  Due to the on-sky orientation
+of KOI-7913 A and KOI-7913 B, the default ``optimal aperture''
+selected in quarters 3, 7, 11, and 15 in fact included both stars,
+while for the remaining quarters KOI-7913 B was excluded from the
+optimal aperture but was included as part of the halo (see pages 35
+through 71 of the data validation reports.)
+
+Given the orientation of the stars and the $\approx$1.5 pixel FWHM of
+the Kepler pixel response function, some blending between the two
+stars is present.  The pointing geometries from quarters 3, 7, 11, and
+15 however did not noticeably affect the observed transit depths,
+which is an indication that the crowding metric applied in the data
+products accurately correct the mean flux level
+\citep{2017ksci.rept....6M}.  Analysis of the target-pixel data that
+was acquired separately for KOI-7913~B also reveals a clearly
+different stellar rotation period, and no hint of the transit signal.
+
 \input{source_Kepler_1643_simpletransit_posteriortable.tex}

### 2022-04-10 papers/ms.tex
update ms
cleanup tex

@@
 \begin{abstract}
-  Stellar positions and velocities from Gaia are yielding a 
-  refined view of open cluster dispersal.
-  Here we present an analysis of a group stars spanning Cepheus
-  ($l=100^\circ$) to Hercules ($l=40^\circ$), hereafter the Cep-Her
-  complex.
+  Stellar positions and velocities from Gaia are yielding a refined
+  view of open cluster dispersal.  Here we present an analysis of a
+  group stars spanning Cepheus ($l=100^\circ$) to Hercules
+  ($l=40^\circ$), hereafter the Cep-Her complex.
   The group includes four Kepler Objects of Interest:
@@
   Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
-  The latter Neptune-sized planet is in a component of the
-  Cep-Her complex called the $\delta$\ Lyr\ cluster
-  \citep{bouma_kep1627_2022}.
-  Here we focus on the former three systems, which are in other
-  regions of the association.
-  Based on kinematic evidence from Gaia, stellar rotation periods from
-  TESS, and spectroscopy, these three systems are also $\approx$40
-  million years (Myr) old.
-  More specifically, we find that Kepler-1643 is $46^{+9}_{-7}$\,Myr
-  old, based on its membership in a dense
-  sub-cluster of the complex called RSG-5.
-  KOI-7368 and KOI-7913 are $36^{+10}_{-8}$\,Myr old, and are in a
-  diffuse region that we call CH-2.
-  Based on the transit shapes and high resolution imaging, all three
-  objects are most likely planets, with false positive probabilities
-  of $6\times10^{-9}$, $4\times10^{-3}$, and $1\times10^{-4}$ for
-  Kepler-1643, KOI-7368, and KOI-7913 respectively.
-  These planets empirically demonstrate that mini-Neptunes with sizes
-  of $\approx$2 Earth radii exist at ages of $\approx$40 million
-  years.
+  The latter Neptune-sized planet is in a component of the Cep-Her
+  complex called the $\delta$\ Lyr\ cluster
+  \citep{bouma_kep1627_2022}.  Here we focus on the former three
+  systems, which are in other regions of the association.  Based on
+  kinematic evidence from Gaia, stellar rotation periods from TESS,
+  and spectroscopy, these three systems are also $\approx$40 million
+  years (Myr) old.  More specifically, we find that Kepler-1643 is
+  $46^{+9}_{-7}$\,Myr old, based on its membership in a dense
+  sub-cluster of the complex called RSG-5.  KOI-7368 and KOI-7913 are
+  $36^{+10}_{-8}$\,Myr old, and are in a diffuse region that we call
+  CH-2.  Based on the transit shapes and high resolution imaging, all
+  three objects are most likely planets, with false positive
+  probabilities of $6\times10^{-9}$, $4\times10^{-3}$, and
+  $1\times10^{-4}$ for Kepler-1643, KOI-7368, and KOI-7913
+  respectively.  These planets empirically demonstrate that
+  mini-Neptunes with sizes of $\approx$2 Earth radii exist at ages of
+  $\approx$40 million years.
 \end{abstract}
@@
 photometry from K2 and TESS has yielded a number of young planets for
-which the planetary evidence comes from transits, and the age 
-is based on either cluster membership
+which the planetary evidence comes from transits, and the age is based
+on either cluster membership
 \citep{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
@@
 previously known cores of nearby open clusters
-\citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  
-Even though these streams can be spread out over tens to hundreds of parsecs,
-their velocity dispersions can remain coherent at the $\sim$1\kms\ level.
-They can also be much larger:
-as an extreme example, internal dynamics and projection effects
-in the Hyades drive stars in the tidal tails span up to $\pm 40\mkms$
-in relative velocity when projected against the cluster center \citep{jerabkova_800_2021}.  
-The stars
-in such diffuse regions can be verified to be the same age as the core
-members ({\it i.e.}, coeval) through analyses of color-absolute
-magnitude diagrams \citep{KounkelCovey2019}, stellar rotation periods
-\citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
-\citep{hawkins_2020}.  While there are many implications for our
-understanding of star formation and cluster evolution
-\citep{dinnbier_tidal_2020}, a separate consequence is that we now
-know the ages of many more stars, including previously known planet
-hosts.
+\citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  Even
+though these streams can be spread out over tens to hundreds of
+parsecs, their velocity dispersions can remain coherent at the
+$\sim$1\kms\ level.  They can also be much larger: as an extreme
+example, internal dynamics and projection effects in the Hyades drive
+stars in the tidal tails span up to $\pm 40\mkms$ in relative velocity
+when projected against the cluster center \citep{jerabkova_800_2021}.
+The stars in such diffuse regions can be verified to be the same age
+as the core members ({\it i.e.}, coeval) through analyses of
+color-absolute magnitude diagrams \citep{KounkelCovey2019}, stellar
+rotation periods \citep{curtis_tess_2019,bouma_2021_ngc2516}, and
+chemical abundances \citep{hawkins_2020}.  While there are many
+implications for our understanding of star formation and cluster
+evolution \citep{dinnbier_tidal_2020}, a separate consequence is that
+we now know the ages of many more stars, including previously known
+planet hosts.
 
@@
   points are stars for which group membership is more secure than for
-  gray points.
-  Kepler-1627 is in the outskirts of the $\delta$ Lyr cluster
-  \citep{bouma_kep1627_2022}, which is centered at $\{ l, b\} \approx
-  \{ 66^\circ, 12^\circ\}$.
+  gray points.  Kepler-1627 is in the outskirts of the $\delta$ Lyr
+  cluster \citep{bouma_kep1627_2022}, which is centered at $\{ l, b\}
+  \approx \{ 66^\circ, 12^\circ\}$.
   {\it Middle row}: Galactic positions.  The Sun is at $\{X, Y, Z\} =
@@
 
-% {\bf TODO RONAN: PLEASE VET ENTIRE SECTION, AND UPDATE WHERE
-% APPROPRIATE!}.
-
 The possibility that the $\delta$~Lyr cluster, RSG-5, and the
@@
 
-The second step is to perform an unsupervised HDBScan
-clustering on the photometrically selected population
+The second step is to perform an unsupervised HDBScan clustering on
+the photometrically selected population
 \citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The
-parameters we use in the clustering are $\{ X, Y, Z, c v_b,
-c v_{l^*} \} $, where $c$ is the size-velocity corrective factor,
-which is taken as $c=6\,{\rm pc / km\,s}^{-1}$ to ensure that the
-spatial and velocity scales have identical standard deviations.
-Positions are computed assuming the \texttt{astropy v4.0} coordinate
-standard \citep{astropy_2018}, which places the Sun $8122$ pc from the
-galactic center, and assumes the solar velocity with respect to the
-local standard of rest from \citet{schonrich_local_2010}.  As input
+parameters we use in the clustering are $\{ X, Y, Z, c v_b, c v_{l^*}
+\} $, where $c$ is the size-velocity corrective factor, which is taken
+as $c=6\,{\rm pc / km\,s}^{-1}$ to ensure that the spatial and
+velocity scales have identical standard deviations.  Positions are
+computed assuming the \texttt{astropy v4.0} coordinate standard
+\citep{astropy_2018}, which places the Sun $8122$ pc from the galactic
+center, and assumes the solar velocity with respect to the local
+standard of rest from \citet{schonrich_local_2010}.  As input
 parameters to HDBScan, we set the minimum $\epsilon$ threshold past
-which clusters cannot be fragmented as $25$\,pc in physical space,
-and $c$\,\kms\ in velocity.  The minimum cluster size $N$ is set to 10,
-as is $k$, the parameter used to define the ``core distance'' density
+which clusters cannot be fragmented as $25$\,pc in physical space, and
+$c$\,\kms\ in velocity.  The minimum cluster size $N$ is set to 10, as
+is $k$, the parameter used to define the ``core distance'' density
 metric.
@@
 and overplot 4{,}763 objects with weights exceeding 0.10 as black
-points.\footnote{These counts only include objects with reliable astrometry
-	and photometry: $\varpi/\sigma_\varpi>5$;
-	$G/\sigma_{G}>50$;
-	$G_{\rm RP}/\sigma_{G_{RP}}>20$;
-	$G_{\rm BP}/\sigma_{G_{BP}}>20$.}
-The previously known $\delta$~Lyr\ cluster is visible at
-$(l,b)=(68^\circ,15^\circ$) and
-$(v_{l'}, v_b)=(-4.5 \mkms,-4 \mkms)$.  RSG-5 is visible at
-$(l,b)=(83^\circ,6^\circ)$,  $(v_{l'}, v_b)=(5.5 \mkms,-3.5 \mkms)$.
-Most of the other subclusters, including in Cep-Cyg
-($l,b=90^\circ,7^\circ$) and Cerberus ($l,b=48^\circ,18^\circ$) are
-too small or dispersed to have previously been analyzed in great
-detail.
+points.\footnote{These counts only include objects with reliable
+astrometry and photometry: $\varpi/\sigma_\varpi>5$;
+$G/\sigma_{G}>50$; $G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm
+BP}/\sigma_{G_{BP}}>20$.} The previously known $\delta$~Lyr\ cluster
+is visible at $(l,b)=(68^\circ,15^\circ$) and $(v_{l'}, v_b)=(-4.5
+\mkms,-4 \mkms)$.  RSG-5 is visible at $(l,b)=(83^\circ,6^\circ)$,
+$(v_{l'}, v_b)=(5.5 \mkms,-3.5 \mkms)$.  Most of the other
+subclusters, including in Cep-Cyg ($l,b=90^\circ,7^\circ$) and
+Cerberus ($l,b=48^\circ,18^\circ$) are too small or dispersed to have
+previously been analyzed in great detail.
 
@@
 correction to refine the target light curves.
-%Like for TESS, we also
-%calculated periods with Lomb-Scargle periodograms using an interactive
-%program that allows us to visually validate our measurements.
-%{\bf How many stars with periods from TESS? ZTF? Overlap?
-%Consistency? }
 
@@
 in the preceding numerator, are M-dwarfs with rotation periods between
-10 and 12.5 days.  As just discussed, the age interpretation for these M-dwarfs
-is not obvious.  Regardless, the $\approx$15\% false
-positive rate determined from the CAMD seems consistent with our
-fraction of detected rotation periods, given that RSG-5 was also
-missing rotation period detections for $\approx$15\% of its candidate
-members, which all seemed photometrically consistent with being part
-of a single pre-main-sequence locus.
-
-It is challenging to convert these stellar rotation periods
-to a precise age estimate, since on the pre-main-sequence
-the stars are spinning up due to thermal contraction
-rather than down due to magnetized braking.  Regardless, the rotation
-period distributions of both CH-2 and RSG-5 seem consistent with other
-30\,Myr to 50\,Myr clusters ({\it e.g.}, IC\,2602 and IC\,2391;
-\citealt{douglas_stephanie_t_2021_5131306}).
-They also seem consistent with the estimated false positive rates
-determined from the color-absolute magnitude diagrams.
+10 and 12.5 days.  As just discussed, the age interpretation for these
+M-dwarfs is not obvious.  Regardless, the $\approx$15\% false positive
+rate determined from the CAMD seems consistent with our fraction of
+detected rotation periods, given that RSG-5 was also missing rotation
+period detections for $\approx$15\% of its candidate members, which
+all seemed photometrically consistent with being part of a single
+pre-main-sequence locus.
+
+It is challenging to convert these stellar rotation periods to a
+precise age estimate, since on the pre-main-sequence the stars are
+spinning up due to thermal contraction rather than down due to
+magnetized braking.  Regardless, the rotation period distributions of
+both CH-2 and RSG-5 seem consistent with other 30\,Myr to 50\,Myr
+clusters ({\it e.g.}, IC\,2602 and IC\,2391;
+\citealt{douglas_stephanie_t_2021_5131306}).  They also seem
+consistent with the estimated false positive rates determined from the
+color-absolute magnitude diagrams.
 
@@
 
-Some salient stellar properties of the Kepler objects in Cep-Her can be gleaned
-from Figure~\ref{fig:age}.  They span spectral types of G8V
+Some salient stellar properties of the Kepler objects in Cep-Her can
+be gleaned from Figure~\ref{fig:age}.  They span spectral types of G8V
 (Kepler-1627) to K6V (KOI-7913 A).  The secondary in the KOI-7913
-system has spectral type $\approx$K8V.  And since a
-Solar-mass star with solar metallicity arrives at the zero-age main
-sequence at $t\approx40$ Myr \citep{choi_mesa_2016}, these
-stars are all in the late stages of their pre-main-sequence
-contraction.  Their heliocentric distances are $\approx$260\,pc
-(KOI-7913 and KOI-7368, in CH-2) and $\approx$340\,pc (Kepler-1643, in
-RSG-5).
+system has spectral type $\approx$K8V.  And since a Solar-mass star
+with solar metallicity arrives at the zero-age main sequence at
+$t\approx40$ Myr \citep{choi_mesa_2016}, these stars are all in the
+late stages of their pre-main-sequence contraction.  Their
+heliocentric distances are $\approx$260\,pc (KOI-7913 and KOI-7368, in
+CH-2) and $\approx$340\,pc (Kepler-1643, in RSG-5).
 
 The adopted stellar parameters are listed in
-Table~\ref{tab:sysparams}.  The stellar surface gravity, radius,
-mass, and density are found by interpolating against the MIST
-isochrones \citep{choi_mesa_2016}.  The statistical uncertainties from
-this technique mostly originate from the parallax uncertaintes; the
+Table~\ref{tab:sysparams}.  The stellar surface gravity, radius, mass,
+and density are found by interpolating against the MIST isochrones
+\citep{choi_mesa_2016}.  The statistical uncertainties from this
+technique mostly originate from the parallax uncertaintes; the
 systematic uncertainties are taken to be the absolute difference
@@
 For Kepler-1643, we acquired two iodine-free spectra from Keck/HIRES
-on the nights of 2020 Aug 16 and 2021 Oct 25.  The acquisition
-and analysis followed the usual techniques of the
-California Planet Survey \citep{howard_cps_2010}.  We derived the
-stellar parameters ($T_{\rm eff}, \log g, R_\star$) using
-\texttt{SpecMatch-Emp} \citep{yee_SM_2017}, which yielded values in
-$<1$-$\sigma$ agreement with those from the cluster-isochrone
-method.  This approach also yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.
-Using the broadened synthetic templates from \texttt{SpecMatch-Synth}
-\citep{petigura_cksi_2017}, we found $v\sin i = 9.3 \pm 1.0\,\mkms$.
-The systemic radial velocity at the two sequence epochs was $-9.1 \pm
-1.9\,\mkms$ and $-7.8\pm 1.2\,\mkms$ respectively.  To infer the
-equivalent width of the Li\,\textsc{I} 6708\,\AA\ doublet, we followed
-the procedure described by \citet{bouma_2021_ngc2516}.  This yielded a
-strong detection: ${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with
-values consistent at $<1$-$\sigma$ between the two epochs.   The
-quoted value does not correct for the Fe\,\textsc{I} blend at
-6707.44\AA, because HIRES resolves the two
-features.
-
-% {\bf TODO verify continuum normalization in Li EW measurement...}
-% Given the reddening-corrected color of the star this is
-% consistent(ish) with the Pleiades/NGC2516 based on my NGC2516 figure.
-% But it looks sub-IC2602 based on the lithium.png figure, which uses
-% the Randich2018 values.
-% It's pretty clearly super-field.
-% I really need to verify the continuum normalization.
-% COMPARE vs:
-% \citealt{randich_gaiaeso_2018}).  
-% \citet{berger_identifying_2018}
+on the nights of 2020 Aug 16 and 2021 Oct 25.  The acquisition and
+analysis followed the usual techniques of the California Planet Survey
+\citep{howard_cps_2010}.  We derived the stellar parameters ($T_{\rm
+eff}, \log g, R_\star$) using \texttt{SpecMatch-Emp}
+\citep{yee_SM_2017}, which yielded values in $<1$-$\sigma$ agreement
+with those from the cluster-isochrone method.  This approach also
+yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.  Using the broadened synthetic
+templates from \texttt{SpecMatch-Synth} \citep{petigura_cksi_2017}, we
+found $v\sin i = 9.3 \pm 1.0\,\mkms$.  The systemic radial velocity at
+the two sequence epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm
+1.2\,\mkms$ respectively.  To infer the equivalent width of the
+Li\,\textsc{I} 6708\,\AA\ doublet, we followed the procedure described
+by \citet{bouma_2021_ngc2516}.  This yielded a strong detection: ${\rm
+EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with values consistent at
+$<1$-$\sigma$ between the two epochs.   The quoted value does not
+correct for the Fe\,\textsc{I} blend at 6707.44\AA, because HIRES
+resolves the two features.
 
 \paragraph{High-Resolution Imaging}
-% KOI-6186 & 58662.62 &   Kp     &  4 &   80.00 &  4.1 &  4.8 &  5.4 &  5.8 &  6.5 &  6.8 &  7.6 &  8.3 &  8.2 &  8.3 &             Petigura \\ % N2.20190628.53389
 We acquired adaptive optics imaging of Kepler-1643 on the night of
-2019 June 28 using the NIRC2 imager on Keck-II.
-Using the narrow camera (FOV = 10.2\arcsec), we obtained 4 images in
-the $K'$ filter ($\lambda = 2.12\,\mu$m) with a total exposure time of
-320\,s. We analyzed these data following \citet{kraus_impact_2016},
-and determined the detection limits to visual companions by analyzing
-the residuals after subtracting an empirical PSF template. 
-This procedure yielded contrast limits of $\Delta K' = 4.1$ mag at
-$\rho = 150$ mas, $\Delta K' = 5.8$ mag at $\rho = 300$ mas, and
-$\Delta K' = 8.3$ mag at $\rho > 1000$ mas.
+2019 June 28 using the NIRC2 imager on Keck-II.  Using the narrow
+camera (FOV = 10.2\arcsec), we obtained 4 images in the $K'$ filter
+($\lambda = 2.12\,\mu$m) with a total exposure time of 320\,s. We
+analyzed these data following \citet{kraus_impact_2016}, and
+determined the detection limits to visual companions by analyzing the
+residuals after subtracting an empirical PSF template.  This procedure
+yielded contrast limits of $\Delta K' = 4.1$ mag at $\rho = 150$ mas,
+$\Delta K' = 5.8$ mag at $\rho = 300$ mas, and $\Delta K' = 8.3$ mag
+at $\rho > 1000$ mas.
 
@@
 spectrograph (TRES; \citealt{furesz_tres_2008}) mounted at the
-Tillinghast 1.5m at the Fred Lawrence Whipple Observatory.
-The stellar parameter classification pipeline for the
-TRES spectra has been described by \citet{2021tsc2.confE.124B}.  It is
-based on the synthetic template library originally constructed by
+Tillinghast 1.5m at the Fred Lawrence Whipple Observatory.  The
+stellar parameter classification pipeline for the TRES spectra has
+been described by \citet{2021tsc2.confE.124B}.  It is based on the
+synthetic template library originally constructed by
 \citet{buchhave_hatp16b_class_2010}.  The resulting stellar parameters
@@
 \paragraph{High-Resolution Imaging}
-% KOI-7368 & 58646.57 &   Kp     &  4 &   80.00 &  5.2 &  5.8 &  6.4 &  6.7 &  7.1 &  7.5 &  8.1 &  8.7 &  8.7 &  8.8 &                Huber \\ % N2.20190612.49290
 We acquired adaptive optics imaging of KOI-7368 on the night of 2019
-June 12, again using NIRC2.  The
-observational configuration and reduction was identical as for
-Kepler-1643.  No companions were detected, and the analysis of the
-image residuals yielded contrast limits of $\Delta K' = 5.2$ mag at
-$\rho = 150$ mas, $\Delta K' = 6.7$ mag at $\rho = 300$ mas, and
-$\Delta K' = 8.7$ mag at $\rho > 1000$ mas.
+June 12, again using NIRC2.  The observational configuration and
+reduction was identical as for Kepler-1643.  No companions were
+detected, and the analysis of the image residuals yielded contrast
+limits of $\Delta K' = 5.2$ mag at $\rho = 150$ mas, $\Delta K' = 6.7$
+mag at $\rho = 300$ mas, and $\Delta K' = 8.7$ mag at $\rho > 1000$
+mas.
 
@@
 \paragraph{Binarity}
-
-%(e.g., Jenkins KSCI-19081-002 KPDH pp 129, or many other pages)
-KOI-7913 is almost certainly a binary. %\footnote{Relevant identifiers for the
-%primary are KIC 8873450 = Gaia DR2/EDR3 2106235301785454208.  For the
-%secondary, they are KIC 8873448 = Gaia DR2/EDR3 2106235301785453824.}
-The north-west primary is $\approx$0.5 magnitudes brighter than the
-south-east secondary in optical passbands.  The two stars are
-separated in Gaia EDR3 by $3\farcs5$ on-sky, and have parallaxes
-consistent within $1$-$\sigma$ (with an average $\varpi=3.66 \pm
-0.01$\,mas).  The apparent on-sky separation is
-$959 \pm 2$ au.  
-The Gaia EDR3 proper motions are also nearly identical.
-%$(\mu_{\alpha'},\mu_\delta)=(2.439\pm0.015, 1.009\pm 0.016)\,{\rm
-%mas}\,{\rm yr}^{-1}$ for the primary, and
-%$(\mu_{\alpha'},\mu_\delta)=(2.696\pm0.020, 0.650 \pm 0.021) \,{\rm
-%mas}\,{\rm yr}^{-1}$ for the secondary.  
-Since two stars were resolved
-in the Kepler Input Catalog and are roughly one Kepler pixel apart, an
-accurate crowding metric has already been applied in the NASA Ames
-data products to correct the mean flux level
-\citep{2017ksci.rept....6M}.  This is important for deriving accurate
-transit depths.
+KOI-7913 is almost certainly a binary.  The north-west primary is
+$\approx$0.5 magnitudes brighter than the south-east secondary in
+optical passbands.  The two stars are separated in Gaia EDR3 by
+$3\farcs5$ on-sky, and have parallaxes consistent within $1$-$\sigma$
+(with an average $\varpi=3.66 \pm 0.01$\,mas).  The apparent on-sky
+separation is $959 \pm 2$ au.  The Gaia EDR3 proper motions are also
+nearly identical.  Since two stars were resolved in the Kepler Input
+Catalog and are roughly one Kepler pixel apart, an accurate crowding
+metric has already been applied in the NASA Ames data products to
+correct the mean flux level \citep{2017ksci.rept....6M}.  This is
+important for deriving accurate transit depths.
 
 \paragraph{Spectra}
-
 We acquired Keck/HIRES spectra for KOI-7913 A on the night of 2021 Nov
@@
 -0.06 \pm 0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm
-sys} = -17.8 \pm 1.1\,\mkms$.  For the secondary, these same parameters
-were $[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm 1.0\,\mkms$,
-and ${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.
-Neither of the KOI-7913 components shows strong lithium absorption,
-but this is expected given their $\approx$K6V and $\approx$K8V
-spectral types.
-They do however both show H$\alpha$ in emission.
-We discuss the significant in greater depth in Appendix~\ref{app:spectra},
-but broadly speaking both observations are consistent with
-a $\approx$40 Myr age for KOI-7913.
+sys} = -17.8 \pm 1.1\,\mkms$.  For the secondary, these same
+parameters were $[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm
+1.0\,\mkms$, and ${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.  Neither
+of the KOI-7913 components shows strong lithium absorption, but this
+is expected given their $\approx$K6V and $\approx$K8V spectral types.
+They do however both show H$\alpha$ in emission.  We discuss the
+significant in greater depth in Appendix~\ref{app:spectra}, but
+broadly speaking both observations are consistent with a $\approx$40
+Myr age for KOI-7913.
 
 \paragraph{High-Resolution Imaging}
-
 We acquired adaptive optics imaging of KOI-7913 on the night of 2020
-Aug 27 using the NIRC2 imager.  The
-observational configuration and reduction was identical as before.
-The images showed KOI-7913 A, KOI-7913 B, and an additional faint star
-$\approx$$0\farcs99$ due East of KOI-7913 B.  Applying the PSF-fitting
-routines from \citet{kraus_impact_2016}, the tertiary star has a
-separation $\rho = 4397 \pm 3$\,mas from the primary, at a position
-angle $231.17^\circ \pm 0.02^\circ$, with a $\Delta K' = 6.97 \pm
-0.04$.  While it is too faint to affect the interpretation of the KOI
-itself, it would be amusing if the faint neighbor were also comoving
-and young -- at the age of the system, it would have a mass roughly
-between 10 and 15\,M$_{\rm Jup}$.  Additional imaging epochs will 
-tell.
+Aug 27 using the NIRC2 imager.  The observational configuration and
+reduction was identical as before.  The images showed KOI-7913 A,
+KOI-7913 B, and an additional faint star $\approx$$0\farcs99$ due East
+of KOI-7913 B.  Applying the PSF-fitting routines from
+\citet{kraus_impact_2016}, the tertiary star has a separation $\rho =
+4397 \pm 3$\,mas from the primary, at a position angle $231.17^\circ
+\pm 0.02^\circ$, with a $\Delta K' = 6.97 \pm 0.04$.  While it is too
+faint to affect the interpretation of the KOI itself, it would be
+amusing if the faint neighbor were also comoving and young -- at the
+age of the system, it would have a mass roughly between 10 and
+15\,M$_{\rm Jup}$.  Additional imaging epochs will tell.
 
@@
 KOI-7368 at a 30-minute cadence between May 2009 and April 2013.  For
-all three systems quarters 1 through 17 were observed with minimal  data
-gaps.  The top panel of
-Figure~\ref{fig:planets} shows a 50-day slice of the \texttt{PDCSAP}
-light curves for the three new Cep-Her candidates, along with
-Kepler-1627.  In \texttt{PDCSAP}, nonastrophysical variability is
-removed through a cotrending approach that uses a set of basis vectors
-derived by applying singular value decomposition to a set of
-systematics-dominated light curves
+all three systems quarters 1 through 17 were observed with minimal
+data gaps.  The top panel of Figure~\ref{fig:planets} shows a 50-day
+slice of the \texttt{PDCSAP} light curves for the three new Cep-Her
+candidates, along with Kepler-1627.  In \texttt{PDCSAP},
+nonastrophysical variability is removed through a cotrending approach
+that uses a set of basis vectors derived by applying singular value
+decomposition to a set of systematics-dominated light curves
 \citep{smith_kepler_PDC_2017}.  In our analysis, we used the
@@
 signals are much larger than the transits, which have depth
-$\approx$0.1\%.
-To quantify the stellar rotation periods, we calculated the
-Lomb-Scargle periodogram for each Kepler quarter independently.  The
-resulting means and standard deviations are in
+$\approx$0.1\%.  To quantify the stellar rotation periods, we
+calculated the Lomb-Scargle periodogram for each Kepler quarter
+independently.  The resulting means and standard deviations are in
 Table~\ref{tab:sysparams}.
@@
 polynomial, which was then divided out from the light curve.  The
-resulting flattened transits were then fitted with a 
-planetary transit under the assumption of quadratic limb darkening.
-Our model therefore included 8 free parameters for the transit ($\{P,
-t_0, \log R_{\rm p}/R_\star, b, u_1 ,u_2, R_\star, \log g\}$), 2 free
-parameters for the light curve normalization and a white noise jitter
-($\{\langle f \rangle, \sigma_f \}$), and an additional 5 implicit
-parameters per transit through the polynomial.
+resulting flattened transits were then fitted with a planetary transit
+under the assumption of quadratic limb darkening.  Our model therefore
+included 8 free parameters for the transit ($\{P, t_0, \log R_{\rm
+p}/R_\star, b, u_1 ,u_2, R_\star, \log g\}$), 2 free parameters for
+the light curve normalization and a white noise jitter ($\{\langle f
+\rangle, \sigma_f \}$), and an additional 5 implicit parameters per
+transit through the polynomial.
 
@@
 of the hyperparameter priors, otherwise the GP would tend to
-incorporate variability that was not part of the transit, even if it was
-short-term variability attributable to flares, or simply
+incorporate variability that was not part of the transit, even if it
+was short-term variability attributable to flares, or simply
 instrumental noise.  This failure mode is somewhat pernicious, in that
@@
 \paragraph{Kepler-1643}
-  %Based on the Q1-Q17 DR25 False Positive Probability table hosted at
-%the NASA Exoplanet Archive, 
 Kepler-1643~b (KOI-6186.01) was already validated as a transiting
@@
 of the transiting source was measured as $1\farcs0$ at 4.4-$\sigma$.
-The reports show however that two outlying quarters (2 and 6) 
-drive the offset -- the centroid locations from the other Kepler
-quarters are consistent at $\lesssim 0\farcs4$.  Nonetheless, this
-is an instructive exercise in how stellar variability complicates
+The reports show however that two outlying quarters (2 and 6) drive
+the offset -- the centroid locations from the other Kepler quarters
+are consistent at $\lesssim 0\farcs4$.  Nonetheless, this is an
+instructive exercise in how stellar variability complicates
 centroid-based vetting tests.  The centroid shifts measured by these
 tests are determined from the in- and out-of-transit flux-weighted
-centroids.  For stars with significant spot-induced
-variability there is no static baseline in either the in- or
-out-of-transit phases.
+centroids.  For stars with significant spot-induced variability there
+is no static baseline in either the in- or out-of-transit phases.
 
@@
 probability for the system because their default trapezoidal fitting
-routine failed, presumably due to the spot-induced variability.
-%The Q1-Q17 DR25 False Positive Probability table however lists its
-%probability of being an astrophysical false positive as ``1.0''. Given
-%that none of the other columns in the table are populated, this seems
-%likely to be a database entry error ({\bf ask Jessie}). 
-Our fitting approach rectifies this point, and our new NIRC2 images
+routine failed, presumably due to the spot-induced variability.  Our
+fitting approach rectifies this point, and our new NIRC2 images
 revealed no new stellar companions.  Performing the validation
@@
 systematic noise in the light curve (see Figure~\ref{fig:planets}).
-The positional probability score also meets the requirements
-for transiting sources thought to share positions with their
-target stars \citep{morton_false_2016,2017ksci.rept...16B}. 
-%A visual inspection of
-%each individual transit also shows a weak notch at the predicted
-%transit times; over the 198 observed transits, the resulting signal
-%seems sensible.
+The positional probability score also meets the requirements for
+transiting sources thought to share positions with their target stars
+\citep{morton_false_2016,2017ksci.rept...16B}.
 
@@
 Kepler-1643.  For KOI-7368, the reported offset is smaller, and less
-statistically significant ($0\farcs2$; 3.0-$\sigma$).  Again,
-the data validation reports show that the shift is caused by a few
-outlying quarters (4, 5, 8, and 12).  The fact that the remaining data
-show a consistent scatter between centroid locations suggests that
-these outliers are also caused by the stellar variability.  Independently, our
-imaging shows that there are no neighboring sources that could
+statistically significant ($0\farcs2$; 3.0-$\sigma$).  Again, the data
+validation reports show that the shift is caused by a few outlying
+quarters (4, 5, 8, and 12).  The fact that the remaining data show a
+consistent scatter between centroid locations suggests that these
+outliers are also caused by the stellar variability.  Independently,
+our imaging shows that there are no neighboring sources that could
 cause an offset of the observed amplitude.
@@
 KOI-7913.01 is also listed on the NASA Exoplanet Archive as a
-``candidate'' planet.  The canonical \citet{morton_false_2016} analysis
-was of Q1-Q17 KOIs from DR24, and therefore spanned KOI-1.01 to
-KOI-7620.01 and omitted KOI-7913.01.  However the results of the
+``candidate'' planet.  The canonical \citet{morton_false_2016}
+analysis was of Q1-Q17 KOIs from DR24, and therefore spanned KOI-1.01
+to KOI-7620.01 and omitted KOI-7913.01.  However the results of the
 subsequent DR25 analysis by Morton et al.\ are listed at the NASA
@@
 to ``candidate''.  The most likely explanation is the presence of
-KOI-7913~B, which is located approximately one Kepler pixel away from Kepler-7913 A (see
-Appendix~\ref{app:koi7913}).  Our pixel-level analysis of the target
-pixel files did not show any evidence to support the idea that
-KOI-7913~B could host the transiting planet.  Our NIRC2 imaging
-similarly ruled out the possibility of any other plausible host star
-for the planet other than KOI-7913~A.
+KOI-7913~B, which is located approximately one Kepler pixel away from
+Kepler-7913 A (see Appendix~\ref{app:koi7913}).  Our pixel-level
+analysis of the target pixel files did not show any evidence to
+support the idea that KOI-7913~B could host the transiting planet.
+Our NIRC2 imaging similarly ruled out the possibility of any other
+plausible host star for the planet other than KOI-7913~A.
 

### 2022-04-11 papers/ms.tex
update ms

@@
 They do however both show H$\alpha$ in emission.  We discuss the
-significant in greater depth in Appendix~\ref{app:spectra}, but
+significance in greater depth in Appendix~\ref{app:spectra}, but
 broadly speaking both observations are consistent with a $\approx$40
@@
 Aug 27 using the NIRC2 imager.  The observational configuration and
-reduction was identical as before.  The images showed KOI-7913 A,
+reduction were identical as before.  The images showed KOI-7913 A,
 KOI-7913 B, and an additional faint star $\approx$$0\farcs99$ due East
@@
 4397 \pm 3$\,mas from the primary, at a position angle $231.17^\circ
-\pm 0.02^\circ$, with a $\Delta K' = 6.97 \pm 0.04$.  While it is too
+\pm 0.02^\circ$, with $\Delta K' = 6.97 \pm 0.04$.  While it is too
 faint to affect the interpretation of the KOI itself, it would be
@@
 also computationally efficient, and it captures the planetary
-parameters that we actually care about.
+parameters about which we care the most.
 
@@
 most likely path would be through transit observations using a
-high-resolution spectrograph \citep[see][]{bouma_kep1627_2022}.  In
+high-resolution spectrograph.  In
 the interim, it is of interest whether the transit signals could be
@@
 contrast curves.  Figure~\ref{fig:planets} shows the justification:
-the transit is flat and has a high S/N ($\approx 47$), which is
-difficult to reproduce with eclipsing binary models.
+the transit is flat and has a high S/N ($\approx 47$).  The shape is
+therefore difficult to reproduce with eclipsing binary models.
 
@@
 fitting approach rectifies this point, and our new NIRC2 images
-revealed no new stellar companions.  Performing the validation
+revealed no new stellar companions.  Performing the relevant
 calculation, we find ${\rm FPP} \approx 4\times10^{-3}$.  Though not
@@
 transiting sources thought to share positions with their target stars
-\citep{morton_false_2016,2017ksci.rept...16B}.
+\citep{2017ksci.rept...16B}.
 
@@
 outliers are also caused by the stellar variability.  Independently,
-our imaging shows that there are no neighboring sources that could
-cause an offset of the observed amplitude.
+our NIRC2 imaging shows that there are no neighboring sources that
+could cause an offset of the observed amplitude.
 
 \paragraph{KOI-7913}
-KOI-7913.01 is also listed on the NASA Exoplanet Archive as a
+KOI-7913.01 is also currently listed on the NASA Exoplanet Archive as a
 ``candidate'' planet.  The canonical \citet{morton_false_2016}
@@
 \subsection{Normal-Sized Mini-Neptunes Exist at 40$\,$Myr}
-
-Current thinking about the origins and evolution of lose-in
-mini-Neptunes has recently been reviewed by \citet{bean_nature_2021}.
-In the context of the existing exoplanet population, the most
-significant novelty about Kepler-1643~b, KOI-7368.01, and KOI-7913 is
-that they have normal sizes.
-More specifically, the size distribution of mini-Neptunes spans
-$\approx$1.8\,$R_\oplus$ to $\approx$3.6\,$R_\oplus$, with a peak at
-$\approx 2.4\,R_\oplus$ \citep[][Figure~7]{Fulton_et_al_2017}.
-Figure~\ref{fig:rp_period_age} shows the sizes, orbital periods, and ages of
-known transiting planets, emphasizing those with precise ages.
-Considering planets younger than $10^8$ years, the smallest previously
-known were AU~Mic~c ($\approx3.0\pm0.2\,R_\oplus$, averaging the
-results of \citealt{martioli_aumicbc_2021} and
-\citealt{gilbert_flares_2022}), Kepler-1627~Ab ($3.8\pm0.2\,R_\oplus$;
-\citealt{bouma_kep1627_2022}), and AU~Mic~d ($4.2\pm0.2\,R_\oplus$;
-\citealt{plavchan_planet_2020}).
-The other transiting planets younger than $10^8$ years are much
-larger, with sizes between 4 and 10$\,R_\oplus$
-\citep{Mann_K2_33b_2016,David_et_al_2017,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020}.
-
-%TODO: I think we might need James Owen to contribute something
-%similar to the TOI-1227 models?
-{\bf Run some evolutionary models, make some comments about the core
-	mass, KH timescale, etc}.
-
-\subsection{Is CH-2 really a star cluster?}
-
-RSG-5, and Kepler-1643's membership inside it, clearly meet the
-typical expectations of a star claimed to be in an open cluster.
-RSG-5 shows an obvious overdensity relative to the local field
-population ({\it e.g.}, Figure~\ref{fig:XYZvtang}), and our membership
-selection easily produces a clean pre-main-sequence locus in 
-color-absolute magnitude space (Figure~\ref{fig:age}).
-CH-2, and KOI-7913 and KOI-7368's membership inside it, do not meet
-those expectations in as obvious a manner.  This is because this
-association of stars is diffuse.
-
-%really 80:9
-To quantify the density discrepancy, we can compare the spatial and
-velocity volumes searched to select candidate members of each cluster.
+\label{subsec:sizes}
+
+The most significant novelty about
+Kepler-1643~b, KOI-7368.01, and KOI-7913~Ab is that their sizes are
+normal, relative to the known population of gigayear-old
+mini-Neptunes.  At field star ages, mini-Neptune sizes span
+1.8\,$R_\oplus$ to 3.6\,$R_\oplus$, with the most
+common size being $\approx 2.4\,R_\oplus$
+\citep[][Figure~7]{Fulton_et_al_2017}.  The known planets younger than
+$10^8$ years are almost all larger, with sizes between 4 and
+10$\,R_\oplus$
+\citep{Mann_K2_33b_2016,David_et_al_2017,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020,mann_toi1227_2022}.
+Figure~\ref{fig:rp_period_age} explores this in more detail by showing
+the sizes, orbital periods, and ages of the known transiting planets,
+emphasizing planets with precise ages.  The smallest previously known planets
+comparable to the new Cep-Her mini-Neptunes
+are AU~Mic~c ($3.0\pm0.2\,R_\oplus$, see
+\citealt{martioli_aumicbc_2021} and \citealt{gilbert_flares_2022}),
+Kepler-1627~Ab ($3.8\pm0.2\,R_\oplus$; \citealt{bouma_kep1627_2022}),
+and AU~Mic~d ($4.2\pm0.2\,R_\oplus$; \citealt{plavchan_planet_2020}).
+
+The theoretical expectation is that mini-Neptunes with sizes of 2 to
+3\,$R_\oplus$ should be common at ages of $10^7$ to $10^8$ years.
+This expectation is tied to inferences about the
+initial distributions of planetary core mass, core composition, and
+atmospheric mass fraction \citep{owen_evaporation_2017}.  The
+Kelvin-Helmholtz cooling timescale, which is tied to the entropy of
+the planetary interior once the disk disperses, could also play a
+significant role \citep{owen_constraining_2020}.  As an example,
+\citet{rogers_unveiling_2021} predicted that given a core-mass
+distribution peaked at $\approx$$4$\,M$_\oplus$, an ice-poor rock/iron
+core composition, and a typical H/He mass fraction of $\approx$4\%,
+there should be a maximum in planet occurrence at 2 to 3\,$R_\oplus$,
+at times between 10 and 100\,Myr
+(their Figure~16).  The models advanced by
+\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} agree;
+the theoretical differences lie in their explanations
+for the origin of the radius valley.  If the radius gap is primordial
+and created by gas accretion physics, then super-Earths
+($<$$1.8$\,$R_\oplus$) would be common as early as $10^7$ years
+\citep{lee_primordial_2021}.   The photoevaporation-driven models more
+often quote $10^8$ year timescales for production of
+atmosphere-stripped super-Earths, while the models with the mass-loss
+driven by a slow escape of heat from the core predict longer
+$\sim$$10^9$ year timescales (but see \citealt{king_euv_2020}).
+
+Systems such as K2-25, V1298~Tau, HIP-67522, TOI-837,
+and TOI-1227 have sizes that are much larger than the picture we have
+just described.  However, their large sizes can be accommodated
+by existing models by invoking any of {\it i)} larger core masses,
+{\it ii)} more volatile-rich compositions, {\it iii)}  larger initial
+atmospheric mass fractions, or {\it iv)} longer thermal cooling times.
+For instance, perhaps all five of the the aforementioned systems are
+large simply because the planets are more massive than most
+mini-Neptunes.  Secure mass measurements would help constrain this
+parameter space, but the $\sim$1\,\kms\ spot-induced radial velocity
+semi-amplitudes makes measuring the Doppler orbits incredibly
+challenging.  \citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.
+Regardless, the new Kepler-1643, KOI-7368, and KOI-7913 systems do
+demonstrate that at least some planets at 40\,Myr have sizes that are
+consistent with theoretical expectations for mini-Neptunes.  While
+selection effects imposed by spot-induced photometric variability are
+a likely explanation for why planets this small have not previously
+been identified \citep[{\it e.g.},][]{zhou_2021_tois}, future work
+should quantify this bias more carefully, in order to enable empirical
+studies of how the planetary size distribution changes at early times.
+
+\subsection{Is CH-2 a real cluster?}
+\label{subsec:ch2}
+
+RSG-5, and Kepler-1643's membership inside it, meet typical
+expectations for a star claimed to be in an open cluster.  RSG-5 is an
+obvious overdensity relative to the local field, and our
+membership selection easily produces a clean pre-main-sequence locus
+in color versus absolute magnitude (Figure~\ref{fig:age}).  CH-2, and
+KOI-7913 and KOI-7368's membership inside it, do not meet these
+expectations in as obvious a manner.  This is because the CH-2
+association is diffuse.
+
+To quantify the density difference between CH-2 and RSG-5, we can compare the spatial and
+velocity volumes searched for each group.
 For RSG-5, we drew 141 candidate members from a $30\,{\rm
-pc}\times30\,{\rm pc}\times30\,{\rm pc}$ spatial cube, given a $1 {\rm \mkms} \times 2 {\rm \mkms }$ rectangle in apparent galactic velocity.
-For CH-2, our 37 candidate members came from a spatial cube of
-dimension $50\,{\rm pc}\times40\,{\rm pc}\times30\,{\rm pc}$, and the
-velocity rectangle of $2 {\rm \mkms} \times 4 {\rm \mkms}$.  If we
-define the ``searched volume'' in units of ${\rm pc}^3 {\rm
-(\mkms)}^2$, then the volume ratio of CH-2:RSG-5 is $\approx$9:1.  The
-density (number of stars per unit searched volume) within RSG-5 relative to
-CH-2 similarly comes out to 34 to 1.
-
-So what makes a star cluster?  Historic answers to this question have
-recently by reviewed by \citet{krumholz_star_2019}: some definitions
-that have been offered include criteria such as being gravitationally
-bound, and having a mass density that significantly exceeds the mean
-in a cluster's galactic neighborhood.  We prefer a modified version of
-the definition adopted by \citet{krumholz_star_2019}: for our purposes
-a star cluster is a group of at least 12 stars that was physically
-associated at its time of formation.  The somewhat arbitrary ``12'' is
-set to distinguish clusters from high-order multiple star systems.  We
-therefore explicitly include dissolved clusters and their tidal tails
-in our concept of clusters.  We also explicitly exclude the idea that
-a particular number of stars per unit spatial volume is required to
-define a cluster.  The latter point acknowledges the fact that an
-important factor in cluster identification is now also the number per
-unit velocity volume, whether in 2-dimensional tangential velocity, or
-when including the third radial component.  Perhaps
-once stellar rotation periods and chemical abundances reach the same
-level of ubiquity as stellar proper motions, they might enable further
-refinement of our ability in cluster discovery.
-
-From a data-driven perspective, how do we demonstrate that a star is
-in a cluster, {\it i.e.}, that it is part of a group of stars that was
-physically associated at its time of formation?
-Back-integrating the orbits is one convincing approach, but it does
-not always work {\bf (CITE)}.
-The relatively minimal approach suggested by \citet{tofflemire_2021}
-is intriguing: search for coeval, phase-space neighbors, measure their
-ages, and determine if they share a common age.
-This approach is more accurately be described as a method for
-determining whether a star is currently associated with a set of coeval
-stars, which is much easier to determine than what the association
-looked like in the past.
-From this standard, our analysis thus far of CH-2 has already
-demonstrated the existence of such an assocation.
-
-A crucial logical step in this method however is to ensure that the
-(automated) search process for coeval phase-space neighbors in fact
-produces neighbors at a rate different from what it would for field
-stars.
-
-% TODO
-And so we ask: what is the density of field stars in the CH-2 region?
-If we had applied Ronan's pre-main-sequence selection, + HDBScan, +
-local neighbors... what would we have gotten?
-Similarly... if you just applied Adam Kraus' Comove 
-What is the likelihood we are fooling ourselves?
-
-\subsection{Further Work on Cep-Her}
-
-Disentangling its kinematics
+pc}\times30\,{\rm pc}\times30\,{\rm pc}$ spatial cube, given a $1 {\rm
+\mkms} \times 2 {\rm \mkms }$ rectangle in apparent galactic velocity.
+For CH-2, our 37 candidate members came from a rectangular prism of
+dimension $50\,{\rm pc}\times40\,{\rm pc}\times30\,{\rm pc}$, and a
+rectangular box of $2 {\rm \mkms} \times 4 {\rm \mkms}$.  If we define
+the searched volume in units of ${\rm pc}^3\,{\rm km^2}\,{\rm s^{-2}}$, then
+the volume ratio of CH-2 to RSG-5 is 9 to 1.  The number of candidate
+members per unit searched volume in RSG-5 relative to CH-2 is
+therefore 34 to 1.
+
+Given its low density, in CH-2 really a star cluster?  For this
+discussion, let us adopt the definition that a star cluster is a group
+of at least 12 stars that was physically associated at its time of
+formation.  The somewhat arbitrary ``12'' is set to distinguish star
+clusters from high-order multiples \citep[see][]{krumholz_star_2019}.
+Dissolved clusters and their tidal tails are therefore included in our
+concept of clusters.  We do not require however that a particular number of stars per
+unit spatial volume is required for a group to be a ``cluster''.
+The latter point acknowledges that an important
+factor in cluster identification is also the density in velocity
+space.  Perhaps once stellar rotation periods and chemical abundances
+reach the same level of ubiquity as stellar proper motions, they might
+enable further refinement of our ability to discover stars that formed
+together.
+
+From a data-driven perspective, demonstrating that a group of stars
+was physically associated at its time of formation is challenging.
+While some young groups show kinematic evidence for expansion
+\citep{kuhn_kinematics_2019}, many, including Sco-Cen, do not
+\citep{wright_kinematics_2018}.  This complicates the feasibility of
+deriving kinematic ages through traceback, as
+well as through the expansion itself
+\citep[see][]{crundall_chronostar_2019}.  A more minimal approach is
+that suggested by \citet{tofflemire_2021}: search for coeval,
+phase-space neighbors, measure their ages, and determine if they share
+a common age.  This approach can demonstrate whether a star is {\it
+currently associated} with a set of coeval stars, which is much easier
+than determining what the association looked like in the past.  Our
+analysis of CH-2 meets the latter standard for the existence of a
+$\approx$40\,Myr assocation today.  It would be a worthy exercise for
+future work to perform a similar search for coeval phase-space
+neighbors on the entire dataset of known exoplanet hosts.  For the
+time being, we can offer the anecdotal point that in our experience,
+most stars do not have dozens of $\sim$40\,Myr neighbors within a
+local volume of a few \kms\ and tens of parsecs.
+
+\subsection{Future work}
+
+\subsubsection{Cep-Her}
+\paragraph{Kinematics}
 * Need an RV survey.
 
+\paragraph{Age(s)}
 Benchmarking its age(s):
@@
 
-Finding more interesting objects in it:
+\paragraph{More objects}
+
+* TESS data: giant planets in Cep-Her.  We need more 1e7-1e8 year old
+stars for the HJ occurrence versus time question.  This will help.
+
+* Kepler data: exocomets?  Search for smaller planets?
+
+\subsubsection{Occ rates}
 * Kepler data... presumably $\sim$200 stars in the cluster were
 observed, in order to have detected 4 planets. 
-
-* TESS data
+Occurrence rates?
 

### 2022-04-12 papers/ms.tex
update ms

@@
 
-Our adopted approach was as follows.  Given an initial guess for the
-transit ephemeris from the \texttt{q1\_q17\_dr25} analysis
-\citep{thompson_planetary_2018}, the light curve was first trimmed to
-a local window around each transit, spanning $\pm 3\times t_{14}$.
-The local out-of-transit points were then fitted with a fourth-order
-polynomial, which was then divided out from the light curve.  The
-resulting flattened transits were then fitted with a planetary transit
-under the assumption of quadratic limb darkening.  Our model therefore
-included 8 free parameters for the transit ($\{P, t_0, \log R_{\rm
-p}/R_\star, b, u_1 ,u_2, R_\star, \log g\}$), 2 free parameters for
-the light curve normalization and a white noise jitter ($\{\langle f
-\rangle, \sigma_f \}$), and an additional 5 implicit parameters per
-transit through the polynomial.
+We fitted the data as follows.  Given an initial guess for the transit
+ephemeris from \citet{thompson_planetary_2018}, we first trimmed the
+light curve to a local window around each transit that spanned three
+transit durations before and after each mid-transit time.  The
+out-of-transit points in each local window were then fitted with a
+fourth-order polynomial, which was then divided out from the light
+curve.  The resulting flattened transits were then fitted with a
+planetary transit under the assumption of quadratic limb darkening.
+Our model therefore included 8 free parameters for the transit ($\{P,
+t_0, \log R_{\rm p}/R_\star, b, u_1 ,u_2, R_\star, \log g\}$), 2 free
+parameters for the light curve normalization and a white noise jitter
+($\{\langle f \rangle, \sigma_f \}$), and an additional 5 implicit
+parameters for each transit.
 
@@
 sampled using \texttt{PyMC3}'s No-U-Turn Sampler
-\citep{hoffman_no-u-turn_2014}, after having initialized using the
+\citep{hoffman_no-u-turn_2014}, after having initialized to the the
 maximum {\it a posteriori} (MAP) model.  We used the
@@
 
-One weakness of our fitting approach is that we have fixed 5 implicit
+One weakness of our approach is that we have fixed 5 implicit
 parameters per individual transit to their MAP values to remove the
@@
 incorporate variability that was not part of the transit, even if it
-was short-term variability attributable to flares, or simply
-instrumental noise.  This failure mode is somewhat pernicious, in that
-it yields an ill-founded sense of confidence in an ``overly clean''
-transit model fit.  While our model is simpler, it has the benefit
-that the white noise jitter never trades off with any parameter
-equivalent to a damping timescale for the coherence of the GP.  It is
-also computationally efficient, and it captures the planetary
-parameters about which we care the most.
+was short-term variability attributable to flares or instrumental
+noise.  This failure mode is pernicious, in that it yields an
+ill-founded sense of confidence in an ``overly clean'' transit model
+fit.  Our model is simpler, and it has the benefit that the white
+noise jitter never trades off with any parameter equivalent to a
+damping timescale for the coherence of the GP.  It is also
+computationally efficient, and it captures the planetary parameters
+about which we care the most.
 
@@
 In the future, it may be possible to obtain independent evidence for
-the Cep-Her planets to directly confirm their planetary nature.  The
-most likely path would be through transit observations using a
-high-resolution spectrograph.  In
-the interim, it is of interest whether the transit signals could be
-astrophysical false positives, or whether they can be argued to be
-planetary on a statistical basis.  We adopt the Bayesian framework
-implemented in \texttt{VESPA} for this purpose
-\citep{morton_efficient_2012,vespa_2015}.  Briefly summarized, the
-priors in \texttt{VESPA} take as given the binary star occurrence rate
-from \citet{raghavan_survey_2010}, direction-specific star counts from
-\citet{girardi_star_2005}, and planet occurrence rates as described by
-\citet[][Section~3.4]{morton_efficient_2012}.  The likelihoods are
-then evaluated by forward-modeling a synthetic population of eclipsing
-bodies for each possible astrophysical model class, in which each
-population member has a particular trapezoidal eclipse depth, total
-duration, and ingress duration.  These summary statistics are then
-compared against the actual photometric data to evaluate the
+the Cep-Her planets to directly confirm their planetary nature, for
+instance through transit observations using a high-resolution
+spectrograph.  For now, it is of interest whether the transit signals
+might be astrophysical false positives, or whether they can be argued
+to be planetary on a statistical basis.  We adopt the Bayesian
+framework implemented in \texttt{VESPA} to assess the relevant
+probabilities \citep{morton_efficient_2012,vespa_2015}.  Briefly
+summarized, the priors in \texttt{VESPA} assume the binary star
+occurrence rate from \citet{raghavan_survey_2010}, direction-specific
+star counts from \citet{girardi_star_2005}, and planet occurrence
+rates as described by \citet[][Section~3.4]{morton_efficient_2012}.
+The likelihoods are then evaluated by forward-modeling a synthetic
+population of eclipsing bodies for each astrophysical model class, in
+which each population member has a known trapezoidal eclipse depth,
+total duration, and ingress duration.  These summary statistics are
+then compared against the actual photometric data to evaluate the
 probabilities of false positive scenarios such as foreground eclipsing
@@
 and the new NIRC2 imaging constraints, we find ${\rm FPP} =
-6\times10^{-9}$, where the improvement is due to the new NIRC2
-contrast curves.  Figure~\ref{fig:planets} shows the justification:
+6\times10^{-9}$.  Figure~\ref{fig:planets} shows the justification:
 the transit is flat and has a high S/N ($\approx 47$).  The shape is
-therefore difficult to reproduce with eclipsing binary models.
-
-Intriguingly, Kepler-1643 formally failed one of the centroid shift
-tests (see the \texttt{q1\_q17\_dr25\_koi} data release): the angular
-distance between the target star's catalog position and the position
-of the transiting source was measured as $1\farcs0$ at 4.4-$\sigma$.
-The reports show however that two outlying quarters (2 and 6) drive
-the offset -- the centroid locations from the other Kepler quarters
-are consistent at $\lesssim 0\farcs4$.  Nonetheless, this is an
-instructive exercise in how stellar variability complicates
-centroid-based vetting tests.  The centroid shifts measured by these
-tests are determined from the in- and out-of-transit flux-weighted
-centroids.  For stars with significant spot-induced variability there
-is no static baseline in either the in- or out-of-transit phases.
+therefore nearly impossible to reproduce with eclipsing binary models.
+
+Intriguingly, Kepler-1643 formally failed one of the data validation
+centroid shift tests (see the \texttt{q1\_q17\_dr25\_koi} data
+release): the angular distance between the target star's catalog
+position and the position of the transiting source was measured as
+$1\farcs0$ at 4.4-$\sigma$.  The reports show however that two
+outlying quarters (2 and 6) drive the offset -- the centroid locations
+from the other Kepler quarters are consistent at $\lesssim 0\farcs4$.
+Nonetheless, this is an instructive exercise in how stellar
+variability complicates centroid-based vetting tests.  The centroid
+shifts measured by these tests are determined from the in- and
+out-of-transit flux-weighted centroids.  For stars with significant
+spot-induced variability there is no static baseline in either the in-
+or out-of-transit phases.
 

### 2022-04-12 papers/ms.tex
update ms

@@
   Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
-  The latter Neptune-sized planet is in a component of the Cep-Her
+  The latter Neptune-sized planet is in part of the Cep-Her
   complex called the $\delta$\ Lyr\ cluster
@@
   kinematic evidence from Gaia, stellar rotation periods from TESS,
-  and spectroscopy, these three systems are also $\approx$40 million
+  and spectroscopy, these three objects are also $\approx$40 million
   years (Myr) old.  More specifically, we find that Kepler-1643 is
@@
   respectively.  These planets empirically demonstrate that
-  mini-Neptunes with sizes of $\approx$2 Earth radii exist at ages of
-  $\approx$40 million years.
+  mini-Neptunes with sizes of $\sim$2 Earth radii exist at ages of
+  40 million years.
 \end{abstract}
@@
 The reason is that the properties of young planets provide benchmarks
-for studies of planetary evolution.  For instance, there are the
-questions of when hot Jupiters arrive on their close-in orbits
+for studies of planetary evolution.  For instance, young planets can
+help inform our understanding
+of when hot Jupiters arrive on their close-in orbits
 \citep{dawson_johnson_2018}, how the sizes of planets with massive
@@
 the planet must exist, and its age must be secured.  Spaced-based
-photometry from K2 and TESS has yielded a number of young planets for
+photometry from K2 and TESS has yielded a number of exemplars for
 which the planetary evidence comes from transits, and the age is based
@@
 analyses have mostly clustered on the stellar positions and 2-D
-velocities measured by Gaia.  One important result is the discovery of
+velocities.  One important result is the discovery of
 diffuse streams and tidal tails comparable in stellar mass to the
@@
 \citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  Even
-though these streams can be spread out over tens to hundreds of
+though these streams are spread over tens to hundreds of
 parsecs, their velocity dispersions can remain coherent at the
@@
 example, internal dynamics and projection effects in the Hyades drive
-stars in the tidal tails span up to $\pm 40\mkms$ in relative velocity
+stars in the tidal tails to $\pm 40\mkms$ in relative velocity
 when projected against the cluster center \citep{jerabkova_800_2021}.
@@
 rotation periods \citep{curtis_tess_2019,bouma_2021_ngc2516}, and
-chemical abundances \citep{hawkins_2020}.  While there are many
+chemical abundances \citep{hawkins_2020}.  While there are 
 implications for our understanding of star formation and cluster
@@
 
-In this work, we expand on our previous study of a $38^{+7}_{-6}$ Myr
+In this work, we expand on our earlier study of a $38^{+7}_{-6}$ Myr
 old Neptune-sized planet in the Kepler field (Kepler-1627~Ab;
@@
 based on its host star's membership in the $\delta$\ Lyr\ cluster.
-Our analysis of the cluster focused on the immediate spatial and
-kinematic vicinity of Kepler-1627~A in order to secure the planet's
-age.  However it became clear that the $\delta$\ Lyr\ cluster seems to
+Our analysis of the cluster focused on the immediate vicinity of
+Kepler-1627 in order to secure the planet's age.  However it became
+clear that the $\delta$\ Lyr\ cluster seems to
 also be part of a much larger group of similarly aged stars.  This
-group, which is at a distance of $\sim$330\,pc from the Sun, appears
-to span Cepheus to Hercules (galactic longitudes, $l$, between
+group, which is at a distance of $\sim$330\,pc from the Sun, 
+spans Cepheus to Hercules (galactic longitudes, $l$, between
 40$^\circ$ and 100$^\circ$), at galactic latitudes roughly between
@@
 Here, our focus is on the intersection of the Cep-Her complex with the
-Kepler field.  Cross-matching the full set of candidate Cep-Her
-members against known Kepler Objects of Interest (KOIs)
-\citep{thompson_planetary_2018} yielded four candidate cluster
-members: Kepler-1627, Kepler-1643, KOI-7368, and KOI-7913.  Given our
-previous analysis of Kepler-1627, we will mostly focus on the latter
-three.  After analyzing the relevant properties of Cep-Her
-(Section~\ref{sec:cluster}), we derive the stellar properties
+Kepler field.  Cross-matching our Cep-Her members against known Kepler
+Objects of Interest (KOIs; \citealt{thompson_planetary_2018}) yielded
+four candidate cluster members: Kepler-1627, Kepler-1643, KOI-7368,
+and KOI-7913.  Given our earlier analysis of Kepler-1627, we focus
+here on the latter three.  After analyzing the relevant properties of
+Cep-Her (Section~\ref{sec:cluster}), we derive the stellar properties
 (Section~\ref{sec:stars}) and validate the planetary nature of each
@@
 high-resolution imaging (Section~\ref{sec:planets}).  We conclude with
-a few caveats, and discuss implications for the size-evolution of
-close-in mini-Neptunes (Section~\ref{sec:disc_conc}).
+a discussion of mini-Neptune size evolution, and point out possible
+directions of future work (Section~\ref{sec:disc_conc}).
 
@@
 KOI-7913.\footnote{These figures
-in fact give an estimate of the size of Cep-Her: if our 4/14
+yield an estimate of the size of Cep-Her: if our 4/14
 possible matches represent the true positive rate, then
@@
 out-of-transit points in each local window were then fitted with a
-fourth-order polynomial, which was then divided out from the light
+fourth-order polynomial, which was divided out from the light
 curve.  The resulting flattened transits were then fitted with a
-planetary transit under the assumption of quadratic limb darkening.
-Our model therefore included 8 free parameters for the transit ($\{P,
+transit model that assumed quadratic limb darkening.
+The model therefore included 8 free parameters for the transit ($\{P,
 t_0, \log R_{\rm p}/R_\star, b, u_1 ,u_2, R_\star, \log g\}$), 2 free
 parameters for the light curve normalization and a white noise jitter
-($\{\langle f \rangle, \sigma_f \}$), and an additional 5 implicit
+($\{\langle f \rangle, \sigma_f \}$), and 5 fixed
 parameters for each transit.
@@
 
-One weakness of our approach is that we have fixed 5 implicit
+One drawback of our approach is that we have fixed 5 implicit
 parameters per individual transit to their MAP values to remove the
-stellar variability signal.  An alternative could be to fit the
+starspot-induced variability.  An alternative could be to fit the
 planetary transits simultaneously with the starspot-induced
-variability using, say, a quasiperiodic Gaussian process (GP).  We
-explored this approach, but found that it required careful fine-tuning
-of the hyperparameter priors, otherwise the GP would tend to
-incorporate variability that was not part of the transit, even if it
-was short-term variability attributable to flares or instrumental
-noise.  This failure mode is pernicious, in that it yields an
-ill-founded sense of confidence in an ``overly clean'' transit model
+variability using a quasiperiodic Gaussian process (GP).  We
+explored this approach, but found that it sometimes required fine-tuning
+of the hyperparameter priors, otherwise the GP would 
+overfit flares, instrumental noise, and any other variability that was not
+part of the transit.  This failure mode is pernicious, in that it yields an
+ill-founded sense of confidence in a visually clean
 fit.  Our model is simpler, and it has the benefit that the white
@@
 	\end{center}
-	\vspace{-0.7cm}
+	\vspace{-0.6cm}
 	\caption{
@@
 		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
-    Planets younger than a gigayear with ${\rm \tau}/\sigma_{\tau} >
-    3$ are emphasized, where $\tau$ is the age and $\sigma_{\tau}$ is
-    its uncertainty. Kepler-1627 (+), KOI-7368 (down-triangle),
-    KOI-7913 (X), Kepler-1643 (diamond).  The large sizes of the
-    youngest transiting planets could be tied to a selection effect.
-    Kepler-1643, KOI-7368, and KOI-7913 have normal mini-Neptune sizes
-    of 2 to 3\,$R_\oplus$, which is a novelty given their ages.
+    Planets younger than a gigayear with ages more precise than a factor of
+    three are emphasized. The Cep-Her planets are
+    Kepler-1643~b ($\square$), KOI-7368~b ($\triangledown$),
+    KOI-7913~Ab (X), and Kepler-1627~Ab (+).  
+    Interesting trends in the population of young planets currently include
+    {\it i)} their large sizes 
+    and {\it ii)} the lack of young hot Jupiters.
+    The new objects of interest in Cep-Her have normal mini-Neptune sizes
+    between 2 and 3\,$R_\oplus$, which is a novelty given their ages.
     Parameters are from the NASA Exoplanet Archive (2022 Apr 5).
@@
 \citet{rogers_unveiling_2021} predicted that given a core-mass
-distribution peaked at $\approx$$4$\,M$_\oplus$, an ice-poor rock/iron
+distribution peaked at $\approx$$4$\,$M_\oplus$, an ice-poor rock/iron
 core composition, and a typical H/He mass fraction of $\approx$4\%,
-there should be a maximum in planet occurrence at 2 to 3\,$R_\oplus$,
+there should be a maximum in planet occurrence at 2 to 3\,$R_\oplus$
 at times between 10 and 100\,Myr
@@
 the theoretical differences lie in their explanations
-for the origin of the radius valley.  If the radius gap is primordial
+for the radius valley.  If the radius gap is primordial
 and created by gas accretion physics, then super-Earths
@@
 time being, we can offer the anecdotal point that in our experience,
-most stars do not have dozens of $\sim$40\,Myr neighbors within a
+most stars do not have dozens of 40\,Myr neighbors within a
 local volume of a few \kms\ and tens of parsecs.
@@
 
-\subsubsection{Cep-Her}
-\paragraph{Kinematics}
-* Need an RV survey.
-
-\paragraph{Age(s)}
-Benchmarking its age(s):
-* Lithium analysis of FGK depletion (bright enough to be feasible).
-
-* Asteroseismology of the delta Sct stars (TESS data are there, group is
-rich enough).
-
-\paragraph{More objects}
-
-* TESS data: giant planets in Cep-Her.  We need more 1e7-1e8 year old
-stars for the HJ occurrence versus time question.  This will help.
-
-* Kepler data: exocomets?  Search for smaller planets?
-
-\subsubsection{Occ rates}
-* Kepler data... presumably $\sim$200 stars in the cluster were
-observed, in order to have detected 4 planets. 
-Occurrence rates?
+\paragraph{Cep-Her}
+Our analysis to date has focused only on portions of Cep-Her
+that were observed by Kepler: namely RSG-5, CH-2, and the $\delta$~Lyr
+cluster.
+The membership and kinematics of the other components shown 
+in Figure~\ref{fig:XYZvtang} also deserve attention.
+R.~Kerr and collaborators have acquired
+{\bf N spectra} from which radial velocities will be derived in their upcoming study, but if the estimate that
+Cep-Her has $\approx$3{,}500 stars is correct then far more might be required
+to fully untangle the kinematics of the various groups.
+Wide-field spectroscopic surveys such as LAMOST ({\bf cite}) or
+APOGEE ({\bf cite}) could provide sensitivity to the Li 6708\,\AA\ line,
+which would enable a more detailed analysis of its abundance across
+FGK stars.
+Gaia DR3 may also provide RVS spectra
+for some of the brighter F and G dwarfs, which could provide the calcium
+infrared triplet as another age indicator.
+
+A number of worthy photometric projects also seem possible
+given the new understanding of the complex.
+One is asteroseismology of the $\delta$~Sct stars,
+using either TESS or Kepler data.
+For cases in which the modes are resolved, this might yield
+age estimates for subgroups independent from other
+approaches.
+Other projects could include searches of the Kepler data
+for exocomets (CITE Zieba), or for missed planets around the
+most rapid rotators.
+
+\paragraph{Exoplanet demographics at early times}
+Our main motivation for finding new young planets is to help
+benchmark models for planetary evolution.
+However demographic analyses of the known planets between $10^7$ and $10^9$ years
+(Figure~\ref{fig:age}) have so far been somewhat limited.
+Approximately 40 such planets
+are now known.
+The majority of published cases ($\approx$20) come from
+K2, though now $\approx$10 come from Kepler, and $\approx$10 
+from TESS.
+
+Given these detections, 
+a few reflections regarding the experimental design of any
+demographic survey might be helpful.
+The first is that the list of target stars is an incredibly important
+ingredient, and the best way to compile relevant stars right
+now is likely through a combination of automated spatio-kinematic clustering
+from Gaia and rotation-period measurement with TESS.
+The second is that the planets $<$$3$\,$R_\oplus$ in size
+all come from either K2 or Kepler.
+So demographic inferences based on TESS
+for the time being are limited to 
+$4$ to $10$\,$R_\oplus$, close-in to the host stars.  
+It seems worthwhile to comparable
+the absolute occurrence rate of such planets to those derived from
+Kepler (CITE Petigura).
+One question that seems like it should be within reach would be to clarify
+whether enough young stars have been searched for the dearth of
+young hot Jupiters shown in Figure~\ref{fig:age} to be significant.
+Since the hot Jupiter occurrence rate is strongly dependent on stellar
+mass and metallicity (CITE), particular care would be needed to select
+a sample of FGK dwarfs for the measurement, likely going out
+to distances of $\approx$500\,pc to get enough stars.
+The major sources of stars in the youngest age bins would
+need to include Sco OB2, Cep-Her, and Orion, among others.
+For studies of how mini-Neptune sizes evolve, the combined K2 and Kepler
+dataset might be the better primary source.
+Though the selection function is more stringent, for a given star
+Kepler acquired $\sim$100$\times$ more photons.  Given the
+star-count limited nature of a cluster-based exoplanet demographic
+survey, every star counts.
 
@@
 
-We have argued that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab {\it i)}
-are 40 to 50 million years old, and {\it ii)} that each signal
+We have showed that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab
+are 40 to 50 million years old, and that each signal
 is most likely planetary.  The evidence for the planetary
@@
 Both components also show H$\alpha$ in emission, which for
-the $\approx$K6V primary is a good indicator that the star is
+the $\approx$K6V primary is a strong indicator that the star is
 $\lesssim$100$\,$Myr old.  For KOI-7368, which is slightly more
@@
 star's youth.
+
+The astrophysical significance of these somewhat detailed considerations
+is that planets $\sim$2 Earth radii in size exist at ages of 40 million years.
+It will be interesting to continue the push down to smaller planetary
+sizes at comparable ages -- the planetary detections we have
+presented are well above the average detection significance of Kepler 
+planets.  There may well still be room at the bottom.
 

### 2022-04-12 papers/ms.tex
update ms

@@
 
-The discovery and characterization of transiting planets younger than
+The discovery and characterization of planets younger than
 a billion years is a major frontier in current exoplanet research.
@@
 for studies of planetary evolution.  For instance, young planets can
-help inform our understanding
+inform our understanding
 of when hot Jupiters arrive on their close-in orbits
 \citep{dawson_johnson_2018}, how the sizes of planets with massive
-gaseous envelopes evolve \citep{rizzuto_tess_2020}, when and if
-close-in multiplanet systems fall out of resonance
+gaseous envelopes evolve \citep{rizzuto_tess_2020}, the timescales for
+close-in multiplanet systems to fall out of resonance
 \citep{arevalo_stability_2022,goldberg_architectures_2022}, and
@@
 
-The discovery of a young planet requires two claims to be fulfilled:
+The discovery of a young planet requires two claims to be true:
 the planet must exist, and its age must be secured.  Spaced-based
@@
 or else on correlates of youth such as stellar rotation, photospheric
-lithium content, x-ray activity, or emission line strength
+lithium content, x-ray activity, and emission line strength
 \citep{zhou_2021_tois,hedges_toi-2076_2021}.
@@
 e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To date these
-analyses have mostly clustered on the stellar positions and 2-D
-velocities.  One important result is the discovery of
+analyses have mostly clustered on stellar positions and 2-D
+velocities.  One important result has been the discovery of
 diffuse streams and tidal tails comparable in stellar mass to the
@@
 parsecs, their velocity dispersions can remain coherent at the
-$\sim$1\kms\ level.  They can also be much larger: as an extreme
-example, internal dynamics and projection effects in the Hyades drive
-stars in the tidal tails to $\pm 40\mkms$ in relative velocity
-when projected against the cluster center \citep{jerabkova_800_2021}.
+$\sim$1\kms\ level.  Internal dynamics and projection effects can also
+drive them to be much larger: in the Hyades,
+stars in the tidal tails are expected to span up to $\pm 40\mkms$ in velocity
+relative to the cluster center \citep{jerabkova_800_2021}.
 The stars in such diffuse regions can be verified to be the same age
-as the core members ({\it i.e.}, coeval) through analyses of
+as the core members through analyses of
 color-absolute magnitude diagrams \citep{KounkelCovey2019}, stellar
@@
 old Neptune-sized planet in the Kepler field (Kepler-1627~Ab;
-\citealt{bouma_kep1627_2022}).  The age of this planet was derived
+\citealt{bouma_kep1627_2022}).  This planet's age was derived
 based on its host star's membership in the $\delta$\ Lyr\ cluster.
-Our analysis of the cluster focused on the immediate vicinity of
-Kepler-1627 in order to secure the planet's age.  However it became
+While our analysis of the cluster focused on the immediate vicinity of
+Kepler-1627 in order to have a reasonable scope,  it became
 clear that the $\delta$\ Lyr\ cluster seems to
@@
 spans Cepheus to Hercules (galactic longitudes, $l$, between
-40$^\circ$ and 100$^\circ$), at galactic latitudes roughly between
+40$^\circ$ and 100$^\circ$), at galactic latitudes between
 0$^\circ$ and 20$^\circ$.  We therefore refer to it as the Cep-Her
@@
 and KOI-7913.  Given our earlier analysis of Kepler-1627, we focus
-here on the latter three.  After analyzing the relevant properties of
+here on the latter three objects.  After analyzing the relevant properties of
 Cep-Her (Section~\ref{sec:cluster}), we derive the stellar properties
@@
 a discussion of mini-Neptune size evolution, and point out possible
-directions of future work (Section~\ref{sec:disc_conc}).
+directions for future work (Section~\ref{sec:disc_conc}).
 
@@
 from the Sun, above the galactic plane, and spanning galactic
-longitudes of roughly $40^\circ$ to $100^\circ$ degrees.  Two rich
+longitudes of roughly $40^\circ$ to $100^\circ$.  Two rich
 clusters in this region are the $\delta$~Lyra cluster
 \citep{stephenson_possible_1959} and RSG-5 \citep{roser_nine_2016}.
-Each of these clusters was known before Gaia.  They have reported ages
+Each of these clusters was known before Gaia.  Their reported ages
 between 30 and 60\,Myr.  Early empirical
@@
 part of an unsupervised clustering analysis of the Gaia DR2 positions
-and tangential velocities with a subsequent manual ``stitching'' step.
+and on-sky velocities with a subsequent manual ``stitching'' step.
 Their results support the idea that there is an overdensity of
@@
 is visible at $(l,b)=(68^\circ,15^\circ$) and $(v_{l'}, v_b)=(-4.5
-\mkms,-4 \mkms)$.  RSG-5 is visible at $(l,b)=(83^\circ,6^\circ)$,
-$(v_{l'}, v_b)=(5.5 \mkms,-3.5 \mkms)$.  Most of the other
+,-4 )\mkms$.  RSG-5 is visible at $(l,b)=(83^\circ,6^\circ)$,
+$(v_{l'}, v_b)=(5.5 ,-3.5 )\mkms$.  Most of the other
 subclusters, including in Cep-Cyg ($l,b=90^\circ,7^\circ$) and
@@
 \paragraph{Cep-Her}
-Our analysis to date has focused only on portions of Cep-Her
-that were observed by Kepler: namely RSG-5, CH-2, and the $\delta$~Lyr
-cluster.
-The membership and kinematics of the other components shown 
-in Figure~\ref{fig:XYZvtang} also deserve attention.
-R.~Kerr and collaborators have acquired
-{\bf N spectra} from which radial velocities will be derived in their upcoming study, but if the estimate that
-Cep-Her has $\approx$3{,}500 stars is correct then far more might be required
-to fully untangle the kinematics of the various groups.
-Wide-field spectroscopic surveys such as LAMOST ({\bf cite}) or
-APOGEE ({\bf cite}) could provide sensitivity to the Li 6708\,\AA\ line,
-which would enable a more detailed analysis of its abundance across
-FGK stars.
-Gaia DR3 may also provide RVS spectra
-for some of the brighter F and G dwarfs, which could provide the calcium
-infrared triplet as another age indicator.
-
-A number of worthy photometric projects also seem possible
-given the new understanding of the complex.
-One is asteroseismology of the $\delta$~Sct stars,
-using either TESS or Kepler data.
-For cases in which the modes are resolved, this might yield
-age estimates for subgroups independent from other
-approaches.
-Other projects could include searches of the Kepler data
-for exocomets (CITE Zieba), or for missed planets around the
-most rapid rotators.
+Our analysis to date has focused only on portions of Cep-Her that were
+observed by Kepler: RSG-5, CH-2, and the $\delta$~Lyr cluster.  The
+membership and kinematics of the other subgroups shown in
+Figure~\ref{fig:XYZvtang} deserve independent attention.  R.~Kerr and
+collaborators have {\bf acquired N spectra} from which radial
+velocities will be derived in their upcoming study, but if the
+estimate that Cep-Her has $\approx$3{,}500 stars is correct then more
+might be required to fully untangle the kinematics of the various
+groups.  Wide-field spectroscopic surveys such as LAMOST
+\citep{zhao_2012_LAMOST} or SDSS-V \citep{kollmeier_2017} could
+provide sensitivity to the Li 6708\,\AA\ line, which would enable a
+more detailed analysis of its abundance across the FGK members.  Gaia
+DR3 may also provide RVS spectra for some of the brighter F and G
+dwarfs, which would contain the calcium infrared triplet as another
+age indicator.
+
+A number of worthy photometric projects also seem possible given the
+new understanding of the complex.  One is asteroseismology of the
+$\delta$~Sct stars, using either TESS or Kepler data
+\citep{bedding_very_2020}.  For cases in which the modes are resolved,
+this might yield age estimates for subgroups independent from other
+approaches.  Other projects could include searches of the Kepler light
+curves for exocomets \citep{zieba_transiting_2019}, and for missed
+planets around the most rapid rotators.
 
 \paragraph{Exoplanet demographics at early times}
-Our main motivation for finding new young planets is to help
-benchmark models for planetary evolution.
-However demographic analyses of the known planets between $10^7$ and $10^9$ years
-(Figure~\ref{fig:age}) have so far been somewhat limited.
-Approximately 40 such planets
-are now known.
-The majority of published cases ($\approx$20) come from
-K2, though now $\approx$10 come from Kepler, and $\approx$10 
-from TESS.
-
-Given these detections, 
-a few reflections regarding the experimental design of any
-demographic survey might be helpful.
-The first is that the list of target stars is an incredibly important
-ingredient, and the best way to compile relevant stars right
-now is likely through a combination of automated spatio-kinematic clustering
-from Gaia and rotation-period measurement with TESS.
-The second is that the planets $<$$3$\,$R_\oplus$ in size
-all come from either K2 or Kepler.
-So demographic inferences based on TESS
-for the time being are limited to 
-$4$ to $10$\,$R_\oplus$, close-in to the host stars.  
-It seems worthwhile to comparable
-the absolute occurrence rate of such planets to those derived from
-Kepler (CITE Petigura).
-One question that seems like it should be within reach would be to clarify
-whether enough young stars have been searched for the dearth of
-young hot Jupiters shown in Figure~\ref{fig:age} to be significant.
-Since the hot Jupiter occurrence rate is strongly dependent on stellar
-mass and metallicity (CITE), particular care would be needed to select
-a sample of FGK dwarfs for the measurement, likely going out
-to distances of $\approx$500\,pc to get enough stars.
-The major sources of stars in the youngest age bins would
-need to include Sco OB2, Cep-Her, and Orion, among others.
-For studies of how mini-Neptune sizes evolve, the combined K2 and Kepler
-dataset might be the better primary source.
-Though the selection function is more stringent, for a given star
-Kepler acquired $\sim$100$\times$ more photons.  Given the
-star-count limited nature of a cluster-based exoplanet demographic
-survey, every star counts.
+Our main motivation for finding new young planets is to help benchmark models
+for planetary evolution.  However demographic analyses of the known planets
+between $10^7$ and $10^9$ years have so far been somewhat limited.
+Approximately 40 such planets are now known (Figure~\ref{fig:age}).  The
+majority of published cases ($\approx$20) come from K2, though now $\approx$10
+come from Kepler, and $\approx$10 from TESS.
+
+Given the current state of the field, a few reflections regarding experimental
+design of a demographic survey might be useful.  The first is that the list of
+target stars with known ages is an essential ingredient.  The best way to
+compile relevant stars right now is likely by combining automated
+spatio-kinematic clustering from Gaia with rotation periods measured using TESS
+\citep[see the appendices of][]{bouma_kep1627_2022}.  The second consideration
+is that known young planets smaller than $3$\,$R_\oplus$ all come from either
+K2 or Kepler.  Demographic inferences based on TESS are therefore limited to
+sizes $\gtrsim4$\,$R_\oplus$, for planets close-in to their host stars.  It
+seems worthwhile to compare the absolute occurrence rate of such planets to
+those observed by Kepler.  One specific question that seems within reach would
+be to clarify whether enough young stars have been searched for the dearth of
+young hot Jupiters shown in Figure~\ref{fig:age} to be significant.  Since the
+hot Jupiter occurrence rate is strongly dependent on stellar mass and
+metallicity \citep{petigura_metallicity_2018,petigura_cksX_2022}, particular
+care would be needed to select a sample of well-studied FGK dwarfs for the
+measurement, likely including stars in Sco OB2, Cep-Her, and Orion.  For
+demographic studies focused on how mini-Neptune sizes evolve, the combined K2
+and Kepler dataset might be the better primary source.
 
@@
 L.G.B{.} acknowledges support from the TESS GI Program (NASA grants
-80NSSC19K0386 and 80NSSC19K1728) and the Heising-Simons Foundation (51 Pegasi~b
+{\bf double check 80NSSC19K0386 and 80NSSC19K1728}) and the Heising-Simons Foundation (51 Pegasi~b
 Fellowship).
@@
 %FIXME
-Keck/NIRC2 imaging was acquired by program 2015A/N301N2L
-(PI: A.~Kraus). % and 2019A/N069 (PI: E.~Petigura).
+Keck/NIRC2 imaging was acquired by programs
+2019X
+{\bf  2019A/N069 (PI: E.~Petigura)},
+{\bf  20NNA/N0XX (PI: D.~Huber)}
+%2015A/N301N2L (PI: A.~Kraus).
 %
@@
 
-\clearpage
+% \clearpage
 \bibliographystyle{yahapj}                            
@@
 
-Tables~\ref{tab:kepler1643}, \ref{tab:koi7368}, \ref{tab:koi7913} give
+%Tables~\ref{tab:kepler1643}, \ref{tab:koi7368}, \ref{tab:koi7913} give
+Table~\ref{tab:kois} gives
 the full set of fitted and derived parameters from the model described in
@@
 
-\input{source_Kepler_1643_simpletransit_posteriortable.tex}
-\input{source_KOI_7368_simpletransit_posteriortable.tex}
-\input{source_KOI_7913_simpletransit_posteriortable.tex}
+\input{source_KOIs_simpletransit_posteriortable.tex}
+%\input{source_Kepler_1643_simpletransit_posteriortable.tex}
+%\input{source_KOI_7368_simpletransit_posteriortable.tex}
+%\input{source_KOI_7913_simpletransit_posteriortable.tex}
 

### 2022-04-13 papers/ms.tex
update ms

@@
 %% Add "Submitted to " argument.
-\received{\today}
+\received{---}
 \revised{---}
@@
 \title{
-  %Three Mini-Neptunes in a 40 Million Year Old Behemoth
   Kepler and the Behemoth: Three Mini-Neptunes in a 40 Million Year Old Association
-  %Three Mini-Neptunes, 40 to 50 Million Years Old 
-  %A Trio of 40 to 50 Million Year Old Mini-Neptunes
 }
@@
 \begin{abstract}
-  Stellar positions and velocities from Gaia are yielding a refined
-  view of open cluster dispersal.  Here we present an analysis of a
+  Stellar positions and velocities from Gaia are yielding a new window
+  on
+  open cluster dispersal.  Here we present an analysis of a
   group stars spanning Cepheus ($l=100^\circ$) to Hercules
@@
 
-% * Main text <3500 words (not including acknowledgements, appendices, or other
-%   supplementary)
-
 \section{Introduction}
@@
 The stars in such diffuse regions can be verified to be the same age
-as the core members through analyses of
+as the core cluster members through analyses of
 color-absolute magnitude diagrams \citep{KounkelCovey2019}, stellar
@@
   diffuse -- we call this region ``CH-2'' (lime-green).
-  The selection method for the black, gray, purple, and green points
+  The selection method for all of the stars displayed
   is described in Section~\ref{subsec:members}.
@@
 Each of these clusters was known before Gaia.  Their reported ages
-between 30 and 60\,Myr.  Early empirical
+are between 30 and 60\,Myr.  Early empirical
 evidence that these two clusters could be part of a large and more
@@
 visibility periods.  We then corrected for extinction using the
-\citet{lallement_threedimensional_2018}\footnote{\url{https://stilism.obspm.fr/}}
-dust maps and the extinction coefficients $k_X\equiv A_X/A_0$ from
+\citet{lallement_threedimensional_2018}
+dust maps\footnote{\url{https://stilism.obspm.fr/}} and the extinction coefficients $k_X\equiv A_X/A_0$ from
 \citet{GaiaCollaboration2018}, assuming that $A_0 = 3.1 E(B-V)$.  This
@@
 same query for the $\delta$~Lyr cluster yielded
-$E(B-V)=0.032\pm0.006$.  Finally, for the plots we set the $\bpmrpo$
+$E(B-V)=0.032\pm0.006$.  Finally, for the plots we set the
 color range to best visualize the region of maximal age information
@@
 
-The CAMDs show that for RSG-5, almost all of our candidate members are
+The CAMDs show that for RSG-5, nearly all of our candidate members are
 on a tight pre-main-sequence locus.  This implies a false positive
@@
 $\approx$12\%, based on the number of stars that photometrically
-overlap with the field population, rather than the bulk cluster
+overlap with the field population, while being outliers from the bulk cluster
 population.  For CH-2, our membership selection gives 27 objects in
@@
 CH-2 members overlap with the $\delta$~Lyr cluster, and that the groups are therefore
-roughly the same age.  To quantify this, we use the empirical method
-introduced by \citet[][their Section~6.3]{gagne_mutau_2020}.  This
+roughly the same age.  To quantify this, we use the method
+introduced by \citet[][their Section~6.3]{gagne_mutau_2020}.  The
 idea is to fit the pre-main-sequence loci of a set of
@@
 uncertainty of the resulting age posterior are $46^{+9}_{-7}$\,Myr for
-RSG-5, and $36^{+10}_{-8}$\,Myr for CH-2.  For comparison, the this
+RSG-5, and $36^{+10}_{-8}$\,Myr for CH-2.  For comparison, this
 procedure yields an age for the $\delta$~Lyr cluster of
@@
 are bright enough to extract usable light curves from the full-frame
-images (FFIs).  The magnitude cut corresponds to $\bpmrpo < 2.6$
+images.  The magnitude cut corresponds to $\bpmrpo < 2.6$
 ($\sim$M3V) at the relevant distances.  These cuts gave 19 stars in
@@
 M-dwarfs is not obvious.  Regardless, the $\approx$15\% false positive
-rate determined from the CAMD seems consistent with our fraction of
-detected rotation periods, given that RSG-5 was also missing rotation
-period detections for $\approx$15\% of its candidate members, which
-all seemed photometrically consistent with being part of a single
-pre-main-sequence locus.
+rate for CH-2 determined from the CAMD seems consistent with our
+fraction of detected rotation periods; RSG-5 was missing
+rotation periods for $\approx$15\% of its candidate members, even
+though all of its stars appear photometrically consistent with being
+on a single pre-main-sequence locus.
 
@@
 \citealt{douglas_stephanie_t_2021_5131306}).  They also seem
-consistent with the estimated false positive rates determined from the
+consistent with the false positive rates estimated from the
 color-absolute magnitude diagrams.
@@
 
-Some salient stellar properties of the Kepler objects in Cep-Her can
-be gleaned from Figure~\ref{fig:age}.  They span spectral types of G8V
-(Kepler-1627) to K6V (KOI-7913 A).  The secondary in the KOI-7913
-system has spectral type $\approx$K8V.  And since a Solar-mass star
-with solar metallicity arrives at the zero-age main sequence at
+A few salient properties of the Kepler objects in Cep-Her can be
+gleaned from Figure~\ref{fig:age}.  The stars span spectral types of
+G8V (Kepler-1627) to K6V (KOI-7913 A).  The secondary in the KOI-7913
+system has spectral type $\approx$K8V.  And given that a Solar-mass
+star with solar metallicity arrives at the zero-age main sequence at
 $t\approx40$ Myr \citep{choi_mesa_2016}, these stars are all in the
-late stages of their pre-main-sequence contraction.  Their
-heliocentric distances are $\approx$260\,pc (KOI-7913 and KOI-7368, in
-CH-2) and $\approx$340\,pc (Kepler-1643, in RSG-5).
+late stages of their pre-main-sequence contraction.  
+% Their heliocentric distances are $\approx$260\,pc (KOI-7913 and
+% KOI-7368, in CH-2) and $\approx$340\,pc (Kepler-1643, in RSG-5).
 
@@
 To verify these parameters and to also analyze spectroscopic youth
-proxies such as the Li\,\textsc{I} 6708\,\AA\ doublet and H$\alpha$,
+proxies such as the Li 6708\,\AA\ doublet and H$\alpha$,
 we acquired spectra.  We also acquired high resolution imaging for
@@
 eff}, \log g, R_\star$) using \texttt{SpecMatch-Emp}
-\citep{yee_SM_2017}, which yielded values in $<1$-$\sigma$ agreement
+\citep{yee_SM_2017}, which yielded values in $<$$1$-$\sigma$ agreement
 with those from the cluster-isochrone method.  This approach also
@@
 found $v\sin i = 9.3 \pm 1.0\,\mkms$.  The systemic radial velocity at
-the two sequence epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm
+the two epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm
 1.2\,\mkms$ respectively.  To infer the equivalent width of the
@@
 EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with values consistent at
-$<1$-$\sigma$ between the two epochs.   The quoted value does not
+$<$$1$-$\sigma$ between the two epochs.   The quoted value does not
 correct for the Fe\,\textsc{I} blend at 6707.44\AA, because HIRES
 resolves the two features.  
+Given the purported age and effective temperature of the star, the
+lithium equivalent width is somewhat low.  We discuss this in greater
+depth in Appendix~\ref{app:spectra}.
 
@@
 Tillinghast 1.5m at the Fred Lawrence Whipple Observatory.  The
-stellar parameter classification pipeline for the TRES spectra has
+stellar parameter classification pipeline for TRES has
 been described by \citet{2021tsc2.confE.124B}.  It is based on the
-synthetic template library originally constructed by
+synthetic template library constructed by
 \citet{buchhave_hatp16b_class_2010}.  The resulting stellar parameters
@@
 June 12, again using NIRC2.  The observational configuration and
-reduction was identical as for Kepler-1643.  No companions were
+reduction were identical as for Kepler-1643.  No companions were
 detected, and the analysis of the image residuals yielded contrast
@@
 13, and KOI-7913 B on the night of 2021 Oct 26.  The
-\texttt{SpecMatch-Emp} \citep{yee_SM_2017} machinery yielded $T_{\rm
-eff,A} = 4324 \pm 70\,{\rm K}$, $T_{\rm eff,B} = 4038 \pm 70\,{\rm
+\texttt{SpecMatch-Emp} machinery yielded $T_{\rm
+eff,A} = 4324 \pm 70\,{\rm K}$, and $T_{\rm eff,B} = 4038 \pm 70\,{\rm
 K}$.  The remaining parameters were in agreement with those from the
@@
 is expected given their $\approx$K6V and $\approx$K8V spectral types.
-They do however both show H$\alpha$ in emission.  We discuss the
-significance in greater depth in Appendix~\ref{app:spectra}, but
-broadly speaking both observations are consistent with a $\approx$40
-Myr age for KOI-7913.
+They do however both show H$\alpha$ in emission.  Broadly speaking,
+both observations are consistent with a $\approx$40 Myr age for
+KOI-7913 (see Appendix~\ref{app:spectra}).
 
@@
 	\caption{
-		%    namelist = ['Kepler-1627 A', 'KOI-7368', 'KOI-7913 A', 'KOI-7913 B', 'Kepler-1643']
-		%    markers = ['P', 'v', 'X', 'X', 's']
 		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
@@
 We have showed that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab
-are 40 to 50 million years old, and that each signal
+are 40 to 50 million years old, and that each system
 is most likely planetary.  The evidence for the planetary
 interpretation comes from an application of \texttt{VESPA} to the
-Kepler data, alongside new adaptive optics imaging from NIRC2.  The
+Kepler data, alongside new imaging from NIRC2.  The
 validity of the \texttt{VESPA} framework rests on the premise that
-non-astrophysical false positives, including false alarms and known
-neighboring eclipsing binaries, can be rejected.  Kepler-1643 and
-KOI-7368 both pass all but one of the canonical vetting checks: the
+non-astrophysical false positives can be rejected.  Kepler-1643 and
+KOI-7368 both pass all but one of the canonical vetting tests: the
 check for centroid offsets between the in and out-of-transit phases.
-For both cases, we have argued the weak observed shifts seem to be
+For both cases, the weak observed shifts seem to be
 caused by starspot-induced variability in specific quarters
-shifting the stellar center-of-light.
-In addition, our imaging rules out companions stars with the necessary
-brightness and position to explain the reported shifts.
+spuriously moving the stellar center-of-light.
+Independently, our imaging rules out companion stars at the brightness
+and positions that would be needed to explain the reported shifts.
 
 Each system has multiple indicators of youth that support the
-reported ages.  For Kepler-1643, perhaps the strongest youth indicator
+reported ages.  For Kepler-1643, the strongest youth indicator
 is its physical and kinematic association with RSG-5.  Based on the
 color-absolute magnitude diagram, we are able to select members
-of this cluster with a false positive rate of $\ll 5\%$
-(Figure~\ref{fig:age}); Kepler-1643 is one such member.  While the stellar
-rotation period period agrees with this assessment, its lithium
-content is somewhat low, which might motivate future exploration
+of this cluster with a false positive rate of a few percent
+(Figure~\ref{fig:age}).  Kepler-1643 is one such member.  While the stellar
+rotation period period agrees with this assessment, the star's lithium
+equivalent width is somewhat low, which might motivate future exploration
 of lithium depletion across FGKM stars in RSG-5 (see Appendix~\ref{app:spectra})
@@
 the $\approx$K6V primary is a strong indicator that the star is
-$\lesssim$100$\,$Myr old.  For KOI-7368, which is slightly more
-massive, the Li 6708\AA\ EW measurement and stellar rotation period
-are more constraining, and provide independent verification of the
+$\lesssim$100$\,$Myr old.  KOI-7368 is slightly more
+massive, and the Li 6708\,\AA\ measurement and stellar rotation period
+provide independent verification of the
 star's youth.
 
-The astrophysical significance of these somewhat detailed considerations
+The astrophysical significance of these considerations
 is that planets $\sim$2 Earth radii in size exist at ages of 40 million years.
@@
 sizes at comparable ages -- the planetary detections we have
-presented are well above the average detection significance of Kepler 
-planets.  There may well still be room at the bottom.
+presented are well above the average detection significance for Kepler 
+planets.  There may still be room at the bottom.
 
@@
 % 	SOAR (SOI),
- 	  TESS \citep{ricker_transiting_2015}.
+ 	  TESS \citep{ricker_transiting_2015},
 % 	TRAPPIST \citep{jehin_trappist_2011},
 % 	VLT:Antu (FORS2).
+    ZTF \citep{bellm_zwicky_2019}.
 }
@@
 
-% \clearpage
+\clearpage
 \bibliographystyle{yahapj}                            
@@
 
-We measured rotation periods for 53 stars with ZTF and 51 stars with
-TESS, for a total of 71 periods reported for unique stars and 32
-overlapping between the surveys. We adopted the ZTF period over the
-TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the
-Lomb-Scargle periodogram favored 6.67 days, consistent with the ZTF
-period of 6.61 days; however, we flagged it as a candidate
-double-dipper, which appears to have inaccurately doubled the TESS
-period to 13.34 days; (2) Gaia EDR3 2081737529891330560: we found 3.06
-days with TESS and 6.64 days with ZTF; we suspect that TESS captured
-the 1/2-period harmonic and adopt the approximately double value from
-ZTF; (3) 2134851775526125696: for this star, we measured 1.91 days
-with TESS from Cycle 2, but noted that the signal appeared to be
-missing in Cycle 4; ZTF found a strong signal at 12.23 days and we
-adopt this as the star's period. In the remaining overlap cases, we
-adopted the average between TESS and ZTF as the final period. For
-these overlap stars, the median absolute deviation is 0.01 days,
-showing remarkable consistency between the surveys. For three stars,
-we failed to detect a period in TESS but recovered one from ZTF; in
-all cases the periods appear to be 13--16 days, pushing the limit of
-what is detectable with TESS with our procedure: (1) Gaia EDR3
-2129930258400157440, TESS showed a flat light curve while ZTF yielded
-a 15.3-day period; (2) Gaia EDR3 2082376861542398336, LS found a
-7.6-day period which we rejected during visual validation; we found
-15.4 days with ZTF, and we suspect that the weak/rejected singal form
-TESS might have been a 1/2 period harmonic; (3) Gaia EDR3
-2082397099429013120, similar to the previous case, we rejected a
-6.7-day signal from TESS and recovered a 12.8-day period with ZTF.
+% We measured rotation periods for 53 stars with ZTF and 51 stars with
+% TESS, for a total of 71 periods reported for unique stars and 32
+% overlapping between the surveys. We adopted the ZTF period over the
+% TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the
+% Lomb-Scargle periodogram favored 6.67 days, consistent with the ZTF
+% period of 6.61 days; however, we flagged it as a candidate
+% double-dipper, which appears to have inaccurately doubled the TESS
+% period to 13.34 days; (2) Gaia EDR3 2081737529891330560: we found 3.06
+% days with TESS and 6.64 days with ZTF; we suspect that TESS captured
+% the 1/2-period harmonic and adopt the approximately double value from
+% ZTF; (3) 2134851775526125696: for this star, we measured 1.91 days
+% with TESS from Cycle 2, but noted that the signal appeared to be
+% missing in Cycle 4; ZTF found a strong signal at 12.23 days and we
+% adopt this as the star's period. In the remaining overlap cases, we
+% adopted the average between TESS and ZTF as the final period. For
+% these overlap stars, the median absolute deviation is 0.01 days,
+% showing remarkable consistency between the surveys. For three stars,
+% we failed to detect a period in TESS but recovered one from ZTF; in
+% all cases the periods appear to be 13--16 days, pushing the limit of
+% what is detectable with TESS with our procedure: (1) Gaia EDR3
+% 2129930258400157440, TESS showed a flat light curve while ZTF yielded
+% a 15.3-day period; (2) Gaia EDR3 2082376861542398336, LS found a
+% 7.6-day period which we rejected during visual validation; we found
+% 15.4 days with ZTF, and we suspect that the weak/rejected singal form
+% TESS might have been a 1/2 period harmonic; (3) Gaia EDR3
+% 2082397099429013120, similar to the previous case, we rejected a
+% 6.7-day signal from TESS and recovered a 12.8-day period with ZTF.
 
@@
     Kepler-1643, and KOI-7913 AB. }
-    The spectra are shown in the observed frame, and
-    the stars are sorted left-to-right in order of decreasing effective temperature.
+    The spectra are shown in the observed frame, and the stars are
+    sorted left-to-right in order of decreasing effective temperature.
     \label{fig:koiyouthindicators}
@@
 Figure~\ref{fig:koiyouthindicators} shows key portions of the HIRES
-and TRES spectra for the Kepler objects in Cep-Her.
-Lithium absorption is obvious at 6708\AA\ in all stars except KOI-7913 B, where it is marginal.
-H$\alpha$ is in emission for both components of KOI-7913,
-and in absorption for the hotter stars.
-In the following, 
-we compare these observations against stars in benchmark
-open clusters.
+and TRES spectra for the Kepler objects in Cep-Her.  Lithium
+absorption is obvious at 6708\AA\ in all stars except KOI-7913 B,
+where it is marginal.  H$\alpha$ is in emission for both components of
+KOI-7913, and in absorption for the hotter stars.  In the following,
+we compare these observations against stars in benchmark open
+clusters.
 
@@
     {\bf Lithium 6708\AA\ and H$\alpha$ equivalent widths for the
-    objects of interest compared to young open clusters and
-    field stars. }
-	Positive equivalent width means absorption; negative
-	equivalent width means emission.
-    The statistical uncertainties on the newly measured equivalent
-    widths are shown, or else are smaller than the markers.
+    objects of interest compared to young open clusters and field
+    stars. } Positive equivalent width means absorption; negative
+    equivalent width means emission.  The statistical uncertainties on
+    the newly measured equivalent widths are shown, or else are
+    smaller than the markers.
     {\it Top}:
-    The field stars are stars with Kepler planets from \citet{berger_identifying_2018}.
-    The ``40-50 Myr'' reference stars (left) are from
-		IC\,2602 \citep{randich_gaiaeso_2018} and Tuc-Hor
-		\citep{kraus_stellar_2014}.
-    The ``112 Myr'' stars (right) are from the Pleiades
+    The field stars are stars with Kepler planets from
+    \citet{berger_identifying_2018}.  The ``40-50 Myr'' reference
+    stars (left) are from IC\,2602 \citep{randich_gaiaeso_2018} and
+    Tuc-Hor \citep{kraus_stellar_2014}.  The ``112 Myr'' stars (right)
+    are from the Pleiades
     \citep{soderblom_evolution_1993,jones_evolution_1996,bouvier_pleiades_lirot_2018}.
@@
 Figure~\ref{fig:lithium} compares the measured lithium equivalent
-widths of the Kepler objects against a few reference populations.
-We selected studies from the literature only where
-upper limits were explicitly reported.  KOI-7368 and
-KOI-7913~A have secure lithium detections, and the KOI-7913~B
-detection is marginal (${\rm EW}_{\rm Li} = 42^{+12}_{-19}$).  For all
-three stars, as well as for Kepler-1627~A, the observed lithium
-equivalent width is consistent with the stellar effective temperatures
-and $\approx40$\,Myr ages.
-
-Kepler-1643, in RSG-5, is conspicuously below the 40-50$\,$Myr sequence
-in the left-panel of Figure~\ref{fig:lithium}, while still being above
-the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$).
+widths of the Kepler objects against a few reference populations.  We
+selected studies from the literature only where upper limits were
+explicitly reported.  KOI-7368 and KOI-7913~A have secure lithium
+detections, and the KOI-7913~B detection is marginal (${\rm EW}_{\rm
+Li} = 42^{+12}_{-19}$).  For all three stars, as well as for
+Kepler-1627~A, the observed lithium equivalent width is consistent
+with the stellar effective temperatures and $\approx40$\,Myr ages.
+
+Kepler-1643, in RSG-5, is conspicuously below the 40-50$\,$Myr
+sequence in the left-panel of Figure~\ref{fig:lithium}, while still
+being above the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$).
 The right panel shows the comparison against the Pleiades, where
-Kepler-1643 is more consistent with the observed dispersion in lithium.
+Kepler-1643 is more consistent with the observed dispersion in
+lithium.
 
@@
 members display a gyrochronal sequence consistent with being younger
-than the Pleaides (Figure~\ref{fig:age} lower left).
-We do not have an explanation, but in the following paragraph
-speculate on one possible line of future work that might help resolve the
-discrepancy.
+than the Pleaides (Figure~\ref{fig:age} lower left).  We do not have
+an explanation, but in the following paragraph speculate on one
+possible line of future work that might help resolve the discrepancy.
 
@@
 
-%Regarding indicators of youth, the primary has a spectral type
-%$\approx$K6V, while the secondary is $\approx$K8V.  At $\approx$40
-%Myr, lithium 6708\,\AA\ absorption is not expected for the secondary,
-%but might be present in the primary \citep[{\it
-%	e.g.},][Figure~8]{soderblom_ages_2014}.  Conversely, H$\alpha$ is
-%expected to be in emission for for all M-dwarfs younger than
-%$\approx150$\,Myr \citep[{\it e.g.},][]{kiman_calibration_2021}.  For
-%the primary, H$\alpha$ emission is also expected based on
-%previous spectroscopic studies of the Pleiades, IC~2602, and IC~2391
-%\citep[{\it e.g.},][their Figure~8]{stauffer_rotational_1997}.  Based
-%on the data from K.~Rocio et al's analysis\footnote{Available at
-%	\url{https://zenodo.org/record/4659293\#.YkuKh27MKDV}}, a $\approx$K6V
-%star should be emitting in H$\alpha$ at 40\,Myr, but not typically
-%past $\approx200\,{\rm Myr}$.  This presents a fairly direct test for
-%the star's age.
-%We
-
 \section{Table of Transit Fit Parameters}
@@
 
-%Tables~\ref{tab:kepler1643}, \ref{tab:koi7368}, \ref{tab:koi7913} give
-Table~\ref{tab:kois} gives
-the full set of fitted and derived parameters from the model described in
-Section~\ref{sec:fitting}.
+Table~2 gives the full set of fitted and derived
+parameters from the model described in Section~\ref{sec:fitting}.
 Priors and convergence statistics are also listed.
@@
 stars directly overlaps with the PRF of the target star (see
-\citealt{thompson_planetary_2018}, Section A.5.2).
-The most obvious explanation for KOI-7913 is the latter case, given
-that KOI-7913 B is $\approx0.9$ Kepler pixels away from Kepler-7913 A
-and so it usually part of the ``halo''.  Due to the on-sky orientation
-of KOI-7913 A and KOI-7913 B, the default ``optimal aperture''
-selected in quarters 3, 7, 11, and 15 in fact included both stars,
-while for the remaining quarters KOI-7913 B was excluded from the
-optimal aperture but was included as part of the halo (see pages 35
-through 71 of the data validation reports.)
+\citealt{thompson_planetary_2018}, Section A.5.2).  The most obvious
+explanation for KOI-7913 is the latter case, given that KOI-7913 B is
+$\approx0.9$ Kepler pixels away from Kepler-7913 A and so it usually
+part of the ``halo''.  Due to the on-sky orientation of KOI-7913 A and
+KOI-7913 B, the default ``optimal aperture'' selected in quarters 3,
+7, 11, and 15 in fact included both stars, while for the remaining
+quarters KOI-7913 B was excluded from the optimal aperture but was
+included as part of the halo (see pages 35 through 71 of the data
+validation reports.)
 
@@
 \input{source_KOIs_simpletransit_posteriortable.tex}
-%\input{source_Kepler_1643_simpletransit_posteriortable.tex}
-%\input{source_KOI_7368_simpletransit_posteriortable.tex}
-%\input{source_KOI_7913_simpletransit_posteriortable.tex}
 

### 2022-04-13 rudolf/plotting.py

@@ -845,9 +845,9 @@ def plot_XYZvtang(outdir, show_1627=0, save_candcomovers=1, save_allphys=1,
             namelist = ['Kepler-1627 A', 'KOI-7368', 'KOI-7913 A', 'Kepler-1643']
             markers = ['P', 'v', 'X', 's']
             # lime: CH-2 (KOI-7913, KOI-7368)
-            # magenta: RSG5 (Kepler-1643)
+            # #ff6eff: RSG5 (Kepler-1643)
             # gray/black: del Lyr cluster (Kepler-1627)
-            mfcs = ['white', 'lime', 'lime', 'magenta']
+            mfcs = ['white', 'lime', 'lime', '#ff6eff']
 
             # drops KOI-7913 B
             koi_df_dict = {k:v for k,v in koi_df_dict.items() if k in namelist}
@@ -872,7 +872,7 @@ def plot_XYZvtang(outdir, show_1627=0, save_candcomovers=1, save_allphys=1,
 
         if show_rsg5:
             axd[k].scatter(
-                df_rsg5_edr3[xv], df_rsg5_edr3[yv], c='magenta', alpha=1, zorder=6,
+                df_rsg5_edr3[xv], df_rsg5_edr3[yv], c='#ff6eff', alpha=1, zorder=6,
                 s=2, edgecolors='none', rasterized=True, marker='.',
                 label='KC19 RSG_5'
             )
@@ -1978,7 +1978,7 @@ def plot_hr(
 
     s = 4
     if smalllims:
-        s = 6.5
+        s = 9
 
     # mixed rasterizing along layers b/c we keep the loading times nice
     l0 = '$\delta$ Lyr candidates'
@@ -2055,9 +2055,9 @@ def plot_hr(
         namelist = ['Kepler-1627 A', 'KOI-7368', 'KOI-7913 A', 'KOI-7913 B', 'Kepler-1643']
         markers = ['P', 'v', 'X', 'X', 's']
         # lime: CH-2 (KOI-7913, KOI-7368)
-        # magenta: RSG5 (Kepler-1643)
+        # #ff6eff: RSG5 (Kepler-1643)
         # gray/black: del Lyr cluster (Kepler-1627)
-        mfcs = ['white', 'lime', 'lime', 'lime', 'magenta']
+        mfcs = ['white', 'lime', 'lime', 'lime', '#ff6eff']
 
         for mfc, marker, (name,_kdf) in zip(mfcs, markers, koi_df_dict.items()):
 
[…]

### 2022-04-20 papers/ms.tex
update ms

@@
 In this work, we leverage recent analyses of the Gaia data, which have
-greatly expanded our knowledge of stellar groups \citep[{\it
-e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To date these
-analyses have mostly clustered on stellar positions and 2-D
+greatly expanded our knowledge of stellar groups \citep[{e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To date these
+analyses have mostly clustered on stellar positions and 2D
 velocities.  One important result has been the discovery of
@@
 Gaia.  It is therefore sensible to revisit the Kepler field, given our
-new knowlege of the stellar ages.
-
-In this work, we expand on our earlier study of a $38^{+7}_{-6}$ Myr
+new knowledge of the stellar ages.
+
+Here, we expand on our earlier study of a $38^{+7}_{-6}$ Myr
 old Neptune-sized planet in the Kepler field (Kepler-1627~Ab;
@@
 
-Here, our focus is on the intersection of the Cep-Her complex with the
+Our focus is on the intersection of the Cep-Her complex with the
 Kepler field.  Cross-matching our Cep-Her members against known Kepler
@@
   gray points.  Kepler-1627 is in the outskirts of the $\delta$ Lyr
-  cluster \citep{bouma_kep1627_2022}, which is centered at $\{ l, b\}
-  \approx \{ 66^\circ, 12^\circ\}$.
-  {\it Middle row}: Galactic positions.  The Sun is at $\{X, Y, Z\} =
-  \{0, 0, 20.8\}$\,pc; lines of constant heliocentric distance are
+  cluster \citep{bouma_kep1627_2022}, which is centered at $(l,b)
+  \approx (66^\circ, 12^\circ)$.
+  {\it Middle row}: Galactic positions.  The Sun is at $(X, Y, Z) =
+  (0, 0, 20.8)$\,pc; lines of constant heliocentric distance are
   shown between 250 and 400\,pc, spaced by 50\,pc.
@@
 and overplot 4{,}763 objects with weights exceeding 0.10 as black
-points.\footnote{These counts only include objects with reliable
+points ({\bf @LGB: slightly more description of the ``weight'' parameter
+is needed}).\footnote{These counts only include objects with reliable
 astrometry and photometry: $\varpi/\sigma_\varpi>5$;
@@
 $\lesssim 10^8$ years old, based on the presence of rotational
-modulation at the expected period and amplitude \citep[{\it
-e.g.},][Figure~9]{rebull_rotation_2020}.  Four objects remained after
+modulation at the expected period and amplitude \citep[{e.g.},][Figure~9]{rebull_rotation_2020}.  Four objects remained after
 this inspection: Kepler-1627, Kepler-1643, KOI-7368, and
@@
 yield an estimate of the size of Cep-Her: if our 4/14
-possible matches represent the true positive rate, then
+matches represent the true positive rate, then
 multiplying by the number of stars above our weight threshold
-(12{,}436) yields $\approx$3{,}500 stars in the complex.}
+(12{,}436) yields $\approx$3{,}500 stars in the complex.  
+This might however be an under-count,
+since Kepler observed sparse 
+regions of the complex, which are expected to have higher
+contamination rates (Figure~\ref{fig:XYZvtang}, top panel).}
 
@@
 candidate RSG-5 members, and 37 candidate CH-2 members.
-Tables of these stars are enumerated in Appendix~\ref{app:members}.
+These stars are listed in Appendix~\ref{app:members}, as are
+the Cep-Her candidates that were observed by
+Kepler.
 An important consideration, especially for CH-2, is the
@@
 it leverages a storied tradition of measuring rotation periods of stars in
-benchmark open clusters \citep[see {\it
-e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data in
+benchmark open clusters \citep[see {e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data in
 our case are especially useful, since they provide 3 to 5 lunar months
@@
 perhaps be taken as an indication for the fraction of stars for
-which rotation periods might not be detectable, due to {\it e.g.}, pole-on
+which rotation periods might not be detectable, due to {e.g.}, pole-on
 stars having lower amplitude starspot modulation.
@@
 both CH-2 and RSG-5 seem consistent with other 30\,Myr to 50\,Myr
-clusters ({\it e.g.}, IC\,2602 and IC\,2391;
+clusters ({e.g.}, IC\,2602 and IC\,2391;
 \citealt{douglas_stephanie_t_2021_5131306}).  They also seem
@@
 G8V (Kepler-1627) to K6V (KOI-7913 A).  The secondary in the KOI-7913
-system has spectral type $\approx$K8V.  And given that a Solar-mass
-star with solar metallicity arrives at the zero-age main sequence at
+system has spectral type $\approx$K8V.  And given that a star with Solar mass
+and metallicity arrives at the zero-age main sequence at
 $t\approx40$ Myr \citep{choi_mesa_2016}, these stars are all in the
@@
 Tillinghast 1.5m at the Fred Lawrence Whipple Observatory.  The
-stellar parameter classification pipeline for TRES has
+Stellar Parameter Classification pipeline for TRES has
 been described by \citet{2021tsc2.confE.124B}.  It is based on the
 synthetic template library constructed by
-\citet{buchhave_hatp16b_class_2010}.  The resulting stellar parameters
+\citet{Buchhave2012}.  The resulting stellar parameters
 ($T_{\rm eff}, \log g, R_\star$) agreed with those from the
@@
 a likely explanation for why planets this small have not previously
-been identified \citep[{\it e.g.},][]{zhou_2021_tois}, future work
+been identified \citep[{e.g.},][]{zhou_2021_tois}, future work
 should quantify this bias more carefully, in order to enable empirical
@@
 the searched volume in units of ${\rm pc}^3\,{\rm km^2}\,{\rm s^{-2}}$, then
-the volume ratio of CH-2 to RSG-5 is 9 to 1.  The number of candidate
-members per unit searched volume in RSG-5 relative to CH-2 is
-therefore 34 to 1.
+the volume ratio of CH-2 to RSG-5 is 9 to 1.  The ratio of number
+densities (candidate members per unit searched volume) in RSG-5
+relative to CH-2 is 34 to 1.
 
 Given its low density, in CH-2 really a star cluster?  For this
-discussion, let us adopt the definition that a star cluster is a group
+discussion, we adopt the definition that a star cluster is a group
 of at least 12 stars that was physically associated at its time of
@@
 well as through the expansion itself
-\citep[see][]{crundall_chronostar_2019}.  A more minimal approach is
+\citep[see][]{crundall_chronostar_2019}.  
+A more minimal approach is
 that suggested by \citet{tofflemire_2021}: search for coeval,
@@
 collaborators have {\bf acquired N spectra} from which radial
-velocities will be derived in their upcoming study, but if the
-estimate that Cep-Her has $\approx$3{,}500 stars is correct then more
-might be required to fully untangle the kinematics of the various
+velocities will be derived in their upcoming study.
+One important aspect of these analyses will be to assess whether the 
+traceback approach could be applicable.
+However if the
+estimate that Cep-Her has $\gtrsim$3{,}500 stars is correct then more
+data might be required to fully untangle the kinematics of the various
 groups.  Wide-field spectroscopic surveys such as LAMOST
 \citep{zhao_2012_LAMOST} or SDSS-V \citep{kollmeier_2017} could
-provide sensitivity to the Li 6708\,\AA\ line, which would enable a
-more detailed analysis of its abundance across the FGK members.  Gaia
+enable such analyses, while also
+providing sensitivity to the Li 6708\,\AA\ line.  Gaia
 DR3 may also provide RVS spectra for some of the brighter F and G
@@
 \citep{bedding_very_2020}.  For cases in which the modes are resolved,
-this might yield age estimates for subgroups independent from other
+this might yield age estimates for the subgroups independent of other
 approaches.  Other projects could include searches of the Kepler light
@@
 between $10^7$ and $10^9$ years have so far been somewhat limited.
-Approximately 40 such planets are now known (Figure~\ref{fig:age}).  The
-majority of published cases ($\approx$20) come from K2, though now $\approx$10
-come from Kepler, and $\approx$10 from TESS.
+Approximately 40 such planets are now known (Figure~\ref{fig:age}).
+About half come from K2, a quarter from TESS, and now a quarter from
+Kepler.
 
 Given the current state of the field, a few reflections regarding experimental
-design of a demographic survey might be useful.  The first is that the list of
-target stars with known ages is an essential ingredient.  The best way to
+design of a demographic survey focused on planetary evolution during
+the first gigayear might be useful.  The first is that such a project
+requires a set of
+target stars with known ages.  The best way to
 compile relevant stars right now is likely by combining automated
@@
 \citep[see the appendices of][]{bouma_kep1627_2022}.  The second consideration
-is that known young planets smaller than $3$\,$R_\oplus$ all come from either
+is that the known young planets smaller than $3$\,$R_\oplus$ all come from either
 K2 or Kepler.  Demographic inferences based on TESS are therefore limited to
-sizes $\gtrsim4$\,$R_\oplus$, for planets close-in to their host stars.  It
+planetary sizes $\gtrsim4$\,$R_\oplus$, for planets close-in to their host stars.  It
 seems worthwhile to compare the absolute occurrence rate of such planets to
@@
 be to clarify whether enough young stars have been searched for the dearth of
-young hot Jupiters shown in Figure~\ref{fig:age} to be significant.  Since the
+young hot Jupiters to be significant.  Since the
 hot Jupiter occurrence rate is strongly dependent on stellar mass and
@@
 care would be needed to select a sample of well-studied FGK dwarfs for the
-measurement, likely including stars in Sco OB2, Cep-Her, and Orion.  For
+measurement, likely using stars in Sco OB2, Cep-Her, and Orion.  For
 demographic studies focused on how mini-Neptune sizes evolve, the combined K2
@@
 of this cluster with a false positive rate of a few percent
-(Figure~\ref{fig:age}).  Kepler-1643 is one such member.  While the stellar
-rotation period period agrees with this assessment, the star's lithium
-equivalent width is somewhat low, which might motivate future exploration
-of lithium depletion across FGKM stars in RSG-5 (see Appendix~\ref{app:spectra})
+(Figure~\ref{fig:age}).  Kepler-1643 is one such member.  While the
+stellar rotation period period agrees with this assessment, the star's
+lithium equivalent width is somewhat low, which might motivate future
+exploration of lithium depletion across FGKM stars in RSG-5 (see
+Appendix~\ref{app:spectra}).
 
@@
 \raggedbottom
-%
-L.G.B{.} acknowledges support from the TESS GI Program (NASA grants
-{\bf double check 80NSSC19K0386 and 80NSSC19K1728}) and the Heising-Simons Foundation (51 Pegasi~b
-Fellowship).
-% programs G011103 and G022117, through 
-%
-%
-%FIXME
-Keck/NIRC2 imaging was acquired by programs
-2019X
-{\bf  2019A/N069 (PI: E.~Petigura)},
-{\bf  20NNA/N0XX (PI: D.~Huber)}
-%2015A/N301N2L (PI: A.~Kraus).
+% Cycle 3 and 4
+L.G.B{.} acknowledges support from the Heising-Simons Foundation (51 Pegasi~b
+Fellowship)
+and the TESS GI Program (NASA grants
+80NSSC21K0335 and 80NSSC22K0298).
 %
@@
 %
-We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
-S.~Seager, J.~Jenkins) and the many TESS team members for their
+We thank the TESS Architects
+%(G.~Ricker, R.~Vanderspek, D.~Latham, S.~Seager, J.~Jenkins)
+and the many TESS team members for their
 efforts to make the mission a continued success.
@@
 \appendix
-\section{Table of Candidate RSG-5 and CH-2 Members}
+\section{Tables of Candidate Cep-Her Members}
 \label{app:members}
 
-{\bf LGB TODO: add Gaia source lists and abbreviated info.  Include rotation periods.}
+{\bf LGB TODO: add Tables}
+Three tables:
+1) All RSG-5 candidates (with rotation periods)
+2) All CH-2 candidates (with rotation periods)
+3) Cep-Her candidates observed by Kepler
 
@@
 members display a gyrochronal sequence consistent with being younger
-than the Pleaides (Figure~\ref{fig:age} lower left).  We do not have
-an explanation, but in the following paragraph speculate on one
-possible line of future work that might help resolve the discrepancy.
-
-Analyzing Figure~\ref{fig:lithium}, it is remarkable that in just 50
-million years, stars between $4500$\,K and $5200$\,K go from having
-such a tight lithium sequence to one with a dispersion
+than the Pleaides (Figure~\ref{fig:age} lower left).  
+There is a $\approx$$1\%$ chance of being a field interloper based on
+the spatio-kinematic selection, and an additional low probability
+($\approx1\%$) of a field K2V star having a rotation period below the
+Pleiades \citep{mcquillan_rotation_2014}.
+The extra observation of a strong lithium detection at these
+effective temperatures yields a very small probability for Kepler-1643 
+being a field star.
+It also seems unlikely that RSG-5 could be much older than 50\,Myr,
+based on its proximity to the $\delta$~Lyr cluster and IC\,2602 in the
+CAMD, as well as the data shown in rotation versus color diagram.
+
+A third possible explanation for Kepler-1643's odd lithium equivalent
+width could be tied to the calibration sample of IC\,2602 and Tuc-Hor
+stars.  Analyzing Figure~\ref{fig:lithium}, it is remarkable that in
+just 50 million years, stars between $4500$\,K and $5200$\,K go from
+having such a tight lithium sequence to one with a dispersion
 $\approx10\times$ greater.  The existence of such a dispersion in
-Pleiades-age K-dwarfs has been known for longer than some of the
-present authors have been alive; it has also been known that the stars
-with the most lithium are also the most rapidly rotating
+Pleiades-age K-dwarfs has been known for decades; it has also been
+known that the stars with the largest lithium abundances are also the
+most rapidly rotating
 \citep{butler_pleiades_1987,soderblom_evolution_1993}.  More recent
@@
 $\delta$~Lyr cluster (and therefore IC\,2602
-\citealt{bouma_kep1627_2022}), it might be a worthy cluster for such
-an analysis.
+\citealt{bouma_kep1627_2022}), it seems like it would be a worthy
+cluster for such an analysis.
 

### 2022-04-20 rudolf/plotting.py

@@ -3764,7 +3764,10 @@ def plot_koiyouthlines(outdir):
             wav[sel], fn(norm(flx[sel])), c='k', zorder=3, lw=0.2
         )
 
-        ax0.set_title(starname.replace("_"," "), fontsize='small')
+        titlestr = starname.replace("_"," ")
+        if starname == 'Kepler-1627':
+            titlestr += ' A'
+        ax0.set_title(titlestr, fontsize='small')
         ax0.set_xlim(xlims[0])
         ax1.set_xlim(xlims[1])
 
@@ -4685,7 +4688,7 @@ def plot_halpha(outdir, reference='TucHor'):
             marker=m,
             c=mfc,
             label=n,
-            elinewidth=1, capsize=0, lw=0, mew=0.5, markersize=6,
+            elinewidth=1, capsize=0, lw=0, mew=0.5, markersize=10,
             mec='k',
             zorder=5
         )
@@ -4818,6 +4821,7 @@ def plot_lithium(outdir, reference='Randich18'):
     #NGC$\,$2547 & IC$\,$2602
     label = '40-50 Myr' if reference != 'Pleiades' else '112 Myr'
     labels = [label, 'Kepler Field']
+    alpha = 1
 
     # plot vals
     for _cls, _col, z, m, l, s in zip(classes, colors, zorders, markers,
@@ -4825,19 +4829,21 @@ def plot_lithium(outdir, reference='Randich18'):
         if _cls == 'young':
             ax.errorbar(
                 d[f'val_teff_{_cls}'], d[f'val_li_ew_{_cls}'],
-                yerr=d[f'val_li_ew_err_{_cls}'], c=_col, alpha=1,
-                zorder=z, markersize=s, elinewidth=0.5, capsize=0, mew=0.5,
+                yerr=d[f'val_li_ew_err_{_cls}'], c=_col, alpha=alpha,
+                zorder=z, markersize=s, elinewidth=0.35, capsize=0, mew=0.5,
                 rasterized=False, label=l, marker=m, lw=0
             )
 
             ax.scatter(
-                d[f'lim_teff_{_cls}'], d[f'lim_li_ew_{_cls}'], c=_col, alpha=1,
+                d[f'lim_teff_{_cls}'], d[f'lim_li_ew_{_cls}'], c=_col,
+                alpha=alpha,
                 zorder=z+1, s=5*s, rasterized=False, linewidths=0, marker='v'
[…]

### 2022-04-20 papers/ms.tex
update ms

@@
 sequence in the left-panel of Figure~\ref{fig:lithium}, while still
-being above the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$).
+being above the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA).
 The right panel shows the comparison against the Pleiades, where
@@
 There is a $\approx$$1\%$ chance of being a field interloper based on
-the spatio-kinematic selection, and an additional low probability
+the spatio-kinematic selection, and a similar independent chance
 ($\approx1\%$) of a field K2V star having a rotation period below the
@@
 effective temperatures yields a very small probability for Kepler-1643 
-being a field star.
+being a field interloper.
 It also seems unlikely that RSG-5 could be much older than 50\,Myr,
@@
 having such a tight lithium sequence to one with a dispersion
-$\approx10\times$ greater.  The existence of such a dispersion in
+$\approx10\times$ greater.  The existence of the Li dispersion in
 Pleiades-age K-dwarfs has been known for decades; it has also been
@@
 Given that RSG-5 seems somewhat isochronally older than the
-$\delta$~Lyr cluster (and therefore IC\,2602
+$\delta$~Lyr cluster (and therefore IC\,2602;
 \citealt{bouma_kep1627_2022}), it seems like it would be a worthy
@@
 hotter stars.  Additionally, the emission appears double-peaked for
-both of the KOI-7913 components.  We note that KOI-7913~A and
+both of the KOI-7913 components.  An important note is that KOI-7913~A and
 KOI-7913~B were spatially resolved from each other during data
@@
 of these flags (which do not apply to the system, according to the
-latest analysis).  We also discuss how the relative locations of
+latest analysis).  We also discuss how the relative on-sky positions of
 KOI-7913~A and KOI-7913~B affect the interpretation of the Kepler
@@
 \citep{2017ksci.rept....6M}.  Analysis of the target-pixel data that
-was acquired separately for KOI-7913~B also reveals a clearly
+was separately acquired for KOI-7913~B also reveals a
 different stellar rotation period, and no hint of the transit signal.

### 2022-04-21 papers/ms.tex
update ms

@@
 
-\newcommand{\npms}{1097} % 20220311_Kerr_SPYGLASS205_Members_All.csv
+\newcommand{\npms}{1{,}097} % 20220311_Kerr_SPYGLASS205_Members_All.csv
 
@@
 in Figure~\ref{fig:XYZvtang}.  To enable a selection cut that filters
-out field-star contaminants, we also compute a weight metric, defined
-such that the group member with the smallest core distance has a
-weight of 1, the group member with the greatest core distance has a
-weight of 0, and weights for the other group members are log-normally
-distributed between these two extremes.  In Figure~\ref{fig:XYZvtang},
-we show 12{,}436 objects with weight exceeding 0.02 as gray points,
-and overplot 4{,}763 objects with weights exceeding 0.10 as black
-points ({\bf @LGB: slightly more description of the ``weight'' parameter
-is needed}).\footnote{These counts only include objects with reliable
-astrometry and photometry: $\varpi/\sigma_\varpi>5$;
-$G/\sigma_{G}>50$; $G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm
-BP}/\sigma_{G_{BP}}>20$.} The previously known $\delta$~Lyr\ cluster
-is visible at $(l,b)=(68^\circ,15^\circ$) and $(v_{l'}, v_b)=(-4.5
-,-4 )\mkms$.  RSG-5 is visible at $(l,b)=(83^\circ,6^\circ)$,
-$(v_{l'}, v_b)=(5.5 ,-3.5 )\mkms$.  Most of the other
-subclusters, including in Cep-Cyg ($l,b=90^\circ,7^\circ$) and
-Cerberus ($l,b=48^\circ,18^\circ$) are too small or dispersed to have
-previously been analyzed in great detail.
+out field-star contaminants, we also compute a weight metric,
+$\alpha$, defined such that the group member with the smallest core
+distance has $\alpha=1$, the group member with the greatest core
+distance has $\alpha=0$, and weights for the other group members are
+log-normally distributed between these two extremes.  After applying a
+set of quality cuts on the astrometry and
+photometry,\footnote{$\varpi/\sigma_\varpi>5$; $G/\sigma_{G}>50$;
+$G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm BP}/\sigma_{G_{BP}}>20$.}
+this procedure yields $\alpha \sim \log_{10}\mathcal{N}(-1.55,0.61)$.
+To visualize the results, in Figure~\ref{fig:XYZvtang} we show
+12{,}436 objects with weight exceeding 0.02 as gray points, and
+overplot 4{,}763 objects with weights exceeding 0.10 as black points.
+The previously known $\delta$~Lyr\ cluster is visible at
+$(l,b)=(68^\circ,15^\circ$) and $(v_{l'}, v_b)=(-4.5 ,-4 )\mkms$.
+RSG-5 is visible at $(l,b)=(83^\circ,6^\circ)$, $(v_{l'}, v_b)=(5.5
+,-3.5 )\mkms$.  Most of the other subclusters, including in Cep-Cyg
+($l,b=90^\circ,7^\circ$) and Cerberus ($l,b=48^\circ,18^\circ$) are
+too small or dispersed to have previously been analyzed in great
+detail.
 

### 2022-04-22 papers/ms.tex
update ms

@@
 \citep{thompson_planetary_2018}.  From the candidate members
-with weights exceeding 0.02, this yielded
- yielded 11 were known false positives, 6 ``confirmed'' planets, and 8
+with weights exceeding 0.02, this 
+ yielded 11 known false positives, 6 ``confirmed'' planets, and 8
  ``candidate'' planets.  
@@
 2) All CH-2 candidates (with rotation periods)
-3) Cep-Her candidates observed by Kepler
+
+Table~\ref{tab:cepherkepler} contains candidate Cep-Her members with
+weights $\alpha>0.02$ observed by Kepler.  A cursory inspection of the
+color absolute-magnitude diagram suggests that more restrictive weight
+cuts would be needed to remove the majority of field star interlopers.
+Table~\ref{tab:cepherkepler}  was created by cross-matching candidate
+Cep-Her members (selected using Gaia EDR3;
+Section~\ref{subsec:members}) against a Kepler to Gaia DR2 cross-match
+(the \texttt{gaia-kepler.fun} crossmatch database created by Megan
+Bedell).  The \texttt{kic_\dr2\_ang\_dist} column is from the latter
+table.  The EDR3 to DR2 match was performed using the
+\texttt{gaiaedr3.dr2_neighbourhood} table, and the closest (proper
+motion and epoch-corrected) angular distance neighbor was taken as the
+as single single match.  The \texttt{edr3\_dr2\_mag\_diff} column
+gives some indication of the reliability of this EDR3 to DR2
+conversion, as there are likely a few cases where partially resolved
+binaries became fully resolved.
 

### 2022-04-22 papers/tab_supp_CepHer_X_Kepler.tex
update ms

--- 2021-09-07 paper/t1_appendix.tex
@@
 %% Use Default (12pt)
-\tabletypesize{\scriptsize}
-%\tabletypesize{\small}
+%\tabletypesize{\scriptsize}
+\tabletypesize{\small}
 
@@
 %% This is the title of the table.
-\tablecaption{Young, Age-dated, and Age-dateable Stars Within the
-  Nearest Few Kiloparsecs.}
-\label{tab:v06}
-
+\tablecaption{Candidate Cep-Her members observed by Kepler}
+\label{tab:cepherkepler}
 %% This command over-rides LaTeX's natural table count
@@
 % paste from
-% /Users/luke/Dropbox/proj/rudolf/results/tables/v06_main_tableheader.tex
-% via drivers/write_v06_main_tableheader.py
+% /Users/luke/Dropbox/proj/earhart/results/tables/NGC_2516_Prot_cleaned_header.tex
+% via drivers/write_NGC2516_main_table.py
 \startdata
-          \texttt{source\_id} &                                          1709456705329541504 &                                              Gaia DR2 source identifier. \\
-                  \texttt{ra} &                                                      247.826 &                                          Gaia DR2 right ascension [deg]. \\
-                 \texttt{dec} &                                                       79.789 &                                              Gaia DR2 declination [deg]. \\
-            \texttt{parallax} &                                                       35.345 &                                                 Gaia DR2 parallax [mas]. \\
-     \texttt{parallax\_error} &                                                        0.028 &                                     Gaia DR2 parallax uncertainty [mas]. \\
-                \texttt{pmra} &                                                       94.884 &      Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr$^{-1}$]. \\
-               \texttt{pmdec} &                                                      -86.971 &                  Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr$^{-1}$]. \\
-  \texttt{phot\_g\_mean\_mag} &                                                         6.85 &                                                  Gaia DR2 $G$ magnitude. \\
- \texttt{phot\_bp\_mean\_mag} &                                                        6.409 &                                      Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
- \texttt{phot\_rp\_mean\_mag} &                                                        7.189 &                                      Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
-             \texttt{cluster} &                  NASAExoArchive\_ps\_20210506,Uma,IR\_excess &                                   Comma-separated cluster or group name. \\
-                 \texttt{age} &                                                 9.48,nan,nan &  Comma-separated logarithm (base-10) of reported$^{\rm a}$ age in years. \\
-           \texttt{mean\_age} &                                                         9.48 &                           Mean (ignoring NaNs) of $\texttt{age}$ column. \\
-       \texttt{reference\_id} &       NASAExoArchive\_ps\_20210506,Ujjwal2020,CottenSong2016 &                           Comma-separted provenance of group membership. \\
-  \texttt{reference\_bibcode} &  2013PASP..125..989A,2020AJ....159..166U,2016ApJS..225...15C &                   ADS bibcode corresponding to $\texttt{reference\_id}$. \\
+\texttt{dr2\_source\_id}      & 2073765172933035008 & Gaia DR2 source identifier. \\
+\texttt{dr3\_source\_id}      & 2073765172933035008 & Gaia (E)DR3 source identifier. \\
+\texttt{kepid} & 5641711      & KIC identifier. \\
+\texttt{ra} &    297.4099     & Gaia EDR3 right ascension [deg]. \\
+\texttt{dec} &   40.8972      & Gaia EDR3 declination [deg]. \\
+  \texttt{weight} & 0.0409    & Strength of the connectivity to other candidate cluster members. \\
+  \texttt{v\_l} & -0.5061     & Longitudinal galactic velocity [km\,s$^{-1}$]. \\
+  \texttt{v\_b} & -8.2328     & Latitudinal galactic velocity [km\,s$^{-1}$]. \\
+  \texttt{x\_pc} & -8035.42&  & Galactocentric $X$ position coordinate [pc]. \\
+  \texttt{y\_pc} & 331.44     & Galactocentric $Y$ position coordinate [pc]. \\
+  \texttt{z\_pc} & 65.32      & Galactocentric $Z$ position coordinate [pc]. \\
+  \texttt{kic\_dr2\_ang\_dist} & 0.298 & Separation between KIC and Gaia DR2 positions [arcsec]. \\
+  \texttt{edr3\_dr2\_mag\_diff} & 0.002 & $G$-band difference between EDR3 and DR2 source match [mag]. \\
 \enddata
@@
 \tablecomments{
-Table~\ref{tab:v06} is published in its entirety in a machine-readable
-format.   This table is a concatenation of the studies listed in
-Table~\ref{tab:metadata}.  One entry is shown for guidance regarding
-form and content.  In this particular example, the star has a cold
-Jupiter on a 16 year orbit, HD 150706b \citep{2012AA...545A..55B}.  An
-infrared excess has been reported \citep{CottenSong2016}, and the star
-was identified by \citet{Ujjwal2020} as a candidate UMa moving group
-member ($\approx 400\,{\rm Myr}$; \citealt{mann_tess_2020}).  The
-star's RV activity and TESS rotation period corroborate its youth.
+Table~\ref{tab:cepherkepler} is published in its entirety in a machine-readable
+format.  One entry is shown for guidance regarding form and content.
+This involved a conversion based
+
+is a concatenation of all candidate
+NGC\,2516 members reported by \citetalias{cantatgaudin_gaia_2018},
+\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
+based on the Gaia DR2 data.  Different levels of purity and
+completeness can be achieved using different cuts on photometric
+periods, periodogram powers, and lithium eqiuvalent widths.  Sets
+$\mathcal{A}$ and $\mathcal{B}$ provide two possible levels of
+cleaning (see Section~\ref{subsubsec:cluster}).  When the
+target star is the only star present in the TESS aperture,
+\texttt{nequal}$=0$, \texttt{nclose}$=1$, and \texttt{nfaint}$=1$.
+The light curves are available at
+\url{https://archive.stsci.edu/hlsp/cdips}.
+Supplementary plots enabling analyses of individual stars
+are available at \url{https://lgbouma.com/notes}.
 }

### 2022-04-22 papers/ms.tex
update ms / appendix

--- 2022-04-21 papers/ms.tex
@@
 
-{\bf LGB TODO: add Tables}
-Three tables:
-1) All RSG-5 candidates (with rotation periods)
-2) All CH-2 candidates (with rotation periods)
-3) Cep-Her candidates observed by Kepler
-
-% We measured rotation periods for 53 stars with ZTF and 51 stars with
-% TESS, for a total of 71 periods reported for unique stars and 32
-% overlapping between the surveys. We adopted the ZTF period over the
-% TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the
-% Lomb-Scargle periodogram favored 6.67 days, consistent with the ZTF
-% period of 6.61 days; however, we flagged it as a candidate
-% double-dipper, which appears to have inaccurately doubled the TESS
-% period to 13.34 days; (2) Gaia EDR3 2081737529891330560: we found 3.06
-% days with TESS and 6.64 days with ZTF; we suspect that TESS captured
-% the 1/2-period harmonic and adopt the approximately double value from
-% ZTF; (3) 2134851775526125696: for this star, we measured 1.91 days
-% with TESS from Cycle 2, but noted that the signal appeared to be
-% missing in Cycle 4; ZTF found a strong signal at 12.23 days and we
-% adopt this as the star's period. In the remaining overlap cases, we
-% adopted the average between TESS and ZTF as the final period. For
-% these overlap stars, the median absolute deviation is 0.01 days,
-% showing remarkable consistency between the surveys. For three stars,
-% we failed to detect a period in TESS but recovered one from ZTF; in
-% all cases the periods appear to be 13--16 days, pushing the limit of
-% what is detectable with TESS with our procedure: (1) Gaia EDR3
-% 2129930258400157440, TESS showed a flat light curve while ZTF yielded
-% a 15.3-day period; (2) Gaia EDR3 2082376861542398336, LS found a
-% 7.6-day period which we rejected during visual validation; we found
-% 15.4 days with ZTF, and we suspect that the weak/rejected singal form
-% TESS might have been a 1/2 period harmonic; (3) Gaia EDR3
-% 2082397099429013120, similar to the previous case, we rejected a
-% 6.7-day signal from TESS and recovered a 12.8-day period with ZTF.
+\paragraph{Table~2} contains candidate Cep-Her members with weights $\alpha>0.02$
+observed by Kepler.  More restrictive weight cuts
+would be needed if one wanted to remove the majority of field star interlopers: at
+least half of the stars included in the table are unlikely
+to be bonafide members of Cep-Her.  Table~2  was created by
+cross-matching candidate Cep-Her members (selected using Gaia EDR3;
+Section~\ref{subsec:members}) against a Kepler to Gaia DR2 cross-match
+(the \texttt{gaia-kepler.fun} crossmatch database created by Megan
+Bedell).  The \texttt{kic\_dr2\_ang\_dist} column is from the latter
+table.  The EDR3 to DR2 match was performed using the
+\texttt{gaiaedr3.dr2\_neighbourhood} table, and the closest proper
+motion and epoch-corrected angular distance neighbor was taken as the
+single best match.  The \texttt{edr3\_dr2\_mag\_diff} column
+gives some indication of the reliability of this EDR3 to DR2
+conversion, as there are a few cases between Gaia DR2 and EDR3 where
+partially resolved binaries became fully resolved.
+
+\paragraph{Table~3} contains spatial, kinematic, astrometric, and rotation period
+information for the 141 candidate RSG-5 members and 37 candidate CH-2
+members described in Section~\ref{subsec:members}.
+These are the data used to make the lower panels of
+Figure~\ref{fig:age}.
+Table~3 includes rotation periods for 53 stars with ZTF and 51 stars with
+ TESS, for a total of 71 periods reported for unique stars and 32
+ overlapping between the surveys. We adopted the ZTF period over the
+ TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the
+ Lomb-Scargle periodogram favored 6.67 days, consistent with the ZTF
+ period of 6.61 days; however, we flagged it as a candidate
+ double-dipper, which appears to have inaccurately doubled the TESS
+ period to 13.34 days; (2) Gaia EDR3 2081737529891330560: we found 3.06
+ days with TESS and 6.64 days with ZTF; we suspect that TESS captured
+ the 1/2-period harmonic and adopt the approximately double value from
+ ZTF; (3) 2134851775526125696: for this star, we measured 1.91 days
+ with TESS from Cycle 2, but noted that the signal appeared to be
+ missing in Cycle 4; ZTF found a strong signal at 12.23 days and we
+ adopt this as the star's period. In the remaining overlap cases, we
+ adopted the average between TESS and ZTF as the final period. For
+ these overlap stars, the median absolute deviation is 0.01 days,
+ showing remarkable consistency between the surveys. For three stars,
+ we failed to detect a period in TESS but recovered one from ZTF; in
+ all cases the periods appear to be 13--16 days.
+ These stars were: (1) Gaia EDR3
+ 2129930258400157440, for which TESS showed a flat light curve while ZTF yielded
+ a 15.3-day period; (2) Gaia EDR3 2082376861542398336, LS found a
+ 7.6-day period which we rejected during visual validation; we found
+ 15.4 days with ZTF, and we suspect that the weak/rejected signal form
+ TESS might have been a 1/2 period harmonic; (3) Gaia EDR3
+ 2082397099429013120, similar to the previous case, we rejected a
+ 6.7-day signal from TESS and recovered a 12.8-day period with ZTF.
 
@@
 
-\begin{figure*}[tp]
+\begin{figure*}[b]
 	\begin{center}
@@
 explicitly reported.  KOI-7368 and KOI-7913~A have secure lithium
-detections, and the KOI-7913~B detection is marginal (${\rm EW}_{\rm
-Li} = 42^{+12}_{-19}$).  For all three stars, as well as for
+detections, while for KOI-7913~B the detection is marginal (${\rm EW}_{\rm
+Li} = 42^{+12}_{-19}$\,m\AA).  For all three stars, as well as for
 Kepler-1627~A, the observed lithium equivalent width is consistent
@@
 One explanation for this could be that Kepler-1643 is a field
-interloper; another could be that RSG-5 is older than 50\,Myr.  We do
-not favor the former, given that {\it i)} the star was selected based
+interloper; another could be that RSG-5 is much older than 50\,Myr.  We do
+not favor either, given that {\it i)} the star was selected based
 on its spatio-kinematic proximity to other RSG-5 members, nearly all
@@
 the spatio-kinematic selection, and a similar independent chance
-($\approx1\%$) of a field K2V star having a rotation period below the
+($\approx$$1\%$) of a field K2V star having a rotation period below the
 Pleiades \citep{mcquillan_rotation_2014}.
@@
 
-A third possible explanation for Kepler-1643's odd lithium equivalent
-width could be tied to the calibration sample of IC\,2602 and Tuc-Hor
-stars.  Analyzing Figure~\ref{fig:lithium}, it is remarkable that in
-just 50 million years, stars between $4500$\,K and $5200$\,K go from
-having such a tight lithium sequence to one with a dispersion
+Our preferred explanation for Kepler-1643's lithium equivalent
+width is tied to the reference sample of IC\,2602 and Tuc-Hor
+stars not fully exploring the possible stellar rotation
+periods and lithium equivalent widths. 
+Analyzing Figure~\ref{fig:lithium}, it is remarkable that in
+just 50 million years, K-dwarfs between $4500$\,K and $5200$\,K go from
+having a relatively tight lithium sequence to one with a dispersion
 $\approx10\times$ greater.  The existence of the Li dispersion in
@@
 within pre-main-sequence stellar physics.  If so, one would expect the
-IC\,2062 and Tuc-Hor K-dwarfs to show a larger intrinsic lithium
-dispersion, which they do not.  It seems therefore that the exact
-onset time of the K-dwarf lithium dispersion might be worth
-clarifying.  \citet{messina_rotation-lithium_2016} for instance do not
-obviously see it in the $\beta$~Pic moving group ($25\pm3$\,Myr).
-Given that RSG-5 seems somewhat isochronally older than the
-$\delta$~Lyr cluster (and therefore IC\,2602;
-\citealt{bouma_kep1627_2022}), it seems like it would be a worthy
-cluster for such an analysis.
+IC\,2602 and Tuc-Hor K-dwarfs to show a larger intrinsic lithium
+dispersion.  A recent analysis of the $\approx$40\,Myr
+NGC~2547 by \citet{binks_2022} suggests that this may
+be the case, though it only had $\approx$10 stars in the relevant
+effective temperature range.
+This suggests that
+RSG-5 and Cep-Her could be
+worthy objects for a closer analysis of the lithium-rotation correlation
+near the zero-age main sequence.
 
@@
 
-Table~2 gives the full set of fitted and derived
+Table~4 gives the full set of fitted and derived
 parameters from the model described in Section~\ref{sec:fitting}.
@@
 
+%\clearpage
+\input{tab_supp_CepHer_X_Kepler.tex}
+\input{tab_supp_RSG5_CH2_Prot.tex}
 \input{source_KOIs_simpletransit_posteriortable.tex}

### 2022-04-22 papers/tab_supp_CepHer_X_Kepler.tex
update ms / appendix

@@
 %% all other tables after this table based on this number
-%\tablenum{3}
+\tablenum{2}
 
@@
 }
-
 %% All data must appear between the \startdata and \enddata commands
-%
-% paste from
-% /Users/luke/Dropbox/proj/earhart/results/tables/NGC_2516_Prot_cleaned_header.tex
-% via drivers/write_NGC2516_main_table.py
 \startdata
@@
 \texttt{dec} &   40.8972      & Gaia EDR3 declination [deg]. \\
-  \texttt{weight} & 0.0409    & Strength of the connectivity to other candidate cluster members. \\
+\texttt{weight} & 0.0409    & Strength of connectivity to other candidate cluster members. \\
 \texttt{v\_l} & -0.5061     & Longitudinal galactic velocity [km\,s$^{-1}$]. \\
 \texttt{v\_b} & -8.2328     & Latitudinal galactic velocity [km\,s$^{-1}$]. \\
-  \texttt{x\_pc} & -8035.42&  & Galactocentric $X$ position coordinate [pc]. \\
+\texttt{x\_pc} & -8035.42  & Galactocentric $X$ position coordinate [pc]. \\
 \texttt{y\_pc} & 331.44     & Galactocentric $Y$ position coordinate [pc]. \\
@@
 \enddata
-
-%% Include any \tablenotetext{key}{text}, \tablerefs{ref list},
-%% or \tablecomments{text} between the \enddata and 
-%% \end{deluxetable} commands
-
 %% General table comment marker
-\tablecomments{
-Table~\ref{tab:cepherkepler} is published in its entirety in a machine-readable
-format.  One entry is shown for guidance regarding form and content.
-This involved a conversion based
-
-is a concatenation of all candidate
-NGC\,2516 members reported by \citetalias{cantatgaudin_gaia_2018},
-\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
-based on the Gaia DR2 data.  Different levels of purity and
-completeness can be achieved using different cuts on photometric
-periods, periodogram powers, and lithium eqiuvalent widths.  Sets
-$\mathcal{A}$ and $\mathcal{B}$ provide two possible levels of
-cleaning (see Section~\ref{subsubsec:cluster}).  When the
-target star is the only star present in the TESS aperture,
-\texttt{nequal}$=0$, \texttt{nclose}$=1$, and \texttt{nfaint}$=1$.
-The light curves are available at
-\url{https://archive.stsci.edu/hlsp/cdips}.
-Supplementary plots enabling analyses of individual stars
-are available at \url{https://lgbouma.com/notes}.
-}
+\tablecomments{Table~2 is published in its entirety in a machine-readable
+format.  One entry is shown for guidance regarding form and content.}
 \vspace{-0.5cm}

### 2022-04-22 papers/tab_supp_RSG5_CH2_Prot.tex
update ms / appendix

--- 2021-09-07 paper/t1_appendix.tex
@@
 %% This is the title of the table.
-\tablecaption{Young, Age-dated, and Age-dateable Stars Within the
-  Nearest Few Kiloparsecs.}
-\label{tab:v06}
-
+\tablecaption{Rotation periods and kinematics for candidate RSG-5 and
+  CH-2 members.}
+\label{tab:rot}
 %% This command over-rides LaTeX's natural table count
@@
 %% all other tables after this table based on this number
-%\tablenum{3}
+\tablenum{3}
 
@@
 }
-
 %% All data must appear between the \startdata and \enddata commands
 %
-% paste from
-% /Users/luke/Dropbox/proj/rudolf/results/tables/v06_main_tableheader.tex
-% via drivers/write_v06_main_tableheader.py
 \startdata
-          \texttt{source\_id} &                                          1709456705329541504 &                                              Gaia DR2 source identifier. \\
-                  \texttt{ra} &                                                      247.826 &                                          Gaia DR2 right ascension [deg]. \\
-                 \texttt{dec} &                                                       79.789 &                                              Gaia DR2 declination [deg]. \\
-            \texttt{parallax} &                                                       35.345 &                                                 Gaia DR2 parallax [mas]. \\
-     \texttt{parallax\_error} &                                                        0.028 &                                     Gaia DR2 parallax uncertainty [mas]. \\
-                \texttt{pmra} &                                                       94.884 &      Gaia DR2 proper motion $\mu_\alpha \cos \delta$ [mas$\,$yr$^{-1}$]. \\
-               \texttt{pmdec} &                                                      -86.971 &                  Gaia DR2 proper motion $\mu_\delta$ [mas$\,$yr$^{-1}$]. \\
-  \texttt{phot\_g\_mean\_mag} &                                                         6.85 &                                                  Gaia DR2 $G$ magnitude. \\
- \texttt{phot\_bp\_mean\_mag} &                                                        6.409 &                                      Gaia DR2 $G_\mathrm{BP}$ magnitude. \\
- \texttt{phot\_rp\_mean\_mag} &                                                        7.189 &                                      Gaia DR2 $G_\mathrm{RP}$ magnitude. \\
-             \texttt{cluster} &                  NASAExoArchive\_ps\_20210506,Uma,IR\_excess &                                   Comma-separated cluster or group name. \\
-                 \texttt{age} &                                                 9.48,nan,nan &  Comma-separated logarithm (base-10) of reported$^{\rm a}$ age in years. \\
-           \texttt{mean\_age} &                                                         9.48 &                           Mean (ignoring NaNs) of $\texttt{age}$ column. \\
-       \texttt{reference\_id} &       NASAExoArchive\_ps\_20210506,Ujjwal2020,CottenSong2016 &                           Comma-separted provenance of group membership. \\
-  \texttt{reference\_bibcode} &  2013PASP..125..989A,2020AJ....159..166U,2016ApJS..225...15C &                   ADS bibcode corresponding to $\texttt{reference\_id}$. \\
+\texttt{dr3\_source\_id}      & 2127562009133684480 & Gaia (E)DR3 source identifier. \\
+\texttt{ra} &    291.0231     & Gaia EDR3 right ascension [deg]. \\
+\texttt{dec} &  46.4384      & Gaia EDR3 declination [deg]. \\
+\texttt{parallax} &  3.7099      & Gaia EDR3 parallax [milliarcsec]. \\
+\texttt{ruwe} &  0.981      & Gaia EDR3 renormalized unit weight error. \\
+\texttt{weight} & 0.0870    & Strength of connectivity to other candidate cluster members. \\
+\texttt{v\_l} & 2.7793     & Longitudinal galactic velocity [km\,s$^{-1}$]. \\
+\texttt{v\_b} & -2.8675     & Latitudinal galactic velocity [km\,s$^{-1}$]. \\
+\texttt{x\_pc} & -8068.52  & Galactocentric $X$ position coordinate [pc]. \\
+\texttt{y\_pc} & 255.96     & Galactocentric $Y$ position coordinate [pc]. \\
+\texttt{z\_pc} & 86.25      & Galactocentric $Z$ position coordinate [pc]. \\
+\texttt{(BP-RP)0} & -0.115 &  Gaia $G_\mathrm{BP}$-$G_\mathrm{RP}$   color, minus $E$($G_\mathrm{BP}$-$G_\mathrm{RP}$). \\
+\texttt{(M\_G)0} & 0.442 & Absolute $G$-band magnitude, corrected for extinction. \\
+\texttt{cluster} & CH-2 & RSG-5 or CH-2. \\
+\texttt{Prot\_Adopted} & NaN & Adopted rotation period [days]. \\
+\texttt{Prot\_TESS} & NaN & TESS rotation period [days]. \\
+\texttt{Prot\_ZTF}  & NaN & ZTF rotation period [days]. \\
+\texttt{Prot\_Confused} & NaN & Boolean flag; true when stars are photometrically blended. \\
 \enddata
-
-%% Include any \tablenotetext{key}{text}, \tablerefs{ref list},
-%% or \tablecomments{text} between the \enddata and 
-%% \end{deluxetable} commands
-
 %% General table comment marker
-\tablecomments{
-Table~\ref{tab:v06} is published in its entirety in a machine-readable
-format.   This table is a concatenation of the studies listed in
-Table~\ref{tab:metadata}.  One entry is shown for guidance regarding
-form and content.  In this particular example, the star has a cold
-Jupiter on a 16 year orbit, HD 150706b \citep{2012AA...545A..55B}.  An
-infrared excess has been reported \citep{CottenSong2016}, and the star
-was identified by \citet{Ujjwal2020} as a candidate UMa moving group
-member ($\approx 400\,{\rm Myr}$; \citealt{mann_tess_2020}).  The
-star's RV activity and TESS rotation period corroborate its youth.
+\tablecomments{Table~3 is published in its entirety in a machine-readable
+format.  One entry is shown for guidance regarding form and content.
 }

### 2022-04-22 drivers/write_RSG5_CH2_prot_table.py

@@ -1,0 +1,100 @@
+"""
+Write the RSG-5 and CH-2 membership + rotation period table.
+"""
+import os
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from astropy.io import fits
+from astropy.table import Table
+from rudolf.paths import DATADIR, RESULTSDIR, TABLEDIR, LOCALDIR
+from rudolf.helpers import get_ronan_cepher_augmented
+
+# __df: after Ronan's flags
+# df: after Ronan's + my flags
+
+df, __df = get_ronan_cepher_augmented()
+fig1_grey = (df.strengths > 0.02)
+fig1_black = (df.strengths > 0.10)
+sdf = df[fig1_grey]
+
+tabpath = os.path.join(TABLEDIR, "RSG-5_auto_withreddening_gaia2018.csv")
+_df1 = pd.read_csv(tabpath)
+tabpath = os.path.join(TABLEDIR, "CH-2_auto_withreddening_gaia2018.csv")
+_df2 = pd.read_csv(tabpath)
+df1 = pd.concat((_df1,_df2))
+
+
+csvpath = os.path.join(DATADIR, "rotation",
+                       "CepHer_RSG5_EDR3_v2_Curtis_20220326.csv")
+rot_df = pd.read_csv(csvpath)
+
+rot_df['Source_Input'] = rot_df.Source_Input.astype(str)
+df1['source_id'] = df1.source_id.astype(str)
+
+mdf = rot_df.merge(df1, left_on='Source_Input', right_on='source_id',
+                   suffixes=('_rot', '_EDR3'))
+
+mdf = mdf.rename({'source_id_EDR3':'dr3_source_id'}, axis='columns')
+
+reddening_corr = 1
+cstr = '_corr' if reddening_corr else ''
+color0 = 'phot_bp_mean_mag'
+get_yval = (
+    lambda _df: np.array(
+        _df['phot_g_mean_mag'+cstr] + 5*np.log10(_df['parallax_EDR3']/1e3) + 5
+    )
+)
+get_xval = (
+    lambda _df: np.array(
+        _df[color0+cstr] - _df['phot_rp_mean_mag'+cstr]
+    )
+)
+
+mdf['(BP-RP)0'] = get_xval(mdf)
[…]

### 2022-04-22 papers/ms.tex
update ms

@@
 To visualize the results, in Figure~\ref{fig:XYZvtang} we show
-12{,}436 objects with weight exceeding 0.02 as gray points, and
-overplot 4{,}763 objects with weights exceeding 0.10 as black points.
+12{,}436 objects with $\alpha>0.02$ as gray points, and
+overplot 4{,}763 objects with $\alpha>0.10$ as black points.
 The previously known $\delta$~Lyr\ cluster is visible at
@@
 this inspection: Kepler-1627, Kepler-1643, KOI-7368, and
-KOI-7913.\footnote{These figures
-yield an estimate of the size of Cep-Her: if our 4/14
-matches represent the true positive rate, then
-multiplying by the number of stars above our weight threshold
-(12{,}436) yields $\approx$3{,}500 stars in the complex.  
-This might however be an under-count,
-since Kepler observed sparse 
-regions of the complex, which are expected to have higher
-contamination rates (Figure~\ref{fig:XYZvtang}, top panel).}
+KOI-7913.
+%\footnote{These figures
+%yield an estimate of the size of Cep-Her: if our 4/14
+%matches represent the true positive rate, then
+%multiplying by the number of stars above our weight threshold
+%(12{,}436) yields $\approx$3{,}500 stars in the complex.  
+%This might however be an under-count,
+%since Kepler observed sparse 
+%regions of the complex, which are expected to have higher
+%contamination rates (Figure~\ref{fig:XYZvtang}, top panel).}
 

### 2022-04-23 papers/ms.tex
update ms
bug-fix auto cluster viz in fig1, fig2 CAMDs.
still need to propagate to rotation period plots and tables.

@@
 \newcommand{\npms}{1{,}097} % 20220311_Kerr_SPYGLASS205_Members_All.csv
+\newcommand{\nchtwo}{37} % CH-2_auto_XYZ_vl_vb_cut.csv
+\newcommand{\nrsgfive}{173} % RSG-5_auto_XYZ_vl_vb_cut.csv
 
@@
   Y/{\rm pc} &\in [320, 350] \nonumber \\
-  Z/{\rm pc} &\in [40, 70] \nonumber \\
-  v_b/{\rm km\,s^{-1}} &\in [-4, -3] \nonumber \\
-  v_{l^*}/{\rm km\,s^{-1}} &\in [4, 6] \nonumber
+  Z/{\rm pc} &\in [40, 80] \nonumber \\
+  v_b/{\rm km\,s^{-1}} &\in [-4, -2.5] \nonumber \\
+  v_{l^*}/{\rm km\,s^{-1}} &\in [3.5, 6], \nonumber
 \end{align}
+though the cluster does have a greater
+spatial extent toward smaller $X$
+(Figure~\ref{fig:XYZvtang}, middle panels).
 For the diffuse stars near KOI-7368 and KOI-7913, we require
@@
 \end{align}
-and we call this latter set of stars ``CH-2''.  These cuts yield 141
-candidate RSG-5 members, and 37 candidate CH-2 members.
-These stars are listed in Appendix~\ref{app:members}, as are
-the Cep-Her candidates that were observed by
-Kepler.
+and we call this latter set of stars ``CH-2''.  These cuts yield
+\nrsgfive\ candidate RSG-5 members, and \nchtwo\ candidate CH-2
+members.  These stars are listed in Appendix~\ref{app:members}, as are
+the Cep-Her candidates that were observed by Kepler.
 An important consideration, especially for CH-2, is the
@@
 
-The CAMDs show that for RSG-5, nearly all of our candidate members are
+The CAMDs show that for RSG-5, practically all of our candidate members are
 on a tight pre-main-sequence locus.  This implies a false positive
@@
 velocity volumes searched for each group.
-For RSG-5, we drew 141 candidate members from a $30\,{\rm
+For RSG-5, we drew \nrsgfive\ candidate members from a $30\,{\rm
 pc}\times30\,{\rm pc}\times30\,{\rm pc}$ spatial cube, given a $1 {\rm
 \mkms} \times 2 {\rm \mkms }$ rectangle in apparent galactic velocity.
-For CH-2, our 37 candidate members came from a rectangular prism of
+For CH-2, our \nchtwo\ candidate members came from a rectangular prism of
 dimension $50\,{\rm pc}\times40\,{\rm pc}\times30\,{\rm pc}$, and a
@@
 \paragraph{Table~3} contains spatial, kinematic, astrometric, and rotation period
-information for the 141 candidate RSG-5 members and 37 candidate CH-2
+information for the \nrsgfive\ candidate RSG-5 members and \nchtwo\ candidate CH-2
 members described in Section~\ref{subsec:members}.

### 2022-04-24 papers/ms.tex
update ms
propagated updated rotation periods for RSG-5

fix bibliography

@@
   Stellar positions and velocities from Gaia are yielding a new window
-  on
-  open cluster dispersal.  Here we present an analysis of a
-  group stars spanning Cepheus ($l=100^\circ$) to Hercules
-  ($l=40^\circ$), hereafter the Cep-Her complex.
-  The group includes four Kepler Objects of Interest:
+  on open cluster dispersal.  Here we present an analysis of a group
+  stars spanning Cepheus ($l=100^\circ$) to Hercules ($l=40^\circ$),
+  hereafter the Cep-Her complex.  The group includes four Kepler
+  Objects of Interest:
   Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
@@
   Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
-  The latter Neptune-sized planet is in part of the Cep-Her
-  complex called the $\delta$\ Lyr\ cluster
-  \citep{bouma_kep1627_2022}.  Here we focus on the former three
-  systems, which are in other regions of the association.  Based on
-  kinematic evidence from Gaia, stellar rotation periods from TESS,
-  and spectroscopy, these three objects are also $\approx$40 million
-  years (Myr) old.  More specifically, we find that Kepler-1643 is
-  $46^{+9}_{-7}$\,Myr old, based on its membership in a dense
-  sub-cluster of the complex called RSG-5.  KOI-7368 and KOI-7913 are
-  $36^{+10}_{-8}$\,Myr old, and are in a diffuse region that we call
-  CH-2.  Based on the transit shapes and high resolution imaging, all
-  three objects are most likely planets, with false positive
-  probabilities of $6\times10^{-9}$, $4\times10^{-3}$, and
-  $1\times10^{-4}$ for Kepler-1643, KOI-7368, and KOI-7913
-  respectively.  These planets empirically demonstrate that
-  mini-Neptunes with sizes of $\sim$2 Earth radii exist at ages of
-  40 million years.
+  The latter Neptune-sized planet is in part of the Cep-Her complex
+  called the $\delta$\ Lyr\ cluster \citep{bouma_kep1627_2022}.  Here
+  we focus on the former three systems, which are in other regions of
+  the association.  Based on kinematic evidence from Gaia, stellar
+  rotation periods from TESS, and spectroscopy, these three objects
+  are also $\approx$40 million years (Myr) old.  More specifically, we
+  find that Kepler-1643 is $46^{+9}_{-7}$\,Myr old, based on its
+  membership in a dense sub-cluster of the complex called RSG-5.
+  KOI-7368 and KOI-7913 are $36^{+10}_{-8}$\,Myr old, and are in a
+  diffuse region that we call CH-2.  Based on the transit shapes and
+  high resolution imaging, all three objects are most likely planets,
+  with false positive probabilities of $6\times10^{-9}$,
+  $4\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368,
+  and KOI-7913 respectively.  These planets empirically demonstrate
+  that mini-Neptunes with sizes of $\sim$2 Earth radii exist at ages
+  of 40 million years.
 \end{abstract}
@@
 
-The discovery and characterization of planets younger than
-a billion years is a major frontier in current exoplanet research.
-The reason is that the properties of young planets provide benchmarks
-for studies of planetary evolution.  For instance, young planets can
-inform our understanding
-of when hot Jupiters arrive on their close-in orbits
+The discovery and characterization of planets younger than a billion
+years is a major frontier in current exoplanet research.  The reason
+is that the properties of young planets provide benchmarks for studies
+of planetary evolution.  For instance, young planets can inform our
+understanding of when hot Jupiters arrive on their close-in orbits
 \citep{dawson_johnson_2018}, how the sizes of planets with massive
@@
 
-The discovery of a young planet requires two claims to be true:
-the planet must exist, and its age must be secured.  Spaced-based
+The discovery of a young planet requires two claims to be true: the
+planet must exist, and its age must be secured.  Spaced-based
 photometry from K2 and TESS has yielded a number of exemplars for
@@
 In this work, we leverage recent analyses of the Gaia data, which have
-greatly expanded our knowledge of stellar groups \citep[{e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To date these
-analyses have mostly clustered on stellar positions and 2D
-velocities.  One important result has been the discovery of
-diffuse streams and tidal tails comparable in stellar mass to the
-previously known cores of nearby open clusters
+greatly expanded our knowledge of stellar groups
+\citep[{e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To
+date these analyses have mostly clustered on stellar positions and 2D
+velocities.  One important result has been the discovery of diffuse
+streams and tidal tails comparable in stellar mass to the previously
+known cores of nearby open clusters
 \citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  Even
-though these streams are spread over tens to hundreds of
-parsecs, their velocity dispersions can remain coherent at the
-$\sim$1\kms\ level.  Internal dynamics and projection effects can also
-drive them to be much larger: in the Hyades,
-stars in the tidal tails are expected to span up to $\pm 40\mkms$ in velocity
-relative to the cluster center \citep{jerabkova_800_2021}.
-The stars in such diffuse regions can be verified to be the same age
-as the core cluster members through analyses of
-color-absolute magnitude diagrams \citep{KounkelCovey2019}, stellar
-rotation periods \citep{curtis_tess_2019,bouma_2021_ngc2516}, and
-chemical abundances \citep{hawkins_2020}.  While there are 
-implications for our understanding of star formation and cluster
-evolution \citep{dinnbier_tidal_2020}, a separate consequence is that
-we now know the ages of many more stars, including previously known
-planet hosts.
+though these streams are spread over tens to hundreds of parsecs,
+their velocity dispersions can remain coherent at the $\sim$1\kms\
+level.  Internal dynamics and projection effects can also drive them
+to be much larger: in the Hyades, stars in the tidal tails are
+expected to span up to $\pm 40\mkms$ in velocity relative to the
+cluster center \citep{jerabkova_800_2021}.  The stars in such diffuse
+regions can be verified to be the same age as the core cluster members
+through analyses of color-absolute magnitude diagrams
+\citep{KounkelCovey2019}, stellar rotation periods
+\citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
+\citep{hawkins_2020}.  While there are implications for our
+understanding of star formation and cluster evolution
+\citep{dinnbier_tidal_2020}, a separate consequence is that we now
+know the ages of many more stars, including previously known planet
+hosts.
 
@@
 
-Here, we expand on our earlier study of a $38^{+7}_{-6}$ Myr
-old Neptune-sized planet in the Kepler field (Kepler-1627~Ab;
-\citealt{bouma_kep1627_2022}).  This planet's age was derived
-based on its host star's membership in the $\delta$\ Lyr\ cluster.
-While our analysis of the cluster focused on the immediate vicinity of
-Kepler-1627 in order to have a reasonable scope,  it became
-clear that the $\delta$\ Lyr\ cluster seems to
-also be part of a much larger group of similarly aged stars.  This
-group, which is at a distance of $\sim$330\,pc from the Sun, 
-spans Cepheus to Hercules (galactic longitudes, $l$, between
-40$^\circ$ and 100$^\circ$), at galactic latitudes between
-0$^\circ$ and 20$^\circ$.  We therefore refer to it as the Cep-Her
-complex.  It exhibits significant sub-structure over its $\approx$250
-parsec length, and a detailed analysis of its memberships, kinematics,
-and possible origin is currently being prepared by R.~Kerr and
-collaborators.
+Here, we expand on our earlier study of a $38^{+7}_{-6}$ Myr old
+Neptune-sized planet in the Kepler field (Kepler-1627~Ab;
+\citealt{bouma_kep1627_2022}).  This planet's age was derived based on
+its host star's membership in the $\delta$\ Lyr\ cluster.  While our
+analysis of the cluster focused on the immediate vicinity of
+Kepler-1627 in order to have a reasonable scope,  it became clear that
+the $\delta$\ Lyr\ cluster seems to also be part of a much larger
+group of similarly aged stars.  This group, which is at a distance of
+$\sim$330\,pc from the Sun, spans Cepheus to Hercules (galactic
+longitudes, $l$, between 40$^\circ$ and 100$^\circ$), at galactic
+latitudes between 0$^\circ$ and 20$^\circ$.  We therefore refer to it
+as the Cep-Her complex.  It exhibits significant sub-structure over
+its $\approx$250 parsec length, and a detailed analysis of its
+memberships, kinematics, and possible origin is currently being
+prepared by R.~Kerr and collaborators.
 
@@
 and KOI-7913.  Given our earlier analysis of Kepler-1627, we focus
-here on the latter three objects.  After analyzing the relevant properties of
-Cep-Her (Section~\ref{sec:cluster}), we derive the stellar properties
-(Section~\ref{sec:stars}) and validate the planetary nature of each
-system using a combination of the Kepler photometry and
-high-resolution imaging (Section~\ref{sec:planets}).  We conclude with
-a discussion of mini-Neptune size evolution, and point out possible
-directions for future work (Section~\ref{sec:disc_conc}).
+here on the latter three objects.  After analyzing the relevant
+properties of Cep-Her (Section~\ref{sec:cluster}), we derive the
+stellar properties (Section~\ref{sec:stars}) and validate the
+planetary nature of each system using a combination of the Kepler
+photometry and high-resolution imaging (Section~\ref{sec:planets}).
+We conclude with a discussion of mini-Neptune size evolution, and
+point out possible directions for future work
+(Section~\ref{sec:disc_conc}).
 
@@
 from the Sun, above the galactic plane, and spanning galactic
-longitudes of roughly $40^\circ$ to $100^\circ$.  Two rich
-clusters in this region are the $\delta$~Lyra cluster
+longitudes of roughly $40^\circ$ to $100^\circ$.  Two rich clusters in
+this region are the $\delta$~Lyra cluster
 \citep{stephenson_possible_1959} and RSG-5 \citep{roser_nine_2016}.
-Each of these clusters was known before Gaia.  Their reported ages
-are between 30 and 60\,Myr.  Early empirical
-evidence that these two clusters could be part of a large and more
-diffuse population was apparent in the Gaia-based photometric analysis
-of pre-main-sequence stars by \citet[][see their Figures~11
-and~13]{Zari2018}.  Further kinematic connections and complexity were
-highlighted by \citet{KounkelCovey2019}, who included these previously
-known groups in the larger structures dubbed ``Theia~73'' and
+Each of these clusters was known before Gaia.  Their reported ages are
+between 30 and 60\,Myr.  Early empirical evidence that these two
+clusters could be part of a large and more diffuse population was
+apparent in the Gaia-based photometric analysis of pre-main-sequence
+stars by \citet[][see their Figures~11 and~13]{Zari2018}.  Further
+kinematic connections and complexity were highlighted by
+\citet{KounkelCovey2019}, who included these previously known groups
+in the larger structures dubbed ``Theia~73'' and
 ``Theia~96''\footnote{See their visualization online at
@@
 and on-sky velocities with a subsequent manual ``stitching'' step.
-Their results support the idea that there is an overdensity of
-30 to 60\,Myr old stars in this region of
-the Galaxy.  \citet{Kerr2021}, in a volume-limited analysis of the
-Gaia DR2 point-source catalog out to one third of a kiloparsec,
-identified three of the nearest sub-populations, dubbed
-``Cepheus-Cygnus'', ``Lyra'', and ``Cerberus''.  \citet{Kerr2021}
-reported ages for each of these subgroups between 30 and 35 Myr.
+Their results support the idea that there is an overdensity of 30 to
+60\,Myr old stars in this region of the Galaxy.  \citet{Kerr2021}, in
+a volume-limited analysis of the Gaia DR2 point-source catalog out to
+one third of a kiloparsec, identified three of the nearest
+sub-populations, dubbed ``Cepheus-Cygnus'', ``Lyra'', and
+``Cerberus''.  \citet{Kerr2021} reported ages for each of these
+subgroups between 30 and 35 Myr.
 
@@
 To visualize the results, in Figure~\ref{fig:XYZvtang} we show
-12{,}436 objects with $\alpha>0.02$ as gray points, and
-overplot 4{,}763 objects with $\alpha>0.10$ as black points.
-The previously known $\delta$~Lyr\ cluster is visible at
-$(l,b)=(68^\circ,15^\circ$) and $(v_{l'}, v_b)=(-4.5 ,-4 )\mkms$.
-RSG-5 is visible at $(l,b)=(83^\circ,6^\circ)$, $(v_{l'}, v_b)=(5.5
-,-3.5 )\mkms$.  Most of the other subclusters, including in Cep-Cyg
+12{,}436 objects with $\alpha>0.02$ as gray points, and overplot
+4{,}763 objects with $\alpha>0.10$ as black points.  The previously
+known $\delta$~Lyr\ cluster is visible at $(l,b)=(68^\circ,15^\circ$)
+and $(v_{l'}, v_b)=(-4.5 ,-4 )\mkms$.  RSG-5 is visible at
+$(l,b)=(83^\circ,6^\circ)$, $(v_{l'}, v_b)=(5.5 ,-3.5 )\mkms$.  Most
+of the other subclusters, including in Cep-Cyg
 ($l,b=90^\circ,7^\circ$) and Cerberus ($l,b=48^\circ,18^\circ$) are
@@
 2022, and also compared against the \texttt{q1\_q17\_dr25} table
-\citep{thompson_planetary_2018}.  From the candidate members
-with weights exceeding 0.02, this 
- yielded 11 known false positives, 6 ``confirmed'' planets, and 8
- ``candidate'' planets.  
-Inspection of the Kepler data validation summaries and Robovetter
-classifications for these objects showed whether they were
-potentially consistent with being {\it i)} planets, and {\it ii)}
-$\lesssim 10^8$ years old, based on the presence of rotational
-modulation at the expected period and amplitude \citep[{e.g.},][Figure~9]{rebull_rotation_2020}.  Four objects remained after
-this inspection: Kepler-1627, Kepler-1643, KOI-7368, and
-KOI-7913.
+\citep{thompson_planetary_2018}.  From the candidate members with
+weights exceeding 0.02, this yielded 11 known false positives, 6
+``confirmed'' planets, and 8 ``candidate'' planets.  Inspection of the
+Kepler data validation summaries and Robovetter classifications for
+these objects showed whether they were potentially consistent with
+being {\it i)} planets, and {\it ii)} $\lesssim 10^8$ years old, based
+on the presence of rotational modulation at the expected period and
+amplitude \citep[{e.g.},][Figure~9]{rebull_rotation_2020}.  Four
+objects remained after this inspection: Kepler-1627, Kepler-1643,
+KOI-7368, and KOI-7913.
 %\footnote{These figures
@@
 closer to the Sun in $Y$.  In tangential galactic velocity space,
-there is some kinematic overlap between the region the
-latter two KOIs are in and the main RSG-5 group.
+there is some kinematic overlap between the region the latter two KOIs
+are in and the main RSG-5 group.
 
@@
 \end{align}
-though the cluster does have a greater
-spatial extent toward smaller $X$
-(Figure~\ref{fig:XYZvtang}, middle panels).
-For the diffuse stars near KOI-7368 and KOI-7913, we require
+though the cluster does have a greater spatial extent toward smaller
+$X$ (Figure~\ref{fig:XYZvtang}, middle panels).  For the diffuse stars
+near KOI-7368 and KOI-7913, we require
 \begin{align}
@@
 members.  These stars are listed in Appendix~\ref{app:members}, as are
-the Cep-Her candidates that were observed by Kepler.
-An important consideration, especially for CH-2, is the
-contamination rate by field stars.  We assess this in the following
-section.
+the Cep-Her candidates that were observed by Kepler.  An important
+consideration, especially for CH-2, is the contamination rate by field
+stars.  We assess this in the following section.
 
@@
     are 40 to 50 million years old.} 
-    {\it Top row}: 
-    Color-absolute magnitude diagram of candidate Cep-Her members, plotted
-    over candidate members of the $\delta$~Lyr~cluster
-    ($\approx38$\,Myr; \citealt{bouma_kep1627_2022}) and the Gaia
-    EDR3 Catalog of Nearby Stars (gray background).  The left and
-    right columns shows stars in RSG-5 and CH-2, respectively.  The
-    range of colors is truncated to emphasize the pre-main-sequence.
-    Stars that fall far below the cluster sequences are field
-    interlopers.
-    {\it Bottom row}:
-    TESS and ZTF-derived stellar rotation periods, with the Pleiades
-    ($\approx 112$\,Myr) and Praesepe ($\approx 650$\,Myr) shown for reference
-    \citep{rebull_rotation_2016a,douglas_poking_2017}.
-    The detection efficiency for reliable rotation periods falls off
+    {\it Top row}: Color-absolute magnitude diagram of candidate
+    Cep-Her members, plotted over candidate members of the
+    $\delta$~Lyr~cluster ($\approx38$\,Myr;
+    \citealt{bouma_kep1627_2022}) and the Gaia EDR3 Catalog of Nearby
+    Stars (gray background).  The left and right columns shows stars
+    in RSG-5 and CH-2, respectively.  The range of colors is truncated
+    to emphasize the pre-main-sequence.  Stars that fall far below the
+    cluster sequences are field interlopers.  {\it Bottom row}: TESS
+    and ZTF-derived stellar rotation periods, with the Pleiades
+    ($\approx 112$\,Myr) and Praesepe ($\approx 650$\,Myr) shown for
+    reference \citep{rebull_rotation_2016a,douglas_poking_2017}.  The
+    detection efficiency for reliable rotation periods falls off
     beyond $\bpmrpo \gtrsim 2.6$.
@@
 diagrams, we imposed the data filtering criteria from
-\citet[][Appendix B]{GaiaCollaboration2018}, which 
-include binaries while omitting instrumental artifacts from for
-instance low photometric signal to noise, or a small number of
-visibility periods.  We then corrected for extinction using the
-\citet{lallement_threedimensional_2018}
-dust maps\footnote{\url{https://stilism.obspm.fr/}} and the extinction coefficients $k_X\equiv A_X/A_0$ from
-\citet{GaiaCollaboration2018}, assuming that $A_0 = 3.1 E(B-V)$.  This
-yielded a mean and standard deviation for the reddening of
-$E(B-V)=0.036\pm0.002$ for RSG-5, and $E(B-V)=0.017\pm0.001$
-for CH-2.  By way of comparison, in \citet{bouma_kep1627_2022} the
-same query for the $\delta$~Lyr cluster yielded
-$E(B-V)=0.032\pm0.006$.  Finally, for the plots we set the
+\citet[][Appendix B]{GaiaCollaboration2018}, which include binaries
+while omitting instrumental artifacts from for instance low
+photometric signal to noise, or a small number of visibility periods.
+We then corrected for extinction using the
+\citet{lallement_threedimensional_2018} dust
+maps and the extinction
+coefficients $k_X\equiv A_X/A_0$ from \citet{GaiaCollaboration2018},
+assuming that $A_0 = 3.1 E(B-V)$.  This yielded a mean and standard
+deviation for the reddening of $E(B-V)=0.036\pm0.002$ for RSG-5, and
+$E(B-V)=0.017\pm0.001$ for CH-2.  By way of comparison, in
+\citet{bouma_kep1627_2022} the same query for the $\delta$~Lyr cluster
+yielded $E(B-V)=0.032\pm0.006$.  Finally, for the plots we set the
 color range to best visualize the region of maximal age information
@@
 
-The CAMDs show that for RSG-5, practically all of our candidate members are
-on a tight pre-main-sequence locus.  This implies a false positive
-rate of a few percent, at most.  By comparison, our control sample
-(the $\delta$~Lyr candidates) has a false positive rate of
+The CAMDs show that for RSG-5, practically all of our candidate
+members are on a tight pre-main-sequence locus.  This implies a false
+positive rate of a few percent, at most.  By comparison, our control
+sample (the $\delta$~Lyr candidates) has a false positive rate of
 $\approx$12\%, based on the number of stars that photometrically
-overlap with the field population, while being outliers from the bulk cluster
-population.  For CH-2, our membership selection gives 27 objects in
-the color range displayed, and 23 of them appear to be consistent with
-being on the pre-main-sequence.  This implies a false positive rate in
-CH-2 of $\approx$15\%.
-
-Figure~\ref{fig:age} also shows that most RSG-5 and
-CH-2 members overlap with the $\delta$~Lyr cluster, and that the groups are therefore
-roughly the same age.  To quantify this, we use the method
-introduced by \citet[][their Section~6.3]{gagne_mutau_2020}.  The
-idea is to fit the pre-main-sequence loci of a set of
-reference clusters, and to then model the locus of the target cluster
-as a linear combination of these reference cluster loci.  For our
-reference clusters, we used UCL, IC\,2602, and the Pleiades, with the
+overlap with the field population, while being outliers from the bulk
+cluster population.  For CH-2, our membership selection gives 27
+objects in the color range displayed, and 23 of them appear to be
+consistent with being on the pre-main-sequence.  This implies a false
+positive rate in CH-2 of $\approx$15\%.
+
+Figure~\ref{fig:age} also shows that most RSG-5 and CH-2 members
+overlap with the $\delta$~Lyr cluster, and that the groups are
+therefore roughly the same age.  To quantify this, we use the method
+introduced by \citet[][their Section~6.3]{gagne_mutau_2020}.  The idea
+is to fit the pre-main-sequence loci of a set of reference clusters,
+and to then model the locus of the target cluster as a linear
+combination of these reference cluster loci.  For our reference
+clusters, we used UCL, IC\,2602, and the Pleiades, with the
 memberships reported by \citet{Damiani2019} and
 \cite{CantatGaudin2018a} respectively.  We adopted ages of 16\,Myr for
-UCL \citep{pecaut_star_2016}, 38\,Myr for IC\,2602\footnote{ Ages for
-IC\,2602 vary from 40 to 46\,Myr based on lithium-depletion-boundary
-(LDB) measurements \citep{dobbie_ic_2010,randich_gaiaeso_2018}, and
-from 30 to 46\,Myr based on isochronal analyses
-\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.},
+UCL \citep{pecaut_star_2016}, 38\,Myr for IC\,2602
+\citep{david_ages_2015,randich_gaiaeso_2018}
+%\footnote{ Ages for
+%IC\,2602 vary from 40 to 46\,Myr based on lithium-depletion-boundary
+%(LDB) measurements \citep{dobbie_ic_2010,randich_gaiaeso_2018}, and
+%from 30 to 46\,Myr based on isochronal analyses
+%\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.},
 and 112\,Myr for the Pleiades \citep{dahm_2015}.  These assumptions
-and the subsequent processing steps taken to exclude field stars and binaries
-were identical to those
-described in \citet{bouma_kep1627_2022}.  The mean and 
-uncertainty of the resulting age posterior are $46^{+9}_{-7}$\,Myr for
-RSG-5, and $36^{+10}_{-8}$\,Myr for CH-2.  For comparison, this
-procedure yields an age for the $\delta$~Lyr cluster of
-$38^{+6}_{-5}$\,Myr.  
-The slightly older isochronal age of RSG-5 is
-consistent with its location
-relative to the $\delta$~Lyr cluster in
-the upper left panel of Figure~\ref{fig:age}.
+and the subsequent processing steps taken to exclude field stars and
+binaries were identical to those described in
+\citet{bouma_kep1627_2022}.  The mean and uncertainty of the resulting
+age posterior are $46^{+9}_{-7}$\,Myr for RSG-5, and
+$36^{+10}_{-8}$\,Myr for CH-2.  For comparison, this procedure yields
+an age for the $\delta$~Lyr cluster of $38^{+6}_{-5}$\,Myr.  The
+slightly older isochronal age of RSG-5 is consistent with its location
+relative to the $\delta$~Lyr cluster in the upper left panel of
+Figure~\ref{fig:age}.
 
@@
 achieved using surveys such as TESS \citep{ricker_transiting_2015} and
-the Zwicky Transient Facility (ZTF, \citealt{bellm_zwicky_2019});
-it leverages a storied tradition of measuring rotation periods of stars in
-benchmark open clusters \citep[see {e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data in
+the Zwicky Transient Facility (ZTF, \citealt{bellm_zwicky_2019}); it
+leverages a storied tradition of measuring rotation periods of stars
+in benchmark open clusters \citep[see
+{e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data in
 our case are especially useful, since they provide 3 to 5 lunar months
@@
 ($\sim$M3V) at the relevant distances.  These cuts gave 19 stars in
-CH-2 and 42 stars in RSG-5.  We extracted light curves from the TESS
-images using the \texttt{unpopular} package
-\citep{hattorio_2021_cpm}, and regressed them against systematics with
-its causal pixel model.  We measured rotation periods using
-Lomb-Scargle periodograms and visually vetted the results using an
-interactive program that allows us to switch between TESS Cycles,
-select particular sectors, flag stars with multiple periods, and
-correct half-period harmonics. For ZTF, we used the same color cut to
-focus on FGKM stars, but restricted the sample to $13 < G < 18$ to
-avoid the saturation limit on the bright end and ensure sufficient
-photometric precision at the faint end. We followed the procedure
-outlined in \citet{curtis_rup147_2020}: we downloaded $8'\times8'$
-image cutouts, ran aperture photometry for the target and neighboring
-stars identified with Gaia, and used them to define a systematics
-correction to refine the target light curves.
+CH-2 and 47 stars in RSG-5.  We extracted light curves from the TESS
+images using the \texttt{unpopular} package \citep{hattorio_2021_cpm},
+and regressed them against systematics with its causal pixel model.
+We measured rotation periods using Lomb-Scargle periodograms and
+visually vetted the results using an interactive program that allows
+us to switch between TESS Cycles, select particular sectors, flag
+stars with multiple periods, and correct half-period harmonics. For
+ZTF, we used the same color cut to focus on FGKM stars, but restricted
+the sample to $13 < G < 18$ to avoid the saturation limit on the
+bright end and ensure sufficient photometric precision at the faint
+end. We followed the procedure outlined in \citet{curtis_rup147_2020}:
+we downloaded $8'\times8'$ image cutouts, ran aperture photometry for
+the target and neighboring stars identified with Gaia, and used them
+to define a systematics correction to refine the target light curves.
 
 The lower panels of Figure~\ref{fig:age} show the results.  In RSG-5,
-36/42 stars have rotation periods faster than the Pleiades (86\%).
-This numerator omits the two stars with periods $>$$12$\,days visible in
-the lower-left panel of Figure~\ref{fig:age}.  The age interpretation
-for these latter stars, particularly the $\approx$M2.5 dwarf, is not
-obvious.  \citet{rebull_usco_2018} for instance have found numerous
-M-dwarfs with 10-12 day rotation periods at ages of USco ($\sim$$8$\,Myr),
-and some may still exist at ages of LCC ($\sim$$16$\,Myr;
-L.~Rebull in preparation).  Regardless, given that nearly no field
-star outliers seem to be present on the RSG-5 CAMD, the fact that we
-do not detect rotation periods for $\approx$14\% of stars should
-perhaps be taken as an indication for the fraction of stars for
-which rotation periods might not be detectable, due to {e.g.}, pole-on
-stars having lower amplitude starspot modulation.
-
-%
-% CH-2
-% 5 stars with Prot>10 days (from the Prot versus color plot):
-%                      DESIGNATION  TESS_Data   bp_rp_x     period
-%25  Gaia EDR3 2129930258400157440          4  1.909693  15.309000
-%24  Gaia EDR3 2129417404945830784          4  2.040988  10.981582
-%4   Gaia EDR3 2107169165113864064          4  2.059904  10.580494
-%9   Gaia EDR3 2134851775526125696          4  2.326998  12.227900
-%1   Gaia EDR3 2127726081184588160          4  2.347244  11.543281
-%
-% CH-2
-% Photometric field stars (from Glue -- one of these is omitted due to phot S/N cut, 
-% presumably 2126008609657702784 )
-%
-%In [9]: df[['DESIGNATION', 'bp_rp', 'M_G']]
-%Out[9]:
-%                     DESIGNATION     bp_rp        M_G
-%0  Gaia EDR3 2126529468937923456  2.518077   9.844733 --> NaN period, G=16.9
-%1  Gaia EDR3 2126008609657702784  2.746735  11.409737 --> NaN, G=18.4
-%2  Gaia EDR3 2133330051428541952  2.499458   9.459546 --> 15.7 d (ZTF), G=16.5
-%3  Gaia EDR3 2133510921093964928  2.720177  10.923209 --> NaN, G=17.9
-%4  Gaia EDR3 2139369183468814464  2.929506  11.170221 --> NaN, G=18.2
-%5  Gaia EDR3 2129930258400157440  1.909693   8.200652 --> P=15.3 (ZTF), G=15.2
+41/47 stars have rotation periods faster than the Pleiades (87\%).
+This numerator omits the two stars with periods $>$$12$\,days visible
+in the lower-left panel of Figure~\ref{fig:age}.  The age
+interpretation for these latter stars, particularly the $\approx$M2.5
+dwarf, is not obvious.  \citet{rebull_usco_2018} for instance have
+found numerous M-dwarfs with 10-12 day rotation periods at ages of
+USco ($\sim$$8$\,Myr), and some may still exist at ages of LCC
+($\sim$$16$\,Myr; L.~Rebull in preparation).  Regardless, given that
+nearly no field star outliers seem to be present on the RSG-5 CAMD,
+the fact that we do not detect rotation periods for $\approx$13\% of
+stars should perhaps be taken as an indication for the fraction of
+stars for which rotation periods might not be detectable, due to
+{e.g.}, pole-on stars having lower amplitude starspot modulation.
 
@@
 rate for CH-2 determined from the CAMD seems consistent with our
-fraction of detected rotation periods; RSG-5 was missing
-rotation periods for $\approx$15\% of its candidate members, even
-though all of its stars appear photometrically consistent with being
-on a single pre-main-sequence locus.
+fraction of detected rotation periods; RSG-5 was missing rotation
+periods for $\approx$15\% of its candidate members, even though all of
+its stars appear photometrically consistent with being on a single
+pre-main-sequence locus.
 
@@
 G8V (Kepler-1627) to K6V (KOI-7913 A).  The secondary in the KOI-7913
-system has spectral type $\approx$K8V.  And given that a star with Solar mass
-and metallicity arrives at the zero-age main sequence at
+system has spectral type $\approx$K8V.  And given that a star with
+Solar mass and metallicity arrives at the zero-age main sequence at
 $t\approx40$ Myr \citep{choi_mesa_2016}, these stars are all in the
 late stages of their pre-main-sequence contraction.
-% Their heliocentric distances are $\approx$260\,pc (KOI-7913 and
-% KOI-7368, in CH-2) and $\approx$340\,pc (Kepler-1643, in RSG-5).
 
@@
 To verify these parameters and to also analyze spectroscopic youth
-proxies such as the Li 6708\,\AA\ doublet and H$\alpha$,
-we acquired spectra.  We also acquired high resolution imaging for
-each system, to constrain the existence of visual companions,
-including possible bound binaries.  The system-by-system details
-follow, and the relevant results are also summarized in
-Table~\ref{tab:sysparams}.
+proxies such as the Li 6708\,\AA\ doublet and H$\alpha$, we acquired
+spectra.  We also acquired high resolution imaging for each system, to
+constrain the existence of visual companions, including possible bound
+binaries.  The system-by-system details follow, and the relevant
+results are also summarized in Table~\ref{tab:sysparams}.
 
@@
 found $v\sin i = 9.3 \pm 1.0\,\mkms$.  The systemic radial velocity at
-the two epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm
-1.2\,\mkms$ respectively.  To infer the equivalent width of the
-Li\,\textsc{I} 6708\,\AA\ doublet, we followed the procedure described
-by \citet{bouma_2021_ngc2516}.  This yielded a strong detection: ${\rm
+the two epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm 1.2\,\mkms$
+respectively.  To infer the equivalent width of the Li\,\textsc{I}
+6708\,\AA\ doublet, we followed the procedure described by
+\citet{bouma_2021_ngc2516}.  This yielded a strong detection: ${\rm
 EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with values consistent at
@@
 correct for the Fe\,\textsc{I} blend at 6707.44\AA, because HIRES
-resolves the two features.  
-Given the purported age and effective temperature of the star, the
-lithium equivalent width is somewhat low.  We discuss this in greater
-depth in Appendix~\ref{app:spectra}.
+resolves the two features.  Given the purported age and effective
+temperature of the star, the lithium equivalent width is somewhat low.
+We discuss this in greater depth in Appendix~\ref{app:spectra}.
 
@@
 Tillinghast 1.5m at the Fred Lawrence Whipple Observatory.  The
-Stellar Parameter Classification pipeline for TRES has
-been described by \citet{2021tsc2.confE.124B}.  It is based on the
-synthetic template library constructed by
-\citet{Buchhave2012}.  The resulting stellar parameters
-($T_{\rm eff}, \log g, R_\star$) agreed with those from the
+Stellar Parameter Classification pipeline for TRES has been described
+by \citet{2021tsc2.confE.124B}.  It is based on the synthetic template
+library constructed by \citet{Buchhave2012}.  The resulting stellar
+parameters ($T_{\rm eff}, \log g, R_\star$) agreed with those from the
 cluster-isochrone method within $1$-$\sigma$, though the effective
@@
 13, and KOI-7913 B on the night of 2021 Oct 26.  The
-\texttt{SpecMatch-Emp} machinery yielded $T_{\rm
-eff,A} = 4324 \pm 70\,{\rm K}$, and $T_{\rm eff,B} = 4038 \pm 70\,{\rm
-K}$.  The remaining parameters were in agreement with those from the
-cluster isochrone.  For the primary, we also found $[{\rm Fe/H}]=
--0.06 \pm 0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm
-sys} = -17.8 \pm 1.1\,\mkms$.  For the secondary, these same
-parameters were $[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm
-1.0\,\mkms$, and ${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.  Neither
-of the KOI-7913 components shows strong lithium absorption, but this
-is expected given their $\approx$K6V and $\approx$K8V spectral types.
-They do however both show H$\alpha$ in emission.  Broadly speaking,
-both observations are consistent with a $\approx$40 Myr age for
-KOI-7913 (see Appendix~\ref{app:spectra}).
+\texttt{SpecMatch-Emp} machinery yielded $T_{\rm eff,A} = 4324 \pm
+70\,{\rm K}$, and $T_{\rm eff,B} = 4038 \pm 70\,{\rm K}$.  The
+remaining parameters were in agreement with those from the cluster
+isochrone.  For the primary, we also found $[{\rm Fe/H}]= -0.06 \pm
+0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm sys} =
+-17.8 \pm 1.1\,\mkms$.  For the secondary, these same parameters were
+$[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm 1.0\,\mkms$, and
+${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.  Neither of the KOI-7913
+components shows strong lithium absorption, but this is expected given
+their $\approx$K6V and $\approx$K8V spectral types.  They do however
+both show H$\alpha$ in emission.  Broadly speaking, both observations
+are consistent with a $\approx$40 Myr age for KOI-7913 (see
+Appendix~\ref{app:spectra}).
 
@@
     Interest in Cep-Her.}   
-    {\it Top}:
-    50 day light curve segment from the 3.9 years of Kepler data.  The
-    ordinate shows the \texttt{PDCSAP} median-subtracted flux in units
-    of parts-per-thousand ($\times 10^{-3}$).  The dominant signal is
-    induced by starspots.  
-    {\it Bottom}:
+    {\it Top}: 50 day light curve segment from the 3.9 years of Kepler
+    data.  The ordinate shows the \texttt{PDCSAP} median-subtracted
+    flux in units of parts-per-thousand ($\times 10^{-3}$).  The
+    dominant signal is induced by starspots.  {\it Bottom}:
     Phase-folded transits of Kepler-1643, KOI-7913, KOI-7368, and
@@
 
-We fitted the data as follows.  Given an initial guess for the transit
+We fitted the data as follows.  Given the transit
 ephemeris from \citet{thompson_planetary_2018}, we first trimmed the
@@
 out-of-transit points in each local window were then fitted with a
-fourth-order polynomial, which was divided out from the light
-curve.  The resulting flattened transits were then fitted with a
-transit model that assumed quadratic limb darkening.
-The model therefore included 8 free parameters for the transit ($\{P,
-t_0, \log R_{\rm p}/R_\star, b, u_1 ,u_2, R_\star, \log g\}$), 2 free
-parameters for the light curve normalization and a white noise jitter
-($\{\langle f \rangle, \sigma_f \}$), and 5 fixed
-parameters for each transit.
+fourth-order polynomial, which was divided out from the light curve.
+The resulting flattened transits were then fitted with a transit model
+that assumed quadratic limb darkening.  The model therefore included 8
+free parameters for the transit ($\{P, t_0, \log R_{\rm p}/R_\star, b,
+u_1 ,u_2, R_\star, \log g\}$), 2 free parameters for the light curve
+normalization and a white noise jitter ($\{\langle f \rangle, \sigma_f
+\}$), and 5 fixed parameters for each transit.
 
@@
 
-One drawback of our approach is that we have fixed 5 implicit
-parameters per individual transit to their MAP values to remove the
-starspot-induced variability.  An alternative could be to fit the
+A potential drawback of our approach is that to remove
+the starspot-induced variability, we fixed 5 
+parameters per transit to their MAP values.
+An alternative could be to fit the
 planetary transits simultaneously with the starspot-induced
-variability using a quasiperiodic Gaussian process (GP).  We
-explored this approach, but found that it sometimes required fine-tuning
-of the hyperparameter priors, otherwise the GP would 
-overfit flares, instrumental noise, and any other variability that was not
-part of the transit.  This failure mode is pernicious, in that it yields an
-ill-founded sense of confidence in a visually clean
-fit.  Our model is simpler, and it has the benefit that the white
-noise jitter never trades off with any parameter equivalent to a
-damping timescale for the coherence of the GP.  It is also
-computationally efficient, and it captures the planetary parameters
-about which we care the most.
+variability using a quasiperiodic Gaussian process (GP).  We explored
+this approach, but found that it sometimes required fine-tuning of the
+hyperparameter priors, otherwise the GP would overfit flares,
+instrumental noise, and any other variability that was not part of the
+transit.  This failure mode is pernicious, in that it yields an
+ill-founded sense of confidence in a visually clean fit.  Our model is
+simpler, and it has the benefit that the white noise jitter never
+trades off with any parameter equivalent to a damping timescale for
+the coherence of the GP.  It is also computationally efficient, and it
+captures the planetary parameters about which we care the most.
 
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f4.pdf}
+		\includegraphics[width=0.8\textwidth]{f4.pdf}
 	\end{center}
@@
 		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
-    Planets younger than a gigayear with ages more precise than a factor of
-    three are emphasized. The Cep-Her planets are
+    Planets younger than a gigayear with ages more precise than a
+    factor of three are emphasized. The Cep-Her planets are
     Kepler-1643~b ($\square$), KOI-7368~b ($\triangledown$),
-    KOI-7913~Ab (X), and Kepler-1627~Ab (+).  
-    Interesting trends in the population of young planets currently include
-    {\it i)} their large sizes 
-    and {\it ii)} the lack of young hot Jupiters.
-    The new objects of interest in Cep-Her have normal mini-Neptune sizes
+    KOI-7913~Ab (X), and Kepler-1627~Ab (+).  Interesting trends in
+    the population of young planets currently include {\it i)} their
+    large sizes and {\it ii)} the lack of young hot Jupiters.  The new
+    objects of interest in Cep-Her have normal mini-Neptune sizes
     between 2 and 3\,$R_\oplus$, which is a novelty given their ages.
@@
 
-The most significant novelty about
-Kepler-1643~b, KOI-7368.01, and KOI-7913~Ab is that their sizes are
-normal, relative to the known population of gigayear-old
-mini-Neptunes.  At field star ages, mini-Neptune sizes span
-1.8\,$R_\oplus$ to 3.6\,$R_\oplus$, with the most
-common size being $\approx 2.4\,R_\oplus$
+The most significant novelty about Kepler-1643~b, KOI-7368.01, and
+KOI-7913~Ab is that their sizes are normal, relative to the known
+population of gigayear-old mini-Neptunes.  At field star ages,
+mini-Neptune sizes span 1.8\,$R_\oplus$ to 3.6\,$R_\oplus$, with the
+most common size being $\approx 2.4\,R_\oplus$
 \citep[][Figure~7]{Fulton_et_al_2017}.  The known planets younger than
@@
 10$\,R_\oplus$
-\citep{Mann_K2_33b_2016,David_et_al_2017,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020,mann_toi1227_2022}.
+\citep{Mann_K2_33b_2016,David_et_al_2016,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020,mann_toi1227_2022}.
 Figure~\ref{fig:rp_period_age} explores this in more detail by showing
 the sizes, orbital periods, and ages of the known transiting planets,
-emphasizing planets with precise ages.  The smallest previously known planets
-comparable to the new Cep-Her mini-Neptunes
-are AU~Mic~c ($3.0\pm0.2\,R_\oplus$, see
-\citealt{martioli_aumicbc_2021} and \citealt{gilbert_flares_2022}),
-Kepler-1627~Ab ($3.8\pm0.2\,R_\oplus$; \citealt{bouma_kep1627_2022}),
-and AU~Mic~d ($4.2\pm0.2\,R_\oplus$; \citealt{plavchan_planet_2020}).
+emphasizing planets with precise ages.  The smallest previously known
+planets comparable to the new Cep-Her mini-Neptunes are AU~Mic~c
+($3.0\pm0.2\,R_\oplus$, see \citealt{martioli_aumicbc_2021} and
+\citealt{gilbert_flares_2022}), Kepler-1627~Ab ($3.8\pm0.2\,R_\oplus$;
+\citealt{bouma_kep1627_2022}), and AU~Mic~d ($4.2\pm0.2\,R_\oplus$;
+\citealt{plavchan_planet_2020}).
 
@@
 3\,$R_\oplus$ should be common at ages of $10^7$ to $10^8$ years.
-This expectation is tied to inferences about the
-initial distributions of planetary core mass, core composition, and
-atmospheric mass fraction \citep{owen_evaporation_2017}.  The
-Kelvin-Helmholtz cooling timescale, which is tied to the entropy of
-the planetary interior once the disk disperses, could also play a
-significant role \citep{owen_constraining_2020}.  As an example,
+This expectation is tied to inferences about the initial distributions
+of planetary core mass, core composition, and atmospheric mass
+fraction \citep{owen_evaporation_2017}.  The Kelvin-Helmholtz cooling
+timescale, which is tied to the entropy of the planetary interior once
+the disk disperses, could also play a significant role
+\citep{owen_constraining_2020}.  As an example,
 \citet{rogers_unveiling_2021} predicted that given a core-mass
@@
 there should be a maximum in planet occurrence at 2 to 3\,$R_\oplus$
-at times between 10 and 100\,Myr
-(their Figure~16).  The models advanced by
-\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} agree;
-the theoretical differences lie in their explanations
-for the radius valley.  If the radius gap is primordial
-and created by gas accretion physics, then super-Earths
+at times between 10 and 100\,Myr.  The models
+advanced by \citet{gupta_signatures_2020} and
+\citet{lee_primordial_2021} agree; their differences lie in the mechanism
+for producing the radius valley.  If the radius valley is
+primordial and created by gas accretion physics, then super-Earths
 ($<$$1.8$\,$R_\oplus$) would be common as early as $10^7$ years
@@
 driven by a slow escape of heat from the core predict longer
-$\sim$$10^9$ year timescales (but see \citealt{king_euv_2020}).
-
-Systems such as K2-25, V1298~Tau, HIP-67522, TOI-837,
-and TOI-1227 have sizes that are much larger than the picture we have
-just described.  However, their large sizes can be accommodated
-by existing models by invoking any of {\it i)} larger core masses,
-{\it ii)} more volatile-rich compositions, {\it iii)}  larger initial
-atmospheric mass fractions, or {\it iv)} longer thermal cooling times.
-For instance, perhaps all five of the the aforementioned systems are
-large simply because the planets are more massive than most
-mini-Neptunes.  Secure mass measurements would help constrain this
-parameter space, but the $\sim$1\,\kms\ spot-induced radial velocity
-semi-amplitudes makes measuring the Doppler orbits incredibly
-challenging.  \citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.
+$\sim$$10^9$ year timescales (but see \citealt{king_2021}).
+
+Systems such as K2-25, V1298~Tau, HIP-67522, TOI-837, and TOI-1227
+have sizes that are much larger than the picture we have just
+described.  However, their large sizes can be accommodated by existing
+models by invoking any of {\it i)} larger core masses, {\it ii)} more
+volatile-rich compositions, {\it iii)}  larger initial atmospheric
+mass fractions, or {\it iv)} longer thermal cooling times.  For
+instance, perhaps all five of the the aforementioned systems are large
+simply because the planets are more massive than most mini-Neptunes.
+Secure mass measurements would help constrain this parameter space,
+but the $\sim$1\,\kms\ spot-induced radial velocity semi-amplitudes
+makes measuring the Doppler orbits incredibly challenging.
+\citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.
 Regardless, the new Kepler-1643, KOI-7368, and KOI-7913 systems do
@@
 a likely explanation for why planets this small have not previously
-been identified \citep[{e.g.},][]{zhou_2021_tois}, future work
-should quantify this bias more carefully, in order to enable empirical
+been identified \citep[{e.g.},][]{zhou_2021_tois}, future work should
+quantify this bias more carefully, in order to enable empirical
 studies of how the planetary size distribution changes at early times.
@@
 expectations for a star claimed to be in an open cluster.  RSG-5 is an
-obvious overdensity relative to the local field, and our
-membership selection easily produces a clean pre-main-sequence locus
-in color versus absolute magnitude (Figure~\ref{fig:age}).  CH-2, and
-KOI-7913 and KOI-7368's membership inside it, do not meet these
-expectations in as obvious a manner.  This is because the CH-2
-association is diffuse.
-
-To quantify the density difference between CH-2 and RSG-5, we can compare the spatial and
-velocity volumes searched for each group.
-For RSG-5, we drew \nrsgfive\ candidate members from a $30\,{\rm
-pc}\times30\,{\rm pc}\times30\,{\rm pc}$ spatial cube, given a $1 {\rm
-\mkms} \times 2 {\rm \mkms }$ rectangle in apparent galactic velocity.
-For CH-2, our \nchtwo\ candidate members came from a rectangular prism of
-dimension $50\,{\rm pc}\times40\,{\rm pc}\times30\,{\rm pc}$, and a
-rectangular box of $2 {\rm \mkms} \times 4 {\rm \mkms}$.  If we define
-the searched volume in units of ${\rm pc}^3\,{\rm km^2}\,{\rm s^{-2}}$, then
-the volume ratio of CH-2 to RSG-5 is 9 to 1.  The ratio of number
-densities (candidate members per unit searched volume) in RSG-5
-relative to CH-2 is 34 to 1.
+obvious overdensity relative to the local field, and our membership
+selection easily produces a clean pre-main-sequence locus in color
+versus absolute magnitude (Figure~\ref{fig:age}).  CH-2, and KOI-7913
+and KOI-7368's membership inside it, do not meet these expectations in
+as obvious a manner.  This is because the CH-2 association is diffuse.
+
+To quantify the density difference between CH-2 and RSG-5, we can
+compare the spatial and velocity volumes searched for each group.  For
+RSG-5, we drew \nrsgfive\ candidate members from a $30\,{\rm
+pc}\times30\,{\rm pc}\times40\,{\rm pc}$ rectangular prism, given a
+$1.5 {\rm \mkms} \times 2.5 {\rm \mkms }$ rectangle in apparent
+galactic velocity.  For CH-2, our \nchtwo\ candidate members came from
+a rectangular prism of dimension $50\,{\rm pc}\times40\,{\rm
+pc}\times30\,{\rm pc}$, and a rectangular box of $2 {\rm \mkms} \times
+4 {\rm \mkms}$.  If we define the searched volume in units of ${\rm
+pc}^3\,{\rm km^2}\,{\rm s^{-2}}$, then the volume ratio of CH-2 to
+RSG-5 is 3.5 to 1.  The ratio of number densities (candidate members
+per unit searched volume) in RSG-5 relative to CH-2 is 16 to 1.
 
 Given its low density, in CH-2 really a star cluster?  For this
-discussion, we adopt the definition that a star cluster is a group
-of at least 12 stars that was physically associated at its time of
-formation.  The somewhat arbitrary ``12'' is set to distinguish star
-clusters from high-order multiples \citep[see][]{krumholz_star_2019}.
-Dissolved clusters and their tidal tails are therefore included in our
-concept of clusters.  We do not require however that a particular number of stars per
-unit spatial volume is required for a group to be a ``cluster''.
-The latter point acknowledges that an important
-factor in cluster identification is also the density in velocity
-space.  Perhaps once stellar rotation periods and chemical abundances
-reach the same level of ubiquity as stellar proper motions, they might
-enable further refinement of our ability to discover stars that formed
-together.
+discussion, we adopt the definition that a star cluster is a group of
+at least 12 stars that was physically associated at its time of
+formation.  The ``12'' is set to distinguish star clusters from
+high-order multiples \citep[see][]{krumholz_star_2019}.  Dissolved
+clusters and their tidal tails are therefore included in our concept
+of clusters.  We do not however require a threshold number of stars
+per unit spatial volume for a group to be a ``cluster''.  The latter
+point acknowledges that an important factor in cluster identification
+is also the density in velocity space.  Perhaps once stellar rotation
+periods and chemical abundances reach the same level of ubiquity as
+stellar proper motions, they might enable further refinement of our
+ability to discover stars that formed together.
 
@@
 \citep{wright_kinematics_2018}.  This complicates the feasibility of
-deriving kinematic ages through traceback, as
-well as through the expansion itself
-\citep[see][]{crundall_chronostar_2019}.  
-A more minimal approach is
-that suggested by \citet{tofflemire_2021}: search for coeval,
-phase-space neighbors, measure their ages, and determine if they share
-a common age.  This approach can demonstrate whether a star is {\it
-currently associated} with a set of coeval stars, which is much easier
-than determining what the association looked like in the past.  Our
-analysis of CH-2 meets the latter standard for the existence of a
-$\approx$40\,Myr assocation today.  It would be a worthy exercise for
-future work to perform a similar search for coeval phase-space
-neighbors on the entire dataset of known exoplanet hosts.  For the
-time being, we can offer the anecdotal point that in our experience,
-most stars do not have dozens of 40\,Myr neighbors within a
-local volume of a few \kms\ and tens of parsecs.
+deriving kinematic ages through traceback, as well as through the
+expansion itself \citep[see][]{crundall_chronostar_2019}.  A more
+minimal approach is that suggested by \citet{tofflemire_2021}: search
+for coeval, phase-space neighbors, measure their ages, and determine
+if they share a common age.  This approach can demonstrate whether a
+star is currently associated with a set of coeval stars, which
+is much easier than determining what the association looked like in
+the past.  Our analysis of CH-2 meets the latter standard for the
+existence of a $\approx$40\,Myr assocation.  It would be a worthy
+exercise for future work to perform a similar search for coeval
+phase-space neighbors on the entire dataset of known exoplanet hosts.
+For the time being, we can offer the anecdotal point that in our
+experience, most stars do not have dozens of 40\,Myr neighbors within
+a local volume of a few \kms\ and tens of parsecs.
 
@@
 Our analysis to date has focused only on portions of Cep-Her that were
-observed by Kepler: RSG-5, CH-2, and the $\delta$~Lyr cluster.  The
-membership and kinematics of the other subgroups shown in
-Figure~\ref{fig:XYZvtang} deserve independent attention.  R.~Kerr and
-collaborators have {\bf acquired N spectra} from which radial
-velocities will be derived in their upcoming study.
-One important aspect of these analyses will be to assess whether the 
-traceback approach could be applicable.
-However if the
-estimate that Cep-Her has $\gtrsim$3{,}500 stars is correct then more
-data might be required to fully untangle the kinematics of the various
-groups.  Wide-field spectroscopic surveys such as LAMOST
-\citep{zhao_2012_LAMOST} or SDSS-V \citep{kollmeier_2017} could
-enable such analyses, while also
-providing sensitivity to the Li 6708\,\AA\ line.  Gaia
-DR3 may also provide RVS spectra for some of the brighter F and G
-dwarfs, which would contain the calcium infrared triplet as another
-age indicator.
+observed by Kepler: RSG-5, CH-2, and the $\delta$~Lyr cluster.  In
+\citet{bouma_kep1627_2022} as well as this work, we have shown that
+these groups share similar ages, and have kinematic correlations that
+suggest a common origin.  With that said, the membership and
+kinematics of the other subgroups shown in Figure~\ref{fig:XYZvtang}
+deserve independent attention.  An important aspect of the work that
+remains will be to acquire radial velocities of a large subset of the
+stars, to assess whether the traceback approach could be applicable.
+Wide-field spectroscopic surveys such as LAMOST
+\citep{zhao_2012_LAMOST} or SDSS-V \citep{kollmeier_2017} could enable
+such analyses, while also providing sensitivity to the Li 6708\,\AA\
+line.  Gaia DR3 may also provide RVS spectra for the brighter F and G
+dwarfs; these would contain the calcium infrared triplet as another
+age indicator.  These indicators, combined with better kinematics,
+will be needed to definitively unravel the formation history of the
+complex.
 
 A number of worthy photometric projects also seem possible given the
-new understanding of the complex.  One is asteroseismology of the
+new understanding of Cep-Her.  One is asteroseismology of the
 $\delta$~Sct stars, using either TESS or Kepler data
 \citep{bedding_very_2020}.  For cases in which the modes are resolved,
-this might yield age estimates for the subgroups independent of other
-approaches.  Other projects could include searches of the Kepler light
-curves for exocomets \citep{zieba_transiting_2019}, and for missed
-planets around the most rapid rotators.
+this might yield age or metallicity estimates for the subgroups
+independent of other approaches.  Other projects could include
+searches of the Kepler light curves for exocomets
+\citep{zieba_transiting_2019}, and for missed planets around the most
+rapid rotators.
 
 \paragraph{Exoplanet demographics at early times}
-Our main motivation for finding new young planets is to help benchmark models
-for planetary evolution.  However demographic analyses of the known planets
-between $10^7$ and $10^9$ years have so far been somewhat limited.
-Approximately 40 such planets are now known (Figure~\ref{fig:age}).
-About half come from K2, a quarter from TESS, and now a quarter from
-Kepler.
-
-Given the current state of the field, a few reflections regarding experimental
-design of a demographic survey focused on planetary evolution during
-the first gigayear might be useful.  The first is that such a project
-requires a set of
-target stars with known ages.  The best way to
-compile relevant stars right now is likely by combining automated
-spatio-kinematic clustering from Gaia with rotation periods measured using TESS
-\citep[see the appendices of][]{bouma_kep1627_2022}.  The second consideration
-is that the known young planets smaller than $3$\,$R_\oplus$ all come from either
-K2 or Kepler.  Demographic inferences based on TESS are therefore limited to
-planetary sizes $\gtrsim4$\,$R_\oplus$, for planets close-in to their host stars.  It
-seems worthwhile to compare the absolute occurrence rate of such planets to
-those observed by Kepler.  One specific question that seems within reach would
-be to clarify whether enough young stars have been searched for the dearth of
-young hot Jupiters to be significant.  Since the
-hot Jupiter occurrence rate is strongly dependent on stellar mass and
-metallicity \citep{petigura_metallicity_2018,petigura_cksX_2022}, particular
-care would be needed to select a sample of well-studied FGK dwarfs for the
+Our main motivation for finding new young planets is to help benchmark
+models for planetary evolution.  However demographic analyses of the
+known planets between $10^7$ and $10^9$ years have so far been
+somewhat limited.  Approximately 40 such planets are now known
+(Figure~\ref{fig:age}).  About half come from K2, a quarter from TESS,
+and now a quarter from Kepler.
+
+Given the current state of the field, a few reflections regarding
+experimental design of a demographic survey focused on planetary
+evolution during the first gigayear might be useful.  The first is
+that such a project requires a set of target stars with known ages.
+The best way to compile relevant stars right now is likely by
+combining automated spatio-kinematic clustering from Gaia with
+rotation periods measured using TESS \citep[see the appendices
+of][]{bouma_kep1627_2022}.  The second consideration is that the known
+young planets smaller than $3$\,$R_\oplus$ all come from either K2 or
+Kepler.  Demographic inferences based on TESS are therefore limited to
+planetary sizes $\gtrsim4$\,$R_\oplus$, for planets close-in to their
+host stars.  It seems worthwhile to compare the absolute occurrence
+rate of such planets to those observed by Kepler.  One specific
+question that seems within reach would be to clarify whether enough
+young stars have been searched for the dearth of young hot Jupiters to
+be significant.  Since the hot Jupiter occurrence rate is strongly
+dependent on stellar mass and metallicity
+\citep{petigura_metallicity_2018,petigura_cksX_2022}, particular care
+would be needed to select a sample of well-studied FGK dwarfs for the
 measurement, likely using stars in Sco OB2, Cep-Her, and Orion.  For
-demographic studies focused on how mini-Neptune sizes evolve, the combined K2
-and Kepler dataset might be the better primary source.
+demographic studies focused on how mini-Neptune sizes evolve, the
+combined K2 and Kepler dataset might be the better primary source.
 
@@
 
-We have showed that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab
-are 40 to 50 million years old, and that each system
-is most likely planetary.  The evidence for the planetary
-interpretation comes from an application of \texttt{VESPA} to the
-Kepler data, alongside new imaging from NIRC2.  The
-validity of the \texttt{VESPA} framework rests on the premise that
-non-astrophysical false positives can be rejected.  Kepler-1643 and
-KOI-7368 both pass all but one of the canonical vetting tests: the
-check for centroid offsets between the in and out-of-transit phases.
-For both cases, the weak observed shifts seem to be
-caused by starspot-induced variability in specific quarters
-spuriously moving the stellar center-of-light.
-Independently, our imaging rules out companion stars at the brightness
-and positions that would be needed to explain the reported shifts.
-
-Each system has multiple indicators of youth that support the
-reported ages.  For Kepler-1643, the strongest youth indicator
-is its physical and kinematic association with RSG-5.  Based on the
-color-absolute magnitude diagram, we are able to select members
-of this cluster with a false positive rate of a few percent
-(Figure~\ref{fig:age}).  Kepler-1643 is one such member.  While the
-stellar rotation period period agrees with this assessment, the star's
-lithium equivalent width is somewhat low, which might motivate future
-exploration of lithium depletion across FGKM stars in RSG-5 (see
+We have showed that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab are 40
+to 50 million years old, and that each system is most likely
+planetary.  The evidence for the planetary interpretation comes from
+an application of \texttt{VESPA} to the Kepler data, alongside new
+imaging from NIRC2.  The validity of the \texttt{VESPA} framework
+rests on the premise that non-astrophysical false positives can be
+rejected.  Kepler-1643 and KOI-7368 both pass all but one of the
+canonical vetting tests: the check for centroid offsets between the in
+and out-of-transit phases.  For both cases, the weak observed shifts
+seem to be caused by starspot-induced variability in specific quarters
+spuriously moving the stellar center-of-light.  Independently, our
+imaging rules out companion stars at the brightness and positions that
+would be needed to explain the reported shifts.
+
+Each system has multiple indicators of youth that support the reported
+ages.  For Kepler-1643, the strongest youth indicator is its physical
+and kinematic association with RSG-5.  Based on the color-absolute
+magnitude diagram, we are able to select members of this cluster with
+a false positive rate of a few percent (Figure~\ref{fig:age}).
+Kepler-1643 is one such member.  While the stellar rotation period
+period agrees with this assessment, the star's lithium equivalent
+width is somewhat low, which might motivate future exploration of
+lithium depletion across FGKM stars in RSG-5 (see
 Appendix~\ref{app:spectra}).
 
-The spatio-kinematic argument for the youth of KOI-7368 and
-KOI-7913 is weaker because they are in an association of stars, CH-2,
-that is more diffuse.  For KOI-7913, stronger indicators of its age
-come from its binary.  Both stellar components in KOI-7913
-have isochronal ages consistent with $40$\,Myr.
-Both components also show H$\alpha$ in emission, which for
-the $\approx$K6V primary is a strong indicator that the star is
-$\lesssim$100$\,$Myr old.  KOI-7368 is slightly more
+The spatio-kinematic argument for the youth of KOI-7368 and KOI-7913
+is weaker because they are in an association of stars, CH-2, that is
+more diffuse.  For KOI-7913, stronger indicators of its age come from
+its binary.  Both stellar components in KOI-7913 have isochronal ages
+consistent with $40$\,Myr.  Both components also show H$\alpha$ in
+emission, which for the $\approx$K6V primary is a strong indicator
+that the star is $\lesssim$100$\,$Myr old.  KOI-7368 is slightly more
 massive, and the Li 6708\,\AA\ measurement and stellar rotation period
-provide independent verification of the
-star's youth.
-
-The astrophysical significance of these considerations
-is that planets $\sim$2 Earth radii in size exist at ages of 40 million years.
-It will be interesting to continue the push down to smaller planetary
-sizes at comparable ages -- the planetary detections we have
-presented are well above the average detection significance for Kepler 
-planets.  There may still be room at the bottom.
+provide independent verification of the star's youth.
+
+The astrophysical significance of these considerations is that planets
+$\sim$2 Earth radii in size exist at ages of 40 million years.  It
+will be interesting to continue the push down to smaller planetary
+sizes at comparable ages -- the planetary detections we have presented
+are well above the average detection significance for Kepler planets.
+There may still be room at the bottom.
 
@@
   %\texttt{altaipony} \citep{ilin_flares_2021},
-  \texttt{astrobase} \citep{bhatti_astrobase_2018},
+  %\texttt{astrobase} \citep{bhatti_astrobase_2018},
   %\texttt{astroplan} \citep{astroplan2018},
 	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
-  \texttt{astropy} \citep{astropy_2018},
-  \texttt{astroquery} \citep{astroquery_2018},
+  %\texttt{astropy} \citep{astropy_2018},
+  %\texttt{astroquery} \citep{astroquery_2018},
   %\texttt{BATMAN} \citep{kreidberg_batman_2015},
@@
   %\texttt{scikit-learn} \citep{scikit-learn},
-  \texttt{scipy} \citep{jones_scipy_2001},
+  %\texttt{scipy} \citep{jones_scipy_2001},
   %\texttt{TESS-point}  \citep{burke_2020},
@@
 
-\paragraph{Table~2} contains candidate Cep-Her members with weights $\alpha>0.02$
-observed by Kepler.  More restrictive weight cuts
-would be needed if one wanted to remove the majority of field star interlopers: at
-least half of the stars included in the table are unlikely
-to be bonafide members of Cep-Her.  Table~2  was created by
+\paragraph{Table~2} contains candidate Cep-Her members with weights
+$\alpha>0.02$ observed by Kepler.  More restrictive weight cuts would
+be needed if one wanted to remove the majority of field star
+interlopers; at least half of the stars included in the table are
+unlikely to be bonafide Cep-Her members.  Table~2  was created by
 cross-matching candidate Cep-Her members (selected using Gaia EDR3;
@@
 motion and epoch-corrected angular distance neighbor was taken as the
-single best match.  The \texttt{edr3\_dr2\_mag\_diff} column
-gives some indication of the reliability of this EDR3 to DR2
-conversion, as there are a few cases between Gaia DR2 and EDR3 where
-partially resolved binaries became fully resolved.
-
-\paragraph{Table~3} contains spatial, kinematic, astrometric, and rotation period
-information for the \nrsgfive\ candidate RSG-5 members and \nchtwo\ candidate CH-2
-members described in Section~\ref{subsec:members}.
-These are the data used to make the lower panels of
-Figure~\ref{fig:age}.
-Table~3 includes rotation periods for 53 stars with ZTF and 51 stars with
- TESS, for a total of 71 periods reported for unique stars and 32
- overlapping between the surveys. We adopted the ZTF period over the
+single best match.  The \texttt{edr3\_dr2\_mag\_diff} column gives
+some indication of the reliability of this EDR3 to DR2 conversion, as
+there are a few cases between Gaia DR2 and EDR3 where partially
+resolved binaries became fully resolved.
+
+\paragraph{Table~3} contains spatial, kinematic, astrometric, and
+rotation period information for the \nrsgfive\ candidate RSG-5 members
+and \nchtwo\ candidate CH-2 members described in
+Section~\ref{subsec:members}.  These are the data used to make the
+lower panels of Figure~\ref{fig:age}. 
+% Table~3 includes rotation periods for 53 stars with ZTF and 51 stars
+% with TESS, for a total of 71 periods reported for unique stars and 32
+% overlapping between the surveys.
+We adopted the ZTF period over the
 TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the
@@
 we failed to detect a period in TESS but recovered one from ZTF; in
- all cases the periods appear to be 13--16 days.
- These stars were: (1) Gaia EDR3
- 2129930258400157440, for which TESS showed a flat light curve while ZTF yielded
- a 15.3-day period; (2) Gaia EDR3 2082376861542398336, LS found a
- 7.6-day period which we rejected during visual validation; we found
- 15.4 days with ZTF, and we suspect that the weak/rejected signal form
- TESS might have been a 1/2 period harmonic; (3) Gaia EDR3
- 2082397099429013120, similar to the previous case, we rejected a
- 6.7-day signal from TESS and recovered a 12.8-day period with ZTF.
+all cases the periods appear to be 13--16 days.  These stars were: (1)
+Gaia EDR3 2129930258400157440, for which TESS showed a flat light
+curve while ZTF yielded a 15.3-day period; (2) Gaia EDR3
+2082376861542398336, LS found a 7.6-day period which we rejected
+during visual validation; we found 15.4 days with ZTF, and we suspect
+that the weak/rejected signal form TESS might have been a 1/2 period
+harmonic; (3) Gaia EDR3 2082397099429013120, similar to the previous
+case, we rejected a 6.7-day signal from TESS and recovered a 12.8-day
+period with ZTF.
 
@@
 measured in the ring of pixels around said aperture (the ``halo'').
-Generally one expects the transit signal to be strongest in the
+One usually expects the transit signal to be strongest in the
 central aperture, rather than the halo.  Two types of false positive
@@
 
-%\clearpage
+\clearpage
 \input{tab_supp_CepHer_X_Kepler.tex}
+\vspace{-6cm}
 \input{tab_supp_RSG5_CH2_Prot.tex}

### 2022-04-24 drivers/plot_hr.py

@@ -11,12 +11,6 @@ PLOTDIR = os.path.join(RESULTSDIR, 'hr')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
-# exo4
-rp.plot_hr(PLOTDIR, isochrone=None, color0='phot_bp_mean_mag', show100pc=1,
-           clusters=['RSG-5'], reddening_corr=1,
-           overplotkep1627=0, show_allknown=0, overplotkep1643=1,
-           smalllims=1)
-
 # show all the known KOIs that are ~40 Myr old
 for smalllims in [1,0]:
 

### 2022-04-24 papers/ms.tex
update ms

@@
   $4\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368,
-  and KOI-7913 respectively.  These planets empirically demonstrate
+  and KOI-7913 respectively.  These planets demonstrate
   that mini-Neptunes with sizes of $\sim$2 Earth radii exist at ages
@@
 
-Our focus is on a region of the Galaxy approximately 200 to 500\,pc
+Our focus is on a region of the Galaxy 200 to 500\,pc
 from the Sun, above the galactic plane, and spanning galactic
-longitudes of roughly $40^\circ$ to $100^\circ$.  Two rich clusters in
+longitudes of $40^\circ$ to $100^\circ$.  Two rich clusters in
 this region are the $\delta$~Lyra cluster
@@
 To visualize the results, in Figure~\ref{fig:XYZvtang} we show
-12{,}436 objects with $\alpha>0.02$ as gray points, and overplot
+12{,}436 objects with $\alpha>0.02$ as gray points, and 
 4{,}763 objects with $\alpha>0.10$ as black points.  The previously
@@
 \citet[][Appendix B]{GaiaCollaboration2018}, which include binaries
-while omitting instrumental artifacts from for instance low
+while omitting artifacts from for instance low
 photometric signal to noise, or a small number of visibility periods.
@@
 
-The CAMDs show that for RSG-5, practically all of our candidate
+The CAMDs show that for RSG-5, all but one of the candidate
 members are on a tight pre-main-sequence locus.  This implies a false
@@
 an age for the $\delta$~Lyr cluster of $38^{+6}_{-5}$\,Myr.  The
-slightly older isochronal age of RSG-5 is consistent with its location
+older isochronal age of RSG-5 is consistent with its location
 relative to the $\delta$~Lyr cluster in the upper left panel of
@@
 USco ($\sim$$8$\,Myr), and some may still exist at ages of LCC
-($\sim$$16$\,Myr; L.~Rebull in preparation).  Regardless, given that
+($\sim$$16$\,Myr; L.~Rebull in preparation).  Regardless, since
 nearly no field star outliers seem to be present on the RSG-5 CAMD,
@@
 G8V (Kepler-1627) to K6V (KOI-7913 A).  The secondary in the KOI-7913
-system has spectral type $\approx$K8V.  And given that a star with
+system has spectral type $\approx$K8V.  And since a star with
 Solar mass and metallicity arrives at the zero-age main sequence at
@@
 
-To verify these parameters and to also analyze spectroscopic youth
+To verify these parameters and to analyze youth
 proxies such as the Li 6708\,\AA\ doublet and H$\alpha$, we acquired
@@
 \paragraph{Binarity}
-KOI-7913 is almost certainly a binary.  The north-west primary is
+KOI-7913 is a binary.  The north-west primary is
 $\approx$0.5 magnitudes brighter than the south-east secondary in
@@
 amusing if the faint neighbor were also comoving and young -- at the
-age of the system, it would have a mass roughly between 10 and
+age of the system, it would have a mass between 10 and
 15\,M$_{\rm Jup}$.  Additional imaging epochs will tell.
@@
 In the future, it may be possible to obtain independent evidence for
-the Cep-Her planets to directly confirm their planetary nature, for
-instance through transit observations using a high-resolution
-spectrograph.  For now, it is of interest whether the transit signals
-might be astrophysical false positives, or whether they can be argued
-to be planetary on a statistical basis.  We adopt the Bayesian
-framework implemented in \texttt{VESPA} to assess the relevant
-probabilities \citep{morton_efficient_2012,vespa_2015}.  Briefly
-summarized, the priors in \texttt{VESPA} assume the binary star
-occurrence rate from \citet{raghavan_survey_2010}, direction-specific
-star counts from \citet{girardi_star_2005}, and planet occurrence
-rates as described by \citet[][Section~3.4]{morton_efficient_2012}.
-The likelihoods are then evaluated by forward-modeling a synthetic
-population of eclipsing bodies for each astrophysical model class, in
-which each population member has a known trapezoidal eclipse depth,
-total duration, and ingress duration.  These summary statistics are
-then compared against the actual photometric data to evaluate the
-probabilities of false positive scenarios such as foreground eclipsing
-binaries, hierarchical eclipsing binaries, and background eclipsing
-binaries.
+the planetary nature of the Cep-Her planets, for instance through
+transit observations using a high-resolution spectrograph.  For now,
+it is of interest whether the transit signals might be astrophysical
+false positives, or whether they are statistically more likely to be
+planetary.  We adopt the Bayesian framework implemented in
+\texttt{VESPA} to assess the relevant probabilities
+\citep{morton_efficient_2012,vespa_2015}.  Briefly summarized, the
+priors in \texttt{VESPA} assume the binary star occurrence rate from
+\citet{raghavan_survey_2010}, direction-specific star counts from
+\citet{girardi_star_2005}, and planet occurrence rates as described by
+\citet[][Section~3.4]{morton_efficient_2012}.  The likelihoods are
+then evaluated by forward-modeling a synthetic population of eclipsing
+bodies for each astrophysical model class, in which each population
+member has a known trapezoidal eclipse depth, total duration, and
+ingress duration.  These summary statistics are then compared against
+the actual photometric data to evaluate the probabilities of false
+positive scenarios such as foreground eclipsing binaries, hierarchical
+eclipsing binaries, and background eclipsing binaries.
 
@@
 
-Intriguingly, Kepler-1643 formally failed one of the data validation
+Intriguingly, Kepler-1643 failed one of the data validation
 centroid shift tests (see the \texttt{q1\_q17\_dr25\_koi} data
@@
 calculation, we find ${\rm FPP} \approx 4\times10^{-3}$.  Though not
-as convincing as Kepler-1643, this formally clears the usual threshold
+as convincing as Kepler-1643, this clears the usual threshold
 for calling the planet statistically validated
@@
 the objects discusssed (${\rm S/N}=14$), its lower FPP relative to
-KOI-7368 can be understood through its relatively flat-bottomed shape,
+KOI-7368 can be understood through its flat-bottomed shape,
 combined with its long transit duration relative to plausible
@@
     KOI-7913~Ab (X), and Kepler-1627~Ab (+).  Interesting trends in
-    the population of young planets currently include {\it i)} their
+    the population of young planets include {\it i)} their
     large sizes and {\it ii)} the lack of young hot Jupiters.  The new
@@
 instance, perhaps all five of the the aforementioned systems are large
-simply because the planets are more massive than most mini-Neptunes.
+because the planets are more massive than most mini-Neptunes.
 Secure mass measurements would help constrain this parameter space,
 but the $\sim$1\,\kms\ spot-induced radial velocity semi-amplitudes
-makes measuring the Doppler orbits incredibly challenging.
+makes measuring the Doppler orbits very difficult.
 \citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.
@@
 experimental design of a demographic survey focused on planetary
-evolution during the first gigayear might be useful.  The first is
+evolution over the first gigayear might be useful.  The first is
 that such a project requires a set of target stars with known ages.
-The best way to compile relevant stars right now is likely by
-combining automated spatio-kinematic clustering from Gaia with
+A promising way to compile relevant stars could be to
+combine automated spatio-kinematic clustering from Gaia with
 rotation periods measured using TESS \citep[see the appendices
@@
 
-We have showed that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab are 40
+We have shown that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab are 40
 to 50 million years old, and that each system is most likely
@@
 just 50 million years, K-dwarfs between $4500$\,K and $5200$\,K go from
-having a relatively tight lithium sequence to one with a dispersion
+having a tight lithium sequence to one with a dispersion
 $\approx10\times$ greater.  The existence of the Li dispersion in
@@
 stars of this age.  \citet{kraus_stellar_2014} for instance, in their
-survey of Tuc-Hor ($\approx$40\,Myr), observed that essentially all
+survey of Tuc-Hor ($\approx$40\,Myr), observed that all
 cluster members with spectral type $>$${\rm K4.5V}$ had H$\alpha$ in
@@
 for all of our other Kepler objects.  The double-peaked nature of the
-emission, though not always present, is also not particularly rare for
+emission, though not always present, is also common for
 active stars.  Proxima Centauri, for instance, has double-peaked
@@
 stars is present.  The pointing geometries from quarters 3, 7, 11, and
-15 however did not noticeably affect the observed transit depths,
+15 however did not affect the observed transit depths,
 which is an indication that the crowding metric applied in the data

### 2022-04-25 papers/ms.tex
update ms

@@
 want to derive accurate planetary sizes and orbital properties.
-Second, we want to remove the stellar variability signal to enable a
+Second, we want to remove the spot-induced variability signal to enable a
 statistical assessment of the probability that the transit signals are
@@
 light curve to a local window around each transit that spanned three
-transit durations before and after each mid-transit time.  The
+transit durations before and after each transit midpoint.  The
 out-of-transit points in each local window were then fitted with a
@@
 
-We fitted the models using \texttt{exoplanet}
+We fitted the data using \texttt{exoplanet}
 \citep{exoplanet:exoplanet}.  We assumed a Gaussian likelihood, and
@@
 maximum {\it a posteriori} (MAP) model.  We used the
-\citet{gelman_inference_1992} statistic ($\hat{R}$) as our convergence
+\citet{gelman_inference_1992} statistic, $\hat{R}$, as our convergence
 diagnostic.  The resulting fits are shown in the lower panels of
 Figure~\ref{fig:planets}, and the important derived parameters are in
-Table~\ref{tab:sysparams}.  The set of full parameters with their
+Table~\ref{tab:sysparams}.  The set of full parameters and their
 priors are given in Appendix~\ref{app:transit}.
@@
 variability using a quasiperiodic Gaussian process (GP).  We explored
-this approach, but found that it sometimes required fine-tuning of the
+this approach, but found that it often required fine-tuning of the
 hyperparameter priors, otherwise the GP would overfit flares,
@@
 In the future, it may be possible to obtain independent evidence for
-the planetary nature of the Cep-Her planets, for instance through
-transit observations using a high-resolution spectrograph.  For now,
+the planetary nature of the Cep-Her planets, for instance by observing
+spectroscopic transits.  For now,
 it is of interest whether the transit signals might be astrophysical
@@
 6\times10^{-9}$.  Figure~\ref{fig:planets} shows the justification:
-the transit is flat and has a high S/N ($\approx 47$).  The shape is
+the transit is flat and has a high S/N ($\approx$$47$).  The shape is
 therefore nearly impossible to reproduce with eclipsing binary models.
@@
 from the other Kepler quarters are consistent at $\lesssim 0\farcs4$.
-Nonetheless, this is an instructive exercise in how stellar
-variability complicates centroid-based vetting tests.  The centroid
+This is an instructive exercise in how stellar
+variability complicates centroid-based vetting tests.  The 
 shifts measured by these tests are determined from the in- and
@@
 revealed no new stellar companions.  Performing the relevant
-calculation, we find ${\rm FPP} \approx 4\times10^{-3}$.  Though not
+calculation, we find ${\rm FPP} = 4\times10^{-3}$.  Though not
 as convincing as Kepler-1643, this clears the usual threshold
@@
 Kepler-1643.  For KOI-7368, the reported offset is smaller, and less
-statistically significant ($0\farcs2$; 3.0-$\sigma$).  Again, the data
+significant ($0\farcs2$; 3.0-$\sigma$).  Again, the data
 validation reports show that the shift is caused by a few outlying
@@
 consistent scatter between centroid locations suggests that these
-outliers are also caused by the stellar variability.  Independently,
-our NIRC2 imaging shows that there are no neighboring sources that
-could cause an offset of the observed amplitude.
+outliers are also caused by the stellar variability.  
+Our NIRC2 imaging independently shows that there are no neighboring
+sources that could cause an offset of the observed amplitude.
 
@@
 KOI-7913.01 is also currently listed on the NASA Exoplanet Archive as a
-``candidate'' planet.  The canonical \citet{morton_false_2016}
+``candidate'' planet.  The \citet{morton_false_2016}
 analysis was of Q1-Q17 KOIs from DR24, and therefore spanned KOI-1.01
-to KOI-7620.01 and omitted KOI-7913.01.  However the results of the
+to KOI-7620.01 (omitting KOI-7913.01).  However the results of the
 subsequent DR25 analysis by Morton et al.\ are listed at the NASA
@@
 1.3\times10^{-4}$.  Though the transit has the lowest S/N of any of
-the objects discusssed (${\rm S/N}=14$), its lower FPP relative to
+the objects discusssed ($\approx$$14$), its lower FPP relative to
 KOI-7368 can be understood through its flat-bottomed shape,
@@
 probability usability score \citep{2017ksci.rept...16B} also meets the
-usual threshold, and so the planet seems to be validated.  Its
+usual threshold, and so the planet is statistically validated.  Its
 disposition has however previously fluctuated from ``false positive''
-to ``candidate''.  The most likely explanation is the presence of
-KOI-7913~B, which is located approximately one Kepler pixel away from
-Kepler-7913 A (see Appendix~\ref{app:koi7913}).  Our pixel-level
-analysis of the target pixel files did not show any evidence to
-support the idea that KOI-7913~B could host the transiting planet.
-Our NIRC2 imaging similarly ruled out the possibility of any other
-plausible host star for the planet other than KOI-7913~A.
+to ``candidate'' (see Appendix~\ref{app:koi7913}).  The most likely
+explanation is the presence of KOI-7913~B, which is located
+approximately one Kepler pixel away from Kepler-7913 A.  Our
+pixel-level analysis of the target pixel files did not show any
+evidence to support the idea that KOI-7913~B could host the transiting
+planet.  Our NIRC2 imaging similarly ruled out the possibility of any
+other plausible host star for the planet other than KOI-7913~A.
 
@@
     KOI-7913~Ab (X), and Kepler-1627~Ab (+).  Interesting trends in
-    the population of young planets include {\it i)} their
-    large sizes and {\it ii)} the lack of young hot Jupiters.  The new
+    the population of $<10^8$ year old planets include {\it i)} their
+    large sizes and {\it ii)} the lack of hot Jupiters.  The new
     objects of interest in Cep-Her have normal mini-Neptune sizes
@@
 
-The most significant novelty about Kepler-1643~b, KOI-7368.01, and
-KOI-7913~Ab is that their sizes are normal, relative to the known
-population of gigayear-old mini-Neptunes.  At field star ages,
+The most significant novelty about the planets in Kepler-1643, KOI-7368, and
+KOI-7913 is that their sizes are normal relative to the known
+population of mini-Neptunes from Kepler.  At field star ages,
 mini-Neptune sizes span 1.8\,$R_\oplus$ to 3.6\,$R_\oplus$, with the
@@
 fraction \citep{owen_evaporation_2017}.  The Kelvin-Helmholtz cooling
-timescale, which is tied to the entropy of the planetary interior once
-the disk disperses, could also play a significant role
+timescale, which is tied to the entropy of the planetary interior
+shortly after
+disk dispersal, also plays a significant role
 \citep{owen_constraining_2020}.  As an example,
@@
 \citet{lee_primordial_2021} agree; their differences lie in the mechanism
-for producing the radius valley.  If the radius valley is
-primordial and created by gas accretion physics, then super-Earths
-($<$$1.8$\,$R_\oplus$) would be common as early as $10^7$ years
-\citep{lee_primordial_2021}.   The photoevaporation-driven models more
-often quote $10^8$ year timescales for production of
-atmosphere-stripped super-Earths, while the models with the mass-loss
-driven by a slow escape of heat from the core predict longer
-$\sim$$10^9$ year timescales (but see \citealt{king_2021}).
+for producing the radius valley. 
+% If the radius valley is
+% primordial and created by gas accretion physics, then super-Earths
+% ($<$$1.8$\,$R_\oplus$) would be common as early as $10^7$ years
+% \citep{lee_primordial_2021}.   The photoevaporation-driven models more
+% often quote $10^8$ year timescales for production of
+% atmosphere-stripped super-Earths, while the models with the mass-loss
+% driven by a slow escape of heat from the core predict longer
+% $\sim$$10^9$ year timescales (but see \citealt{king_2021}).
 
 Systems such as K2-25, V1298~Tau, HIP-67522, TOI-837, and TOI-1227
-have sizes that are much larger than the picture we have just
-described.  However, their large sizes can be accommodated by existing
-models by invoking any of {\it i)} larger core masses, {\it ii)} more
+have sizes that are anomalously large relative to the predicted peak
+in planet occurrence at 2 to 3\,$R_\oplus$.
+However, their large sizes can be accommodated 
+by invoking any of {\it i)} larger core masses, {\it ii)} more
 volatile-rich compositions, {\it iii)}  larger initial atmospheric
-mass fractions, or {\it iv)} longer thermal cooling times.  For
-instance, perhaps all five of the the aforementioned systems are large
-because the planets are more massive than most mini-Neptunes.
+mass fractions, or {\it iv)} longer thermal cooling times.  
 Secure mass measurements would help constrain this parameter space,
@@
 suggest a common origin.  With that said, the membership and
-kinematics of the other subgroups shown in Figure~\ref{fig:XYZvtang}
+kinematics of the other Cep-Her groups shown in Figure~\ref{fig:XYZvtang}
 deserve independent attention.  An important aspect of the work that
@@
 such analyses, while also providing sensitivity to the Li 6708\,\AA\
-line.  Gaia DR3 may also provide RVS spectra for the brighter F and G
-dwarfs; these would contain the calcium infrared triplet as another
+line.  If Gaia DR3 provides RVS spectra for the brighter F and G
+dwarfs, these would also contain the calcium infrared triplet as another
 age indicator.  These indicators, combined with better kinematics,
-will be needed to definitively unravel the formation history of the
+would help in definitively unravelling the formation history of the
 complex.
@@
 this might yield age or metallicity estimates for the subgroups
-independent of other approaches.  Other projects could include
+independent of other methods.  Other projects could include
 searches of the Kepler light curves for exocomets
@@
 known planets between $10^7$ and $10^9$ years have so far been
-somewhat limited.  Approximately 40 such planets are now known
+rather limited.  Approximately 40 such planets are now known
 (Figure~\ref{fig:age}).  About half come from K2, a quarter from TESS,
@@
 rotation periods measured using TESS \citep[see the appendices
-of][]{bouma_kep1627_2022}.  The second consideration is that the known
-young planets smaller than $3$\,$R_\oplus$ all come from either K2 or
+of][]{bouma_kep1627_2022}.  The second consideration is that all the known
+young planets smaller than $3$\,$R_\oplus$ come from either K2 or
 Kepler.  Demographic inferences based on TESS are therefore limited to
 planetary sizes $\gtrsim4$\,$R_\oplus$, for planets close-in to their
-host stars.  It seems worthwhile to compare the absolute occurrence
-rate of such planets to those observed by Kepler.  One specific
+host stars.  It would be worthwhile to compare the occurrence
+rates of both types of planets with those from the main Kepler sample.  One specific
 question that seems within reach would be to clarify whether enough
@@
 demographic studies focused on how mini-Neptune sizes evolve, the
-combined K2 and Kepler dataset might be the better primary source.
+combined K2 and Kepler dataset would be the better primary source.
 
@@
 and out-of-transit phases.  For both cases, the weak observed shifts
-seem to be caused by starspot-induced variability in specific quarters
-spuriously moving the stellar center-of-light.  Independently, our
-imaging rules out companion stars at the brightness and positions that
-would be needed to explain the reported shifts.
+are consistent with being caused by starspot-induced variability in
+specific quarters spuriously moving the stellar center-of-light.
+Independently, our imaging rules out companion stars with the brightness
+and positions that would be needed to otherwise explain the reported shifts.
 
@@
 period agrees with this assessment, the star's lithium equivalent
-width is somewhat low, which might motivate future exploration of
+width is marginally low, which might motivate future exploration of
 lithium depletion across FGKM stars in RSG-5 (see
@@
 
-The astrophysical significance of these considerations is that planets
-$\sim$2 Earth radii in size exist at ages of 40 million years.  It
+The astrophysical implicatoin of these considerations is that planets
+$\approx$2 Earth radii in size exist at ages of 40 million years.  It
 will be interesting to continue the push down to smaller planetary
@@
 
-\clearpage
+%\clearpage
 \bibliographystyle{yahapj}                            

### 2022-04-26 papers/ms.tex
update ms

@@
 regions can be verified to be the same age as the core cluster members
-through analyses of color-absolute magnitude diagrams
+through analyses of color--absolute magnitude diagrams
 \citep{KounkelCovey2019}, stellar rotation periods
@@
 close to the photometrically young stars, but which cannot be
-identified as young based on their positions in the color-absolute
+identified as young based on their positions in the color--absolute
 magnitude diagram.
@@
     are 40 to 50 million years old.} 
-    {\it Top row}: Color-absolute magnitude diagram of candidate
+    {\it Top row}: Color--absolute magnitude diagram of candidate
     Cep-Her members, plotted over candidate members of the
@@
 
-\subsubsection{Color-Absolute Magnitude Diagram}
+\subsubsection{Color--Absolute Magnitude Diagram}
 \label{sec:camd}
 
-Color-absolute magnitude diagrams (CAMDs) of the candidate RSG-5 and
+Color--absolute magnitude diagrams (CAMDs) of the candidate RSG-5 and
 CH-2 members are shown in the upper row of Figure~\ref{fig:age}.  The
@@
 consistent with the false positive rates estimated from the
-color-absolute magnitude diagrams.
+color--absolute magnitude diagrams.
 
@@
 the two epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm 1.2\,\mkms$
-respectively.  To infer the equivalent width of the Li\,\textsc{I}
+respectively.  To infer the equivalent width of the \ion{Li}{1}
 6708\,\AA\ doublet, we followed the procedure described by
@@
 $<$$1$-$\sigma$ between the two epochs.   The quoted value does not
-correct for the Fe\,\textsc{I} blend at 6707.44\AA, because HIRES
+correct for the \ion{Fe}{1} blend at 6707.44\,\AA, because HIRES
 resolves the two features.  Given the purported age and effective
@@
 $1\farcs0$ at 4.4-$\sigma$.  The reports show however that two
-outlying quarters (2 and 6) drive the offset -- the centroid locations
+outlying quarters (2 and 6) drive the offset --- the centroid locations
 from the other Kepler quarters are consistent at $\lesssim 0\farcs4$.
@@
     KOI-7913~Ab (X), and Kepler-1627~Ab (+).  Interesting trends in
-    the population of $<10^8$ year old planets include {\it i)} their
+    the population of planets younger than $10^8$ years old include {\it i)} their
     large sizes and {\it ii)} the lack of hot Jupiters.  The new
@@
 kinematics of the other Cep-Her groups shown in Figure~\ref{fig:XYZvtang}
-deserve independent attention.  An important aspect of the work that
-remains will be to acquire radial velocities of a large subset of the
+deserve independent attention.  An important aspect of the remaining work 
+will be to acquire radial velocities of a large subset of the
 stars, to assess whether the traceback approach could be applicable.
@@
 ages.  For Kepler-1643, the strongest youth indicator is its physical
-and kinematic association with RSG-5.  Based on the color-absolute
+and kinematic association with RSG-5.  Based on the color--absolute
 magnitude diagram, we are able to select members of this cluster with
@@
 % Cycle 3 and 4
-L.G.B{.} acknowledges support from the Heising-Simons Foundation (51 Pegasi~b
-Fellowship)
+L.G.B{.} acknowledges support from the Heising-Simons Foundation 51 Pegasi~b
+Fellowship
 and the TESS GI Program (NASA grants
 80NSSC21K0335 and 80NSSC22K0298).
+J.L.C{.} is supported by NSF AST-2009840 and NASA TESS G04217. 
 %
@@
 	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
-  %\texttt{astropy} \citep{astropy_2018},
-  %\texttt{astroquery} \citep{astroquery_2018},
+  \texttt{astropy} \citep{astropy_2018},
+  \texttt{astroquery} \citep{astroquery_2018},
   %\texttt{BATMAN} \citep{kreidberg_batman_2015},
@@
   %\texttt{TESS-point}  \citep{burke_2020},
-  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
+  \texttt{tesscut} \citep{brasseur_astrocut_2019},
+  \texttt{unpopular} \citep{hattorio_2021_cpm},
 \texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},

### 2022-04-29 papers/ms.tex
update ms

@@
 close-in multiplanet systems to fall out of resonance
-\citep{arevalo_stability_2022,goldberg_architectures_2022}, and
+\citep{izidoro_breaking_2017,arevalo_stability_2022,goldberg_architectures_2022}, and
 whether and how mass-loss explains the radius valley
@@
 memberships, kinematics, and possible origin is currently being
-prepared by R.~Kerr and collaborators.
+prepared as part of the 1\,kpc expansion of the SPYGLASS project
+(R.~Kerr et al.\ in prep).
 
@@
 sub-populations identified by \citet{Kerr2021} share a common origin
-has yet to be fully substantiated, and is the subject of the upcoming
-study by R.~Kerr and collaborators.  Our primary interest in the
-region stems from the fact that a portion of it was observed by Kepler
-(Figure~\ref{fig:XYZvtang}, top panel).  To further explore the
-population of stars that were observed, we select candidate Cep-Her
-members through four steps, the first three being identical to those
-described in Section~3 of \citet{Kerr2021}.  We briefly summarize them
-here.
+has yet to be fully substantiated, but preliminary clustering results
+from the 1\,kpc SPYGLASS analysis (R.~Kerr et al.\ in prep) suggest
+the presence of contiguous stellar populations connecting each of
+these groups.  In this work, our primary interest in the region stems
+from the fact that a portion of it was observed by Kepler
+(Figure~\ref{fig:XYZvtang}, top panel).  To further explore this
+sub-population, we select candidate Cep-Her members through four
+steps, the first three being identical to those described in Section~3
+of \citet{Kerr2021}.  We briefly summarize them here.
 
@@
 which clusters cannot be fragmented as $25$\,pc in physical space, and
-$c$\,\kms\ in velocity.  The minimum cluster size $N$ is set to 10, as
+$25/c$\,\kms\ in velocity.  The minimum cluster size $N$ is set to 10, as
 is $k$, the parameter used to define the ``core distance'' density
@@
 
-This unsupervised clustering in our case yielded 8 distinct groups.
+This unsupervised clustering in this case yielded 8 distinct subgroups.
 These groups are then used as the ``seed'' populations for the third
@@
 out field-star contaminants, we also compute a weight metric,
-$\alpha$, defined such that the group member with the smallest core
-distance has $\alpha=1$, the group member with the greatest core
-distance has $\alpha=0$, and weights for the other group members are
+$D$, defined such that the group member with the smallest core
+distance has $D=1$, the group member with the greatest core
+distance has $D=0$, and weights for the other group members are
 log-normally distributed between these two extremes.  After applying a
@@
 $G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm BP}/\sigma_{G_{BP}}>20$.}
-this procedure yields $\alpha \sim \log_{10}\mathcal{N}(-1.55,0.61)$.
+this procedure yields $D \sim \log_{10}\mathcal{N}(-1.55,0.61)$.
 To visualize the results, in Figure~\ref{fig:XYZvtang} we show
-12{,}436 objects with $\alpha>0.02$ as gray points, and 
-4{,}763 objects with $\alpha>0.10$ as black points.  The previously
+12{,}436 objects with $D>0.02$ as gray points, and 
+4{,}763 objects with $D>0.10$ as black points.  The previously
 known $\delta$~Lyr\ cluster is visible at $(l,b)=(68^\circ,15^\circ$)
@@
 \end{align}
-and we call this latter set of stars ``CH-2''.  These cuts yield
+and we call this latter set of stars ``CH-2'', using the preliminary 
+Cep-Her (CH) subgroup identifier from R.~Kerr et al.\ (in prep).
+These cuts yield
 \nrsgfive\ candidate RSG-5 members, and \nchtwo\ candidate CH-2
@@
 USco ($\sim$$8$\,Myr), and some may still exist at ages of LCC
-($\sim$$16$\,Myr; L.~Rebull in preparation).  Regardless, since
+($\sim$$16$\,Myr; L.~Rebull submitted).  Regardless, since
 nearly no field star outliers seem to be present on the RSG-5 CAMD,
@@
 the two epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm 1.2\,\mkms$
-respectively.  To infer the equivalent width of the \ion{Li}{1}
-6708\,\AA\ doublet, we followed the procedure described by
-\citet{bouma_2021_ngc2516}.  This yielded a strong detection: ${\rm
-EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with values consistent at
-$<$$1$-$\sigma$ between the two epochs.   The quoted value does not
-correct for the \ion{Fe}{1} blend at 6707.44\,\AA, because HIRES
-resolves the two features.  Given the purported age and effective
-temperature of the star, the lithium equivalent width is somewhat low.
-We discuss this in greater depth in Appendix~\ref{app:spectra}.
+respectively \citep{chubak_2012}.  To infer the equivalent width of
+the \ion{Li}{1} 6708\,\AA\ doublet, we followed the procedure
+described by \citet{bouma_2021_ngc2516}.  This yielded a strong
+detection: ${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with values
+consistent at $<$$1$-$\sigma$ between the two epochs.   The quoted
+value does not correct for the \ion{Fe}{1} blend at 6707.44\,\AA,
+because HIRES resolves the two features.  Given the purported age and
+effective temperature of the star, the lithium equivalent width is
+somewhat low.  We discuss this in greater depth in
+Appendix~\ref{app:spectra}.
 
@@
     Phase-folded transits of Kepler-1643, KOI-7913, KOI-7368, and
-    Kepler-1627 with stellar variability removed.  The maximum {\it a
-    posteriori} transit model is shown with the gray line, and the
+    Kepler-1627 with stellar variability removed.  The maximum a
+    posteriori transit model is shown with the gray line, and the
     residual after subtracting the transit model is vertially
@@
 \citep{hoffman_no-u-turn_2014}, after having initialized to the the
-maximum {\it a posteriori} (MAP) model.  We used the
+maximum a posteriori (MAP) model.  We used the
 \citet{gelman_inference_1992} statistic, $\hat{R}$, as our convergence
@@
 
-\subsection{Is CH-2 a real cluster?}
+\subsection{Is CH-2 real?}
 \label{subsec:ch2}
@@
 formation.  The ``12'' is set to distinguish star clusters from
-high-order multiples \citep[see][]{krumholz_star_2019}.  Dissolved
-clusters and their tidal tails are therefore included in our concept
-of clusters.  We do not however require a threshold number of stars
-per unit spatial volume for a group to be a ``cluster''.  The latter
-point acknowledges that an important factor in cluster identification
-is also the density in velocity space.  Perhaps once stellar rotation
+high-order multiples \citep[see][]{krumholz_star_2019}.  We explicitly
+do not require a ``star cluster'' to be gravitationally bound:
+dissolved clusters as well as their tidal tails are included in our
+adopted definition of ``clusters''.  We similarly do not require a
+threshold number of stars per unit spatial volume.  The latter point
+acknowledges that an important factor in cluster identification is
+also density in velocity space.  Perhaps once stellar rotation
 periods and chemical abundances reach the same level of ubiquity as
 stellar proper motions, they might enable further refinement of our
-ability to discover stars that formed together.
+ability to discover stars that formed as part of the same event.
 
@@
 if they share a common age.  This approach can demonstrate whether a
-star is currently associated with a set of coeval stars, which
-is much easier than determining what the association looked like in
+star is currently associated with a set of coeval stars, though it
+falls short of determining what the association looked like in
 the past.  Our analysis of CH-2 meets the latter standard for the
-existence of a $\approx$40\,Myr assocation.  It would be a worthy
-exercise for future work to perform a similar search for coeval
+existence of a $\approx$40\,Myr stellar assocation.  It would be a
+worthy exercise for future work to perform a similar search for coeval
 phase-space neighbors on the entire dataset of known exoplanet hosts.
@@
 % Cycle 3 and 4
-L.G.B{.} acknowledges support from the Heising-Simons Foundation 51 Pegasi~b
-Fellowship
-and the TESS GI Program (NASA grants
-80NSSC21K0335 and 80NSSC22K0298).
-J.L.C{.} is supported by NSF AST-2009840 and NASA TESS G04217. 
+L.G.B{.} acknowledges support from the Heising-Simons Foundation 51
+Pegasi~b Fellowship and the TESS GI Program (NASA grants 80NSSC21K0335
+and 80NSSC22K0298).
+J.L.C{.} is supported by NSF AST-2009840 and the TESS GI Program (NASA
+grant 80NSSC22K0299).
 %
@@
 %
-This paper also includes data collected by the TESS mission, which are
-publicly available from the Mikulski Archive for Space Telescopes
-(MAST).
-%
-Funding for the TESS mission is provided by NASA's Science Mission
-directorate.
-%
-We thank the TESS Architects
-%(G.~Ricker, R.~Vanderspek, D.~Latham, S.~Seager, J.~Jenkins)
-and the many TESS team members for their
-efforts to make the mission a continued success.
+%This paper includes data collected by the TESS mission, which are
+%publicly available from the Mikulski Archive for Space Telescopes
+%(MAST).
+%%
+%Funding for the TESS mission is provided by NASA's Science Mission
+%directorate.
+%%
+%We thank the TESS architects
+%%(G.~Ricker, R.~Vanderspek, D.~Latham, S.~Seager, J.~Jenkins)
+%and the many TESS team members for their
+%efforts to make the mission a continued success.
 %
@@
 %Aeronautics and Space Administration.
-Finally, we also thank the Keck Observatory staff for their support of
-HIRES and remote observing.  We recognize the importance that the
-summit of Maunakea has within the indigenous Hawaiian community, and
-are deeply grateful to have the opportunity to conduct observations
-from this mountain.
+%We also thank the Keck Observatory staff for their support of
+%HIRES and remote observing.  We recognize the importance that the
+%summit of Maunakea has within the indigenous Hawaiian community, and
+%are deeply grateful to have the opportunity to conduct observations
+%from this mountain.
 %
@@
 
-% \input{starparams.tex}
-% \input{model_posterior_table.tex}
-% \input{t1_appendix.tex}
-% \input{t2_appendix.tex}
-
 %\clearpage
@@
 \paragraph{Table~2} contains candidate Cep-Her members with weights
-$\alpha>0.02$ observed by Kepler.  More restrictive weight cuts would
-be needed if one wanted to remove the majority of field star
-interlopers; at least half of the stars included in the table are
-unlikely to be bonafide Cep-Her members.  Table~2  was created by
-cross-matching candidate Cep-Her members (selected using Gaia EDR3;
-Section~\ref{subsec:members}) against a Kepler to Gaia DR2 cross-match
-(the \texttt{gaia-kepler.fun} crossmatch database created by Megan
-Bedell).  The \texttt{kic\_dr2\_ang\_dist} column is from the latter
-table.  The EDR3 to DR2 match was performed using the
+$D>0.02$ observed by Kepler.  The final catalog of candidate
+Cep-Her members will be provided by R.~Kerr et al.\ in prep.; Table~2 is
+from a preliminary version of that same catalog.  Note that more
+restrictive weight cuts should be imposed should one wish to remove
+the majority of field star interlopers; at least half of the stars
+included in the table are unlikely to be bonafide Cep-Her members.
+Table~2  was created by cross-matching candidate Cep-Her members
+(selected using Gaia EDR3; Section~\ref{subsec:members}) against a
+Kepler to Gaia DR2 cross-match (the \texttt{gaia-kepler.fun}
+crossmatch database created by Megan Bedell).  The
+\texttt{kic\_dr2\_ang\_dist} column is from the latter table.  The
+EDR3 to DR2 match was performed using the
 \texttt{gaiaedr3.dr2\_neighbourhood} table, and the closest proper
@@
 Section~\ref{subsec:members}.  These are the data used to make the
-lower panels of Figure~\ref{fig:age}. 
+lower panels of Figure~\ref{fig:age}; as with Table~2, these are
+from a preliminary version of the SPYGLASS 1\,kpc expansion (R. Kerr et
+al.\ in prep). 
 % Table~3 includes rotation periods for 53 stars with ZTF and 51 stars

### 2022-04-29 papers/ms.tex
update ms

@@
 KOI-7368, and KOI-7913.
-%\footnote{These figures
-%yield an estimate of the size of Cep-Her: if our 4/14
-%matches represent the true positive rate, then
-%multiplying by the number of stars above our weight threshold
-%(12{,}436) yields $\approx$3{,}500 stars in the complex.  
-%This might however be an under-count,
-%since Kepler observed sparse 
-%regions of the complex, which are expected to have higher
-%contamination rates (Figure~\ref{fig:XYZvtang}, top panel).}
 
@@
 UCL \citep{pecaut_star_2016}, 38\,Myr for IC\,2602
-\citep{david_ages_2015,randich_gaiaeso_2018}
-%\footnote{ Ages for
-%IC\,2602 vary from 40 to 46\,Myr based on lithium-depletion-boundary
-%(LDB) measurements \citep{dobbie_ic_2010,randich_gaiaeso_2018}, and
-%from 30 to 46\,Myr based on isochronal analyses
-%\citep{stauffer_rotational_1997,david_ages_2015,bossini_age_2019}.},
-and 112\,Myr for the Pleiades \citep{dahm_2015}.  These assumptions
-and the subsequent processing steps taken to exclude field stars and
-binaries were identical to those described in
-\citet{bouma_kep1627_2022}.  The mean and uncertainty of the resulting
-age posterior are $46^{+9}_{-7}$\,Myr for RSG-5, and
-$36^{+10}_{-8}$\,Myr for CH-2.  For comparison, this procedure yields
-an age for the $\delta$~Lyr cluster of $38^{+6}_{-5}$\,Myr.  The
-older isochronal age of RSG-5 is consistent with its location
-relative to the $\delta$~Lyr cluster in the upper left panel of
-Figure~\ref{fig:age}.
+\citep{david_ages_2015,randich_gaiaeso_2018} and 112\,Myr for the
+Pleiades \citep{dahm_2015}.  These assumptions and the subsequent
+processing steps taken to exclude field stars and binaries were
+identical to those described in \citet{bouma_kep1627_2022}.  The mean
+and uncertainty of the resulting age posterior are $46^{+9}_{-7}$\,Myr
+for RSG-5, and $36^{+10}_{-8}$\,Myr for CH-2.  For comparison, this
+procedure yields an age for the $\delta$~Lyr cluster of
+$38^{+6}_{-5}$\,Myr.  The older isochronal age of RSG-5 is consistent
+with its location relative to the $\delta$~Lyr cluster in the upper
+left panel of Figure~\ref{fig:age}.
 
@@
 camera (FOV = 10.2\arcsec), we obtained 4 images in the $K'$ filter
-($\lambda = 2.12\,\mu$m) with a total exposure time of 320\,s. We
-analyzed these data following \citet{kraus_impact_2016}, and
-determined the detection limits to visual companions by analyzing the
-residuals after subtracting an empirical PSF template.  This procedure
-yielded contrast limits of $\Delta K' = 4.1$ mag at $\rho = 150$ mas,
-$\Delta K' = 5.8$ mag at $\rho = 300$ mas, and $\Delta K' = 8.3$ mag
+($\lambda = 2.12\,\mu$m) with a total exposure time of 320\,s. 
+The images did not show any additional visual companions.
+We analyzed the data following \citet{kraus_impact_2016}, and
+determined the detection limits by analyzing the residuals after
+subtracting an empirical PSF template.  This procedure yielded
+contrast limits of $\Delta K' = 4.1$ mag at $\rho = 150$ mas, $\Delta
+K' = 5.8$ mag at $\rho = 300$ mas, and $\Delta K' = 8.3$ mag
 at $\rho > 1000$ mas.
@@
 separation is $959 \pm 2$ au.  The Gaia EDR3 proper motions are also
-nearly identical.  Since two stars were resolved in the Kepler Input
+very similar.  Since two stars were resolved in the Kepler Input
 Catalog and are roughly one Kepler pixel apart, an accurate crowding
@@
 
-The most significant novelty about the planets in Kepler-1643, KOI-7368, and
-KOI-7913 is that their sizes are normal relative to the known
-population of mini-Neptunes from Kepler.  At field star ages,
+The most significant novelty about the planets in Kepler-1643,
+KOI-7368, and KOI-7913 is that their sizes are normal relative to the
+known population of mini-Neptunes from Kepler.  At field star ages,
 mini-Neptune sizes span 1.8\,$R_\oplus$ to 3.6\,$R_\oplus$, with the
@@
 timescale, which is tied to the entropy of the planetary interior
-shortly after
-disk dispersal, also plays a significant role
+shortly after disk dispersal, also plays a significant role
 \citep{owen_constraining_2020}.  As an example,
@@
 there should be a maximum in planet occurrence at 2 to 3\,$R_\oplus$
-at times between 10 and 100\,Myr.  The models
-advanced by \citet{gupta_signatures_2020} and
-\citet{lee_primordial_2021} agree; their differences lie in the mechanism
-for producing the radius valley. 
-% If the radius valley is
-% primordial and created by gas accretion physics, then super-Earths
-% ($<$$1.8$\,$R_\oplus$) would be common as early as $10^7$ years
-% \citep{lee_primordial_2021}.   The photoevaporation-driven models more
-% often quote $10^8$ year timescales for production of
-% atmosphere-stripped super-Earths, while the models with the mass-loss
-% driven by a slow escape of heat from the core predict longer
-% $\sim$$10^9$ year timescales (but see \citealt{king_2021}).
+at times between 10 and 100\,Myr.  The models advanced by
+\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} agree;
+their differences lie in the mechanism for producing the radius
+valley.
 
@@
 have sizes that are anomalously large relative to the predicted peak
-in planet occurrence at 2 to 3\,$R_\oplus$.
-However, their large sizes can be accommodated 
-by invoking any of {\it i)} larger core masses, {\it ii)} more
-volatile-rich compositions, {\it iii)}  larger initial atmospheric
-mass fractions, or {\it iv)} longer thermal cooling times.  
-Secure mass measurements would help constrain this parameter space,
-but the $\sim$1\,\kms\ spot-induced radial velocity semi-amplitudes
-makes measuring the Doppler orbits very difficult.
+in planet occurrence at 2 to 3\,$R_\oplus$.  However, their large
+sizes can be accommodated by invoking any of {\it i)} larger core
+masses, {\it ii)} more volatile-rich compositions, {\it iii)}  larger
+initial atmospheric mass fractions, or {\it iv)} longer thermal
+cooling times.  Secure mass measurements would help constrain this
+parameter space, but the $\sim$1\,\kms\ spot-induced radial velocity
+semi-amplitudes makes measuring the Doppler orbits very difficult.
 \citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.
@@
 acknowledges that an important factor in cluster identification is
-also density in velocity space.  Perhaps once stellar rotation
-periods and chemical abundances reach the same level of ubiquity as
-stellar proper motions, they might enable further refinement of our
-ability to discover stars that formed as part of the same event.
+also density in velocity space.  Perhaps once stellar rotation periods
+and chemical abundances reach the same level of ubiquity as stellar
+proper motions, they might enable further refinement of our ability to
+discover stars that formed as part of the same event.
 
@@
 star is currently associated with a set of coeval stars, though it
-falls short of determining what the association looked like in
-the past.  Our analysis of CH-2 meets the latter standard for the
+falls short of determining what the association looked like in the
+past.  Our analysis of CH-2 meets the latter standard for the
 existence of a $\approx$40\,Myr stellar assocation.  It would be a
@@
 suggest a common origin.  With that said, the membership and
-kinematics of the other Cep-Her groups shown in Figure~\ref{fig:XYZvtang}
-deserve independent attention.  An important aspect of the remaining work 
-will be to acquire radial velocities of a large subset of the
-stars, to assess whether the traceback approach could be applicable.
-Wide-field spectroscopic surveys such as LAMOST
+kinematics of the other Cep-Her groups shown in
+Figure~\ref{fig:XYZvtang} deserve independent attention.  An important
+aspect of the remaining work will be to acquire radial velocities of a
+large subset of the stars, to assess whether the traceback approach
+could be applicable.  Wide-field spectroscopic surveys such as LAMOST
 \citep{zhao_2012_LAMOST} or SDSS-V \citep{kollmeier_2017} could enable
@@
 line.  If Gaia DR3 provides RVS spectra for the brighter F and G
-dwarfs, these would also contain the calcium infrared triplet as another
-age indicator.  These indicators, combined with better kinematics,
-would help in definitively unravelling the formation history of the
-complex.
+dwarfs, these would also contain the calcium infrared triplet as
+another age indicator.  These indicators, combined with better
+kinematics, would help in definitively unravelling the formation
+history of the complex.
 
@@
 this might yield age or metallicity estimates for the subgroups
-independent of other methods.  Other projects could include
+independent of other methods.  Other projects could include a more
+comprehensive analysis of the stellar rotation periods,
 searches of the Kepler light curves for exocomets
-\citep{zieba_transiting_2019}, and for missed planets around the most
-rapid rotators.
+\citep{zieba_transiting_2019}, and searches for missed planets around
+the most rapid rotators.
 
@@
 models for planetary evolution.  However demographic analyses of the
-known planets between $10^7$ and $10^9$ years have so far been
-rather limited.  Approximately 40 such planets are now known
+known planets between $10^7$ and $10^9$ years have so far been rather
+limited.  Approximately 40 such planets are now known
 (Figure~\ref{fig:age}).  About half come from K2, a quarter from TESS,
@@
 experimental design of a demographic survey focused on planetary
-evolution over the first gigayear might be useful.  The first is
-that such a project requires a set of target stars with known ages.
-A promising way to compile relevant stars could be to
-combine automated spatio-kinematic clustering from Gaia with
-rotation periods measured using TESS \citep[see the appendices
-of][]{bouma_kep1627_2022}.  The second consideration is that all the known
-young planets smaller than $3$\,$R_\oplus$ come from either K2 or
-Kepler.  Demographic inferences based on TESS are therefore limited to
-planetary sizes $\gtrsim4$\,$R_\oplus$, for planets close-in to their
-host stars.  It would be worthwhile to compare the occurrence
-rates of both types of planets with those from the main Kepler sample.  One specific
-question that seems within reach would be to clarify whether enough
-young stars have been searched for the dearth of young hot Jupiters to
-be significant.  Since the hot Jupiter occurrence rate is strongly
+evolution over the first gigayear might be useful.  The first is that
+such a project requires a set of target stars with known ages.  A
+promising way to compile relevant stars could be to combine automated
+spatio-kinematic clustering from Gaia with rotation periods measured
+using TESS \citep[see the appendices of][]{bouma_kep1627_2022}.  The
+second consideration is that all the known young planets smaller than
+$3$\,$R_\oplus$ come from either K2 or Kepler.  Demographic inferences
+based on TESS are therefore limited to planetary sizes
+$\gtrsim4$\,$R_\oplus$, for planets close-in to their host stars.  It
+would be worthwhile to compare the occurrence rates of both types of
+planets with those from the main Kepler sample.  One specific question
+that seems within reach would be to clarify whether enough young stars
+have been searched for the dearth of young hot Jupiters to be
+significant.  Since the hot Jupiter occurrence rate is strongly
 dependent on stellar mass and metallicity
@@
 specific quarters spuriously moving the stellar center-of-light.
-Independently, our imaging rules out companion stars with the brightness
-and positions that would be needed to otherwise explain the reported shifts.
+Independently, our imaging rules out companion stars with the
+brightness and positions that would be needed to otherwise explain the
+reported shifts.
 

### 2022-04-29 papers/ms.tex
update ms
Adam Kraus comments round 1:

Add description of all the KOI matches.

Clarify Fig 3 upper

@@
   on open cluster dispersal.  Here we present an analysis of a group
-  stars spanning Cepheus ($l=100^\circ$) to Hercules ($l=40^\circ$),
+  of stars spanning Cepheus ($l=100^\circ$) to Hercules ($l=40^\circ$),
   hereafter the Cep-Her complex.  The group includes four Kepler
@@
   \approx (66^\circ, 12^\circ)$.
+  The Kepler footprint is shown in gray.
   {\it Middle row}: Galactic positions.  The Sun is at $(X, Y, Z) =
@@
 step, which is to search for objects at least as close to the 10$^{\rm
-th}$ nearest HDBSCAN-identified member in space-velocity coordinates.
+th}$ nearest HDBScan-identified member in space-velocity coordinates.
 This third step yields stars that are spatially and kinematically
@@
 members.  These stars are listed in Appendix~\ref{app:members}, as are
-the Cep-Her candidates that were observed by Kepler.  An important
+the Cep-Her candidates that were observed by Kepler, and all of our
+candidate matches against the Kepler Objects of Interest.  An important
 consideration, especially for CH-2, is the contamination rate by field
@@
 \citep{choi_mesa_2016}.  The statistical uncertainties from this
-technique mostly originate from the parallax uncertaintes; the
+technique mostly originate from the parallax uncertainties; the
 systematic uncertainties are taken to be the absolute difference
@@
     flux in units of parts-per-thousand ($\times 10^{-3}$).  The
-    dominant signal is induced by starspots.  {\it Bottom}:
+    dominant signal is from starspots; planetary transit times
+    are indicated but the individual transits are not visible at this scale.
+    {\it Bottom}:
     Phase-folded transits of Kepler-1643, KOI-7913, KOI-7368, and
     Kepler-1627 with stellar variability removed.  The maximum a
-    posteriori transit model is shown with the gray line, and the
-    residual after subtracting the transit model is vertially
+    posteriori model is shown with the gray line, and the
+    residual after subtracting the transit model is vertically
     displaced.  Windows over 10 hours are shown.  Gray points are
@@
 candidates, along with Kepler-1627.  In \texttt{PDCSAP},
-nonastrophysical variability is removed through a cotrending approach
+non-astrophysical variability is removed through a cotrending approach
 that uses a set of basis vectors derived by applying singular value
@@
 1.3\times10^{-4}$.  Though the transit has the lowest S/N of any of
-the objects discusssed ($\approx$$14$), its lower FPP relative to
+the objects discussed ($\approx$$14$), its lower FPP relative to
 KOI-7368 can be understood through its flat-bottomed shape,
@@
 past.  Our analysis of CH-2 meets the latter standard for the
-existence of a $\approx$40\,Myr stellar assocation.  It would be a
+existence of a $\approx$40\,Myr stellar association.  It would be a
 worthy exercise for future work to perform a similar search for coeval
@@
 another age indicator.  These indicators, combined with better
-kinematics, would help in definitively unravelling the formation
+kinematics, would help in definitively unraveling the formation
 history of the complex.
@@
 
-The astrophysical implicatoin of these considerations is that planets
+The astrophysical implication of these considerations is that planets
 $\approx$2 Earth radii in size exist at ages of 40 million years.  It
@@
 
-%\clearpage
+\clearpage
 \bibliographystyle{yahapj}                            
@@
 \appendix
-\section{Tables of Candidate Cep-Her Members}
+\section{Candidate Cep-Her Members}
 \label{app:members}
@@
 $D>0.02$ observed by Kepler.  The final catalog of candidate
-Cep-Her members will be provided by R.~Kerr et al.\ in prep.; Table~2 is
-from a preliminary version of that same catalog.  Note that more
+Cep-Her members will be provided by R.~Kerr et al.\ in prep; Table~2 is
+from a preliminary version of that work.  Note that more
 restrictive weight cuts should be imposed should one wish to remove
@@
 resolved binaries became fully resolved.
+
+\paragraph{Candidate matches between Cep-Her and the Kepler Objects of
+Interest:}
+The full list of candidate matches between Cep-Her and the Kepler
+objects of interest is as follows -- the objects are listed in order
+of descending weights, $D$.
+Objects designated as confirmed planets included
+Kepler-1627,
+Kepler-1643,
+Kepler-1331,
+Kepler-1062, and
+Kepler-1933.  % Newly validated by Valizadegan+2022
+Objects designated as candidate planets included
+KOI-5264,
+KOI-8007,
+KOI-7572,
+KOI-7375,
+KOI-7368,
+KOI-7638,
+KOI-5632, and
+KOI-7913.
+Objects designated known false positive planet candidates included
+KOI-6437, 
+KOI-5988, 
+KOI-7871,
+KOI-7655,
+KOI-5024,
+KOI-61,
+KOI-4336,
+KOI-6812,
+KOI-3399, and
+KOI-6277.
+Finally, Kepler-1902 (KOI-3090) has one confirmed planet
+(KOI-3090.02), and one false positive (KOI-3090.01).
+Of these objects, only Kepler-1627, Kepler-1643, KOI-7368, and
+KOI-7913 met our requirements for potentially
+both {\it i)} having real planets, and {\it ii)} being $\lesssim 10^8$
+years old, based on the presence of rotational modulation at the
+expected period and amplitude.  One object was ambiguous: Kepler-1933.
+This system has a confirmed $\approx$$1.4\,R_\oplus$ planet, a stellar
+rotation period of 6.5\ days, and an effective temperature of $\approx$$5750\,{\rm K}$.  This places it near the upper envelope of the
+rotation period vs.\ color distribution for the Pleiades, making
+it unlikely to be $\approx$40\,Myr old.  Nonetheless, we acquired a
+reconnaissance HIRES spectrum, and it yielded ${\rm EW}_{\rm Li} = 93 \pm
+5$\,m\AA.  Combined with the rotation period, this suggests an age
+between 100 and 200\,Myr for Kepler-1933, but the system is
+unlikely to be part of Cep-Her.
 
@@
 
-\begin{figure*}[b]
+\begin{figure*}[t]
 	\begin{center}

### 2022-04-29 drivers/plot_koikeplerlc.py

@@ -7,6 +7,7 @@ PLOTDIR = os.path.join(RESULTSDIR, 'koikeplerlc')
 if not os.path.exists(PLOTDIR):
     os.mkdir(PLOTDIR)
 
+rp.plot_koikeplerlc(PLOTDIR, [550,600])
 for t0 in np.arange(0,1500,50):
     t1 = t0+50
     rp.plot_koikeplerlc(PLOTDIR, [t0,t1])

### 2022-04-30 papers/ms.tex
update ms

@@
 consistent at $<$$1$-$\sigma$ between the two epochs.   The quoted
-value does not correct for the \ion{Fe}{1} blend at 6707.44\,\AA,
-because HIRES resolves the two features.  Given the purported age and
+value does not correct for the \ion{Fe}{1} blend at 6707.44\,\AA.  Given the purported age and
 effective temperature of the star, the lithium equivalent width is
@@
 		\subfloat{
-			\includegraphics[width=0.85\textwidth]{f3e.pdf}
+			\includegraphics[width=0.85\textwidth]{f3a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.33\textwidth]{f3a.pdf}
 			\includegraphics[width=0.33\textwidth]{f3b.pdf}
+			\includegraphics[width=0.33\textwidth]{f3c.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.33\textwidth]{f3c.pdf}
 			\includegraphics[width=0.33\textwidth]{f3d.pdf}
+			\includegraphics[width=0.33\textwidth]{f3e.pdf}
 		}
@@
 
-It bears mentioning that KOI-7368 shows a weak centroid shift in the
+It bears mentioning that KOI-7368 shows a centroid shift in the
 \texttt{q1\_q17\_dr25\_koi} validation reports, similar to
@@
 validation reports show that the shift is caused by a few outlying
-quarters (4, 5, 8, and 12).  The fact that the remaining data show a
-consistent scatter between centroid locations suggests that these
-outliers are also caused by the stellar variability.  
+quarters (4, 5, 8, and 12).  Since the remaining data show
+consistent scatter in their centroid locations, these
+outlying quarters are likely also caused by the stellar variability.  
 Our NIRC2 imaging independently shows that there are no neighboring
@@
 binary.  Repeating the calculation with our new detrending and
-contrast curves, we find a similar result: ${\rm FPP} =
+NIRC2 contrast curves, we find a similar result: ${\rm FPP} =
 1.3\times10^{-4}$.  Though the transit has the lowest S/N of any of
@@
 KOI-7368 can be understood through its flat-bottomed shape,
-combined with its long transit duration relative to plausible
+combined with its long transit duration relative to most
 eclipsing binary models (Figure~\ref{fig:planets}).  The host star
@@
 explanation is the presence of KOI-7913~B, which is located
-approximately one Kepler pixel away from Kepler-7913 A.  Our
-pixel-level analysis of the target pixel files did not show any
-evidence to support the idea that KOI-7913~B could host the transiting
-planet.  Our NIRC2 imaging similarly ruled out the possibility of any
-other plausible host star for the planet other than KOI-7913~A.
+$\approx0.9$ Kepler pixels away from Kepler-7913 A.  
+While the $\approx$1.5 pixel FWHM of the Kepler pixel response function
+implies that there is blending between the two stars,
+the target-pixel level data for KOI-7913 B reveals an
+entirely different stellar rotation
+period (Table~\ref{tab:sysparams}), and no hint of the transit signal.
+This implies that KOI-7913 B cannot host the planet.
 
@@
 		\leavevmode
-		\includegraphics[width=0.8\textwidth]{f4.pdf}
+		\includegraphics[width=0.9\textwidth]{f4.pdf}
 	\end{center}
@@
 % Cycle 3 and 4
-L.G.B{.} acknowledges support from the Heising-Simons Foundation 51
+L.G.B{.} is supported by the Heising-Simons Foundation 51
 Pegasi~b Fellowship and the TESS GI Program (NASA grants 80NSSC21K0335
 and 80NSSC22K0298).
+R.K{.} is supported by the Heising-Simons Foundation.
 J.L.C{.} is supported by NSF AST-2009840 and the TESS GI Program (NASA

### 2022-04-30 papers/ms.tex
update ms

@@
 \begin{abstract}
+  %The best way to measure a planet's age is to ignore the planet and
+  %pay attention to the star.
   Stellar positions and velocities from Gaia are yielding a new window
-  on open cluster dispersal.  Here we present an analysis of a group
+  on open cluster dispersal.  
+  Here we present an analysis of a group
   of stars spanning Cepheus ($l=100^\circ$) to Hercules ($l=40^\circ$),
@@
 on the presence of rotational modulation at the expected period and
-amplitude \citep[{e.g.},][Figure~9]{rebull_rotation_2020}.  Four
-objects remained after this inspection: Kepler-1627, Kepler-1643,
+amplitude \citep[{e.g.},][Figure~9]{rebull_rotation_2020}. 
+Four
+objects passed both cuts: Kepler-1627, Kepler-1643,
 KOI-7368, and KOI-7913.
@@
 parameters ($T_{\rm eff}, \log g, R_\star$) agreed with those from the
-cluster-isochrone method within $1$-$\sigma$, though the effective
-temperature was more precise by a factor of three.  Auxiliary
+cluster-isochrone method within $1$-$\sigma$.  Auxiliary
 spectroscopic parameters included the metallicity $[{\rm Fe/H}]= -0.02
@@
 \citep{owen_constraining_2020}.  As an example,
-\citet{rogers_unveiling_2021} predicted that given a core-mass
+\citet{rogers_unveiling_2021} predicted that given a core mass
 distribution peaked at $\approx$$4$\,$M_\oplus$, an ice-poor rock/iron
@@
 at times between 10 and 100\,Myr.  The models advanced by
-\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} agree;
+\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} broadly agree;
 their differences lie in the mechanism for producing the radius
-valley.
+valley, and in their assumed initial core mass distributions.
 
@@
 parameter space, but the $\sim$1\,\kms\ spot-induced radial velocity
-semi-amplitudes makes measuring the Doppler orbits very difficult.
+semi-amplitudes make measuring the Doppler orbits very difficult.
 \citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.
@@
 rests on the premise that non-astrophysical false positives can be
-rejected.  Kepler-1643 and KOI-7368 both pass all but one of the
-canonical vetting tests: the check for centroid offsets between the in
+rejected.  
+This seems to be the case for all three objects, even though
+Kepler-1643 and KOI-7368 
+both show weak centroid offsets between the in
 and out-of-transit phases.  For both cases, the weak observed shifts
@@
 brightness and positions that would be needed to otherwise explain the
-reported shifts.
+reported shifts.  All three objects are therefore validated planets.
 
@@
 L.G.B{.} is supported by the Heising-Simons Foundation 51
-Pegasi~b Fellowship and the TESS GI Program (NASA grants 80NSSC21K0335
+Pegasi~b Fellowship and the NASA TESS GI Program (80NSSC21K0335
 and 80NSSC22K0298).
 R.K{.} is supported by the Heising-Simons Foundation.
-J.L.C{.} is supported by NSF AST-2009840 and the TESS GI Program (NASA
-grant 80NSSC22K0299).
+J.L.C{.} is supported by NSF AST-2009840 and the NASA TESS GI Program
+(80NSSC22K0299).
+D.H. is supported by the Alfred P. Sloan Foundation and NASA (80NSSC19K0597).
 %
@@
 
-\paragraph{Table~2} contains candidate Cep-Her members with weights
+\paragraph{Table~2} contains 338 candidate Cep-Her members with weights
 $D>0.02$ observed by Kepler.  The final catalog of candidate
@@
 from a preliminary version of that work.  Note that more
-restrictive weight cuts should be imposed should one wish to remove
-the majority of field star interlopers; at least half of the stars
-included in the table are unlikely to be bonafide Cep-Her members.
+restrictive weight cuts should be imposed if one wishes to remove
+the majority of field star interlopers.
 Table~2 was created by cross-matching candidate Cep-Her members
@@
 The full list of candidate matches between Cep-Her and the Kepler
-objects of interest is as follows -- the objects are listed in order
+Objects of Interest is as follows -- the objects are listed in order
 of descending weights, $D$.
@@
 years old, based on the presence of rotational modulation at the
-expected period and amplitude.  One object was ambiguous: Kepler-1933.
+expected period and amplitude.  
+Of the
+14 confirmed and candidate planets,
+6 failed first filter, and 7 independently failed the second.
+One object was ambiguous: Kepler-1933.
 This system has a confirmed $\approx$$1.4\,R_\oplus$ planet, a stellar
@@
 5$\,m\AA.  Combined with the rotation period, this suggests an age
-between 100 and 200\,Myr for Kepler-1933, but the system is
-unlikely to be part of Cep-Her.
+for Kepler-1933 between 100 and 300\,Myr.  Based on these indicators, the system is
+unlikely to be part of Cep-Her, but could merit further study.
 

### 2022-04-30 papers/ms.tex
update ms

@@
 
-\subsection{Is CH-2 real?}
+\subsection{Is CH-2 a Coeval Population?}
 \label{subsec:ch2}
@@
 
-Given its low density, in CH-2 really a star cluster?  For this
+Given its low density, in CH-2 a real star cluster?  For this
 discussion, we adopt the definition that a star cluster is a group of
@@
 falls short of determining what the association looked like in the
-past.  Our analysis of CH-2 meets the latter standard for the
-existence of a $\approx$40\,Myr stellar association.  It would be a
-worthy exercise for future work to perform a similar search for coeval
-phase-space neighbors on the entire dataset of known exoplanet hosts.
-For the time being, we can offer the anecdotal point that in our
-experience, most stars do not have dozens of 40\,Myr neighbors within
-a local volume of a few \kms\ and tens of parsecs.
+past.  While our analysis of CH-2 meets the latter standard for
+demonstrating the
+existence of a $\approx$40\,Myr stellar association, it does not meet
+the standard of demonstrating that its stars were associated at birth. 
+It would be a worthy exercise for future work to perform a similar
+search for coeval phase-space neighbors on the entire dataset of known
+exoplanet hosts.  For the time being, we can offer the anecdotal point
+that in our experience, most stars do not have dozens of 40\,Myr
+neighbors within a local volume of a few \kms\ and tens of parsecs.
 

### 2022-04-30 papers/ms.tex
update ms

@@
   %pay attention to the star.
-  Stellar positions and velocities from Gaia are yielding a new window
-  on open cluster dispersal.  
+  Stellar positions and velocities from Gaia are yielding a new view
+  of open cluster dispersal.  
   Here we present an analysis of a group
@@
 Gaia.  It is therefore sensible to revisit the Kepler field, given our
-new knowledge of the stellar ages.
+new constraints on the stellar ages.
 
@@
 	\caption{
-		{\bf The stellar groups near KOI-7368, KOI-7913, and Kepler-1643
-    are 40 to 50 million years old.} 
+    {\bf Age-diagnostic diagrams from the stellar groups near
+    Kepler-1643, KOI-7368, and KOI-7913.} 
     {\it Top row}: Color--absolute magnitude diagram of candidate
@@
     to emphasize the pre-main-sequence.  Stars that fall far below the
-    cluster sequences are field interlopers.  {\it Bottom row}: TESS
-    and ZTF-derived stellar rotation periods, with the Pleiades
+    cluster sequences are field interlopers.  {\it Bottom row}:
+    TESS and ZTF-derived stellar rotation periods, with the Pleiades
     ($\approx 112$\,Myr) and Praesepe ($\approx 650$\,Myr) shown for
@@
 diagrams, we imposed the data filtering criteria from
-\citet[][Appendix B]{GaiaCollaboration2018}, which include binaries
+\citet{GaiaCollaboration2018}, which include binaries
 while omitting artifacts from for instance low
@@
 falls short of determining what the association looked like in the
-past.  While our analysis of CH-2 meets the latter standard for
+past.  Our analysis of CH-2 meets the latter standard for
 demonstrating the
-existence of a $\approx$40\,Myr stellar association, it does not meet
-the standard of demonstrating that its stars were associated at birth. 
+existence of a $\approx$40\,Myr stellar association.
+
 It would be a worthy exercise for future work to perform a similar
@@
 that in our experience, most stars do not have dozens of 40\,Myr
-neighbors within a local volume of a few \kms\ and tens of parsecs.
+neighbors within a local volume of a few \!\kms\ and tens of parsecs.
 
@@
  	{\it Astrometry}:
- 	Gaia \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2021_edr3}.
+ 	Gaia. % \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2021_edr3}.
  	{\it Imaging}:
@@
     %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
- 	Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
+ 	Keck:II~(NIRC2).
  	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
@@
   %PFS ({\bf CITE}),
-	Tillinghast:1.5m~(TRES; \citealt{furesz_tres_2008}).
+	Tillinghast:1.5m~(TRES).
   %MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
@@
 	%AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
- 	Keck:I~(HIRES; \citealt{vogt_hires_1994}).
+ 	Keck:I~(HIRES).
  	%VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
@@
 % 	Euler1.2m (EulerCam),
-	  Kepler \citep{borucki_kepler_2010},
+	  Kepler,
 % 	Magellan:Baade (MagIC),
@@
 % 	SOAR (SOI),
- 	  TESS \citep{ricker_transiting_2015},
+ 	  TESS,
 % 	TRAPPIST \citep{jehin_trappist_2011},
 % 	VLT:Antu (FORS2).
-    ZTF \citep{bellm_zwicky_2019}.
+    ZTF.
 }
@@
 \paragraph{Table~2} contains 338 candidate Cep-Her members with weights
-$D>0.02$ observed by Kepler.  The final catalog of candidate
-Cep-Her members will be provided by R.~Kerr et al.\ in prep; Table~2 is
-from a preliminary version of that work.  Note that more
+$D>0.02$ observed by Kepler.  The complete catalog of candidate
+Cep-Her members will be provided by R.~Kerr et al.\ in prep.\ using Gaia DR3; Table~2 is
+from an early version of that analysis based on Gaia EDR3.  Note that more
 restrictive weight cuts should be imposed if one wishes to remove

### 2022-04-30 drivers/write_RSG5_CH2_prot_table.py

@@ -95,16 +95,16 @@ rename_dict = {
 smdf = smdf.rename(rename_dict, axis='columns')
 
 round_dict = {
-    'ra_EDR3': 4,
-    'dec_EDR3': 4,
+    'ra_EDR3': 5,
+    'dec_EDR3': 5,
     'parallax_EDR3': 4,
     'ruwe_EDR3': 3,
-    'weight': 5,
-    'v_l': 4,
-    'v_b': 4,
-    'x_pc': 2,
-    'y_pc': 2,
-    'z_pc': 2,
+    'weight': 3,
+    'v_l': 2,
+    'v_b': 2,
+    'x_pc': 1,
+    'y_pc': 1,
+    'z_pc': 1,
     '(M_G)0': 3,
     '(BP-RP)0': 3,
     'Prot_Adopted': 3,

### 2022-04-30 rudolf/plotting.py

@@ -5394,6 +5394,7 @@ def plot_CepHer_XYZvtang_sky(outdir, showgroups=0):
             )
 
             # Schonrich+2010 solar velocity wrt local standard of rest
+            # note the systemic uncertainties are not included
             U = 11.10 * u.km/u.s
             U_hi = 0.69 * u.km/u.s
             U_lo = 0.75 * u.km/u.s

### 2022-05-01 papers/ms.tex
update ms

@@
 the $\delta$\ Lyr\ cluster seems to also be part of a much larger
-group of similarly aged stars.  This group, which is at a distance of
-$\sim$330\,pc from the Sun, spans Cepheus to Hercules (galactic
-longitudes, $l$, between 40$^\circ$ and 100$^\circ$), at galactic
-latitudes between 0$^\circ$ and 20$^\circ$.  We therefore refer to it
-as the Cep-Her complex.  It exhibits significant sub-structure over
-its $\approx$250 parsec length, and a detailed analysis of its
-memberships, kinematics, and possible origin is currently being
-prepared as part of the 1\,kpc expansion of the SPYGLASS project
-(R.~Kerr et al.\ in prep).
+group of similarly aged stars.  This association, which is at an
+average distance of 330\,pc from the Sun, spans Cepheus to Hercules
+(galactic longitudes, $l$, between 40$^\circ$ and 100$^\circ$), at
+galactic latitudes between 0$^\circ$ and 20$^\circ$.  An assessment of
+its membership, substructure, and age distribution will be provided as
+part of the 1\,kpc expansion of the SPYGLASS project (R. Kerr et al.\
+in prep), where it is given the name Cep-Her, after the endpoint
+constellations.
 
 Our focus is on the intersection of the Cep-Her complex with the
-Kepler field.  Cross-matching our Cep-Her members against known Kepler
+Kepler field.  Cross-matching the stars thought to be in Cep-Her against known Kepler
 Objects of Interest (KOIs; \citealt{thompson_planetary_2018}) yielded
@@
   points are stars for which group membership is more secure than for
-  gray points.  Kepler-1627 is in the outskirts of the $\delta$ Lyr
+  gray points.  Kepler-1627 is in the outskirts of the $\delta$\ Lyr
   cluster \citep{bouma_kep1627_2022}, which is centered at $(l,b)
@@
   projection of the Solar velocity with respect to the local standard
-  of rest.  There is a strong spatial and kinematic overlap between
+  of rest \citep{schonrich_local_2010}.  There is a strong spatial and kinematic overlap between
   Kepler-1643 and RSG-5 (magenta).  The local population
@@
 longitudes of $40^\circ$ to $100^\circ$.  Two rich clusters in
-this region are the $\delta$~Lyra cluster
+this region are the $\delta$\ Lyr cluster
 \citep{stephenson_possible_1959} and RSG-5 \citep{roser_nine_2016}.
@@
 the presence of contiguous stellar populations connecting each of
-these groups.  In this work, our primary interest in the region stems
+these groups in space-velocity coordinates.  
+In other words, the stars appear to be co-moving, though with a
+continuous gradient in velocity as a function of position
+(Figure~\ref{fig:XYZvtang}, lower panels).
+In this work, our primary interest in the region stems
 from the fact that a portion of it was observed by Kepler
@@
 
-The second step is to perform an unsupervised HDBScan clustering on
+The second step is to perform an unsupervised HDBSCAN clustering on
 the photometrically selected population
@@
 computed assuming the \texttt{astropy v4.0} coordinate standard
-\citep{astropy_2018}, which places the Sun $8122$ pc from the galactic
-center, and assumes the solar velocity with respect to the local
-standard of rest from \citet{schonrich_local_2010}.  As input
-parameters to HDBScan, we set the minimum $\epsilon$ threshold past
+\citep{astropy_2018}. %, which places the Sun $8122$ pc from the galactic
+%center, and assumes the solar velocity with respect to the local
+%standard of rest from \citet{schonrich_local_2010}.  
+As input
+parameters to HDBSCAN, we set the minimum $\epsilon$ threshold past
 which clusters cannot be fragmented as $25$\,pc in physical space, and
@@
 
-This unsupervised clustering in this case yielded 8 distinct subgroups.
+The unsupervised clustering in this case yielded 8 distinct subgroups.
 These groups are then used as the ``seed'' populations for the third
 step, which is to search for objects at least as close to the 10$^{\rm
-th}$ nearest HDBScan-identified member in space-velocity coordinates.
+th}$ nearest HDBSCAN-identified member in space-velocity coordinates.
 This third step yields stars that are spatially and kinematically
@@
 set of quality cuts on the astrometry and
-photometry,\footnote{$\varpi/\sigma_\varpi>5$; $G/\sigma_{G}>50$;
-$G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm BP}/\sigma_{G_{BP}}>20$.}
+photometry\footnote{$\varpi/\sigma_\varpi>5$; $G/\sigma_{G}>50$;
+$G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm BP}/\sigma_{G_{BP}}>20$}
 this procedure yields $D \sim \log_{10}\mathcal{N}(-1.55,0.61)$.
@@
 12{,}436 objects with $D>0.02$ as gray points, and 
-4{,}763 objects with $D>0.10$ as black points.  The previously
-known $\delta$~Lyr\ cluster is visible at $(l,b)=(68^\circ,15^\circ$)
+4{,}763 objects with $D>0.10$ as black points.  The 
+$\delta$~Lyr\ cluster is visible at $(l,b)=(68^\circ,15^\circ$)
 and $(v_{l'}, v_b)=(-4.5 ,-4 )\mkms$.  RSG-5 is visible at
@@
 weights exceeding 0.02, this yielded 11 known false positives, 6
-``confirmed'' planets, and 8 ``candidate'' planets.  Inspection of the
-Kepler data validation summaries and Robovetter classifications for
-these objects showed whether they were potentially consistent with
-being {\it i)} planets, and {\it ii)} $\lesssim 10^8$ years old, based
-on the presence of rotational modulation at the expected period and
-amplitude \citep[{e.g.},][Figure~9]{rebull_rotation_2020}. 
-Four
-objects passed both cuts: Kepler-1627, Kepler-1643,
-KOI-7368, and KOI-7913.
-
-Figure~\ref{fig:XYZvtang} shows the positions of the KOIs along
+confirmed planets, and 8 candidate planets (see Appendix~\ref{app:members}).
+To determine whether these objects were
+potentially consistent with being {\it i)} planets, and {\it ii)}
+$\lesssim 10^8$ years old, we inspected the Kepler data validation
+reports and Robovetter classifications.  Youth was assessed based on
+the presence of rotational modulation at the expected period and
+amplitude \citep[{e.g.},][]{rebull_rotation_2020}. 
+Planetary status was assessed through the Robovetter flags, and by
+requiring non-grazing transits with ${\rm S/N}>10$.
+Four objects passed both cuts: Kepler-1627, Kepler-1643, KOI-7368, and
+KOI-7913.
+% FIXME: if you need to add lines, "One object (Kepler-1933) was ambiguous,
+
+Figure~\ref{fig:XYZvtang} shows the positions of these KOIs along
 various projections.  Kepler-1643 is near the core RSG-5 population
@@
 closer to the Sun in $Y$.  In tangential galactic velocity space,
-there is some kinematic overlap between the region the latter two KOIs
-are in and the main RSG-5 group.
+there is some kinematic overlap between the region containing the latter two KOIs
+and the main RSG-5 group.
 
@@
 We define two sets of stars in the local vicinity of our objects of
-interest.  For candidate RSG-5 members, we require:
+interest.  For candidate RSG-5 members near Kepler-1643, we require:
 \begin{align}
@@
 \end{align}
-though the cluster does have a greater spatial extent toward smaller
+though RSG-5 does have a greater spatial extent toward smaller
 $X$ (Figure~\ref{fig:XYZvtang}, middle panels).  For the diffuse stars
@@
 Cep-Her (CH) subgroup identifier from R.~Kerr et al.\ (in prep).
-These cuts yield
+These cuts yielded
 \nrsgfive\ candidate RSG-5 members, and \nchtwo\ candidate CH-2
 members.  These stars are listed in Appendix~\ref{app:members}, as are
-the Cep-Her candidates that were observed by Kepler, and all of our
-candidate matches against the Kepler Objects of Interest.  An important
-consideration, especially for CH-2, is the contamination rate by field
-stars.  We assess this in the following section.
+the Cep-Her candidates that were observed by Kepler. 
+%An important
+%consideration, especially for CH-2, is the contamination rate by field
+%stars.  We assess this in the following section.
 
@@
 left panel of Figure~\ref{fig:age}.
+Generally speaking, this method is expected to be accurate provided
+that the metallicities of IC\,2602 and the Cep-Her groups (RSG-5,
+CH-2, and the $\delta$~Lyr cluster) are roughly identical.  The
+spectroscopic metallicities that we find in Section~\ref{sec:stars}
+suggest that this is indeed the case.
 
@@
 in the preceding numerator, are M-dwarfs with rotation periods between
-10 and 12.5 days.  As just discussed, the age interpretation for these
+10 and 12.5 days.  As previously noted, the age interpretation for these
 M-dwarfs is not obvious.  Regardless, the $\approx$15\% false positive
@@
 both show weak centroid offsets between the in
-and out-of-transit phases.  For both cases, the weak observed shifts
+and out-of-transit phases.  For both cases, the observed shifts
 are consistent with being caused by starspot-induced variability in
@@
 Independently, our imaging rules out companion stars with the
-brightness and positions that would be needed to otherwise explain the
-reported shifts.  All three objects are therefore validated planets.
+brightnesses and positions that would be needed to explain the
+reported shifts.  All three objects are therefore most likely planets.
 

### 2022-05-01 papers/ms.tex
update ms

@@
   diffuse -- we call this region ``CH-2'' (lime-green).
-  The selection method for all of the stars displayed
+  The selection method for all of the stars 
   is described in Section~\ref{subsec:members}.
@@
 these groups in space-velocity coordinates.  
-In other words, the stars appear to be co-moving, though with a
+In other words, the stars appear to be comoving, though with a
 continuous gradient in velocity as a function of position
@@
 reduction were identical as before.  The images showed KOI-7913 A,
-KOI-7913 B, and an additional faint star $\approx$$0\farcs99$ due East
+KOI-7913 B, and an additional faint neighbor $\approx$$0\farcs99$ due East
 of KOI-7913 B.  Applying the PSF-fitting routines from
-\citet{kraus_impact_2016}, the tertiary star has a separation $\rho =
+\citet{kraus_impact_2016}, the tertiary object has a separation $\rho =
 4397 \pm 3$\,mas from the primary, at a position angle $231.17^\circ
 \pm 0.02^\circ$, with $\Delta K' = 6.97 \pm 0.04$.  While it is too
-faint to affect the interpretation of the KOI itself, it would be
-amusing if the faint neighbor were also comoving and young -- at the
-age of the system, it would have a mass between 10 and
+faint to affect the interpretation of the transit signal, it would be
+amusing if this faint neighbor were comoving and therefore part of the system, 
+because it would have a mass between 10 and
 15\,M$_{\rm Jup}$.  Additional imaging epochs will tell.
@@
 core composition, and a typical H/He mass fraction of $\approx$4\%,
-there should be a maximum in planet occurrence at 2 to 3\,$R_\oplus$
+there should be a local maximum in planet occurrence rates at 2 to 3\,$R_\oplus$,
 at times between 10 and 100\,Myr.  The models advanced by
-\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} broadly agree;
+\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} agree;
 their differences lie in the mechanism for producing the radius
-valley, and in their assumed initial core mass distributions.
+valley, and in whether a population of rocky planets is predicted to exist
+at the time of disk dispersal.
 
@@
 parameter space, but the $\sim$1\,\kms\ spot-induced radial velocity
-semi-amplitudes make measuring the Doppler orbits very difficult.
+semi-amplitudes make measuring the Doppler orbits very difficult
 \citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.

### 2022-05-01 papers/ms.tex
update ms

@@
 %% Add "Submitted to " argument.
-\received{---}
+\received{2022 May 2}
 \revised{---}
@@
   and KOI-7913 respectively.  These planets demonstrate
-  that mini-Neptunes with sizes of $\sim$2 Earth radii exist at ages
+  that mini-Neptunes with sizes of $\approx$2 Earth radii exist at ages
   of 40 million years.
@@
 
-A few salient properties of the Kepler objects in Cep-Her can be
+Many of the salient properties of the Kepler objects of interest in Cep-Her can be
 gleaned from Figure~\ref{fig:age}.  The stars span spectral types of
@@
 constrain the existence of visual companions, including possible bound
-binaries.  The system-by-system details follow, and the relevant
-results are also summarized in Table~\ref{tab:sysparams}.
+binaries.  We give the system-by-system details in Sections~\ref{subsec:kep1643}
+through~\ref{subsec:koi7913}, and summarize their implications for
+the youth of the stars in Section~\ref{subsec:specyouth}.
 
 \subsection{Kepler\,1643}
+\label{subsec:kep1643}
 
@@
 somewhat low.  We discuss this in greater depth in
-Appendix~\ref{app:spectra}.
+Section~\ref{subsec:specyouth}.
 
@@
 \subsection{KOI-7913}
+\label{subsec:koi7913}
 
@@
 are consistent with a $\approx$40 Myr age for KOI-7913 (see
-Appendix~\ref{app:spectra}).
+Section~\ref{subsec:specyouth}).
 
@@
 15\,M$_{\rm Jup}$.  Additional imaging epochs will tell.
+
+\subsection{Spectroscopic Youth Indicators}
+\label{subsec:specyouth}
+
+\begin{figure*}[t]
+	\begin{center}
+		\leavevmode
+			\includegraphics[width=0.9\textwidth]{f5.pdf}
+	\end{center}
+	\vspace{-0.6cm}
+	\caption{
+    {\bf Spectroscopic youth diagnostics for Kepler-1627, KOI-7368,
+    Kepler-1643, and KOI-7913 AB. }
+    The spectra are shown in the observed frame, and the stars are
+    sorted left-to-right in order of decreasing effective temperature.
+    \label{fig:koiyouthindicators}
+	}
+\end{figure*}
+
+Figure~\ref{fig:koiyouthindicators} shows key portions of the HIRES
+and TRES spectra for the Kepler objects in Cep-Her.  Lithium
+absorption is obvious at 6708\AA\ in all stars except KOI-7913 B.  
+H$\alpha$ is in emission for both components of
+KOI-7913, and in absorption for the hotter stars.  Here, we compare
+these observations against benchmark open clusters in order to
+assess their implications for the stellar ages.
+
+\subsubsection{Lithium}
+
+\begin{figure}[tp]
+	\begin{center}
+		\leavevmode
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f6a.pdf}
+		}
+
+		\vspace{-0.35cm}
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f6b.pdf}
+		}
+
+		\vspace{-0.35cm}
+		\subfloat{
+			\includegraphics[width=0.49\textwidth]{f6c.pdf}
+		}
+	\end{center}
+	\vspace{-0.6cm}
+	\caption{
+    {\bf Lithium 6708\AA\ and H$\alpha$ equivalent widths for the
+    objects of interest compared to young open clusters and field
+    stars. } Positive equivalent width means absorption; negative
+    equivalent width means emission.  
+    {\it Top \& middle}:
+    The field stars are KOIs from
+    \citet{berger_identifying_2018}.  The ``40-50 Myr'' reference
+    stars ({\it left}) are from IC\,2602 \citep{randich_gaiaeso_2018} and
+    Tuc-Hor \citep{kraus_stellar_2014}.  The ``112 Myr'' stars
+    are from the Pleiades
+    \citep{soderblom_evolution_1993,jones_evolution_1996,bouvier_pleiades_lirot_2018}.
+    The statistical uncertainties on
+    the equivalent widths are shown, or else are
+    smaller than the markers.
+    {\it Bottom}:
+    The H$\alpha$ comparison is against Tuc-Hor
+    \citep[$\approx$$40$\,Myr;][]{kraus_stellar_2014}
+    \label{fig:lithium}
+	}
+\end{figure}
+
+Figure~\ref{fig:lithium} compares the measured lithium equivalent
+widths of the Kepler objects against a few reference populations.  We
+selected reference studies from the literature only when upper limits were
+explicitly reported.  KOI-7368 and KOI-7913~A have secure lithium
+detections, while for KOI-7913~B the detection is marginal (${\rm EW}_{\rm
+Li} = 42^{+12}_{-19}$\,m\AA).  For all three stars, as well as for
+Kepler-1627~A, the observed lithium equivalent width is consistent
+with the stellar effective temperatures and a $\approx40$\,Myr age.
+
+Kepler-1643, in RSG-5, is conspicuously below the 40-50$\,$Myr
+sequence in the left-panel of Figure~\ref{fig:lithium}, though
+above the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA).
+The right panel shows the comparison against the Pleiades, where
+Kepler-1643 is more consistent with the observed dispersion in
+lithium.
+
+One explanation for the lack of Li in Kepler-1643 could be that it is a field
+interloper; another could be that RSG-5 is much older than 50\,Myr.  We do
+not favor either explanation.  
+RSG-5 cannot be much older than 50\,Myr
+based on its proximity to the $\delta$~Lyr cluster and IC\,2602 in the
+CAMD, and because it is below the Pleiades in the 
+rotation versus color diagram (Figure~\ref{fig:age}).
+Kepler-1643 also seems highly unlikely to be a field interloper, 
+because we demonstrated a $\approx$$1\%$ false positive probability
+in our spatio-kinematic selection of RSG-5 members, and there is a 
+similar independent chance
+($\approx$$1\%$) of a field K2V star having a rotation period below the
+Pleiades \citep{mcquillan_rotation_2014}.
+This yields a puzzle: how could a star have spatial, kinematic, and rotational
+evidence consistent with being in a $\approx$50\,Myr cluster,
+but a low lithium content?
+
+Our preferred explanation for Kepler-1643's meager lithium 
+content is that the reference sample of IC\,2602 and Tuc-Hor
+stars may not fully explore all possible stellar rotation
+periods and lithium equivalent widths at this age. 
+Considering the top panels of Figure~\ref{fig:lithium}, it is remarkable that in
+50 million years, stars between $4500$\,K and $5200$\,K go from
+having a tight lithium sequence to one with a dispersion
+$\approx10\times$ greater.  The existence of the Li dispersion in
+Pleiades-age K-dwarfs has been known for decades; it has also been
+known that the stars with the largest lithium abundances are also the
+most rapidly rotating
+\citep{butler_pleiades_1987,soderblom_evolution_1993}.  More recent
+analyses of this correlation have been reviewed by
+\citet{bouvier_lithium-rotation_2020}.  The conclusion of that work
+was that the origin of the rotation-lithium correlation likely lies
+within pre-main-sequence stellar physics.  If so, one would expect the
+IC\,2602 and Tuc-Hor K-dwarfs to show a larger intrinsic lithium
+dispersion.  A recent analysis of the $\approx$40\,Myr
+NGC~2547 by \citet{binks_2022} suggests that this may
+be the case, though that study only had $\approx$10 stars in the relevant
+effective temperature range.
+This suggests that
+RSG-5 and Cep-Her could be
+worthy objects for a closer analysis of the lithium-rotation correlation
+near the zero-age main sequence.
+
+\subsubsection{H$\alpha$}
+
+As shown in Figure~\ref{fig:koiyouthindicators}, H$\alpha$ is in
+emission for both components of KOI-7913, and in absorption for the
+hotter stars.  Additionally, the emission appears double-peaked for
+both of the KOI-7913 components.  An important note is that KOI-7913~A and
+KOI-7913~B were spatially resolved from each other during data
+acquisition.  Performing a cross-correlation between each of the stars
+and the nearest matches in the Keck/HIRES template library, we also
+found that the CCFs for both components of KOI-7913 showed no
+indications of double-lined binarity \citep{kolbl_detection_2015}.
+
+Balmer line emission, particularly in H$\alpha$, is expected for
+low-mass stars of this age.  \citet{kraus_stellar_2014} for instance, in their
+survey of Tuc-Hor ($\approx$40\,Myr), observed that all
+cluster members with spectral types $>$${\rm K4.5V}$ had H$\alpha$ in
+emission.  Ths is consistent with our observations: KOI-7913 shows
+H$\alpha$ in emission for both components, and in absorption
+for all of our other Kepler objects (Figure~\ref{fig:koiyouthindicators}, lower panel).  The double-peaked nature of the
+emission, though not always present, is also common for
+active stars.  Proxima Centauri, for instance, has double-peaked
+H$\alpha$ emission \citep{collins_calculations_2017}.  Given
+that we have ruled out spectroscopic binarity, the most
+likely
+explanation is self-absorption: photons near the center of the line
+see a greater optical depth from higher layers of the chromosphere,
+while photons on the wings are too far from the
+rest-wavelength to excite electrons and be re-absorbed in the upper
+layers.  The exact details of when a star's atmosphere reaches the
+conditions for such self-absorption require non-local thermal
+equilibrium models of the chromosphere
+\citep{short_chromospheric_1998,2005A&A...439.1137F}.
 
@@
 		\leavevmode
-		\includegraphics[width=0.9\textwidth]{f4.pdf}
+		\includegraphics[width=0.93\textwidth]{f4.pdf}
 	\end{center}
@@
 The most significant novelty about the planets in Kepler-1643,
-KOI-7368, and KOI-7913 is that their sizes are normal relative to the
-known population of mini-Neptunes from Kepler.  At field star ages,
-mini-Neptune sizes span 1.8\,$R_\oplus$ to 3.6\,$R_\oplus$, with the
-most common size being $\approx 2.4\,R_\oplus$
-\citep[][Figure~7]{Fulton_et_al_2017}.  The known planets younger than
+KOI-7368, and KOI-7913 is that their sizes (2.2 to 2.3\,$R_\oplus$)
+are normal relative to the known population of mini-Neptunes from
+Kepler.  At field star ages, mini-Neptune sizes span 1.8\,$R_\oplus$
+to 3.6\,$R_\oplus$, with the most common size being $\approx
+2.4\,R_\oplus$
+\citep{Fulton_et_al_2017}.  The known planets younger than
 $10^8$ years are almost all larger, with sizes between 4 and
@@
 \citep{Mann_K2_33b_2016,David_et_al_2016,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020,mann_toi1227_2022}.
-Figure~\ref{fig:rp_period_age} explores this in more detail by showing
+Figure~\ref{fig:rp_period_age} explores this by showing
 the sizes, orbital periods, and ages of the known transiting planets,
@@
 obvious overdensity relative to the local field, and our membership
-selection easily produces a clean pre-main-sequence locus in color
-versus absolute magnitude (Figure~\ref{fig:age}).  CH-2, and KOI-7913
-and KOI-7368's membership inside it, do not meet these expectations in
-as obvious a manner.  This is because the CH-2 association is diffuse.
+selection easily yielded a clean pre-main-sequence locus
+(Figure~\ref{fig:age}).  CH-2, and KOI-7913 and KOI-7368's membership
+inside it, do not meet these expectations in as obvious a manner.
+This is because the CH-2 association is diffuse.
 
@@
 
-Given its low density, in CH-2 a real star cluster?  For this
+Given its low density, in CH-2 truly a star cluster?  For this
 discussion, we adopt the definition that a star cluster is a group of
@@
 past.  Our analysis of CH-2 meets the latter standard for
-demonstrating the
-existence of a $\approx$40\,Myr stellar association.
-
-It would be a worthy exercise for future work to perform a similar
+demonstrating the existence of a $\approx$40\,Myr stellar association.
+
+It would be a worthy exercise to perform a similar
 search for coeval phase-space neighbors on the entire dataset of known
@@
 aspect of the remaining work will be to acquire radial velocities of a
-large subset of the stars, to assess whether the traceback approach
+large subset of the stars, and to determine whether the traceback approach
 could be applicable.  Wide-field spectroscopic surveys such as LAMOST
@@
 lithium depletion across FGKM stars in RSG-5 (see
-Appendix~\ref{app:spectra}).
+Section~\ref{subsec:specyouth}).
 
@@
 \acknowledgements
-\raggedbottom
+%\raggedbottom
 % Cycle 3 and 4
@@
 
-\section{Spectroscopic Youth Indicators}
-\label{app:spectra}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-			\includegraphics[width=0.99\textwidth]{f5.pdf}
-	\end{center}
-	\vspace{-0.3cm}
-	\caption{
-    {\bf Spectroscopic youth diagnostics for Kepler-1627, KOI-7368,
-    Kepler-1643, and KOI-7913 AB. }
-    The spectra are shown in the observed frame, and the stars are
-    sorted left-to-right in order of decreasing effective temperature.
-    \label{fig:koiyouthindicators}
-	}
-\end{figure*}
-
-Figure~\ref{fig:koiyouthindicators} shows key portions of the HIRES
-and TRES spectra for the Kepler objects in Cep-Her.  Lithium
-absorption is obvious at 6708\AA\ in all stars except KOI-7913 B,
-where it is marginal.  H$\alpha$ is in emission for both components of
-KOI-7913, and in absorption for the hotter stars.  In the following,
-we compare these observations against stars in benchmark open
-clusters.
-
-\subsection{Lithium}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f6a.pdf}
-			\includegraphics[width=0.49\textwidth]{f6b.pdf}
-		}
-
-		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f6c.pdf}
-		}
-	\end{center}
-	\vspace{-0.3cm}
-	\caption{
-    {\bf Lithium 6708\AA\ and H$\alpha$ equivalent widths for the
-    objects of interest compared to young open clusters and field
-    stars. } Positive equivalent width means absorption; negative
-    equivalent width means emission.  The statistical uncertainties on
-    the newly measured equivalent widths are shown, or else are
-    smaller than the markers.
-    {\it Top}:
-    The field stars are stars with Kepler planets from
-    \citet{berger_identifying_2018}.  The ``40-50 Myr'' reference
-    stars (left) are from IC\,2602 \citep{randich_gaiaeso_2018} and
-    Tuc-Hor \citep{kraus_stellar_2014}.  The ``112 Myr'' stars (right)
-    are from the Pleiades
-    \citep{soderblom_evolution_1993,jones_evolution_1996,bouvier_pleiades_lirot_2018}.
-    {\it Bottom}:
-    The H$\alpha$ comparison is against Tuc-Hor
-    \citep[$\approx$$40$\,Myr;][]{kraus_stellar_2014}
-    \label{fig:lithium}
-	}
-\end{figure*}
-
-Figure~\ref{fig:lithium} compares the measured lithium equivalent
-widths of the Kepler objects against a few reference populations.  We
-selected studies from the literature only where upper limits were
-explicitly reported.  KOI-7368 and KOI-7913~A have secure lithium
-detections, while for KOI-7913~B the detection is marginal (${\rm EW}_{\rm
-Li} = 42^{+12}_{-19}$\,m\AA).  For all three stars, as well as for
-Kepler-1627~A, the observed lithium equivalent width is consistent
-with the stellar effective temperatures and $\approx40$\,Myr ages.
-
-Kepler-1643, in RSG-5, is conspicuously below the 40-50$\,$Myr
-sequence in the left-panel of Figure~\ref{fig:lithium}, while still
-being above the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA).
-The right panel shows the comparison against the Pleiades, where
-Kepler-1643 is more consistent with the observed dispersion in
-lithium.
-
-One explanation for this could be that Kepler-1643 is a field
-interloper; another could be that RSG-5 is much older than 50\,Myr.  We do
-not favor either, given that {\it i)} the star was selected based
-on its spatio-kinematic proximity to other RSG-5 members, nearly all
-of which appear consistent with being $\approx$50\,Myr old
-(Figure~\ref{fig:age} upper left), and {\it ii)} the same RSG-5
-members display a gyrochronal sequence consistent with being younger
-than the Pleaides (Figure~\ref{fig:age} lower left).  
-There is a $\approx$$1\%$ chance of being a field interloper based on
-the spatio-kinematic selection, and a similar independent chance
-($\approx$$1\%$) of a field K2V star having a rotation period below the
-Pleiades \citep{mcquillan_rotation_2014}.
-The extra observation of a strong lithium detection at these
-effective temperatures yields a very small probability for Kepler-1643 
-being a field interloper.
-It also seems unlikely that RSG-5 could be much older than 50\,Myr,
-based on its proximity to the $\delta$~Lyr cluster and IC\,2602 in the
-CAMD, as well as the data shown in rotation versus color diagram.
-
-Our preferred explanation for Kepler-1643's lithium equivalent
-width is tied to the reference sample of IC\,2602 and Tuc-Hor
-stars not fully exploring the possible stellar rotation
-periods and lithium equivalent widths. 
-Analyzing Figure~\ref{fig:lithium}, it is remarkable that in
-just 50 million years, K-dwarfs between $4500$\,K and $5200$\,K go from
-having a tight lithium sequence to one with a dispersion
-$\approx10\times$ greater.  The existence of the Li dispersion in
-Pleiades-age K-dwarfs has been known for decades; it has also been
-known that the stars with the largest lithium abundances are also the
-most rapidly rotating
-\citep{butler_pleiades_1987,soderblom_evolution_1993}.  More recent
-analyses of this correlation have been reviewed by
-\citet{bouvier_lithium-rotation_2020}.  The conclusion of that work
-was that the origin of the rotation-lithium correlation likely lies
-within pre-main-sequence stellar physics.  If so, one would expect the
-IC\,2602 and Tuc-Hor K-dwarfs to show a larger intrinsic lithium
-dispersion.  A recent analysis of the $\approx$40\,Myr
-NGC~2547 by \citet{binks_2022} suggests that this may
-be the case, though it only had $\approx$10 stars in the relevant
-effective temperature range.
-This suggests that
-RSG-5 and Cep-Her could be
-worthy objects for a closer analysis of the lithium-rotation correlation
-near the zero-age main sequence.
-
-\subsection{H$\alpha$}
-
-As shown in Figure~\ref{fig:koiyouthindicators}, H$\alpha$ is in
-emission for both components of KOI-7913, and in absorption for the
-hotter stars.  Additionally, the emission appears double-peaked for
-both of the KOI-7913 components.  An important note is that KOI-7913~A and
-KOI-7913~B were spatially resolved from each other during data
-acquisition.  Performing a cross-correlation between each of the stars
-and the nearest matches in the Keck/HIRES template library, we also
-found that the CCFs for both components of KOI-7913 showed no
-indications of double-lined binarity \citep{kolbl_detection_2015}.
-
-Balmer line emission, particularly for H$\alpha$, is expected for
-stars of this age.  \citet{kraus_stellar_2014} for instance, in their
-survey of Tuc-Hor ($\approx$40\,Myr), observed that all
-cluster members with spectral type $>$${\rm K4.5V}$ had H$\alpha$ in
-emission.  The agrees with our observations that KOI-7913 shows
-H$\alpha$ in emission for both components, while it is in absorption
-for all of our other Kepler objects.  The double-peaked nature of the
-emission, though not always present, is also common for
-active stars.  Proxima Centauri, for instance, has double-peaked
-H$\alpha$ emission \citep{collins_calculations_2017}.  A simple
-explanation is self-absorption: photons nearest the center of the line
-see a greater optical depth from higher layers of the chromosphere,
-while photons on the wings are on average too far from the
-rest-wavelength to excite electrons and be re-absorbed in the upper
-layers.  The exact details of when a star's atmosphere reaches the
-conditions for such self-absorption require non-local thermal
-equilibrium models of the chromosphere
-\citep{short_chromospheric_1998,2005A&A...439.1137F}.
-
 \section{Table of Transit Fit Parameters}
@@
 
-\clearpage
+% \clearpage
 \input{tab_supp_CepHer_X_Kepler.tex}

### 2022-07-22 papers/paper/authors.tex
mv Cep-Her paper to new subdirectory

--- 2022-03-14 papers/authors.tex
@@
 % ... Kinematics
-\author[0000-0002-6549-9792]{R.~Kerr}
+\author[0000-0002-6549-9792]{R.~Kerr} % Y
 \affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}% % ... Kepler correlations
@@
 % ... stellar rotation & the initial crossmatch
-\author[0000-0002-2792-134X]{J. L. Curtis}
+\author[0000-0002-2792-134X]{J. L. Curtis} % Y
 \affiliation{Department of Astronomy, Columbia University, 550 West 120th Street, New York, NY 10027, USA}
-\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY 10024, USA}
 %
 % HIRES
-\author[0000-0002-0531-1073]{H. Isaacson}
+\author[0000-0002-0531-1073]{H. Isaacson} % Y
 \affiliation{Astronomy Department, University of California, Berkeley, CA 94720, USA}
@@
 % planet-fitting, kinematics, cluster.
-\author{L. A. Hillenbrand}
+\author{L. A. Hillenbrand} % R
 \affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
@@
 % HIRES Collaborators
-\author[0000-0001-8638-0320]{A. W. Howard}
+\author[0000-0001-8638-0320]{A. W. Howard} % Y
 \affiliation{Cahill Center for Astrophysics, California Institute of Technology, Pasadena, CA 91125, USA}
@@
 % AO IMAGING
-\author[0000-0001-9811-568X]{A.~L.~Kraus}
-\affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}% % ... Kepler correlations
+\author[0000-0001-9811-568X]{A.~L.~Kraus} % Y
+\affiliation{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}
 %
 % TRES
-\author[0000-0001-6637-5401]{A. Bieryla}
+\author[0000-0001-6637-5401]{A. Bieryla} % Y
 \affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
@@
 % TRES
-\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu
+\author[0000-0001-9911-7388]{D. W.~Latham} % Y
 \affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
@@
 % AO / NIRC2
-\author[0000-0003-0967-2893]{E. A.~Petigura}
+\author[0000-0003-0967-2893]{E. A.~Petigura} % Y
 \affiliation{Department of Physics \& Astronomy, University of California Los Angeles, Los Angeles, CA 90095, USA}
@@
 % AO / NIRC2
-\author[0000-0001-8832-4488]{D. Huber}
-\affiliation{Institute for Astronomy, University of Hawaii, 2680 Woodlawn Drive, Honolulu, HI 96822, USA}
+\author[0000-0001-8832-4488]{D. Huber} % R
+\affiliation{Institute for Astronomy, University of Hawai`i, 2680 Woodlawn Drive, Honolulu, HI 96822, USA}

### 2022-07-22 papers/paper/tab_supp_RSG5_CH2_Prot.tex
mv Cep-Her paper to new subdirectory

--- 2022-04-22 papers/tab_supp_RSG5_CH2_Prot.tex
@@
 %\startlongtable
-\begin{deluxetable*}{lll}
+\begin{deluxetable}{lll}
 
@@
 %\tabletypesize{\scriptsize}
-\tabletypesize{\small}
+\tabletypesize{\footnotesize}
 
@@
 \texttt{dr3\_source\_id}      & 2127562009133684480 & Gaia (E)DR3 source identifier. \\
-\texttt{ra} &    291.0231     & Gaia EDR3 right ascension [deg]. \\
-\texttt{dec} &  46.4384      & Gaia EDR3 declination [deg]. \\
+\texttt{ra} &    291.02306     & Gaia EDR3 right ascension [deg]. \\
+\texttt{dec} &  46.43843      & Gaia EDR3 declination [deg]. \\
 \texttt{parallax} &  3.7099      & Gaia EDR3 parallax [milliarcsec]. \\
 \texttt{ruwe} &  0.981      & Gaia EDR3 renormalized unit weight error. \\
-\texttt{weight} & 0.0870    & Strength of connectivity to other candidate cluster members. \\
-\texttt{v\_l} & 2.7793     & Longitudinal galactic velocity [km\,s$^{-1}$]. \\
-\texttt{v\_b} & -2.8675     & Latitudinal galactic velocity [km\,s$^{-1}$]. \\
-\texttt{x\_pc} & -8068.52  & Galactocentric $X$ position coordinate [pc]. \\
-\texttt{y\_pc} & 255.96     & Galactocentric $Y$ position coordinate [pc]. \\
-\texttt{z\_pc} & 86.25      & Galactocentric $Z$ position coordinate [pc]. \\
+\texttt{weight} & 0.087    & Strength of connectivity to other candidate cluster members. \\
+\texttt{v\_l} & 2.78     & Longitudinal galactic velocity [km\,s$^{-1}$]. \\
+\texttt{v\_b} & -2.87     & Latitudinal galactic velocity [km\,s$^{-1}$]. \\
+\texttt{x\_pc} & -8068.5  & Galactocentric $X$ position coordinate [pc]. \\
+\texttt{y\_pc} & 256.0     & Galactocentric $Y$ position coordinate [pc]. \\
+\texttt{z\_pc} & 86.3      & Galactocentric $Z$ position coordinate [pc]. \\
 \texttt{(BP-RP)0} & -0.115 &  Gaia $G_\mathrm{BP}$-$G_\mathrm{RP}$   color, minus $E$($G_\mathrm{BP}$-$G_\mathrm{RP}$). \\
@@
 \texttt{cluster} & CH-2 & RSG-5 or CH-2. \\
-\texttt{Prot\_Adopted} & NaN & Adopted rotation period [days]. \\
-\texttt{Prot\_TESS} & NaN & TESS rotation period [days]. \\
-\texttt{Prot\_ZTF}  & NaN & ZTF rotation period [days]. \\
+\texttt{Prot\_Adopted} & NaN & Adopted rotation period if available, else NaN [days]. \\
+\texttt{Prot\_TESS} & NaN & TESS rotation period if available, else NaN [days]. \\
+\texttt{Prot\_ZTF}  & NaN & ZTF rotation period if available, else NaN [days]. \\
 \texttt{Prot\_Confused} & NaN & Boolean flag; true when stars are photometrically blended. \\
@@
 \vspace{-0.5cm}
-\end{deluxetable*}
+\end{deluxetable}

### 2022-08-10 papers/paper_review/response_1.txt
[WIP] stage ~20% referee response

Dear Dr. Endl,

We thank you for organizing the review of our paper, and also thank the referee for constructive criticism.  We have considered the remarks of the reviewer carefully and revised our manuscript accordingly.

Below, we have reproduced the relevant portions of the reviewer's report, alongside our responses. The changes are bolded in the attached ms_boldchanges.pdf file, using the trackchanges AASTeX macros.

Sincerely,

Luke Bouma

******************************************
Reviewer #1 MODERATE CONCERNS

REVIEWER COMMENT
----------
Figure 1:
- It is interesting to note that all four planets are concentrated in X-Z, vb-vl*, and vl*-l. This could be because of the small fraction of space that the Kepler field surveyed. Is that accurate? Is there any way to indicated which stars were observed by Kepler in these figures?
- Note that in B&W only the + symbol has a different color than the rest, although since the symbols are different this wasn't much of an issue. The exception is that I can't tell which group is magenta and which lime green, so specifying which are the larger filled circles and which are the smaller would be helpful.
- Why is it important to show the projection of the solar velocity?
RESPONSE
----------
>

REVIEWER COMMENT
----------
If at all possible, please assure in Kerr+in prep that the nomenclature does not change such that CH-2 remains CH-2, or update this paper prior to publication.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 248: I am skeptical of the few % false positive rate (later referenced as 1%). Once stars get hotter than K2, the young and field sequences overlap significantly and so field stars wouldn't likely be outliers in the CAMD. I'm also wondering if not all of the stars are shown on the CAMD; the text states 173 stars in RSG-5; it doesn't quite look to me like there are 173 stars on the upper left panel, but I could be wrong.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Figure 2:
- I don't see any squares for the field sample, is this just the background of small indistinct points?
- Very hard to see Praesepe in black and white
- Praesepe isn't very useful as a reference since the figure is considering things younger than Pleiades. However, it would be useful to see a younger comparison cluster more similar in age to Cep-Her.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 280, "Generally speaking, this method is expected": It is not clear that this method of interpolation should be accurate, since it relies on changes occurring linearly in time. Since the ages are quite close to one of the anchors it is probably fine in any case. Please provide a citation or more discussion.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Overall, the discussion of rotation discrepancies in Sec 2.3.2 is not clear and has the feel of the authors trying overly hard to justify their argument (I do agree with the argument, however).
- It would generally be much easier to follow this section if Fig 2 contained a younger cluster for comparison, such as the two cited at the end of this section.
- After some reflection, I think the argument about the M dwarfs is that both very young M dwarfs still in the process of contracting, and moderately young M dwarfs that have just begun to spin down, can have periods of 10-12 days. I suggest simply stating this up front, and then only consider FGK stars in the subsequent discussion.
- Line 328: "no field star outliers" - but the authors have previously stated that there is one outlier.
- 13% seems like a lot of pole-on stars for random chance, and at these young ages the amplitudes should almost always be quite large. Rampalli+2021 detects a higher fraction of rotators amongst older Praesepe stars. It seems more likely that the false positive rate is similar to the other sub-associations considered.
RESPONSE
----------
>

REVIEWER COMMENT
----------
The discussion of vsini measurements did not provide adequate details, such as macroturbulence and potential impact of the targets being very young.
RESPONSE
----------
>

REVIEWER COMMENT
----------
The discussion of Li could use further clarity.
- Identification of which stars have Li is inconsistent. Line 456 says neither component of 7913 has strong Li, but at line 477 Li is obvious in all but one star, and at line 487 Li is noted as secure in 7913A.
- Line 497: "Lack of Li" implies there's no Li, not that there's not as much as expected
- As I mentioned previously, I am skeptical of the 1% false positive rate for RSG-5
- The authors point out an interesting possibility about why the Li in 1643 could be low, namely that the dispersion could be higher than expected. Though I am not an expert in Li, this seems reasonable. However, I think it should be qualified a bit more: how many stars were previously studied; few enough that scatter could be missed? Why might RSG-5 be different?
- Did the authors consider initial Li abundance, or differences in Fe abundance, as a possible reason for differences in current Li? 1643 certainly stands out the most, but I might be able to argue from Fig 4 that on average, the observed planet hosts in Cep-Her are all a bit lower than expected from the reference sequences.

RESPONSE
----------
>

REVIEWER COMMENT
----------
Figure 4, H-alpha plot:
- Only the part of the plot showing M0 and earlier is really useful here, perhaps limit the axis range.
- It is difficult to distinguish the field stars and TucHor stars in black and white
- There must be more field K stars with H-a measurements for comparison; the two showing in the late K dwarf regime don't do a good job of illustrating how H-a inactive field K dwarfs are. It can be inferred from Rauscher & Marcy (2006), although I couldn't really find an appropriate reference for field K dwarfs since 1990 (https://iopscience.iop.org/article/10.1086/132624). However, https://academic.oup.com/mnras/article/476/1/908/4828379#112872390 includes hotter stars in numerous clusters.

RESPONSE
----------
>

REVIEWER COMMENT
----------
The discussion of centroid offsets could use further elaboration.
- Line 664: how many sigma is 0.4"? Is this generally considered a good value?
- 1643 and 7368 both have centroid shifts that are larger than normal for Kepler, and it is stated that stellar variability is the likely cause. But not all young stars show centroid shifts, and presumably some level of centroid shift would still be concerning. Is the impact of stellar variability well established in the literature? Can the amplitude of stellar variability be used to inform one about the centroid offsets?
- The discussion for 7368 indicates that the centroid shift can't be due to another star due to the high contrast imaging constraints. Is there an established method for this? Can the same analysis be done for 1643?

RESPONSE
----------
>

******************************************
Reviewer #1 CLARITY, CITATIONS, ETC.

REVIEWER COMMENT
----------
Line 54, "To date...": Sentence meaning unclear. These analyses have focused on positions & velocities? Or have used clustering based on?
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 63: Unclear whether "larger" refers to size or kinematic spread, perhaps "more diffuse kinematically" if referring to kinematic spread.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 70: Consider also citing Arancibia-Silva et al. (2020) which also looked at Li in Psc-Eri
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 114: Consider adding citations for ages for delta Lyr and RSG-5.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 118: From the information given so far about delta Lyr and RSG-5, I do not know where to look in Zari+2018 figures 11 and 13, so this is not a very useful reference without additional information on the velocities and XY positions.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 122: It is worth noting that Zucker+2022 found that many of the "strings" from Kounkel & Covey are not likely real associations. Obviously, some of the associations K&C identify are real, but the extended groups that are more similar to Cep-Her may more often be artifacts.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 130, "identified three of the nearest subpopulations": subpopulations of Cep-Her?
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 140: Figure 1 reference seems to be specifically to the lower right panel, not both lower panels.
RESPONSE
----------
>

REVIEWER COMMENT
----------
*Line 162: vb and vl* not defined
RESPONSE
----------
>

REVIEWER COMMENT
----------
*Line 171: core distance metric is not defined, but is quite important per the subsequent paragraph and so should be explained. It would also be helpful to state that larger values of D are better.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 174-175, "at least as close to the 10th nearest": Not clear what this means.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 188: I understand the notation for a normal distribution, but perhaps worth just writing in words instead. I also don't know why D is Gaussian in the first place, but this would presumably be explained by defining the core distance.
RESPONSE
----------
>

REVIEWER COMMENT
----------
*Line 190-191: Please explain how thresholds of 0.02 and 0.10 were selected.
RESPONSE
----------
>

REVIEWER COMMENT
----------
*Line 192: vl' not defined
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 209: The expected period and amplitude based on stars of what age?
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 225: Phrasing makes it sound like Cep-Her candidates observed by Kepler are distinct from those in RSG-5 and CH-2, but I assume they overlap?
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 235: May be useful to readers to specify which part of Gaia Collaboration + 2018 the data filters are listed.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 246: "set the color range" had me looking for a color map, suggest saying "color axis" instead
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 251: This struck me as an odd use of "control sample" since delta Lyra seems like just another example equivalent to RSG-5 and CH-2
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 253-254: If photometrically overlapping with the field, by definition a star must be an outlier from the cluster sequence, correct? Current wording suggests otherwise, suggest rephrasing.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 260: Specify overlap is on CAMD.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 359: Stars are referred to by spectral types, but then define ZAMS in terms of mass so it is not possible to compare without additional information about the spectral type of the sun on the ZAMS.
RESPONSE
----------
>

REVIEWER COMMENT
----------
*Table 1: Please specify which component of the binary KOI 7913 orbits
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 364: From later comparisons with the spectroscopy, it seems that the MIST analysis also returns Teff.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 365: Clarify what is interpolating against the MIST isochrones.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 372: Should be stated Teff is also adopted from spectra, not just used to verify previous values.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 373, "we acquired spectra": might as well specify they are high resolution optical spectra!
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 392, citation to Chubak et al.: were these values determined following the methods of those authors, or by comparison to templates from that work?
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 394: Would be helpful to include a one-sentence description of Li measurement.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 413: Just wanted to confirm the 2015 date, presumably this was an archival spectrum from Kepler days.
RESPONSE
----------
>

REVIEWER COMMENT
----------
*Line 449: Teff values are only quoted for these two stars, and were not in previous sections, and the subsequent text also implies that for these stars, the spectra-based Teff values disagreed with the isochrone-based values. Either re-phrase, or discuss disagreement.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 471: "amusing" ?
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 472: mass of X - at the assumed age of 40Myr?
RESPONSE
----------
>

REVIEWER COMMENT
----------
Section 3.4.2.: This section makes a big deal out of H-a self-reversals, but to my knowledge these are well-known and expected since the 80s. Admittedly the references I can think of are all for M dwarfs, but I wouldn't expect a late K dwarf to be fundamentally different than an early M. The section also implies it's a possible (though ruled out) concern for binarity, but since no other spectral lines in Fig 3 are doubled this does not need to be a concern.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Line 618: for a GP regression, it is necessary to remove outliers, so this outcome is expected.
RESPONSE
----------
>

REVIEWER COMMENT
----------
Figure 5: "planetary transit times are indicated" add "by dashed vertical lines"
RESPONSE
----------
> Done, thank you.

REVIEWER COMMENT
----------
Line 678: "clears the usual threshold" - what is the usual threshold?
RESPONSE
----------
> We've updated the sentence as follows:
> "Though not as convincing as Kepler-1643, this clears the \replaced{usual threshold for calling the planet statistically validated \citep{morton_efficient_2012}}{threshold probability of 1 in 100 suggested by \citet{morton_false_2016} for calling a planet statistically validated}."
> Note the updated citation is to Morton+2016, rather than Morton 2012.  The relevant portion from Morton 2012 that we had previously considered (Section 3) has important discussion about this boundary, but the Morton+2016 citation is more relevant as the practical adoption of the 2012 method.

REVIEWER COMMENT
----------
Line 682: Please define positional probability.
RESPONSE
----------
> We have added a footnote pointing to the relevant columns at the NASA Exoplanet Archive, which contains more extensive documentation on this calculation (as does the referenced memo by Bryson & Morton 2017).  We have also reworded the text in favor of a plain-english interpretation, which is that the probability they calculated indicates that the transit signal shares its position with the target star.

REVIEWER COMMENT
----------
Line 707: The FPP is only barely lower than 7368.
RESPONSE
----------
> Good point -- the relevant text has been reworded to omit the comparison against 7368.

REVIEWER COMMENT
----------
Line 710: Please define probability usability score.
RESPONSE
----------
> We have reworded the relevant sentence to emphasize the interpretation: the positional probability calculation performed by Bryson & Morton yielded a near-unity probability that the transit event is at the same location as the host star.  We have omitted discussion of the "usability score", since the discussion in the final sentences of Section 4 provide more salient information than this metric -- the pixel level data show two distinct stellar rotation signals from the two marginally resolved stars, and no evidence for the transiting planet signal in the pixels that show the KOI-7913B rotation signal.

REVIEWER COMMENT
----------
In the discussion in 5.2, it may be worth noting that in recent years difuse associations have become more common. For example, what is the spatial density of Psc-Eri?
RESPONSE
----------
> We have added the following sentences in paragraph 3: \added{For instance, the Psc-Eri stream, which has a shape that can be approximated as a 600 parsec-long cylinder with a radius of 30 parsecs, has a number density roughly a factor of three times lower than even CH-2 \citep{roser_psceri_2020}.  However its existence is discernible because of the $\lesssim 2.5$\kms\ scatter in its cylindrical velocities. }

> This estimate imposed a G<16.5 magnitude cut to the 1387 Psc-Eri members from Roser & Schilbach 2020, which yielded 858 stars.  The shape of the stream was approximated as a 600 pc long cylinder, with radius 30 pc, and tangential velocity radius 2.5km/s, based on the figures and discussion in that analysis.  The resulting relative number density is as quoted.

REVIEWER COMMENT
----------
Line 804: Would "association" be a useful term here?
RESPONSE
----------
> The term is used at a few points throughout Section 5.2 already, and it is not clear that its usage at this exact line would necessarily add anything that it hasn't already (e.g., in the first paragraph, or second-to-last paragraph, of Section 5.2).

REVIEWER COMMENT
----------
Line 847: I am not aware of precise RVs coming from low resolution optical spectra for low mass stars. Perhaps for higher mass stars they would be useful, but in these cases I imagine DR3 will provide.
RESPONSE
----------
> We have added the qualification that LAMOST and SDSS-V spectra would likely only be useful for velocities for the brightest (higher mass) members.

REVIEWER COMMENT
----------
Line 849: I realize this is my fault since this report is quite delayed, but DR3 now is released and it should be possibly to determine (broadly, I am certainly not asking this paper to look at the DR3 data) whether DR3 RVs are going to be important.
RESPONSE
----------
> We have reworded these sentences to point out that DR3 contains reported velocities down to G_RVS of 14, or spectral types of ~K5V.

REVIEWER COMMENT
----------
*Line 925: "come from its binary" add "companion" but it would also be helpful to remind readers which component the KOI orbits.
RESPONSE
----------
> Thank you for this catch.  "binary" was a typo, and has been replaced by "binarity".  We have added the phrase "transit-hosting" when describing the primary for the suggested reminder.

REVIEWER COMMENT
----------
Table 4: is the median value the adopted value?
RESPONSE
----------
> Posterior values quoted in the text are means and standard deviations for symmetric distributions, and are otherwise medians bracketed by the upper and lower 84.1 and 15.9 percentile deviations.  We have added a table note to this effect.

REVIEWER COMMENT
----------
Line 1182: I'm confused by the period(s) for this star. "The L-S favored 6.67 days..." L-S as applied to what data? Both the ZTF period and the TESS period are quoted subsequently. Perhaps I am confused by the concept of "double dipper" but it seems unlikely for spots to make a signal appear at 2P in the TESS light curve.
RESPONSE
----------
> Thank you for this suggestion -- we've reworded this sentence for clarity.

******************************************

REVIEWER COMMENT
----------
Apologies to the authors for the very tardy report, I was sick for several weeks.
RESPONSE
----------
Thank you for the helpful report!  Health is more important than science.

******************************************

Other minor changes made to the manuscript include:

* The radius of Kepler-1643b was previously stated in the abstract to be 2.32 +/- 0.14 R_earth.  This was a typo, it is +/- 0.13 R_earth.

* We added the follow text as a note to table 2: "\added{Users who wish to minimize field star contamination should apply more restrictive weight cuts, {\it e.g.}, $\texttt{weight}>0.1$.}"

### 2022-08-10 papers/paper/ms.tex
mv Cep-Her paper to new subdirectory
[WIP] stage ~20% referee response

--- 2022-05-01 papers/ms.tex
@@
 %\documentclass[12pt,modern,twocolumn,tighten]{aastex63}
-\documentclass[12pt,twocolumn]{aastex63}
+%\documentclass[12pt,twocolumn]{aastex63}
 %\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
-%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
+\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
 \usepackage{amsmath,amstext,amssymb}
@@
 \received{2022 May 2}
-\revised{---}
+\revised{2022 Sep 6}
 \accepted{---}
@@
   Objects of Interest:
-  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.14\,R_\oplus$, $P = 5.3\ {\rm days}$),
+  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.13\,R_\oplus$, $P = 5.3\ {\rm days}$),
   KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), 
@@
     dominant signal is from starspots; planetary transit times
-    are indicated but the individual transits are not visible at this scale.
+    are indicated \added{with vertical dashed lines }but the individual transits are not visible at this scale.
     {\it Bottom}:
@@
 calculation, we find ${\rm FPP} = 4\times10^{-3}$.  Though not
-as convincing as Kepler-1643, this clears the usual threshold
+as convincing as Kepler-1643, this clears the \replaced{usual threshold
 for calling the planet statistically validated
-\citep{morton_efficient_2012}.  The S/N of the transit is
+\citep{morton_efficient_2012}}{threshold probability of 1 in 100 suggested by \citet{morton_false_2016} for calling a planet statistically validated}.  The S/N of the transit is
 $\approx$$32$, which indicates that it is unlikely to be caused by
 systematic noise in the light curve (see Figure~\ref{fig:planets}).
-The positional probability score also meets the requirements for
-transiting sources thought to share positions with their target stars
-\citep{2017ksci.rept...16B}.
+The positional probability\deleted{ score}\added{\footnote{Columns $\textit{pp\_host\_rel\_prob}$ and $\textit{pp\_host\_prob\_score}$ on the KOI Positional Probabilities table at the NASA Exoplanet Archive \citep{Akeson13}.}}\added{ calculated by \citet{2017ksci.rept...16B}} also \replaced{meets the requirements for transiting sources thought to share positions with their target stars \citep{2017ksci.rept...16B}}{indicates that the transit signal shares its position with the target star.}
 
@@
 1.3\times10^{-4}$.  Though the transit has the lowest S/N of any of
-the objects discussed ($\approx$$14$), its lower FPP relative to
-KOI-7368 can be understood through its flat-bottomed shape,
+the objects discussed ($\approx$$14$), its low\deleted{er} FPP \deleted{relative to
+KOI-7368} can be understood through its flat-bottomed shape,
 combined with its long transit duration relative to most
-eclipsing binary models (Figure~\ref{fig:planets}).  The host star
+eclipsing binary models (Figure~\ref{fig:planets}).  \replaced{The host star
 probability usability score \citep{2017ksci.rept...16B} also meets the
-usual threshold, and so the planet is statistically validated.  Its
+usual threshold, and so the planet is statistically validated}{The positional probability calculation performed by \citet{2017ksci.rept...16B} yielded a near-unity probability that the transit event is at the same location as the host star, and so the cumulative evidence suggests that KOI-7913~Ab is indeed a statistically validated planet}.
+Its
 disposition has however previously fluctuated from ``false positive''
@@
 
-Given its low density, in CH-2 truly a star cluster?  For this
+Given its low density, is CH-2 truly a star cluster?  For this
 discussion, we adopt the definition that a star cluster is a group of
@@
 acknowledges that an important factor in cluster identification is
-also density in velocity space.  Perhaps once stellar rotation periods
+also \replaced{density}{coherence} in velocity space.
+\added{For instance, the Psc-Eri stream, which has a shape that can be approximated as a 600 parsec-long cylinder with a radius of 30 parsecs, has a number density roughly a factor of three times lower than even CH-2 \citep{roser_psceri_2020}.
+However its existence is discernible because of the $\lesssim 2.5$\kms\ scatter in its cylindrical velocities.
+}Perhaps once stellar rotation periods
 and chemical abundances reach the same level of ubiquity as stellar
-proper motions, they might enable further refinement of our ability to
+proper motions, they might enable further refinement in our ability to
 discover stars that formed as part of the same event.
@@
 Figure~\ref{fig:XYZvtang} deserve independent attention.  An important
-aspect of the remaining work will be to acquire radial velocities of a
-large subset of the stars, and to determine whether the traceback approach
+aspect of the remaining work will be to acquire radial velocities \replaced{of a
+large}{for a larger} subset of the stars, and to determine whether the traceback approach
 could be applicable.  Wide-field spectroscopic surveys such as LAMOST
 \citep{zhao_2012_LAMOST} or SDSS-V \citep{kollmeier_2017} could enable
-such analyses, while also providing sensitivity to the Li 6708\,\AA\
-line.  If Gaia DR3 provides RVS spectra for the brighter F and G
+such analyses\added{ for the brightest members}, while also providing sensitivity to the Li 6708\,\AA\
+line.  \replaced{If Gaia DR3 provides RVS spectra for the brighter F and G
 dwarfs, these would also contain the calcium infrared triplet as
-another age indicator.  These indicators, combined with better
-kinematics, would help in definitively unraveling the formation
+another age indicator}{The Gaia DR3 RVS spectra (released during review of this manuscript) could contain similar velocity information down to spectral types of $\approx$K5V ($G_{\rm RVS}\lesssim 14$), and perhaps also enable analyses of the calcium infrared triplet as a youth indicator}.
+\replaced{These indicators, combined with better
+kinematics,}{The combination of more complete kinematics and youth indicators} would help in definitively unraveling the formation
 history of the complex.
@@
 more diffuse.  For KOI-7913, stronger indicators of its age come from
-its binary.  Both stellar components in KOI-7913 have isochronal ages
+its \replaced{binary}{binarity}.  Both stellar components in KOI-7913 have isochronal ages
 consistent with $40$\,Myr.  Both components also show H$\alpha$ in
-emission, which for the $\approx$K6V primary is a strong indicator
-that the star is $\lesssim$100$\,$Myr old.  KOI-7368 is slightly more
-massive, and the Li 6708\,\AA\ measurement and stellar rotation period
+emission, which for the \added{transit-hosting} $\approx$K6V primary is a strong indicator
+that the star is $\lesssim$100$\,$Myr old.  KOI-7368 is \deleted{slightly} more
+massive, and \replaced{the}{its} Li 6708\,\AA\ measurement and stellar rotation period
 provide independent verification of the star's youth.
@@
 We adopted the ZTF period over the
-TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the
+TESS period in three cases: (1) Gaia EDR3 2081755809272821248: \replaced{the
 Lomb-Scargle periodogram favored 6.67 days, consistent with the ZTF
@@
 double-dipper, which appears to have inaccurately doubled the TESS
-period to 13.34 days; (2) Gaia EDR3 2081737529891330560: we found 3.06
+period to 13.34 days}{the top ZTF Lomb-Scargle peak gave 6.61 days, while our default pipeline favored a TESS peak of 13.34 days;  manual inspection of the light curve favors the former}; (2) Gaia EDR3 2081737529891330560: we found 3.06
 days with TESS and 6.64 days with ZTF; we suspect that TESS captured
@@
 
-%\listofchanges
+\clearpage
+\listofchanges
 %\allauthors

### 2022-08-10 papers/paper/tab_supp_CepHer_X_Kepler.tex
[WIP] stage ~20% referee response

--- 2022-04-22 papers/tab_supp_CepHer_X_Kepler.tex
@@
 \texttt{kepid} & 5641711      & KIC identifier. \\
-\texttt{ra} &    297.4099     & Gaia EDR3 right ascension [deg]. \\
-\texttt{dec} &   40.8972      & Gaia EDR3 declination [deg]. \\
-\texttt{weight} & 0.0409    & Strength of connectivity to other candidate cluster members. \\
-\texttt{v\_l} & -0.5061     & Longitudinal galactic velocity [km\,s$^{-1}$]. \\
-\texttt{v\_b} & -8.2328     & Latitudinal galactic velocity [km\,s$^{-1}$]. \\
-\texttt{x\_pc} & -8035.42  & Galactocentric $X$ position coordinate [pc]. \\
-\texttt{y\_pc} & 331.44     & Galactocentric $Y$ position coordinate [pc]. \\
-\texttt{z\_pc} & 65.32      & Galactocentric $Z$ position coordinate [pc]. \\
+\texttt{ra} &    297.40986     & Gaia EDR3 right ascension [deg]. \\
+\texttt{dec} &   40.89719      & Gaia EDR3 declination [deg]. \\
+\texttt{weight} & 0.041    & Strength of connectivity to other candidate cluster members. \\
+\texttt{v\_l} & -0.51     & Longitudinal galactic velocity [km\,s$^{-1}$]. \\
+\texttt{v\_b} & -8.23     & Latitudinal galactic velocity [km\,s$^{-1}$]. \\
+\texttt{x\_pc} & -8035.4  & Galactocentric $X$ position coordinate [pc]. \\
+\texttt{y\_pc} & 331.4     & Galactocentric $Y$ position coordinate [pc]. \\
+\texttt{z\_pc} & 65.3      & Galactocentric $Z$ position coordinate [pc]. \\
 \texttt{kic\_dr2\_ang\_dist} & 0.298 & Separation between KIC and Gaia DR2 positions [arcsec]. \\
@@
 \tablecomments{Table~2 is published in its entirety in a machine-readable
-format.  One entry is shown for guidance regarding form and content.}
+format.
+One entry is shown for guidance regarding form and content.
+\added{Users who wish to minimize field star contamination should apply more restrictive weight cuts, {\it e.g.}, $\texttt{weight}>0.1$.}}
 \vspace{-0.5cm}
-\end{deluxetable*}
+\end{deluxetable}

### 2022-08-10 papers/paper_review/response_1.txt
[WIP] referee response contd

@@
 ----------
->
+> Confirmed.
 
@@
 ----------
->
+> Thank you for this wording catch -- the text has been updated to clarify that the temperatures as well as the other spectroscopic parameters agreed with the isochrone parameters within 1-sigma.
+This is a rather brief way of hiding what in our view is more complicated that what needs to be discussed.
+Table 1 states the adopted (spectroscopic) effective temperatures for all the stars.
+We reported the spectroscopic effective temperatures in all cases because their formal uncertainties were smaller than those from the isochrone method, and they seemed less model-dependent.
+The isochrone method yielded the following effective temperatures for the stars in Table 1:
+> Kepler-1643: 4878 +/- 140 K
+> KOI-7368: 5135 +/- 162 K
+> KOI-7913 A: 4286 +/- 144 K
+> KOI-7913 B 4006 +/- 335 K.
+> As discussed in the second paragraph of Section 3, the quoted values are from MIST, while the uncertainties are a quadrature sum of a statistical and systematic component (the absolute difference between the PARSEC and MIST isochrones).
+For KOI-7913 B, this systematic component is higher than for any other stars, because MIST gives Teff_MIST = 4005 +75 -73 K, while PARSEC gives Teff_PARSEC = 3679 +95 -60 K.
+This is most likely due to different assumptions in the MIST and PARSEC models for how to treat young M-dwarfs.
+The MIST models have known discrepancies between their synthetic photometry and observed open cluster data (e.g., Choi+2016 Section 8).
+Possible explanations include starspots, or incomplete molecular line lists (e.g., Mann+2013, Rajpurohit+2013).
+The PARSEC isochrones might be a better choice for the lowest mass stars (KOI-7913A and KOI-7913B) for this reason, since Chen et al (2014) implemented an empirical temperature–opacity calibration into the PARSEC models to create synthetic photometry that matches available data.
+See for instance "hr_parsec_7913b.png" and "hr_mist_7913b.png" attached -- the 39.8 Myr MIST model diverges from the IC2602 and delta-Lyr loci at BP-RP>~1.6.
+In comparison, the 39.8 Myr PARSEC model fits the observed locus down to BP-RP~=2.0.
+> Three additional plots are also included in the response: smemp_koi7913B.png, smemp_koi7913A.png, and smemp_kepler1643.png.
+These plots give some diagnostic understanding of how SpecMatch-Empirical performs for these stars.
+Note that a chi^2 of 5 corresponds to a reduced chi^2 of 1 in this model.
+While the KOI-7913B spectra is still the worst fit(!), our prefernce is for calibrating our reported effective temperatures to this spectral library, rather than to the isochronal method.
+Any suggestions on possible methodological improvements would be appreciated.
 
@@
 ----------
->
+> Depending on one's sense of humor.
 
@@
 ----------
->
+> Yes -- we've added this to the manuscript to make it explicit.
 
@@
 ----------
->
+> Enough of our team did not know about these H-alpha self-reversals that it merited some internal discussion, which yielded the second paragraph of text, including the comparison to proxima Cen and the references to NLTE models (Short & Doyle 1998) explaining the effect.
+Sometimes things that are "well-known" are worth repeating!
+Regarding binarity, yes, the first paragraph of Section 3.4.2 makes the point about double-lined binarity being ruled out more directly through analysis of the CCFs between KOI-7913A/B and the HIRES template matches.
+This is a stronger test than the visual inspection of the Li and Fe lines in the upper-right panels of Figure 3.
 
@@
 ----------
->
+> Thank you for this point -- it is true that we did not pre-mask the flares in our exploratory GP analysis, and so they may have contributed to the overfitting that we encountered.
+This does seem qualitatively surprising, since "flare-contaminated" data comprise ~0.1% of the 30-minute cadence data points (e.g., Appendix D of Bouma+22, AJ, 163, 121 found a total duration of 6.5 hours of flares over the single 90 day quarter of available 1-minute cadence data).
+Nonetheless, we've adjusted the discussion to omit the claim that fine-tuning of the hyperparameter priors would be strictly necessary, and instead focus on the reasons that we ultimately adopted our simpler model.
 

### 2022-08-10 papers/paper/ms.tex
[WIP] referee response contd

@@
 \texttt{SpecMatch-Emp} machinery yielded $T_{\rm eff,A} = 4324 \pm
-70\,{\rm K}$, and $T_{\rm eff,B} = 4038 \pm 70\,{\rm K}$.  The
-remaining parameters were in agreement with those from the cluster
-isochrone.  For the primary, we also found $[{\rm Fe/H}]= -0.06 \pm
+70\,{\rm K}$, and $T_{\rm eff,B} = 4038 \pm 70\,{\rm K}$.  \replaced{The
+remaining parameters were in agreement}{These temperatures as well as the other spectroscopic parameters agreed} with those from the cluster
+isochrone\added{ method within 1-$\sigma$}.  For the primary, we also found $[{\rm Fe/H}]= -0.06 \pm
 0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm sys} =
@@
 because it would have a mass between 10 and
-15\,M$_{\rm Jup}$.  Additional imaging epochs will tell.
+15\,M$_{\rm Jup}$\added{ at an assumed age of 40\,Myr}.  Additional imaging epochs will tell.
 
@@
 variability using a quasiperiodic Gaussian process (GP).  We explored
-this approach, but found that it often required fine-tuning of the
+this approach, but \replaced{found that it often required fine-tuning of the
 hyperparameter priors, otherwise the GP would overfit flares,
@@
 ill-founded sense of confidence in a visually clean fit.  Our model is
-simpler, and it has the benefit that the white noise jitter never
+simpler, and it has}{ultimately prefer our model for its simplicity, and for} the benefit that the white noise jitter never
 trades off with any parameter equivalent to a damping timescale for

### 2022-09-05 papers/paper_review/response_1.txt
update response

@@
 ----------
+> - The Kepler field is shown in the top panel of the figure, between roughly l of 70 to 85, and b of 5 to 20.
+This roughly corresponds to a cone in XYZ -- in case it helps intution, (l,b)=(90,0) corresponds to the line of X=0.
+The resulting concentration of the planets in vl* we believe follows from the clustering in l, which is a direct result of the fraction of space that Kepler surveyed, combined with the solar velocity projection effect (described in point three below).
 >
+> - We have specified in the caption which are the larger and smaller filled circles.
+>
+> - We have briefly summarized in the first paragraph of section 2.2 why we have shown the projection of the solar velocity: it is because we are interested in how much of the spread in v_l* it might explain.
+The short answer is "some, but not all of it".
+To give a more complete explanation: the Sun moves at around 16 km/s toward (l,b)=(53deg,23deg), relative to a hypothetical circular orbit around the center of the galaxy.
+This means that in our observed frame, nearby stars on average appear to move opposite this direction.
+At l=53 and 53+180 degrees, the effect does not influence vl.
+At l=53+90 degrees, the effect on vl (or equivalently vl*) is maximized, with amplitude = v_LSR = 16 km/s.
+The minimum is at l=53+270 degrees.
+Figure 11 of Zari, Brown et al 2018 provides some helpful intuition for how much of the variance in observed vl can be ascribed to this projection effect for stars near the galactic plane.
+The lower-right panel of our Figure 1 however shows that while the LSR projection accounts for some of the change in observed v_l* in Cep-Her, the data and the gray band do not match!
+There therefore needs to be some intrinsic differences in the UVW velocities of these stars (though v_r, v_phi, and v_z is perhaps the better coordinate system).
 
@@
 ----------
->
+> The sub-group identifier will remain the same.
 
@@
 ----------
+> Thank you for this comment -- it has helped us assess the false positive rates in RSG-5 more closely.
 >
+> In the color versus absolute magnitude diagram (CAMD), let us use (BP-RP)0 = 1.5 as the dividing line, redward of which stars can be easily seen to be on the pre-main-sequence (or at least elevated relative to typical field stars of the same color).
+For RSG-5, there are 89 stars with (BP-RP)0 > 1.5.
+One of them falls on the "field star" locus; the others fall on the pre-main-sequence locus.
+This implies that, photometrically, 88/89= 98.9% of the candidate RSG-5 members redward of (BP-RP)0=1.5 are photometrically consistent with being on the pre-main-sequence.
+>
+> Regarding more massive stars, yes, there are plenty: see hr_rsg5.png attached to the response.
+The caption of Figure 2 already explicitly states "The range of colors is truncated to emphasize the pre-main-sequence."  This visualization point is necessary because the full dynamic range in magnitude and color.
+But yes, we cannot determine whether these stars earlier than ~K2 are young based on their location in the CAMD.
+We can and do use stellar rotation.
+>
+> Using stellar rotation, we previously quoted in Section 2.3.2 that in RSG-5 41/47 stars have rotation periods consistent with being at least as fast as the Pleiades (87%).
+Of the six inconsistent stars, two are the M-dwarfs previously discussed in the text.
+The new point that emerged from our reanalysis concerns the other four stars without any detectable rotation period: three of them were the bluest three stars in our sample!
+This implies that our blue cutoff was too lenient -- the amplitudes of spot-induced modulation plummet for young stars blueward of the Kraft break (Figure 10 in Rebull+2020, 2020AJ....159..273R).
+We have propagated this through the manuscript: we now require (BP-RP)0 > 0.6, and G < 16.
+The corresponding fraction of RSG-5 stars for which detected rotation periods is 39 detections / 42 expected, i.e.
+93%.
+>
+> The manuscript has been adjusted for this discussion, including the quantitative statements regarding the pre-main-sequence K and M dwarfs, as well as the rapidly-rotating FGK stars.
 
@@
 ----------
->
+> $v_b$ is the distance-corrected proper motion in the direction of increasing galactic latitude, and $v_{l^*}=v_l \cos b$ is the same in the direction of increasing galactic longitude after accounting for the local tangent plane correction.
+We have added verbiage to this effect.
 
@@
 > Thank you for this wording catch -- the text has been updated to clarify that the temperatures as well as the other spectroscopic parameters agreed with the isochrone parameters within 1-sigma.
-This is a rather brief way of hiding what in our view is more complicated that what needs to be discussed.
+This is a rather brief way of hiding what in our view is more complicated than what needs to be discussed.
 Table 1 states the adopted (spectroscopic) effective temperatures for all the stars.
@@
 > KOI-7913 B 4006 +/- 335 K.
-> As discussed in the second paragraph of Section 3, the quoted values are from MIST, while the uncertainties are a quadrature sum of a statistical and systematic component (the absolute difference between the PARSEC and MIST isochrones).
+> As discussed in the second paragraph of Section 3, the quoted values are from MIST, while the uncertainties are a quadrature sum of a statistical and systematic component (where the latter is from the absolute difference between the PARSEC and MIST isochrones).
 For KOI-7913 B, this systematic component is higher than for any other stars, because MIST gives Teff_MIST = 4005 +75 -73 K, while PARSEC gives Teff_PARSEC = 3679 +95 -60 K.
@@
 Possible explanations include starspots, or incomplete molecular line lists (e.g., Mann+2013, Rajpurohit+2013).
-The PARSEC isochrones might be a better choice for the lowest mass stars (KOI-7913A and KOI-7913B) for this reason, since Chen et al (2014) implemented an empirical temperature–opacity calibration into the PARSEC models to create synthetic photometry that matches available data.
+The PARSEC isochrones might be a better choice for the lowest mass stars (KOI-7913A and KOI-7913B), since Chen et al (2014) implemented an empirical temperature-opacity calibration into the PARSEC models to create synthetic photometry that matches available data.
 See for instance "hr_parsec_7913b.png" and "hr_mist_7913b.png" attached -- the 39.8 Myr MIST model diverges from the IC2602 and delta-Lyr loci at BP-RP>~1.6.
@@
 Note that a chi^2 of 5 corresponds to a reduced chi^2 of 1 in this model.
-While the KOI-7913B spectra is still the worst fit(!), our prefernce is for calibrating our reported effective temperatures to this spectral library, rather than to the isochronal method.
+While the KOI-7913B spectrum is still the worst fit(!), our prefernce is for calibrating our reported effective temperatures to the HIRES spectral library, rather than to the isochronal method.
 Any suggestions on possible methodological improvements would be appreciated.

### 2022-09-05 papers/paper/ms.tex
update response

@@
   of rest \citep{schonrich_local_2010}.  There is a strong spatial and kinematic overlap between
-  Kepler-1643 and RSG-5 (magenta).  The local population
+  Kepler-1643 and RSG-5 (magenta\added{; smaller circles}).  The local population
   of candidate young stars around KOI-7368 and KOI-7913 is more
-  diffuse -- we call this region ``CH-2'' (lime-green).
-  The selection method for all of the stars 
+  diffuse -- we call this region ``CH-2'' (lime-green\added{; larger
+  circles}).
+  The selection method for \replaced{all of the stars}{these groups}
   is described in Section~\ref{subsec:members}.
@@
 the presence of contiguous stellar populations connecting each of
-these groups in space-velocity coordinates.  
-In other words, the stars appear to be comoving, though with a
-continuous gradient in velocity as a function of position
-(Figure~\ref{fig:XYZvtang}, lower panels).
-In this work, our primary interest in the region stems
+these groups in space-velocity coordinates.  In other words, the stars
+appear to be comoving, though with a continuous gradient in velocity
+as a function of position\replaced{ (Figure~\ref{fig:XYZvtang}, lower
+panels).}{. The lower panels of Figure~\ref{fig:XYZvtang} show this in
+detail, where $v_b$ is the distance-corrected proper motion in the
+direction of increasing galactic latitude, and $v_{l^*}=v_l \cos b$ is
+the distance-corrected proper motion in the direction of increasing
+galactic longitude after accounting for the local tangent plane
+correction.  Some, but not all, of the gradient in the $v_{l^*}$-$l$
+plane can be understood through a projection effect stemming from the
+Sun's motion with respect to the local standard of rest (see also
+Figure 11 by \citet{Zari2018}).  }
+In this work, our primary interest in \replaced{the region}{this
+region of sky} stems
 from the fact that a portion of it was observed by Kepler
@@
 parameters we use in the clustering are $\{ X, Y, Z, c v_b, c v_{l^*}
-\} $, where $c$ is the size-velocity corrective factor, which is taken
-as $c=6\,{\rm pc / km\,s}^{-1}$ to ensure that the spatial and
-velocity scales have identical standard deviations.  Positions are
-computed assuming the \texttt{astropy v4.0} coordinate standard
+\} $, where $c$ is
+the size-velocity corrective factor, which is taken as $c=6\,{\rm pc /
+km\,s}^{-1}$ to ensure that the spatial and velocity scales have
+identical standard deviations.  Positions are computed assuming the
+\texttt{astropy v4.0} coordinate standard
 \citep{astropy_2018}. %, which places the Sun $8122$ pc from the galactic
@@
 The CAMDs show that for RSG-5, all but one of the candidate
-members are on a tight pre-main-sequence locus.  This implies a false
+members are on a tight pre-main-sequence locus.  \added{Quantiatively,
+88/89 stars with $\bpmrpo \geq 1.5$ are consistent with being on the
+pre-main-sequence.} This implies a false
 positive rate of a few percent, at most.  By comparison, our control
@@
 
-We selected stars suitable for gyrochronology by requiring $\bpmrpo
-\geq 0.5$ to focus on FGKM stars that experience magnetic braking.
+We selected stars suitable for gyrochronology by requiring 
+\replaced{$\bpmrpo \geq 0.5$}{$\bpmrpo \geq 0.6$} to focus on FGKM stars that experience magnetic
+braking.
 For TESS, we also restricted our sample to $G<16$, to ensure the stars
@@
 ($\sim$M3V) at the relevant distances.  These cuts gave 19 stars in
-CH-2 and 47 stars in RSG-5.  We extracted light curves from the TESS
+CH-2 and \replaced{47}{42} stars in RSG-5.  We extracted light curves from the TESS
 images using the \texttt{unpopular} package \citep{hattorio_2021_cpm},
@@
 The lower panels of Figure~\ref{fig:age} show the results.  In RSG-5,
-41/47 stars have rotation periods faster than the Pleiades (87\%).
-This numerator omits the two stars with periods $>$$12$\,days visible
-in the lower-left panel of Figure~\ref{fig:age}.  The age
-interpretation for these latter stars, particularly the $\approx$M2.5
-dwarf, is not obvious.  \citet{rebull_usco_2018} for instance have
-found numerous M-dwarfs with 10-12 day rotation periods at ages of
-USco ($\sim$$8$\,Myr), and some may still exist at ages of LCC
-($\sim$$16$\,Myr; L.~Rebull submitted).  Regardless, since
-nearly no field star outliers seem to be present on the RSG-5 CAMD,
-the fact that we do not detect rotation periods for $\approx$13\% of
-stars should perhaps be taken as an indication for the fraction of
+\replaced{41/47}{39/42} stars have rotation periods \replaced{faster
+than}{at least as fast as} the Pleiades (\replaced{87}{93}\%).  This
+numerator omits the two stars with periods $>$$12$\,days visible in
+the lower-left panel of Figure~\ref{fig:age}.  The age interpretation
+for these latter stars, particularly the $\approx$M2.5 dwarf, is not
+obvious.  \citet{rebull_usco_2018} for instance have found numerous
+M-dwarfs with 10-12 day rotation periods at ages of USco
+($\sim$$8$\,Myr), and some \replaced{may}{do} still exist at ages of
+LCC ($\sim$$16$\,Myr; \replaced{L.~Rebull
+submitted}{\citealt{2022AJ....164...80R}}).  Regardless, since nearly
+no field star outliers seem to be present on the RSG-5 CAMD, the fact
+that we do not detect rotation periods for $\approx$\replaced{13}{7}\%
+of stars should perhaps be taken as an indication for the fraction of
 stars for which rotation periods might not be detectable, due to
@@
 For CH-2, 13/19 stars have rotation periods that are obviously faster
-than their counterparts in the Pleiades (68\%).  4 stars, not included
+than their counterparts in the Pleiades\deleted{ (68\%)}.  4 stars, not included
 in the preceding numerator, are M-dwarfs with rotation periods between
-10 and 12.5 days.  As previously noted, the age interpretation for these
-M-dwarfs is not obvious.  Regardless, the $\approx$15\% false positive
-rate for CH-2 determined from the CAMD seems consistent with our
-fraction of detected rotation periods; RSG-5 was missing rotation
+10 and 12.5 days.  As previously noted, the age interpretation for
+these M-dwarfs is not obvious.  \added{If none are cluster members,
+the rotation period detection fraction is 68\%; if all are members, it
+is 89\%}.
+\replaced{Regardless, the $\approx$15\% false
+positive rate for CH-2 determined from the CAMD seems consistent with
+our fraction of detected rotation periods; RSG-5 was missing rotation
 periods for $\approx$15\% of its candidate members, even though all of
 its stars appear photometrically consistent with being on a single
-pre-main-sequence locus.
+pre-main-sequence locus.}{
+This sets an upper bound on the contaminant fraction in our candidate
+CH-2 members at about one in three.  The true value may be as low as 
+one in six, based on our earlier CAMD analysis.
+}
 

### 2022-09-05 papers/paper/ms.tex
update figure stlye

@@
 
-		\vspace{-0.6cm}
+		\vspace{-0.5cm}
 		\subfloat{
@@
 pre-main-sequence locus.}{
-This sets an upper bound on the contaminant fraction in our candidate
+This sets an upper bound on the contamination fraction in our candidate
 CH-2 members at about one in three.  The true value may be as low as 
-one in six, based on our earlier CAMD analysis.
+one in six, based on the earlier CAMD analysis.
 }
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f6a.pdf}
+			\includegraphics[width=0.48\textwidth]{f6a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f6b.pdf}
+			\includegraphics[width=0.48\textwidth]{f6b.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.49\textwidth]{f6c.pdf}
+			\includegraphics[width=0.48\textwidth]{f6c.pdf}
 		}
@@
     dominant signal is from starspots; planetary transit times
-    are indicated \added{with vertical dashed lines }but the individual transits are not visible at this scale.
+    are indicated \added{with vertical dashed lines, }but the individual transits are not visible at this scale.
     {\it Bottom}:

### 2022-09-05 papers/paper_review/response_1.txt
update response

@@
 ----------
+> Yes, the field sample is the gray indistinct background.
 >
+> We have updated the Praesepe markers to be thicker, which improves their visibility.
+>
+> The attached figure rotn_ic2602_ic2391_reference.png shows a concatenation from S.
+Douglas (private communication) of TESS-derived rotation periods from IC-2602 and IC-2391, which are both similar in age to Cep-Her/RSG-5.
+These are preliminary rotation period measurements, from the same Douglas+2021 work referenced in the manuscript, which is not yet published.
+The IC-2602 and IC-2391 rotation period distributions look quite similar to those from RSG-5, though their work a) extends to lower masses, b) appears to show slightly more super-Pleiades rotation periods from F5 to G0 spectral types, and c) appears to show more stars on the fast sequence, which are lacking in the RSG-5 sample.
+Given the unpublished nature of this comparison sample, and the fact that our aim is not to firmly establish the intrinsic rotational distribution of stars at this age, we prefer to leave the reference to Douglas+2021 as it is and to not include these data in the manuscript.
 
@@
 * We added the follow text as a note to table 2: "\added{Users who wish to minimize field star contamination should apply more restrictive weight cuts, {\it e.g.}, $\texttt{weight}>0.1$.}"
+
+* We have modified the axis width on all figures for visual clarity.

### 2022-09-05 drivers/plot_rotation_vs_color.py

@@ -9,12 +9,13 @@ if not os.path.exists(PLOTDIR):
 from earhart.priors import AVG_EBpmRp
 
 # add all KOIs
-for runid in ['CH-2', 'RSG-5', 'deltaLyrCluster']:
-    rp.plot_rotationperiod_vs_color(
-        PLOTDIR, runid, yscale='linear', cleaning='defaultcleaning',
-        emph_binaries=0, talk_aspect=1, kinematic_selection=1,
-        overplotkep1627=0, show_allknown=1
-    )
+for runid in ['RSG-5', 'CH-2', 'deltaLyrCluster']:
+    for show_douglas in [1,0]:
+        rp.plot_rotationperiod_vs_color(
+            PLOTDIR, runid, yscale='linear', cleaning='defaultcleaning',
+            emph_binaries=0, talk_aspect=1, kinematic_selection=1,
+            overplotkep1627=0, show_allknown=1, show_douglas=show_douglas
+        )
 
 # eres-vii
 rp.plot_rotationperiod_vs_color(

### 2022-09-06 papers/paper_review/response_1.txt
update response

@@
 ----------
->
+> We have added the following text: 
+> \added{While in reality stellar populations do not evolve linearly in the dimensions of absolute magnitude versus color, in our case the Cep-Her loci are nearly indistinguishable from IC\,2602 (e.g., Figure~3 of \citet{bouma_kep1627_2022}).
+Systematic errors incurred in the age from the non-linear evolution are therefore likely much smaller than the $\approx$10\,Myr systematic uncertainty in the absolute reference age for IC\,2602 itself \citep{david_ages_2015,randich_gaiaeso_2018}.
 
@@
 ----------
+> - Regarding younger clusters for comparison, please see our response regarding Figure 2, which contains the comparison against IC-2391 and IC-2602, and our reason for not including it in the manuscript.
 >
+> - Regarding M dwarfs, the suggestion to exclude them by cutting at say (BP-RP)0=1.84 (Mamajek's M0V main sequence boundary) would have the following effect.
+In RSG-5, it would omit 13 stars currently in the 42 star sample, including the one non-detection, and the one 12.7 day rotator.
+The net result would be that 28 stars in RSG-5 would have rotation period detections consistent with being younger than the Pleaides, out of 29 expected.
+In CH-2, this cut would omit 10 stars currently in the 19 star sample.
+The net result would be that 8 stars in CH-2 would have rotation period detections consistent with being younger than the Pleiades, out of 9 stars expected.
+We would prefer to not make this cut, chiefly because of the factor of 2x reduction in the CH-2 rotation sample.
+In addition, in our view the rotation periods of the early M-dwarfs are still informative, particularly in the sense that non-detections immediately imply inconsistency with the expected >1% amplitude spot-induced modulations (e.g.
+Figure 10 of Rebull 2020 2020AJ....159..273R).
+We have clarified the wording in the relevant paragraphs of Section 2.3.2 to emphasize the main points, which concern the range of plausible false positive rates from this analysis.
+They seem to be at most one in three for CH-2 candidate members, and one in fifteen for candidate RSG-5 members.
+>
+>
+> - Regarding "nearly no field star outliers", which was the relevant phrase, the wording has been adjusted to "only one field star outlier".  Thank you for this helpful comment.
+>
+> - Regarding 87/13%, please see the earlier discussion regarding the sensitivity of this fraction to our blue cutoff boundary, and the adjustment to (BP-RP)0 > 0.6, which yielded a 93% detection rate.
+Given the small amplitudes of rotation-induced signals in the bluer regime, this seems like the wiser approach.
+Based on the color absolute magnitude diagrams, the false positive rate in the candidate RSG-5 members is quite likely lower than in the candidate δ Lyr or CH-2 members.
 
@@
 ----------
->
+>  FIXME FIXME TODO TODO GET THE MACROTURBULENCE ESTIMATES FOR MAIN SEQUENCE STARS...
 

### 2022-09-06 papers/paper/ms.tex
update response

@@
 $\delta$~Lyr\ cluster is visible at $(l,b)=(68^\circ,15^\circ$)
-and $(v_{l'}, v_b)=(-4.5 ,-4 )\mkms$.  RSG-5 is visible at
-$(l,b)=(83^\circ,6^\circ)$, $(v_{l'}, v_b)=(5.5 ,-3.5 )\mkms$.  Most
+and $(v_{l^*}, v_b)=(-4.5 ,-4 )\mkms$.  RSG-5 is visible at
+$(l,b)=(83^\circ,6^\circ)$, $(v_{l^*}, v_b)=(5.5 ,-3.5 )\mkms$.  Most
 of the other subclusters, including in Cep-Cyg
@@
 The CAMDs show that for RSG-5, all but one of the candidate
-members are on a tight pre-main-sequence locus.  \added{Quantiatively,
+members are on a tight pre-main-sequence locus.  \added{Quantitatively,
 88/89 stars with $\bpmrpo \geq 1.5$ are consistent with being on the
 pre-main-sequence.} This implies a false
-positive rate of a few percent, at most.  By comparison, our control
+positive rate of a few percent, at most.  In comparison, our control
 sample (the $\delta$~Lyr candidates) has a false positive rate of
@@
 objects in the color range displayed, and 23 of them appear to be
-consistent with being on the pre-main-sequence.  This implies a false
+consistent with being on the pre-main-sequence.  This
+\replaced{implies}{would imply} a false
 positive rate in CH-2 of $\approx$15\%.
@@
 suggest that this is indeed the case.
+\added{While in reality stellar populations do not evolve linearly in
+the dimensions of absolute magnitude versus color, in our case the
+Cep-Her loci are nearly indistinguishable from IC\,2602 (e.g.,
+Figure~3 of \citet{bouma_kep1627_2022}).  Systematic errors incurred
+in the age from the non-linear evolution are therefore likely much
+smaller than the $\approx$10\,Myr systematic uncertainty in the
+absolute reference age for IC\,2602 itself
+\citep{david_ages_2015,randich_gaiaeso_2018}.
 
@@
 M-dwarfs with 10-12 day rotation periods at ages of USco
-($\sim$$8$\,Myr), and some \replaced{may}{do} still exist at ages of
+($\sim$$8$\,Myr), and some \deleted{may} still exist at ages of
 LCC ($\sim$$16$\,Myr; \replaced{L.~Rebull
-submitted}{\citealt{2022AJ....164...80R}}).  Regardless, since nearly
-no field star outliers seem to be present on the RSG-5 CAMD, the fact
+submitted}{\citealt{2022AJ....164...80R}}).  Regardless, since
+\replaced{nearly no field star outliers seem}{only one field star
+outlier seems} to be present on the RSG-5 CAMD,
+the fact
 that we do not detect rotation periods for $\approx$\replaced{13}{7}\%
@@
 {e.g.}, pole-on stars having lower amplitude starspot modulation.
+\added{Field star contamination is another possible contributor.}
 
@@
 10 and 12.5 days.  As previously noted, the age interpretation for
-these M-dwarfs is not obvious.  \added{If none are cluster members,
+these M-dwarfs is \replaced{not obvious}{ambiguous}.  \added{If none are cluster members,
 the rotation period detection fraction is 68\%; if all are members, it
@@
 This sets an upper bound on the contamination fraction in our candidate
-CH-2 members at about one in three.  The true value may be as low as 
-one in six, based on the earlier CAMD analysis.
+CH-2 members at about one in three.  Combined with the roughly one in
+six contaminant rate implied by the earlier CAMD analysis, this
+suggests that the sample of candidate CH-2 members is more polluted by
+field stars than the RSG-5 sample.
 }

### 2022-09-06 papers/paper_review/response_1.txt
update response

@@
 ----------
-> - The Kepler field is shown in the top panel of the figure, between roughly l of 70 to 85, and b of 5 to 20.
-This roughly corresponds to a cone in XYZ -- in case it helps intution, (l,b)=(90,0) corresponds to the line of X=0.
+> - The Kepler field is shown in the top panel of the figure, between l of 70 to 85, and b of 5 to 20.
+This approximately corresponds to a cone in XYZ -- in case it helps intution, (l,b)=(90,0) corresponds to the line of X=0.
 The resulting concentration of the planets in vl* we believe follows from the clustering in l, which is a direct result of the fraction of space that Kepler surveyed, combined with the solar velocity projection effect (described in point three below).
@@
 > - We have briefly summarized in the first paragraph of section 2.2 why we have shown the projection of the solar velocity: it is because we are interested in how much of the spread in v_l* it might explain.
-The short answer is "some, but not all of it".
-To give a more complete explanation: the Sun moves at around 16 km/s toward (l,b)=(53deg,23deg), relative to a hypothetical circular orbit around the center of the galaxy.
+The short answer is "some of it, but not all of it".
+To give a more complete explanation: the Sun moves at ~17 km/s toward (l,b)=(53deg,23deg), relative to a hypothetical circular orbit around the center of the galaxy.
 This means that in our observed frame, nearby stars on average appear to move opposite this direction.
-At l=53 and 53+180 degrees, the effect does not influence vl.
-At l=53+90 degrees, the effect on vl (or equivalently vl*) is maximized, with amplitude = v_LSR = 16 km/s.
+At l=53 and 53+180 degrees, the effect does not influence v_l.
+At l=53+90 degrees, the effect on v_l is maximized, with amplitude = v_LSR = 17 km/s.
 The minimum is at l=53+270 degrees.
-Figure 11 of Zari, Brown et al 2018 provides some helpful intuition for how much of the variance in observed vl can be ascribed to this projection effect for stars near the galactic plane.
+Figure 11 of Zari, Brown et al 2018 provides an example for how much of the variance in observed v_l can be ascribed to this projection effect for stars near the galactic plane.
 The lower-right panel of our Figure 1 however shows that while the LSR projection accounts for some of the change in observed v_l* in Cep-Her, the data and the gray band do not match!
@@
 > Regarding more massive stars, yes, there are plenty: see hr_rsg5.png attached to the response.
-The caption of Figure 2 already explicitly states "The range of colors is truncated to emphasize the pre-main-sequence."  This visualization point is necessary because the full dynamic range in magnitude and color.
-But yes, we cannot determine whether these stars earlier than ~K2 are young based on their location in the CAMD.
+The caption of Figure 2 already explicitly states "The range of colors is truncated to emphasize the pre-main-sequence."  The manuscript also has a sentence "Finally, for the plots we set the color range to best visualize the region of maximal age information content: the pre-main-sequence."  But yes, we cannot determine whether these stars more massive than ~K2V are young based on their location in the CAMD.
 We can and do use stellar rotation.
 >
-> Using stellar rotation, we previously quoted in Section 2.3.2 that in RSG-5 41/47 stars have rotation periods consistent with being at least as fast as the Pleiades (87%).
+> Using stellar rotation, we previously wrote in Section 2.3.2 that in RSG-5 41/47 stars have rotation periods consistent with being at least as fast as the Pleiades (87%).
 Of the six inconsistent stars, two are the M-dwarfs previously discussed in the text.
 The new point that emerged from our reanalysis concerns the other four stars without any detectable rotation period: three of them were the bluest three stars in our sample!
-This implies that our blue cutoff was too lenient -- the amplitudes of spot-induced modulation plummet for young stars blueward of the Kraft break (Figure 10 in Rebull+2020, 2020AJ....159..273R).
-We have propagated this through the manuscript: we now require (BP-RP)0 > 0.6, and G < 16.
+This implies that our blue cutoff was too lenient -- the amplitudes of spot-induced modulation plummet for young stars blueward of the Kraft break (eg.
+Figure 10 in Rebull+2020, 2020AJ....159..273R).
+We have propagated this through the manuscript: we now require (BP-RP)0 > 0.6, and G < 16; the previous cutoff was (BP-RP)0 > 0.5.
 The corresponding fraction of RSG-5 stars for which detected rotation periods is 39 detections / 42 expected, i.e.
@@
 >
-> The manuscript has been adjusted for this discussion, including the quantitative statements regarding the pre-main-sequence K and M dwarfs, as well as the rapidly-rotating FGK stars.
+> The manuscript has been adjusted following this discussion, including the quantitative statements regarding the pre-main-sequence K and M dwarfs, as well as the rapidly-rotating FGK stars.
 
@@
 These are preliminary rotation period measurements, from the same Douglas+2021 work referenced in the manuscript, which is not yet published.
-The IC-2602 and IC-2391 rotation period distributions look quite similar to those from RSG-5, though their work a) extends to lower masses, b) appears to show slightly more super-Pleiades rotation periods from F5 to G0 spectral types, and c) appears to show more stars on the fast sequence, which are lacking in the RSG-5 sample.
-Given the unpublished nature of this comparison sample, and the fact that our aim is not to firmly establish the intrinsic rotational distribution of stars at this age, we prefer to leave the reference to Douglas+2021 as it is and to not include these data in the manuscript.
+The IC-2602 and IC-2391 rotation period distributions look quite similar to those from RSG-5, though their work a) extends to lower masses, b) appears to show somewhat more super-Pleiades rotation periods from F5 to G0 spectral types, and c) appears to show more stars on the fast sequence, which are lacking in the RSG-5 sample for unknown reasons.
+Given the unpublished nature of this comparison IC-2602 & IC-2391 sample, and the fact that our aim is to simply check for rotational consistency with the stars being at least as young as the Pleiades, we prefer to leave the reference to Douglas+2021 as it is and to not include these data in the manuscript.
 
@@
 ----------
-> - Regarding younger clusters for comparison, please see our response regarding Figure 2, which contains the comparison against IC-2391 and IC-2602, and our reason for not including it in the manuscript.
+> - Regarding younger clusters for comparison, please see our response regarding Figure 2, which contains the comparison against IC-2391 and IC-2602, and our reasons for not including it in the manuscript.
 >
@@
 The net result would be that 8 stars in CH-2 would have rotation period detections consistent with being younger than the Pleiades, out of 9 stars expected.
-We would prefer to not make this cut, chiefly because of the factor of 2x reduction in the CH-2 rotation sample.
-In addition, in our view the rotation periods of the early M-dwarfs are still informative, particularly in the sense that non-detections immediately imply inconsistency with the expected >1% amplitude spot-induced modulations (e.g.
+We would prefer to not make this cut, chiefly because of the factor of 2x reduction in the CH-2 rotation sample, which would weaken its statistical power.
+In our view, the rotation periods of the early M-dwarfs are still informative, particularly in the sense that non-detections would imply inconsistency with the expected >1% amplitude spot-induced modulations (e.g.
 Figure 10 of Rebull 2020 2020AJ....159..273R).
-We have clarified the wording in the relevant paragraphs of Section 2.3.2 to emphasize the main points, which concern the range of plausible false positive rates from this analysis.
-They seem to be at most one in three for CH-2 candidate members, and one in fifteen for candidate RSG-5 members.
+We have clarified the wording in the relevant paragraphs of Section 2.3.2 to emphasize the main points, which concern the range of plausible false positive rates based on the rotation period analysis.
+The contamination rates seem to be at most one in three for CH-2 candidate members, and one in fifteen for candidate RSG-5 members.
 >
->
-> - Regarding "nearly no field star outliers", which was the relevant phrase, the wording has been adjusted to "only one field star outlier".  Thank you for this helpful comment.
+> - Regarding "nearly no field star outliers", which was the relevant phrase, the wording has been adjusted to "only one field star outlier".
 >
@@
 ----------
->  FIXME FIXME TODO TODO GET THE MACROTURBULENCE ESTIMATES FOR MAIN SEQUENCE STARS...
+> We have added the following footnote to the manuscript:
+>
+> \footnote{The broadening is calculated using the joint rotational and macroturbulent broadening kernel from \citet{hirano_2011}, assuming that the macroturbulent velocity scales with effective temperature similar to the prescription from \citet{2014MNRAS.444.3592D}.
+The latter assumption could be a source of systematic uncertainty in our equatorial velocity measurements, since the macroturbulent velocity could be systematically higher (or lower) on the pre-main-sequence than it is for more slowly rotating field stars.}
+>
+> We have also corrected the number of significant figures in the KOI-7368 vsini, and imposed an uncertainty floor of 1 km/s.
 
@@
 ----------
-Thank you for the helpful report!  Health is more important than science.
+Thank you for the helpful report!  Health is more important than science.  I hope everything is OK.
 

### 2022-09-06 papers/paper/ms.tex
update response

@@
 absolute reference age for IC\,2602 itself
-\citep{david_ages_2015,randich_gaiaeso_2018}.
+\citep{david_ages_2015,randich_gaiaeso_2018}.}
 
@@
 yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.  Using the broadened synthetic
-templates from \texttt{SpecMatch-Synth} \citep{petigura_cksi_2017}, we
+templates\added{\footnote{The broadening is calculated using the joint
+rotational and macroturbulent broadening kernel from \citet{hirano_2011},
+assuming that the macroturbulent velocity scales with effective temperature similar
+to the prescription from \citet{2014MNRAS.444.3592D}.  The latter assumption could be a 
+source of systematic uncertainty in our equatorial velocity measurements, since
+the macroturbulent velocity could be systematically higher (or lower) on
+the pre-main-sequence than it is for more slowly rotating field stars.}}
+ from \texttt{SpecMatch-Synth} \citep{petigura_cksi_2017}, we
 found $v\sin i = 9.3 \pm 1.0\,\mkms$.  The systemic radial velocity at
@@
 spectroscopic parameters included the metallicity $[{\rm Fe/H}]= -0.02
-\pm 0.08$, the equatorial velocity $v\sin i = 20.21 \pm 0.50\,\mkms$,
+\pm 0.08$, the equatorial velocity \replaced{$v\sin i = 20.21 \pm 0.50\,\mkms$}{$v\sin i = 20.2 \pm 1.0\,\mkms$},
 and the systemic velocity ${\rm RV}_{\rm sys} = -10.9 \pm 0.2\,\mkms$.

### 2022-09-06 papers/paper_review/response_1.txt
update response

@@
 ----------
+> - The wording at L456 has been updated as follows: "The primary showed lithium in absorption with ${\rm EW}_{\rm Li} = 65^{+8}_{-6}$\,m\AA, while the secondary had a marginal detection of ${\rm EW}_{\rm Li} = 42^{+12}_{-19}$\,m\AA."
 >
+> - The wording at L497 has similary been adjusted.
+>
+> - Regarding the 1% false postive rate, we have previously addressed that photometrically, 88/89= 98.9% of the candidate RSG-5 members redward of (BP-RP)0=1.5 are consistent with being on the pre-main-sequence.
+Based on rotation periods, the relevant fraction was 39/42 (93%).
+The detection process for stellar rotation periods is more complicated than that for measuring a star's brightness and color, so our initial wording of "$\approx 1\%$ false positive probability" seems accurate.
+We have nonetheless replaced the text with the term "few-percent false positive probability".
+This does not change the conclusion, which is based on both the FPP from the spatio-kinematic selection, as well as the chance of selecting an interloping field star at this color with a sub-Pleiades rotation period.
+>
+> - We have added a sentence quantifying the discrepancy: "Quantitatively, there are 14 reference stars within $\pm$150\,K of Kepler-1643.
+The mean and standard deviation of their lithium EWs is $255 \pm 31$\,m\AA, which implies that Kepler-1643 is $4.0$-$\sigma$ discrepant from expectations."  We have also added sentences indicating that given these ~dozen stars, an anomalous Li EW scatter would indeed be surprising.
+However, in our view it remains the most plausible interpretation.
+The suggested alternative of differences in overall metallicity seems unlikely because the [Fe/H] abundances measured for the Kepler Objects of Interest are all near-solar.
+>
+> - In constructing this response, a typo was correct: the manuscript previously referenced Randich+2018 for the lithium sample in IC-2602.
+The correct reference is Randich+2001.
+We did consider the Randich+2018 study, but the results were qualitatively the same, but with less clearly stated uncertainties on the equivalent width measurements, and a selection function (their Figure 1b) that we did not want to impose.
 

### 2022-09-06 papers/paper/ms.tex
update response

@@
 $[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm 1.0\,\mkms$, and
-${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.  Neither of the KOI-7913
+${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.  \replaced{Neither of the KOI-7913
 components shows strong lithium absorption, but this is expected given
-their $\approx$K6V and $\approx$K8V spectral types.  They do however
-both show H$\alpha$ in emission.  Broadly speaking, both observations
+their $\approx$K6V and $\approx$K8V spectral types}{The primary showed lithium in absorption
+with ${\rm EW}_{\rm Li} =
+65^{+8}_{-6}$\,m\AA, while the secondary had a marginal detection of ${\rm EW}_{\rm Li} =
+42^{+12}_{-19}$\,m\AA. }.  \replaced{They do however
+both show H$\alpha$ in emission.  Broadly speaking, both}{Both components displayed 
+H$\alpha$ in emission.  Given the spectral types of the stars, these} observations
 are consistent with a $\approx$40 Myr age for KOI-7913 (see
@@
     \citet{berger_identifying_2018}.  The ``40-50 Myr'' reference
-    stars ({\it left}) are from IC\,2602 \citep{randich_gaiaeso_2018} and
+    stars ({\it left}) are from IC\,2602 \replaced{\citep{randich_gaiaeso_2018}}{\citep{randich_2001}} and
     Tuc-Hor \citep{kraus_stellar_2014}.  The ``112 Myr'' stars
@@
 above the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA).
+added{
+Quantitatively, there are 14 reference stars within $\pm$150\,K of Kepler-1643.
+The mean and standard deviation of their lithium EWs is $255 \pm 31$\,m\AA,
+which implies that Kepler-1643 is $4.0$-$\sigma$ discrepant from expectations.
+}
 The right panel shows the comparison against the Pleiades, where
@@
 
-One explanation for the lack of Li in Kepler-1643 could be that it is a field
+One explanation for the \replaced{lack of Li in Kepler-1643}{low Li equivalent width
+in Kepler-1643 relative to the comparison stars} could be that it is a field
 interloper; another could be that RSG-5 is much older than 50\,Myr.  We do
@@
 Kepler-1643 also seems highly unlikely to be a field interloper, 
-because we demonstrated a $\approx$$1\%$ false positive probability
+because we demonstrated a \replaced{$\approx$$1\%$}{few-percent} false positive probability
 in our spatio-kinematic selection of RSG-5 members, and there is a 
@@
 Our preferred explanation for Kepler-1643's meager lithium 
-content is that the reference sample of IC\,2602 and Tuc-Hor
+content is that the reference sample\added{s} of IC\,2602 and Tuc-Hor
 stars may not fully explore all possible stellar rotation
 periods and lithium equivalent widths at this age.
-Considering the top panels of Figure~\ref{fig:lithium}, it is remarkable that in
+\added{This would be somewhat surprising since over a dozen
+stars have already been analyzed in the relevant effective temperature range. However, considering}
+the top panels of Figure~\ref{fig:lithium}, it is remarkable that in
 50 million years, stars between $4500$\,K and $5200$\,K go from
@@
 effective temperature range.
-This suggests that
+\added{An alternative explanation could be that the overall metallicity of the cluster 
+is different from Tuc-Hor and IC\,2602, but this seems unlikely given the near-solar
+metallicities we have measured for the Kepler Objects of Interest.}
+\replaced{This suggests}{Broadly, these considerations suggest} that
 RSG-5 and Cep-Her could be
-worthy objects for a closer analysis of the lithium-rotation correlation
+worthy objects for a \replaced{closer analysis of the lithium-rotation correlation}{spectroscopic analysis of lithium}
 near the zero-age main sequence.

### 2022-09-06 papers/paper/ms.tex
update response (H-a, Li)

@@
     The H$\alpha$ comparison is against Tuc-Hor
-    \citep[$\approx$$40$\,Myr;][]{kraus_stellar_2014}
+    \citep[$\approx$$40$\,Myr;][]{kraus_stellar_2014}\added{ and the Pleiades \citep{fang_2018}.}
     \label{fig:lithium}
@@
 periods and lithium equivalent widths at this age.
-\added{This would be somewhat surprising since over a dozen
-stars have already been analyzed in the relevant effective temperature range. However, considering}
-the top panels of Figure~\ref{fig:lithium}, it is remarkable that in
+\added{This would be somewhat surprising since over a dozen stars have
+already been analyzed in the relevant effective temperature range.
+However, considering}
+the top panels of Figure~\ref{fig:lithium}, it is \added{also} remarkable that in
 50 million years, stars between $4500$\,K and $5200$\,K go from
@@
 effective temperature range.
-\added{An alternative explanation could be that the overall metallicity of the cluster 
+\added{An alternative explanation could be that the overall metallicity of Cep-Her
 is different from Tuc-Hor and IC\,2602, but this seems unlikely given the near-solar
 metallicities we have measured for the Kepler Objects of Interest.}
-\replaced{This suggests}{Broadly, these considerations suggest} that
-RSG-5 and Cep-Her could be
-worthy objects for a \replaced{closer analysis of the lithium-rotation correlation}{spectroscopic analysis of lithium}
-near the zero-age main sequence.
+\replaced{This suggests that RSG-5 and Cep-Her could be worthy objects for a
+closer analysis of the lithium-rotation correlation near the zero-age
+main sequence.}{Broadly, these considerations suggest that Cep-Her is
+a worthy object for further spectroscopic analyses of lithium near the
+zero-age main sequence.}
 
@@
 cluster members with spectral types $>$${\rm K4.5V}$ had H$\alpha$ in
-emission.  Ths is consistent with our observations: KOI-7913 shows
+emission.  This is consistent with our observations: KOI-7913 shows
 H$\alpha$ in emission for both components, and in absorption

### 2022-09-06 papers/paper_review/response_1.txt
update response (H-a, Li)
update response

@@
 ----------
+> - We have limited the axis range.
 >
+> - We have updated the markers to be more distinguishable.
+>
+> - Good point regarding the field comparison.  We have replaced it for the Pleaides using the data from Fang+2018, as suggested.
 
@@
 ----------
+> The Data Validation Summaries for the KOIs are available at the following links, for reference.
+> Kepler-1643: https://exoplanetarchive.ipac.caltech.edu/data/KeplerData/008/008653/008653134/dv/kplr008653134-001-20160209194854_dvs.pdf
+> KOI-7368: https://exoplanetarchive.ipac.caltech.edu/data/KeplerData/010/010736/010736489/dv/kplr010736489-001-20160209194854_dvs.pdf
+> KOI-7913: https://exoplanetarchive.ipac.caltech.edu/data/KeplerData/008/008873/008873450/dv/kplr008873450-006-20160209194854_dvs.pdf
 >
+> - Re: 0.4" and "good values", \citet{bryson_2013} showed that for typical field star KOIs without centroid offsets, the mean offset distribution peaks at 0.3$''$ (their Figure 23).
+By comparison, stars with centroid offsets that can be localized to nearby stars have a distribution that peaks at 7$''$ (their Figure 32).
+The stellar variability in Kepler-1643 complicates the centroid-based vetting tests, because the shifts measured by these tests are determined from the in- and out-of-transit flux-weighted centroids.
+For stars with significant spot-induced variability there is no static baseline in either the in- or out-of-transit , and so the centroid location may shift depending on the rotational phase combined with the local scene.
+Nonetheless, it appears that the centroid-level diagnostics for Kepler-1643 are consistent with the transit signal being localized to the target star, though only at moderate confidence.
+We have modified the text to include these points.
+>
+> - Re: what level of centroid shift is concerning -- if the centroid shifts across quarters were consistently >~ 1 arcsecond (and measured at ~0.1 arcsecond precision), and in the same direction across many Kepler quarters, it would be much more concerning.
+However the outlier centroid shifts in Kepler-1643 and KOI-7368 appear to be in effectively random directions, which suggests that the stellar rotation and systematic errors in the NASA SPOC pipeline's whitening are more likely to be the root cause.
+A few sentences have been modified to this effect.
+>
+> - Re: methods for establishing whether other stars can create the centroid shift, and Kepler-1643, yes, the same logic applies.
+We have added a comment to the manuscript to this effect.
+The "established method" for this is the logic laid out by Bryson+2013.
+For instance, given the target star T and a neighbor star N separated by distance D, one can calculate the expected centroid shifts for the various scenarios where the target T hosts an eclipse signal of depth δ, and also when the neighbor N hosts the signal.
+We do not present these calculations in the text, because there is no evidence that the offset itself is statistically signficant.
+Nonetheless, the absence of companions for Kepler-1643 and KOI-7368 does help bolster the interpretation.
 

### 2022-09-06 papers/paper/ms.tex
update response

@@
 centroid shift tests (see the \texttt{q1\_q17\_dr25\_koi} data
-release): the angular distance between the target star's catalog
+release): the angular distance between the target star's \added{KIC }catalog
 position and the position of the transiting source was measured as
@@
 outlying quarters (2 and 6) drive the offset --- the centroid locations
-from the other Kepler quarters are consistent at $\lesssim 0\farcs4$.
-This is an instructive exercise in how stellar
-variability complicates centroid-based vetting tests.  The 
-shifts measured by these tests are determined from the in- and
-out-of-transit flux-weighted centroids.  For stars with significant
-spot-induced variability there is no static baseline in either the in-
-or out-of-transit phases.
+from the other Kepler quarters are consistent at $\lesssim 0\farcs4$\added{ (3-$\sigma$)}.
+\added{\citet{bryson_2013} showed that for typical field star KOIs without
+centroid offsets, the mean offset distribution peaks at 0.3$''$ (their
+Figure 23).  By comparison, stars with centroid offsets that can be
+localized to nearby stars have a distribution that peaks at 7$''$ (their
+Figure 32). } \replaced{This is an instructive exercise in how }{The }stellar
+variability \added{in Kepler-1643} complicates \added{the} centroid-based
+vetting tests\replaced{. The}{, because the} shifts measured by these tests are
+determined from the in- and out-of-transit flux-weighted centroids.  For stars
+with significant spot-induced variability there is no static baseline in either
+the in- or out-of-transit phases\added{, and so the centroid location may shift
+depending on the combination of the rotation phase combined and the local scene.  Based on
+these considerations, the centroid-level diagnostics for Kepler-1643 appear to
+be consistent with the transit signal being localized to the target star}.
 
@@
 Kepler-1643.  For KOI-7368, the reported offset is smaller, and less
-significant ($0\farcs2$; 3.0-$\sigma$).  Again, the data
-validation reports show that the shift is caused by a few outlying
-quarters (4, 5, 8, and 12).  Since the remaining data show
-consistent scatter in their centroid locations, these
-outlying quarters are likely also caused by the stellar variability.  
-Our NIRC2 imaging independently shows that there are no neighboring
-sources that could cause an offset of the observed amplitude.
+\added{formally }significant ($0\farcs2$; 3.0-$\sigma$).  Again, the
+data validation reports show that the shift is caused by a few
+outlying quarters (4, 5, 8, and 12).  Since the remaining data show
+consistent scatter in their centroid locations, these outlying
+quarters are likely also caused by the stellar variability\added{,
+since their directions across different quarters are inconsistent}.
+Our NIRC2 imaging independently shows that there are no \added{known
+}neighboring sources that could cause an offset of the observed
+amplitude\added{, as is also the case for Kepler-1643}.
 

### 2022-09-07 papers/paper_review/response_1.txt
update response

@@
 ----------
->
+> Sentence has been reworded: \replaced{To date these analyses have mostly clustered on stellar
+positions and 2D velocities}{So far, these analyses have mostly leveraged 3D stellar positions and 2D on-sky tangential velocities.}.
 
@@
 ----------
->
+> Sentence has been reworded as suggested.
 
@@
 ----------
->
+> Done.
 
@@
 ----------
->
+> The end of Section 2.1 already cites Kerr+21's ages for some of the subgroups.
+The relevant citations other than Bouma+2022 (Kepler-1627 / delta Lyr cluster) would be Kounkel & Covey 2019 and Roser+2016.
+The Roser+2016 reference is already made in the preceding sentence.
+The KC2019 reference we prefer to avoid in reference to the age based on the manual stitching step they used which yielded a very diferent "delta Lyra cluster" (Stephenson-1) than that originally described by Stephenson (1959).
+Given the existing citations to the relevant studies, these age citations are not needed.
 
@@
 ----------
->
+> This Zari+2018 reference has been updated to refer to Figure 1, to help the reader know where to look.
 
@@
 ----------
->
+> The Kounkel & Covey 2019 study indeed included many false positives, both in the sense of "clusters" that do not correspond to stars that were ever physically associated, and also in the sense of clusters that do have some "true" core population, but many erroneous field star interlopers.
+Zucker et al's argument however was not particularly helpful for our purposes of identifying which are which.
+One reason is that for the youngest (<100 Myr) cluster candidates from KC2019, most of the HR diagrams independently show the expected pre-main-sequence M-dwarfs as well as massive stars (http://mkounkel.com/mw3d/hr.html).
+It is hard to fake such stellar populations.
+Nonetheless, Zucker+2022 did nicely demonstrate the extreme nature of some of the groups in e.g., XY space (https://faun.rc.fas.harvard.edu/czucker/Paper_Figures/String_Gallery_Interactive.html), and for the extended groups older than a few hundred million years, there is almost no independent evidence for their coevality.
+Despite this point, for the youngest groups, the KC2019 study to us seems to be more likely an example of why the manual stitching must be treated cautiously, and also for why there are better dimensions to cluster in than those selected in the original study.
+A reference has been added to Zucker et al's rebuttal.
 
@@
 ----------
->
+> Yes, clarification made.
 
@@
 ----------
->
+> The wording in this paragraph has been clarified.
 
@@
 ----------
->
+> RONAN TODO
 
@@
 ----------
->
+> Done; these thresholds were selected visually based on the apparent purity with which they yielded pre-main-sequence stars on a color--absolute magnitude diagram.
 
@@
 ----------
->
+> Thank you for the typo catch: v_l' was equivalent to v_l*, and has been relabelled as such.  v_l* is also now defined (inre: Line 162 comment).
 
@@
 ----------
->
+> For stars at least as young as the Pleiades -- words added.
 
@@
 ----------
->
+> Yes, there is overlap.  But the set of Cep-Her stars that Kepler observed is a distinct set from the described sets of candidate stars in RSG-5 and CH-2.  The wording has been adjusted.
 
@@
 ----------
->
+> We now include a reference to appendix B of Gaia Collaboration+2018.
 
@@
 ----------
->
+> Done
 
@@
 ----------
->
+> We now call it a "comparison sample".
+It is another example, equivalent to RSG-5 and CH-2, except that it was studied in our earlier Kepler-1627 paper, where the explicit comparison against IC-2602 is made.
 
@@
 ----------
->
+> Wording simplified, thank you.
 
@@
 ----------
->
+> We do not entirely understand what, if any, change is requested:  the purpose of these sentences is to give a broad-brush overview of the stellar properties.
+The rough implicit understanding that the Sun would be "near" G2V at ZAMS (while in reality perhaps being a little less bright, and a little lower surface gravity) provides the necessary context that the stars being discussed spanning G8V to K6V are less massive than the Sun, and given ~40 Myr ages, therefore on the pre-main-sequence.
 
@@
 ----------
->
+> A table note has been added.
 
@@
 ----------
->
+> Yes -- please see the response to Line 449 below.
 
@@
 ----------
->
+> The interpolation is performed in the dimensions of absolute reddening-corrected $G$-band magnitude versus (BP-RP)0 color.  This clarification has been added.
 
@@
 ----------
->
+> Following their methods; the manuscript has been updated accordingly.
 
@@
 Note that a chi^2 of 5 corresponds to a reduced chi^2 of 1 in this model.
-While the KOI-7913B spectrum is still the worst fit(!), our prefernce is for calibrating our reported effective temperatures to the HIRES spectral library, rather than to the isochronal method.
+While the KOI-7913B spectrum is still the worst fit(!), our preference is for calibrating our reported effective temperatures to the HIRES spectral library, rather than to the isochronal method.
 Any suggestions on possible methodological improvements would be appreciated.

### 2022-09-07 papers/paper/ms.tex
update response

@@
 %\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
-%\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
-\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
+\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
+%\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
 \usepackage{amsmath,amstext,amssymb}
@@
 greatly expanded our knowledge of stellar groups
-\citep[{e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.  To
-date these analyses have mostly clustered on stellar positions and 2D
-velocities.  One important result has been the discovery of diffuse
+\citep[{e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.
+\replaced{To date these analyses have mostly clustered on stellar
+positions and 2D velocities}{So far, these analyses have mostly
+leveraged 3D stellar positions and 2D on-sky tangential velocities.}.
+One important result has been the discovery of diffuse
 streams and tidal tails comparable in stellar mass to the previously
@@
 level.  Internal dynamics and projection effects can also drive them
-to be much larger: in the Hyades, stars in the tidal tails are
+to be much \replaced{larger}{more kinematically diffuse}: in the Hyades, stars in the tidal tails are
 expected to span up to $\pm 40\mkms$ in velocity relative to the
@@
 \citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
-\citep{hawkins_2020}.  While there are implications for our
+\citep{hawkins_2020,2020A&A...635L..13A}.  While there are implications for our
 understanding of star formation and cluster evolution
@@
 apparent in the Gaia-based photometric analysis of pre-main-sequence
-stars by \citet[][see their Figures~11 and~13]{Zari2018}.  Further
-kinematic connections and complexity were highlighted by
-\citet{KounkelCovey2019}, who included these previously known groups
-in the larger structures dubbed ``Theia~73'' and
-``Theia~96''\footnote{See their visualization online at
-\url{http://mkounkel.com/mw3d/mw2d.html} (accessed 15 March 2022)}.
+stars by \citet[][compare their Figures~11 and~13 to our
+Figure~\ref{fig:XYZvtang}]{Zari2018}.  Further kinematic connections
+and complexity were highlighted by \citet{KounkelCovey2019}, who
+included these previously known groups in the larger structures dubbed
+``Theia~73'' and ``Theia~96''\footnote{See their visualization online
+at \url{http://mkounkel.com/mw3d/mw2d.html} (accessed 15 March
+2022)\added{. Important caveats, particularly for extended groups $\gtrsim 100$\,Myr
+old, were presented by \citet{2022arXiv220514160Z}.}}.
 The connection made by \citet{KounkelCovey2019} between the previously
@@
 one third of a kiloparsec, identified three of the nearest
-sub-populations, dubbed ``Cepheus-Cygnus'', ``Lyra'', and
+sub-populations\added{ of Cep-Her}, dubbed ``Cepheus-Cygnus'', ``Lyra'', and
 ``Cerberus''.  \citet{Kerr2021} reported ages for each of these
@@
 the presence of contiguous stellar populations connecting each of
-these groups in space-velocity coordinates.  In other words, the stars
+these groups in \replaced{space-velocity}{both space and velocity} coordinates.  In other words, the stars
 appear to be comoving, though with a continuous gradient in velocity
@@
 galactic longitude after accounting for the local tangent plane
-correction.  Some, but not all, of the gradient in the $v_{l^*}$-$l$
+correction.  Some, but not all, of the gradient in the $v_{l^*}$ {\it vs.}~$l$
 plane can be understood through a projection effect stemming from the
@@
 12{,}436 objects with $D>0.02$ as gray points, and 
-4{,}763 objects with $D>0.10$ as black points.  The 
-$\delta$~Lyr\ cluster is visible at $(l,b)=(68^\circ,15^\circ$)
-and $(v_{l^*}, v_b)=(-4.5 ,-4 )\mkms$.  RSG-5 is visible at
-$(l,b)=(83^\circ,6^\circ)$, $(v_{l^*}, v_b)=(5.5 ,-3.5 )\mkms$.  Most
-of the other subclusters, including in Cep-Cyg
-($l,b=90^\circ,7^\circ$) and Cerberus ($l,b=48^\circ,18^\circ$) are
-too small or dispersed to have previously been analyzed in great
-detail.
+4{,}763 objects with $D>0.10$ as black points.  \added{These thresholds were
+selected visually based on the apparent purity with which they yielded
+pre-main-sequence stars on a color--absolute magnitude diagram.} The
+$\delta$~Lyr\ cluster is visible at $(l,b)=(68^\circ,15^\circ$) and $(v_{l^*},
+v_b)=(-4.5 ,-4 )\mkms$.  RSG-5 is visible at $(l,b)=(83^\circ,6^\circ)$,
+$(v_{l^*}, v_b)=(5.5 ,-3.5 )\mkms$.  Most of the other subclusters, including
+in Cep-Cyg ($l,b=90^\circ,7^\circ$) and Cerberus ($l,b=48^\circ,18^\circ$) are
+too small or dispersed to have previously been analyzed in great detail.
 
@@
 the presence of rotational modulation at the expected period and
-amplitude \citep[{e.g.},][]{rebull_rotation_2020}. 
+amplitude \added{for stars at least as young as the Pleiades} \citep[{e.g.},][]{rebull_rotation_2020}. 
 Planetary status was assessed through the Robovetter flags, and by
@@
 \nrsgfive\ candidate RSG-5 members, and \nchtwo\ candidate CH-2
-members.  These stars are listed in Appendix~\ref{app:members}, as are
-the Cep-Her candidates that were observed by Kepler. 
+members.  These stars are listed in Appendix~\ref{app:members}, \replaced{as are
+the Cep-Her candidates that were observed by Kepler}{as is the set of Cep-Her
+candidates that was observed by Kepler}.
 %An important
@@
 diagrams, we imposed the data filtering criteria from
-\citet{GaiaCollaboration2018}, which include binaries
+\citet[][Appendix~B]{GaiaCollaboration2018}, which include binaries
 while omitting artifacts from for instance low
@@
 yielded $E(B-V)=0.032\pm0.006$.  Finally, for the plots we set the
-color range to best visualize the region of maximal age information
+color \replaced{range}{axis} to best visualize the region of maximal age information
 content: the pre-main-sequence.
@@
 pre-main-sequence.} This implies a false
-positive rate of a few percent, at most.  In comparison, our control
+positive rate of a few percent, at most.  In comparison, our \replaced{control}{comparison}
 sample (the $\delta$~Lyr candidates) has a false positive rate of
 $\approx$12\%, based on the number of stars that photometrically
-overlap with the field population, while being outliers from the bulk
-cluster population.  For CH-2, our membership selection gives 27
+overlap with the field population\deleted{, while being outliers from the bulk
+cluster population}.  For CH-2, our membership selection gives 27
 objects in the color range displayed, and 23 of them appear to be
@@
 Figure~\ref{fig:age} also shows that most RSG-5 and CH-2 members
-overlap with the $\delta$~Lyr cluster, and that the groups are
+overlap with the $\delta$~Lyr cluster\added{ on the CAMD}, and that the groups are
 therefore roughly the same age.  To quantify this, we use the method
@@
 and density are found by interpolating against the MIST isochrones
+\added{in absolute reddening-corrected $G$-band magnitude as a function of $\bpmrpo$\ color}
 \citep{choi_mesa_2016}.  The statistical uncertainties from this
@@
 
-To verify these parameters and to analyze youth
+To verify these parameters\added{, determine the stellar effective temperatures,} and to analyze youth
 proxies such as the Li 6708\,\AA\ doublet and H$\alpha$, we acquired
-spectra.  We also acquired high resolution imaging for each system, to
+\added{high resolution optical } spectra.  We also acquired high resolution imaging for each system, to
 constrain the existence of visual companions, including possible bound
@@
 the two epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm 1.2\,\mkms$
-respectively \citep{chubak_2012}.  To infer the equivalent width of
+respectively\replaced{ \citep{chubak_2012}}{, and was calculated
+following the methods of \citet{chubak_2012}}.  To infer the equivalent width of
 the \ion{Li}{1} 6708\,\AA\ doublet, we followed the procedure
-described by \citet{bouma_2021_ngc2516}.  This yielded a strong
-detection: ${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with values
-consistent at $<$$1$-$\sigma$ between the two epochs.   The quoted
-value does not correct for the \ion{Fe}{1} blend at 6707.44\,\AA.  Given the purported age and
-effective temperature of the star, the lithium equivalent width is
-somewhat low.  We discuss this in greater depth in
-Section~\ref{subsec:specyouth}.
+described by \citet{bouma_2021_ngc2516}. \added{In brief, this
+involved a Monte Carlo procedure in which the continuum normalization
+was allowed to vary through a bootstrap approach based on the local
+scatter, and the line width itself was calculated by integrating a
+single Gaussian over a local window.}  For Kepler-1643, this yielded a
+strong detection: ${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with
+values consistent at $<$$1$-$\sigma$ between the two epochs.   The
+quoted value does not correct for the \ion{Fe}{1} blend at
+6707.44\,\AA.  Given the purported age and effective temperature of
+the star, the lithium equivalent width is somewhat low.  We discuss
+this in greater depth in Section~\ref{subsec:specyouth}.
 
@@
 	\end{center}
-	\vspace{-0.6cm}
+	\vspace{-0.5cm}
 	\caption{
@@
 		\subfloat{
-			\includegraphics[width=0.48\textwidth]{f6a.pdf}
+			\includegraphics[width=0.47\textwidth]{f6a.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.48\textwidth]{f6b.pdf}
+			\includegraphics[width=0.47\textwidth]{f6b.pdf}
 		}
@@
 		\subfloat{
-			\includegraphics[width=0.48\textwidth]{f6c.pdf}
+			\includegraphics[width=0.47\textwidth]{f6c.pdf}
 		}
@@
 \subsection{Planet Validation}
+\label{subsec:validation}
 
@@
 the in- or out-of-transit phases\added{, and so the centroid location may shift
-depending on the combination of the rotation phase combined and the local scene.  Based on
+depending on the rotation phase combined with the local scene.  Based on
 these considerations, the centroid-level diagnostics for Kepler-1643 appear to
@@
 Kepler-1643 and KOI-7368 
-both show weak centroid offsets between the in
-and out-of-transit phases.  For both cases, the observed shifts
+both show weak centroid offsets \replaced{between the in
+and out-of-transit phases}{in specific quarters}.  For both \replaced{cases}{systems}, the observed shifts
 are consistent with being caused by starspot-induced variability in

### 2022-09-07 papers/paper_review/response_1.txt
correct typos

@@
 ----------
-> We now call it a "comparison sample".
-It is another example, equivalent to RSG-5 and CH-2, except that it was studied in our earlier Kepler-1627 paper, where the explicit comparison against IC-2602 is made.
+> We now call it a "reference sample".
+It is indeed another example, equivalent to RSG-5 and CH-2.
+The only difference is that it was studied in our earlier Kepler-1627 paper, where the explicit comparison against IC-2602 was made (Figure 3 of that work).
 
@@
 ----------
-> The interpolation is performed in the dimensions of absolute reddening-corrected $G$-band magnitude versus (BP-RP)0 color.  This clarification has been added.
+> The interpolation is performed in the dimensions of reddening-corrected absolute $G$-band magnitude versus (BP-RP)0 color.  This clarification has been added.
 

### 2022-09-07 papers/paper/ms.tex
correct typos

@@
 positions and 2D velocities}{So far, these analyses have mostly
-leveraged 3D stellar positions and 2D on-sky tangential velocities.}.
+leveraged 3D stellar positions and 2D on-sky tangential velocities}.
 One important result has been the discovery of diffuse
@@
 galactic longitude after accounting for the local tangent plane
-correction.  Some, but not all, of the gradient in the $v_{l^*}$ {\it vs.}~$l$
+correction.  Some, but not all, of the gradient in the $v_{l^*}$ vs.~$l$
 plane can be understood through a projection effect stemming from the
 Sun's motion with respect to the local standard of rest (see also
-Figure 11 by \citet{Zari2018}).  }
+Figure 11 by \citealt{Zari2018}).  }
 In this work, our primary interest in \replaced{the region}{this
@@
 pre-main-sequence.} This implies a false
-positive rate of a few percent, at most.  In comparison, our \replaced{control}{comparison}
+positive rate of a few percent, at most.  In comparison, our
+\replaced{control}{reference}
 sample (the $\delta$~Lyr candidates) has a false positive rate of
@@
 Cep-Her loci are nearly indistinguishable from IC\,2602 (e.g.,
-Figure~3 of \citet{bouma_kep1627_2022}).  Systematic errors incurred
+Figure~3 of \citealt{bouma_kep1627_2022}).  Systematic errors incurred
 in the age from the non-linear evolution are therefore likely much
@@
 M-dwarfs with 10-12 day rotation periods at ages of USco
-($\sim$$8$\,Myr), and some \deleted{may} still exist at ages of
+($\sim$$8$\,Myr), and some \deleted{may} still exist at the age of
 LCC ($\sim$$16$\,Myr; \replaced{L.~Rebull
@@
 and density are found by interpolating against the MIST isochrones
-\added{in absolute reddening-corrected $G$-band magnitude as a function of $\bpmrpo$\ color}
+\added{in reddening-corrected absolute $G$-band magnitude as a function of $\bpmrpo$\ color}
 \citep{choi_mesa_2016}.  The statistical uncertainties from this
@@
 described by \citet{bouma_2021_ngc2516}. \added{In brief, this
-involved a Monte Carlo procedure in which the continuum normalization
-was allowed to vary through a bootstrap approach based on the local
-scatter, and the line width itself was calculated by integrating a
-single Gaussian over a local window.}  For Kepler-1643, this yielded a
+involved calculating the line width by numerically integrating a
+single best-fit Gaussian over a local window, and estimating the
+uncertainties through a Monte Carlo procedure in which the continuum
+normalization was allowed to vary through a bootstrap approach based
+on the local scatter in the spectra.}  For Kepler-1643, this yielded a
 strong detection: ${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with
@@
 separation is $959 \pm 2$ au.  The Gaia EDR3 proper motions are also
-very similar.  Since two stars were resolved in the Kepler Input
+very similar.  Since the two stars were resolved in the Kepler Input
 Catalog and are roughly one Kepler pixel apart, an accurate crowding
@@
 Kepler-1643, in RSG-5, is conspicuously below the 40-50$\,$Myr
-sequence in the left-panel of Figure~\ref{fig:lithium}, though
+sequence in the \replaced{left-panel}{top panel} of Figure~\ref{fig:lithium}, though
 above the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA).
-added{
+\added{
 Quantitatively, there are 14 reference stars within $\pm$150\,K of Kepler-1643.
@@
 }
-The right panel shows the comparison against the Pleiades, where
+The \replaced{right}{middle} panel shows \replaced{the}{a} comparison against the Pleiades, where
 Kepler-1643 is more consistent with the observed dispersion in
@@
 content is that the reference sample\added{s} of IC\,2602 and Tuc-Hor
-stars may not fully explore all possible stellar rotation
-periods and lithium equivalent widths at this age.
+stars may not fully explore all possible \deleted{stellar rotation
+periods and }lithium equivalent widths at this age.
 \added{This would be somewhat surprising since over a dozen stars have
@@
 having a tight lithium sequence to one with a dispersion
-$\approx10\times$ greater.  The existence of the Li dispersion in
+$\approx$$10\times$ greater.  The existence of the Li dispersion in
 Pleiades-age K-dwarfs has been known for decades; it has also been
@@
 while photons on the wings are too far from the
-rest-wavelength to excite electrons and be re-absorbed in the upper
+rest wavelength to excite electrons and be re-absorbed in the upper
 layers.  The exact details of when a star's atmosphere reaches the
@@
 data validation reports show that the shift is caused by a few
-outlying quarters (4, 5, 8, and 12).  Since the remaining data show
+outlying quarters (4, 5, 8, and 12).  Since the remaining
+\replaced{data}{quarters} show
 consistent scatter in their centroid locations, these outlying
 quarters are likely also caused by the stellar variability\added{,
-since their directions across different quarters are inconsistent}.
+because their directions are inconsistent across different quarters}.
 Our NIRC2 imaging independently shows that there are no \added{known
@@
 core composition, and a typical H/He mass fraction of $\approx$4\%,
-there should be a local maximum in planet occurrence rates at 2 to 3\,$R_\oplus$,
-at times between 10 and 100\,Myr.  The models advanced by
-\citet{gupta_signatures_2020} and \citet{lee_primordial_2021} agree;
+there should be a \added{single} local maximum in planet occurrence rates at 2 to 3\,$R_\oplus$,
+at times between 10 and 100\,Myr. \added{In other words,
+\citet{rogers_unveiling_2021} predict the existence of a ``radius
+mountain'' at these early times, rather than a ``radius valley''.}  The models advanced by
+\citet{gupta_signatures_2020} and \citet{lee_primordial_2021}
+agree\added{ that this local maximum should exist};
 their differences lie in the mechanism for producing the radius
@@
 %
-% \newline
+%\newline
 %
@@
 14 confirmed and candidate planets,
-6 failed first filter, and 7 independently failed the second.
+6 failed the first filter, and 7 independently failed the second.
 One object was ambiguous: Kepler-1933.

### 2022-09-07 papers/paper_review/response_1.txt
wordsmith response

@@
 
-Below, we have reproduced the relevant portions of the reviewer's report, alongside our responses. The changes are bolded in the attached ms_boldchanges.pdf file, using the trackchanges AASTeX macros.
+Below, we have reproduced the relevant portions of the reviewer's report, alongside our responses.
+The changes are bolded in the attached ms_boldchanges.pdf file, using the trackchanges AASTeX macros.
+The revised text is in the ms.pdf file.
 
@@
 > - The Kepler field is shown in the top panel of the figure, between l of 70 to 85, and b of 5 to 20.
-This approximately corresponds to a cone in XYZ -- in case it helps intution, (l,b)=(90,0) corresponds to the line of X=0.
-The resulting concentration of the planets in vl* we believe follows from the clustering in l, which is a direct result of the fraction of space that Kepler surveyed, combined with the solar velocity projection effect (described in point three below).
+This approximately corresponds to a cone in XYZ -- in case it helps intuition, (l,b)=(90,0) corresponds to the line of X=Z=0.
+The resulting concentration of the planets in v_l* follows from the narrow range of galactic longitudes that Kepler looked at, and the narrow spread in v_l* at a given galactic longitude.
 >
-> - We have specified in the caption which are the larger and smaller filled circles.
+> - We now specify in the caption which are the larger and smaller filled circles.
 >
@@
 This means that in our observed frame, nearby stars on average appear to move opposite this direction.
-At l=53 and 53+180 degrees, the effect does not influence v_l.
-At l=53+90 degrees, the effect on v_l is maximized, with amplitude = v_LSR = 17 km/s.
+At l=53 and 53+180 degrees, the effect does not influence v_l*.
+At l=53+90 degrees, the effect on v_l* is maximized, with amplitude = v_LSR = 17 km/s.
 The minimum is at l=53+270 degrees.
-Figure 11 of Zari, Brown et al 2018 provides an example for how much of the variance in observed v_l can be ascribed to this projection effect for stars near the galactic plane.
+Figure 11 of Zari, Brown et al 2018 provides an example for how much of the variance in observed v_l* can be ascribed to this projection effect for stars near the galactic plane.
 The lower-right panel of our Figure 1 however shows that while the LSR projection accounts for some of the change in observed v_l* in Cep-Her, the data and the gray band do not match!
@@
 >
-> In the color versus absolute magnitude diagram (CAMD), let us use (BP-RP)0 = 1.5 as the dividing line, redward of which stars can be easily seen to be on the pre-main-sequence (or at least elevated relative to typical field stars of the same color).
+> In the color versus absolute magnitude diagram (CAMD), let us use (BP-RP)0 = 1.5 as the dividing line, red-ward of which stars can be easily seen to be on the pre-main-sequence (or at least elevated relative to typical field stars of the same color).
 For RSG-5, there are 89 stars with (BP-RP)0 > 1.5.
 One of them falls on the "field star" locus; the others fall on the pre-main-sequence locus.
-This implies that, photometrically, 88/89= 98.9% of the candidate RSG-5 members redward of (BP-RP)0=1.5 are photometrically consistent with being on the pre-main-sequence.
+This implies that, photometrically, 88/89= 98.9% of the candidate RSG-5 members red-ward of (BP-RP)0=1.5 are photometrically consistent with being on the pre-main-sequence.
 >
 > Regarding more massive stars, yes, there are plenty: see hr_rsg5.png attached to the response.
-The caption of Figure 2 already explicitly states "The range of colors is truncated to emphasize the pre-main-sequence."  The manuscript also has a sentence "Finally, for the plots we set the color range to best visualize the region of maximal age information content: the pre-main-sequence."  But yes, we cannot determine whether these stars more massive than ~K2V are young based on their location in the CAMD.
+The caption of Figure 2 already explicitly states "The range of colors is truncated to emphasize the pre-main-sequence."  The manuscript also has a sentence "Finally, for the plots we set the color [axis] to best visualize the region of maximal age information content: the pre-main-sequence."  But yes, we cannot determine whether these stars more massive than ~K2V are young based on their location in the CAMD.
 We can and do use stellar rotation.
@@
 Of the six inconsistent stars, two are the M-dwarfs previously discussed in the text.
-The new point that emerged from our reanalysis concerns the other four stars without any detectable rotation period: three of them were the bluest three stars in our sample!
-This implies that our blue cutoff was too lenient -- the amplitudes of spot-induced modulation plummet for young stars blueward of the Kraft break (eg.
+The new point that emerged from our reanalysis concerns the other four stars without any detectable rotation period: three of them were the bluest three stars in the sample!
+This implies that our blue cutoff was too lenient -- the amplitudes of spot-induced modulation plummet for young stars blue-ward of the Kraft break (eg.
 Figure 10 in Rebull+2020, 2020AJ....159..273R).
@@
 > We have added the following text: 
-> \added{While in reality stellar populations do not evolve linearly in the dimensions of absolute magnitude versus color, in our case the Cep-Her loci are nearly indistinguishable from IC\,2602 (e.g., Figure~3 of \citet{bouma_kep1627_2022}).
+> \added{While in reality stellar populations do not evolve linearly in the dimensions of absolute magnitude versus color, in our case the Cep-Her loci are nearly indistinguishable from IC\,2602 (e.g., Figure~3 of \citealtt{bouma_kep1627_2022}).
 Systematic errors incurred in the age from the non-linear evolution are therefore likely much smaller than the $\approx$10\,Myr systematic uncertainty in the absolute reference age for IC\,2602 itself \citep{david_ages_2015,randich_gaiaeso_2018}.
@@
 The net result would be that 8 stars in CH-2 would have rotation period detections consistent with being younger than the Pleiades, out of 9 stars expected.
-We would prefer to not make this cut, chiefly because of the factor of 2x reduction in the CH-2 rotation sample, which would weaken its statistical power.
-In our view, the rotation periods of the early M-dwarfs are still informative, particularly in the sense that non-detections would imply inconsistency with the expected >1% amplitude spot-induced modulations (e.g.
-Figure 10 of Rebull 2020 2020AJ....159..273R).
+We would prefer to not make this cut, chiefly because of the factor of 2x reduction in the CH-2 rotation sample, which would weaken its statistical power more than necessary.
+In our view, the rotation periods of the early M-dwarfs are still informative, particularly in the sense that non-detections would imply inconsistency with the expected >1% amplitude spot-induced modulations (see again Figure 10 of Rebull 2020 2020AJ....159..273R).
 We have clarified the wording in the relevant paragraphs of Section 2.3.2 to emphasize the main points, which concern the range of plausible false positive rates based on the rotation period analysis.
@@
 >
-> We have also corrected the number of significant figures in the KOI-7368 vsini, and imposed an uncertainty floor of 1 km/s.
+> Relatedly, we have corrected the number of significant figures in the KOI-7368 vsini, and imposed an uncertainty floor of 1 km/s.
 
@@
 >
-> - The wording at L497 has similary been adjusted.
+> - The wording at L497 has similarly been adjusted.
 >
-> - Regarding the 1% false postive rate, we have previously addressed that photometrically, 88/89= 98.9% of the candidate RSG-5 members redward of (BP-RP)0=1.5 are consistent with being on the pre-main-sequence.
+> - Regarding the 1% false positive rate, we have previously addressed that photometrically, 88/89= 98.9% of the candidate RSG-5 members red-ward of (BP-RP)0=1.5 are consistent with being on the pre-main-sequence.
 Based on rotation periods, the relevant fraction was 39/42 (93%).
 The detection process for stellar rotation periods is more complicated than that for measuring a star's brightness and color, so our initial wording of "$\approx 1\%$ false positive probability" seems accurate.
-We have nonetheless replaced the text with the term "few-percent false positive probability".
-This does not change the conclusion, which is based on both the FPP from the spatio-kinematic selection, as well as the chance of selecting an interloping field star at this color with a sub-Pleiades rotation period.
+We have nonetheless updated the text with the term "few-percent false positive probability".
+This does not change the conclusion, which is based on both the FPP from the spatio-kinematic selection, as well as the chance of randomly selecting an interloping field star at this color with a sub-Pleiades rotation period.
 >
@@
 >
-> - In constructing this response, a typo was correct: the manuscript previously referenced Randich+2018 for the lithium sample in IC-2602.
+> - In constructing this response, a typo was noticed and corrected: the manuscript previously referenced Randich+2018 for the lithium sample in IC-2602.
 The correct reference is Randich+2001.
-We did consider the Randich+2018 study, but the results were qualitatively the same, but with less clearly stated uncertainties on the equivalent width measurements, and a selection function (their Figure 1b) that we did not want to impose.
+We did consider the Randich+2018 study, but the results were more challenging to interpret because that study had less clearly stated uncertainties on the equivalent width measurements, and a selection function (their Figure 1b) that we did not want to impose.
 
@@
 >
-> - Good point regarding the field comparison.  We have replaced it for the Pleaides using the data from Fang+2018, as suggested.
+> - Good point regarding the field comparison.  We have replaced it for the Pleiades using the data from Fang+2018, as suggested.
 
@@
 ----------
-> The Data Validation Summaries for the KOIs are available at the following links, for reference.
+> For reference, the relevant Data Validation Summaries for the KOIs are available at the following links.
 > Kepler-1643: https://exoplanetarchive.ipac.caltech.edu/data/KeplerData/008/008653/008653134/dv/kplr008653134-001-20160209194854_dvs.pdf
@@
 For stars with significant spot-induced variability there is no static baseline in either the in- or out-of-transit , and so the centroid location may shift depending on the rotational phase combined with the local scene.
-Nonetheless, it appears that the centroid-level diagnostics for Kepler-1643 are consistent with the transit signal being localized to the target star, though only at moderate confidence.
+Nonetheless, it appears that the centroid-level diagnostics for Kepler-1643 are consistent with the transit signal being localized to the target star.
 We have modified the text to include these points.
 >
-> - Re: what level of centroid shift is concerning -- if the centroid shifts across quarters were consistently >~ 1 arcsecond (and measured at ~0.1 arcsecond precision), and in the same direction across many Kepler quarters, it would be much more concerning.
-However the outlier centroid shifts in Kepler-1643 and KOI-7368 appear to be in effectively random directions, which suggests that the stellar rotation and systematic errors in the NASA SPOC pipeline's whitening are more likely to be the root cause.
-A few sentences have been modified to this effect.
+> - Re: what level of centroid shift is concerning -- if the centroid shifts across quarters were consistently >~ 1 arcsecond (and measured at ~0.1 arcsecond precision), and in the same direction across many Kepler quarters, it would be concerning.
+The measured centroid shifts in Kepler-1643 and KOI-7368 are not that large, and appear to be in effectively random directions, which suggests that the stellar rotation and systematic errors in the NASA SPOC pipeline's whitening are more likely to be the root cause.
+A few sentences in the manuscript have been modified to this effect.
 >
@@
 We have added a comment to the manuscript to this effect.
-The "established method" for this is the logic laid out by Bryson+2013.
-For instance, given the target star T and a neighbor star N separated by distance D, one can calculate the expected centroid shifts for the various scenarios where the target T hosts an eclipse signal of depth δ, and also when the neighbor N hosts the signal.
-We do not present these calculations in the text, because there is no evidence that the offset itself is statistically signficant.
-Nonetheless, the absence of companions for Kepler-1643 and KOI-7368 does help bolster the interpretation.
+The "established method" for this is the logic laid out by Bryson+2013 (which unfortunately excludes considerations of intrinsic stellar variability).
+For instance, given the target star T and a neighbor star N separated by distance D with known flux ratio, one can calculate the expected centroid shifts for the various scenarios where the target T hosts an eclipse signal of depth δ, and also when the neighbor N hosts the signal.
+We do not present these calculations in the text, because they would not improve the clarity, and there is no evidence that the offsets themselves are statistically significant.
+Nonetheless, the absence of directly resolved companions for Kepler-1643 and KOI-7368 does help bolster the interpretation.
 
@@
 The Roser+2016 reference is already made in the preceding sentence.
-The KC2019 reference we prefer to avoid in reference to the age based on the manual stitching step they used which yielded a very diferent "delta Lyra cluster" (Stephenson-1) than that originally described by Stephenson (1959).
+The KC2019 reference we prefer to avoid in reference to the age based on the manual stitching step they used which yielded a very different "delta Lyra cluster" (Stephenson-1) than that originally described by Stephenson (1959).
 Given the existing citations to the relevant studies, these age citations are not needed.
@@
 ----------
-> This Zari+2018 reference has been updated to refer to Figure 1, to help the reader know where to look.
+> This Zari+2018 reference has been updated to also refer to our Figure 1, to help the reader know where to look.
 
@@
 ----------
-> The Kounkel & Covey 2019 study indeed included many false positives, both in the sense of "clusters" that do not correspond to stars that were ever physically associated, and also in the sense of clusters that do have some "true" core population, but many erroneous field star interlopers.
+> A reference has been added to Zucker et al's rebuttal.
+>
+> In more detail, the Kounkel & Covey 2019 study indeed included many false positives, both in the sense of "clusters" that do not correspond to stars that were ever physically associated, and also in the sense of clusters that do have some "true" core population, but many erroneous field star interlopers.
 Zucker et al's argument however was not particularly helpful for our purposes of identifying which are which.
@@
 It is hard to fake such stellar populations.
-Nonetheless, Zucker+2022 did nicely demonstrate the extreme nature of some of the groups in e.g., XY space (https://faun.rc.fas.harvard.edu/czucker/Paper_Figures/String_Gallery_Interactive.html), and for the extended groups older than a few hundred million years, there is almost no independent evidence for their coevality.
-Despite this point, for the youngest groups, the KC2019 study to us seems to be more likely an example of why the manual stitching must be treated cautiously, and also for why there are better dimensions to cluster in than those selected in the original study.
-A reference has been added to Zucker et al's rebuttal.
+Nonetheless, Zucker+2022 did nicely demonstrate the extreme nature of some of the groups in e.g., XY space (https://faun.rc.fas.harvard.edu/czucker/Paper_Figures/String_Gallery_Interactive.html).
+And yes, for the extended groups older than a few hundred million years, there is almost no independent evidence for their coevality.
+Despite these points, for the youngest groups, the KC2019 study to us seems to be more likely an example of why the manual stitching must be treated cautiously, and also for why there are better dimensions to cluster in than those selected in the original KC2019 study.
 
@@
 ----------
-> $v_b$ is the distance-corrected proper motion in the direction of increasing galactic latitude, and $v_{l^*}=v_l \cos b$ is the same in the direction of increasing galactic longitude after accounting for the local tangent plane correction.
+> Thank you for this catch -- $v_b$ is the distance-corrected proper motion in the direction of increasing galactic latitude, and $v_{l^*}=v_l \cos b$ is the same in the direction of increasing galactic longitude after accounting for the local tangent plane correction.
 We have added verbiage to this effect.
@@
 > We do not entirely understand what, if any, change is requested:  the purpose of these sentences is to give a broad-brush overview of the stellar properties.
-The rough implicit understanding that the Sun would be "near" G2V at ZAMS (while in reality perhaps being a little less bright, and a little lower surface gravity) provides the necessary context that the stars being discussed spanning G8V to K6V are less massive than the Sun, and given ~40 Myr ages, therefore on the pre-main-sequence.
+The implicit understanding that the Sun would be "near" G2V at ZAMS (while in reality perhaps being a little less bright, and a little lower surface gravity) provides the necessary context that the stars being discussed spanning "G8V to K6V" are less massive than the Sun, and given their ~40 Myr ages, on the late stages of the pre-main-sequence.
 
@@
 > Thank you for this wording catch -- the text has been updated to clarify that the temperatures as well as the other spectroscopic parameters agreed with the isochrone parameters within 1-sigma.
-This is a rather brief way of hiding what in our view is more complicated than what needs to be discussed.
+>
+> This is a rather brief way of hiding what in our view is more complicated than what needs to be discussed in the manuscript.
 Table 1 states the adopted (spectroscopic) effective temperatures for all the stars.

### 2022-09-07 papers/paper_review/response_1.txt
update response
v0 response ready for coauthor review

@@
 ----------
-> RONAN TODO
+> We have added the following clarifying sentence to the manuscript:
+> \added{Core distance is the distance to the $k^{\rm th}$ nearest star, and therefore $k$ acts as a smoothing parameter, where a larger value reduces the influence of local overdensities smaller than the scale that interests us.}
 
@@
 ----------
-> RONAN TODO
+> The relevant sentence has been reworded for clarity.
 
@@
 ----------
-> RONAN TODO
+> Thank you for this catch -- the wording in the original text had a subtle error. The revised wording is as follows.
+>
+> "To enable a selection cut that filters out field-star contaminants, we also compute a weight metric, $D$, defined such that the group member with the smallest core distance has $D=1$, the group member with the greatest core distance has $D=0$, and the metric $D$ scales linearly between the two extremes.
+After applying a set of quality cuts on the astrometry and photometry [...] this procedure yields a distribution of weights $D$ that is well described by a log-normal distribution with \log_{10}\mathcal{N}(-1.55,0.61)$."
+>
+> The original wording incorrectly stated that the "weights for the other group members are log-normally distributed" between the two extremes, while instead the scaling is linear.
+The fact that the resulting distributon of weights is well-fit by a log-normal is an emergent property of the HDBScan clustering and our definition of D -- the attached hist_weight.png substantiates the claim that the distribution is "well-described" by such a functional form.
 

### 2022-09-07 papers/paper/ms.tex
update response
v0 response ready for coauthor review

@@
 metric. 
+\added{Core distance is the distance to the $k^{\rm th}$ nearest star,
+and therefore $k$ acts as a smoothing parameter, where a larger value
+reduces the influence of local overdensities smaller than the
+scale that interests us.}
 
 The unsupervised clustering in this case yielded 8 distinct subgroups.
-These groups are then used as the ``seed'' populations for the third
-step, which is to search for objects at least as close to the 10$^{\rm
-th}$ nearest HDBSCAN-identified member in space-velocity coordinates.
-This third step yields stars that are spatially and kinematically
-close to the photometrically young stars, but which cannot be
-identified as young based on their positions in the color--absolute
-magnitude diagram.
+These groups are then used as the ``seed'' populations\replaced{ for
+the third step, which is to search for objects at least as close to
+the 10$^{\rm th}$ nearest HDBSCAN-identified member in space-velocity
+coordinates}{, in which the stellar members each have their own
+individually-assigned distances to their tenth-nearest
+photometrically-young neighbor.  Using those distances, we search the
+entire Gaia EDR3 point source catalog for stars that fall within each
+star's 10$^\th$ nearest-neighbor distance}.  This third step yields
+stars that are spatially and kinematically close to the
+photometrically young stars, but which cannot be identified as young
+based on their positions in the color--absolute magnitude diagram.
 
@@
 distance has $D=1$, the group member with the greatest core
-distance has $D=0$, and weights for the other group members are
-log-normally distributed between these two extremes.  After applying a
+distance has $D=0$, and \replaced{weights for the other group members are
+log-normally distributed between these two extremes}{the weight $D$
+scales linearly between the two extremes}.  After applying a
 set of quality cuts on the astrometry and
@@
 $G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm BP}/\sigma_{G_{BP}}>20$}
-this procedure yields $D \sim \log_{10}\mathcal{N}(-1.55,0.61)$.
+this procedure yields \replaced{$D \sim
+\log_{10}\mathcal{N}(-1.55,0.61)$}{a distribution of weights $D$
+that is well described by a log-normal distribution with
+\log_{10}\mathcal{N}(-1.55,0.61)$}.
 To visualize the results, in Figure~\ref{fig:XYZvtang} we show

### 2022-09-08 papers/paper_review/response_1.txt
update response

@@
 ----------
-> - The Kepler field is shown in the top panel of the figure, between l of 70 to 85, and b of 5 to 20.
+> - The Kepler field is shown with the large gray squares in the top panel of the figure, between l of 70 to 85, and b of 5 to 20.
 This approximately corresponds to a cone in XYZ -- in case it helps intuition, (l,b)=(90,0) corresponds to the line of X=Z=0.
@@
 The lower-right panel of our Figure 1 however shows that while the LSR projection accounts for some of the change in observed v_l* in Cep-Her, the data and the gray band do not match!
-There therefore needs to be some intrinsic differences in the UVW velocities of these stars (though v_r, v_phi, and v_z is perhaps the better coordinate system).
+There therefore needs to be some intrinsic differences in the UVW velocities of these stars (though v_r, v_phi, and v_z may perhaps the better coordinate system).
 
@@
 > The original wording incorrectly stated that the "weights for the other group members are log-normally distributed" between the two extremes, while instead the scaling is linear.
-The fact that the resulting distributon of weights is well-fit by a log-normal is an emergent property of the HDBScan clustering and our definition of D -- the attached hist_weight.png substantiates the claim that the distribution is "well-described" by such a functional form.
+The fact that the resulting distribution of weights is well-fit by a log-normal is an emergent property of the HDBScan clustering and our definition of D -- the attached hist_weight.png substantiates the claim that the distribution is "well-described" by such a functional form.
 
@@
 
-* We added the follow text as a note to table 2: "\added{Users who wish to minimize field star contamination should apply more restrictive weight cuts, {\it e.g.}, $\texttt{weight}>0.1$.}"
+* We added the following text as a note to table 2: "\added{Users who wish to minimize field star contamination should apply more restrictive weight cuts, {\it e.g.}, $\texttt{weight}>0.1$.}"
 
-* We have modified the axis width on all figures for visual clarity.
+* We reduced the axis width on all figures for visual clarity.

### 2022-09-08 papers/paper/ms.tex
update response

@@
 %\submitjournal{AAS Journals}
-\shorttitle{Trio of Mini-Neptunes in Cep-Her}
+\shorttitle{Kepler and the Behemoth}
 
@@
 entire Gaia EDR3 point source catalog for stars that fall within each
-star's 10$^\th$ nearest-neighbor distance}.  This third step yields
+star's 10$^{\rm th}$ nearest-neighbor distance}.  This third step yields
 stars that are spatially and kinematically close to the
@@
 $G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm BP}/\sigma_{G_{BP}}>20$}
-this procedure yields \replaced{$D \sim
-\log_{10}\mathcal{N}(-1.55,0.61)$}{a distribution of weights $D$
+this procedure yields \replaced{$D \sim \log_{10}\mathcal{N}(-1.55,0.61)$}{a distribution of weights $D$
 that is well described by a log-normal distribution with
-\log_{10}\mathcal{N}(-1.55,0.61)$}.
+$\log_{10}\mathcal{N}(-1.55,0.61)$}.
 To visualize the results, in Figure~\ref{fig:XYZvtang} we show

### 2022-09-17 papers/paper_arxiv_v2/collect.sh
AJ: accepted

@@ -1,0 +1,37 @@
+# Cep-Her XYZ
+cp /Users/luke/Dropbox/proj/rudolf/results/CepHer_XYZvtang_sky/CepHer_XYZvtang_sky_showgroups.pdf f1.pdf
+
+# rotation, CAMD, spectra (Li/H-alpha)
+# CAMD:
+
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_RSG-5_cleanhrcut_gaia2018_smalllims_overplotkep1643.pdf f2a.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/hr/hr_Bp_m_Rp_show100pc_redcorr_δ_Lyr_cluster_CH-2_cleanhrcut_gaia2018_smalllims_overplotkoi7368_overplotkoi7913.pdf f2b.pdf
+
+# rotation:
+cp /Users/luke/Dropbox/proj/rudolf/results/rotation/RSG-5_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_defaultcleaning_allknown.pdf f2c.pdf
+cp /Users/luke/Dropbox/proj/rudolf/results/rotation/CH-2_rotation_vs_BpmRp_talkaspect_kinematicspatialselected_linear_defaultcleaning_allknown.pdf f2d.pdf
+
+# phased and sliced LCs
+cp /Users/luke/Dropbox/proj/rudolf/results/koikeplerlc/koikeplerlc_xlim0550_0600.pdf f3a.pdf
+cp ~/Dropbox/proj/rudolf/results/run_simpletransit_Kepler_1643/*singlepanel*plot.pdf f3b.pdf
+cp ~/Dropbox/proj/rudolf/results/run_simpletransit_KOI_7913/*singlepanel*plot.pdf f3c.pdf
+cp ~/Dropbox/proj/rudolf/results/run_simpletransit_KOI_7368/*singlepanel*plot.pdf f3d.pdf
+cp ~/Dropbox/proj/rudolf/results/run_simpletransit_Kepler_1627/*singlepanel*plot.pdf f3e.pdf
+
+# Rp, Period, Age
+cp ~/Dropbox/proj/rudolf/results/rp_vs_period_scatter/rp_vs_period_scatter_20220405_colorbyage_showCepHer.pdf f4.pdf
+
+# spectra
[…]

### 2022-09-22 papers/paper_arxiv_v2/ms.tex
AJ: accepted
Update arxiv upload

--- 2022-09-07 papers/paper/ms.tex
@@
 %\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
-\documentclass[12pt,twocolumn,tighten,linenumbers]{aastex63}
+\documentclass[12pt,twocolumn,tighten]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
@@
 \received{2022 May 2}
-\revised{2022 Sep 6}
-\accepted{---}
-%\submitjournal{AAS Journals}
-\shorttitle{Trio of Mini-Neptunes in Cep-Her}
+\revised{2022 Sep 14}
+\accepted{2022 Sep 17}
+\submitjournal{Astronomical Journal}
+\shorttitle{Kepler and the Behemoth}
 
@@
 \citep[{e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.
-\replaced{To date these analyses have mostly clustered on stellar
-positions and 2D velocities}{So far, these analyses have mostly
-leveraged 3D stellar positions and 2D on-sky tangential velocities}.
+So far, these analyses have mostly
+leveraged 3D stellar positions and 2D on-sky tangential velocities.
 One important result has been the discovery of diffuse
@@
 level.  Internal dynamics and projection effects can also drive them
-to be much \replaced{larger}{more kinematically diffuse}: in the Hyades, stars in the tidal tails are
+to be much more kinematically diffuse: in the Hyades, stars in the tidal tails are
 expected to span up to $\pm 40\mkms$ in velocity relative to the
@@
   of rest \citep{schonrich_local_2010}.  There is a strong spatial and kinematic overlap between
-  Kepler-1643 and RSG-5 (magenta\added{; smaller circles}).  The local population
+  Kepler-1643 and RSG-5 (magenta; smaller circles).  The local population
   of candidate young stars around KOI-7368 and KOI-7913 is more
-  diffuse -- we call this region ``CH-2'' (lime-green\added{; larger
-  circles}).
-  The selection method for \replaced{all of the stars}{these groups}
+  diffuse -- we call this region ``CH-2'' (lime-green; larger
+  circles).
+  The selection method for these groups
   is described in Section~\ref{subsec:members}.
@@
 at \url{http://mkounkel.com/mw3d/mw2d.html} (accessed 15 March
-2022)\added{. Important caveats, particularly for extended groups $\gtrsim 100$\,Myr
-old, were presented by \citet{2022arXiv220514160Z}.}}.
+2022). Important caveats, particularly for extended groups $\gtrsim 100$\,Myr
+old, were presented by \citet{2022arXiv220514160Z}.}.
 The connection made by \citet{KounkelCovey2019} between the previously
@@
 one third of a kiloparsec, identified three of the nearest
-sub-populations\added{ of Cep-Her}, dubbed ``Cepheus-Cygnus'', ``Lyra'', and
+sub-populations of Cep-Her, dubbed ``Cepheus-Cygnus'', ``Lyra'', and
 ``Cerberus''.  \citet{Kerr2021} reported ages for each of these
@@
 the presence of contiguous stellar populations connecting each of
-these groups in \replaced{space-velocity}{both space and velocity} coordinates.  In other words, the stars
+these groups in both space and velocity coordinates.  In other words, the stars
 appear to be comoving, though with a continuous gradient in velocity
-as a function of position\replaced{ (Figure~\ref{fig:XYZvtang}, lower
-panels).}{. The lower panels of Figure~\ref{fig:XYZvtang} show this in
+as a function of position. The lower panels of Figure~\ref{fig:XYZvtang} show this in
 detail, where $v_b$ is the distance-corrected proper motion in the
@@
 Sun's motion with respect to the local standard of rest (see also
-Figure 11 by \citealt{Zari2018}).  }
-In this work, our primary interest in \replaced{the region}{this
-region of sky} stems
+Figure 11 by \citealt{Zari2018}).  
+In this work, our primary interest in this
+region of sky stems
 from the fact that a portion of it was observed by Kepler
@@
 metric. 
+Core distance is the distance to the $k^{\rm th}$ nearest star,
+and therefore $k$ acts as a smoothing parameter, where a larger value
+reduces the influence of local overdensities smaller than the
+scale that interests us.
 
 The unsupervised clustering in this case yielded 8 distinct subgroups.
-These groups are then used as the ``seed'' populations for the third
-step, which is to search for objects at least as close to the 10$^{\rm
-th}$ nearest HDBSCAN-identified member in space-velocity coordinates.
-This third step yields stars that are spatially and kinematically
-close to the photometrically young stars, but which cannot be
-identified as young based on their positions in the color--absolute
-magnitude diagram.
+These groups are then used as the ``seed'' populations, in which the stellar members each have their own
+individually-assigned distances to their tenth-nearest
+photometrically-young neighbor.  Using those distances, we search the
+entire Gaia EDR3 point source catalog for stars that fall within each
+star's 10$^{\rm th}$ nearest-neighbor distance.  This third step yields
+stars that are spatially and kinematically close to the
+photometrically young stars, but which cannot be identified as young
+based on their positions in the color--absolute magnitude diagram.
 
@@
 distance has $D=1$, the group member with the greatest core
-distance has $D=0$, and weights for the other group members are
-log-normally distributed between these two extremes.  After applying a
+distance has $D=0$, and the weight $D$
+scales linearly between the two extremes.  After applying a
 set of quality cuts on the astrometry and
@@
 $G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm BP}/\sigma_{G_{BP}}>20$}
-this procedure yields $D \sim \log_{10}\mathcal{N}(-1.55,0.61)$.
+this procedure yields a distribution of weights $D$
+that is well described by a log-normal distribution with
+$\log_{10}\mathcal{N}(-1.55,0.61)$.
 To visualize the results, in Figure~\ref{fig:XYZvtang} we show
 12{,}436 objects with $D>0.02$ as gray points, and 
-4{,}763 objects with $D>0.10$ as black points.  \added{These thresholds were
+4{,}763 objects with $D>0.10$ as black points.  These thresholds were
 selected visually based on the apparent purity with which they yielded
-pre-main-sequence stars on a color--absolute magnitude diagram.} The
+pre-main-sequence stars on a color--absolute magnitude diagram. The
 $\delta$~Lyr\ cluster is visible at $(l,b)=(68^\circ,15^\circ$) and $(v_{l^*},
@@
 the presence of rotational modulation at the expected period and
-amplitude \added{for stars at least as young as the Pleiades} \citep[{e.g.},][]{rebull_rotation_2020}. 
+amplitude for stars at least as young as the Pleiades \citep[{e.g.},][]{rebull_rotation_2020}. 
 Planetary status was assessed through the Robovetter flags, and by
@@
 \nrsgfive\ candidate RSG-5 members, and \nchtwo\ candidate CH-2
-members.  These stars are listed in Appendix~\ref{app:members}, \replaced{as are
-the Cep-Her candidates that were observed by Kepler}{as is the set of Cep-Her
-candidates that was observed by Kepler}.
+members.  These stars are listed in Appendix~\ref{app:members}, as is the set of Cep-Her
+candidates that was observed by Kepler.
 %An important
@@
 
-		\vspace{-0.5cm}
+		\vspace{-0cm}
 		\subfloat{
@@
     in RSG-5 and CH-2, respectively.  The range of colors is truncated
-    to emphasize the pre-main-sequence.  Stars that fall far below the
+    to emphasize the pre-main-sequence; approximate spectral types are shown on the upper axes.  Stars that fall far below the
     cluster sequences are field interlopers.  {\it Bottom row}:
@@
 yielded $E(B-V)=0.032\pm0.006$.  Finally, for the plots we set the
-color \replaced{range}{axis} to best visualize the region of maximal age information
+color axis to best visualize the region of maximal age information
 content: the pre-main-sequence.
@@
 The CAMDs show that for RSG-5, all but one of the candidate
-members are on a tight pre-main-sequence locus.  \added{Quantitatively,
+members are on a tight pre-main-sequence locus.  Quantitatively,
 88/89 stars with $\bpmrpo \geq 1.5$ are consistent with being on the
-pre-main-sequence.} This implies a false
+pre-main-sequence. This implies a false
 positive rate of a few percent, at most.  In comparison, our
-\replaced{control}{reference}
+reference
 sample (the $\delta$~Lyr candidates) has a false positive rate of
 $\approx$12\%, based on the number of stars that photometrically
-overlap with the field population\deleted{, while being outliers from the bulk
-cluster population}.  For CH-2, our membership selection gives 27
+overlap with the field population.  For CH-2, our membership selection gives 27
 objects in the color range displayed, and 23 of them appear to be
 consistent with being on the pre-main-sequence.  This
-\replaced{implies}{would imply} a false
+would imply a false
 positive rate in CH-2 of $\approx$15\%.
@@
 Figure~\ref{fig:age} also shows that most RSG-5 and CH-2 members
-overlap with the $\delta$~Lyr cluster\added{ on the CAMD}, and that the groups are
+overlap with the $\delta$~Lyr cluster on the CAMD, and that the groups are
 therefore roughly the same age.  To quantify this, we use the method
@@
 suggest that this is indeed the case.
-\added{While in reality stellar populations do not evolve linearly in
+While in reality stellar populations do not evolve linearly in
 the dimensions of absolute magnitude versus color, in our case the
@@
 absolute reference age for IC\,2602 itself
-\citep{david_ages_2015,randich_gaiaeso_2018}.}
+\citep{david_ages_2015,randich_gaiaeso_2018}.
 
@@
 We selected stars suitable for gyrochronology by requiring 
-\replaced{$\bpmrpo \geq 0.5$}{$\bpmrpo \geq 0.6$} to focus on FGKM stars that experience magnetic
+$\bpmrpo \geq 0.6$ to focus on FGKM stars that experience magnetic
 braking.
@@
 ($\sim$M3V) at the relevant distances.  These cuts gave 19 stars in
-CH-2 and \replaced{47}{42} stars in RSG-5.  We extracted light curves from the TESS
+CH-2 and 42 stars in RSG-5.  We extracted light curves from the TESS
 images using the \texttt{unpopular} package \citep{hattorio_2021_cpm},
@@
 The lower panels of Figure~\ref{fig:age} show the results.  In RSG-5,
-\replaced{41/47}{39/42} stars have rotation periods \replaced{faster
-than}{at least as fast as} the Pleiades (\replaced{87}{93}\%).  This
+39/42 stars have rotation periods at least as fast as the Pleiades (93\%).  This
 numerator omits the two stars with periods $>$$12$\,days visible in
@@
 M-dwarfs with 10-12 day rotation periods at ages of USco
-($\sim$$8$\,Myr), and some \deleted{may} still exist at the age of
-LCC ($\sim$$16$\,Myr; \replaced{L.~Rebull
-submitted}{\citealt{2022AJ....164...80R}}).  Regardless, since
-\replaced{nearly no field star outliers seem}{only one field star
-outlier seems} to be present on the RSG-5 CAMD,
+($\sim$$8$\,Myr), and some  still exist at the age of
+LCC ($\sim$$16$\,Myr; \citealt{2022AJ....164...80R}).  Regardless, since
+only one field star
+outlier seems to be present on the RSG-5 CAMD,
 the fact
-that we do not detect rotation periods for $\approx$\replaced{13}{7}\%
+that we do not detect rotation periods for $\approx$7\%
 of stars should perhaps be taken as an indication for the fraction of
@@
 {e.g.}, pole-on stars having lower amplitude starspot modulation.
-\added{Field star contamination is another possible contributor.}
+Field star contamination is another possible contributor.
 
 For CH-2, 13/19 stars have rotation periods that are obviously faster
-than their counterparts in the Pleiades\deleted{ (68\%)}.  4 stars, not included
+than their counterparts in the Pleiades.  4 stars, not included
 in the preceding numerator, are M-dwarfs with rotation periods between
 10 and 12.5 days.  As previously noted, the age interpretation for
-these M-dwarfs is \replaced{not obvious}{ambiguous}.  \added{If none are cluster members,
+these M-dwarfs is ambiguous.  If none are cluster members,
 the rotation period detection fraction is 68\%; if all are members, it
-is 89\%}.
-\replaced{Regardless, the $\approx$15\% false
-positive rate for CH-2 determined from the CAMD seems consistent with
-our fraction of detected rotation periods; RSG-5 was missing rotation
-periods for $\approx$15\% of its candidate members, even though all of
-its stars appear photometrically consistent with being on a single
-pre-main-sequence locus.}{
+is 89\%.
+
 This sets an upper bound on the contamination fraction in our candidate
@@
 field stars than the RSG-5 sample.
-}
 
@@
 and density are found by interpolating against the MIST isochrones
-\added{in reddening-corrected absolute $G$-band magnitude as a function of $\bpmrpo$\ color}
+in reddening-corrected absolute $G$-band magnitude as a function of $\bpmrpo$\ color
 \citep{choi_mesa_2016}.  The statistical uncertainties from this
@@
 
-To verify these parameters\added{, determine the stellar effective temperatures,} and to analyze youth
+To verify these parameters, determine the stellar effective temperatures, and to analyze youth
 proxies such as the Li 6708\,\AA\ doublet and H$\alpha$, we acquired
-\added{high resolution optical } spectra.  We also acquired high resolution imaging for each system, to
+high resolution optical  spectra.  We also acquired high resolution imaging for each system, to
 constrain the existence of visual companions, including possible bound
@@
 yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.  Using the broadened synthetic
-templates\added{\footnote{The broadening is calculated using the joint
+templates\footnote{The broadening is calculated using the joint
 rotational and macroturbulent broadening kernel from \citet{hirano_2011},
@@
 the macroturbulent velocity could be systematically higher (or lower) on
-the pre-main-sequence than it is for more slowly rotating field stars.}}
+the pre-main-sequence than it is for more slowly rotating field stars.}
  from \texttt{SpecMatch-Synth} \citep{petigura_cksi_2017}, we
@@
 the two epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm 1.2\,\mkms$
-respectively\replaced{ \citep{chubak_2012}}{, and was calculated
-following the methods of \citet{chubak_2012}}.  To infer the equivalent width of
+respectively, and was calculated
+following the methods of \citet{chubak_2012}.  To infer the equivalent width of
 the \ion{Li}{1} 6708\,\AA\ doublet, we followed the procedure
-described by \citet{bouma_2021_ngc2516}. \added{In brief, this
+described by \citet{bouma_2021_ngc2516}. In brief, this
 involved calculating the line width by numerically integrating a
@@
 normalization was allowed to vary through a bootstrap approach based
-on the local scatter in the spectra.}  For Kepler-1643, this yielded a
+on the local scatter in the spectra.  For Kepler-1643, this yielded a
 strong detection: ${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with
@@
 spectroscopic parameters included the metallicity $[{\rm Fe/H}]= -0.02
-\pm 0.08$, the equatorial velocity \replaced{$v\sin i = 20.21 \pm 0.50\,\mkms$}{$v\sin i = 20.2 \pm 1.0\,\mkms$},
+\pm 0.08$, the equatorial velocity $v\sin i = 20.2 \pm 1.0\,\mkms$,
 and the systemic velocity ${\rm RV}_{\rm sys} = -10.9 \pm 0.2\,\mkms$.
@@
 \texttt{SpecMatch-Emp} machinery yielded $T_{\rm eff,A} = 4324 \pm
-70\,{\rm K}$, and $T_{\rm eff,B} = 4038 \pm 70\,{\rm K}$.  \replaced{The
-remaining parameters were in agreement}{These temperatures as well as the other spectroscopic parameters agreed} with those from the cluster
-isochrone\added{ method within 1-$\sigma$}.  For the primary, we also found $[{\rm Fe/H}]= -0.06 \pm
+70\,{\rm K}$, and $T_{\rm eff,B} = 4038 \pm 70\,{\rm K}$.  These temperatures as well as the other spectroscopic parameters agreed with those from the cluster
+isochrone method within 1-$\sigma$.  For the primary, we also found $[{\rm Fe/H}]= -0.06 \pm
 0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm sys} =
@@
 $[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm 1.0\,\mkms$, and
-${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.  \replaced{Neither of the KOI-7913
-components shows strong lithium absorption, but this is expected given
-their $\approx$K6V and $\approx$K8V spectral types}{The primary showed lithium in absorption
+${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.  The primary showed lithium in absorption
 with ${\rm EW}_{\rm Li} =
 65^{+8}_{-6}$\,m\AA, while the secondary had a marginal detection of ${\rm EW}_{\rm Li} =
-42^{+12}_{-19}$\,m\AA. }.  \replaced{They do however
-both show H$\alpha$ in emission.  Broadly speaking, both}{Both components displayed 
-H$\alpha$ in emission.  Given the spectral types of the stars, these} observations
+42^{+12}_{-19}$\,m\AA. .  Both components displayed 
+H$\alpha$ in emission.  Given the spectral types of the stars, these observations
 are consistent with a $\approx$40 Myr age for KOI-7913 (see
@@
 because it would have a mass between 10 and
-15\,M$_{\rm Jup}$\added{ at an assumed age of 40\,Myr}.  Additional imaging epochs will tell.
+15\,M$_{\rm Jup}$ at an assumed age of 40\,Myr.  Additional imaging epochs will tell.
 
@@
     \citet{berger_identifying_2018}.  The ``40-50 Myr'' reference
-    stars ({\it left}) are from IC\,2602 \replaced{\citep{randich_gaiaeso_2018}}{\citep{randich_2001}} and
+    stars ({\it left}) are from IC\,2602 \citep{randich_2001} and
     Tuc-Hor \citep{kraus_stellar_2014}.  The ``112 Myr'' stars
@@
     The H$\alpha$ comparison is against Tuc-Hor
-    \citep[$\approx$$40$\,Myr;][]{kraus_stellar_2014}\added{ and the Pleiades \citep{fang_2018}.}
+    \citep[$\approx$$40$\,Myr;][]{kraus_stellar_2014} and the Pleiades \citep{fang_2018}.
     \label{fig:lithium}
@@
 Kepler-1643, in RSG-5, is conspicuously below the 40-50$\,$Myr
-sequence in the \replaced{left-panel}{top panel} of Figure~\ref{fig:lithium}, though
+sequence in the top panel of Figure~\ref{fig:lithium}, though
 above the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA).
-\added{
+
 Quantitatively, there are 14 reference stars within $\pm$150\,K of Kepler-1643.
@@
 which implies that Kepler-1643 is $4.0$-$\sigma$ discrepant from expectations.
-}
-The \replaced{right}{middle} panel shows \replaced{the}{a} comparison against the Pleiades, where
+
+The middle panel shows a comparison against the Pleiades, where
 Kepler-1643 is more consistent with the observed dispersion in
@@
 
-One explanation for the \replaced{lack of Li in Kepler-1643}{low Li equivalent width
-in Kepler-1643 relative to the comparison stars} could be that it is a field
+One explanation for the low Li equivalent width
+in Kepler-1643 relative to the comparison stars could be that it is a field
 interloper; another could be that RSG-5 is much older than 50\,Myr.  We do
@@
 Kepler-1643 also seems highly unlikely to be a field interloper, 
-because we demonstrated a \replaced{$\approx$$1\%$}{few-percent} false positive probability
+because we demonstrated a few-percent false positive probability
 in our spatio-kinematic selection of RSG-5 members, and there is a 
@@
 Our preferred explanation for Kepler-1643's meager lithium 
-content is that the reference sample\added{s} of IC\,2602 and Tuc-Hor
-stars may not fully explore all possible \deleted{stellar rotation
-periods and }lithium equivalent widths at this age.
-\added{This would be somewhat surprising since over a dozen stars have
+content is that the reference samples of IC\,2602 and Tuc-Hor
+stars may not fully explore all possible lithium equivalent widths at this age.
+This would be somewhat surprising since over a dozen stars have
 already been analyzed in the relevant effective temperature range.
-However, considering}
-the top panels of Figure~\ref{fig:lithium}, it is \added{also} remarkable that in
+However, considering
+the top panels of Figure~\ref{fig:lithium}, it is also remarkable that in
 50 million years, stars between $4500$\,K and $5200$\,K go from
@@
 effective temperature range.
-\added{An alternative explanation could be that the overall metallicity of Cep-Her
+An alternative explanation could be that the overall metallicity of Cep-Her
 is different from Tuc-Hor and IC\,2602, but this seems unlikely given the near-solar
-metallicities we have measured for the Kepler Objects of Interest.}
-\replaced{This suggests that RSG-5 and Cep-Her could be worthy objects for a
-closer analysis of the lithium-rotation correlation near the zero-age
-main sequence.}{Broadly, these considerations suggest that Cep-Her is
+metallicities we have measured for the Kepler Objects of Interest.
+Broadly, these considerations suggest that Cep-Her is
 a worthy object for further spectroscopic analyses of lithium near the
-zero-age main sequence.}
+zero-age main sequence.
 
@@
     dominant signal is from starspots; planetary transit times
-    are indicated \added{with vertical dashed lines, }but the individual transits are not visible at this scale.
+    are indicated with vertical dashed lines, but the individual transits are not visible at this scale.
     {\it Bottom}:
@@
 variability using a quasiperiodic Gaussian process (GP).  We explored
-this approach, but \replaced{found that it often required fine-tuning of the
-hyperparameter priors, otherwise the GP would overfit flares,
-instrumental noise, and any other variability that was not part of the
-transit.  This failure mode is pernicious, in that it yields an
-ill-founded sense of confidence in a visually clean fit.  Our model is
-simpler, and it has}{ultimately prefer our model for its simplicity, and for} the benefit that the white noise jitter never
+this approach, but ultimately prefer our model for its simplicity, and for the benefit that the white noise jitter never
 trades off with any parameter equivalent to a damping timescale for
@@
 centroid shift tests (see the \texttt{q1\_q17\_dr25\_koi} data
-release): the angular distance between the target star's \added{KIC }catalog
+release): the angular distance between the target star's KIC catalog
 position and the position of the transiting source was measured as
@@
 outlying quarters (2 and 6) drive the offset --- the centroid locations
-from the other Kepler quarters are consistent at $\lesssim 0\farcs4$\added{ (3-$\sigma$)}.
-\added{\citet{bryson_2013} showed that for typical field star KOIs without
+from the other Kepler quarters are consistent at $\lesssim 0\farcs4$ (3-$\sigma$).
+\citet{bryson_2013} showed that for typical field star KOIs without
 centroid offsets, the mean offset distribution peaks at 0.3$''$ (their
@@
 localized to nearby stars have a distribution that peaks at 7$''$ (their
-Figure 32). } \replaced{This is an instructive exercise in how }{The }stellar
-variability \added{in Kepler-1643} complicates \added{the} centroid-based
-vetting tests\replaced{. The}{, because the} shifts measured by these tests are
+Figure 32).  The stellar
+variability in Kepler-1643 complicates the centroid-based
+vetting tests, because the shifts measured by these tests are
 determined from the in- and out-of-transit flux-weighted centroids.  For stars
 with significant spot-induced variability there is no static baseline in either
-the in- or out-of-transit phases\added{, and so the centroid location may shift
+the in- or out-of-transit phases, and so the centroid location may shift
 depending on the rotation phase combined with the local scene.  Based on
 these considerations, the centroid-level diagnostics for Kepler-1643 appear to
-be consistent with the transit signal being localized to the target star}.
+be consistent with the transit signal being localized to the target star.
 
@@
 calculation, we find ${\rm FPP} = 4\times10^{-3}$.  Though not
-as convincing as Kepler-1643, this clears the \replaced{usual threshold
-for calling the planet statistically validated
-\citep{morton_efficient_2012}}{threshold probability of 1 in 100 suggested by \citet{morton_false_2016} for calling a planet statistically validated}.  The S/N of the transit is
+as convincing as Kepler-1643, this clears the threshold probability of 1 in 100 suggested by \citet{morton_false_2016} for calling a planet statistically validated.  The S/N of the transit is
 $\approx$$32$, which indicates that it is unlikely to be caused by
 systematic noise in the light curve (see Figure~\ref{fig:planets}).
-The positional probability\deleted{ score}\added{\footnote{Columns $\textit{pp\_host\_rel\_prob}$ and $\textit{pp\_host\_prob\_score}$ on the KOI Positional Probabilities table at the NASA Exoplanet Archive \citep{Akeson13}.}}\added{ calculated by \citet{2017ksci.rept...16B}} also \replaced{meets the requirements for transiting sources thought to share positions with their target stars \citep{2017ksci.rept...16B}}{indicates that the transit signal shares its position with the target star.}
+The positional probability\footnote{Columns $\textit{pp\_host\_rel\_prob}$ and $\textit{pp\_host\_prob\_score}$ on the KOI Positional Probabilities table at the NASA Exoplanet Archive \citep{Akeson13}.} calculated by \citet{2017ksci.rept...16B} also indicates that the transit signal shares its position with the target star.
 
@@
 Kepler-1643.  For KOI-7368, the reported offset is smaller, and less
-\added{formally }significant ($0\farcs2$; 3.0-$\sigma$).  Again, the
+formally significant ($0\farcs2$; 3.0-$\sigma$).  Again, the
 data validation reports show that the shift is caused by a few
 outlying quarters (4, 5, 8, and 12).  Since the remaining
-\replaced{data}{quarters} show
+quarters show
 consistent scatter in their centroid locations, these outlying
-quarters are likely also caused by the stellar variability\added{,
-because their directions are inconsistent across different quarters}.
-Our NIRC2 imaging independently shows that there are no \added{known
-}neighboring sources that could cause an offset of the observed
-amplitude\added{, as is also the case for Kepler-1643}.
+quarters are likely also caused by the stellar variability,
+because their directions are inconsistent across different quarters.
+Our NIRC2 imaging independently shows that there are no known
+neighboring sources that could cause an offset of the observed
+amplitude, as is also the case for Kepler-1643.
 
@@
 1.3\times10^{-4}$.  Though the transit has the lowest S/N of any of
-the objects discussed ($\approx$$14$), its low\deleted{er} FPP \deleted{relative to
-KOI-7368} can be understood through its flat-bottomed shape,
+the objects discussed ($\approx$$14$), its low FPP  can be understood through its flat-bottomed shape,
 combined with its long transit duration relative to most
-eclipsing binary models (Figure~\ref{fig:planets}).  \replaced{The host star
-probability usability score \citep{2017ksci.rept...16B} also meets the
-usual threshold, and so the planet is statistically validated}{The positional probability calculation performed by \citet{2017ksci.rept...16B} yielded a near-unity probability that the transit event is at the same location as the host star, and so the cumulative evidence suggests that KOI-7913~Ab is indeed a statistically validated planet}.
+eclipsing binary models (Figure~\ref{fig:planets}).
+The positional probability calculation performed by \citet{2017ksci.rept...16B} yielded a near-unity probability that the transit event is at the same location as the host star, and so the cumulative evidence suggests that KOI-7913~Ab is indeed a statistically validated planet.
 Its
@@
     between 2 and 3\,$R_\oplus$, which is a novelty given their ages.
-    Parameters are from the NASA Exoplanet Archive (2022 Apr 5).
+    Parameters are from the \citet{PSCompPars}.
 		\label{fig:rp_period_age}
@@
 core composition, and a typical H/He mass fraction of $\approx$4\%,
-there should be a \added{single} local maximum in planet occurrence rates at 2 to 3\,$R_\oplus$,
-at times between 10 and 100\,Myr. \added{In other words,
+there should be a single local maximum in planet occurrence rates at 2 to 3\,$R_\oplus$,
+at times between 10 and 100\,Myr. In other words,
 \citet{rogers_unveiling_2021} predict the existence of a ``radius
-mountain'' at these early times, rather than a ``radius valley''.}  The models advanced by
+mountain'' at these early times, rather than a ``radius valley''.  The models advanced by
 \citet{gupta_signatures_2020} and \citet{lee_primordial_2021}
-agree\added{ that this local maximum should exist};
+agree that this local maximum should exist;
 their differences lie in the mechanism for producing the radius
@@
 acknowledges that an important factor in cluster identification is
-also \replaced{density}{coherence} in velocity space.
-\added{For instance, the Psc-Eri stream, which has a shape that can be approximated as a 600 parsec-long cylinder with a radius of 30 parsecs, has a number density roughly a factor of three times lower than even CH-2 \citep{roser_psceri_2020}.
+also coherence in velocity space.
+For instance, the Psc-Eri stream, which has a shape that can be approximated as a 600 parsec-long cylinder with a radius of 30 parsecs, has a number density roughly a factor of three times lower than even CH-2 \citep{roser_psceri_2020}.
 However its existence is discernible because of the $\lesssim 2.5$\kms\ scatter in its cylindrical velocities.
-}Perhaps once stellar rotation periods
+Perhaps once stellar rotation periods
 and chemical abundances reach the same level of ubiquity as stellar
@@
 Figure~\ref{fig:XYZvtang} deserve independent attention.  An important
-aspect of the remaining work will be to acquire radial velocities \replaced{of a
-large}{for a larger} subset of the stars, and to determine whether the traceback approach
+aspect of the remaining work will be to acquire radial velocities for a larger subset of the stars, and to determine whether the traceback approach
 could be applicable.  Wide-field spectroscopic surveys such as LAMOST
 \citep{zhao_2012_LAMOST} or SDSS-V \citep{kollmeier_2017} could enable
-such analyses\added{ for the brightest members}, while also providing sensitivity to the Li 6708\,\AA\
-line.  \replaced{If Gaia DR3 provides RVS spectra for the brighter F and G
-dwarfs, these would also contain the calcium infrared triplet as
-another age indicator}{The Gaia DR3 RVS spectra (released during review of this manuscript) could contain similar velocity information down to spectral types of $\approx$K5V ($G_{\rm RVS}\lesssim 14$), and perhaps also enable analyses of the calcium infrared triplet as a youth indicator}.
-\replaced{These indicators, combined with better
-kinematics,}{The combination of more complete kinematics and youth indicators} would help in definitively unraveling the formation
+such analyses for the brightest members, while also providing sensitivity to the Li 6708\,\AA\
+line.
+The Gaia DR3 RVS spectra (released during review of this manuscript) could contain similar velocity information down to spectral types of $\approx$K5V ($G_{\rm RVS}\lesssim 14$), and perhaps also enable analyses of the calcium infrared triplet as a youth indicator.
+The combination of more complete kinematics and youth indicators would help in definitively unraveling the formation
 history of the complex.
@@
 Kepler-1643 and KOI-7368 
-both show weak centroid offsets \replaced{between the in
-and out-of-transit phases}{in specific quarters}.  For both \replaced{cases}{systems}, the observed shifts
+both show weak centroid offsets in specific quarters.  For both systems, the observed shifts
 are consistent with being caused by starspot-induced variability in
@@
 more diffuse.  For KOI-7913, stronger indicators of its age come from
-its \replaced{binary}{binarity}.  Both stellar components in KOI-7913 have isochronal ages
+its binarity.  Both stellar components in KOI-7913 have isochronal ages
 consistent with $40$\,Myr.  Both components also show H$\alpha$ in
-emission, which for the \added{transit-hosting} $\approx$K6V primary is a strong indicator
-that the star is $\lesssim$100$\,$Myr old.  KOI-7368 is \deleted{slightly} more
-massive, and \replaced{the}{its} Li 6708\,\AA\ measurement and stellar rotation period
+emission, which for the transit-hosting $\approx$K6V primary is a strong indicator
+that the star is $\lesssim$100$\,$Myr old.  KOI-7368 is  more
+massive, and its Li 6708\,\AA\ measurement and stellar rotation period
 provide independent verification of the star's youth.
@@
 We adopted the ZTF period over the
-TESS period in three cases: (1) Gaia EDR3 2081755809272821248: \replaced{the
-Lomb-Scargle periodogram favored 6.67 days, consistent with the ZTF
-period of 6.61 days; however, we flagged it as a candidate
-double-dipper, which appears to have inaccurately doubled the TESS
-period to 13.34 days}{the top ZTF Lomb-Scargle peak gave 6.61 days, while our default pipeline favored a TESS peak of 13.34 days;  manual inspection of the light curve favors the former}; (2) Gaia EDR3 2081737529891330560: we found 3.06
+TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the top ZTF Lomb-Scargle peak gave 6.61 days, while our default pipeline favored a TESS peak of 13.34 days;  manual inspection of the light curve favors the former; (2) Gaia EDR3 2081737529891330560: we found 3.06
 days with TESS and 6.64 days with ZTF; we suspect that TESS captured

### 2023-09-22 erratum/ms.tex
stage erratum

--- 2022-09-22 papers/paper_arxiv_v2/ms.tex
@@
 %\documentclass[12pt,modern,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
-\documentclass[12pt,twocolumn,tighten]{aastex63}
+\documentclass[12pt,tighten,linenumbers]{aastex63}
 %\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
@@
 %% Add "Submitted to " argument.
-\received{2022 May 2}
-\revised{2022 Sep 14}
-\accepted{2022 Sep 17}
+%\received{2022 May 2}
+%\revised{2022 Sep 14}
+%\accepted{2022 Sep 17}
 \submitjournal{Astronomical Journal}
@@
 \title{
-  Kepler and the Behemoth: Three Mini-Neptunes in a 40 Million Year Old Association
+  Erratum: ``Kepler and the Behemoth: Three Mini-Neptunes in a 40 Million Year Old Association'' (2022, AJ, 164, 215)
 }
@@
 
-% 208 words (250 max)
-\begin{abstract}
-  %The best way to measure a planet's age is to ignore the planet and
-  %pay attention to the star.
-  Stellar positions and velocities from Gaia are yielding a new view
-  of open cluster dispersal.  
-  Here we present an analysis of a group
-  of stars spanning Cepheus ($l=100^\circ$) to Hercules ($l=40^\circ$),
-  hereafter the Cep-Her complex.  The group includes four Kepler
-  Objects of Interest:
-  Kepler-1643 b ($R_{\rm p} = 2.32 \pm 0.13\,R_\oplus$, $P = 5.3\ {\rm days}$),
-  KOI-7368 b ($R_{\rm p} = 2.22 \pm 0.12\,R_\oplus$, $P = 6.8\ {\rm days}$), 
-  KOI-7913 Ab ($R_{\rm p} = 2.34 \pm 0.18\,R_\oplus$, $P = 24.2\ {\rm days}$), and
-  Kepler-1627 Ab ($R_{\rm p} = 3.85 \pm 0.11\,R_\oplus$, $P = 7.2\ {\rm days}$).
-  The latter Neptune-sized planet is in part of the Cep-Her complex
-  called the $\delta$\ Lyr\ cluster \citep{bouma_kep1627_2022}.  Here
-  we focus on the former three systems, which are in other regions of
-  the association.  Based on kinematic evidence from Gaia, stellar
-  rotation periods from TESS, and spectroscopy, these three objects
-  are also $\approx$40 million years (Myr) old.  More specifically, we
-  find that Kepler-1643 is $46^{+9}_{-7}$\,Myr old, based on its
-  membership in a dense sub-cluster of the complex called RSG-5.
-  KOI-7368 and KOI-7913 are $36^{+10}_{-8}$\,Myr old, and are in a
-  diffuse region that we call CH-2.  Based on the transit shapes and
-  high resolution imaging, all three objects are most likely planets,
-  with false positive probabilities of $6\times10^{-9}$,
-  $4\times10^{-3}$, and $1\times10^{-4}$ for Kepler-1643, KOI-7368,
-  and KOI-7913 respectively.  These planets demonstrate
-  that mini-Neptunes with sizes of $\approx$2 Earth radii exist at ages
-  of 40 million years.
-\end{abstract}
-
-\keywords{
-  exoplanet evolution (491),
-  open star clusters (1160),
-	stellar ages (1581)
-}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Introduction}
+\section*{ }
 
-The discovery and characterization of planets younger than a billion
-years is a major frontier in current exoplanet research.  The reason
-is that the properties of young planets provide benchmarks for studies
-of planetary evolution.  For instance, young planets can inform our
-understanding of when hot Jupiters arrive on their close-in orbits
-\citep{dawson_johnson_2018}, how the sizes of planets with massive
-gaseous envelopes evolve \citep{rizzuto_tess_2020}, the timescales for
-close-in multiplanet systems to fall out of resonance
-\citep{izidoro_breaking_2017,arevalo_stability_2022,goldberg_architectures_2022}, and
-whether and how mass-loss explains the radius valley
-\citep{lopez_how_2012,Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018,lee_primordial_2021}.
-
-The discovery of a young planet requires two claims to be true: the
-planet must exist, and its age must be secured.  Spaced-based
-photometry from K2 and TESS has yielded a number of exemplars for
-which the planetary evidence comes from transits, and the age is based
-on either cluster membership
-\citep{Mann_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,nardiello_pathosII_2020}
-or else on correlates of youth such as stellar rotation, photospheric
-lithium content, x-ray activity, and emission line strength
-\citep{zhou_2021_tois,hedges_toi-2076_2021}.
-
-In this work, we leverage recent analyses of the Gaia data, which have
-greatly expanded our knowledge of stellar groups
-\citep[{e.g.},][]{CantatGaudin2018a,KounkelCovey2019,Kerr2021}.
-So far, these analyses have mostly
-leveraged 3D stellar positions and 2D on-sky tangential velocities.
-One important result has been the discovery of diffuse
-streams and tidal tails comparable in stellar mass to the previously
-known cores of nearby open clusters
-\citep{meingast_psceri_2019,Meingast2021,gagne_number_2021}.  Even
-though these streams are spread over tens to hundreds of parsecs,
-their velocity dispersions can remain coherent at the $\sim$1\kms\
-level.  Internal dynamics and projection effects can also drive them
-to be much more kinematically diffuse: in the Hyades, stars in the tidal tails are
-expected to span up to $\pm 40\mkms$ in velocity relative to the
-cluster center \citep{jerabkova_800_2021}.  The stars in such diffuse
-regions can be verified to be the same age as the core cluster members
-through analyses of color--absolute magnitude diagrams
-\citep{KounkelCovey2019}, stellar rotation periods
-\citep{curtis_tess_2019,bouma_2021_ngc2516}, and chemical abundances
-\citep{hawkins_2020,2020A&A...635L..13A}.  While there are implications for our
-understanding of star formation and cluster evolution
-\citep{dinnbier_tidal_2020}, a separate consequence is that we now
-know the ages of many more stars, including previously known planet
-hosts.
-
-The prime Kepler mission \citep{borucki_kepler_2010} found most of the
-currently known transiting exoplanets, and it was conducted before
-Gaia.  It is therefore sensible to revisit the Kepler field, given our
-new constraints on the stellar ages.
-
-Here, we expand on our earlier study of a $38^{+7}_{-6}$ Myr old
-Neptune-sized planet in the Kepler field (Kepler-1627~Ab;
-\citealt{bouma_kep1627_2022}).  This planet's age was derived based on
-its host star's membership in the $\delta$\ Lyr\ cluster.  While our
-analysis of the cluster focused on the immediate vicinity of
-Kepler-1627 in order to have a reasonable scope,  it became clear that
-the $\delta$\ Lyr\ cluster seems to also be part of a much larger
-group of similarly aged stars.  This association, which is at an
-average distance of 330\,pc from the Sun, spans Cepheus to Hercules
-(galactic longitudes, $l$, between 40$^\circ$ and 100$^\circ$), at
-galactic latitudes between 0$^\circ$ and 20$^\circ$.  An assessment of
-its membership, substructure, and age distribution will be provided as
-part of the 1\,kpc expansion of the SPYGLASS project (R. Kerr et al.\
-in prep), where it is given the name Cep-Her, after the endpoint
-constellations.
-
-Our focus is on the intersection of the Cep-Her complex with the
-Kepler field.  Cross-matching the stars thought to be in Cep-Her against known Kepler
-Objects of Interest (KOIs; \citealt{thompson_planetary_2018}) yielded
-four candidate cluster members: Kepler-1627, Kepler-1643, KOI-7368,
-and KOI-7913.  Given our earlier analysis of Kepler-1627, we focus
-here on the latter three objects.  After analyzing the relevant
-properties of Cep-Her (Section~\ref{sec:cluster}), we derive the
-stellar properties (Section~\ref{sec:stars}) and validate the
-planetary nature of each system using a combination of the Kepler
-photometry and high-resolution imaging (Section~\ref{sec:planets}).
-We conclude with a discussion of mini-Neptune size evolution, and
-point out possible directions for future work
-(Section~\ref{sec:disc_conc}).
-
-\section{The Cep-Her Complex}
-\label{sec:cluster}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.99\textwidth]{f1.pdf}
-	\end{center}
-	\vspace{-0.6cm}
-	\caption{
-  {\bf Positions and velocities of candidate members of the Cep-Her
-  complex.}
-  {\it Top row}: On-sky positions in galactic coordinates.  Black
-  points are stars for which group membership is more secure than for
-  gray points.  Kepler-1627 is in the outskirts of the $\delta$\ Lyr
-  cluster \citep{bouma_kep1627_2022}, which is centered at $(l,b)
-  \approx (66^\circ, 12^\circ)$.
-  The Kepler footprint is shown in gray.
-  {\it Middle row}: Galactic positions.  The Sun is at $(X, Y, Z) =
-  (0, 0, 20.8)$\,pc; lines of constant heliocentric distance are
-  shown between 250 and 400\,pc, spaced by 50\,pc.
-  {\it Bottom row}: Galactic tangential velocities (left) and
-  galactic longitudinal velocity versus galactic longitude (right).
-  The gray band in the lower-right shows the $\pm$1-$\sigma$
-  projection of the Solar velocity with respect to the local standard
-  of rest \citep{schonrich_local_2010}.  There is a strong spatial and kinematic overlap between
-  Kepler-1643 and RSG-5 (magenta; smaller circles).  The local population
-  of candidate young stars around KOI-7368 and KOI-7913 is more
-  diffuse -- we call this region ``CH-2'' (lime-green; larger
-  circles).
-  The selection method for these groups
-  is described in Section~\ref{subsec:members}.
-	\label{fig:XYZvtang}
-	}
-\end{figure*}
-
-\subsection{Previous Related Work}
-
-Our focus is on a region of the Galaxy 200 to 500\,pc
-from the Sun, above the galactic plane, and spanning galactic
-longitudes of $40^\circ$ to $100^\circ$.  Two rich clusters in
-this region are the $\delta$\ Lyr cluster
-\citep{stephenson_possible_1959} and RSG-5 \citep{roser_nine_2016}.
-Each of these clusters was known before Gaia.  Their reported ages are
-between 30 and 60\,Myr.  Early empirical evidence that these two
-clusters could be part of a large and more diffuse population was
-apparent in the Gaia-based photometric analysis of pre-main-sequence
-stars by \citet[][compare their Figures~11 and~13 to our
-Figure~\ref{fig:XYZvtang}]{Zari2018}.  Further kinematic connections
-and complexity were highlighted by \citet{KounkelCovey2019}, who
-included these previously known groups in the larger structures dubbed
-``Theia~73'' and ``Theia~96''\footnote{See their visualization online
-at \url{http://mkounkel.com/mw3d/mw2d.html} (accessed 15 March
-2022). Important caveats, particularly for extended groups $\gtrsim 100$\,Myr
-old, were presented by \citet{2022arXiv220514160Z}.}.
-The connection made by \citet{KounkelCovey2019} between the previously
-known open clusters and the other groups in the region was made as
-part of an unsupervised clustering analysis of the Gaia DR2 positions
-and on-sky velocities with a subsequent manual ``stitching'' step.
-Their results support the idea that there is an overdensity of 30 to
-60\,Myr old stars in this region of the Galaxy.  \citet{Kerr2021}, in
-a volume-limited analysis of the Gaia DR2 point-source catalog out to
-one third of a kiloparsec, identified three of the nearest
-sub-populations of Cep-Her, dubbed ``Cepheus-Cygnus'', ``Lyra'', and
-``Cerberus''.  \citet{Kerr2021} reported ages for each of these
-subgroups between 30 and 35 Myr.
-
-\subsection{Member Selection}
-\label{subsec:members}
-
-The possibility that the $\delta$~Lyr cluster, RSG-5, and the
-sub-populations identified by \citet{Kerr2021} share a common origin
-has yet to be fully substantiated, but preliminary clustering results
-from the 1\,kpc SPYGLASS analysis (R.~Kerr et al.\ in prep) suggest
-the presence of contiguous stellar populations connecting each of
-these groups in both space and velocity coordinates.  In other words, the stars
-appear to be comoving, though with a continuous gradient in velocity
-as a function of position. The lower panels of Figure~\ref{fig:XYZvtang} show this in
-detail, where $v_b$ is the distance-corrected proper motion in the
-direction of increasing galactic latitude, and $v_{l^*}=v_l \cos b$ is
-the distance-corrected proper motion in the direction of increasing
-galactic longitude after accounting for the local tangent plane
-correction.  Some, but not all, of the gradient in the $v_{l^*}$ vs.~$l$
-plane can be understood through a projection effect stemming from the
-Sun's motion with respect to the local standard of rest (see also
-Figure 11 by \citealt{Zari2018}).  
-In this work, our primary interest in this
-region of sky stems
-from the fact that a portion of it was observed by Kepler
-(Figure~\ref{fig:XYZvtang}, top panel).  To further explore this
-sub-population, we select candidate Cep-Her members through four
-steps, the first three being identical to those described in Section~3
-of \citet{Kerr2021}.  We briefly summarize them here.
-
-The first step is to select stars that are photometrically distinct
-from the field star population based on Gaia EDR3 magnitudes $\{G,
-G_{\rm RP}, G_{\rm BP}\}$, parallaxes and auxiliary reddening
-estimates \citep{lallement_gaia-2mass_2019}.  This step yielded \npms\
-stars with high-quality photometry and astrometry.  These stars are
-either pre-main-sequence K and M dwarfs due to their long contraction
-timescales, or massive stars near the zero-age main sequence due to
-their rapid evolutionary timescales.
-
-The second step is to perform an unsupervised HDBSCAN clustering on
-the photometrically selected population
-\citep{campello_hierarchical_2015,mcinnes_hdbscan_2017}.  The
-parameters we use in the clustering are $\{ X, Y, Z, c v_b, c v_{l^*}
-\} $, where $c$ is
-the size-velocity corrective factor, which is taken as $c=6\,{\rm pc /
-km\,s}^{-1}$ to ensure that the spatial and velocity scales have
-identical standard deviations.  Positions are computed assuming the
-\texttt{astropy v4.0} coordinate standard
-\citep{astropy_2018}. %, which places the Sun $8122$ pc from the galactic
-%center, and assumes the solar velocity with respect to the local
-%standard of rest from \citet{schonrich_local_2010}.  
-As input
-parameters to HDBSCAN, we set the minimum $\epsilon$ threshold past
-which clusters cannot be fragmented as $25$\,pc in physical space, and
-$25/c$\,\kms\ in velocity.  The minimum cluster size $N$ is set to 10, as
-is $k$, the parameter used to define the ``core distance'' density
-metric. 
-Core distance is the distance to the $k^{\rm th}$ nearest star,
-and therefore $k$ acts as a smoothing parameter, where a larger value
-reduces the influence of local overdensities smaller than the
-scale that interests us.
-
-The unsupervised clustering in this case yielded 8 distinct subgroups.
-These groups are then used as the ``seed'' populations, in which the stellar members each have their own
-individually-assigned distances to their tenth-nearest
-photometrically-young neighbor.  Using those distances, we search the
-entire Gaia EDR3 point source catalog for stars that fall within each
-star's 10$^{\rm th}$ nearest-neighbor distance.  This third step yields
-stars that are spatially and kinematically close to the
-photometrically young stars, but which cannot be identified as young
-based on their positions in the color--absolute magnitude diagram.
-
-The outcome of the analysis up to the point of the third step is shown
-in Figure~\ref{fig:XYZvtang}.  To enable a selection cut that filters
-out field-star contaminants, we also compute a weight metric,
-$D$, defined such that the group member with the smallest core
-distance has $D=1$, the group member with the greatest core
-distance has $D=0$, and the weight $D$
-scales linearly between the two extremes.  After applying a
-set of quality cuts on the astrometry and
-photometry\footnote{$\varpi/\sigma_\varpi>5$; $G/\sigma_{G}>50$;
-$G_{\rm RP}/\sigma_{G_{RP}}>20$; $G_{\rm BP}/\sigma_{G_{BP}}>20$}
-this procedure yields a distribution of weights $D$
-that is well described by a log-normal distribution with
-$\log_{10}\mathcal{N}(-1.55,0.61)$.
-To visualize the results, in Figure~\ref{fig:XYZvtang} we show
-12{,}436 objects with $D>0.02$ as gray points, and 
-4{,}763 objects with $D>0.10$ as black points.  These thresholds were
-selected visually based on the apparent purity with which they yielded
-pre-main-sequence stars on a color--absolute magnitude diagram. The
-$\delta$~Lyr\ cluster is visible at $(l,b)=(68^\circ,15^\circ$) and $(v_{l^*},
-v_b)=(-4.5 ,-4 )\mkms$.  RSG-5 is visible at $(l,b)=(83^\circ,6^\circ)$,
-$(v_{l^*}, v_b)=(5.5 ,-3.5 )\mkms$.  Most of the other subclusters, including
-in Cep-Cyg ($l,b=90^\circ,7^\circ$) and Cerberus ($l,b=48^\circ,18^\circ$) are
-too small or dispersed to have previously been analyzed in great detail.
-
-%
-% KOI match numbers from calc_CepHer_group_numbers.py
-%
-Our fourth and final step was to cross-match the candidate Cep-Her
-member list against all known Kepler Objects of Interest.  We used the
-Cumulative KOI table from the NASA Exoplanet Archive from 27 March
-2022, and also compared against the \texttt{q1\_q17\_dr25} table
-\citep{thompson_planetary_2018}.  From the candidate members with
-weights exceeding 0.02, this yielded 11 known false positives, 6
-confirmed planets, and 8 candidate planets (see Appendix~\ref{app:members}).
-To determine whether these objects were
-potentially consistent with being {\it i)} planets, and {\it ii)}
-$\lesssim 10^8$ years old, we inspected the Kepler data validation
-reports and Robovetter classifications.  Youth was assessed based on
-the presence of rotational modulation at the expected period and
-amplitude for stars at least as young as the Pleiades \citep[{e.g.},][]{rebull_rotation_2020}. 
-Planetary status was assessed through the Robovetter flags, and by
-requiring non-grazing transits with ${\rm S/N}>10$.
-Four objects passed both cuts: Kepler-1627, Kepler-1643, KOI-7368, and
-KOI-7913.
-% FIXME: if you need to add lines, "One object (Kepler-1933) was ambiguous,
-
-Figure~\ref{fig:XYZvtang} shows the positions of these KOIs along
-various projections.  Kepler-1643 is near the core RSG-5 population
-both spatially and kinematically.  KOI-7368 and KOI-7913 are in a
-diffuse region $\approx$40\,pc above RSG-5 in $Z$ and $\approx$100\,pc
-closer to the Sun in $Y$.  In tangential galactic velocity space,
-there is some kinematic overlap between the region containing the latter two KOIs
-and the main RSG-5 group.
-
-%
-% Numbers are from RSG-5_auto_XYZ_vl_vb_cut.csv,
-% CH-2_auto_XYZ_vl_vb_cut.csv
-%
-We define two sets of stars in the local vicinity of our objects of
-interest.  For candidate RSG-5 members near Kepler-1643, we require:
-\begin{align}
-  X/{\rm pc} &\in [45, 75] \nonumber \\
-  Y/{\rm pc} &\in [320, 350] \nonumber \\
-  Z/{\rm pc} &\in [40, 80] \nonumber \\
-  v_b/{\rm km\,s^{-1}} &\in [-4, -2.5] \nonumber \\
-  v_{l^*}/{\rm km\,s^{-1}} &\in [3.5, 6], \nonumber
-\end{align}
-though RSG-5 does have a greater spatial extent toward smaller
-$X$ (Figure~\ref{fig:XYZvtang}, middle panels).  For the diffuse stars
-near KOI-7368 and KOI-7913, we require
-\begin{align}
-  X/{\rm pc} &\in [20, 70] \nonumber \\
-  Y/{\rm pc} &\in [230, 270] \nonumber \\
-  Z/{\rm pc} &\in [75, 105] \nonumber \\
-  v_b/{\rm km\,s^{-1}} &\in [-3.5, -1.5] \nonumber \\
-  v_{l^*}/{\rm km\,s^{-1}} &\in [2, 6] \nonumber
-\end{align}
-and we call this latter set of stars ``CH-2'', using the preliminary 
-Cep-Her (CH) subgroup identifier from R.~Kerr et al.\ (in prep).
-These cuts yielded
-\nrsgfive\ candidate RSG-5 members, and \nchtwo\ candidate CH-2
-members.  These stars are listed in Appendix~\ref{app:members}, as is the set of Cep-Her
-candidates that was observed by Kepler.
-%An important
-%consideration, especially for CH-2, is the contamination rate by field
-%stars.  We assess this in the following section.
-
-\subsection{The Cluster's Age}
-\label{sec:clusterage}
+In the original manuscript, the top-right panel of Figure~2
+erroneously omitted KOI-7913~B due to an error in the plotting script.
+Figure~\ref{fig:age} in this erratum corrects the omission.
 
@@
 			\includegraphics[width=0.49\textwidth]{f2a.pdf}
-			\includegraphics[width=0.49\textwidth]{f2b.pdf}
+			\includegraphics[width=0.49\textwidth]{f2b_corrected.pdf}
 		}
@@
     in RSG-5 and CH-2, respectively.  The range of colors is truncated
-    to emphasize the pre-main-sequence; approximate spectral types are shown on the upper axes.  Stars that fall far below the
-    cluster sequences are field interlopers.  {\it Bottom row}:
-    TESS and ZTF-derived stellar rotation periods, with the Pleiades
-    ($\approx 112$\,Myr) and Praesepe ($\approx 650$\,Myr) shown for
-    reference \citep{rebull_rotation_2016a,douglas_poking_2017}.  The
-    detection efficiency for reliable rotation periods falls off
-    beyond $\bpmrpo \gtrsim 2.6$.
+    to emphasize the pre-main-sequence; approximate spectral types are
+    shown on the upper axes.  Stars that fall far below the cluster
+    sequences are field interlopers.  {\it Bottom row}: TESS and
+    ZTF-derived stellar rotation periods, with the Pleiades ($\approx
+    112$\,Myr) and Praesepe ($\approx 650$\,Myr) shown for reference
+    \citep{rebull_rotation_2016a,douglas_poking_2017}.  The detection
+    efficiency for reliable rotation periods falls off beyond $\bpmrpo
+    \gtrsim 2.6$.
 	\label{fig:age}
@@
 \end{figure*}
-
-\subsubsection{Color--Absolute Magnitude Diagram}
-\label{sec:camd}
-
-Color--absolute magnitude diagrams (CAMDs) of the candidate RSG-5 and
-CH-2 members are shown in the upper row of Figure~\ref{fig:age}.  The
-stars from the $\delta$~Lyr cluster are from
-\citet{bouma_kep1627_2022}, and the field stars are from the Gaia EDR3
-Catalog of Nearby Stars \citep{gaia_gcns_2021}.  To make these
-diagrams, we imposed the data filtering criteria from
-\citet[][Appendix~B]{GaiaCollaboration2018}, which include binaries
-while omitting artifacts from for instance low
-photometric signal to noise, or a small number of visibility periods.
-We then corrected for extinction using the
-\citet{lallement_threedimensional_2018} dust
-maps and the extinction
-coefficients $k_X\equiv A_X/A_0$ from \citet{GaiaCollaboration2018},
-assuming that $A_0 = 3.1 E(B-V)$.  This yielded a mean and standard
-deviation for the reddening of $E(B-V)=0.036\pm0.002$ for RSG-5, and
-$E(B-V)=0.017\pm0.001$ for CH-2.  By way of comparison, in
-\citet{bouma_kep1627_2022} the same query for the $\delta$~Lyr cluster
-yielded $E(B-V)=0.032\pm0.006$.  Finally, for the plots we set the
-color axis to best visualize the region of maximal age information
-content: the pre-main-sequence.
-
-The CAMDs show that for RSG-5, all but one of the candidate
-members are on a tight pre-main-sequence locus.  Quantitatively,
-88/89 stars with $\bpmrpo \geq 1.5$ are consistent with being on the
-pre-main-sequence. This implies a false
-positive rate of a few percent, at most.  In comparison, our
-reference
-sample (the $\delta$~Lyr candidates) has a false positive rate of
-$\approx$12\%, based on the number of stars that photometrically
-overlap with the field population.  For CH-2, our membership selection gives 27
-objects in the color range displayed, and 23 of them appear to be
-consistent with being on the pre-main-sequence.  This
-would imply a false
-positive rate in CH-2 of $\approx$15\%.
-
-Figure~\ref{fig:age} also shows that most RSG-5 and CH-2 members
-overlap with the $\delta$~Lyr cluster on the CAMD, and that the groups are
-therefore roughly the same age.  To quantify this, we use the method
-introduced by \citet[][their Section~6.3]{gagne_mutau_2020}.  The idea
-is to fit the pre-main-sequence loci of a set of reference clusters,
-and to then model the locus of the target cluster as a linear
-combination of these reference cluster loci.  For our reference
-clusters, we used UCL, IC\,2602, and the Pleiades, with the
-memberships reported by \citet{Damiani2019} and
-\cite{CantatGaudin2018a} respectively.  We adopted ages of 16\,Myr for
-UCL \citep{pecaut_star_2016}, 38\,Myr for IC\,2602
-\citep{david_ages_2015,randich_gaiaeso_2018} and 112\,Myr for the
-Pleiades \citep{dahm_2015}.  These assumptions and the subsequent
-processing steps taken to exclude field stars and binaries were
-identical to those described in \citet{bouma_kep1627_2022}.  The mean
-and uncertainty of the resulting age posterior are $46^{+9}_{-7}$\,Myr
-for RSG-5, and $36^{+10}_{-8}$\,Myr for CH-2.  For comparison, this
-procedure yields an age for the $\delta$~Lyr cluster of
-$38^{+6}_{-5}$\,Myr.  The older isochronal age of RSG-5 is consistent
-with its location relative to the $\delta$~Lyr cluster in the upper
-left panel of Figure~\ref{fig:age}.
-Generally speaking, this method is expected to be accurate provided
-that the metallicities of IC\,2602 and the Cep-Her groups (RSG-5,
-CH-2, and the $\delta$~Lyr cluster) are roughly identical.  The
-spectroscopic metallicities that we find in Section~\ref{sec:stars}
-suggest that this is indeed the case.
-While in reality stellar populations do not evolve linearly in
-the dimensions of absolute magnitude versus color, in our case the
-Cep-Her loci are nearly indistinguishable from IC\,2602 (e.g.,
-Figure~3 of \citealt{bouma_kep1627_2022}).  Systematic errors incurred
-in the age from the non-linear evolution are therefore likely much
-smaller than the $\approx$10\,Myr systematic uncertainty in the
-absolute reference age for IC\,2602 itself
-\citep{david_ages_2015,randich_gaiaeso_2018}.
-
-\subsubsection{Stellar Rotation Periods}
-\label{sec:rotation}
-
-An independent way to assess the age of the candidate cluster members
-is to measure their stellar rotation periods.  This approach can be
-achieved using surveys such as TESS \citep{ricker_transiting_2015} and
-the Zwicky Transient Facility (ZTF, \citealt{bellm_zwicky_2019}); it
-leverages a storied tradition of measuring rotation periods of stars
-in benchmark open clusters \citep[see
-{e.g.},][]{skumanich_time_1972,curtis_rup147_2020}.  The TESS data in
-our case are especially useful, since they provide 3 to 5 lunar months
-of photometry for all of our candidate CH-2 and RSG-5 members.
-
-We selected stars suitable for gyrochronology by requiring 
-$\bpmrpo \geq 0.6$ to focus on FGKM stars that experience magnetic
-braking.
-For TESS, we also restricted our sample to $G<16$, to ensure the stars
-are bright enough to extract usable light curves from the full-frame
-images.  The magnitude cut corresponds to $\bpmrpo < 2.6$
-($\sim$M3V) at the relevant distances.  These cuts gave 19 stars in
-CH-2 and 42 stars in RSG-5.  We extracted light curves from the TESS
-images using the \texttt{unpopular} package \citep{hattorio_2021_cpm},
-and regressed them against systematics with its causal pixel model.
-We measured rotation periods using Lomb-Scargle periodograms and
-visually vetted the results using an interactive program that allows
-us to switch between TESS Cycles, select particular sectors, flag
-stars with multiple periods, and correct half-period harmonics. For
-ZTF, we used the same color cut to focus on FGKM stars, but restricted
-the sample to $13 < G < 18$ to avoid the saturation limit on the
-bright end and ensure sufficient photometric precision at the faint
-end. We followed the procedure outlined in \citet{curtis_rup147_2020}:
-we downloaded $8'\times8'$ image cutouts, ran aperture photometry for
-the target and neighboring stars identified with Gaia, and used them
-to define a systematics correction to refine the target light curves.
-
-The lower panels of Figure~\ref{fig:age} show the results.  In RSG-5,
-39/42 stars have rotation periods at least as fast as the Pleiades (93\%).  This
-numerator omits the two stars with periods $>$$12$\,days visible in
-the lower-left panel of Figure~\ref{fig:age}.  The age interpretation
-for these latter stars, particularly the $\approx$M2.5 dwarf, is not
-obvious.  \citet{rebull_usco_2018} for instance have found numerous
-M-dwarfs with 10-12 day rotation periods at ages of USco
-($\sim$$8$\,Myr), and some  still exist at the age of
-LCC ($\sim$$16$\,Myr; \citealt{2022AJ....164...80R}).  Regardless, since
-only one field star
-outlier seems to be present on the RSG-5 CAMD,
-the fact
-that we do not detect rotation periods for $\approx$7\%
-of stars should perhaps be taken as an indication for the fraction of
-stars for which rotation periods might not be detectable, due to
-{e.g.}, pole-on stars having lower amplitude starspot modulation.
-Field star contamination is another possible contributor.
-
-For CH-2, 13/19 stars have rotation periods that are obviously faster
-than their counterparts in the Pleiades.  4 stars, not included
-in the preceding numerator, are M-dwarfs with rotation periods between
-10 and 12.5 days.  As previously noted, the age interpretation for
-these M-dwarfs is ambiguous.  If none are cluster members,
-the rotation period detection fraction is 68\%; if all are members, it
-is 89\%.
-
-This sets an upper bound on the contamination fraction in our candidate
-CH-2 members at about one in three.  Combined with the roughly one in
-six contaminant rate implied by the earlier CAMD analysis, this
-suggests that the sample of candidate CH-2 members is more polluted by
-field stars than the RSG-5 sample.
-
-It is challenging to convert these stellar rotation periods to a
-precise age estimate, since on the pre-main-sequence the stars are
-spinning up due to thermal contraction rather than down due to
-magnetized braking.  Regardless, the rotation period distributions of
-both CH-2 and RSG-5 seem consistent with other 30\,Myr to 50\,Myr
-clusters ({e.g.}, IC\,2602 and IC\,2391;
-\citealt{douglas_stephanie_t_2021_5131306}).  They also seem
-consistent with the false positive rates estimated from the
-color--absolute magnitude diagrams.
-
-\section{The Stars}
-\label{sec:stars}
-
-\input{tab_sysparams.tex}
-
-Many of the salient properties of the Kepler objects of interest in Cep-Her can be
-gleaned from Figure~\ref{fig:age}.  The stars span spectral types of
-G8V (Kepler-1627) to K6V (KOI-7913 A).  The secondary in the KOI-7913
-system has spectral type $\approx$K8V.  And since a star with
-Solar mass and metallicity arrives at the zero-age main sequence at
-$t\approx40$ Myr \citep{choi_mesa_2016}, these stars are all in the
-late stages of their pre-main-sequence contraction.
-
-The adopted stellar parameters are listed in
-Table~\ref{tab:sysparams}.  The stellar surface gravity, radius, mass,
-and density are found by interpolating against the MIST isochrones
-in reddening-corrected absolute $G$-band magnitude as a function of $\bpmrpo$\ color
-\citep{choi_mesa_2016}.  The statistical uncertainties from this
-technique mostly originate from the parallax uncertainties; the
-systematic uncertainties are taken to be the absolute difference
-between the PARSEC \citep{bressan_parsec_2012} and MIST isochrones.
-Reported uncertainties are a quadrature sum of the statistical and
-systematic components.
-
-To verify these parameters, determine the stellar effective temperatures, and to analyze youth
-proxies such as the Li 6708\,\AA\ doublet and H$\alpha$, we acquired
-high resolution optical  spectra.  We also acquired high resolution imaging for each system, to
-constrain the existence of visual companions, including possible bound
-binaries.  We give the system-by-system details in Sections~\ref{subsec:kep1643}
-through~\ref{subsec:koi7913}, and summarize their implications for
-the youth of the stars in Section~\ref{subsec:specyouth}.
-
-\subsection{Kepler\,1643}
-\label{subsec:kep1643}
-
-\paragraph{Spectra}
-For Kepler-1643, we acquired two iodine-free spectra from Keck/HIRES
-on the nights of 2020 Aug 16 and 2021 Oct 25.  The acquisition and
-analysis followed the usual techniques of the California Planet Survey
-\citep{howard_cps_2010}.  We derived the stellar parameters ($T_{\rm
-eff}, \log g, R_\star$) using \texttt{SpecMatch-Emp}
-\citep{yee_SM_2017}, which yielded values in $<$$1$-$\sigma$ agreement
-with those from the cluster-isochrone method.  This approach also
-yielded $[{\rm Fe/H}]=0.13 \pm 0.09$.  Using the broadened synthetic
-templates\footnote{The broadening is calculated using the joint
-rotational and macroturbulent broadening kernel from \citet{hirano_2011},
-assuming that the macroturbulent velocity scales with effective temperature similar
-to the prescription from \citet{2014MNRAS.444.3592D}.  The latter assumption could be a 
-source of systematic uncertainty in our equatorial velocity measurements, since
-the macroturbulent velocity could be systematically higher (or lower) on
-the pre-main-sequence than it is for more slowly rotating field stars.}
- from \texttt{SpecMatch-Synth} \citep{petigura_cksi_2017}, we
-found $v\sin i = 9.3 \pm 1.0\,\mkms$.  The systemic radial velocity at
-the two epochs was $-9.1 \pm 1.9\,\mkms$ and $-7.8\pm 1.2\,\mkms$
-respectively, and was calculated
-following the methods of \citet{chubak_2012}.  To infer the equivalent width of
-the \ion{Li}{1} 6708\,\AA\ doublet, we followed the procedure
-described by \citet{bouma_2021_ngc2516}. In brief, this
-involved calculating the line width by numerically integrating a
-single best-fit Gaussian over a local window, and estimating the
-uncertainties through a Monte Carlo procedure in which the continuum
-normalization was allowed to vary through a bootstrap approach based
-on the local scatter in the spectra.  For Kepler-1643, this yielded a
-strong detection: ${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA, with
-values consistent at $<$$1$-$\sigma$ between the two epochs.   The
-quoted value does not correct for the \ion{Fe}{1} blend at
-6707.44\,\AA.  Given the purported age and effective temperature of
-the star, the lithium equivalent width is somewhat low.  We discuss
-this in greater depth in Section~\ref{subsec:specyouth}.
-
-\paragraph{High-Resolution Imaging}
-We acquired adaptive optics imaging of Kepler-1643 on the night of
-2019 June 28 using the NIRC2 imager on Keck-II.  Using the narrow
-camera (FOV = 10.2\arcsec), we obtained 4 images in the $K'$ filter
-($\lambda = 2.12\,\mu$m) with a total exposure time of 320\,s. 
-The images did not show any additional visual companions.
-We analyzed the data following \citet{kraus_impact_2016}, and
-determined the detection limits by analyzing the residuals after
-subtracting an empirical PSF template.  This procedure yielded
-contrast limits of $\Delta K' = 4.1$ mag at $\rho = 150$ mas, $\Delta
-K' = 5.8$ mag at $\rho = 300$ mas, and $\Delta K' = 8.3$ mag
-at $\rho > 1000$ mas.
-
-\subsection{KOI-7368}
-\paragraph{Spectra}
-For KOI-7368, we acquired a spectrum on 2015 June 1 using the echelle
-spectrograph (TRES; \citealt{furesz_tres_2008}) mounted at the
-Tillinghast 1.5m at the Fred Lawrence Whipple Observatory.  The
-Stellar Parameter Classification pipeline for TRES has been described
-by \citet{2021tsc2.confE.124B}.  It is based on the synthetic template
-library constructed by \citet{Buchhave2012}.  The resulting stellar
-parameters ($T_{\rm eff}, \log g, R_\star$) agreed with those from the
-cluster-isochrone method within $1$-$\sigma$.  Auxiliary
-spectroscopic parameters included the metallicity $[{\rm Fe/H}]= -0.02
-\pm 0.08$, the equatorial velocity $v\sin i = 20.2 \pm 1.0\,\mkms$,
-and the systemic velocity ${\rm RV}_{\rm sys} = -10.9 \pm 0.2\,\mkms$.
-The Li 6708\AA\ EW measurement procedure yielded ${\rm EW}_{\rm Li} =
-236^{+16}_{-14}$\,m\AA.
-
-\paragraph{High-Resolution Imaging}
-We acquired adaptive optics imaging of KOI-7368 on the night of 2019
-June 12, again using NIRC2.  The observational configuration and
-reduction were identical as for Kepler-1643.  No companions were
-detected, and the analysis of the image residuals yielded contrast
-limits of $\Delta K' = 5.2$ mag at $\rho = 150$ mas, $\Delta K' = 6.7$
-mag at $\rho = 300$ mas, and $\Delta K' = 8.7$ mag at $\rho > 1000$
-mas.
-
-\subsection{KOI-7913}
-\label{subsec:koi7913}
-
-\paragraph{Binarity}
-KOI-7913 is a binary.  The north-west primary is
-$\approx$0.5 magnitudes brighter than the south-east secondary in
-optical passbands.  The two stars are separated in Gaia EDR3 by
-$3\farcs5$ on-sky, and have parallaxes consistent within $1$-$\sigma$
-(with an average $\varpi=3.66 \pm 0.01$\,mas).  The apparent on-sky
-separation is $959 \pm 2$ au.  The Gaia EDR3 proper motions are also
-very similar.  Since the two stars were resolved in the Kepler Input
-Catalog and are roughly one Kepler pixel apart, an accurate crowding
-metric has already been applied in the NASA Ames data products to
-correct the mean flux level \citep{2017ksci.rept....6M}.  This is
-important for deriving accurate transit depths.
-
-\paragraph{Spectra}
-We acquired Keck/HIRES spectra for KOI-7913 A on the night of 2021 Nov
-13, and KOI-7913 B on the night of 2021 Oct 26.  The
-\texttt{SpecMatch-Emp} machinery yielded $T_{\rm eff,A} = 4324 \pm
-70\,{\rm K}$, and $T_{\rm eff,B} = 4038 \pm 70\,{\rm K}$.  These temperatures as well as the other spectroscopic parameters agreed with those from the cluster
-isochrone method within 1-$\sigma$.  For the primary, we also found $[{\rm Fe/H}]= -0.06 \pm
-0.09$, $v\sin i = 13.3 \pm 1.0\,\mkms$, and ${\rm RV}_{\rm sys} =
--17.8 \pm 1.1\,\mkms$.  For the secondary, these same parameters were
-$[{\rm Fe/H}]= -0.01 \pm 0.09$, $v\sin i = 10.7 \pm 1.0\,\mkms$, and
-${\rm RV}_{\rm sys} = -18.8 \pm 1.1\,\mkms$.  The primary showed lithium in absorption
-with ${\rm EW}_{\rm Li} =
-65^{+8}_{-6}$\,m\AA, while the secondary had a marginal detection of ${\rm EW}_{\rm Li} =
-42^{+12}_{-19}$\,m\AA. .  Both components displayed 
-H$\alpha$ in emission.  Given the spectral types of the stars, these observations
-are consistent with a $\approx$40 Myr age for KOI-7913 (see
-Section~\ref{subsec:specyouth}).
-
-\paragraph{High-Resolution Imaging}
-We acquired adaptive optics imaging of KOI-7913 on the night of 2020
-Aug 27 using the NIRC2 imager.  The observational configuration and
-reduction were identical as before.  The images showed KOI-7913 A,
-KOI-7913 B, and an additional faint neighbor $\approx$$0\farcs99$ due East
-of KOI-7913 B.  Applying the PSF-fitting routines from
-\citet{kraus_impact_2016}, the tertiary object has a separation $\rho =
-4397 \pm 3$\,mas from the primary, at a position angle $231.17^\circ
-\pm 0.02^\circ$, with $\Delta K' = 6.97 \pm 0.04$.  While it is too
-faint to affect the interpretation of the transit signal, it would be
-amusing if this faint neighbor were comoving and therefore part of the system, 
-because it would have a mass between 10 and
-15\,M$_{\rm Jup}$ at an assumed age of 40\,Myr.  Additional imaging epochs will tell.
-
-\subsection{Spectroscopic Youth Indicators}
-\label{subsec:specyouth}
-
-\begin{figure*}[t]
-	\begin{center}
-		\leavevmode
-			\includegraphics[width=0.9\textwidth]{f5.pdf}
-	\end{center}
-	\vspace{-0.5cm}
-	\caption{
-    {\bf Spectroscopic youth diagnostics for Kepler-1627, KOI-7368,
-    Kepler-1643, and KOI-7913 AB. }
-    The spectra are shown in the observed frame, and the stars are
-    sorted left-to-right in order of decreasing effective temperature.
-    \label{fig:koiyouthindicators}
-	}
-\end{figure*}
-
-Figure~\ref{fig:koiyouthindicators} shows key portions of the HIRES
-and TRES spectra for the Kepler objects in Cep-Her.  Lithium
-absorption is obvious at 6708\AA\ in all stars except KOI-7913 B.  
-H$\alpha$ is in emission for both components of
-KOI-7913, and in absorption for the hotter stars.  Here, we compare
-these observations against benchmark open clusters in order to
-assess their implications for the stellar ages.
-
-\subsubsection{Lithium}
-
-\begin{figure}[tp]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.47\textwidth]{f6a.pdf}
-		}
-
-		\vspace{-0.35cm}
-		\subfloat{
-			\includegraphics[width=0.47\textwidth]{f6b.pdf}
-		}
-
-		\vspace{-0.35cm}
-		\subfloat{
-			\includegraphics[width=0.47\textwidth]{f6c.pdf}
-		}
-	\end{center}
-	\vspace{-0.6cm}
-	\caption{
-    {\bf Lithium 6708\AA\ and H$\alpha$ equivalent widths for the
-    objects of interest compared to young open clusters and field
-    stars. } Positive equivalent width means absorption; negative
-    equivalent width means emission.  
-    {\it Top \& middle}:
-    The field stars are KOIs from
-    \citet{berger_identifying_2018}.  The ``40-50 Myr'' reference
-    stars ({\it left}) are from IC\,2602 \citep{randich_2001} and
-    Tuc-Hor \citep{kraus_stellar_2014}.  The ``112 Myr'' stars
-    are from the Pleiades
-    \citep{soderblom_evolution_1993,jones_evolution_1996,bouvier_pleiades_lirot_2018}.
-    The statistical uncertainties on
-    the equivalent widths are shown, or else are
-    smaller than the markers.
-    {\it Bottom}:
-    The H$\alpha$ comparison is against Tuc-Hor
-    \citep[$\approx$$40$\,Myr;][]{kraus_stellar_2014} and the Pleiades \citep{fang_2018}.
-    \label{fig:lithium}
-	}
-\end{figure}
-
-Figure~\ref{fig:lithium} compares the measured lithium equivalent
-widths of the Kepler objects against a few reference populations.  We
-selected reference studies from the literature only when upper limits were
-explicitly reported.  KOI-7368 and KOI-7913~A have secure lithium
-detections, while for KOI-7913~B the detection is marginal (${\rm EW}_{\rm
-Li} = 42^{+12}_{-19}$\,m\AA).  For all three stars, as well as for
-Kepler-1627~A, the observed lithium equivalent width is consistent
-with the stellar effective temperatures and a $\approx40$\,Myr age.
-
-Kepler-1643, in RSG-5, is conspicuously below the 40-50$\,$Myr
-sequence in the top panel of Figure~\ref{fig:lithium}, though
-above the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$\,m\AA).
-
-Quantitatively, there are 14 reference stars within $\pm$150\,K of Kepler-1643.
-The mean and standard deviation of their lithium EWs is $255 \pm 31$\,m\AA,
-which implies that Kepler-1643 is $4.0$-$\sigma$ discrepant from expectations.
-
-The middle panel shows a comparison against the Pleiades, where
-Kepler-1643 is more consistent with the observed dispersion in
-lithium.
-
-One explanation for the low Li equivalent width
-in Kepler-1643 relative to the comparison stars could be that it is a field
-interloper; another could be that RSG-5 is much older than 50\,Myr.  We do
-not favor either explanation.  
-RSG-5 cannot be much older than 50\,Myr
-based on its proximity to the $\delta$~Lyr cluster and IC\,2602 in the
-CAMD, and because it is below the Pleiades in the 
-rotation versus color diagram (Figure~\ref{fig:age}).
-Kepler-1643 also seems highly unlikely to be a field interloper, 
-because we demonstrated a few-percent false positive probability
-in our spatio-kinematic selection of RSG-5 members, and there is a 
-similar independent chance
-($\approx$$1\%$) of a field K2V star having a rotation period below the
-Pleiades \citep{mcquillan_rotation_2014}.
-This yields a puzzle: how could a star have spatial, kinematic, and rotational
-evidence consistent with being in a $\approx$50\,Myr cluster,
-but a low lithium content?
-
-Our preferred explanation for Kepler-1643's meager lithium 
-content is that the reference samples of IC\,2602 and Tuc-Hor
-stars may not fully explore all possible lithium equivalent widths at this age.
-This would be somewhat surprising since over a dozen stars have
-already been analyzed in the relevant effective temperature range.
-However, considering
-the top panels of Figure~\ref{fig:lithium}, it is also remarkable that in
-50 million years, stars between $4500$\,K and $5200$\,K go from
-having a tight lithium sequence to one with a dispersion
-$\approx$$10\times$ greater.  The existence of the Li dispersion in
-Pleiades-age K-dwarfs has been known for decades; it has also been
-known that the stars with the largest lithium abundances are also the
-most rapidly rotating
-\citep{butler_pleiades_1987,soderblom_evolution_1993}.  More recent
-analyses of this correlation have been reviewed by
-\citet{bouvier_lithium-rotation_2020}.  The conclusion of that work
-was that the origin of the rotation-lithium correlation likely lies
-within pre-main-sequence stellar physics.  If so, one would expect the
-IC\,2602 and Tuc-Hor K-dwarfs to show a larger intrinsic lithium
-dispersion.  A recent analysis of the $\approx$40\,Myr
-NGC~2547 by \citet{binks_2022} suggests that this may
-be the case, though that study only had $\approx$10 stars in the relevant
-effective temperature range.
-An alternative explanation could be that the overall metallicity of Cep-Her
-is different from Tuc-Hor and IC\,2602, but this seems unlikely given the near-solar
-metallicities we have measured for the Kepler Objects of Interest.
-Broadly, these considerations suggest that Cep-Her is
-a worthy object for further spectroscopic analyses of lithium near the
-zero-age main sequence.
-
-\subsubsection{H$\alpha$}
-
-As shown in Figure~\ref{fig:koiyouthindicators}, H$\alpha$ is in
-emission for both components of KOI-7913, and in absorption for the
-hotter stars.  Additionally, the emission appears double-peaked for
-both of the KOI-7913 components.  An important note is that KOI-7913~A and
-KOI-7913~B were spatially resolved from each other during data
-acquisition.  Performing a cross-correlation between each of the stars
-and the nearest matches in the Keck/HIRES template library, we also
-found that the CCFs for both components of KOI-7913 showed no
-indications of double-lined binarity \citep{kolbl_detection_2015}.
-
-Balmer line emission, particularly in H$\alpha$, is expected for
-low-mass stars of this age.  \citet{kraus_stellar_2014} for instance, in their
-survey of Tuc-Hor ($\approx$40\,Myr), observed that all
-cluster members with spectral types $>$${\rm K4.5V}$ had H$\alpha$ in
-emission.  This is consistent with our observations: KOI-7913 shows
-H$\alpha$ in emission for both components, and in absorption
-for all of our other Kepler objects (Figure~\ref{fig:koiyouthindicators}, lower panel).  The double-peaked nature of the
-emission, though not always present, is also common for
-active stars.  Proxima Centauri, for instance, has double-peaked
-H$\alpha$ emission \citep{collins_calculations_2017}.  Given
-that we have ruled out spectroscopic binarity, the most
-likely
-explanation is self-absorption: photons near the center of the line
-see a greater optical depth from higher layers of the chromosphere,
-while photons on the wings are too far from the
-rest wavelength to excite electrons and be re-absorbed in the upper
-layers.  The exact details of when a star's atmosphere reaches the
-conditions for such self-absorption require non-local thermal
-equilibrium models of the chromosphere
-\citep{short_chromospheric_1998,2005A&A...439.1137F}.
-
-\section{The Planets}
-\label{sec:planets}
-
-\begin{figure*}[tp]
-	\begin{center}
-		\leavevmode
-		\subfloat{
-			\includegraphics[width=0.85\textwidth]{f3a.pdf}
-		}
-
-		\vspace{-0.5cm}
-		\subfloat{
-			\includegraphics[width=0.33\textwidth]{f3b.pdf}
-			\includegraphics[width=0.33\textwidth]{f3c.pdf}
-		}
-
-		\vspace{-1.27cm}	
-		\subfloat{
-			\includegraphics[width=0.33\textwidth]{f3d.pdf}
-			\includegraphics[width=0.33\textwidth]{f3e.pdf}
-		}
-	\end{center}
-	\vspace{-0.5cm}
-	\caption{
-		{\bf Raw and processed light curves for the Kepler Objects of
-    Interest in Cep-Her.}   
-    {\it Top}: 50 day light curve segment from the 3.9 years of Kepler
-    data.  The ordinate shows the \texttt{PDCSAP} median-subtracted
-    flux in units of parts-per-thousand ($\times 10^{-3}$).  The
-    dominant signal is from starspots; planetary transit times
-    are indicated with vertical dashed lines, but the individual transits are not visible at this scale.
-    {\it Bottom}:
-    Phase-folded transits of Kepler-1643, KOI-7913, KOI-7368, and
-    Kepler-1627 with stellar variability removed.  The maximum a
-    posteriori model is shown with the gray line, and the
-    residual after subtracting the transit model is vertically
-    displaced.  Windows over 10 hours are shown.  Gray points are
-    individual flux measurements; black points are binned to 20 minute
-    intervals, and have a representative 1-$\sigma$ error bar in the
-    center-right of each panel. 
-		\label{fig:planets}
-	}
-\end{figure*}
-
-\subsection{Kepler Data}
-
-The Kepler space telescope observed Kepler-1643, KOI-7913, and
-KOI-7368 at a 30-minute cadence between May 2009 and April 2013.  For
-all three systems quarters 1 through 17 were observed with minimal
-data gaps.  The top panel of Figure~\ref{fig:planets} shows a 50-day
-slice of the \texttt{PDCSAP} light curves for the three new Cep-Her
-candidates, along with Kepler-1627.  In \texttt{PDCSAP},
-non-astrophysical variability is removed through a cotrending approach
-that uses a set of basis vectors derived by applying singular value
-decomposition to a set of systematics-dominated light curves
-\citep{smith_kepler_PDC_2017}.  In our analysis, we used the
-\texttt{PDCSAP} light curves with the default optimal aperture
-\citep{smith_finding_2016}.  Cadences with non-zero quality flags were
-omitted.  In all cases, the stars are dominated by spot-induced
-modulation with peak-to-peak variability between 2\% and 10\%.  These
-signals are much larger than the transits, which have depth
-$\approx$0.1\%.  To quantify the stellar rotation periods, we
-calculated the Lomb-Scargle periodogram for each Kepler quarter
-independently.  The resulting means and standard deviations are in
-Table~\ref{tab:sysparams}.
-
-\subsection{Transit and Stellar Variability Model}
-\label{sec:fitting}
-
-Our goals in fitting the Kepler light curves are twofold.  First, we
-want to derive accurate planetary sizes and orbital properties.
-Second, we want to remove the spot-induced variability signal to enable a
-statistical assessment of the probability that the transit signals are
-planetary.
-
-We fitted the data as follows.  Given the transit
-ephemeris from \citet{thompson_planetary_2018}, we first trimmed the
-light curve to a local window around each transit that spanned three
-transit durations before and after each transit midpoint.  The
-out-of-transit points in each local window were then fitted with a
-fourth-order polynomial, which was divided out from the light curve.
-The resulting flattened transits were then fitted with a transit model
-that assumed quadratic limb darkening.  The model therefore included 8
-free parameters for the transit ($\{P, t_0, \log R_{\rm p}/R_\star, b,
-u_1 ,u_2, R_\star, \log g\}$), 2 free parameters for the light curve
-normalization and a white noise jitter ($\{\langle f \rangle, \sigma_f
-\}$), and 5 fixed parameters for each transit.
-
-We fitted the data using \texttt{exoplanet}
-\citep{exoplanet:exoplanet}.  We assumed a Gaussian likelihood, and
-sampled using \texttt{PyMC3}'s No-U-Turn Sampler
-\citep{hoffman_no-u-turn_2014}, after having initialized to the the
-maximum a posteriori (MAP) model.  We used the
-\citet{gelman_inference_1992} statistic, $\hat{R}$, as our convergence
-diagnostic.  The resulting fits are shown in the lower panels of
-Figure~\ref{fig:planets}, and the important derived parameters are in
-Table~\ref{tab:sysparams}.  The set of full parameters and their
-priors are given in Appendix~\ref{app:transit}.
-
-A potential drawback of our approach is that to remove
-the starspot-induced variability, we fixed 5 
-parameters per transit to their MAP values.
-An alternative could be to fit the
-planetary transits simultaneously with the starspot-induced
-variability using a quasiperiodic Gaussian process (GP).  We explored
-this approach, but ultimately prefer our model for its simplicity, and for the benefit that the white noise jitter never
-trades off with any parameter equivalent to a damping timescale for
-the coherence of the GP.  It is also computationally efficient, and it
-captures the planetary parameters about which we care the most.
-
-\subsection{Planet Validation}
-\label{subsec:validation}
-
-In the future, it may be possible to obtain independent evidence for
-the planetary nature of the Cep-Her planets, for instance by observing
-spectroscopic transits.  For now,
-it is of interest whether the transit signals might be astrophysical
-false positives, or whether they are statistically more likely to be
-planetary.  We adopt the Bayesian framework implemented in
-\texttt{VESPA} to assess the relevant probabilities
-\citep{morton_efficient_2012,vespa_2015}.  Briefly summarized, the
-priors in \texttt{VESPA} assume the binary star occurrence rate from
-\citet{raghavan_survey_2010}, direction-specific star counts from
-\citet{girardi_star_2005}, and planet occurrence rates as described by
-\citet[][Section~3.4]{morton_efficient_2012}.  The likelihoods are
-then evaluated by forward-modeling a synthetic population of eclipsing
-bodies for each astrophysical model class, in which each population
-member has a known trapezoidal eclipse depth, total duration, and
-ingress duration.  These summary statistics are then compared against
-the actual photometric data to evaluate the probabilities of false
-positive scenarios such as foreground eclipsing binaries, hierarchical
-eclipsing binaries, and background eclipsing binaries.
-
-\paragraph{Kepler-1643}
-Kepler-1643~b (KOI-6186.01) was already validated as a transiting
-planet by \citet{morton_false_2016}, who found a probability for any
-of the aforementioned false positive scenarios of 9$\times$10$^{-6}$.
-Repeating the calculation with our own stellar-variability correction
-and the new NIRC2 imaging constraints, we find ${\rm FPP} =
-6\times10^{-9}$.  Figure~\ref{fig:planets} shows the justification:
-the transit is flat and has a high S/N ($\approx$$47$).  The shape is
-therefore nearly impossible to reproduce with eclipsing binary models.
-
-Intriguingly, Kepler-1643 failed one of the data validation
-centroid shift tests (see the \texttt{q1\_q17\_dr25\_koi} data
-release): the angular distance between the target star's KIC catalog
-position and the position of the transiting source was measured as
-$1\farcs0$ at 4.4-$\sigma$.  The reports show however that two
-outlying quarters (2 and 6) drive the offset --- the centroid locations
-from the other Kepler quarters are consistent at $\lesssim 0\farcs4$ (3-$\sigma$).
-\citet{bryson_2013} showed that for typical field star KOIs without
-centroid offsets, the mean offset distribution peaks at 0.3$''$ (their
-Figure 23).  By comparison, stars with centroid offsets that can be
-localized to nearby stars have a distribution that peaks at 7$''$ (their
-Figure 32).  The stellar
-variability in Kepler-1643 complicates the centroid-based
-vetting tests, because the shifts measured by these tests are
-determined from the in- and out-of-transit flux-weighted centroids.  For stars
-with significant spot-induced variability there is no static baseline in either
-the in- or out-of-transit phases, and so the centroid location may shift
-depending on the rotation phase combined with the local scene.  Based on
-these considerations, the centroid-level diagnostics for Kepler-1643 appear to
-be consistent with the transit signal being localized to the target star.
-
-\paragraph{KOI-7368}
-KOI-7368.01 is listed on the NASA Exoplanet Archive as a ``candidate''
-planet.  \citet{morton_false_2016} did not compute a false positive
-probability for the system because their default trapezoidal fitting
-routine failed, presumably due to the spot-induced variability.  Our
-fitting approach rectifies this point, and our new NIRC2 images
-revealed no new stellar companions.  Performing the relevant
-calculation, we find ${\rm FPP} = 4\times10^{-3}$.  Though not
-as convincing as Kepler-1643, this clears the threshold probability of 1 in 100 suggested by \citet{morton_false_2016} for calling a planet statistically validated.  The S/N of the transit is
-$\approx$$32$, which indicates that it is unlikely to be caused by
-systematic noise in the light curve (see Figure~\ref{fig:planets}).
-The positional probability\footnote{Columns $\textit{pp\_host\_rel\_prob}$ and $\textit{pp\_host\_prob\_score}$ on the KOI Positional Probabilities table at the NASA Exoplanet Archive \citep{Akeson13}.} calculated by \citet{2017ksci.rept...16B} also indicates that the transit signal shares its position with the target star.
-
-It bears mentioning that KOI-7368 shows a centroid shift in the
-\texttt{q1\_q17\_dr25\_koi} validation reports, similar to
-Kepler-1643.  For KOI-7368, the reported offset is smaller, and less
-formally significant ($0\farcs2$; 3.0-$\sigma$).  Again, the
-data validation reports show that the shift is caused by a few
-outlying quarters (4, 5, 8, and 12).  Since the remaining
-quarters show
-consistent scatter in their centroid locations, these outlying
-quarters are likely also caused by the stellar variability,
-because their directions are inconsistent across different quarters.
-Our NIRC2 imaging independently shows that there are no known
-neighboring sources that could cause an offset of the observed
-amplitude, as is also the case for Kepler-1643.
-
-\paragraph{KOI-7913}
-KOI-7913.01 is also currently listed on the NASA Exoplanet Archive as a
-``candidate'' planet.  The \citet{morton_false_2016}
-analysis was of Q1-Q17 KOIs from DR24, and therefore spanned KOI-1.01
-to KOI-7620.01 (omitting KOI-7913.01).  However the results of the
-subsequent DR25 analysis by Morton et al.\ are listed at the NASA
-Exoplanet Archive.  The relevant table gives a probability for the
-system being an astrophysical false positive of $1.4\times10^{-4}$,
-with the most likely false positive scenario being a blended eclipsing
-binary.  Repeating the calculation with our new detrending and
-NIRC2 contrast curves, we find a similar result: ${\rm FPP} =
-1.3\times10^{-4}$.  Though the transit has the lowest S/N of any of
-the objects discussed ($\approx$$14$), its low FPP  can be understood through its flat-bottomed shape,
-combined with its long transit duration relative to most
-eclipsing binary models (Figure~\ref{fig:planets}).
-The positional probability calculation performed by \citet{2017ksci.rept...16B} yielded a near-unity probability that the transit event is at the same location as the host star, and so the cumulative evidence suggests that KOI-7913~Ab is indeed a statistically validated planet.
-Its
-disposition has however previously fluctuated from ``false positive''
-to ``candidate'' (see Appendix~\ref{app:koi7913}).  The most likely
-explanation is the presence of KOI-7913~B, which is located
-$\approx0.9$ Kepler pixels away from Kepler-7913 A.  
-While the $\approx$1.5 pixel FWHM of the Kepler pixel response function
-implies that there is blending between the two stars,
-the target-pixel level data for KOI-7913 B reveals an
-entirely different stellar rotation
-period (Table~\ref{tab:sysparams}), and no hint of the transit signal.
-This implies that KOI-7913 B cannot host the planet.
-
-\section{Discussion \& Conclusion}
-\label{sec:disc_conc}
-
-\begin{figure*}[!t]
-	\begin{center}
-		\leavevmode
-		\includegraphics[width=0.93\textwidth]{f4.pdf}
-	\end{center}
-	\vspace{-0.6cm}
-	\caption{
-		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
-    Planets younger than a gigayear with ages more precise than a
-    factor of three are emphasized. The Cep-Her planets are
-    Kepler-1643~b ($\square$), KOI-7368~b ($\triangledown$),
-    KOI-7913~Ab (X), and Kepler-1627~Ab (+).  Interesting trends in
-    the population of planets younger than $10^8$ years old include {\it i)} their
-    large sizes and {\it ii)} the lack of hot Jupiters.  The new
-    objects of interest in Cep-Her have normal mini-Neptune sizes
-    between 2 and 3\,$R_\oplus$, which is a novelty given their ages.
-    Parameters are from the \citet{PSCompPars}.
-		\label{fig:rp_period_age}
-	}
-\end{figure*}
-
-\subsection{Normal-Sized Mini-Neptunes Exist at 40$\,$Myr}
-\label{subsec:sizes}
-
-The most significant novelty about the planets in Kepler-1643,
-KOI-7368, and KOI-7913 is that their sizes (2.2 to 2.3\,$R_\oplus$)
-are normal relative to the known population of mini-Neptunes from
-Kepler.  At field star ages, mini-Neptune sizes span 1.8\,$R_\oplus$
-to 3.6\,$R_\oplus$, with the most common size being $\approx
-2.4\,R_\oplus$
-\citep{Fulton_et_al_2017}.  The known planets younger than
-$10^8$ years are almost all larger, with sizes between 4 and
-10$\,R_\oplus$
-\citep{Mann_K2_33b_2016,David_et_al_2016,benatti_possibly_2019,david_four_2019,newton_tess_2019,rizzuto_tess_2020,bouma_cluster_2020,mann_toi1227_2022}.
-Figure~\ref{fig:rp_period_age} explores this by showing
-the sizes, orbital periods, and ages of the known transiting planets,
-emphasizing planets with precise ages.  The smallest previously known
-planets comparable to the new Cep-Her mini-Neptunes are AU~Mic~c
-($3.0\pm0.2\,R_\oplus$, see \citealt{martioli_aumicbc_2021} and
-\citealt{gilbert_flares_2022}), Kepler-1627~Ab ($3.8\pm0.2\,R_\oplus$;
-\citealt{bouma_kep1627_2022}), and AU~Mic~d ($4.2\pm0.2\,R_\oplus$;
-\citealt{plavchan_planet_2020}).
-
-The theoretical expectation is that mini-Neptunes with sizes of 2 to
-3\,$R_\oplus$ should be common at ages of $10^7$ to $10^8$ years.
-This expectation is tied to inferences about the initial distributions
-of planetary core mass, core composition, and atmospheric mass
-fraction \citep{owen_evaporation_2017}.  The Kelvin-Helmholtz cooling
-timescale, which is tied to the entropy of the planetary interior
-shortly after disk dispersal, also plays a significant role
-\citep{owen_constraining_2020}.  As an example,
-\citet{rogers_unveiling_2021} predicted that given a core mass
-distribution peaked at $\approx$$4$\,$M_\oplus$, an ice-poor rock/iron
-core composition, and a typical H/He mass fraction of $\approx$4\%,
-there should be a single local maximum in planet occurrence rates at 2 to 3\,$R_\oplus$,
-at times between 10 and 100\,Myr. In other words,
-\citet{rogers_unveiling_2021} predict the existence of a ``radius
-mountain'' at these early times, rather than a ``radius valley''.  The models advanced by
-\citet{gupta_signatures_2020} and \citet{lee_primordial_2021}
-agree that this local maximum should exist;
-their differences lie in the mechanism for producing the radius
-valley, and in whether a population of rocky planets is predicted to exist
-at the time of disk dispersal.
-
-Systems such as K2-25, V1298~Tau, HIP-67522, TOI-837, and TOI-1227
-have sizes that are anomalously large relative to the predicted peak
-in planet occurrence at 2 to 3\,$R_\oplus$.  However, their large
-sizes can be accommodated by invoking any of {\it i)} larger core
-masses, {\it ii)} more volatile-rich compositions, {\it iii)}  larger
-initial atmospheric mass fractions, or {\it iv)} longer thermal
-cooling times.  Secure mass measurements would help constrain this
-parameter space, but the $\sim$1\,\kms\ spot-induced radial velocity
-semi-amplitudes make measuring the Doppler orbits very difficult
-\citep[][]{cale_diving_2021,zicher_one_2022,klein_one_2022}.
-Regardless, the new Kepler-1643, KOI-7368, and KOI-7913 systems do
-demonstrate that at least some planets at 40\,Myr have sizes that are
-consistent with theoretical expectations for mini-Neptunes.  While
-selection effects imposed by spot-induced photometric variability are
-a likely explanation for why planets this small have not previously
-been identified \citep[{e.g.},][]{zhou_2021_tois}, future work should
-quantify this bias more carefully, in order to enable empirical
-studies of how the planetary size distribution changes at early times.
-
-\subsection{Is CH-2 a Coeval Population?}
-\label{subsec:ch2}
-
-RSG-5, and Kepler-1643's membership inside it, meet typical
-expectations for a star claimed to be in an open cluster.  RSG-5 is an
-obvious overdensity relative to the local field, and our membership
-selection easily yielded a clean pre-main-sequence locus
-(Figure~\ref{fig:age}).  CH-2, and KOI-7913 and KOI-7368's membership
-inside it, do not meet these expectations in as obvious a manner.
-This is because the CH-2 association is diffuse.
-
-To quantify the density difference between CH-2 and RSG-5, we can
-compare the spatial and velocity volumes searched for each group.  For
-RSG-5, we drew \nrsgfive\ candidate members from a $30\,{\rm
-pc}\times30\,{\rm pc}\times40\,{\rm pc}$ rectangular prism, given a
-$1.5 {\rm \mkms} \times 2.5 {\rm \mkms }$ rectangle in apparent
-galactic velocity.  For CH-2, our \nchtwo\ candidate members came from
-a rectangular prism of dimension $50\,{\rm pc}\times40\,{\rm
-pc}\times30\,{\rm pc}$, and a rectangular box of $2 {\rm \mkms} \times
-4 {\rm \mkms}$.  If we define the searched volume in units of ${\rm
-pc}^3\,{\rm km^2}\,{\rm s^{-2}}$, then the volume ratio of CH-2 to
-RSG-5 is 3.5 to 1.  The ratio of number densities (candidate members
-per unit searched volume) in RSG-5 relative to CH-2 is 16 to 1.
-
-Given its low density, is CH-2 truly a star cluster?  For this
-discussion, we adopt the definition that a star cluster is a group of
-at least 12 stars that was physically associated at its time of
-formation.  The ``12'' is set to distinguish star clusters from
-high-order multiples \citep[see][]{krumholz_star_2019}.  We explicitly
-do not require a ``star cluster'' to be gravitationally bound:
-dissolved clusters as well as their tidal tails are included in our
-adopted definition of ``clusters''.  We similarly do not require a
-threshold number of stars per unit spatial volume.  The latter point
-acknowledges that an important factor in cluster identification is
-also coherence in velocity space.
-For instance, the Psc-Eri stream, which has a shape that can be approximated as a 600 parsec-long cylinder with a radius of 30 parsecs, has a number density roughly a factor of three times lower than even CH-2 \citep{roser_psceri_2020}.
-However its existence is discernible because of the $\lesssim 2.5$\kms\ scatter in its cylindrical velocities.
-Perhaps once stellar rotation periods
-and chemical abundances reach the same level of ubiquity as stellar
-proper motions, they might enable further refinement in our ability to
-discover stars that formed as part of the same event.
-
-From a data-driven perspective, demonstrating that a group of stars
-was physically associated at its time of formation is challenging.
-While some young groups show kinematic evidence for expansion
-\citep{kuhn_kinematics_2019}, many, including Sco-Cen, do not
-\citep{wright_kinematics_2018}.  This complicates the feasibility of
-deriving kinematic ages through traceback, as well as through the
-expansion itself \citep[see][]{crundall_chronostar_2019}.  A more
-minimal approach is that suggested by \citet{tofflemire_2021}: search
-for coeval, phase-space neighbors, measure their ages, and determine
-if they share a common age.  This approach can demonstrate whether a
-star is currently associated with a set of coeval stars, though it
-falls short of determining what the association looked like in the
-past.  Our analysis of CH-2 meets the latter standard for
-demonstrating the existence of a $\approx$40\,Myr stellar association.
-
-It would be a worthy exercise to perform a similar
-search for coeval phase-space neighbors on the entire dataset of known
-exoplanet hosts.  For the time being, we can offer the anecdotal point
-that in our experience, most stars do not have dozens of 40\,Myr
-neighbors within a local volume of a few \!\kms\ and tens of parsecs.
-
-\subsection{Future work}
-
-\paragraph{Cep-Her}
-Our analysis to date has focused only on portions of Cep-Her that were
-observed by Kepler: RSG-5, CH-2, and the $\delta$~Lyr cluster.  In
-\citet{bouma_kep1627_2022} as well as this work, we have shown that
-these groups share similar ages, and have kinematic correlations that
-suggest a common origin.  With that said, the membership and
-kinematics of the other Cep-Her groups shown in
-Figure~\ref{fig:XYZvtang} deserve independent attention.  An important
-aspect of the remaining work will be to acquire radial velocities for a larger subset of the stars, and to determine whether the traceback approach
-could be applicable.  Wide-field spectroscopic surveys such as LAMOST
-\citep{zhao_2012_LAMOST} or SDSS-V \citep{kollmeier_2017} could enable
-such analyses for the brightest members, while also providing sensitivity to the Li 6708\,\AA\
-line.
-The Gaia DR3 RVS spectra (released during review of this manuscript) could contain similar velocity information down to spectral types of $\approx$K5V ($G_{\rm RVS}\lesssim 14$), and perhaps also enable analyses of the calcium infrared triplet as a youth indicator.
-The combination of more complete kinematics and youth indicators would help in definitively unraveling the formation
-history of the complex.
-
-A number of worthy photometric projects also seem possible given the
-new understanding of Cep-Her.  One is asteroseismology of the
-$\delta$~Sct stars, using either TESS or Kepler data
-\citep{bedding_very_2020}.  For cases in which the modes are resolved,
-this might yield age or metallicity estimates for the subgroups
-independent of other methods.  Other projects could include a more
-comprehensive analysis of the stellar rotation periods,
-searches of the Kepler light curves for exocomets
-\citep{zieba_transiting_2019}, and searches for missed planets around
-the most rapid rotators.
-
-\paragraph{Exoplanet demographics at early times}
-Our main motivation for finding new young planets is to help benchmark
-models for planetary evolution.  However demographic analyses of the
-known planets between $10^7$ and $10^9$ years have so far been rather
-limited.  Approximately 40 such planets are now known
-(Figure~\ref{fig:age}).  About half come from K2, a quarter from TESS,
-and now a quarter from Kepler.
-
-Given the current state of the field, a few reflections regarding
-experimental design of a demographic survey focused on planetary
-evolution over the first gigayear might be useful.  The first is that
-such a project requires a set of target stars with known ages.  A
-promising way to compile relevant stars could be to combine automated
-spatio-kinematic clustering from Gaia with rotation periods measured
-using TESS \citep[see the appendices of][]{bouma_kep1627_2022}.  The
-second consideration is that all the known young planets smaller than
-$3$\,$R_\oplus$ come from either K2 or Kepler.  Demographic inferences
-based on TESS are therefore limited to planetary sizes
-$\gtrsim4$\,$R_\oplus$, for planets close-in to their host stars.  It
-would be worthwhile to compare the occurrence rates of both types of
-planets with those from the main Kepler sample.  One specific question
-that seems within reach would be to clarify whether enough young stars
-have been searched for the dearth of young hot Jupiters to be
-significant.  Since the hot Jupiter occurrence rate is strongly
-dependent on stellar mass and metallicity
-\citep{petigura_metallicity_2018,petigura_cksX_2022}, particular care
-would be needed to select a sample of well-studied FGK dwarfs for the
-measurement, likely using stars in Sco OB2, Cep-Her, and Orion.  For
-demographic studies focused on how mini-Neptune sizes evolve, the
-combined K2 and Kepler dataset would be the better primary source.
-
-\subsection{Summary}
-
-We have shown that Kepler-1643~b, KOI-7368~b, and KOI-7913~Ab are 40
-to 50 million years old, and that each system is most likely
-planetary.  The evidence for the planetary interpretation comes from
-an application of \texttt{VESPA} to the Kepler data, alongside new
-imaging from NIRC2.  The validity of the \texttt{VESPA} framework
-rests on the premise that non-astrophysical false positives can be
-rejected.  
-This seems to be the case for all three objects, even though
-Kepler-1643 and KOI-7368 
-both show weak centroid offsets in specific quarters.  For both systems, the observed shifts
-are consistent with being caused by starspot-induced variability in
-specific quarters spuriously moving the stellar center-of-light.
-Independently, our imaging rules out companion stars with the
-brightnesses and positions that would be needed to explain the
-reported shifts.  All three objects are therefore most likely planets.
-
-Each system has multiple indicators of youth that support the reported
-ages.  For Kepler-1643, the strongest youth indicator is its physical
-and kinematic association with RSG-5.  Based on the color--absolute
-magnitude diagram, we are able to select members of this cluster with
-a false positive rate of a few percent (Figure~\ref{fig:age}).
-Kepler-1643 is one such member.  While the stellar rotation period
-period agrees with this assessment, the star's lithium equivalent
-width is marginally low, which might motivate future exploration of
-lithium depletion across FGKM stars in RSG-5 (see
-Section~\ref{subsec:specyouth}).
-
-The spatio-kinematic argument for the youth of KOI-7368 and KOI-7913
-is weaker because they are in an association of stars, CH-2, that is
-more diffuse.  For KOI-7913, stronger indicators of its age come from
-its binarity.  Both stellar components in KOI-7913 have isochronal ages
-consistent with $40$\,Myr.  Both components also show H$\alpha$ in
-emission, which for the transit-hosting $\approx$K6V primary is a strong indicator
-that the star is $\lesssim$100$\,$Myr old.  KOI-7368 is  more
-massive, and its Li 6708\,\AA\ measurement and stellar rotation period
-provide independent verification of the star's youth.
-
-The astrophysical implication of these considerations is that planets
-$\approx$2 Earth radii in size exist at ages of 40 million years.  It
-will be interesting to continue the push down to smaller planetary
-sizes at comparable ages -- the planetary detections we have presented
-are well above the average detection significance for Kepler planets.
-There may still be room at the bottom.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\clearpage
-\acknowledgements
-%\raggedbottom
-% Cycle 3 and 4
-L.G.B{.} is supported by the Heising-Simons Foundation 51
-Pegasi~b Fellowship and the NASA TESS GI Program (80NSSC21K0335
-and 80NSSC22K0298).
-R.K{.} is supported by the Heising-Simons Foundation.
-J.L.C{.} is supported by NSF AST-2009840 and the NASA TESS GI Program
-(80NSSC22K0299).
-D.H. is supported by the Alfred P. Sloan Foundation and NASA (80NSSC19K0597).
-%
-% ACKNOWLEDGE PFS / CAMPANAS.
-%
-%This paper includes data collected by the TESS mission, which are
-%publicly available from the Mikulski Archive for Space Telescopes
-%(MAST).
-%%
-%Funding for the TESS mission is provided by NASA's Science Mission
-%directorate.
-%%
-%We thank the TESS architects
-%%(G.~Ricker, R.~Vanderspek, D.~Latham, S.~Seager, J.~Jenkins)
-%and the many TESS team members for their
-%efforts to make the mission a continued success.
-%
-%
-%This study was based in part on observations at Cerro Tololo
-%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
-%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
-%Association of Universities for Research in Astronomy (AURA) under a
-%cooperative agreement with the National Science Foundation.
-%
-%
-%Finally, this research has made use of the Keck Observatory Archive (KOA),
-%which is operated by the W. M. Keck Observatory and the NASA Exoplanet
-%Science Institute (NExScI), under contract with the National
-%Aeronautics and Space Administration.
-%We also thank the Keck Observatory staff for their support of
-%HIRES and remote observing.  We recognize the importance that the
-%summit of Maunakea has within the indigenous Hawaiian community, and
-%are deeply grateful to have the opportunity to conduct observations
-%from this mountain.
-%
-% The Digitized Sky Survey was produced at the Space Telescope Science
-% Institute under U.S. Government grant NAG W-2166.
-% Figure~\ref{fig:scene} is based on photographic data obtained using
-% the Oschin Schmidt Telescope on Palomar Mountain.
-%
-
-% %
-% This research made use of the NASA Exoplanet Archive, which is
-% operated by the California Institute of Technology, under contract
-% with the National Aeronautics and Space Administration under the
-% Exoplanet Exploration Program.
-% %
-
-% Resources supporting this work were provided by the NASA High-End
-% Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
-% Division at Ames Research Center for the production of the SPOC data
-% products.
-%
-
-% A.J.\ and R.B.\ acknowledge support from project IC120009 ``Millennium
-% Institute of Astrophysics (MAS)'' of the Millenium Science Initiative,
-% Chilean Ministry of Economy. A.J.\ acknowledges additional support
-% from FONDECYT project 1171208.  J.I.V\ acknowledges support from
-% CONICYT-PFCHA/Doctorado Nacional-21191829.  R.B.\ acknowledges support
-% from FONDECYT Post-doctoral Fellowship Project 3180246.
-% %
-% C.T.\ and C.B\ acknowledge support from Australian Research Council
-% grants LE150100087, LE160100014, LE180100165, DP170103491 and
-% DP190103688.
-% %
-% C.Z.\ is supported by a Dunlap Fellowship at the Dunlap Institute for
-% Astronomy \& Astrophysics, funded through an endowment established by
-% the Dunlap family and the University of Toronto.
-% %
-% D.D.\ acknowledges support through the TESS Guest Investigator Program
-% Grant 80NSSC19K1727.
-%
-%
-%
-% %
-% Based on observations obtained at the Gemini Observatory, which is
-% operated by the Association of Universities for Research in Astronomy,
-% Inc., under a cooperative agreement with the NSF on behalf of the
-% Gemini partnership: the National Science Foundation (United States),
-% National Research Council (Canada), CONICYT (Chile), Ministerio de
-% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
-% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
-% and Korea Astronomy and Space Science Institute (Republic of Korea).
-% %
-% Observations in the paper made use of the High-Resolution Imaging
-% instrument Zorro at Gemini-South. Zorro was funded by the NASA
-% Exoplanet Exploration Program and built at the NASA Ames Research
-% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
-% Quigley.
-% %
-% This research has made use of the VizieR catalogue access tool, CDS,
-% Strasbourg, France. The original description of the VizieR service was
-% published in A\&AS 143, 23.
-% %
-% This work has made use of data from the European Space Agency (ESA)
-% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
-% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
-% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
-% for the DPAC has been provided by national institutions, in particular
-% the institutions participating in the {\it Gaia} Multilateral
-% Agreement.
-%
-% (Some of) The data presented herein were obtained at the W. M. Keck
-% Observatory, which is operated as a scientific partnership among the
-% California Institute of Technology, the University of California and
-% the National Aeronautics and Space Administration. The Observatory was
-% made possible by the generous financial support of the W. M. Keck
-% Foundation.
-% The authors wish to recognize and acknowledge the very significant
-% cultural role and reverence that the summit of Maunakea has always had
-% within the indigenous Hawaiian community.  We are most fortunate to
-% have the opportunity to conduct observations from this mountain.
-%
-%\newline
-%
-
-\software{
-  %\texttt{arviz} \citep{arviz_2019},
-  %\texttt{altaipony} \citep{ilin_flares_2021},
-  %\texttt{astrobase} \citep{bhatti_astrobase_2018},
-  %\texttt{astroplan} \citep{astroplan2018},
-	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
-  \texttt{astropy} \citep{astropy_2018},
-  \texttt{astroquery} \citep{astroquery_2018},
-  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
-  %\texttt{ceres} \citep{brahm_2017_ceres},
-  %\texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
-  %\texttt{corner} \citep{corner_2016},
-  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
-  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
-  dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
-   	exoplanet:theano},
-	%\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
-	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
-  %\texttt{IPython} \citep{perez_2007},
-	%\texttt{isochrones} \citep{morton_2015_isochrones},
-	%\texttt{lightkurve} \citep{lightkurve_2018},
-  %\texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
-  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
-  %\texttt{numpy} \citep{walt_numpy_2011}, 
-  %\texttt{pandas} \citep{mckinney-proc-scipy-2010},
-  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
-  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
-  %\texttt{radvel} \citep{fulton_radvel_2018},
-  %\texttt{scikit-learn} \citep{scikit-learn},
-  %\texttt{scipy} \citep{jones_scipy_2001},
-  %\texttt{TESS-point}  \citep{burke_2020},
-  \texttt{tesscut} \citep{brasseur_astrocut_2019},
-  \texttt{unpopular} \citep{hattorio_2021_cpm},
-\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
-  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
-  %\texttt{wotan} \citep{hippke_wotan_2019}.
-}
-\
-
-\facilities{
- 	{\it Astrometry}:
- 	Gaia. % \citep{gaia_collaboration_gaia_2018,gaia_collaboration_2021_edr3}.
- 	{\it Imaging}:
-    Second Generation Digitized Sky Survey. %,
-    %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
- 	Keck:II~(NIRC2).
- 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
- 	%Gemini:North~(`Alopeke; \citealt{scott_nessi_2018,scott_twin_2021}.
- 	{\it Spectroscopy}:
-	%CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
-  %PFS ({\bf CITE}),
-	Tillinghast:1.5m~(TRES).
-  %MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
-	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
-	%AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
- 	Keck:I~(HIRES).
- 	%VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
-% 	Euler1.2m~(CORALIE),
-% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
- 	{\it Photometry}:
-%	  ASTEP:0.40$\,$m (ASTEP400),
-% 	CTIO:1.0m (Y4KCam),
-% 	Danish 1.54m Telescope,
-%	  El Sauce:0.356$\,$m,
-% 	Elizabeth 1.0m at SAAO,
-% 	Euler1.2m (EulerCam),
-	  Kepler,
-% 	Magellan:Baade (MagIC),
-% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
-%   MuSCAT3 \citep{Narita_2020},
-% 	NTT,
-% 	SOAR (SOI),
- 	  TESS,
-% 	TRAPPIST \citep{jehin_trappist_2011},
-% 	VLT:Antu (FORS2).
-    ZTF.
-}
 
@@
 
-\appendix
-\section{Candidate Cep-Her Members}
-\label{app:members}
-
-\paragraph{Table~2} contains 338 candidate Cep-Her members with weights
-$D>0.02$ observed by Kepler.  The complete catalog of candidate
-Cep-Her members will be provided by R.~Kerr et al.\ in prep.\ using Gaia DR3; Table~2 is
-from an early version of that analysis based on Gaia EDR3.  Note that more
-restrictive weight cuts should be imposed if one wishes to remove
-the majority of field star interlopers.
-Table~2 was created by cross-matching candidate Cep-Her members
-(selected using Gaia EDR3; Section~\ref{subsec:members}) against a
-Kepler to Gaia DR2 cross-match (the \texttt{gaia-kepler.fun}
-crossmatch database created by Megan Bedell).  The
-\texttt{kic\_dr2\_ang\_dist} column is from the latter table.  The
-EDR3 to DR2 match was performed using the
-\texttt{gaiaedr3.dr2\_neighbourhood} table, and the closest proper
-motion and epoch-corrected angular distance neighbor was taken as the
-single best match.  The \texttt{edr3\_dr2\_mag\_diff} column gives
-some indication of the reliability of this EDR3 to DR2 conversion, as
-there are a few cases between Gaia DR2 and EDR3 where partially
-resolved binaries became fully resolved.
-
-\paragraph{Candidate matches between Cep-Her and the Kepler Objects of
-Interest:}
-The full list of candidate matches between Cep-Her and the Kepler
-Objects of Interest is as follows -- the objects are listed in order
-of descending weights, $D$.
-Objects designated as confirmed planets included
-Kepler-1627,
-Kepler-1643,
-Kepler-1331,
-Kepler-1062, and
-Kepler-1933.  % Newly validated by Valizadegan+2022
-Objects designated as candidate planets included
-KOI-5264,
-KOI-8007,
-KOI-7572,
-KOI-7375,
-KOI-7368,
-KOI-7638,
-KOI-5632, and
-KOI-7913.
-Objects designated known false positive planet candidates included
-KOI-6437, 
-KOI-5988, 
-KOI-7871,
-KOI-7655,
-KOI-5024,
-KOI-61,
-KOI-4336,
-KOI-6812,
-KOI-3399, and
-KOI-6277.
-Finally, Kepler-1902 (KOI-3090) has one confirmed planet
-(KOI-3090.02), and one false positive (KOI-3090.01).
-Of these objects, only Kepler-1627, Kepler-1643, KOI-7368, and
-KOI-7913 met our requirements for potentially
-both {\it i)} having real planets, and {\it ii)} being $\lesssim 10^8$
-years old, based on the presence of rotational modulation at the
-expected period and amplitude.  
-Of the
-14 confirmed and candidate planets,
-6 failed the first filter, and 7 independently failed the second.
-One object was ambiguous: Kepler-1933.
-This system has a confirmed $\approx$$1.4\,R_\oplus$ planet, a stellar
-rotation period of 6.5\ days, and an effective temperature of $\approx$$5750\,{\rm K}$.  This places it near the upper envelope of the
-rotation period vs.\ color distribution for the Pleiades, making
-it unlikely to be $\approx$40\,Myr old.  Nonetheless, we acquired a
-reconnaissance HIRES spectrum, and it yielded ${\rm EW}_{\rm Li} = 93 \pm
-5$\,m\AA.  Combined with the rotation period, this suggests an age
-for Kepler-1933 between 100 and 300\,Myr.  Based on these indicators, the system is
-unlikely to be part of Cep-Her, but could merit further study.
-
-\paragraph{Table~3} contains spatial, kinematic, astrometric, and
-rotation period information for the \nrsgfive\ candidate RSG-5 members
-and \nchtwo\ candidate CH-2 members described in
-Section~\ref{subsec:members}.  These are the data used to make the
-lower panels of Figure~\ref{fig:age}; as with Table~2, these are
-from a preliminary version of the SPYGLASS 1\,kpc expansion (R. Kerr et
-al.\ in prep). 
-% Table~3 includes rotation periods for 53 stars with ZTF and 51 stars
-% with TESS, for a total of 71 periods reported for unique stars and 32
-% overlapping between the surveys.
-We adopted the ZTF period over the
-TESS period in three cases: (1) Gaia EDR3 2081755809272821248: the top ZTF Lomb-Scargle peak gave 6.61 days, while our default pipeline favored a TESS peak of 13.34 days;  manual inspection of the light curve favors the former; (2) Gaia EDR3 2081737529891330560: we found 3.06
-days with TESS and 6.64 days with ZTF; we suspect that TESS captured
-the 1/2-period harmonic and adopt the approximately double value from
-ZTF; (3) 2134851775526125696: for this star, we measured 1.91 days
-with TESS from Cycle 2, but noted that the signal appeared to be
-missing in Cycle 4; ZTF found a strong signal at 12.23 days and we
-adopt this as the star's period. In the remaining overlap cases, we
-adopted the average between TESS and ZTF as the final period. For
-these overlap stars, the median absolute deviation is 0.01 days,
-showing remarkable consistency between the surveys. For three stars,
-we failed to detect a period in TESS but recovered one from ZTF; in
-all cases the periods appear to be 13--16 days.  These stars were: (1)
-Gaia EDR3 2129930258400157440, for which TESS showed a flat light
-curve while ZTF yielded a 15.3-day period; (2) Gaia EDR3
-2082376861542398336, LS found a 7.6-day period which we rejected
-during visual validation; we found 15.4 days with ZTF, and we suspect
-that the weak/rejected signal form TESS might have been a 1/2 period
-harmonic; (3) Gaia EDR3 2082397099429013120, similar to the previous
-case, we rejected a 6.7-day signal from TESS and recovered a 12.8-day
-period with ZTF.
-
-\section{Table of Transit Fit Parameters}
-\label{app:transit}
-
-Table~4 gives the full set of fitted and derived
-parameters from the model described in Section~\ref{sec:fitting}.
-Priors and convergence statistics are also listed.
-
-\section{Disposition History of KOI-7913}
-\label{app:koi7913}
-
-The disposition of KOI-7913.01 has been debated: in
-\texttt{q1\_q17\_dr25\_koi} the source was flagged as a false
-positive, with the comment ``cent\_kic\_pos---halo\_ghost''.  This
-comment and disposition were removed in the
-\texttt{q1\_q17\_dr25\_sup\_koi} data release, which renamed the
-planet a ``candidate''.  In this note, we discuss the interpretation
-of these flags (which do not apply to the system, according to the
-latest analysis).  We also discuss how the relative on-sky positions of
-KOI-7913~A and KOI-7913~B affect the interpretation of the Kepler
-data.
-
-As described by \citet{thompson_planetary_2018}, the
-``cent\_kic\_pos'' flag is an indication that the measured source
-centroid is offset from its expected location in the Kepler Input
-Catalog.  The final Kepler data validation reports, generated 2016 Jan
-30, do not show this to be the case for KOI-7913.  Moreover, the
-statistical significance of any centroid offset is lower than for
-KOI-7368 and Kepler-1643 (which both show centroid offsets that are
-likely explained by the stellar variability).
-
-What of the ``halo\_ghost'' flag?  This test measures the transit
-strength for the pixels inside the aperture, and compares it to that
-measured in the ring of pixels around said aperture (the ``halo'').
-One usually expects the transit signal to be strongest in the
-central aperture, rather than the halo.  Two types of false positive
-scenarios can change this and trigger the flag: the first is when
-optical ghosts from bright eclipsing binaries reflect off the CCD, and
-contaminate the target star.  The second is when the PRF of nearby
-stars directly overlaps with the PRF of the target star (see
-\citealt{thompson_planetary_2018}, Section A.5.2).  The most obvious
-explanation for KOI-7913 is the latter case, given that KOI-7913 B is
-$\approx0.9$ Kepler pixels away from Kepler-7913 A and so it usually
-part of the ``halo''.  Due to the on-sky orientation of KOI-7913 A and
-KOI-7913 B, the default ``optimal aperture'' selected in quarters 3,
-7, 11, and 15 in fact included both stars, while for the remaining
-quarters KOI-7913 B was excluded from the optimal aperture but was
-included as part of the halo (see pages 35 through 71 of the data
-validation reports.)
-
-Given the orientation of the stars and the $\approx$1.5 pixel FWHM of
-the Kepler pixel response function, some blending between the two
-stars is present.  The pointing geometries from quarters 3, 7, 11, and
-15 however did not affect the observed transit depths,
-which is an indication that the crowding metric applied in the data
-products accurately correct the mean flux level
-\citep{2017ksci.rept....6M}.  Analysis of the target-pixel data that
-was separately acquired for KOI-7913~B also reveals a
-different stellar rotation period, and no hint of the transit signal.
-
-% \clearpage
-\input{tab_supp_CepHer_X_Kepler.tex}
-\vspace{-6cm}
-\input{tab_supp_RSG5_CH2_Prot.tex}
-\input{source_KOIs_simpletransit_posteriortable.tex}
-
-\clearpage
-\listofchanges
+%\clearpage
+%\listofchanges
 %\allauthors
